### 2011-11-21 wp1description/src/w1p.description.tex
Sections 1 and 2

%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{article}

\title{Work Package 1 detailed description}
\date{\today}
\author{Peter Bloem}

\begin{document}

\maketitle

\begin{abstract}
A concise, non-formal description of the aims of Work Package 1 of Project 23
(Data2Semantics) of the COMMIT project. This documents provides the basic
intuitions required to understand the questions which this projects hopes to
answer.
\end{abstract}

Consider what we might mean by a `dataset'. A fMRI image of a subject's brain,
taken while the subject performs a specific task, a large body of annotated
language, TIFF images of the Taj Mahal, a collection of chess positions which
end in a draw. To be able to talk about data in the most general sense, we
suggest that any dataset can be translated to a single string of ones and zeroes
(as it is when stored on a computer system). We then build our basic concepts on
this description of data, so that we can be sure that they apply to everything
that we call data.

Any process, from simple operations like sorting a list of numbers, to learning
natural languages or discovering scientific knowledge, can be thought of as a
procedure on such a string of ones and zeroes. Any procedure that can be
described unambiguously and performed independent of any human interpretation or
ill-understood processes is called \emph{effective}. The notions of data encoded
as ones and zeroes, and effective procedures on such data have been thoroughly
studied over the last century.

This project will ask the following questions of such data:
\begin{itemize}
  \item Can we say, a priori, that some datasets are more interesting than
  others
  \item Can we describe a context in the broadest possible terms and define
  the interestingness or utility of a dataset withing such a context?
  \item What happens to the scientific process as we move into an era of
  increasingly large datasets? Will all scientific problems be solved simply by
  fast enough computer programs and large enough datasets, or will we find new
  limits that will defeat supercomputers and petabyte-sized datasets?
\end{itemize}

\section{Question 1: Complexity and Interestingness}

Consider the  following two datasets:
\indent101010101010101010101010101010101010101010101010101010101010101010101010101010\ldots
\indent101000110111111001110001011111010000001000101000110000111101010100101001100001\ldots
These examples represent extremes on a spectrum of interestingness. They are
sometimes called \emph{the clock} and \emph{the coinflip}. The first (the clock)
shows a high degree of regularity. It's perfectly predictable. The second shows
the highest possible degree of `chaos'. It's perfectly unpredictable.

There are many well defined notions of complexity. These measures tend to consider the coinflip maximally
complex and the clock maximally simple. While this certainly captures, for
instance, the effort required to describe the data, intuitively, it leaves
something to be desired.

Consider an old-fashioned black-and-white television. If we leave the
television off, the screen remains black. This is analogous to the clock above.
If we don't tune the receiver to a specific channel, we get a uniform static. By
most notions, he most complex signal the television is capable of displaying. It
is also commonly said that this signal contains the highest possible level of
information. That is were our intuition starts to confict with the theory.

Because while the static contains a lot of information in the sense that it is
difficult to describe exactly in simple terms, it is of no use to us. We are
much more interested in the television programs we see when we tune the dial to
a specific frequency. These are not entirely predictable and not entirely
unpredictable. We tend to be interested in the phenomena that lie somewhere
between the clock and the coinflip. Maximal information isn't enough, there must
be such a thing as `useful information'.

Intuitively, we guess that any datasets near the extremes will never be
interesting to anyone. How far can we push this? Can we narrow the band of
strings that are likely to be interesting? Can we suggest an optimum of a priori
interestingness somewhere between the clock and the coinflip? How far can we
take this without defining any kind of a context against which to measure
interestingness.

\section{Question 2: Conditional Complexity}

A priori interestingness can only go so far. A TIFF image of the Mona Lisa
may have a high a priori interestingness, but it's unlikely to to be of any use
to a theoretical physicist.

Determining interestingness on the basis of a context is easy to do in domain
specific terms. Given a well-formulated research question in molecular biology,
it should be simple enough to evaluate the utility of any dataset in answering
the research question, provided that the dataset is written in the same
`language' as the question. If we take the notion of `well-formed' questions to
its logical extreme, we end up with a computer program to be run on a dataset. A
query.

Can we generalize again? Ignore domain specific coding and programming and
simply consider the landscape of all datasets and all queries? The notions of
\epmh{data} and \emph{computer program} have been thoroughly generalized. The
concept of a question, or a query has (to our knowledge) not received the same
treatment.

\section{Question 3: Convergence}

\end{document}

### 2011-11-21 wp1description/src/w1p.description.tex
First draft finished

@@
-%%This is a very basic article template.
-%%There is just one section and two subsections.
 \documentclass{article}
 
-\title{Work Package 1 detailed description}
+\usepackage{charter}
+
+\title{Work Package 1: Description}
 \date{\today}
 \author{Peter Bloem}
+
+% non-indented, spaced paragraphs
+\setlength{\parindent}{0.0in}
+\setlength{\parskip}{0.1in}
 
@@
 \begin{abstract}
-A concise, non-formal description of the aims of Work Package 1 of Project 23
-(Data2Semantics) of the COMMIT project. This documents provides the basic
-intuitions required to understand the questions which this projects hopes to
-answer.
+\noindent A concise, non-formal description of the aims of Work~Package~1 of
+Project~23 (Data2Semantics) of the {\sc commit} project.
 \end{abstract}
 
-Consider what we might mean by a `dataset'. A fMRI image of a subject's brain,
-taken while the subject performs a specific task, a large body of annotated
-language, TIFF images of the Taj Mahal, a collection of chess positions which
-end in a draw. To be able to talk about data in the most general sense, we
-suggest that any dataset can be translated to a single string of ones and zeroes
-(as it is when stored on a computer system). We then build our basic concepts on
-this description of data, so that we can be sure that they apply to everything
-that we call data.
-
-Any process, from simple operations like sorting a list of numbers, to learning
-natural languages or discovering scientific knowledge, can be thought of as a
-procedure on such a string of ones and zeroes. Any procedure that can be
-described unambiguously and performed independent of any human interpretation or
-ill-understood processes is called \emph{effective}. The notions of data encoded
-as ones and zeroes, and effective procedures on such data have been thoroughly
-studied over the last century.
-
-This project will ask the following questions of such data:
+This project will ask the following questions:
 \begin{itemize}
-  \item Can we say, a priori, that some datasets are more interesting than
-  others
-  \item Can we describe a context in the broadest possible terms and define
-  the interestingness or utility of a dataset withing such a context?
+  \item Can we say, without relying on context, that some datasets are more interesting than others?
+  \item Can we describe the notion of a context in the broadest possible terms and define the interestingness or utility of a dataset given such a description?
   \item What happens to the scientific process as we move into an era of
@@
 
-\section{Question 1: Complexity and Interestingness}
+The first two questions serve to create a framework to pose and answer the third in a rigorous manner.
+Within context of Project~23, the first two questions are most likely to produce techniques that may benefit the other work packages, while the third question relates more to the enterprise of e-science as a whole.
+
+\section*{Framework}
+
+Consider what we might mean by a `dataset'.
+A large body of annotated language.
+An medical scan of a patient's brain.
+Photographs of the Taj Mahal.
+A collection of chess positions which end in a draw.
+To be able to talk about data in the most general sense, we suggest that anything we might call data can be translated to a single string of ones and zeroes.
+If it can't be, then it cannot be stored on a computer.
+We then build our theoretical constructs on only this assumption about the structure of the data.
+That way we can be sure that dataset will be subject to our conclusions.
+
+Any process, from simple operations like sorting a list of numbers, to learning natural languages or discovering scientific knowledge, can be thought of as a procedure on such a string of ones and zeroes.
+Any procedure that can be described unambiguously and performed independently of any human interpretation or ill-understood processes is called \emph{effective}.
+It is a deep and open question whether processes that cannot be described by an effective procedure can exist at all.
+It is less controversial by far to say that science deals only with effective procedures.
+
+Thus, we will investigate effective procedures on strings of ones and zeroes.
+This field of mathematics is called Computer Science, and it has been thoroughly studied for over a century.
+Anything that holds true in this framework of an effective procedure on a string of ones and zeroes will be true for all scientific process.
+
+\pagebreak[0]
+
+\section*{{\sc Question 1}: Complexity and Interestingness}
 
 Consider the  following two datasets:
-\indent101010101010101010101010101010101010101010101010101010101010101010101010101010\ldots
-\indent101000110111111001110001011111010000001000101000110000111101010100101001100001\ldots
+\begin{center}
+\indent101010101010101010101010101010101010101010101010101\ldots
+
+\indent101000110111111001110001011111010000001000101000110\ldots
+
+\end{center}
 These examples represent extremes on a spectrum of interestingness. They are
 sometimes called \emph{the clock} and \emph{the coinflip}. The first (the clock)
-shows a high degree of regularity. It's perfectly predictable. The second shows
-the highest possible degree of `chaos'. It's perfectly unpredictable.
+shows a high degree of regularity. It's perfectly predictable. The second
+(the coinflip) shows the highest possible degree of `chaos'. It's perfectly
+unpredictable.
 
-There are many well defined notions of complexity. These measures tend to consider the coinflip maximally
-complex and the clock maximally simple. While this certainly captures, for
-instance, the effort required to describe the data, intuitively, it leaves
-something to be desired.
+There are many well defined notions of complexity. These measures tend to
+consider the coinflip maximally complex and the clock maximally simple. While 
+this certainly captures, for instance, the effort required to describe the data,
+intuitively, it leaves something to be desired.
 
@@
 
-Because while the static contains a lot of information in the sense that it is
-difficult to describe exactly in simple terms, it is of no use to us. We are
-much more interested in the television programs we see when we tune the dial to
-a specific frequency. These are not entirely predictable and not entirely
-unpredictable. We tend to be interested in the phenomena that lie somewhere
+The static may be complex, or complicated, it is of no use to us. We are
+much more interested in the television programs we see when we tune the
+television to a specific frequency. These are not entirely predictable and not
+entirely unpredictable. We tend to be interested in the phenomena that lie somewhere
 between the clock and the coinflip. Maximal information isn't enough, there must
@@
 
-\section{Question 2: Conditional Complexity}
+\section*{Question 2: Conditional Complexity}
 
@@
 simply consider the landscape of all datasets and all queries? The notions of
-\epmh{data} and \emph{computer program} have been thoroughly generalized. The
-concept of a question, or a query has (to our knowledge) not received the same
-treatment.
+\emph{data} and \emph{computer program} have been thoroughly generalized. The
+concept of a question, or a query has (to our knowledge) not yet received the
+same treatment.
 
-\section{Question 3: Convergence}
+Consider a simple question like ``When does the supermarket close''.
+This question defines a hierachy of three classes.
+The class of \emph{replies}, which is simply anything that can be said in response.
+A subset of that is the class of \emph{answers}, which is anything that fits the question.
+A subset of answers is the set of \emph{correct answers}.
+
+An reply like ``blue'' can be rejected without any verification based on data.
+I's a reply, but it's not an answer.
+The reply ``three o'clock'' is an an answer.
+It satisfies the expectations defined by the question.
+To check whether or not it's the correct answer, we need consult the data (ie.
+the table of closing times on the store's website).
+Depending on the dataset, the correct answer may be different.
+
+*** Comments:
+I don't know if this will pan out, but I can see connections to two-part coding, with the query serving the part of the model (defining a set of answers), and the correct answer serving the part of the data given the model.
+
+Either way, I think the possiblity of `conditional interestingness' is very relevant, and the connection to queries on data ties it in to the original work package description.
+***
+\section*{Question 3: Convergence}
+
+Consider a linguist trying to build a model of the English language.
+This model (a grammar) is basically an effective preocedure for determining whether sentences are grammatically correct.
+The linguist finds that a simple grammar, taking up 1 kbit on his laptop, will capture about half of the sentences he finds on the internet.
+He wishes to increase this coverage, so he keeps working on the grammar.
+By the time it captures 3/4 of sentences correctly, its size has grown to 10 kBit.
+The linguist finds that to capture an additional 1/8th, the model size needs to increase to 100 kBit.
+the next 1/16 of sentences requires another ten-fold increase in complexity.
+
+At this point the linguist will likely lose his enthusiasm for the project.
+It has long been a property of scientific research that models converge.
+Physics is a prime example.
+Newtons laws were built on a minimum of data and survived for centuries of additional experimentation.
+When it did finally show its weaknesses, The Theory of Relativity was discovered and survived the explosion of data-volume that the invention of the computer brought us.
+
+If the linguist were as lucky as the physicist, his 1 kB grammar would have been more than enough to capture all the language he could record.
+And if it did eventually fail after he had collected ten times as many sentences as he had started with, a new model of 2 kB would suffice to model the new data as well.
+
+It would seem that not all sciences are created equal.
+While the linguist struggles with his growing grammars, the biologist next door must come to terms with the fact that when he doubles his sample size, he will see twice as many diseases.
+For some fields of science, increasing the amount of data will only increase the number of questions.
+And while the computer revolution has so far given us exponentially increasing power, we are fast approaching the physical limits of computation.
+
+At some point the computational power required to infer some answer from a dataset will simply not be sustainable in our universe.
+We hope to build a framework to formalize such statements, and possibly estimate where various datasets fall in the landscape of convergent and non-convergent models.
+
+\subsection*{Power Laws}
+
+Question 3 relates to 1 and 2, mostly through the notion of a \emph{power law}.
+Power laws relate the probability that some event will occur, to some property of that event.
+For instance, there is a power law relating the probability of an earthquake to its size.
+A large magnitude earthquake is more rare than a small magnitude earthquake, and the relation between probability and magnitude follows roughly the following relation
+\[
+p = m^{-\alpha}
+\] 
+where $\alpha$ is the \emph{exponent}, the most important parameter of the distribution.
+Other phenomena which follow a power law distribution are the size of an individual's wealth, the population size of cities and the size of craters on the moon.
+
+Intuitively, the example of the linguist building progressively more complex grammars calls to mind a power law.
+The relation here is between probability of seeing a sentence, and the complexity of the model required to capture this sentence.
+
+If we can formalize this picture, we can relate probability of events to required model complexity, and make suggestions for which fields will benefit from increased data acquisition, and which will will need to manage their expectations.
 

### 2011-11-24 random/src/random_sets.tex
Stub

--- 2011-11-21 wp1description/src/w1p.description.tex
@@
-%%This is a very basic article template.
-%%There is just one section and two subsections.
 \documentclass{article}
 
-\title{Work Package 1 detailed description}
+\usepackage{charter}
+
+\title{Investigating the random set}
 \date{\today}
 \author{Peter Bloem}
+
+% non-indented, spaced paragraphs
+\setlength{\parindent}{0.0in}
+\setlength{\parskip}{0.1in}
 
@@
 \begin{abstract}
-A concise, non-formal description of the aims of Work Package 1 of Project 23
-(Data2Semantics) of the COMMIT project. This documents provides the basic
-intuitions required to understand the questions which this projects hopes to
-answer.
+\noindent Brief, formal definitions of various flavors of Kolmogorov Complexity, and an investigation of the set of random strings it defines.
 \end{abstract}
 
-Consider what we might mean by a `dataset'. A fMRI image of a subject's brain,
-taken while the subject performs a specific task, a large body of annotated
-language, TIFF images of the Taj Mahal, a collection of chess positions which
-end in a draw. To be able to talk about data in the most general sense, we
-suggest that any dataset can be translated to a single string of ones and zeroes
-(as it is when stored on a computer system). We then build our basic concepts on
-this description of data, so that we can be sure that they apply to everything
-that we call data.
-
-Any process, from simple operations like sorting a list of numbers, to learning
-natural languages or discovering scientific knowledge, can be thought of as a
-procedure on such a string of ones and zeroes. Any procedure that can be
-described unambiguously and performed independent of any human interpretation or
-ill-understood processes is called \emph{effective}. The notions of data encoded
-as ones and zeroes, and effective procedures on such data have been thoroughly
-studied over the last century.
-
-This project will ask the following questions of such data:
-\begin{itemize}
-  \item Can we say, a priori, that some datasets are more interesting than
-  others
-  \item Can we describe a context in the broadest possible terms and define
-  the interestingness or utility of a dataset withing such a context?
-  \item What happens to the scientific process as we move into an era of
-  increasingly large datasets? Will all scientific problems be solved simply by
-  fast enough computer programs and large enough datasets, or will we find new
-  limits that will defeat supercomputers and petabyte-sized datasets?
-\end{itemize}
-
-\section{Question 1: Complexity and Interestingness}
-
-Consider the  following two datasets:
-\indent101010101010101010101010101010101010101010101010101010101010101010101010101010\ldots
-\indent101000110111111001110001011111010000001000101000110000111101010100101001100001\ldots
-These examples represent extremes on a spectrum of interestingness. They are
-sometimes called \emph{the clock} and \emph{the coinflip}. The first (the clock)
-shows a high degree of regularity. It's perfectly predictable. The second shows
-the highest possible degree of `chaos'. It's perfectly unpredictable.
-
-There are many well defined notions of complexity. These measures tend to consider the coinflip maximally
-complex and the clock maximally simple. While this certainly captures, for
-instance, the effort required to describe the data, intuitively, it leaves
-something to be desired.
-
-Consider an old-fashioned black-and-white television. If we leave the
-television off, the screen remains black. This is analogous to the clock above.
-If we don't tune the receiver to a specific channel, we get a uniform static. By
-most notions, he most complex signal the television is capable of displaying. It
-is also commonly said that this signal contains the highest possible level of
-information. That is were our intuition starts to confict with the theory.
-
-Because while the static contains a lot of information in the sense that it is
-difficult to describe exactly in simple terms, it is of no use to us. We are
-much more interested in the television programs we see when we tune the dial to
-a specific frequency. These are not entirely predictable and not entirely
-unpredictable. We tend to be interested in the phenomena that lie somewhere
-between the clock and the coinflip. Maximal information isn't enough, there must
-be such a thing as `useful information'.
-
-Intuitively, we guess that any datasets near the extremes will never be
-interesting to anyone. How far can we push this? Can we narrow the band of
-strings that are likely to be interesting? Can we suggest an optimum of a priori
-interestingness somewhere between the clock and the coinflip? How far can we
-take this without defining any kind of a context against which to measure
-interestingness.
-
-\section{Question 2: Conditional Complexity}
-
-A priori interestingness can only go so far. A TIFF image of the Mona Lisa
-may have a high a priori interestingness, but it's unlikely to to be of any use
-to a theoretical physicist.
-
-Determining interestingness on the basis of a context is easy to do in domain
-specific terms. Given a well-formulated research question in molecular biology,
-it should be simple enough to evaluate the utility of any dataset in answering
-the research question, provided that the dataset is written in the same
-`language' as the question. If we take the notion of `well-formed' questions to
-its logical extreme, we end up with a computer program to be run on a dataset. A
-query.
-
-Can we generalize again? Ignore domain specific coding and programming and
-simply consider the landscape of all datasets and all queries? The notions of
-\epmh{data} and \emph{computer program} have been thoroughly generalized. The
-concept of a question, or a query has (to our knowledge) not received the same
-treatment.
-
-\section{Question 3: Convergence}
-
 \end{document}

### 2011-11-24 power laws/src/power_laws.tex
Stub
Outline

--- 2011-11-24 random/src/random_sets.tex
@@
 \usepackage{charter}
+\usepackage{eulervm}
 
-\title{Investigating the random set}
+\title{Power laws and other fat tails}
 \date{\today}
@@
 \begin{abstract}
-\noindent Brief, formal definitions of various flavors of Kolmogorov Complexity, and an investigation of the set of random strings it defines.
+\noindent A thorough definition of the various types of power laws that exist, their realtions to one another and the general properties.
 \end{abstract}
 
+\section*{Power laws in general: functions with negative exponentials}
+\subsection*{Pure power laws}
+
+\[
+f(x) = x^{-\alpha}
+\]
+
+\subsection*{Approximate power laws}
+
+\[
+f(x) = \phi(x)\;\cdot\;x^{-\alpha}
+\]
+
+\subsection*{Broken power law}
+
+\subsection*{Exponential cutoff}
+
+\section*{Power law distributions}
+
+\subsection*{Attribute-based}
+
+\subsubsection*{Pareto distribution}
+
+\subsection*{Rank: probability as an attribute}
+
+\subsubsection*{Zipf distribution}
+
+\section*{Testing for power laws}
+
+\section*{Fat tailed distributions}
+
+\section*{Examples and applications}
+
+This section provides some examples of power law and fat-tail phenomena and classifies them more exactly according to the definitions in this document.
+
+\section*{Dimension}
+
 \end{document}

### 2011-12-05 random/src/random_sets.tex
Added timesheet

@@
 \usepackage{charter}
+\usepackage{eulervm}
+\usepackage{amsmath, amsthm, amssymb}
 
-\title{Investigating the random set}
+\theoremstyle{definition}
+\newtheorem*{thm}{Definition}
+\newtheorem*{dfn}{Definition}
+\newtheorem*{exm}{Example}
+
+\title{Kolmogorov Complexity}
 \date{\today}
@@
 \begin{abstract}
-\noindent Brief, formal definitions of various flavors of Kolmogorov Complexity, and an investigation of the set of random strings it defines.
+\noindent This document provides a brief description of Kolmogorov Complexity and Marting L\o:f randomness.
+It may do some experimental tweking of notation and traditional definition, to highlight the important points in the development of Kolmogorov complexity that are often glossed over.
 \end{abstract}
 
+\section*{Notation and preliminaries}
+
+$\cal{B}$ is the set of finite binary strings, including the empty string $\epsilon$. Members of $\cal B$ tend to be named $x$, $y$ and $z$.
+
+A subset of $\cal B$ has the \textbf{prefix-property} if no member of the set is a prefix of another member.
+We fix one such subset $\cal P$ as a canonical example.
+Members of $\cal{P}$ tend to be named $\bar{x}$, $\bar{y}$, $\bar{z}$.
+\footnote{Note that this is different from the convention of having $\bar{x}$ be the prefix free encoding of $x$.}
+
+If we fix some canonical ordering of $\cal B$, we can match each finite binary string to a natural number, thus ordering the binary string.
+We will sometimes consider the two interchangeable.
+Note that this also serves as an ordering for all subsets of $\cal A$.
+Let ${\cal A}_i$ be the $i$-th member of $\cal A$.
+For any binary string $x$, let ${\mathbb N}_x$ be the corresponding natural number by the agreed ordering.
+
+Each subset of $\cal{B}$ encodes binary strings, we name this function after its set as follows: if ${\cal A} \subseteq {\cal B}$ then ${\cal A}({\cal B}_i) = {\cal A}_i$.
+Thus, ${\cal P}(x)$ is the prefix encoding of $x$.
+
+If $x$ and $y$ are binary strings, then $xy$ represents the concatenation of the two.
+
+\section*{The universal Turing machine}
+
+\begin{dfn}
+A \textbf{pairing function} $\langle\,\,\cdot,\,\cdot\,\rangle_f$ is any function $f: \mathbb{N} \times \mathbb{N} \rightarrow \mathbb{N}$ which is \emph{primitive recursive} and \emph{invertible}.
+\end{dfn}
+
+\begin{exm} The G\"{o}del pairing function is defined as $\langle a, b \rangle_g = 2^a3^b$.
+Since $2$ and $3$ are prime, each $a$ and $b$ will result in a product of a different sequence of primes, and thus a unique integer.
+\end{exm}
+
+\begin{exm} The \em{prefix pairing function} $\langle a, b \rangle_p = {\cal P}(a)b$ for any set ${\cal P}$ with the prefix property.
+Since a is prefix encoded, there is only one pair of inputs to each output.
+[Is this actually primitive recursive?]
+\end{exm}
+
+\begin{thm}
+The set of all pairing functions is recursively enumerable.
+\end{thm}
+
+\begin{dfn}
+A Turing machine $U$ is called universal if there is some pairing function $f$ and enumeration ${T_i}$ of Turing machines, such that if $x = \langle i, p\rangle$ then $U(x) = T_i(p)$.
+That is, a universal Turing machine can simulate any other turing machine running any other program.
+\end{dfn}
+
+If we think of Turing machines as computing a sequence of elements from ${\mathbb N} \cup \{\infty\}$, then a universal Turing machine weaves all these sequences together in a way that is easily decoded.
+
 \end{document}

### 2012-02-15 Sawtooth/src/sawtooth.tex
Stuff

%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{article}

\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath, amsthm, amssymb}

\theoremstyle{definition}
\newtheorem{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem*{dfn}{Definition}
\newtheorem*{exm}{Example}

\title{Kolmogorov Complexity}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle

\begin{abstract}
\noindent This article aims to highlight some of the basic properties of the sawtooth diagram for facticity versus complexity.
\end{abstract}

\section*{The random strings}

As is well known, the majority of strings is random. Thus, their minimal model is that of all strings (ie. the copying Turing Machine). As n goes to infinity, the set has density 1. That means that if we create a density plot of an approximation of $K(x)$ versus $\phi(x)$ for all strings of a reasonable length, we will only see a single point in the bottom right of the diagram.

If on the other hand, we do a scatter plot we are likely to see the rough structure of the facticity curve, but in the limit, the whole area below $K(x) = \phi(x)$ and $\phi(x) = \frac{1}{2}$ will be filled, obscuring most of the interesting detail.

To study the structure of this diagram in detail, we need to restrict ourselves to a subsets of strings.

\section*{The simple Shannon models}

To start with, we can assume that we are studying compressible strings with an 
uneven number of ones and zeroes. For such strings, we have a simple compression scheme: store the frequency of ones in a prefix code and then code the rest of the string in $nH(x)$ bits.

\section*{Complex Shannon models}

\end{document}

### 2012-02-15 opmerkingen/src/opmerkingen.tex
Stuff

%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{article}

\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath, amsthm, amssymb}

\theoremstyle{definition}
\newtheorem{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem*{dfn}{Definition}
\newtheorem*{exm}{Example}

\title{Facticity: opmerkingen}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\DeclareMathOperator*{\argmin}{arg\,min}

\begin{document}

\maketitle

\section{Lemma 3}

Volgens mij klopt dit bewijs niet. I neem aan dat ``$|\bar{\imath}| + |p| > |p|$'' zoiets moet zijn als ``$|\bar{\imath}| + |p| > |q|$''. De $p$ van $K_2$ is immers niet dezelfde als die van $C$.

Vervolgens zeg je $U(\bar{\epsilon}p) = x$. Volgens een eerdere definitie is $U(\bar{\epsilon}x) = x$. Het werkt als je de prefix code van de $U$ van $C$ zelf neemt (zeg $u$): $U(\bar{u}p) = x$. In dat geval stelt het lemma $C(x) + |\bar{u}| \geq K_2(x)$.

Overigens is dit volgens mij alleen nodig als $C$ en $K_2$ gedefinieerd zijn tov verschillende $U$'s. Als ze allebei gebruik maken van de prefix UTM, dan heb je altijd $\bar{\imath}p = q$.

\section{Nickname problem}

Ik vraag me af of deze oplossing vh. nickname problem werkt. Specifiek het idee om uit te gaan van de Kolmogorov Complexiteit van de canonieke beschrijving van de TM.

Ik kan een beschrijving van een Turing Machine arbitrair complexer maken door een routine toe te voegen die 100000 random bits print en weer verwijdert. Dit draagt niet bij aan de modelinformatie (dus de beschrijving is 100k bits minder 'faithful'), maar de Kolmogorov complexiteit groeit wel met 100000 bits. De Kolmogorov complexiteit van T is dus niet per se een goed ideaal van de model informatie.

Ik ben er nog niet helemaal uit, maar het lijkt me dat je dit probleem op moet lossen op basis van functionele eigenschappen ipv. op basis van de programma's.

In Lemma 6 heb je het over een optimaal programma, maar ik kan geen definitie vinden? Is dit een standaard term? Ik heb het in deze context nog nooit gehoord.

Ik vraag me bij Lemma 6 ook af of het wel terecht is om te zeggen dat de dichtheid van random strings in de limiet 1 is, omdat beschrijvingen niet random gegenereerd worden. Als ik een enumeratie maak op basis van bijv C programma's, dan zijn de beschrijvingen allemaal sterk comprimeerbaar.

\section{Definitie 8}

Volgens werkt de definitie zoals hij er nu staat alleen als je de notatie niet helemaal letterlijk neemt. Volgens wikipedia is een argmin altijd over een functie (ipv een set) en in het predicaat is $i$ geen vrije variabele.

Dit is het beste wat ik kon bedenken:
\[
\varphi(x) = \min \left \{|i|\ |\ \exists{p} : |\bar{\imath}p| = K_2(x), \,U(\bar{\imath}p) = x\right\}
\]

of misschien
\[
P(x) = \left\{(i,p)\ |\ U(\bar{\imath}p) = x\right\}
\]
\[
k_2(x) = \argmin_{(i, p) \in P(x)} |\bar{\imath}p|
\]
\[
\varphi(x) = \min \left \{ i \ |\ (i, \cdot) \in k_2(x)\right\}
\]

\end{document}

### 2012-02-15 random/src/random_sets.tex
Stuff

@@
 \theoremstyle{definition}
-\newtheorem*{thm}{Definition}
+\newtheorem{thm}{Theorem}
+\newtheorem{lma}{Lemma}
 \newtheorem*{dfn}{Definition}
@@
 \begin{abstract}
-\noindent This document provides a brief description of Kolmogorov Complexity and Marting L\o:f randomness.
-It may do some experimental tweking of notation and traditional definition, to highlight the important points in the development of Kolmogorov complexity that are often glossed over.
+\noindent This document provides a brief description of Kolmogorov Complexity and Marting L\"of randomness.
+It may do some experimental tweaking of notation and traditional definition, to highlight the important points in the development of Kolmogorov complexity that are often glossed over.
 \end{abstract}
@@
 
-\section*{The universal Turing machine}
+\section*{Turing Machines}
+
+\ldots
+
+It is common to distinguish explicity between a Turing machine $T_i$ and the function which it computes. In this article we will forego this distinction and refer to both as $T_i$.
+
+For some purposes it is important to analyse the internal workings of Turing Machines as well as their output.
+For instance, for those Turing machines which never finish, but print out some infinite sequence in the limit, or for those problems where we are interested in running time as well as output.
+For the current article, we will only look at the input and output values (with $\infty$ as the `output' value for a non-halting Truing machine).
+If two machines have idental outputs for all inputs, then we shall consider them functionally equal.
+
+\section*{Universality}
+
+Informally, a universal Turing Machine (UTM) is a machine which combines the computational power of all other turing machines.
+Usually this means that the universal turing machine is given a description of another turing machine and a program, and then simulates the running of that machine internally and gives the output that the target machine would give for the program.
+
+Usually a reference UTM is described by giving a simple pairing functions to compose the two inputs and some proof is given that a UTM exists that decomposes the input, and simulates the target machine running the program.
+
+For the purposes of this article we would like to be able to decide which Turing Machines have this property of universality. There are different ways to defines this, and we will compare these.
 
@@
 
-\begin{thm}
+\begin{lma}
 The set of all pairing functions is recursively enumerable.
-\end{thm}
+\end{lma}
 
-\begin{dfn}
-A Turing machine $U$ is called universal if there is some pairing function $f$ and enumeration ${T_i}$ of Turing machines, such that if $x = \langle i, p\rangle$ then $U(x) = T_i(p)$.
+\begin{dfn}[Universally Simulating Turing Machine]
+A Turing machine $U$ is called \textbf{universally simulating} if there is some pairing function $f$ and enumeration ${T_i}$ of Turing machines, such that if $x = \langle i, p\rangle$ then $U(x) = T_i(p)$.
 That is, a universal Turing machine can simulate any other turing machine running any other program.
@@
 
+We call $U$'s specific pairing function $\langle \cdot, \cdot \rangle_U$ and drop the index if it is clear from context.
+
+This definition comes from the standard notion of a UTM simulating other turing machines.
+This is usually evaluated with respect to some specific pairing function.
+If we want to determine the notion of universality without reference to an arbitrary pairing function, then the question becomes whether a given Turing machine becomes universal if we decode the inputs with some pairing function.
+
+We might ask, for instance whether the identity function $T_I(x) = x$ is universal.
+If so, there should be a pairing function which separates all $x$ into a pair $\langle T_i, p\rangle$ such that $T_i(p) = x$.
+One of the many reasons that this function cannot exist is that each possible output $x$ occurs only once in the output of $T_i$, when is should be occuring for every $\langle i, p\rangle$ where $T_i(p) = x$.
+
 If we think of Turing machines as computing a sequence of elements from ${\mathbb N} \cup \{\infty\}$, then a universal Turing machine weaves all these sequences together in a way that is easily decoded.
 
+\begin{dfn}[Universally Minimizing Turing Machine]
+Consider the smallest program which produces $x$ on some Turing Machine $T$.
+Call this program $X^*_T$.
+For some $x$, $T$ might have the shortest program for $x$ over all Turing Machines.
+If, for all $x$, $T$'s shortest program is never more than a constant additive term away from any other Turing machine, T is \textbf{universally minimizing}.
+
+That is, T is universally minimizing if for all $x$ and any other Turing Machine $Q$:
+\[
+\left | l(x^*_T) - l(x^*_Q) \right | \leq c_{Q T}
+\]
+Where $c_{Q T}$ does not depend on $x$.
+
+This is also known as \em{universality by adjunction}
+\end{dfn}
+
+If $T$ does not produce $x$ for any input, then we say that $l(x^*_T) = \infty$.
+
+It is simple to see that a universally simulating Turing machine must be a universally minimizing Turing machine.
+If $p$ is the shortest program for $x$ on $Q$ then any universally simulating Turing machine $U$ can produce $Q$ as $U(\langle Q, p\rangle)$.
+Specifying $Q$ will cost $c_Q$ extra bits, independent of $x$, thus satisfying the requirement for a universal minimizing turing machine.
+
+The reverse---is every universally minimizing TM also universally simulating?---is more complicated.
+Any universally minimizing TM must at least be able to produce all possible outputs in ${\cal B} \cup \infty$.
+If not, the distance to a program that does produce this output would be infinite, and thus not bounded by a constant.
+
 \end{document}

### 2012-02-16 Nickname problem/src/nickname.tex
Nickname problem

%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{article}

\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath, amsthm, amssymb}

\theoremstyle{definition}
\newtheorem{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem*{dfn}{Definition}
\newtheorem*{exm}{Example}

\title{The Nickname Problem}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle
The \textbf{nickname problem} arises when we use an indexing of the computable functions as a mathematical representation of all possible models, and their complexity.

Consider, for instance, the string $\sigma$ representing the collected works of Shakespeare in binary form. We can define a simple enumeration of the recursive functions based on Turing Machines as usual, with the single change that a Turing Machine which outputs $\sigma$ (disregarding its input) is given index 20. While pathological, this index is perfectly computable, and $\sigma$ would be seen as a highly compressible string, consisting purely of model information.

We might also construct an enumeration which places a Turing machine at index 0 which produces the collected works of Shakespeare only for a high input with slightly smaller than any other description of $\sigma$. Under this enumeration, the string would be seen as reasonably compressible, but with only one bit of model information.

The nickname problem is inescapable. Kolmogorov Complexity itself suffers from a similar problem, where for a given string a pathological universal machine can be constructed which compresses that string to a single bit. Like with Kolmogorov complexity, we would like some assurance that this kind of pathological fiddling can only affect a limited number of string in a limited way. In other words, we can change the nature of facticity by switching to a different universal machine, but only to a limited extent.

\begin{dfn}
For a given index of the computable functions $\{\phi_i\}$ we define $i^* = \min\left\{\phi_j : \phi_j \simeq \phi_i\right\}$. That is, the minimal index of a function that is equivalent to $\phi_i$
\end{dfn}

\begin{dfn}
An index of the computable functions $\{\phi_i\}$ is called \textbf{compact} if for every 
$\phi_i(\bar{x}y)$ we have a $\phi_j$ such that $\phi_i(\bar{x}y) \simeq \phi_j(y)$ and  $|j| - |i| \leq |x| + O(1)$
\end{dfn}

Note that this is a natural property of basic computational metaphors. We can always take part of the argument of one function and hardcode it into the program at the cost of its length, and some additional bits to add it to the argument before processing.

A \textbf{non-compact} index would imply a computational metaphor which does not allow this type of manipulation.

\begin{lma}
Let $\{\phi_i\}$ and $\{\psi_j\}$ be two compact, computable enumerations of the computable functions. For a given computable function $\phi_i \simeq \psi_j$,  we have $\mbox{abs}(i^* - j^*) \leq O(1)$ 
\end{lma}
\begin{proof}
Let $\phi_u$ be a universal function for $\psi$, such that $\phi_u(\bar{\imath}x) \simeq \psi_i(x)$. Since $\psi$ is computable this function exists in $\phi$. We have $\psi_{j^*}(x) \simeq \phi_u(\overline{\jmath^*}x)$. By the compactness of $\phi$, we have some $h$ with $|h| \leq |u| + |\overline{\jmath^*}| + O(1)$ such that $\phi_{h}(x) \simeq \phi_u(\overline{\jmath^*}x) \simeq \psi_{j^*}(x)$.

\end{proof}

\end{document}

### 2012-02-24 Nickname problem/src/nickname.tex
Stuffstuffstuff

@@
 \begin{dfn}
-For a given index of the computable functions $\{\phi_i\}$ we define $i^* = \min\left\{\phi_j : \phi_j \simeq \phi_i\right\}$. That is, the minimal index of a function that is equivalent to $\phi_i$
+For a given index of the computable functions $\{\phi_i\}$ we define $i^*_\phi = \min\left\{\phi_j : \phi_j \simeq \phi_i\right\}$.
+That is, the minimal index of a function that is equivalent to $\phi_i$.
+The subscript may be dropped when it is clear from context.
 \end{dfn}
@@
 An index of the computable functions $\{\phi_i\}$ is called \textbf{compact} if for every 
-$\phi_i(\bar{x}y)$ we have a $\phi_j$ such that $\phi_i(\bar{x}y) \simeq \phi_j(y)$ and  $|j| - |i| \leq |x| + O(1)$
+$\phi_i(\bar{x}y)$ we have a $\phi_j$ such that $\phi_i(\bar{x}y) \simeq \phi_j(y)$ and  $|j| - |i| \leq |\bar{x}| + O(1)$
 \end{dfn}
@@
 
-A \textbf{non-compact} index would imply a computational metaphor which does not allow this type of manipulation.
+We can construct a non-compact enumeration by starting with a compact one, and inserting after function $i$ a series of $f(i)$ copies of $\phi_0$, where $f(i)$ is some fast growing function like $2^{i!}$.
 
@@
 Let $\{\phi_i\}$ and $\{\psi_j\}$ be two compact, computable enumerations of the computable functions.
-For a given computable function $\phi_i \simeq \psi_j$,  we have $\mbox{abs}(i^* - j^*) \leq O(1)$
+For a given computable function $\phi_i \simeq \psi_j$,  we have $\mbox{abs}(|i^*| - |j^*|) \leq c(\max(i^*, j^*)) + O(1)$.
+Where $c(x) \leq 2 \log(|x|) + 1$ is the overhead of prefix encoding $x$
 \end{lma}
@@
 
+Thus $|i^*| \leq |h| = |u| + |\overline{\jmath^*}| + O(1) = |j*| + c(\jmath^*) + O(1)$.
+Reversing $\phi$ and $\psi$ gives us $|j^*| \leq |i^*| + c(i^*) + O(1)$.
+Combining these, we get the desired result.
 \end{proof}
 
+This lemma tells us that so long as we restrict our numberings to be compact and computable, we can never gain more than a logarithmic difference between model sizes.
+In the limit as $i \rightarrow \infty$, this term represents a vanishing proportion.
+
 \end{document}

### 2012-02-24 Referee/document.tex
Stuffstuffstuff

\documentclass{article}

\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath, amsthm, amssymb}

\theoremstyle{definition}
\newtheorem*{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem*{dfn}{Definition}
\newtheorem*{exm}{Example}

\title{Review: {\em Lossless compressors: Degree of compression and optimality}}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle

\section{Global review}

The paper aims to analyse compression from an Algorithmic Information point of view. Its three main contributions are:
\begin{itemize}
  \item A definition of a normalizatio procedure for compression algorithms, which ensures that the compressed string is always shorter than or equal to the or length of the original string plus one, at the cost of one bit. (section 2)
  \item An analysis, mostly by counting arguments, of the expected compression under the uniform distribution. (section 3)
  \item An application of Levin's theorem to compression, suggesting that there is an algorithm which can output any . (section 4)
\end{itemize}

Overall, the paper is clearly written, and easy to understand for readers with a reasonable background.

It is difficult for me to asses to what extent the paper presents a novel or relevant view. The results discussed in section 3, for example, have clear analogues in Kolmogorov Complexity, as does the result of section 4. This leaves me wondering why the authors have opted to invent their own framework, which is only a slight modification of that of Kolmogorov Complexity.

It is of course possible that I am misinterpreting these results and the way they fit into existing research. Without that information, however, I am
forced to assume that this paper does not contain a great deal of novel or relevant research.

To remedy this, the paper would benefit from a conclusions section at the end, summarizing the results and the way in which they tie together.

To be considered for acceptance the paper should clearly state (better yet, demonstrate) the relevance and novelty of its results. The latter is particularly important with respect to the field of Kolmogorov Complexity where many of the methods and results that the authors seem to present as novel are in fact commonplace.

\section{Specific review}
\subsection*{Section 1}

The introduction is clear, and well written, although the reference to the linear congruential method (which I needer to look up) seems obscure and technical, where a simpler example would suffice.

\subsection*{Section 2}

The idea that compressors may actually inflate, rather than compress seems like an important theoretical detail. The method of normalization is helpful for theoretical purposes.

The normalization of the universal Turing Machine on page 2 can also be achieved by setting an empty Turing Machine $T(x) = x$ at index 0 in a prefix coded enumeration. This method does not have the drawback of adding a bit of complexity to all compressible strings as well. Both methods, however, trade universality for specificity. $C(x)$ is more well-behaved, but holds for fewer computational metaphors. For this reason it is general asserted that $C(x) \leq |x| + O(1)$ instead of replacing the constant order with a $1$.

The phrasing in the abstract that the authors ``reformulate a particular result of Algorithmic Information theory'' attaches too much importance to this method.

\subsection*{Section 3}
My main issue with this section is that (as the authors themselves assert at the top of page 8) compression over the uniform distribution is largely useless, as the uncompressible strings will quickly grow towards density one. Thus the expected compression under this distribution will be 0 for any computable method.

Secondly, when quantifiying over all compressors (like in section 3.1 definition 1), the set over which this quantification happens is never defined.  One assume the computable compressors (also an undefined notion), but then

While Theorem 1 is quite clear, its proff is difficult to follow. In particular the step that starts with ``there must be some maximum value\ldots'' could do with greater explanation.

The last line of the derivation in the middle of the proof to Theorem 2 should read $= 2^n(n-2) + 2 + n$ rather than $= 2^n(n-2) + 2$.

The relevance of Theorems 3 and 5 is not clear to me. If we are free to choose any values for $\epsilon_i$, the result does not appear that remarkable. Perhaps an example of the values of $\{\epsilon_i\}$ would make things more clear.

Overall this section seems to present some refinements of the incompressibility results of Li \& Vitanyi, section 2.2. If they were presented as such (without the reliance on probability distributions and the compressor/decompressor framework) there may be some novelty.

\subsection*{Section 4}

\section{Minor issues}

Here I present some specific stylistic errors and remarks for the benefit of the authors. Overall the paper did suffer from any great number of such problems, but a second proofreading would not be an unnecessary luxury.

\begin{itemize}
  \item The document lacks page numbers. I number the title page as 1.
  \item The notation is a little confusing with respect to the standard notations of Kolmogorov Complexity. $C(x)$ of course has a different meaning from it's normal use (which is not diffcult to overcome), but it is also define to return the compressed program rather than the lentgh of that program, as the definition is in Kolmogorov Complexity. I would recommend using lowercase letters $c(x)$ and $d(x)$ to avoid confusion.
  \item On page 2: ``The set of words with length $n$ is $\Sigma^n$ and will be denoted by $S_n$''. It seems to me that not much is won by referring to this set as $S_n$ rather than $\Sigma^n$.
  \item ``Relatively to an optimal compressor \ldots'' should be ``Relative to an optimal compressor\ldots''
\end{itemize}

\end{document}

### 2012-04-10 Facticity opmerkingen/src/opmerkingen.tex
Stub

\documentclass{article}

\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath, amsthm, amssymb}

\theoremstyle{definition}
\newtheorem*{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem*{dfn}{Definition}
\newtheorem*{exm}{Example}

\title{Opmerkingen Facticity}
\author{Peter Bloem}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle

\subsubsection*{algemeen}
\begin{itemize}

    \item Met de latex code \begin{verbatim}\bar{\imath} of \overline{\imath}\end{verbatim} komt het streepje over de $i$ en $j$ in plaats van het puntje: $\overline{\imath}$. Het is een kwestie van smaak, maar ik vind het zelf overzichtelijker.

    \item Je zegt een aantal keer "Interesting/Important is the following lemma...". Volgens mij is dit ongebruikelijk Engels. Ik zou eerder zeggen "The following lemma is interesting..."
\end{itemize}

\subsubsection*{pg 1}
\begin{itemize}

    \item Laatste zin abstract: ``to define the meaningful or usefulness of a dataset'' $\rightarrow$ ``to the define the meaningfulness or usefulness of a dataset''
\end{itemize}

\subsubsection*{pg 2}
\begin{itemize}

    \item Definitie self-delimiting code: ``$\bar{x} = 0^{log n}1cx$'' die $n$ moet een $c$ zijn, denk ik.

    \item ``A Turing Machine is prefix-free if it only accepts programs in prefix-free code'' Dit is voor geen van de TM's in dit paper het geval. De UTM accepteert alleen TM indices in prefix vorm, maar de complete programma's zijn niet prefix-free.

   \item In de definitie heb je het over self-delimiting codes, maar in het voorbeeld over prefix-free codes. 
\end{itemize}

\subsubsection*{pg 3}
\begin{itemize}

   \item In de definitie van klassieke Kolmogorov complexiteit en het kortste programma is het subscript na de ``$\min$'' niet nodig. Het is het minimum van een set van getallen.

   \item     Vlak voor lemma 2: ``the following lemma's''  $\rightarrow$  ``the following lemmas''
\end{itemize}

\subsubsection*{pg 3}

\begin{itemize}
  \item   De inverse entropie is niet een functie van $S$, maar van the entropie: $H(S) = h$, $H'(h) = S$.
  \item        De $p$'s in de formule moeten denk ik ook $h$'s zijn
  \item        Ik zou zelf de $^{-1}$ achter de $H$ zetten, ipv achter de haakjes: $H^{-1(h)}$.
  \item        Dit komt terug op pagina's 7, 8 \& 9
  \item    Bewijs van Lemma 6: ``Since x stochastic...''  $\rightarrow$  ``Since x is stochastic...''
  \item    Eerste paragraaf sectie 3: ``The following lemma's...''  $\rightarrow$  ``The following lemmas...''
  \item    Definitie 10: ``for all elements of set of descriptions''  $\rightarrow$  ``for all elements of a set of descriptions''
\end{itemize}

\subsubsection*{pg 5}
\begin{itemize}

  \item      Bewijs lemma 8: ``q reads in i converts it to...''  $\rightarrow$  ``q reads in i, converts it to...''
  \item      De laatste stap van het bewijs van Lemma 8: $|\bar{q}i|$ valt binnen een constante term vanaf $C(t)$, maar bewijst dat dat $|\bar{q}i|$ niet berekenbaar is? In principe kan een niet-recursieve functie best tot binnen een constante term benaderd worden door een recursieve functie, lijkt me.
\end{itemize}

\subsubsection*{pg 6}
\begin{itemize}

  \item      Bewijs lemma 11: ``Suppose $x$ is compressible and $\phi(x) > 0$.''  $\rightarrow$  ``Suppose $x$ is compressible and $\phi(x) = 0$.''
\end{itemize}

\subsubsection*{pg 7}
\begin{itemize}

  \item      ``In the paragraph...'' -> ``In this paragraph''
\end{itemize}

\subsubsection*{pg 8}
\begin{itemize}

  \item      Definitie 14: ``the minimal number of trials a collector of n coupons has to make to collect all n coupons as n log(n)''  $\rightarrow$  ``the number of trials a collector of n coupons can expect to make as n log(x)''. Het minimum is n, uiteraard.
  \item      Definitie Collapse point for binary stochastic strings: Latex fout
\end{itemize}

\subsubsection*{pg 9}
\begin{itemize}

  \item      Lemma 14: Een kleine letter $u$ voor de index vd UTM is misschien meer in de lijn van je andere indices, zoals $s$, $i$, $q$, etc.
  \item      Lemma 14: In mijn afleiding valt de $+ 1$ aan het einde weg en is de basis wortel 2 ipv 2:
\begin{align*}
		U(\bar{u}p) &= U(\bar{p}\epsilon) = x \\
        |\bar{u}p| &> |\bar{p}\epsilon| \\
        |\bar{u}| + |p| &> |p| + 2 \log (|p|) + 1 \\
        |\bar{u}| &> 2 \log(|p|) + 1 \\
        2\log\varphi(x) &< |\bar{u}| - 1 \\
		\log\varphi(x) &< \frac{1}{2}(|\bar{u}| - 1) \\
		\varphi(x) &< \sqrt{2}^{|u| + 2\log(u)}  
\end{align*}
\end{itemize}

\subsubsection*{pg 11}
\begin{itemize}
  \item De notatie $d$ geldt strikt genomen niet voor integers. Misschien dat $\Delta$ toepasselijker is.
  \item Een omkeerbaar/deterministisch proces bevat altijd de informatie over de huidige tijdsstap, dus volgens heb je dan (voor de meeste waardes van $t$) 
\[
\frac{dK(x+t)}{dt} = \frac{d(log(t) + O(1))}{dt} = \frac{1}{x \ln(2)}
\]
\item 
\end{itemize}

\subsubsection*{pg 12}
\begin{itemize}
  \item Laatste zin sectie 3: ``are maximally instable'' $\rightarrow$ ``are maximally unstable''.
\end{itemize}

\subsubsection*{pg 13}
\begin{itemize}
  \item De definitie van resource-bounded facticity gebruikt nog de oude formulering.
  \item ``Theorem 1 shows that the definition of facticity in terms of $K^*(M)$\ldots'': is dit nog een verwijzing naar gen. Kolmogorov complexity? 
\end{itemize}

\subsubsection*{pg 14}
\begin{itemize}
  \item Halverwege: ``Gell-Mann and Lloye [17] \ldots''  $\rightarrow$ ``Gell-Mann and Llody [17] \ldots''
\end{itemize}

\subsubsection*{pg 15}
\begin{itemize}
  \item Objection 1: Volgens mij mis je een slash voor de argmax/argmin in de MDL formules.
\end{itemize}

\subsubsection*{pg 16}
\begin{itemize}
	\item ``The reason is that $K$ does not observe ICP.'' In deze versie van het paper is het ICP hier nog niet genoemd. Een referentie/uitleg zou nuttig zijn.
	\item Clearly we have to put constraints of\ldots''  $\rightarrow$ ``Clearly we have to put constraints on\ldots''
\end{itemize}

\end{document}

### 2012-04-10 Referee/ref.tex
Stub

\documentclass{article}

\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath, amsthm, amssymb}

\theoremstyle{definition}
\newtheorem*{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem*{dfn}{Definition}
\newtheorem*{exm}{Example}

\title{Review: {\em Lossless compressors: Degree of compression and optimality}}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle

\section{Global review}

This paper aims to analyse compression from an algorithmic information point of view. Its three main contributions are:
\begin{itemize}
  \item A definition of a normalization procedure for compression algorithms, which ensures that the compressed string is always shorter than or equal to the length of the original string (plus one bit), at the cost of one bit. (section 2)
  \item An analysis, mostly by counting arguments, of the expected compression under the uniform distribution. (section 3)
  \item An application of Levin's theorem to compression, suggesting that there is an algorithm which can match the performance of any compressor to within a multiplicative constant. (section 4)
\end{itemize}

Overall, the paper is clearly written, and easy to understand for readers with a reasonable background.

It is difficult for me to asses to what extent the paper presents novel or relevant results. The results discussed in section 3, for example, have clear analogues in Kolmogorov complexity, as does the result of section 4. Specifically the fact that the authors have presented their results almost, but not quite in the framework of algorithmic information theory makes it difficult to place the paper in a greater context.

To remedy this, the paper would benefit from an explicit explanation of how these results are relevant, how they differ from existing research, and why they are beneficial to ongoing efforts. Preferably, this should be done in bookending fashion, by extending the introduction and finishing with a \emph{conclusions} section.

The existing section ``Relationship with algorithmic information theory and with Koppel's sophistication'' underplays the relationship of these results to algorithmic information theory.

My intuition is that these results, if they haven't been published explicitly yet, are at least part of the `folklore' of the field of Kolmogorov complexity. If this is the case, it wouldn't hurt to make them explicit, but by themselves they lack sufficient impact for a publication of this kind.

My overall impression is that the authors are on the road to interesting results for which these are lemmas, but that in its current form the paper only presents reformulations of well known properties. If this impression is unjust, the paper would be improved by making it explicit why, to prevent other readers from forming the same impression.

\section{Specific review}
\subsection*{Section 1}

The introduction is clear and well written.

\subsection*{Section 2}

The idea that compressors may actually inflate, rather than compress, seems like an important theoretical detail. The method of normalization is helpful for theoretical purposes.

The normalization of the universal Turing Machine on page 2 can also be achieved by setting an empty Turing Machine $T(x) = x$ at index 0 in a prefix coded enumeration. This method does not have the drawback of adding a bit of complexity to all compressible strings as well. Both methods, however, trade universality for specificity. $C(x)$ becomes more well-behaved, but holds for fewer computational metaphors. This is the reason that it is generally asserted that $C(x) \leq |x| + O(1)$ instead of replacing the constant order with a $1$.

The phrasing in the abstract that the authors ``reformulate a particular result of Algorithmic Information theory'' attaches too much importance to this method.

\pagebreak[4]

\subsection*{Section 3}
My main issue with this section is that (as the authors themselves assert at the top of page 8) compression over the uniform distribution is largely useless, as the uncompressible strings will quickly grow towards density one. Thus the expected compression under this distribution will be 0 for any computable method.

When quantifiying over all compressors (like in section 3.1 definition 1), the set over which this quantification happens is never defined.

While Theorem 1 is quite clear, its proof is difficult to follow. In particular the step that starts with ``there must be some maximum value \ldots'' could do with greater explanation.

The last line of the derivation in the middle of the proof to Theorem 2 should read $= 2^n(n-2) + 2 + n$ rather than $= 2^n(n-2) + 2$, I believe.

The relevance of Theorems 3 and 5 is not clear to me. If we are free to choose any values for $\{\epsilon_i\}$, the result does not appear that remarkable. Perhaps an example of the values of $\{\epsilon_i\}$ would make things more clear.

Overall this section seems to present some refinements of the incompressibility results as presented (for instance) in  Li \& Vitanyi, section 2.2. If they were presented as such (without the reliance on probability distributions and the compressor/decompressor framework), I would personally consider them of greater value. To clarify, they could be presented as results about the number of compressible strings, and the effect of the universal Turing machine chosen. The results for specific (de)compressors could then be stated as a consequence of this.

\subsection*{Section 4}

The conclusion that Levin's universal search applies to compression algorithms seems a natural one. This seems like a well-established result, although I cannot find any statements of the theorem explicitly in this form. While such a statement is useful to have, by itself it does not seem worthy of publication.

The statement ``typically, the time complexity of the expansion is smaller than the time complexity of the compression'' requires qualification.

\pagebreak[4]

\section{Minor issues}

Here I present some specific stylistic errors and remarks for the consideration of the authors. Overall the paper did not suffer from any great number of such problems. They are included here only in the hope that they may be of some use.

\begin{itemize}
  \item The document lacks page numbers. I number the title page as 1.
  \item The notation is a little confusing with respect to the standard notations of Kolmogorov Complexity. $C(x)$ of course has a different meaning from its normal use (which is not diffcult to overcome), but it is also defined to return the compressed program rather than the length of that program, as the definition is in Kolmogorov Complexity. I would recommend using lowercase letters $c(x)$ and $d(x)$ to avoid confusion.
  \item On page 2: ``The set of words with length $n$ is $\Sigma^n$ and will be denoted by $S_n$''. It seems to me that not much is won by referring to this set as $S_n$ rather than $\Sigma^n$.
  \item ``Relatively to an optimal compressor \ldots'' should be ``Relative to an optimal compressor \ldots''
  \item ``\ldots{ }has the same time order than \ldots'' should be ``\ldots{ }has the same time order as \ldots''
\end{itemize}

\end{document}

### 2012-04-10 fractal paper/src/fractal.tex
Stub

\documentclass{article}

\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath, amsthm, amssymb}

\theoremstyle{definition}
\newtheorem*{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem*{dfn}{Definition}
\newtheorem*{exm}{Example}

\title{Fractal modelling}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle

\begin{abstract}
We present a method for modelling data with fractals, mathematical constructs with certain interesting properties, such as non-smoothness, infinite detail and non-integer dimension. Fractal models can vary their intrinsic dimension smoothly to fit that of the dataset. We use the well known family of Iterated Function Systems as statistical models and present a new algorithm for fitting such models to a given dataset (the so called inverse problem). To show the value of such models, we construct a classifier from them, and compare this classifier against other popular classifiers.
\end{abstract}

\section{Introduction}
\section{Preliminaries}
\subsection{Fractals}
\subsection{Iterated Function Systems}
\section{Methods}
\subsection{The inverse problem}
\subsection{Classification}
\section{Conclusions and Further Work}

One advantage of the IFS model that has been left unexplored in the current work is that it assigns each point in the dataset a $k$-ary string so that a small distance between points suggest simlar strings. Thus, a point pattern in a Euclidean space is naturally mapped to the space of strings in a way that uses the structure of the dataset rather than the structure of the embedding space.

Future investigation might focus on the use of IFS models in clustering, spatial indexing and the induction of random fractals (as discussed in \cite{})
\end{document}

### 2012-04-10 fractal paper/src/fractal.tex
Text

@@
 
-\section{Introduction}
 \section{Preliminaries}
 \subsection{Fractals}
+
+Fractals are mathematical objects (figures, sets or measures) with certain remarkable properties.
+No official definition exists for what does and does not constitute a fractal, but the following properties are central to all treatments:
+\begin{itemize}
+  \item They are selfsimilar.
+A part is a scaled down copy of the whole.
+See for instance the Sierpinski triangle (fig !@), which can be seen as three scaled up copies of itself.
+In some cases the self similarity is statistical, so that the scaled down copy is statistically similar to the whole
+  \item They have infinite detail.
+'Zooming in' reveals ever finer detail.
+In the case of the Sierpinski triangle we will simple see the same structure recurring over and over again, but fractals like the Mandelbrot set reveal a great variety of detail.
+  \item They can have non-integer dimension. This basic and counter-intuitive fact is not easily explained. It will be discussed in section @! 
+\end{itemize}
+
+The earliest examples of fractals were described around the beginning of the twentieth century as counterexamples to conjectures in planar geometry and largely considered pathological.
+It was not until the 1970s that they were considered as a family of mathematical objects that could accurately describe many physical phenomena.
+Coast lines, trees, clouds and lightning bolts could all be described more accurately as fractals, than as lines, spheres or cubes.
+
 \subsection{Iterated Function Systems}
+
 \section{Methods}
 \subsection{The inverse problem}
+One of the main challenges in all forms of fractal modelling has been \textit{the inverse problem}.
+Whereas the chaos game produces a set of points from a fractal model, the inverse problem is to find a fractal for a given set of datapoints.
+In short, fitting a fractal model to data.
+
+So far, two approaches have succesfully tackled the inverse problem in some form:
+\begin{itemize}
+  \item Genetic algorithms.
+A straightforward search through the space of IFSs.
+The drawback of this approach is one of time complexity.
+The running time of these approaches scale poorly in the number of dimensions and components.
+  \item Partitioned iterated function systems.
+This is a modified version of the problem to make the model tractable and feaible in the domina of image compression.
+This approach does not translate well to n-dimensional point patterns.
+\end{itemize}
+
 \subsection{Classification}

### 2012-07-08 P23WP1Q2Deliverable/src/deliverable.p23wp1q2..tex
...

\documentclass{article}

\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{authordate1-4}
\usepackage{url}

\theoremstyle{definition}
\newtheorem*{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem*{dfn}{Definition}
\newtheorem*{exm}{Example}
\usepackage{lscape}

\title{Complexity Measures for Graphs}
\author{Peter Bloem}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle

\begin{abstract}
\noindent This document presents some complexity measures on graph data. We focus specifically on those measures that are designed for `complex networks': large graphs like social networks, technological networks and biological networks.
\end{abstract}

\section*{Complexity measures}
\subsection*{Basic graph measures}
\begin{description}
\item[Number of nodes ($N$)] A count of the number of nodes (or vertices) in the network.
\item[Number of links ($L$)] A count of the number of links (or edges) in the network. 
\item[Mean degree ($\bar{d}$)] The mean number of nodes a node is connected to.
\item[Degree standard deviation ($\sigma_d$)]  The sample standard deviation of the degrees.
\item[Size of the largest component ($W$)] Size of the largest connected subgraph.
\item[Proportion of the largest component ($W_p$)] The size of the largest connected subgraph divided by the size of the total graph. (Note that rounding will sometimes cause this to appear as 1.0, even when the largest component is smaller than the total graph).
\item[Diameter of the largest component ($D$)] If we define the distance between two nodes as the length of the shortest path, then the diameter is the longest posssible distance in the graph. If the graph is not connected, then the diameter is infinite, so we report the diameter of the largest connected component.  
\item[Mean distance ($M$)] The mean distance between two nodes. A large network with a small mean distance is said to have the \textit{small world} property, ie. there is likely to be a short path between any to nodes.
\item[Clustering Coefficient ($C$, $C_{WS}$)] The clustering coefficient measures the level of \textit{transitivity in the network}. For instance, if system a on the internet is connected to system b, and b is connected to system c, with what probability can we say that a connects to c also.

There is also a local clustering coefficient $C_i$, which is the same but only for the neighbouring nodes of node $i$. The mean local clustering coefficient we call $C_{WS}$. The two values  
\end{description}
\subsection*{Scale free networks ($\alpha$,$p_\alpha$)}

The degree distribution (the probability that a node has a given number of neighbours) is one of the most telling attributes of a network. In particular the networks for which the degree distribution follows a power law are interesting. These networks are called \emph{scale free}, and they have many  interesting characteristics, such as a tolerance to random attacks and the small world property. \cite{barabasi1999emergence}

If the degree distribution follows a power law, we refer to its exponent as $\alpha$.
The significance level of the conclusion that the degree distribution follows a power law we call $p_\alpha$. If this value is greater than 0.1 we consider the power law a reasonable model, and report the exponent. All methods for determining significance and exponents were taken from \cite{clauset2007power}.

\subsection*{Assortative mixing ($r$)}

Another important property of a network is whether nodes are likely to be connected to nodes that are similar, for some measure of similarity. Nodes in friendship networks, for instance tend to be connected to nodes with similar properties (gender, height, ethnicity). This is called assortative mixing. An example of disassortative mixing is a sexual contact network, where the majority of links are between different genders.

We can also test for assortative mixing by structural properties of the graph. Most commonly, the degree of nodes is used for this. For instance, do people with many friends tend to be friends with other people with many friends, or do websites with many links often link to other websites with many links.

The most common measure for Assortative mixing by degree is the Pearson correlation coefficient between the degrees on either side of an edge. We call this value $r$. Since $r$ represents a correlation, it ranges from -1 to 1, with -1 being perfect disassortive mixing and 1 being perfect assortative mixing.

A high value of $r$ is associated with resilience to targeted attack.

\subsection*{Self-similarity ($d_B$)}

For data sets like point patterns, metric data of tabular data, the notion of dimension can be a very descriptive property. There are many methods that allow on to measure the intrinsic dimension of the dataset, without considering the specifics of its representation.

Almost all notions of intrinsic dimension rely on so called scaling laws. A scaling law relates a one dimensional measure $x$ (like the width of a box) to some measure over the dataset $y$ (like the number of point that fall in a box). The laws of geometry then tell us that these should be (roughly) related as $y \sim x^\alpha$, where $\alpha$ is the intrinsic dimension of the dataset.

If we try a similar approach with networks we will soon see it fall down. If for instance, we take some random node in a small-world network and ask how many nodes $N(l)$ are within a distance $l$ of it, the small-world property will ensure that most nodes are within a small distance. The growth will be logarithmic rather than polynomial.

To find an analogue to dimension for networks, Song et al @!cite devise the following method: cover the network in clusters of nodes called boxes, such that each box has a maximum distance between any two nodes of $l_B$. If we create a new network with a node for each box and links between nodes if these is a single-step connection between the corresponding boxes, we find that this new (smaller) network will, for many networks, have the same degree distribution as the old. Applying the procedure again will have the same result. In this sense, the network is said to be \emph{self-similar}.

If we apply a single step for different values $l_B$ and count the resulting number of boxes $N_B(l_B)$, we find the following relation for self-similar networks:
\[
\frac{N_B(lb)}{N} \approx l_B^{d_B}
\]
Where $N$ is the number of nodes in the network and $d_B$ is a characteristic scaling exponent of the network, analogous to dimension. The exact properties of this self-fimilar scaling have not been fully investigated, but some relation has been found between robustness and self similarity. @! Cite

\subsection*{Compressibility}
The minimum/maximum compression ratio $\kappa_{\mbox{min}}$, $\kappa_{\mbox{max}}$ and the median $ \bar{\kappa}$

\section*{Introduction}

\section*{Data}

We test these complexity measures on the following datasets.

\subsection*{Technological networks}

\subsubsection*{The internet}

The internet is a classic example of a large, complex network. So large, in fact that we have little hope capturing a full snapshot of the whole thin at once. The best we can do is to capture some part of it using traceroutes. We have two datasets, each representing autonomous systems as vertices, with edges between those that are directly connected.

The large dataset was downloaded from the Stanford Large Network Dataset Collection(\url{http://snap.stanford.edu/data/as-skitter.html}), and it contains roughly 1.7 million nodes. @! Cite

The smaller dataset was downloaded from the personal page of Mark Newman (\url{http://www-personal.umich.edu/~mejn/netdata/}), and it contains @!Cite

\subsubsection*{The world wide web}
\subsection*{Social networks}
\subsubsection*{Epinions}
\subsubsection*{Actor network}

\subsection*{Biological networks}
\subsubsection*{E. Coli Cellular network}
\subsubsection*{C. Elegans Neural Network}

\subsection*{RDF Data}
\subsubsection*{IAFB Dataset}

\section*{Code}

The code is written in Java, using the JUNG graph library. All code is available in GitHub in the following repsitories.
\begin{description}
\item[D2S Tools] (\url{https://github.com/Data2Semantics/d2s-tools}) Common code shared by WP1 and WP2.
\item[Lilian] (\url{https://github.com/pbloem/Lilian}) Implementation of the complexity measures, including some tools for running experiments, and much unrelated code.
\end{description}

\begin{landscape}

\section*{Results}

\begin{tabular}{l | r r r r r r r r r r r r r r r }
  \hline
    & $N$ & $L$ & $W$ & $W_p$ & $D$ & $M$ & $C$ & $C_{\mbox{WS}}$ & $\alpha$ & $p_\alpha$ & $r$ & $d_B$ & $\bar{\kappa}$ & $\kappa_{\mbox{min}}$ & $\kappa_{\mbox{max}}$  \\
    \hline
 Internet (big) & N & L & W & Wp & D & C & C WS & alpha & p alpha & r & dB & kappa & kappa min & kappa max \\
 Internet (small) & N & L &  W & Wp & D & C & C WS & alpha & p alpha & r & dB & kappa & kappa min & kappa max \\    
 WWW & & & & & & & & & & & & & & & km \\
 Epinions & & & & & & & & & & & & & & & km \\ 
 Actors  & & & & & & & & & & & & & & & km \\
 E. Coli  & & & & & & & & & & & & & & & km \\
 C. Elegans  & & & & & & & & & & & & & & & km \\
 IAFB  & & & & & & & & & & & & & & & km \\ 
 \hline

\end{tabular}

\end{landscape}

\bibliographystyle{authordate3}
\bibliography{citations}

\end{document}

### 2012-07-08 fractal paper/src/fractal.tex
...

@@
 
-The earliest examples of fractals were described around the beginning of the twentieth century as counterexamples to conjectures in planar geometry and largely considered pathological.
-It was not until the 1970s that they were considered as a family of mathematical objects that could accurately describe many physical phenomena.
-Coast lines, trees, clouds and lightning bolts could all be described more accurately as fractals, than as lines, spheres or cubes.
+The earliest examples of fractals were described around the beginning of the twentieth century as counterexamples to conjectures in planar geometry and mostly considered pathological.
+It was not until the 1970s that they were began to be seen as a single and useful family of mathematical objects that could accurately describe many physical phenomena.
+Coast lines, trees, clouds and lightning bolts can all be described more accurately as fractals than as lines, spheres or cubes.
 
@@
 \begin{itemize}
-  \item Genetic algorithms.
+  \item Genetic algorithms!@.
 A straightforward search through the space of IFSs.
@@
 The running time of these approaches scale poorly in the number of dimensions and components.
-  \item Partitioned iterated function systems.
-This is a modified version of the problem to make the model tractable and feaible in the domina of image compression.
+  \item Partitioned iterated function systems!@.
+This is a modified version of the problem to make the model tractable and feasible in the domain of image compression.
 This approach does not translate well to n-dimensional point patterns.

### 2012-08-09 Dimension/src/dimension.tex
...

\documentclass{article}

\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath, amsthm, amssymb}

\theoremstyle{definition}
\newtheorem*{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem*{dfn}{Definition}
\newtheorem*{exm}{Example}

\title{Measuring Dimension}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle

\begin{abstract}
The intrinsic dimension of a dataset is one of the most telling statistics that can be calculated, and one that has received a great deal of attentation over the past decades. Most methods for estimating dimension relsy on at least some subjective judgment to make a selection in the data. Adapting methods recently described [@! cite] for the statistical analysis of power laws, we revisit an existing MLE estimator and present a simple, objective procedure to estimate the dimension of a dataset. The procedure runs without human intervention, and produces a value for the dimension, uncertainty bounds and a significance value.\end{abstract}

\section{Introduction}

Definitions of dimension, fractal dimension and measuring dimension.

Previous work. Theiler. Manifold learning. Databases (Cite faloutsos)We will not discuss the mathematical backgrounds of dimension (Cite Edgar). We will not focus greatly on the algorithmic complexity.

\section{Dimension}

Scaling laws.

Box counting dimension, Correlation integral, Haussdorff & Packing dimension.

Box counting estimator, Correlation estimator, Takens estimator.

\subsection{Physics}
\subsection{Statistics and Machine Learning}
\subsection{}

\section{Methods}

\subsection{Maximum Likelihood Estimator}

Derivation of Takens' MLE and its standard error. Comparison to the Hill estimator and discussion of the differences.

\subsection{Estimating the $d_Max$ parameter}

Quick review of the methods described in Clauset. Translation of the method to this estimator. Experiment showing the validity of the method for this MLE.

\subsection{Uncertainty}

Quick review of the approach of Clauset.

\subsection{Significance}

Review of the approach of Clauset, and a description of the method translated to this problem.

\subsection{Metric data}

Discussion of the use of dimension for metric data and the use of the significance measure.

\subsection{Runnning time}

Discussion of the running time an an investigation of the effects of reducing the data size.
\section{Experiments}
\subsection{Data sets}

\begin{itemize}
  \item Swiss roll
  \item 8-loop

  \item Hands
  \item Faces

  \item Sierpinski
  \item Cantor
  \item Menger Sponge

  \item 
\end{itemize}

\subsection{Results}

\section{Conclusion}

We have presented a straightforward, objective method for calculating a maximum likelihood estimate of the intrinsic dimension of a dataset. We have com

\end{document}

### 2012-08-09 P23WP1Q2Deliverable/src/deliverable.p23wp1q2..tex
...

@@
 \usepackage{url}
+\usepackage{booktabs}
 
@@
 
+\nocite{*}
 \begin{abstract}
-\noindent This document presents some complexity measures on graph data.
-We focus specifically on those measures that are designed for `complex networks': large graphs like social networks, technological networks and biological networks.
+\noindent In this quarter, we have implemented various complexity measures for graphs.
+This document provides a description of those methods and the preliminary results of applying the methods to several datasets.
+We focus specifically on those measures that are designed for \emph{complex networks}: large graphs like social, technological and biological networks.
 \end{abstract}
+
+\section*{Code}
+
+The code is written in Java, using the JUNG graph library. All code is available from GitHub in the following repsitories.
+\begin{description}
+\item[D2S Tools] (\url{https://github.com/Data2Semantics/d2s-tools}) Common code shared by WP1 and WP2.
+\item[Lilian] (\url{https://github.com/pbloem/Lilian}) Implementation of the complexity measures, including some tools for running experiments.
+This repository also contains a great deal of code unrelated to this deliverable.
+The relevant code resides mostly in the package \texttt{org.lilian.util.graphs.jung}.
+\end{description}
 
@@
 \begin{description}
-\item[Number of nodes ($N$)] A count of the number of nodes (or vertices) in the network.
-\item[Number of links ($L$)] A count of the number of links (or edges) in the network. 
+\item[Number of nodes ($N$)] A count of the number of nodes in the network.
+\item[Number of links ($L$)] A count of the number of links in the network. 
 \item[Mean degree ($\bar{d}$)] The mean number of nodes a node is connected to.
@@
 (Note that rounding will sometimes cause this to appear as 1.0, even when the largest component is smaller than the total graph).
-\item[Diameter of the largest component ($D$)] If we define the distance between two nodes as the length of the shortest path, then the diameter is the longest posssible distance in the graph.
+\item[Diameter of the largest component ($D$)] If we define the distance between two nodes as the length of the shortest path between them, then the diameter is the longest possible distance in the graph.
 If the graph is not connected, then the diameter is infinite, so we report the diameter of the largest connected component.
@@
 there is likely to be a short path between any to nodes.
-\item[Clustering Coefficient ($C$, $C_{WS}$)] The clustering coefficient measures the level of \textit{transitivity in the network}.
-For instance, if system a on the internet is connected to system b, and b is connected to system c, with what probability can we say that a connects to c also.
+\item[Clustering Coefficient ($C$, $C_{WS}$)] The clustering coefficient measures the level of \textit{transitivity} in the network.
+For instance, if system $a$ on the internet is connected to system $b$, and $b$ is connected to system $c$, with what probability can we say that $a$ connects to $c$ also?
 
-There is also a local clustering coefficient $C_i$, which is the same but only for the neighbouring nodes of node $i$. The mean local clustering coefficient we call $C_{WS}$. The two values  
+There is also a local clustering coefficient $C_i$, which is defined in the same way, but only for the neighbouring nodes of node $i$.
+The mean local clustering coefficient we call $C_{WS}$.
+The values $C$ and $C_{WS}$are not necessarily similar.
+If they diverge, this can be taken as a sign that the clustering coefficient is not a meaning measure for this network.\cite{newman2010networks}
 \end{description}
-\subsection*{Scale free networks ($\alpha$,$p_\alpha$)}
+
+\subsection*{Scale free networks ($\alpha$, $p_\alpha$)}
+
+\textit{The implementation of these measures is still under construction.}
 
@@
 
-Another important property of a network is whether nodes are likely to be connected to nodes that are similar, for some measure of similarity.
-Nodes in friendship networks, for instance tend to be connected to nodes with similar properties (gender, height, ethnicity).
-This is called assortative mixing.
+Another important property of a network is whether nodes are likely to be connected to nodes that are similar to themselves (for some measure of similarity).
+Nodes in friendship networks, for instance, tend to be connected to nodes with similar properties (gender, height, ethnicity).
+This is called \emph{assortative mixing}.
 An example of disassortative mixing is a sexual contact network, where the majority of links are between different genders.
@@
 Most commonly, the degree of nodes is used for this.
-For instance, do people with many friends tend to be friends with other people with many friends, or do websites with many links often link to other websites with many links.
+For instance, do people with many friends tend to be friends with other people with many friends, or do websites with many links often link to other websites with many links?
 
-The most common measure for Assortative mixing by degree is the Pearson correlation coefficient between the degrees on either side of an edge.
-We call this value $r$.
+The most common measure for assortative mixing by degree is the Pearson correlation coefficient between the degrees on either side of an edge.
+\cite{newman2002assortative} We call this value $r$.
 Since $r$ represents a correlation, it ranges from -1 to 1, with -1 being perfect disassortive mixing and 1 being perfect assortative mixing.
 
-A high value of $r$ is associated with resilience to targeted attack.
+A high value of $r$ is associated with resilience to targeted attack.\cite{newman2002assortative}
 
@@
 
-For data sets like point patterns, metric data of tabular data, the notion of dimension can be a very descriptive property.
-There are many methods that allow on to measure the intrinsic dimension of the dataset, without considering the specifics of its representation.
+For data sets like point patterns, metric data or tabular data, the notion of \emph{dimension} can be a very descriptive property.
+There are many methods that allow one to measure the intrinsic dimension of the dataset, without relying on the specifics of its representation.
 
-Almost all notions of intrinsic dimension rely on so called scaling laws.
-A scaling law relates a one dimensional measure $x$ (like the width of a box) to some measure over the dataset $y$ (like the number of point that fall in a box).
+Almost all notions of intrinsic dimension are based on \emph{scaling laws}.
+A scaling law relates a one dimensional measure $x$ (like the width of a box) to some measure over the dataset $y$ (like the number of data points covered by the box).
 The laws of geometry then tell us that these should be (roughly) related as $y \sim x^\alpha$, where $\alpha$ is the intrinsic dimension of the dataset.
@@
 
-To find an analogue to dimension for networks, Song et al @!cite devise the following method: cover the network in clusters of nodes called boxes, such that each box has a maximum distance between any two nodes of $l_B$.
-If we create a new network with a node for each box and links between nodes if these is a single-step connection between the corresponding boxes, we find that this new (smaller) network will, for many networks, have the same degree distribution as the old.
+To find an analogue to dimension for networks, Song et al \cite{song2005self} devise the following method: cover the network in clusters of nodes called boxes, such that each box has a maximum distance between any two nodes of $l_B$.
+If we create a new network with a node for each box and links between nodes where there is a single-step connection between the corresponding boxes, we find that this new (smaller) network will, for many networks, have the same degree distribution as the old.
 Applying the procedure again will have the same result.
@@
 \[
-\frac{N_B(lb)}{N} \approx l_B^{d_B}
+\frac{N_B(l_b)}{N} \approx {l_B}^{d_B}
 \]
@@
 The exact properties of this self-fimilar scaling have not been fully investigated, but some relation has been found between robustness and self similarity.
-@!
-Cite
+\cite{song2006origins}
 
-\subsection*{Compressibility}
-The minimum/maximum compression ratio $\kappa_{\mbox{min}}$, $\kappa_{\mbox{max}}$ and the median $ \bar{\kappa}$
+\subsection*{Compressibility ($\kappa$)}
 
-\section*{Introduction}
+One of the most useful and theoretically well-founded notions of complexity is Kolmogorov Complexity.
+Put simply, the Kolmogorov Complexity of some data is the length of the shortest program that outputs that data on a universal computer.
+Kolmogorov Complexity is often approximated by compressability with some simple compression algorithm (such a GZIP).
+\cite{li2008introduction}
+
+We would like to bring the subject of Graph complexity under the umbrella of Kolmogorov Complexity in some meaningful way.
+To begin with, we study the compressibility of networks.
+We take some random permutation of the nodes, generate the bottom half of the adjacency matrix and then flatten that into a binary string.
+From this representation the network structure can be reconstructed.
+We then chekc the compression ratio we get with GZIP.
+
+Repeating this experiment 1000 times will tell us two things.
+First, how compressible (and thus how random) the structure of the network is and second, how much influence the ordering of the nodes has on the compression.
+
+We call the minimum/maximum compression ratio $\kappa_{\mbox{min}}$, $\kappa_{\mbox{max}}$ and the mean $ \bar{\kappa}$.
+If the network was too large to perform multiple runs, we report only the compressibility for a single run
 
@@
 
-We test these complexity measures on the following datasets.
+We test these complexity measures on the following datasets:
 
@@
 
-The internet is a classic example of a large, complex network.
-So large, in fact that we have little hope capturing a full snapshot of the whole thin at once.
+The internet is a classic example of a large, complex graph.
+So large, in fact that we have little hope of seeing a full snapshot of the whole network at once.
 The best we can do is to capture some part of it using traceroutes.
-We have two datasets, each representing autonomous systems as vertices, with edges between those that are directly connected.
-
-The large dataset was downloaded from the Stanford Large Network Dataset Collection(\url{http://snap.stanford.edu/data/as-skitter.html}), and it contains roughly 1.7 million nodes. @! Cite
-
-The smaller dataset was downloaded from the personal page of Mark Newman (\url{http://www-personal.umich.edu/~mejn/netdata/}), and it contains @!Cite
+We use a dataset downloaded from the personal page of Mark Newman (\url{http://www-personal.umich.edu/~mejn/netdata/}), which contains around 23 thousand nodes.
 
 \subsubsection*{The world wide web}
+
+The world wide web, while closely associated with the internet, is a fundamentally different type of network.
+For example, a link between two systems on the internet will bear some relation to physical proximity of the systems.
+Linking to a website that is hosted far away costs nothing.
+
+The WWW dataset was downloaded from the iCeNSA website \url{icensa.nd.edu} and is the same dataset used in \cite{song2005self}.
+
 \subsection*{Social networks}
 \subsubsection*{Epinions}
+
+Epinions.org is a review site where users can choose to `trust' one another.
+these trust relations are then used to determine the value of a given review.
+The total network, downloaded form \url{http://snap.stanford.edu/data/soc-Epinions1.html} contains about 76 thousand nodes.
+
+Originally used in \cite{richardson2003trust}.
+
 \subsubsection*{Actor network}
+
+The network of which actors have appeared together in films (and the related ``six degrees of Kevin Bacon'' game) is one of the earliest models of the theory of complex networks.
+
+This dataset was downloaded form the iCeNSA website \url{icensa.nd.edu} and generated from data from the Internet Movie database.
 
@@
 \subsubsection*{E. Coli Cellular network}
+
+Interactions between the metabolites of E. Coli in the course of the metabolic cycle. Downloaded from the iCeNSA website \url{icensa.nd.edu}. Also used in \cite{song2005self}.
+
 \subsubsection*{C. Elegans Neural Network}
+
+C. Elegans, a nematode worm, is a model organism for research into biological neural networks. The whole of its neural network (some 307 cells) has been mapped out.
+
+Our dataset (downloaded from \url{http://www-personal.umich.edu/~mejn/netdata/} and verified with other sources) contains only 297 nodes. We assume that this is only the largest connected component.
+
+Citations for the dataset: \cite{watts1998collective,white1986structure}
 
@@
 
-\section*{Code}
-
-The code is written in Java, using the JUNG graph library. All code is available in GitHub in the following repsitories.
-\begin{description}
-\item[D2S Tools] (\url{https://github.com/Data2Semantics/d2s-tools}) Common code shared by WP1 and WP2.
-\item[Lilian] (\url{https://github.com/pbloem/Lilian}) Implementation of the complexity measures, including some tools for running experiments, and much unrelated code.
-\end{description}
+This dataset was included to test our methods on RDF data.
+We take the dataset from \cite{bloehdorn2007kernel} and disregard any predicates that are not part of the domain specific ontology (http://swrc.ontoware.org/ontology).
+Resources are translated to vertices and predicates are translated to edges.
 
@@
 
-\section*{Results}
+\section*{Preliminary results}
+
+These are the results that have been gathered so far.
+They still require further testing and verification with results reported by others.
+Some measures are too intensive to be applied to the larger datasets.
+For others, our implementations can be improved to accomodate larger datasets.
 
@@
   \hline
-    & $N$ & $L$ & $W$ & $W_p$ & $D$ & $M$ & $C$ & $C_{\mbox{WS}}$ & $\alpha$ & $p_\alpha$ & $r$ & $d_B$ & $\bar{\kappa}$ & $\kappa_{\mbox{min}}$ & $\kappa_{\mbox{max}}$  \\
+    & $N$ & $L$ & $W$ & $W_p$ & $D$ & $M$ & $C$ & $C_{WS}$ & $\alpha$ & $p_\alpha$ & $r$ & $d_B$ & $\bar{\kappa}$ & $\kappa_{\mbox{min}}$ & $\kappa_{\mbox{max}}$  \\
     \hline
- Internet (big) & N & L & W & Wp & D & C & C WS & alpha & p alpha & r & dB & kappa & kappa min & kappa max \\
- Internet (small) & N & L &  W & Wp & D & C & C WS & alpha & p alpha & r & dB & kappa & kappa min & kappa max \\    
- WWW & & & & & & & & & & & & & & & km \\
- Epinions & & & & & & & & & & & & & & & km \\ 
- Actors  & & & & & & & & & & & & & & & km \\
- E. Coli  & & & & & & & & & & & & & & & km \\
- C. Elegans  & & & & & & & & & & & & & & & km \\
- IAFB  & & & & & & & & & & & & & & & km \\ 
+ Internet & 22963 & 48436 & 75877.0 & 1.0 & & & & & & & & & & & \\   
+ WWW & 325729 & 1497134 & & & & & & & & & & & & & \\
+ Epinions & 75879 & 508837 & & & & & & & & & -0.01 & & & & \\ 
+ Actors  & 383640 & 1229020 & & & & & & & & & & & & & \\
+ E. Coli  & 2859 & 6890 & 2859 & 1.0 & 18 & 4.70 & 0 & 0 & & & -0.16 & & & & \\
+ C. Elegans  & 297 & 2345 & 297 & 1.0 & 5 & 2.45 & 0.18 & 0.29 & & & -0.15 & -4.5 & 0.39 & 0.38 & 0.39 \\
+ IAFB  & 8318 & 61255 & 8318 & 1.0 & & & & & & & -0.26 & & 0.025 & &  \\ 
  \hline

### 2012-08-09 fractal paper/src/fractal.tex
...

@@
 
-\title{Fractal modelling}
+\title{Fractal Modeling}
 \date{\today}
@@
 \begin{abstract}
-We present a method for modelling data with fractals, mathematical constructs with certain interesting properties, such as non-smoothness, infinite detail and non-integer dimension.
-Fractal models can vary their intrinsic dimension smoothly to fit that of the dataset.
-We use the well known family of Iterated Function Systems as statistical models and present a new algorithm for fitting such models to a given dataset (the so called inverse problem).
-To show the value of such models, we construct a classifier from them, and compare this classifier against other popular classifiers.
+Fractals are mathematical objects that display great detail and complexity from a simple description.
+It has often been suggested that fractals mimic the structures of nature and everyday life better than the shapes of classical geometry \cite{}.
+These properties make fractals interesting candidates for modeling data.
+We present an overview of existing concepts that are relevant to this task, and a new algorithm for fitting fractal models to data.
 \end{abstract}
@@
 \begin{itemize}
-  \item They are selfsimilar.
+  \item They are self-similar.
 A part is a scaled down copy of the whole.
-See for instance the Sierpinski triangle (fig !@), which can be seen as three scaled up copies of itself.
-In some cases the self similarity is statistical, so that the scaled down copy is statistically similar to the whole
+See for instance the Sierpinski triangle (fig !@), which can be seen as three scaled down copies of itself.
+In some cases the self similarity is statistical, so that the scaled down copy is statistically similar to the whole, but not exactly so.
   \item They have infinite detail.
 'Zooming in' reveals ever finer detail.
-In the case of the Sierpinski triangle we will simple see the same structure recurring over and over again, but fractals like the Mandelbrot set reveal a great variety of detail.
-  \item They can have non-integer dimension. This basic and counter-intuitive fact is not easily explained. It will be discussed in section @! 
+In the case of the Sierpinski triangle we will simply see the same shape recurring again and again, but fractals like the Mandelbrot set reveal a great variety of detail.
+  \item They can have non-integer dimension, for instance the Sierpinski triangle has dimension about 1.58.
+This basic and counter-intuitive fact is not easily explained.
+It will be discussed in section @!.
 \end{itemize}
@@
 The earliest examples of fractals were described around the beginning of the twentieth century as counterexamples to conjectures in planar geometry and mostly considered pathological.
-It was not until the 1970s that they were began to be seen as a single and useful family of mathematical objects that could accurately describe many physical phenomena.
-Coast lines, trees, clouds and lightning bolts can all be described more accurately as fractals than as lines, spheres or cubes.
+It was not until the 1970s that they began to be seen as a single and useful family of mathematical objects that could accurately describe many physical phenomena.
+[@!
+Cite mandelbrot] Beno\^it Mandelbrot, commonly regarded as the father of fractal geometry, put it as follows in the \emph{The Fractal Geometry of Nature} [@!
+cite]:
+
+\begin{quotation}
+\noindent Clouds are not spheres, mountains are not cones, coastlines are not circles, and bark is not smooth, nor does lightning travel in a straight line.
+\end{quotation}
 
@@
 \section{Methods}
-\subsection{The inverse problem}
+
+\subsection{Dimension}
+
+\section{Modeling}
+
+There are many well-known algorithms for
+
 One of the main challenges in all forms of fractal modelling has been \textit{the inverse problem}.
-Whereas the chaos game produces a set of points from a fractal model, the inverse problem is to find a fractal for a given set of datapoints.
+Whereas the chaos game produces a set of points from a fractal model, the inverse problem is to find a fractal for a given set of data points.
 In short, fitting a fractal model to data.

### 2012-08-13 fractal paper/src/fractal.tex
A day of writing

@@
-\documentclass{article}
+\documentclass[11pt, twocolumn]{article}
 
@@
 \newtheorem*{exm}{Example}
+
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 \begin{abstract}
-Fractals are mathematical objects that display great detail and complexity from a simple description.
+\noindent Fractals are mathematical objects that display great detail and complexity from a simple description.
 It has often been suggested that fractals mimic the structures of nature and everyday life better than the shapes of classical geometry \cite{}.
@@
 
-Fractals are mathematical objects (figures, sets or measures) with certain remarkable properties.
-No official definition exists for what does and does not constitute a fractal, but the following properties are central to all treatments:
+Fractals are mathematical objects (generally sets or measures) with certain remarkable properties.
+What constitutes a fractal is not precisely defined\footnote{Mandelbrot, who coined the word, originally defined it as a set whose topological dimension differs from its Hausdorff dimension.
+He later retracted this definition, stated that he preferred the word to be not precisely defined.}, but the following properties are common:
+
 \begin{itemize}
-  \item They are self-similar.
-A part is a scaled down copy of the whole.
-See for instance the Sierpinski triangle (fig !@), which can be seen as three scaled down copies of itself.
-In some cases the self similarity is statistical, so that the scaled down copy is statistically similar to the whole, but not exactly so.
-  \item They have infinite detail.
-'Zooming in' reveals ever finer detail.
+  \item Self similarity: A part is a scaled down copy of the whole.
+See for instance the Sierpinski triangle (fig !@).
+It consists of three triangular shapes which are scaled down copies of the whole (as such these three components each also consist of scaled down copies of themselves, and so on).
+  \item Infinitely small detail: 'Zooming in' reveals ever finer detail.
 In the case of the Sierpinski triangle we will simply see the same shape recurring again and again, but fractals like the Mandelbrot set reveal a great variety of detail.
-  \item They can have non-integer dimension, for instance the Sierpinski triangle has dimension about 1.58.
-This basic and counter-intuitive fact is not easily explained.
-It will be discussed in section @!.
+  \item Non-integer dimension: for instance the Sierpinski triangle has dimension about 1.58. This basic and counter-intuitive fact is not easily explained. It will be discussed in section @!.
 \end{itemize}
@@
 \begin{quotation}
+\small
 \noindent Clouds are not spheres, mountains are not cones, coastlines are not circles, and bark is not smooth, nor does lightning travel in a straight line.
@@
 
-\subsection{Iterated Function Systems}
+Fractal geometry has been used in various forms in fields such as physics @!, biology @!
+and economics.
+The first connection between fractals and the natural world was in the form of tangible objetcs such as coast lines, trees or bronchia, but the more hidden aspects of reality were soon shown to have fractal properties as well.
+The path traced out by the Brownian motion of a particle, the price chart for a commodity or the level of floods in a river delta all benefited from a fractal perspective.
+
+One of the problems of fractal analysis has always been the difficulty of finding a fractal model for a given set of data.
+The assumption that data contains fractal properties by itself already allows us to calculate certain statistics (such as its dimension), but the various fractal models that have been proposed are difficult to fit to data.
+This is called the \emph{inverse problem}.
+Instead of generating an image from a given mathematical description of a fractal (a popular programming exercise) we are asked to consider an image or dataset and find a good fractal model for it.
+
+This problem was investigated thoroughly in the context of image compression, but has received relatively little attention in the area of statistics, and fitting a fractal model to data rather than an image.
+In particular the case of data in arbitrary dimensions has been left largely uninvestigated.
 
@@
 
+In this paper we focus on numerical data in many dimensions.
+That is, each instance in our dataset is represented by a fixed number of numerical features.
+Call our dataset $X$, our instance $x \in X$ and $x \in {\mathbb R}^d$.
+We will assume that our dataset represents iid draws from a probability distribution over ${\mathbb R}^d$ which we will try to model.
+
+\subsection{Iterated Function Systems}
+
+If we are to fit fractal models to data then the classic examples of single fractals are of little use to us. We require a family of fractals which is easy to  
+parametrize, and likely to generate a large number of organic and useful patterns.
+We will use the formalism of \emph{Iterated Function Systems} (IFSs) for this purpose.
+We will briefly review the formalism here.
+For a more rigorous treament, we recommend [@!cite hutchinson].
+
+Iterated Function Systems are a simple generalization of many of the simpler fractals.
+Consider the Sierpinski Gasket (fig @!).
+As noted before, it consists of three scaled down versions of itself.
+In mathematical terms, if we take the union of three specific transformations applied to the Sierpinski Gasket, we will get the Sierpinski gasket again.
+The key result in the theory of iterated function systems is that for these three transformations there is only one set that has this property.
+The three transformations together form a scaling law, and we say the the Sierpinski gasket satisfies the scaling law.
+
+If we replace the transformations by others (making sure that they are contractive) the set for which this property holds will still be uniquely determined by the transformations, although it will differ from the Sierpinski gasket.
+Another helpful property of IFSs is that when we start with a different set, and we apply the transformations iteratively, the resulting sequence of sets will converge towards the unique single set that satisfies the scaling law.
+
+We will use the following notation. Let $S_1$, $S_2$, \ldots, $S_n$ be a set of contractive transformations, then there exists a single set $K$ that satisfies the following property:
+\[
+K = S_1(K) \cup S_2(K) \cup \ldots \cup S_n(K)
+\]
+
+If we consider the property a function of set $X \subseteq {\mathbb R}^d$: $S(X) = \bigcup S_i(X)$ with and we iterate the function $S^{n+1}(X) = S(S^n(X))$, then we have 
+\[
+\lim_{n \rightarrow \infty} S^n{X_0} = K
+\]
+With $X_0$ any non-empty set and $K$ the unique set that satisfies $S$'s scaling law.
+
+Of course, in statistics, we are usually interested in finding probability distributions rather than sets.
+The IFS formalism can be extended to measures in a natural way.
+First, we must extend the concept of a map to measures.
+Let $p$ be some probability measure over ${\mathbb R}^d$ and $F:{\mathbb R}^d \rightarrow {\mathbb R}^d$ be some map on ${\mathbb R}^d$.
+Then we define the mapping of $p$ by $F$ as:
+\[
+(F(p))(X) = v(F^{-1}(X))
+\]
+So if $p$ is a probability measure, then $F(p)$ is a probability measure as well, which assigns $X$ the same probability as $p$ would assign to $F^{-1}(X)$, where $F^{-1}$ is the inverse of $F$.
+
+Our IFS over probability measures is made up, as before, of $n$ transformations, but here we also introduce $n$ weights $w_i$ with the constraint that they sum to one.
+The full scaling law then becomes
+\[
+k(X) = \sum_{i\in[1, n]} w_i S_i(k)(X)
+\]
+
+As before, there is a unique probability measure that satisfies this scaling law, and if we use the scaling law as a function on probability measures:
+\[
+p = \sum w_i S_i(p)
+\]
+the iteration of this function will converge to the unique value $k$ as described above.
+
+\subsection{Representing IFSs}
+
+As discussed, an IFs is completely determined by its component transformations and their weights.
+The definition does not specify any kind of transformation, to make fitting IFSs to data feasible we must fix a family of transformations.
+For the purposes of this paper we will use similarities (also known as similitudes), transformations consisting of a translation, a rotation and a uniform scaling.
+
+The method described below can easily be extended to other families of transformations such as affine transformations or neural networks.
+The only requirement is that we must have some way of finding a transformation $F$ given two ordered point sets ${x_i}, {y_i} \subseteq {\mathbb R}^d$ so that $|F(x_i) - y_i|$ is minimized.
+For similarities there is a direct and optimal method.
+[@!
+cite]
+
+\subsection{Drawing points from an IFS distribution}
+
+For various reasons, it can be useful to generate a set of points drawn from an IFS probability distribution.
+The most common method to achieve this is known as the \emph{chaos game}.
+The algorithm is simple: we start with some initial point $x_0$ we choose a random component $S_i$ from the IFS (so that $S_i$ is chosen with probability $w_i$) and and let $x_1 = S_i(x_0)$.
+We iterate this process and discard the first 50 or so points.
+The rest of the sequence $x_n$ can be used as a set of points drawn from the IFS.
+
+It should be mentioned that the points generated in this way are not independently drawn.
+Any two succesive points in the sequence will satisfy the relation $x_{n+1} = S_i(x_{n})$.
+This is not true for two independently drawn points.
+For our experiments we generate a sequence and then shuffle the data to remove this dependency where it is relevant.
+
+\subsection{Code space}
+
+Let $(\sigma_i)$ be a sequence with $\sigma_i \in [1, n]$ for some IFS $S$ with $n$ components.
+We can compose the components of the IFS acording to the elements of $\sigma$.
+Call this function $S_sigma$:
+\[
+S_\sigma = S_{\sigma_1} \circ S_{\sigma_2} \circ \ldots \circ S_{\sigma_m}
+\]
+
+Since each $S_i$ is contractive, we can let the length of $\sigma$ go to infinity and get 
+\[
+S_\sigma(X) = x_\sigma  
+\] 
+That is, the sequence of functions maps each $X$ to a unique point $x_\sigma$.
+In the case of probability measures, every probability measure is mapped to a probability measure that gives $x_\sigma$ full probability mass.
+
+We will require the inverse process of this.
+Given a point on or near the support of some IFS, we wish to find $\sigma$.
+To this end we start with the standard normal multivariate distribution and apply all maps $S_\sigma$ for all $\sigma$ of a given lentgh $k$.
+The $\sigma$ with the greatest probability is used as a code for the point.
+
+Since the variance of the distribution drops off exponentially this method will fail if points are to far off the support of the IFS (even if logarithmic probabilities are used).
+To increase the range of the method we fall back on the distance from the point to the mean of the distribution.
+If the first method cannot distinguish between codes, we chose the code whose probability distribution has its mean closest to the point.
+
+We should provide some caution about the use of the codes, as the ordering is often reversed from what our intuition suggests.
+If we number the three components of the Sierpinski triangle clockwise 0, 1 and 2, we might number a small section of the triangle by first moving into a one of the large subsections 2, then into one of the subsections of subsection 1 and then into subsection 0 of that.
+Intuitively, this is the most natural way of coding sections of the Sierpinski triangle.
+[@!
+fig]
+
+If we wish to map the whole triangle to this subset, we must reverse the codes: the transformation that maps to the smallest subsection must be applied first.
+The sequence of transformations which will map the whole triangle into the required subspace is:
+\[
+S_0(S_1(S_2(x)))
+\]
+and its code $\sigma$ is thus $(0, 1, 2)$. If we wish to travel into the Sierpinski gasket by this code, we must read it right-to-left.
+
+We also note that when we use a simple IFS as a model for the codes---such as the two component IFS that maps the interval $[0,1]$ onto its lower and upper half, or the four component IFS that maps the unit square onto its four quarters---its codes coincide with commonly used spatial indexing functions and string based construction for the continuum.
+In this sense we can see the iterated function system as covering a subspace of ${\mathbb R}^d$ and the codes as describing a kind of metric that is intrinsic to the data rather than the embedding space.
+
 \subsection{Dimension}
 
-\section{Modeling}
+The curse of dimensionality is a blanket term for many effects related the analysis of high dimensional data.
+The main cause of these problems is that as dimensionality increases, the volume of a subspace (such as the unit ball) increases exponentially with it.
+This means, for instance, that vast amounts of samples are required to maintain the same level of coverage in a high dimensional space as in a low dimensional one.
 
-There are many well-known algorithms for
+The solution to many of these problems is that while the data has a high-dimensional representation, it has a lopw intrinsic dimension.
+That is, it can be mapped to a low dimensional space without any oss of information.
 
-One of the main challenges in all forms of fractal modelling has been \textit{the inverse problem}.
-Whereas the chaos game produces a set of points from a fractal model, the inverse problem is to find a fractal for a given set of data points.
-In short, fitting a fractal model to data.
+IFSs have a particular advantage over classical models like Gaussian mixtures in that they can vary their intrinsic dimension over the whole range of positive reals.
 
-So far, two approaches have succesfully tackled the inverse problem in some form:
+Whe measure the dimension of datasets and data generated from models using Takens' estimator:
+\[
+D(d_{\mbox{max}}) = -\frac{|A|}{\sum_{a \in A} \ln a} 
+\]
+Where $A$ is the set of distances between pairs of data points lower than parameter $d_{\mbox{max}}$.
+We choose $d_{\mbox{max}}$ as the value that produces the lowest Kolmogorov Smirnov statistic with the probability distribution underlying the Takens estimator.
+This method is based on the work for general power laws in @!
+and is detailed in @!.
+
+\subsection{Evaluating models}
+
+The standard way of evaluating the fit of a model to a dataset is the likelihood of the model given the data.
+This is often used in optimization algorithms for fitting probability models to data.
+In the case of fractal models, the likelihood function has an important drawback.
+
+Consider a dataset drawn from the Sierpinski gasket with a minute amount of two dimensional noise added.
+Since the Sierpinski gasket has dimension less than two, the probability that the point will still fall within the Sierpinski gasket will be zero.
+Thus all points will have probability zero under the Sierpinski gasket and its likelihood will be zero.
+Intuitively, the Sierpinski gasket is a good model for this data (since the noise term can be made arbitarily small).
+
+For this reason we will often use the Hausdorff distance to evaluate our models. The Hausdorff $H(X, Y)$ distance between two sets $X$ and $Y$ is defined as follows:
+\[ 
+H'(X, Y) = \max_{x\in X} \min_{y\in Y} d(x, y) 
+\]
+\[
+H(X, Y) = \max\left \{H(X, Y), H(Y, X)\right\}
+\]
+ (for infinite sets the $\max$ and $\min$ functions should be replaced with $\sup$ and $\inf$)
+
+The most intuitive way of understanding the Hausdorff distance is as a game where player one must get from one set to the other in as small a distance as possible, and the the second player will choose his starting point to maximize that distance.
+
+Hausdorff distance is a common tool in dealing with the fractal inverse problem [@!
+cite].
+We will use it to evaluate an IFS model by generating a number of points from the model and calculating the Hausdorff distance of that set to the model.
+Since the Hausdorff distance has not been substantiated as a statistical measure, we can only use it as an intermediate indication of model quality.
+
+\subsection{Analogous model: Mixture of Gaussians}
+
+To study the IFS as a statistical model, it helps to have an anlogous model to compare it with.
+We use the Mixture of Gaussians (MoG) model.
+This model consists of $n$ multivariate Gaussians over ${\mathbb R}^d$ each represented by a mean $\mu \in {\mathbb R}^D$ and a $d \times d$ covariance matrix $\Sigma$.
+Each component also has a prior $w_i$.
+The complete model described the following probability distribution:
+\[
+p(X) = \sum_{i \in [1,n]} w_i N(x|\mu_i, \Sigma)
+\]
+Since every multivariate Gaussian can also be described as an Affine transformation of the standard normal distribution (mean at the origin and $\Sigma$ the identity matrix), we can also represent the mixture of Gaussians as $n$ affine transformations and $n$ weights, just as the IFS:
+\[
+p(X) = \sum_{i \in [1,n]} w_i S_i(N(\mu_i, \Sigma))(X)
+\]
+This gives us a classical model with a similar representation.
+Note that the representation would be equal if we used similarities for the MoG model.
+However since an MVN transformed with uniform scaling is rotationally symmetric that would make all the rotational parameters useless.
+For this reason we will allow the MoG model full affine transformations as parameters and the IFS model only similarities.
+Even with this discrepancy, we still consider the MoG to be a reasonable analog.
+
+Unlike the single MVN distribution, there is no closed form solution for the maximum likelihood fit of a MoG model to data.
+We must approximate it.
+The most common method is by the well-known EM algorithm.
+Since our solution to the fractal inverse problem takes inspiration from this algorithm, we describe it here briefly.
+The key intuition behind the EM algorithm is that if we knew which component of the MoG model `generated' each datapoint, we would have a closed form solution for the parameters of the component, and if we knew the parameters of the components we could find out which component was most likely to have generated the data point.
+We turn this circular dependency into an iterative algorithm.
+Starting with some initial MoG model we iterate the following steps
+
 \begin{itemize}
-  \item Genetic algorithms!@.
-A straightforward search through the space of IFSs.
-The drawback of this approach is one of time complexity.
-The running time of these approaches scale poorly in the number of dimensions and components.
-  \item Partitioned iterated function systems!@.
-This is a modified version of the problem to make the model tractable and feasible in the domain of image compression.
-This approach does not translate well to n-dimensional point patterns.
+  \item Expectation: Assign each data point a probability distribution over the components of the MoG model proportional to the likelihood of the model given the datapoint.
+  \item Maximization: Estimate the parameters for each component based on the probability weights calculated in the previous step.  
 \end{itemize}
 
+[@! did I name the steps correctly?]
+
+\section{Fitting IFS models to data}
+
+We can now describe the central algorithm of this paper.
+To make the explanation as clear as possible we will first describe a skeletal version of the algorithm that works in principle and then describe various alterations to this skeleton that are required to improve the performance of the algorithm on real data with finite approximations of the IFS measures.
+
+The results described are always for the full algorithm with all improvements, with the exception of the branching mode (which we test as a separate algorithm).
+An implementation of only the basic algorithm should be enough to retrieve the Sierpinski gasket (or a crude approximation) from data.
+
+\subsection{The basic algorithm}
+
+As with the EM algorithm for MoG models, the main barrier to inducing an IFS model from the data directly is a dependency on latent variables.
+In the MoG case each data point needed to be assigned to a component, in the IFS case we need to know, for each data point the code $\sigma$ that represents the sequence of components.
+
+Given these codes it's a relatively simple matter to find which points maps to which under the data's self-similarity.
+And given two ordered lists of points that are supposed to map into one another we can easily find a map to minimize the error.
+Thus our algorithm works as follows.
+\footnote{the names `expectation' and `maximization' of these steps are only by analogy with the EM algorithm for MoG models. We do not claim any true expectation or maximization procedures.}
+
+\begin{itemize}
+\item We start with in initial IFS $S$, chosen by one of the methods described below. 
+\item (Expectation) We assign each datapoint a code $\sigma$ of length $k$ (where $k$ is a parameter of the algorithm). The codes are provides by the model $S$ as described in section @!.
+\item (Maximization) We know that if a point $x_\sigma$ is mapped using one of our components we have $S_i(x_\sigma) = x_{i\sigma}$.
+That is, the symbol for the mapping component is prepended to the code.
+For each $x_\sigma$ in our dataset we search the data for a point with a code $\tau = i\sigma_{:-1}$, the symbol $i$ followed by the elements of $\sigma$ save for the last.
+For each $i$ we maintain ordered lists of the matches found.
+\item (Maximization) For each component $S_i$ we know have two lists of points such that $S_i$ should map one into the other with minimal squared error.
+We used the method described in @!
+to find the optimal similarity transformation.
+\end{itemize}
+
+\subsection{Improvements}
+
+\subsubsection{Centering and scaling the data}
+
+Iterated function systems are not very good at approximating data that is weighted far from the origin.
+for this reason we find the affine transformation which maps the data to the range $[0,1]$ in each dimension and puts the mean at the origin.
+
+We store this transformation with the model to map the model back to the original space.
+
+\subsubsection{Using all depths}
+
+The algorithm as described here can work at all depths $k$.
+Since each depth may contain information about self similar scaling in the data, we want to consider all depths $\leq k$ rather than just $k$.
+This is easily done by considering codes of all length up to and including $k$.
+We build a code tree of the codes generated from the dataset and search it systematically for the required matches at all levels.
+
+\subsubsection{Multiple matches}
+
+It is prefectly possible that we may get multiple points per code.
+Indeed, at low depths, it is almost certain.
+In this case we have two sets of points $X_\sigma$ and $X_\tau$ such that the points in the formar should map onto the latter.
+But our algorithm for finding the map $S_i$ requires the individual points to be matched to each other.
+If the number of points is low we can simply take the mean of both sets, but if the number of points is high, we risk throwing away a lot of information (in particular about rotation).
+Theoretically each point in the first set should map onto every point in the other but this creates an explosion of possiblities.
+If we pair the points up randomly we introduces artifacts in the induced rotation.
+
+We tackle this problem by finding the MVN for both sets.
+Ideally we would then extend our method for finding maps to paired MVNs rather than paired points, but this is a highly complex problem.
+Instead, we find for each MVN the transformation that transforms the standard distribution into that MVN.
+We then generate a small amount (20 in our experiments) of points from the standard MVN and map these to both sets.
+We add these 20 point pairs to the list of matched points from which the maps will be calculated.
+
+\subsubsection{Weights}
+
+The algorithm as described above does not give us a way to retrieve the component weights.
+For the coding step we already take the weights into accounts, so the only thing we must do is retrieve them from the coded dataset.
+
+If we have a collection of points $X_\sigma$ with code $\sigma$ and a collection of points $X_\tau$ with code $\tau$ such that $S_i$ should satisfy $S_i(x_\sigma) = x_\tau$, we know that if we generate the IFS model, $w_i$ of the probability mass at stage $t$ always gets distributed to component $S_i$ for stage $t + 1$.
+thus we can expect the relative size of these sets to follow $w_i$: $|X_\tau| = w_i |X_\sigma|$.
+Over all such sets $X_\sigma^j$ and $X_\tau^j$ we must minimize the squared error:
+\[
+\overline{w_i} = \argmin_{w_i} \sum_{j} \left( |X_\sigma^j| - w_i |X_\tau^j| \right)^2
+\]
+This works out as 
+\[
+\overline{w_i} = \frac{\sum_{j}|X_\sigma^j||X_\tau^j|}{2\sum_{j}|X_\tau^j|^2}
+\]
+
+\subsubsection{Edge cases}
+
+Some undesired behavior can emerge when the initial model is far from the optimum.
+Consider for instance the case where the Sierpinski triangle is the current model and all data isto the top right of the triangle.
+The symbol 2 would not occur in any of the codes and we would have no basis to find $S_2$ in the maximization step.
+If this occurs, we will `split' one of the components that could be found in two (distributing the weight equally) and perturb each component slightly to ensure that the components differ.
+The idea behind this approach is simple.
+For every place in every code that used to be assigned to the original component there is now around a fifty/fifty probability of the place being assigned to the other code.
+In the next step roughly half of the matches of the component will have been `donated' to the other component.
+
+The perturbation is done by flattening the model into a vector of values in ${\mathbb R}$ and ading a small random value $N(0, \sigma)$ to each. In our experiments $\sigma$ was set to 0.1.
+
+\subsubsection{Branching}
+
+[@! Test if this helps]
+
+As an optimization algorithm, EM is strictly linear in the sense that it only considers a single model at each iteration.
+To avoid local minima it can often help to evaluate multiple models per iteration.
+To test whether our algorithm benefits from such additions we introduce a branching mode which works as follows.
+
+During the iteration of the algorithm we maintain a buffer of models.
+After each iteration, we generate $b$ versions of the current best model perturbed by a parameter $\sigma$ as with the edge cases.
+We add these models to the buffer and sort it by Hausdorff distance with the data (as described in section @!).
+We discard all but the top $h$ models.
+At the next iteration, we continue with the top model from the list and repeat the procedure.
+
+This method is analogous with beam search.
+Its advantage is that we perform the expensive EM step the same number of times as the non-branching model, and the extra cost lies purely in testing evaluating the extra models.
+(This is not inexpensive in itself, but any branching algorithm will need to test candidate models anyway).
+
+\section{Results}
+
+There are three expected benefits to fractal modeling that we wish to expore here.
+First, we expect that the models will scale their dimension to fit that of the data (whereas our comparison model will be unable to do so).
+To test this we will fit models to various datasets and estimate their dimension.
+
+Second, we expect that the IFS distributions make good fits.
+Since we cannot test this directly using likelihood, we will embed the models in a classifier and test its performance on various data sets.
+
+Finally, we expect that the codes described in section @!
+are of actual use outside of the fitting procedure.
+These codes describe a metric space that is much more compact\footnote{in an informal sense} that the Euclidean space in which the data is embedded.
+Furthermore, the transformation to strings allows the use of a whole family of technique that was not previously available.
+
+Finally we give some indication of the time preformance of the algorithm.
+
+\subsection{Dimension scaling}
+
 \subsection{Classification}
-\section{Conclusions and Further Work}
 
-One advantage of the IFS model that has been left unexplored in the current work is that it assigns each point in the dataset a $k$-ary string so that a small distance between points suggest simlar strings.
-Thus, a point pattern in a Euclidean space is naturally mapped to the space of strings in a way that uses the structure of the dataset rather than the structure of the embedding space.
+\subsection{Coding}
 
-Future investigation might focus on the use of IFS models in clustering, spatial indexing and the induction of random fractals (as discussed in \cite{})
+@! Investigate the use of DFAs on the code form of the dataset. Use for classification.
+@! Map the codes back to a space ${\mathbb R}^{d'}$ as a method of dimensionality reduction.
+
+\subsection{Performance}
+
+\section{Conclusions}
+
+We have presented a framework for a fractal based approach to the statistics of point patterns.
+We introduced an EM-style algorithm for fitting fractal models to data.
+Compared to search algorithms for other types of models the algorithm is relatively expensive, but compared to other approaches for fractal modeling the algorithm represents a significant imporvement @!.
+
+We have given some indication of how the study of fractal modeling can open up interesting new prospects in machine learning, for instance by 
 \end{document}

### 2012-08-22 Dimension/src/dimension.tex
...

@@
 
-\theoremstyle{definition}
-\newtheorem*{thm}{Theorem}
+\newtheorem{thm}{Theorem}
 \newtheorem{lma}{Lemma}
-\newtheorem*{dfn}{Definition}
-\newtheorem*{exm}{Example}
+\newtheorem{dfn}{Definition}
+\newtheorem{exm}{Example}
+\usepackage{lscape}
 
-\title{Measuring Dimension}
+\usepackage{float}
+
+\title{Measuring dimension across scales}
 \date{\today}
@@
 \setlength{\parskip}{0.1in}
+
+% pseudocode 
+\floatstyle{ruled}
+\newfloat{pseudo}{h}{lop}
+\floatname{pseudo}{Algorithm}
 
@@
 
+\section{The Takens estimator}
+
+Like the Hill estimator. State assumptions for MLE.
+
+\subsection{Testing the independence assumption}
+
+If we use each point for only one pair, (and our dataset is iid) then the distances for those pairs will be iid samples from $p_D$.
+If we reuse points, we will invalidate the first assumption.
+Consider for instance, a dataset of three points $a$, $b$ and $c$ drawn independently from some distribution.
+If we use only $d(a, b)$ we get a single sample from $p_D$.
+If we use $d(a, b)$ and $d(a, c)$, the two values are dependent the location of $a$.
+we are in fact sampling twice independently from the distribution of $d(a, X)$ rather than $d(X, Y)$.
+
+If we use all pairs from a large dataset, however, the recurrence of single points will simply mean that we are sampling from $d(a, Y)$ averaged over many different $a$.
+Thus we expect that for a large dataset, the distances of all pairs do not form a substantially different distribution than the distances over successive pairs, while giving us a squared increase in the volume of data.
+
+We test this assumption by taking datasets in full (after shuffling the data) and performing the Kolmogorov-Smirnov statistic on them. The results are shown in table @!.
+
+\subsection{Testing the distribution assumption}
+
 \subsection{Physics}
 \subsection{Statistics and Machine Learning}
-\subsection{}
 
-\section{Methods}
+\section{Estimating dimension}
 
@@
 Discussion of the running time an an investigation of the effects of reducing the data size.
+
+\section{A multiscale estimator}
+
+Consider a datset produces by drawing random points in some 1D curve in 2D Euclidean space, but with a small term of 2D Gaussian noise added.
+The theoretical dimension of this process, for any definition of dimension will be 2, since the limit towards the smallest scales determines the dimension.
+However, the shape that we are interested in (the distribution without the noise term) has dimension 1.
+
+This is an example of a distribution with multiple intrinsic dimensions.
+The data is generated by different processes that each create a dimensionality in the data at a particular scale.
+It's not uncommon to see data with more than two intrinsic dimensions.
+In the log-log plot produces by the Grassberger-Procaccia estimator, this behavior will produce linear segments for the relevant length scales.
+
+From a natural science perspective, finding these length scales in data can provide great insight.
+They hint at clearly separable processes producing the data, each with their own intrinsic length scales.
+However as noted before, the Grassberger-Procaccia estimator relies on subjective judgement to find these scales.
+Here, we adapt our method for single-sclae dimension measurement to multi-scale measurements.
+
+To start with we assume that our data has some number of intrinsic scales determined by $\{r_i\} \in {\mathbb R}^+$ and $r_0 = 0$ so that we have an intrinsic dimension $D_t^i$ over distances in the half-open interval $[r_{i-1}, r_{i})$.
+We let the maximum $r_i$ be less than the maximum distance in the dataset, so that as with the single-scale estimator, our distance distributin has a tail that is not covered by the dimensional scaling (and about which we make no assumptions).
+
+We define the correlation integral over an interval simply as $C(r_{i-1}, r_i) = C(r_i) - C(r_{i-1})$. Our assumption that the dimension $D_t^i$ determines scaling in this interval tells us that 
+\[
+p(d \leq r | d \in [r_{i-1}, r_i)) = \frac{p(d \leq r | d \in [r_{i-1}, r_i))}{p(d \d \in [r_{i-1}, r_i)])} = \frac{p(d \in [r_{i-1}, r])}{p(d \in [r_{i_1}, r_i))}
+\]
+\[
+ = \frac{C(r) - C(r_{i-1})}{C(r_i) - C(r_{i-1})} = \frac{cr^{D^i_t} - cr_{i-1}^{D^i_t}}{cr^{D^i_t} - cr{i-1}^{D^i_t}}\
+\]
+
+We turn this into a probability function by taking the derivative over $r$, and we use this to compute the score (the derivative of the log likelihood of the data).
+Setting the score to zero gives us a maximum likelihood estimator for $D^t_i$.
+This process is detailed in appendix @!.
+Unlike the single scale version, we do not find a closed form version of the MLE.
+We are left with:
+\[
+\frac{1}{D^t_i} - \frac{\ln \frac{r_i}{r_{i-1}}}{(\frac{r_i}{r_{i-1}})^{D^t_i}} - \frac{\Sum_{r \in R_i}\ln r - |R_i| \ln r_{i-1}}{|R_i|} = 0
+\]
+
+Which is unlikely to be expressible in $D^t_1$.
+However, as proved in the appendix, the function is strictly decreasing in $D^i_t$, for the required values of the constants, so that we can use a simple binary search to pinpoint its root.
+Call the the left hand side of the equation above $f(x)$.
+Then the search algorithm works as follows
+
+\begin{pseudo}[h]
+	\textbf{starting with $x = 1$ repeat $x \leftarrow 2x$ until $f(x)$ becomes negative}  \\
+	\hspace*{5mm}\textit{\# This gives us an upper bound for $f(x) = 0$}\\
+	\textbf{return find}$(0, x)$ \\
+
+	\textbf{function find}$(x_{\min}, x_{\max})$ \\
+	\hspace*{5mm} $x = (x_{\max} + x_{\min})/2$ \\
+	\hspace*{5mm} \textbf{if} $x_{\max} - x_{\min} < \epsilon$ \\
+	\hspace*{10mm} \textbf{return} $x$\\
+	\hspace*{5mm} \textbf{if} $f(x) < 0$
+	\textbf{return} $\mbox{find}(x_{\min}, x)$ \\ 
+	\hspace*{5mm} \textbf{else} 
+ 	\textbf{return} $\mbox{find}(x, x_{\max})$ \\	
+\end{pseudo}
+
+This algorithm first searches for an $x$ that makes $f(x)$ negative, and then bisects theinterval $[0, x]$ recursively until the interval is smaller tha a given $\epsilon$, at which point the middle of the interval is taken as the estimate of $D_i^t$.
+
+To find the interval limits $r_i$ we proced as we did before with $d_{\max}$ and search for a set of values from the data which minimize the Kolmogorov-Smirnov statistic with the filtered data.
+We compare each interval with the data filtered to that interval and take the sum of all intervals for a given set as the value to minimize.
+As searching a for a single value was already expensive in the single-scale case, here we must give up on the idea of checking all possible values and instead search for a suitable set of ranges.
+
+@!NOTE: Try contiguous and non-contiguous scaling intervals.
+
 \section{Experiments}
@@
 
-\begin{itemize}
-  \item Swiss roll
-  \item 8-loop
+\pagebreak
+\begin{landscape}
 
-  \item Hands
-  \item Faces
+\begin{tabular}{l | r r r r}
+\hline
+data set & size & dimension & significance & true dimension \\
+\hline
+Swiss roll & 1000 & & & \\
+Noisy swiss roll ! & 1000 & & & \\ 
+Sierpinski & 1000 & & & \\
+Cantor & & & & \\
+Menger Sponge ! & & & & \\
+Persistent process ! & & & & \\
 
-  \item Sierpinski
-  \item Cantor
-  \item Menger Sponge
+MVN 50 & 1000 & & & \\ 
+MVN 500 & 1000 & & & \\
 
-  \item 
-\end{itemize}
+British coast line & & & & \\
+
+EEG & & & & \\
+JAFFE expression database & & & & \\
+Currency (Traina/Faloutsos) ! & & & & \\
+Galaxies & & & & \\
+DNA random walk & & & & \\
+Turning cup & & & & \\ 
+Faces & & & & \\
+
+Newsgroup posts (edit distance, possibly normalized by length) & & & & \\
+Protein sequences (mPAM metric) & & & & \\
+Graphs (Weisfeiler Lehman) & & & & \\
+\hline
+\end{tabular}
+
+\end{landscape}
+\pagebreak
 
@@
 
+\section{Appendix}
+
+\subsection{Derivation of the MLE}
+
+We start with the cumulative distribution function below $d_{\max}$:
+\[
+p(d \leq r | d \leq r_{\max}) = \frac{p(d \leq r)}{p(d \leq r_{\max})} = \frac{C(r)}{C(r_{\max})} = \frac{c \cdot r^{D_t}}{c \cdot {r_{\max}} ^ {D_t}} = \left ( \frac{r}{r_{\max}} \right )^{D_t}
+\]
+
+We take the derivative (over $r$) to get the probability density function:
+
+\[
+p(d = r | d \leq r_{\max}) = D_t \frac{r^{D_t-1}}{{r_{\max}}^D_t} = D_t \frac{1}{r} \left (\frac{r}{r_{\max}} \right ) ^ {D_t} 
+\]
+
+Let $R$ be the set of distances in the dataset less than $r_{\max}$. The the likelihood of this dataset is
+\[
+{\cal L}(R|D_t) = \prod_{r \in R} D_t \frac{1}{r} \left (\frac{r}{r_{\max}} \right ) ^ {D_t} 
+\] 
+Which we must maximize. Maximizing the logarithm of the likelihood is equivalent, and easier.
+\begin{align*}
+\ln {\cal L}(R|D_t) &= -\sum \ln r + D_t \sum \ln r - D_t |R| \ln r_{\max} + |R| \ln D_t
+\end{align*}
+
+To maximize, we take the derivate with respect to $D_t$
+
+\begin{align*}
+\frac{\partial \ln{\cal L}(R|D_t)}{\partial D_t} = \sum{\ln r} - |R|\ln r_{\max} + |R|\frac{1}{D_t}
+\end{align*}
+
+and set it equal to zero and solve for $D_t$
+
+\begin{align*}
+ \sum{\ln r} - |R|\ln r_{\max} + |R|\frac{1}{D_t} &= 0 \\
+|R|\frac{1}{D_t} &= - \sum \ln r + |R| \ln r_{\max} \\
+D_t &= \frac{|R|}{- \sum \ln r + |R| \ln r_{\max}} 
+\end{align*}
+
+Many publications assume that the data $|R|$ are normalized by $r_{\max}$ which allows the cleaner result:
+\begin{align*}
+D_t &= \frac{|R|}{- \sum \ln r} 
+\end{align*}
+
+However, this hides the reliance on $r_{\max}$. As we will be iterating over all possible values for $r_{\max}$, it's easier to include in the estimator explicitly.
+
+\subsection{Derivation of the multiscale MLE}
+
+The only difference between the single- and the multiscale MLE is that the latter has a lower bound to its range in addition to an upper bound.
+Unfortunately this small difference causes the derivation of the MLE to become a great deal less elegant.
+
+We start with the cumulative distribution function as described in the text. We use $r_{i-1} = r_j$ to clear up the notation.
+
+\begin{align*}
+p(d \leq r | d \in [r_{i-1}, r_i)) = \frac{r^{D_i} - {r_{j}}^{D_i}}{{r_i}^{D_i} - {r_{j}}^{D_i}} = \frac{\left(\frac{r}{r_j}\right)^{D_i} - 1}{\left(\frac{r_i}{r_j}\right)^{D_i} - 1}
+\end{align*}
+We take the derivative with respect to $r$ to get the probability density function.
+
+\begin{align*}
+p(d = r | \ldots) = {D_i}r^{D_i-1} \frac{1}{\left(\frac{r_i}{r_j}\right)^{D_i} - 1}
+\end{align*}
+
+This gives us the following log likelihood function:
+\begin{align*}
+\ln{\cal L}(R_i|D_i) &= \sum \ln D_i + (D_i - 1)\sum \ln r - \sum \ln \left(\left(\frac{r_i}{r_j}\right)^{D_i} -1\right)\\
+&=  |R_i|\ln D_i + (D_i - 1)\sum \ln r - |R_i| \ln \left(\left(\frac{r_i}{r_j}\right)^{D_i} -1\right) 
+\end{align*}
+
+Where $R_i$ is the subset of distances in the dataset that fall within the interval $[r_{i-1}, r_i)$.
+
+Taking the derivative:
+\begin{align*}
+\frac {\partial \ln{\cal L}(R_i|D_i)}{\partial D_i} &=  \frac{|R_i|}{D_i} + \sum \ln r - |R_i| \frac{\left(\frac{r_i}{r_j}\right)^{D_i} \ln \frac{r_i}{r_j}}{\left(\frac{r_i}{r_j}\right)^{D_i} -1}
+\end{align*}
+
+Setting to zero and solving for $D_t$
+
+\begin{align*}
+\frac{|R_i|}{D_i} + \sum \ln r - |R_i| \frac{\left(\frac{r_i}{r_j}\right)^{D_i} \ln \frac{r_i}{r_j}}{\left(\frac{r_i}{r_j}\right)^{D_i} -1} &= 0 \\
+\frac{1}{D_i}  - \frac{\left(\frac{r_i}{r_j}\right)^{D_i} \ln \frac{r_i}{r_j}}{\left(\frac{r_i}{r_j}\right)^{D_i} -1} &= \frac{-\sum \ln r}{|R_i|}
+\end{align*}
+
+@! If we pre-normalize $R_i$ we may still get a closed form solution!
+
+\subsection{Proof of monotonicity}
+
+The following lemma tells us that the function describing the maximum likelihood estimate for the multiscale case is strictly decreasing over positive values, so that we can search for its root with a simple binary search.
+
+\begin{lma}
+The function
+\[
+f(x) = \frac{1}{x} + \frac{\ln(a)}{a^x - 1} + c
+\]
+is strictly decreasing for $ x \in [0, \infty)$ and $ a \in [0, 1]$ 
+\end{lma}
+\subsection{Algorithms}
+
 \end{document}

### 2012-08-24 Dimension/src/dimension.tex
Writing

@@
-\documentclass{article}
+\documentclass[10pt]{article}
 
@@
 
-\title{Measuring dimension across scales}
+\title{Measuring dimension}
 \date{\today}
@@
 \begin{abstract}
-The intrinsic dimension of a dataset is one of the most telling statistics that can be calculated, and one that has received a great deal of attentation over the past decades.
-Most methods for estimating dimension relsy on at least some subjective judgment to make a selection in the data.
-Adapting methods recently described [@!
-cite] for the statistical analysis of power laws, we revisit an existing MLE estimator and present a simple, objective procedure to estimate the dimension of a dataset.
-The procedure runs without human intervention, and produces a value for the dimension, uncertainty bounds and a significance value.\end{abstract}
+\noindent The intrinsic dimension of a dataset is one of the most telling statistics that can be calculated, and one that has received a great deal of attentation over the past decades.
+Most methods for estimating dimension rely on at least some subjective judgment to make a selection in the data.
+Adapting methods recently described in \cite{clauset2007power} for the statistical analysis of power laws, we revisit an existing MLE estimator and present a simple, objective procedure to estimate the dimension of a dataset.
+The procedure runs without human intervention, and produces a value for the dimension, uncertainty bounds and a significance value.
+We also extend the procedure to extract multiple characteristic dimensions from a dataset and revisit the concept of \textit{lacunarity} in the context of these methods.\end{abstract}
+\section{Introduction}
+The \textit{intrinsic dimension} of a dataset does not refer to how many features are used to represent it, but---loosely speaking---to how few features could be used to represent it.
+It is also the exponent of various power laws that determine the way the internal structure of the dataset scales.
+For many datasets, in particular those with coarse ``fractal'' features or those drawn from a chaotic process, dimension is one of the few well defined statistics available, and usually the most expressive.
 
-\section{Introduction}
+In the natural sciences, the intrinsic dimension takes the role of the observable: predicted by theory, and derived from experiment.
+Fractal phenomena do not respond well to traditional geometric measures.
+We cannot measure the length of a coast line, or the surface area of our lungs.
+In the words of Eugene Stanley \cite{stanley1986growth}:
+
+\begin{quotation}
+\small
+\noindent If you are an experimentalist, you try to measure the fractal dimension of things in nature.
+If you are a theorist, you try to calculate the fractal dimension of of models chosen to describe experimental situations; if there is no agreement then you try another model.
+\end{quotation}
+
+As noted by Stanley, this is only a simplistic view and the study of the dimension of a dataset can also give vital clues to scaling behaviors and critical phenomena in the process from which the data originated.
+Thus the theorist can not only validate his work by the dimension, but also use it as a guide.
+When we investigate the notion of dimension more closely we must give up intuitive notion and the only thing we can say is that the gemotry of a set or space produces various power laws related to scaling and that these power laws will usually have the same exponent in common.
+There is a common intuition in the natural sciences that power law phenomena with the same exponent will likely share some similarity in their underlying components.
+Thus the dimension can be an important guide in finding the universal principles common to  complex phenomana.
+
+The fields of machine learning and statistics have their own interest in the notion of dimension.
+Intrinsic dimension has been shown to be an effective lower bound to the number of features to which a dataset can be reduced without losing information.\cite{kumaraswamy2008fractal} Since many methods and algorithms work faster and more accurately on low dimensional data, dimensionality reduction is an important factor in bringing high dimensional data into the domain of machine learning.
+The field of manifold learning in particular has succeeded in mapping data in many hundreds of dimensions to low dimensional euclidean spaces\cite{roweis2000nonlinear} and often relies on acurate estimates of intrinsic dimension.\cite{levina2004maximum}
+
+Most methods for measuring dimension have relied at least in some part on subjective judgements and processes that make it difficult to capture experimental error.
+A similar situation was present in the study of power law distributions until recently Clauset, Shalizi and Newman published \cite{clauset2007power}, in which they dealt with various problems, and described a simple objective approach for estimating the exponent of such distributions with no subjective elements and a clear indications of uncertainty and significance with respect of the assumptions.
+
+The methods in this paper come from the realization that the relatively unknown Takens estimator for dimension is also based on the assumption of a power law distribution (albeit of a different form than the ones studied by Clauset et al.), allowing these methods can be adapted to the estimation of dimension.
+This process also gives us a means to determine multiple dimensions for datasets that can be considered to have different dimensions for different ranges of scales, and we can give a new measure for the lacunarity, a measure of the `gapiness' or texture of a dataset that is often used in combination with dimension.
+
+There is a wealth of research on the subject of dimension and related subjects.
+For a good overview on definitions of dimension and proposed estimators, we recommend \cite{theiler1990estimating}.
+For a derivation of dimension from basic mathematics (topology and metric spaces) we refer the reader to \cite{edgar2008measure}.
+
+\section{Dimension}
+
+Dimension is usually introduced in layman's terms as the ``amount of numbers required to define a single point on a surface, in a volume or in some other set''.
+This definition, while it explains why a line has lower dimension than a square or the surface of a sphere, breaks down on closer inspection.
+There are may ways to encode three numbers into one (for instance with the use of space-filling curves, or simply by alternating digits) so that we can, theoretically, locate a single point in a volume by only one coordinate.
+
+Nevertheless, something fundamental seems to separate basic curves from figures in the plane from figures in space.
+Some property that we have come to call dimension, before defining it rigorously.
+In fact, investigations show that once notions of dimension become well defined they can be naturally extends to metric spaces, or even topological spaces.
+
+Topology has several (equivalent definitions) defined inductively.
+For instance the covering dimension is based on the idea that a set of dimension $d$ can be covered with open sets so that no more than $d-1$ will overlap anywhere.
+Starting with the convention that the empty set has dimension $-1$, this gives us a complete notion of dimension.
+
+A spanner was thrown in the works with the investigation of \textit{fractals}, Coarse, non-smooth shapes that reveal ever greater detail as one `zooms in'.
+The problem with these figures was that they did not seem to obey the Euclidean scaling laws.
+An example of such a law is that scaling a square so that its sides increase by $s$, will increase its surface by $s^2$.
+Similarly, a cube scaled in the same way will see its volume increased by $s^3$ and its surface area will increase by $s^2$.
+By the seventeenth century, the scaling laws had been generalized to all curves, surfaces and solids.
+Except fractals.
+
+To illustrate how fractal figures seem to break the scaling laws consider figure @!.a (adapted from [@!
+Cite Manfred Schroeder]).
+This figure contains seven hexagon tiles arranged into a larger shape.
+We can now replace each line segment with three smaller line segments as shown in the figure.
+We note that this operation does not change the surface area of the figure.
+Afterwards we are left with another collection of line segments, so we can apply the operation again.
+In the limit we get the figure shows in @!.2, a fractal.
+It can be shown that this fractal shape is invariant to the transformation.
+
+The issue arises when we recognize that the outline of this figure is exactly a scaled up version of one of its tiles.
+This is possible because the outline we started with was a scaled up (and slightly rotated) version of the tiles after one transformation.
+Thus, the outline is always one transformation ahead of the tiles, a difference which disappears in the limit.
+The length scaling between the outline at step $n$ and a tile at step $n+1$ is $3$.
+$3^2 \neq 7$ so it seems like this figure violates the scaling laws.
+
+The situation is resolved when we re-evaluate our notion of dimension.
+Consider the outline of the figure, traditionally seen as a 1 dimensional curve.
+If we measure its length (a 1 dimensional measure) we see that it goes to infinity in the limit of our process.
+If we take from this the hypothesis that its dimension is not $1$, since the 1-dimensional measure fails, but greater than one, we might try to measure its surface area by covering it with ever smaller squares and summing their areas.
+This measure goes to zero.
+We are left with the impression that this shape has a dimension above $1$ and below $2$.
+
+In fact, if we decide to hold to the scaling law and release our grip on the notion of dimension qwe see that the value that makes the scaling law work will satisfy $3^d = 7$, which gives us  $d = \log_3 7 \approx 1.77$.
+A non-integer dimension.
+Upon investigation, the choice was easy, the fractal dimension (properly defined) were consistent and well behaved.
+Our notion of dimension was relaxed, and the scaling laws were upheld.
+
+Not only were these pathological counterexamples given a place in Euclidean geometry, they were connected to many natural phenomena.
+The coast-line of Britain, for example, behaves the same way as the side of our tiles.
+If we try to measure its length, our values for increasing accuracy of measurement will go to infinity.
+The same goes for the surface area of a human lung or a mountain range.
+Luckily, the rate with which conventional these measures fail allows us to get a hint of the intrinsic dimension.
+
+\subsection{Hausdorff dimension}
+
+[@! is this necessary? Maybe move it to appendix and derive the scaling laws for box counting and correlation there]
+
+For theoretical purposes, the Hausdorff dimension is the most popular definition of dimension.
+It's not much in measuring dimension, but it can help us understand where the various scaling laws come from that help us to measure notions of dimension that are more suited to estimation from data.
+Of course, most of the notions of dimension agree with each other for most sets, or have well-understood equalities.
+
+A complete treatment of Hausdorff dimension is beyond the scope of this article (@!
+Edgar provides a reasonably direct overview of the basic requirements for understanding Hausdorff dimension).
+We will give some intuition and the basic definitions.
+
+The basic notion behind Hausdorff dimension is that there are measures that are intrinsic to a given dimension (ie.
+length is related to dimension 1 and area to dimension 2) and that these measures will be $\infty$ or $0$ when they are applied to sets with dimension below or above the true dimension respectively.
+A fractal with dimension 1.5 will have length $\infty$ and area $0$.
+
+The Hausdorff dimension is defined over metric spaces. Euclidean spaces and sets in Euclidean space are specific instances of metric spaces. We will first introduce some basic concepts.
+
+\begin{dfn}
+A covering $E$ of a metric space $S$ is a sets of sets such that $S \subseteq\bigcup_e \in E e$. 
+\end{dfn}
+
+\begin{dfn}
+The diameter of a set $X$ in metric space $S$ with metric $\sigma$ is
+\[
+\mbox{diam}(X) = \sup_{x, y \in X}\sigma(x, y)
+\]
+\end{dfn}
+
+\begin{dfn}
+The quantity $H^d(x, r)$ of a metric space $X$ is the covering of $X$ that minimizes the sum of the diameters raised to $d$:
+\[
+H^d(X, r) = \inf_E \sum_e \in E \mbox{diam}(e)^d
+\]  
+\end{dfn}
+
+Where the infimum is taken over all coverings $E$.
+
+If we let $r$ got to zero, we get the measure mentioned above
+
+\begin{dfn}
+\[
+H^d(X) = \lim_{r \rightarrow 0} H^d(X, r)
+\] 
+\end{dfn}
+
+This measure has an intrinsic dimensionality $d$.
+If it is used on sets with the wrong dimension, it produces $0$ or $\infty$.
+Specifically, it can be shown that for each argument $X$ there is a single unique value $D_H$ such that
+\[
+H^d(X) = \left \{ 
+\begin{array}{rl}
+  \infty &\mbox{if} d < D_H(X) \\
+  0 &\mbox{if} d > D_H(X)
+\end{array} \right.
+\]
+
+This value is the Hausdorff dimension of $X$.
+
+\subsection{Box counting dimension}
+
+The box counting dimension is one of the earlier general-purpose definitions of dimension, and one of the first to be used experimentally.
+Let $X \subseteq {\mathbb R}^d$.
+The we can cover the ${\mathbb R}^d$ with a grid of $d$-dimensional hypercubes with sides of length $\epsilon$.
+We then count how many of these hypercubes contain elements of $X$ and call this value $N_\epsilon$.
+We assume the following relation:
+\begin{align}
+\label{box-counting-assumption}
+N_\epsilon = \left (\frac{1}{\epsilon}\right) ^ d
+\end{align}
+
+If we take the logarithm on both sides, we get
+
+\begin{align*}
+ - d \ln \epsilon = \ln N_\epsilon
+ d = - \frac{\ln \epsilon}{\ln N_\epsilon}
+\end{align*}
+
+In other words a linear relationship in log-log axes. A common approach, especially in the early days of measuring
+
+The logic behind assumption \ref{box-counting-assumption} is detailed in @!. However, this relation is not guaranteed to hold exactly. A common tactic is to assume the relation 
+\begin{align}
+N_\epsilon = \Phi(\frac{1}{\epsilon})\left (\frac{1}{\epsilon}\right) ^ d
+\end{align}
+
+Where $\Phi\left(\epsilon\right)$ is called the \textit{pre-factor}. If we say that the pre-factor satisfies
+\[
+\lim_{\epsilon \rightarrow 0} \frac{\ln \Phi(\frac{1}{\epsilon})}{\ln \frac{1}{\epsilon}} = 0
+\]
+We can estimate the dimension using 
+\[
+D_B = \lim_{\epsilon \rightarrow 0} \frac{\ln N_\epsilon}{ \ln \epsilon}
+\]
+
+\subsection{Correlation dimension}
 
@@
 
-\section{Dimension}
+\subsection{Dimension of measures}
 
-Scaling laws.
+The dimension of a finite set of points, by any definition, is zero.
+Thus the dimension of a dataset will always be zero.
+When we talk about applying some estimator to a dataset we are not measuring the dimension of the dataset, but rather the dimension of the underlying probability distribution.
+We will usually assume that there is some probability measure over our instance space from which the elements in our dataset were independently drawn.
+It is the dimension of this measure that we are estimating.
+But before we do so, we must have some idea of how the dimension of a measure is defined.
 
-Box counting dimension, Correlation integral, Haussdorff & Packing dimension.
+Most commonly, this is done by extension of the box counting dimension.
+If we take the approach to the box counting dimension for sets and instead of counting the boxes which contain points of a set, we count the boxes with non-zero probability under our measure, we are in essence measuring the dimension of the support of the measure.
+However, the dimension of the support will not always capture the scaling we are interested in.
+Consider, for instance, a uniform distribution $U_0$ over the interval $[0, 1]$.
+We can cut this interval in two, and define a new distribution $U_1$ which combines a uniform distribution over the top half of the inteval with three quarters of the probability mass and a uniform distribution over the bottom half with the remaining probability mass.
+We repeat this process so that $U_{n+1}$ takes the uniform segments of $U_n$ and cuts them in half, distributing their probability mass over the new segements with the same ratio as before.
+The limiting distribution of this process ($\lim_{n \rightarrow \infty} U_n$) has the interval $[0, 1]$ as its support, and thus dimension $1$ by the definition above.
+The fractal shape, however, suggests that there are many interesting scaling laws that will give us an insight into its structure.
 
-Box counting estimator, Correlation estimator, Takens estimator.
+We can change the definition of dimension by introducing a parameter $q$ that determines how much more we care about regions with high probability than about regions with low probability.
+We define the quantity $I_q$ as an analgoue of the box counting dimension:
+\[
+I_q(p, \epsilon) = \sum_{c \in C_\epsilon}p(c)^q 
+\]
+Where $C$ is the covering of the probability distribution in boxes of side length $\epsilon$. The \textit{generalized dimension} of $X$ is then defined as
+\[
+D_q(p) = \frac{1}{q-1} \lim_{\epsilon \rightarrow 0} \frac{\ln I_q(X, \epsilon)}{\ln \epsilon}
+\]
 
-\section{The Takens estimator}
+For $D_0$ reduces to the box counting dimension of the support. $D_1$ causes a division by zero, but letting $q \rightarrow 1$ we can formulate 
+\[
+D_1(p) = lim_{\epsilon \rightarrow 0} \frac{1}{\ln \epsilon} \sum_{c \in C_\epsilon} p(c) \ln p(\epsilon)
+\] 
+The right part of this is the negative of the entropy of the distribution over $C$ under $p$. In essence $D_1$ measures how the entropy of a covering scales with the accuracy of that covering.
+
+For $q = 2$ we get 
+\[
+D_2(p) = \lim_{\epsilon \rightarrow 0} \frac{1}{\ln \epsilon}\ln \sum_{c \in C_\epsilon} {p(c)^2}
+\]
+This is called the correlation dimension, and this is the dimension we focus on.
+It has the advantage that it is easy to estimate.
+It should be noted that for many distributions $D_q$ does not vary with $q$, so that we can choose $q$ to suit our needs.
+Where the dimension does depend on $q$ (the so called \textit{multifractals}) $2$ provides reasonable compromise between caprturing the whole support, and capturing the scaling within the probability density function.
+
+The theoretical basis for the generalized dimension is usually based on an analogous treatment of the Hausdorff dimension.
+We will leave the subject at this brief introduction, noting only that we will technically be estimating $D_2$.
+
+\subsection{Non-independent data}
+
+As noted before, we assume that our data is independently sampled from some probability distribution. This may not be the case. In particular time series data\ldots
+
+Read Theiler's EEG paper. Discuss natural measures briefly. Also ergodicity?
+
+\section{Methods}
+
+\subsection{The Takens estimator}
 
@@
 
-\subsection{Physics}
 \subsection{Statistics and Machine Learning}
@@
 
-\subsection{Estimating the $d_Max$ parameter}
+\subsection{Estimating the $d_{\max}$ parameter}
 
@@
 
-Consider a datset produces by drawing random points in some 1D curve in 2D Euclidean space, but with a small term of 2D Gaussian noise added.
-The theoretical dimension of this process, for any definition of dimension will be 2, since the limit towards the smallest scales determines the dimension.
+Consider a dataset produced by drawing random points in some 1D curve in 2D Euclidean space, but with a small term of 2D Gaussian noise added.
+The theoretical dimension of this process, for any definition of dimension, will be 2 since the limit towards the smallest scales determines the dimension.
 However, the shape that we are interested in (the distribution without the noise term) has dimension 1.
@@
 \[
-p(d \leq r | d \in [r_{i-1}, r_i)) = \frac{p(d \leq r | d \in [r_{i-1}, r_i))}{p(d \d \in [r_{i-1}, r_i)])} = \frac{p(d \in [r_{i-1}, r])}{p(d \in [r_{i_1}, r_i))}
+p(d \leq r | d \in [r_{i-1}, r_i)) = \frac{p(d \leq r | d \in [r_{i-1}, r_i))}{p(d \in [r_{i-1}, r_i)])} = \frac{p(d \in [r_{i-1}, r])}{p(d \in [r_{i_1}, r_i))}
 \]
@@
 We are left with:
+
 \[
-\frac{1}{D^t_i} - \frac{\ln \frac{r_i}{r_{i-1}}}{(\frac{r_i}{r_{i-1}})^{D^t_i}} - \frac{\Sum_{r \in R_i}\ln r - |R_i| \ln r_{i-1}}{|R_i|} = 0
+\frac{1}{D^t_i} - \frac{\ln \frac{r_i}{r_{i-1}}}{(\frac{r_i}{r_{i-1}})^{D^t_i}} - \frac{\sum_{r \in R_i}\ln r - |R_i| \ln r_{i-1}}{|R_i|}= 0
 \]
@@
 
-This algorithm first searches for an $x$ that makes $f(x)$ negative, and then bisects theinterval $[0, x]$ recursively until the interval is smaller tha a given $\epsilon$, at which point the middle of the interval is taken as the estimate of $D_i^t$.
+This algorithm first searches for an $x$ that makes $f(x)$ negative, and then bisects the interval $[0, x]$ recursively until the interval is smaller tha a given $\epsilon$, at which point the middle of the interval is taken as the estimate of $D_i^t$.
 
@@
 @!NOTE: Try contiguous and non-contiguous scaling intervals.
+
+\section{Lacunarity}
 
@@
 Faces & & & & \\ 
+Sunspots & & & & \\
 
@@
 
-We start with the cumulative distribution function as described in the text. We use $r_{i-1} = r_j$ to clear up the notation.
+We start with the cumulative distribution function as described in the text. We use $r_{i-1} = r_h$ tohelp with the notation.
 
 \begin{align*}
-p(d \leq r | d \in [r_{i-1}, r_i)) = \frac{r^{D_i} - {r_{j}}^{D_i}}{{r_i}^{D_i} - {r_{j}}^{D_i}} = \frac{\left(\frac{r}{r_j}\right)^{D_i} - 1}{\left(\frac{r_i}{r_j}\right)^{D_i} - 1}
+p\left(d \leq r \mid r \in \left[r_h, r_i\right)\right) = \frac{r^{D_i} - {r_h}^{D_i}}{{r_i}^{D_i} - {r_{h}}^{D_i}} = \frac{\left(\frac{r}{r_h}\right)^{D_i} - 1}{\left(\frac{r_i}{r_h}\right)^{D_i} - 1}
 \end{align*}
@@
 \begin{align*}
-p(d = r | \ldots) = {D_i}r^{D_i-1} \frac{1}{\left(\frac{r_i}{r_j}\right)^{D_i} - 1}
+p(d = r \mid \ldots) = {D_i}r^{D_i-1} \frac{1}{\left(\frac{r_i}{r_j}\right)^{D_i} - 1}
 \end{align*}
@@
 
-Setting to zero and solving for $D_t$
+Setting to zero gives us.
 
@@
 \frac{|R_i|}{D_i} + \sum \ln r - |R_i| \frac{\left(\frac{r_i}{r_j}\right)^{D_i} \ln \frac{r_i}{r_j}}{\left(\frac{r_i}{r_j}\right)^{D_i} -1} &= 0 \\
-\frac{1}{D_i}  - \frac{\left(\frac{r_i}{r_j}\right)^{D_i} \ln \frac{r_i}{r_j}}{\left(\frac{r_i}{r_j}\right)^{D_i} -1} &= \frac{-\sum \ln r}{|R_i|}
+\frac{1}{D_i}  - \frac{\left(\frac{r_i}{r_j}\right)^{D_i} \ln \frac{r_i}{r_j}}{\left(\frac{r_i}{r_j}\right)^{D_i} -1} - \frac{-\sum \ln r}{|R_i|} &= 0  
 \end{align*}
 
-@! If we pre-normalize $R_i$ we may still get a closed form solution!
+Wehave little hope of solving this for $D_i$, but as shown in lemma \ref{monotonicity}, the left of this equation is strictly decreasing in the range over which we are interested.
+This allows us to find the value of $D_i$ by binary search as detailed in the main text.
 
-\subsection{Proof of monotonicity}
+In the single-scale case we could normalize the data by scaling the original data, so that the maximum distance would become one.
+In this case we might be tempted to scale the distances so that we get $r_j' = 0$ and $r_i' = 1$ which would allow us to use the single-scale MLE.
+Unfortunately this transformation, while linear in the space of the distances does not correspond to an affine transformation of the original data \footnote{Or even a bijective transformation, since it maps all point pairs at distance $r_h$ to points with distance $0$, ie.
+a single point.}, and does not preserve $D_i$.
+We must make due with the non-closed form expression of the MLE.
+
+\subsection{Monotonicity of the multi-scale MLE}
 
@@
 \begin{lma}
+\label{monotonicity}
 The function
 \[
-f(x) = \frac{1}{x} + \frac{\ln(a)}{a^x - 1} + c
+f(x) = \frac{1}{x} + \frac{a^x \ln(a)}{a^x - 1} + c
 \]
-is strictly decreasing for $ x \in [0, \infty)$ and $ a \in [0, 1]$ 
+is strictly decreasing for $ x \in (0, \infty)$ and $ a \in (0, 1)$.
 \end{lma}
+\begin{proof}
+Taking the derivative gives us.
+
+\[
+f'(x) = -\frac{1}{x^2} + \left [ \frac{\sqrt{a^x}\ln a}{a^x - 1} \right ]^2
+\]
+
+$f$ is stricly decreasing in the given domain if $f'(x) < 0$ holds. Assume for a contradiction that
+\begin{align*}
+-\frac{1}{x^2} + \left [ \frac{\sqrt{a^x}\ln a}{a^x - 1} \right ]^2 \geq 0
+\end{align*}
+Which rewrites to 
+\begin{align*}
+\frac{1}{x}  &\leq \frac{\sqrt{a^x}\ln a}{a^x - 1}
+x  &\geq \frac{a^x - 1}{\sqrt{a^x}\ln a}
+\end{align*}
+We know that $x > 0$ so that 
+\begin{align*}
+0  &> (a^x - 1)\frac{1}{\sqrt{a^x}\ln a}
+\end{align*}
+
+For the conditions stated $a^x < 1$ so that the left factor is negative. $\sqrt{a^x}$ is positive and $\ln a$ negative so that the right factor is negative as well, giving us a contradiction.
+\end{proof}
+
 \subsection{Algorithms}
 
+\bibliographystyle{siam}
+\bibliography{dimension}
+
 \end{document}

### 2012-08-29 Dimension/src/dimension.tex
...

@@
-\documentclass[10pt]{article}
+\documentclass[11pt]{article}
 
@@
 \begin{align}
-N_\epsilon = \Phi(\frac{1}{\epsilon})\left (\frac{1}{\epsilon}\right) ^ d
+N_\epsilon = \Phi\left(\frac{1}{\epsilon}\right)\left (\frac{1}{\epsilon}\right) ^ d
 \end{align}
@@
 \]
-Where $C$ is the covering of the probability distribution in boxes of side length $\epsilon$. The \textit{generalized dimension} of $X$ is then defined as
+Where $C$ is the covering of the probability distribution in boxes of side length $\epsilon$. The \textit{generalized dimension} of $p$ is then defined as
 \[
@@
 \] 
-The right part of this is the negative of the entropy of the distribution over $C$ under $p$. In essence $D_1$ measures how the entropy of a covering scales with the accuracy of that covering.
+The right part of this is the negative of the entropy of the distribution over $C$ under $p$.
+In essence $D_1$ measures how the entropy of a covering scales with the accuracy of that covering.
+This gives a more rigorous background to the intuitive notion of dimension as the amount of number need to describe a point.
+If we \ldots @!
+see Theiler.
 
@@
 
-As noted before, we assume that our data is independently sampled from some probability distribution. This may not be the case. In particular time series data\ldots
+As noted before, we assume that our data is independently sampled from some probability distribution. This may not be the case. In particular, time series data\ldots
 
-Read Theiler's EEG paper. Discuss natural measures briefly. Also ergodicity?
+[@! Read Theiler's EEG paper. Discuss natural measures briefly. Also ergodicity?]
 
@@
 
-Discussion of the running time an an investigation of the effects of reducing the data size.
+The great benefit of the Takens estimator (and other estimators based on the correlation integral) is that it works in distances between data points rather than data points themselves.
+This has the effect of squaring the volume of available data (barring some reduction of independence).
+
+The drawback is that using all available data can become very expensive, even for small datasets.
+In the basic algorithm taken from \cite{clauset2007power}, we must sum over all distances to find the estimate for the dimension, then repeat this process to find the estimate for the $d_max$ parameter and finally we must repeat the process around 1000 times to get a reasonably accurate estimate for the significance.
+
+In some contexts this problem can be overcome.
+If only a single dimension calculation is required after a long process of generating data, then the experimenter will likely not mind waiting a day or so for the results.
+Alternatively, one could exploit `embarrasingly parallel` nature of the algorithm and spread the search for a fitting $d_max$ or the iterations of the significance calculation over many cores.
+
+If many calculations of dimension are required, or the dataset is so large that even parallelization and patience are insufficient, we can sample the data.
+The simplest approach of course, is to take a sample of the data points and run the procedure on that sample.
+This has two drawbacks.
+First, of we sample from the set of distances instead we will increase the likelihood that our dataset represents a sample of independent distances.
+More importantly, we are using our sample to estimate two parameters $D_t$ and $d_{\max}$ some crude tests will show very quickly that if we want to estimate $D_t$ as accurately as possible, we should put all our computational power in the evaluation of the maximum likelihood estimator rather than considering many different values for $d_{\max}$ in other words, if we calculate the MLE with all available distances for a single randomly selected value of $d_{\max}$ we are generally better off than when we use to random values for $d_{\max}$ and half the distances for the MLE.
+
+Of course, even if the expected accurary is higher, using a single candidate for $d_{\max}$ is still likely to cause a very high variance. We use the following algorithm as a compromise.
+
+We first generate a set of candidates $C$ with sie $c$ for the value of $d_max$.
+These candidates are generated by taking a subsample $R'$ of $s$ of the distances and running the full estimator on this subsample.
+When then add the $d_max$ value of this estimator to list of candidates.
+Since we can make $s$ relatively big (ie.
+$5000$) without incurring a great computational cost, this allows us to generate a much better set of candidates then a simple random sample would give us, while not making any further assumptions about the data and our estimator than that more data improves accurary on average.
+\footnote{It may be of interest to note that if we set $s$ to one, we are essentially sampling $c$ candidates at random for $d_{\max}$ while using all data for the MLE as described before.
+Thus blind subsampling is a special case of this algorithm.}
+
+It is tempting to look toward search or optimization algorithms over the domain of $d_max$ to minimize the KS statistic, but these will often have rndom behaviours and assumptions on smoothness which will reduce the emprical strength of the resulting answer.
+Only the complete estimator will guarantee that the returned values will optimize their respective measures (maximum likelihood for $D_t$ and minimum KS statistic for $d_{\max}$).
+However if we must reduce the size of the observed data, we believe our approach strikes a nice balance between using computational resources efficiently and producing a trustworthy result.
+
+To consider the impact, we test how the accuracy of the estimator and the measure of its significance changes with a reduction in the number of candidates (keeping the number of samples per candidate fixed).
+The result is shown in figure @!
+(averging over 100 repeats of the experiment).
+We see that the $D_t$ estimator is the least affected.
+The $d_{\max}$ estimator suffers more, but as it is of secondary importance, this shouldn't bother us too much.
+The significance clearly becomes unreliable at low values of $c$.
+This is shown not only in the mean value but more importantly in the standard deviation.
+When using the full estimator, we can expect a good estimate of the significance with low variance, but even small subsampling begins to introduce massive vriance and error.
+For this reason we will report results on large datasets (using the method detailed here) but we will always take a subsample of the data to compute the significance.
+We will then make a further assumption that the significance for the subsample is a good estimate for the significance of the estimate on large data.
 
@@
 It's not uncommon to see data with more than two intrinsic dimensions.
-In the log-log plot produces by the Grassberger-Procaccia estimator, this behavior will produce linear segments for the relevant length scales.
+In the log-log plot produced by the Grassberger-Procaccia estimator, this behavior will produce linear segments for the relevant length scales.
 
@@
 To find the interval limits $r_i$ we proced as we did before with $d_{\max}$ and search for a set of values from the data which minimize the Kolmogorov-Smirnov statistic with the filtered data.
-We compare each interval with the data filtered to that interval and take the sum of all intervals for a given set as the value to minimize.
-As searching a for a single value was already expensive in the single-scale case, here we must give up on the idea of checking all possible values and instead search for a suitable set of ranges.
 
-@!NOTE: Try contiguous and non-contiguous scaling intervals.
+To find multiple, non-overlapping intervals, we use the follwing procedure.
+We first test all pairs of distances $r_a, r_b$ with $r_a < r_b$ and to find the pair with the lowest KS-statistic.
+Then, for the ranges to the left and right of $[r_a, r_b)$ we repeat the the procedure on only the data in that range.
+We repeat the procedure resursively until the data runs out or until the significance for the best subrange of any range drops below 0.01.
 
@@
 \begin{align*}
-\ln {\cal L}(R|D_t) &= -\sum \ln r + D_t \sum \ln r - D_t |R| \ln r_{\max} + |R| \ln D_t
+\ln {\cal L}(R|D_t) &= -\sum \ln r + D_t \sum \ln \frac{r}{r_{\max}} + |R| \ln D_t
 \end{align*}
@@
 \begin{align*}
-\frac{\partial \ln{\cal L}(R|D_t)}{\partial D_t} = \sum{\ln r} - |R|\ln r_{\max} + |R|\frac{1}{D_t}
+\frac{\partial \ln{\cal L}(R|D_t)}{\partial D_t} = \sum{\ln \frac{r}{r_{\max}}} + |R|\frac{1}{D_t}
 \end{align*}
@@
 \begin{align*}
- \sum{\ln r} - |R|\ln r_{\max} + |R|\frac{1}{D_t} &= 0 \\
-|R|\frac{1}{D_t} &= - \sum \ln r + |R| \ln r_{\max} \\
-D_t &= \frac{|R|}{- \sum \ln r + |R| \ln r_{\max}} 
+\sum{\ln \frac{r}{r_{\max}}} + |R|\frac{1}{D_t} &= 0 \\
+|R|\frac{1}{D_t} &= - \sum{\ln \frac{r}{r_{\max}}} \\
+D_t &= \frac{|R|}{- \sum{\ln \frac{r}{r_{\max}}}} 
 \end{align*}
-
-Many publications assume that the data $|R|$ are normalized by $r_{\max}$ which allows the cleaner result:
-\begin{align*}
-D_t &= \frac{|R|}{- \sum \ln r} 
-\end{align*}
-
-However, this hides the reliance on $r_{\max}$. As we will be iterating over all possible values for $r_{\max}$, it's easier to include in the estimator explicitly.
 
@@
 
-We start with the cumulative distribution function as described in the text. We use $r_{i-1} = r_h$ tohelp with the notation.
+We start with the cumulative distribution function as described in the text. We use $r_{i-1} = r_h$ to clear up the notation.
 
@@
 \begin{align*}
-p(d = r \mid \ldots) = {D_i}r^{D_i-1} \frac{1}{\left(\frac{r_i}{r_j}\right)^{D_i} - 1}
+p(d = r \mid \ldots) = {D_i} r^{D_i - 1} r_h^{-D_i} \frac{1}{\left(\frac{r_i}{r_h}\right)^{D_i} - 1}
 \end{align*}
@@
 \begin{align*}
-\ln{\cal L}(R_i|D_i) &= \sum \ln D_i + (D_i - 1)\sum \ln r - \sum \ln \left(\left(\frac{r_i}{r_j}\right)^{D_i} -1\right)\\
-&=  |R_i|\ln D_i + (D_i - 1)\sum \ln r - |R_i| \ln \left(\left(\frac{r_i}{r_j}\right)^{D_i} -1\right) 
+\ln{\cal L}(R_i|D_i) &= \sum \ln D_i +  (D_i-1) \sum \ln r - D_i \sum \ln r_h  - \sum \ln \left(\left(\frac{r_i}{r_h}\right)^{D_i} -1\right)\\
+&=  |R_i|\ln D_i + (D_i - 1)\sum \ln r - |R_i| D_i \ln r_h - |R_i| \ln \left(\left(\frac{r_i}{r_h}\right)^{D_i} -1\right) 
 \end{align*}
 
-Where $R_i$ is the subset of distances in the dataset that fall within the interval $[r_{i-1}, r_i)$.
+Where $R_i$ is the subset of distances in the dataset that fall within the interval $[r_h, r_i)$.
 
@@
 \begin{align*}
-\frac {\partial \ln{\cal L}(R_i|D_i)}{\partial D_i} &=  \frac{|R_i|}{D_i} + \sum \ln r - |R_i| \frac{\left(\frac{r_i}{r_j}\right)^{D_i} \ln \frac{r_i}{r_j}}{\left(\frac{r_i}{r_j}\right)^{D_i} -1}
+\frac {\partial \ln{\cal L}(R_i|D_i)}{\partial D_i} &=  \frac{|R_i|}{D_i} + \sum \ln r - |R_i|\ln r_h - |R_i| \frac{\left(\frac{r_i}{r_h}\right)^{D_i} \ln \frac{r_i}{r_h}}{\left(\frac{r_i}{r_h}\right)^{D_i} -1}
 \end{align*}
@@
 \begin{align*}
-\frac{|R_i|}{D_i} + \sum \ln r - |R_i| \frac{\left(\frac{r_i}{r_j}\right)^{D_i} \ln \frac{r_i}{r_j}}{\left(\frac{r_i}{r_j}\right)^{D_i} -1} &= 0 \\
-\frac{1}{D_i}  - \frac{\left(\frac{r_i}{r_j}\right)^{D_i} \ln \frac{r_i}{r_j}}{\left(\frac{r_i}{r_j}\right)^{D_i} -1} - \frac{-\sum \ln r}{|R_i|} &= 0  
+\frac{1}{D_i}  - \frac{\left(\frac{r_i}{r_h}\right)^{D_i} \ln \frac{r_i}{r_h}}{\left(\frac{r_i}{r_h}\right)^{D_i} -1} &= \frac{|R_i| \ln r_h - \sum \ln r}{|R_i|}  
 \end{align*}
 
-Wehave little hope of solving this for $D_i$, but as shown in lemma \ref{monotonicity}, the left of this equation is strictly decreasing in the range over which we are interested.
+We have little hope of solving this for $D_i$, but as shown in lemma \ref{monotonicity}, the function whose root we're looking for is strictly decreasing for the values we are interested in.
 This allows us to find the value of $D_i$ by binary search as detailed in the main text.
+
+For $r_h = 0$ this equation becomes undefined. However, this is just the single scale case, so we can use the basic Takens estimator derived in the previous section.
 
@@
 
+@! TODO Prove that it starts at a positive value
+
 The following lemma tells us that the function describing the maximum likelihood estimate for the multiscale case is strictly decreasing over positive values, so that we can search for its root with a simple binary search.
@@
 \[
-f(x) = \frac{1}{x} + \frac{a^x \ln(a)}{a^x - 1} + c
+f(x) = \frac{1}{x} - \frac{a^x \ln(a)}{a^x - 1} + c
 \]
-is strictly decreasing for $ x \in (0, \infty)$ and $ a \in (0, 1)$.
+is strictly decreasing for $ x > 0$ and $ a > 1$.
 \end{lma}
@@
 \begin{align*}
-\frac{1}{x}  &\leq \frac{\sqrt{a^x}\ln a}{a^x - 1}
-x  &\geq \frac{a^x - 1}{\sqrt{a^x}\ln a}
-\end{align*}
-We know that $x > 0$ so that 
-\begin{align*}
-0  &> (a^x - 1)\frac{1}{\sqrt{a^x}\ln a}
+\frac{\sqrt{a^x}\ln a}{a^x - 1}  &\geq \frac{1}{x}\\
+\frac{a^x - 1}{\sqrt{a^x}\ln a} &\leq x \\
+\frac{a^x}{x\sqrt{a^x}\ln a} - \frac{1}{x\sqrt{a^x}\ln a} &\leq 0 \\
+a^x &\leq 1
 \end{align*}
 
-For the conditions stated $a^x < 1$ so that the left factor is negative. $\sqrt{a^x}$ is positive and $\ln a$ negative so that the right factor is negative as well, giving us a contradiction.
+For $a>1$ and $x>0$ we know that $a^x > 1$ so we have a contradiction.
 \end{proof}
+
+\subsection{The Kolmogorov Smirnov test as an MDL approximation}
 

### 2012-08-30 Dimension/src/dimension.tex
...

@@
 
-As noted by Stanley, this is only a simplistic view and the study of the dimension of a dataset can also give vital clues to scaling behaviors and critical phenomena in the process from which the data originated.
-Thus the theorist can not only validate his work by the dimension, but also use it as a guide.
-When we investigate the notion of dimension more closely we must give up intuitive notion and the only thing we can say is that the gemotry of a set or space produces various power laws related to scaling and that these power laws will usually have the same exponent in common.
-There is a common intuition in the natural sciences that power law phenomena with the same exponent will likely share some similarity in their underlying components.
-Thus the dimension can be an important guide in finding the universal principles common to  complex phenomana.
+For this reason, any scientist who deals with fractal phenomena will want to measure dimension, in order to capture those phenomena in meaningful statistics.
+Examples are biologists studying the emergence of structure, geologists studying the fractal surfaces and outlines of land masses and rock formations or physicists studying the measures of strange attractors.
 
@@
 \[
-p(d = r | d \leq r_{\max}) = D_t \frac{r^{D_t-1}}{{r_{\max}}^D_t} = D_t \frac{1}{r} \left (\frac{r}{r_{\max}} \right ) ^ {D_t} 
+p(d = r | d \leq r_{\max}) = D_t \frac{r^{D_t-1}}{{r_{\max}}^{D_t}} = D_t \frac{1}{r} \left (\frac{r}{r_{\max}} \right ) ^ {D_t} 
 \]

### 2012-09-04 Dimension/src/dimension.tex
...

@@
 The \textit{intrinsic dimension} of a dataset does not refer to how many features are used to represent it, but---loosely speaking---to how few features could be used to represent it.
-It is also the exponent of various power laws that determine the way the internal structure of the dataset scales.
-For many datasets, in particular those with coarse ``fractal'' features or those drawn from a chaotic process, dimension is one of the few well defined statistics available, and usually the most expressive.
+This is expressed in various power laws that determine how attributes of the datasets scale over varying length scales.
+For many datasets, in particular those with coarse ``fractal'' features or those generated by deterministic chaos, dimension is one of the few well defined statistics available, and usually the most expressive.
 
@@
 For this reason, any scientist who deals with fractal phenomena will want to measure dimension, in order to capture those phenomena in meaningful statistics.
-Examples are biologists studying the emergence of structure, geologists studying the fractal surfaces and outlines of land masses and rock formations or physicists studying the measures of strange attractors.
+Scientists encoutenering this problem include biologists studying the emergence of structure, geologists studying the fractal surfaces and outlines of land masses and physicists studying the measures of strange attractors.
 
@@
 Most methods for measuring dimension have relied at least in some part on subjective judgements and processes that make it difficult to capture experimental error.
-A similar situation was present in the study of power law distributions until recently Clauset, Shalizi and Newman published \cite{clauset2007power}, in which they dealt with various problems, and described a simple objective approach for estimating the exponent of such distributions with no subjective elements and a clear indications of uncertainty and significance with respect of the assumptions.
+A similar situation was present in the study of power law distributions until recently Clauset, Shalizi and Newman published \cite{clauset2007power}, in which they dealt with various problems, and described a simple objective approach for estimating the exponent of such distributions with no subjective elements and a clear indications of uncertainty and significance with respect to the assumptions.
 
-The methods in this paper come from the realization that the relatively unknown Takens estimator for dimension is also based on the assumption of a power law distribution (albeit of a different form than the ones studied by Clauset et al.), allowing these methods can be adapted to the estimation of dimension.
+The methods in this paper come from the realization that the relatively unknown Takens estimator for dimension is also based on the assumption of a power law distribution (albeit of a different form than the ones studied by Clauset et al.), allowing these methods to be adapted to the estimation of dimension.
 This process also gives us a means to determine multiple dimensions for datasets that can be considered to have different dimensions for different ranges of scales, and we can give a new measure for the lacunarity, a measure of the `gapiness' or texture of a dataset that is often used in combination with dimension.
@@
 \section{Dimension}
+
+[@! Discuss the range of D for which estimation is reasonable]
 
@@
 
-\subsection{Hausdorff dimension}
+\subsection{A theoretical note}
 
-[@! is this necessary? Maybe move it to appendix and derive the scaling laws for box counting and correlation there]
+All notions and estimators of dimension described in this paper (and almost all such notions in general) require the data to follow a particular power law, usually between osme length scale and some property of the data.
+Often this power law can be proved to hold in the limit as the length scale goes to zero, or the power law is show to hold approximately (perhaps for a certain class of sets).
 
-For theoretical purposes, the Hausdorff dimension is the most popular definition of dimension.
-It's not much in measuring dimension, but it can help us understand where the various scaling laws come from that help us to measure notions of dimension that are more suited to estimation from data.
-Of course, most of the notions of dimension agree with each other for most sets, or have well-understood equalities.
+This approach to dimension has two main drawbacks.
+First, in the limit towards zero for length scales data becomes increasingly sparse .
+If we draw two points uniformly and measure their distance, it will likely be in the larger part of the range, with probability shrinking rapidly for lower distances.
+Secondly, we will often also be interested in datasets that seem to have one dimension at some range of scales and another dimension over another range of scales (imagine, for instance, points sampled from a curved line with some small, two-dimensional noise added).
+In the limit to zero we are likely only measureing the properties of the experimental noise.
+In the larger distance scales we can measure dimensions that actually reveal information about the processes that shaped our data.
 
-A complete treatment of Hausdorff dimension is beyond the scope of this article (@!
-Edgar provides a reasonably direct overview of the basic requirements for understanding Hausdorff dimension).
-We will give some intuition and the basic definitions.
+Our solution is not to reduce the meaning of our statistics down to a mathematical definition of dimension, but to take as an assumption that a certain scaling law holds for some range in our data.
+Data sets, by virtue of filling a geometrical space will generally follow various power laws but they can also deviate from them slightly (see section \ref{lacunarity}).
+If we simply assume that some scaling law holds exactly, we can take a maximum likelihood estimate of the dimension and use a significance measure to test to what extent our assumption is justified.
+We can also say that our curved line with added noise, while it technically has dimension 2, behaves as a clean curved line over certain length scales so that in that regime we can expect a 1 dimensional scaling law to hold.
 
-The basic notion behind Hausdorff dimension is that there are measures that are intrinsic to a given dimension (ie.
-length is related to dimension 1 and area to dimension 2) and that these measures will be $\infty$ or $0$ when they are applied to sets with dimension below or above the true dimension respectively.
-A fractal with dimension 1.5 will have length $\infty$ and area $0$.
-
-The Hausdorff dimension is defined over metric spaces. Euclidean spaces and sets in Euclidean space are specific instances of metric spaces. We will first introduce some basic concepts.
-
-\begin{dfn}
-A covering $E$ of a metric space $S$ is a sets of sets such that $S \subseteq\bigcup_e \in E e$. 
-\end{dfn}
-
-\begin{dfn}
-The diameter of a set $X$ in metric space $S$ with metric $\sigma$ is
-\[
-\mbox{diam}(X) = \sup_{x, y \in X}\sigma(x, y)
-\]
-\end{dfn}
-
-\begin{dfn}
-The quantity $H^d(x, r)$ of a metric space $X$ is the covering of $X$ that minimizes the sum of the diameters raised to $d$:
-\[
-H^d(X, r) = \inf_E \sum_e \in E \mbox{diam}(e)^d
-\]  
-\end{dfn}
-
-Where the infimum is taken over all coverings $E$.
-
-If we let $r$ got to zero, we get the measure mentioned above
-
-\begin{dfn}
-\[
-H^d(X) = \lim_{r \rightarrow 0} H^d(X, r)
-\] 
-\end{dfn}
-
-This measure has an intrinsic dimensionality $d$.
-If it is used on sets with the wrong dimension, it produces $0$ or $\infty$.
-Specifically, it can be shown that for each argument $X$ there is a single unique value $D_H$ such that
-\[
-H^d(X) = \left \{ 
-\begin{array}{rl}
-  \infty &\mbox{if} d < D_H(X) \\
-  0 &\mbox{if} d > D_H(X)
-\end{array} \right.
-\]
-
-This value is the Hausdorff dimension of $X$.
+The distinction is a subtle one, and in most cases people will be satisfied to claim simply that they are measuring dimension, but in some situation the distinction may be important.
 
@@
 
-The box counting dimension is one of the earlier general-purpose definitions of dimension, and one of the first to be used experimentally.
+The box counting dimension is one of the earlier general-purpose definitions of dimension, and one of the first to be used to define an estimator.
 Let $X \subseteq {\mathbb R}^d$.
-The we can cover the ${\mathbb R}^d$ with a grid of $d$-dimensional hypercubes with sides of length $\epsilon$.
+We can cover the ${\mathbb R}^d$ with a grid of $d$-dimensional hypercubes with sides of length $\epsilon$.
 We then count how many of these hypercubes contain elements of $X$ and call this value $N_\epsilon$.
@@
 
-In other words a linear relationship in log-log axes. A common approach, especially in the early days of measuring
+In other words a linear relationship in log-log axes.
+A common approach, especially in the early days of measuring dimension was to use a log-log plot of many values of $N_\epsilon$ and $\epsilon$ and fit a line over a range of $\epsilon$ where the data seems reasonably linear.
+A more refined approach is to find a range of scales in which the log-log plot looks linear and to perform linear regression on the logarithmes of these values.
+apart from the subjective element in this approach, linear regression if logarithmic axes has undesirable statistica properties as detailed in \cite{clauset}.
+[@!
+Review]
 
-The logic behind assumption \ref{box-counting-assumption} is detailed in @!. However, this relation is not guaranteed to hold exactly. A common tactic is to assume the relation 
+There is no guarantee that this relation holds. A common tactic is to reduce the assumed power law to  
 \begin{align}
@@
 
-Where $\Phi\left(\epsilon\right)$ is called the \textit{pre-factor}. If we say that the pre-factor satisfies
+where $\Phi\left(\epsilon\right)$ is called the \textit{pre-factor}. If we say that the pre-factor satisfies
 \[
@@
 \]
-
-\subsection{Correlation dimension}
-
-Definitions of dimension, fractal dimension and measuring dimension.
-
-Previous work.
-Theiler.
-Manifold learning.
-Databases (Cite faloutsos)We will not discuss the mathematical backgrounds of dimension (Cite Edgar).
-We will not focus greatly on the algorithmic complexity.
+This allows many deviations from the power law (such as oscilation) but requires the dimension to be measured in the lower ranges of $\epsilon$ (note that we will need to increse the volume of data for decreasing values of $\epsilon$ if we want to keep getting meaningful measurements).
 
@@
 \]
-This is called the correlation dimension, and this is the dimension we focus on.
+This is called the correlation dimension (\ref{correlation dimension}), and this is the dimension we focus on.
 It has the advantage that it is easy to estimate.
@@
 
+\subsection{Correlation dimension}
+\label{correlation-dimension}
+In 1983 Grassberger and Procaccia \cite{grassberger1983measuring} introduced the Correlation dimension.
+The Correlation dimension is based on the probability that two points, drawn independently wil have a distance less than $r$.
+We call this probability measure over distances the correlation integral $C(r)$ and we assume the following scaling law:
+\begin{align*}
+C(r) = \Phi(r) r^D
+\end{align*}
+Where $Phi(r)$ is again the pre-factor.
+
+The correlation estimator $\widehat{C}(r, X)$ for the correlation integral (over dataset $X$) is straightforward\footnote{$[\ldots]$ are Iverson brackets.
+They are $1$ when the logical formula within them is true, and $0$ otherwise.}:
+\begin{align*}
+\widehat{C}(r, X) = \frac{\sum_{x\neq y \in X} [d(x, y) < r]}{\sum_{x \neq y \in X} 1}
+\end{align*}
+
+@! Mention the statistical properties of the correlation estimator.
+
+We can substitute $\widehat{C}$ for $C$ and follow the same approach as with the box counting estimator. We plot the $\log C$ against $\log r$, and find the slope of the in an appropriate regime.
+
+The great value of the correlation integral is that it turns a dataset of $n$ points into a dataset of $(n^2 - n)/2$ points.
+While the set of distances is derived (and thus technically contains no new information), we can treat it as a a set of independent data at very low cost to accuracy (see \ref{independence}).
+The effect of the correlation integral then, is to give us a polynomial increase in data at almost no cost.
+
+Unfortunately, the Grassberger Procaccia estimtor still suffers form the drawbacks of the box counting estimator.
+A requirement of a subjective judgement and a reliance on a linear fit in logarithmic axes.
+The Takens estimator described in section \ref{methods}, does not suffer these problems, so that it what we will base our method on.
+
 \subsection{Non-independent data}
 
-As noted before, we assume that our data is independently sampled from some probability distribution. This may not be the case. In particular, time series data\ldots
+As noted before, we assume that our data is independently sampled from some probability distribution.
+This may not be the case.
+For instance time series data can be highly autocorrelated.
+In some cases, for instance when data is generated by a strange attractor, the whole time series may be predetermined by the initial state of the system, so that we cannot speak of a real probability measure, let alone dependency on one.
 
-[@! Read Theiler's EEG paper. Discuss natural measures briefly. Also ergodicity?]
+The solution to the last problem is the natural measure. Let $\{x_i\} \in S$ be an infinite timeseries generated by some process. Then we define the natural measure of that process as
+\begin{align*}
+p(S') = \lim_{n \rightarrow \infty} \frac{\sum_{x_i, i < n} [xi \in S']}{n}
+\end{align*} 
+where $S' \subseteq S$ a similar measure can be defined for continuous $i$.
+In the case of ergodic strange attractors a slight uncertainty about the initial condition of the system will quickly spread out to coincide with the systems natural measure, so that for such systems the natural measure actually functions as a kind of probability measure representing uncertainty of the state of the system after a number of iterations.
+
+Even when we take the natural measure as the measure for which we are measuring the dimension, we have no guarantee that time series data will be independent.
+In fac, if it were, our time series wouldn't be very interesting.
+A solution proposed in \cite{theiler1990estimating} is to use for the numerator of the correlation estimator all distances below $r$ except those for which the points are within $w$ time units.
+The division is still by the number of all point pairs.
+
+A good example of the dangers of autocorrelation in data is provided by \cite{theiler1996re}. For the rest of this paper we will assume independently drawn data.
+
+We should also caution against bootstrapping.
+In instances drawn form a distribution with infinite support, the probability that the same instance will occur twice in a finite data set is zero so that the distance 0 should never occur.
+If we boostrap a dataset by sampling with replacement, we will very likely have recurring elements, and thus a number of zero distances.
+This can lead to a dimension estimate of 0.0 (as would be correct for independent draws with recurring elements, since it suggests a finite support).
+For this reason if we want to apply the Takens estimator to a bootstrapped dataset, we recommend bootstrapping over the collection of distances rather than the collection of points.
 
 \section{Methods}
+\label{methods}
 
@@
 
-Like the Hill estimator. State assumptions for MLE.
+The Takens estimator is based on two assumptions about a dataset $X$ drawn from probability distribution $p$:
+\begin{itemize}
+  \item The set of distances $R = \{r : x_i, x_j \in X, i>j, d(x_i, x_j)\}$ can be taken as a set of $(n^2 - n)/2$ independent draws from the probability distribution of te random variable $D = d(X, Y)$ where $X$ and $Y$ are independent random variables distributed according to $p$.
+  \item The correlation integral follows a power law with a constant pre-factor $C(r) = c r^D_t$ for values of $r$ below some value $r_{\max}$.
+\end{itemize}
+
+Under these two assumptions we can derive the following maximum likelihood estimator for $D_t$:
+\begin{align*}
+\widehat{D_t} &= \frac{|R|}{\sum{R} - |R| \ln r_{\max}}
+\end{align*}
+
+The derivation is presented on section \ref{mle-derivation}.
+
+This estimator was first described by Floris Takens in 1985 \cite{takens1985numerical} in relation to measuring dimension, although the estimator for power law distributions (and generic distributions with tail) was already treated in depth by Hill in 1975 \cite{hill1975simple}.
+
+The following assumption discusses the independence asumption and test its influences.
+The second assumption we take as a hypothesis to be validated by a calculation of the significance (see \ref{significance}).
 
 \subsection{Testing the independence assumption}
-
+\label{independence}
 If we use each point for only one pair, (and our dataset is iid) then the distances for those pairs will be iid samples from $p_D$.
@@
 If we use only $d(a, b)$ we get a single sample from $p_D$.
-If we use $d(a, b)$ and $d(a, c)$, the two values are dependent the location of $a$.
-we are in fact sampling twice independently from the distribution of $d(a, X)$ rather than $d(X, Y)$.
+If we use $d(a, b)$ and $d(a, c)$, the two values are dependent on the location of $a$.
+We are in fact sampling twice independently from the distribution of $d(a, X)$ rather than $d(X, Y)$.
 
@@
 Thus we expect that for a large dataset, the distances of all pairs do not form a substantially different distribution than the distances over successive pairs, while giving us a squared increase in the volume of data.
+In \cite{}
 
-We test this assumption by taking datasets in full (after shuffling the data) and performing the Kolmogorov-Smirnov statistic on them. The results are shown in table @!.
+We test this assumption by generating both the set of all distances and the set of distances between successive points (for a shuffled dataset) and calculating the Kolmogorov-Smirnov statistic $k$ for the two datasets.
+To give us a baseline for comparing the values, we also generate (for the datasets based on generating algorothms) two sets of $n/2$ and $(n^2 - n)/2$ values drawn from the true distribution over distances, and report their KS value $k$.
+For generated datasets experiments were repeated 100 times and the sample standard deviation is reported in parenthesis after the mean.
 
-\subsection{Testing the distribution assumption}
+The results show that the difference between $k$ and $\widehat{k}$
 
-\subsection{Statistics and Machine Learning}
+\begin{tabular}{l | r r r r r r}
+\hline
+data set & size & features & $\widehat{k}$  & & $k$&  \\
+\hline
+Swiss roll & 400 & 3 & 0.065 &(0.02) & 0.059&(0.019) \\
+Noisy swiss roll & 400 & 3 & 0.068&(0.022)&  0.062&(0.02)\\ 
+Sierpinski & 400 & 2 & 0.065&(0.23)& 0.063&(0.19)\\
+Cantor & 400 & 1 & 0.06&(0.18) & 0.58&(0.02)\\
+Menger Sponge ! & 400 & 3 & & & & \\
+Persistent process ! & 400 & 2 & & & & \\
 
-\section{Estimating dimension}
+MVN 5 & 400 & 5 & & & & \\
 
-\subsection{Maximum Likelihood Estimator}
+Logistic equation & 400 & 1 & & & & \\
+Lorenz attractor & 400 & 3 & & & & \\
 
-Derivation of Takens' MLE and its standard error. Comparison to the Hill estimator and discussion of the differences.
+British coast line & 400 & 2 & & & & \\   
+Expressions & & & & & & \\
+Currency ! & & & & & & \\
+Galaxies & & & & & & \\
+DNA random walk & & & & & & \\
+Turning cup & 481 & 245760 & 0.067 & & & \\ 
+Faces & & & & & & \\ 
+Sunspots & & & & & & \\
+
+Newsgroup posts & & & & & & \\
+Protein sequences  & & & & & & \\
+Graphs & & & & & & \\
+\hline
+\end{tabular}
+
+\subsection{Bias and variance of the MLE}
+
+Refer to Smith.
+
+In \cite{smith1992estimating}, Smith derives the following formula for the number of points $N$ (below $d_{\max}$) required to measure the intrinsic dimension to an (mean squared) error of $\nu$, if the intrinsic dimension is $D$:
+\begin{align*}
+N = D\left[\frac{4D(D+4)}{(D+2)^2}\right]^{\frac{D}{8}} \left[\frac{D+4}{2}\right]^\frac{1}{2} g^\frac{D}{4} \nu^\frac{-(D+4)}{4}
+\end{align*}
+
+Where $g$ is a constant depending on the shape of the distribution.
 
@@
 
-Quick review of the methods described in Clauset. Translation of the method to this estimator. Experiment showing the validity of the method for this MLE.
+Quick review of the methods described in Clauset. Translation of the method to this estimator.
 
-\subsection{Uncertainty}
+\subsection{Significance and Uncertainty}
+\label{significance}
 
-Quick review of the approach of Clauset.
+Following \cite{clauset2007power}, we provide uncertainty values by bootstrapping over the original data: we sample with replacement a dataset the same size as the original and repeat the experiment.
+We repeat this procedure 100 times and report the sample standard deviation of the outcome.
 
-\subsection{Significance}
+Since we assume the scaling law to be present in the data and make no claims that for its existence a priori, we must consider the case that we are fitting a distribution to data that does not support it.
+Again we turn to \cite{clauset2007power} for our approach.
 
-Review of the approach of Clauset, and a description of the method translated to this problem.
+Having fit a power law to the regime $[0, r_{\max}]$ we can generate a dataset that is like our original given that our assumption is true.
+For the $n$ points in the data w generate from the fitted distribution with probability $\delta$ and sample from the points above $r_{\max}$ with probability $1 - \delta$, where $\delta$ is the proportion of points in the original data below $r_{\max}$.
+This way we generate a new dataset for the head of the distribution parametrically and the taiol is bootstrapped so that we do not have to make a assumptions about its shape.
 
-\subsection{Metric data}
+We then assume that our data has a bad fit to the distribution and ask what the probability is that such a bad fit would be seen if the assumptions about distribution were true.
+We generate $n$ datasets by the methods described above, repeat the experiment on them and calculate the KS statistic (note that we calculate the KS statistic of the generated ddata against the distribution fitted on the generated data).
+The proportion $p$ of trials for which the KS value is above the KS value measured for our original dataset is the probability that such a bad fit would emerge with correct assumptions.
+Thus, the closer $p$ is to one, the higher our significance (contrary to more traditional significance measures where low values of $p$ indicate).
 
-Discussion of the use of dimension for metric data and the use of the significance measure.
+If $p$ is low (for instance below $0.01$) then the scaling law does not hold for the regime used and we cannot trust the result of the maximum likelihood estimator.
+In the case of power law distributions the most likely reason for a low $p$ value is that the data fit a distribution that is similar to a power law distribution (which can then be tested by log-likelihood ratio tests).
+In our case the most likely reason for a low $p$ is hat the power law has a strong non-constant pre-factor which is throwing off the estimator.
+The cantor set, for instance shows a basic scaling law in log-log axes, but with an oscilating component.
+In other words the pre-factor is not constant and does not converge to a constant with zero.
+In these cases the the Takens estimator will fail to converge  as $r_{\max}$ gos to zero (although it oscillates within 0.05 of the true value).
+
+This deviation from the scaling law is called \textit{lacunarity}, and because of the structural properties that sets and measures must posess to deviate from scaling laws, lacunarity is an interesting statistic in itself.
+We discuss this interpretation further in section \ref{lacunarity}.
 
@@
 However if we must reduce the size of the observed data, we believe our approach strikes a nice balance between using computational resources efficiently and producing a trustworthy result.
+
+For very large datasets, we may run in to the problem that it is not economical to store the set of all distances.
+We can store the dataset and compute distances on the fly for all elements of the algorithm, except for the KS test, where we require a sorted list of distances.
+If we are satisfied with a subsample of the set of distances for each run of the KS test, we can then combine this with the method described above to get an approximation to the full estimator with three parameters: the number of candidates, the number of samples per candidate, and the size of the subsample of the data to use for each candidate.
+
+@! Test all three methods
 
@@
 
-\section{Lacunarity}
-
+Results are reported in the next section. 
 \section{Experiments}
@@
 \hline
-Swiss roll & 1000 & & & \\
+Swiss roll & 500 & & & \\
 Noisy swiss roll ! & 1000 & & & \\ 
@@
 
+Logistic equation & & & & \\
+Lorenz attractor & & & & \\
+
 British coast line & & & & \\   
-
-EEG & & & & \\
 JAFFE expression database & & & & \\
@@
 
+\section{Lacunarity}
+\label{lacunarity}
+
 \section{Conclusion}
@@
 \subsection{Derivation of the MLE}
+\label{mle-derivation}
 
@@
 
+\subsection{Hausdorff dimension}
+\label{hausdorff-dimension}
+
+For theoretical purposes, the Hausdorff dimension is the most popular definition of dimension.
+It's not much in measuring dimension, but it can help us understand where the various scaling laws come from that help us to measure notions of dimension that are more suited to estimation from data.
+Of course, most of the notions of dimension agree with each other for most sets, or have well-understood equalities.
+
+A complete treatment of Hausdorff dimension is beyond the scope of this article (@!
+Edgar provides a reasonably direct overview of the basic requirements for understanding Hausdorff dimension).
+We will give some intuition and the basic definitions.
+
+The basic notion behind Hausdorff dimension is that there are measures that are intrinsic to a given dimension (ie.
+length is related to dimension 1 and area to dimension 2) and that these measures will be $\infty$ or $0$ when they are applied to sets with dimension below or above the true dimension respectively.
+A fractal with dimension 1.5 will have length $\infty$ and area $0$.
+
+The Hausdorff dimension is defined over metric spaces. Euclidean spaces and sets in Euclidean space are specific instances of metric spaces. We will first introduce some basic concepts.
+
+\begin{dfn}
+A covering $E$ of a metric space $S$ is a sets of sets such that $S \subseteq\bigcup_e \in E e$. 
+\end{dfn}
+
+\begin{dfn}
+The diameter of a set $X$ in metric space $S$ with metric $\sigma$ is
+\[
+\mbox{diam}(X) = \sup_{x, y \in X}\sigma(x, y)
+\]
+\end{dfn}
+
+\begin{dfn}
+The quantity $H^d(x, r)$ of a metric space $X$ is the covering of $X$ that minimizes the sum of the diameters raised to $d$:
+\[
+H^d(X, r) = \inf_E \sum_e \in E \mbox{diam}(e)^d
+\]  
+\end{dfn}
+
+Where the infimum is taken over all coverings $E$.
+
+If we let $r$ got to zero, we get the measure mentioned above
+
+\begin{dfn}
+\[
+H^d(X) = \lim_{r \rightarrow 0} H^d(X, r)
+\] 
+\end{dfn}
+
+This measure has an intrinsic dimensionality $d$.
+If it is used on sets with the wrong dimension, it produces $0$ or $\infty$.
+Specifically, it can be shown that for each argument $X$ there is a single unique value $D_H$ such that
+\[
+H^d(X) = \left \{ 
+\begin{array}{rl}
+  \infty &\mbox{if} d < D_H(X) \\
+  0 &\mbox{if} d > D_H(X)
+\end{array} \right.
+\]
+
+This value is the Hausdorff dimension of $X$.
+
 \subsection{The Kolmogorov Smirnov test as an MDL approximation}
 
+\subsection{Generating random points}
 \subsection{Algorithms}

### 2012-10-01 fractal paper/src/fractal.tex
...

@@
-\documentclass[11pt, twocolumn]{article}
+\documentclass[11pt]{article}
 
@@
 \begin{abstract}
-\noindent Fractals are mathematical objects that display great detail and complexity from a simple description.
-It has often been suggested that fractals mimic the structures of nature and everyday life better than the shapes of classical geometry \cite{}.
+\noindent Fractals are mathematical objects that display great detail and complexity despite having simple descriptions.
+It has often been suggested that fractals mimic the structures of nature and everyday life better than the shapes of classical geometry @!
+\cite{}.
 These properties make fractals interesting candidates for modeling data.
@@
 Fractals are mathematical objects (generally sets or measures) with certain remarkable properties.
-What constitutes a fractal is not precisely defined\footnote{Mandelbrot, who coined the word, originally defined it as a set whose topological dimension differs from its Hausdorff dimension.
-He later retracted this definition, stated that he preferred the word to be not precisely defined.}, but the following properties are common:
+What constitutes a fractal is not precisely defined\footnote{Mandelbrot originally defined it as a set whose topological dimension differs from its Hausdorff dimension.
+He later retracted this definition, stating that he preferred the word to be not precisely defined.}, but the following properties are common:
 
@@
   \item Infinitely small detail: 'Zooming in' reveals ever finer detail.
-In the case of the Sierpinski triangle we will simply see the same shape recurring again and again, but fractals like the Mandelbrot set reveal a great variety of detail.
-  \item Non-integer dimension: for instance the Sierpinski triangle has dimension about 1.58. This basic and counter-intuitive fact is not easily explained. It will be discussed in section @!.
+In the case of the Sierpinski triangle we will simply see the same shape recurring again and again, but fractals like the Mandelbrot set reveal a great variety of shapes.
+  \item Non-integer dimension: for instance the Sierpinski triangle has dimension about 1.58.
+This is a counter intuitive notion, that is not simple to explain.
+We refer the reader to @!
+for an intuitive explanation.
 \end{itemize}
@@
 The first connection between fractals and the natural world was in the form of tangible objetcs such as coast lines, trees or bronchia, but the more hidden aspects of reality were soon shown to have fractal properties as well.
-The path traced out by the Brownian motion of a particle, the price chart for a commodity or the level of floods in a river delta all benefited from a fractal perspective.
+The path traced out by the Brownian motion of a particle, the price chart for a commodity or the level of floods in a river delta have all been studied as fractals to great effect.
 
 One of the problems of fractal analysis has always been the difficulty of finding a fractal model for a given set of data.
-The assumption that data contains fractal properties by itself already allows us to calculate certain statistics (such as its dimension), but the various fractal models that have been proposed are difficult to fit to data.
+The assumption that data contains fractal properties by itself already allows us to calculate certain statistics (such as its dimension), but the various families of fractals that have been proposed are difficult to fit to data.
 This is called the \emph{inverse problem}.
@@
 Call our dataset $X$, our instance $x \in X$ and $x \in {\mathbb R}^d$.
-We will assume that our dataset represents iid draws from a probability distribution over ${\mathbb R}^d$ which we will try to model.
+We will assume that our dataset represents iid draws from a probability distribution over ${\mathbb R}^d$ which we will attempt to model.
 
@@
 
-Iterated Function Systems are a simple generalization of many of the simpler fractals.
+Iterated Function Systems are a simple generalization of many classic fractals.
 Consider the Sierpinski Gasket (fig @!).
@@
 This is not true for two independently drawn points.
-For our experiments we generate a sequence and then shuffle the data to remove this dependency where it is relevant.
+For our experiments we generate a sequence and then shuffle the data to remove this dependency where it is undesirable.
 
@@
 
-We will require the inverse process of this.
-Given a point on or near the support of some IFS, we wish to find $\sigma$.
+We will require the inverse process: given a point on or near the support of some IFS, we wish to find $\sigma$.
 To this end we start with the standard normal multivariate distribution and apply all maps $S_\sigma$ for all $\sigma$ of a given lentgh $k$.
@@
 
-We should provide some caution about the use of the codes, as the ordering is often reversed from what our intuition suggests.
+A word of caution on codes is appropriate: the ordering is often reversed from what our intuition suggests.
 If we number the three components of the Sierpinski triangle clockwise 0, 1 and 2, we might number a small section of the triangle by first moving into a one of the large subsections 2, then into one of the subsections of subsection 1 and then into subsection 0 of that.
@@
 
-We also note that when we use a simple IFS as a model for the codes---such as the two component IFS that maps the interval $[0,1]$ onto its lower and upper half, or the four component IFS that maps the unit square onto its four quarters---its codes coincide with commonly used spatial indexing functions and string based construction for the continuum.
+We also note that when we use a simple IFS as a model for the codes---such as the two component IFS that maps the interval $[0,1]$ onto its lower and upper half, or the four component IFS that maps the unit square onto its four quarters---its codes coincide with commonly used spatial indexing functions and string-based constructions for the continuum.
 In this sense we can see the iterated function system as covering a subspace of ${\mathbb R}^d$ and the codes as describing a kind of metric that is intrinsic to the data rather than the embedding space.
@@
 The main cause of these problems is that as dimensionality increases, the volume of a subspace (such as the unit ball) increases exponentially with it.
-This means, for instance, that vast amounts of samples are required to maintain the same level of coverage in a high dimensional space as in a low dimensional one.
+This means, for instance, that vast amounts of samples are required to maintain the same level of coverage in a high dimensional space as compared to a low dimensional one.
 
-The solution to many of these problems is that while the data has a high-dimensional representation, it has a lopw intrinsic dimension.
-That is, it can be mapped to a low dimensional space without any oss of information.
+The solution to many of these problems is that while the data has a high-dimensional representation, it has a low intrinsic dimension.
+That is, it can be mapped to a low dimensional space without any loss of information.
 
@@
 
-Whe measure the dimension of datasets and data generated from models using Takens' estimator:
+We measure the dimension of datasets and data generated from models using Takens' estimator:
 \[
@@
 This method is based on the work for general power laws in @!
-and is detailed in @!.
+and is detailed in @!
+(to be published).
 
@@
 If we have a collection of points $X_\sigma$ with code $\sigma$ and a collection of points $X_\tau$ with code $\tau$ such that $S_i$ should satisfy $S_i(x_\sigma) = x_\tau$, we know that if we generate the IFS model, $w_i$ of the probability mass at stage $t$ always gets distributed to component $S_i$ for stage $t + 1$.
-thus we can expect the relative size of these sets to follow $w_i$: $|X_\tau| = w_i |X_\sigma|$.
+Thus we can expect the relative size of these sets to follow $w_i$: $|X_\tau| = w_i |X_\sigma|$.
 Over all such sets $X_\sigma^j$ and $X_\tau^j$ we must minimize the squared error:
 \[
-\overline{w_i} = \argmin_{w_i} \sum_{j} \left( |X_\sigma^j| - w_i |X_\tau^j| \right)^2
+\overline{w_i} = \argmin_{w_i} \sum_{j} \left( |X_\tau^j| - w_i |X_\sigma^j| \right)^2
 \]
@@
 \[
-\overline{w_i} = \frac{\sum_{j}|X_\sigma^j||X_\tau^j|}{2\sum_{j}|X_\tau^j|^2}
+\overline{w_i} = \frac{\sum_{j}|X_\sigma^j||X_\tau^j|}{\sum_{j}|X_\sigma^j|^2}
 \]
@@
 
-Finally we give some indication of the time preformance of the algorithm.
+Finally we give some indication of the time performance of the algorithm.
 
-\subsection{Dimension scaling}
+\subsection{Datasets}
+
+\begin{description}
+  \item[forest]
+  \item[census]
+  \item[abalone]
+\end{description}
+
+\subsection{Model fitting and dimension scaling}
+
+First, we fit IFS models to our datasets (ignoring the class information).
+We report the Haussdorff distance of a sample of 1000 points from the data to 1000 points generated by the model (we consider the iris dataset to be too small for this purpose).
+For this test we test on a random 25\% of the data and repeat the experiment 20 and report the sample standard deviation and the mean.
+
+We are also interested in the dimensional scaling of the model.
+In this case we are interested in the training behavior of the model, so we do not withhold test data.
+We measure the dimension of the dataset on a random sample of 300 points (repeating 20 times to estimate the sample standard deviation) and generate 300 points from the best model and measure the dimension in the same way, taking the standard deviation over 20 separately trained models, so that it represents the combined uncertainty over training the model and estimating dimension.
+
+In all cases we train the model for 300 generations and select the best model by Hausdorff distance.
+
+\begin{table}
+\begin{tabular}{l | r r r r | r |}
+\hline
+ & ifs1 & ifs3 & ifs5 & ifs7 & mog \\
+\hline
+census & 0.02 (0.04, 80) & & & & \\
+forest & & & & & \\
+abalone & & & & & \\
+\hline
+
+\end{tabular}
+\caption{Hausdorff distances for models trained against various datasets.}
+\label{tablelabel}
+\end{table}
+
+\begin{table}
+\begin{tabular}{l | r r r r r | r |}
+\hline
+  & dim & ifs1 & ifs3 & ifs5 & ifs7 & mog \\
+\hline
+census & 3.2 (0.2) & 2.9 (0.45) & & & & \\
+forest &  & & & & & \\
+abalone & &  & & & & \\
+\hline
+
+\end{tabular}
+\caption{Hausdorff distances for models trained against various datasets.}
+\label{tablelabel}
+\end{table}
 

### 2012-10-25 fractal paper/src/fractal.tex
...

@@
 
-\title{Fractal Modeling}
+\title{Fractal Modeling for Machine Learning}
 \date{\today}
@@
 \begin{abstract}
-\noindent Fractals are mathematical objects that display great detail and complexity despite having simple descriptions.
-It has often been suggested that fractals mimic the structures of nature and everyday life better than the shapes of classical geometry @!
-\cite{}.
-These properties make fractals interesting candidates for modeling data.
-We present an overview of existing concepts that are relevant to this task, and a new algorithm for fitting fractal models to data.
+\noindent We present a new algorithm for fitting fractals to data.
+We use this algorithm to investigate the use of fractal models in a machine learning context.
+Fractals are mathematical objects that display great detail and complexity despite having simple descriptions.
+It has often been suggested that fractals model many natural and other real-world structures better than the shapes of classical geometry \cite{mandelbrot1982fractal}.
 \end{abstract}
 
-\section{Preliminaries}
-\subsection{Fractals}
+\section{Introduction}
 
 Fractals are mathematical objects (generally sets or measures) with certain remarkable properties.
-What constitutes a fractal is not precisely defined\footnote{Mandelbrot originally defined it as a set whose topological dimension differs from its Hausdorff dimension.
-He later retracted this definition, stating that he preferred the word to be not precisely defined.}, but the following properties are common:
+What constitutes a fractal is not precisely defined\footnote{Mandelbrot originally defined it as a set whose topological dimension differs from its Hausdorff dimension, but retracted this definition, stating that he preferred the word to be not precisely defined.}, but the following properties are common:
 
@@
 It was not until the 1970s that they began to be seen as a single and useful family of mathematical objects that could accurately describe many physical phenomena.
-[@!
-Cite mandelbrot] Beno\^it Mandelbrot, commonly regarded as the father of fractal geometry, put it as follows in the \emph{The Fractal Geometry of Nature} [@!
-cite]:
+Beno\^it Mandelbrot, commonly regarded as the father of fractal geometry, put it as follows in the \emph{The Fractal Geometry of Nature} \cite{mandelbrot1982fractal}:
 
@@
 
-Fractal geometry has been used in various forms in fields such as physics @!, biology @!
-and economics.
-The first connection between fractals and the natural world was in the form of tangible objetcs such as coast lines, trees or bronchia, but the more hidden aspects of reality were soon shown to have fractal properties as well.
+Fractal geometry has been used in many fields, including physics @!
+\cite{}, geology \cite{cheng1997multifractal}, biology \cite{goldberger1992fractal} and economics \cite{turiel2003multifractal}.
+The first connection between fractals and the natural world was in the form of tangible objects such as coast lines, trees or bronchia, but the more hidden aspects of reality were soon shown to have fractal properties as well.
 The path traced out by the Brownian motion of a particle, the price chart for a commodity or the level of floods in a river delta have all been studied as fractals to great effect.
 
-One of the problems of fractal analysis has always been the difficulty of finding a fractal model for a given set of data.
+The main problem with fractal analysis has always been the difficulty of finding a fractal model for a given set of data.
 The assumption that data contains fractal properties by itself already allows us to calculate certain statistics (such as its dimension), but the various families of fractals that have been proposed are difficult to fit to data.
 This is called the \emph{inverse problem}.
-Instead of generating an image from a given mathematical description of a fractal (a popular programming exercise) we are asked to consider an image or dataset and find a good fractal model for it.
+Instead of generating an image or dataset from a given mathematical description of a fractal, we are asked to consider an image or dataset and find a good fractal model for it.
 
@@
 
+\subsection{Previous research}
+[@! Review the literature that does exist]
+
 \section{Methods}
 
-In this paper we focus on numerical data in many dimensions.
-That is, each instance in our dataset is represented by a fixed number of numerical features.
+In this paper we focus on numerical data an arbitrary number of dimensions.
+Each instance in our dataset is represented by a fixed number of numerical features.
 Call our dataset $X$, our instance $x \in X$ and $x \in {\mathbb R}^d$.
-We will assume that our dataset represents iid draws from a probability distribution over ${\mathbb R}^d$ which we will attempt to model.
+We will assume that our dataset represents iid samples from a probability distribution over ${\mathbb R}^d$ which we will attempt to model.
 
@@
 
-If we are to fit fractal models to data then the classic examples of single fractals are of little use to us. We require a family of fractals which is easy to  
-parametrize, and likely to generate a large number of organic and useful patterns.
+If we are to fit fractal models to data then we require a family of fractals which is easy to parametrize, and likely to encompass a large number of organic and useful patterns.
 We will use the formalism of \emph{Iterated Function Systems} (IFSs) for this purpose.
-We will briefly review the formalism here.
-For a more rigorous treament, we recommend [@!cite hutchinson].
+
+Since the aim of this paper is to present our induction algorithm and to introduce IFS models to a machine learning context, we will provide only an informal treatment of the properties of IFSs that are already well established.
+For a more rigorous treatment, we recommend \cite{hutchinson2000deterministic} and the references therein.
 
@@
 As noted before, it consists of three scaled down versions of itself.
-In mathematical terms, if we take the union of three specific transformations applied to the Sierpinski Gasket, we will get the Sierpinski gasket again.
+In mathematical terms, if we view the Sierpinski gasket as a set in ${\mathbb R}^2$ and take the union of three specific transformations applied to it, we will get the Sierpinski gasket again.
 The key result in the theory of iterated function systems is that for these three transformations there is only one set that has this property.
@@
 
-If we replace the transformations by others (making sure that they are contractive) the set for which this property holds will still be uniquely determined by the transformations, although it will differ from the Sierpinski gasket.
-Another helpful property of IFSs is that when we start with a different set, and we apply the transformations iteratively, the resulting sequence of sets will converge towards the unique single set that satisfies the scaling law.
+We will use the following notation.
+Let $S_1$, $S_2$, \ldots, $S_n$ be a set of contractive transformations on ${\mathbb R}^d$.
+(We will later limit the family of allowable functions, but from a mathematical point of view we only require them to be contractive).
+There exists a single set $Q$ that satisfies the following property:
+\begin{align}
+Q = S_1(Q) \cup S_2(Q) \cup \ldots \cup S_k(Q) \label{scaling law}
+\end{align}
 
-We will use the following notation. Let $S_1$, $S_2$, \ldots, $S_n$ be a set of contractive transformations, then there exists a single set $K$ that satisfies the following property:
+We call $\{S_i\}$ a \textit{scaling law}, and we say that $Q$ satisfies it.
+
+If we apply the transformations $S_i$ to a set $X$ that doesn't satisfy the scaling law (and take the union of the results), we get a different set $X'$.
+Call this function $S$: $S(X) = \bigcup S_i(X)$.
+One of the core results about IFSs is that when we iterate $S$ as $S^{n+1}(X) = S(S^n(X))$, then we get
 \[
-K = S_1(K) \cup S_2(K) \cup \ldots \cup S_n(K)
+\lim_{n \rightarrow \infty} S^n(X_0) = Q
 \]
+With $X_0$ any non-empty set and $Q$ the unique set that satisfies the scaling law $\{S_i\}$.
 
-If we consider the property a function of set $X \subseteq {\mathbb R}^d$: $S(X) = \bigcup S_i(X)$ with and we iterate the function $S^{n+1}(X) = S(S^n(X))$, then we have 
-\[
-\lim_{n \rightarrow \infty} S^n{X_0} = K
-\]
-With $X_0$ any non-empty set and $K$ the unique set that satisfies $S$'s scaling law.
-
-Of course, in statistics, we are usually interested in finding probability distributions rather than sets.
+Of course, in statistics, we are interested in finding probability distributions rather than sets.
 The IFS formalism can be extended to measures in a natural way.
@@
 \]
-So if $p$ is a probability measure, then $F(p)$ is a probability measure as well, which assigns $X$ the same probability as $p$ would assign to $F^{-1}(X)$, where $F^{-1}$ is the inverse of $F$.
+So if $p$ is a probability measure, then $F(p)$ is also a probability measure and $F(p)$ assigns $X$ the same probability as $p$ would assign to $F^{-1}(X)$, where $F^{-1}$ is the inverse of $F$.
 
-Our IFS over probability measures is made up, as before, of $n$ transformations, but here we also introduce $n$ weights $w_i$ with the constraint that they sum to one.
-The full scaling law then becomes
+Our IFS over probability measures consists, as before, of $k$ transformations, but here we also introduce $k$ weights $w_i$ with the constraint that they sum to one.
+Let $q$ be a probability distribution and $S = \{(S_i, w_i)\}$ be a scaling law.
+We say that $q$ satifies $S$ iff:
 \[
-k(X) = \sum_{i\in[1, n]} w_i S_i(k)(X)
+q(X) = \sum_{i\in[1, k]} w_i S_i(q)(X)
 \]
 
-As before, there is a unique probability measure that satisfies this scaling law, and if we use the scaling law as a function on probability measures:
-\[
-p = \sum w_i S_i(p)
-\]
-the iteration of this function will converge to the unique value $k$ as described above.
+As before q is uniquely defined by $S$. If we apply the transformations and weights to a different probability distribution iteratively, we get a sequence which converges to $q$:
+\begin{align*}
+p_{n+1} &= \sum w_i S_i(p_n) \\
+\lim_{n \rightarrow \infty} p_n &= q 
+\end{align*}
+
+This type of IFS is what we will use as a formalism for modeling datasets with fractals.
 
@@
 
-As discussed, an IFs is completely determined by its component transformations and their weights.
-The definition does not specify any kind of transformation, to make fitting IFSs to data feasible we must fix a family of transformations.
-For the purposes of this paper we will use similarities (also known as similitudes), transformations consisting of a translation, a rotation and a uniform scaling.
+As discussed, an IFS is completely determined by its component transformations and their weights.
+The definition does not specify any kind of transformation, so to make fitting IFSs to data feasible we must fix a family of transformations which we can search.
+For the purposes of this paper we will use \emph{similarities} (also known as \emph{similitudes}): maps on ${\mathbb R}^d$ consisting of a translation, a rotation and a uniform scaling.
 
@@
 
-It should be mentioned that the points generated in this way are not independently drawn.
-Any two succesive points in the sequence will satisfy the relation $x_{n+1} = S_i(x_{n})$.
+It should be mentioned that the sequence of points generated this way dos not represent a sequence of independent draws.
+Any two successive points in the sequence will satisfy the relation $x_{n+1} = S_i(x_{n})$ for some $i$.
 This is not true for two independently drawn points.
@@
 
+A more correct (and more expensive) way to generate random data from an IFS is to start with a draw from some initial distribution (eg.
+a multivariate normal) and apply $n$ transformations randomly chosen from $S$.
+The distribution of the resulting random variable will converge to $S$ with $n$.
+For multiple points the process can be repeated.
+This process adds a factor of $n$ to the time complexity of the chaos game, so it is not commonly used.
+
 \subsection{Code space}
+
+[@! picture]
 
@@
 We can compose the components of the IFS acording to the elements of $\sigma$.
-Call this function $S_sigma$:
+Call this function $S_\sigma$:
 \[
@@
 That is, the sequence of functions maps each $X$ to a unique point $x_\sigma$.
-In the case of probability measures, every probability measure is mapped to a probability measure that gives $x_\sigma$ full probability mass.
+In the case of probability measures, every probability measure is mapped to a probability measure that gives the single point $x_\sigma$ full probability mass.
+Figure [@!] illustrates the concept, an Iterated Function System assigns a code to each point on its support.
+Finite codes map to subsets of the support.
+Since the size of the subsets decreases exponentially we can talk about the endpoint of a finite coee without losing much accuracy.
+We will take the point to which the origin is mapped by $S_\sigma$ as the canonical choice of the endpoint of finite $\sigma$.
 
-We will require the inverse process: given a point on or near the support of some IFS, we wish to find $\sigma$.
-To this end we start with the standard normal multivariate distribution and apply all maps $S_\sigma$ for all $\sigma$ of a given lentgh $k$.
-The $\sigma$ with the greatest probability is used as a code for the point.
+Finding the endpoint for given code is straightforward, but we will also require the inverse process: given a point $x$ on or near the support of some IFS, we wish to find $\sigma$.
+To this end we start with the standard normal multivariate distribution and apply all maps $S_\sigma$ for all $\sigma$ of a given length $l$ giving us a set of $k^l$ normal distributions.
+The $\sigma$ which assigns $x$ the greatest probability is chosen.
 
-Since the variance of the distribution drops off exponentially this method will fail if points are to far off the support of the IFS (even if logarithmic probabilities are used).
+Since the variance of the distribution drops off exponentially as the length $l$ is increased, this method will fail if points are too far off the support of the IFS (even if logarithmic probabilities are used).
 To increase the range of the method we fall back on the distance from the point to the mean of the distribution.
-If the first method cannot distinguish between codes, we chose the code whose probability distribution has its mean closest to the point.
+If the first method cannot distinguish between codes, we choose the code whose probability distribution has its mean closest to the point.
 
-A word of caution on codes is appropriate: the ordering is often reversed from what our intuition suggests.
-If we number the three components of the Sierpinski triangle clockwise 0, 1 and 2, we might number a small section of the triangle by first moving into a one of the large subsections 2, then into one of the subsections of subsection 1 and then into subsection 0 of that.
+A word of caution on codes is appropriate: the ordering is often the reverse of what our intuition suggests.
+If we number the three components of the Sierpinski triangle clockwise from the top 0, 1 and 2, we might number a small section of the triangle by first moving into the top subtriangle (applying transformation 0), then into the left subtriangle (2) of that and then into the right subtriangle (1) of that.
 Intuitively, this is the most natural way of coding sections of the Sierpinski triangle.
@@
 \]
-and its code $\sigma$ is thus $(0, 1, 2)$. If we wish to travel into the Sierpinski gasket by this code, we must read it right-to-left.
+so its code $\sigma$ is $(0, 1, 2)$. If we wish to `travel into' the Sierpinski gasket by this code, we must read it right-to-left.
 
-We also note that when we use a simple IFS as a model for the codes---such as the two component IFS that maps the interval $[0,1]$ onto its lower and upper half, or the four component IFS that maps the unit square onto its four quarters---its codes coincide with commonly used spatial indexing functions and string-based constructions for the continuum.
-In this sense we can see the iterated function system as covering a subspace of ${\mathbb R}^d$ and the codes as describing a kind of metric that is intrinsic to the data rather than the embedding space.
+Finally, we wish to note that a line segment, the square, the cube and higher dimensional analogues can be represented as simple IFSs of $2^d$ components.
+For instance, for the unit line, we map the whole onto its lower and upper half with two components.
+The codes of this IFS are binary strings and the subsets they code are the diadic intervals.
+For the square, we can map the whole onto each one of its quadrants and get quaternary codes.
+Using this mechanism we can always map any dataset into the bi-unit hypercube and find a straightforward coding fot the instances in $2^d$-ary codes.
+What fractal fitting offers us is a potential for a subspace that fits the data more compactly and gives us a good code of lower arity.
+
+\subsection{Evaluating models}
+
+The standard way of evaluating the fit of a model to a dataset is the likelihood of the model given the data.
+This is often used in optimization algorithms for fitting probability models to data.
+In the case of fractal models, the likelihood function has an important drawback.
+
+Consider a dataset drawn from the Sierpinski gasket with a minute amount of two dimensional noise added.
+Since the Sierpinski gasket has dimension less than two, the probability that the point will still fall within the Sierpinski gasket will be zero.
+Thus all points will have probability zero under the Sierpinski gasket and its likelihood will be zero.
+Intuitively, the Sierpinski gasket is a good model for this data.
+
+For this reason we will often use the Hausdorff distance to evaluate our models. The Hausdorff distance  $H(X, Y)$ between two sets $X$ and $Y$ is defined as follows:
+\[ 
+H'(X, Y) = \max_{x\in X} \min_{y\in Y} d(x, y) 
+\]
+\[
+H(X, Y) = \max\left \{H(X, Y), H(Y, X)\right\}
+\]
+ (for infinite sets the $\max$ and $\min$ functions should be replaced with $\sup$ and $\inf$)
+
+The most intuitive way of understanding the Hausdorff distance is as a game where the first player must get from one set to the other in as small a distance as possible, and the second player will choose his starting point to maximize that distance.
+
+Hausdorff distance is a common tool in dealing with the fractal inverse problem [@!
+cite].
+We will use it to evaluate an IFS model by generating a number of points from the model and calculating the Hausdorff distance of that set to the model.
+Since the Hausdorff distance has not been substantiated as a statistical measure, we can only use it as an intermediate indication of model quality.
+To prove the utility of IFS models we will use them to solve three classical machine learning problems (classification, dimension reduction and clustering).
+
+We estimate the uncertainty in the estimate of the Hausdorff distance by repeating the process $50$ times and calculating the sample standard deviation.
 
@@
 That is, it can be mapped to a low dimensional space without any loss of information.
+A simple example is that of all points lying on a line in some high dimensional space.
 
-IFSs have a particular advantage over classical models like Gaussian mixtures in that they can vary their intrinsic dimension over the whole range of positive reals.
+IFSs have a particular advantage over classical models like Gaussian mixtures: their intrinsic dimension varies over the whole range of positive reals.
+A search through the space of IFSs can return models of any dimension.
 
@@
 
-\subsection{Evaluating models}
-
-The standard way of evaluating the fit of a model to a dataset is the likelihood of the model given the data.
-This is often used in optimization algorithms for fitting probability models to data.
-In the case of fractal models, the likelihood function has an important drawback.
-
-Consider a dataset drawn from the Sierpinski gasket with a minute amount of two dimensional noise added.
-Since the Sierpinski gasket has dimension less than two, the probability that the point will still fall within the Sierpinski gasket will be zero.
-Thus all points will have probability zero under the Sierpinski gasket and its likelihood will be zero.
-Intuitively, the Sierpinski gasket is a good model for this data (since the noise term can be made arbitarily small).
-
-For this reason we will often use the Hausdorff distance to evaluate our models. The Hausdorff $H(X, Y)$ distance between two sets $X$ and $Y$ is defined as follows:
-\[ 
-H'(X, Y) = \max_{x\in X} \min_{y\in Y} d(x, y) 
-\]
-\[
-H(X, Y) = \max\left \{H(X, Y), H(Y, X)\right\}
-\]
- (for infinite sets the $\max$ and $\min$ functions should be replaced with $\sup$ and $\inf$)
-
-The most intuitive way of understanding the Hausdorff distance is as a game where player one must get from one set to the other in as small a distance as possible, and the the second player will choose his starting point to maximize that distance.
-
-Hausdorff distance is a common tool in dealing with the fractal inverse problem [@!
-cite].
-We will use it to evaluate an IFS model by generating a number of points from the model and calculating the Hausdorff distance of that set to the model.
-Since the Hausdorff distance has not been substantiated as a statistical measure, we can only use it as an intermediate indication of model quality.
+We estimate the uncertainty of this estimate as follows: bootstrap a version of the dataset by sampling to the same size with replacement, and estimate the dimension of that dataset.
+We repeat the process 50 times and take the sample standard deviation as an estimate of the uncertainty.
 
@@
 
-To study the IFS as a statistical model, it helps to have an anlogous model to compare it with.
+To study the IFS as a statistical model, it helps to have a commonly used analogous model to compare it with.
 We use the Mixture of Gaussians (MoG) model.
@@
 Each component also has a prior $w_i$.
-The complete model described the following probability distribution:
+The complete model describes the following probability distribution:
 \[
@@
 \item We start with in initial IFS $S$, chosen by one of the methods described below. 
-\item (Expectation) We assign each datapoint a code $\sigma$ of length $k$ (where $k$ is a parameter of the algorithm). The codes are provides by the model $S$ as described in section @!.
+\item (Expectation) We assign each datapoint a code $\sigma$ of length $k$ (where $k$ is a parameter of the algorithm). The codes are provided by the model $S$ as described in section @!.
 \item (Maximization) We know that if a point $x_\sigma$ is mapped using one of our components we have $S_i(x_\sigma) = x_{i\sigma}$.
 That is, the symbol for the mapping component is prepended to the code.
-For each $x_\sigma$ in our dataset we search the data for a point with a code $\tau = i\sigma_{:-1}$, the symbol $i$ followed by the elements of $\sigma$ save for the last.
+For each $x_\sigma$ in our dataset we search the data for a point with a code $\tau = i\sigma_{1:k-1}$, the symbol $i$ followed by the elements of $\sigma$ save for the last.
 For each $i$ we maintain ordered lists of the matches found.
-\item (Maximization) For each component $S_i$ we know have two lists of points such that $S_i$ should map one into the other with minimal squared error.
+\item (Maximization) For each component $S_i$ we now have two lists of points such that $S_i$ should map one into the other with minimal squared error.
 We used the method described in @!
-to find the optimal similarity transformation.
+to find $S_i$.
 \end{itemize}
@@
 
+% This means that for each $i$ we have two sets of points at each depth.
+If the components are equally weighted, we will have a set of about $1/n$ of the data, mapping into a set of about $1/{n^2}$ of the data.
+At depth two the ratios become $1/{n^2}$ and $1/{n^3}$ and so on.
+For large depths, only specific codes will have matching points.
+
 \subsubsection{Multiple matches}
 
-It is prefectly possible that we may get multiple points per code.
+It is perfectly possible that we get multiple points per code.
 Indeed, at low depths, it is almost certain.
-In this case we have two sets of points $X_\sigma$ and $X_\tau$ such that the points in the formar should map onto the latter.
+In this case we have two sets of points $X_\sigma$ and $X_\tau$ such that the points in the former should map onto the latter.
 But our algorithm for finding the map $S_i$ requires the individual points to be matched to each other.
@@
 Ideally we would then extend our method for finding maps to paired MVNs rather than paired points, but this is a highly complex problem.
-Instead, we find for each MVN the transformation that transforms the standard distribution into that MVN.
-We then generate a small amount (20 in our experiments) of points from the standard MVN and map these to both sets.
+Instead, we find for each MVN the transformation that transforms the standard distribution ($N(0, I)$) into that MVN.
+We then generate a small amount of points (20 in our experiments) from the standard MVN and map these to both sets.
 We add these 20 point pairs to the list of matched points from which the maps will be calculated.
@@
 Some undesired behavior can emerge when the initial model is far from the optimum.
-Consider for instance the case where the Sierpinski triangle is the current model and all data isto the top right of the triangle.
+Consider for instance the case where the Sierpinski triangle is the current model and all data is to the top right of the triangle.
 The symbol 2 would not occur in any of the codes and we would have no basis to find $S_2$ in the maximization step.
@@
 
-\subsubsection{Branching}
+% \subsubsection{Branching}
 
-[@! Test if this helps]
+% [@! Test if this helps]
 
-As an optimization algorithm, EM is strictly linear in the sense that it only considers a single model at each iteration.
+%As an optimization algorithm, EM is strictly linear in the sense that it only considers a single model at each iteration.
 To avoid local minima it can often help to evaluate multiple models per iteration.
@@
 
-During the iteration of the algorithm we maintain a buffer of models.
+%During the iteration of the algorithm we maintain a buffer of models.
 After each iteration, we generate $b$ versions of the current best model perturbed by a parameter $\sigma$ as with the edge cases.
@@
 
-This method is analogous with beam search.
+%This method is analogous with beam search.
 Its advantage is that we perform the expensive EM step the same number of times as the non-branching model, and the extra cost lies purely in testing evaluating the extra models.
@@
 
-There are three expected benefits to fractal modeling that we wish to expore here.
+There are three expected benefits to fractal modeling that we wish to explore here.
 First, we expect that the models will scale their dimension to fit that of the data (whereas our comparison model will be unable to do so).
@@
 are of actual use outside of the fitting procedure.
-These codes describe a metric space that is much more compact\footnote{in an informal sense} that the Euclidean space in which the data is embedded.
-Furthermore, the transformation to strings allows the use of a whole family of technique that was not previously available.
+These codes describe a metric space that is much more compact\footnote{in an informal sense} than the Euclidean space in which the data is embedded.
+Furthermore, the transformation to strings promises the use of a whole family of techniques that was not previously available.
 
@@
 
+[@! EXPERIMENT Fit models to data ]
+
+[@!
+EXPERIMENT for various datasets, for various depths and component sizes, scatter plot the difference between the model dimension and the data dimension against the model accuracy (by hausdorff distance) show uncertainty with error bars]
+
 First, we fit IFS models to our datasets (ignoring the class information).
 We report the Haussdorff distance of a sample of 1000 points from the data to 1000 points generated by the model (we consider the iris dataset to be too small for this purpose).
-For this test we test on a random 25\% of the data and repeat the experiment 20 and report the sample standard deviation and the mean.
+For this test we test on a random 25\% of the data, repeat the experiment 20 times and report the sample standard deviation and the mean.
 
@@
 In this case we are interested in the training behavior of the model, so we do not withhold test data.
-We measure the dimension of the dataset on a random sample of 300 points (repeating 20 times to estimate the sample standard deviation) and generate 300 points from the best model and measure the dimension in the same way, taking the standard deviation over 20 separately trained models, so that it represents the combined uncertainty over training the model and estimating dimension.
+We measure the dimension of the dataset on a random sample of 300 points (repeating 50 times to measure the uncertainty) and generate 300 points from the best model and measure the dimension in the same way.
+We expect that the difference between the model dimension and the data dimension depends on the quality of the model.
+To investigate we produce a scatterplot of the absolute difference between the two ($\delta d$) and the quality of the model fit by Hausdorff distance ($d_H$).
 
@@
 
-\subsection{Coding}
+\subsubsection{Bayesian}
+\subsubsection{Single model}
 
-@! Investigate the use of DFAs on the code form of the dataset. Use for classification.
-@! Map the codes back to a space ${\mathbb R}^{d'}$ as a method of dimensionality reduction.
+\begin{table}
+\begin{tabular}{l r | r r | r r r r r |}
+\hline
+  & dim & ifs bayes & ifs single & mog & ann & tree & svm & knn \\
+\hline
+census & 3.2 (0.2) & 2.9 (0.45) & & & & \\
+forest & & & & & & \\
+abalone & & & & & & \\
+\hline
 
+\end{tabular}
+\caption{Hausdorff distances for models trained against various datasets.}
+\label{tablelabel}
+\end{table}
 \subsection{Performance}
@@
 
-We have presented a framework for a fractal based approach to the statistics of point patterns.
+We have presented a framework for a fractal approach to the statistics of point patterns.
 We introduced an EM-style algorithm for fitting fractal models to data.
-Compared to search algorithms for other types of models the algorithm is relatively expensive, but compared to other approaches for fractal modeling the algorithm represents a significant imporvement @!.
+Compared to search algorithms for other types of models the algorithm is relatively expensive, but compared to other approaches for fractal modeling the algorithm represents a significant improvement @!.
 
-We have given some indication of how the study of fractal modeling can open up interesting new prospects in machine learning, for instance by 
+Further, we have given some indication of how the study of fractal modeling can open up interesting new prospects in machine learning.
+We expect that these methods will allow many avenues of future research.
+One option would be to investigate the use of more expressive maps.
+The similitudes used here could be replaced by neural networks, for instance.
+The EM search algorithm is fairly simple, and could be extended to try various branches and backtrack if it deviates too far from a good model.
+
+Finally, there is the task of learning random fractals.
+The random fractal model is an extension of the Iterated Function System which produces not one single set but a distribution over all sets.
+The dataset for which to learn a random fractal would not consist of one list of vectors, but of a list of list of vector.
+For instance, a collection of timeseries preceding a critical event, or a collection of images of snowflakes.
+In [@!] we showed a proof-of-concept, indicating that, in principle, such models can be learned by evolutionary algorithm, but we did not advance beyond the simplest conceivable model.
+If the EM algorithm described here could be extended to random fractals, it would open up a new type of machine learning task.
+
+\bibliographystyle{siam}
+\bibliography{fractal}
+
 \end{document}

### 2012-10-31 fractal paper/src/fractal.tex
....

@@
-\documentclass[11pt]{article}
+\documentclass[10pt]{article}
 
@@
 \usepackage{amsmath, amsthm, amssymb}
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 \newtheorem*{exm}{Example}
+
+\usepackage{color}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
 
@@
 
-\begin{itemize}
-  \item Self similarity: A part is a scaled down copy of the whole.
+\begin{description}
+  \item[Self similarity] A part is a scaled down copy of the whole.
 See for instance the Sierpinski triangle (fig !@).
 It consists of three triangular shapes which are scaled down copies of the whole (as such these three components each also consist of scaled down copies of themselves, and so on).
-  \item Infinitely small detail: 'Zooming in' reveals ever finer detail.
+  \item[Infinitely small detail] 'Zooming in' reveals ever finer detail.
 In the case of the Sierpinski triangle we will simply see the same shape recurring again and again, but fractals like the Mandelbrot set reveal a great variety of shapes.
-  \item Non-integer dimension: for instance the Sierpinski triangle has dimension about 1.58.
-This is a counter intuitive notion, that is not simple to explain.
-We refer the reader to @!
-for an intuitive explanation.
-\end{itemize}
+  \item[Non-integer dimension] for instance the Sierpinski triangle has dimension of about $1.58$.
+This is a counter-intuitive notion, that is not simple to explain.
+We refer the reader to \cite{schroeder2009fractals} for an intuitive explanation.
+\end{description}
+
+\begin{figure}[h!]
+  \centering
+  \begin{subfigure}[b]{\textwidth}@!
+    %\includegraphics[width=\textwidth]{}
+    \label{fig:sierpinski}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{\textwidth}@!
+   % \includegraphics[width=\textwidth]{}
+    \label{fig:sierpinski-measure}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{\textwidth}@!
+    %\includegraphics[width=\textwidth]{}
+    \label{fig:ifs-measure}
+  \end{subfigure}
+  \caption{(\ref{fig:sierpinski}) The Sierpinski gasket, a simple example fractal.
+Starting with any initial image, we can compose three transformations of the original.
+The iteration of this map converges to the Sierpinski gasket.
+The initial image does not influence the end result.
+(\ref{fig:sierpinski-measure}) Starting with probability measures rather then images, we can define a fractal probability measure.
+Again the choice of initial distribution doesn't matter.
+In this example we have assigned one of the transformations a higher weight than the other two.
+(\ref{fig:ifs-measure})If we change the transformations, the limit set changes too.
+Here we show some probability distributions that are described in this way.
+Note that simple shapes such as the triangle, the square and the line can also be described like this.}
+\end{figure}
 
@@
 It was not until the 1970s that they began to be seen as a single and useful family of mathematical objects that could accurately describe many physical phenomena.
-Beno\^it Mandelbrot, commonly regarded as the father of fractal geometry, put it as follows in the \emph{The Fractal Geometry of Nature} \cite{mandelbrot1982fractal}:
+Mandelbrot put it as follows in the \emph{The Fractal Geometry of Nature} \cite{mandelbrot1982fractal}:
 
@@
 
-This problem was investigated thoroughly in the context of image compression, but has received relatively little attention in the area of statistics, and fitting a fractal model to data rather than an image.
+This problem was investigated thoroughly in the context of image compression \cite{hart1996fractal}, but has received relatively little attention in the area of statistics, and fitting a fractal model to data rather than an image.
 In particular the case of data in arbitrary dimensions has been left largely uninvestigated.
@@
 \subsection{Code space}
-
+\label{sec:code-space}
 [@! picture]
@@
 \item We start with in initial IFS $S$, chosen by one of the methods described below. 
-\item (Expectation) We assign each datapoint a code $\sigma$ of length $k$ (where $k$ is a parameter of the algorithm). The codes are provided by the model $S$ as described in section @!.
+\item (Expectation) We assign each datapoint a code $\sigma$ of length $d$ (where $d$ is a parameter of the algorithm).
+The codes are provided by the model $S$ as described in section \ref{sec:code-space}.
 \item (Maximization) We know that if a point $x_\sigma$ is mapped using one of our components we have $S_i(x_\sigma) = x_{i\sigma}$.
@@
 \item (Maximization) For each component $S_i$ we now have two lists of points such that $S_i$ should map one into the other with minimal squared error.
-We used the method described in @!
-to find $S_i$.
+We used the method described in \cite{umeyama1991least} to find $S_i$.
 \end{itemize}
+
+We will often use a subsample of the data, to allow us to increase the depth or number of components. In such cases, we reample at each EM iteration.
 
@@
 Iterated function systems are not very good at approximating data that is weighted far from the origin.
-for this reason we find the affine transformation which maps the data to the range $[0,1]$ in each dimension and puts the mean at the origin.
+For this reason we find the affine transformation which maps the data to the range $[0,1]$ in each dimension and puts the mean at the origin.
 
-We store this transformation with the model to map the model back to the original space.
+We store this transformation with the model to map the model back to the data space as needed.
 
@@
 
-The algorithm as described here can work at all depths $k$.
-Since each depth may contain information about self similar scaling in the data, we want to consider all depths $\leq k$ rather than just $k$.
+The algorithm as described here can work at all depths $d$.
+Since each depth may contain information about self similar scaling in the data, we want to consider all depths $\leq d$ rather than just $d$.
 This is easily done by considering codes of all length up to and including $k$.
@@
 We report the Haussdorff distance of a sample of 1000 points from the data to 1000 points generated by the model (we consider the iris dataset to be too small for this purpose).
-For this test we test on a random 25\% of the data, repeat the experiment 20 times and report the sample standard deviation and the mean.
+We withhold a random 25\% of the data for testing, repeat the experiment 20 times and report the sample standard deviation and the mean of the resulting values for symmetric error.
 
@@
 In this case we are interested in the training behavior of the model, so we do not withhold test data.
-We measure the dimension of the dataset on a random sample of 300 points (repeating 50 times to measure the uncertainty) and generate 300 points from the best model and measure the dimension in the same way.
+We measure the dimension of the dataset on a random sample of 300 points (repeating 25 times to measure the uncertainty) and generate 300 points from the best model and measure the dimension in the same way.
 We expect that the difference between the model dimension and the data dimension depends on the quality of the model.
-To investigate we produce a scatterplot of the absolute difference between the two ($\delta d$) and the quality of the model fit by Hausdorff distance ($d_H$).
+Figure @!
+plots the data dimension against the model dimension.
 
@@
 \hline
- & ifs1 & ifs3 & ifs5 & ifs7 & mog \\
+ & ifs1 & ifs3 & ifs7 & mog \\
 \hline
@@
 
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\textwidth]{../img/out-table.pdf}
+\caption{Dimensional scaling. The marker size and color both indicate the quality of the model fit by Hausdorff distance. }
+\label{fig:dim_scale}
+\end{figure}
+
 \subsection{Classification}
@@
 \subsubsection{Bayesian}
+
 \subsubsection{Single model}
 
+An alternative way to construct a classifier with IUFS modelling is to learn a single model for the whole dataset, ignoring class labels.
+For some datasets, the set of all points may present a more coherent and simple distribution than the sets split by class label.
+Another advantage of this approach is that we only have to learn one model, which is cheaper both in training and in evaluation.
+
+The model provides us with a mapping from the data space (${\mathbb R}^d$) to code space (${1, \ldots, k}^d$).
+To turn this model into a classifier we proceed as follows.
+We build a code tree such that each possible code represents a path from the root to the leaf node.
+We then traverse this tree for each point in the training set, maintaining a frequency count fora each class observed at each node.
+Thus, the root node gives us the frequency distribution over the classes for the whole dataset, and each node below it refines that to a distribution over a section of the dataset whose code starts with a given prefix.
+This turns each node into a very basic classifier, which simply returns the class with the highest frequency count.
+The root node will give us the classifier that uses nothing mode than class frequencies and each node below it will use more information about the neighbourhood of the point.
+
+For each point to be classified we take the probabilities at each node in the path and average them evenly.
+
 \begin{table}
-\begin{tabular}{l r | r r | r r r r r |}
+\begin{tabular}{| l | r l r l | r l r l  r l r l |}
 \hline
-  & dim & ifs bayes & ifs single & mog & ann & tree & svm & knn \\
+  & ifs bayes & & ifs single & & mog & & ann & & knn & & tree & \\
 \hline
-census & 3.2 (0.2) & 2.9 (0.45) & & & & \\
-forest & & & & & & \\
-abalone & & & & & & \\
+census  & & & & & & & 0.054 &\color{gray}(0) & 0.06 & \color{gray} (0.002) & 0.05 & \color{gray} (0.001) \\
+forest  & & & & & & & 0.163 &\color{gray}(0.001) & 0.138 &\color{gray}(0.001) & 0.12&\color{gray}(0.003) \\
+abalone & & & & & & & 0.47 & \color{gray}(0.008) & 0.47 & \color{gray}(0.006) & 0.48 & \color{gray}(0.01) \\
+mnist   & & & & & & & & & & & & \\
 \hline
@@
 \caption{Hausdorff distances for models trained against various datasets.}
-\label{tablelabel}
+\label{classification-results}
 \end{table}
-\subsection{Performance}
+\subsection{Time performance}
+
+An informal analysis of the algorithm shows that the expectation step costs $O(k^dn)$, with the maximization step taking roughly $O(n^2)$.
+This falsely suggests that the expectation step, being linear in the number of data points, is negligable compared to the maximization step, which is quadratic.
+Of course, when viewed in terms of the parameters the expectation step is exponential and this is what causes the expectation step to become the bottleneck for anything but very low values of $k$ and $d$.
 
@@
 For instance, a collection of timeseries preceding a critical event, or a collection of images of snowflakes.
-In [@!] we showed a proof-of-concept, indicating that, in principle, such models can be learned by evolutionary algorithm, but we did not advance beyond the simplest conceivable model.
-If the EM algorithm described here could be extended to random fractals, it would open up a new type of machine learning task.
+In \cite{bloem2010fractal} we showed a proof-of-concept, indicating that such models can be learned by evolutionary algorithm in principle, but we did not advance beyond the simplest conceivable model.
+If the EM algorithm described here could be extended to random fractals, it would open up a new type of machine learning task, where instances ae represented by point patterns, and the model to be learned is a distribution over distributions.
 

### 2012-11-05 P23WP1Q2Deliverable/src/deliverable.p23wp1q3.tex
...

--- 2012-08-09 P23WP1Q2Deliverable/src/deliverable.p23wp1q2..tex
@@
 \usepackage{lscape}
+
+\usepackage{color}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
 
@@
     \hline
- Internet & 22963 & 48436 & 75877.0 & 1.0 & & & & & & & & & & & \\   
- WWW & 325729 & 1497134 & & & & & & & & & & & & & \\
- Epinions & 75879 & 508837 & & & & & & & & & -0.01 & & & & \\ 
- Actors  & 383640 & 1229020 & & & & & & & & & & & & & \\
- E. Coli  & 2859 & 6890 & 2859 & 1.0 & 18 & 4.70 & 0 & 0 & & & -0.16 & & & & \\
- C. Elegans  & 297 & 2345 & 297 & 1.0 & 5 & 2.45 & 0.18 & 0.29 & & & -0.15 & -4.5 & 0.39 & 0.38 & 0.39 \\
- IAFB  & 8318 & 61255 & 8318 & 1.0 & & & & & & & -0.26 & & 0.025 & &  \\ 
+ Internet & 22963 & 48436 & 75877.0 & 1.0 & & & & & 2.1 & 0.5 & & & & & \\   
+ WWW & 325729 & 1497134 & & & & & & & \color{gray} 2.5 & 0.0 & & & & & \\
+ Epinions & 75879 & 508837 & & & & & & & 3.8 & 0.6 & -0.01 & & & & \\ 
+ Actors  & 383640 & 1229020 & & & & & & & \color{gray}2.5 & 0.0 & -0.12 & & & & \\
+ E. Coli  & 2859 & 6890 & 2859 & 1.0 & 18 & 4.70 & 0 & 0 & \color{gray} 2.3 & 0.04 & -0.16 & & & & \\
+ C. Elegans  & 297 & 2345 & 297 & 1.0 & 5 & 2.45 & 0.18 & 0.29 & 3.3 & 0.33 & -0.15 & -4.5 & 0.39 & 0.38 & 0.39 \\
+ IAFB  & 8318 & 61255 & 8318 & 1.0 & & & & & \color{gray} 2.2 & 0.0 & -0.26 & & 0.025 & &  \\ 
  \hline

### 2012-11-05 fractal paper/src/fractal.tex
...

@@
-\documentclass[10pt]{article}
+\documentclass[10pt,a4paper,oneside]{article}
 
@@
 \DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argmax}{argmax}
 
@@
 \begin{abstract}
-\noindent We present a new algorithm for fitting fractals to data.
-We use this algorithm to investigate the use of fractal models in a machine learning context.
+\noindent We present an algorithm for fitting fractals to data and we investigate the use of fractal models in machine learning.
 Fractals are mathematical objects that display great detail and complexity despite having simple descriptions.
-It has often been suggested that fractals model many natural and other real-world structures better than the shapes of classical geometry \cite{mandelbrot1982fractal}.
+It has been suggested that fractals model many natural and other real-world structures better than the shapes of classical geometry \cite{mandelbrot1982fractal}.
 \end{abstract}
 
-\section{Introduction}
+\begin{figure}[b]
+  \centering
+  \begin{subfigure}[b]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../img/sierpinski.pdf}
+    \caption{}
+    \label{fig:sierpinski}
+  \end{subfigure}  
+  \hspace{0.03\textwidth}
+  \begin{subfigure}[b]{0.3\textwidth}
+   \includegraphics[width=\textwidth]{../img/sierpinski-evo.pdf}
+    \caption{}
+  \end{subfigure}
+  \hspace{0.03\textwidth}
+  \begin{subfigure}[b]{0.3\textwidth}
+    \includegraphics[width=0.45\textwidth]{../img/sierpinski-off.png}
+    \includegraphics[width=0.45\textwidth]{../img/random013.png} \\
+    \includegraphics[width=0.45\textwidth]{../img/random010.png}
+    \includegraphics[width=0.45\textwidth]{../img/mvn1k.png}
+    \caption{}
+  \end{subfigure}
+  \caption{\small (a) The Sierpinski triangle, a simple example fractal. (b) Constructing \textit{(a)} from an arbitrary initial image. (c) Examples of IFS probability distributions.}
+  \label{fig:examples}
+\end{figure}
 
-Fractals are mathematical objects (generally sets or measures) with certain remarkable properties.
-What constitutes a fractal is not precisely defined\footnote{Mandelbrot originally defined it as a set whose topological dimension differs from its Hausdorff dimension, but retracted this definition, stating that he preferred the word to be not precisely defined.}, but the following properties are common:
+Fractals are mathematical objects with certain remarkable properties. What constitutes a fractal is not precisely defined\footnotemark, but the following properties are common:
 
@@
   \item[Self similarity] A part is a scaled down copy of the whole.
-See for instance the Sierpinski triangle (fig !@).
-It consists of three triangular shapes which are scaled down copies of the whole (as such these three components each also consist of scaled down copies of themselves, and so on).
+See for instance the Sierpinski triangle (fig.
+\ref{fig:sierpinski}).
+It consists of three triangular shapes which are scaled down copies of itself.
   \item[Infinitely small detail] 'Zooming in' reveals ever finer detail.
-In the case of the Sierpinski triangle we will simply see the same shape recurring again and again, but fractals like the Mandelbrot set reveal a great variety of shapes.
-  \item[Non-integer dimension] for instance the Sierpinski triangle has dimension of about $1.58$.
+In the case of the Sierpinski triangle we will see the same shape recurring again and again, but fractals like the Mandelbrot set reveal a great variety of shapes.
+  \item[Non-integer dimension] The Sierpinski triangle, for example, has a dimension of $1.58$.
 This is a counter-intuitive notion, that is not simple to explain.
@@
 
-\begin{figure}[h!]
-  \centering
-  \begin{subfigure}[b]{\textwidth}@!
-    %\includegraphics[width=\textwidth]{}
-    \label{fig:sierpinski}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{\textwidth}@!
-   % \includegraphics[width=\textwidth]{}
-    \label{fig:sierpinski-measure}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{\textwidth}@!
-    %\includegraphics[width=\textwidth]{}
-    \label{fig:ifs-measure}
-  \end{subfigure}
-  \caption{(\ref{fig:sierpinski}) The Sierpinski gasket, a simple example fractal.
-Starting with any initial image, we can compose three transformations of the original.
-The iteration of this map converges to the Sierpinski gasket.
-The initial image does not influence the end result.
-(\ref{fig:sierpinski-measure}) Starting with probability measures rather then images, we can define a fractal probability measure.
-Again the choice of initial distribution doesn't matter.
-In this example we have assigned one of the transformations a higher weight than the other two.
-(\ref{fig:ifs-measure})If we change the transformations, the limit set changes too.
-Here we show some probability distributions that are described in this way.
-Note that simple shapes such as the triangle, the square and the line can also be described like this.}
-\end{figure}
-
-The earliest examples of fractals were described around the beginning of the twentieth century as counterexamples to conjectures in planar geometry and mostly considered pathological.
+The earliest examples of fractals were described around the beginning of the twentieth century as counterexamples to conjectures in planar geometry and analysis and mostly considered pathological.
 It was not until the 1970s that they began to be seen as a single and useful family of mathematical objects that could accurately describe many physical phenomena.
@@
 
-Fractal geometry has been used in many fields, including physics @!
-\cite{}, geology \cite{cheng1997multifractal}, biology \cite{goldberger1992fractal} and economics \cite{turiel2003multifractal}.
+Fractal geometry has been used in many fields, including physics \cite{mandelbrot1984fractals}, geology \cite{cheng1997multifractal}, biology \cite{goldberger1992fractal} and economics \cite{turiel2003multifractal}.
 The first connection between fractals and the natural world was in the form of tangible objects such as coast lines, trees or bronchia, but the more hidden aspects of reality were soon shown to have fractal properties as well.
@@
 
-This problem was investigated thoroughly in the context of image compression \cite{hart1996fractal}, but has received relatively little attention in the area of statistics, and fitting a fractal model to data rather than an image.
+This problem was investigated thoroughly in the context of image compression (see for instance \cite{hart1996fractal}), but has received relatively little attention in the area of statistics, and fitting a fractal model to data rather than an image.
 In particular the case of data in arbitrary dimensions has been left largely uninvestigated.
 
-\subsection{Previous research}
-[@! Review the literature that does exist]
+\footnotetext{Mandelbrot originally defined it as a set whose topological dimension differs from its Hausdorff dimension, but retracted this definition, stating that he preferred the word to be not precisely defined.}
 
-\section{Methods}
+\subsection*{Previous research}
 
-In this paper we focus on numerical data an arbitrary number of dimensions.
+The most fruitful algorithms for the induction of fractal models so far have been partitioned iterated function systems and evolutionary algorithms.
+
+Partitioned iterated function systems \cite{hart1996fractal} are a subset of iterated function systems (described in the next section) designed for fast image compression by fractal models.
+The compression algorithm segments an image in to two tilings of different size, and then attempts to map the larger tiles to the smaller tiles.
+The resulting set of contraction mappings can be iterated to approximate the original image.
+The approach works well for two dimensional data stored in binned form, but does not translate well to higher dimensions.
+
+Another approach is to represent a fractal model as a point in some space (eg.
+a vector or a bitstring) and to search the space for a model which approximates the data by some target function using generic search algorithms.
+Fractal models tend to be difficult to optimize for and only specific methods show promising results.
+For instance, the use of evolutionary strategies combined with Hausdorff distance as a target function has been shown to be able to reconstruct fractals from data, and to fit models to common datasets.
+
+Nevertheless, the generic nature of these search methods mean that they do not exploit any knowledge of the fractal models.
+We expect that the induction of fractal models can be sped up and made more accurate if we tailor the search algorithm to the model.
+
+The investigation of iterated function systems has yielded some results that are promising, but have not resulted in an algorithm for the induction of fractal models in arbitrary dimensions.
+Examples include the method of moments \cite{rinaldo1994inverse}, similarity hashing \cite{hart1997similarity} and wavelet analysis \cite{struzik1996coastline}.
+
+\section*{Methods}
+
+In this paper we focus on numerical data in an arbitrary number of dimensions.
 Each instance in our dataset is represented by a fixed number of numerical features.
@@
 
-\subsection{Iterated Function Systems}
+\subsection*{Iterated Function Systems}
 
@@
 We will use the following notation.
-Let $S_1$, $S_2$, \ldots, $S_n$ be a set of contractive transformations on ${\mathbb R}^d$.
+Let $S_1$, $S_2$, \ldots, $S_k$ be a set of contractive transformations on ${\mathbb R}^d$.
 (We will later limit the family of allowable functions, but from a mathematical point of view we only require them to be contractive).
@@
 
-\subsection{Representing IFSs}
+\subsection*{Representing IFSs}
 
@@
 The method described below can easily be extended to other families of transformations such as affine transformations or neural networks.
-The only requirement is that we must have some way of finding a transformation $F$ given two ordered point sets ${x_i}, {y_i} \subseteq {\mathbb R}^d$ so that $|F(x_i) - y_i|$ is minimized.
+The only requirement is that we must have some way of finding a transformation $F$ within this family given two ordered point sets $\{x_i\}, \{y_i\} \subseteq {\mathbb R}^d$ so that the error is minimized.
 For similarities there is a direct and optimal method.
-[@!
-cite]
+\cite{umeyama1991least}
 
-\subsection{Drawing points from an IFS distribution}
+\subsection*{Drawing points from an IFS distribution}
 
@@
 
-It should be mentioned that the sequence of points generated this way dos not represent a sequence of independent draws.
+It should be mentioned that the sequence of points generated this way does not represent a sequence of independent draws.
 Any two successive points in the sequence will satisfy the relation $x_{n+1} = S_i(x_{n})$ for some $i$.
@@
 
-\subsection{Code space}
+\subsection*{Code space}
 \label{sec:code-space}
-[@! picture]
 
@@
 Finite codes map to subsets of the support.
-Since the size of the subsets decreases exponentially we can talk about the endpoint of a finite coee without losing much accuracy.
+Since the size of the subsets decreases exponentially we can talk about the endpoint of a finite code without losing much accuracy.
 We will take the point to which the origin is mapped by $S_\sigma$ as the canonical choice of the endpoint of finite $\sigma$.
@@
 Finding the endpoint for given code is straightforward, but we will also require the inverse process: given a point $x$ on or near the support of some IFS, we wish to find $\sigma$.
-To this end we start with the standard normal multivariate distribution and apply all maps $S_\sigma$ for all $\sigma$ of a given length $l$ giving us a set of $k^l$ normal distributions.
+To this end we start with the standard normal multivariate distribution and apply all maps $S_\sigma$ for all $\sigma$ of a given length $l$ giving us a set of $k^l$ normal distributions (since an affine transformation of a normal distribution is also a normal distribution).
 The $\sigma$ which assigns $x$ the greatest probability is chosen.
@@
 Since the variance of the distribution drops off exponentially as the length $l$ is increased, this method will fail if points are too far off the support of the IFS (even if logarithmic probabilities are used).
-To increase the range of the method we fall back on the distance from the point to the mean of the distribution.
+To increase the range of the method, we fall back on the distance from the point to the mean of the distribution.
 If the first method cannot distinguish between codes, we choose the code whose probability distribution has its mean closest to the point.
@@
 A word of caution on codes is appropriate: the ordering is often the reverse of what our intuition suggests.
-If we number the three components of the Sierpinski triangle clockwise from the top 0, 1 and 2, we might number a small section of the triangle by first moving into the top subtriangle (applying transformation 0), then into the left subtriangle (2) of that and then into the right subtriangle (1) of that.
+If we number the three components of the Sierpinski triangle clockwise from the top 0, 1 and 2, we might number a small section of the triangle by first moving into the top subtriangle, then into the left subtriangle (2) of that and then into the right subtriangle (1) of that.
 Intuitively, this is the most natural way of coding sections of the Sierpinski triangle.
-[@!
-fig]
 
@@
 For the square, we can map the whole onto each one of its quadrants and get quaternary codes.
-Using this mechanism we can always map any dataset into the bi-unit hypercube and find a straightforward coding fot the instances in $2^d$-ary codes.
-What fractal fitting offers us is a potential for a subspace that fits the data more compactly and gives us a good code of lower arity.
+Using this mechanism we can always map any dataset into the bi-unit hypercube and find a straightforward coding for the instances in $2^d$-ary codes.
+What fractal fitting offers us is a potential for a subspace that fits the data more compactly (ie.
+using fewer components) and gives us a good code of lower arity.
 
-\subsection{Evaluating models}
+\begin{figure}[b!]
+  \centering
+  \begin{subfigure}[b]{0.48\textwidth}
+    \includegraphics[width=\textwidth]{../img/sierpinski-codes.pdf}
+    \caption{}
+    \label{fig:sierpinski-codes}
+  \end{subfigure}  
+  \hspace{0.015\textwidth}
+  \begin{subfigure}[b]{0.48\textwidth}
+    \includegraphics[width=\textwidth]{../img/code-construction.pdf}
+    \caption{}
+    \label{fig:code-construction}
+  \end{subfigure}
+
+  \caption{\small (a) Length-three codes on the Sierpinski triangle and the subsets they code for. (b) The construction of a subset from its code.}
+  \label{fig:codes}
+\end{figure}
+
+\subsection*{Evaluating models}
 
@@
 Thus all points will have probability zero under the Sierpinski gasket and its likelihood will be zero.
-Intuitively, the Sierpinski gasket is a good model for this data.
+Intuitively however, the Sierpinski gasket is still a good model for this data.
 
@@
 \[ 
-H'(X, Y) = \max_{x\in X} \min_{y\in Y} d(x, y) 
+H'(X, Y) = \max_{x\in X} \, \min_{y\in Y} \, d(x, y) 
 \]
 \[
-H(X, Y) = \max\left \{H(X, Y), H(Y, X)\right\}
+H(X, Y) = \max\left (H'(X, Y), H'(Y, X)\right )
 \]
@@
 
-The most intuitive way of understanding the Hausdorff distance is as a game where the first player must get from one set to the other in as small a distance as possible, and the second player will choose his starting point to maximize that distance.
+The most intuitive way of understanding the Hausdorff distance is as a game where the first player must get from one set to the other in as small a distance as possible, and the malevolant second player will choose his starting point to maximize that distance.
 
-Hausdorff distance is a common tool in dealing with the fractal inverse problem [@!
-cite].
+Hausdorff distance is a common tool in dealing with the fractal inverse problem \cite{nettleton1994evolutionary}.
 We will use it to evaluate an IFS model by generating a number of points from the model and calculating the Hausdorff distance of that set to the model.
 Since the Hausdorff distance has not been substantiated as a statistical measure, we can only use it as an intermediate indication of model quality.
-To prove the utility of IFS models we will use them to solve three classical machine learning problems (classification, dimension reduction and clustering).
+To prove the utility of IFS models we will test them on classification and dimensionality reduction tasks.
 
-We estimate the uncertainty in the estimate of the Hausdorff distance by repeating the process $50$ times and calculating the sample standard deviation.
+\subsection*{Dimension}
 
-\subsection{Dimension}
-
-The curse of dimensionality is a blanket term for many effects related the analysis of high dimensional data.
+The \textit{curse of dimensionality} is a blanket term for many effects related the analysis of high dimensional data.
 The main cause of these problems is that as dimensionality increases, the volume of a subspace (such as the unit ball) increases exponentially with it.
@@
 
-\subsection{Analogous model: Mixture of Gaussians}
+\subsection*{Analogous model: Mixture of Gaussians}
 
@@
 \]
-Since every multivariate Gaussian can also be described as an Affine transformation of the standard normal distribution (mean at the origin and $\Sigma$ the identity matrix), we can also represent the mixture of Gaussians as $n$ affine transformations and $n$ weights, just as the IFS:
+Since every multivariate Gaussian can also be described as an affine transformation of the standard normal distribution (mean at the origin and $\Sigma$ the identity matrix), we can also represent the mixture of Gaussians as $n$ affine transformations and $n$ weights, just as the IFS:
 \[
@@
 
-\begin{itemize}
-  \item Expectation: Assign each data point a probability distribution over the components of the MoG model proportional to the likelihood of the model given the datapoint.
-  \item Maximization: Estimate the parameters for each component based on the probability weights calculated in the previous step.  
-\end{itemize}
+\begin{description}
+  \item[Expectation] Assign each data point a probability distribution over the components of the MoG model proportional to the likelihood of the model given the datapoint.
+In other words, each component takes a certain amount of `responsibility' for the data point.
+  \item[Maximization] Estimate the parameters for each component based on the probability weights calculated in the previous step.  
+\end{description}
 
@@
 
-\section{Fitting IFS models to data}
+\section*{Fitting IFS models to data}
 
@@
 
-\subsection{The basic algorithm}
+\subsection*{The basic algorithm}
 
@@
 
-\subsection{Improvements}
+\subsection*{Improvements}
 
-\subsubsection{Centering and scaling the data}
+\subsubsection*{Centering and scaling the data}
 
@@
 
-\subsubsection{Using all depths}
+\subsubsection*{Using all depths}
 
@@
 
-\subsubsection{Multiple matches}
+\subsubsection*{Multiple matches}
 
@@
 
-\subsubsection{Weights}
+\subsubsection*{Weights}
 
@@
 
-\subsubsection{Edge cases}
+\subsubsection*{Edge cases}
 
@@
 
-\section{Results}
+\section*{Results}
 
@@
 
-\subsection{Datasets}
+\subsection*{Datasets}
 
@@
 
-\subsection{Model fitting and dimension scaling}
+\subsection*{Model fitting and dimension scaling}
 
@@
 
-\subsection{Classification}
+\subsection*{Classification}
 
-\subsubsection{Bayesian}
+We explore two methods for using IFS models for classification.
+In the first, we train a single model for each class and combine in a straightforward manner using Bayes' rule.
+In the second, we learn a single model for the whole set of points, and use the codes generated by the model to perform classification.
 
-\subsubsection{Single model}
+\subsubsection*{Bayesian}
 
-An alternative way to construct a classifier with IUFS modelling is to learn a single model for the whole dataset, ignoring class labels.
+Given some data set $X = \{x_i\} \subseteq {\mathbb R}^d$ with associated classes $c_i \in \{1, 2, \ldots, m\}$, we split the data by class: $X_c = \{ x \in X \mid c_i = c\}$.
+We train a probability model $p_c$ for each $X_c$ and classify each new point with
+\[
+C(x) = \argmax_{c} p_c(x) p(c)
+\]
+The class prior $p(c)$ is estimated as $\frac{|X_c|}{|X|}$.
+
+\subsubsection*{Single model}
+
+An alternative way to construct a classifier with IFS modelling is to learn a single model for the whole dataset, ignoring class labels.
 For some datasets, the set of all points may present a more coherent and simple distribution than the sets split by class label.
@@
 
-The model provides us with a mapping from the data space (${\mathbb R}^d$) to code space (${1, \ldots, k}^d$).
+The model provides us with a mapping from the data space (${\mathbb R}^d$) to code space ($\{1, \ldots, k\}^d$).
 To turn this model into a classifier we proceed as follows.
 We build a code tree such that each possible code represents a path from the root to the leaf node.
-We then traverse this tree for each point in the training set, maintaining a frequency count fora each class observed at each node.
-Thus, the root node gives us the frequency distribution over the classes for the whole dataset, and each node below it refines that to a distribution over a section of the dataset whose code starts with a given prefix.
+We then traverse this tree for each point in the training set, maintaining a frequency count for each class observed at each node.
+Thus, the root node gives us the frequency distribution over the classes for the whole dataset and each node below it refines that to a distribution over a section of the dataset whose code starts with a given prefix.
 This turns each node into a very basic classifier, which simply returns the class with the highest frequency count.
-The root node will give us the classifier that uses nothing mode than class frequencies and each node below it will use more information about the neighbourhood of the point.
+The root node will give us the classifier that uses nothing more than class frequencies and each node below it will use more information about the neighbourhood of the point.
 
@@
 
-\section{Conclusions}
+\section*{Conclusions}
 

### 2012-11-07 fractal paper/src/fractal.tex
...

@@
 For instance, the use of evolutionary strategies combined with Hausdorff distance as a target function has been shown to be able to reconstruct fractals from data, and to fit models to common datasets.
+These approaches can be found in \cite{bloem2010fractal,nettleton1994evolutionary} and many others.
 
-Nevertheless, the generic nature of these search methods mean that they do not exploit any knowledge of the fractal models.
+The generic nature of these search methods mean that they do not exploit any knowledge of the fractal models.
 We expect that the induction of fractal models can be sped up and made more accurate if we tailor the search algorithm to the model.
@@
 
+\begin{figure}[b!]
+  \centering
+  \begin{subfigure}[b]{0.48\textwidth}
+    \includegraphics[width=\textwidth]{../img/sierpinski-codes.pdf}
+    \caption{}
+    \label{fig:sierpinski-codes}
+  \end{subfigure}  
+  \hspace{0.015\textwidth}
+  \begin{subfigure}[b]{0.48\textwidth}
+    \includegraphics[width=\textwidth]{../img/code-construction.pdf}
+    \caption{}
+    \label{fig:code-construction}
+  \end{subfigure}
+
+  \caption{\small (a) Codes of length three on the Sierpinski triangle and the subsets they code for. (b) The construction of a subset from its code.}
+  \label{fig:codes}
+\end{figure}
+
 Let $(\sigma_i)$ be a sequence with $\sigma_i \in [1, n]$ for some IFS $S$ with $n$ components.
@@
 
-\begin{figure}[b!]
-  \centering
-  \begin{subfigure}[b]{0.48\textwidth}
-    \includegraphics[width=\textwidth]{../img/sierpinski-codes.pdf}
-    \caption{}
-    \label{fig:sierpinski-codes}
-  \end{subfigure}  
-  \hspace{0.015\textwidth}
-  \begin{subfigure}[b]{0.48\textwidth}
-    \includegraphics[width=\textwidth]{../img/code-construction.pdf}
-    \caption{}
-    \label{fig:code-construction}
-  \end{subfigure}
-
-  \caption{\small (a) Length-three codes on the Sierpinski triangle and the subsets they code for. (b) The construction of a subset from its code.}
-  \label{fig:codes}
-\end{figure}
-
 \subsection*{Evaluating models}
@@
 
-We measure the dimension of datasets and data generated from models using Takens' estimator:
+We measure the dimension of datasets and data generated from models using Takens' estimator \cite{takens1985numerical}:
 \[
@@
 We choose $d_{\mbox{max}}$ as the value that produces the lowest Kolmogorov Smirnov statistic with the probability distribution underlying the Takens estimator.
-This method is based on the work for general power laws in @!
-and is detailed in @!
-(to be published).
+This method is based on the work for general power laws in \cite{clauset2009power} and will be detailed in a later publication.
 
@@
 We use the Mixture of Gaussians (MoG) model.
-This model consists of $n$ multivariate Gaussians over ${\mathbb R}^d$ each represented by a mean $\mu \in {\mathbb R}^D$ and a $d \times d$ covariance matrix $\Sigma$.
+This model consists of $n$ multivariate Gaussians over ${\mathbb R}^d$ each represented by a mean $\mu \in {\mathbb R}^d$ and a $d \times d$ covariance matrix $\Sigma$.
 Each component also has a prior $w_i$.
@@
 \[
-p(X) = \sum_{i \in [1,n]} w_i N(x|\mu_i, \Sigma)
+p(x) = \sum_{i \in [1,n]} w_i N(x|\mu_i, \Sigma)
 \]
@@
 \[
-p(X) = \sum_{i \in [1,n]} w_i S_i(N(\mu_i, \Sigma))(X)
+p(x) = \sum_{i \in [1,n]} w_i S_i(N(0, I))(x)
 \]
@@
 
-Some undesired behavior can emerge when the initial model is far from the optimum.
+As with standard EM algorithms, Some undesired behavior can emerge when the initial model is far from the optimum.
 Consider for instance the case where the Sierpinski triangle is the current model and all data is to the top right of the triangle.
@@
 
-The perturbation is done by flattening the model into a vector of values in ${\mathbb R}$ and ading a small random value $N(0, \sigma)$ to each. In our experiments $\sigma$ was set to 0.1.
+The perturbation is done by flattening the model into a vector of values in ${\mathbb R}$ and adding a small random value $N(0, \sigma)$ to each.
+In our experiments $\sigma$ was set to 0.3.
+After perturbation, we check if the model is still contractive, and invert it if not.
+This last step avoids some undersirable looping behavior that occurs when the IFS model is not contractive on average.
+
+\subsubsection*{Use the data as a basis}
+
+[@! This is not yet implemented properly]
+
+When we generate a sample from the model (to a fixed depth, rather than with the chaos game), and when we compute the probability density to a fixed point, we make use of some initial normal distribution.
+As the depth increases the influence of ythe choice of initial distribution increases exponentially.
+For low depths, however, our choice can affect the quality of our model.
+
+To determine our initial model, we examine the situation for depth one.
+In this case, our data points get a conde of one symbol, and the transformations induced are those that optimally map the whole dataset to these subsets.
+The equivalent MOG model fits a normal distribution to each subset.
+For these normal distributions to be the result of a depth one IFS model, we must choose as its initial distribution the MVN fit to the data.
+With this choice, the MOG model and the MVN model are equivalent (so long as they use the same class of transformations).
+
+To summarize, we choose as our initial distribution an MVN fit for the data restricted to our transformation class.
+Since we are using Similitudes, we fit a symmetrical MVN to our data, and use that as the initial distribution.
 

### 2012-12-04 fractal paper/src/fractal.tex
...

@@
 \begin{abstract}
-\noindent We present an algorithm for fitting fractals to data and we investigate the use of fractal models in machine learning.
+\noindent We present an algorithm for fitting IFS fractals to data and we investigate the use of fractal models in machine learning.
 Fractals are mathematical objects that display great detail and complexity despite having simple descriptions.
@@
     \caption{}
+    \label{fig:measures}
+
   \end{subfigure}
@@
 
-This problem was investigated thoroughly in the context of image compression (see for instance \cite{hart1996fractal}), but has received relatively little attention in the area of statistics, and fitting a fractal model to data rather than an image.
+This problem was investigated thoroughly in the context of image compression (see for instance \cite{hart1996fractal}), but has received relatively little attention in the area of statistics.
 In particular the case of data in arbitrary dimensions has been left largely uninvestigated.
@@
 
-The generic nature of these search methods mean that they do not exploit any knowledge of the fractal models.
+The generic nature of these search methods means that they do not exploit any knowledge of the fractal models.
 We expect that the induction of fractal models can be sped up and made more accurate if we tailor the search algorithm to the model.
@@
 Iterated Function Systems are a simple generalization of many classic fractals.
-Consider the Sierpinski Gasket (fig @!).
+Consider the Sierpinski Gasket (fig.
+\ref{fig:sierpinski}).
 As noted before, it consists of three scaled down versions of itself.
@@
 The key result in the theory of iterated function systems is that for these three transformations there is only one set that has this property.
-The three transformations together form a scaling law, and we say the the Sierpinski gasket satisfies the scaling law.
+The three transformations together form a scaling law, and we say the Sierpinski gasket satisfies the scaling law.
 
@@
 Call this function $S$: $S(X) = \bigcup S_i(X)$.
-One of the core results about IFSs is that when we iterate $S$ as $S^{n+1}(X) = S(S^n(X))$, then we get
+Another important result about IFSs is that when we iterate $S$ as $S^{n+1}(X) = S(S^n(X))$ we get
 \[
@@
 \]
-With $X_0$ any non-empty set and $Q$ the unique set that satisfies the scaling law $\{S_i\}$.
+with $X_0$ any non-empty set and $Q$ the unique set that satisfies the scaling law $\{S_i\}$.
 
@@
 
-This type of IFS is what we will use as a formalism for modeling datasets with fractals.
+This type of IFS is what we will use as a formalism for modeling datasets with fractals. See figure\ref{fig:measures} for some examples IFS measures.
 
@@
 
-It should be mentioned that the sequence of points generated this way does not represent a sequence of independent draws.
-Any two successive points in the sequence will satisfy the relation $x_{n+1} = S_i(x_{n})$ for some $i$.
+It should be mentioned that the sequence of points generated this way does not strictly represent a sequence of independent draws.
+We can only say that the natural measure of this process coincides with the IFS.
+However, any two successive points in the sequence will satisfy the relation $x_{n+1} = S_i(x_{n})$ for some $i$.
 This is not true for two independently drawn points.
@@
 \] 
-That is, the sequence of functions maps each $X$ to a unique point $x_\sigma$.
+That is, the sequence of functions maps any non-empty $X$ to a unique point $x_\sigma$.
 In the case of probability measures, every probability measure is mapped to a probability measure that gives the single point $x_\sigma$ full probability mass.
-Figure [@!] illustrates the concept, an Iterated Function System assigns a code to each point on its support.
+Figure \ref{fig:codes} illustrates the concept: an Iterated Function System assigns a code to each point on its support.
 Finite codes map to subsets of the support.
@@
 Finding the endpoint for given code is straightforward, but we will also require the inverse process: given a point $x$ on or near the support of some IFS, we wish to find $\sigma$.
-To this end we start with the standard normal multivariate distribution and apply all maps $S_\sigma$ for all $\sigma$ of a given length $l$ giving us a set of $k^l$ normal distributions (since an affine transformation of a normal distribution is also a normal distribution).
-The $\sigma$ which assigns $x$ the greatest probability is chosen.
+To this end we start with the standard  multivariate normal distribution and apply all maps $S_\sigma$ for all $\sigma$ of a given length $l$ giving us a set of $k^l$ normal distributions (since an affine transformation of a normal distribution is also a normal distribution).
+The $\sigma$ which assigns $x$ the greatest probability density is chosen.
 
@@
 To increase the range of the method, we fall back on the distance from the point to the mean of the distribution.
-If the first method cannot distinguish between codes, we choose the code whose probability distribution has its mean closest to the point.
+If the first method cannot distinguish between codes, we choose the code whose probability distribution has its mean closest to $x$.
 
@@
 \[
-S_0(S_1(S_2(x)))
+S_1(S_2(S_0(x)))
 \]
-so its code $\sigma$ is $(0, 1, 2)$. If we wish to `travel into' the Sierpinski gasket by this code, we must read it right-to-left.
+so its code $\sigma$ is $(1, 2, 0)$. If we wish to `travel into' the Sierpinski gasket by this code, we must read it right-to-left.
 
@@
 \]
- (for infinite sets the $\max$ and $\min$ functions should be replaced with $\sup$ and $\inf$)
+where $d(\cdot, \cdot)$ is the Euclidean distance. For infinite sets the $\max$ and $\min$ functions should be replaced with $\sup$ and $\inf$.
 
@@
 Where $A$ is the set of distances between pairs of data points lower than parameter $d_{\mbox{max}}$.
-We choose $d_{\mbox{max}}$ as the value that produces the lowest Kolmogorov Smirnov statistic with the probability distribution underlying the Takens estimator.
+We choose $d_{\mbox{max}}$ as the value that produces the lowest Kolmogorov-Smirnov statistic with the probability distribution underlying the Takens estimator.
 This method is based on the work for general power laws in \cite{clauset2009power} and will be detailed in a later publication.
@@
 
-[@! did I name the steps correctly?]
-
 \section*{Fitting IFS models to data}
@@
 
-The results described are always for the full algorithm with all improvements, with the exception of the branching mode (which we test as a separate algorithm).
+The results described are always for the full algorithm with all improvements.
 An implementation of only the basic algorithm should be enough to retrieve the Sierpinski gasket (or a crude approximation) from data.
@@
 \item We start with in initial IFS $S$, chosen by one of the methods described below. 
-\item (Expectation) We assign each datapoint a code $\sigma$ of length $d$ (where $d$ is a parameter of the algorithm).
-The codes are provided by the model $S$ as described in section \ref{sec:code-space}.
+\item (Expectation) We assign each datapoint a code $\sigma$ of length $d$ (where $d$ is a parameter of the algorithm). The codes are provided by the model $S$ as described above. 
 \item (Maximization) We know that if a point $x_\sigma$ is mapped using one of our components we have $S_i(x_\sigma) = x_{i\sigma}$.
@@
 Iterated function systems are not very good at approximating data that is weighted far from the origin.
-For this reason we find the affine transformation which maps the data to the range $[0,1]$ in each dimension and puts the mean at the origin.
+For this reason we find the affine transformation which maps the data to the range $[0,1]$ in each dimension.
 
@@
 The algorithm as described here can work at all depths $d$.
-Since each depth may contain information about self similar scaling in the data, we want to consider all depths $\leq d$ rather than just $d$.
+Since each depth may contain information about self similar scaling in the data, we also want to consider all depths less than $d$ rather than just $d$.
 This is easily done by considering codes of all length up to and including $k$.
@@
 
-\subsubsection*{Multiple matches}
-
-It is perfectly possible that we get multiple points per code.
-Indeed, at low depths, it is almost certain.
-In this case we have two sets of points $X_\sigma$ and $X_\tau$ such that the points in the former should map onto the latter.
-But our algorithm for finding the map $S_i$ requires the individual points to be matched to each other.
-If the number of points is low we can simply take the mean of both sets, but if the number of points is high, we risk throwing away a lot of information (in particular about rotation).
-Theoretically each point in the first set should map onto every point in the other but this creates an explosion of possiblities.
-If we pair the points up randomly we introduces artifacts in the induced rotation.
-
-We tackle this problem by finding the MVN for both sets.
-Ideally we would then extend our method for finding maps to paired MVNs rather than paired points, but this is a highly complex problem.
-Instead, we find for each MVN the transformation that transforms the standard distribution ($N(0, I)$) into that MVN.
-We then generate a small amount of points (20 in our experiments) from the standard MVN and map these to both sets.
-We add these 20 point pairs to the list of matched points from which the maps will be calculated.
-
 \subsubsection*{Weights}
@@
 The algorithm as described above does not give us a way to retrieve the component weights.
-For the coding step we already take the weights into accounts, so the only thing we must do is retrieve them from the coded dataset.
+For the coding step we already take the weights into account, so the only thing we must do is retrieve them from the set of codes with which we start the maximization step.
 
@@
 
+\subsubsection*{Multiple matches ($\sigma$)}
+
+It is perfectly possible that we get multiple points per code.
+Indeed, at low depths, it is almost certain.
+In this case we have two sets of points $X_\sigma$ and $X_\tau$ such that the points in the former should map onto the latter.
+But our algorithm for finding the map $S_i$ requires the individual points to be matched to each other.
+If the number of points is low we can simply take the mean of both sets, but if the number of points is high, we risk throwing away a lot of information (in particular about rotation).
+Theoretically each point in the first set should map onto every point in the other but this creates an explosion of possiblities.
+If we pair the points up randomly we introduce artifacts in the induced rotation.
+
+We tackle this problem by finding the MVN for both sets.
+Ideally we would then extend our method for finding maps to paired MVNs rather than paired points, but this is a highly complex problem.
+Instead, we make use of the correspondence between MVNs and affine maps.
+We find for each MVN the transformation that transforms the standard distribution ($N(0, I)$) into that MVN.
+We then generate $|X_\tau|$ `spanning points' from some initial distribution and map these to both sets.
+We add these point pairs to the list of matched points from which the maps will be calculated.
+
+To generate the spanning points we use a spherical MVN with variance $\sigma$.
+This value becomes an important parameter in the algorithm.
+If the spanning points have a high variance, then the rotation becomes a more important part of the found affine map.
+The limiting case with $\sigma \rightarrow 0$ is essentially equal to just using the mean of all points with a given code and getting no rotation information from the mapping between single code pairs.
+
+The behaviour of the algorithm under varying $\sigma$ is somewhat mysterious, but trial-and-error testing suggests that for high values the algorithm is more likely to converge.
+This comes at the cost of some solutions, as with $\sigma = 1.0$, the algorithm is unable to find the two-component Koch curve.
+For all experiments reported in this paper we use $\sigma = 0.1$.
+
 \subsubsection*{Edge cases}
 
-As with standard EM algorithms, Some undesired behavior can emerge when the initial model is far from the optimum.
+As with standard EM algorithms, some undesired behavior can emerge when the initial model is far from the optimum.
 Consider for instance the case where the Sierpinski triangle is the current model and all data is to the top right of the triangle.
 The symbol 2 would not occur in any of the codes and we would have no basis to find $S_2$ in the maximization step.
-If this occurs, we will `split' one of the components that could be found in two (distributing the weight equally) and perturb each component slightly to ensure that the components differ.
+If this occurs, we will `split' one of the components that could be found  (distributing the weight equally) and perturb each component slightly to ensure that the components differ.
 The idea behind this approach is simple.
@@
 
-The perturbation is done by flattening the model into a vector of values in ${\mathbb R}$ and adding a small random value $N(0, \sigma)$ to each.
-In our experiments $\sigma$ was set to 0.3.
-After perturbation, we check if the model is still contractive, and invert it if not.
-This last step avoids some undersirable looping behavior that occurs when the IFS model is not contractive on average.
+The perturbation is done by flattening the model into a vector of values in ${\mathbb R}$ and adding a small random value $N(0, \pi)$ to each. In our experiments $\pi$ was set to 0.3.
 
@@
 
-[@! This is not yet implemented properly]
-
 When we generate a sample from the model (to a fixed depth, rather than with the chaos game), and when we compute the probability density to a fixed point, we make use of some initial normal distribution.
-As the depth increases the influence of ythe choice of initial distribution increases exponentially.
+As the depth increases the influence of the choice of initial distribution increases exponentially.
 For low depths, however, our choice can affect the quality of our model.
 
-To determine our initial model, we examine the situation for depth one.
-In this case, our data points get a conde of one symbol, and the transformations induced are those that optimally map the whole dataset to these subsets.
+To determine our what our initial model should be, we examine the situation for depth one.
+In this case, our data points get a code of one symbol, and the transformations induced are those that optimally map the whole dataset to these subsets.
 The equivalent MOG model fits a normal distribution to each subset.
@@
 
+\subsection*{Choosing the initial model}
+
+The choice of initial model is crucial for achieving good performance
+
+\subsection*{Convergence}
+
 \section*{Results}
 
-There are three expected benefits to fractal modeling that we wish to explore here.
-First, we expect that the models will scale their dimension to fit that of the data (whereas our comparison model will be unable to do so).
-To test this we will fit models to various datasets and estimate their dimension.
+We perform a number of experiments to give an impression of the behavior of IFSs as statistical models, and the EM algorithm as a learning method.
 
-Second, we expect that the IFS distributions make good fits.
-Since we cannot test this directly using likelihood, we will embed the models in a classifier and test its performance on various data sets.
+First, we show a selection of results on 2D datasets to give a visual impression of the behavior of the algorithm.
+These images represent the value of the EM algorithm over earlier approaches, as methods based on evolutionary algorithms commonly fail to induce fractals from a fractal dataset with this speed or accuracy.
 
-Finally, we expect that the codes described in section @!
-are of actual use outside of the fitting procedure.
-These codes describe a metric space that is much more compact\footnote{in an informal sense} than the Euclidean space in which the data is embedded.
-Furthermore, the transformation to strings promises the use of a whole family of techniques that was not previously available.
-
-Finally we give some indication of the time performance of the algorithm.
+We then investigate the use of IFS models on higher dimensional data.
+To attempt to visualize the way an iterated function system models a high dimensional dataset, we use a four component model so that we can map the data back to two dimensions and plot it.
 
@@
 
-[@! EXPERIMENT Fit models to data ]
+\begin{table}[p]
+\begin{tabular}{c c c c c}
+\hline
+data & initial & & & best model \\
+\hline
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/data.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/generation0000.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/generation0001.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/generation0010.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/generation0032.png}
+\\
+ & & 6 & 48 & 83 \\
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/data.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/generation0000.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/generation0003.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/generation0020.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/generation0079.png}
+\\    
+ &  & 3 & 20 & 79 \\
+	\includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/data.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/generation0000.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/generation0008.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/generation0022.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/generation0078.png}
+\\    
+ &  & 8 & 22 & 78 \\
+	\includegraphics[width=0.2\textwidth]{../img/runs2d/ball/data.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/ball/generation0000.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/ball/generation0003.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/ball/generation0024.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/ball/generation0072.png}
+\\   
+ &   & 3  & 24 & 72 \\
+	\includegraphics[width=0.2\textwidth]{../img/runs2d/coast/data.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/coast/generation0000.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/coast/generation0008.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/coast/generation0020.png}
+& 
+    \includegraphics[width=0.2\textwidth]{../img/runs2d/coast/generation0069.png}
+\\
+ &  & 8 & 20 & 69 \\
+\hline
+
+\end{tabular}
+\caption{Selected training steps of the algorithm for various data sets.
+All models were trained with a dataset of 100000 points, sampling 300 per iteration for 100 iterations.
+Model depth was 6 for all runs.
+1000 points were sampled from each model and from the dataset to estimate the Hausdorff distance (and determine the `best model').
+The blue squares indicate the transformations (of the red square) that make up the model.
+The green bar indicates component weight.
+By row: (1) The sierpinski triangle, $\sigma = 0.1$ (2) The sierpinski triangle with unequal weights.
+$\sigma = 0.1$ (3) The 2-component koch curve.
+$\sigma = 0.001$ (4) A uniform distribution over the bi-unit ball (3 components, $\sigma = 0.1$) (5) Points sampled from a small stretch of coast line (3 components, $sigma = 0.1$).
+}
+\label{tablelabel}
+\end{table}
 
@@
 
-\begin{table}
-\begin{tabular}{l | r r r r | r |}
-\hline
- & ifs1 & ifs3 & ifs7 & mog \\
-\hline
-census & 0.02 (0.04, 80) & & & & \\
-forest & & & & & \\
-abalone & & & & & \\
-\hline
-
-\end{tabular}
-\caption{Hausdorff distances for models trained against various datasets.}
-\label{tablelabel}
-\end{table}
-
-\begin{table}
-\begin{tabular}{l | r r r r r | r |}
-\hline
-  & dim & ifs1 & ifs3 & ifs5 & ifs7 & mog \\
-\hline
-census & 3.2 (0.2) & 2.9 (0.45) & & & & \\
-forest & & & & & & \\
-abalone & & & & & & \\
-\hline
-
-\end{tabular}
-\caption{Hausdorff distances for models trained against various datasets.}
-\label{tablelabel}
-\end{table}
-
 \begin{figure}[htb]
 \centering
-\includegraphics[width=\textwidth]{../img/out-table.pdf}
+
+  \begin{subfigure}[b]{0.66\textwidth}
+	\includegraphics[width=\textwidth]{../img/dimension/out-table.pdf}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.33\textwidth}
+    \begin{tabular}{c c }
+    data & model \\
+	\hline
+		\includegraphics[width=0.5\textwidth]{../img/dimension/data-henon.png} & 
+		\includegraphics[width=0.5\textwidth]{../img/dimension/model-henon.png} \\
+		\includegraphics[width=0.5\textwidth]{../img/dimension/data-ikeda.png} & 
+		\includegraphics[width=0.5\textwidth]{../img/dimension/model-henon.png} \\
+		\includegraphics[width=0.5\textwidth]{../img/dimension/data-henon.png} & 
+		\includegraphics[width=0.5\textwidth]{../img/dimension/model-henon.png} \\
+	\end{tabular}
+  \end{subfigure}
   \caption{Dimensional scaling. The marker size and color both indicate the quality of the model fit by Hausdorff distance. }
@@
 
-\begin{table}
+\begin{table}[h]
 \begin{tabular}{| l | r l r l | r l r l  r l r l |}
@@
 abalone & & & & & & & 0.47 & \color{gray}(0.008) & 0.47 & \color{gray}(0.006) & 0.48 & \color{gray}(0.01) \\
-mnist   & & & & & & & & & & & & \\
+mnist   & & & & & & & 0.09 & & 0.05 & & 0.20 & \\
 \hline
@@
 
-Finally, there is the task of learning random fractals.
-The random fractal model is an extension of the Iterated Function System which produces not one single set but a distribution over all sets.
-The dataset for which to learn a random fractal would not consist of one list of vectors, but of a list of list of vector.
+\subsubsection*{Other fractals}
+
+It is important to note that there are many other families of fractals besides iterated function systems.
+They are not covered by our method, and cannot be properly represented by Iterated Function Systems.
+Examples include Julia sets and dynamical systems.
+
+There is one family of fractals in particular that would be very interesting to explore in light of this research, and that is the family of random fractals.
+When we think of natural phenomena that seem to resemble iterated function systems---coast lines, trees, the veins of a leaf---there is always a combination of fractal scaling and some random component.
+The framework most likely to capture these phenomena is that of random Iterated Function Systems, as described in \cite{hutchinson2000deterministic}.
+
+The random IFS model is an extension of the Iterated Function System which produces not one single set but a distribution over all sets.
+The dataset for which to learn a random fractal would not consist of one list of vectors, but of a list of lists of vectors.
 For instance, a collection of timeseries preceding a critical event, or a collection of images of snowflakes.

### 2012-12-07 fractal paper/src/fractal.tex
...

@@
 
-\begin{figure}[htb]
+\begin{figure}[t]
 \centering
 
-  \begin{subfigure}[b]{0.66\textwidth}
+  \begin{subfigure}{0.66\textwidth}
 	\includegraphics[width=\textwidth]{../img/dimension/out-table.pdf}
   \end{subfigure}
-  \begin{subfigure}[b]{0.33\textwidth}
+  \begin{subfigure}{0.33\textwidth}
     \begin{tabular}{c c }
     data & model \\
-	\hline
 		\includegraphics[width=0.5\textwidth]{../img/dimension/data-henon.png} & 
@@
   \end{subfigure}
-  \caption{Dimensional scaling. The marker size and color both indicate the quality of the model fit by Hausdorff distance. }
+  \caption{To investigate the dimensional scaling of IFS models, we train them on three generated datasets.
+The datasets were chosen for a) their low dimension so that we can acurately measure dimension, b) their synthetic nature so that we hav an analytic value of their dimension and c) their lack of an accurate IFS approximation, so that the models will not fit perfectly.
+The datasets are shown on the left with their respective IFS model.
+They are, from top to bottom: the H\'enon attractor, the Ikeda attractor, and the R\"ossler attractor.}
   \label{fig:dim_scale}

### 2012-12-14 fractal paper/src/fractal.tex
...

@@
 This is not true for two independently drawn points.
-For our experiments we generate a sequence and then shuffle the data to remove this dependency where it is undesirable.
 
@@
 For multiple points the process can be repeated.
-This process adds a factor of $n$ to the time complexity of the chaos game, so it is not commonly used.
+This process adds a factor of $n$ to the time complexity of the chaos game, but it has the advantage of drawing indepent points and being able to draw points from a model with a fixed depths.
+Though the process converges to the IFS with the depth, we will often learn models with a fixed depth, so drawing from a fixed-depth model keeps the various views of a single model consistent with one another.
 
@@
 To increase the range of the method, we fall back on the distance from the point to the mean of the distribution.
-If the first method cannot distinguish between codes, we choose the code whose probability distribution has its mean closest to $x$.
+If the first method cannot distinguish between codes, we choose the code whose probability distribution has its mean closest to $x$ (this is equivalent to the approximation method described in section \ref{sec:density}).
 
@@
 
-\subsection*{Evaluating models}
+\subsection*{Calculating density}
+\label{sec:density}
 
-The standard way of evaluating the fit of a model to a dataset is the likelihood of the model given the data.
+The probability density of a given point $x$ under a given IFS model $S$ is straighforward to calculate. For a given depth, we start with a basis MVN $N(x)$ , and calculate the following sum:
+\[
+\sum_{\sigma \in \{1, \ldots, k\}^l} w_\sigma S(N)(x)
+\]
+
+That is, we transform the basis function by $S_\sigma$, the composition of the components of $S$ in the order indicated by $\sigma$ (the endpoint distribution), we weight this value by the $w_\sigma = \prod_{i \in \sigma} w_i$ (the endpoint prior) and sum over all possible $\sigma$.
+
+Clearly, this succession will quickly foce each endpoint distribution to a single point, and only points near one of the models' end points will have a probability that is distinguishable from zero in standard 64-bit floating point representation.
+
+To deal with this situation, we also use an approximated density, which can be used when the correct density approximation yields zero.
+Of course, we must be careful only ever to compare two correct densities, or two approximated densities, but never to mix the two.
+
+For the approximated density, we only visit the end points $S_\sigma(\mathbf{0})$ of the IFS model, forgetting the endpoint distributions.
+We place an MVN at this endpoint with a variance of $\rho^l$.
+We then sum the densities under these distributions, weighted by the endpoint priors $w_\sigma$.
+So that the approxiamted density becomes:
+
+\[
+\sum_{\sigma \in \{1,\ldots,k\}^l} w_\sigma N(x | S_\sigma(\mathbf{0}), I\rho^l)
+\]
+
+For our experiments we use $\rho = 0.9$.
+
+% The standard way of evaluating the fit of a model to a dataset is the likelihood of the model given the data.
 This is often used in optimization algorithms for fitting probability models to data.
@@
 
-Consider a dataset drawn from the Sierpinski gasket with a minute amount of two dimensional noise added.
+% Consider a dataset drawn from the Sierpinski gasket with a minute amount of two dimensional noise added.
 Since the Sierpinski gasket has dimension less than two, the probability that the point will still fall within the Sierpinski gasket will be zero.
@@
 
-For this reason we will often use the Hausdorff distance to evaluate our models. The Hausdorff distance  $H(X, Y)$ between two sets $X$ and $Y$ is defined as follows:
-\[ 
-H'(X, Y) = \max_{x\in X} \, \min_{y\in Y} \, d(x, y) 
-\]
-\[
-H(X, Y) = \max\left (H'(X, Y), H'(Y, X)\right )
-\]
-where $d(\cdot, \cdot)$ is the Euclidean distance. For infinite sets the $\max$ and $\min$ functions should be replaced with $\sup$ and $\inf$.
+% For this reason we will often use the Hausdorff distance to evaluate our models. The Hausdorff distance  $H(X, Y)$ between two sets $X$ and $Y$ is defined as follows:
+%\[ 
+%H'(X, Y) = \max_{x\in X} \, \min_{y\in Y} \, d(x, y) 
+%\]
+%\[
+%H(X, Y) = \max\left (H'(X, Y), H'(Y, X)\right )
+%\]
+% where $d(\cdot, \cdot)$ is the Euclidean distance. For infinite sets the $\max$ and $\min$ functions should be replaced with $\sup$ and $\inf$.
 
-The most intuitive way of understanding the Hausdorff distance is as a game where the first player must get from one set to the other in as small a distance as possible, and the malevolant second player will choose his starting point to maximize that distance.
+% The most intuitive way of understanding the Hausdorff distance is as a game where the first player must get from one set to the other in as small a distance as possible, and the malevolant second player will choose his starting point to maximize that distance.
 
-Hausdorff distance is a common tool in dealing with the fractal inverse problem \cite{nettleton1994evolutionary}.
+% Hausdorff distance is a common tool in dealing with the fractal inverse problem \cite{nettleton1994evolutionary}.
 We will use it to evaluate an IFS model by generating a number of points from the model and calculating the Hausdorff distance of that set to the model.
@@
 
-The solution to many of these problems is that while the data has a high-dimensional representation, it has a low intrinsic dimension.
+The solution to many of these problems is that while the data has a high-dimensional representation, it has a low intrinsic dimension ('The blessing of self-similarity').
 That is, it can be mapped to a low dimensional space without any loss of information.
@@
 \begin{itemize}
-\item We start with in initial IFS $S$, chosen by one of the methods described below. 
+\item We start with an initial IFS $S$, chosen by one of the methods described below. 
 \item (Expectation) We assign each datapoint a code $\sigma$ of length $d$ (where $d$ is a parameter of the algorithm). The codes are provided by the model $S$ as described above. 
@@
 
-We will often use a subsample of the data, to allow us to increase the depth or number of components. In such cases, we reample at each EM iteration.
+We will often use a subsample of the data, for performance reasons. In such cases, we resample for each EM iteration.
 
@@
 
-It is perfectly possible that we get multiple points per code.
+It is perfectly possible that we get multiple data points which are assigned the same code.
 Indeed, at low depths, it is almost certain.
@@
 For all experiments reported in this paper we use $\sigma = 0.1$.
+[@!
+No we don't]
 
@@
 
-The choice of initial model is crucial for achieving good performance
+The choice of initial model is crucial for achieving good performance. The following approaches showed promise in our tests:
+
+\begin{description}
+  \item[points] For a model with $k$ components, choose $k$ points from the training data at random.
+For point $x$, find the transformation which contracts all points to $x$ with a contraction ratio of $0.1$.
+  \item[sphere] Choose $k$ random points on the bi-unit sphere in ${\mathbb R}^d$. Continue as with the \textbf{points} method.
+  \item[spread] Search (using for instance evolutionary strategies) for $k$ points on the bi-unit sphere so that the sum of their distances is maximized. Continue as before.
+  \item[identity] Sample $k$ small perturbations of the identity transformation. Invert if the transformation is not contractive. 
+\end{description}
+
+Tests on our datasets show that \textbf{spread} and \textbf{sphere} tend to perform best. Since \textbf{sphere} is the simplest to implement  we use that in all reported results.
 
@@
   \caption{To investigate the dimensional scaling of IFS models, we train them on three generated datasets.
-The datasets were chosen for a) their low dimension so that we can acurately measure dimension, b) their synthetic nature so that we hav an analytic value of their dimension and c) their lack of an accurate IFS approximation, so that the models will not fit perfectly.
+The datasets were chosen for a) their low intrinsic dimension, which is easy to measure b) their synthetic nature so that we have an analytic value of their dimension and c) their lack of an accurate IFS approximation, so that the models will not fit perfectly.
 The datasets are shown on the left with their respective IFS model.

### 2013-02-08 fractal paper/src/fractal.old.tex
...

--- 2012-12-14 fractal paper/src/fractal.tex
@@
 
-This type of IFS is what we will use as a formalism for modeling datasets with fractals. See figure\ref{fig:measures} for some examples IFS measures.
+This is the type of IFS model we will attempt to learn from data. See figure \ref{fig:measures} for some examples IFS measures.
 
@@
 
+The single model classifier also gives us a good baseline to evaluate how well the model fitting works on natural datasets.
+As noted before, for any dimension $d$ we can define an IFS model in $2^d$ components which defines a uniform distribution of the $d$-dimensional hypercube.
+For example, in two dimensions we can start with a square and use the four transformations that map the square onto each of its quadrants.
+The codes assigned by this model are strongly related to basic spatial indexing techniques, and can be calculated very efficiently.
+
+The single classifier described above, when used with such codes, classifies based on histograms over a simple binning of the feature space (averaging over various levels of coarseness).
+It does not perform any dimensionality reduction and very little generalization.
+We will refer to this method as a spatial index classifier.
+
+If we wish to assert that the EM algorithm for IFS induction results in a meaningful subspace, the single model classifier with an optimized model should perform significantly better than the spatial index classifier.
+
 \begin{table}[h]

### 2013-03-10 Benelearn/src/benelearn.tex
...

--- 2012-12-07 fractal paper/src/fractal.tex
@@
 
-\title{Fractal Modeling for Machine Learning}
+\title{Compression-based inference on graph data}
 \date{\today}
@@
 \begin{abstract}
-\noindent We present an algorithm for fitting IFS fractals to data and we investigate the use of fractal models in machine learning.
-Fractals are mathematical objects that display great detail and complexity despite having simple descriptions.
-It has been suggested that fractals model many natural and other real-world structures better than the shapes of classical geometry \cite{mandelbrot1982fractal}.
+\noindent 
 \end{abstract}
 
-\begin{figure}[b]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \includegraphics[width=\textwidth]{../img/sierpinski.pdf}
-    \caption{}
-    \label{fig:sierpinski}
-  \end{subfigure}  
-  \hspace{0.03\textwidth}
-  \begin{subfigure}[b]{0.3\textwidth}
-   \includegraphics[width=\textwidth]{../img/sierpinski-evo.pdf}
-    \caption{}
-  \end{subfigure}
-  \hspace{0.03\textwidth}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \includegraphics[width=0.45\textwidth]{../img/sierpinski-off.png}
-    \includegraphics[width=0.45\textwidth]{../img/random013.png} \\
-    \includegraphics[width=0.45\textwidth]{../img/random010.png}
-    \includegraphics[width=0.45\textwidth]{../img/mvn1k.png}
-    \caption{}
-    \label{fig:measures}
-
-  \end{subfigure}
-  \caption{\small (a) The Sierpinski triangle, a simple example fractal. (b) Constructing \textit{(a)} from an arbitrary initial image. (c) Examples of IFS probability distributions.}
-  \label{fig:examples}
-\end{figure}
-
-Fractals are mathematical objects with certain remarkable properties. What constitutes a fractal is not precisely defined\footnotemark, but the following properties are common:
-
-\begin{description}
-  \item[Self similarity] A part is a scaled down copy of the whole.
-See for instance the Sierpinski triangle (fig.
-\ref{fig:sierpinski}).
-It consists of three triangular shapes which are scaled down copies of itself.
-  \item[Infinitely small detail] 'Zooming in' reveals ever finer detail.
-In the case of the Sierpinski triangle we will see the same shape recurring again and again, but fractals like the Mandelbrot set reveal a great variety of shapes.
-  \item[Non-integer dimension] The Sierpinski triangle, for example, has a dimension of $1.58$.
-This is a counter-intuitive notion, that is not simple to explain.
-We refer the reader to \cite{schroeder2009fractals} for an intuitive explanation.
-\end{description}
-
-The earliest examples of fractals were described around the beginning of the twentieth century as counterexamples to conjectures in planar geometry and analysis and mostly considered pathological.
-It was not until the 1970s that they began to be seen as a single and useful family of mathematical objects that could accurately describe many physical phenomena.
-Mandelbrot put it as follows in the \emph{The Fractal Geometry of Nature} \cite{mandelbrot1982fractal}:
-
-\begin{quotation}
-\small
-\noindent Clouds are not spheres, mountains are not cones, coastlines are not circles, and bark is not smooth, nor does lightning travel in a straight line.
-\end{quotation}
-
-Fractal geometry has been used in many fields, including physics \cite{mandelbrot1984fractals}, geology \cite{cheng1997multifractal}, biology \cite{goldberger1992fractal} and economics \cite{turiel2003multifractal}.
-The first connection between fractals and the natural world was in the form of tangible objects such as coast lines, trees or bronchia, but the more hidden aspects of reality were soon shown to have fractal properties as well.
-The path traced out by the Brownian motion of a particle, the price chart for a commodity or the level of floods in a river delta have all been studied as fractals to great effect.
-
-The main problem with fractal analysis has always been the difficulty of finding a fractal model for a given set of data.
-The assumption that data contains fractal properties by itself already allows us to calculate certain statistics (such as its dimension), but the various families of fractals that have been proposed are difficult to fit to data.
-This is called the \emph{inverse problem}.
-Instead of generating an image or dataset from a given mathematical description of a fractal, we are asked to consider an image or dataset and find a good fractal model for it.
-
-This problem was investigated thoroughly in the context of image compression (see for instance \cite{hart1996fractal}), but has received relatively little attention in the area of statistics.
-In particular the case of data in arbitrary dimensions has been left largely uninvestigated.
-
-\footnotetext{Mandelbrot originally defined it as a set whose topological dimension differs from its Hausdorff dimension, but retracted this definition, stating that he preferred the word to be not precisely defined.}
-
-\subsection*{Previous research}
-
-The most fruitful algorithms for the induction of fractal models so far have been partitioned iterated function systems and evolutionary algorithms.
-
-Partitioned iterated function systems \cite{hart1996fractal} are a subset of iterated function systems (described in the next section) designed for fast image compression by fractal models.
-The compression algorithm segments an image in to two tilings of different size, and then attempts to map the larger tiles to the smaller tiles.
-The resulting set of contraction mappings can be iterated to approximate the original image.
-The approach works well for two dimensional data stored in binned form, but does not translate well to higher dimensions.
-
-Another approach is to represent a fractal model as a point in some space (eg.
-a vector or a bitstring) and to search the space for a model which approximates the data by some target function using generic search algorithms.
-Fractal models tend to be difficult to optimize for and only specific methods show promising results.
-For instance, the use of evolutionary strategies combined with Hausdorff distance as a target function has been shown to be able to reconstruct fractals from data, and to fit models to common datasets.
-These approaches can be found in \cite{bloem2010fractal,nettleton1994evolutionary} and many others.
-
-The generic nature of these search methods means that they do not exploit any knowledge of the fractal models.
-We expect that the induction of fractal models can be sped up and made more accurate if we tailor the search algorithm to the model.
-
-The investigation of iterated function systems has yielded some results that are promising, but have not resulted in an algorithm for the induction of fractal models in arbitrary dimensions.
-Examples include the method of moments \cite{rinaldo1994inverse}, similarity hashing \cite{hart1997similarity} and wavelet analysis \cite{struzik1996coastline}.
-
-\section*{Methods}
-
-In this paper we focus on numerical data in an arbitrary number of dimensions.
-Each instance in our dataset is represented by a fixed number of numerical features.
-Call our dataset $X$, our instance $x \in X$ and $x \in {\mathbb R}^d$.
-We will assume that our dataset represents iid samples from a probability distribution over ${\mathbb R}^d$ which we will attempt to model.
-
-\subsection*{Iterated Function Systems}
-
-If we are to fit fractal models to data then we require a family of fractals which is easy to parametrize, and likely to encompass a large number of organic and useful patterns.
-We will use the formalism of \emph{Iterated Function Systems} (IFSs) for this purpose.
-
-Since the aim of this paper is to present our induction algorithm and to introduce IFS models to a machine learning context, we will provide only an informal treatment of the properties of IFSs that are already well established.
-For a more rigorous treatment, we recommend \cite{hutchinson2000deterministic} and the references therein.
-
-Iterated Function Systems are a simple generalization of many classic fractals.
-Consider the Sierpinski Gasket (fig.
-\ref{fig:sierpinski}).
-As noted before, it consists of three scaled down versions of itself.
-In mathematical terms, if we view the Sierpinski gasket as a set in ${\mathbb R}^2$ and take the union of three specific transformations applied to it, we will get the Sierpinski gasket again.
-The key result in the theory of iterated function systems is that for these three transformations there is only one set that has this property.
-The three transformations together form a scaling law, and we say the Sierpinski gasket satisfies the scaling law.
-
-We will use the following notation.
-Let $S_1$, $S_2$, \ldots, $S_k$ be a set of contractive transformations on ${\mathbb R}^d$.
-(We will later limit the family of allowable functions, but from a mathematical point of view we only require them to be contractive).
-There exists a single set $Q$ that satisfies the following property:
-\begin{align}
-Q = S_1(Q) \cup S_2(Q) \cup \ldots \cup S_k(Q) \label{scaling law}
-\end{align}
-
-We call $\{S_i\}$ a \textit{scaling law}, and we say that $Q$ satisfies it.
-
-If we apply the transformations $S_i$ to a set $X$ that doesn't satisfy the scaling law (and take the union of the results), we get a different set $X'$.
-Call this function $S$: $S(X) = \bigcup S_i(X)$.
-Another important result about IFSs is that when we iterate $S$ as $S^{n+1}(X) = S(S^n(X))$ we get
-\[
-\lim_{n \rightarrow \infty} S^n(X_0) = Q
-\]
-with $X_0$ any non-empty set and $Q$ the unique set that satisfies the scaling law $\{S_i\}$.
-
-Of course, in statistics, we are interested in finding probability distributions rather than sets.
-The IFS formalism can be extended to measures in a natural way.
-First, we must extend the concept of a map to measures.
-Let $p$ be some probability measure over ${\mathbb R}^d$ and $F:{\mathbb R}^d \rightarrow {\mathbb R}^d$ be some map on ${\mathbb R}^d$.
-Then we define the mapping of $p$ by $F$ as:
-\[
-(F(p))(X) = v(F^{-1}(X))
-\]
-So if $p$ is a probability measure, then $F(p)$ is also a probability measure and $F(p)$ assigns $X$ the same probability as $p$ would assign to $F^{-1}(X)$, where $F^{-1}$ is the inverse of $F$.
-
-Our IFS over probability measures consists, as before, of $k$ transformations, but here we also introduce $k$ weights $w_i$ with the constraint that they sum to one.
-Let $q$ be a probability distribution and $S = \{(S_i, w_i)\}$ be a scaling law.
-We say that $q$ satifies $S$ iff:
-\[
-q(X) = \sum_{i\in[1, k]} w_i S_i(q)(X)
-\]
-
-As before q is uniquely defined by $S$. If we apply the transformations and weights to a different probability distribution iteratively, we get a sequence which converges to $q$:
-\begin{align*}
-p_{n+1} &= \sum w_i S_i(p_n) \\
-\lim_{n \rightarrow \infty} p_n &= q 
-\end{align*}
-
-This type of IFS is what we will use as a formalism for modeling datasets with fractals. See figure\ref{fig:measures} for some examples IFS measures.
-
-\subsection*{Representing IFSs}
-
-As discussed, an IFS is completely determined by its component transformations and their weights.
-The definition does not specify any kind of transformation, so to make fitting IFSs to data feasible we must fix a family of transformations which we can search.
-For the purposes of this paper we will use \emph{similarities} (also known as \emph{similitudes}): maps on ${\mathbb R}^d$ consisting of a translation, a rotation and a uniform scaling.
-
-The method described below can easily be extended to other families of transformations such as affine transformations or neural networks.
-The only requirement is that we must have some way of finding a transformation $F$ within this family given two ordered point sets $\{x_i\}, \{y_i\} \subseteq {\mathbb R}^d$ so that the error is minimized.
-For similarities there is a direct and optimal method.
-\cite{umeyama1991least}
-
-\subsection*{Drawing points from an IFS distribution}
-
-For various reasons, it can be useful to generate a set of points drawn from an IFS probability distribution.
-The most common method to achieve this is known as the \emph{chaos game}.
-The algorithm is simple: we start with some initial point $x_0$ we choose a random component $S_i$ from the IFS (so that $S_i$ is chosen with probability $w_i$) and and let $x_1 = S_i(x_0)$.
-We iterate this process and discard the first 50 or so points.
-The rest of the sequence $x_n$ can be used as a set of points drawn from the IFS.
-
-It should be mentioned that the sequence of points generated this way does not strictly represent a sequence of independent draws.
-We can only say that the natural measure of this process coincides with the IFS.
-However, any two successive points in the sequence will satisfy the relation $x_{n+1} = S_i(x_{n})$ for some $i$.
-This is not true for two independently drawn points.
-For our experiments we generate a sequence and then shuffle the data to remove this dependency where it is undesirable.
-
-A more correct (and more expensive) way to generate random data from an IFS is to start with a draw from some initial distribution (eg.
-a multivariate normal) and apply $n$ transformations randomly chosen from $S$.
-The distribution of the resulting random variable will converge to $S$ with $n$.
-For multiple points the process can be repeated.
-This process adds a factor of $n$ to the time complexity of the chaos game, so it is not commonly used.
-
-\subsection*{Code space}
-\label{sec:code-space}
-
-\begin{figure}[b!]
-  \centering
-  \begin{subfigure}[b]{0.48\textwidth}
-    \includegraphics[width=\textwidth]{../img/sierpinski-codes.pdf}
-    \caption{}
-    \label{fig:sierpinski-codes}
-  \end{subfigure}  
-  \hspace{0.015\textwidth}
-  \begin{subfigure}[b]{0.48\textwidth}
-    \includegraphics[width=\textwidth]{../img/code-construction.pdf}
-    \caption{}
-    \label{fig:code-construction}
-  \end{subfigure}
-
-  \caption{\small (a) Codes of length three on the Sierpinski triangle and the subsets they code for. (b) The construction of a subset from its code.}
-  \label{fig:codes}
-\end{figure}
-
-Let $(\sigma_i)$ be a sequence with $\sigma_i \in [1, n]$ for some IFS $S$ with $n$ components.
-We can compose the components of the IFS acording to the elements of $\sigma$.
-Call this function $S_\sigma$:
-\[
-S_\sigma = S_{\sigma_1} \circ S_{\sigma_2} \circ \ldots \circ S_{\sigma_m}
-\]
-
-Since each $S_i$ is contractive, we can let the length of $\sigma$ go to infinity and get 
-\[
-S_\sigma(X) = x_\sigma  
-\] 
-That is, the sequence of functions maps any non-empty $X$ to a unique point $x_\sigma$.
-In the case of probability measures, every probability measure is mapped to a probability measure that gives the single point $x_\sigma$ full probability mass.
-Figure \ref{fig:codes} illustrates the concept: an Iterated Function System assigns a code to each point on its support.
-Finite codes map to subsets of the support.
-Since the size of the subsets decreases exponentially we can talk about the endpoint of a finite code without losing much accuracy.
-We will take the point to which the origin is mapped by $S_\sigma$ as the canonical choice of the endpoint of finite $\sigma$.
-
-Finding the endpoint for given code is straightforward, but we will also require the inverse process: given a point $x$ on or near the support of some IFS, we wish to find $\sigma$.
-To this end we start with the standard  multivariate normal distribution and apply all maps $S_\sigma$ for all $\sigma$ of a given length $l$ giving us a set of $k^l$ normal distributions (since an affine transformation of a normal distribution is also a normal distribution).
-The $\sigma$ which assigns $x$ the greatest probability density is chosen.
-
-Since the variance of the distribution drops off exponentially as the length $l$ is increased, this method will fail if points are too far off the support of the IFS (even if logarithmic probabilities are used).
-To increase the range of the method, we fall back on the distance from the point to the mean of the distribution.
-If the first method cannot distinguish between codes, we choose the code whose probability distribution has its mean closest to $x$.
-
-A word of caution on codes is appropriate: the ordering is often the reverse of what our intuition suggests.
-If we number the three components of the Sierpinski triangle clockwise from the top 0, 1 and 2, we might number a small section of the triangle by first moving into the top subtriangle, then into the left subtriangle (2) of that and then into the right subtriangle (1) of that.
-Intuitively, this is the most natural way of coding sections of the Sierpinski triangle.
-
-If we wish to map the whole triangle to this subset, we must reverse the codes: the transformation that maps to the smallest subsection must be applied first.
-The sequence of transformations which will map the whole triangle into the required subspace is:
-\[
-S_1(S_2(S_0(x)))
-\]
-so its code $\sigma$ is $(1, 2, 0)$. If we wish to `travel into' the Sierpinski gasket by this code, we must read it right-to-left.
-
-Finally, we wish to note that a line segment, the square, the cube and higher dimensional analogues can be represented as simple IFSs of $2^d$ components.
-For instance, for the unit line, we map the whole onto its lower and upper half with two components.
-The codes of this IFS are binary strings and the subsets they code are the diadic intervals.
-For the square, we can map the whole onto each one of its quadrants and get quaternary codes.
-Using this mechanism we can always map any dataset into the bi-unit hypercube and find a straightforward coding for the instances in $2^d$-ary codes.
-What fractal fitting offers us is a potential for a subspace that fits the data more compactly (ie.
-using fewer components) and gives us a good code of lower arity.
-
-\subsection*{Evaluating models}
-
-The standard way of evaluating the fit of a model to a dataset is the likelihood of the model given the data.
-This is often used in optimization algorithms for fitting probability models to data.
-In the case of fractal models, the likelihood function has an important drawback.
-
-Consider a dataset drawn from the Sierpinski gasket with a minute amount of two dimensional noise added.
-Since the Sierpinski gasket has dimension less than two, the probability that the point will still fall within the Sierpinski gasket will be zero.
-Thus all points will have probability zero under the Sierpinski gasket and its likelihood will be zero.
-Intuitively however, the Sierpinski gasket is still a good model for this data.
-
-For this reason we will often use the Hausdorff distance to evaluate our models. The Hausdorff distance  $H(X, Y)$ between two sets $X$ and $Y$ is defined as follows:
-\[ 
-H'(X, Y) = \max_{x\in X} \, \min_{y\in Y} \, d(x, y) 
-\]
-\[
-H(X, Y) = \max\left (H'(X, Y), H'(Y, X)\right )
-\]
-where $d(\cdot, \cdot)$ is the Euclidean distance. For infinite sets the $\max$ and $\min$ functions should be replaced with $\sup$ and $\inf$.
-
-The most intuitive way of understanding the Hausdorff distance is as a game where the first player must get from one set to the other in as small a distance as possible, and the malevolant second player will choose his starting point to maximize that distance.
-
-Hausdorff distance is a common tool in dealing with the fractal inverse problem \cite{nettleton1994evolutionary}.
-We will use it to evaluate an IFS model by generating a number of points from the model and calculating the Hausdorff distance of that set to the model.
-Since the Hausdorff distance has not been substantiated as a statistical measure, we can only use it as an intermediate indication of model quality.
-To prove the utility of IFS models we will test them on classification and dimensionality reduction tasks.
-
-\subsection*{Dimension}
-
-The \textit{curse of dimensionality} is a blanket term for many effects related the analysis of high dimensional data.
-The main cause of these problems is that as dimensionality increases, the volume of a subspace (such as the unit ball) increases exponentially with it.
-This means, for instance, that vast amounts of samples are required to maintain the same level of coverage in a high dimensional space as compared to a low dimensional one.
-
-The solution to many of these problems is that while the data has a high-dimensional representation, it has a low intrinsic dimension.
-That is, it can be mapped to a low dimensional space without any loss of information.
-A simple example is that of all points lying on a line in some high dimensional space.
-
-IFSs have a particular advantage over classical models like Gaussian mixtures: their intrinsic dimension varies over the whole range of positive reals.
-A search through the space of IFSs can return models of any dimension.
-
-We measure the dimension of datasets and data generated from models using Takens' estimator \cite{takens1985numerical}:
-\[
-D(d_{\mbox{max}}) = -\frac{|A|}{\sum_{a \in A} \ln a} 
-\]
-Where $A$ is the set of distances between pairs of data points lower than parameter $d_{\mbox{max}}$.
-We choose $d_{\mbox{max}}$ as the value that produces the lowest Kolmogorov-Smirnov statistic with the probability distribution underlying the Takens estimator.
-This method is based on the work for general power laws in \cite{clauset2009power} and will be detailed in a later publication.
-
-We estimate the uncertainty of this estimate as follows: bootstrap a version of the dataset by sampling to the same size with replacement, and estimate the dimension of that dataset.
-We repeat the process 50 times and take the sample standard deviation as an estimate of the uncertainty.
-
-\subsection*{Analogous model: Mixture of Gaussians}
-
-To study the IFS as a statistical model, it helps to have a commonly used analogous model to compare it with.
-We use the Mixture of Gaussians (MoG) model.
-This model consists of $n$ multivariate Gaussians over ${\mathbb R}^d$ each represented by a mean $\mu \in {\mathbb R}^d$ and a $d \times d$ covariance matrix $\Sigma$.
-Each component also has a prior $w_i$.
-The complete model describes the following probability distribution:
-\[
-p(x) = \sum_{i \in [1,n]} w_i N(x|\mu_i, \Sigma)
-\]
-Since every multivariate Gaussian can also be described as an affine transformation of the standard normal distribution (mean at the origin and $\Sigma$ the identity matrix), we can also represent the mixture of Gaussians as $n$ affine transformations and $n$ weights, just as the IFS:
-\[
-p(x) = \sum_{i \in [1,n]} w_i S_i(N(0, I))(x)
-\]
-This gives us a classical model with a similar representation.
-Note that the representation would be equal if we used similarities for the MoG model.
-However since an MVN transformed with uniform scaling is rotationally symmetric that would make all the rotational parameters useless.
-For this reason we will allow the MoG model full affine transformations as parameters and the IFS model only similarities.
-Even with this discrepancy, we still consider the MoG to be a reasonable analog.
-
-Unlike the single MVN distribution, there is no closed form solution for the maximum likelihood fit of a MoG model to data.
-We must approximate it.
-The most common method is by the well-known EM algorithm.
-Since our solution to the fractal inverse problem takes inspiration from this algorithm, we describe it here briefly.
-The key intuition behind the EM algorithm is that if we knew which component of the MoG model `generated' each datapoint, we would have a closed form solution for the parameters of the component, and if we knew the parameters of the components we could find out which component was most likely to have generated the data point.
-We turn this circular dependency into an iterative algorithm.
-Starting with some initial MoG model we iterate the following steps
-
-\begin{description}
-  \item[Expectation] Assign each data point a probability distribution over the components of the MoG model proportional to the likelihood of the model given the datapoint.
-In other words, each component takes a certain amount of `responsibility' for the data point.
-  \item[Maximization] Estimate the parameters for each component based on the probability weights calculated in the previous step.
-\end{description}
-
-\section*{Fitting IFS models to data}
-
-We can now describe the central algorithm of this paper.
-To make the explanation as clear as possible we will first describe a skeletal version of the algorithm that works in principle and then describe various alterations to this skeleton that are required to improve the performance of the algorithm on real data with finite approximations of the IFS measures.
-
-The results described are always for the full algorithm with all improvements.
-An implementation of only the basic algorithm should be enough to retrieve the Sierpinski gasket (or a crude approximation) from data.
-
-\subsection*{The basic algorithm}
-
-As with the EM algorithm for MoG models, the main barrier to inducing an IFS model from the data directly is a dependency on latent variables.
-In the MoG case each data point needed to be assigned to a component, in the IFS case we need to know, for each data point the code $\sigma$ that represents the sequence of components.
-
-Given these codes it's a relatively simple matter to find which points maps to which under the data's self-similarity.
-And given two ordered lists of points that are supposed to map into one another we can easily find a map to minimize the error.
-Thus our algorithm works as follows.
-\footnote{the names `expectation' and `maximization' of these steps are only by analogy with the EM algorithm for MoG models. We do not claim any true expectation or maximization procedures.}
-
-\begin{itemize}
-\item We start with in initial IFS $S$, chosen by one of the methods described below. 
-\item (Expectation) We assign each datapoint a code $\sigma$ of length $d$ (where $d$ is a parameter of the algorithm). The codes are provided by the model $S$ as described above. 
-\item (Maximization) We know that if a point $x_\sigma$ is mapped using one of our components we have $S_i(x_\sigma) = x_{i\sigma}$.
-That is, the symbol for the mapping component is prepended to the code.
-For each $x_\sigma$ in our dataset we search the data for a point with a code $\tau = i\sigma_{1:k-1}$, the symbol $i$ followed by the elements of $\sigma$ save for the last.
-For each $i$ we maintain ordered lists of the matches found.
-\item (Maximization) For each component $S_i$ we now have two lists of points such that $S_i$ should map one into the other with minimal squared error.
-We used the method described in \cite{umeyama1991least} to find $S_i$.
-\end{itemize}
-
-We will often use a subsample of the data, to allow us to increase the depth or number of components. In such cases, we reample at each EM iteration.
-
-\subsection*{Improvements}
-
-\subsubsection*{Centering and scaling the data}
-
-Iterated function systems are not very good at approximating data that is weighted far from the origin.
-For this reason we find the affine transformation which maps the data to the range $[0,1]$ in each dimension.
-
-We store this transformation with the model to map the model back to the data space as needed.
-
-\subsubsection*{Using all depths}
-
-The algorithm as described here can work at all depths $d$.
-Since each depth may contain information about self similar scaling in the data, we also want to consider all depths less than $d$ rather than just $d$.
-This is easily done by considering codes of all length up to and including $k$.
-We build a code tree of the codes generated from the dataset and search it systematically for the required matches at all levels.
-
-% This means that for each $i$ we have two sets of points at each depth.
-If the components are equally weighted, we will have a set of about $1/n$ of the data, mapping into a set of about $1/{n^2}$ of the data.
-At depth two the ratios become $1/{n^2}$ and $1/{n^3}$ and so on.
-For large depths, only specific codes will have matching points.
-
-\subsubsection*{Weights}
-
-The algorithm as described above does not give us a way to retrieve the component weights.
-For the coding step we already take the weights into account, so the only thing we must do is retrieve them from the set of codes with which we start the maximization step.
-
-If we have a collection of points $X_\sigma$ with code $\sigma$ and a collection of points $X_\tau$ with code $\tau$ such that $S_i$ should satisfy $S_i(x_\sigma) = x_\tau$, we know that if we generate the IFS model, $w_i$ of the probability mass at stage $t$ always gets distributed to component $S_i$ for stage $t + 1$.
-Thus we can expect the relative size of these sets to follow $w_i$: $|X_\tau| = w_i |X_\sigma|$.
-Over all such sets $X_\sigma^j$ and $X_\tau^j$ we must minimize the squared error:
-\[
-\overline{w_i} = \argmin_{w_i} \sum_{j} \left( |X_\tau^j| - w_i |X_\sigma^j| \right)^2
-\]
-This works out as 
-\[
-\overline{w_i} = \frac{\sum_{j}|X_\sigma^j||X_\tau^j|}{\sum_{j}|X_\sigma^j|^2}
-\]
-
-\subsubsection*{Multiple matches ($\sigma$)}
-
-It is perfectly possible that we get multiple points per code.
-Indeed, at low depths, it is almost certain.
-In this case we have two sets of points $X_\sigma$ and $X_\tau$ such that the points in the former should map onto the latter.
-But our algorithm for finding the map $S_i$ requires the individual points to be matched to each other.
-If the number of points is low we can simply take the mean of both sets, but if the number of points is high, we risk throwing away a lot of information (in particular about rotation).
-Theoretically each point in the first set should map onto every point in the other but this creates an explosion of possiblities.
-If we pair the points up randomly we introduce artifacts in the induced rotation.
-
-We tackle this problem by finding the MVN for both sets.
-Ideally we would then extend our method for finding maps to paired MVNs rather than paired points, but this is a highly complex problem.
-Instead, we make use of the correspondence between MVNs and affine maps.
-We find for each MVN the transformation that transforms the standard distribution ($N(0, I)$) into that MVN.
-We then generate $|X_\tau|$ `spanning points' from some initial distribution and map these to both sets.
-We add these point pairs to the list of matched points from which the maps will be calculated.
-
-To generate the spanning points we use a spherical MVN with variance $\sigma$.
-This value becomes an important parameter in the algorithm.
-If the spanning points have a high variance, then the rotation becomes a more important part of the found affine map.
-The limiting case with $\sigma \rightarrow 0$ is essentially equal to just using the mean of all points with a given code and getting no rotation information from the mapping between single code pairs.
-
-The behaviour of the algorithm under varying $\sigma$ is somewhat mysterious, but trial-and-error testing suggests that for high values the algorithm is more likely to converge.
-This comes at the cost of some solutions, as with $\sigma = 1.0$, the algorithm is unable to find the two-component Koch curve.
-For all experiments reported in this paper we use $\sigma = 0.1$.
-
-\subsubsection*{Edge cases}
-
-As with standard EM algorithms, some undesired behavior can emerge when the initial model is far from the optimum.
-Consider for instance the case where the Sierpinski triangle is the current model and all data is to the top right of the triangle.
-The symbol 2 would not occur in any of the codes and we would have no basis to find $S_2$ in the maximization step.
-If this occurs, we will `split' one of the components that could be found  (distributing the weight equally) and perturb each component slightly to ensure that the components differ.
-The idea behind this approach is simple.
-For every place in every code that used to be assigned to the original component there is now around a fifty/fifty probability of the place being assigned to the other code.
-In the next step roughly half of the matches of the component will have been `donated' to the other component.
-
-The perturbation is done by flattening the model into a vector of values in ${\mathbb R}$ and adding a small random value $N(0, \pi)$ to each. In our experiments $\pi$ was set to 0.3.
-
-\subsubsection*{Use the data as a basis}
-
-When we generate a sample from the model (to a fixed depth, rather than with the chaos game), and when we compute the probability density to a fixed point, we make use of some initial normal distribution.
-As the depth increases the influence of the choice of initial distribution increases exponentially.
-For low depths, however, our choice can affect the quality of our model.
-
-To determine our what our initial model should be, we examine the situation for depth one.
-In this case, our data points get a code of one symbol, and the transformations induced are those that optimally map the whole dataset to these subsets.
-The equivalent MOG model fits a normal distribution to each subset.
-For these normal distributions to be the result of a depth one IFS model, we must choose as its initial distribution the MVN fit to the data.
-With this choice, the MOG model and the MVN model are equivalent (so long as they use the same class of transformations).
-
-To summarize, we choose as our initial distribution an MVN fit for the data restricted to our transformation class.
-Since we are using Similitudes, we fit a symmetrical MVN to our data, and use that as the initial distribution.
-
-% \subsubsection{Branching}
-
-% [@! Test if this helps]
-
-%As an optimization algorithm, EM is strictly linear in the sense that it only considers a single model at each iteration.
-To avoid local minima it can often help to evaluate multiple models per iteration.
-To test whether our algorithm benefits from such additions we introduce a branching mode which works as follows.
-
-%During the iteration of the algorithm we maintain a buffer of models.
-After each iteration, we generate $b$ versions of the current best model perturbed by a parameter $\sigma$ as with the edge cases.
-We add these models to the buffer and sort it by Hausdorff distance with the data (as described in section @!).
-We discard all but the top $h$ models.
-At the next iteration, we continue with the top model from the list and repeat the procedure.
-
-%This method is analogous with beam search.
-Its advantage is that we perform the expensive EM step the same number of times as the non-branching model, and the extra cost lies purely in testing evaluating the extra models.
-(This is not inexpensive in itself, but any branching algorithm will need to test candidate models anyway).
-
-\subsection*{Choosing the initial model}
-
-The choice of initial model is crucial for achieving good performance
-
-\subsection*{Convergence}
-
-\section*{Results}
-
-We perform a number of experiments to give an impression of the behavior of IFSs as statistical models, and the EM algorithm as a learning method.
-
-First, we show a selection of results on 2D datasets to give a visual impression of the behavior of the algorithm.
-These images represent the value of the EM algorithm over earlier approaches, as methods based on evolutionary algorithms commonly fail to induce fractals from a fractal dataset with this speed or accuracy.
-
-We then investigate the use of IFS models on higher dimensional data.
-To attempt to visualize the way an iterated function system models a high dimensional dataset, we use a four component model so that we can map the data back to two dimensions and plot it.
-
-\subsection*{Datasets}
-
-\begin{description}
-  \item[forest]
-  \item[census]
-  \item[abalone]
-\end{description}
-
-\subsection*{Model fitting and dimension scaling}
-
-\begin{table}[p]
-\begin{tabular}{c c c c c}
-\hline
-data & initial & & & best model \\
-\hline
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/data.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/generation0000.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/generation0001.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/generation0010.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/generation0032.png}
-\\
- & & 6 & 48 & 83 \\
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/data.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/generation0000.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/generation0003.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/generation0020.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/generation0079.png}
-\\    
- &  & 3 & 20 & 79 \\
-	\includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/data.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/generation0000.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/generation0008.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/generation0022.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/generation0078.png}
-\\    
- &  & 8 & 22 & 78 \\
-	\includegraphics[width=0.2\textwidth]{../img/runs2d/ball/data.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/ball/generation0000.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/ball/generation0003.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/ball/generation0024.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/ball/generation0072.png}
-\\   
- &   & 3  & 24 & 72 \\
-	\includegraphics[width=0.2\textwidth]{../img/runs2d/coast/data.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/coast/generation0000.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/coast/generation0008.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/coast/generation0020.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/coast/generation0069.png}
-\\
- &  & 8 & 20 & 69 \\
-\hline
-
-\end{tabular}
-\caption{Selected training steps of the algorithm for various data sets.
-All models were trained with a dataset of 100000 points, sampling 300 per iteration for 100 iterations.
-Model depth was 6 for all runs.
-1000 points were sampled from each model and from the dataset to estimate the Hausdorff distance (and determine the `best model').
-The blue squares indicate the transformations (of the red square) that make up the model.
-The green bar indicates component weight.
-By row: (1) The sierpinski triangle, $\sigma = 0.1$ (2) The sierpinski triangle with unequal weights.
-$\sigma = 0.1$ (3) The 2-component koch curve.
-$\sigma = 0.001$ (4) A uniform distribution over the bi-unit ball (3 components, $\sigma = 0.1$) (5) Points sampled from a small stretch of coast line (3 components, $sigma = 0.1$).
-}
-\label{tablelabel}
-\end{table}
-
-[@!
-EXPERIMENT for various datasets, for various depths and component sizes, scatter plot the difference between the model dimension and the data dimension against the model accuracy (by hausdorff distance) show uncertainty with error bars]
-
-First, we fit IFS models to our datasets (ignoring the class information).
-We report the Haussdorff distance of a sample of 1000 points from the data to 1000 points generated by the model (we consider the iris dataset to be too small for this purpose).
-We withhold a random 25\% of the data for testing, repeat the experiment 20 times and report the sample standard deviation and the mean of the resulting values for symmetric error.
-
-We are also interested in the dimensional scaling of the model.
-In this case we are interested in the training behavior of the model, so we do not withhold test data.
-We measure the dimension of the dataset on a random sample of 300 points (repeating 25 times to measure the uncertainty) and generate 300 points from the best model and measure the dimension in the same way.
-We expect that the difference between the model dimension and the data dimension depends on the quality of the model.
-Figure @!
-plots the data dimension against the model dimension.
-
-In all cases we train the model for 300 generations and select the best model by Hausdorff distance.
-
-\begin{figure}[t]
-\centering
-
-  \begin{subfigure}{0.66\textwidth}
-	\includegraphics[width=\textwidth]{../img/dimension/out-table.pdf}
-  \end{subfigure}
-  \begin{subfigure}{0.33\textwidth}
-    \begin{tabular}{c c }
-    data & model \\
-		\includegraphics[width=0.5\textwidth]{../img/dimension/data-henon.png} & 
-		\includegraphics[width=0.5\textwidth]{../img/dimension/model-henon.png} \\
-		\includegraphics[width=0.5\textwidth]{../img/dimension/data-ikeda.png} & 
-		\includegraphics[width=0.5\textwidth]{../img/dimension/model-henon.png} \\
-		\includegraphics[width=0.5\textwidth]{../img/dimension/data-henon.png} & 
-		\includegraphics[width=0.5\textwidth]{../img/dimension/model-henon.png} \\
-	\end{tabular}
-  \end{subfigure}
-  \caption{To investigate the dimensional scaling of IFS models, we train them on three generated datasets.
-The datasets were chosen for a) their low dimension so that we can acurately measure dimension, b) their synthetic nature so that we hav an analytic value of their dimension and c) their lack of an accurate IFS approximation, so that the models will not fit perfectly.
-The datasets are shown on the left with their respective IFS model.
-They are, from top to bottom: the H\'enon attractor, the Ikeda attractor, and the R\"ossler attractor.}
-  \label{fig:dim_scale}
-\end{figure}
-
-\subsection*{Classification}
-
-We explore two methods for using IFS models for classification.
-In the first, we train a single model for each class and combine in a straightforward manner using Bayes' rule.
-In the second, we learn a single model for the whole set of points, and use the codes generated by the model to perform classification.
-
-\subsubsection*{Bayesian}
-
-Given some data set $X = \{x_i\} \subseteq {\mathbb R}^d$ with associated classes $c_i \in \{1, 2, \ldots, m\}$, we split the data by class: $X_c = \{ x \in X \mid c_i = c\}$.
-We train a probability model $p_c$ for each $X_c$ and classify each new point with
-\[
-C(x) = \argmax_{c} p_c(x) p(c)
-\]
-The class prior $p(c)$ is estimated as $\frac{|X_c|}{|X|}$.
-
-\subsubsection*{Single model}
-
-An alternative way to construct a classifier with IFS modelling is to learn a single model for the whole dataset, ignoring class labels.
-For some datasets, the set of all points may present a more coherent and simple distribution than the sets split by class label.
-Another advantage of this approach is that we only have to learn one model, which is cheaper both in training and in evaluation.
-
-The model provides us with a mapping from the data space (${\mathbb R}^d$) to code space ($\{1, \ldots, k\}^d$).
-To turn this model into a classifier we proceed as follows.
-We build a code tree such that each possible code represents a path from the root to the leaf node.
-We then traverse this tree for each point in the training set, maintaining a frequency count for each class observed at each node.
-Thus, the root node gives us the frequency distribution over the classes for the whole dataset and each node below it refines that to a distribution over a section of the dataset whose code starts with a given prefix.
-This turns each node into a very basic classifier, which simply returns the class with the highest frequency count.
-The root node will give us the classifier that uses nothing more than class frequencies and each node below it will use more information about the neighbourhood of the point.
-
-For each point to be classified we take the probabilities at each node in the path and average them evenly.
-
-\begin{table}[h]
-\begin{tabular}{| l | r l r l | r l r l  r l r l |}
-\hline
-  & ifs bayes & & ifs single & & mog & & ann & & knn & & tree & \\
-\hline
-census  & & & & & & & 0.054 &\color{gray}(0) & 0.06 & \color{gray} (0.002) & 0.05 & \color{gray} (0.001) \\
-forest  & & & & & & & 0.163 &\color{gray}(0.001) & 0.138 &\color{gray}(0.001) & 0.12&\color{gray}(0.003) \\
-abalone & & & & & & & 0.47 & \color{gray}(0.008) & 0.47 & \color{gray}(0.006) & 0.48 & \color{gray}(0.01) \\
-mnist   & & & & & & & 0.09 & & 0.05 & & 0.20 & \\
-\hline
-
-\end{tabular}
-\caption{Hausdorff distances for models trained against various datasets.}
-\label{classification-results}
-\end{table}
-\subsection{Time performance}
-
-An informal analysis of the algorithm shows that the expectation step costs $O(k^dn)$, with the maximization step taking roughly $O(n^2)$.
-This falsely suggests that the expectation step, being linear in the number of data points, is negligable compared to the maximization step, which is quadratic.
-Of course, when viewed in terms of the parameters the expectation step is exponential and this is what causes the expectation step to become the bottleneck for anything but very low values of $k$ and $d$.
-
-\section*{Conclusions}
-
-We have presented a framework for a fractal approach to the statistics of point patterns.
-We introduced an EM-style algorithm for fitting fractal models to data.
-Compared to search algorithms for other types of models the algorithm is relatively expensive, but compared to other approaches for fractal modeling the algorithm represents a significant improvement @!.
-
-Further, we have given some indication of how the study of fractal modeling can open up interesting new prospects in machine learning.
-We expect that these methods will allow many avenues of future research.
-One option would be to investigate the use of more expressive maps.
-The similitudes used here could be replaced by neural networks, for instance.
-The EM search algorithm is fairly simple, and could be extended to try various branches and backtrack if it deviates too far from a good model.
-
-\subsubsection*{Other fractals}
-
-It is important to note that there are many other families of fractals besides iterated function systems.
-They are not covered by our method, and cannot be properly represented by Iterated Function Systems.
-Examples include Julia sets and dynamical systems.
-
-There is one family of fractals in particular that would be very interesting to explore in light of this research, and that is the family of random fractals.
-When we think of natural phenomena that seem to resemble iterated function systems---coast lines, trees, the veins of a leaf---there is always a combination of fractal scaling and some random component.
-The framework most likely to capture these phenomena is that of random Iterated Function Systems, as described in \cite{hutchinson2000deterministic}.
-
-The random IFS model is an extension of the Iterated Function System which produces not one single set but a distribution over all sets.
-The dataset for which to learn a random fractal would not consist of one list of vectors, but of a list of lists of vectors.
-For instance, a collection of timeseries preceding a critical event, or a collection of images of snowflakes.
-In \cite{bloem2010fractal} we showed a proof-of-concept, indicating that such models can be learned by evolutionary algorithm in principle, but we did not advance beyond the simplest conceivable model.
-If the EM algorithm described here could be extended to random fractals, it would open up a new type of machine learning task, where instances ae represented by point patterns, and the model to be learned is a distribution over distributions.
-
-\bibliographystyle{siam}
-\bibliography{fractal}
-
 \end{document}

### 2013-03-15 Benelearn/src/benelearn.tex
...

@@
 
+\usepackage{style/benelearn2013}
 \usepackage{charter}
 \usepackage{eulervm}
+% \usepackage{graphicx}
 \usepackage{amsmath, amsthm, amssymb}
-\usepackage{graphicx}
 \usepackage{caption}
 \usepackage{subcaption}
-
-\theoremstyle{definition}
-\newtheorem*{thm}{Theorem}
-\newtheorem{lma}{Lemma}
-\newtheorem*{dfn}{Definition}
-\newtheorem*{exm}{Example}
 
@@
 
-\DeclareMathOperator*{\argmin}{arg\,min}
-\DeclareMathOperator*{\argmax}{argmax}
+% \DeclareMathOperator*{\argmin}{arg\,min}
+% \DeclareMathOperator*{\argmax}{argmax}
 
@@
 % non-indented, spaced paragraphs
-\setlength{\parindent}{0.0in}
-\setlength{\parskip}{0.1in}
+% \setlength{\parindent}{0.0in}
+% \setlength{\parskip}{0.1in}
 
@@
 \begin{abstract}
-\noindent 
+We investigate the use of compression-based learning on graph data.
+We hypothesize that the sequential representation of data on which general purpose compressors operate, hampers compression-based learning, as a graph can be represented sequentially in many different, equivalent ways.
+Using Normalized Compression Distance (NCD), we test various compressors that operate directly on the graph representation against compressors that operate on a bitstring representation.
+We use both synthetic, randomly generated graphs and real-life datasets.
 \end{abstract}
 
+\emph{Compression equals learning} is the mantra of the field of compression-based learning.
+If we can compress data, we must have found some structure, we must have learned something.
+Conversely, if we have learned something about data, we should be able to use that knowledge to represent our data more succinctly.
+
+The idea of compression-based learning is expressed in two strongly related fields: \emph{Minimum Description Length} and \emph{Kolmogorov Complexity}.
+Minimum Description Length \cite{}@!
+builds a statistical framework on the principle that a good model is one that can be used to describe the data in as few bits as possible.
+Kolmogorov Complexity is concerned with the mathematical expression of the \emph{information content} of a piece of data.
+If we can find some short description of a dataset (ie.
+compress it) then the total information content must be below the length of that description.
+
+In this paper we investigate the use of these compression-based methods on graph data.
+Examples of such data are social graphs, transportation graphs, trade networks or semantic graphs.
+The graph is a powerful and versatile representation.
+Most applications of compression-based learning choose sequential models, such as deterministic finite automata, or general purpose compressors, which operate on bitstrings.
+If we have data in the form of a graph, we can translate it to a bitstring, of course, but in this transformation we complicate our problem.
+For a different ordering of nodes, the graph is the same but the bitstring changes radically.
+To a simple compressor, the two bitstrings may not share very much structure, even though they represent the same graph.
+
+If this problem holds, we should investigate compressors that operate on the level of the graph representation, for instance by finding frequent subgraphs or clustering the graph.
+The methods do not suffer the problem of isomorphism, but as a result, they are more expensive than their sequential cousings.
+In this paper we hope to provide a first indication of how far the sequential approach will go, and whether the native approach will let us continue on from there.
+
+\section*{Normalized Compression Distance}
+
+From the many methods born from Kolmogorov Complexity and Minimum Description Length, we choose \emph{Normalized Compression Distance} (NCD) as a representative method of compression-based learning.
+The choice is a practical one, NCD is a simple method which requires nothing more than a compressor fit for the type of data under question.
+Any domain-specific knowledge we wish to use about our data (eg.
+it represents a graph) must be placed in the compressor.
+We proceed under the assumption that conclusions reached about the performance of NCD on graph data will translate to MDL and other frameworks.
+
+We will provide a brief, intuitive explanation of the principles involved, sufficient to understand the ideas presented in this paper.
+For a more in-depth and rigorous treatment we refer the reader to @!
+and @!.
+
+\subsection*{Kolmogorov Complexity}
+
+Kolmogorov complexity is a notion of information content, based on two principles: (a) all data can be represented as a bitstring (b) the shorter this bitstring can be described, the less information is contained in it.
+
+The second principle is formalized in two ways.
+First, by `description' we mean a description in a formal language that is maximally expressive.
+To formalize expressivenes, we require that the method of description is Turing-complete (of equivalent strength to a Turing machine).
+By the (strong) Church-Turing thesis, this suggests that there is no reasonable way of defining a more expressive method of description.
+
+We choose the Turing machine as a canonical method of description, and we fix some enumeration of all Turing machines $\{T_i\}$. There exists a Universal Turing machine $U$ that is defined as follows:
+\[
+U(\langle i, p\rangle) \simeq T_i(p)
+\] 
+That is, if $U$'s input consists two bitstring arguments $i$ and $p$, combined with some computable pairing function $\langle \cdot, \cdot\rangle$, then $U$ computes the same function as $T_i$ on input $p$ if $T_i(p)$ halts, or fails to halt if $T_i$ fails to halt.
+$U$ provides our formalization of 'description'.
+If our data is $x$, and for some $y = \langle i, p\rangle$, $U(y) = x$, we say that $y$ is a description of $x$ on our reference universal Turing machine.
+
+We can now say, that with respect to $U$, there must be a minimal description for any given data:
+\[
+K_U(x) = \min\left \{|y| : U(y) = x\right\} 
+\]
+
+In some contexts, it is desirable to distinguish between the classical Kolmogorov complexity and the prefix-free Kolmogorov complexity.
+In the context of Normalized Compression Distance the distinction does not matter.
+
+A complete treatment of Kolmogorov Complexity is outside the scope of this paper, but the following properties are important to understand.
+
+\begin{description}
+\item[$K(\cdot)$ is uncomputable.] There can be no algorithm which computes the Kolmogorov Complexity of $x$ for all $x$.
+It can however be bounded from above, and for every algorithm which bounds it, there is an algorithm which provides a better bound.
+\item[$K(\cdot)$ is approximated by all computable compressors.] If we have some compressor for our data $x$ (say GZIP) we can find the decompression algorithm somewhere in $\{T_i\}$, say as $T_g$.
+We can have a description on $U$ as $U(\langle g, z\rangle) = x$ so that $K_U(x)\leq |z| + O(1)$.
+In this way, any computational structure in $x$ is taken into account in $K(x)$, and $K(\cdot)$ can be approximated by any computable compressor.
+\end{description}
+
+This gives us the basic philosophy behind all translations of Kolmogorov complexity to the realm of practical applications: we approximate Kolmogorov complexity by some learning algorithm or compressor.
+
+Finally, we define conditional Kolmogorov Complexity $K(x| y)$.
+Where regular Kolmogorov Complexity is defined as the shortest program which produces $x$, the conditional variant is defined as the shortest program which produces $x$ when provided with $y$.
+A complete treatment is available in \cite{}.
+
+\subsection*{Normalized Information Distance (NID)}
+
+The length of the shortest program to get from $y$ to $x$ intuitively suggests that $K(\cdot\mid \cdot)$ can be seen as a similarity measure.
+Clearly, very little is required to transform a string into itself, or a very similar string, whereas for two random strings, only a program that stores the second in its entirety can make the transformation.
+
+This intuition prompted Li and Vitanyi @!
+\cite{} to investigate the use of Kolmogorov Complexity as a metric.
+The first issue is that $K(\cdot\mid \cdot)$ is not symmetric.
+It takes a small program to blank out the collected works of Shakespeare, but the reverse is a more complex operation.
+The first step is thus to define the (symmetric) Information Distance:
+\[
+ID(x, y) = \max \left [K(x\mid y),K(y\mid x) \right ] 
+\]
+
+The second issue is one of scale.
+If two strings of a million bits differ by 1000 bits, we might consider them quite similar, whereas two strings of 1000 bits that differ by that amount could not be more different.
+Thus, we would like to take the length of the strings into account.
+This gives us the Normalized information Distance (NID)
+
+\[
+NID(x, y) = \frac{\max \left [K(x \mid y),K(y \mid x) \right ] }{\max \left [K(x), K(y) \right ]}
+\]
+
+We would like to approximate this by replacing each occurence of the Kolmogorov complexity with an approximation by a compressor, which we will call $C$.
+As most compressors do work on a conditional basis (expressing data given some existing data), we want to rewrite the condition $K$'s as nonconditional ones.
+In doing this, we make first accept beyond the constant term inaccuracy that is innate to Kolmogorov Complexity, a further logarithmic inaccuracy so that we can rewrite as
+
+\begin{align*}
+NID(x, y)	 &= \frac{\max \left [K(x, y) - K(x),K(y, x) - K(y)\right ] }{\max \left [K(x), K(y) \right ]} \\ 
+	&= \frac{\max \left [K(xy) - K(x),K(yx) - K(y)\right ] } {\max \left [K(x), K(y) \right ]} \\
+	&= \frac{\min \left [K(xy), K(yx)\right ] - \min \left[K(x), K(y)\right]}{\max \left [K(x), K(y) \right ]} 
+\end{align*}
+
+If we replace the Kolmogorov complexity with a compressor $C$, we get the normalized compression distance
+\begin{align*}
+NCD(x, y) 
+	&= \frac{\min \left [C(xy), C(yx)\right ] - \min \left[C(x), C(y)\right]}{\max \left [C(x), C(y) \right ]}\\
+	&= \frac{C(xy) - \min \left[C(x), C(y)\right]}{\max \left [C(x), C(y) \right ]} 
+\end{align*}
+
+Where the last step follows from the assumption that our compressor is roughly symmetric ($C(xy) = C(yx)$).
+
+\section*{Method}
+
+Our aim is to test a selection of compressors on a variety of graph data.
+To ascertain the performance of the compressors, we generate graphs from different sources, calculate their distances and see whether a clustering algorithm can reconstruct the original sources as clusters.
+
+\subsection*{Experiment 1: Synthetic data}
+
+We generate graphs from two models.
+The first is the classic Erd\H{o}s-R\'enyi model, where a uniform random choice is made from all graphs with $n$ nodes and $m$ links.
+The second is the Barab\'asi-Albert model \cite{}@!, which grows a graph from a set of $n_0$ unconnected nodes, one node at a time, connecting each new node to $k$ distinct existing nodes where the probability that a given existing node is chosen for a connection is its degree, divided by the sum of the degrees of all nodes.
+Thus, under the BA model the more links a node has, the higher the probability that it will accrue even more.
+This effect causes the degree distribution of a BA network to form a scale free distribution.
+
+Since we want there to be some challenge in separating the two classes of network, we, ensure that they have the same number of nodes and links.
+To achieve this, we first generate the BA networks, count their nodes and links and use these as parameters for the ER model.
+
+Once we have this dataset, calculate the NCD with a given compressor for every pair of graphs in the dataset, giving us a symmetric matrix.
+We use the k-medioid algorithm to cluster this set into two clusters, and we check whether graphs from a given source tend to end up in a specific cluster.
+
+\subsection*{Experiment 2: Real-life data}
+
+In this experiment we sample subgraphs from large, existing graphs.
+Again, to make the problem challenging enough to be meaningful, we want the subgraphs to have identical size in both nodes and links.
+To accomplish this, we proceed as follows.
+
+\begin{itemize}
+  \item Choose a random subgraph by choosing a random node uniformly, and including all nodes within distance $d$. Generate $N$ such graphs for each source.
+  \item Find the subgraph with the smallest number of nodes $n_{\mbox{min}}$ over all subgraphs and, for every other subgraph, remove randomly chosen nodes until its number of nodes matches $n_{\mbox{min}}$
+  \item Find the subgraph with the smallest number of links $l_{\mbox{min}}$ and, for every other subgraph, remove randomly chosen links until the number of links matches $l_{\mbox{min}}$ 
+\end{itemize}
+
+With this dataset of subsets, we proceed as before, calculating the distances between the subgraphs and clustering them into as many clusters as we have sources, to see whether the resulting clusters match the sources.
+
+We use the following datasets:
+\begin{description}
+	\item[web]
+	\item[metabolic] 
+	\item[semantic A] A semantic description of 
+	\item[semantic B]
+\end{description}
+
+Some basic properties of these graphs are given in tabel \ref{table:properties}.
+
+\begin{table}
+\label{table:properties}
+\begin{tabular}{ l | r r r r r}
+\hline
+    & nodes & links & C & $alpha$ & $p_\alpha$ \\
+\hline
+  web & & & & & \\
+  metabolic & & & & & \\
+  semantic A & & & & & \\
+  semantic B & & & & & \\
+
+\hline
+\end{tabular}
+\caption{The basic properties of our natural datasets.
+We report the number of nodes, the number of links, the number of nodes in the largest component (C), the estimates power law exponent ($\alpha$), and the significance for the power law model ($p_\alpha$, significant for values \emph{above} 0.01).}
+
+\end{table}
+
+\subsection*{Compressors}
+
+\subsubsection*{GZIP}
+
+We use GZIP as our general purpose compressor.
+Specifically, in our experiments, we use the implementation of GZIP that is part of the standard Java SDK.
+To store a graph with GZIP, we flatten the lower half of its adjacency matrix into a bitstring and store this together with a list of the node and link labels.
+We use Java object serialization to take care of delimiting the data and translating it to bits.
+
+\subsubsection*{Subdue}
+
+Subdue \cite{} is an algorithm for finding frequent subgraphs in graph data. The algorithm searches through all possible subgraphs, searching for the  one that maximally compresses the data.
+
+\section*{Results}
+
+\subsection*{Synthetic data}
+
+\subsubsection*{Small graphs}
+
+\begin{table}[h]
+\label{table:synthetic-small}
+
+\begin{subfigure}[b]{0.5\columnwidth}
+	\label{table:gzip-small}
+	\centering
+	\begin{tabular}{l | r | r r}
+	\hline
+	  ER & @!/@! & @! & @! \\
+	  BA & @!/@! & @! & @! \\
+	\hline
+	\end{tabular}
+	\caption{GZIP: error @!}
+\end{subfigure}\begin{subfigure}[b]{0.5\columnwidth}
+	\label{table:subdue-small}
+	\centering
+	\begin{tabular}{l| r | r r}
+	\hline
+	  ER & @!/@! & @! & @! \\
+	  BA & @!/@! & @! & @! \\
+	\hline
+	\end{tabular}
+	\caption{Subdue: error @!}
+\end{subfigure}
+
+\caption{Confusion matrices for the synthetic data.
+Results for synthetic graphs of exactly @!
+nodes and @!
+links generated by the ER and the BA algorithms.
+For each source we generated @!
+such graphs and clustered the results with different compressors.
+The error is computed by assuming an optimal labeling and computing the symmetric error as in classification.}
+\end{table}
+
+\subsubsection*{Large graphs}
+
+For larger graphs, the problem becomes easier (as there is more structure), but we can add the fractal generator.
+This generator can only produce larger graphs, and with its inclusion we can no longer ensure that all graphs have equal size.
+\begin{table}[h]
+\label{table:synthetic-large}
+\begin{subfigure}[b]{1\columnwidth}
+\begin{tabular}{l | r | r r r r}
+\hline
+  ER      & @!/@!  & @! & @! & @! & @! \\
+  BA      & @!/@!  & @! & @! & @! & @!\\
+  fractal (pure) & @!/@!  & @! & @! & @! & @!\\
+  fractal (small world) & @!/@!  & @! & @! & @! & @!\\
+\hline
+\end{tabular}
+\caption{GZIP: error @!}
+\end{subfigure}
+\begin{subfigure}[b]{1\columnwidth}
+\begin{tabular}{l | r | r r r r}
+\hline
+  ER      & @!/@!  & @! & @! & @! & @! \\
+  BA      & @!/@!  & @! & @! & @! & @!\\
+  fractal (pure) & @!/@!  & @! & @! & @! & @!\\
+  fractal (small world)& @!/@! & @! & @! & @! & @!\\
+\hline
+\end{tabular}
+\caption{Subdue: error @!}
+\end{subfigure}
+\caption{\ldots}
+\end{table}
+
+\subsection*{Real-life data}
+
+\begin{table}[h]
+\label{table:real-life}
+\begin{subfigure}[b]{1\columnwidth}
+\begin{tabular}{l | r | r r r r}
+\hline
+  ER      & @!/@!  & @! & @! & @! & @! \\
+  BA      & @!/@!  & @! & @! & @! & @!\\
+  fractal (pure) & @!/@!  & @! & @! & @! & @!\\
+  fractal (small world) & @!/@!  & @! & @! & @! & @!\\
+\hline
+\end{tabular}
+\caption{GZIP: error @!}
+\end{subfigure}
+\begin{subfigure}[b]{1\columnwidth}
+\begin{tabular}{l | r | r r r r}
+\hline
+  ER      & @!/@!  & @! & @! & @! & @! \\
+  BA      & @!/@!  & @! & @! & @! & @!\\
+  fractal (pure) & @!/@!  & @! & @! & @! & @!\\
+  fractal (small world)& @!/@! & @! & @! & @! & @!\\
+\hline
+\end{tabular}
+\caption{Subdue: error @!}
+\end{subfigure}
+\caption{\ldots}
+\end{table}
+
+\subsection*{The models}
+
+One advantage of the model-based compressors over the general purpose ones is that we use the model as an indication of the structure in the data.
+In the case of graph data in particular, we can expect compressors that operate directly on the graph to produce models that are easier to relate to the semantics of the original data than those of the sequential compressors.
+
+To evaluate this we concatenate all graphs from a single source and report the model (ie. the frequent subgraph) found by Subdue.
+
+\section*{Conclusions and future work}
+
+Our experiments show that sequential general purpose compressors are far better at performing graph inference than expected.
+Despite the random ordering of the nodes, the bitstring contains enough shallow patterns that a compressor like GZIP can tell a random graph, a scale free one and a two fractal graphs apart.
+Even for graphs with the same number of nodes and links.
+This result is particularly interesting considering the high resource reuirements put on most algorithms for graph inference.
+
+This result suggests that inference on graphs can be performed by sequential algorithms on a sequential representation in linear time, with decent results.
+
+As for the graph-compressors, we see a small improvement relative to the sequential compressors for a string trade-off in computational resources.
+Subdue as used in this paper is a relatively simple compressor, which isolates only a single subgraph for compression and we tested it only at modest parameters.
+The publications surreounding subdue offer much more complex solutions (such as the induction of graph grammars) and there may be some promise in parallelizing the algorithm.
+
+The notion of compression is a good framework within which to combine many approaches to inference from the most general to the most domain specific.
+The Minimum Description Length principle and its associated techniques, which have not been investigated yet for reasons of scope, offer the promise of an even broader field of approaches to the analysis of graph data.
+
+\section*{Acknowledgements}
+
+This publication was supported by the Dutch national program COMMIT. The author would like to thank Leen Torenvliet, Eugenio Bargiacchi and Eugenio di Leo for comments and preliminary research.
+
 \end{document}

### 2013-03-15 Benelearn/src/style/benelearn2013abstract.tex
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% BENELEARN 2013 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Use the following line _only_ if you're still using LaTeX 2.09.
%\documentstyle[benelearn2013,epsf,mlapa]{article}
% If you rely on Latex2e packages, like most moden people use this:
\documentclass{article}

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''

\usepackage{benelearn2013}

% Employ this version of the ``usepackage'' statement after the paper has
% been accepted, when creating the final version.  This will set the
% note in the first column to ``Appearing in''
%\usepackage[accepted]{benelearn2013}

% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}

% For citations
\usepackage{mlapa}

% The \benelearntitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\benelearntitlerunning{Formatting Instructions for BENELEARN 2013 Abstracts}

\begin{document}

\twocolumn[
\benelearntitle{Formatting Instructions for BENELEARN 2013 Abstracts \\
           \texttt{http://www.benelearn2013.org/}}

 \benelearnauthor{Your Name}{email@yourdomain.edu}
 \benelearnaddress{Your Fantastic Institute,
             van Goghstraat 7, 2387, Baarle-Hertog, Belgium}
 \benelearnauthor{Your CoAuthor's Name}{email@coauthordomain.edu}
 \benelearnaddress{Their Fantastic Institute,
             Magrittelaan 42, 5110 AB Baarle-Nassau, The Netherlands}
\benelearnaddress{{\bf Keywords}: submission information, deadline data,
	formatting information
          }
\vskip 0.3in
]

Abstracts and papers should be submitted electronically, no later than March 12, 2013. Notification of acceptance is on April 16.

BENELEARN 2013 will rely exclusively on electronic formats for submission and review. Authors are expected to write their paper with the \textbf{\LaTeX} editor, for which electronic templates are accessible at the BENELEARN website. If you are using the templates, the formatting instructions below will already be enforced.

References must include page numbers whenever possible and be as complete as possible. Place multiple citations in chronological order. For abstracts dealing with published work, please indicate this in the text and provide the appropriate references. The review process will be single blind, so also mention your name and affiliation. Please consult the template for full papers to format references. If the authors' names are included in the sentence, place only
the year in parentheses, for example when referencing Rob Schapire's seminal result \yrcite{schapire90}. Otherwise place the entire reference in parentheses with the authors and year separated by a comma \cite{schapire90,kearns89}.

An abstract should not be longer than one page. Submission to BENELEARN 2013 will be entirely electronic, through Easychair.  The URL and information about the submission process will appear on the conference web site. To ensure our ability to print submissions, authors must provide their manuscripts in PDF format.

Authors need to pay close attention to the typefaces used. More information about this can be found in the template for full papers. Problems can be avoided by using the \textbf{pdflatex} program instead of straight \LaTeX. This program avoids the Type-3 font problem, however you must ensure that all of the fonts are embedded (use {\tt
pdffonts}).

\section*{Acknowledgments}
\begin{scriptsize}
{\footnotesize •}
\end{scriptsize}We acknowledge that this template is a modification of the template for BeneLearn 2012.

% In the unusual situation where you want a paper to appear in the
% references without citing it in the main text, use \nocite
%\nocite{zinkevich03}

\bibliography{example_paper}
\bibliographystyle{mlapa}

\end{document}

% Version History
% Document updated by Florian Kunneman for BENELEARN 2013
% Document updated by Willem Waegeman for BeneLearn 2012
% Document updated by Peter van der Putten for BeneLearn 2011
% This document was modified by Kurt Driessens for BeneLearn 2010.
% Before it was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This 2009 version was
% created by Kiri Wagstaff, slightly modified from Sam Roweis's 2008
% version, which is slightly modified from Prasad Tadepalli's 2007
% version which is a lightly changed version of the previous year's
% version by Andrew Moore, which was in turn edited from those of
% Kristian Kersting and Codrina Lauth. Alex Smola contributed to the
% algorithmic style files.

### 2013-03-15 Benelearn/src/style/benelearn2013fullPaper.tex
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% BENELEARN 2013 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Use the following line _only_ if you're still using LaTeX 2.09.
%\documentstyle[benelearn2012,epsf,mlapa]{article}
% If you rely on Latex2e packages, like most moden people use this:
\documentclass{article}

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''

\usepackage{benelearn2013}

% Employ this version of the ``usepackage'' statement after the paper has
% been accepted, when creating the final version.  This will set the
% note in the first column to ``Appearing in''
%\usepackage[accepted]{benelearn2013}

% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}

% For citations
\usepackage{mlapa}

% The \benelearntitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\benelearntitlerunning{Formatting Instructions for BENELEARN 2013 Papers}

\begin{document}

\twocolumn[
\benelearntitle{Formatting Instructions for BENELEARN 2013 Papers \\
           \texttt{http://www.benelearn2013.org/}}

 \benelearnauthor{Your Name}{email@yourdomain.edu}
 \benelearnaddress{Your Fantastic Institute,
             van Goghstraat 7, 2387, Baarle-Hertog, Belgium}
 \benelearnauthor{Your CoAuthor's Name}{email@coauthordomain.edu}
 \benelearnaddress{Their Fantastic Institute,
             Magrittelaan 42, 5110 AB Baarle-Nassau, The Netherlands}
\benelearnaddress{{\bf Keywords}: submission information, deadline data,
	formatting information
          }
\vskip 0.3in
]

\begin{abstract}

In the following, you will find formatting guidelines for the submission of a BENELEARN 2013 long paper.

\end{abstract}

\section{Electronic Submission}
\label{submission}

\subsection{Templates for Papers}

Electronic templates for producing papers for submission are available for \LaTeX\/ at the BENELEARN website.

If you are using the templates, the formatting instructions below will already be enforced. The content rules you must follow are:
\begin{itemize}
\item The maximum paper length for full papers is 8 pages (references included).  For abstracts, it is 1 page.
\item You are allowed to include author information or acknowledgments in your initial submission. The review process will be single blind. Do include keywords.
\item Place figure captions {\em under} the figure (and omit titles from inside the graphic file itself). Place table captions {\em over} the table.
\item References must include page numbers whenever possible and be as complete as possible. Place multiple citations in chronological order.
\end{itemize}
Please see below for details on each of these items.

\subsection{Submitting Papers}

Submission to BENELEARN will be entirely electronic, by means of easychair (follow the link on the website).

{\bf Paper Deadline:} The deadline for paper submission to BENELEARN 2013 is March 12th, 2013.

To ensure our ability to print submissions, authors must provide their manuscripts in \textbf{PDF} format. Furthermore, please make sure that files contain only Type-1 fonts (e.g.,~using the program {\tt pdffonts} in linux or using File/DocumentProperties/Fonts in Acrobat).  Other fonts (like Type-3) might come from graphics files imported into the document.

Authors should pay close attention to the \textbf{\LaTeX} typefaces used.  Specifically, when converting the dvi output of \LaTeX to Postscript the default behavior is to use non-scalable Type-3 PostScript bitmap fonts to represent the standard \LaTeX fonts. The resulting document is difficult to read in electronic form; the type appears fuzzy. To avoid this problem, dvips must be instructed to use an alternative font map.  This can be achieved with something like the following commands:\\ {\bf dvips -Ppdf -tletter -G0 -o paper.ps paper.dvi}\\ {\bf ps2pdf paper.ps}\\ Note that it is a zero following the ``-G''. This tells dvips to use the config.pdf file (and this file refers to a better font mapping).

Another alternative is to use the \textbf{pdflatex} program instead of straight \LaTeX. This program avoids the Type-3 font problem, however you must ensure that all of the fonts are embedded (use {\tt pdffonts}). If they are not, you need to configure pdflatex to use a font map file that specifies that the fonts be embedded. Also you should ensure that images are not downsampled or otherwise compressed in a lossy way.

\subsection{Submitting Final Camera-Ready Copy}

The final versions of papers accepted for publication should follow the same format and naming convention as initial submissions. See Section~\ref{final author} for details of how to format this.

The footnote, ``Preliminary work.  Under review by the annual machine learning conference of Belgium and The Netherlands (BENELEARN). Do not distribute.'' must be
modified to ``Appearing in \textit{Proceedings of BENELEARN 2013}. Copyright 2013 by the author(s)/owner(s).'' In your .tex file, simply change
$\mathtt{\backslash usepackage\{benelearn2013\}}$ to $\mathtt{\backslash
usepackage[accepted]\{benelearn2013\}}$.

Camera-ready copies should have the title of the paper as running head on each page except the first one.  The running title consists of a single line centered above a horizontal rule which is $1$ point thick. The running head should be centered, bold and in $9$ point type.  The rule should be $10$ points above the main text.  For the camera-ready copies, the original title is automatically set as running head using the {\tt fancyhdr} package which can be obtained at the BENELEARN 2013 web site. In cases where the original title exceeds the size restrictions, a shorter form can be supplied by using
$$\mathtt{\backslash benelearntitlerunning\{\ldots\}}$$
just before $\mathtt{\backslash begin\{document\}}$.

\section{Format of the Paper}

All submissions should follow the same format to ensure the printer can reproduce them without problems and to let readers more easily find the information that they desire.

\subsection{Length and Dimensions}

Papers must not exceed 8 pages (1 page for abstracts).
including all figures, tables, references, and appendices. We will return to the authors any submissions that exceed this page limit or that diverge significantly from the format specified herein.

The text of the paper should be formatted in two columns, with an overall width of 6.75 inches, height of 9.0 inches, and 0.25 inches between the columns. The left margin should be 0.75 inches and the top margin 1.0 inch (2.54~cm). The right and bottom margins will depend on whether you print on US letter or A4 paper, but all final versions must be produced for A4 paper.

The paper body should be set in 10~point type with a vertical spacing of 11~points. Please use Times Roman typeface throughout the text.

\subsection{Title}

The paper title should be set in 14~point bold type and centered between two horizontal rules that are 1~point thick, with 1.0~inch between the top rule and the top edge of the page. Capitalize the first letter of content words and put the rest of the title in lower case.

\subsubsection{Author Information}
\label{final author}
The review process will be single blind. Author information should start 0.3~inches below the bottom rule surrounding the title. The authors' names should appear in 10~point bold type, electronic mail addresses in 10~point small capitals, and physical addresses in ordinary 10~point type. Each author's name should be flush left, whereas the email address should be flush right on the same line. The author's physical address should appear flush left on the ensuing line, on a single line if possible. If successive authors have the same affiliation, then give their physical address only once.

\subsection{Abstract}

The paper abstract should begin in the left column, 0.4~inches below the final address. The heading `Abstract' should be centered, bold, and in 11~point type. The abstract body should use 10~point type, with a vertical spacing of 11~points, and should be indented 0.25~inches more than normal on left-hand and right-hand margins. Insert 0.4~inches of blank space after the body. Keep your abstract brief, limiting it to one paragraph and no more than six or seven sentences.

\subsubsection{Sections and Subsections}

Section headings should be numbered, flush left, and set in 11~pt bold type with the content words capitalized. Leave 0.25~inches of space before the heading and 0.15~inches after the heading.

Similarly, subsection headings should be numbered, flush left, and set in 10~pt bold type with the content words capitalized. Leave 0.2~inches of space before the heading and 0.13~inches afterward.

Finally, subsubsection headings should be numbered, flush left, and set in 10~pt small caps with the content words capitalized. Leave 0.18~inches of space before the heading and 0.1~inches after the heading. Please use no more than three levels of headings.

\subsubsection{Paragraphs and Footnotes}

Within each section or subsection, you should further partition the paper into paragraphs. Do not indent the first line of a given paragraph, but insert a blank line between succeeding ones.

You can use footnotes\footnote{For the sake of readability, footnotes should be complete sentences.} to provide readers with additional information about a topic without interrupting the flow of the paper. Indicate footnotes with a number in the text where the point is most relevant. Place the footnote in 9~point type at the bottom of the column in which it appears. Precede the first footnote in a column with a horizontal rule of 0.8~inches.\footnote{Multiple footnotes can appear in each column, in the same order as they appear in the text, but spread them across columns and pages if possible.}

\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{icml_numpapers}}
\caption{Historical locations and number of accepted papers for International
  Machine Learning Conferences (ICML 1993 -- ICML 2008) and
  International Workshops on Machine Learning (ML 1988 -- ML
  1992). At the time this figure was produced, the number of
  accepted papers for ICML 2008 was unknown and instead estimated.}
\label{icml-historical}
\end{center}
\vskip -0.2in
\end{figure}

\subsection{Figures}

You may want to include figures in the paper to help readers visualize your approach and your results. Such artwork should be centered, legible, and separated from the text. Lines should be dark and at least 0.5~points thick for purposes of reproduction, and text should not appear on a gray background.

Label all distinct components of each figure. If the figure takes the form of a graph, then give a name for each axis and include a legend that briefly describes each curve. Do not include a title inside the figure; instead, the caption should serve this function.

Number figures sequentially, placing the figure number and caption {\it after\/} the graphics, with at least 0.1~inches of space before the caption and 0.1~inches after it, as in
Figure~\ref{icml-historical}.  The figure caption should be set in 9~point type and centered unless it runs two or more lines, in which case it should be flush left.  You may float figures to the top or bottom of a column, and you may set wide figures across both columns (use the environment {\tt figure*}), but always place two-column figures at the top or bottom of the page.

\subsection{Algorithms}

Please use the ``algorithm'' and ``algorithmic''
environments to format pseudocode in \LaTeX\/. These require
the corresponding stylefiles, algorithm.sty and
algorithmic.sty, which are supplied with this package.
Algorithm~\ref{alg:example} shows an example.

\begin{algorithm}[tb]
   \caption{Bubble Sort}
   \label{alg:example}
\begin{algorithmic}
   \STATE {\bfseries Input:} data $x_i$, size $m$
   \REPEAT
   \STATE Initialize $noChange = true$.
   \FOR{$i=1$ {\bfseries to} $m-1$}
   \IF{$x_i > x_{i+1}$}
   \STATE Swap $x_i$ and $x_{i+1}$
   \STATE $noChange = false$
   \ENDIF
   \ENDFOR
   \UNTIL{$noChange$ is $true$}
\end{algorithmic}
\end{algorithm}

\subsection{Tables}

You may also want to include tables that summarize material. Like figures, these should be centered, legible, and numbered consecutively. However, place the title {\it above\/} the table with at least 0.1~inches of space before the title and the same after it, as in Table~\ref{sample-table}. The table title should be set in 9~point type and centered unless it runs two or more lines, in which case it should be flush left.

% Note use of \abovespace and \belowspace to get reasonable spacing
% above and below tabular lines.

\begin{table}[t]
\caption{Classification accuracies for naive Bayes and flexible
Bayes on various data sets.}
\label{sample-table}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}
\begin{tabular}{lcccr}
\hline
\abovespace\belowspace
Data set & Naive & Flexible & Better? \\
\hline
\abovespace
Breast    & 95.9$\pm$ 0.2& 96.7$\pm$ 0.2& $\surd$ \\
Cleveland & 83.3$\pm$ 0.6& 80.0$\pm$ 0.6& $\times$\\
Glass2    & 61.9$\pm$ 1.4& 83.8$\pm$ 0.7& $\surd$ \\
Credit    & 74.8$\pm$ 0.5& 78.3$\pm$ 0.6&         \\
Horse     & 73.3$\pm$ 0.9& 69.7$\pm$ 1.0& $\times$\\
Meta      & 67.1$\pm$ 0.6& 76.5$\pm$ 0.5& $\surd$ \\
Pima      & 75.1$\pm$ 0.6& 73.9$\pm$ 0.5&         \\
\belowspace
Vehicle   & 44.9$\pm$ 0.6& 61.5$\pm$ 0.4& $\surd$ \\
\hline
\end{tabular}
\end{sc}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

Tables contain textual material that can be typeset, as contrasted with figures, which contain graphical material that must be drawn. Specify the contents of each row and column in the table's topmost row. Again, you may float tables to a column's top or bottom, and set wide tables across both columns, but place two-column tables at the top or bottom of the page.

\subsection{Citations and References}
Please use APA reference format. Within the \LaTeX\/ bibliographic facility, use {\tt mlapa.sty} and {\tt mlapa.bst} to obtain this format.

Citations within the text should include the authors' last names and year. If the authors' names are included in the sentence, place only the year in parentheses, for example when referencing Rob Schapire's seminal result \yrcite{schapire90}. Otherwise place the entire reference in parentheses with the authors and year separated by a comma \cite{schapire90}.

List multiple references separated by semicolons \cite{kearns89,schapire90,neal93}. Use the `et~al.' construct only for citations with four or more authors or after listing all authors to a publication in an earlier reference.

Use an unnumbered first-level section heading for the references, and use a hanging indent style, with the first line of the reference flush against the left margin and subsequent lines indented by 10 points. The references at the end of this document give examples for journal articles, conference publications, book chapters, books, edited volumes,
technical reports, and dissertations.

Alphabetize references by the surnames of the first authors, with single author entries preceding multiple author entries. Order references for the same authors by year of publication, with the earliest first.

\section*{Acknowledgments}
If a paper is accepted, the final camera-ready version can (and
probably should) include acknowledgements. In this case, please place such acknowledgements in an unnumbered section at the end of the paper. Typically, this will include thanks to reviewers who gave useful comments, to colleagues who contributed to the ideas, and to funding agencies and corporate sponsors that provided financial support.

% In the unusual situation where you want a paper to appear in the
% references without citing it in the main text, use \nocite
\nocite{zinkevich03}

\bibliography{example_paper}
\bibliographystyle{mlapa}

\end{document}

% Version History
% Document updated by Florian Kunneman for BENELEARN 2013
% Document updated by Willem Waegeman for BeneLearn 2012
% Document updated by Peter van der Putten for BeneLearn 2011
% This document was modified by Kurt Driessens for BeneLearn 2010.
% Before it was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This 2009 version was
% created by Kiri Wagstaff, slightly modified from Sam Roweis's 2008
% version, which is slightly modified from Prasad Tadepalli's 2007
% version which is a lightly changed version of the previous year's
% version by Andrew Moore, which was in turn edited from those of
% Kristian Kersting and Codrina Lauth. Alex Smola contributed to the
% algorithmic style files.

### 2013-03-17 Benelearn/src/benelearn.tex
...

@@
 \usepackage{subcaption}
+\usepackage{float}
 
@@
 
+\newfloat{pseudo}{h}{lop}
+\floatname{pseudo}{Algorithm}
+
 \begin{document}
@@
 We investigate the use of compression-based learning on graph data.
-We hypothesize that the sequential representation of data on which general purpose compressors operate, hampers compression-based learning, as a graph can be represented sequentially in many different, equivalent ways.
+Our hypothesis is that the sequential representation of data on which general purpose compressors operate, hampers compression-based learning, as a graph can be represented sequentially in many different, equivalent ways.
 Using Normalized Compression Distance (NCD), we test various compressors that operate directly on the graph representation against compressors that operate on a bitstring representation.
@@
 
-The idea of compression-based learning is expressed in two strongly related fields: \emph{Minimum Description Length} and \emph{Kolmogorov Complexity}.
-Minimum Description Length \cite{}@!
-builds a statistical framework on the principle that a good model is one that can be used to describe the data in as few bits as possible.
-Kolmogorov Complexity is concerned with the mathematical expression of the \emph{information content} of a piece of data.
+The idea of compression-based learning is expressed in two related fields: \emph{Minimum Description Length} and \emph{Kolmogorov Complexity}.
+Minimum Description Length \cite{grunwald2007minimum} builds a statistical framework on the principle that a good model is one that can be used to describe the data in as few bits as possible.
+Kolmogorov Complexity \cite{li1997introduction} is concerned with the mathematical expression of the \emph{information content} of  data.
 If we can find some short description of a dataset (ie.
@@
 From the many methods born from Kolmogorov Complexity and Minimum Description Length, we choose \emph{Normalized Compression Distance} (NCD) as a representative method of compression-based learning.
-The choice is a practical one, NCD is a simple method which requires nothing more than a compressor fit for the type of data under question.
+The choice is a practical one: NCD is a simple method which requires nothing more than a general purpose compressor.
 Any domain-specific knowledge we wish to use about our data (eg.
-it represents a graph) must be placed in the compressor.
+it represents a graph) can be placed in the compressor.
 We proceed under the assumption that conclusions reached about the performance of NCD on graph data will translate to MDL and other frameworks.
@@
 We will provide a brief, intuitive explanation of the principles involved, sufficient to understand the ideas presented in this paper.
-For a more in-depth and rigorous treatment we refer the reader to @!
-and @!.
+For a more in-depth and rigorous treatment we refer the reader to \cite{li2004similarity} and \cite{cilibrasi2005clustering}.
+For a general introduction to Kolmogorov complexity, see \cite{li1997introduction}.
 
@@
 
-Kolmogorov complexity is a notion of information content, based on two principles: (a) all data can be represented as a bitstring (b) the shorter this bitstring can be described, the less information is contained in it.
+Kolmogorov complexity is a notion of information content, based on two principles: (a) all data can be represented as a bitstring (b) the shorter this string can be described, the less information is contained in it.
 
@@
 
-This intuition prompted Li and Vitanyi @!
-\cite{} to investigate the use of Kolmogorov Complexity as a metric.
-The first issue is that $K(\cdot\mid \cdot)$ is not symmetric.
-It takes a small program to blank out the collected works of Shakespeare, but the reverse is a more complex operation.
+This intuition prompted Li and Vit{\'a}nyi \cite{li2004similarity} to investigate the use of Kolmogorov Complexity as a metric.
+The first issue is that $K(\cdot\mid \cdot)$ is not symmetric: it takes a small program to blank out the collected works of Shakespeare, but the reverse is a more complex operation.
 The first step is thus to define the (symmetric) Information Distance:
@@
 
+When our data is represented as a graph, rather than a string, we replace the notion of concatenation strings by concatenation of graphs.
+That is, we combine the graphs $x$ and $y$ into a single (disconnected) graph.
+
 \section*{Method}
@@
 To ascertain the performance of the compressors, we generate graphs from different sources, calculate their distances and see whether a clustering algorithm can reconstruct the original sources as clusters.
+Source code for these experiments is available \footnote{@!@!}
 
@@
 We generate graphs from two models.
-The first is the classic Erd\H{o}s-R\'enyi model, where a uniform random choice is made from all graphs with $n$ nodes and $m$ links.
-The second is the Barab\'asi-Albert model \cite{}@!, which grows a graph from a set of $n_0$ unconnected nodes, one node at a time, connecting each new node to $k$ distinct existing nodes where the probability that a given existing node is chosen for a connection is its degree, divided by the sum of the degrees of all nodes.
+The first is the classic Erd\H{o}s-R\'enyi (ER) model, where a uniform random choice is made from all graphs with $n$ nodes and $m$ links.
+The second is the Barab\'asi-Albert (BA) model \cite{albert2002statistical}, which grows a graph from a set of $n_0$ unconnected nodes, one node at a time, connecting each new node to $k$ distinct existing nodes where the probability that a given existing node is chosen for a connection is its degree, divided by the sum of the degrees of all nodes.
 Thus, under the BA model the more links a node has, the higher the probability that it will accrue even more.
@@
 
-Since we want there to be some challenge in separating the two classes of network, we, ensure that they have the same number of nodes and links.
+Since we want there to be some challenge in separating the two classes of network, we ensure that they have the same number of nodes and links.
 To achieve this, we first generate the BA networks, count their nodes and links and use these as parameters for the ER model.
@@
 Once we have this dataset, calculate the NCD with a given compressor for every pair of graphs in the dataset, giving us a symmetric matrix.
-We use the k-medioid algorithm to cluster this set into two clusters, and we check whether graphs from a given source tend to end up in a specific cluster.
+We use the k-medoids algorithm to cluster this set into two clusters, and we check whether graphs from a given source tend to end up in a specific cluster.
+
+For second experiment with larger graphs, we can include the fractal graph generation algorithm from \cite{song2006origins}.
+We set the hub-parameter which determines the level of fractality (as a trade-off with the level of small worldness) to $0.1$ (for a small world network) and to $0.9$ (for a mostly fractal network).
+
+\subsubsection*{Parameter settings}
+
+@!
+
+For all graphs we shuffle the node order before adding them to the dataset.
 
@@
 \begin{description}
-	\item[web]
-	\item[metabolic] 
-	\item[semantic A] A semantic description of 
+	\item[web] A graph snapshot of a section of the world wide web.
+All nodes and edges have the same label.
+(This is roughly the same as not having labels, in terms of compression.
+To make the compressor as general as possible we choose some canonical string as the label for all nodes and links).
+	\item[metabolic] A metabolic network. @!
+	\item[semantic A] A semantic description of researchers at the @! institute \ldots 
 	\item[semantic B]
@@
 
-Some basic properties of these graphs are given in tabel \ref{table:properties}.
+Some basic properties of these graphs are given in table \ref{table:properties}.
 
@@
 
-Subdue \cite{} is an algorithm for finding frequent subgraphs in graph data. The algorithm searches through all possible subgraphs, searching for the  one that maximally compresses the data.
+Subdue \cite{ketkar2005subdue} is an algorithm for finding frequent subgraphs in graph data.
+The algorithm searches through all possible subgraphs, searching for the  one that maximally compresses the data.
+The algorithm is essentially a beam search through the space of subgraphs.
+
+Each subgraph serves as a model for compression by finding all occurrences of the subgraph in the data.
+Each occurrence is replaced by a single node, with the edges to the nodes containing annotations on how the edge shouldbe connected within the subgraph.
+We then store the subgrap once, together with the edited datagraph, to represent the original data.
+If the subgraph occurs often enough to offset the extra bits required to annotate the edges in the datagraph, the Subdue representation will be more efficient.
+
+The following is a rough description of the algorithm:
+
+\begin{pseudo}[h]
+{
+	$G$: the data graph \\
+	$b$: the beam size \\
+	\\
+	$S \leftarrow [K_1]$ \textit{\#  initialize the list of substructures with a graph of a single node} \\
+	\\
+	\textbf{loop} \\
+	\hspace*{5mm} $s \leftarrow $ \textit{head element of} $S$ \\
+	\hspace*{5mm} \textit{add all extensions of} $s$ \textbf{to} $S$ \\
+
+	\hspace*{5mm}\textit{sort $S$ by score($s', G$)} for $s' \in S$\\
+	\hspace*{5mm}\textit{remove all but the first $b$ elements of $S$} \\
+	\\
+	\textbf{function } $(s', G)$\\
+	\hspace*{5mm} \textit{replace occurrences of $s'$ in $G$ with node $N$}\\
+	\hspace*{5mm} \textit{annotate links to $N$ with the nodes in $s'$}\\
+	\hspace*{5mm} \textbf{return} \textit{number of bits to store the edited $G$ and $s'$ }\\
+}
+	\caption{Pseudocode for the Subdue algorithm}
+\end{pseudo}
 
@@
 \end{subfigure}
+\vspace{3mm}
+
 \begin{subfigure}[b]{1\columnwidth}
@@
 \begin{subfigure}[b]{1\columnwidth}
+\centering
 \begin{tabular}{l | r | r r r r}
 \hline
-  ER      & @!/@!  & @! & @! & @! & @! \\
-  BA      & @!/@!  & @! & @! & @! & @!\\
-  fractal (pure) & @!/@!  & @! & @! & @! & @!\\
-  fractal (small world) & @!/@!  & @! & @! & @! & @!\\
+  Web      & @!/@!  & @! & @! & @! & @! \\
+  Metabolic     & @!/@!  & @! & @! & @! & @!\\
+  Semantic A & @!/@!  & @! & @! & @! & @!\\
+  Semantic B & @!/@!  & @! & @! & @! & @!\\
 \hline
 \end{tabular}
-\caption{GZIP: error @!}
+\caption{GZIP: error @! (@!)}
 \end{subfigure}
+\vspace{3mm}
+
 \begin{subfigure}[b]{1\columnwidth}
+\centering
 \begin{tabular}{l | r | r r r r}
 \hline
-  ER      & @!/@!  & @! & @! & @! & @! \\
-  BA      & @!/@!  & @! & @! & @! & @!\\
-  fractal (pure) & @!/@!  & @! & @! & @! & @!\\
-  fractal (small world)& @!/@! & @! & @! & @! & @!\\
+  Web      & @!/@!  & @! & @! & @! & @! \\
+  Metabolic     & @!/@!  & @! & @! & @! & @!\\
+  Semantic A & @!/@!  & @! & @! & @! & @!\\
+  Semantic B & @!/@! & @! & @! & @! & @!\\
 \hline
 \end{tabular}
-\caption{Subdue: error @!}
+\caption{Subdue: error @! (@!)}
 \end{subfigure}
@@
 
-This result suggests that inference on graphs can be performed by sequential algorithms on a sequential representation in linear time, with decent results.
+This result suggests that inference on graphs can be performed by sequential algorithms on a sequential representation in linear time, with decent results, depending on the complexity of the task.
 
@@
 
+\bibliographystyle{plain} @!
+\bibliography{benelearn}
+
 \end{document}

### 2013-04-19 Overview/src/overview.tex
...

\documentclass[10pt,a4paper,oneside]{article}

\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

\theoremstyle{definition}
\newtheorem*{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem*{dfn}{Definition}
\newtheorem*{exm}{Example}

\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{argmax}

\title{The question of interesting information}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle

\begin{abstract}
\noindent Ever since the introduction of absolute measures of information content---such as Information Entropy and Kolmogorov Complexity---there have been  criticisms of the idea that these measures capture the informal notion of  information. The argument is that the signals which, under these measures, contain the highest amount of information are by definition completely unpredictable, and therefore cannot contain any interesting information. There is an informal notion of information content, which is not captured by these absolutes, a kind of useful or meaningful information content. We review the literature available on this topic and provide a distillation of common themes and possible connections.
\end{abstract}

\section*{Introduction}

Consider the following, illustrative example. An old-fashioned, black-and-white television set is turned off. The screen is black. This signal represents a complete lack of information. It is highly predictable. At every point we can predict exactly the next state of the signal. Describing this signal would be very simple.

We turn the television on, and are faced with white noise. This represents the other extreme, of a signal as rich in information as possible. for each point on the screen at each point in time, we have no wa of prediction its next value. To describe this signal, we have no other option than to describe every value of every point of the screen for every point in time.

Finally, we tune the television to a news cast. This signal lies somewhere in between the two extremes. Most of the signal is precictable in some way. It contains many patterns. We aid prediction by modeling the average of each grayscale value. We can model common elements in the signal like faces, or the news channel's logo. We can segment the signal and model sections of the news anchor speaking to camera separately from reports on location. All these options showthat we have an almost infite variety of patterns to exploit, but also that there are many sudden shifts in how well each strategy will work, and that some part of the signal will still be unpredictable.

For the first two situations in this example, we have solid, well understood mathematical formalizations. If the signal consists of single elements which are transmitted one after the other from an iid source, we can use Information Entropy \cite{}@!, which will tell us the exact number of bits we will need for each element transmitted. If a finite sequence of elements is viewed as a whole, with all dependencies between transmitted elements considered, we use Kolmogorov Complexity \cite{} @!. In both cases, the predictability of the signal gives us a spectrum, with the blank screen on one end, and white noise on the other.

The third example has no formalization that is well-understood and broadly agreed on. There is a strong intuition pervading the literature that signals or data in between the two extremes maximize some mixture of predicatbility and unpredictability. That both the presence of useful patterns and the sudden disappearance of these patterns indicate that the data is in some way interesting. There are many open questions. Can this intuitive notion be effectively formalized? Is there a single, objective function that is maximized by the most interesting data. Is there a single measure that can  be convincingly argued to supercede or encompass all competing notions? What reasons do we have to suggest that `interestingness' can be captured in an objective manner?

\begin{quotation}
A fundamentalproblem for statistical mechanics is to explain why dissipative systems (those in which entropy is continually being produced and removed to the surroundings) tend to undergo ``self-organisation,'' a spontaneous increase of structural complexity, of which the most extreme example is the origin and evolution of life. The converse principle, namely that nothing very interesting is likely to happen in a system in thermal equilibrium, is reflected in the term ``heat death.'' In the modern world view, thermodynamic driving forces, such as the temperature difference between the hot sun and the cold night sky, have taken over one of the functions of God: they make matter transcend its clod-like nature andbehave instead in dramatic and unforeseen ways, for example molding itself into thunderstorms, people, and umbrellas.
\end{quotation}

\subsection*{A note on terminology}

Complexity, order, etc.
\section*{Preliminaries}

\subsection*{Information Entropy}

\subsection*{Kolmogorov Complexity}

\subsection*{Two-part coding}

\subsection*{Dynamical systems}

\section*{Questions to ask}

\begin{itemize}
  \item Level of formalization
  \item The objectivity of interestingness
  \item Broadness. What types of data are covered, and what models and patterns are covered? 
  \item Context-dependence. Context plays an undeniable part in any notion of beauty or interestingness. Are contextual considerations acknowledged and given a place in the framework?
\end{itemize}

\section*{Measures of interesting information}

\subsection*{The Kolmogorov structure functions}
[!@ Chronological order. Each presented in an intuitive manner first, followed by the formalization.]

\subsection*{Aesthetic measures}

Some of the earliest ideas that relate to the question of interesting information have come from the field of generative art. This should not be surprising: if we can reverse a measure for interestingness, we get an algorithm for generating interesting data. Formalizing the aspects that make art interesting, can help us achieve both a measure of interestingness and a generative algorithm for producing interesting things.

The formal study of aesthetics is a sprawling field, a complete review of which is out of scope for this paper. While aethetics is related to the notion of interesting information, they do not mean the exact same thing. Specifically, aesthetics is primarly concerned with a formalization of human cognition, whereas one of the assumptions underlying the question of interesting information seems to be that the notion of aesthetic beauty emerges from a more general objective princpiple of interestingness. We mention aesthetics here mainly to provide historical context, and to show some of the recurring themes appearing in early work.

One interesting line of research started with G.D. Birkhoffs suggestion of an aesthetic measure. He proposed the following pseudoformula to express the level of aesthetic quality of an object:
\[
M\; \mbox{aesthetic measure}= \frac{O\; \mbox{order}}{C\; \mbox{complexity}}
\]
The specific methods used to measure each value differ per domain. Birkhoff himself gave the example of regular and irregular polygons as an example. In this case the polygons with low complexity were those with a low number of sides and high order was measured by a high number of symmetries. Using these principles Birkhoff calculated that the square was the polygon with the greatest aesthetic measure.

Attempts to verify this hypothesis have revealed some interesting results. \cite{mcwhinnie1968review} provides a review of such studies, showing, among other things, that

law of Prägnanz

\cite{bense1969einfuhrung, birkhoff1950collected, mcwhinnie1968review}

\subsection*{Incomplete sequence}

1977 Levin \& V'jugin

\subsection*{Hitting time}

1984 Levin

\subsection*{Generalized Kolmogorov Complexity and Algorithmic information}

1983 Hartmanis

\subsection*{Self-generated complexity and computation at the onset of chaos}

One of the earliest suggestions of a balance between highly structured and highly unstructured signals appears in literature from the field of dynamical systems. In \cite{}, Crutchfield characterizes the two extremes as \emph{the clock} and \emph{the coin}, with a clock providing a highly regular and structured signal and a coinflip being the classic example of a fully random process.

1986 Peter Grassberger

1990 Crutchfield \& Young

\cite{crutchfield1989inferring, crutchfield1990computation, feldman1998measures, crutchfield1983symbolic}

A review by Cructhfield: \cite{crutchfield2011between}

\cite{grassberger1986toward}

\subsection*{Sophistication}

1987 \cite{kopel1987complexity}

\subsection*{Logical depth}

1995 Bennett

\cite{bennett1995logical}

\subsection*{Effective Complexity}

2003 Gell-Man \& Lloyd

\cite{gell1996information}

\subsection*{Meaningful information}

2006 Vitanyi

\cite{vitanyi2006meaningful}

\subsection*{Computational depth}

2006 Antunes \& Fortnow

\cite{antunes2006computational, antunes2003sophistication}

\subsection*{Self-dissimilarity}

2007 Wolpert \& Macready 
\cite{wolpert2000self, wolpert2007using}

\subsection*{Facticity}

2011
\cite{adriaans2012facticity}

\subsection*{Algorithmic rate distortion}

2012 de Rooij \& Vitanyi

\cite{de2012approximating}

\section*{Common themes and connections}

Effective complexity and its relation to logical depth \cite{ay2010effective}

\subsection*{Common themes}
\subsection*{The connection between model size and depth}

\section*{Future directions}

\subsection*{Applications}

@! `philosophical' implications
@! Statistics
@! practical applications, ranking, search target. Data with `interesting' ad-hoc codes.

\nocite{*}

\bibliographystyle{siam}
\bibliography{overview}

\end{document}

### 2013-04-23 Bastards/src/bastards.tex
...

\documentclass[10pt,a4paper,oneside]{article}

\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

%\theoremstyle{definition}
\newtheorem{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem{dfn}{Definition}
\newtheorem{exm}{Example}

\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{argmax}

\title{Bounds on the approximation of Kolmogorov complexity}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle

\section*{preliminaries}
Let $\mathbb B$ be the set of all finite binary strings $\{0,1\}^*$.

Let {\cal T} be the set of all finite turing machines.

\subsection*{probability distributions}

We concern ourselves with probability distributions on the set of finite strings. All measure discussed are discrete and have $\mathbb B$ or a subset as support.

\begin{dfn}
A \emph{semimeasure} is a function $p : {\mathbb B} \rightarrow {\mathbb R}$ with $\sum_{x \in \mathbb B} x \leq 1$. If this sum is equal to $1$, $p$ is a \emph{measure}, which we will refer to a a probability distribution.
\end{dfn}

\begin{dfn}
Semicomputability. A function $f : {\mathbb B} \rightarrow {\mathbb R}$ is \emph{lower semicomputable} iff there exists a computable two-argument function $f': {\mathbb B} \times {\mathbb N} \rightarrow \mathbb Q$ such that  
\[
\lim_{i \rightarrow \infty} f'(x, i) = f(x)
\] and $f'(x, i + 1) \geq f'(x, i)$, ie. $f'$ never decreases with $i$. If the first condition is the same, but $f'$ never increases with $i$, the function is \emph{upper semicomputable}.
\end{dfn}

Put simply, we can approximate the semicomputable $f(\cdot)$ to any precision by computing $f'$ with high enough $i$. Any function that is both upper- and lower semicomputable is computable.

Since probability distributions are functions from $\mathbb B$ into $\mathBB R$, we talk about (semi)computable probability distributions and (semi)computable semimeasures.

\subsection*{Kolmogorov Complexity}

We define $q^*$ as the shortest program implementing this function \footnote{Note that $|q^*|$ is different from $C(q)$. For instance, a program may encode the collected works of Shakespeare in padding. $C(q)$ must reflect this, whereas to $q^*$ this is non-functional code which can be removed.} and $\tilde{q} = |q| - |q^*|$ as the penalty we face for not using the optimal program.

\section*{sampling computable distributions}

\begin{dfn}
A \emph{probabilistic Turing machine} (PTM) is defined as a regular Turing machine with the addition of a tape filled with random bits. This tape can only be read, and the tape head can only read in one direction.
\end{dfn}

A given PTM thus produces a random output for no input. In this sense we can see it as a machine for sampling from some distribution $p_q$.

However, the constraint that the head moves only in one direction also allows us to see the random tape as an input tape. After each bit read from the random tape, the machine either continues to another bit, or halts. This means that the inputs for which the machine halts form a set of prefix codes, making every PTM a prefix Turing machine.

We say that under the prefix-machine perspective, the Turing machine does not halt if it is presented with fewer bits input than it requires. If it is presented with more bits than required, the remaining bits are not counted as part of the input.

\begin{lma}
Every PTM samples from a lower semicomputable measure or semimeasure.
\end{lma}
\begin{proof}
Let $q$ be any PTM. We will define a program $p_q'(y, i)$ to approximate the distribution $p_q(y)$ from which $q$ samples.

Dovetail the computation of $q$ on all inputs $x \in \mathbb B$ for $i$ cycles. Calculate $\sum_{x : q(x) \downharpoonright^i y}2^{-|x|}$. Ie. the sum of the probabilities for all inputs that have halted so far on output $y$. Return this sum.

Clearly this value can never decrease with $i$. To show that this function reaches $p_q(y)$ in the limit, we note that $p_q(y)$ is the sum of the probabilities of all inputs that produce $y$,
\[
p_q(y) = \sum_{x : q(x) = y} 2^{-|x|}
\]

whereas $p_q(y, i)$ is the sum of all probabilities of all inputs that produce $y$ within $i$ cycles of a dovetailing computation.
\[
p_q'(y, i) = \sum_{x : q(x) \downharpoonright^i y} 2^{-|x|}
\]

For each $i$ we can expand $p_q(x) - p_q'(x, i)$ into its sums and cross out terms where the $x$s match. Let $\epsilon$ be some positive value. If we order all inputs by length, and sum the probabilities of the first $j$, we know that every next code will add as much as or less than the last one. The total sum, we know, converges to a finite value $t_q$. \ldots

\end{proof}

\begin{lma}
Every lower semicomputable measure or semimeasure can be sampled by a PTM.
\end{lma}
\begin{proof}
We will construct a program to sample from a semicomputable distribution $p_q$.

Let $p_q'(y, i)$ be the function that approximates $p_q(y)$. We construct a binary tree $B$, where we will assign nodes $x$ to output values $y$. If $x$ is assigned to $y$, we say $B(x) = $. We want the tree to converge to a prefix tree, so once a node is assigned, all nodes below it become unavailable.

We dovetail the computation of $p_q'(y, i)$. After each step we calculate the value $A(y) = \sum_{x \in B} B(x)$ for all $y$. If $A(y)$ is less than the last value of $p_q(y, i)$ calculated, we calculate the difference $\rho_y$ and find the first available  node $x$ in $B$ with depth $|x|$ such that $2^{-|x|} < \rho_y$.

In the limit, this process generates a many-to-one prefix coding that maps the distribution $p(x) = 2^{-|x|}$ to $p_q(y)$.

Since the assignments made are never reversed, we can run the process until an assigned node matches the prefix code on our input tape, and return the assigned $y$.

[@! If we want to prove correctness also, we'll need to clean this up. I think there was a section in Cover \& Thomas about simulating any distribution with a coinflip. May save some time to refer to that.]
\end{proof}

@! With high likelihood, the sampled string will have low randomness deficiency for $q$ as a model

\begin{dfn}
The randomness-deficiency of a finite set $A$ is 
\[
\delta(x|A) = \log |A| - C(x|A)
\]
\end{dfn}

\begin{lma}[{\cite{li1993introduction}[p 120]}]
For any finite set $A$, the proportion of strings with randomness deficieny above $k$ is less than $2^-k$. 
\end{lma}

\begin{proof}
\begin{align*}
\log |A| - C(x|A) &< k \\
C(x|A) &> \log |A| - k 
\end{align*}

Since there are only $2^m-1$ descriptions of length below $m$, There can be no more than $2^m-1$ strings x with $C(x|\cdot) < m$. This means that the number of strings $N_k$ with randomness-deficiency below $k$ is

\begin{align*}
N_k &< 2^{\log |A| - k}-1 \\
&= \frac{|A|}{2^k} - 1
\end{align*}

Finally, we bound the proportion:
\begin{align*}
\frac{N_k}{|A|} < \frac{|A|/2^k}{|A|} =  2^{-k}
\end{align*}
\end{proof}

As a sanity check, we note that all strings have randomness deficiency at least $0$. The exponential decay with $k$ tells us that if we sample a string at random from $A$, we will likely end up with a string with low $k$. In other words, sampling from $A$ will likely give us a string that is typical for $A$.

\begin{dfn}[{\cite{antunes2009depth}}]
The randomness-deficiency of a string $x$ given a probability distribution $p_q$ is 
\begin{align*}
\delta(x|p_q) &= \log \frac{p_u(x)}{p_q(x)}
\end{align*} 
Where $p_u$ is the universal distribution.

We can rewrite this to 
\begin{align*}
\delta(x|p_q) &= - \log(p_q(x)) - C(x) 
\end{align*}

(Note that unlike the finite set case, here we do not make the complexity term conditional on the model)

\end{dfn}

We would like to show the same property we proved for the randomness deficiency of finite sets. That sampling from $p_q$ will likely produce a string that is typical for $p_q$, ie. has low randomness deficiency.

\begin{lma}
The probability of all strings with randomness deficiency above $k$ is less than $2^-k$
\end{lma}

\begin{proof}
Let 
\begin{align*}
X_k^q &= \left\{x : \delta(x|q) > k \right\}
\end{align*}
ie. the set of all strings with randomness deficiency larger than $k$. We have 
\begin{align*}
p_q(X_k^q) &= \sum_{x:\delta(x|q)>k}p_q(x)\\
    &= \sum_{x: -\log p_q(x) - K(x) > k}p_q(x)\\
    &= \sum_{x:  p_q(x) < 2^{-k}2^{-K(x)}} p_q(x)\\
    &< \sum_{x:\delta(x|q) > k} 2^{-k} 2^{-K(x)}\\
    &= 2^{-k}\sum_{x:\delta(x|q) > k} 2^{-K(x)}
\end{align*}

Since the sum in the last line is less than or equal to $1$ (as the same sum over all $x$ sums to 1), we obtain

\begin{align*}
p_q(X_k^q) &< 2^{-k}
\end{align*}
\end{proof}

\begin{exm}
What is the probability that we sample a string $x$ with a randomness deficiency of at least half its length? In other words, if we use the optimal description, we gain $|x|/2$ over describing it in terms of $p_q$.
\end{exm}

\subsection*{the universal distribution}

\section*{resource bounded probability distributions}

We define the complexity class of a PTM as normal, as a function of the input length. In the prefix-machine perspective, this suggests that the computational power represented by a probability distribution is a function of the number of `bits of randomness' it uses.

Let $\cal C$ be the set of all Turing machines. Let ${\cal C}^t$ be the set of all Turing machines whose running time is bounded by $t(n)$ where $n$ is the length of the input.

\subsection*{}

\section*{Inference from bounded distributions}
\nocite{*}
\bibliographystyle{siam}
\bibliography{bastards}

\end{document}

### 2013-04-23 Benelearn/src/benelearn.tex
...

@@
-\documentclass[10pt,a4paper,oneside]{article}
+\documentclass{article}
 
@@
 \usepackage{caption}
+\DeclareCaptionType{copyrightbox}
 \usepackage{subcaption}
 \usepackage{float}
+\usepackage{style/mlapa}
+\usepackage{style/benelearn2013}
 
@@
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+\usepackage{url}
 
@@
 
-\title{Compression-based inference on graph data}
-\date{\today}
+%\date{\today}
 
@@
 
+\benelearntitlerunning{Compression-based inference on graph data}
+
 \begin{document}
 
-\maketitle
+\twocolumn[
+  \benelearntitle{Compression-based inference on graph data}
+
+  \benelearnauthor{Peter Bloem}{p@peterbloem.nl}
+  \benelearnaddress{System and Network Engineering Group, University of Amsterdam,
+   Science Park 904, 1098 XH Amsterdam, The Netherlands}
+
+  \benelearnaddress{{\bf Keywords}: 
+	compression, machine learning, graphs, Kolmogorov complexity, minimum description length, normalized compression distance, subdue}
+\vskip 0.3in
+]
 
 \begin{abstract}
+
 We investigate the use of compression-based learning on graph data.
@@
 We use both synthetic, randomly generated graphs and real-life datasets.
+
 \end{abstract}
@@
 
-The idea of compression-based learning is expressed in two related fields: \emph{Minimum Description Length} and \emph{Kolmogorov Complexity}.
+The idea of compression-based learning is expressed in two related subjects: \emph{Minimum Description Length} and \emph{Kolmogorov Complexity}.
 Minimum Description Length \cite{grunwald2007minimum} builds a statistical framework on the principle that a good model is one that can be used to describe the data in as few bits as possible.
 Kolmogorov Complexity \cite{li1997introduction} is concerned with the mathematical expression of the \emph{information content} of  data.
-If we can find some short description of a dataset (ie.
+It states that if we can find some short description of a dataset (ie.
 compress it) then the total information content must be below the length of that description.
@@
 The graph is a powerful and versatile representation.
-Most applications of compression-based learning choose sequential models, such as deterministic finite automata, or general purpose compressors, which operate on bitstrings.
-If we have data in the form of a graph, we can translate it to a bitstring, of course, but in this transformation we complicate our problem.
-For a different ordering of nodes, the graph is the same but the bitstring changes radically.
+Most applications of compression-based learning choose sequential models, such as deterministic finite automata or block-sorting compressors, which operate on bitstrings.
+If we have data in the form of a graph, we can translate it to a bitstring, but in this transformation we complicate our problem.
+For a different ordering of nodes, the graph is the same, but the bitstring changes radically.
 To a simple compressor, the two bitstrings may not share very much structure, even though they represent the same graph.
@@
 If this problem holds, we should investigate compressors that operate on the level of the graph representation, for instance by finding frequent subgraphs or clustering the graph.
-The methods do not suffer the problem of isomorphism, but as a result, they are more expensive than their sequential cousings.
+The methods do not suffer this problem of isomorphism, but as a result, they are more expensive than their sequential cousins.
 In this paper we hope to provide a first indication of how far the sequential approach will go, and whether the native approach will let us continue on from there.
@@
 Any domain-specific knowledge we wish to use about our data (eg.
-it represents a graph) can be placed in the compressor.
+it represents a graph) can be added to the compressor.
 We proceed under the assumption that conclusions reached about the performance of NCD on graph data will translate to MDL and other frameworks.
@@
 \] 
-That is, if $U$'s input consists two bitstring arguments $i$ and $p$, combined with some computable pairing function $\langle \cdot, \cdot\rangle$, then $U$ computes the same function as $T_i$ on input $p$ if $T_i(p)$ halts, or fails to halt if $T_i$ fails to halt.
-$U$ provides our formalization of 'description'.
+That is, if $U$'s input consists of two bitstring arguments $i$ and $p$, combined with some computable pairing function $\langle \cdot, \cdot\rangle$, then $U$ computes the same function as $T_i$ on input $p$ if $T_i(p)$ halts, or fails to halt if $T_i$ fails to halt.
+$U$ provides our true formalization of `description'.
 If our data is $x$, and for some $y = \langle i, p\rangle$, $U(y) = x$, we say that $y$ is a description of $x$ on our reference universal Turing machine.
@@
 \]
+
+The result that makes Kolmogorov Complexity a useful measure of information content is that $K_U(\cdot)$ is only marginally dependent on the choice of $U$.
+If we suppose that there is another universal description method, $V$, we might ask what the expected difference is between $K_U(x)$ and $K_U(v)$.
+Let $k_V(x)$ be the shortest program for $x$ on $V$.
+Since, $U$ is universal, we know that is can compute $k_v$ by simulating $V$.
+This simulation is a program for $x$ on $U$ which is a bound for the shortest program on $U$:
+\begin{align*}
+K_U(x) 	&\leq |\langle T_v, k_v(x)\rangle|\\ 
+		&=  |k_v(x)| + |v| + p(v) \\
+		&=  K_V(x) + O(1)
+\end{align*}
+
+Where $p(v)$ is the penalty (ie.
+the additional bits) that the pairing function reuires to store its two arguments in a separable way.
+We require that this is only dependent on $v$.
+The final line shows that $K_U(\cdot)$ and $K_V(\cdot)$ differ only by a constant term, which is independent of $x$.
+To summarize: we may differ in opinion on how much information our data contains, but only by a constant amount.
 
@@
 \item[$K(\cdot)$ is uncomputable.] There can be no algorithm which computes the Kolmogorov Complexity of $x$ for all $x$.
-It can however be bounded from above, and for every algorithm which bounds it, there is an algorithm which provides a better bound.
+It can, however, be bounded from above, and for every algorithm which bounds it, there is an algorithm which provides a better bound.
 \item[$K(\cdot)$ is approximated by all computable compressors.] If we have some compressor for our data $x$ (say GZIP) we can find the decompression algorithm somewhere in $\{T_i\}$, say as $T_g$.
@@
 Where regular Kolmogorov Complexity is defined as the shortest program which produces $x$, the conditional variant is defined as the shortest program which produces $x$ when provided with $y$.
-A complete treatment is available in \cite{}.
+A complete treatment is available in \cite{li1997introduction}.
 
@@
 
-This intuition prompted Li and Vit{\'a}nyi \cite{li2004similarity} to investigate the use of Kolmogorov Complexity as a metric.
-The first issue is that $K(\cdot\mid \cdot)$ is not symmetric: it takes a small program to blank out the collected works of Shakespeare, but the reverse is a more complex operation.
-The first step is thus to define the (symmetric) Information Distance:
+This intuition prompted Li and Vit{\'a}nyi \cite{li2004similarity} to investigate the use of Kolmogorov Complexity as a metric of computational similarity.
+To acquire a true metric, some problems have to be solved.
+The first is that $K(\cdot\mid \cdot)$ is not symmetric: it takes a small program to blank out the collected works of Shakespeare, but the reverse is a more complex operation.
+The first step, then, is to define the (symmetric) Information Distance:
 \[
@@
 We would like to approximate this by replacing each occurence of the Kolmogorov complexity with an approximation by a compressor, which we will call $C$.
-As most compressors do work on a conditional basis (expressing data given some existing data), we want to rewrite the condition $K$'s as nonconditional ones.
-In doing this, we make first accept beyond the constant term inaccuracy that is innate to Kolmogorov Complexity, a further logarithmic inaccuracy so that we can rewrite as
+As most compressors do not work on a conditional basis (expressing data given some existing data), we want to rewrite the conditional $K$'s as nonconditional ones.
+To achieve this, we accept beyond the constant term uncertainty that is innate to Kolmogorov Complexity, a further logarithmic inaccuracy.
+This allows us to rewrite as
 
@@
 
-\section*{Method}
+\section*{Methods}
 
-Our aim is to test a selection of compressors on a variety of graph data.
+Our aim is to test a sequential and graph-based compressor on a variety of graph data.
 To ascertain the performance of the compressors, we generate graphs from different sources, calculate their distances and see whether a clustering algorithm can reconstruct the original sources as clusters.
-Source code for these experiments is available \footnote{@!@!}
+Datasets and source code for these experiments are available.
+\footnote{\url{http://www.peterbloem.nl/benelearn2013}}
 
@@
 
-Once we have this dataset, calculate the NCD with a given compressor for every pair of graphs in the dataset, giving us a symmetric matrix.
+We also include graphs from the fractal graph generation algorithm from \cite{song2006origins}.
+We set the hub-parameter which determines the level of fractality (as a trade-off with the level of small worldness) to $0.0$ (for a small world network) and to $1.0$ (for a fractal network).
+
+Once we have this dataset, consisting of four gold clusters, we calculate the NCD with a given compressor for every pair of graphs in the dataset, giving us a symmetric matrix.
 We use the k-medoids algorithm to cluster this set into two clusters, and we check whether graphs from a given source tend to end up in a specific cluster.
-
-For second experiment with larger graphs, we can include the fractal graph generation algorithm from \cite{song2006origins}.
-We set the hub-parameter which determines the level of fractality (as a trade-off with the level of small worldness) to $0.1$ (for a small world network) and to $0.9$ (for a mostly fractal network).
-
-\subsubsection*{Parameter settings}
-
-@!
-
-For all graphs we shuffle the node order before adding them to the dataset.
+As a random baseline, we generate a random distance matrix with every distance a uniform random value in $[0, 1)$, and run the clustering algorithm on that.
 
@@
 In this experiment we sample subgraphs from large, existing graphs.
-Again, to make the problem challenging enough to be meaningful, we want the subgraphs to have identical size in both nodes and links.
-To accomplish this, we proceed as follows.
+We sample by choosing a uniform random node, and performing a random walk of length $n$.
+We then extract a subgraph containing the nodes encountered and any links connecting two encountered nodes.
+We replace all node  and link labels with a single canonical symbol.
 
-\begin{itemize}
-  \item Choose a random subgraph by choosing a random node uniformly, and including all nodes within distance $d$. Generate $N$ such graphs for each source.
-  \item Find the subgraph with the smallest number of nodes $n_{\mbox{min}}$ over all subgraphs and, for every other subgraph, remove randomly chosen nodes until its number of nodes matches $n_{\mbox{min}}$
-  \item Find the subgraph with the smallest number of links $l_{\mbox{min}}$ and, for every other subgraph, remove randomly chosen links until the number of links matches $l_{\mbox{min}}$ 
-\end{itemize}
-
-With this dataset of subsets, we proceed as before, calculating the distances between the subgraphs and clustering them into as many clusters as we have sources, to see whether the resulting clusters match the sources.
+With this dataset of subgraphs, we proceed as before, calculating the distances between the subgraphs and clustering them into as many clusters as we have sources, to see whether the resulting clusters match the sources.
 
@@
 \begin{description}
-	\item[web] A graph snapshot of a section of the world wide web.
-All nodes and edges have the same label.
-(This is roughly the same as not having labels, in terms of compression.
-To make the compressor as general as possible we choose some canonical string as the label for all nodes and links).
-	\item[metabolic] A metabolic network. @!
-	\item[semantic A] A semantic description of researchers at the @! institute \ldots 
-	\item[semantic B]
+	\item[cellular] The cellular network of the E. Coli bacterium. \cite{jeong2000large} Acquired from \url{http://www.nd.edu/~networks/resources/metabolic/index.html}
+	\item[neural] The neural network of the C. Elegans nematode worm (ignoring link directions). \cite{achacoso1991ay, watts1998small} Acquired from \url{http://toreopsahl.com/datasets/#celegans}
+	\item[co-purchase] A graph of items commonly purchased together on internet retailer Amazon.com. \cite{leskovec2007dynamics} Acquired from \url{http://snap.stanford.edu/data/amazon0302.html}
 \end{description}
 
-Some basic properties of these graphs are given in table \ref{table:properties}.
+% Some basic properties of these graphs are given in table \ref{table:properties}.
 
-\begin{table}
-\label{table:properties}
-\begin{tabular}{ l | r r r r r}
-\hline
-    & nodes & links & C & $alpha$ & $p_\alpha$ \\
-\hline
-  web & & & & & \\
-  metabolic & & & & & \\
-  semantic A & & & & & \\
-  semantic B & & & & & \\
-
-\hline
-\end{tabular}
-\caption{The basic properties of our natural datasets.
+%\begin{table}
+%\label{table:properties}
+%\begin{tabular}{ l | r r r r r}
+%\hline
+%    & nodes & links & C & $alpha$ & $p_\alpha$ \\
+%\hline
+%  web & & & & & \\
+%  metabolic & & & & & \\
+%  semantic A & & & & & \\
+%  semantic B & & & & & \\
+%  
+%\hline
+%\end{tabular}
+%\caption{The basic properties of our natural datasets.
 We report the number of nodes, the number of links, the number of nodes in the largest component (C), the estimates power law exponent ($\alpha$), and the significance for the power law model ($p_\alpha$, significant for values \emph{above} 0.01).}
 
-\end{table}
+%\end{table}
 
@@
 To store a graph with GZIP, we flatten the lower half of its adjacency matrix into a bitstring and store this together with a list of the node and link labels.
-We use Java object serialization to take care of delimiting the data and translating it to bits.
+We use Java object serialization to take care of delimiting the label data and translating it to bits.
+(Since all graphs in our experiments have a single label, this is unlikely to affect the outcome).
 
@@
 	\\
-	\textbf{function } $(s', G)$\\
+	\textbf{function score} $(s', G)$\\
 	\hspace*{5mm} \textit{replace occurrences of $s'$ in $G$ with node $N$}\\
@@
 
+The graph matching search (the first line of the \textbf{score} function) allows for inexact matches of the subgraph.
+In these cases, use a rough estimate of the number of bits required to transform the stored subgraph into the subgraph that is actually present in the data.
+
+To make the calculation of a distance matrix feasible, we must find a way to reduce the computational costs of the Subdue algorithm.
+Our solution is to replace the exhaustive graph matching search by a non-exhaustive beam search.
+With this change, we cannot be sure that the score assigned to the given subgraph describes the optimal way to compress the data with that subgraph, but it does provide an upper bound.
+
+\subsubsection*{Parameter settings and specifics}
+
+All graphs generated contain $100$ nodes.
+In the BA-model, we attach one node each step, giving $100$ links also.
+Note that this makes the BA graphs UAGs.
+For $k = 2$ the clustering problem would be more difficult.
+The fractal graphs we generate to depth $2$ by adding $4$ ancestors at each side of each link and $1$ extra link between the groups of ancestors.
+This results in networks of $90$ nodes and $100$ links.
+
+For each source, we generate $3$ graphs.
+
+The Subdue algorithm has a lot of parameters.
+Wuring the search we return only one best subgraph.
+Our beamwidth at the top level is set to $5$.
+The beam width in the graph matching routine is set to $10$.
+
+We let the k-medoids algorithm run for $20$ iterations. This is more than enough for convergence in all experiments.
+
 \section*{Results}
 
-\subsection*{Synthetic data}
-
-\subsubsection*{Small graphs}
+Table \ref{table:synthetic-large} shows the results on randomly generated graphs. Table \ref{table:real-life} shows the results for subgraphs sample from real-life datasets.
 
 \begin{table}[h]
-\label{table:synthetic-small}
 
-\begin{subfigure}[b]{0.5\columnwidth}
-	\label{table:gzip-small}
-	\centering
-	\begin{tabular}{l | r | r r}
+\begin{subfigure}[b]{1\columnwidth}
+\begin{tabular}{l |  r r r r}
 \hline
-	  ER & @!/@! & @! & @! \\
-	  BA & @!/@! & @! & @! \\
+  ER                    &  0.083 & 0    & 0.17  & 0 \\
+  BA                    &  0.16  & 0    & 0.083 & 0 \\
+  fractal (pure)        &  0     & 0    & 0.083 & 0.17 \\
+  fractal (small world) &  0     & 0.25 & 0     & 0 \\
 \hline
 \end{tabular}
-	\caption{GZIP: error @!}
-\end{subfigure}\begin{subfigure}[b]{0.5\columnwidth}
-	\label{table:subdue-small}
-	\centering
-	\begin{tabular}{l| r | r r}
-	\hline
-	  ER & @!/@! & @! & @! \\
-	  BA & @!/@! & @! & @! \\
-	\hline
-	\end{tabular}
-	\caption{Subdue: error @!}
-\end{subfigure}
-
-\caption{Confusion matrices for the synthetic data.
-Results for synthetic graphs of exactly @!
-nodes and @!
-links generated by the ER and the BA algorithms.
-For each source we generated @!
-such graphs and clustered the results with different compressors.
-The error is computed by assuming an optimal labeling and computing the symmetric error as in classification.}
-\end{table}
-
-\subsubsection*{Large graphs}
-
-For larger graphs, the problem becomes easier (as there is more structure), but we can add the fractal generator.
-This generator can only produce larger graphs, and with its inclusion we can no longer ensure that all graphs have equal size.
-\begin{table}[h]
-\label{table:synthetic-large}
-\begin{subfigure}[b]{1\columnwidth}
-\begin{tabular}{l | r | r r r r}
-\hline
-  ER      & @!/@!  & @! & @! & @! & @! \\
-  BA      & @!/@!  & @! & @! & @! & @!\\
-  fractal (pure) & @!/@!  & @! & @! & @! & @!\\
-  fractal (small world) & @!/@!  & @! & @! & @! & @!\\
-\hline
-\end{tabular}
-\caption{GZIP: error @!}
+\caption{Random baseline: error 0.46 (0.11)}
 \end{subfigure}
@@
 \begin{subfigure}[b]{1\columnwidth}
-\begin{tabular}{l | r | r r r r}
+\begin{tabular}{l |  r r r r}
 \hline
-  ER      & @!/@!  & @! & @! & @! & @! \\
-  BA      & @!/@!  & @! & @! & @! & @!\\
-  fractal (pure) & @!/@!  & @! & @! & @! & @!\\
-  fractal (small world)& @!/@! & @! & @! & @! & @!\\
+  ER                    &  0    & 0.25 & 0 & 0 \\
+  BA                    &  0    & 0.25 & 0 & 0 \\
+  fractal (pure)        &  0    & 0    & 0 & 0.25 \\
+  fractal (small world) &  0.25 & 0    & 0 & 0 \\
 \hline
 \end{tabular}
-\caption{Subdue: error @!}
+\caption{GZIP: error 0.27 (0.12)}
 \end{subfigure}
-\caption{\ldots}
+\vspace{3mm}
+
+\begin{subfigure}[b]{1\columnwidth}
+\begin{tabular}{l |  r r r r}
+\hline
+  ER                    &  0    & 0.25 & 0    & 0    \\
+  BA                    &  0    & 0    & 0.25 & 0    \\
+  fractal (pure)        &  0    & 0    & 0    & 0.25 \\
+  fractal (small world) &  0.25 & 0    & 0    & 0    \\
+\hline
+\end{tabular}
+\caption{Subdue: error 0.14 (0.14)}
+\end{subfigure}
+\caption{Confusion matrices for vearious compressors.
+If we label the resulting clusters with so that the error is minimized (ie.
+reorder the columns of the confusion matrix to maximize the sum of the diagonal) we get an indication of the performance of the compressor captured in a single number.
+We report the mean over 10 experiments (and the standard deviation in brackets) below each confusion matrix.
+The confusion matrix is always for the first experiment.}
+
+\label{table:synthetic-large}
+
 \end{table}
 
-\subsection*{Real-life data}
+\begin{table}[h]
 
-\begin{table}[h]
-\label{table:real-life}
 \begin{subfigure}[b]{1\columnwidth}
 \centering
-\begin{tabular}{l | r | r r r r}
+\begin{tabular}{l |  r r r}
 \hline
-  Web      & @!/@!  & @! & @! & @! & @! \\
-  Metabolic     & @!/@!  & @! & @! & @! & @!\\
-  Semantic A & @!/@!  & @! & @! & @! & @!\\
-  Semantic B & @!/@!  & @! & @! & @! & @!\\
+  cellular      &  0.11 & 0.11 & 0.11 \\
+  neural        &  0.11 & 0.11 & 0.11 \\
+  co-purchase   &  0.22 & 0    & 0.11 \\
 \hline
 \end{tabular}
-\caption{GZIP: error @! (@!)}
+\caption{Random baseline: error 0.43 (0.11)}
 \end{subfigure}
@@
 \centering
-\begin{tabular}{l | r | r r r r}
+\begin{tabular}{l |  r r r}
 \hline
-  Web      & @!/@!  & @! & @! & @! & @! \\
-  Metabolic     & @!/@!  & @! & @! & @! & @!\\
-  Semantic A & @!/@!  & @! & @! & @! & @!\\
-  Semantic B & @!/@! & @! & @! & @! & @!\\
+  cellular      &  0 & 0.33 & 0 \\
+  neural        &  0.11 & 0.22 & 0 \\
+  co-purchase   &  0.22 & 0 & 0.11 \\
 \hline
 \end{tabular}
-\caption{Subdue: error @! (@!)}
+\caption{GZIP: error 0.28 (0.17)}
 \end{subfigure}
-\caption{\ldots}
+\vspace{3mm}
+
+\begin{subfigure}[b]{1\columnwidth}
+\centering
+\begin{tabular}{l | r r r}
+\hline
+  cellular      &  0    & 0.33 & 0 \\
+  neural        &  0.33 & 0    & 0 \\
+  co-purchase   &  0    & 0.33 & 0 \\
+\hline
+\end{tabular}
+\caption{Subdue: error 0.34 (0.17) }
+\end{subfigure}
+\caption{Results for the experiment on natural datasets.}
+
+\label{table:real-life}
+
 \end{table}
 
-\subsection*{The models}
+% \subsection*{The models}
 
-One advantage of the model-based compressors over the general purpose ones is that we use the model as an indication of the structure in the data.
+% One advantage of the model-based compressors over the general purpose ones is that we use the model as an indication of the structure in the data.
 In the case of graph data in particular, we can expect compressors that operate directly on the graph to produce models that are easier to relate to the semantics of the original data than those of the sequential compressors.
 
-To evaluate this we concatenate all graphs from a single source and report the model (ie. the frequent subgraph) found by Subdue.
+% To evaluate this we concatenate all graphs from a single source and report the model (ie. the frequent subgraph) found by Subdue.
 
@@
 
-Our experiments show that sequential general purpose compressors are far better at performing graph inference than expected.
-Despite the random ordering of the nodes, the bitstring contains enough shallow patterns that a compressor like GZIP can tell a random graph, a scale free one and a two fractal graphs apart.
-Even for graphs with the same number of nodes and links.
-This result is particularly interesting considering the high resource reuirements put on most algorithms for graph inference.
+Our experiments show that sequential, general purpose compressors are better at performing graph inference than expected.
+Despite the random ordering of the nodes, the bitstring contains enough shallow patterns that a compressor like GZIP can tell the two types of fractal graphs apart, and only struggles with the difference between the random and BA graphs.
+This is particularly interesting considering the high resource requirements of most algorithms for graph inference, and the low resource use of general purpose compressors.
 
-This result suggests that inference on graphs can be performed by sequential algorithms on a sequential representation in linear time, with decent results, depending on the complexity of the task.
+This result suggests that inference on graphs can be performed by sequential algorithms on a sequential representation in linear time, with decent results.
 
-As for the graph-compressors, we see a small improvement relative to the sequential compressors for a string trade-off in computational resources.
+As for the graph-compressors, we see a small improvement relative to the sequential compressors for a strong increase in computational resources.
 Subdue as used in this paper is a relatively simple compressor, which isolates only a single subgraph for compression and we tested it only at modest parameters.
-The publications surreounding subdue offer much more complex solutions (such as the induction of graph grammars) and there may be some promise in parallelizing the algorithm.
+The publications surrounding Subdue offer much more complex solutions (such as the induction of graph grammars \cite{jonyer2004mdl}).
+To investigate the promise of these models as compressors further, it will be necessary to investigate both parallelized versions of these algorithms and a relaxation of the exhaustive nature of their components.
 
@@
 The Minimum Description Length principle and its associated techniques, which have not been investigated yet for reasons of scope, offer the promise of an even broader field of approaches to the analysis of graph data.
+
+\pagebreak
 
@@
 
-\bibliographystyle{plain} @!
+\bibliographystyle{style/mlapa} 
 \bibliography{benelearn}

### 2013-04-23 Bastards/src/bastards.tex
...

@@
 $C(q)$ must reflect this, whereas to $q^*$ this is non-functional code which can be removed.} and $\tilde{q} = |q| - |q^*|$ as the penalty we face for not using the optimal program.
+
+\begin{dfn}
+The \emph{resource bounded Kolmogoroc complexity} of a string $x$ is defined as
+\begin{align*}
+K_U^t(x) = \min{|p| : U^t(p) = x}
+\end{align*}
+Where $t$ is a time-constructible fuction $t(n)$ and $U^t$ is a prefix Turing machine that only halts within $t(|x|)$ timesteps.
+\end{dfn}
+
+It is important to note that the invariance of the kolmogorov complexity under resource bounds is weaker than the unbounded version.
+Specifically, if we have some prefix turing machine $q$ which computes $x$ within bounds for an input of length $K^t_q(x)$, then we can simulate that Turing machine for a constant cost as normal, but only with a multiplicative constant penalty and an additive logarithmic one:
+\[
+K^{c \cdot t + \log t} \leq K^t_q(x)
+\] 
+For the following, we assume that our resource classes are sufficiently crude that this distinction doesnt matter.\footnote{For a more complete treatment of this issue, see \cite{li1993introduction}[pages 532--536].
+The form of the penalty may change, depending on the computational metaphor.}
 
@@
 For any finite set $A$, the proportion of strings with randomness deficieny above $k$ is less than $2^-k$.
+
+\label{lma:rd-finitesets}
 \end{lma}
@@
 
-(Note that unlike the finite set case, here we do not make the complexity term conditional on the model)
+(Note that unlike the finite set case, here we do not make the complexity term conditional on the model) [@! does this matter?]
 
@@
 \begin{lma}
-The probability of all strings with randomness deficiency above $k$ is less than $2^-k$
+The probability of all strings with randomness deficiency above $k$ is less than $2^{-k}$
+\label{lma:rd-distributions}
 \end{lma}
@@
 
-\begin{exm}
-What is the probability that we sample a string $x$ with a randomness deficiency of at least half its length?
-In other words, if we use the optimal description, we gain $|x|/2$ over describing it in terms of $p_q$.
-\end{exm}
+Note that if we see a finite set as a uniform distribution over its elements, this proof provides a more general version of lemma \ref{lma:rd-finitesets}
 
 \subsection*{the universal distribution}
+
+\begin{dfn}
+Let $u$ be the prefix-free universal Turing machine.
+$p_u(x)$ is a \emph{universal distribution}.
+It dominates every computable distribution $q$ up to a multiplicative constant (dependent on $q$ but not $x$):
+\[
+p_q(x) \leq c_q p_u(x)
+\]
+\label{dfn:universal-distribution}
+\end{dfn}
+
+If we bound the resources of the universal Turing machine, we get a resource bounded universal distribution:
+
+\begin{dfn}
+Let $u^t$ be the prefix-free universal Turing machine bounded by $t$.
+$p_{u^t}(x)$ is a \emph{resource-bounded universal distribution}.
+It dominates every computable distribution $q$ up to a multiplicative constant (dependent on $q$ but not $x$):
+\[
+p_q(x) \leq c_q p_{u^t}(x)
+\]
+
+\label{dfn:resource-bounded-universal-distribution}
+\end{dfn}
+
+Note again, that this only holds for a certain level of courseness in our resource-classes.
 
@@
 
-\subsection*{}
+\section*{Inference from bounded distributions}
 
-\section*{Inference from bounded distributions}
+\subsection*{Bastards}
+
+We call a string a \emph{bastard} if resource-bounded Kolmogorov complexity shows it to be random, but under relaxation of these bounds, the string turns out to be compressible.
+This way, the probability of a bastard tells us the probability that we will make a mistake about the level of randomness in our data if we use a resource-bounded version of Kolmogorov complexity rather than the incomputable version.
+
+\begin{dfn}
+A \emph{$(t, k)$-bastard} is a string for which $K^t(x) - K(x) > k$
+\end{dfn}
+
+\begin{thm}
+Let $q \in {\cal C}^t$. The probability of sampling a $(t, k)$-bastard from $q$ is negligable for large $k$.
+\end{thm}
+\begin{proof}
+
+Let $B^t_k$ be the set of $(t, k)$-bastards. By definition \ref{dfn:resource-bounded-universal-distribution}, we have $p_q(B^t_k) \leq c_q p_{u^t}(B^t_k)$.
+
+\begin{align*}
+B^t_k &= \{x : K^t(x) - K(x) > k\} \\ 
+	&\subseteq \{x : - \log p_{u^t}(x) - K(x) > k\} \\
+	&= \{x : \delta(x | u^t) > k\}
+\end{align*}
+
+So the set of strings with randomness deficiency larger than $k$ for $u^t$ is a superset of the set of bastards.
+Since we sample string with such randomness deficiency with probability below $2^{-k}$, we have
+\begin{align*}
+p_{u^t}(B^t_k) < 2^{-k} 
+\end{align*}
+and 
+\begin{align*}
+p_q(B^t_k) < c_q \cdot 2^{-k} 
+\end{align*}
+\end{proof}
+
+\begin{thm}
+Let $q \in {\cal C}^u$ and $t < u$. The probability of sampling a $(t, k)$-bastard from $q$ is not negligable.
+\end{thm}
+\begin{proof}
+\ldots
+\end{proof}
+
 \nocite{*}

### 2013-04-25 Bastards/src/bastards.tex
...

--- 2013-04-23 Bastards/src/bastards.tex
@@
 
-\title{Bounds on the approximation of Kolmogorov complexity}
+\title{Bounds on the probability of erroneous inference with resource-bounded Kolmogorov complexity}
 \date{\today}
@@
 
-\section*{preliminaries}
+\begin{abstract}
+\ldots
+\end{abstract}
+
+It is well know that Kolmogorov complexity, though perhaps the most conceptually sound notion of information content, is incomputable.
+We can approximate it from above, which tells us that any model we find which compresses a given dataset provides an upper bound for the amount of data it contains, but we cannot approximate it from below.
+If we find no compression after a finite amount of learning we have no guarantee that the next cycle of learning won't suddenly prove our data to be highly compressible.
+
+We show that if we use a resource bounded (and thus computable) version of Kolmogorov complexity, and if we ensure that the resources match or exceed those of the source of our data, the probability of concluding that our data is random while it is actually nonrandom is negligable.
+
+We also show that if our source has strictly more computational power than our resource bounded Kolmogorov complexity, there is always a non-zero probability of seeing a compressible string and colncluding that it is incompressible.
+
+\section*{Preliminaries}
 Let $\mathbb B$ be the set of all finite binary strings $\{0,1\}^*$.
@@
 
-\subsection*{probability distributions}
+All logartihms are base $2$.
 
-We concern ourselves with probability distributions on the set of finite strings. All measure discussed are discrete and have $\mathbb B$ or a subset as support.
+\subsection*{Probability distributions}
+
+We concern ourselves with probability distributions on the set of finite strings. All measures discussed are discrete and have $\mathbb B$ or a subset as support.
 
@@
 A \emph{semimeasure} is a function $p : {\mathbb B} \rightarrow {\mathbb R}$ with $\sum_{x \in \mathbb B} x \leq 1$.
-If this sum is equal to $1$, $p$ is a \emph{measure}, which we will refer to a a probability distribution.
+If this sum is equal to $1$, $p$ is a \emph{measure}, which we will refer to as a probability distribution.
 \end{dfn}
@@
 \lim_{i \rightarrow \infty} f'(x, i) = f(x)
-\] and $f'(x, i + 1) \geq f'(x, i)$, ie. $f'$ never decreases with $i$. If the first condition is the same, but $f'$ never increases with $i$, the function is \emph{upper semicomputable}.
+\] and $f'(x, i + 1) \geq f'(x, i)$, ie. $f'$ never decreases with $i$. If the first condition is the same, but $f'$ never \emph{increases} with $i$, the function is \emph{upper semicomputable}.
 \end{dfn}
@@
 
-Since probability distributions are functions from $\mathbb B$ into $\mathBB R$, we talk about (semi)computable probability distributions and (semi)computable semimeasures.
+Since probability distributions are functions from $\mathbb B$ into $\mathbb R$, we talk about (semi)computable probability distributions and (semi)computable semimeasures.
 
@@
 
-We define $q^*$ as the shortest program implementing this function \footnote{Note that $|q^*|$ is different from $C(q)$.
+We define $q^*$ as the shortest program implementing this function \footnote{Note that $|q^*|$ is different from $K(q)$.
 For instance, a program may encode the collected works of Shakespeare in padding.
-$C(q)$ must reflect this, whereas to $q^*$ this is non-functional code which can be removed.} and $\tilde{q} = |q| - |q^*|$ as the penalty we face for not using the optimal program.
+$K(q)$ must reflect this, whereas to $q^*$ this is non-functional code which can be removed.} and $\tilde{q} = |q| - |q^*|$ as the penalty we face for not using the optimal program.
 
-\section*{sampling computable distributions}
+\begin{dfn}
+The \emph{resource-bounded Kolmogoroc complexity} of a string $x$ is defined as
+\begin{align*}
+K_U^t(x) = \min{|p| : U^t(p) = x}
+\end{align*}
+Where $t$ is a time-constructible fuction $t(n)$ and $U^t$ is a prefix Turing machine that only halts within $t(|x|)$ timesteps.
+\end{dfn}
+
+It is important to note that the invariance of the Kolmogorov complexity under resource bounds is weaker than the unbounded version.
+Specifically, if we have some prefix Turing machine $q$ which computes $x$ within bounds for an input of length $K^t_q(x)$, then we can simulate that Turing machine for a constant cost as normal, but only with a multiplicative logarithmic penalty:
+\[
+K^{c \cdot t + \log t}(x) \leq K^t_q(x) + O(1)
+\] 
+For the following, we assume that our resource classes are sufficiently crude that this distinction doesn't matter and we can say that $u^t \in {\cal C}^t$ \footnote{For a more complete treatment of this issue, see \cite{li1993introduction}[pages 532--536].
+The form of the penalty may change, depending on the computational metaphor.} [@!
+This is a little handwavy\ldots]
+
+\section*{Sampling computable distributions}
 
@@
 
-A given PTM thus produces a random output for no input. In this sense we can see it as a machine for sampling from some distribution $p_q$.
+A given PTM $q$ thus produces a random output for no input. In this sense we can see it as a machine for sampling from some distribution $p_q$.
 
@@
 We say that under the prefix-machine perspective, the Turing machine does not halt if it is presented with fewer bits input than it requires.
-If it is presented with more bits than required, the remaining bits are not counted as part of the input.
+If it is presented with more bits than required, the remaining bits are ignored and not counted as part of the input.
+
+We now show that the set of PTMs is equal to the set of semicomputable measures and semimeasures.
 
@@
 We construct a binary tree $B$, where we will assign nodes $x$ to output values $y$.
-If $x$ is assigned to $y$, we say $B(x) = $.
+If $x$ is assigned to $y$, we say $B(x) = y$.
 We want the tree to converge to a prefix tree, so once a node is assigned, all nodes below it become unavailable.
@@
 We dovetail the computation of $p_q'(y, i)$.
-After each step we calculate the value $A(y) = \sum_{x \in B} B(x)$ for all $y$.
-If $A(y)$ is less than the last value of $p_q(y, i)$ calculated, we calculate the difference $\rho_y$ and find the first available  node $x$ in $B$ with depth $|x|$ such that $2^{-|x|} < \rho_y$.
+After each step we calculate the value $A(y) = \sum_{x \in B} B(x) = y$ for all $y$.
+If $A(y)$ is less than the best approximation of $p_q(y)$ (ie.
+the highest $i$ for which we've seen $p_q(y, i)$ halt), we calculate the difference $\rho_y$ and find the first available node $x$ in $B$ with depth $|x|$ such that $2^{-|x|} < \rho_y$.
 
@@
 
-@! With high likelihood, the sampled string will have low randomness deficiency for $q$ as a model
+\subsection*{Randomness deficiency}
 
@@
 \begin{lma}[{\cite{li1993introduction}[p 120]}]
-For any finite set $A$, the proportion of strings with randomness deficieny above $k$ is less than $2^-k$. 
+For any finite set $A$, the proportion of strings with randomness deficieny above $k$ is less than $2^{-k}$.
+
+\label{lma:rd-finitesets}
 \end{lma}
@@
 
-Finally, we bound the proportion:
+We divide by $|A|$ to get the proportion:
+
 \begin{align*}
@@
 As a sanity check, we note that all strings have randomness deficiency at least $0$.
-The exponential decay with $k$ tells us that if we sample a string at random from $A$, we will likely end up with a string with low $k$.
+The exponential decay with $k$ tells us that if we draw a string at random from $A$, we will likely end up with a string with low $k$.
 In other words, sampling from $A$ will likely give us a string that is typical for $A$.
@@
 \begin{align*}
-\delta(x|p_q) &= \log \frac{p_u(x)}{p_q(x)}
+\delta(x|p_q) &= \log \frac{m(x)}{p_q(x)}
 \end{align*} 
-Where $p_u$ is the universal distribution.
+Where $m(x)$ is the universal distribution.
 
@@
 \begin{align*}
-\delta(x|p_q) &= - \log(p_q(x)) - C(x) 
+\delta(x|p_q) &= - \log p_q(x) - K(x) 
 \end{align*}
 
-(Note that unlike the finite set case, here we do not make the complexity term conditional on the model)
+(Note that unlike the finite set case, here we do not make the complexity term conditional on the model) [@! does this matter?]
 
@@
 
-We would like to show the same property we proved for the randomness deficiency of finite sets.
-That sampling from $p_q$ will likely produce a string that is typical for $p_q$, ie.
-has low randomness deficiency.
+We would like to show the same property we proved for the randomness deficiency of finite sets: that sampling from $p_q$ will likely produce a string that is typical for $p_q$, ie.
+a string that has low randomness deficiency.
 
 \begin{lma}
-The probability of all strings with randomness deficiency above $k$ is less than $2^-k$
+The probability of all strings with randomness deficiency above $k$ is less than $2^{-k}$
+\label{lma:rd-distributions}
 \end{lma}
@@
 ie. the set of all strings with randomness deficiency larger than $k$. We have 
-\begin{align*}
-p_q(X_k^q) &= \sum_{x:\delta(x|q)>k}p_q(x)\\
-    &= \sum_{x: -\log p_q(x) - K(x) > k}p_q(x)\\
-    &= \sum_{x:  p_q(x) < 2^{-k}2^{-K(x)}} p_q(x)\\
-    &< \sum_{x:\delta(x|q) > k} 2^{-k} 2^{-K(x)}\\
-    &= 2^{-k}\sum_{x:\delta(x|q) > k} 2^{-K(x)}
-\end{align*}
+\begin{alignat*}{2}
+p_q(X_k^q) &= \sum\nolimits_{x\;:\;\delta(x|q)>k} &p_q(x)\\
+    &= \sum\nolimits_{x\;:\; -\log p_q(x) - K(x) > k} &p_q(x)\\
+    &= \sum\nolimits_{x\;:\;  p_q(x) < 2^{-k}2^{-K(x)}} &p_q(x)\\
+    &< \sum\nolimits_{x\;:\;\delta(x|q) > k} &2^{-k} 2^{-K(x)}\\
+    &= 2^{-k}\sum\nolimits_{x\;:\;\delta(x|q) > k} &2^{-K(x)}
+\end{alignat*}
 
-Since the sum in the last line is less than or equal to $1$ (as the same sum over all $x$ sums to 1), we obtain
+Since the sum in the last line is less than or equal to $1$ (as the same over all $x$ sums to 1), we obtain
 
@@
 
-\begin{exm}
-What is the probability that we sample a string $x$ with a randomness deficiency of at least half its length?
-In other words, if we use the optimal description, we gain $|x|/2$ over describing it in terms of $p_q$.
-\end{exm}
+Note that if we see a finite set as a uniform distribution over its elements, this proof can be seen as a generalization of lemma \ref{lma:rd-finitesets}.
 
-\subsection*{the universal distribution}
+\subsection*{The universal distribution}
 
-\section*{resource bounded probability distributions}
+\begin{dfn}
+Let $u$ be the prefix-free universal Turing machine.
+$p_u(x) = m(x)$ is a \emph{universal distribution}.
+It dominates every computable distribution $q$ up to a multiplicative constant (dependent on $q$ but not $x$):
+\[
+p_q(x) \leq c_q m(x)
+\]
+\label{dfn:universal-distribution}
+\end{dfn}
+
+If we bound the resources of the universal Turing machine, we get a resource bounded universal distribution:
+
+\begin{dfn}
+Let $u^t$ be the prefix-free universal Turing machine bounded by $t$.
+$p_{u^t}(x) = m^t(x)$ is a \emph{resource-bounded universal distribution}.
+It dominates every computable distribution $q$ up to a multiplicative constant (dependent on $q$ but not $x$):
+\[
+p_q(x) \leq c_q m^t(x)
+\]
+
+\label{dfn:resource-bounded-universal-distribution}
+\end{dfn}
+
+Note again, that this only holds for a certain level of courseness in our resource classes.
+
+Another definition of the universal distribution is as $2^{-K^t(x)}$.
+In the non-bounded case the two definitions are equal up to a mulitplicative constant.
+In the recource bounded variant, we can only be certain of the following inequality:
+\begin{align*}
+2^{-K^t(x)} &\leq \sum_{p\;:\;u^t(p) = x} 2^{-|p|}  \\
+2^{-K^t(x)} &\leq m^t(x) \\
+-K^t(x) &\leq \log m^t(x) \\ 
+K^t(x) &\geq \log m^t(x)
+\end{align*}
+
+Where the first line follows from the fact that the left side of the inequality is a term in the sum on the right (and all terms are positive).
+
+\section*{Resource-bounded probability distributions}
 
@@
 
-Let $\cal C$ be the set of all Turing machines. Let ${\cal C}^t$ be the set of all Turing machines whose running time is bounded by $t(n)$ where $n$ is the length of the input.
+Let $\cal C$ be the set of all Turing machines. Let ${\cal C}^t$ be the set of all Turing machines whose running time for input $x$is bounded by $t(|x|)$.
 
-\subsection*{}
+We must take some care in defining what it means for a Turing machine to be outside ${\cal C}^t$.
+If, for instance, a Turing machine $q$ has no equal in ${\cal C}^t$ but only has a finite number of inputs for which it runs longer than $t(|x|)$, we can `hardcode' these inputs at a linear penatly, which would give the Turing machine an equivalent in ${\cal C}^t$, so long as the linear penalty keeps it there.
+
+To have a strong definition of Turing machines outside ${\cal C}^t$, we take a defining property of ${\cal C}^t$, the fact that $K^t$ minorizes it, and invert it.
+
+\begin{dfn}
+A Turing machine $q$ dominates ${\cal C}^t$ iff:
+\begin{align*}
+\forall c \exists x : K^t(x) \geq -\log p_q(x) + c
+\end{align*}
+\label{dfn:domination}
+\end{dfn}
+
+We don't know whether all Turing machines outside ${\cal C}^t$ dominate, but we do know that some do (including any $K^u$ with $u > t$) and that no Turing machine in ${\cal C}^t$ dominates it.
+
+%%\begin{dfn}
+%%We say that a probability distribution $q$ is outside a class ${\cal C}^t$ iff there is no $q' \in {\cal C}^t with \forall x : p_q(x) = p_{q'}(x)$
+%%\label{dfn:outside}
+%%\end{dfn}
+
+%%\begin{dfn}
+%%We say that a probability distribution $q$ is weakly outside ${\cal C}^t$ iff it is outside ${\cal C}^t$ and there exists at least one $x$ for which there is no $q' \in {\cal C}^t$ and $p$ such that $-\log p_{q'}(x) \leq -\log p_q(x)$.
+%%\label{dfn:weakly-outside}
+%%\end{dfn}
+
+%%\begin{dfn}
+%We say that a probability distribution $q$ is strongly outside ${\cal C}^t$ iff it is outside ${\cal C}^t$ and for all $q' \in {\cal C}^t$ there exist an infinite number of $x$  such that $-\log p_{q}(x) \leq -\log p_{q'}(x)$.
+%\label{dfn:strongly-outside}
+%\end{dfn}
+
+%The distinction is important for the following reason.
+If $q$ is weakly but not strongly outside ${\cal C}^t$, there will be a finite set $X$ for which $-\log p_q(x) \leq -\log p_{q'}(x)$ for any $q' \in {\cal C}^t$.
+For all other $x$, $K^t$ will minorize $q'$.
+We can define a Turing machine which `hardcodes' $X$ for the inputs that they have in $q$.
+This machine will run in linear time.
+If this machine is in ${\cal C}^t$, $q$ must be in ${\cal C}^t$ as well, and any machine that is outside ${\cal C}^t$ must be strongly outside ${\cal C}^t$.
+
+%Whether we can truly eliminate distributions that are weakly outside a resource class depends on the specifics of the function that defines it and the details of the computational metaphor (in our case Turing machines).
+To allow our results to be independent of both, we will retain the distinction between weak and strong `outsideness'.
+
+%\begin{lma}
+%If $q$ is strongly outside ${\cal C}^t$ we have:
+%\begin{align*}
+%\forall {c \exists x : K^t(x) \geq - \log p_q(x) + c}
+%\end{align*}
+%\label{lma:strongly-outside}
+%\end{lma}
+%\begin{proof}
+%Assume for a contradiction
+%\begin{align*}
+%\exists c \forall x : K^t(x) < -\log p_q(x) + c
+%\end{align*}
+%Since $u^t \in {\cal C}^t$, we have infinitely many $x$, such that 
+%\begin{align*}
+%-\log p_q(x) &\leq - \log m^t(x) \\
+%-\log p_q(x) &\leq K^t(x) \\
+%-\log p_q(x) &\leq K^t(x) < -\log p_q(x) +c
+%\end{align*}
+%\end{proof}
 
 \section*{Inference from bounded distributions}
+
+\subsection*{Bastards}
+
+We call a string a \emph{bastard} if resource-bounded Kolmogorov complexity shows it to be random, but under relaxation of these bounds, the string turns out to be compressible.
+This way, the probability of a bastard tells us the probability that we will make a mistake about the level of randomness in our data if we use a resource-bounded version of Kolmogorov complexity to approximate the incomputable version.
+
+\begin{dfn}
+A \emph{$(t, k)$-bastard} is a string for which $K^t(x) - K(x) > k$
+\end{dfn}
+
+\begin{thm}
+Let $q \in {\cal C}^t$. The probability of sampling a $(t, k)$-bastard from $q$ is negligable for large $k$.
+\end{thm}
+\begin{proof}
+
+Let $B^t_k$ be the set of $(t, k)$-bastards. By definition \ref{dfn:resource-bounded-universal-distribution}, we have $p_q(B^t_k) \leq c_q m^t(B^t_k)$.
+
+\begin{align*}
+B^t_k &= \left\{x : K^t(x) - K(x) > k\right\} \\ 
+	&\subseteq \left\{x : - \log m^t(x) - K(x) > k\right\} \\
+	&= \left\{x : \delta(x | m^t) > k\right\}
+\end{align*}
+
+So the set of strings with randomness deficiency larger than $k$ for $u^t$ is a superset of the set of bastards.
+Since we sample strings with such randomness deficiency with probability below $2^{-k}$, we have
+\begin{align*}
+m^t\left(B^t_k\right) < 2^{-k} 
+\end{align*}
+and 
+\begin{align*}
+p_q\left(B^t_k\right) < c_q \cdot 2^{-k} 
+\end{align*}
+\end{proof}
+
+\begin{thm}
+Let $q$ dominate ${\cal C}^t$. The probability of sampling a $(t, k)$-bastard from $q$ is bounded below by a constant value, independent of $k$ (and therefore, not negligable).
+\end{thm}
+
+\begin{proof}
+By definition ref{def:domination}, we have 
+\begin{align}
+\forall {c \exists x : K^t(x) \geq - \log p_q(x) + c} \label{line:non-invariance}
+\end{align}
+
+This rewrites to
+\begin{align*}
+\forall c \exists x : \frac{2^{-K^t(x)}}{p_q(x)} \leq 2^{-c}
+\end{align*}
+
+Since we know that $2^{-K^u(x)} \geq p_q(x) c_q$ we get
+\begin{align}
+\forall c \exists x : \frac{2^{-K^t(x)}}{2^{-K^u(x)}} \leq \frac{2^{-K^t(x)}}{p_q(x)c_q} \leq 2^{-c}\frac{1}{c_q}
+\label{line:prelim}
+\end{align}
+
+The statement to be proved asks for the probability of seeing a $(t, k)$-bastard. We get
+\begin{align*}
+p_q\left(B^t_k\right) &\geq p_q\left (\left\{x: K^t(x) - K(x) > k\right\}\right) \\
+	&\geq p_q\left(\left\{x: K^t(x) - K^u(x) > k\right\}\right) \\
+	&= p_q\left(\left\{x : \frac{2^{-K^t(x)}}{2^{-K^u(x)}} < 2^{-k}\right\}\right) \\
+\end{align*}
+
+From (\ref{line:prelim}), we know that there is at least one element in the set in the last line, whatever $k$. Call this element $x'$, we have:
+\begin{align*}
+p_q\left(B^t_k\right) \geq p_q\left(x'\right)
+\end{align*}
+We know that $p_q\left(x'\right)$ cannot be zero, because, then by (\ref{line:non-invariance}), its code length under $K^t$ would be infinite.
+Thus, there is always a lower bound, independent of $k$ on the probability of seeing a $(k, t)$-bastard.
+\end{proof}
+
+\section*{Discussion}
+
 \nocite{*}

### 2013-04-29 Benelearn/src/benelearn.tex
...

@@
 
-\usepackage{style/benelearn2013}
-\usepackage{charter}
+\usepackage[accepted]{style/benelearn2013}
+%\usepackage{charter}
 \usepackage{eulervm}
-% \usepackage{graphicx}
+%\usepackage{graphicx}
 \usepackage{amsmath, amsthm, amssymb}
@@
 \usepackage{style/mlapa}
-\usepackage{style/benelearn2013}
 
@@
 
-\benelearntitlerunning{Compression-based inference on graph data}
+%\benelearntitlerunning{Compression}
 
@@
 \begin{abstract}
-
 We investigate the use of compression-based learning on graph data.
-Our hypothesis is that the sequential representation of data on which general purpose compressors operate, hampers compression-based learning, as a graph can be represented sequentially in many different, equivalent ways.
-Using Normalized Compression Distance (NCD), we test various compressors that operate directly on the graph representation against compressors that operate on a bitstring representation.
+General purpose compressors operate on bitstrings or other sequential representations.
+Since a single graph can be represented sequentially in many ways, we study the effect of this projection.
+Using Normalized Compression Distance (NCD), we test a sequential compressor versus a native graph compressor.
 We use both synthetic, randomly generated graphs and real-life datasets.
-
+We conclude that, even under adverse circumstances, sequential representations contain enough structure for shallow algorithms to perform inference successfully.
+Algorithms that operate directly on the graph representation require a considerable increase in resources, but do allow for an increase in performance also.
 \end{abstract}
 
-\emph{Compression equals learning} is the mantra of the field of compression-based learning.
+The mantra of compression-based inference is that \emph{compression equals learning}.
 If we can compress data, we must have found some structure, we must have learned something.
 Conversely, if we have learned something about data, we should be able to use that knowledge to represent our data more succinctly.
+\footnote{Other features commonly associated with learning, such as generalization will, so the adherents of compression-based learning argue, be optimized along with the optimization of compression.
+See, for instance \cite{grunwald2005tutorial,grunwald2007minimum}.}
 
@@
 The graph is a powerful and versatile representation.
-Most applications of compression-based learning choose sequential models, such as deterministic finite automata or block-sorting compressors, which operate on bitstrings.
-If we have data in the form of a graph, we can translate it to a bitstring, but in this transformation we complicate our problem.
-For a different ordering of nodes, the graph is the same, but the bitstring changes radically.
+Most applications of compression-based learning use sequential models, such as deterministic finite automata or block-sorting compressors, which operate on bitstrings.
+If we have data in the form of a graph, we can translate it to a bitstring, of course, but in this transformation we complicate our problem.
+For different orderings of nodes, the graph is the same, but the bitstring changes radically.
 To a simple compressor, the two bitstrings may not share very much structure, even though they represent the same graph.
+We will call this the problem of isomorphism.
 
-If this problem holds, we should investigate compressors that operate on the level of the graph representation, for instance by finding frequent subgraphs or clustering the graph.
+If this issue really `blinds' the sequential compressors to the structures in the graph, one option is to investigate compressors that operate on the level of the graph representation, for instance by finding frequent subgraphs or clustering the graph.
 The methods do not suffer this problem of isomorphism, but as a result, they are more expensive than their sequential cousins.
-In this paper we hope to provide a first indication of how far the sequential approach will go, and whether the native approach will let us continue on from there.
+In this paper we hope to provide a first indication of how far \emph{the sequential approach} will go, and whether \emph{the native approach} will let us continue on from there.
+
+There is a wealth of research available on machine learning and data mining both within single graphs and on sets of graphs.
+See for instance, the book \cite{cook2006mining}.
+In this paper we do not share the data mining goal of extracting interesting features from the data, but only the goal of performing inference, using the compressors and their learning techniques as black boxes, and evaluating the the results of a chosen inference task.
 
@@
 
-From the many methods born from Kolmogorov Complexity and Minimum Description Length, we choose \emph{Normalized Compression Distance} (NCD) as a representative method of compression-based learning.
+From the many machine learning methods based on principles of compression, we choose \emph{Normalized Compression Distance} (NCD) as a representative method of compression-based learning.
 The choice is a practical one: NCD is a simple method which requires nothing more than a general purpose compressor.
@@
 
-The second principle is formalized in two ways.
+The second principle is formalized in two steps.
 First, by `description' we mean a description in a formal language that is maximally expressive.
 To formalize expressivenes, we require that the method of description is Turing-complete (of equivalent strength to a Turing machine).
-By the (strong) Church-Turing thesis, this suggests that there is no reasonable way of defining a more expressive method of description.
+By the strong Church-Turing thesis, this suggests that there is no reasonable way of defining a more expressive method of description.
 
@@
 Let $k_V(x)$ be the shortest program for $x$ on $V$.
-Since, $U$ is universal, we know that is can compute $k_v$ by simulating $V$.
+Since, $U$ is universal, we know that it can compute $k_v$ by simulating $V$.
 This simulation is a program for $x$ on $U$ which is a bound for the shortest program on $U$:
@@
 Where $p(v)$ is the penalty (ie.
-the additional bits) that the pairing function reuires to store its two arguments in a separable way.
+the additional bits) that the pairing function requires to store its two arguments in a separable way.
 We require that this is only dependent on $v$.
@@
 The length of the shortest program to get from $y$ to $x$ intuitively suggests that $K(\cdot\mid \cdot)$ can be seen as a similarity measure.
-Clearly, very little is required to transform a string into itself, or a very similar string, whereas for two random strings, only a program that stores the second in its entirety can make the transformation.
+Clearly, very little is required to transform a string into itself, or into a very similar string, whereas for two random strings, only a program that stores the second in its entirety can make the transformation.
 
@@
 The second issue is one of scale.
-If two strings of a million bits differ by 1000 bits, we might consider them quite similar, whereas two strings of 1000 bits that differ by that amount could not be more different.
-Thus, we would like to take the length of the strings into account.
+If two strings of a million bits differ by $1000$ bits, we might consider them quite similar, whereas two strings of $1000$ bits that differ by that amount could not be more different.\footnote{Note that this is only an intuitive example.
+If two strings differ in exactly every bit,a very short program transforms one into the other, so by NID, they are very similar.} Therefore, we would like to take the length of the strings into account.
 This gives us the Normalized information Distance (NID)
@@
 
-Our aim is to test a sequential and graph-based compressor on a variety of graph data.
-To ascertain the performance of the compressors, we generate graphs from different sources, calculate their distances and see whether a clustering algorithm can reconstruct the original sources as clusters.
+Our aim is to test a sequential and a graph-based compressor on an inference tasks for a variety of graph data.
+To ascertain the performance of the compressors, we generate graphs from different sources, calculate their NCD distances and see whether a clustering algorithm can reconstruct the original sources as clusters.
 Datasets and source code for these experiments are available.
@@
 
+\subsection*{Node ordering}
+
+An important and subtle concern is the ordering of nodes in the sequential representation of our graphs.
+This issue is detailed very well in \cite{kang2011beyond}.
+As shown, there are various algorithms to determine node orderings that bring out a lot of the graph's inherent structure in the adjacency matrix, allowing a sequential compressor to use it.
+We can use substantial resources to find a good ordering of nodes to improve the performance of the sequential compressor, but if we did, we could no longer consider it a `shallow model'.
+
+Since we are testing the capacity of the general purpose compressor to perform inference despite the problem of isomorphism, we will present it with a worst case scenario.
+We will use a random ordering of nodes for all graphs.
+If the general purpose compressor still outperforms the random baseline under these conditions, it will tell us that it is, at least in part, resistant to the problem of isomorphism.
+
 \subsection*{Experiment 1: Synthetic data}
 
-We generate graphs from two models.
+We generate graphs from four models.
+
 The first is the classic Erd\H{o}s-R\'enyi (ER) model, where a uniform random choice is made from all graphs with $n$ nodes and $m$ links.
@@
 Once we have this dataset, consisting of four gold clusters, we calculate the NCD with a given compressor for every pair of graphs in the dataset, giving us a symmetric matrix.
-We use the k-medoids algorithm to cluster this set into two clusters, and we check whether graphs from a given source tend to end up in a specific cluster.
+We use the k-medoids algorithm to cluster this set into four clusters.
+
+To assess the performance of the clustering we label the clusters so that the accuracy is maximized (essentially assigning the optimal labels).
+Clearly, this would be cheating when testing a classifier, but since we are onnly interested in the clustering aspect, it gives us a straightforward performance measure.
+
 As a random baseline, we generate a random distance matrix with every distance a uniform random value in $[0, 1)$, and run the clustering algorithm on that.
@@
 	\item[cellular] The cellular network of the E. Coli bacterium. \cite{jeong2000large} Acquired from \url{http://www.nd.edu/~networks/resources/metabolic/index.html}
-	\item[neural] The neural network of the C. Elegans nematode worm (ignoring link directions). \cite{achacoso1991ay, watts1998small} Acquired from \url{http://toreopsahl.com/datasets/#celegans}
+	\item[neural] The neural network of the C. Elegans nematode worm (ignoring link directions). \cite{achacoso1991ay,watts1998small} Acquired from \url{http://toreopsahl.com/datasets/#celegans}
 	\item[co-purchase] A graph of items commonly purchased together on internet retailer Amazon.com. \cite{leskovec2007dynamics} Acquired from \url{http://snap.stanford.edu/data/amazon0302.html}
 \end{description}
-
-% Some basic properties of these graphs are given in table \ref{table:properties}.
-
-%\begin{table}
-%\label{table:properties}
-%\begin{tabular}{ l | r r r r r}
-%\hline
-%    & nodes & links & C & $alpha$ & $p_\alpha$ \\
-%\hline
-%  web & & & & & \\
-%  metabolic & & & & & \\
-%  semantic A & & & & & \\
-%  semantic B & & & & & \\
-%  
-%\hline
-%\end{tabular}
-%\caption{The basic properties of our natural datasets.
-We report the number of nodes, the number of links, the number of nodes in the largest component (C), the estimates power law exponent ($\alpha$), and the significance for the power law model ($p_\alpha$, significant for values \emph{above} 0.01).}
-
-%\end{table}
 
@@
 The algorithm is essentially a beam search through the space of subgraphs.
+If consists of three main routines:
 
-Each subgraph serves as a model for compression by finding all occurrences of the subgraph in the data.
-Each occurrence is replaced by a single node, with the edges to the nodes containing annotations on how the edge shouldbe connected within the subgraph.
-We then store the subgrap once, together with the edited datagraph, to represent the original data.
-If the subgraph occurs often enough to offset the extra bits required to annotate the edges in the datagraph, the Subdue representation will be more efficient.
+\begin{description}
+\item[Subgraph matching] This is essentially an algorithm for finding the occurrences of a given subgraph in a graph.
+The algorithm used is detailed in \cite{bunke1983inexact}.
+Since this is essentially a semi-exhaustive search for the solution to the NP-Hard problem of subgraph isomorphism, the matching can only be solved for very small subgraphs.
+Unfortunately, weven with subgraphs of four or five nodes, the matching is too slow in combination with the number of times it is executed to calculate a full distance matrix.
+To combat this issue, we remove all but the first $b_{\mbox{inner}}$ elements from the search queue after it is sorted at ech iteration, effectively turning the algorithm into a beam search.
+\item[MDL Scoring]
+This routine takes a subgraph, finds its occurrences in the data by the previous routine and replaces these occurrences with a single node with a new label, which we call a \emph{symbol node}.
+All links that connected to nodes in the subgraph are now connected to the symbol node.
+The links connecting to the symbol node are annotated with the indices of the nodes in the subgraph they were connected to.
+Storing the subgraph once together with the reduced data graph is enough to reconstruct the original data, and, if the subgraph occurs often, will lead to a significant compression.
+\item[Subgraph search]
+This is the `outer loop' of the algorithm.
+Starting with a graph of a single link beteen two nodes, it searches through the space of all connected graphs by extending each current candidate by one link at a time (possibly by adding a new node as well).
+The buffer of current candidates is sorted by MDL score, and the candidate with the highest score is extended to create new candidates.
+At each iteration all but the top $b_{\mbox{outer}}$ candidates are removed, turning the search into a beam search.
+\end{description}
 
-The following is a rough description of the algorithm:
+Algorithm \ref{pseudo:algorithm} shows a broad description of the whole algorithm.
 
@@
 	$G$: the data graph \\
-	$b$: the beam size \\
+	$b_{\mbox{outer}}$: the beam size \\
 	\\
@@
 	\hspace*{5mm} \textit{annotate links to $N$ with the nodes in $s'$}\\
-	\hspace*{5mm} \textbf{return} \textit{number of bits to store the edited $G$ and $s'$ }\\
+	\hspace*{5mm} \textbf{return} \textit{\# of bits to store the edited $G$ and $s'$ }\\
 }
 	\caption{Pseudocode for the Subdue algorithm}
+
+	\label{pseudo:algorithm}
 \end{pseudo}
@@
 The graph matching search (the first line of the \textbf{score} function) allows for inexact matches of the subgraph.
-In these cases, use a rough estimate of the number of bits required to transform the stored subgraph into the subgraph that is actually present in the data.
-
-To make the calculation of a distance matrix feasible, we must find a way to reduce the computational costs of the Subdue algorithm.
-Our solution is to replace the exhaustive graph matching search by a non-exhaustive beam search.
-With this change, we cannot be sure that the score assigned to the given subgraph describes the optimal way to compress the data with that subgraph, but it does provide an upper bound.
+In these cases, we use a rough estimate of the number of bits required to transform the stored subgraph into the subgraph that is actually present in the data.
 
@@
 For $k = 2$ the clustering problem would be more difficult.
+The random graphs are generated with the exact same number of nodes and links.
 The fractal graphs we generate to depth $2$ by adding $4$ ancestors at each side of each link and $1$ extra link between the groups of ancestors.
@@
 The Subdue algorithm has a lot of parameters.
-Wuring the search we return only one best subgraph.
-Our beamwidth at the top level is set to $5$.
-The beam width in the graph matching routine is set to $10$.
+During the search we return only one best subgraph.
+Our beamwidth at the top level ($b_{\mbox{outer}}$) is set to $5$.
+The beam width in the graph matching routine ($b_{\mbox{inner}}$) is set to $10$.
 
@@
 \begin{subfigure}[b]{1\columnwidth}
-\begin{tabular}{l |  r r r r}
+\begin{tabular}{l |  l l l l}
 \hline
-  ER                    &  0.083 & 0    & 0.17  & 0 \\
-  BA                    &  0.16  & 0    & 0.083 & 0 \\
-  fractal (pure)        &  0     & 0    & 0.083 & 0.17 \\
-  fractal (small world) &  0     & 0.25 & 0     & 0 \\
+  ER                    &  0.17  & 0.083   & 0     & 0 \\
+  BA                    &  0.083 & 0.17    & 0     & 0 \\
+  fractal (pure)        &  0.083 & 0       & 0.17  & 0 \\
+  fractal (small world) &  0     & 0       & 0     & 0.25 \\
 \hline
@@
 \begin{subfigure}[b]{1\columnwidth}
-\begin{tabular}{l |  r r r r}
+\begin{tabular}{l |   l l l l}
 \hline
@@
 \begin{subfigure}[b]{1\columnwidth}
-\begin{tabular}{l |  r r r r}
+\begin{tabular}{l |  l l l l}
 \hline
@@
 \end{subfigure}
-\caption{Confusion matrices for vearious compressors.
-If we label the resulting clusters with so that the error is minimized (ie.
-reorder the columns of the confusion matrix to maximize the sum of the diagonal) we get an indication of the performance of the compressor captured in a single number.
-We report the mean over 10 experiments (and the standard deviation in brackets) below each confusion matrix.
-The confusion matrix is always for the first experiment.}
-
+\caption{Confusion matrices for various compressors.
+Columns represent the clusters found by the k-medioids algorithm.
+We have labeled the resulting clusters so that the error is minimized (ie.
+reordered the columns of the confusion matrix to maximize the sum of the diagonal).
+We report the mean error (1 - the sum of the diagonal) over 10 experiments (and the standard deviation in brackets) below each confusion matrix.
+The confusion matrix shown is always for the first experiment in the run.}
 \label{table:synthetic-large}
-
 \end{table}
 
-\begin{table}[h]
+\begin{table}
 
@@
 \centering
-\begin{tabular}{l |  r r r}
+\begin{tabular}{l | l l l }
 \hline
@@
 \centering
-\begin{tabular}{l | r r r}
+\begin{tabular}{l | l l l }
 \hline
@@
 \caption{Results for the experiment on natural datasets.}
-
 \label{table:real-life}
-
 \end{table}
-
-% \subsection*{The models}
-
-% One advantage of the model-based compressors over the general purpose ones is that we use the model as an indication of the structure in the data.
-In the case of graph data in particular, we can expect compressors that operate directly on the graph to produce models that are easier to relate to the semantics of the original data than those of the sequential compressors.
-
-% To evaluate this we concatenate all graphs from a single source and report the model (ie. the frequent subgraph) found by Subdue.
 
@@
 
-The notion of compression is a good framework within which to combine many approaches to inference from the most general to the most domain specific.
+The notion of compression-based learning is a good framework within which to combine many approaches to inference from the most general to the most domain specific.
 The Minimum Description Length principle and its associated techniques, which have not been investigated yet for reasons of scope, offer the promise of an even broader field of approaches to the analysis of graph data.
-
-\pagebreak
 
@@
 
+We would like to thank the reviewers for many insightful comments.
+
+\appendix
+
+\section*{Appendix: Graph coding}
+
+The method of coding data is always a sensitive point in compression-based learning.
+The precise choices made in translating the data to a string of bits can greatly affect which patterns are picked up, or ignored by the subsequent inference procedure.
+Here we detail the exact procedure used to encode the graphs in both compressors.
+
+\subsection*{GZIP}
+We take half of the adjacency matrix ($(n^2 + n)/2$ bits), and store it in flattened form as an array of java byte primitives, together with the size of the string (since the stored representation is padded to a multiple of eight).
+We then serialize this representation into a java GZIPOutputStream.
+To make our implementation generic, we assume that the nodes and links are labeled, and serialize the labels in a fixed order after the adjacency matrix.
+In the graphs mentioned in this paper, there is a single label assigned to all elements, so inference shouldn;t be influenced.
+
+\subsection*{Subdue}
+
+In the subdue case we do our coding in a more precise way, without relying on platform-specific functions.
+More importantly, we only count the bits required to store our graph, rather than actually constructing the representation itself.
+
+\subsubsection*{Plain graph}
+
+To store a plain graph, we follow roughly the coding strategy outlined in \cite{holder1994substructure}.
+
+We first store the number of nodes, and the maximum number of neighbours for a node in the graph $n_{\mbox{max}}$. We then store the adjacency matrix row by row.
+
+For node $i$, we only need to store the connections to the $i$ nodes below it including itself.
+We use $l(n_{\mbox{max}})$ bits to store the number of such neighbours $n_i$, and $\log{i \choose n_i}$ bits to store the configuration of those neighbours.
+
+We then account for the possibility of multiple links between node pairs.
+Since we expect this to be unlikely, we want to ensure that we do not waste bits coding for this occasion if it does not occur.
+We store the pairs for which there are multiple links as a list consisting of two numbers per element: the first is the index of the pair in a canonical ordering, the second is the number of extra links that exist for that pair (in prefix-coding).
+We use a single bit to indicate whether this list will follow, and we reserver a codeword to signal the end of the list.
+
+After this, all that remains is to encode the labels of the nodes and links.
+We assume that the sender and receiver in our coding scheme posess a codebook that is efficient for the data given, so that if a label $l$ occurs with frequency $\#l$, we can encode it in $-\log \frac{\#l}{\sum_k \#k}$ bits.
+
+We assume that there is some canonical ordering among the nodes and links, and store them as a stream. Since the number of labels is known from the adjacency matrix, this code is self-delimiting.
+
+\subsubsection*{Graph with substructure}
+
+To store the graph with a substructure, we first store the substructure itself.
+Since this contains no special symbols, we can store it simply using the above method (except we use the codebook based on the whole graph rather than the substructure to encode the labels).
+This is a prefix code, so we can start encoding the rest of the graph right after.
+
+In the rest of the graph, we replace the occurrences of the substructure with a symbol node and store this graph by the method detailed for plain graphs.
+We store the number of symbol nodes in prefix coding, and only encode labels for the rest of the nodes (again, by the codebook of the full data).
+
+Finally, we must store, for every link leading into a symbol node, which node in the substructure it should connect to to restore the original data.
+Since we know the graph and all the symbol nodes, we can presume there to be a canonical ordering in these links, and we use $\log n_s$ bits for each, where $n_s$ is the size of the substructure in nodes.
+
 \bibliographystyle{style/mlapa} 

### 2013-04-30 Bastards/src/bastards.tex
...

@@
 
-We also show that if our source has strictly more computational power than our resource bounded Kolmogorov complexity, there is always a non-zero probability of seeing a compressible string and colncluding that it is incompressible.
+We also show that if our source has strictly more computational power than our resource bounded Kolmogorov complexity, there is always a non-zero probability of seeing a compressible string and concluding that it is incompressible.
 
@@
 
-All logartihms are base $2$.
+All logarithms are base $2$.
 
@@
 -K^t(x) &\leq \log m^t(x) \\ 
-K^t(x) &\geq \log m^t(x)
+K^t(x) &\geq -\log m^t(x)
 \end{align*}
@@
 \begin{proof}
-By definition ref{def:domination}, we have 
+By definition \ref{dfn:domination}, we have 
 \begin{align}

### 2013-04-30 Benelearn/src/benelearn.tex
...

@@
 \documentclass{article}
+
+% \usepackage{natbib}
 
@@
 \usepackage{float}
+
 \usepackage{style/mlapa}
@@
 General purpose compressors operate on bitstrings or other sequential representations.
-Since a single graph can be represented sequentially in many ways, we study the effect of this projection.
+A single graph can be represented sequentially in many ways, which may influence the performance of sequential compressors.
 Using Normalized Compression Distance (NCD), we test a sequential compressor versus a native graph compressor.
@@
 We conclude that, even under adverse circumstances, sequential representations contain enough structure for shallow algorithms to perform inference successfully.
-Algorithms that operate directly on the graph representation require a considerable increase in resources, but do allow for an increase in performance also.
+Algorithms that operate directly on the graph representation likely require a considerable increase in resources, but do allow for an increase in performance also.
 \end{abstract}
@@
 \footnote{Other features commonly associated with learning, such as generalization will, so the adherents of compression-based learning argue, be optimized along with the optimization of compression.
-See, for instance \cite{grunwald2005tutorial,grunwald2007minimum}.}
+See, for instance \npcite{grunwald2005tutorial,grunwald2007minimum}.}
 
@@
 
-We can now say, that with respect to $U$, there must be a minimal description for any given data:
+We can now say that with respect to $U$, there must be a minimal description for any given data:
 \[
@@
 
-The result that makes Kolmogorov Complexity a useful measure of information content is that $K_U(\cdot)$ is only marginally dependent on the choice of $U$.
+The result that makes Kolmogorov complexity a useful measure of information content is that $K_U(\cdot)$ is only marginally dependent on the choice of $U$.
 If we suppose that there is another universal description method, $V$, we might ask what the expected difference is between $K_U(x)$ and $K_U(v)$.
@@
 If two strings of a million bits differ by $1000$ bits, we might consider them quite similar, whereas two strings of $1000$ bits that differ by that amount could not be more different.\footnote{Note that this is only an intuitive example.
-If two strings differ in exactly every bit,a very short program transforms one into the other, so by NID, they are very similar.} Therefore, we would like to take the length of the strings into account.
+If two strings differ in exactly every bit, a very short program transforms one into the other, so by NID, they are very similar.} Therefore, we would like to take the length of the strings into account.
 This gives us the Normalized information Distance (NID)
@@
 
-When our data is represented as a graph, rather than a string, we replace the notion of concatenation strings by concatenation of graphs.
+When our data is represented as a graph, rather than a string, we replace the notion of concatenation strings by of concatenation of graphs.
 That is, we combine the graphs $x$ and $y$ into a single (disconnected) graph.
@@
 As shown, there are various algorithms to determine node orderings that bring out a lot of the graph's inherent structure in the adjacency matrix, allowing a sequential compressor to use it.
-We can use substantial resources to find a good ordering of nodes to improve the performance of the sequential compressor, but if we did, we could no longer consider it a `shallow model'.
+We could use substantial resources to find a good ordering of nodes to improve the performance of the sequential compressor.
+I we did, however, we might lose the ability to consider the sequential compressor a shallow model.
+To maintain the sequential compressor as a representative example of shallow models, the node ordering should be cheap to establish from a random ordering, preferably in linear time.
 
-Since we are testing the capacity of the general purpose compressor to perform inference despite the problem of isomorphism, we will present it with a worst case scenario.
+Since we are testing the capacity of the general purpose compressor to perform inference despite the problem of isomorphism, we will actually present it with a worst case scenario.
 We will use a random ordering of nodes for all graphs.
@@
 Thus, under the BA model the more links a node has, the higher the probability that it will accrue even more.
-This effect causes the degree distribution of a BA network to form a scale free distribution.
+This effect causes the degree distribution of a BA network to become scale-free (ie.
+it follows a power law).
 
@@
 To assess the performance of the clustering we label the clusters so that the accuracy is maximized (essentially assigning the optimal labels).
-Clearly, this would be cheating when testing a classifier, but since we are onnly interested in the clustering aspect, it gives us a straightforward performance measure.
+Clearly, this would be cheating when testing a classifier, but since we are only interested in the clustering aspect, it gives us a straightforward performance measure.
 
@@
 Subdue \cite{ketkar2005subdue} is an algorithm for finding frequent subgraphs in graph data.
-The algorithm searches through all possible subgraphs, searching for the  one that maximally compresses the data.
-The algorithm is essentially a beam search through the space of subgraphs.
-If consists of three main routines:
+The algorithm searches for the subgraph that maximally compresses the data.
+The body of the algorithm is essentially a beam search through the space of subgraphs.
+It consists of three main routines:
 
 \begin{description}
-\item[Subgraph matching] This is essentially an algorithm for finding the occurrences of a given subgraph in a graph.
-The algorithm used is detailed in \cite{bunke1983inexact}.
-Since this is essentially a semi-exhaustive search for the solution to the NP-Hard problem of subgraph isomorphism, the matching can only be solved for very small subgraphs.
-Unfortunately, weven with subgraphs of four or five nodes, the matching is too slow in combination with the number of times it is executed to calculate a full distance matrix.
+\item[Subgraph matching] This is an algorithm for finding the occurrences of a given subgraph in a graph.
+The method used is detailed in \cite{bunke1983inexact}.
+Since this is  a semi-exhaustive search for the solution to the NP-complete problem of subgraph isomorphism, the matching can only be solved for very small subgraphs.
+Unfortunately, even with subgraphs of four or five nodes, the matching is too slow in combination with the number of times it is executed to calculate a full distance matrix.
 To combat this issue, we remove all but the first $b_{\mbox{inner}}$ elements from the search queue after it is sorted at ech iteration, effectively turning the algorithm into a beam search.
@@
 This is the `outer loop' of the algorithm.
-Starting with a graph of a single link beteen two nodes, it searches through the space of all connected graphs by extending each current candidate by one link at a time (possibly by adding a new node as well).
+Starting with a graph of a single link between two nodes, it searches through the space of all connected graphs by extending each current candidate by one link at a time (possibly by adding a new node as well).
 The buffer of current candidates is sorted by MDL score, and the candidate with the highest score is extended to create new candidates.
@@
 
-	\hspace*{5mm}\textit{sort $S$ by score($s', G$)} for $s' \in S$\\
+	\hspace*{5mm}\textit{sort $S$ by} score($s', G$) \textit{for} $s' \in S$\\
 	\hspace*{5mm}\textit{remove all but the first $b$ elements of $S$} \\
 	\\
-	\textbf{function score} $(s', G)$\\
+	\textbf{function} score$(s', G)$\\
 	\hspace*{5mm} \textit{replace occurrences of $s'$ in $G$ with node $N$}\\
 	\hspace*{5mm} \textit{annotate links to $N$ with the nodes in $s'$}\\
-	\hspace*{5mm} \textbf{return} \textit{\# of bits to store the edited $G$ and $s'$ }\\
+	\hspace*{5mm} \textbf{return} \textit{nr of bits to store the edited $G$ and $s'$ }\\
 }
@@
 
-Table \ref{table:synthetic-large} shows the results on randomly generated graphs. Table \ref{table:real-life} shows the results for subgraphs sample from real-life datasets.
+Table \ref{table:synthetic-large} shows the results on randomly generated graphs. Table \ref{table:real-life} shows the results for subgraphs sampled from real-life datasets.
 
@@
 The publications surrounding Subdue offer much more complex solutions (such as the induction of graph grammars \cite{jonyer2004mdl}).
-To investigate the promise of these models as compressors further, it will be necessary to investigate both parallelized versions of these algorithms and a relaxation of the exhaustive nature of their components.
+To investigate the promise of these models as compressors further, it will be necessary to investigate both parallelized versions of these algorithms and a more elegant relaxation of the exhaustive nature of their components.
 
@@
 The precise choices made in translating the data to a string of bits can greatly affect which patterns are picked up, or ignored by the subsequent inference procedure.
-Here we detail the exact procedure used to encode the graphs in both compressors.
+Here we detail the procedure used to encode the graphs in both compressors.
 
@@
 To make our implementation generic, we assume that the nodes and links are labeled, and serialize the labels in a fixed order after the adjacency matrix.
-In the graphs mentioned in this paper, there is a single label assigned to all elements, so inference shouldn;t be influenced.
+In the graphs mentioned in this paper, there is a single label assigned to all elements, so inference shouldn't be affected by the labels.
 
@@
 
-To store a plain graph, we follow roughly the coding strategy outlined in \cite{holder1994substructure}.
+To store a plain graph, we follow roughly the coding strategy outlined in \npcite{holder1994substructure}.
 
-We first store the number of nodes, and the maximum number of neighbours for a node in the graph $n_{\mbox{max}}$. We then store the adjacency matrix row by row.
+We first store the number of nodes $n$ in preefirx free coding, and the maximum number of neighbours for a node in the graph $n_{\mbox{max}}$ (int $\log n$ bits).
+We then store the lower half of the adjacency matrix (including the diagonal) row by row.
 
@@
 We then account for the possibility of multiple links between node pairs.
-Since we expect this to be unlikely, we want to ensure that we do not waste bits coding for this occasion if it does not occur.
+Since we expect miultiple links to be unlikely, we want to ensure that we do not waste many bits coding for this occasion if it does not occur.
 We store the pairs for which there are multiple links as a list consisting of two numbers per element: the first is the index of the pair in a canonical ordering, the second is the number of extra links that exist for that pair (in prefix-coding).
-We use a single bit to indicate whether this list will follow, and we reserver a codeword to signal the end of the list.
+We use a single bit to indicate whether this list will follow, and we reserve a codeword to signal the end of the list.
 
 After this, all that remains is to encode the labels of the nodes and links.
-We assume that the sender and receiver in our coding scheme posess a codebook that is efficient for the data given, so that if a label $l$ occurs with frequency $\#l$, we can encode it in $-\log \frac{\#l}{\sum_k \#k}$ bits.
+We assume that the sender and receiver in our coding scheme possess a codebook that is efficient for the data given, so that if a label $l$ occurs with frequency $\#l$, we can encode it in $-\log \frac{\#l}{\sum_k \#k}$ bits.
 
-We assume that there is some canonical ordering among the nodes and links, and store them as a stream. Since the number of labels is known from the adjacency matrix, this code is self-delimiting.
+We assume that there is some canonical ordering among the nodes and links, and store them as a stream.
+Since the number of labels is known from the adjacency matrix, the code for the entire graph is self-delimiting.
 
@@
 
-Finally, we must store, for every link leading into a symbol node, which node in the substructure it should connect to to restore the original data.
+Finally, we must store for every link leading into a symbol node which node in the substructure it should connect to to restore the original data.
 Since we know the graph and all the symbol nodes, we can presume there to be a canonical ordering in these links, and we use $\log n_s$ bits for each, where $n_s$ is the size of the substructure in nodes.

### 2013-05-07 Bastards/src/bastards.tex
...

@@
 
-\title{Bounds on the probability of erroneous inference with resource-bounded Kolmogorov complexity}
+\title{Bounds for non-approximable quantities in Kolmogorov complexity}
 \date{\today}
@@
 
-We show that if we use a resource bounded (and thus computable) version of Kolmogorov complexity, and if we ensure that the resources match or exceed those of the source of our data, the probability of concluding that our data is random while it is actually nonrandom is negligable.
+We show that if we use a resource bounded (and thus computable) version of Kolmogorov complexity, and if we ensure that the resources match or exceed those of the source of our data, the probability of concluding that our data is random while it is actually nonrandom is negligible.
 
@@
 
-Let {\cal T} be the set of all finite turing machines.
+Let {\cal C} be the set of all finite Turing machines.
 
@@
 \end{align*}
-Where $t$ is a time-constructible fuction $t(n)$ and $U^t$ is a prefix Turing machine that only halts within $t(|x|)$ timesteps.
+Where $t$ is a time-constructible fuction $t(n)$ and $U^t$ is a prefix Turing machine that only halts within $t(|x|)$ steps.
 \end{dfn}
@@
 For the following, we assume that our resource classes are sufficiently crude that this distinction doesn't matter and we can say that $u^t \in {\cal C}^t$ \footnote{For a more complete treatment of this issue, see \cite{li1993introduction}[pages 532--536].
-The form of the penalty may change, depending on the computational metaphor.} [@!
-This is a little handwavy\ldots]
-
-\section*{Sampling computable distributions}
+The form of the penalty may change, depending on the computational metaphor.}
+\subsection*{Sampling computable distributions}
 
@@
 
-Clearly this value can never decrease with $i$. To show that this function reaches $p_q(y)$ in the limit, we note that $p_q(y)$ is the sum of the probabilities of all inputs that produce $y$,
-\[
-p_q(y) = \sum_{x : q(x) = y} 2^{-|x|}
-\]
+%Clearly this value can never decrease with $i$. To show that this function reaches $p_q(y)$ in the limit, we note that $p_q(y)$ is the sum of the probabilities of all inputs that produce $y$,
+%\[
+%p_q(y) = \sum_{x : q(x) = y} 2^{-|x|}
+%\]
 
-whereas $p_q(y, i)$ is the sum of all probabilities of all inputs that produce $y$ within $i$ cycles of a dovetailing computation.
-\[
-p_q'(y, i) = \sum_{x : q(x) \downharpoonright^i y} 2^{-|x|}
-\]
+%Whereas $p_q(y, i)$ is the sum of all probabilities of all inputs that produce $y$ within $i$ cycles of a dovetailing computation.
+%\[
+%p_q'(y, i) = \sum_{x : q(x) \downharpoonright^i y} 2^{-|x|}
+%\]
 
-For each $i$ we can expand $p_q(x) - p_q'(x, i)$ into its sums and cross out terms where the $x$s match.
+%For each $i$ we can expand $p_q(x) - p_q'(x, i)$ into its sums and cross out terms where the $x$s match.
 Let $\epsilon$ be some positive value.
@@
 
-[@!
+%[@!
 If we want to prove correctness also, we'll need to clean this up.
@@
 \begin{lma}[{\cite{li1993introduction}[p 120]}]
-For any finite set $A$, the proportion of strings with randomness deficieny above $k$ is less than $2^{-k}$.
+For any finite set $A$, the proportion of strings with randomness deficiency above $k$ is less than $2^{-k}$.
 
@@
 \begin{align*}
-\delta(x|p_q) &= \log \frac{m(x)}{p_q(x)}
+\delta(x|p_q) &= \log \frac{2^{-K(x|q)}}{p_q(x)}
 \end{align*}
-Where $m(x)$ is the universal distribution.
 
@@
 \begin{align*}
-\delta(x|p_q) &= - \log p_q(x) - K(x) 
+\delta(x|p_q) &= - \log p_q(x) - K(x|q) 
 \end{align*}
+\end{dfn}
 
-(Note that unlike the finite set case, here we do not make the complexity term conditional on the model) [@! does this matter?]
-
-\end{dfn}
+Note that sometimes the randomness deficiency for measures is defined without the conditional on $q$ in the last term.
+For the sake of consistency with the set-based version, we include the conditional here.
 
@@
 
-Note again, that this only holds for a certain level of courseness in our resource classes.
+Note again, that this only holds for a certain level of coarseness in our resource classes.
 
 Another definition of the universal distribution is as $2^{-K^t(x)}$.
-In the non-bounded case the two definitions are equal up to a mulitplicative constant.
-In the recource bounded variant, we can only be certain of the following inequality:
+In the non-bounded case the two definitions are equal up to a multiplicative constant.
+In the resource bounded variant, we can only be certain of the following inequality:
 \begin{align*}
@@
 
-\section*{Resource-bounded probability distributions}
+\subsection*{Resource-bounded probability distributions}
 
@@
 
-%%\begin{dfn}
-%%We say that a probability distribution $q$ is outside a class ${\cal C}^t$ iff there is no $q' \in {\cal C}^t with \forall x : p_q(x) = p_{q'}(x)$
-%%\label{dfn:outside}
-%%\end{dfn}
-
-%%\begin{dfn}
-%%We say that a probability distribution $q$ is weakly outside ${\cal C}^t$ iff it is outside ${\cal C}^t$ and there exists at least one $x$ for which there is no $q' \in {\cal C}^t$ and $p$ such that $-\log p_{q'}(x) \leq -\log p_q(x)$.
-%%\label{dfn:weakly-outside}
-%%\end{dfn}
-
-%%\begin{dfn}
-%We say that a probability distribution $q$ is strongly outside ${\cal C}^t$ iff it is outside ${\cal C}^t$ and for all $q' \in {\cal C}^t$ there exist an infinite number of $x$  such that $-\log p_{q}(x) \leq -\log p_{q'}(x)$.
-%\label{dfn:strongly-outside}
-%\end{dfn}
-
-%The distinction is important for the following reason.
-If $q$ is weakly but not strongly outside ${\cal C}^t$, there will be a finite set $X$ for which $-\log p_q(x) \leq -\log p_{q'}(x)$ for any $q' \in {\cal C}^t$.
-For all other $x$, $K^t$ will minorize $q'$.
-We can define a Turing machine which `hardcodes' $X$ for the inputs that they have in $q$.
-This machine will run in linear time.
-If this machine is in ${\cal C}^t$, $q$ must be in ${\cal C}^t$ as well, and any machine that is outside ${\cal C}^t$ must be strongly outside ${\cal C}^t$.
-
-%Whether we can truly eliminate distributions that are weakly outside a resource class depends on the specifics of the function that defines it and the details of the computational metaphor (in our case Turing machines).
-To allow our results to be independent of both, we will retain the distinction between weak and strong `outsideness'.
-
-%\begin{lma}
-%If $q$ is strongly outside ${\cal C}^t$ we have:
-%\begin{align*}
-%\forall {c \exists x : K^t(x) \geq - \log p_q(x) + c}
-%\end{align*}
-%\label{lma:strongly-outside}
-%\end{lma}
-%\begin{proof}
-%Assume for a contradiction
-%\begin{align*}
-%\exists c \forall x : K^t(x) < -\log p_q(x) + c
-%\end{align*}
-%Since $u^t \in {\cal C}^t$, we have infinitely many $x$, such that 
-%\begin{align*}
-%-\log p_q(x) &\leq - \log m^t(x) \\
-%-\log p_q(x) &\leq K^t(x) \\
-%-\log p_q(x) &\leq K^t(x) < -\log p_q(x) +c
-%\end{align*}
-%\end{proof}
-
-\section*{Inference from bounded distributions}
+\section*{Approximating randomness}
 
@@
 \begin{thm}
-Let $q \in {\cal C}^t$. The probability of sampling a $(t, k)$-bastard from $q$ is negligable for large $k$.
+Let $q \in {\cal C}^t$. The probability of sampling a $(t, k)$-bastard from $q$ is negligible for large $k$.
 \end{thm}
@@
 
+Considering this result, we may ask whether there exist distributions for which bastards are non-negligible. We show the following:
+
 \begin{thm}
-Let $q$ dominate ${\cal C}^t$. The probability of sampling a $(t, k)$-bastard from $q$ is bounded below by a constant value, independent of $k$ (and therefore, not negligable).
+There exists a probability distribution $q$ outside ${\cal C}^t$, such that the probability of sampling a $(t, k)$-bastard from $q$ is at least the inverse of a polynomial over $k$.
 \end{thm}
+\begin{proof}
+Consider the following algorithm.
+given $n \in {\mathbb N}$:
+\begin{itemize}
+  \item Calculate $K^t(x)$ for all $x$ of length $n$ 
+  \item Assign the first $x$ such that $K^t(x) = |x|$ the probability $c \cdot n^{-2}$
+  \item Assign all other $x$ probability zero.
+\end{itemize}
+If we set $c$ to $6 / \pi^2$, the probabilities converge to $1$. For each string produced with nonzero probability, we have $K^t(x) = |x|$:
 
-\begin{proof}
-By definition \ref{dfn:domination}, we have 
-\begin{align}
-\forall {c \exists x : K^t(x) \geq - \log p_q(x) + c} \label{line:non-invariance}
-\end{align}
+So
+\begin{align*}
+p_q(B^t_k) &= \sum_{n : n - 2\log n + O(1) \geq k} cn^{-2}\\
+p_q(B^t_k) &> cn^{-2} \quad\text{with}\quad n -2\log n + O(1) = k\\
+\end{align*}
+Taking a rough bound $n/2 < k$
+\begin{align*}
+p_q(B^t_k) &> cn^{-2} > c4 \;k^{-2} = c' k^{-2}\\
+\end{align*}
+\end{proof}
 
-This rewrites to
-\begin{align*}
-\forall c \exists x : \frac{2^{-K^t(x)}}{p_q(x)} \leq 2^{-c}
-\end{align*}
+Not that we can compute this $q$ in $2^n+1\cdot t(n)$. Thus if the computational power of our source is large enough, the probability of bastards becomes non-negligible.
 
-Since we know that $2^{-K^u(x)} \geq p_q(x) c_q$ we get
-\begin{align}
-\forall c \exists x : \frac{2^{-K^t(x)}}{2^{-K^u(x)}} \leq \frac{2^{-K^t(x)}}{p_q(x)c_q} \leq 2^{-c}\frac{1}{c_q}
-\label{line:prelim}
-\end{align}
+\section*{Approximating model size}
 
-The statement to be proved asks for the probability of seeing a $(t, k)$-bastard. We get
-\begin{align*}
-p_q\left(B^t_k\right) &\geq p_q\left (\left\{x: K^t(x) - K(x) > k\right\}\right) \\
-	&\geq p_q\left(\left\{x: K^t(x) - K^u(x) > k\right\}\right) \\
-	&= p_q\left(\left\{x : \frac{2^{-K^t(x)}}{2^{-K^u(x)}} < 2^{-k}\right\}\right) \\
-\end{align*}
-
-From (\ref{line:prelim}), we know that there is at least one element in the set in the last line, whatever $k$. Call this element $x'$, we have:
-\begin{align*}
-p_q\left(B^t_k\right) \geq p_q\left(x'\right)
-\end{align*}
-We know that $p_q\left(x'\right)$ cannot be zero, because, then by (\ref{line:non-invariance}), its code length under $K^t$ would be infinite.
-Thus, there is always a lower bound, independent of $k$ on the probability of seeing a $(k, t)$-bastard.
-\end{proof}
+\section*{Approximating Normalized information distance}
 

### 2013-07-02 Bastards/src/bastards.tex
...

@@
 \usepackage{subcaption}
+\usepackage{enumerate}
 
@@
 
-We must take some care in defining what it means for a Turing machine to be outside ${\cal C}^t$.
-If, for instance, a Turing machine $q$ has no equal in ${\cal C}^t$ but only has a finite number of inputs for which it runs longer than $t(|x|)$, we can `hardcode' these inputs at a linear penatly, which would give the Turing machine an equivalent in ${\cal C}^t$, so long as the linear penalty keeps it there.
+\section*{Generalizing bounded Kolmogorov complexity}
 
-To have a strong definition of Turing machines outside ${\cal C}^t$, we take a defining property of ${\cal C}^t$, the fact that $K^t$ minorizes it, and invert it.
+We note that in many applications of resource-bounded Kolmogorov complexity, the only properties of the resource bounds that are used, are the fact that Kolmogorov complexity is reduced as the resources are increased, and the fact that there exist universal Turing machines that simulate all Turing machines in a given resource class.
+This leads us to generalize beyond the resource bounds to a general bounded Kolmogorov complexity.
 
 \begin{dfn}
-A Turing machine $q$ dominates ${\cal C}^t$ iff:
-\begin{align*}
-\forall c \exists x : K^t(x) \geq -\log p_q(x) + c
-\end{align*}
-\label{dfn:domination}
+A \emph{model class} $\{T_i\} \subseteq \cal C$ is a computably enumerable set of Turing machines.
+A \emph{universal model} for $C$ is a Turing machine such that $U^C(\langle i, p\rangle) = T_i(p)$.
+If a model class includes a universal model for itself, we call it \emph{inclusive}.
+If all models in a model class are total, we call the model class \emph{total}.
 \end{dfn}
 
-We don't know whether all Turing machines outside ${\cal C}^t$ dominate, but we do know that some do (including any $K^u$ with $u > t$) and that no Turing machine in ${\cal C}^t$ dominates it.
+\begin{dfn}
+The model-bounded Kolmogorov complexity $K^C(x)$ for the model class $C$ with universal model $U^C$ is defined as
+\[
+K^C(x) = \min \{|p| \mid U^C(p) = x\}
+\]  
+\end{dfn}
 
 \section*{Approximating randomness}
+
+\begin{lma}
+We have the following properties
+\begin{enumerate}[(a)]
+  \item For any $T_m \in C$ we have $K^C(x) \leq c_m + K_m(x)$
+  \item For model classed $C_0$ and $C_1$, with $C_0 \subseteq C_1$, we have $K^{C_1}(x) \leq K^{C_0}(x)$
+  \item $K^C(x)$ is computable for any total model class.
+\end{enumerate}
+\end{lma}
+
+The Turing machines bounded by $t$ form a total, inclusive model class. Another example is the model class of read-only Turing machines (or DFAs), which is total, but not inclusive.
+
+We define $\text{depth}^C$ and $m^C$ analogous to the time-bounded versions.
 
@@
 
+This result generalizes to model any model class without change.
+
 Considering this result, we may ask whether there exist distributions for which bastards are non-negligible. We show the following:
@@
 
-Not that we can compute this $q$ in $2^n+1\cdot t(n)$. Thus if the computational power of our source is large enough, the probability of bastards becomes non-negligible.
+Note that we can compute this $q$ in $2^{|p|+1}\cdot t(|p|)$, with $p = 2 \log(|x|) + O(1)$.
+Thus if the computational power of our source is large enough, the probability of bastards becomes non-negligible.
 
-\section*{Approximating model size}
+The result generalizes to any total model class $C$. For non-total model classes (such as $\cal C$), the first line of the algorithm may not be computable.
 
-\section*{Approximating Normalized information distance}
+\section*{Approximating Model Size}
+
+There are various measures of information that equate the size of an optimal model in two-part coding with the level `interestingness' or meaningful information in a string.
+The include the Kolmogorov structure functions, coarse sophistication and facticity.
+
+\begin{dfn}{\cite{}}
+The \emph{coarse sophistication} of a string $x$ is 
+\[
+	\text{soph}_c(x) = \min \left \{ K(f) : K(f) + K_f(x) \leq c \right \}
+\]
+\end{dfn}
+
+\begin{dfn}{\cite{}}
+The \emph{sophistication} of a string $x$ is 
+\[
+	\text{soph}(x) = \min \left \{ K(f) : K(f) + K_f(x) \overset{\underset{+}{}}{=} K(x) \right \}
+\]
+\end{dfn}
+
+\begin{dfn}{\cite{}}
+The \emph{facticity}  of a string $x$ is 
+\[
+\text{fact}(x) = \min \left\{ |\bar{\imath}| : \exists p \;\text{st.}\; |\bar{\imath}p| = K_2(x) \;\&\; U(\bar{\imath}p) = x\right\}
+\] with $U$ a universal Turing machine with a prefix-based pairing function, and $K_2 = \min \{|\bar{\imath}p| : U|\bar{\imath}p| = x\}$
+\end{dfn}
+
+\section*{Approximating Normalized Information Distance}
+
+\begin{dfn}
+The normalized information distance between two string $x$ ad $y$ is 
+\[
+NID(x, y) = \frac{\max \left [ K(x\mid y), K(y\mid x)\right]}{\left[ K(x), K(y)\right]} 
+\]
+\end{dfn}
+
+As shown in \cite{}, this function is neither lower nor upper semicomputable.
+An approximation nevertheless exists in the form of the Normalized Compression Distance \cite{}.
+This function (which replaces the $K$'s with general purpose compressors) has been used to great effect with good experimental results without a sound theoretical reason for why it should work.
 

### 2013-07-02 Benelearn/src/benelearn.tex
...

@@
 We conclude that, even under adverse circumstances, sequential representations contain enough structure for shallow algorithms to perform inference successfully.
-Algorithms that operate directly on the graph representation likely require a considerable increase in resources, but do allow for an increase in performance also.
+Algorithms that operate directly on the graph representation usually require a considerable increase in resources, but do allow for an increase in performance also.
 \end{abstract}
@@
 If this issue really `blinds' the sequential compressors to the structures in the graph, one option is to investigate compressors that operate on the level of the graph representation, for instance by finding frequent subgraphs or clustering the graph.
-The methods do not suffer this problem of isomorphism, but as a result, they are more expensive than their sequential cousins.
-In this paper we hope to provide a first indication of how far \emph{the sequential approach} will go, and whether \emph{the native approach} will let us continue on from there.
+The methods do not suffer this problem of isomorphism, but as a result they are more expensive than their sequential cousins.
+In this paper we hope to provide a first indication of how far the sequential approach will go, and whether the native approach will let us continue on from there.
 
-There is a wealth of research available on machine learning and data mining both within single graphs and on sets of graphs.
-See for instance, the book \cite{cook2006mining}.
-In this paper we do not share the data mining goal of extracting interesting features from the data, but only the goal of performing inference, using the compressors and their learning techniques as black boxes, and evaluating the the results of a chosen inference task.
+There is a wealth of research available on machine learning and data mining both within single graphs and on sets of graphs (see for instance \npcite{cook2006mining}).
+In this paper we do not share the data mining goal of extracting interesting features and models from the data, but only the goal of performing inference, using the compressors and their learning techniques as black boxes, and evaluating the the results of a chosen inference task.
 
@@
 
-\subsection*{Kolmogorov Complexity}
+\subsection*{Kolmogorov complexity}
 
-Kolmogorov complexity is a notion of information content, based on two principles: (a) all data can be represented as a bitstring (b) the shorter this string can be described, the less information is contained in it.
+Kolmogorov complexity is a notion of information content based on two principles: (a) all data can be represented as a bitstring (b) the shorter this string can be described, the less information is contained in it.
 
@@
 First, by `description' we mean a description in a formal language that is maximally expressive.
-To formalize expressivenes, we require that the method of description is Turing-complete (of equivalent strength to a Turing machine).
+To formalize expressiveness, we require that the method of description is Turing-complete (of equivalent strength to a Turing machine).
 By the strong Church-Turing thesis, this suggests that there is no reasonable way of defining a more expressive method of description.
@@
 The result that makes Kolmogorov complexity a useful measure of information content is that $K_U(\cdot)$ is only marginally dependent on the choice of $U$.
-If we suppose that there is another universal description method, $V$, we might ask what the expected difference is between $K_U(x)$ and $K_U(v)$.
-Let $k_V(x)$ be the shortest program for $x$ on $V$.
-Since, $U$ is universal, we know that it can compute $k_v$ by simulating $V$.
+If we suppose that there is another universal description method, $V$, we might ask what the expected difference is between $K_U(x)$ and $K_V(x)$.
+Let $k_V(x)$ is the shortest program for $x$ on $V$.
+Since, $U$ is universal, we know that it can compute $k_V$ by simulating $V$.
+Somewhere in our enumeration of Turing machines $\{T_i\}$, there is a $T_v$ which computes the same function as $V$.
 This simulation is a program for $x$ on $U$ which is a bound for the shortest program on $U$:
 \begin{align*}
-K_U(x) 	&\leq |\langle T_v, k_v(x)\rangle|\\ 
-		&=  |k_v(x)| + |v| + p(v) \\
+K_U(x) 	&\leq |\langle T_v, k_V(x)\rangle|\\ 
+		&=  |k_V(x)| + |v| + p(v) \\
 		&=  K_V(x) + O(1)
@@
 \begin{description}
-\item[$K(\cdot)$ is uncomputable.] There can be no algorithm which computes the Kolmogorov Complexity of $x$ for all $x$.
-It can, however, be bounded from above, and for every algorithm which bounds it, there is an algorithm which provides a better bound.
-\item[$K(\cdot)$ is approximated by all computable compressors.] If we have some compressor for our data $x$ (say GZIP) we can find the decompression algorithm somewhere in $\{T_i\}$, say as $T_g$.
+\item[$K(\cdot)$ is uncomputable] There can be no algorithm which computes the Kolmogorov complexity of $x$ for all $x$.
+It can, however, be bounded from above, and for every algorithm which bounds it, there is another algorithm which provides a better bound.
+\item[All computable compressors approximate $K(\cdot)$] If we have some compressor for our data $x$ (say GZIP) we can find the decompression algorithm somewhere in $\{T_i\}$, say as $T_g$.
 We can have a description on $U$ as $U(\langle g, z\rangle) = x$ so that $K_U(x)\leq |z| + O(1)$.
@@
 
-Finally, we define conditional Kolmogorov Complexity $K(x| y)$.
+Finally, we define conditional Kolmogorov Complexity $K(x \mid y)$.
 Where regular Kolmogorov Complexity is defined as the shortest program which produces $x$, the conditional variant is defined as the shortest program which produces $x$ when provided with $y$.
@@
 
-We would like to approximate this by replacing each occurence of the Kolmogorov complexity with an approximation by a compressor, which we will call $C$.
+We would like to approximate this by replacing each occurrence of the Kolmogorov complexity with an approximation by a compressor, which we will call $C$.
 As most compressors do not work on a conditional basis (expressing data given some existing data), we want to rewrite the conditional $K$'s as nonconditional ones.
@@
 NID(x, y)	 &= \frac{\max \left [K(x, y) - K(x),K(y, x) - K(y)\right ] }{\max \left [K(x), K(y) \right ]} \\ 
-	&= \frac{\max \left [K(xy) - K(x),K(yx) - K(y)\right ] } {\max \left [K(x), K(y) \right ]} \\
-	&= \frac{\min \left [K(xy), K(yx)\right ] - \min \left[K(x), K(y)\right]}{\max \left [K(x), K(y) \right ]} 
+	&= \frac{\max \left [K(xy) - K(x),K(yx) - K(y)\right ] } {\max \left [K(x), K(y) \right ]}
+%	&= \frac{\min \left [K(xy), K(yx)\right ] - \min \left[K(x), K(y)\right]}{\max \left [K(x), K(y) \right ]} 
 \end{align*}
@@
 NCD(x, y) 
-	&= \frac{\min \left [C(xy), C(yx)\right ] - \min \left[C(x), C(y)\right]}{\max \left [C(x), C(y) \right ]}\\
+%	&= \frac{\min \left [C(xy), C(yx)\right ] - \min \left[C(x), C(y)\right]}{\max \left [C(x), C(y) \right ]}
 	&= \frac{C(xy) - \min \left[C(x), C(y)\right]}{\max \left [C(x), C(y) \right ]} 
@@
 
-Where the last step follows from the assumption that our compressor is roughly symmetric ($C(xy) = C(yx)$).
+This step also includes the assumption that our compressor is roughly symmetric ($C(xy) = C(yx)$).
 
-When our data is represented as a graph, rather than a string, we replace the notion of concatenation strings by of concatenation of graphs.
+When our data is represented as a graph, rather than a string, we replace the notion of concatenation of strings by concatenation of graphs.
 That is, we combine the graphs $x$ and $y$ into a single (disconnected) graph.
@@
 
-Our aim is to test a sequential and a graph-based compressor on an inference tasks for a variety of graph data.
+Our aim is to test a sequential and a graph-based compressor on an inference task for a variety of graph data.
 To ascertain the performance of the compressors, we generate graphs from different sources, calculate their NCD distances and see whether a clustering algorithm can reconstruct the original sources as clusters.
@@
 An important and subtle concern is the ordering of nodes in the sequential representation of our graphs.
-This issue is detailed very well in \cite{kang2011beyond}.
-As shown, there are various algorithms to determine node orderings that bring out a lot of the graph's inherent structure in the adjacency matrix, allowing a sequential compressor to use it.
+This issue is detailed very well by \npcite{kang2011beyond}.
+As shown, there are various algorithms to determine node orderings that bring out a lot of the graph's inherent structure in the adjacency matrix, allowing a sequential compressor to exploit it.
 We could use substantial resources to find a good ordering of nodes to improve the performance of the sequential compressor.
-I we did, however, we might lose the ability to consider the sequential compressor a shallow model.
+If we did, however, the extra computation might mean that the compressor is no longer a shallow model.
 To maintain the sequential compressor as a representative example of shallow models, the node ordering should be cheap to establish from a random ordering, preferably in linear time.
@@
 Since we are testing the capacity of the general purpose compressor to perform inference despite the problem of isomorphism, we will actually present it with a worst case scenario.
-We will use a random ordering of nodes for all graphs.
+We use a random ordering of nodes for all graphs.
 If the general purpose compressor still outperforms the random baseline under these conditions, it will tell us that it is, at least in part, resistant to the problem of isomorphism.
@@
 We also include graphs from the fractal graph generation algorithm from \cite{song2006origins}.
-We set the hub-parameter which determines the level of fractality (as a trade-off with the level of small worldness) to $0.0$ (for a small world network) and to $1.0$ (for a fractal network).
+We set the hub-parameter which determines the level of fractality (as a trade-off with the level of small-worldness) to $0.0$ (for a small world network) and to $1.0$ (for a fractal network).
 
@@
 In this experiment we sample subgraphs from large, existing graphs.
-We sample by choosing a uniform random node, and performing a random walk of length $n$.
+We sample by choosing a  random node uniformly from all nodes and performing a random walk of length $n$.
 We then extract a subgraph containing the nodes encountered and any links connecting two encountered nodes.
@@
 	\item[cellular] The cellular network of the E. Coli bacterium. \cite{jeong2000large} Acquired from \url{http://www.nd.edu/~networks/resources/metabolic/index.html}
-	\item[neural] The neural network of the C. Elegans nematode worm (ignoring link directions). \cite{achacoso1991ay,watts1998small} Acquired from \url{http://toreopsahl.com/datasets/#celegans}
+	\item[neural] The neural network of the C. Elegans nematode worm (ignoring link directions). \cite{achacoso1991ay,watts1998collective} Acquired from \url{http://toreopsahl.com/datasets/#celegans}
 	\item[co-purchase] A graph of items commonly purchased together on internet retailer Amazon.com. \cite{leskovec2007dynamics} Acquired from \url{http://snap.stanford.edu/data/amazon0302.html}
@@
 
-Subdue \cite{ketkar2005subdue} is an algorithm for finding frequent subgraphs in graph data.
+Subdue \cite{jonyer2004mdl,ketkar2005subdue} is an algorithm for finding frequent subgraphs in graph data.
 The algorithm searches for the subgraph that maximally compresses the data.
@@
 \item[Subgraph matching] This is an algorithm for finding the occurrences of a given subgraph in a graph.
-The method used is detailed in \cite{bunke1983inexact}.
+The method used is detailed by \npcite{bunke1983inexact}.
 Since this is  a semi-exhaustive search for the solution to the NP-complete problem of subgraph isomorphism, the matching can only be solved for very small subgraphs.
 Unfortunately, even with subgraphs of four or five nodes, the matching is too slow in combination with the number of times it is executed to calculate a full distance matrix.
-To combat this issue, we remove all but the first $b_{\mbox{inner}}$ elements from the search queue after it is sorted at ech iteration, effectively turning the algorithm into a beam search.
+To combat this issue, we remove all but the first $b_{\mbox{inner}}$ elements from the search queue after it is sorted at each iteration, effectively turning the algorithm into a beam search.
 \item[MDL Scoring]
-This routine takes a subgraph, finds its occurrences in the data by the previous routine and replaces these occurrences with a single node with a new label, which we call a \emph{symbol node}.
-All links that connected to nodes in the subgraph are now connected to the symbol node.
-The links connecting to the symbol node are annotated with the indices of the nodes in the subgraph they were connected to.
-Storing the subgraph once together with the reduced data graph is enough to reconstruct the original data, and, if the subgraph occurs often, will lead to a significant compression.
+This routine takes a subgraph, finds its occurrences in the data by the previous routine and deletes these from the data.
+The subgraph is then stored once, together with the remainder of the data and a list of where the subgraph should be attached to reconstruct the original data.
+See the appendix for an exact description.
 \item[Subgraph search]
@@
 
-Algorithm \ref{pseudo:algorithm} shows a broad description of the whole algorithm.
+Algorithm \ref{pseudo:algorithm} shows a broad description of the whole procedure.
 
-\begin{pseudo}[h]
+\begin{pseudo}
 {
@@
 	\hspace*{5mm} $s \leftarrow $ \textit{head element of} $S$ \\
-	\hspace*{5mm} \textit{add all extensions of} $s$ \textbf{to} $S$ \\
-
+	\hspace*{5mm} \textit{add all extensions of} $s$ \textbf{to} $S$ 
+	\\
 	\hspace*{5mm}\textit{sort $S$ by} score($s', G$) \textit{for} $s' \in S$\\
@@
 	\caption{Pseudocode for the Subdue algorithm}
-
 	\label{pseudo:algorithm}
@@
 The graph matching search (the first line of the \textbf{score} function) allows for inexact matches of the subgraph.
-In these cases, we use a rough estimate of the number of bits required to transform the stored subgraph into the subgraph that is actually present in the data.
+In these cases, we use a rough upper bound of the number of bits required to transform the stored subgraph into the subgraph that is actually present in the data.
 
@@
 Note that this makes the BA graphs UAGs.
-For $k = 2$ the clustering problem would be more difficult.
+For more nodes attached per step the clustering problem would become more difficult.
 The random graphs are generated with the exact same number of nodes and links.
@@
 During the search we return only one best subgraph.
-Our beamwidth at the top level ($b_{\mbox{outer}}$) is set to $5$.
+Our beam width at the top level ($b_{\mbox{outer}}$) is set to $5$.
 The beam width in the graph matching routine ($b_{\mbox{inner}}$) is set to $10$.
+We run the search for $10$ iterations, limiting the size of the subgraph used to $5$.
 
@@
 
-Table \ref{table:synthetic-large} shows the results on randomly generated graphs. Table \ref{table:real-life} shows the results for subgraphs sampled from real-life datasets.
+\begin{table}
 
-\begin{table}[h]
+\caption{Confusion matrices for various compressors.
+Columns represent the clusters found by the k-medoids algorithm.
+To calculate the error, we label the resulting clusters so that the error is minimized (ie.
+reorder the columns of the confusion matrix to maximize the sum of the diagonal).
+We report the mean error (1 - the sum of the diagonal) over 10 experiments (and the standard deviation in brackets) below each confusion matrix.
+The confusion matrix shown is always for the first experiment in the run.}
 
@@
   BA                    &  0    & 0.25 & 0    & 0 \\
-  fractal (pure)        &  0    & 0    & 0 & 0.25 \\
-  fractal (small world) &  0.25 & 0    & 0 & 0 \\
+  fractal (pure)        &  0    & 0    & 0.25 & 0 \\
+  fractal (small world) &  0    & 0    & 0    & 0.25 \\
 \hline
@@
 \hline
-  ER                    &  0    & 0.25 & 0    & 0    \\
-  BA                    &  0    & 0    & 0.25 & 0    \\
-  fractal (pure)        &  0    & 0    & 0    & 0.25 \\
-  fractal (small world) &  0.25 & 0    & 0    & 0    \\
+  ER                     & 0.25 & 0    & 0    & 0\\
+  BA                     & 0    & 0.25 & 0    & 0\\
+  fractal (pure)         & 0    & 0    & 0.25 & 0\\
+  fractal (small world)  & 0    & 0    & 0    & 025\\
 \hline
@@
 \end{subfigure}
-\caption{Confusion matrices for various compressors.
-Columns represent the clusters found by the k-medioids algorithm.
-We have labeled the resulting clusters so that the error is minimized (ie.
-reordered the columns of the confusion matrix to maximize the sum of the diagonal).
-We report the mean error (1 - the sum of the diagonal) over 10 experiments (and the standard deviation in brackets) below each confusion matrix.
-The confusion matrix shown is always for the first experiment in the run.}
 \label{table:synthetic-large}
@@
 \begin{table}
+
+\caption{Results for the experiment on natural datasets.}
 
@@
   neural        &  0.11 & 0.11 & 0.11 \\
-  co-purchase   &  0.22 & 0    & 0.11 \\
+  co-purchase   &  0.11 & 0    & 0.22 \\
 \hline
@@
 \hline
-  cellular      &  0 & 0.33 & 0 \\
-  neural        &  0.11 & 0.22 & 0 \\
-  co-purchase   &  0.22 & 0 & 0.11 \\
+  cellular      & 0.33 & 0 & 0 \\
+  neural        & 0.22 & 0.11 & 0 \\
+  co-purchase   & 0    & 0.22 & 0.11 \\
 \hline
@@
 \hline
-  cellular      &  0    & 0.33 & 0 \\
-  neural        &  0.33 & 0    & 0 \\
-  co-purchase   &  0    & 0.33 & 0 \\
+  cellular     & 0.33 & 0 & 0 \\
+  neural       & 0    & 0.33 & 0 \\
+  co-purchase  & 0.33 & 0 & 0 \\
 \hline
@@
 \end{subfigure}
-\caption{Results for the experiment on natural datasets.}
 \label{table:real-life}
 \end{table}
+
+Table \ref{table:synthetic-large} shows the results on randomly generated graphs. Table \ref{table:real-life} shows the results for subgraphs sampled from real-life datasets.
 
@@
 
-This result suggests that inference on graphs can be performed by sequential algorithms on a sequential representation in linear time, with decent results.
+This result suggests that at least some inference on graphs can be performed by sequential algorithms on a sequential representation in linear time, with decent results.
 
@@
 To investigate the promise of these models as compressors further, it will be necessary to investigate both parallelized versions of these algorithms and a more elegant relaxation of the exhaustive nature of their components.
+Subgraph sampling methods like the ones detailed by \npcite{kashtan2004efficient}, may be able to provide a significant increase in performance.
 
@@
 
-We first store the number of nodes $n$ in preefirx free coding, and the maximum number of neighbours for a node in the graph $n_{\mbox{max}}$ (int $\log n$ bits).
+We first store the number of nodes $n$ in prefix free coding, and the maximum number of neighbours for a node in the graph $n_{\mbox{max}}$ (in $\log n$ bits).
 We then store the lower half of the adjacency matrix (including the diagonal) row by row.
@@
 For node $i$, we only need to store the connections to the $i$ nodes below it including itself.
-We use $l(n_{\mbox{max}})$ bits to store the number of such neighbours $n_i$, and $\log{i \choose n_i}$ bits to store the configuration of those neighbours.
+We use $\log n_{\mbox{max}}$ bits to store the number of such neighbours $n_i$, and $\log{i \choose n_i}$ bits to store the configuration of those neighbours.
 
-We then account for the possibility of multiple links between node pairs.
-Since we expect miultiple links to be unlikely, we want to ensure that we do not waste many bits coding for this occasion if it does not occur.
+% We then account for the possibility of multiple links between node pairs.
+Since we expect multiple links to be unlikely, we want to ensure that we do not waste many bits coding for this occasion if it does not occur.
 We store the pairs for which there are multiple links as a list consisting of two numbers per element: the first is the index of the pair in a canonical ordering, the second is the number of extra links that exist for that pair (in prefix-coding).
@@
 
-In the rest of the graph, we replace the occurrences of the substructure with a symbol node and store this graph by the method detailed for plain graphs.
-We store the number of symbol nodes in prefix coding, and only encode labels for the rest of the nodes (again, by the codebook of the full data).
+In the rest of the graph, we remove the nodes matched to the substructure and all links connecting to them.
+We store the `silhouette' of the substructure as a plain graph (again with the codebooks of the whole graph).
 
-Finally, we must store for every link leading into a symbol node which node in the substructure it should connect to to restore the original data.
-Since we know the graph and all the symbol nodes, we can presume there to be a canonical ordering in these links, and we use $\log n_s$ bits for each, where $n_s$ is the size of the substructure in nodes.
+We then store the way the substructures should be connected into the silhouette to reconstruct the original graph.
+For each substructure we first store the transformation cost (if the substructure was an inexact match), with a prefix penalty to make the description self-delimiting, we then store the number of links connecting the substructure un prefix-coded form, and then for each link we take $\log s$ bits to encode how to connect it in the substructure and $\log d$ to connect it in the rest of the graph, where $s$ is the number of nodes in the substructure and $d$ is the number of nodes in the silhouette.
+
+In later versions of our code, the occurrences of the subgraph are replaced by symbol nodes, but the version used to perform these experiments uses the silhouette method.
 

### 2013-07-02 Deliverable 13Q2/src/13q2.tex
...

\documentclass{article}

%\usepackage{amsthm}
\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{enumerate}
\usepackage{cancel}
\usepackage{hyperref}

\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\DeclareMathOperator*{\argmin}{\arg\,\min}
\DeclareMathOperator*{\argmax}{\arg\,\max}

\newcommand{\sdr}[1]{\textcolor{blue}{\small #1\textsuperscript{[sdr]} }}
\newcommand{\pb}[1]{\textcolor{OliveGreen}{\small #1 \textsuperscript{[pb]} }}

\newcommand{\p}{\mbox{\,.}}
\newcommand{\fl}[1]{\left \lfloor #1 \right \rfloor}
\newcommand{\g}[1]{\color{gray} #1 \color{black}}
\newcommand{\hide}[1]{}

\author{Peter Bloem \and Pieter Adriaans}

\title{Empirical survey of complexity measures on graphs, focusing on RDF data and Algorithmic Statistics}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

We have scaled up the results from the previous work package in data volume and added various complexity measures. We will process more datasets, and attempt to establish interesting empirical results based on compression and complexity analysis. The existing implementations of complexity measures from 2012Q3 are included as a comparison.

The purpose of this deliverable is threefold. It validates that our methods, such as they are, can be used to provide useful empirical results. It will serve as a test of our software and make sure that it is maintained (preventing 'code rust'). And finally, it will provide us with feedback for which features of our code are interesting enough to scale up in Q3.

\section{Datasets}

\section{Complexity measures}

\section{Results}

\section{Interpretation}

\end{document}

### 2014-01-16 PhD Thesis/src/thesis.tex
...

\documentclass{thesis}

%\usepackage{amsthm}
%\usepackage{charter}
%\usepackage{eulervm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{enumerate}
\usepackage{cancel}
\usepackage{hyperref}

%\theoremstyle{definition}
\newtheorem{thm}{Theorem}
\newtheorem{crl}{Corollary}
\newtheorem{lma}{Lemma}
\newtheorem{dfn}{Definition}
\newtheorem{exm}{Example}

\newenvironment{out}
 {\emph{Proof outline.\;}}
 {}

%\let\doendproof\endproof
%\renewcommand\endproof{~\hfill\qed\doendproof}

\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\DeclareMathOperator*{\argmin}{\arg\,\min}
\DeclareMathOperator*{\argmax}{\arg\,\max}
\DeclareMathOperator*{\nid}{NID}
\DeclareMathOperator*{\id}{ID}

\newcommand{\sdr}[1]{\textcolor{blue}{\small #1\textsuperscript{[sdr]} }}
\newcommand{\pb}[1]{\textcolor{OliveGreen}{\small #1 \textsuperscript{[pb]} }}

\newcommand{\p}{\mbox{\,.}}
\newcommand{\fl}[1]{\left \lfloor #1 \right \rfloor}
\newcommand{\g}[1]{\color{gray} #1 \color{black}}
\newcommand{\B}{{\mathbb B}}
\newcommand{\R}{{\mathbb R}}
\newcommand{\N}{{\mathbb N}}
\newcommand{\m}{{\overline{m}}}
\newcommand{\ok}{{\overline{\kappa}}}
\newcommand{\mmid}{\;\middle|\;}

\newcommand{\hide}[1]{}

\title{Statistics for complex graphs: an algorithmic approach}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\chapter{Introduction}

% Start with a story
% Mention RDF: the future of data storage

\section{The complex graph}

% Complex graphs are not just funny things to look at, the are the fundamental unit of science in the 21st century.

\section{Computer programs as statistical models}

\chapter{Statistical analyses of complex graphs}

\begin{summary}We review the literature on complex graphs, and the approaches that have been used so far to provide statistical analyses on them.
\end{summary}

% Algorithmic statistics is how you stud graphs. Nothing else makes sense.

\chapter{Algorithmic Statistics: Bridging the gap between Kolmogorov complexity and MDL}

\begin{summary}MDL and Kolmogorov complexity are two offshoots of the same basic idea: that good compression makes good statistics. MDL takes this as an axiom and builds on it, while Kolmogorov complexity makes only the assumption that the data is produced by a computational process. The two methods are not incompatible, but they have produced very different results and techniques. In order to have access to the best of both worlds, we start with the rigorous theoretical foundations of Kolmogorov complexity and build a platform that will allow us to fold in the techniques of MDL. We will revisit the principle of two-part coding, considered to be a usable, but cumbersome technique in MDL and show that when combined with Kolmogorov complexity it has some deep advantages over other methods.
\end{summary}

\chapter{What makes data interesting: two-part coding in action}

\begin{summary}The question of what makes data complex is has been decisively answered. First for IID data in, as maximal Entropy, and then for any data as maximal Kolmogorov Complexity. But there's some dissonance in calling this data information-rich. We are not very interested in watching TV broadcasts of static, or recordings of white noise. This may contain maximal information to a computer, but to us, it's not very interesting. In fact the most interesting data seems to be somewhere in between the zero complexity signal (s tv which is turned off) and the maximum complexity one (analog noise). The question of whether we can capture this notion of `interestingness' in a metric analogous to information entropy and Kolmogorov complexity has occupied research for over half a century, and has produced some surprisingly deep results. This chapter provides a review.
\end{summary}

\section{The question of interestingness}

\chapter{Graph modeling: extracting motifs}

\begin{summary}Modeling complex objects starts with finding out what the components are. If we know the letters, we can find the words. If we know the words, we can find the sentences. Graphs give us nodes and links for free. How do they combine into functional, recurring units? How can we build up to the global structure?
\end{summary}

\chapter{Graph modeling with rewrite grammars}

\begin{summary}In last chapter we induced motifs for various types of complex networks. We ended by applying the procedure recursively, collapsing upwards across several levels to compress the whole structure.
\end{summary}

\chapter{Graph modeling with deep neural networks}

\begin{summary}To model graphs in an algorithmic setting we can use any family of computational structures that can turn randomness into graph structures. One recently popular model that has been shown to be effective at producing random examples of any given model is the Deep Belief Network. We investigate their potential to model graph structure, and fit them in to the framework of algorithmic statistics.
\end{summary}

\chapter{Machine learning on graphs}

% This does not fit the line too well, refocus?

\begin{summary}The classic view of offline machine learning starts with a list of instances from which to learn. These instances can then be represented in some way---preferably a vector of numeric features---and analyzed statistically. In the past decade a situation has emerged and re-emerged that is an ill fit to this paradigm. We are faced with information in a graph: a social network, a web graph, an RDF dataset. We may know our instances, if they correspond to specific nodes in the graph, but the things we know about them are spread throughout the graph. We investigate the problem and provide a bridge to traditional machine learning.
\end{summary}

\section{A pipeline}
\subsection{Instance extraction}
\subsection{Feature extraction}
\section{Methods}
\subsection{Classification}
\subsection{Clustering}
\subsection{Regression}

\chapter{Citation networks, a use case}

\begin{summary}Throughout the previous chapters, we have performed many quick tests on small datasets, to illustrate and validate our points. In this chapter we take one use case and investigate more thoroughly, in order to show the full potential of our methods. We choose the domain of citation networks, and investigate various questions.
\end{summary}

\begin{itemize}
  \item Who are the hubs?
  \item H-index: can we do better?
  \item Multi-resolution analysis
  \item Visualization
\end{itemize}

\chapter{Statistics for complex graphs}

\begin{summary}\ldots
\end{summary}

Conclusatory chapter.

\appendix
\chapter{Tutorials}
\section{Kolmogorov Complexity}
\section{Minimum Description Length}
\section{Graphs and RDF}
\chapter{Algorithms and techniques}
\chapter{Proofs}

\end{document}

### 2014-10-09 Facticity/Facticity/facticity.tex
...

\documentclass{style/llncs}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{enumerate}
\usepackage{cancel}
\usepackage{hyperref}
\usepackage{footmisc}
\usepackage{mathrsfs}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\title{Ideal model selection is impossible}

% Steven's comments
\newcommand{\sdr}[1]{\textcolor{blue}{\small #1\textsuperscript{[sdr]} }}
% Peter's comments 
\newcommand{\pb}[1]{\textcolor{OliveGreen}{\small #1 \textsuperscript{[pb]} }}
% Pieter's comments 
\newcommand{\pa}[1]{\textcolor{Red}{\small #1 \textsuperscript{[pa]} }}

\newcommand{\p}{\mbox{\,.}}
\newcommand{\fl}[1]{\left \lfloor #1 \right \rfloor}
\newcommand{\g}[1]{\color{gray} #1 \color{black}}
\newcommand{\B}{{\mathbb B}}
\newcommand{\Rbb}{{\mathbb R}}
\newcommand{\N}{{\mathbb N}}
\newcommand{\C}{{\mathscr C}}
\newcommand{\Co}{{\overline{\mathscr C}}}
\newcommand{\R}{{\mathscr R}}
\newcommand{\m}{{\overline{m}}}
\newcommand{\ok}{{\overline{\kappa}}}
\newcommand{\s}[1]{{\overline{#1}}}
\newcommand{\mmid}{\;\middle|\;}

\begin{document}

\maketitle

\abstract{The idea behind Sophistication is that a dataset can be expressed in a single, canonical two-part coding: a model, and a code given that model. When the best compressing model is chosen over all computable models, the result is a representation that can be seen as an inherent property of the data. The data belongs to a single model. This can be seen as a kind of MDL model selection in the limit. We show that there are serious issues with this notion, and that a single sample from a probabilistic data source is not enough to distinguish between a set of models ranging in size from a small constant number of bits, to nearly the size of the data. Only if repeated samples are available can we pin down an intrinsic model for the data. One consequence of these results is that if we use MDL with increasingly general model classes, our ability to perform accurate model selection will diminish.}

\section{Introduction}

* Definitie consistent two-part model selection (protocol) \\
* Eisen aan two-part coding \\
** Invariante modelrepresentatie
**\\ 
* De boundary is het beste dat je kunt identificeren, als je definities goed zijn, dwz voor elk model kunnen we meten hoeveel counterevidence de data geeft, en die meting is niet heel anders met een andere UTM. Modellen met kleine counterevidence zijn de ``boundary" en die zijn dus interessant als beste verklaring voor de data.\\
** Bij de definitie van Koppel (inefficiente modelbeschrijving) gaat het ernstig fout omdat modellen in de boundary van een goede methode een super inefficiente index kunnen hebben bij Koppel (en Adriaans): er is geen invariantie voor de modelcodelengte. Verwarring van de twee geneste toepassingen van two-part coding?\\
* Koppel faalt harder dan Vitanyi, hoewel het allebei ``sophistication'' heet

\section{\ldots}

\section{Sandbox}

\begin{definition}
Let $M_r: \C \rightarrow \N$ be defined as follows $M_r(\phi) = \min\left \{|i| : r(i) \simeq \phi\right \}$. That is, it returns the length of the shortest function (under $r$) computing $f$.
\end{definition}

Another tack. Complexity of a function:

\begin{definition}
Let $f \in \C$, ie. the abstract object representing a computable function. We define its complexity in bits as
\[
	K^\phi(f) = \min \left\{ |\s{i}p| : \forall_y \phi_i(p ; y) = f(y) \right\}
\] where $\{\phi_i\}$ is an effective enumeration of the computable functions.  
\end{definition}

\pb{This definition guards against inefficient indices, by using a kind of hidden three-part coding. If the basic indexing is inefficient, we can find a Turing machine with an efficient index and use $p$ on that to encode any string efficiently.}

\begin{theorem}
$K^\phi(f)$ is invariant to the choice of enumeration $\phi$, up to an additive constant. Let $\{\phi_i\}$ and $\{\psi_i\}$ represent two effective enumerations of the computable functions.
\end{theorem} 
\begin{proof}
Let $\psi_u(\s{\imath}p; y) = \phi_i(y)$. Since both enumerations are effective, this function must exist in $\{\psi_i\}$.
\begin{align*}
K^\psi(f) &= \min_{\s{j}p}\left\{|\s{j}p| : \forall \psi_j(p ; y) = f(y)\right\} \\
&\leq  \min_{p}\left\{|\s{u}p| : \forall \psi_u(p ; y) = f(y)\right\} & \text{taking $j = u$}\\
&= \min_{p}\left\{|p| : \forall \psi_u(p ; y) = f(y)\right\} + |\s{u}| \\
&= \min_{\s{i}q}\left\{|\s{i}q| : \forall \phi_i(q ; y) = f(y)\right\} + |\s{u}| \\
&= K^\phi(f) + O(1)
\end{align*}
We can reverse the proof to find the bound in the other direction.
\end{proof}

\begin{definition}
For index functions $r$ and $q$, we say that $r$ logaritheoremically dominates $q$ if $\forall f \in \C: M_r(f) \leq M_q(f) + O(\log(|x|))$.
\end{definition}

It is important to note that for some $f \in H$ computing $\phi$, $K(f)$ can differ by arbitrary amounts from $M(\phi)$. $K(f)$ works in the structure function, because it considers all $f$ in a given model class, but as a function on its own, it is an awkward representation of the complexity of a function.

\begin{theorem}
There exists a function in $\R$ which log-dominates every other element in $r$. 
\end{theorem}

\begin{theorem}
$M$ and $K$ are in the second level of the Turing hierarchy of incomputability. 
\end{theorem}

\begin{theorem}
\cite{DBLP:journals/cj/Shen99}

\end{theorem}

\bibliography{facticity}

\end{document}

### 2014-10-17 Facticity/Facticity/facticity-deliverable.tex
stuff

\documentclass{style/llncs}

\usepackage{eulervm}
\usepackage{charter}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{enumerate}
\usepackage{cancel}
\usepackage{hyperref}
\usepackage{footmisc}
\usepackage{mathrsfs}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\title{Ideal model selection is impossible}

\author{Peter Bloem and Steven de Rooij}

% Steven's comments
\newcommand{\sdr}[1]{\textcolor{blue}{\small #1\textsuperscript{[sdr]} }}
% Peter's comments 
\newcommand{\pb}[1]{\textcolor{OliveGreen}{\small #1 \textsuperscript{[pb]} }}
% Pieter's comments 
\newcommand{\pa}[1]{\textcolor{Red}{\small #1 \textsuperscript{[pa]} }}

\newcommand{\p}{\mbox{\,.}}
\newcommand{\fl}[1]{\left \lfloor #1 \right \rfloor}
\newcommand{\g}[1]{\color{gray} #1 \color{black}}
\newcommand{\B}{{\mathbb B}}
\newcommand{\Rbb}{{\mathbb R}}
\newcommand{\N}{{\mathbb N}}
\newcommand{\C}{{\mathscr C}}
\newcommand{\Co}{{\overline{\mathscr C}}}
\newcommand{\R}{{\mathscr R}}
\newcommand{\m}{{\overline{m}}}
\newcommand{\ok}{{\overline{\kappa}}}
\newcommand{\s}[1]{{\overline{#1}}}
\newcommand{\mmid}{\;\middle|\;}
\newcommand{\hide}[1]{}

\begin{document}

\maketitle

\abstract{The basic idea behind sophistication is this: firstly, a dataset has an implicit model, which is objectively superior to any other model. Secondly, the complexity of this model expresses how interesting the model is. There are many formalizations of this idea, mostly based on computability theory, forumlated in the hope that this quantity can be expressed with robustness similar to that of Kolmogorov Complexity. We review existing definitions, and place them in a general and precise framework. We then show that these approaches are doomed to fail, due to what we call a 'collapse' induced by the including of universal models. This effect has been mentioned before, but it has always been ascribed to the power of the model class, and most definition attempt to circumvent it by reducing the power of the model class to something less than Turing completeness. We show that the existence of models dominating other models is sufficient for the sophistication functions, and thus model selection in general, to be ill defined.}

\section{Introduction}

Consider the following example: a television which can be in three states. Switched off, its screen shows nothing, it's tuned to a television broadcast, or it shows white noise. Kolmogorov complexity allows us to make precise statements about their information content, which increases with each example. One fundamental criticism of the idea that Kolmogorov complexity represents an amount of information has always been that the objects with the greatest amount of information (ie. the white noise) are meaningless and uninteresting. `Real' information, in the informal sense of the word, seems to be maximized more by the television broadcast than by the noise.

This idea has led to a long search for a notion of what is variously called sophistication\cite{koppelSoph1988}, facticity\cite{adriaans2012facticity} or meaningful information. \footnotemark While there are important differences, the principle behind them is always the same. We have some two-part coding of a given data set. Any such coding which is less efficient than the Kolmogorov complexity can be discarded. We choose one of the remaining set of two part descriptions (for instance the one with the smallest model) and take the complexity of the model part. This value should represent the amount of `interestingness' in the data.

\footnotetext{There are also other approaches such as logical depth\cite{bennett1998information}, which we will not consider here.}

Intuitively, this seems to hold water. For both the random noise and the simple blank signal, there are only small, and simple models. But to compress the broadcast, with it's human faces, three-dimensional spaces, outdoor locations and scene changes, we can benefit from many highly complex models to represent such information efficiently.

Unfortunately, any such approach must fulfill two desiderata:
\begin{itemize}
  \item The length of the complete description of the data must be the same (up to a constant) as the Kolmogorov complexity. If this is not the case, we cannot see the description an optimal explanation of the data.
  \item The result must be invariant to ad-hoc choices in the construction of the function. If this is not the case, the value returned cannot be seen as an inherent property of the data, as equivalent functions return different answers.
\end{itemize}

Our results show that following this basic principle cannot satisfy these desiderata. As a corollary, we note that model selection, using the model class of all partial computable functions is not viable. Any such method will always have to choose between a range of equally plausible models, with no means of making a distinction.

\section{Review}
In this section, we review some of the common approaches to defining sophistication, and 
the ways in which they succeed and fail.

\subsection{Two perspectives}

The suggestions for defining sophistication can be divided in two approaches.

\begin{description}
\item[The UTM view] In this view, we use that fact that the Kolmogoirov complexity is defined with a two-part coding. The aim is to simply take the program that $K$ `selects' for $x$, and measure the size of its model. This view places certain constraints on the construction of the UTM, so that not any definition of Kolmogorov complexity will work. This is the approach taken by Koppel \cite{koppelSoph1988} and by Adriaans \cite{adriaans2012facticity}. 
\item[The black box view] In this view, we see the UTM as a black box, and do not attempt to `open it up'. Instead, we define a separate two part coding, and merely require that it is as efficient a description of the data as the most efficient description of the data (that is, its Kolmogorov complexity). In this view, taken for instance by Vitanyi\cite{vitanyi2006meaningful} we require that the the representation of the model is as small as the smallest effective representation of that function.
\end{description}

\subsubsection{Compact indices}
In the UTM view, we require that the index used for the model is compact:

\begin{definition}
For a given index of the computable functions $\{\phi_i\}$ we define $i^*_\phi = \min\left\{\phi_j : \phi_j \simeq \phi_i\right\}$. That is, the minimal index of a function that is equivalent to $\phi_i$. The subscript may be dropped when it is clear from context.
\end{definition}

\begin{definition}
An index of the computable functions $\{\phi_i\}$ is called \textbf{compact} if for every 
$\phi_i(\bar{x}y)$ we have a $\phi_j$ such that $\phi_i(\bar{x}y) \simeq \phi_j(y)$ and  $|j| - |i| \leq |\bar{x}| + O(1)$
\end{definition}

Note that this is a natural property of basic computational metaphors. We can always take part of the argument of one function and hardcode it into the program at the cost of its prefix-coded length, and some additional bits to add it to the argument before processing.

We can construct a non-compact enumeration by starting with a compact one, and inserting after function $i$ a series of $f(i)$ copies of $\phi_0$, where $f(i)$ is some fast growing function like $2^{i!}$.

\begin{lemma}
Let $\{\phi_i\}$ and $\{\psi_j\}$ be two compact, computable enumerations of the computable functions. For a given computable function $\phi_i \simeq \psi_j$,  we have $\mbox{abs}(|i^*| - |j^*|) \leq \max(|\overline{\imath^*}|, |\overline{\jmath^*}|) + O(1)$. That is, the difference between the sizes of the minimal models is in $O(\log(\max(i, j)))$
\end{lemma}
\begin{proof}
Let $\phi_u$ be a universal function for $\psi$, such that $\phi_u(\bar{\imath}x) \simeq \psi_i(x)$. Since $\psi$ is computable this function exists in $\phi$. We have $\psi_{j^*}(x) \simeq \phi_u(\overline{\jmath^*}x)$. By the compactness of $\phi$, we have some $h$ with $|h| \leq |u| + |\overline{\jmath^*}| + O(1)$ such that $\phi_{h}(x) \simeq \phi_u(\overline{\jmath^*}x) \simeq \psi_{j^*}(x)$.

Thus $|i^*| \leq |h| = |u| + |\overline{\jmath^*}| + O(1)$. Reversing $\phi$ and $\psi$ gives us $|j^*| \leq |h'| = |u'| + |\overline{\imath^*}| + O(1)$. Combining these, we get the desired result.
\end{proof}

This lemma tells us that so long as we restrict our indices to compact and effective, we can never gain more than a logarithmic difference between model sizes. In the limit as $i \rightarrow \infty$, this term represents a vanishing proportion.

\begin{theorem}
$K^\phi(f)$ is invariant to the choice of enumeration $\phi$, up to an additive constant. Let $\{\phi_i\}$ and $\{\psi_i\}$ represent two effective enumerations of the computable functions.
\end{theorem} 
\begin{proof}
Let $\psi_u(\s{\imath}p; y) = \phi_i(y)$. Since both enumerations are effective, this function must exist in $\{\psi_i\}$.
\begin{align*}
K^\psi(f) &= \min_{\s{j}p}\left\{|\s{j}p| : \forall \psi_j(p ; y) = f(y)\right\} \\
&\leq  \min_{p}\left\{|\s{u}p| : \forall \psi_u(p ; y) = f(y)\right\} & \text{taking $j = u$}\\
&= \min_{p}\left\{|p| : \forall \psi_u(p ; y) = f(y)\right\} + |\s{u}| \\
&= \min_{\s{i}q}\left\{|\s{i}q| : \forall \phi_i(q ; y) = f(y)\right\} + |\s{u}| \\
&= K^\phi(f) + O(1)
\end{align*}
We can reverse the proof to find the bound in the other direction.
\end{proof}

\begin{definition}
For index functions $r$ and $q$, we say that $r$ logarithmically dominates $q$ if $\forall f \in \C: M_r(f) \leq M_q(f) + O(\log(|x|))$.
\end{definition}

\subsection{The complexity of computable functions}

Whichever perspective we use, we must be sure that the representation of our models is efficient. If it is inefficient---ie. there is another effective representation, in some other `language' which expresses the same model more succinctly, then an efficient way to represent our data in with our inefficient index, would be to use a universal model for the efficient index, and select the relevant model in that representation. This means that our model would be a universal Turing machine for another language, rather than the model that describes the family to which our data belongs.

The best approach is to define a function $K(f)$\footnotemark, which is analogous to the Kolmogorov complexity. We want this function to represent the minimum number of bits in which we can specify $f$ effectively, that is, if we specify $f$ and some input $x$, a second party can compute $f(x)$. We would like to show an invariance similar to that used with Kolmogorov complexity: that the arbitrary choices used to construct $K(f)$ can only lead to a constant difference in the values it produces.

\begin{definition}[{\cite[p13]{grunwald2004shannon}}]
\[
K(f) = \min_i \{ K(i) : T_i\,\text{computes}\,f\}
\]
where $T_i$ is the $i$th Turing machine in a canonical enumeration.
\end{definition}

\footnotetext{Note that $K$ is a function from the set of computable functions to $\N$. The argument is not a representation of $f$ but the function itself.}

To show that this definition fits our expectations, we must show two things. That no effective way to express $f$ is significantly smaller than $K(f)$ and that $K(f)$ is invariant to the choices made.

\section{Issues}

The first issue is that many definitions in the UTM view, are not careful with their indices. Koppel for instance, in \cite{koppelSoph1988}, places no restrictions on the index used. In such cases, we can construct an index with a great deal of `air'. This means that if some model compresses a string best using an efficient index, the UTM can compress very efficiently by first switching to a UTM with an efficient index, and then using that to express the string (a kind of three-part coding). This means that the first part of the description (our model length) will always be constant for long enough data.

This issue is solved in the black box perspective, where the model is always described as efficiently as possible. In the UTM view, this issue can be resolved by requiring that the UTM uses an efficient index.

\subsection{Domination}

A greater issue is that of models in our model class dominating others. If our model class contains models $a$ and $b$ and $a$ dominates $b$ (ie. $\exists c \forall x :K_a(x) \leq K_b(x) + c $), then data sampled from $b$ can always be compressed equally well (up to a constant) by $c$.

If we view this from a statistical point of view, we see that if we have data sampled from $b$, we will always have models $a$ and $b$ as viable candidates, no matter how large our dataset (ie. any model selection method is inconsistent for this model class).

This issue has been identified in the literature as a `collapse' that occurs when the universal Turing machine becomes part of the model class (since it dominates all other models). To remedy it, authors usually place a small restriction on the model class, such as requiring that all models are total, or that models are finite sets. The assumption seems to be that the collapse is caused by the model class being too powerful. As we have shown, it is not the power of the model class but its structure. We can define a very low-powered model class, but as soon as we insert a Bayesian mixture ove the class, there is a collapse: a constant model size is enough to compress the data as well as the Kolmogorov complexity.

Put simply, the size of the best model representing the data cannot be made invariant to the universal Turing machine.

\begin{definition}[Kolmogorov complexity]
\[
K^\psi(x) = \min\left\{|\bar\imath p| : \psi_i(p) = x\right\}
\]
\end{definition}

\begin{definition}[Complexity of a function]
\[
K^{\phi\psi}(f) = \min\left\{|\bar\imath p| : \psi_i(p) = j, \forall_x\phi_j(x) = f(x)\right\}
\]
Where $\psi$ and $\psi$ represent acceptable numberings, and $f$ is some partial computable function.
\end{definition}

\begin{definition}[Specific Kolmogorov complexity]
\[
K^{\phi\psi}(x) = \min\left\{K^{\phi\psi}(f) + |p|: f(p) = x\right\}
\]
\end{definition}
This quantity is the same as the Kolmogorov complexity up to a constant. It also provides the value which all minimal two-part descriptions in our framework `witness', ie. it is the length of the shortest description in our two-part coding framework.

\begin{definition}[Sophistication]
\[
F^{\phi\psi}(x) = \min\left\{K^{\phi\psi}(f) : \exists_p f(p) = x, K^{\phi\psi}(f)+|p| \leq K^{\phi\psi}(x)\right\}
\]
where $\phi$ and $\psi$ represent acceptable numberings.
\end{definition}

\begin{lemma}
There exist numberings $\phi$ and $\psi$ such that $F^{\phi\psi}(x) \leq c$ for all $x$ and some constant $c$ independent of $x$. \label{lemma:constant}
\end{lemma}
\begin{proof}
Let $\psi$ be some acceptable numbering. We then define $\phi$ as
\begin{align*}
\phi_0(\bar\imath\jmath p) &= \psi_{\psi_i(j)}(p) \\
\phi_{11r\bar\imath}(p) &= \psi_(p)
\end{align*}
with $r$ a random string. For any $\phi_i$ where $i$ does not fit either of these cases, we define $\phi_i(x) = \infty$.

Let $x$ be any binary string. Let $\psi_i$ be some function with minimal $i$ such that $\psi_i(p) = x$. This gives us two representations in $\phi$: $\phi_0(\bar\imath j p)$ and $\phi_{ri}(p)$. These have the following description lengths:

\begin{description}
\item[$\phi_0(\imath j p)$:] $K^{\phi\psi}(\phi_0) + |\overline{a}bp| = K^\psi(0) + K^\psi(i) + |p|$
\item[$\phi_{ri}(p)$:] $K^{\phi\psi}(\phi_{ri}) + |p| = K^{\psi}(ri) + |p|$
\end{description}

The difference between the two is $K^{\psi}(11ri) - K^\psi(0) - K^\psi(i)$. For some $r$ large enough this will always be larger than $0$, which means that the two part coding using $\phi_0$ will always provide a smaller description. 
This means that $F^{\psi\phi}(x)$ will be $K^{\phi\psi}(\phi_0)$ for any $x$.
\end{proof}

\begin{lemma}
There exist a $\phi$ and $\psi$ such that for all $x$ with $K^\psi(x) > |x|$ (with $c$ some constant), we have $F^{\phi\psi}(x) > |x| - O(1)$. \label{lemma:unbounded}
\end{lemma}
\begin{proof}
Let $\psi$ be some acceptable numbering. We then define $\phi$ as
\begin{align*}
\phi_0(\bar\imath\jmath p) &= \psi_{\psi_i(j)}(p) \\
\phi_{11ri}(p) &= \psi_i(p)
\end{align*}

with $r$ a random string. For any $\phi_i$ where $i$ does not fit either of these cases, we define $\phi_i(x) = \infty$.

Let $x$ be a string with $K^\psi(x) > |x|$ as in the statement of the lemma. Let $\psi_i(p)$ be a smallest description under $\psi$. We have two descriptions in $\phi$:

\begin{description}
\item[$\phi_{0\bar\imath p}(\epsilon)$:] $K^{\phi\psi}(\phi_{0\bar\imath p}) K^{\phi}(0\bar\imath p) \leq |0\bar\imath p| + c$
\item[$\phi_{11ri}(p)$] $K^{\phi\psi}(11ri) + |p| = $  
\end{description}

Again, if $r$ is large enough, the first description will be shorter. Since $K^\psi(x) > |x|$, so isthe length of the model: $|0\bar\imath p | > |x|$, which means that the sophistication $F^{\psi\phi}$ grows linearly with $|x|$.

\end{proof}
\begin{theorem}
There exist pairs of acceptable numbers $\phi$, $\psi$ and $\eta$, $\mu$ such that for any constant $c$ there exist $x$ for which $\left|F^{\psi\phi}(x) -F^{\phi\psi}(x)\right| > c$ 
\end{theorem}
\begin{proof}
This follows directly from Lemmas~\ref{lemma:constant} and \ref{lemma:unbounded}.
\end{proof}

\section{Solutions}

\subsection{Multiple samples}

Model selection becomes possible again, as soon as we are allowed to take multiple, independent samples from the same source. If the course is, for instance, the universal Turing machine, we will see a variety of datasets, whereas a specific model, will produce multiple sets with the same structure.

\subsection{The boundary}

Another solution is to look at what structure we \emph{can} define that \emph{is} invariant to the choice of UTM. Each UTM provides a partial ordering on two part codings, in terms of the total efficiency of the description. The minimal elements of this semilattice are the candidates for the minimal sufficient statistic. We have shown that this set itself is not invariant, but it is likely that, for instance, a measure on the space of two part codings can be defined which is invariant.

If such an object can be found, we may be able to determine a funcion on it (to determine its variance, perhaps) which will suit our desiderata for the sophistication function.

\section{Conclusion}

We have shown that given a dataset, no matter how large, we cannot perform model selection in the limit. That is, if we consider all computable models as potential sources of the data, we will always end up with a set of models that the data does not allow us to distinguish between. There may be small differences in total compression, but these are not invariant to our choice of universal Turing machine. If make small changes to arbitrary choices in our construction, the outcome of our model selection algorithm can change radically (from a small  constant model, to a model which is the same size as the data).

The property that creates this ambiguity is the presence of models that dominate others in the model class. While we have phrased our results in terms of model selection in the limit, even simple model classes encounter this problem. For instance, taking a typical MDL model class like finite order Markov chains, and adding to it a Bayesian mixture over all these models, with some suitable prior, creates a model class where one model dominates all others. This means that we can never distinguish between the Bayesian mixture or an individual markov chain. Which of the two provides the more efficient explanation is an artifact of arbitrary choices in our framework of represenation (ie. our choice of universal Turing machine).

In the limit of all computable models, this effect becomes so strong that for any given integer $s$ between 0 and the length of the data, we can find a model whose size is a constant from $s$ and whose compression beats all other models in the $c$-boundary for a given choice of $U$.

With this type of reasoning it is of course not impossible that a work-around can be established in some unforeseen way, but it seems unlikely.

While the idea of a measure of Sophistication based on model selection seems to be impossible, there is still the intuition that there should be an objective means to distinguish between simple data (the blank screen), random data (the white noise) and interesting data (the television broadcast). We have shown that looking at the complexity of the inherent model of the data is a fruitless approach, since such a model does not exist. However the boundary \emph{can} be taken as an inherent property of the data, and an investigation of the properties of this boundary may allow us to define more robust notions of data ``interestingness''.

In broad terms, one may not be able to say definitively whether something is a painting rather than an object, or an impressionist painitng rather than a painting, or a Monet rather than an impressionist painting, but this chain of increasingly specific models \emph{does} tell us something about the data, and that may be enough to distinguish our random noise from a television broadcast.

\subsubsection{Acknowledgements} This publication was supported by the Dutch national program COMMIT.

\hide{
\section{Sandbox}

* Definitie consistent two-part model selection (protocol) \\
* Eisen aan two-part coding \\
** Invariante modelrepresentatie
**\\ 
* De boundary is het beste dat je kunt identificeren, als je definities goed zijn, dwz voor elk model kunnen we meten hoeveel counterevidence de data geeft, en die meting is niet heel anders met een andere UTM. Modellen met kleine counterevidence zijn de ``boundary" en die zijn dus interessant als beste verklaring voor de data.\\
** Bij de definitie van Koppel (inefficiente modelbeschrijving) gaat het ernstig fout omdat modellen in de boundary van een goede methode een super inefficiente index kunnen hebben bij Koppel (en Adriaans): er is geen invariantie voor de modelcodelengte. Verwarring van de twee geneste toepassingen van two-part coding?\\
* Koppel faalt harder dan Vitanyi, hoewel het allebei ``sophistication'' heet

\pb{Losse dingen, probeersels, oude dingen alles wat weg moest maar niet per se helemaal weg.}

\begin{definition}
Let $M_r: \C \rightarrow \N$ be defined as follows $M_r(\phi) = \min\left \{|i| : r(i) \simeq \phi\right \}$. That is, it returns the length of the shortest function (under $r$) computing $f$.
\end{definition}

Another tack. Complexity of a function:

\begin{definition}
Let $f \in \C$, ie. the abstract object representing a computable function. We define its complexity in bits as
\[
	K^\phi(f) = \min \left\{ |\s{i}p| : \forall_y \phi_i(p ; y) = f(y) \right\}
\] where $\{\phi_i\}$ is an effective enumeration of the computable functions.  
\end{definition}

\pb{This definition guards against inefficient indices, by using a kind of hidden three-part coding. If the basic indexing is inefficient, we can find a Turing machine with an efficient index and use $p$ on that to encode any string efficiently.}

It is important to note that for some $f \in H$ computing $\phi$, $K(f)$ can differ by arbitrary amounts from $M(\phi)$. $K(f)$ works in the structure function, because it considers all $f$ in a given model class, but as a function on its own, it is an awkward representation of the complexity of a function.

\begin{theorem}
There exists a function in $\R$ which log-dominates every other element in $r$. 
\end{theorem}

\begin{theorem}
$M$ and $K$ are in the second level of the Turing hierarchy of incomputability. 
\end{theorem}

\begin{theorem}
\cite{DBLP:journals/cj/Shen99}

\end{theorem}
}

\bibliographystyle{style/splncs}
\bibliography{facticity}

\end{document}

### 2014-10-17 Facticity/Facticity/facticity.tex
stuff

@@
 \documentclass{style/llncs}
+
+\usepackage{eulervm}
+\usepackage{charter}
 
@@
 \title{Ideal model selection is impossible}
+
+\author{Peter Bloem and Steven de Rooij}
 
@@
 \newcommand{\mmid}{\;\middle|\;}
+\newcommand{\hide}[1]{}
 
@@
 
-\abstract{The idea behind Sophistication is that a dataset can be expressed in a single, canonical two-part coding: a model, and a code given that model.
-When the best compressing model is chosen over all computable models, the result is a representation that can be seen as an inherent property of the data.
-The data belongs to a single model.
-This can be seen as a kind of MDL model selection in the limit.
-We show that there are serious issues with this notion, and that a single sample from a probabilistic data source is not enough to distinguish between a set of models ranging in size from a small constant number of bits, to nearly the size of the data.
-Only if repeated samples are available can we pin down an intrinsic model for the data.
-One consequence of these results is that if we use MDL with increasingly general model classes, our ability to perform accurate model selection will diminish.}
-
-\section{Introduction}
-
-* Definitie consistent two-part model selection (protocol) \\
-* Eisen aan two-part coding \\
-** Invariante modelrepresentatie
-**\\ 
-* De boundary is het beste dat je kunt identificeren, als je definities goed zijn, dwz voor elk model kunnen we meten hoeveel counterevidence de data geeft, en die meting is niet heel anders met een andere UTM.
-Modellen met kleine counterevidence zijn de ``boundary" en die zijn dus interessant als beste verklaring voor de data.\\
-** Bij de definitie van Koppel (inefficiente modelbeschrijving) gaat het ernstig fout omdat modellen in de boundary van een goede methode een super inefficiente index kunnen hebben bij Koppel (en Adriaans): er is geen invariantie voor de modelcodelengte.
-Verwarring van de twee geneste toepassingen van two-part coding?\\
-* Koppel faalt harder dan Vitanyi, hoewel het allebei ``sophistication'' heet
-
-\section{\ldots}
-
-\section{Sandbox}
-
-\begin{definition}
-Let $M_r: \C \rightarrow \N$ be defined as follows $M_r(\phi) = \min\left \{|i| : r(i) \simeq \phi\right \}$. That is, it returns the length of the shortest function (under $r$) computing $f$.
+\begin{definition}[Kolmogorov complexity]
+\[
+K^\psi(x) = \min\left\{|\bar\imath p| : \psi_i(p) = x\right\}
+\]
 \end{definition}
 
-Another tack. Complexity of a function:
-
-\begin{definition}
-Let $f \in \C$, ie. the abstract object representing a computable function. We define its complexity in bits as
+\begin{definition}[Complexity of a function]
 \[
-	K^\phi(f) = \min \left\{ |\s{i}p| : \forall_y \phi_i(p ; y) = f(y) \right\}
-\] where $\{\phi_i\}$ is an effective enumeration of the computable functions.  
+K^{\phi\psi}(f) = \min\left\{|\bar\imath p| : \psi_i(p) = j, \forall_x\phi_j(x) = f(x)\right\}
+\]
+Where $\psi$ and $\psi$ represent acceptable numberings, and $f$ is some partial computable function.
 \end{definition}
 
-\pb{This definition guards against inefficient indices, by using a kind of hidden three-part coding.
-If the basic indexing is inefficient, we can find a Turing machine with an efficient index and use $p$ on that to encode any string efficiently.}
+\begin{definition}[Specific Kolmogorov complexity]
+\[
+K^{\phi\psi}(x) = \min\left\{K^{\phi\psi}(f) + |p|: f(p) = x\right\}
+\]
+\end{definition}
+This quantity is the same as the Kolmogorov complexity up to a constant.
+It also provides the value which all minimal two-part descriptions in our framework `witness', ie.
+it is the length of the shortest description in our two-part coding framework.
 
+\begin{definition}[Sophistication]
+\[
+F^{\phi\psi}(x) = \min\left\{K^{\phi\psi}(f) : \exists_p f(p) = x, K^{\phi\psi}(f)+|p| \leq K^{\phi\psi}(x)\right\}
+\]
+where $\phi$ and $\psi$ represent acceptable numberings.
+\end{definition}
+
+\begin{lemma}
+There exist numberings $\phi$ and $\psi$ such that $F^{\phi\psi}(x) \leq c$ for all $x$ and some constant $c$ independent of $x$. \label{lemma:constant}
+\end{lemma}
+\begin{proof}
+Let $\psi$ be some acceptable numbering. We then define $\phi$ as
+\begin{align*}
+\phi_0(\bar\imath\jmath p) &= \psi_{\psi_i(j)}(p) \\
+\phi_{11r\bar\imath}(p) &= \psi_(p)
+\end{align*}
+with $r$ a random string. For any $\phi_i$ where $i$ does not fit either of these cases, we define $\phi_i(x) = \infty$.
+
+Let $x$ be any binary string.
+Let $\psi_i$ be some function with minimal $i$ such that $\psi_i(p) = x$.
+This gives us two representations in $\phi$: $\phi_0(\bar\imath j p)$ and $\phi_{ri}(p)$.
+These have the following description lengths:
+
+\begin{description}
+\item[$\phi_0(\imath j p)$:] $K^{\phi\psi}(\phi_0) + |\overline{a}bp| = K^\psi(0) + K^\psi(i) + |p|$
+\item[$\phi_{ri}(p)$:] $K^{\phi\psi}(\phi_{ri}) + |p| = K^{\psi}(ri) + |p|$
+\end{description}
+
+The difference between the two is $K^{\psi}(11ri) - K^\psi(0) - K^\psi(i)$.
+For some $r$ large enough this will always be larger than $0$, which means that the two part coding using $\phi_0$ will always provide a smaller description.
+This means that $F^{\psi\phi}(x)$ will be $K^{\phi\psi}(\phi_0)$ for any $x$.
+\end{proof}
+
+\begin{lemma}
+There exist a $\phi$ and $\psi$ such that for all $x$ with $K^\psi(x) > |x|$ (with $c$ some constant), we have $F^{\phi\psi}(x) > |x| - O(1)$. \label{lemma:unbounded}
+\end{lemma}
+\begin{proof}
+Let $\psi$ be some acceptable numbering. We then define $\phi$ as
+\begin{align*}
+\phi_0(\bar\imath\jmath p) &= \psi_{\psi_i(j)}(p) \\
+\phi_{11ri}(p) &= \psi_i(p)
+\end{align*}
+
+with $r$ a random string. For any $\phi_i$ where $i$ does not fit either of these cases, we define $\phi_i(x) = \infty$.
+
+Let $x$ be a string with $K^\psi(x) > |x|$ as in the statement of the lemma. Let $\psi_i(p)$ be a smallest description under $\psi$. We have two descriptions in $\phi$:
+
+\begin{description}
+\item[$\phi_{0\bar\imath p}(\epsilon)$:] $K^{\phi\psi}(\phi_{0\bar\imath p}) K^{\phi}(0\bar\imath p) \leq |0\bar\imath p| + c$
+\item[$\phi_{11ri}(p)$] $K^{\phi\psi}(11ri) + |p| = $  
+\end{description}
+
+Again, if $r$ is large enough, the first description will be shorter.
+Since $K^\psi(x) > |x|$, so isthe length of the model: $|0\bar\imath p | > |x|$, which means that the sophistication $F^{\psi\phi}$ grows linearly with $|x|$.
+
+\end{proof}
 \begin{theorem}
-$K^\phi(f)$ is invariant to the choice of enumeration $\phi$, up to an additive constant. Let $\{\phi_i\}$ and $\{\psi_i\}$ represent two effective enumerations of the computable functions.
+There exist pairs of acceptable numbers $\phi$, $\psi$ and $\eta$, $\mu$ such that for any constant $c$ there exist $x$ for which $\left|F^{\psi\phi}(x) -F^{\phi\psi}(x)\right| > c$ 
 \end{theorem}
 \begin{proof}
-Let $\psi_u(\s{\imath}p; y) = \phi_i(y)$. Since both enumerations are effective, this function must exist in $\{\psi_i\}$.
-\begin{align*}
-K^\psi(f) &= \min_{\s{j}p}\left\{|\s{j}p| : \forall \psi_j(p ; y) = f(y)\right\} \\
-&\leq  \min_{p}\left\{|\s{u}p| : \forall \psi_u(p ; y) = f(y)\right\} & \text{taking $j = u$}\\
-&= \min_{p}\left\{|p| : \forall \psi_u(p ; y) = f(y)\right\} + |\s{u}| \\
-&= \min_{\s{i}q}\left\{|\s{i}q| : \forall \phi_i(q ; y) = f(y)\right\} + |\s{u}| \\
-&= K^\phi(f) + O(1)
-\end{align*}
-We can reverse the proof to find the bound in the other direction.
+This follows directly from Lemmas~\ref{lemma:constant} and \ref{lemma:unbounded}.
 \end{proof}
 
-\begin{definition}
-For index functions $r$ and $q$, we say that $r$ logaritheoremically dominates $q$ if $\forall f \in \C: M_r(f) \leq M_q(f) + O(\log(|x|))$.
-\end{definition}
-
-It is important to note that for some $f \in H$ computing $\phi$, $K(f)$ can differ by arbitrary amounts from $M(\phi)$.
-$K(f)$ works in the structure function, because it considers all $f$ in a given model class, but as a function on its own, it is an awkward representation of the complexity of a function.
-
-\begin{theorem}
-There exists a function in $\R$ which log-dominates every other element in $r$. 
-\end{theorem}
-
-\begin{theorem}
-$M$ and $K$ are in the second level of the Turing hierarchy of incomputability. 
-\end{theorem}
-
-\begin{theorem}
-\cite{DBLP:journals/cj/Shen99}
-
-\end{theorem}
-
+\bibliographystyle{style/splncs}
 \bibliography{facticity}

### 2014-10-21 Facticity/Facticity/paper.tex
beginnetje paper

\documentclass{article}
\usepackage{amsmath,amsfonts,amsthm,fullpage}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}

\newcommand{\M}{\mathcal M}
\newcommand{\K}{\mathcal K}
\newcommand{\X}{\mathcal X}
\newcommand{\C}{\mathcal C}
\newcommand{\D}{\Delta}
\newcommand{\N}{\mathbb N}
\newcommand{\tn}[1]{\textnormal{#1}}
\newcommand{\pair}[1]{\left\langle{#1}\right\rangle}
\newcommand{\concat}{\oplus}
\newcommand{\symb}[1]{\texttt{#1}}

\title{Ideal model selection is impossible}
\author{Peter Bloem and Steven de Rooij}

\begin{document}
\maketitle

\begin{abstract}
Kolmogorov complexity provides an objective quantification of the amount of information in an individual binary string, i.e. the number of bits required to effectively describe it, without making any assumptions about the source of that string. However, not all information is created equal -- over the years, there have been many attempts that use two-part codes in order to separate the amount of ``meaningful information'', ``sophistication'' or ``facticity'' of the data, from information representing its random properties. How wonderful would it be if we could quantify not just the syntactic information content of the data  not only syntactically, but the meaningful information as well!

In this paper we show this to be an impossible dream: in general, additional assumptions are required if one wishes to find the true model, or its complexity. This is true regardless of the length or Kolmogorov complexity of the input data. In effect, this means that ideal model selection based on a single data object with no further knowledge is impossible. We unify and contrast the several existing approaches and describe what they can and cannot do. 
\end{abstract}

\section{Notes}

Do we have the citation ``Paul M. B. Vit\'anyi: Meaningful Information. IEEE Transactions on Information Theory 52(10): 4617-4626 (2006)''?

(From Koppel, Complexity, Depth and Sophistication:)

``Sophistication is a generalization of the "H-function" or "minimal sufficient statistic" of Cover and Kolmogoroff, using the motonic complexity of Levin: is Cover the original source for the idea of a ``minimal sufficient statistic''? Also Koppel does indeed use monotonic complexity as in Levin. We need to know at least a little about the difference between monotonic and regular KC: is it less than a logarithmic additive term? If so, we can conveniently sweep it under the rug.

\section{Case against index length as model complexity}
All definitions of meaningful information are based on a two-part description of the data: the first part describes a \emph{model} for the data, and the second part describes an input to the model that makes it reconstruct $x$ exactly. However there is a fundamental dichotomy in way the various authors work out this idea technically.

The aproach taken in the works of Kolmogorov, Vit\'anyi, (Cover? Shen? etc) is sound (apart from the issues discussed elsewhere in this paper), but the other approach, adopted by Koppel, Adriaans, (others), while very intuitive, has a fundamental problem that seems to originate in a misunderstanding of the role Kolmorov complexity has to play in the two part code.

The misunderstanding, which appears to be shared by authors on both sides of the dichotomy, is that the two components of the Kolmogorov complexity (the index of a Turing machine and its input) \emph{are} the two-part code used to separate the meaningful information from the noise. This idea certainly seems reasonable, as it is expressed not only by Koppel, who writes

[insert quote here: In his 1988 chapter ``Structure'', Koppel has a very good quote about how it is such a pity that program and data are not considered separately anymore.]

Perhaps surprisingly, it is also reinforced by Vit\'anyi, when he writes : (From Vereshchagin and Vit\'anyi, ``Kolmogorov’s Structure Functions and Model Selection'':)
\begin{quotation}
  This expression emphasizes the two-part code nature of Kolmogorov complexity. In the example
\[x=10101010101010101010101010\]
we can encode $x$ by a small Turing machine printing a specified number of copies of the pattern `10' which computes $x$ from the program `13'.” This way, $K(x)$ is viewed as the shortest length of a two-part code for $x$, one part describing a Turing machine, or model, for the regular aspects of $x$, and the second part describing the irregular aspects of $x$ in the form of a program to be interpreted by $T$. The regular, or “valuable,” information in  is constituted by the bits in the “model” while the random or ``useless'' information of $x$ constitutes the remainder.
\end{quotation}

The problem of this two-part code lies with the first part: the index of the Turing machine depends on the used universal Turing machine, and the same Turing machine may well have, say, an exponentially higher index on one UTM than on another. (This is sometimes referred to as the ``nickname problem'', \cite{TODO}.) As such the index is \emph{not} a good measure of the model complexity.

To illustrate what can go wrong, consider all TM's $\{T_{i_1}, \ldots, T_{i_k}\}$  that are $c$-sufficient statistics for $x$, i.e. they have $\overline{i_j}+\min\{|y|:T_{i_j}(y)=x\}\le K(x)+c$. If we were to switch to a different universal Turing machine that uses a different index function, any or all of these Turing machines may suddenly yield a two-part codelength that is much larger than the Kolmogorov complexity -- and the potential difference in two-part codelength is not bounded by a constant.

Adriaans takes this problem seriously and attempts to address it by imposing additional restrictions on the ordering of TMs implied by the used UTM \cite{TODO}. We are skeptical that such measures can effectively eliminate this problem without resorting to orderings of the TMs that are not admissible [TODO ref].

But there is already a simple way to avoid this pitfall, which uses Kolmogorov complexity as a \emph{building block} of the two-part code: the first part of the code identifies a model, using its shortest possible description, whose length is measured by Kolmogorov complexity. The second part of the code provides an input to the model for which it outputs $x$.

The model may be a Turing machine, but can also be used with other model classes; for example in Kolmogorov's original formulation the model is a finite set of objects that contains $x$. Since the models are not usually represented as binary strings, the definition of Kolmogorov complexity has to be suitably extended to the considered model classes, but this is not a fundamental problem, [refer to appropriate definitions].

\section{Thoughts and questions}

\begin{itemize}

\item Kolmogorov complexity measures the shortest effective description of a binary sequence $x$. As such it is an \emph{objective} property of $x$ in the sense that changing around the enumeration of Turing machines in its definition only changes the Kolmogorov complexity by a constant independent of $x$.

The same should hold for all definitions:
\begin{itemize}
\item The complexity of any model should not change by more than a constant when the ordering of TMs is changed.
\item The same holds for the measure of meaningful information/sophistication
\end{itemize}

\item Iets uitgebreidere versie van bovenstaande, uit mijn mailtje: definitie 1: twee frameworks met complexiteitsmaten $K_1$ en $K_2$ en modelklassen $\M_1$ en $\M_2$ zijn equivalent als voor alle representaties $r$ in $\M_1$ er een $s$ in $\M_2$ is zodanig dat 
\begin{itemize}
\item $K_1(r)-K_2(s) = O(\log(K_1(r)+K_2(s)))$
\item $K_1(r)+L_r(x) - (K_2(s)+L_s(x)) = O(1)$
\end{itemize}

\item We waren bezig met operationaliseren wanneer een methode nu eigenlijk de juiste hypothese identificeert, met protocollen en asymptotisch gedrag enzo. Maar kunnen we het niet veel eenvoudiger houden? Ik ben er erg voor om het technisch zo simpel mogelijk te houden! Elke redelijke two-part code voor modelselectie moet onder wijziging van de enumeratie van TMs aan de volgende eigenschap voldoen:
(a) de twee versies moeten equivalent zijn onder definitie 1, want anders is de maat namelijk arbitrair. Dit geldt voor zover ik begrijp niet voor de Koppel/Adriaans/Antunez versie van sophistication, maar wel voor die van Paul. Hiermee kunnen we dus leuk het eerste probleem aanduiden. 
Deze eigenschap zegt iets over de two part codelengtes voor de hele model class. Maar als we bovendien willen aannemen dat de lengte van het geselecteerde model iets zegt over de data,
(b) dan moet definitie 1 bovendien gelden voor de specifieke r en s die worden geselecteerd! Hiervan kunnen we bewijzen dat het onmogelijk is.

\item What is an ``acceptable enumeration'', is it really required in the standard definition of KC or is there a difference of opinion? Does this mean that it is impossible to do facticity in the standard definition of KC? That would be a nice claim to make!

\item Is it possible to require that the enumeration of TM's is such that any TM can also efficiently store input data to the TM? That is, is there an acceptable numbering such that for all $i$, all $p$, there is a $j$ with $|j|\le|i|+|p|$ such that $T_j(\epsilon)=T_i(p)$. This enumeration is potentially still very simple, the programming language just has a provision for storing data inside the program. If there are acceptable enumerations with this property, and I suspect there are, then perhaps it could be reasonable to require this, and use it as a motivation to forget about the distinction between the index of a program and its input. Perhaps some authors silently assume the enumeration to have this property? This would make it valid to do what some authors, like Paul, are doing: to identify the two-part code \emph{within} the definition of KC, with the two-part code used to define meaningful information. Or am I now giving these authors too much credit?

\item Why the ``up to log-terms''? Is it because of algorithmic mutual information holds up to log terms? Can we find references motivating the decision to work up to log terms? Is it possible to define a two part code where the length of both parts varies by only a constant as the ordering of Turing machines is changed? Why (not)?

\item Paul argues that the randomness deficiency of a model is a measure of the amount of structure left in the noise part. I feel that it is also a measure of the amount of counterevidence against a model in the statistical sense, as follows: the randomness deficiency of $x$ given model $M$ is $\delta(x|M)=L_M(x)-K(x|M)$. Suppose that model $M$ is true. We can then use the no hypercompression inequality to rewrite:
%
\[
P_{x\sim M}(\delta(x|M)\ge c) = P_{x\sim M}(L_M(x) - K(x|M) \ge c)\le 2^{-c}
\]
Thus, if $M$ is true the probability of a large randomness deficiency is exponentially small! This amount of evidence is robust in the sense that changing the UTM around will change the amount of evidence by at most a constant. If the structure function is going to give us a \emph{cloud} of candidate models instead of just a single one, I think this is a really nice interpretation of the meaning of that cloud: we simply measure the available evidence against all individual models.

\item Shen explains the collapse of the structure function if you use enumerating complexity rather than listing complexity. How can we most cleanly explain that this collapse does not vanish for weaker model classes?

\item Properties of representations on $K(M)$ vs $L_M(x)$ graph: diagonals identify representations with the same two-part codelength. The randomness deficiency is the distance to the minimum, which is the $K(x)$ diagonal:
\[L(x;M)=K(M)+L_M(x) =K(M)+K(x|M)+L_M(x)-K(x|M)  = K(x)+\delta(x|M).\]
 Models move horizontally by a constant when the ordering of TMs is changed. Thus the randomness deficiency and two part code are also changed by at most a constant. In contrast the \emph{minimum} can now be achieved by a very different representation.

\item Discuss worries about overfitting in the literature, explain that actually the collapse problem is a case of severe \emph{underfitting} and that both can occur. Any attempts to penalize the model complexity in the two part code will exacerbate the underfitting problem.

\item Ik denk dat we in ons eigen paper minder hoog van de toren moeten blazen dan in het stukje dat ik je laatst stuurde, misschien meer zoiets zeggen als "er is een verschil tussen het kortste programma dat een string produceert, die mogelijk veel langer is dan de string zelf als data niet efficient kan worden gerepresenteerd in de gebruikte programmeertaal, en de korste beschrijving van de string, die mag bestaan uit een programma plus de input voor dat programma. Dit verschil is cruciaal voor een juist begrip van de verschillende aanpakken voor meaningful information, maar in veel publicaties over het onderwerp wordt dit onderscheid niet expliciet gemaakt en is het dus niet duidelijk of de besproken complexiteitsmaten nu wel of niet afhangen van de gekozen universele Turingmachine."

\item Paul-Pieter: model $M$ is eindige set, data $D$ is een subset ervan, dus meerdere binaire strings. Two part code is $K(M|d)+\log\binom{m}{d}$, met $d=|D|$ en $m=|M|$. Vraag, waarom zit $d$ in de conditional? Heeft dit te maken met andere keuzes over waar je de hoeveelheid noise stopt?

\item ``Power and peril of MDL'': bespreken ook de ``collapse of the structure function'' als de model class te sterk is, dwz sets worden te efficient gecodeerd (wsch equivalent met Shen's enumerating complexity).

\item Conjecture: to define valid two-part code inside KC, need efficient numbering: data can be inserted into TM index efficiently. In general, KC with admissible numbering will not work.

\item Idea: two distinct models are impossible to separate if the probability they give to the data remains within a constant. This can happen even if they do not dominate each other for other data sequences. Sufficient condition: one dominates the other.
\end{itemize}

\begin{definition} An enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions is called \emph{$\gamma$-compact} if for all $i,y$ there is a $j$ such that $|j|\le|i y|+\gamma(|y|)$ and $\phi_i(y z)=\phi_j(z)$ for all $z$.
\end{definition}

\begin{lemma}
There is a $0$-compact acceptable numbering.
\end{lemma}
\begin{proof}
  Let $\phi_1,\phi_2,\ldots$ be any acceptable numbering. Let $\psi_1,\psi_2,\ldots$ be a new acceptable numbering defined by $\psi_j(x)=\phi_i(y x)$ if $j$ is a pair satisfying $j=\bar\imath y$, and $\psi_j(x)=\infty$ otherwise.  We will show that $\psi$ is both $0$-compact and acceptable.

To show compactness, pick any index $i$ and binary string $y$. First assume $i$ is a valid pair $i=\bar a b$. Then  $\psi_i(y x) = \psi_{\bar a b}(y x)=\phi_a(b y x)$, but also $\psi_{\bar a b y}(x)=\phi_a(b y x)$, so $\psi_i(y x) = \psi_j(x)$ for $j=\bar a b y$; so $|j|=|iy|+0$. On the other hand, if $i$ is not a valid pair, we can simply use $j=i$ so that $\psi_i(y x)=\psi_j(x)=\infty$ for all $x$.

To show that $\psi_1,\psi_2,\ldots$ is acceptable, we define total computable functions $f$ and $g$ such that $\psi_i(x)=\phi_{f(i)}(x)$ and $\phi_j(x)=\psi_{g(j)}(x)$. The easy direction is $g(j)=\bar\jmath\epsilon$. The definition of the mapping in the other direction $f(i)$ depends on whether $i$ is a valid pair. If it is not, then $f(i)=h$ for some $h$ such that $\phi_h(x)=\infty$ for all $x$. On the other hand if $i=\bar\jmath y$, then $f(i)$ is the index of a partial recursive function $\phi_{f(i)}(x)$ defined by a Turing machine that simulates the machine for $\phi_j$ but with $yx$ on the input tape.
\end{proof}

\begin{lemma}[Framework equivalence]
Let $\phi_1,\phi_2,\ldots$ be a $\gamma$-compact effective enumeration. There is a constant $c$ such that for any index $i$ there is another index $j$ with $\phi_j(x)=\phi_i(x)$ for all $x$, and
\[|j|\le K(i)+2\log(K(i))+\gamma(K(i)+2\log K(i)).\]
\end{lemma}
\begin{proof}
First, let $v$ be the index of the partial recursive function given by $\phi_v(\bar k\bar p x)=\phi_{\phi_k(p)}(x)$. Let $\bar k p$ be a shortest program for $i$, that is $|\bar k p| = K(i)$ and $\phi_k(p)=i$. Then $\phi_i(x)=\phi_v(\bar k \bar p x)$. We now use $\gamma$-compactness to find an index $j$ with $\phi_v(\bar k \bar p x)=\phi_j(x)$ and length
\[\begin{split}
|j|&\le |v\bar k\bar p|+\gamma(\bar k\bar p)\\
&\le|v|+|\bar k p|+2\log|\bar k p|+\gamma(|\bar k p|+2\log|\bar k p|)\\
&=K(i)+|v|+2\log K(i)+\gamma(K(i)+2\log K(i)),
\end{split}\]
as required.
\end{proof}
\begin{corollary}
If $\gamma(x)=O(\log x)$, then $|j|=K(i)+O(\log K(i))$.
\end{corollary}

We need a definition of $K(f)$. We adopt the definition given in \cite{TODO}:

\begin{definition}
  The Kolmogorov complexity of a partial recursive function is defined as
  \[
  K(f) = \min\{K(i):\forall_x \phi_i(x)=f(x).\}
  \]
\end{definition}

It is not immediately clear that this is a reasonable definition of the shortest effective description of a function, but fortunately this is indeed the case:

\begin{lemma}
Let $\psi_1,\psi_2,\ldots$ be any acceptable numbering. There is a constant $c$ such that for any $i$ with $\psi_i(x)=f(x)$ for all $x$, the description length satisfies $K(f)\le |i|+c$.
\end{lemma}
\begin{proof}
The numbering is acceptable, so there is a total computable function $\phi_r$ that maps indices in the $\psi$-enumeration to indices in the $\phi$-enumeration. Let $j=\phi_r(i)$. Then $\bar r i$ is a program for $j$, so $K(j)\le |\bar r i| = |i|+c$. Now,
\[
K(f)=\min\{K(k):\forall_x\phi_k(x)=f(x)\}\le K(j)\le|i|+c
\]
as required.
\end{proof}

\begin{lemma}
There is an acceptable numbering $\psi_1,\psi_2,\ldots$ such that $F^\psi(x)=|\bar 1|$ for all $x$.
\end{lemma}
\begin{proof}
Let $\phi_1,\phi_2,\ldots$ be another acceptable numbering given by
\[
\phi_i(x)=\begin{cases}
u&\tn{if $i=1$}\\
\infty&\tn{if $1<i<r$}\\
\psi_{i-r+1}(x)&\tn{if $i\ge r$}.
\end{cases}\]
Here $u\ge r$ is the index of the universal element satisfying $\phi_u(\bar\jmath p)=\psi_j(p)$.
Now the two-part codelength of $x$ using representation $\phi_u$ is
\[\begin{split}
K(\phi_u)+\min\{|y|:\phi_u(y)=x\}&=\min\{K(i):\forall_x\phi_i(x)=\phi_u(x)\}+\min\{|y|:\phi_u(y)=x\}\\
&=|\bar 1|+\min\{|\bar\jmath p|:\phi_u(\bar\jmath p)=x\}\\
&=|\bar 1|+\min\{|\bar\jmath p|:\psi_j(p)=x\}\\
&=|\bar 1|+K^\psi(x).
\end{split}\]
On the other hand, the two-part codelength of $x$ using a representation $s\ne u$ is
\[\begin{split}
K(\phi_s)+\min\{|y|:\phi_s(y)=x\}&=\min\{K(i):\forall_x\phi_i(x)=\phi_s(x)\}+\min\{|y|:\phi_s(y)=x\}\\
&=c+\min\{|y|:\psi_{s-r+1}(y)=x\}\\
\end{split}\]

\end{proof}

\begin{theorem}
  There are acceptable numberings $\phi_1,\phi_2,\ldots$ and $\psi_1,\psi_2,\ldots$ such that for every constant $c$ there is a [are lots of] string $x$ with $|F^\phi(x)-F^\psi(x)|>c$ [regardless of penalties and possibly generalized to more model classes].
\end{theorem}

Conjectures:
\begin{itemize}
\item Als er een unbounded penalty zit op modelcomplexiteit dan is de facticity bounded door een constante.
\end{itemize}

% idea 0: show collapse of structure function

% idea 1: ideal model selection is impossible
% procedure selects model m1
% there is another model m2 that dominates m1
% there is a numbering such that L(m1)+L_m1(data) > L(m2)+L_m2(data), 
% in the other numbering m2 is preferred
% limitation of this approach: you have no control over the difference in model complexity,
% it may not be very large.

% idea 2: model complexity is not invariant
% there are numberings \phi,\psi such that for all c there is an x such that |C_1(x)-C_2(x)|>=c
% prove this for total recursive functions?
% to make the result stronger:
% - show that it does not go away with penalties on model complexity
% - show that it holds for *all* numberings that are sufficiently distinct in some sense
% - show that it holds for more model classes
% - show that it occurs for most strings
%
% proof idea: truth is bernoulli with a complicated but computable parameter value, K(theta) is large
% for small x, the print function will outperform the true model.
% for large x, the true model gives best overall code length.
% somewhere in between, the model complexity jumps from the one to the other. But that will not happen
% at the same time for the two numberings. When one has switched and the other has not, the difference
% in model complexity is about K(theta), which can be made arbitrarily large.
% problem: what about in-between models that approximate theta?

\end{document}

### 2014-10-21 Facticity/Facticity/paper.tex
wip

@@
 
+\begin{definition}[Facticity]
+  Let $\phi_1,\phi_2,\ldots$ be any effective enumeration of the partial recursive functions. Then
+  \[F^\phi(x)=|\bar\imath|\qquad\tn{where}\qquad \bar\imath p=\min\{\bar\jmath q:\phi_j(q)=x\}.
+  \]
+  Here the minimum is taken with respect to the standard ordering of binary sequences.
+\end{definition}
+
+\begin{definition}
+Given an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions, and an enumeration $\psi_1,\psi_2,\ldots$ of the partial recursive prefix functions.
+Let $\psi_u$ be a universal element with property $\psi_u(\bar\imath p)=\psi_i(p)$.
+Then define $v$ by
+\[
+v(a p)=\begin{cases}\phi_{\psi_u(a)}(p)&\tn{if $a$ is in the domain of $\psi_u$,}\\
+\infty&\tn{otherwise}.\end{cases}
+\]
+\end{definition}
+\begin{lemma}
+The function $v=\phi_v$ as defined above is partial recursive.
+\end{lemma}
+\begin{proof}
+Let $r$ be the input of $v$.
+Simulate $\psi_u$ using a TM with a one-way read-only input tape.
+If the TM does not halt, then enter an infinite loop.
+Otherwise, let $r=ap$ where $a$ is the part of the input tape that has been read when the TM halts, and let $i$ be the output of the TM.
+Now simulate $\phi_i$ on input $p$.
+\end{proof}
+
+\begin{lemma}
+  Facticity is bounded.
+\end{lemma}
+\begin{proof}
+Let $\bar\imath p$ be any two-part representation for the data $x$, i.e.
+$\phi_i(p)=x$.
+Then construct an alternative two-part representation $\bar vi^* p$, where $i^*$ is the shortest $\psi_u$-program for $i$ such that $\phi_v(i^* p)=\phi_{\psi_u(i^*)}(p) = \phi_i(p)=x$.
+We compare the lengths of these two representations.
+Note that $|i^*|=K^\psi(i)$.
+Therefore,
+\[
+|\bar\imath p|-|\bar v i^* p| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-K(i).
+\]
+Since $|\bar\imath|$ is a computable codelength function, there must be an $i_0$ such that $|\bar\imath|-K(i)>|v|$ for all $i>i_0$ (see~\cite[Theorem~2.3.1]{li1993introduction}).
+From this $i$ onwards, the second representation (using $v$) will have a shorter code length, so any two-part representation $\bar ip$ cannot achieve the minimum in the definition of the facticity if $i>i_0$.
+Consequently, the facticity is bounded by $F^\phi(x)<|\overline{\imath_0}|$ for all $x$.
+\end{proof}
+
 % idea 0: show collapse of structure function
@@
 
+\bibliographystyle{plain}
+\bibliography{facticity}
+
 \end{document}

### 2014-10-21 Facticity/Facticity/paper.tex
allemaal stuff joh

@@
 \newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{conjecture}[theorem]{Conjecture}
 \newtheorem{definition}{Definition}
@@
 \newcommand{\symb}[1]{\texttt{#1}}
+
+\newcommand{\argmin}{\mathop{\arg\min}}
 
@@
 
+\[
+b(i_0)=\min_{i\ge i_0}(L(i)-K(i))
+\]
+
 \begin{lemma}
-  Facticity is bounded.
+  If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then facticity is bounded.
 \end{lemma}
@@
 \]
-Since $|\bar\imath|$ is a computable codelength function, there must be an $i_0$ such that $|\bar\imath|-K(i)>|v|$ for all $i>i_0$ (see~\cite[Theorem~2.3.1]{li1993introduction}).
-From this $i$ onwards, the second representation (using $v$) will have a shorter code length, so any two-part representation $\bar ip$ cannot achieve the minimum in the definition of the facticity if $i>i_0$.
+By assumption there must be an $i_0$ such that the above expression is positive for all $i>i_0$.
+From this $i$ onwards, the second representation (using $v$) will have a shorter code length, so $\bar\imath p$ cannot achieve the minimum in the definition of the facticity.
 Consequently, the facticity is bounded by $F^\phi(x)<|\overline{\imath_0}|$ for all $x$.
 \end{proof}
+Note that the condition for this lemma applies to all known partial recursive prefix functions:
+
+\begin{conjecture}
+For any partial recursive prefix code length function $L$, the function
+\[
+\min\{L(i)-K(i):i\ge i_0\}
+\]
+is unbounded in $i_0$.
+\end{conjecture}
+
+\begin{definition}[Model complexity]
+Given a model class $\M$, the model complexity of a data sequence $x$ is given by
+\[G(x)=K(f^*)\qquad\tn{where}\qquad (f^*,p^*)=\argmin_{f,p:f\in\M, f(p)=x}(K(f)+|p|),
+\]
+where ties are broken by picking a model $f^*$ of minimal complexity.
+\end{definition}
+
+\begin{theorem}
+There are enumerations $\phi_1,\phi_2,\ldots$ and $\psi_1,\psi_2,\ldots$ of the partial recursive (prefix) functions such that for every constant $c$ there is a binary sequence $x$ with
+\[
+|G^\phi(x)-G^\psi(x)|>c
+\]
+\end{theorem}
+
+\begin{lemma}
+There is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive (prefix) functions, and a constant $c$ such that for all $n$ we have $G^\phi(x)<c$ for more than half of all strings $x$ of length $n$.
+\end{lemma}
+
+\begin{lemma}
+There is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive (prefix) functions such that for every constant $c$ there is an $n_0$ such that for all $n>n_0$ we have $G^\phi(x)>c$ for more than half of all strings $x$ of length $n$.
+\end{lemma}
+
+\begin{theorem}
+There are enumerations $\phi_1,\phi_2,\ldots$ and $\psi_1,\psi_2,\ldots$ of the partial recursive (prefix) functions such that for every constant $c$ there is a binary sequence $x$ with
+\[
+|G^\phi(x)-G^\psi(x)|>c
+\]
+\end{theorem}
 

### 2014-10-22 Facticity/Facticity/paper.tex
blub

@@
 \begin{lemma}
-There is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive (prefix) functions, and a constant $c$ such that for all $n$ we have $G^\phi(x)<c$ for more than half of all strings $x$ of length $n$.
+There is an acceptable numbering $\phi_1,\phi_2,\ldots$ of all partial recursive prefix functions, and a constant $c$ such that for all $n$ we have $G^\phi(x)<c$ for more than half of all strings $x$ of length $n$ with respect to the model class of all partial recursive prefix functions.
 \end{lemma}
@@
 \begin{lemma}
-There is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive (prefix) functions such that for every constant $c$ there is an $n_0$ such that for all $n>n_0$ we have $G^\phi(x)>c$ for more than half of all strings $x$ of length $n$.
+There is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive prefix functions such that for every constant $c$ there is an $n_0$ such that for all $n>n_0$ we have $G^\phi(x)>c$ for more than half of all strings $x$ of length $n$ provided the model class contains programs for every singleton $x$.
 \end{lemma}
@@
 \begin{theorem}
-There are enumerations $\phi_1,\phi_2,\ldots$ and $\psi_1,\psi_2,\ldots$ of the partial recursive (prefix) functions such that for every constant $c$ there is a binary sequence $x$ with
+There are enumerations $\phi_1,\phi_2,\ldots$ and $\psi_1,\psi_2,\ldots$ of the partial recursive prefix functions such that for every constant $c$ there is a binary sequence $x$ with
 \[
@@
 \]
+with respect to the model class of all partial recursive prefix functions.
 \end{theorem}
 
-% idea 0: show collapse of structure function
+\begin{definition}
+Let $\psi_1,\psi_2,\ldots$ be an arbitrary acceptable numbering of the partial recursive functions. Define
+\[
+B^n = \{ x : |x|=n~\tn{and}~n-c \le K^\psi(x) \le n+c \}
+\]
+\end{definition}
 
-% idea 1: ideal model selection is impossible
-% procedure selects model m1
-% there is another model m2 that dominates m1
-% there is a numbering such that L(m1)+L_m1(data) > L(m2)+L_m2(data), 
-% in the other numbering m2 is preferred
-% limitation of this approach: you have no control over the difference in model complexity,
-% it may not be very large.
+\begin{lemma}
+  For sufficiently large $c$ the set $B^n$ is nonempty for all $n$.
+\end{lemma}
 
-% idea 2: model complexity is not invariant
-% there are numberings \phi,\psi such that for all c there is an x such that |C_1(x)-C_2(x)|>=c
-% prove this for total recursive functions?
-% to make the result stronger:
-% - show that it does not go away with penalties on model complexity
-% - show that it holds for *all* numberings that are sufficiently distinct in some sense
-% - show that it holds for more model classes
-% - show that it occurs for most strings
-%
-% proof idea: truth is bernoulli with a complicated but computable parameter value, K(theta) is large
-% for small x, the print function will outperform the true model.
-% for large x, the true model gives best overall code length.
-% somewhere in between, the model complexity jumps from the one to the other. But that will not happen
-% at the same time for the two numberings. When one has switched and the other has not, the difference
-% in model complexity is about K(theta), which can be made arbitrarily large.
-% problem: what about in-between models that approximate theta?
+\begin{lemma}[Print function]
+  Let $\M$ be any model class that contains the identity function. There is an acceptable numbering $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that for every $n$, all $x\in B^n$, 
+we have $G^\phi(x)=1$.
+\end{lemma}
+
+\begin{lemma}[Building block]
+Given $\phi_1,\phi_2,\ldots$ for all $c$ there is $\psi_1,\psi_2,\ldots$ such that for all $f$ we have
+\[
+K^\phi(f) +c \le K^\psi(f)
+\]
+\end{lemma}
 

### 2014-10-22 Facticity/Facticity/paper.tex
dit is t denk ik

@@
 
-\begin{lemma}[Building block]
-Given $\phi_1,\phi_2,\ldots$ for all $c$ there is $\psi_1,\psi_2,\ldots$ such that for all $f$ we have
+\begin{lemma}[I think this is it]
+Let $\psi_1,\psi_2,\ldots$ be any acceptable numbering.
+For every $c$ there is another acceptable numbering $\phi_1,\phi_2,\ldots$ such that for all partial recursive functions $f$ we have
 \[
-K^\phi(f) +c \le K^\psi(f)
+K^\phi(f)-K^\psi(f) \ge c.
 \]
 \end{lemma}
+\begin{proof}
+Define $\tau$ by $\psi_\tau(\bar k p) = j$ where $j$ is given by $1^r 0 j = \psi_k(p)$.
+In other words, it is a universal element for $\psi$ that subsequently strips the prefix $1^r 0 $ from the output; this operation is clearly still partial recursive.
+Also note that this function has a fixed index -- it does not have to know $r$.
+For completeness define $\psi_\tau$ to infinity if the output is not in the correct format.
+
+Now define a new acceptable numbering $\phi_1,\phi_2,\ldots$ as follows:
+\[
+\phi_{1^r0 j}(p)=\psi_j(p)
+\]
+For any given partial recursive function $f$, we now compare $K^\phi(f)$ and $K^\psi(f)$.
+For $\phi$ it is:
+\[
+K^\phi(f) = \min\{K^\phi(i):\phi_i = f\}.
+\]
+Let $i^-$ be the index that achieves the minimum, and let $i^*$ be the shortest $\phi$-program that computes it, that is, $i^*=\bar a b$ such that $\phi_a(b)=i^-$.
+Then the codelength is $K^\phi(f)=K^\phi(i^-)=|i^*|$. Moreover, we must have $a=1^r0 k$ where $k$ is the index of the corresponding function in the $\psi$-enumeration: $\psi_k(b)=i^-$.
+
+To find an index of $f$ in the $\psi$-enumeration, note that $f=\phi_{i^-}=\phi_{1^r 0 j^-} = \psi_{j^-}$.
+We have $K^\psi(f) \le K^\psi(j^-)$; we proceed to provide a small $\psi$-program for $j^-$.
+This can be done because we have set things up so that $\psi_\tau(\bar k b)=j^-$.
+Thus, the $\psi$-program is $j^*=\bar\tau\bar k b$.
+The difference in length is
+%
+\[
+|i^*|-|j^*| = |\bar a b|-|\bar\tau\bar k b| = |\overline{1^r 0 k}| - |\bar\tau| - |\bar k| \approx r-|\bar\tau|.
+\]
+%
+(Here the $\approx$ is used because the prefix function has a slightly different effect on the two terms; however this is an unimportant detail that can be ironed out later).
+
+Thus, by choosing $r>|\bar\tau|+c$ we can guarantee that $K^\phi$ sucks by at least $c$ bits. Yay!
+\end{proof}
 

### 2014-10-22 Facticity/Facticity/paper.tex
too late...

--- 2014-10-22 Facticity/Facticity/paper.tex
@@
 \[
-K^\phi(f) +c \le K^\psi(f)
+K^\psi(f) - K^\phi(f) \le c
 \]
 \end{lemma}
+\begin{proof}
+We first note that $K(x) \leq K(yx) + K(|y|) + c$ for some constant $c$, since we can combine the shortest program to produce $yx$ and the shortest program for the length of $y$ to create a program which produces $yx$ and removes the first $|y|$ bits.
+
+Let $r$ be a random string, and define psi as $\psi_{ri}(x) = \phi_{i}(x)$, with $\psi_{j}(x) = \infty$ if $j$ cannot be read as $ri$.
+Note that this gives us $K^\psi(x) = K^\phi(x) + |r|$ for all $x$.
+
+We can now state the following:
+\[\begin{split}
+K^\phi(f) &= \min\left\{ K^\phi(i) : \phi_i(x) = f(x) \right\} \\
+       &= \min\left\{ K^\phi(i) : \psi_{ri}(x) = f(x) \right\} \\
+       &\leq \min\left\{ K^\phi(ri) + K^\phi(|r|) + c : \psi_{ri}(x) = f(x) \right\} \\
+       &= \min\left\{ K^\phi(ri) : \psi_{ri}(x) = f(x) \right\} + K^phi(|r|) + c \\
+       &= \min\left\{ K^\phi(j) : \psi_{j}(x) = f(x) \right\} + K^phi(|r|) + c \\
+       &= \min\left\{ K^\psi(j) + c' : \psi_{j}(x) = f(x) \right\} + K^phi(|r|) + c \\
+       &= \min\left\{ K^\psi(j) : \psi_{j}(x) = f(x) \right\} + K^phi(|r|) + c + |1r|\\
+       &= K^\psi(f) + K^\phi(|1r|) + c + |r|\\
+\end{split}\]
+Which gives us $K^\phi(f) - K^\psi(f) \leq c$ with some constant $d$
+\end{proof}
 

### 2014-10-22 Deliverable 13Q2/src/13q2.tex

@@
 \usepackage{hyperref}
+\usepackage{lscape}
+
+\usepackage{authordate1-4}
 
@@
 
-We have scaled up the results from the previous work package in data volume and added various complexity measures.
+\maketitle
+
+We have scaled up the results from the previous deliverable in data volume and added various complexity measures.
 We will process more datasets, and attempt to establish interesting empirical results based on compression and complexity analysis.
@@
 It validates that our methods, such as they are, can be used to provide useful empirical results.
-It will serve as a test of our software and make sure that it is maintained (preventing 'code rust').
-And finally, it will provide us with feedback for which features of our code are interesting enough to scale up in Q3.
+It will serve as a test of our software and make sure that it is maintained (preventing `code rust').
+Finally, it will provide us with feedback for which features of our code are interesting enough to scale up in quarter 3.
 
-\section{Datasets}
+Parts of this document have been taken and adapted from the deliverables of 2012 quarter 2 and quarter 3.
 
-\section{Complexity measures}
+\section*{Code}
 
-\section{Results}
+The code is written in Java.
+The JUNG graph library has been replaced with an alternative implemented for the project.
+All code is available from GitHub in the following repsitory.
+It is released under an MIT open source license.
+\begin{description}
+\item[Lilian] (\url{https://github.com/pbloem/Lilian}) Implementation of the complexity measures, including some tools for running experiments.
+This repository also contains a great deal of code unrelated to this deliverable.
+The relevant code resides mostly in the package \texttt{org.lilian.graphs}.
+\end{description}
 
-\section{Interpretation}
+\section{Data}
+
+We test these complexity measures on the following datasets:
+
+\subsection*{Technological networks}
+
+\subsection{An Erdos-Renyi random graph}
+
+A random graph consisting of 1000 nodes, with connection probability 0.5.
+
+\subsection{A preferential attachment random graph}
+
+A random graph consisting of 1000 nodes, generated by the Barabasi Albert algorithm, with 2 connection added per iteration.
+
+\subsubsection*{The internet}
+
+The internet is a classic example of a large, complex graph.
+So large, in fact that we have little hope of seeing a full snapshot of the whole network at once.
+The best we can do is to capture some part of it using traceroutes.
+We use a dataset downloaded from the personal page of Mark Newman (\url{http://www-personal.umich.edu/~mejn/netdata/}), which contains around 23 thousand nodes.
+
+\subsubsection*{Commit social web}
+
+This dataset was produced by the Data2Semantics project to map out the relations within the commit project.
+
+\subsubsection*{IAFB Dataset}
+
+This dataset was included to test our methods on RDF data.
+We take the dataset from \cite{bloehdorn2007kernel} and disregard any predicates that are not part of the domain specific ontology (http://swrc.ontoware.org/ontology).
+Named resources, blank nodes and literals are translated to vertices and predicates are translated to edges.
+
+\subsubsection*{The world wide web}
+
+The world wide web, while closely associated with the internet, is a fundamentally different type of network.
+For example, a link between two systems on the internet will bear some relation to physical proximity of the systems.
+Linking to a website that is hosted far away costs nothing.
+
+The WWW dataset is know as www-barabasi, and is a common dataset in graph analysis.
+
+\subsection*{Social networks}
+\subsubsection*{Epinions}
+
+Epinions.org is a review site where users can choose to `trust' one another.
+these trust relations are then used to determine the value of a given review.
+The total network, downloaded form \url{http://snap.stanford.edu/data/soc-Epinions1.html} contains about 76 thousand nodes.
+
+Originally used in \cite{richardson2003trust}.
+
+\subsubsection*{Pokec relationship network}
+
+Pokec is a Slovakian social networking website. This dataset describes the social relations on the site.
+
+Downloaded from \url{http://snap.stanford.edu/data/soc-pokec.html}. Citations: \cite{takacdata}
+
+\section*{Experiments}
+
+\subsubsection*{Basic complexity measures}
+
+To show the scale of dataset that we can analyse, we perform some basic graph analyses on an datasets of increasing size.
+
+\begin{tabular}{|l | r r r r r r r r |}
+  \hline                       
+  & $n$ & $l$ & $d_\mu$ & $d_\sigma$ & $W$ & $\omega$ & $\alpha$ & $t$ \\ 
+  \hline
+er  & $10000$ & $50045$ & $10.009$ & $3.14$ & $10000$ & $0.68$ & $0.0$ & $13.7 s$ \\ 
+pa  & $10000$ & $19997$ & $4.0$ & $6.09$ & $10000$ & $0.145$ & $-0.06$ & $11.2 s$ \\ 
+internet  & $22963$ & $48436$ & $4.29$ & $32.94$ & $22963$ & $0.030$ & $-0.20$ & $28.7 s$ \\ 
+commit  & $18117$ & $56031$ & $6.19$ & $74.83$ & $18117$ & $0.065$ & $-0.13$ & $43.4 s$ \\ 
+iafb  & $8318$ & $61255$ & $14.73$ & $69.87$ & $8318$ & $0.13$ & $-0.26$ & $20.6 s$ \\ 
+epinions  & $75877$ & $508837$ & $13.41$ & $52.66$ & $75877$ & $0.11$ & $-0.011$ & $15.6 s$ \\ 
+barabasi  & $325729$ & $1497135$ & $9.19$ & $48.38$ & $325729$ & $0.050$ & $-0.043$ & $30.0 s$ \\ 
+skitter  & $1696415$ & $11095298$ & $13.08$ & $136.86$ & $1694616$ & $0.11$ & $-0.081$ & $7m 36 s$ \\ 
+patents  & $6009555$ & $16518948$ & $5.5$ & $9.33$ & $3764117$ & $0.23$ & $0.17$ & $13h 27m 27s$ \\ 
+social  & $1632804$ & $30622564$ & $37.51$ & $59.52$ & $1632803$ & $0.49$ & $ 5.3628-4 $ & $33m 34s$ \\ 
+  \hline  
+\end{tabular}
+
+With the following meanings:
+\begin{description}
+\item[$n$] The number of nodes in the graph.
+\item[$l$] The number of links in the graph.
+\item[$d_\mu$] The mean degree.
+\item[$d_\sigma$] The standard deviation of the degree distribution.
+\item[$W$] The size of the largest weakly connected component.
+\item[$\omega$] The wing width ratio \cite{kang2011beyond}.
+\item[$\alpha$] The Assortativity
+\item[$t$] The total running time of all analyses run for this report. \footnote{note that for graphs with more than 50000 nodes, the AM compressor wasn't run).}
+\end{description}
+
+\subsubsection*{Slash-and-burn and spyplots}
+
+We implemented and studied the Slash-and-burn algorithm \cite{kang2011beyond}, and the spyplot method of visualization.
+This algorithm calculates an order of nodes that is likely to benefit compression (and computation) on real-world graphs.
+
+The following plots show adjacency matrices (spyplots) for the datasets mentioned above.
+
+\begin{table}[h]
+\begin{tabular}{l | c c c c}
+\hline
+ & random & given & degree & sb \\
+\hline
+er
+&
+    \includegraphics[width=0.23\textwidth]{../img/er/adjacency-matrix-random-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/er/adjacency-matrix-given-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/er/adjacency-matrix-degree-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/er/adjacency-matrix-slashburn-ordering.png} \\
+
+\hline
+pa
+&
+    \includegraphics[width=0.23\textwidth]{../img/pa/adjacency-matrix-random-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/pa/adjacency-matrix-given-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/pa/adjacency-matrix-degree-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/pa/adjacency-matrix-slashburn-ordering.png} \\
+
+\hline
+internet
+&
+    \includegraphics[width=0.23\textwidth]{../img/internet/adjacency-matrix-random-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/internet/adjacency-matrix-given-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/internet/adjacency-matrix-degree-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/internet/adjacency-matrix-slashburn-ordering.png} \\
+
+\hline
+commit
+&
+    \includegraphics[width=0.23\textwidth]{../img/commit/adjacency-matrix-random-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/commit/adjacency-matrix-given-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/commit/adjacency-matrix-degree-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/commit/adjacency-matrix-slashburn-ordering.png} \\
+
+\hline
+aifb
+&
+    \includegraphics[width=0.23\textwidth]{../img/aifb/adjacency-matrix-random-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/aifb/adjacency-matrix-given-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/aifb/adjacency-matrix-degree-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/aifb/adjacency-matrix-slashburn-ordering.png} \\
+
+\hline
+epinions
+&
+    \includegraphics[width=0.23\textwidth]{../img/epinions/adjacency-matrix-random-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/epinions/adjacency-matrix-given-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/epinions/adjacency-matrix-degree-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/epinions/adjacency-matrix-slashburn-ordering.png} \\
+
+\hline
+barabasi
+&
+    \includegraphics[width=0.23\textwidth]{../img/barabasi/adjacency-matrix-random-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/barabasi/adjacency-matrix-given-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/barabasi/adjacency-matrix-degree-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/barabasi/adjacency-matrix-slashburn-ordering.png} \\
+    \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[h]
+\begin{tabular}{l | c c c c}
+\hline
+ & random & given & degree & sb \\
+\hline
+skitter
+&
+    \includegraphics[width=0.23\textwidth]{../img/skitter/adjacency-matrix-random-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/skitter/adjacency-matrix-given-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/skitter/adjacency-matrix-degree-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/skitter/adjacency-matrix-slashburn-ordering.png} \\
+
+\hline
+patents
+&
+    \includegraphics[width=0.23\textwidth]{../img/patents/adjacency-matrix-random-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/patents/adjacency-matrix-given-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/patents/adjacency-matrix-degree-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/patents/adjacency-matrix-slashburn-ordering.png} \\
+
+\hline
+pokec
+&
+    \includegraphics[width=0.23\textwidth]{../img/pokec/adjacency-matrix-random-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/pokec/adjacency-matrix-given-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/pokec/adjacency-matrix-degree-ordering.png}
+& 
+    \includegraphics[width=0.23\textwidth]{../img/pokec/adjacency-matrix-slashburn-ordering.png} \\
+
+\hline
+
+\end{tabular}
+\end{table}
+
+\subsubsection*{Compression tests}
+
+Our ultimate goal is to study compression of graphs as a means of performing inference and complexity-analysis.\footnote{See also \cite{bloemcompression}, which was submitted as a deliverable this quarter.} To this end, we study the basic methods of representing graphs in a bitstring, and the effect that node ordsering has on the performance.
+We test the following methods
+\begin{description}
+\item[edge list (el)] The graph is stored as a list of prefix-coded pairs of integers representing the graph's edges. 
+\item[neighbor list (nl)] The graph is stored as a list of lists of indices.
+Each list represents the neighbors of a given node.
+The indices are incremented by one so that the index 0 can be used as a delimiter.
+\item[adjacency matrix (am)] We generate an adjacency matrix of $\frac{n^2 + n}{2}$ cells for an undirected graph or $n^2$ cells for a directed graph and flatten this into a bitstring.
+We zip the resulting string to exploit simple structures within it.
+We add a single bit to distinguish between directed and underected graphs so that the code is self-delimiting.
+\end{description}
+
+\begin{landscape}
+\hspace{-1cm}
+\begin{tabular}{|l | r r r | r r r | r r r | r r r |}
+\hline
+\multicolumn{1}{|c}{} & \multicolumn{3}{c}{natural} & \multicolumn{3}{c}{random} & \multicolumn{3}{c}{degree} & \multicolumn{3}{c|}{slashburn} \\
+  \hline                       
+  & el & nl & am & el & nl & am & el & nl & am & el & nl & am \\ 
+  \hline
+ er & 1538162 & 816534 & 146338 & 1537724 & 816010 & 146685 & 1494122 & 800123 & 144941 & 1495829 & 801555 & 143144 \\ 
+ pa & 558505 & 317266 & 63798 & 615444 & 332357 & 68439 & 546885 & 318000 & 61711 & 551301 & 318136 & 58245 \\
+ internet & 1329380 & 808430 & & 1594604 & 860025 & 165759 & 1190776 & 777404 & 114297 & 1224515 & 798488 &  \\
+ commit & 1609616 & 673296 &  & 1817950 & 876699 & 215356 & 1370638 & 556954 & 135640 & 1411396 & 558630 &  \\
+ aifb & 1583268 & 745875 & 70518 & 1844193 & 927728 & 155225 & 1366585 & 591342 & 69044 & 1403070 & 595136 & 69086 \\
+ epinions & 1.5760E7 & 8091701 &  & 1.9181E7 & 9670775 &  & 1.4272E7 & 7530653 &  & 1.4720E7 & 7642574 &  \\
+ barabasi & 6.3445E7 & 3.2238E7 & & 6.3769E7 & 3.2175E7 & & 5.0114E7 & 2.5458E7 & & 5.5466E7 & 2.83587E7 & \\
+ skitter & 4.84613E8 & 2.2962E8 & & 5.2691E8 & 2.6509E8 &  & 4.2595E8 & 1.8885E8 & & 4.3375E8 & 1.9100E8 &  \\
+ patents &  8.8071E8 &  4.5034E8  & & 8.4574E8 & 4.2886E8 &  & 7.6607E8 & 3.8617E8 &  & 7.6810E8 & 3.8725E8 &  \\
+ pokec &  1.3859E9 &  6.9489E8  &  & 1.4513E9 & 7.2733E8 & & 1.3180E9 & 6.6017E8 &  & 1.3200E9 & 6.6129E8 &  \\
+  \hline  
+\end{tabular}
+\end{landscape}
+\section{Acknowledgements}
+This research was supported by the Dutch national program COMMIT.
+
+\bibliographystyle{authordate3}
+\bibliography{citations}
 

### 2014-10-22 Nickname problem/src/nickname.tex

@@
 \title{The Nickname Problem}
-\date{\today}
-
+\date{\ldots}
 % non-indented, spaced paragraphs
@@
 Note that this is a natural property of basic computational metaphors.
-We can always take part of the argument of one function and hardcode it into the program at the cost of its length, and some additional bits to add it to the argument before processing.
+We can always take part of the argument of one function and hardcode it into the program at the cost of its prefix-coded length, and some additional bits to add it to the argument before processing.
 
@@
 Let $\{\phi_i\}$ and $\{\psi_j\}$ be two compact, computable enumerations of the computable functions.
-For a given computable function $\phi_i \simeq \psi_j$,  we have $\mbox{abs}(|i^*| - |j^*|) \leq c(\max(i^*, j^*)) + O(1)$.
-Where $c(x) \leq 2 \log(|x|) + 1$ is the overhead of prefix encoding $x$
+For a given computable function $\phi_i \simeq \psi_j$,  we have $\mbox{abs}(|i^*| - |j^*|) \leq \max(|\overline{\imath^*}|, |\overline{\jmath^*}|) + O(1)$.
+That is, the difference between the sizes of the minimal models is in $O(\log(\max(i, j)))$
 \end{lma}
@@
 
-Thus $|i^*| \leq |h| = |u| + |\overline{\jmath^*}| + O(1) = |j*| + c(\jmath^*) + O(1)$.
-Reversing $\phi$ and $\psi$ gives us $|j^*| \leq |i^*| + c(i^*) + O(1)$.
-Combining these, we get the desired result.
+Thus $|i^*| \leq |h| = |u| + |\overline{\jmath^*}| + O(1)$. Reversing $\phi$ and $\psi$ gives us $|j^*| \leq |h'| = |u'| + |\overline{\imath^*}| + O(1)$. Combining these, we get the desired result.
 \end{proof}
 
-This lemma tells us that so long as we restrict our numberings to be compact and computable, we can never gain more than a logarithmic difference between model sizes.
+This lemma tells us that so long as we restrict our indices to compact and effective, we can never gain more than a logarithmic difference between model sizes.
 In the limit as $i \rightarrow \infty$, this term represents a vanishing proportion.

### 2014-10-22 Overview/src/overview.tex

@@
 
-\title{The question of interesting information}
+\title{The question of interestingness}
 \date{\today}
@@
 \noindent Ever since the introduction of absolute measures of information content---such as Information Entropy and Kolmogorov Complexity---there have been  criticisms of the idea that these measures capture the informal notion of  information.
-The argument is that the signals which, under these measures, contain the highest amount of information are by definition completely unpredictable, and therefore cannot contain any interesting information.
+The argument is that the signals which, under these measures, contain the highest amount of information are of absolutely no use or interest to any observer.
 There is an informal notion of information content, which is not captured by these absolutes, a kind of useful or meaningful information content.
@@
 
-Consider the following, illustrative example.
+Consider the following example.
 An old-fashioned, black-and-white television set is turned off.
@@
 It is highly predictable.
-At every point we can predict exactly the next state of the signal.
 Describing this signal would be very simple.
@@
 \begin{quotation}
-A fundamentalproblem for statistical mechanics is to explain why dissipative systems (those in which entropy is continually being produced and removed to the surroundings) tend to undergo ``self-organisation,'' a spontaneous increase of structural complexity, of which the most extreme example is the origin and evolution of life.
+A fundamental problem for statistical mechanics is to explain why dissipative systems (those in which entropy is continually being produced and removed to the surroundings) tend to undergo ``self-organisation,'' a spontaneous increase of structural complexity, of which the most extreme example is the origin and evolution of life.
 The converse principle, namely that nothing very interesting is likely to happen in a system in thermal equilibrium, is reflected in the term ``heat death.'' In the modern world view, thermodynamic driving forces, such as the temperature difference between the hot sun and the cold night sky, have taken over one of the functions of God: they make matter transcend its clod-like nature andbehave instead in dramatic and unforeseen ways, for example molding itself into thunderstorms, people, and umbrellas.
 \end{quotation}
+
+\section{Preliminaries: The question of complexity}
+
+This section discusses the matter of what makes data complex.
+Most approaches to what makes data interesting build on these concepts.
+We review them here briefly, to briefly discuss our preliminaries and to define a uniform notation for the rest of the paper.
+
+\subsection{Entropy}
+
+In physical systems; in probability and coding.
+
+\subsection{Kolmogorov Complexity}
+
+Conditional complexity, Resource-bounded, MDL, two-part coding.
+
+\section{Approaches to Interestingness}
 
@@
 Complexity, order, etc.
-\section*{Preliminaries}
-
-\subsection*{Information Entropy}
-
-\subsection*{Kolmogorov Complexity}
-
-\subsection*{Two-part coding}
-
-\subsection*{Dynamical systems}
 
@@
 
+What questions should we ask of a proposed measure?
 \begin{itemize}
@@
   \item Context-dependence. Context plays an undeniable part in any notion of beauty or interestingness. Are contextual considerations acknowledged and given a place in the framework?
+\end{itemize}
+
+checklist per paper:
+\begin{itemize}
+  \item Check google scholar for citations to the paper.
+  \item Walk through the citations in the paper.
 \end{itemize}

### 2014-10-22 PhD Thesis/src/thesis.tex

@@
 
+\maketitle
+
+\tableofcontents
+
 \chapter{Introduction}
+\pb{General explanation of graphs, RDF, Kolmogorov complexity and two-part coding that my mother can understand.}
 
@@
 
-\section{The complex graph}
+\section{Complex graphs}
 
-% Complex graphs are not just funny things to look at, the are the fundamental unit of science in the 21st century.
+\pb{Complex graphs are not just funny things to look at, they are one of the most robust ways of storing data and communicating concepts.
+They should/will take a central role in science in the future.
+Moreso than tabular data.}
 
@@
 
-\begin{summary}We review the literature on complex graphs, and the approaches that have been used so far to provide statistical analyses on them.
+\begin{summary} We review the literature on complex graphs, and the approaches that have been used so far to provide statistical analyses on them.
+We highlight some commond pitfalls in reasoning, and we contrast the status quo with statistical practice on more conventional data.
 \end{summary}
 
-% Algorithmic statistics is how you stud graphs. Nothing else makes sense.
+\pb{Review paper on complex graph analysis.
+The current approach to statistics on complex graphs is to propose a an algorithm which generates random graphs, and to show that these graphs have some property in common with the graphs of interest (cf.
+the Barabasi algorithm and scale freeness).
+We can do better.}
 
-\chapter{Algorithmic Statistics: Bridging the gap between Kolmogorov complexity and MDL}
+\chapter{Algorithmic Statistics: Kolmogorov complexity and MDL}
 
-\begin{summary}MDL and Kolmogorov complexity are two offshoots of the same basic idea: that good compression makes good statistics.
+\begin{summary}Minimum Description Length (MDL) and Kolmogorov complexity are two offshoots of the same idea: good compression makes for good statistics.
 MDL takes this as an axiom and builds on it, while Kolmogorov complexity makes only the assumption that the data is produced by a computational process.
@@
 In order to have access to the best of both worlds, we start with the rigorous theoretical foundations of Kolmogorov complexity and build a platform that will allow us to fold in the techniques of MDL.
-We will revisit the principle of two-part coding, considered to be a usable, but cumbersome technique in MDL and show that when combined with Kolmogorov complexity it has some deep advantages over other methods.
+We will revisit the principle of two-part coding, considered to be a usable, but often crude, technique in MDL and show that when combined with Kolmogorov complexity it has some deep advantages over other methods.
 \end{summary}
+
+\pb{ICALP paper and possible continuation.
+The end result of this section is that (a) we can use techniques and results from Kolmogorov complexity or MDL in the same context (b) we can use computable approximations and (under resonable assumptions) assume that we are approximating the platonic ideals of Kolmogorov Complexity.}
+
+\section{Algorithmic statistics for graphs}
+
+\pb{Interprets the common practices of analysing complex graphs in terms of algorithmic statistics on a theoretical level. Running example: the Barabasi algorithm and the property of scale freeness.}
 
@@
 \begin{summary}The question of what makes data complex is has been decisively answered.
-First for IID data in, as maximal Entropy, and then for any data as maximal Kolmogorov Complexity.
+First for IID data---being the datasets with maximal Shannon Entropy---and then for any data as maximal Kolmogorov Complexity.
 But there's some dissonance in calling this data information-rich.
@@
 
+\pb{Review paper on facticity.}
+
+\section{A well-founded, universal approach}
+
+\pb{Our take on facticity.}
+
 \chapter{Graph modeling: extracting motifs}
 
-\begin{summary}Modeling complex objects starts with finding out what the components are.
+\begin{summary} Modeling complex objects starts with finding out what the components are.
 If we know the letters, we can find the words.
@@
 
-\begin{summary}In last chapter we induced motifs for various types of complex networks.
-We ended by applying the procedure recursively, collapsing upwards across several levels to compress the whole structure.
+\begin{summary} In the last chapter, we induced motifs for various types of complex networks.
+We ended by applying the procedure recursively, collapsing upwards across several levels to compress the whole structure into rewrite rules.
+From this, we can induce recursive rules, and with a rewrite grammar for a given graph.
 \end{summary}
@@
 
-\begin{summary}To model graphs in an algorithmic setting we can use any family of computational structures that can turn randomness into graph structures.
-One recently popular model that has been shown to be effective at producing random examples of any given model is the Deep Belief Network.
+\begin{summary}mTo model graphs in an algorithmic setting we can use any family of computational structures that can turn randomness into graphs.
+One recently popular model that has been shown to be effective at generating random examples is the Deep Belief Network.
 We investigate their potential to model graph structure, and fit them in to the framework of algorithmic statistics.
@@
 These instances can then be represented in some way---preferably a vector of numeric features---and analyzed statistically.
-In the past decade a situation has emerged and re-emerged that is an ill fit to this paradigm.
-We are faced with information in a graph: a social network, a web graph, an RDF dataset.
+In the past decade a situation has emerged and re-emerged that makes for an ill fit to this paradigm.
+We are faced with information in a graph; perhaps a social network, a web graph or an RDF dataset.
 We may know our instances, if they correspond to specific nodes in the graph, but the things we know about them are spread throughout the graph.
+There is no hard line between which parts of our data correspond to which instance.
+Als data belongs to all instances to verying degrees.
 We investigate the problem and provide a bridge to traditional machine learning.
 \end{summary}
+
+\pb{ESWC paper, of het vervolg daar op. Past niet helemaal in de lijn.}
 
@@
 
-\begin{summary}Throughout the previous chapters, we have performed many quick tests on small datasets, to illustrate and validate our points.
+\begin{summary} Throughout the previous chapters, we have performed many quick tests on small datasets, to illustrate and validate our points.
 In this chapter we take one use case and investigate more thoroughly, in order to show the full potential of our methods.
-We choose the domain of citation networks, and investigate various questions.
+We choose the domain of citation networks, and investigate various practical questions.
 \end{summary}
@@
 
-Conclusatory chapter.
+\pb{Conclusatory chapter. Tells the whole story again, but in terms of the concepts we've introduced. Shows that we did what we set out to do.}
 
@@
 \chapter{Tutorials}
-\section{Kolmogorov Complexity}
-\section{Minimum Description Length}
-\section{Graphs and RDF}
+\begin{summary}
+This appendix provides brief introductions to the subjects that are required to understand this thesis.
+They illustrate the use of our notation and provide a basis for readers who are well-versed in mathematics and computer science, but lack knowledge of these specific subjects.
+\end{summary}
+
+\section{Kolmogorov complexity}
+\section{Minimum description length}
+\section{Graphs and RDF-data}
+\section{Recommended literature}
+
+\chapter{Symbols and Terms}
+
+\begin{description}
+ \item[$K()$]
+ \item[$k()$]
+ \item[Shannon Entropy]
+\end{description}
+
 \chapter{Algorithms and techniques}

### 2014-10-22 Facticity/Facticity/paper.tex
Mijn versie...

--- 2014-10-22 Facticity/Facticity/paper.tex
@@
 
+\begin{proof}
+We first note that 
+\begin{equation}
+	K(yx) + K(|y|) + c' \geq K(x) \label{line:prelim1}
+\end{equation}
+for some constant $c$, since we can combine the shortest program to produce $yx$ and the shortest program for the length of $y$ to create a program which produces $yx$ and removes the first $|y|$ bits.
+
+Let $r$ be some string such that 
+\begin{equation}
+|r| - K^\phi(|r|) - c' > c \quad \text{,} \label{line:prelim2}\\
+\end{equation}
+
+with $c$ as given in the statement of the lemma. Define psi as $\psi_{ri}(x) = \phi_{i}(x)$, with $\psi_{j}(x) = \infty$ if $j$ cannot be read as $ri$. Note that this gives us 
+\begin{equation}
+K^\psi(x) \geq K^\phi(x) + |r| \label{line:prelim3}
+\end{equation} 
+for all $x$.
+
+We can now state the following:
+\begin{align*}
+K^\psi(f) &= \min\left\{ K^\psi(j) : \psi_j = f \right\}& \\
+       &= \min\left\{ K^\psi(ri) : \psi_{ri} = f \right\}& \\
+       &\geq \min\left\{ K^\phi(ri) + |r| : \psi_{ri} = f \right\}& \text{by \ref{line:prelim3}}\\
+       &= \min\left\{ K^\phi(ri) : \phi_{i} = f \right\} + |r| & \\
+       &\geq \min\left\{ K^\phi(i) - K^\phi(|r|) - c': \phi_{i} = f \right\} + |r| & \text{by \ref{line:prelim1}}\\
+       &= \min\left\{ K^\phi(i) : \phi_{i} = f \right\} + |r| - K^\phi(|r|) - c' &\\
+       &= \min\left\{ K^\phi(i) : \phi_{i} = f \right\} + |r| - K^\phi(|r|) - c' &\\
+       &= K^\phi(f) + |r| - K^\phi(|r|) - c' &  \\
+       &\geq K^\phi(f) + c & \text{by \ref{line:prelim2}} \\
+\end{align*}
+
+\end{proof}
+
 \bibliographystyle{plain}

### 2014-10-23 Facticity/Facticity/paper.tex
best of both worlds proof

--- 2014-10-22 Facticity/Facticity/paper.tex
@@
 
+\begin{proof}[Proof (combined)]
+First note that there is a $c'$ such that 
+\begin{equation}
+	K(1^r 0  x) + c' \geq K(x) \label{eq:prelim1}
+\end{equation}
+for all $x$, since we can create a program that first runs the shortest program for $1^r 0 x$ and then strips the prefix $1^r 0$ from the result.
+
+Define $\phi_{1^r0i}(x) = \psi_{i}(x)$, with $\phi_{j}(x) = \infty$ if $j$ does not contain a zero.
+Intuitively, any $\phi$-program is $r+1$ bits longer than the corresponding $\psi$-program.
+However, the used prefix function complicates this slightly.
+[TODO assume the prefix function to be convex.] But all we need is the bound
+\begin{equation}\label{eq:prelim2}\begin{split}
+K^\phi(x) &= \min \{ |\overline{1^r 0 \imath}| y : \phi_{1^r 0 i}(y) = x \}\\
+&  \ge \min\{|1^r0\bar\imath y| : \phi_{1^r 0 i}(y) = x \}\qquad\tn{(using convexity of $\bar\cdot$)}\\
+& = r+1 +\min\{|\bar\imath y| : \psi_i(y) = x \}\\
+& = r+1+K^\psi(x).
+\end{split}\end{equation}
+
+We can now relate the complexity of $f$ under the two enumerations as follows:
+\begin{align*}
+K^\phi(f) &= \min\left\{ K^\phi(j) : \phi_j = f \right\}& \\
+       &= \min\left\{ K^\phi(1^r0i) : \phi_{ri} = f \right\}& \\
+       &\ge  \min\left\{ K^\phi(i) : \phi_{ri} = f \right\}-c'&\text{by~\eqref{eq:prelim1}}\\
+       &\geq \min\left\{ K^\psi(i)  : \phi_{ri} = f \right\}+r+1-c'& \text{by \eqref{eq:prelim2}}\\
+       &= \min\left\{ K^\psi(i) : \psi_{i} = f \right\} + r+1 -c'& \\
+       &= K^\psi(f) + r + 1 - c'.
+\end{align*}
+The proof is completed by choosing $r=c+c'-1$.
+\end{proof}
+
 \bibliographystyle{plain}

### 2014-10-23 Facticity/Facticity/paper.tex
typo

@@
 K^\phi(f) &= \min\left\{ K^\phi(j) : \phi_j = f \right\}& \\
-       &= \min\left\{ K^\phi(1^r0i) : \phi_{ri} = f \right\}& \\
-       &\ge  \min\left\{ K^\phi(i) : \phi_{ri} = f \right\}-c'&\text{by~\eqref{eq:prelim1}}\\
-       &\geq \min\left\{ K^\psi(i)  : \phi_{ri} = f \right\}+r+1-c'& \text{by \eqref{eq:prelim2}}\\
+       &= \min\left\{ K^\phi(1^r0i) : \phi_{1^r0i} = f \right\}& \\
+       &\ge  \min\left\{ K^\phi(i) : \phi_{1^r0i} = f \right\}-c'&\text{by~\eqref{eq:prelim1}}\\
+       &\geq \min\left\{ K^\psi(i)  : \phi_{1^r0i} = f \right\}+r+1-c'& \text{by \eqref{eq:prelim2}}\\
        &= \min\left\{ K^\psi(i) : \psi_{i} = f \right\} + r+1 -c'& \\

### 2014-10-23 Facticity/Facticity/paper.tex
nog iets preciezer over die prefix functie

@@
 \begin{lemma}
-  For sufficiently large $c$ the set $B^n$ is nonempty for all $n$.
+  For sufficiently large $c$ the sets $B^n$ are nonempty for all $n$.
 \end{lemma}
@@
 
-\begin{lemma}[I think this is it]
+\begin{lemma}
+Assume the prefix function $\bar\cdot$ used in the definition of complexity is such that $\bar x-x$ is monotonically increasing.
 Let $\psi_1,\psi_2,\ldots$ be any acceptable numbering.
-For every $c$ there is another acceptable numbering $\phi_1,\phi_2,\ldots$ such that for all partial recursive functions $f$ we have
+Then for every $c$ there is another acceptable numbering $\phi_1,\phi_2,\ldots$ such that for all partial recursive functions $f$ we have
 \[
@@
 \begin{proof}
-Define $\tau$ by $\psi_\tau(\bar k p) = j$ where $j$ is given by $1^r 0 j = \psi_k(p)$.
-In other words, it is a universal element for $\psi$ that subsequently strips the prefix $1^r 0 $ from the output; this operation is clearly still partial recursive.
-Also note that this function has a fixed index -- it does not have to know $r$.
-For completeness define $\psi_\tau$ to infinity if the output is not in the correct format.
-
-Now define a new acceptable numbering $\phi_1,\phi_2,\ldots$ as follows:
-\[
-\phi_{1^r0 j}(p)=\psi_j(p)
-\]
-For any given partial recursive function $f$, we now compare $K^\phi(f)$ and $K^\psi(f)$.
-For $\phi$ it is:
-\[
-K^\phi(f) = \min\{K^\phi(i):\phi_i = f\}.
-\]
-Let $i^-$ be the index that achieves the minimum, and let $i^*$ be the shortest $\phi$-program that computes it, that is, $i^*=\bar a b$ such that $\phi_a(b)=i^-$.
-Then the codelength is $K^\phi(f)=K^\phi(i^-)=|i^*|$. Moreover, we must have $a=1^r0 k$ where $k$ is the index of the corresponding function in the $\psi$-enumeration: $\psi_k(b)=i^-$.
-
-To find an index of $f$ in the $\psi$-enumeration, note that $f=\phi_{i^-}=\phi_{1^r 0 j^-} = \psi_{j^-}$.
-We have $K^\psi(f) \le K^\psi(j^-)$; we proceed to provide a small $\psi$-program for $j^-$.
-This can be done because we have set things up so that $\psi_\tau(\bar k b)=j^-$.
-Thus, the $\psi$-program is $j^*=\bar\tau\bar k b$.
-The difference in length is
-%
-\[
-|i^*|-|j^*| = |\bar a b|-|\bar\tau\bar k b| = |\overline{1^r 0 k}| - |\bar\tau| - |\bar k| \approx r-|\bar\tau|.
-\]
-%
-(Here the $\approx$ is used because the prefix function has a slightly different effect on the two terms; however this is an unimportant detail that can be ironed out later).
-
-Thus, by choosing $r>|\bar\tau|+c$ we can guarantee that $K^\phi$ sucks by at least $c$ bits. Yay!
-\end{proof}
-
-\begin{proof}[Proof (combined)]
 First note that there is a $c'$ such that 
@@
 However, the used prefix function complicates this slightly.
-[TODO assume the prefix function to be convex.] But all we need is the bound
+But all we need is the bound
 \begin{equation}\label{eq:prelim2}\begin{split}
 K^\phi(x) &= \min \{ |\overline{1^r 0 \imath}| y : \phi_{1^r 0 i}(y) = x \}\\
-&  \ge \min\{|1^r0\bar\imath y| : \phi_{1^r 0 i}(y) = x \}\qquad\tn{(using convexity of $\bar\cdot$)}\\
+&  \ge \min\{|1^r0\bar\imath y| : \phi_{1^r 0 i}(y) = x \}\qquad\tn{(using assumption on $\bar\cdot$)}\\
 & = r+1 +\min\{|\bar\imath y| : \psi_i(y) = x \}\\

### 2014-10-23 Facticity/Facticity/paper.tex
Kleine stapjes

@@
 \documentclass{article}
+
 \usepackage{amsmath,amsfonts,amsthm,fullpage}
+\usepackage{color}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
 
@@
 \newcommand{\C}{\mathcal C}
+\newcommand{\B}{\mathbb B}
 \newcommand{\D}{\Delta}
@@
 \newcommand{\symb}[1]{\texttt{#1}}
+\newcommand{\p}{\quad\text{.}}
+
+\newcommand{\sdr}[1]{\textcolor{blue}{\small #1\textsuperscript{[sdr]} }}
+\newcommand{\pb}[1]{\textcolor{OliveGreen}{\small #1 \textsuperscript{[pb]} }}
 
@@
   Let $\M$ be any model class that contains the identity function. There is an acceptable numbering $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that for every $n$, all $x\in B^n$, 
-we have $G^\phi(x)=1$.
+we have $G^\phi_M(x)=1$.
 \end{lemma}
@@
 K^\phi(f)-K^\psi(f) \ge c.
-\]
+\] \label{lemma:building-block}
 \end{lemma}
@@
 
+\begin{lemma}
+Given $n_0$ and $c_0$, there exist $y \in \B$ and $h \in \N$ such that 
+\begin{itemize}
+  \item $|0^hy| \geq n$,
+  \item $|0^hy| - c \leq K(0^hy) \leq |0^hy| + c$,
+  \item and $C(0^hy) + c' \geq K(h) + C(y)$ for some constant $c'$.
+  \item $K(h) \geq c'' log(|0^hy|)$ for some positive $c''$
+\end{itemize}
+\end{lemma}
+
+\noindent \pb{Ik denk dat we dit kunnen bewijzen, maar het leidt uiteindelijk maar tot een logterm variantie tussen verschillende versies van G.
+Dat betekent dat ze asymptotisch niet echt verschillen.
+Wat we nodig hebben is een groot model $f$ waarvan we weten dat er een $x$ in $B^n$ zit, die met $f$ $C(x)$ to op een constant benaderd.
+Een familie van modellen $f$ kan ook.
+Het probleem is dat het bij zulke lage compressie moeilijk is om een compressiemethode met een groot model te bedenken\ldots}
+
+\begin{lemma}
+There exists an acceptable numbering $\phi1, \phi2, \ldots$ such that for all $n_0$, $c_0$, there exist $x, n,c$ with $c > c_0$, $n> n_0$ and $x \in B^n_c$ so that $G^\phi(x) > c'\log(|x|)$ for some positive $c'$.
+\end{lemma}
+
+\begin{theorem}
+The exist acceptable numberings $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2, \ldots$ such that for any given constant $c$ there exist $x$ such that $|G_M^\phi(x) - G_M^\psi(x)| > c$
+\end{theorem}
+
 \bibliographystyle{plain}

### 2014-10-27 Facticity/Facticity/paper.tex
...

@@
 
-\newcommand{\sdr}[1]{\textcolor{blue}{\small #1\textsuperscript{[sdr]} }}
-\newcommand{\pb}[1]{\textcolor{OliveGreen}{\small #1 \textsuperscript{[pb]} }}
+\newcommand{\sdr}[1]{\textcolor{blue}{\small #1\textsuperscript{[Steven]} }}
+\newcommand{\pb}[1]{\textcolor{OliveGreen}{\small #1 \textsuperscript{[Peter]} }}
 
@@
 
+\section{Unbounded and top-heavy sophistications}
+
+\begin{definition}
+We call a version of Sophistication $F^\phi$ \emph{bounded} if its ad-hoc choices (the enumeration $\phi$) can be chosen so that
+\[
+	\exists c \forall x: F^\phi(x) \leq c \p
+\]
+Similarly, we call $F^\phi(x)$ \emph{top-heavy} if there exists a $\phi_1, \phi_2, \ldots$ such that
+\[
+	\exists c \forall x: F^\phi(x) \geq K^F(x) - c 
+\]
+Where $K^F$ is the two-part coding minimized within $F$.
+\end{definition}
+
+If some $F$ is either top-heavy, or unbounded we can show that it must either lack invariance, or be useless as a measure of model complexity.
+If the property proved holds for all enumerations $\phi$, and not just a single one, the sophistication is either bounded, or the same as the Kolmogorov complexity.
+If the measure is not bounded or top-heavy for other enumerations, the difference between the two is unbounded, and $F$ is not invariant.
+
+One of the earliest proposals to separate structure and content was the Kolmogorov structure function \cite{TODO}. The following is the version of sophistication it leads to. 
+\begin{definition}
+\[
+	S^\phi(x) = K(S^*) \,\tn{where }\, (S, i) = \argmin_{(S, i)\,:\,\text{$S$ is a finite set},\,S_i = x} K(S) + |i|
+\]
+\end{definition}
+
+\begin{theorem}
+There exist acceptable numberings $\phi_1, \phi_2, \ldots$ for which $S^\phi$ is top-heavy.
+\end{theorem}
+\begin{proof}
+Let $\psi_1, \psi_2, \ldots$ be some acceptable numbering. Define $\psi$ as 
+\begin{align*}
+\phi_{0\bar\imath p}(\epsilon) &= \{\psi_i(p)\} \\
+\phi_{1ri}(p) &= \psi_i(p)\\ 
+\phi_j(x) = \infty \,\text{if $j$ cannot be decoded as above}
+\end{align*}
+For any $x$ with $\psi$-description $(i, p)$, that is is $\psi_i(p) = x$, we have two representations in $\phi$.
+Firstly, we have $\{x\}, 1$ with description length $K^\psi(x) + 2$.
+Secondly, we have a representation in $K^\psi{x} + r$.
+By choosing $r$ sufficiently large, the first representation will always win.
+\end{proof}
+
+\section{Vitanyi's Sophistication}
+
 \begin{lemma}
@@
 Dat betekent dat ze asymptotisch niet echt verschillen.
-Wat we nodig hebben is een groot model $f$ waarvan we weten dat er een $x$ in $B^n$ zit, die met $f$ $C(x)$ to op een constant benaderd.
+Wat we nodig hebben is een groot model $f$ waarvan we weten dat er een $x$ in $B^n$ zit, die met $f$ $C(x)$ to op een constante benaderd.
 Een familie van modellen $f$ kan ook.

### 2014-10-29 Facticity/Facticity/paper.tex
fuck
mooi maar beetje abstract lemma; prove is nog todo

@@
 \newcommand{\p}{\quad\text{.}}
+\newcommand{\tuple}[1]{\left\langle{#1}\right\rangle}
 
@@
 \begin{definition}
+  Given a model $f$ and model class $\M$, define model-based complexities by
+  \begin{align*}
+  C^f(x) &= \min\{|y|:f(y)=x\}\\
+  L^f(x) &= K(f)+C^f(x)\\
+  L^{\M}(x) &= \min\{L^f(x)\mid f\in\M\}
+  \end{align*}
+\end{definition}
+
+\hrule
+
+\begin{lemma}
+  Let $\M$ be any model class and let $\M'\subseteq \M$ be a sub-model with the following properties:
+\begin{enumerate}
+  \item The function $g(x):=\argmin_{f\in\M'}L^f(x)$ is computable and uniquely defined for all $x$
+  \item $L^{\M'}(x)-L^{\M}(x)\le c$ for some constant $c$, all $x$.
+\end{enumerate}
+Then there is an enumeration $\phi_1,\phi_2,\ldots$ such that $F^{\M,\phi}(x)=K^\phi(g(x))+1$ for all $x$.
+\end{lemma}
+\begin{proof}
+
+Construct $\phi_1,\phi_2,\ldots$ such that
+\begin{enumerate}
+\item For all $f\in\M'$ have $K^\phi(f) = K^\psi(f)+1$ (zelf bewijzen)
+\item For all $f\not\in\M'$ have $K^\phi(f)>K^\psi(f)+c$.
+\end{enumerate}
+
+Let $\psi_1,\psi_2,\ldots$ be an acceptable numbering of the partial recursive functions, and let $g'(x)=i$ such that $\psi_i=g(x)$. Now define a new ordering $\phi_1,\phi_2,\ldots$ by
+\[
+\begin{cases}
+\phi_{0i}(y)&=\psi_i(y),\qquad\tn{if $i=g'(x)$ for some $x$, and $\infty$ otherwise}\\
+\phi_{1^r0i}(y)&=\psi_i(y)\\
+\phi_{\tn{all-ones}}(y)&=\infty
+\end{cases}
+\]
+
+Voor alle $x$ kiezen we een makkelijk te vinden, en goedkope representatie met input $y$. Die maken we extra goedkoop in de nieuwe enumeratie. Dan wordt ie altijd gekozen.
+\end{proof}
+\hrule
+\begin{lemma}
+  Let $\M$ be a model class with the property that there is a constant $c$ and a computable function $g$ such that for all $x$ we have $\gamma(x\mid g(x))\le c$.
+  For such model classes, there are acceptable enumerations $\phi_1,\phi_2,\ldots$ of the partial recursive functions for which $K^{\M,\phi}(x)-F^{\M,\phi}(x)$ is bounded.
+\end{lemma}
+\begin{proof}
+
+\end{proof}
+
+\begin{lemma}
+  Let $\M$ be a model class with the property that there is a constant $c$ and a model $f$ such that for all $x$ the counterevidence satisfies $\gamma(x\mid f)\le c$.
+  For such model classes, there are acceptable enumerations $\phi_1,\phi_2,\ldots$ of the partial recursive functions for which $F^{\M,\phi}(x)$ is bounded.
+\end{lemma}
+
+\begin{definition}
+  Given a model class $\M$, a representation $(f,y)$ with $f\in\M$ is said to be \emph{$c$-optimal} if
+\[
+K(f)+|y|\le \min_{f',y':f'(y')=f(y)}\left(K(f')+|y'|\right)+c.
+\]
+\end{definition}
+
+\begin{lemma}
+  Let $\M$ be a model class with the property that there are $c,y$ such that for all $x$ there is an $f$ such that $f(y)=x$ and $(f,y)$ is $c$-optimal.
+  For such model classes, there are acceptable enumerations $\phi_1,\phi_2,\ldots$ of the partial recursive functions for which $K^{\M,\phi}(x)-F^{\M,\phi}(x)$ is bounded.
+\end{lemma}
+
+\begin{lemma}
+  Let $\M$ be a model class with the property that there are $c,f$ such that for all $x$ there is an $y$ such that $f(y)=x$ and $(f,y)$ is $c$-optimal.
+For such model classes, there are acceptable enumerations $\phi_1,\phi_2,\ldots$ of the partial recursive functions for which $F^{\M,\phi}(x)$ is bounded.
+\end{lemma}
+
+\begin{definition}
 We call a version of Sophistication $F^\phi$ \emph{bounded} if its ad-hoc choices (the enumeration $\phi$) can be chosen so that
@@
 
-\section{Vitanyi's Sophistication}
-
-\begin{lemma}
-Given $n_0$ and $c_0$, there exist $y \in \B$ and $h \in \N$ such that 
-\begin{itemize}
-  \item $|0^hy| \geq n$,
-  \item $|0^hy| - c \leq K(0^hy) \leq |0^hy| + c$,
-  \item and $C(0^hy) + c' \geq K(h) + C(y)$ for some constant $c'$.
-  \item $K(h) \geq c'' log(|0^hy|)$ for some positive $c''$
-\end{itemize}
-\end{lemma}
-
-\noindent \pb{Ik denk dat we dit kunnen bewijzen, maar het leidt uiteindelijk maar tot een logterm variantie tussen verschillende versies van G.
-Dat betekent dat ze asymptotisch niet echt verschillen.
-Wat we nodig hebben is een groot model $f$ waarvan we weten dat er een $x$ in $B^n$ zit, die met $f$ $C(x)$ to op een constante benaderd.
-Een familie van modellen $f$ kan ook.
-Het probleem is dat het bij zulke lage compressie moeilijk is om een compressiemethode met een groot model te bedenken\ldots}
-
-\begin{lemma}
-There exists an acceptable numbering $\phi1, \phi2, \ldots$ such that for all $n_0$, $c_0$, there exist $x, n,c$ with $c > c_0$, $n> n_0$ and $x \in B^n_c$ so that $G^\phi(x) > c'\log(|x|)$ for some positive $c'$.
-\end{lemma}
-
-\begin{theorem}
-The exist acceptable numberings $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2, \ldots$ such that for any given constant $c$ there exist $x$ such that $|G_M^\phi(x) - G_M^\psi(x)| > c$
-\end{theorem}
-
 \bibliographystyle{plain}

### 2014-10-30 Facticity/Facticity/paper.tex
...

@@
 \newcommand{\symb}[1]{\texttt{#1}}
+
 \newcommand{\p}{\quad\text{.}}
+
 \newcommand{\tuple}[1]{\left\langle{#1}\right\rangle}
@@
 
+\begin{lemma}[Invariance]
+Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be two acceptable numberings.
+There exists a constant $c$ such that $\left| K^\phi(f) - K^\psi(f)\right | \leq c$ for all $f$.
+\label{lemma:invariance}
+\end{lemma}
+\begin{proof}
+Let $g(i)$ be the function which which translates from $\psi$ to $\phi$.
+\begin{align*}
+K^\phi(f) &= \min\left\{ K^\phi(i) : \phi_i= f\right\} \\
+&\geq \min\left\{ K^\psi(i) : \phi_i= f\right\} - c\\
+&= \min\left\{ K^\psi(i) : \psi_{g(i)}= f\right\} - c\\
+&\geq \min\left\{ K^\psi(i) : \psi_i= f\right\} - c\\
+= K^\psi(f)
+\end{align*}
+We can reverse $\phi$ and $\psi$ without loss of generality to achieve the same inequality the other way around, completing the proof.
+\end{proof}
+
 \begin{lemma}
@@
 \begin{definition}
+We call $\phi_1, \phi_2, \ldots$ a \emph{partial computable numbering} of some subset of the partial computable functions, if there exists a partial computable function $u: \N \times \B \to \B \cup \infty$ such that $u(i, p) = \phi_i(p)$.
+We call $u$ the universal model for the numbering.
+We call a numbering \emph{optimal} if there is a constant $c$ such that for all $i$, we have $|i| \leq K(\phi_i) + c$.
+\end{definition}
+
+\begin{lemma}
+There exists an acceptable, optimal partial computable numbering of the partial computable functions.
+\label{lemma:optimal-numbering}
+\end{lemma}
+\begin{proof}
+Let $\psi_1, \psi_2, \ldots$ be some acceptable numbering, with some $\psi_{\text{div}}$ with is undefined for all its arguments. Let 
+\begin{align*}
+	\phi_{\bar\imath p} &= \psi_{\psi_i(p)} \qquad\text{if $\psi_i(p)$ is defined}\\
+	\phi_{\bar\imath p} &= \psi_{\text{div}} \qquad\text{otherwise}\\
+	\phi_{j} &= \psi_{\text{div}} \qquad\text{if $j$ can not be read as $\bar\imath p$}
+\end{align*}
+
+\noindent To show that this numbering is acceptable.
+We must provide two computable function to translate t and from a given acceptable numbering.
+We use $\psi$ for this purpose.
+To find a $\phi_j$ which corresponds to a given $\psi_i$, we can simply use the identity function $\psi_d(x) = x$.
+For a given $\psi_i$, the function $\phi_{\overline{d}i}$ computes the same function.
+
+To translate from $\phi_j$ to $\phi_i$, we are faced with a problem.
+If we are given $j = \bar\imath p$, we cannot decide algorithmically whether $\psi_i(p)$ is defined.
+However, we can construct the function $\psi_r(\bar\imath p, q)$ such that it computes $\psi_r{\psi_i(p)}(q)$ if $\psi_i(p)$ is defined and is undefined otherwise, or if the first argument cannot be read as $\bar\imath p$.
+For a given $j$ we can also find an $r'$ such that $\psi_{r'}(q) = \psi_(j, q)$.
+Thus, given any $\phi_i$ we can construct and return this function.
+
+Finally, we'll show that the minimum length of an index for function $f$ in $\{\phi_i\}$ is equal to $K(f)$ up to a constant:
+\begin{align*}
+K^{\psi}(f) &= \min\left\{K^{\psi}(j)\,:\,\psi_j = f \right\} \\
+&= \min_j\left\{\min_{i, p} \{|\bar\imath p| \,:\,\psi_i(p) = j\} \,:\,\psi_j = f\right\} \\
+&= \min_{i, p} \left\{|\bar\imath p| \,:\,\psi_i(p) = j, \psi_j = f\right\} \\
+&= \min_j \left\{|j|\,:\, \phi_j = f \right\}
+\end{align*}
+\end{proof}
+
+\begin{lemma}
+Let $M$ be the class of \emph{singleton functions}, that is, the functions which produce a single output for all arguments, including the function which is undefined for all its arguments.
+There exists an optimal partial computable numbering of $M$.
+\end{lemma}
+\begin{proof}
+Let $\psi_1, \psi_2, \ldots$ be some acceptable numbering. Let $\phi_{\bar\imath p}(\cdot) = \psi_i(p)$ \ldots. 
+\end{proof}
+
+\begin{definition}
   Given a model $f$ and model class $\M$, define model-based complexities by
   \begin{align*}
-  C^f(x) &= \min\{|y|:f(y)=x\}\\
-  L^f(x) &= K(f)+C^f(x)\\
-  L^{\M}(x) &= \min\{L^f(x)\mid f\in\M\}
+  C^f(x) &= \min\{|y|:f(y)=x\} \text{,}\\
+  L^f(x) &= K(f)+C^f(x) \text{,}\\
+  L^{\M}(x) &= \min\{L^f(x)\mid f\in\M\}\text{.}
   \end{align*}
 \end{definition}
-
-\hrule
 
@@
 \begin{enumerate}
-  \item The function $g(x):=\argmin_{f\in\M'}L^f(x)$ is computable and uniquely defined for all $x$
+  \item There exists an optimal partial computable numbering $\eta_1, \eta_2, \ldots$ for $M'$,
   \item $L^{\M'}(x)-L^{\M}(x)\le c$ for some constant $c$, all $x$.
 \end{enumerate}
-Then there is an enumeration $\phi_1,\phi_2,\ldots$ such that $F^{\M,\phi}(x)=K^\phi(g(x))+1$ for all $x$.
+There is an enumeration $\phi_1,\phi_2,\ldots$ of $\M$, such that $S^\phi_{\M}(x) = S^\eta_{\M'}(x)+1$ for all $x$.
+\end{lemma}
+\begin{proof}
+Let $\psi_1, \psi_2, \ldots$ be some acceptable numbering. Let $u$ be the universal model for the numbering $\eta_1, \eta_2, \ldots$. We define the numbering $\phi$ as follows:
+\begin{align}
+\phi_{0}i(p) &= u(i, p) \\
+\phi1^r0i(p) &= \psi_i(p) \\
+\phi_j(\cdot) &= \infty \quad\text{if $j$ contains no zeroes.}
+\end{align}
+Note that with this definition, we have $K^\phi(f) \leq K^\eta(f) \leq K^\psi(f) + c$ if $f \in \M'$.
+If $f$ is not in $\M'$, we have $K^\phi(f) = K^\psi(f) + r + 1 - c'$, by the same argument as used in the proof of Lemma~\ref{lemma:building-block}.
+
+Let $x$ be some finite string, and let $(i, p)$ be the pair such that $\psi_i(p)$ witnesses $L^M_\psi(x)$.
+By assumption, $L^{\M'}_\psi \leq L^M(x) +c$.
+Thus, we have a pair $(j, q)$, with $\eta_j(q) = x$ and $K^\eta(j) + |q| \leq K^\phi(i) + |p| + c$.
+If we choose $r = c' + c+ 1$, we know that $\phi$ will have representations $(1^r0i, q)$ with code length $L^M_\psi(x) + c+ 1 $, and $(0j, q)$ with code length less than or equal too $L^{\M}_\psi(x) +c$.
+This ensures that the sophistication under $\phi$ will be $K^\phi(\eta_j) = K^\eta(\eta_j) + O(1) = F^{\M'}_\eta(j)$.
+\end{proof}
+
+\begin{lemma}
+  Let $\M$ be any model class and let $\M'\subseteq \M$ be a sub-model with the following properties:
+\begin{enumerate}
+  \item There exists a (non-optimal) computable numbering $\psi_1, \psi_2, \ldots$ for $M'$,
+  \item $L^{\M'}(x)-L^{\M}(x)\le c$ for some constant $c$, all $x$.
+\end{enumerate}
+Then there is an enumeration $\phi_1,\phi_2,\ldots$ such that $|F^{\psi}_{\M'} - \leq F^{\phi}_{\M}| \leq c$ for infinitely many $x$.
 \end{lemma}
@@
 
-Construct $\phi_1,\phi_2,\ldots$ such that
-\begin{enumerate}
-\item For all $f\in\M'$ have $K^\phi(f) = K^\psi(f)+1$ (zelf bewijzen)
-\item For all $f\not\in\M'$ have $K^\phi(f)>K^\psi(f)+c$.
-\end{enumerate}
+\end{proof}
 
-Let $\psi_1,\psi_2,\ldots$ be an acceptable numbering of the partial recursive functions, and let $g'(x)=i$ such that $\psi_i=g(x)$. Now define a new ordering $\phi_1,\phi_2,\ldots$ by
-\[
-\begin{cases}
-\phi_{0i}(y)&=\psi_i(y),\qquad\tn{if $i=g'(x)$ for some $x$, and $\infty$ otherwise}\\
-\phi_{1^r0i}(y)&=\psi_i(y)\\
-\phi_{\tn{all-ones}}(y)&=\infty
-\end{cases}
-\]
-
-Voor alle $x$ kiezen we een makkelijk te vinden, en goedkope representatie met input $y$. Die maken we extra goedkoop in de nieuwe enumeratie. Dan wordt ie altijd gekozen.
-\end{proof}
 \hrule

### 2014-10-31 Facticity/Facticity/paper.tex
lemma aant fixen

@@
 \begin{definition}
-We call $\phi_1, \phi_2, \ldots$ a \emph{partial computable numbering} of some subset of the partial computable functions, if there exists a partial computable function $u: \N \times \B \to \B \cup \infty$ such that $u(i, p) = \phi_i(p)$.
-We call $u$ the universal model for the numbering.
-We call a numbering \emph{optimal} if there is a constant $c$ such that for all $i$, we have $|i| \leq K(\phi_i) + c$.
+We call a numbering $\eta_1,\eta_2,\ldots$ \emph{optimal} if there is a constant $c$ such that for all $i$, we have $\min\{|j|:\eta_i=\eta_j\} \leq K(\eta_i) + c$.
 \end{definition}
@@
 \begin{lemma}
-There exists an acceptable, optimal partial computable numbering of the partial computable functions.
+There exists an acceptable, optimal numbering of the partial computable functions.
 \label{lemma:optimal-numbering}
@@
 Let $\psi_1, \psi_2, \ldots$ be some acceptable numbering, with some $\psi_{\text{div}}$ with is undefined for all its arguments. Let 
-\begin{align*}
-	\phi_{\bar\imath p} &= \psi_{\psi_i(p)} \qquad\text{if $\psi_i(p)$ is defined}\\
-	\phi_{\bar\imath p} &= \psi_{\text{div}} \qquad\text{otherwise}\\
-	\phi_{j} &= \psi_{\text{div}} \qquad\text{if $j$ can not be read as $\bar\imath p$}
-\end{align*}
+\[\begin{cases}
+	\phi_{j} = \psi_{\text{div}} &\text{if $j$ can not be read as $\bar\imath p$}\\
+	\phi_{\bar\imath p} = \psi_{\psi_i(p)} &\text{if $\psi_i(p)$ is defined}\\
+	\phi_{\bar\imath p} = \psi_{\text{div}} & \text{otherwise}
+\end{cases}\]
 
-\noindent To show that this numbering is acceptable.
-We must provide two computable function to translate t and from a given acceptable numbering.
+\noindent To show that this numbering is acceptable, we must provide two computable functions to translate to and from a given acceptable numbering.
 We use $\psi$ for this purpose.
@@
 \begin{lemma}
-  Let $\M$ be any model class and let $\M'\subseteq \M$ be a sub-model with the following properties:
+  Let $\psi_1,\psi_2,\ldots$ be any acceptable enumeration of the partial recursive functions.
+  Let $\M$ be any model class and let $D:\{0,1\}^*\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e.
+if $f=\psi_{D(p)}$ then $p$ is a $D$-description of $f$.
+Let $\M'=\{\psi_i:i\in\tn{range}(D)\}$.
+Further assume there is a constant $c$ such that
 \begin{enumerate}
-  \item There exists an optimal partial computable numbering $\eta_1, \eta_2, \ldots$ for $M'$,
-  \item $L^{\M'}(x)-L^{\M}(x)\le c$ for some constant $c$, all $x$.
+  \item $\forall_{f\in\M'}:\min\{|p|:\psi_{D(p)}=f\}\le K^\psi(f)+c$
+  \item $\forall_x:L^{\M'}(x)-L^{\M}(x)\le c$.
 \end{enumerate}
-There is an enumeration $\phi_1,\phi_2,\ldots$ of $\M$, such that $S^\phi_{\M}(x) = S^\eta_{\M'}(x)+1$ for all $x$.
+Then is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that $S^\phi_{\M}(x) = 1$ for all $x$.
 \end{lemma}
 \begin{proof}
-Let $\psi_1, \psi_2, \ldots$ be some acceptable numbering. Let $u$ be the universal model for the numbering $\eta_1, \eta_2, \ldots$. We define the numbering $\phi$ as follows:
-\begin{align}
-\phi_{0}i(p) &= u(i, p) \\
-\phi1^r0i(p) &= \psi_i(p) \\
-\phi_j(\cdot) &= \infty \quad\text{if $j$ contains no zeroes.}
-\end{align}
+We define the numbering $\phi$ as follows:
+\[\begin{cases}
+\phi_0(p) = 1^r 0 D(p) \\
+\phi_{1^r0i}(p) = \psi_i(p) \\
+\phi_j(\cdot) = \infty &\text{if $j$ contains no zeroes.}
+\end{cases}\]
+First suppose $f\in\M'$. Then
+\[\begin{split}
+K^\phi(f)&=\min\{K^\phi(i):\phi_i=f\}\\
+&=\min\{|\bar\jmath q|:\phi_j(q)=i, \phi_i=f\}\\
+&=\min\{|\bar\jmath q|:\phi_{\phi_j(q)}=f\}\\
+&\le\min\{|\bar0 q|:\phi_{\phi_0(q)}=f\}~\text{(using $j=0$)}\\
+&=\min\{|\bar0 q|:\phi_{1^r0 D(q)}=f\}\\
+&=|\bar 0|+\min\{|q|:\psi_{D(q)}=f\}\\
+&\le K^\psi(f)+c+|\bar 0|,
+\end{split}
+\]
+where the last inequality uses the first assumption.
+
+\hrule
+
 Note that with this definition, we have $K^\phi(f) \leq K^\eta(f) \leq K^\psi(f) + c$ if $f \in \M'$.

### 2014-11-04 Facticity/Facticity/paper.tex
...

@@
 
-Adriaans takes this problem seriously and attempts to address it by imposing additional restrictions on the ordering of TMs implied by the used UTM \cite{TODO}.
+Adriaans takes this problem seriously and attempts to address it by imposing additional restrictions on the ordering of TMs implied by the used UTM \cite{adriaans2012facticity}.
 We are skeptical that such measures can effectively eliminate this problem without resorting to orderings of the TMs that are not admissible [TODO ref].

### 2014-11-12 Facticity/Facticity/paper.tex
beetje zitten plannen

--- 2014-10-31 Facticity/Facticity/paper.tex
@@
 
-\section{Case against index length as model complexity}
-All definitions of meaningful information are based on a two-part description of the data: the first part describes a \emph{model} for the data, and the second part describes an input to the model that makes it reconstruct $x$ exactly.
-However there is a fundamental dichotomy in way the various authors work out this idea technically.
+\section{Plan}
 
-The aproach taken in the works of Kolmogorov, Vit\'anyi, (Cover?
-Shen?
-etc) is sound (apart from the issues discussed elsewhere in this paper), but the other approach, adopted by Koppel, Adriaans, (others), while very intuitive, has a fundamental problem that seems to originate in a misunderstanding of the role Kolmorov complexity has to play in the two part code.
-
-The misunderstanding, which appears to be shared by authors on both sides of the dichotomy, is that the two components of the Kolmogorov complexity (the index of a Turing machine and its input) \emph{are} the two-part code used to separate the meaningful information from the noise.
-This idea certainly seems reasonable, as it is expressed not only by Koppel, who writes
-
-[insert quote here: In his 1988 chapter ``Structure'', Koppel has a very good quote about how it is such a pity that program and data are not considered separately anymore.]
+Idee van Sophistication: gebruik two-part coding om een scheiding te maken tussen structuur van de data (modelinformatie) en de ruis.
 
@@
 
-The problem of this two-part code lies with the first part: the index of the Turing machine depends on the used universal Turing machine, and the same Turing machine may well have, say, an exponentially higher index on one UTM than on another.
-(This is sometimes referred to as the ``nickname problem'', \cite{TODO}.) As such the index is \emph{not} a good measure of the model complexity.
+Inherent property of the data: 
+\begin{itemize}
+\item Should be a two-part effective description of $x$
+\item Invariant: $\exists c:\forall x:|S^1(x)-S^2(x)|\le c$
+\item Both structure and noise are unbounded
+\end{itemize}
 
-To illustrate what can go wrong, consider all TM's $\{T_{i_1}, \ldots, T_{i_k}\}$  that are $c$-sufficient statistics for $x$, i.e.
-they have $\overline{i_j}+\min\{|y|:T_{i_j}(y)=x\}\le K(x)+c$.
-If we were to switch to a different universal Turing machine that uses a different index function, any or all of these Turing machines may suddenly yield a two-part codelength that is much larger than the Kolmogorov complexity -- and the potential difference in two-part codelength is not bounded by a constant.
+Conjecture: er is geen functie van de data die aan bovenstaande eisen voldoet.
 
-Adriaans takes this problem seriously and attempts to address it by imposing additional restrictions on the ordering of TMs implied by the used UTM \cite{TODO}.
-We are skeptical that such measures can effectively eliminate this problem without resorting to orderings of the TMs that are not admissible [TODO ref].
+\subsection{Review}
+Onderscheid op basis van drie properties:
+\begin{itemize}
+\item Model class: total / partial. Reden voor total: bij partial rec. is de UTM in het model, collapse, citaten. Nadeel: arbitraire beperking op de beschrijvingskracht, dus minder universeel
+\item Models are prefix functions / not prefix functions: viz $C$ vs $K$.
+Als wel prefix, dan gaat alles makkelijker kapot.
+Als niet prefix dan wordt de balans verstoord, het wordt moeilijk als de balans ook de andere kant op wordt verstoord door de modelklasse te beperken: die twee handicaps zijn moeilijk te vergelijken.
+\item Model length is prefix encoded index length / $K(f)$: compact numberings, bespiegeling over of het gat tussen $|\bar\imath|$ en $K(\phi_i)$ unbounded is.
+Wat is nu eigenlijk de two-part code precies, is dat $K$ als white box of is $K$ gebruikt voor het eerste deel?
+Ook beetje vreemd om in de eerste visie modelklassen te introduceren.
+\end{itemize}
 
-But there is already a simple way to avoid this pitfall, which uses Kolmogorov complexity as a \emph{building block} of the two-part code: the first part of the code identifies a model, using its shortest possible description, whose length is measured by Kolmogorov complexity.
-The second part of the code provides an input to the model for which it outputs $x$.
+Bestaande aanpakken:
+\begin{itemize}
 
-The model may be a Turing machine, but can also be used with other model classes; for example in Kolmogorov's original formulation the model is a finite set of objects that contains $x$.
-Since the models are not usually represented as binary strings, the definition of Kolmogorov complexity has to be suitably extended to the considered model classes, but this is not a fundamental problem, [refer to appropriate definitions].
+\item Facticity: partial / niet prefix: bounded
+
+\item Paul's sophistication: Sufficient statistic: alles wat $K$-complexiteit haalt tot op slack $c$.
+Minimal sufficient statistic minimaliseert daarover de modelgrootte.
+Model bevat totale prefixfuncties.
+Noise is bounded.
+
+\item total / niet prefix, o.a.
+coarse sophistication.
+We geloven invariantie niet.
+Voor alle shallow strings bounded (nog niet bewezen, maar werkt door alles total te maken met vaste grote maar computable time bound)
+
+\end{itemize}
+
+\subsection{Conclusion}
+Stel onze conjecture is waar, is het dan helemaal nutteloos?
+Nee, want lengte van two-part representations is nog steeds een redelijke maat voor counterevidence against a model.
+No-hypercompression, $p$-values, randomness deficiency.
+Probleem zit in harde cutoffs zoals minimalisatie en een vaste constante waarboven representaties niet meer meedoen.
+Daar gaat invariantie van kapot.
+
+Onze mening: als onze conjecture waar is, dan moet je constraints weglaten: maak het makkelijk en doe prefix functies en alle partial recursive functions, en accepteer de collapse gewoon.
+Maar dat betekent weer dat je een hele wolk van two-part optimale representaties hebt waar je niet tussen moet willen kiezen.
+
+Er is een gat tussen $K(f)+|y|-K(x)$ en $\delta(x\mid f)$. De laatste is de juiste maat voor de counterevidence tegen $f$.
 
@@
 \begin{itemize}
-
-\item Kolmogorov complexity measures the shortest effective description of a binary sequence $x$.
-As such it is an \emph{objective} property of $x$ in the sense that changing around the enumeration of Turing machines in its definition only changes the Kolmogorov complexity by a constant independent of $x$.
-
-The same should hold for all definitions:
-\begin{itemize}
-\item The complexity of any model should not change by more than a constant when the ordering of TMs is changed.
-\item The same holds for the measure of meaningful information/sophistication
-\end{itemize}
-
-\item Iets uitgebreidere versie van bovenstaande, uit mijn mailtje: definitie 1: twee frameworks met complexiteitsmaten $K_1$ en $K_2$ en modelklassen $\M_1$ en $\M_2$ zijn equivalent als voor alle representaties $r$ in $\M_1$ er een $s$ in $\M_2$ is zodanig dat
-\begin{itemize}
-\item $K_1(r)-K_2(s) = O(\log(K_1(r)+K_2(s)))$
-\item $K_1(r)+L_r(x) - (K_2(s)+L_s(x)) = O(1)$
-\end{itemize}
-
-\item We waren bezig met operationaliseren wanneer een methode nu eigenlijk de juiste hypothese identificeert, met protocollen en asymptotisch gedrag enzo.
-Maar kunnen we het niet veel eenvoudiger houden?
-Ik ben er erg voor om het technisch zo simpel mogelijk te houden!
-Elke redelijke two-part code voor modelselectie moet onder wijziging van de enumeratie van TMs aan de volgende eigenschap voldoen:
-(a) de twee versies moeten equivalent zijn onder definitie 1, want anders is de maat namelijk arbitrair.
-Dit geldt voor zover ik begrijp niet voor de Koppel/Adriaans/Antunez versie van sophistication, maar wel voor die van Paul.
-Hiermee kunnen we dus leuk het eerste probleem aanduiden.
-Deze eigenschap zegt iets over de two part codelengtes voor de hele model class. Maar als we bovendien willen aannemen dat de lengte van het geselecteerde model iets zegt over de data,
-(b) dan moet definitie 1 bovendien gelden voor de specifieke r en s die worden geselecteerd! Hiervan kunnen we bewijzen dat het onmogelijk is.
-
-\item What is an ``acceptable enumeration'', is it really required in the standard definition of KC or is there a difference of opinion?
-Does this mean that it is impossible to do facticity in the standard definition of KC?
-That would be a nice claim to make!
-
-\item Is it possible to require that the enumeration of TM's is such that any TM can also efficiently store input data to the TM?
-That is, is there an acceptable numbering such that for all $i$, all $p$, there is a $j$ with $|j|\le|i|+|p|$ such that $T_j(\epsilon)=T_i(p)$.
-This enumeration is potentially still very simple, the programming language just has a provision for storing data inside the program.
-If there are acceptable enumerations with this property, and I suspect there are, then perhaps it could be reasonable to require this, and use it as a motivation to forget about the distinction between the index of a program and its input.
-Perhaps some authors silently assume the enumeration to have this property?
-This would make it valid to do what some authors, like Paul, are doing: to identify the two-part code \emph{within} the definition of KC, with the two-part code used to define meaningful information.
-Or am I now giving these authors too much credit?
-
-\item Why the ``up to log-terms''?
-Is it because of algorithmic mutual information holds up to log terms?
-Can we find references motivating the decision to work up to log terms?
-Is it possible to define a two part code where the length of both parts varies by only a constant as the ordering of Turing machines is changed?
-Why (not)?
-
 \item Paul argues that the randomness deficiency of a model is a measure of the amount of structure left in the noise part.
@@
 
-\item Discuss worries about overfitting in the literature, explain that actually the collapse problem is a case of severe \emph{underfitting} and that both can occur.
-Any attempts to penalize the model complexity in the two part code will exacerbate the underfitting problem.
-
-\item Ik denk dat we in ons eigen paper minder hoog van de toren moeten blazen dan in het stukje dat ik je laatst stuurde, misschien meer zoiets zeggen als "er is een verschil tussen het kortste programma dat een string produceert, die mogelijk veel langer is dan de string zelf als data niet efficient kan worden gerepresenteerd in de gebruikte programmeertaal, en de korste beschrijving van de string, die mag bestaan uit een programma plus de input voor dat programma.
-Dit verschil is cruciaal voor een juist begrip van de verschillende aanpakken voor meaningful information, maar in veel publicaties over het onderwerp wordt dit onderscheid niet expliciet gemaakt en is het dus niet duidelijk of de besproken complexiteitsmaten nu wel of niet afhangen van de gekozen universele Turingmachine."
-
-\item Paul-Pieter: model $M$ is eindige set, data $D$ is een subset ervan, dus meerdere binaire strings.
-Two part code is $K(M|d)+\log\binom{m}{d}$, met $d=|D|$ en $m=|M|$.
-Vraag, waarom zit $d$ in de conditional?
-Heeft dit te maken met andere keuzes over waar je de hoeveelheid noise stopt?
-
 \item ``Power and peril of MDL'': bespreken ook de ``collapse of the structure function'' als de model class te sterk is, dwz sets worden te efficient gecodeerd (wsch equivalent met Shen's enumerating complexity).
 
-\item Conjecture: to define valid two-part code inside KC, need efficient numbering: data can be inserted into TM index efficiently. In general, KC with admissible numbering will not work.
-
-\item Idea: two distinct models are impossible to separate if the probability they give to the data remains within a constant.
-This can happen even if they do not dominate each other for other data sequences.
-Sufficient condition: one dominates the other.
 \end{itemize}
 
-\begin{definition} An enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions is called \emph{$\gamma$-compact} if for all $i,y$ there is a $j$ such that $|j|\le|i y|+\gamma(|y|)$ and $\phi_i(y z)=\phi_j(z)$ for all $z$.
+\begin{definition} An enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions is called \emph{compact} if for all $i,y$ there is a $j$ such that $|j|\le|i y|$ and $\phi_i(y z)=\phi_j(z)$ for all $z$.
 \end{definition}
@@
 \begin{lemma}
-There is a $0$-compact acceptable numbering.
+There is a compact acceptable numbering.
 \end{lemma}
@@
 \begin{lemma}[Framework equivalence]
-Let $\phi_1,\phi_2,\ldots$ be a $\gamma$-compact effective enumeration. There is a constant $c$ such that for any index $i$ there is another index $j$ with $\phi_j(x)=\phi_i(x)$ for all $x$, and
-\[|j|\le K(i)+2\log(K(i))+\gamma(K(i)+2\log K(i)).\]
+Let $\phi_1,\phi_2,\ldots$ be a compact effective enumeration. There is a constant $c$ such that for any index $i$ there is another index $j$ with $\phi_j(x)=\phi_i(x)$ for all $x$, and
+\[|j|\le K(i)+2\log(K(i))+c.\]
 \end{lemma}
@@
 Then $\phi_i(x)=\phi_v(\bar k \bar p x)$.
-We now use $\gamma$-compactness to find an index $j$ with $\phi_v(\bar k \bar p x)=\phi_j(x)$ and length
-\[\begin{split}
-|j|&\le |v\bar k\bar p|+\gamma(\bar k\bar p)\\
-&\le|v|+|\bar k p|+2\log|\bar k p|+\gamma(|\bar k p|+2\log|\bar k p|)\\
-&=K(i)+|v|+2\log K(i)+\gamma(K(i)+2\log K(i)),
-\end{split}\]
+We now use compactness to find an index $j$ with $\phi_v(\bar k \bar p x)=\phi_j(x)$ and length
+\[|j|\le |v\bar k\bar p|\le|v|+|\bar k p|+2\log|\bar k p|=K(i)+|v|+2\log K(i),\]
 as required.
 \end{proof}
-\begin{corollary}
-If $\gamma(x)=O(\log x)$, then $|j|=K(i)+O(\log K(i))$.
-\end{corollary}
 
@@
 
-\begin{lemma}
-Let $\psi_1,\psi_2,\ldots$ be any acceptable numbering. There is a constant $c$ such that for any $i$ with $\psi_i(x)=f(x)$ for all $x$, the description length satisfies $K(f)\le |i|+c$.
+\begin{lemma}[TODO: $i$ moet prefix encoded zijn, anders klopt t niet. Ook beetje moeilijk te begrijpen]
+Let $\psi_1,\psi_2,\ldots$ be any acceptable numbering. There is a constant $c$ such that for any $i$ with $\psi_i(x)=f(x)$ for all $x$, the description length satisfies $K^\phi(f)\le |i|+c$.
 \end{lemma}
@@
 
-\begin{lemma}
+\begin{lemma}[TODO: is het niet leuker om te zeggen dat het altijd bounded is? Krijgen we dat kado van een later lemma?]
 There is an acceptable numbering $\psi_1,\psi_2,\ldots$ such that $F^\psi(x)=|\bar 1|$ for all $x$.
@@
 \end{proof}
-
-\begin{theorem}
-  There are acceptable numberings $\phi_1,\phi_2,\ldots$ and $\psi_1,\psi_2,\ldots$ such that for every constant $c$ there is a [are lots of] string $x$ with $|F^\phi(x)-F^\psi(x)|>c$ [regardless of penalties and possibly generalized to more model classes].
-\end{theorem}
-
-Conjectures:
-\begin{itemize}
-\item Als er een unbounded penalty zit op modelcomplexiteit dan is de facticity bounded door een constante.
-\end{itemize}
 
@@
 
-\begin{lemma}
+\begin{lemma}\label{lem:thecoolone}
   Let $\psi_1,\psi_2,\ldots$ be any acceptable enumeration of the partial recursive functions.
-  Let $\M$ be any model class and let $D:\{0,1\}^*\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e.
+  Let $\M$ be any model class, let $\X$ be any set of binary sequences and let $D:\{0,1\}^*\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e.
 if $f=\psi_{D(p)}$ then $p$ is a $D$-description of $f$.
@@
   \item $\forall_{f\in\M'}:\min\{|p|:\psi_{D(p)}=f\}\le K^\psi(f)+c$
-  \item $\forall_x:L^{\M'}(x)-L^{\M}(x)\le c$.
+  \item $\forall_{x\in\X}:L^{\M',\psi}(x)-L^{\M,\psi}(x)\le c$.
 \end{enumerate}
-Then is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that $S^\phi_{\M}(x) = 1$ for all $x$.
+Then is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
 \end{lemma}
@@
 \end{cases}\]
+We will show that under the $\phi$-enumeration, the best representation for $x$ using a model $f\in\M'$ is always better than the best representation using some $f\not\in\M'$.
+
 First suppose $f\in\M'$. Then
@@
 
-\hrule
+Now assume that the best model $f$ for $x$ is not in $\M'$. 
+Let $i$ be the index of $f$ with the shortest description, i.e. it achieves the minimum in
+\[
+K^\phi(f)=\min\{K^\phi(i):\phi_i=f\}.
+\]
+There are two possibilities. Either $i=0$, in which case we have $K^\phi(f)\ge r$ because $\phi_0$ cannot output zero and all other $\phi$-programs are at least $r$ bits long.
 
-Note that with this definition, we have $K^\phi(f) \leq K^\eta(f) \leq K^\psi(f) + c$ if $f \in \M'$.
-If $f$ is not in $\M'$, we have $K^\phi(f) = K^\psi(f) + r + 1 - c'$, by the same argument as used in the proof of Lemma~\ref{lemma:building-block}.
+Otherwise, we can bound
+\[\begin{split}
+K^\phi(f)&=K^\phi(i)=K^\phi(1^r0j)\\
+&\ge K^\phi(j)-c'\\
+&\ge K^\psi(j)+r+1-c'\\
+&\ge\min\{K^\psi(j):\phi_j=f\}+r+1-c'\\
+&=K^\psi(f)+r+1-c'.
+\end{split}
+\]
+Now choose $r=c''+\max\{K^\psi(\phi_0),c'-1\}$. Then substitution yields, for both cases, 
+$K^\phi(f)\ge c''+K^\psi(f)$.
 
-Let $x$ be some finite string, and let $(i, p)$ be the pair such that $\psi_i(p)$ witnesses $L^M_\psi(x)$.
-By assumption, $L^{\M'}_\psi \leq L^M(x) +c$.
-Thus, we have a pair $(j, q)$, with $\eta_j(q) = x$ and $K^\eta(j) + |q| \leq K^\phi(i) + |p| + c$.
-If we choose $r = c' + c+ 1$, we know that $\phi$ will have representations $(1^r0i, q)$ with code length $L^M_\psi(x) + c+ 1 $, and $(0j, q)$ with code length less than or equal too $L^{\M}_\psi(x) +c$.
-This ensures that the sophistication under $\phi$ will be $K^\phi(\eta_j) = K^\eta(\eta_j) + O(1) = F^{\M'}_\eta(j)$.
+Combining the inequalities above, for any model $g\not\in\M'$, there is a model $f\in\M$ such that
+\[\begin{split}
+K^\phi(g)+C^g(x) &\ge K^\psi(g)+C^g(x)+c''\\
+&\ge K^\psi(f)+C^f(x)-c+c''\\
+&\ge K^\phi(f)+C^f(x) -2c+c'' -|\bar0|.
+\end{split}\]
+
+By choosing $c''$ sufficiently large, we can ensure that the best representation is in $\M'$ for all $x\in\X$, which completes the proof.
 \end{proof}
@@
 \begin{lemma}
-  Let $\M$ be any model class and let $\M'\subseteq \M$ be a sub-model with the following properties:
-\begin{enumerate}
-  \item There exists a (non-optimal) computable numbering $\psi_1, \psi_2, \ldots$ for $M'$,
-  \item $L^{\M'}(x)-L^{\M}(x)\le c$ for some constant $c$, all $x$.
-\end{enumerate}
-Then there is an enumeration $\phi_1,\phi_2,\ldots$ such that $|F^{\psi}_{\M'} - \leq F^{\phi}_{\M}| \leq c$ for infinitely many $x$.
+Let $\M$ be a prefix model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$.
+Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the prefix partial recursive functions, and a constant $c$, such that
+\[
+K(x)-S^{\M,\phi}(x)\le c
+\]
+for all $x\in\X$.
 \end{lemma}
 \begin{proof}
+Let $\psi_1,\psi_2,\ldots$ be any default enumeration of the partial recursive prefix functions.
+Note that since $f$ is a prefix function, if $f$ is defined for input $\epsilon$ then it cannot be defined for any other input.
+Pick $f,x$ with $f(\epsilon)=x$.
+Note that $x$ can be computed from $f$ and a fixed program, so there is a $c$ such that $K(x)\le K(f)+c$.
+Vice versa, given any $x$ we can construct an index of $f$, since $\psi$ is an acceptable numbering.
+Therefore $|K(f)-K(x)|\le c$.
 
+We now define a computable function $D$ by $D(\bar\imath p)=j$ where $\psi_j(\epsilon) = \psi_i(p)$.
+We will show that the two conditions of Lemma~\ref{lem:thecoolone} hold for the prefix function $D$.
+
+\begin{enumerate}
+\item Let $f$ be any function in the range of $D$, and $x$ its output. Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\psi_i(q)=x\}=K(x)\le K(f)+c$.
+\item On the one hand $L^{\M',\psi}(x)\le K(f)+|\epsilon|\le K(x)+c$. On the other hand, $L^{\M,\psi}(x)$ is an effective description of $x$, so $K(x)$ is at most a constant larger.
+Together, these inequalities establish the second condition.
+\end{enumerate}
+
+Then by Lemma~\ref{lem:thecoolone} there is an enumeration $\phi_1,\phi_2,\ldots$ such that $S^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
+We observed that $|K(f)-K(x)|\le c$ for any $f\in\M'$, so $S^{\M',\psi}(x)\ge K(x)+c$.
+This proves the lemma.
 \end{proof}
 
-\hrule
-\begin{lemma}
-  Let $\M$ be a model class with the property that there is a constant $c$ and a computable function $g$ such that for all $x$ we have $\gamma(x\mid g(x))\le c$.
-  For such model classes, there are acceptable enumerations $\phi_1,\phi_2,\ldots$ of the partial recursive functions for which $K^{\M,\phi}(x)-F^{\M,\phi}(x)$ is bounded.
-\end{lemma}
-\begin{proof}
-
-\end{proof}
-
-\begin{lemma}
-  Let $\M$ be a model class with the property that there is a constant $c$ and a model $f$ such that for all $x$ the counterevidence satisfies $\gamma(x\mid f)\le c$.
-  For such model classes, there are acceptable enumerations $\phi_1,\phi_2,\ldots$ of the partial recursive functions for which $F^{\M,\phi}(x)$ is bounded.
-\end{lemma}
-
-\begin{definition}
-  Given a model class $\M$, a representation $(f,y)$ with $f\in\M$ is said to be \emph{$c$-optimal} if
-\[
-K(f)+|y|\le \min_{f',y':f'(y')=f(y)}\left(K(f')+|y'|\right)+c.
-\]
-\end{definition}
-
-\begin{lemma}
-  Let $\M$ be a model class with the property that there are $c,y$ such that for all $x$ there is an $f$ such that $f(y)=x$ and $(f,y)$ is $c$-optimal.
-  For such model classes, there are acceptable enumerations $\phi_1,\phi_2,\ldots$ of the partial recursive functions for which $K^{\M,\phi}(x)-F^{\M,\phi}(x)$ is bounded.
-\end{lemma}
-
-\begin{lemma}
-  Let $\M$ be a model class with the property that there are $c,f$ such that for all $x$ there is an $y$ such that $f(y)=x$ and $(f,y)$ is $c$-optimal.
-For such model classes, there are acceptable enumerations $\phi_1,\phi_2,\ldots$ of the partial recursive functions for which $F^{\M,\phi}(x)$ is bounded.
-\end{lemma}
-
-\begin{definition}
-We call a version of Sophistication $F^\phi$ \emph{bounded} if its ad-hoc choices (the enumeration $\phi$) can be chosen so that
-\[
-	\exists c \forall x: F^\phi(x) \leq c \p
-\]
-Similarly, we call $F^\phi(x)$ \emph{top-heavy} if there exists a $\phi_1, \phi_2, \ldots$ such that
-\[
-	\exists c \forall x: F^\phi(x) \geq K^F(x) - c 
-\]
-Where $K^F$ is the two-part coding minimized within $F$.
-\end{definition}
-
-If some $F$ is either top-heavy, or unbounded we can show that it must either lack invariance, or be useless as a measure of model complexity.
-If the property proved holds for all enumerations $\phi$, and not just a single one, the sophistication is either bounded, or the same as the Kolmogorov complexity.
-If the measure is not bounded or top-heavy for other enumerations, the difference between the two is unbounded, and $F$ is not invariant.
-
-One of the earliest proposals to separate structure and content was the Kolmogorov structure function \cite{TODO}. The following is the version of sophistication it leads to. 
-\begin{definition}
-\[
-	S^\phi(x) = K(S^*) \,\tn{where }\, (S, i) = \argmin_{(S, i)\,:\,\text{$S$ is a finite set},\,S_i = x} K(S) + |i|
-\]
-\end{definition}
-
-\begin{theorem}
-There exist acceptable numberings $\phi_1, \phi_2, \ldots$ for which $S^\phi$ is top-heavy.
-\end{theorem}
-\begin{proof}
-Let $\psi_1, \psi_2, \ldots$ be some acceptable numbering. Define $\psi$ as 
-\begin{align*}
-\phi_{0\bar\imath p}(\epsilon) &= \{\psi_i(p)\} \\
-\phi_{1ri}(p) &= \psi_i(p)\\ 
-\phi_j(x) = \infty \,\text{if $j$ cannot be decoded as above}
-\end{align*}
-For any $x$ with $\psi$-description $(i, p)$, that is is $\psi_i(p) = x$, we have two representations in $\phi$.
-Firstly, we have $\{x\}, 1$ with description length $K^\psi(x) + 2$.
-Secondly, we have a representation in $K^\psi{x} + r$.
-By choosing $r$ sufficiently large, the first representation will always win.
-\end{proof}
+Plan: een lemma voor total/niet prefix dat laat zien mbv pigeonhole dat facticity juist bounded of juist unbounded kan zijn.
+De ene kant op gaat dat met de print functie (klein model, grote data).
+De andere kant op is moeilijk
 

### 2014-11-13 Facticity/Facticity/paper.tex
poep

@@
 \newcommand{\M}{\mathcal M}
+\newcommand{\C}{\mathcal C}
+\newcommand{\T}{\mathcal T}
+\newcommand{\F}{\mathcal F}
+\renewcommand{\P}{\mathcal P}
 \newcommand{\K}{\mathcal K}
 \newcommand{\X}{\mathcal X}
-\newcommand{\C}{\mathcal C}
 \newcommand{\B}{\mathbb B}
@@
 \end{abstract}
+
+\section{Definitions}
+
+\begin{definition}[Models and model classes]
+  A \emph{model} is a function $f:\{0,1\}^*\to\{0,1\}^*$. A
+  \emph{model class} is a set of models. We distinguish four important
+  model classes:
+  \begin{itemize}
+  \item $\C$ is the set of all partial recursive functions,
+  \item $\K\subset\C$ is the set of all partial recursive prefix
+    functions,
+  \item $\T\subset\C$ is the set of all total recursive functions,
+  \item $\P\subset\K$ is the set of all prefix functions such that for
+    every infinite binary string the function is defined for a
+    prefix. This is the natural analogue of totality in the prefix
+  \item $\F\subset\P$ contains, for each finite set $S$, a computable
+    surjective function maping the binary sequences of length
+    $\lceil\log S\rceil$ onto $S$.
+  \end{itemize}
+\end{definition}
+
+\begin{definition}[Enumerations]
+  Let $\phi_1,\phi_2,\ldots$ be a fixed acceptable enumeration of all partial
+  recursive functions, e.g. the enumeration described in \cite{TODO}.  
+\end{definition}
+
+\begin{definition}[Representations]
+  A pair $(m,y)$ is a \emph{representation} for $x$ if $m$ is a model
+  and $m(y)=x$. Given a model class $\M$, define the set of all
+  representations for $x$:
+  \[
+  R^\M(x):=\{(m,y):m\in\M, m(y)=x\}.
+  \]
+\end{definition}
+
+\begin{definition}[Code length functions]
+Define the following code length functions for the representations:
+\begin{itemize}
+\item $L(m)$ is the length of a self-delimiting description of
+  the model $m$, i.e. it is the length function of  a prefix code,
+\item $L^m(x)=L(m)+\min\{|y|:m(y)=x\}$ measures the number of bits required to describe both the model $m$ and
+  the data $x$ using the model,
+\item $L^\M(x)=\min\{L^m(x):m\in\M\}$ measures the number of bits
+  required to describe $x$ using the best model in the model class.
+\end{itemize}
+\end{definition}
+
+\begin{definition}[Candidate set]
+  The \emph{candidate models} for $x$ from model class $\M$ are
+  \[
+  C^\M_c(x):=\{m\in\M:L^m(x)\le L^\M(x)+c\}.
+  \]
+\end{definition}
+
+\begin{definition}[Sophistication]
+  The model complexity, sophistication, or facticity, of a string $x$
+  given a model class $\M$ and a constant $c$ is defined by
+  \[
+  S^\M_c(x):=\min\{L(m):m\in C^\M_c(x)\}.
+  \]
+\end{definition}
+
+\begin{definition}[Prefix encoding function]
+  A prefix encoding function $f:\{0,1\}^*\to\{0,1\}^*$ is an injective
+  function that maps the binary strings to a prefix-free set. There
+  are prefix encoding functions $g$ with the property that
+  $|g(x)|\le|x|+2\log(|x|+2)|$ \cite{TODO}. We fix such a function and
+  denote it by $\bar x=g(x)$.
+\end{definition}
+
+\paragraph{Kolmogorov's Structure Function}
+
+\begin{itemize}
+\item Model class is $\F$
+\item Description method for models $L(m)=K(S)$ where $S$ is the range
+  of $m$. There is some ambiguity as to the definition of prefix
+  complexity for finite sets; discuss and refer to Shen \cite{TODO}.
+\end{itemize}
+
+\paragraph{Koppel Sophistication}
+
+\begin{itemize}
+\item Model class is $\T$
+\item Description method for models is required to be prefix-free, but
+  can otherwise be chosen arbitrarily.
+\item Koppel uses monotone complexity and discusses infinite strings,
+  TODO figure out if and how this affects our results
+\end{itemize}
+
+\paragraph{Vit\'anyi's Sophistication / Meaningful Information}
+
+\begin{itemize}
+\item Model class is $\T$
+\item Description method for models is $L(m)=K(m)$
+\item Candidate set is defined differently: $C_c=\{m\in\T:|L^m(x)-K(x)|\le
+  c\}$. There appears to be a problem with this definition, because
+  the two-part descriptions of $x$ may be shorter than $K(x)$ by
+  exploiting the fact that the models do not need to be prefix functions.
+\end{itemize}
+
+\paragraph{Antunes's Coarse Sophistication}
+
+\begin{itemize}
+\item Model class is $\T$
+\item Description method for models is unclear, but appears to be
+  taken from Koppel, i.e. arbitrary but self-delimiting.
+\item Candidate set is defined differently:
+  $C^\T(x)=\{m^*\}$ where $m^*=\argmin_m L(m)+L^m(x)-C(x)$. Thus, there is no slack
+  variable and $C^\T$ contains only a single model.
+\end{itemize}
+
+\paragraph{Adriaans' Facticity}
+
+\begin{itemize}
+\item Model class is $\C$
+\item Description method for models is $L(m)=|\bar\imath|$ where
+  $\phi_i=m$. The index function is assumed to be \emph{faithful},
+  i.e. there is a constant $c$ such that $|i|-C(i)\le c|$ for all $i$.
+\item Candidate set is $C^\C_0$.
+\item Facticity is actually defined as index length without prefix
+  encoding:  $S^\C_0(x):=\min\{|i|:\phi_i\in C^\C_0(x)\}$. But this
+  difference is not important for the discussion.
+\end{itemize}
 
@@
 \end{proof}
-
-\[
-b(i_0)=\min_{i\ge i_0}(L(i)-K(i))
-\]
 

### 2014-11-24 Facticity/Facticity/paper.tex

@@
 \end{quotation}
+
+Adriaans takes this problem seriously and attempts to address it by imposing additional restrictions on the ordering of TMs implied by the used UTM \cite{adriaans2012facticity}.
+We are skeptical that such measures can effectively eliminate this problem without resorting to orderings of the TMs that are not admissible [TODO ref].
 

### 2014-11-24 Facticity/Facticity/paper.tex
reshuffle

--- 2014-11-13 Facticity/Facticity/paper.tex
@@
 
+All numberings should be acceptable!
+
 \begin{definition}[Models and model classes]
@@
 
-\begin{definition}[Enumerations]
+\begin{definition}[Prefix encoding function]
+  A prefix encoding function $f:\{0,1\}^*\to\{0,1\}^*$ is an injective
+  function that maps the binary strings to a prefix-free set. There
+  are prefix encoding functions $g$ with the property that
+  $|g(x)|\le|x|+2\log(|x|+2)|$ \cite{TODO}. We fix such a function and
+  denote it by $\bar x=g(x)$.
+\end{definition}
+
+\begin{definition}[Complexity]
   Let $\phi_1,\phi_2,\ldots$ be a fixed acceptable enumeration of all partial
   recursive functions, e.g. the enumeration described in \cite{TODO}.
+  Let $\Pi$ be a computable subset of the natural numbers such that
+  $\{\phi_i:i\in\Pi\}$ contains all computable prefix functions. Now
+  define Kolmogorov complexity by
+\begin{align*}
+C(x)&=\min\{|\bar\imath| y:\phi_i(y)=x\},\\
+K(x)&=\min\{|\bar\imath| y:\phi_i(y)=x, i\in\Pi\}.\\
+\end{align*}
 \end{definition}
+This definition of $K$ is equivalent to the definition in the
+textbook~\cite{TODO}, but it is more convenient for defining the
+$K$-complexity of non-prefix functions:
 
@@
   A pair $(m,y)$ is a \emph{representation} for $x$ if $m$ is a model
-  and $m(y)=x$. Given a model class $\M$, define the set of all
-  representations for $x$:
-  \[
-  R^\M(x):=\{(m,y):m\in\M, m(y)=x\}.
-  \]
+  and $m(y)=x$.
 \end{definition}
@@
 \begin{definition}[Candidate set]
-  The \emph{candidate models} for $x$ from model class $\M$ are
+  The \emph{candidate models} for $x$ from model class $\M$ and
+  threshold function $\tau$ are
   \[
-  C^\M_c(x):=\{m\in\M:L^m(x)\le L^\M(x)+c\}.
+  C^\M_\tau(x):=\{m\in\M:L^m(x)\le \tau(x)\}.
   \]
@@
   The model complexity, sophistication, or facticity, of a string $x$
-  given a model class $\M$ and a constant $c$ is defined by
+  given a model class $\M$ and threshold function $\tau$ is defined by
   \[
-  S^\M_c(x):=\min\{L(m):m\in C^\M_c(x)\}.
+  S^\M_\tau(x):=\min\{L(m):m\in C^\M_\tau(x)\}.
   \]
@@
 
-\begin{definition}[Prefix encoding function]
-  A prefix encoding function $f:\{0,1\}^*\to\{0,1\}^*$ is an injective
-  function that maps the binary strings to a prefix-free set. There
-  are prefix encoding functions $g$ with the property that
-  $|g(x)|\le|x|+2\log(|x|+2)|$ \cite{TODO}. We fix such a function and
-  denote it by $\bar x=g(x)$.
+\section{Measures of meaningful information}
+
+\subsection{Koppel Sophistication}
+
+\begin{itemize}
+\item Model class is $\T$
+\item Threshold function $\tau(x)=L^\M(x)+c$
+\item Description method for models is required to be prefix-free, but
+  can otherwise be chosen arbitrarily.
+\item Koppel uses monotone complexity and discusses infinite strings,
+  TODO figure out if and how this affects our results
+\end{itemize}
+
+This does not work because of the nickname problem! To fix this:
+
+\begin{definition}
+  The complexity of a partial recursive function is defined by
+  \begin{align*}
+    C(f) &:= \min\{C(i):\phi_i=f\}\\
+    K(f) &:= \min\{K(i):\phi_i=f\}.
+  \end{align*}
 \end{definition}
 
-\paragraph{Kolmogorov's Structure Function}
+These definitions make sense because:
+
+\begin{lemma}[Invariance]
+Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be two acceptable numberings.
+There exists a constant $c$ such that $\left| K^\phi(f) - K^\psi(f)\right | \leq c$ for all $f$.
+\label{lemma:invariance}
+\end{lemma}
+\begin{proof}
+Let $g(i)$ be the function which which translates from $\psi$ to $\phi$.
+\begin{align*}
+K^\phi(f) &= \min\left\{ K^\phi(i) : \phi_i= f\right\} \\
+&\geq \min\left\{ K^\psi(i) : \phi_i= f\right\} - c\\
+&= \min\left\{ K^\psi(i) : \psi_{g(i)}= f\right\} - c\\
+&\geq \min\left\{ K^\psi(i) : \psi_i= f\right\} - c\\
+&= K^\psi(f).
+\end{align*}
+We can reverse $\phi$ and $\psi$ without loss of generality to achieve the same inequality the other way around, completing the proof.
+\end{proof}
+
+Thus, the complexity of a function depends on the choice of
+enumeration by no more than a constant term. Note that this implies
+that for any enumeration $\psi_1,\psi_2,\ldots$, the complexity $K(\psi_i)$
+is less than the literal description length of the index
+$|\bar\imath|$ up to a constant. For some numberings, there is a gap
+between these two description methods; in those cases, the index is an
+inefficient representation of the corresponding function. However it
+is possible to construct numberings that are \emph{faithful} in the
+sense that no such gap exists:
+
+\begin{definition}[Faithful Numbering]\label{def:faithful}
+  A numbering $\psi_1,\psi_2,\ldots$ of the partial recursive
+  functions is \emph{faithful} if there is a constant $c$ such that
+  for all indices $i$ there is a $j$ such that $\psi_i=\psi_j$ and
+  $|j|\le C(\psi_j)+c$.
+\end{definition}
+
+\begin{lemma}
+  There is a faithful acceptable numbering.
+\end{lemma}
+\begin{proof}
+Let $i_\tn{div}(y)$ be an index such that $\phi_{i_\tn{div}}(y)=\infty$ for all $y$. Define
+  \[\psi_q=\begin{cases}
+    \phi_{\phi_i(p)}&\tn{if $q$ can be written as $\bar\imath p$ and $\phi_i(p)<\infty$,}\\
+    \phi_{i_\tn{div}}&\tn{otherwise.}\end{cases}
+  \]
+  To show that $\psi$ is faithful, pick any function $f$. Then
+\[\begin{split}
+C(f)&=\min\{C(i):\phi_i=f\}\\
+&=\min\{\min\{|\bar a b|:\phi_a(b)=i\}:\phi_i=f\}\\
+&=\min\{|\bar a b|:\phi_{\phi_a(b)}=f\}\\
+&=\min\{|\bar a b|:\psi_{\bar a b}=f\}.
+\end{split}\]
+This shows there is a sufficiently small $\psi$ index.
+
+To show that $\psi$ is acceptable, let $\phi_j$ denote the identity
+function. Then a $\phi$-index $i$ can be mapped to a $\psi$-index
+using the computable function $r(i)=\bar\jmath i$, so that
+$\psi_{r(i)}(y)=\psi_{\bar\jmath i}(y)=\phi_i(y)$. For the reverse,
+define $\phi_v(\bar\imath p, y)=\phi_{\phi_i(p)}(y)$. For fixed
+$\bar\imath p$, the 
+$s^n_m$-theorem (see~\cite{TODO}) states that we can compute the $h$
+such that $\phi_h(y)=\phi_v(\bar\imath p,y)$. Let $h(\bar\imath p)$
+denote this index as a function of the program; further define
+$h(q)=i_\tn{div}$ if $q$ cannot be expressed as $\bar\imath p$. By
+construction $h$ is total and computable. To check that the mapping
+returns the correct function, rewrite $\phi_{h(\bar\imath
+  p)}(y)=\phi_v(\bar\imath p,y)=\phi_{\phi_i(p)}(y)=\psi_{\bar\imath p}(y)$.
+\end{proof}
+
+\subsection{Antunes's Coarse Sophistication}
+
+\begin{itemize}
+\item Model class is $\T$
+\item Description method $L(m)$ for models is unclear, but appears to be
+  taken from Koppel, i.e. arbitrary but self-delimiting.
+\item Sophistication is defined slightly differently: $S^\T(x)=L(m^*)$ where $m^*=\argmin_m L(m)+L^m(x)-C(x)$. Thus, there is no slack.
+\end{itemize}
+
+\subsection{Adriaans' Facticity}
+
+\begin{itemize}
+\item Model class is $\C$
+\item Description method for models is $L(m)=|\bar\imath|$ where
+  $\phi_i=m$. The index function is assumed to be \emph{faithful}, see
+  Definition~\ref{def:faithful}
+\item Threshold function is $\tau(x)=L^m(x)$.
+\item Facticity is actually defined as index length without prefix
+  encoding:  $S^\C_0(x):=\min\{|i|:\phi_i\in C^\C_0(x)\}$. But this
+  difference is not important for the discussion.
+\end{itemize}
+
+\begin{lemma}
+  If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then facticity is bounded.
+\end{lemma}
+\begin{proof}
+Let $\bar\imath p$ be any two-part representation for the data $x$, i.e.
+$\phi_i(p)=x$.
+Then construct an alternative two-part representation $\bar vi^* p$, where $i^*$ is the shortest $\psi_u$-program for $i$ such that $\phi_v(i^* p)=\phi_{\psi_u(i^*)}(p) = \phi_i(p)=x$.
+We compare the lengths of these two representations.
+Note that $|i^*|=K^\psi(i)$.
+Therefore,
+\[
+|\bar\imath p|-|\bar v i^* p| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-K(i).
+\]
+By assumption there must be an $i_0$ such that the above expression is positive for all $i>i_0$.
+From this $i$ onwards, the second representation (using $v$) will have a shorter code length, so $\bar\imath p$ cannot achieve the minimum in the definition of the facticity.
+Consequently, the facticity is bounded by $F^\phi(x)<|\overline{\imath_0}|$ for all $x$.
+\end{proof}
+Note that the condition for this lemma applies to all known partial recursive prefix functions:
+
+\begin{conjecture}
+For any partial recursive prefix code length function $L$, the function
+\[
+\min\{L(i)-K(i):i\ge i_0\}
+\]
+is unbounded in $i_0$.
+\end{conjecture}
+
+\subsection{Roll your own}
+
+\begin{itemize}
+\item Model class is either $\C$ or $\K$
+\item Description method for models is $L(m)=K(m)$.
+\item Threshold function is $\tau(x)=L^m(x)+c$, which matches $C(x)$
+  resp. $K(x)$ up to a constant.
+\end{itemize}
+
+\begin{lemma}\label{lem:thecoolone}
+  Let $\psi_1,\psi_2,\ldots$ be any acceptable enumeration of the partial recursive functions.
+  Let $\M$ be any model class, let $\X$ be any set of binary sequences and let $D:\{0,1\}^*\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e.
+if $f=\psi_{D(p)}$ then $p$ is a $D$-description of $f$.
+Let $\M'=\{\psi_i:i\in\tn{range}(D)\}$.
+Further assume there is a constant $c$ such that
+\begin{enumerate}
+  \item $\forall_{f\in\M'}:\min\{|p|:\psi_{D(p)}=f\}\le K^\psi(f)+c$
+  \item $\forall_{x\in\X}:L^{\M',\psi}(x)-L^{\M,\psi}(x)\le c$.
+\end{enumerate}
+Then is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
+\end{lemma}
+\begin{proof}
+We define the numbering $\phi$ as follows:
+\[\begin{cases}
+\phi_0(p) = 1^r 0 D(p) \\
+\phi_{1^r0i}(p) = \psi_i(p) \\
+\phi_j(\cdot) = \infty &\text{if $j$ contains no zeroes.}
+\end{cases}\]
+We will show that under the $\phi$-enumeration, the best representation for $x$ using a model $f\in\M'$ is always better than the best representation using some $f\not\in\M'$.
+
+First suppose $f\in\M'$. Then
+\[\begin{split}
+K^\phi(f)&=\min\{K^\phi(i):\phi_i=f\}\\
+&=\min\{|\bar\jmath q|:\phi_j(q)=i, \phi_i=f\}\\
+&=\min\{|\bar\jmath q|:\phi_{\phi_j(q)}=f\}\\
+&\le\min\{|\bar0 q|:\phi_{\phi_0(q)}=f\}~\text{(using $j=0$)}\\
+&=\min\{|\bar0 q|:\phi_{1^r0 D(q)}=f\}\\
+&=|\bar 0|+\min\{|q|:\psi_{D(q)}=f\}\\
+&\le K^\psi(f)+c+|\bar 0|,
+\end{split}
+\]
+where the last inequality uses the first assumption.
+
+Now assume that the best model $f$ for $x$ is not in $\M'$. 
+Let $i$ be the index of $f$ with the shortest description, i.e. it achieves the minimum in
+\[
+K^\phi(f)=\min\{K^\phi(i):\phi_i=f\}.
+\]
+There are two possibilities. Either $i=0$, in which case we have $K^\phi(f)\ge r$ because $\phi_0$ cannot output zero and all other $\phi$-programs are at least $r$ bits long.
+
+Otherwise, we can bound
+\[\begin{split}
+K^\phi(f)&=K^\phi(i)=K^\phi(1^r0j)\\
+&\ge K^\phi(j)-c'\\
+&\ge K^\psi(j)+r+1-c'\\
+&\ge\min\{K^\psi(j):\phi_j=f\}+r+1-c'\\
+&=K^\psi(f)+r+1-c'.
+\end{split}
+\]
+Now choose $r=c''+\max\{K^\psi(\phi_0),c'-1\}$. Then substitution yields, for both cases, 
+$K^\phi(f)\ge c''+K^\psi(f)$.
+
+Combining the inequalities above, for any model $g\not\in\M'$, there is a model $f\in\M$ such that
+\[\begin{split}
+K^\phi(g)+C^g(x) &\ge K^\psi(g)+C^g(x)+c''\\
+&\ge K^\psi(f)+C^f(x)-c+c''\\
+&\ge K^\phi(f)+C^f(x) -2c+c'' -|\bar0|.
+\end{split}\]
+
+By choosing $c''$ sufficiently large, we can ensure that the best representation is in $\M'$ for all $x\in\X$, which completes the proof.
+\end{proof}
+
+\paragraph{TODO: instantiate for UTM}
+
+\begin{itemize}
+\item Model class is $\K$
+\item Description method for models is $L(m)=K(m)$.
+\item Threshold function is $\tau(x)=L^m(x)+c$, which matches $K(x)$
+  up to a constant.
+\end{itemize}
+
+\begin{lemma}
+Let $\M$ be a prefix model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$.
+Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the prefix partial recursive functions, and a constant $c$, such that
+\[
+K(x)-S^{\M,\phi}(x)\le c
+\]
+for all $x\in\X$.
+\end{lemma}
+\begin{proof}
+Let $\psi_1,\psi_2,\ldots$ be any default enumeration of the partial recursive prefix functions.
+Note that since $f$ is a prefix function, if $f$ is defined for input $\epsilon$ then it cannot be defined for any other input.
+Pick $f,x$ with $f(\epsilon)=x$.
+Note that $x$ can be computed from $f$ and a fixed program, so there is a $c$ such that $K(x)\le K(f)+c$.
+Vice versa, given any $x$ we can construct an index of $f$, since $\psi$ is an acceptable numbering.
+Therefore $|K(f)-K(x)|\le c$.
+
+We now define a computable function $D$ by $D(\bar\imath p)=j$ where $\psi_j(\epsilon) = \psi_i(p)$.
+We will show that the two conditions of Lemma~\ref{lem:thecoolone} hold for the prefix function $D$.
+
+\begin{enumerate}
+\item Let $f$ be any function in the range of $D$, and $x$ its output. Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\psi_i(q)=x\}=K(x)\le K(f)+c$.
+\item On the one hand $L^{\M',\psi}(x)\le K(f)+|\epsilon|\le K(x)+c$. On the other hand, $L^{\M,\psi}(x)$ is an effective description of $x$, so $K(x)$ is at most a constant larger.
+Together, these inequalities establish the second condition.
+\end{enumerate}
+
+Then by Lemma~\ref{lem:thecoolone} there is an enumeration $\phi_1,\phi_2,\ldots$ such that $S^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
+We observed that $|K(f)-K(x)|\le c$ for any $f\in\M'$, so $S^{\M',\psi}(x)\ge K(x)+c$.
+This proves the lemma.
+\end{proof}
+
+\subsection{Kolmogorov's Structure Function}
 
@@
 
-\paragraph{Koppel Sophistication}
-
-\begin{itemize}
-\item Model class is $\T$
-\item Description method for models is required to be prefix-free, but
-  can otherwise be chosen arbitrarily.
-\item Koppel uses monotone complexity and discusses infinite strings,
-  TODO figure out if and how this affects our results
-\end{itemize}
-
-\paragraph{Vit\'anyi's Sophistication / Meaningful Information}
+\subsection{Vit\'anyi's Sophistication / Meaningful Information}
 
@@
 
-\paragraph{Antunes's Coarse Sophistication}
-
-\begin{itemize}
-\item Model class is $\T$
-\item Description method for models is unclear, but appears to be
-  taken from Koppel, i.e. arbitrary but self-delimiting.
-\item Candidate set is defined differently:
-  $C^\T(x)=\{m^*\}$ where $m^*=\argmin_m L(m)+L^m(x)-C(x)$. Thus, there is no slack
-  variable and $C^\T$ contains only a single model.
-\end{itemize}
-
-\paragraph{Adriaans' Facticity}
-
-\begin{itemize}
-\item Model class is $\C$
-\item Description method for models is $L(m)=|\bar\imath|$ where
-  $\phi_i=m$. The index function is assumed to be \emph{faithful},
-  i.e. there is a constant $c$ such that $|i|-C(i)\le c|$ for all $i$.
-\item Candidate set is $C^\C_0$.
-\item Facticity is actually defined as index length without prefix
-  encoding:  $S^\C_0(x):=\min\{|i|:\phi_i\in C^\C_0(x)\}$. But this
-  difference is not important for the discussion.
-\end{itemize}
+% stru fu: model class is all finite sets
+% code length L(m)+L_m(x)
+% acceptability criterion acc(m) = 2L(m)+L_m(x)-C(x)
+%
 
@@
 \begin{itemize}
-\item Paul argues that the randomness deficiency of a model is a measure of the amount of structure left in the noise part.
+\item Paul argues bleurg that the randomness deficiency of a model is a measure of the amount of structure left in the noise part.
 I feel that it is also a measure of the amount of counterevidence against a model in the statistical sense, as follows: the randomness deficiency of $x$ given model $M$ is $\delta(x|M)=L_M(x)-K(x|M)$.
@@
 
-\begin{definition} An enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions is called \emph{compact} if for all $i,y$ there is a $j$ such that $|j|\le|i y|$ and $\phi_i(y z)=\phi_j(z)$ for all $z$.
-\end{definition}
-
-\begin{lemma}
-There is a compact acceptable numbering.
-\end{lemma}
-\begin{proof}
-  Let $\phi_1,\phi_2,\ldots$ be any acceptable numbering.
-Let $\psi_1,\psi_2,\ldots$ be a new acceptable numbering defined by $\psi_j(x)=\phi_i(y x)$ if $j$ is a pair satisfying $j=\bar\imath y$, and $\psi_j(x)=\infty$ otherwise.
-We will show that $\psi$ is both $0$-compact and acceptable.
-
-To show compactness, pick any index $i$ and binary string $y$.
-First assume $i$ is a valid pair $i=\bar a b$.
-Then  $\psi_i(y x) = \psi_{\bar a b}(y x)=\phi_a(b y x)$, but also $\psi_{\bar a b y}(x)=\phi_a(b y x)$, so $\psi_i(y x) = \psi_j(x)$ for $j=\bar a b y$; so $|j|=|iy|+0$.
-On the other hand, if $i$ is not a valid pair, we can simply use $j=i$ so that $\psi_i(y x)=\psi_j(x)=\infty$ for all $x$.
-
-To show that $\psi_1,\psi_2,\ldots$ is acceptable, we define total computable functions $f$ and $g$ such that $\psi_i(x)=\phi_{f(i)}(x)$ and $\phi_j(x)=\psi_{g(j)}(x)$.
-The easy direction is $g(j)=\bar\jmath\epsilon$.
-The definition of the mapping in the other direction $f(i)$ depends on whether $i$ is a valid pair.
-If it is not, then $f(i)=h$ for some $h$ such that $\phi_h(x)=\infty$ for all $x$.
-On the other hand if $i=\bar\jmath y$, then $f(i)$ is the index of a partial recursive function $\phi_{f(i)}(x)$ defined by a Turing machine that simulates the machine for $\phi_j$ but with $yx$ on the input tape.
-\end{proof}
-
-\begin{lemma}[Framework equivalence]
-Let $\phi_1,\phi_2,\ldots$ be a compact effective enumeration. There is a constant $c$ such that for any index $i$ there is another index $j$ with $\phi_j(x)=\phi_i(x)$ for all $x$, and
-\[|j|\le K(i)+2\log(K(i))+c.\]
-\end{lemma}
-\begin{proof}
-First, let $v$ be the index of the partial recursive function given by $\phi_v(\bar k\bar p x)=\phi_{\phi_k(p)}(x)$.
-Let $\bar k p$ be a shortest program for $i$, that is $|\bar k p| = K(i)$ and $\phi_k(p)=i$.
-Then $\phi_i(x)=\phi_v(\bar k \bar p x)$.
-We now use compactness to find an index $j$ with $\phi_v(\bar k \bar p x)=\phi_j(x)$ and length
-\[|j|\le |v\bar k\bar p|\le|v|+|\bar k p|+2\log|\bar k p|=K(i)+|v|+2\log K(i),\]
-as required.
-\end{proof}
-
-We need a definition of $K(f)$. We adopt the definition given in \cite{TODO}:
-
-\begin{definition}
-  The Kolmogorov complexity of a partial recursive function is defined as
-  \[
-  K(f) = \min\{K(i):\forall_x \phi_i(x)=f(x).\}
-  \]
-\end{definition}
-
-It is not immediately clear that this is a reasonable definition of the shortest effective description of a function, but fortunately this is indeed the case:
-
-\begin{lemma}[TODO: $i$ moet prefix encoded zijn, anders klopt t niet. Ook beetje moeilijk te begrijpen]
-Let $\psi_1,\psi_2,\ldots$ be any acceptable numbering. There is a constant $c$ such that for any $i$ with $\psi_i(x)=f(x)$ for all $x$, the description length satisfies $K^\phi(f)\le |i|+c$.
-\end{lemma}
-\begin{proof}
-The numbering is acceptable, so there is a total computable function $\phi_r$ that maps indices in the $\psi$-enumeration to indices in the $\phi$-enumeration.
-Let $j=\phi_r(i)$.
-Then $\bar r i$ is a program for $j$, so $K(j)\le |\bar r i| = |i|+c$.
-Now,
-\[
-K(f)=\min\{K(k):\forall_x\phi_k(x)=f(x)\}\le K(j)\le|i|+c
-\]
-as required.
-\end{proof}
-
-\begin{lemma}[Invariance]
-Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be two acceptable numberings.
-There exists a constant $c$ such that $\left| K^\phi(f) - K^\psi(f)\right | \leq c$ for all $f$.
-\label{lemma:invariance}
-\end{lemma}
-\begin{proof}
-Let $g(i)$ be the function which which translates from $\psi$ to $\phi$.
-\begin{align*}
-K^\phi(f) &= \min\left\{ K^\phi(i) : \phi_i= f\right\} \\
-&\geq \min\left\{ K^\psi(i) : \phi_i= f\right\} - c\\
-&= \min\left\{ K^\psi(i) : \psi_{g(i)}= f\right\} - c\\
-&\geq \min\left\{ K^\psi(i) : \psi_i= f\right\} - c\\
-= K^\psi(f)
-\end{align*}
-We can reverse $\phi$ and $\psi$ without loss of generality to achieve the same inequality the other way around, completing the proof.
-\end{proof}
-
-\begin{lemma}[TODO: is het niet leuker om te zeggen dat het altijd bounded is? Krijgen we dat kado van een later lemma?]
-There is an acceptable numbering $\psi_1,\psi_2,\ldots$ such that $F^\psi(x)=|\bar 1|$ for all $x$.
-\end{lemma}
-\begin{proof}
-Let $\phi_1,\phi_2,\ldots$ be another acceptable numbering given by
-\[
-\phi_i(x)=\begin{cases}
-u&\tn{if $i=1$}\\
-\infty&\tn{if $1<i<r$}\\
-\psi_{i-r+1}(x)&\tn{if $i\ge r$}.
-\end{cases}\]
-Here $u\ge r$ is the index of the universal element satisfying $\phi_u(\bar\jmath p)=\psi_j(p)$.
-Now the two-part codelength of $x$ using representation $\phi_u$ is
-\[\begin{split}
-K(\phi_u)+\min\{|y|:\phi_u(y)=x\}&=\min\{K(i):\forall_x\phi_i(x)=\phi_u(x)\}+\min\{|y|:\phi_u(y)=x\}\\
-&=|\bar 1|+\min\{|\bar\jmath p|:\phi_u(\bar\jmath p)=x\}\\
-&=|\bar 1|+\min\{|\bar\jmath p|:\psi_j(p)=x\}\\
-&=|\bar 1|+K^\psi(x).
-\end{split}\]
-On the other hand, the two-part codelength of $x$ using a representation $s\ne u$ is
-\[\begin{split}
-K(\phi_s)+\min\{|y|:\phi_s(y)=x\}&=\min\{K(i):\forall_x\phi_i(x)=\phi_s(x)\}+\min\{|y|:\phi_s(y)=x\}\\
-&=c+\min\{|y|:\psi_{s-r+1}(y)=x\}\\
-\end{split}\]
-
-\end{proof}
-
-\begin{definition}[Facticity]
-  Let $\phi_1,\phi_2,\ldots$ be any effective enumeration of the partial recursive functions. Then
-  \[F^\phi(x)=|\bar\imath|\qquad\tn{where}\qquad \bar\imath p=\min\{\bar\jmath q:\phi_j(q)=x\}.
-  \]
-  Here the minimum is taken with respect to the standard ordering of binary sequences.
-\end{definition}
+\appendix
+\section{Dead darlings}
 
@@
 \end{proof}
-
-\begin{lemma}
-  If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then facticity is bounded.
-\end{lemma}
-\begin{proof}
-Let $\bar\imath p$ be any two-part representation for the data $x$, i.e.
-$\phi_i(p)=x$.
-Then construct an alternative two-part representation $\bar vi^* p$, where $i^*$ is the shortest $\psi_u$-program for $i$ such that $\phi_v(i^* p)=\phi_{\psi_u(i^*)}(p) = \phi_i(p)=x$.
-We compare the lengths of these two representations.
-Note that $|i^*|=K^\psi(i)$.
-Therefore,
-\[
-|\bar\imath p|-|\bar v i^* p| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-K(i).
-\]
-By assumption there must be an $i_0$ such that the above expression is positive for all $i>i_0$.
-From this $i$ onwards, the second representation (using $v$) will have a shorter code length, so $\bar\imath p$ cannot achieve the minimum in the definition of the facticity.
-Consequently, the facticity is bounded by $F^\phi(x)<|\overline{\imath_0}|$ for all $x$.
-\end{proof}
-Note that the condition for this lemma applies to all known partial recursive prefix functions:
-
-\begin{conjecture}
-For any partial recursive prefix code length function $L$, the function
-\[
-\min\{L(i)-K(i):i\ge i_0\}
-\]
-is unbounded in $i_0$.
-\end{conjecture}
-
-\begin{definition}[Model complexity]
-Given a model class $\M$, the model complexity of a data sequence $x$ is given by
-\[G(x)=K(f^*)\qquad\tn{where}\qquad (f^*,p^*)=\argmin_{f,p:f\in\M, f(p)=x}(K(f)+|p|),
-\]
-where ties are broken by picking a model $f^*$ of minimal complexity.
-\end{definition}
-
-\begin{theorem}
-There are enumerations $\phi_1,\phi_2,\ldots$ and $\psi_1,\psi_2,\ldots$ of the partial recursive (prefix) functions such that for every constant $c$ there is a binary sequence $x$ with
-\[
-|G^\phi(x)-G^\psi(x)|>c
-\]
-\end{theorem}
-
-\begin{lemma}
-There is an acceptable numbering $\phi_1,\phi_2,\ldots$ of all partial recursive prefix functions, and a constant $c$ such that for all $n$ we have $G^\phi(x)<c$ for more than half of all strings $x$ of length $n$ with respect to the model class of all partial recursive prefix functions.
-\end{lemma}
-
-\begin{lemma}
-There is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive prefix functions such that for every constant $c$ there is an $n_0$ such that for all $n>n_0$ we have $G^\phi(x)>c$ for more than half of all strings $x$ of length $n$ provided the model class contains programs for every singleton $x$.
-\end{lemma}
-
-\begin{theorem}
-There are enumerations $\phi_1,\phi_2,\ldots$ and $\psi_1,\psi_2,\ldots$ of the partial recursive prefix functions such that for every constant $c$ there is a binary sequence $x$ with
-\[
-|G^\phi(x)-G^\psi(x)|>c
-\]
-with respect to the model class of all partial recursive prefix functions.
-\end{theorem}
-
-\begin{definition}
-Let $\psi_1,\psi_2,\ldots$ be an arbitrary acceptable numbering of the partial recursive functions. Define
-\[
-B^n = \{ x : |x|=n~\tn{and}~n-c \le K^\psi(x) \le n+c \}
-\]
-\end{definition}
-
-\begin{lemma}
-  For sufficiently large $c$ the sets $B^n$ are nonempty for all $n$.
-\end{lemma}
-
-\begin{lemma}[Print function]
-  Let $\M$ be any model class that contains the identity function. There is an acceptable numbering $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that for every $n$, all $x\in B^n$, 
-we have $G^\phi_M(x)=1$.
-\end{lemma}
 
@@
 
-\section{Unbounded and top-heavy sophistications}
-
 \begin{definition}
-We call a numbering $\eta_1,\eta_2,\ldots$ \emph{optimal} if there is a constant $c$ such that for all $i$, we have $\min\{|j|:\eta_i=\eta_j\} \leq K(\eta_i) + c$.
+  An enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive
+  functions is called \emph{compact} if for all $i,y$ there is a $j$
+  such that $|j|\le|i y|$ and $\phi_i(y z)=\phi_j(z)$ for all $z$.
 \end{definition}
@@
 \begin{lemma}
-There exists an acceptable, optimal numbering of the partial computable functions.
-\label{lemma:optimal-numbering}
+There is a compact acceptable numbering.
 \end{lemma}
 \begin{proof}
-Let $\psi_1, \psi_2, \ldots$ be some acceptable numbering, with some $\psi_{\text{div}}$ with is undefined for all its arguments. Let 
-\[\begin{cases}
-	\phi_{j} = \psi_{\text{div}} &\text{if $j$ can not be read as $\bar\imath p$}\\
-	\phi_{\bar\imath p} = \psi_{\psi_i(p)} &\text{if $\psi_i(p)$ is defined}\\
-	\phi_{\bar\imath p} = \psi_{\text{div}} & \text{otherwise}
-\end{cases}\]
+  Let $\phi_1,\phi_2,\ldots$ be any acceptable numbering.
+Let $\psi_1,\psi_2,\ldots$ be a new acceptable numbering defined by $\psi_j(x)=\phi_i(y x)$ if $j$ is a pair satisfying $j=\bar\imath y$, and $\psi_j(x)=\infty$ otherwise.
+We will show that $\psi$ is both $0$-compact and acceptable.
 
-\noindent To show that this numbering is acceptable, we must provide two computable functions to translate to and from a given acceptable numbering.
-We use $\psi$ for this purpose.
-To find a $\phi_j$ which corresponds to a given $\psi_i$, we can simply use the identity function $\psi_d(x) = x$.
-For a given $\psi_i$, the function $\phi_{\overline{d}i}$ computes the same function.
+To show compactness, pick any index $i$ and binary string $y$.
+First assume $i$ is a valid pair $i=\bar a b$.
+Then  $\psi_i(y x) = \psi_{\bar a b}(y x)=\phi_a(b y x)$, but also $\psi_{\bar a b y}(x)=\phi_a(b y x)$, so $\psi_i(y x) = \psi_j(x)$ for $j=\bar a b y$; so $|j|=|iy|+0$.
+On the other hand, if $i$ is not a valid pair, we can simply use $j=i$ so that $\psi_i(y x)=\psi_j(x)=\infty$ for all $x$.
 
-To translate from $\phi_j$ to $\phi_i$, we are faced with a problem.
-If we are given $j = \bar\imath p$, we cannot decide algorithmically whether $\psi_i(p)$ is defined.
-However, we can construct the function $\psi_r(\bar\imath p, q)$ such that it computes $\psi_r{\psi_i(p)}(q)$ if $\psi_i(p)$ is defined and is undefined otherwise, or if the first argument cannot be read as $\bar\imath p$.
-For a given $j$ we can also find an $r'$ such that $\psi_{r'}(q) = \psi_(j, q)$.
-Thus, given any $\phi_i$ we can construct and return this function.
-
-Finally, we'll show that the minimum length of an index for function $f$ in $\{\phi_i\}$ is equal to $K(f)$ up to a constant:
-\begin{align*}
-K^{\psi}(f) &= \min\left\{K^{\psi}(j)\,:\,\psi_j = f \right\} \\
-&= \min_j\left\{\min_{i, p} \{|\bar\imath p| \,:\,\psi_i(p) = j\} \,:\,\psi_j = f\right\} \\
-&= \min_{i, p} \left\{|\bar\imath p| \,:\,\psi_i(p) = j, \psi_j = f\right\} \\
-&= \min_j \left\{|j|\,:\, \phi_j = f \right\}
-\end{align*}
+To show that $\psi_1,\psi_2,\ldots$ is acceptable, we define total
+computable functions $f$ and $g$ such that $\psi_i(x)=\phi_{f(i)}(x)$
+and $\phi_j(x)=\psi_{g(j)}(x)$. The easy direction is
+$g(j)=\bar\jmath\epsilon$. The definition of the mapping in the other
+direction $f(i)$ depends on whether $i$ is a valid pair. If it is not,
+then $f(i)=h$ for some $h$ such that $\phi_h(x)=\infty$ for all $x$.
+On the other hand if $i=\bar\jmath y$, then $f(i)$ is the index of a partial recursive function $\phi_{f(i)}(x)$ defined by a Turing machine that simulates the machine for $\phi_j$ but with $yx$ on the input tape.
 \end{proof}
@@
 \begin{lemma}
-Let $M$ be the class of \emph{singleton functions}, that is, the functions which produce a single output for all arguments, including the function which is undefined for all its arguments.
-There exists an optimal partial computable numbering of $M$.
+Compact numberings are sort of faithful.
 \end{lemma}
 \begin{proof}
-Let $\psi_1, \psi_2, \ldots$ be some acceptable numbering. Let $\phi_{\bar\imath p}(\cdot) = \psi_i(p)$ \ldots. 
+First, let $v$ be the index of the partial recursive function given by $\phi_v(\bar k p x)=\phi_{\phi_k(p)}(x)$.
+Let $\bar k p$ be a shortest program for $i$, that is $|\bar k p| = K(i)$ and $\phi_k(p)=i$.
+Then $\phi_i(x)=\phi_v(\bar k \bar p x)$.
+We now use compactness to find an index $j$ with $\phi_v(\bar k \bar p x)=\phi_j(x)$ and length
+\[|j|\le |v\bar k\bar p|\le|v|+|\bar k p|+2\log|\bar k p|=K(i)+|v|+2\log K(i),\]
+as required.
 \end{proof}
-
-\begin{definition}
-  Given a model $f$ and model class $\M$, define model-based complexities by
-  \begin{align*}
-  C^f(x) &= \min\{|y|:f(y)=x\} \text{,}\\
-  L^f(x) &= K(f)+C^f(x) \text{,}\\
-  L^{\M}(x) &= \min\{L^f(x)\mid f\in\M\}\text{.}
-  \end{align*}
-\end{definition}
-
-\begin{lemma}\label{lem:thecoolone}
-  Let $\psi_1,\psi_2,\ldots$ be any acceptable enumeration of the partial recursive functions.
-  Let $\M$ be any model class, let $\X$ be any set of binary sequences and let $D:\{0,1\}^*\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e.
-if $f=\psi_{D(p)}$ then $p$ is a $D$-description of $f$.
-Let $\M'=\{\psi_i:i\in\tn{range}(D)\}$.
-Further assume there is a constant $c$ such that
-\begin{enumerate}
-  \item $\forall_{f\in\M'}:\min\{|p|:\psi_{D(p)}=f\}\le K^\psi(f)+c$
-  \item $\forall_{x\in\X}:L^{\M',\psi}(x)-L^{\M,\psi}(x)\le c$.
-\end{enumerate}
-Then is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
-\end{lemma}
-\begin{proof}
-We define the numbering $\phi$ as follows:
-\[\begin{cases}
-\phi_0(p) = 1^r 0 D(p) \\
-\phi_{1^r0i}(p) = \psi_i(p) \\
-\phi_j(\cdot) = \infty &\text{if $j$ contains no zeroes.}
-\end{cases}\]
-We will show that under the $\phi$-enumeration, the best representation for $x$ using a model $f\in\M'$ is always better than the best representation using some $f\not\in\M'$.
-
-First suppose $f\in\M'$. Then
-\[\begin{split}
-K^\phi(f)&=\min\{K^\phi(i):\phi_i=f\}\\
-&=\min\{|\bar\jmath q|:\phi_j(q)=i, \phi_i=f\}\\
-&=\min\{|\bar\jmath q|:\phi_{\phi_j(q)}=f\}\\
-&\le\min\{|\bar0 q|:\phi_{\phi_0(q)}=f\}~\text{(using $j=0$)}\\
-&=\min\{|\bar0 q|:\phi_{1^r0 D(q)}=f\}\\
-&=|\bar 0|+\min\{|q|:\psi_{D(q)}=f\}\\
-&\le K^\psi(f)+c+|\bar 0|,
-\end{split}
-\]
-where the last inequality uses the first assumption.
-
-Now assume that the best model $f$ for $x$ is not in $\M'$. 
-Let $i$ be the index of $f$ with the shortest description, i.e. it achieves the minimum in
-\[
-K^\phi(f)=\min\{K^\phi(i):\phi_i=f\}.
-\]
-There are two possibilities. Either $i=0$, in which case we have $K^\phi(f)\ge r$ because $\phi_0$ cannot output zero and all other $\phi$-programs are at least $r$ bits long.
-
-Otherwise, we can bound
-\[\begin{split}
-K^\phi(f)&=K^\phi(i)=K^\phi(1^r0j)\\
-&\ge K^\phi(j)-c'\\
-&\ge K^\psi(j)+r+1-c'\\
-&\ge\min\{K^\psi(j):\phi_j=f\}+r+1-c'\\
-&=K^\psi(f)+r+1-c'.
-\end{split}
-\]
-Now choose $r=c''+\max\{K^\psi(\phi_0),c'-1\}$. Then substitution yields, for both cases, 
-$K^\phi(f)\ge c''+K^\psi(f)$.
-
-Combining the inequalities above, for any model $g\not\in\M'$, there is a model $f\in\M$ such that
-\[\begin{split}
-K^\phi(g)+C^g(x) &\ge K^\psi(g)+C^g(x)+c''\\
-&\ge K^\psi(f)+C^f(x)-c+c''\\
-&\ge K^\phi(f)+C^f(x) -2c+c'' -|\bar0|.
-\end{split}\]
-
-By choosing $c''$ sufficiently large, we can ensure that the best representation is in $\M'$ for all $x\in\X$, which completes the proof.
-\end{proof}
-
-\begin{lemma}
-Let $\M$ be a prefix model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$.
-Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the prefix partial recursive functions, and a constant $c$, such that
-\[
-K(x)-S^{\M,\phi}(x)\le c
-\]
-for all $x\in\X$.
-\end{lemma}
-\begin{proof}
-Let $\psi_1,\psi_2,\ldots$ be any default enumeration of the partial recursive prefix functions.
-Note that since $f$ is a prefix function, if $f$ is defined for input $\epsilon$ then it cannot be defined for any other input.
-Pick $f,x$ with $f(\epsilon)=x$.
-Note that $x$ can be computed from $f$ and a fixed program, so there is a $c$ such that $K(x)\le K(f)+c$.
-Vice versa, given any $x$ we can construct an index of $f$, since $\psi$ is an acceptable numbering.
-Therefore $|K(f)-K(x)|\le c$.
-
-We now define a computable function $D$ by $D(\bar\imath p)=j$ where $\psi_j(\epsilon) = \psi_i(p)$.
-We will show that the two conditions of Lemma~\ref{lem:thecoolone} hold for the prefix function $D$.
-
-\begin{enumerate}
-\item Let $f$ be any function in the range of $D$, and $x$ its output. Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\psi_i(q)=x\}=K(x)\le K(f)+c$.
-\item On the one hand $L^{\M',\psi}(x)\le K(f)+|\epsilon|\le K(x)+c$. On the other hand, $L^{\M,\psi}(x)$ is an effective description of $x$, so $K(x)$ is at most a constant larger.
-Together, these inequalities establish the second condition.
-\end{enumerate}
-
-Then by Lemma~\ref{lem:thecoolone} there is an enumeration $\phi_1,\phi_2,\ldots$ such that $S^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
-We observed that $|K(f)-K(x)|\le c$ for any $f\in\M'$, so $S^{\M',\psi}(x)\ge K(x)+c$.
-This proves the lemma.
-\end{proof}
-
-Plan: een lemma voor total/niet prefix dat laat zien mbv pigeonhole dat facticity juist bounded of juist unbounded kan zijn.
-De ene kant op gaat dat met de print functie (klein model, grote data).
-De andere kant op is moeilijk
 

### 2014-11-24 Facticity/Facticity/paper.tex

@@
 
+Adriaans takes this problem seriously and attempts to address it by imposing additional restrictions on the ordering of TMs implied by the used UTM \cite{adriaans2012facticity}.
+We are skeptical that such measures can effectively eliminate this problem without resorting to orderings of the TMs that are not admissible [TODO ref].
+
 Inherent property of the data: 

### 2014-11-26 Facticity/Facticity/paper.tex
...

@@
-\documentclass{article}
+\documentclass{style/llncs}
 
-\usepackage{amsmath,amsfonts,amsthm,fullpage}
+\usepackage{amsmath,amsfonts,fullpage}
 \usepackage{color}
@@
 
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{conjecture}[theorem]{Conjecture}
-\newtheorem{definition}{Definition}
+%\newtheorem{theorem}{Theorem}
+%\newtheorem{lemma}[theorem]{Lemma}
+%\newtheorem{corollary}[theorem]{Corollary}
+%\newtheorem{conjecture}[theorem]{Conjecture}
+%\newtheorem{definition}{Definition}
 
@@
 
-\title{Ideal model selection is impossible}
+\title{A Critical Review of Sophistication and Model Selection in Algorithmic Statistics}
 \author{Peter Bloem and Steven de Rooij}
@@
 \begin{abstract}
-Kolmogorov complexity provides an objective quantification of the amount of information in an individual binary string, i.e.
-the number of bits required to effectively describe it, without making any assumptions about the source of that string.
-However, not all information is created equal -- over the years, there have been many attempts that use two-part codes in order to separate the amount of ``meaningful information'', ``sophistication'' or ``facticity'' of the data, from information representing its random properties.
-How wonderful would it be if we could quantify not just the syntactic information content of the data  not only syntactically, but the meaningful information as well!
+\noindent Kolmogorov complexity is a sounds expression of the information of the information content of a binary string, but it leaves open the issue of \emph{meaningful information}.
+The strings with maximal information content are not the ones we would regard as interesting, or meaningful.
+One attempt to formalize a notion of meaningful information is sophistication.
+It is a simple idea with many attempted formalizations: describe the string efficiently with a two-part coding approach, and take the complexity of the model as a measure of the amount of meaningful information.
+We review the variety of attempts to formalize this approach and highlight their shortcomings.
+We provide a critical perspective: we question whether any such formalization can truly be successful as an objective notion of meaningful information.
+A true property of the data, like Kolmogorov complexity.
+We show that no formalization currently proposed in the literature satisfies this demand.
+All sophistication functions are either bounded by a constant value, or are not invariant to the choice of universal Turing machine used in their construction.
+\end{abstract}
+\section{Introduction}
 
-In this paper we show this to be an impossible dream: in general, additional assumptions are required if one wishes to find the true model, or its complexity.
-This is true regardless of the length or Kolmogorov complexity of the input data.
-In effect, this means that ideal model selection based on a single data object with no further knowledge is impossible.
-We unify and contrast the several existing approaches and describe what they can and cannot do.
-\end{abstract}
+\noindent Kolmogorov complexity provides an objective measurement of the information content of a string: it is a property of the data, which does not depend on ad-hoc choices in its definition.
+Unfortunateldoes not answer the question of meaningful or interesting data.
+The strings that have minimal and maximal Kolmogorov complexity are the least interesting in an intuitive sense.
+The strings that contain useful information are somewhere in between.
+
+Consider the following, illustrative example: an old-fashioned, analog television set which can be tuned to three frequencies.
+The first shows a blank screen, the second a regular program, and the third white noise.
+By the Kolmogorov complexity, these  broadcasts have increasing information content.
+The first is easiest to describe, the second contains a great deal of regularity, and the third cannot be compressed.
+However, in terms of meaningful information, information that we find interesting, the second signal is by far the richest.
+This notion has proved far harder to capture in formal terms than the simple notion of `raw' information content.
+One of the most frequently explored ideas is \emph{sophistication}.
+
+Sophistication select a model for the data by two-part coding, and to take the complexity of the model as a measure of the meaningful information.
+The assumption is that both simple and random data have small models, and the more interesting a dataset is, the more complex its model will be.
+A key assumption of the approach of sophistication is that we can perform objective model selection for a given string.
+That is: there is a function that, given a string, selects a single unique model that can be used to compress this string optimally.
+Moreover, any elements of this function that are chosen ad hoc can be changed without significantly affecting the outcome.
+
+In the case of the Kolmogorov complexity, this holds true; we can change the universal Turing machine used in its definition and the complexities returned will not change by more than a constant.
+This is known as \emph{invariance}.
+For this reason, we view the Kolmogorov complexity as an intrinsic property of the data, rather than simply a function computed on it.
+If a formalization of sophistication is ever to be held in the same regard as the Kolmogorov complexity, it must be subject to a similar principle of invariance.
+
+The unifying principle over all the notions we discuss in this paper is simple.
+We start with an object expressed as a finite binary string.
+We represent this object, hoping to compress it, in two parts, the first is a model, and the second an `input' to this model.
+How the two parts are encoded, and what the class of models is are details that differ between variants.
+We know, however, that for succesful representations, those that capture all structure in the string, the length of the two-part code should approach the Kolmogorov complexity.
+We take the set of all \emph{sufficient statistics}---those representations that are `close enough' to the Kolmogorov complexity---and call the one with the lowest model complexity the \emph{minimal sufficient statistic}.
+The size of the minimal sufficient statistic is the sophistication.
+
+The intuition is that both very simple and very random data have small, simple programs: one with a small input and one with a large, random input.
+The strings with much structure and many interesting properties are those for which the best model is large.
+Two assumptions are crucial to this view:
+\begin{itemize}
+  \item This principle separates the structure and the noise into distinct parts.
+In the optimal representation the model contains only `structural information' and the noise contains only the `uninteresting' information.
+  \item This principle is invariant. If we change, for instance, the universal Turing machine used for our representations, the minimal sufficient statistic will not change significantly.
+\end{itemize}
+
+\subsection{Sophistication, Minimal Sufficient Statistic, Effective Complexity and Facticity}
+
+We will first briefly discuss the different variants of sophistication in a historical context, to give the reader a simple overview, without discussing too much technical detail.
+In the main body of the paper, we will define each variant rigorously, and in non-chronological order.
+
+The first suggestion that a representation of a string can be separated cleanly into a structural part, containing a description of its properties, and a noise part, containing the residual information, came from Kolmogorov himself, in a talk given at a conference in Tallinn in 1973 (first committed to paper by Cover \cite{todo} and later more extensively discussed by Vitanyi \cite{todo}).
+Here, he proposed what has become known as the \emph{Kolmogorov structure function}.
+Briefly, this function operates on a space of two part representations for a given string $x$, consisting of some finite set $S$ (containing $x$) and the index of $x$ in $S$.
+The \emph{minimal sufficient statistic}, it was proposed, is the set with the lowest complexity for which the length of the two-part representation reaches the prefix-free Kolmogorov complexity.
+\pb{Did Komogorov already talk about the MSS?}
+
+Koppel \cite{} built on this notion, coining the phrase \emph{sophistication}.
+He changed the class of `models' from finite sets to total functions and used Levin's monotone complexity, instead of prefix-free Kolmogorov complexity.
+One interesting difference is that Koppel takes the description of the data under a variant of Kolmogorov complexity, which consists of a description of a model and that model's input, and takes the size of the model part as the sophistication.
+In contrast Kolmogorov's minimal sufficient statistic is built on a second, two-part, representation in addition to the Kolmogorov complexity's one-part representation.
+This difference in perspective is subtle but relevant to the discussion ahead.
+
+Separately, G\'acs, Tromp and Vit\'anyi \cite{}, built on the idea of the structure function to establish a framework of \emph{algorithmic statistics}, in the hope of founding a statistical theory not on probabilistic notions, but on combinatorical ones.
+This echoes Kolmogorv's own hope for the theory of information:
+\begin{quotation}
+\noindent Information theory must precede probability theory and not be based on it.
+By the very essence of this discipline, the foundations of information theory must have a finite combinatorial character
+\end{quotation}
+
+\pb{emphasize the different aims}
+\pb{effective complexity}
 

### 2014-12-02 Facticity/Facticity/paper.tex
Effective complexity

@@
 Separately, G\'acs, Tromp and Vit\'anyi \cite{}, built on the idea of the structure function to establish a framework of \emph{algorithmic statistics}, in the hope of founding a statistical theory not on probabilistic notions, but on combinatorical ones.
-This echoes Kolmogorv's own hope for the theory of information:
+This echoes Kolmogorov's own hope for the theory of information:
 \begin{quotation}
@@
 \end{conjecture}
+
+\subsection{Effective complexity}
+
+\begin{itemize}
+\item Model class is \emph{ensembles}. Equivalent to $\K$.
+\item Description method for models is $L(m) = K(m)$ (the minimization is not mentioned explicitly, but we'll give them the benefit of the doubt).
+\item The threshold function is equal to the minimum two-part code: $\tau(x)=L^m(x)$.
+\item Gell-mann and Lloyd acknowledge that representations exist with small and large models.
+The choice from the candidate set is made by applying a time-bound to the computation from the string to the model, but not to the decoding of the model itself.
+The idea is that this forces `deep' information into the model, and shallow information into the data.
+Presumably, the representation with the smallest model code under this constraint is chosen.
+\end{itemize} 
+\pb{We can still manipulate the UTM so that all non-singleton representations are pushed out of the candidate set.
+Since the model side is allowed unbounded decoding time, this invalidates effective complexity.}
 

### 2014-12-08 Facticity/Facticity/paper.tex
gacs

--- 2014-11-24 Facticity/Facticity/paper.tex
@@
 
+\section{Paul's sophistication (total, non-prefix) is \emph{not}  bounded}
+
+Proof sketch:
+
+\begin{itemize}
+  \item There is a $C$ such that for every $n$ there are strings $x$
+    with $K(S)+\log|S|\ge K(x)+(3/4)n+C$ for all $K(S)>1/4$.
+
+  \item For all $C$ there is an $n$ such that there are $x$ with
+    $K(S)+\log(S)<K(x)+C$ only if $K(S)>1/4$.
+\end{itemize}
+
 \bibliographystyle{plain}

### 2014-12-08 Facticity/Facticity/paper.tex
flupsie

--- 2014-12-02 Facticity/Facticity/paper.tex
@@
 
+\section{Paul's sophistication (total, non-prefix) is \emph{not}  bounded}
+
+Proof sketch:
+
+\begin{itemize}
+  \item From \cite[Proposition~I.3 (b)]{gacs2001algorithmic} we know
+    that for sufficiently large $n$ there are strings of length less
+    than $n$ that are not $(n/4, n/4)$-stochastic.
+  \item By definition of $(\alpha,\beta)$-stochasticity, if a string
+    is not $(n/4,n/4)$-stochastic, then for every finite set $S$,
+    either (a) $\log|S|\ge K(x)+n/4$, so $S$ cannot be a candidate
+    representation for $x$, or (b) $K(S)>n/4$. For sufficiently large
+    $n$, we have $n/4>K(x)/5$, so we also know $K(S)>K(x)/5$.
+  \item Neem nonstochastic string zoals hierboven beschreven. Stel dat
+    die een total candidate heeft met klein model. Dan kun je de
+    reductie naar sets doen ter betaling van $\log(n)$ additieve term,
+    en dan heb je dus ook een kleine finite-set representatie, maar
+    dat is een tegenspraak. QED
+  \item De reductie van total naar finite sets staat (met typo's) in
+    \cite[Lemma~7.2]{vitanyi2004meaningful}. 
+\end{itemize}
+
 \bibliographystyle{plain}

### 2014-12-08 Facticity/Facticity/paper.tex
...

@@
 
+\begin{lemma}
+Let $f$ be a total, computable function, such that for some $d$, $f(d) = x$ and let $k = K(f) + |d|$.
+Then there exists a finite set $S$ containing $x$, such that we have $K(S) \leq K(f) + K(|d|)$ and $\log |S| \leq |d|$.\label{lemma:total-to-sets}
+\end{lemma}
+\begin{proof}
+Let $S = \{f(y) : |y| = |d|\}$.
+Since $f$ is total, this set can be explicity computed from a description of $f$ and a description of $|d|$, this tells us that there is a constant such that $K(S) \leq K(f) + K(|d|) + c$.
+
+Note that the inputs $y$ used to construct all have length $|d|$, so that $S$ contains at most $2^{|d|}$ strings. Therefore, we have $\log |S| \leq |d|$.
+\end{proof}
+This lemma is an extension of \cite[Lemma~7.2]{vitanyi2004meaningful} and uses approximately the same proof.
+
+\begin{theorem}
+There are $x$, such that $S^\T(x) \geq \tfrac{1}{4}K(x)$.
+\end{theorem}
+\begin{proof}
+Let $x\geq n$ be ($\tfrac{1}{4n}$, $\tfrac{1}{4}$)-stochastic. This means that for any $S$ containing $x$ with $K(S) \leq \tfrac{1}{4}$ we have $K(S) + \log|S| \geq K(x) + \tfrac{3}{4}n$. \ldots
+\end{proof}
+
 \bibliographystyle{plain}

### 2014-12-09 Facticity/Facticity/paper.tex
...

--- 2014-12-08 Facticity/Facticity/paper.tex
@@
   \[
-  C^\M_\tau(x):=\{m\in\M:L^m(x)\le \tau(x)\}.
+  C^\M_\tau(x):=\{m\in\M:L^m(x)\le \tau(x) + c\}
   \]
+  with $c$ a constant independent of $x$ which may be chosen arbitrarily. The default threshold function is $\tau^\circ(x)=L^\M(x)$.
 \end{definition}
@@
   Definition~\ref{def:faithful}
-\item Threshold function is $\tau(x)=L^m(x)$.
+\item Threshold function is $\tau(x)=L^\M(x)$, the constant in the candidate set is $0$.
 \item Facticity is actually defined as index length without prefix
@@
 %
+
+\section{Paul's sophistication (total, non-prefix) is \emph{not}  bounded}
+
+\begin{lemma}
+Let $f$ be a total, computable function, such that for some $d$, $f(d) = x$ and let $k = K(f) + |d|$.
+Then there exists a finite set $S$ containing $x$ and a constant $c$, such that $K(S) \leq K(f) + K(|d|) + c$ and $\log |S| \leq |d|$.\label{lemma:total-to-sets}
+\end{lemma}
+\begin{proof}
+Let $S = f\left(\{0,1\}^{|d|}\right)$.
+Since $f$ is total, this set can be explicity computed from a description of $f$ and a description of $|d|$, which tells us that there is a constant $c$ such that $K(S) \leq K(f) + K(|d|) + c$.
+
+Since $S$ is the image of a set of size $2^{|d|}$ under $f$, we have $\log |S| \leq |d|$.
+\end{proof}
+This lemma is a variation on \cite[Lemma~7.2]{vitanyi2004meaningful}.
+
+\begin{theorem}
+There are infinitely many $x$ with 
+\begin{enumerate}
+  \item $S^\F_K(x) \geq K(x)/5$,\label{eq:item1}
+  \item $S^\T_K(x) \geq K(x)/6$, {and}\label{eq:poezenvoer}
+  \item $S^\T_{\tau^\circ}(x) \geq K(x)/6$. \label{eq:hondevoer}
+\end{enumerate}
+\end{theorem}
+\begin{proof}
+From \cite[Proposition~I.3 (b)]{gacs2001algorithmic} we know that for sufficiently large $i$ there are strings of length less than $i$ that are not $(i/4, i/4)$-stochastic.
+For each $i$, let $x_i$ be the smallest binary string that is not $(i/4, i/4)$-stochastic.
+Since every string is $(\alpha,\alpha)$-stochastic for sufficiently large $\alpha$, this sequence contains infinitely many distinct binary strings.
+Now pick any $i$ that is large enough that (a) $|x_i|<i$ and (b) $K(x_i)/5 < i/4$.
+By definition of $(\alpha,\beta)$-stochasticity, if a string is not $(i/4,i/4)$-stochastic, then for every finite set $S$ containing $x$, either (a) $\log|S|\ge K(x_i)+i/4$, in which case $S$ cannot be a candidate representation for $x$ in $C^\F_K$, or (b) $K(S)>i/4>K(x_i)/5$.
+This proves Item~\ref{eq:item1}.
+
+For Items~\ref{eq:poezenvoer} and~\ref{eq:hondevoer}, suppose towards contradiction that either $C^\T_K(x)$ or $C^T_{\tau^\circ}(x)$ contains a candidate $f$ with $K(f)\le i/5$.
+Then by Lemma~\ref{lemma:total-to-sets}, there is a set $S$ with $\log|S|\le|d|\le\le\tau^\circ(x)\le K(x)$, so (a) above cannot be the case, and $K(S)\le K(f)+K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) for large enough $i$.
+Therefore such candidates $f$ cannot exist and all candidates must satisfy $K(f)>i/5>K(x_i)/6$.
+\end{proof}
+Note that we favoured simplicity of the proof over the strength of the bounds: the constants can be improved.
 
@@
 
-\section{Paul's sophistication (total, non-prefix) is \emph{not}  bounded}
-
-Proof sketch:
-
-\begin{itemize}
-  \item There is a $C$ such that for every $n$ there are strings $x$
-    with $K(S)+\log|S|\ge K(x)+(3/4)n+C$ for all $K(S)>1/4$.
-
-  \item For all $C$ there is an $n$ such that there are $x$ with
-    $K(S)+\log(S)<K(x)+C$ only if $K(S)>1/4$.
-\end{itemize}
-
 \bibliographystyle{plain}

### 2014-12-10 Facticity/Facticity/paper.tex
...

@@
 \newcommand{\symb}[1]{\texttt{#1}}
+
+\let\doendproof\endproof
+\renewcommand\endproof{~\hfill\qed\doendproof}
 
@@
 
+\begin{theorem}
+There are infinitely many $x$ such that $S^\F_K(x) \geq K(x)/5$. \label{theorem:structure-function-is-not-bounded}
+\end{theorem}
+\begin{proof}
+From \cite[Proposition~I.3 (b)]{gacs2001algorithmic} we know that for sufficiently large $i$ there are strings of length less than $i$ that are not $(i/4, i/4)$-stochastic.
+For each $i$, let $x_i$ be the smallest binary string that is not $(i/4, i/4)$-stochastic.
+Since every string is $(\alpha,\alpha)$-stochastic for sufficiently large $\alpha$, this sequence contains infinitely many distinct binary strings.
+Now pick any $i$ that is large enough that (a) $|x_i|<i$ and (b) $K(x_i)/5 < i/4$.
+By definition of $(\alpha,\beta)$-stochasticity, if a string is not $(i/4,i/4)$-stochastic, then for every finite set $S$ containing $x$, either (a) $\log|S|\ge K(x_i)+i/4$, in which case $S$ cannot be a candidate representation for $x$ in $C^\F_K$, or (b) $K(S)>i/4>K(x_i)/5$.
+This proves the theorem.
+\end{proof}
+
 \subsection{Vit\'anyi's Sophistication / Meaningful Information}
@@
 
-\section{Paul's sophistication (total, non-prefix) is \emph{not}  bounded}
-
 \begin{lemma}
@@
 
+We use this lemma to transport the result of Theorem~\ref{theorem:structure-function-is-not-bounded} to the model class of total functions:
+
 \begin{theorem}
@@
 \begin{enumerate}
-  \item $S^\F_K(x) \geq K(x)/5$,\label{eq:item1}
   \item $S^\T_K(x) \geq K(x)/6$, {and}\label{eq:poezenvoer}
@@
 \begin{proof}
-From \cite[Proposition~I.3 (b)]{gacs2001algorithmic} we know that for sufficiently large $i$ there are strings of length less than $i$ that are not $(i/4, i/4)$-stochastic.
-For each $i$, let $x_i$ be the smallest binary string that is not $(i/4, i/4)$-stochastic.
-Since every string is $(\alpha,\alpha)$-stochastic for sufficiently large $\alpha$, this sequence contains infinitely many distinct binary strings.
-Now pick any $i$ that is large enough that (a) $|x_i|<i$ and (b) $K(x_i)/5 < i/4$.
-By definition of $(\alpha,\beta)$-stochasticity, if a string is not $(i/4,i/4)$-stochastic, then for every finite set $S$ containing $x$, either (a) $\log|S|\ge K(x_i)+i/4$, in which case $S$ cannot be a candidate representation for $x$ in $C^\F_K$, or (b) $K(S)>i/4>K(x_i)/5$.
-This proves Item~\ref{eq:item1}.
-
-For Items~\ref{eq:poezenvoer} and~\ref{eq:hondevoer}, suppose towards contradiction that either $C^\T_K(x)$ or $C^T_{\tau^\circ}(x)$ contains a candidate $f$ with $K(f)\le i/5$.
-Then by Lemma~\ref{lemma:total-to-sets}, there is a set $S$ with $\log|S|\le|d|\le\le\tau^\circ(x)\le K(x)$, so (a) above cannot be the case, and $K(S)\le K(f)+K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) for large enough $i$.
+We continue from the proof of Theorem~\ref{theorem:structure-function-is-not-bounded}.
+Suppose towards contradiction that either $C^\T_K(x_i)$ or $C^T_{\tau^\circ}(x_i)$ contains a candidate $f$ with $K(f)\le i/5$.
+Then by Lemma~\ref{lemma:total-to-sets}, there is a set $S$ with $\log|S|\le|d|\le\tau^\circ(x)\le K(x)$, so (a) in the proof of Theorem~\ref{theorem:structure-function-is-not-bounded} above cannot be the case, and $K(S)\le K(f)+K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) in the same proof for large enough $i$.
 Therefore such candidates $f$ cannot exist and all candidates must satisfy $K(f)>i/5>K(x_i)/6$.
 \end{proof}
+
 Note that we favoured simplicity of the proof over the strength of the bounds: the constants can be improved.

### 2014-12-16 Facticity/Facticity/paper.tex
...

@@
 
-\usepackage{amsmath,amsfonts,fullpage}
+\usepackage{amsmath,amsfonts}
 \usepackage{color}
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+\usepackage[mathscr]{eucal}
 
-%\newtheorem{theorem}{Theorem}
-%\newtheorem{lemma}[theorem]{Lemma}
-%\newtheorem{corollary}[theorem]{Corollary}
-%\newtheorem{conjecture}[theorem]{Conjecture}
-%\newtheorem{definition}{Definition}
-
-\newcommand{\M}{\mathcal M}
-\newcommand{\C}{\mathcal C}
-\newcommand{\T}{\mathcal T}
-\newcommand{\F}{\mathcal F}
-\renewcommand{\P}{\mathcal P}
-\newcommand{\K}{\mathcal K}
-\newcommand{\X}{\mathcal X}
+\newcommand{\M}{\mathscr M}
+\newcommand{\C}{\mathscr C}
+\newcommand{\T}{\mathscr T}
+\newcommand{\F}{\mathscr F}
+\renewcommand{\P}{\mathscr P}
+\newcommand{\K}{\mathscr K}
+\newcommand{\X}{\mathscr X}
 \newcommand{\B}{\mathbb B}
@@
 \newcommand{\symb}[1]{\texttt{#1}}
+\newcommand{\br}[1]{\overline{#1}}
+\newcommand{\s}{\tn{soph}}
+
+\newtheorem{conj}{Conjecture}
 
@@
 
-\newcommand{\p}{\quad\text{.}}
+\newcommand{\p}{\,\text{.}}
 
 \newcommand{\tuple}[1]{\left\langle{#1}\right\rangle}
+
+\newcommand{\hide}[1]{}
 
@@
 
-\title{A Critical Review of Sophistication and Model Selection in Algorithmic Statistics}
+\title{A Critical Review of Sophistication and Algorithmic Model Selection}
 \author{Peter Bloem and Steven de Rooij}
+
+\institute{
+  System and Network Engineering Group, \\University of Amsterdam, the Netherlands\\
+  \email{uva@peterbloem.nl, steven.de.rooij@gmail.com}
+}
 
@@
 \begin{abstract}
-\noindent Kolmogorov complexity is a sounds expression of the information of the information content of a binary string, but it leaves open the issue of \emph{meaningful information}.
+Kolmogorov complexity is a sound expression of the information content of a binary string, but it leaves open the issue of \emph{meaningful information}.
 The strings with maximal information content are not the ones we would regard as interesting, or meaningful.
-One attempt to formalize a notion of meaningful information is sophistication.
+One attempt to capture the idea of meaningful information is \emph{sophistication}.
 It is a simple idea with many attempted formalizations: describe the string efficiently with a two-part coding approach, and take the complexity of the model as a measure of the amount of meaningful information.
-We review the variety of attempts to formalize this approach and highlight their shortcomings.
-We provide a critical perspective: we question whether any such formalization can truly be successful as an objective notion of meaningful information.
-A true property of the data, like Kolmogorov complexity.
-We show that no formalization currently proposed in the literature satisfies this demand.
-All sophistication functions are either bounded by a constant value, or are not invariant to the choice of universal Turing machine used in their construction.
+We describe three requirements that any formalization must meet in order to satisfy the expectations that come with the idea of sophistication: invariance, unboundedness and efficiency.
+We show that no currently published proposal meets these requirements \pb{TODO}, and conjecture that no formalization of the idea of sophistication can hope to do so.
+We conclude by reviewing the consequences of this conjecture.
 \end{abstract}
@@
 
-\noindent Kolmogorov complexity provides an objective measurement of the information content of a string: it is a property of the data, which does not depend on ad-hoc choices in its definition.
-Unfortunateldoes not answer the question of meaningful or interesting data.
-The strings that have minimal and maximal Kolmogorov complexity are the least interesting in an intuitive sense.
-The strings that contain useful information are somewhere in between.
+\noindent Kolmogorov complexity provides an objective measurement of the information content of an object.
+Any arbitrary choices made in its construction can be changed, without significantly changing the outcome.
+This property is crucial for the measure's success: it allows us to see Kolmogorov complexity as a \emph{property of the data} rather than an arbitary function computed on it.
+This, in turn, allows us to define the intrinsic entropy, or randomness, inherent in a single object expressed as a binary string, without any assumptions on the process that produced it.
+Unfortunately, this idea of complexity does not fit one-on-one to our intuitive idea of which objects are complex.
 
 Consider the following, illustrative example: an old-fashioned, analog television set which can be tuned to three frequencies.
-The first shows a blank screen, the second a regular program, and the third white noise.
+The first produces a blank screen, the second a regular program, and the third white noise.
 By the Kolmogorov complexity, these  broadcasts have increasing information content.
@@
 This notion has proved far harder to capture in formal terms than the simple notion of `raw' information content.
-One of the most frequently explored ideas is \emph{sophistication}.
 
-Sophistication select a model for the data by two-part coding, and to take the complexity of the model as a measure of the meaningful information.
-The assumption is that both simple and random data have small models, and the more interesting a dataset is, the more complex its model will be.
-A key assumption of the approach of sophistication is that we can perform objective model selection for a given string.
+One of the most frequently explored ideas for capturing the notion of meaningful information is \emph{sophistication}.
+The idea behind sophistication is to select a description for the data by two-part coding (describing a model first, and then the data in terms of the model), and to take the complexity of only the model part as a measure of the meaningful information.
+The assumption is that both simple and random data have small models; in the case of the simple data, a small model describes all regularity, and no further information is required, and in the case of the random data, the data itself is its shortest representation, and the model is a `print function' which simply reproduces its input.
+If the dataset is more interesting, however, containing both regularity and randomness, its best representation will have a more complex model.
+A key assumption of the idea of sophistication is that we can perform objective model selection for a given string.
 That is: there is a function that, given a string, selects a single unique model that can be used to compress this string optimally.
@@
 This is known as \emph{invariance}.
-For this reason, we view the Kolmogorov complexity as an intrinsic property of the data, rather than simply a function computed on it.
+Because Kolmogorov complexity is invariant, we can view it as an intrinsic property of the data, rather than simply an arbitrary function computed on it.
 If a formalization of sophistication is ever to be held in the same regard as the Kolmogorov complexity, it must be subject to a similar principle of invariance.
 
-The unifying principle over all the notions we discuss in this paper is simple.
+The unifying principle over all the notions we review in this paper is simple.
 We start with an object expressed as a finite binary string.
@@
 How the two parts are encoded, and what the class of models is are details that differ between variants.
-We know, however, that for succesful representations, those that capture all structure in the string, the length of the two-part code should approach the Kolmogorov complexity.
-We take the set of all \emph{sufficient statistics}---those representations that are `close enough' to the Kolmogorov complexity---and call the one with the lowest model complexity the \emph{minimal sufficient statistic}.
-The size of the minimal sufficient statistic is the sophistication.
+We know, however, that for successful representations, those that capture all structure in the string, the length of the two-part code should approach the Kolmogorov complexity.
+We take the set of all \emph{sufficient statistics}---the models of those representations that are `close enough' to the Kolmogorov complexity---and call the one with the lowest complexity the \emph{minimal sufficient statistic}.
+The complexity of the minimal sufficient statistic is the sophistication.
 
-The intuition is that both very simple and very random data have small, simple programs: one with a small input and one with a large, random input.
+The intuition is that both very simple and very random data have small, simple programs.
 The strings with much structure and many interesting properties are those for which the best model is large.
@@
 \begin{itemize}
-  \item This principle separates the structure and the noise into distinct parts.
+  \item The two-part coding efficiently separates the structure and the noise into distinct parts.
 In the optimal representation the model contains only `structural information' and the noise contains only the `uninteresting' information.
-  \item This principle is invariant. If we change, for instance, the universal Turing machine used for our representations, the minimal sufficient statistic will not change significantly.
+  \item This principle is invariant. If we change, for instance, the universal Turing machine used to unpack our descriptions, the minimal sufficient statistic will not change significantly.
 \end{itemize}
 
-\subsection{Sophistication, Minimal Sufficient Statistic, Effective Complexity and Facticity}
+In addition to
 
-We will first briefly discuss the different variants of sophistication in a historical context, to give the reader a simple overview, without discussing too much technical detail.
+In this article we will show that most proposals do not achieve these goal.
+For one remaining case, we can offer no such proof, but there exists no proof to the contrary either.
+We offer several insights that suggest that the idea of an objective sophistication function is an unattainable ideal.
+
+\subsection{A brief historical overview}
+
+We will first briefly discuss the different variants of sophistication in a historical context, to give the reader a simple, chronological overview, without discussing too much technical detail.
 In the main body of the paper, we will define each variant rigorously, and in non-chronological order.
 
-The first suggestion that a representation of a string can be separated cleanly into a structural part, containing a description of its properties, and a noise part, containing the residual information, came from Kolmogorov himself, in a talk given at a conference in Tallinn in 1973 (first committed to paper by Cover \cite{todo} and later more extensively discussed by Vitanyi \cite{todo}).
+The first suggestion that a representation of a string can be separated cleanly into a structural part, containing a description of its regularities, and a noise part, containing the residual information, came from Kolmogorov himself, in a talk given at a conference in Tallinn in 1973 (first committed to paper by Cover \cite{todo} and later more extensively discussed by Vitanyi \cite{todo}).
 Here, he proposed what has become known as the \emph{Kolmogorov structure function}.
@@
 The \emph{minimal sufficient statistic}, it was proposed, is the set with the lowest complexity for which the length of the two-part representation reaches the prefix-free Kolmogorov complexity.
-\pb{Did Komogorov already talk about the MSS?}
+\pb{TODO: lowest randomness deficienty.}
 
@@
 One interesting difference is that Koppel takes the description of the data under a variant of Kolmogorov complexity, which consists of a description of a model and that model's input, and takes the size of the model part as the sophistication.
-In contrast Kolmogorov's minimal sufficient statistic is built on a second, two-part, representation in addition to the Kolmogorov complexity's one-part representation.
+In contrast, Kolmogorov's minimal sufficient statistic is built on a second, two-part, representation in addition to the Kolmogorov complexity's one-part representation.
 This difference in perspective is subtle but relevant to the discussion ahead.
 
-Separately, G\'acs, Tromp and Vit\'anyi \cite{}, built on the idea of the structure function to establish a framework of \emph{algorithmic statistics}, in the hope of founding a statistical theory not on probabilistic notions, but on combinatorical ones.
+While Koppel expands the model class from finite sets to functions, he maintains a restriction that the functions be total.
+This is a recurring theme: if the model class is expanded to all total functions, the the universal Turing machine itself becomes a model, so that we always have a two-part description with a small model, to within a constant of the Kolmogorov complexity.
+In most treatment this `collapse' to the universal model is avoided by excluding from the model class, using some restriction.
+This acts as a kind of force, pushing information into the model part of the two-part code.
+As we shall see, defining sophistication quickly becomes a balancing act, trying to match two forces which by themselves push information one way or the other.
+
+In 1996 Gell-Mann and Lloyd introduced `Effective complexity'\cite{todo}.
+Working from the perspective of statistical mechanics, they arrived at a notion remarkably similar to Sophistication.
+Effective complexity describes abinary string in terms of a probability distribution over finite strings (an `ensemble' in the terminology of Gell-Mann and Lloyd) and the (Shannon) entropy of the distribution.
+The reasoning is that, if the string is typical for the distribution, the entropy is sufficient to describe the string given the distribution, so that the length of the description of the distribution and the size of its entropy suffices as a two-part description.
+The two-part description which determines the effective complexity is chosen by placing a time bound (taken from the singleton description) on the computation of the string from the distribution.
+The hope is that this will move deep (ie.
+hard to compute) information to the model part, and shallow information to the data.
+In a later treatment\cite{}, Gell-Mann and Lloyd select the decisive two-part representation from those achieving minimal code-length based on ``the average quantities judged to be important'', which would seem to make the value highly subjective and context-dependant.
+
+G\'acs, Tromp and Vit\'anyi \cite{}, built on the idea of the structure function to establish a framework of \emph{algorithmic statistics}, in the hope of founding a statistical theory not on probabilistic notions, but on combinatorial ones.
 This echoes Kolmogorov's own hope for the theory of information:
@@
 \end{quotation}
+Algorithmic statistics studies binary strings and their possible two-part representations.
+In a later paper\cite{gacs2001algorithmic}, Vit\'anyi specifically studies the role of the minimal sufficient statistic as an expression of the `meaningful' information of a string (taking the name sophistication from Koppel).
+Like Koppel, Vit\'anyi restricts the model class to total functions, eliminating the universal model.
+He balances this force, by allowing the total two-part description to be non-self delimiting, so that both the representation with all information in the model part, and the representation with everything in the residual part are less efficient than the best two-part coding.
 
-\pb{emphasize the different aims}
-\pb{effective complexity}
+Antunes et al.
+\cite{} note that the constant cutoff used to separate candidate two-part descriptions from those deemed to be too inefficient, is a source of instability.
+Small changes to this constant could shift the result sophistication by arbitrary amounts.
+They introduce a more elaborate criterion for the cutoff, calling this variant \emph{coarse sophistication}.
+As we shall see, this version behaves very differently from regular sophistication, and is closer to computational depth.
 
-\section{Definitions}
+Adriaans' \emph{Facticity}, introduced in 2012\cite{adriaans2012facticity}, relaxes the model class to include all
+partial recursive functions.
+By constructing a very specific universal Turing machine, and taking constant terms into account, the facticity is expected to become robust against the collapse to a universal model.
 
-All numberings should be acceptable!
+\pb{Criticisms of sophistication}
+
+\pb{Vereshchagin's strongly sufficient statistic}
+
+\pb{Antunes' instability result}
+
+\pb{Diagrammetje}
+
+\section{Preliminaries, definitions and desiderata}
+
+Before we start, we will define a framework into which we can fit all current variants of sophistication.
+The principle is always the same: we start with a class of models, we consider the space of two-part descriptions using these models, we select the set of representations with the shortest code-lengths and we select, by some method, the best model from this set.
+
+Let $\B = {0,1}^*$.
+We deal with partial functions $\phi: \B \to \B$, which we will also call \emph{models}.
+Such functions are \emph{computable} if they can be computed by a Turing machine as described in \cite[Definition~1.7.1]{li1993introduction}.
+A function is prefix-free if its domain, $D_f = \{y : f(y) \neq \infty\}$ is a prefix set.
+A function $f$ is \emph{total} if $D_f = \B$.
+A prefix-free function is \emph{complete} if every infinite binary string has a member of $D_f$ as a prefix.
+\footnotemark
+
+\footnotetext{Completeness is analogous to totality, in that a prefix-free Turing machine that does not halt corresponds to a complete prefix-free function.}
 
 \begin{definition}[Models and model classes]
-  A \emph{model} is a function $f:\{0,1\}^*\to\{0,1\}^*$. A
+  A \emph{model} is a function $\phi:\B\to\B$. A
   \emph{model class} is a set of models. We distinguish four important
@@
   \begin{itemize}
-  \item $\C$ is the set of all partial recursive functions,
-  \item $\K\subset\C$ is the set of all partial recursive prefix
+  \item $\C$ is the set of all partial computable functions,
+  \item $\K\subset\C$ is the set of all prefix-free
     functions,
-  \item $\T\subset\C$ is the set of all total recursive functions,
-  \item $\P\subset\K$ is the set of all prefix functions such that for
-    every infinite binary string the function is defined for a
-    prefix. This is the natural analogue of totality in the prefix
-  \item $\F\subset\P$ contains, for each finite set $S$, a computable
-    surjective function maping the binary sequences of length
-    $\lceil\log S\rceil$ onto $S$.
+  \item $\T\subset\C$ is the set of all total functions,
+  \item $\P\subset\K$ is the set of all complete, prefix-free functions,
+  \item $\F\subset\P$ contains, for every finite set $S$, a computable
+    surjective function mapping the binary sequences of length
+    $\lceil\log S\rceil$ onto $S$. In other words, this is the model class of finite sets, with uniform distributions.
   \end{itemize}
+\end{definition}
+\begin{definition}[Numberings]
+A numbering is an enumeration of $\C$, denoted as $\phi_1, \phi_2, \ldots$ or by a function $g: \N \to \C$.
+We fix one canonical numbering $g^\circ$, which works as an effective description, that is for any $i \in \N, y\in \B$, we can effectively compute $(g^\circ(i))(y)$.
+We call a numbering $g$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ as that for all $i$ $g^\circ(a(i)) = h(i)$ and  $g^\circ(i) = h(b(i))$.
 \end{definition}
@@
 \begin{definition}[Prefix encoding function]
-  A prefix encoding function $f:\{0,1\}^*\to\{0,1\}^*$ is an injective
+  A prefix encoding function $f:\B\to\B$ is an injective
   function that maps the binary strings to a prefix-free set. There
   are prefix encoding functions $g$ with the property that
-  $|g(x)|\le|x|+2\log(|x|+2)|$ \cite{TODO}. We fix such a function and
-  denote it by $\bar x=g(x)$.
+  $|g(x)|\le|x|+2\log(|x|+2)|$ \cite{li1993introduction}. We fix such a function and
+  denote it by $\br{x}=g(x)$.
 \end{definition}
@@
 \begin{definition}[Complexity]
-  Let $\phi_1,\phi_2,\ldots$ be a fixed acceptable enumeration of all partial
-  recursive functions, e.g. the enumeration described in \cite{TODO}.
-  Let $\Pi$ be a computable subset of the natural numbers such that
-  $\{\phi_i:i\in\Pi\}$ contains all computable prefix functions. Now
-  define Kolmogorov complexity by
+Let $\phi_1,\phi_2,\ldots$ be the canonical acceptable enumeration of $\C$.
+Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\} = \K$.
+Now define Kolmogorov complexity by
 \begin{align*}
@@
 \end{definition}
-This definition of $K$ is equivalent to the definition in the
-textbook~\cite{TODO}, but it is more convenient for defining the
-$K$-complexity of non-prefix functions:
+This definition of $K$ is uncommon.
+It is functionally equivalent to the standard definition as used in \cite{li1993introduction}, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).
 
-\begin{definition}[Representations]
-  A pair $(m,y)$ is a \emph{representation} for $x$ if $m$ is a model
-  and $m(y)=x$.
-\end{definition}
+We now have the basic tools with which we set up our generic sophistication function.
+We first define the space of two-part representations for a given $x$, and the code length that each represenrtation achieves:
 
-\begin{definition}[Code length functions]
-Define the following code length functions for the representations:
+\begin{definition}[Representations and code-lengths]
+A pair $(m,y)$ is a \emph{representation} for $x$ if $m$ is a model and $m(y)=x$.
+
+Define the following code-length functions for the representations:
 \begin{itemize}
 \item $L(m)$ is the length of a self-delimiting description of
-  the model $m$, i.e. it is the length function of  a prefix code,
-\item $L^m(x)=L(m)+\min\{|y|:m(y)=x\}$ measures the number of bits required to describe both the model $m$ and
-  the data $x$ using the model,
+  the model $m$, i.e. it is the length function of a prefix code for models. Each variant of sophistication must supply such a function.
+\item $L^m(x)=L(m)+\min\{|y|:m(y)=x\}$ measures the number of bits required to describe both the model $m$ and the data $x$ using the model.
 \item $L^\M(x)=\min\{L^m(x):m\in\M\}$ measures the number of bits
@@
 \end{definition}
-
+We can now define which of these representations express the data most efficiently:
 \begin{definition}[Candidate set]
-  The \emph{candidate models} for $x$ from model class $\M$ and
-  threshold function $\tau$ are
+The \emph{candidate models} for $x$ from model class $\M$ and threshold function $\tau$ are
 \[
@@
 \begin{definition}[Sophistication]
-  The model complexity, sophistication, or facticity, of a string $x$
-  given a model class $\M$ and threshold function $\tau$ is defined by
+  The model complexity, meaningful information, effective complexity, facticity or \emph{sophistication},  of a string $x$ given a model class $\M$ and threshold function $\tau$ is
   \[
-  S^\M_\tau(x):=\min\{L(m):m\in C^\M_\tau(x)\}.
-  \]
+  \s_\tau^{\M}(x):=\min\{L(m):m\in C^\M_\tau(x)\}\p
+  \]\label{definition:sophistication}
 \end{definition}
+
+For a function like sophistication to be meaningful, it must satisfy certain properties. We will check all currently published variants against the following minimal requirements:
+
+\begin{description}
+  \item[Invariance] Like the Kolmogorov complexity, the construction of a sophistication function makes certain ad-hoc choices, which are captured in the choice of an arbitrary universal Turing machine $U$ or an arbitrary numbering of $\C$.
+If we change what choice we make, the Kolmogorov complexity remains invariant up to a constant.
+This is the sole reason we allow ourselves to interpret the Komogorov complexioty as a property of the data, rather than simply the result of some computation on it.
+A similar invariance must hold for any sophistication function.
+   \item[Unboundedness] Both the sophistication and the residual information (roughly the complexity minus the sophistication) must not be bounded by a constant.
+If the sophistication is bekow a constant for all strings it cannot have much meaning.
+Also, if it is always withing a constant of the Kolmogorov complexity, it cannot correspond to the notion we hope to capture with it.
+   \item[Efficiency] The two-part representation $(m, y)$ that witnesses the sophistication (that is the one that reaches the minimum described by $\s^\M_\tau$) must reach the Kolmogorov complexity (or at least the difference should be insignificant).
+If there is an effective one-part representation that is much more efficient than the two-part representation that witnesses the sophistication, we cannot, in all honesty, claim that $(m, y)$ captures all the information in $x$.
+\end{description}
+
+Given the philosophical underpinnings of sophistication, these three requirements are the very least that we could expect.
+We will show that no current proposal satisfies all three.
+In the is no proposal for which invariance has been proved to be satisfied.
+This leads us to the following conjecture:
+
+\begin{conj}
+Let $\s_g$ be a sophistication function constructed as in Definition~\ref{definition:sophistication}, built on some acceptable numbering $g$, and some fixed model class $\M$ and threshold function $\tau$ (left out of the sub/superscript for the sake of clarity).
+\pb{dit kan mooier}
+
+There is no combination of $\M$ and $\tau$ such that the following all hold: 
+\begin{enumerate}
+  \item $\forall x : \left|\s_g(x) - \s_h(x)\right| \leq c$ for some constant $c$. (invariance)
+  \item $\forall c \exists x : \s_g(x) \geq c$ and $\forall c \exists x : L^\M(x) \s_g(x) \geq c$. (unboundedness)
+  \item $forall x : L^\M(x) = K(x)$ if the representations in $\M$ are self-delimiting, or $\forall x : L^\M(x) = C(x)$ otherwise. (efficiency)
+\end{enumerate}
+\end{conj}
+In more informal terms, we conjecture that no approach following the basic principle of sophistication---that is, model selection followed by measuring of the model complexity---can satisfy all three requirements.
+
+In the rest of this paper, we will analyse the existing variants one-by-one.
+We will conclude with the informal reasoning supporting this conjecture and consider its the consequences.
+What does this mean for the problem of meaningful information, what does it mean for algorithmic statistics and most importantly, what are the consequences for the idea of model selection in general?
 
@@
   can otherwise be chosen arbitrarily.
-\item Koppel uses monotone complexity and discusses infinite strings,
-  TODO figure out if and how this affects our results
+\item Koppel uses monotone complexity, and monotone functions and discusses 
+  infinite strings, TODO figure out if and how this affects our results.
 \end{itemize}
@@
     K(f) &:= \min\{K(i):\phi_i=f\}.
-  \end{align*}
+  \end{align*} \label{definition:model-complexity}
 \end{definition}
@@
 \end{proof}
+Note that we favoured simplicity of the proof over the strength of the bounds: the multiplicative constant $1/6$ can be improved with a more detailed treatment.
+\section{A world without sophistication}
 
-Note that we favoured simplicity of the proof over the strength of the bounds: the constants can be improved.
+We have conjectured that sophistication is a lost cause, and that the notion is impossible to formalize in a satisfying manner.
 
-\section{Notes}
+\pb{Procedural view: we cannot between dominating models.
+This may not be the aim of algorithmic statistics, but it suggests that it's kind of unreasonable to ask for model selection without repeated samples, or assumptions on the model class.}
+\pb{The problem with making a fixed choice.}
+\pb{The existence of balanced strings.}
+\pb{Total functions. They ruin efficiency and push sophistication into a land of unprovable claims.}
 
-Do we have the citation ``Paul M. B. Vit\'anyi: Meaningful Information. IEEE Transactions on Information Theory 52(10): 4617-4626 (2006)''?
+However, even if sophistication must be abandoned, that does not mean that the efforts surrounding it are lost. The are many elements of algorithmic statistics that do not rely 
+\pb{Non-stochastic strings are in some sense invariant}
+\pb{We can eliminate almost all models. Complexity is countevidence.No difficult business with total functions.}
 
-(From Koppel, Complexity, Depth and Sophistication:)
+Finally, the question remains how we can solve the matter of meaningful information without the ability to perform model selection. \pb{All candidates likely Take the finite set case. }
 
-``Sophistication is a generalization of the "H-function" or "minimal sufficient statistic" of Cover and Kolmogoroff, using the motonic complexity of Levin: is Cover the original source for the idea of a ``minimal sufficient statistic''?
-Also Koppel does indeed use monotonic complexity as in Levin.
-We need to know at least a little about the difference between monotonic and regular KC: is it less than a logarithmic additive term?
-If so, we can conveniently sweep it under the rug.
+\subsubsection*{\ackname}
 
-\section{Plan}
+This publication was supported by the Dutch national program COMMIT and by the the Netherlands eScience center.
 
-Idee van Sophistication: gebruik two-part coding om een scheiding te maken tussen structuur van de data (modelinformatie) en de ruis.
+\bibliographystyle{plain}
+\bibliography{facticity}
+
+\appendix
+
+\section{Stuff to move up or delete}
+
+\hide{
+
+\subsection{Plan}
 
@@
 \end{quotation}
-
-Adriaans takes this problem seriously and attempts to address it by imposing additional restrictions on the ordering of TMs implied by the used UTM \cite{adriaans2012facticity}.
-We are skeptical that such measures can effectively eliminate this problem without resorting to orderings of the TMs that are not admissible [TODO ref].
-
-Inherent property of the data: 
-\begin{itemize}
-\item Should be a two-part effective description of $x$
-\item Invariant: $\exists c:\forall x:|S^1(x)-S^2(x)|\le c$
-\item Both structure and noise are unbounded
-\end{itemize}
-
-Conjecture: er is geen functie van de data die aan bovenstaande eisen voldoet.
-
-\subsection{Review}
-Onderscheid op basis van drie properties:
-\begin{itemize}
-\item Model class: total / partial. Reden voor total: bij partial rec. is de UTM in het model, collapse, citaten. Nadeel: arbitraire beperking op de beschrijvingskracht, dus minder universeel
-\item Models are prefix functions / not prefix functions: viz $C$ vs $K$.
-Als wel prefix, dan gaat alles makkelijker kapot.
-Als niet prefix dan wordt de balans verstoord, het wordt moeilijk als de balans ook de andere kant op wordt verstoord door de modelklasse te beperken: die twee handicaps zijn moeilijk te vergelijken.
-\item Model length is prefix encoded index length / $K(f)$: compact numberings, bespiegeling over of het gat tussen $|\bar\imath|$ en $K(\phi_i)$ unbounded is.
-Wat is nu eigenlijk de two-part code precies, is dat $K$ als white box of is $K$ gebruikt voor het eerste deel?
-Ook beetje vreemd om in de eerste visie modelklassen te introduceren.
-\end{itemize}
-
-Bestaande aanpakken:
-\begin{itemize}
-
-\item Facticity: partial / niet prefix: bounded
-
-\item Paul's sophistication: Sufficient statistic: alles wat $K$-complexiteit haalt tot op slack $c$.
-Minimal sufficient statistic minimaliseert daarover de modelgrootte.
-Model bevat totale prefixfuncties.
-Noise is bounded.
-
-\item total / niet prefix, o.a.
-coarse sophistication.
-We geloven invariantie niet.
-Voor alle shallow strings bounded (nog niet bewezen, maar werkt door alles total te maken met vaste grote maar computable time bound)
-
-\end{itemize}
 
@@
 
-\section{Thoughts and questions}
+\subsection{Thoughts and questions}
 
@@
 
-\appendix
-\section{Dead darlings}
+\subsection{Dead darlings}
 
@@
 \end{proof}
-
-\bibliographystyle{plain}
-\bibliography{facticity}
+}
 

### 2014-12-17 Facticity/Facticity/paper.tex
iets

--- 2014-12-10 Facticity/Facticity/paper.tex
@@
 Since the model side is allowed unbounded decoding time, this invalidates effective complexity.}
+
+\begin{theorem}[Reduction of Effective Complexity to Structure
+  Function]
+  There is a constant $c$ such that, for every representation of $x$
+  consisting of an ensemble $E$ with $x\in E$ with entropy $H$,
+  yielding a total information of $K(E)+H$, there is a finite set $S$
+  with $x\in S$ such that $K(S)\le K(E)+c$ and $\log|S|\le H$.
+\end{theorem}
+\begin{proof}
+  First note that although the ensemble $E$ may itself be infinite, it
+  is ``typical'', and therefore a valid candidate representation, for
+  only a finite number of its elements, namely the set $S=\{y\in E:-\log
+  P(y)\le H\} = \{y:P(y)\ge 2^{-H}\}$. Since the sum of probabilities must
+  converge, this set cannot be infinite: since $1\ge\sum_{y\in S} P(y)\ge
+  \sum_{y\in S}2^{-H} = |S|2^{-H}$ so $|S|\le 2^H$.
+
+\end{proof}
 

### 2014-12-17 Facticity/Facticity/paper.tex

--- 2014-12-16 Facticity/Facticity/paper.tex
@@
 
+\begin{theorem}[Reduction of Effective Complexity to Structure
+  Function]
+  There is a constant $c$ such that, for every representation of $x$
+  consisting of an ensemble $E$ with $x\in E$ with entropy $H$,
+  yielding a total information of $K(E)+H$, there is a finite set $S$
+  with $x\in S$ such that $K(S)\le K(E)+c$ and $\log|S|\le H$.
+\end{theorem}
+\begin{proof}
+  First note that although the ensemble $E$ may itself be infinite, it
+  is ``typical'', and therefore a valid candidate representation, for
+  only a finite number of its elements, namely the set $S=\{y\in E:-\log
+  P(y)\le H\} = \{y:P(y)\ge 2^{-H}\}$. Since the sum of probabilities must
+  converge, this set cannot be infinite: since $1\ge\sum_{y\in S} P(y)\ge
+  \sum_{y\in S}2^{-H} = |S|2^{-H}$ so $|S|\le 2^H$.
+
+\end{proof}
+
 \subsection{Roll your own}

### 2014-12-17 Facticity/Facticity/paper.tex
gezeur van steven

@@
 
+% in introductie?
+%
+% model, model class
+% two-part code: sophistication is length of first part
+% kolmogorov complexity?
+% KC = two part vs KC in two part
+% collapse, restrictions on model class (eg total instead of partial recursive, see section XX)
+% conjecture
+
 \begin{document}
@@
 \begin{abstract}
+\sdr{Emphasize ongoing attempts to formalize meaningful information, ``recurring dream''. Also in the introduction} \sdr{Don't single out sophistication}
 Kolmogorov complexity is a sound expression of the information content of a binary string, but it leaves open the issue of \emph{meaningful information}.
@@
 \end{abstract}
+
+\sdr{Get conjecture right}
+
 \section{Introduction}
 
-\noindent Kolmogorov complexity provides an objective measurement of the information content of an object.
-Any arbitrary choices made in its construction can be changed, without significantly changing the outcome.
-This property is crucial for the measure's success: it allows us to see Kolmogorov complexity as a \emph{property of the data} rather than an arbitary function computed on it.
+\noindent Kolmogorov complexity is an objective measure of the information content of an object.
+Although its definition contains several arbitrary choices, these only affect its value up to a constant additive term.
+This property is crucial for the measure's success: it allows us to view Kolmogorov complexity as a \emph{property of the data} rather than an arbitary function computed on it.
 This, in turn, allows us to define the intrinsic entropy, or randomness, inherent in a single object expressed as a binary string, without any assumptions on the process that produced it.
-Unfortunately, this idea of complexity does not fit one-on-one to our intuitive idea of which objects are complex.
+However, this notion of complexity does not in all respects match our intuitive idea of which objects are complex, and which objects are not.
 
-Consider the following, illustrative example: an old-fashioned, analog television set which can be tuned to three frequencies.
-The first produces a blank screen, the second a regular program, and the third white noise.
-By the Kolmogorov complexity, these  broadcasts have increasing information content.
-The first is easiest to describe, the second contains a great deal of regularity, and the third cannot be compressed.
-However, in terms of meaningful information, information that we find interesting, the second signal is by far the richest.
+Imagine an old-fashioned, analog television set which can be tuned to three frequencies.
+The first produces a blank screen, the second an ordinary TV show, and the third white noise.
+Judging by their Kolmogorov complexity, these  broadcasts have increasing information content.
+The first is easiest to describe, the second is harder but still contains a great deal of regularity, and the third cannot be compressed.
+However, intuitively speaking the second signal clearly conveys more meaningful information than the other two.
 This notion has proved far harder to capture in formal terms than the simple notion of `raw' information content.
+\sdr{Emphasize recurring dream here}
 
@@
 The idea behind sophistication is to select a description for the data by two-part coding (describing a model first, and then the data in terms of the model), and to take the complexity of only the model part as a measure of the meaningful information.
-The assumption is that both simple and random data have small models; in the case of the simple data, a small model describes all regularity, and no further information is required, and in the case of the random data, the data itself is its shortest representation, and the model is a `print function' which simply reproduces its input.
+The assumption is that both simple and random data have small models: \sdr{the intuition is that small models suffice to describe both simple and random data, but not the meaningful information we are attempting to quantify.} in the case of the simple data, a small model describes all regularity, and no further information is required, and in the case of the random data, the data itself is its shortest representation, and the model is a `print function' which simply reproduces its input.
 If the dataset is more interesting, however, containing both regularity and randomness, its best representation will have a more complex model.
-A key assumption of the idea of sophistication is that we can perform objective model selection for a given string.
+\sdr{slightly repetitive} A key assumption of the idea of sophistication is that we can perform objective model selection for a given string.
 That is: there is a function that, given a string, selects a single unique model that can be used to compress this string optimally.
-Moreover, any elements of this function that are chosen ad hoc can be changed without significantly affecting the outcome.
+Moreover, any elements of this function that are chosen ad hoc can be changed without significantly affecting the size of the preferred model.
 
@@
 If a formalization of sophistication is ever to be held in the same regard as the Kolmogorov complexity, it must be subject to a similar principle of invariance.
+\sdr{I feel it's not about regard}
 
@@
 The complexity of the minimal sufficient statistic is the sophistication.
+\sdr{But I thought in some of our settings we use a different baseline than KC!}
 
 The intuition is that both very simple and very random data have small, simple programs.
-The strings with much structure and many interesting properties are those for which the best model is large.
-Two assumptions are crucial to this view:
+The strings with much structure and many interesting properties are those for which the best model is large.\sdr{Repetitive} Two assumptions are crucial to this view: \sdr{This bit should be more solid}
 \begin{itemize}
-  \item The two-part coding efficiently separates the structure and the noise into distinct parts.
+  \item The two-part coding efficiently \sdr{what does efficiently mean?} separates the structure \sdr{undefined} and the noise into distinct parts.
 In the optimal representation the model contains only `structural information' and the noise contains only the `uninteresting' information.
@@
 
-In addition to
-
-In this article we will show that most proposals do not achieve these goal.
+In this article we will show that most proposals do not achieve these goals.
 For one remaining case, we can offer no such proof, but there exists no proof to the contrary either.
-We offer several insights that suggest that the idea of an objective sophistication function is an unattainable ideal.
+\sdr{Say this more strongly: we believe that ``fixes'' of the problems with the approach are obfuscation rather than real solutions}  We offer several insights that suggest \sdr{insights that suggest stuff sounds very weak} that the idea of an objective sophistication function is an unattainable ideal.
 
@@
 We will first briefly discuss the different variants of sophistication in a historical context, to give the reader a simple, chronological overview, without discussing too much technical detail.
-In the main body of the paper, we will define each variant rigorously, and in non-chronological order.
+\sdr{We first provide a chronological overview of the variants of sophistication and discuss their relationships.} In the main body of the paper, we will define each variant rigorously, and in non-chronological order.
 
-The first suggestion that a representation of a string can be separated cleanly into a structural part, containing a description of its regularities, and a noise part, containing the residual information, came from Kolmogorov himself, in a talk given at a conference in Tallinn in 1973 (first committed to paper by Cover \cite{todo} and later more extensively discussed by Vitanyi \cite{todo}).
+The first suggestion that a representation of a string can be separated cleanly into a structural part, containing a description of its regularities, and a noise part, containing the residual information, came from Kolmogorov himself, in a talk given at a conference in Tallinn in 1973 (first committed to paper by Cover \cite{todo} and later more extensively discussed by Vit\'anyi \cite{todo}).
 Here, he proposed what has become known as the \emph{Kolmogorov structure function}.
 Briefly, this function operates on a space of two part representations for a given string $x$, consisting of some finite set $S$ (containing $x$) and the index of $x$ in $S$.
-The \emph{minimal sufficient statistic}, it was proposed, is the set with the lowest complexity for which the length of the two-part representation reaches the prefix-free Kolmogorov complexity.
+The \emph{minimal sufficient statistic} is the set with the lowest complexity for which the length of the two-part representation reaches the prefix-free Kolmogorov complexity.
 \pb{TODO: lowest randomness deficienty.}
@@
 Koppel \cite{} built on this notion, coining the phrase \emph{sophistication}.
-He changed the class of `models' from finite sets to total functions and used Levin's monotone complexity, instead of prefix-free Kolmogorov complexity.
+He changed the class of `models' from finite sets to total functions and used Levin's monotone complexity, instead of prefix-free Kolmogorov complexity \sdr{why?}.
 One interesting difference is that Koppel takes the description of the data under a variant of Kolmogorov complexity, which consists of a description of a model and that model's input, and takes the size of the model part as the sophistication.
 In contrast, Kolmogorov's minimal sufficient statistic is built on a second, two-part, representation in addition to the Kolmogorov complexity's one-part representation.
-This difference in perspective is subtle but relevant to the discussion ahead.
+This difference in perspective is subtle but relevant to the discussion ahead.\sdr{Good point but currently much too hard to understand}
 
-While Koppel expands the model class from finite sets to functions, he maintains a restriction that the functions be total.
+\sdr{Following bit is currently pretty vague.
+I think the collapse idea deserves a better introduction} While Koppel expands the model class from finite sets to functions, he maintains a restriction that the functions be total.
 This is a recurring theme: if the model class is expanded to all total functions, the the universal Turing machine itself becomes a model, so that we always have a two-part description with a small model, to within a constant of the Kolmogorov complexity.
-In most treatment this `collapse' to the universal model is avoided by excluding from the model class, using some restriction.
+In most treatment this `collapse' to the universal model is avoided by excluding it from the model class, using some restriction.
 This acts as a kind of force, pushing information into the model part of the two-part code.
@@
 Working from the perspective of statistical mechanics, they arrived at a notion remarkably similar to Sophistication.
-Effective complexity describes abinary string in terms of a probability distribution over finite strings (an `ensemble' in the terminology of Gell-Mann and Lloyd) and the (Shannon) entropy of the distribution.
+Effective complexity describes abinary string in terms of a probability distribution over finite strings (an `ensemble' in the terminology of Gell-Mann and Lloyd) \sdr{I really think the ensemble is just the set} and the (Shannon) entropy of the distribution.
 The reasoning is that, if the string is typical for the distribution, the entropy is sufficient to describe the string given the distribution, so that the length of the description of the distribution and the size of its entropy suffices as a two-part description.
@@
 hard to compute) information to the model part, and shallow information to the data.
-In a later treatment\cite{}, Gell-Mann and Lloyd select the decisive two-part representation from those achieving minimal code-length based on ``the average quantities judged to be important'', which would seem to make the value highly subjective and context-dependant.
+In a later treatment\cite{}, Gell-Mann and Lloyd select the decisive two-part representation from those achieving minimal code-length based on ``the average quantities judged to be important'', which would seem to make the value highly subjective and context-dependent.
 
-G\'acs, Tromp and Vit\'anyi \cite{}, built on the idea of the structure function to establish a framework of \emph{algorithmic statistics}, in the hope of founding a statistical theory not on probabilistic notions, but on combinatorial ones.
+G\'acs, Tromp and Vit\'anyi \cite{}, built on the idea of the structure function to establish a framework of \emph{algorithmic statistics}, as an attempt to found a statistical theory not on probabilistic notions, but on combinatorial ones.
 This echoes Kolmogorov's own hope for the theory of information:
@@
 \noindent Information theory must precede probability theory and not be based on it.
-By the very essence of this discipline, the foundations of information theory must have a finite combinatorial character
+By the very essence of this discipline, the foundations of information theory must have a finite combinatorial character \sdr{Source?}
 \end{quotation}
@@
 Like Koppel, Vit\'anyi restricts the model class to total functions, eliminating the universal model.
-He balances this force, by allowing the total two-part description to be non-self delimiting, so that both the representation with all information in the model part, and the representation with everything in the residual part are less efficient than the best two-part coding.
+He balances this force, by allowing the total two-part description to be non-self delimiting, so that both the representation with all information in the model part, and the representation with everything in the residual part are less efficient than the best two-part coding.\sdr{This balancing act is not yet clear to the reader}
 

### 2014-12-19 Facticity/Facticity/paper.tex
...

@@
 \begin{abstract}
-\sdr{Emphasize ongoing attempts to formalize meaningful information, ``recurring dream''. Also in the introduction} \sdr{Don't single out sophistication}
-Kolmogorov complexity is a sound expression of the information content of a binary string, but it leaves open the issue of \emph{meaningful information}.
-The strings with maximal information content are not the ones we would regard as interesting, or meaningful.
-One attempt to capture the idea of meaningful information is \emph{sophistication}.
-It is a simple idea with many attempted formalizations: describe the string efficiently with a two-part coding approach, and take the complexity of the model as a measure of the amount of meaningful information.
-We describe three requirements that any formalization must meet in order to satisfy the expectations that come with the idea of sophistication: invariance, unboundedness and efficiency.
-We show that no currently published proposal meets these requirements \pb{TODO}, and conjecture that no formalization of the idea of sophistication can hope to do so.
-We conclude by reviewing the consequences of this conjecture.
+Kolmogorov complexity is a sound expression of the information content of a binary string, but it clashes with our intuitive notions of information: the strings with maximal information content are not the ones we would regard as interesting, or meaningful.
+One attempt to capture this idea of `interesting' information is \emph{sophistication}.
+It is a simple idea with many attempted formalizations: describe the string efficiently with a two-part coding approach, and take the complexity of the model as a measure of the amount of interesting information.
+Kolmogorov complexity is invariant to the ad-hoc choices in its construction: change them, and the result stays the same, up to a constant.
+We review the history of sophistication critically, asking whether invariance exists for the proposed sophistication functions as well.
+In most cases, we can show that it doesn't.
+We conjecture that no formalization of sophistication can be invariant, and we briefly review the consequences of this conjecture.
 \end{abstract}
-
-\sdr{Get conjecture right}
 
@@
 This notion has proved far harder to capture in formal terms than the simple notion of `raw' information content.
-\sdr{Emphasize recurring dream here}
 
-One of the most frequently explored ideas for capturing the notion of meaningful information is \emph{sophistication}.
-The idea behind sophistication is to select a description for the data by two-part coding (describing a model first, and then the data in terms of the model), and to take the complexity of only the model part as a measure of the meaningful information.
-The assumption is that both simple and random data have small models: \sdr{the intuition is that small models suffice to describe both simple and random data, but not the meaningful information we are attempting to quantify.} in the case of the simple data, a small model describes all regularity, and no further information is required, and in the case of the random data, the data itself is its shortest representation, and the model is a `print function' which simply reproduces its input.
-If the dataset is more interesting, however, containing both regularity and randomness, its best representation will have a more complex model.
-\sdr{slightly repetitive} A key assumption of the idea of sophistication is that we can perform objective model selection for a given string.
-That is: there is a function that, given a string, selects a single unique model that can be used to compress this string optimally.
-Moreover, any elements of this function that are chosen ad hoc can be changed without significantly affecting the size of the preferred model.
+Over the years one idea in particular has been proposed in many different forms.
+We select a description for the data by two-part coding (describing a model first, and then the data in terms of the model), and we take the complexity of only the model part as a measure of the meaningful information.
+If we can find a single two-part description that we can objectively call the optimal description of the data, the complexity of the data part gives us an objective property of the data.
+While this idea has been proposed many times in many different guises, we will, for the purposes of this paper refer to it as \emph{sophistication}\footnotemark.
+The general problem of how to characterize the `richness' of the information in a string rather than the plain information content (to which sophistication is but one proposed solution), we will call the `problem of interesting information'.
 
-In the case of the Kolmogorov complexity, this holds true; we can change the universal Turing machine used in its definition and the complexities returned will not change by more than a constant.
-This is known as \emph{invariance}.
-Because Kolmogorov complexity is invariant, we can view it as an intrinsic property of the data, rather than simply an arbitrary function computed on it.
-If a formalization of sophistication is ever to be held in the same regard as the Kolmogorov complexity, it must be subject to a similar principle of invariance.
-\sdr{I feel it's not about regard}
+\footnotetext{As the historical overview shows, sophistication is only one name for this idea, and not the first.
+The alternatives, however, are more awkward, and it has been used by different authors for different realizations of the same principle.}
 
-The unifying principle over all the notions we review in this paper is simple.
-We start with an object expressed as a finite binary string.
-We represent this object, hoping to compress it, in two parts, the first is a model, and the second an `input' to this model.
-How the two parts are encoded, and what the class of models is are details that differ between variants.
-We know, however, that for successful representations, those that capture all structure in the string, the length of the two-part code should approach the Kolmogorov complexity.
-We take the set of all \emph{sufficient statistics}---the models of those representations that are `close enough' to the Kolmogorov complexity---and call the one with the lowest complexity the \emph{minimal sufficient statistic}.
-The complexity of the minimal sufficient statistic is the sophistication.
-\sdr{But I thought in some of our settings we use a different baseline than KC!}
+The intuitive idea of sophistication makes one crucial demand: we can find an objectively `true' two-part description.
+In the framework of Komogorov complexity two observers cannot disagree by more than a constant about the information content of a dataset.
+In formal terms: the Kolmogorov complexity of a dataset $x$ is defined as the shortest effective description for $x$ in some Turing complete description language.
+This description language is usually expressed as a universal Turing machine $U(\bar\imath y) = \phi_i(y)$ where $\phi_i$ is the $i$'th partial computable function in some effective enumeration, and the $\bar\imath$ is a self-delimiting encoding of $i$.
+If two people disagree about the shortest effective description of $x$, we can interpret this as them using different description languages, ie.
+different universal Turing machines.
+The crucial aspect of Kolmogorov complexity, the starting theorem that makes it interesting as a function at all, is \emph{invariance}: two people using different description languages will get the same value for a string's Kolmogorov complexity, albeit up to a fixed constant.
+Without invariance, Kolmogorov complexity would be an arbitrary function.
+With invariance, we can be sure that we are talking about an intrinsic property of the data.
+The same consideration must be given to each proposal for formalising sophistication: is it invariant?
 
-The intuition is that both very simple and very random data have small, simple programs.
-The strings with much structure and many interesting properties are those for which the best model is large.\sdr{Repetitive} Two assumptions are crucial to this view: \sdr{This bit should be more solid}
-\begin{itemize}
-  \item The two-part coding efficiently \sdr{what does efficiently mean?} separates the structure \sdr{undefined} and the noise into distinct parts.
-In the optimal representation the model contains only `structural information' and the noise contains only the `uninteresting' information.
-  \item This principle is invariant. If we change, for instance, the universal Turing machine used to unpack our descriptions, the minimal sufficient statistic will not change significantly.
-\end{itemize}
+There are many approaches to sophistication, but the following pattern is common to all of them.
+We start with a space of \emph{models}.
+There can be finite sets, computable probability distributions or simple functions from strings to strings.
+The function perspective is the most generic, all other model classes can be translated to it.
+The approach to sophistication is to find a representation for a dataset $x$ in two parts: a computable function $f$ and and input $y$, such that $f(y) = x$.
+As with Kolmogorov complexity, the shorter a description is, the better.
+We take the shortest description and those close to it, and consider them \emph{candidates}.
+How close a description must be to the optimum differs between variants, it is often a pre-determined constant value.
+We then choose a single description from the candidate set and return its \emph{model complexity} as the sophistication.
+The method for determining model complexity varies.
+In almost all variants the candidate with the smallest model complexity determines the sophistication.
 
-In this article we will show that most proposals do not achieve these goals.
-For one remaining case, we can offer no such proof, but there exists no proof to the contrary either.
-\sdr{Say this more strongly: we believe that ``fixes'' of the problems with the approach are obfuscation rather than real solutions}  We offer several insights that suggest \sdr{insights that suggest stuff sounds very weak} that the idea of an objective sophistication function is an unattainable ideal.
+The implicit assumption behind sophistication is that this approach will separate the description of the string into a sturctural part, representing its regularities, and a stochastic part representing its non-deterministic properties.
+In other words, the string is cast as a random choice from a model for which it is `typical'.
+
+In this article we will show that most proposals are not invariant.
+For one remaining case, we can offer no such proof, but no proof of invariance exists either.
+This is likely due to the complex nature of the model class chosen, and the difficulty of resolving the matter of invariance shows how complex such a sophistication measure would be to build on, even if it did prove to be invariant.
 
@@
 
-We will first briefly discuss the different variants of sophistication in a historical context, to give the reader a simple, chronological overview, without discussing too much technical detail.
-\sdr{We first provide a chronological overview of the variants of sophistication and discuss their relationships.} In the main body of the paper, we will define each variant rigorously, and in non-chronological order.
+We first provide a chronological overview of the variants of sophistication. In the body of the paper, we will define each variant rigorously, and in non-chronological order.
 
@@
 Koppel \cite{} built on this notion, coining the phrase \emph{sophistication}.
-He changed the class of `models' from finite sets to total functions and used Levin's monotone complexity, instead of prefix-free Kolmogorov complexity \sdr{why?}.
-One interesting difference is that Koppel takes the description of the data under a variant of Kolmogorov complexity, which consists of a description of a model and that model's input, and takes the size of the model part as the sophistication.
-In contrast, Kolmogorov's minimal sufficient statistic is built on a second, two-part, representation in addition to the Kolmogorov complexity's one-part representation.
-This difference in perspective is subtle but relevant to the discussion ahead.\sdr{Good point but currently much too hard to understand}
+He changed the class of `models' from finite sets to total functions and used Levin's monotone complexity, instead of prefix-free Kolmogorov complexity.
+This is likely to extend the theory to infinte strings, an approach which has been abandoned by subsequent authors.
 
-\sdr{Following bit is currently pretty vague.
-I think the collapse idea deserves a better introduction} While Koppel expands the model class from finite sets to functions, he maintains a restriction that the functions be total.
-This is a recurring theme: if the model class is expanded to all total functions, the the universal Turing machine itself becomes a model, so that we always have a two-part description with a small model, to within a constant of the Kolmogorov complexity.
-In most treatment this `collapse' to the universal model is avoided by excluding it from the model class, using some restriction.
+While Koppel expands the model class from finite sets to functions, he maintains a restriction that the functions be total.
+This is a recurring theme: if the model class is expanded to all computable functions, the the universal Turing machine $U$ itself becomes a model.
+If we assume that $U$ computes the function $\phi_u$ in the standard enumeration of computable functions, and that $\bar\imath y$ is a shortest description for $x$ on $U$, then $\br{u}\br{\imath}y$ is also a description for $x$.
+This description has a constant model size ($|\br{u}|$), ie.
+every $x$ has constant sophistication.
+This `collapse' to a universal modle is the reason that most treatments of sophistication constrain the model class to exclude the universal Turing machine.
 This acts as a kind of force, pushing information into the model part of the two-part code.
-As we shall see, defining sophistication quickly becomes a balancing act, trying to match two forces which by themselves push information one way or the other.
+As we shall see, defining sophistication quickly becomes a balancing act, trying to match two forces which by themselves would push information one way or the other.
 
-In 1996 Gell-Mann and Lloyd introduced `Effective complexity'\cite{todo}.
-Working from the perspective of statistical mechanics, they arrived at a notion remarkably similar to Sophistication.
-Effective complexity describes abinary string in terms of a probability distribution over finite strings (an `ensemble' in the terminology of Gell-Mann and Lloyd) \sdr{I really think the ensemble is just the set} and the (Shannon) entropy of the distribution.
+In 1996 Gell-Mann and Lloyd introduced `Effective complexity'\cite{gellmann1996information}.
+Working from the perspective of statistical mechanics, they arrived (largely independently, it seems) at a notion that firts the mold of sophistication.
+Effective complexity describes a binary string in terms of a probability distribution over finite strings (an `ensemble' in the terminology of Gell-Mann and Lloyd) and the entropy of the distribution.
 The reasoning is that, if the string is typical for the distribution, the entropy is sufficient to describe the string given the distribution, so that the length of the description of the distribution and the size of its entropy suffices as a two-part description.
@@
 hard to compute) information to the model part, and shallow information to the data.
-In a later treatment\cite{}, Gell-Mann and Lloyd select the decisive two-part representation from those achieving minimal code-length based on ``the average quantities judged to be important'', which would seem to make the value highly subjective and context-dependent.
+In a later treatment\cite{gell2004nonextensive}, Gell-Mann and Lloyd select the decisive two-part representation from the candidates based on ``the average quantities judged to be important'', which seems to be a concious choice to make the measure subjective and context-dependant (sacrificing invariance).
 
@@
 \noindent Information theory must precede probability theory and not be based on it.
-By the very essence of this discipline, the foundations of information theory must have a finite combinatorial character \sdr{Source?}
+By the very essence of this discipline, the foundations of information theory must have a finite combinatorial character.
+\cite{kolmogorov1983combinatorial}
 \end{quotation}
@@
 Like Koppel, Vit\'anyi restricts the model class to total functions, eliminating the universal model.
-He balances this force, by allowing the total two-part description to be non-self delimiting, so that both the representation with all information in the model part, and the representation with everything in the residual part are less efficient than the best two-part coding.\sdr{This balancing act is not yet clear to the reader}
+He balances this force by allowing the total two-part description to be non-self delimiting, so that both the representation with all information in the model part, and the representation with everything in the residual part are less efficient than the best two-part coding.\sdr{This balancing act is not yet clear to the reader}
 
@@
 
-Given the philosophical underpinnings of sophistication, these three requirements are the very least that we could expect.
-We will show that no current proposal satisfies all three.
-In the is no proposal for which invariance has been proved to be satisfied.
-This leads us to the following conjecture:
+Given the philosophical underpinnings of sophistication, these three requirements are the very least that we could expect. We will show that no current proposal satisfies all three.
+
+For all variants but one, we have shown that all measures that are invariant are bounded.
+This leads us to the following conjecture.
+Let $\gamma(n)$ be a function representing the extent to which we care about differences between functions.
+That is, we will call two functions $f$ and $g$ $\gamma$-equal if $|f(x) - g(x)| \in O(\gamma(f(x)))$.
 
 \begin{conj}
-Let $\s_g$ be a sophistication function constructed as in Definition~\ref{definition:sophistication}, built on some acceptable numbering $g$, and some fixed model class $\M$ and threshold function $\tau$ (left out of the sub/superscript for the sake of clarity).
-\pb{dit kan mooier}
+Let $\s_g$ be a sophistication function constructed as in Definition~\ref{definition:sophistication}, built on some acceptable numbering $g$, and some fixed model class $\M$ and threshold function $\tau$ (omitted from the sub/superscript for the sake of clarity).
 
-There is no combination of $\M$ and $\tau$ such that the following all hold: 
-\begin{enumerate}
-  \item $\forall x : \left|\s_g(x) - \s_h(x)\right| \leq c$ for some constant $c$. (invariance)
-  \item $\forall c \exists x : \s_g(x) \geq c$ and $\forall c \exists x : L^\M(x) \s_g(x) \geq c$. (unboundedness)
-  \item $forall x : L^\M(x) = K(x)$ if the representations in $\M$ are self-delimiting, or $\forall x : L^\M(x) = C(x)$ otherwise. (efficiency)
-\end{enumerate}
+For any sophistication function so constructed, and any granularity $\gamma$ we have that for any two acceptable numberings $g$ and $h$, either:
+\begin{itemize}
+  \item there exist $s_g$ and $s_h$ which are not are $\gamma$-equal (ie $s$ is not invariant),
+  \item $s_g$ is $\gamma$-equal to $0$ or to the Kolmogorov complexity.
+\end{itemize}
 \end{conj}
-In more informal terms, we conjecture that no approach following the basic principle of sophistication---that is, model selection followed by measuring of the model complexity---can satisfy all three requirements.
+In these terms, all instantiations of the Kolmogorov complexity, using different Universal Turing machines are $1$-equal.
+It may be unreasonable to require the sophistication to attain the same level of invariance (a log-invariant sophistication could still be useful).
+This conjecture allows such relaxations: if it can be shown to be true, sophistication must be considered fundamentally flawed.
+We hope that this conjecture will prompt authors to address the matter of invariance ciritcally, and to consider the possibility that sophistication and its ilk are an untenable proposition.
 
@@
   infinite strings, TODO figure out if and how this affects our results.
+  \item Koppel has a `machine-independence result', but not one that does what we want.
 \end{itemize}
@@
 \begin{itemize}
-\item Model class is \emph{ensembles}. Equivalent to $\K$.
+\item Model class is \emph{ensembles}. Equivalent to $\K$. Subsequent literature suggests that ensembles have finite support. 
 \item Description method for models is $L(m) = K(m)$ (the minimization is not mentioned explicitly, but we'll give them the benefit of the doubt).

### 2014-12-19 Facticity/Facticity/paper.tex
merge
stevens gezeur weer

@@
 % two-part code: sophistication is length of first part
+% KC = two part vs KC in two part 
 % kolmogorov complexity?
-% KC = two part vs KC in two part
 % collapse, restrictions on model class (eg total instead of partial recursive, see section XX)
@@
 Over the years one idea in particular has been proposed in many different forms.
-We select a description for the data by two-part coding (describing a model first, and then the data in terms of the model), and we take the complexity of only the model part as a measure of the meaningful information.
+We select a description for the data by two-part coding (describing a model first, and then the data in terms of the model), and we take the complexity \sdr{number of bits required for} of only the model part as a measure of the meaningful information.
 If we can find a single two-part description that we can objectively call the optimal description of the data, the complexity of the data part gives us an objective property of the data.
-While this idea has been proposed many times in many different guises, we will, for the purposes of this paper refer to it as \emph{sophistication}\footnotemark.
+While this idea has been proposed many times in many different guises, \sdr{Perhaps say here already that these many attempts indicate that many people believe something can be done, and at the same time that it is not easy to get it right.
+Perhaps use word ``dream''} we will, for the purposes of this paper refer to it as \emph{sophistication}\footnotemark.
 The general problem of how to characterize the `richness' of the information in a string rather than the plain information content (to which sophistication is but one proposed solution), we will call the `problem of interesting information'.
+\sdr{I'd not define it as a term here, but rather say that there are other approaches such as blah blah, but that we are not gonna talk about those in this paper.
+That way, we're done with it}
 
@@
 The intuitive idea of sophistication makes one crucial demand: we can find an objectively `true' two-part description.
-In the framework of Komogorov complexity two observers cannot disagree by more than a constant about the information content of a dataset.
+\sdr{This is too strong, it is still okay if many descriptions are identified as long as they have more or less the same model size.
+Also we don't really need to make truth assumptions} In the framework of Komogorov complexity two observers cannot disagree by more than a constant about the information content of a dataset.
 In formal terms: the Kolmogorov complexity of a dataset $x$ is defined as the shortest effective description for $x$ in some Turing complete description language.
@@
 different universal Turing machines.
-The crucial aspect of Kolmogorov complexity, the starting theorem that makes it interesting as a function at all, is \emph{invariance}: two people using different description languages will get the same value for a string's Kolmogorov complexity, albeit up to a fixed constant.
+The crucial aspect of Kolmogorov complexity, the starting theorem that makes it interesting as a function at all, is \emph{invariance}: two people using different description languages will get \sdr{a complexities that are within a fixed constant of each other for \emph{all} strings} the same value for a string's Kolmogorov complexity, albeit up to a fixed constant.
 Without invariance, Kolmogorov complexity would be an arbitrary function.
@@
 The function perspective is the most generic, all other model classes can be translated to it.
-The approach to sophistication is to find a representation for a dataset $x$ in two parts: a computable function $f$ and and input $y$, such that $f(y) = x$.
+The approach to sophistication \sdr{Raar begin van de zin} is to find a representation for a dataset $x$ in two parts: a computable function $f$ and and input $y$, such that $f(y) = x$.
 As with Kolmogorov complexity, the shorter a description is, the better.
 We take the shortest description and those close to it, and consider them \emph{candidates}.
-How close a description must be to the optimum differs between variants, it is often a pre-determined constant value.
+\sdr{We call the set of descriptions that are not much longer than the shortest one \emph{candidates}.} How close a description must be to the optimum differs between variants \sdr{varies}, it is often \sdr{but as a concrete example it could be} a pre-determined constant value.
 We then choose a single description from the candidate set and return its \emph{model complexity} as the sophistication.
 The method for determining model complexity varies.
-In almost all variants the candidate with the smallest model complexity determines the sophistication.
+\sdr{Conceptually it is always the shortest program that implements the model, but technically this is defined in several ways, some of which are buggy.
+I feel we should not suggest that the notion of model complexity is very different in all approaches, because they \emph{try} to do the same thing.} In almost all variants the candidate with the smallest model complexity determines the sophistication.
 
@@
 In this article we will show that most proposals are not invariant.
-For one remaining case, we can offer no such proof, but no proof of invariance exists either.
+\sdr{Ik wil dit nog steeds graag sterker gesteld hebben.} For one remaining case, we can offer no such proof, but no proof of invariance exists either.
 This is likely due to the complex nature of the model class chosen, and the difficulty of resolving the matter of invariance shows how complex such a sophistication measure would be to build on, even if it did prove to be invariant.
@@
 
-We first provide a chronological overview of the variants of sophistication. In the body of the paper, we will define each variant rigorously, and in non-chronological order.
+We first provide a chronological overview of the variants of sophistication.
+In the body of the paper, we will define each variant rigorously \sdr{Don't emphasize distinction between formal and informal too much, it's not interesting.
+Everything we say should be true and clear anyway.
+Just say we'll provide more detailed descriptions.}, and in non-chronological order.
 
-The first suggestion that a representation of a string can be separated cleanly into a structural part, containing a description of its regularities, and a noise part, containing the residual information, came from Kolmogorov himself, in a talk given at a conference in Tallinn in 1973 (first committed to paper by Cover \cite{todo} and later more extensively discussed by Vit\'anyi \cite{todo}).
+The first suggestion that a representation of a string can be separated cleanly into a structural part, containing a description of its regularities, and a noise part, containing the residual information, came from Kolmogorov himself \sdr{remove hint of hero worship}, in a talk given at a conference in Tallinn in 1973 (first committed to paper by Cover \cite{todo} and later more extensively discussed by Vit\'anyi \cite{todo}).
 Here, he proposed what has become known as the \emph{Kolmogorov structure function}.
-Briefly, this function operates on a space of two part representations for a given string $x$, consisting of some finite set $S$ (containing $x$) and the index of $x$ in $S$.
-The \emph{minimal sufficient statistic} is the set with the lowest complexity for which the length of the two-part representation reaches the prefix-free Kolmogorov complexity.
+Briefly, this function operates on a space \sdr{what does ``operates on a space'' mean?
+It is not the domain of the function.
+Perhaps simplify, just say, all finite sets of binary sequences are interpreted as models} of two part representations for a given string $x$, consisting of some finite set $S$ (containing $x$) and the index of $x$ in $S$.
+The \emph{minimal sufficient statistic} is the set with the lowest \sdr{model?} complexity for which the length of the two-part representation reaches the prefix-free Kolmogorov complexity.
 \pb{TODO: lowest randomness deficienty.}
@@
 He changed the class of `models' from finite sets to total functions and used Levin's monotone complexity, instead of prefix-free Kolmogorov complexity.
-This is likely to extend the theory to infinte strings, an approach which has been abandoned by subsequent authors.
+This is likely to \sdr{This was for the purpose of extending} extend the theory to infinite strings, an approach \sdr{attempt} which has been abandoned by subsequent authors.
+\sdr{Dit wordt niet expliciet gemotiveerd maar is waarschijnlijk voor inf strings.
+In dit paper kijken we alleen naar eindige objecten}
 
-While Koppel expands the model class from finite sets to functions, he maintains a restriction that the functions be total.
-This is a recurring theme: if the model class is expanded to all computable functions, the the universal Turing machine $U$ itself becomes a model.
+While Koppel expands the model class from finite sets to functions, he maintains a \sdr{does impose the} restriction that the functions be total.
+\sdr{This collapse problem still deserves a less off-handed introduction somehow} This is a recurring theme: if the model class is expanded to all computable functions, then the universal Turing machine $U$ itself becomes a model.
 If we assume that $U$ computes the function $\phi_u$ in the standard enumeration of computable functions, and that $\bar\imath y$ is a shortest description for $x$ on $U$, then $\br{u}\br{\imath}y$ is also a description for $x$.
@@
 every $x$ has constant sophistication.
-This `collapse' to a universal modle is the reason that most treatments of sophistication constrain the model class to exclude the universal Turing machine.
-This acts as a kind of force, pushing information into the model part of the two-part code.
-As we shall see, defining sophistication quickly becomes a balancing act, trying to match two forces which by themselves would push information one way or the other.
+This `collapse' to a universal model is the reason that most treatments of sophistication constrain the model class to exclude the universal Turing machine.
+This acts as a kind of force \sdr{still don't like explicit force metaphor, although I do like pushing information from side to side} pushing information into the model part of the two-part code.
+As we shall see, defining sophistication quickly becomes \sdr{devolves into?} a balancing act, trying to match two forces which by themselves would push information one way or the other.
 
 In 1996 Gell-Mann and Lloyd introduced `Effective complexity'\cite{gellmann1996information}.
-Working from the perspective of statistical mechanics, they arrived (largely independently, it seems) at a notion that firts the mold of sophistication.
+Working from the perspective of statistical mechanics, they arrived (largely independently, it seems) at a notion that fits the mold of sophistication.
 Effective complexity describes a binary string in terms of a probability distribution over finite strings (an `ensemble' in the terminology of Gell-Mann and Lloyd) and the entropy of the distribution.
-The reasoning is that, if the string is typical for the distribution, the entropy is sufficient to describe the string given the distribution, so that the length of the description of the distribution and the size of its entropy suffices as a two-part description.
-The two-part description which determines the effective complexity is chosen by placing a time bound (taken from the singleton description) on the computation of the string from the distribution.
-The hope is that this will move deep (ie.
+\sdr{Somehow sounds a bit unfocused} The reasoning is that, if the string is typical for the distribution, the entropy is sufficient to describe the string given the distribution, so that the length of the description of the distribution and the size of its entropy suffices as a two-part description.
+The two-part description that determines the effective complexity is chosen by placing a time bound (taken from the singleton description) on the computation of the string from the distribution.
+The hope is that this will move \sdr{just say hard to compute} deep (ie.
 hard to compute) information to the model part, and shallow information to the data.
@@
 In a later paper\cite{gacs2001algorithmic}, Vit\'anyi specifically studies the role of the minimal sufficient statistic as an expression of the `meaningful' information of a string (taking the name sophistication from Koppel).
-Like Koppel, Vit\'anyi restricts the model class to total functions, eliminating the universal model.
-He balances this force by allowing the total two-part description to be non-self delimiting, so that both the representation with all information in the model part, and the representation with everything in the residual part are less efficient than the best two-part coding.\sdr{This balancing act is not yet clear to the reader}
+Like Koppel, Vit\'anyi \sdr{et al?} restricts the model class to total functions, eliminating the universal model.
+He balances this force \sdr{I'm against the use of force} by allowing the total two-part description to be non-self delimiting, so that both the representation with all information in the model part, and the representation with everything in the residual part are less efficient than the best two-part coding.\sdr{Is this balancing act made sufficiently clear?
+I'm not happy yet about the force metaphor and explanation}
 
 Antunes et al.
-\cite{} note that the constant cutoff used to separate candidate two-part descriptions from those deemed to be too inefficient, is a source of instability.
-Small changes to this constant could shift the result sophistication by arbitrary amounts.
+\cite{} note that the constant cutoff used to separate candidate two-part descriptions from those deemed to be too inefficient, is a source of instability.\sdr{Actually their cutoff is logarithmic isn't it?
+Also probably very hard to understand exactly what cutoff you mean to a first time reader} Small changes to this constant could shift the resulting sophistication by arbitrary amounts.
 They introduce a more elaborate criterion for the cutoff, calling this variant \emph{coarse sophistication}.

### 2015-01-05 Facticity/Facticity/paper.tex
....

--- 2014-12-19 Facticity/Facticity/paper.tex
@@
 \newcommand{\s}{\tn{soph}}
+\newcommand{\num}[1]{#1_1, #1_2, \ldots}
 
@@
 Kolmogorov complexity is a sound expression of the information content of a binary string, but it clashes with our intuitive notions of information: the strings with maximal information content are not the ones we would regard as interesting, or meaningful.
-One attempt to capture this idea of `interesting' information is \emph{sophistication}.
+One attempt to capture this idea of `useful' information is \emph{sophistication}.
 It is a simple idea with many attempted formalizations: describe the string efficiently with a two-part coding approach, and take the complexity of the model as a measure of the amount of interesting information.
-Kolmogorov complexity is invariant to the ad-hoc choices in its construction: change them, and the result stays the same, up to a constant.
+Kolmogorov complexity is invariant to the ad-hoc choices in its construction: change them, and the result will not change more than a constant.
 We review the history of sophistication critically, asking whether invariance exists for the proposed sophistication functions as well.
@@
 Although its definition contains several arbitrary choices, these only affect its value up to a constant additive term.
-This property is crucial for the measure's success: it allows us to view Kolmogorov complexity as a \emph{property of the data} rather than an arbitary function computed on it.
+This property is crucial for the measure's success: it allows us to view Kolmogorov complexity as a \emph{property of the data} rather than an arbitrary function computed on it.
 This, in turn, allows us to define the intrinsic entropy, or randomness, inherent in a single object expressed as a binary string, without any assumptions on the process that produced it.
@@
 
-Over the years one idea in particular has been proposed in many different forms.
-We select a description for the data by two-part coding (describing a model first, and then the data in terms of the model), and we take the complexity of only the model part as a measure of the meaningful information.
-If we can find a single two-part description that we can objectively call the optimal description of the data, the complexity of the data part gives us an objective property of the data.
+Over the years one solution in particular has been proposed in many different forms.
+We select a description for the data by two-part coding; describing a model first, and then the data in terms of the model.
+The size in bits, used to describe the model is taken as a measure of the useful information in our string.
 While this idea has been proposed many times in many different guises, we will, for the purposes of this paper refer to it as \emph{sophistication}\footnotemark.
-The general problem of how to characterize the `richness' of the information in a string rather than the plain information content (to which sophistication is but one proposed solution), we will call the `problem of interesting information'.
 
 \footnotetext{As the historical overview shows, sophistication is only one name for this idea, and not the first.
-The alternatives, however, are more awkward, and it has been used by different authors for different realizations of the same principle.}
+The alternatives, however, are more awkward, and sophistication has been used by different authors for different realizations of the same principle.}
 
-The intuitive idea of sophistication makes one crucial demand: we can find an objectively `true' two-part description.
-In the framework of Komogorov complexity two observers cannot disagree by more than a constant about the information content of a dataset.
+The intuitive idea of sophistication makes one crucial demand: we can find an objectively value for the model complexity of a string, either by having a single objectively optimal two-part description, or a set of descriptions with equivalent model complexity.
+In the framework of Kolmogorov complexity, two observers cannot disagree by more than a constant about the information content of a dataset.
 In formal terms: the Kolmogorov complexity of a dataset $x$ is defined as the shortest effective description for $x$ in some Turing complete description language.
@@
 The method for determining model complexity varies.
-In almost all variants the candidate with the smallest model complexity determines the sophistication.
+In almost all variants, the candidate with the smallest model complexity determines the sophistication.
 
-The implicit assumption behind sophistication is that this approach will separate the description of the string into a sturctural part, representing its regularities, and a stochastic part representing its non-deterministic properties.
+The implicit assumption behind sophistication is that this approach will separate the description of the string into a structural part, representing its regularities, and a stochastic part representing its non-deterministic properties.
 In other words, the string is cast as a random choice from a model for which it is `typical'.
@@
 
-We first provide a chronological overview of the variants of sophistication. In the body of the paper, we will define each variant rigorously, and in non-chronological order.
+We first provide a chronological overview of the existing variants of sophistication. In the body of the paper, we will define each variant rigorously, and in non-chronological order.
 
-The first suggestion that a representation of a string can be separated cleanly into a structural part, containing a description of its regularities, and a noise part, containing the residual information, came from Kolmogorov himself, in a talk given at a conference in Tallinn in 1973 (first committed to paper by Cover \cite{todo} and later more extensively discussed by Vit\'anyi \cite{todo}).
+The first suggestion that a representation of a string can be separated cleanly into an encoding of its regular and its stochastic proeperties came from Kolmogorov himself, in a talk given at a conference in Tallinn in 1973 (first committed to paper by Cover \cite{cover1985kolmogorov} and later discussed more extensively discussed by Vereshchagin and Vit\'anyi \cite{vereshchagin2004kolmogorov}).
 Here, he proposed what has become known as the \emph{Kolmogorov structure function}.
@@
 The \emph{minimal sufficient statistic} is the set with the lowest complexity for which the length of the two-part representation reaches the prefix-free Kolmogorov complexity.
-\pb{TODO: lowest randomness deficienty.}
+It was first shown by Shen \cite{shen1983concept} that there are strings for which the best model cannot be small.
 
-Koppel \cite{} built on this notion, coining the phrase \emph{sophistication}.
-He changed the class of `models' from finite sets to total functions and used Levin's monotone complexity, instead of prefix-free Kolmogorov complexity.
-This is likely to extend the theory to infinte strings, an approach which has been abandoned by subsequent authors.
+Koppel \cite{koppelSoph1988,koppel1991almost} built on this notion, coining the phrase \emph{sophistication}.
+He changed the class of `models' from finite sets to total functions and used monotone functions, instead of prefix-free Kolmogorov complexity.
+We expect that this change was made to extend the theory to infinite strings, an approach which has been abandoned by subsequent authors.
+We will only consider the finite version (based on regular functions) in this paper.
 
-While Koppel expands the model class from finite sets to functions, he maintains a restriction that the functions be total.
-This is a recurring theme: if the model class is expanded to all computable functions, the the universal Turing machine $U$ itself becomes a model.
-If we assume that $U$ computes the function $\phi_u$ in the standard enumeration of computable functions, and that $\bar\imath y$ is a shortest description for $x$ on $U$, then $\br{u}\br{\imath}y$ is also a description for $x$.
+While Koppel expands the model class from finite sets to functions,he does require that the functions be total.
+This is a recurring theme: if the model class is expanded to all computable functions, the universal Turing machine $U$ itself becomes a model.
+If we assume that $U$ computes the function $\phi_u$ in the standard enumeration of computable functions, and that $\bar\imath y$ is a shortest description for $x$ on $U$ (ie.
+$U(\bar\imath y) = x)$), then $\br{u}\br{\imath}y$ is also a description for $x$ on $\bar u$.
 This description has a constant model size ($|\br{u}|$), ie.
 every $x$ has constant sophistication.
-This `collapse' to a universal modle is the reason that most treatments of sophistication constrain the model class to exclude the universal Turing machine.
-This acts as a kind of force, pushing information into the model part of the two-part code.
-As we shall see, defining sophistication quickly becomes a balancing act, trying to match two forces which by themselves would push information one way or the other.
+This `collapse' to a universal model is the reason that most treatments of sophistication constrain the model class to exclude the universal Turing machine.
+This makes it more expensive to move information out of the model.
+As we shall see, defining sophistication quickly becomes a balancing act, forcing information one way or the other.
 
-In 1996 Gell-Mann and Lloyd introduced `Effective complexity'\cite{gellmann1996information}.
-Working from the perspective of statistical mechanics, they arrived (largely independently, it seems) at a notion that firts the mold of sophistication.
+In 1996 Gell-Mann and Lloyd introduced `Effective complexity' \cite{gellmann1996information}.
+Working from the perspective of statistical mechanics, they arrived (largely independently, it seems) at a notion that fits the mold of sophistication.
 Effective complexity describes a binary string in terms of a probability distribution over finite strings (an `ensemble' in the terminology of Gell-Mann and Lloyd) and the entropy of the distribution.
@@
 hard to compute) information to the model part, and shallow information to the data.
-In a later treatment\cite{gell2004nonextensive}, Gell-Mann and Lloyd select the decisive two-part representation from the candidates based on ``the average quantities judged to be important'', which seems to be a concious choice to make the measure subjective and context-dependant (sacrificing invariance).
+In a later treatment\cite{gell2004nonextensive}, Gell-Mann and Lloyd select the decisive two-part representation from the candidates based on ``the average quantities judged to be important'', which seems to be a conscious choice to make the measure subjective and context-dependant (sacrificing invariance).
 
-G\'acs, Tromp and Vit\'anyi \cite{}, built on the idea of the structure function to establish a framework of \emph{algorithmic statistics}, as an attempt to found a statistical theory not on probabilistic notions, but on combinatorial ones.
+G\'acs, Tromp and Vit\'anyi \cite{gacs2001algorithmic}, built on the idea of the structure function to establish a framework of \emph{algorithmic statistics}, as an attempt to found a statistical theory not on probabilistic notions, but on combinatorial ones.
 This echoes Kolmogorov's own hope for the theory of information:
@@
 Algorithmic statistics studies binary strings and their possible two-part representations.
-In a later paper\cite{gacs2001algorithmic}, Vit\'anyi specifically studies the role of the minimal sufficient statistic as an expression of the `meaningful' information of a string (taking the name sophistication from Koppel).
+In a later paper \cite{gacs2001algorithmic}, Vit\'anyi specifically studies the role of the minimal sufficient statistic as an expression of the `meaningful' information of a string (taking the name sophistication from Koppel).
 Like Koppel, Vit\'anyi restricts the model class to total functions, eliminating the universal model.
-He balances this force by allowing the total two-part description to be non-self delimiting, so that both the representation with all information in the model part, and the representation with everything in the residual part are less efficient than the best two-part coding.\sdr{This balancing act is not yet clear to the reader}
 
 Antunes et al.
-\cite{} note that the constant cutoff used to separate candidate two-part descriptions from those deemed to be too inefficient, is a source of instability.
+\cite{antunes2009sophistication} note that the constant cutoff used to separate candidate two-part descriptions from those deemed to be too inefficient, is a source of instability.
 Small changes to this constant could shift the result sophistication by arbitrary amounts.
@@
 
-Adriaans' \emph{Facticity}, introduced in 2012\cite{adriaans2012facticity}, relaxes the model class to include all
-partial recursive functions.
+Adriaans' \emph{Facticity}, introduced in 2012\cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions.
 By constructing a very specific universal Turing machine, and taking constant terms into account, the facticity is expected to become robust against the collapse to a universal model.
-
-\pb{Criticisms of sophistication}
 
@@
 We now have the basic tools with which we set up our generic sophistication function.
-We first define the space of two-part representations for a given $x$, and the code length that each represenrtation achieves:
+We first define the space of two-part representations for a given string $x$, and the codelength that each representation achieves:
 
@@
 \end{definition}
-
 For a function like sophistication to be meaningful, it must satisfy certain properties. We will check all currently published variants against the following minimal requirements:
@@
 
-For all variants but one, we have shown that all measures that are invariant are bounded.
+For all variants but one, we show that all measures that are invariant are bounded.
 This leads us to the following conjecture.
-Let $\gamma(n)$ be a function representing the extent to which we care about differences between functions.
-That is, we will call two functions $f$ and $g$ $\gamma$-equal if $|f(x) - g(x)| \in O(\gamma(f(x)))$.
+Let $\gamma(n)$ be a function representing a \emph{granularity}: the extent to which we care about differences between functions.
+That is, we will call two functions $f$ and $g$ $\gamma$-equal if $|f(x) - g(x)| \in O\left(\gamma\left(f(x)\right)\right)$.
 
@@
 \begin{itemize}
-  \item there exist $s_g$ and $s_h$ which are not are $\gamma$-equal (ie $s$ is not invariant),
+  \item there exist $\s_g$ and $\s_h$ which are not are $\gamma$-equal (ie $\s$ is not invariant),
   \item $s_g$ is $\gamma$-equal to $0$ or to the Kolmogorov complexity.
@@
 \end{conj}
+
 In these terms, all instantiations of the Kolmogorov complexity, using different Universal Turing machines are $1$-equal.
 It may be unreasonable to require the sophistication to attain the same level of invariance (a log-invariant sophistication could still be useful).
-This conjecture allows such relaxations: if it can be shown to be true, sophistication must be considered fundamentally flawed.
-We hope that this conjecture will prompt authors to address the matter of invariance ciritcally, and to consider the possibility that sophistication and its ilk are an untenable proposition.
+This conjecture allows such relaxations: if it is true, sophistication must be considered fundamentally flawed.
+We hope that this conjecture will prompt authors to address the matter of invariance critically, and to consider the possibility that sophistication and its ilk are an untenable proposition.
 
-In the rest of this paper, we will analyse the existing variants one-by-one.
+In the rest of this paper, we will analyse the existing variants one by one.
 We will conclude with the informal reasoning supporting this conjecture and consider its the consequences.
@@
 
-\subsection{Koppel Sophistication}
+\subsection{Koppel's Sophistication}
 
@@
 
-This does not work because of the nickname problem! To fix this:
+We start with sophistication as originally introduced by Koppel and Atlan \cite{koppelSoph1988,koppel1991almost}.
+Their sophistication is built on a two-argument universal Turing Machine $U(i, y) = \phi_p(y)$.
+They consider a pair $(i, y)$ a representation of a string $x$ if (a) $\phi_i$ is total, and (b) $\phi_i(y)$ has $x$ as a prefix.\footnotemark
+
+\footnotetext{The original sophistication also allows infinite binary strings in the range of the computable functions considered.
+We ignore this aspect for the sake of simplicity, although our conclusions still hold with such functions included.}
+
+This construction shows one of the first pitfalls of sophistication: using an inefficient representation for the model. We show that unless 
+\begin{lemma}
+Koppel and Atlan's sophistication is not both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
+\end{lemma}
+\begin{proof}
+Assume that the sophistication is $\gamma$-invariant, for some $\gamma(n) \in o(n)$. We will show that this implies that it is bounded.
+
+Let $\num{\phi}$ be the canonical acceptable numbering.
+Choose $\alpha < 1$ so that $\gamma(i) < \alpha i$ for sufficiently large $i$.
+Let $s: \N \to \B$ be a computable function such that $|s(i)| = \left\lceil(1-\alpha)^{-i}\right\rceil$.
+
+We now define two new acceptable numberings $\num{\psi}$ and $\num{\xi}$:
+\[
+\psi_j(x) = 
+\begin{cases}
+	\phi_i(x) \;\text{if}\; j = s(2i) \\
+	\infty\;\text{otherwise}
+\end{cases}
+\]
+and 
+\[
+\xi_j(x) = 
+\begin{cases}
+	\phi_i(x) \;\text{if}\; j = s(2i+1) \\
+	\infty \;\text{otherwise}\p
+\end{cases}
+\]
+Let $\s_\psi$ and $\s_\xi$ be the sophistication functions built on these enumerations.
+Choose any $x$ and assume w.l.o.g.
+that $\s_\psi(x) < \s_\xi(x)$.
+By construction of the numberings $\psi$ and $\xi$, we have $\s_\psi(x)/ (1-\alpha)  \leq \s_\xi(x) + 1$, which gives us:
+\begin{align*}
+	\s_\xi(x) - \s_\psi(x) &\geq \alpha\, \s_\xi(x) - 1\\
+	& \in \Omega(\s_\xi(x)) - 1\\
+\end{align*}
+while, by our original assumption $\s_\xi(x) - \s_\psi(x) \in O(\gamma(\s_\xi(x))) \subset o(\s_\xi(x))$, which contradicts the bound above.
+\end{proof}
+
+The main trick we use in this proof is that we can inflate the indexes arbitrarily.
+In short, the representation of the models is highly inefficient.
+For the definition of Kolmogorov complexity, this is not a problem, since we can switch to another universal machine at only a constant cost, but for the definition of sophistication, we must make sure that both parts of our two-part code are efficient.
+
+We define the complexity of a given model as its shortest effective description: a description that can be used to compute the function for a given input:
 
@@
 \end{definition}
-
-These definitions make sense because:
-
+These definitions also appear in \cite{} and \cite{}, but are not discussed in depth.
+We will show that these properties are still invariant up to a constant (ie.
+complexity is a property of a function):
 \begin{lemma}[Invariance]
@@
 
-Thus, the complexity of a function depends on the choice of
+This lemma shows that the complexity of a function depends on the choice of
 enumeration by no more than a constant term. Note that this implies
@@
 \begin{itemize}
-\item Model class is \emph{ensembles}. Equivalent to $\K$. Subsequent literature suggests that ensembles have finite support. 
+\item Model class is \emph{ensembles}.
+Equivalent to $\K$.
+Subsequent literature suggests that ensembles have finite support, which would mean that effective complexity reduces to the structure function.
 \item Description method for models is $L(m) = K(m)$ (the minimization is not mentioned explicitly, but we'll give them the benefit of the doubt).
@@
 \end{itemize}
-\pb{We can still manipulate the UTM so that all non-singleton representations are pushed out of the candidate set.
-Since the model side is allowed unbounded decoding time, this invalidates effective complexity.}
+
+\pb{\cite{ay2010effective} provides a more formal definition.}
+\begin{theorem}
+There exists a UTM so that for all $x$, $\s_\text{EC}(x) = K(x)$, up to a constant. 
+\end{theorem}
+
+Subsequent publications suggest that all ensemebke
 

### 2015-01-05 Facticity/Facticity/paper.tex

@@
 
-% in introductie?
-%
-% model, model class
-% two-part code: sophistication is length of first part
-% kolmogorov complexity?
-% KC = two part vs KC in two part
-% collapse, restrictions on model class (eg total instead of partial recursive, see section XX)
-% conjecture
-
 \begin{document}
@@
 The size in bits, used to describe the model is taken as a measure of the useful information in our string.
-While this idea has been proposed many times in many different guises, we will, for the purposes of this paper refer to it as \emph{sophistication}\footnotemark.
+This idea has been proposed many times in many different guises, which shows that it is both a persistent intuition but also that it is a difficult idea to formalize in a meaningful way.
+For the purposes of this paper, we will refer to this concept as \emph{sophistication}\footnotemark.
 
@@
 We take the shortest description and those close to it, and consider them \emph{candidates}.
-How close a description must be to the optimum differs between variants, it is often a pre-determined constant value.
-We then choose a single description from the candidate set and return its \emph{model complexity} as the sophistication.
-The method for determining model complexity varies.
-In almost all variants, the candidate with the smallest model complexity determines the sophistication.
+How close a description must be to the optimum differs between variants, one popular choice is to use a pre-chosen constant value.
+We then choose a single description from the candidate set and return the description length of the model as the sophistication.
+Usually, the candidate with the smallest model determines the sophistication.
 
@@
 
-We first provide a chronological overview of the existing variants of sophistication. In the body of the paper, we will define each variant rigorously, and in non-chronological order.
+We first provide a brief chronological overview of the existing variants of sophistication. We will return to each in-depth in the body of the paper.
 
 The first suggestion that a representation of a string can be separated cleanly into an encoding of its regular and its stochastic proeperties came from Kolmogorov himself, in a talk given at a conference in Tallinn in 1973 (first committed to paper by Cover \cite{cover1985kolmogorov} and later discussed more extensively discussed by Vereshchagin and Vit\'anyi \cite{vereshchagin2004kolmogorov}).
-Here, he proposed what has become known as the \emph{Kolmogorov structure function}.
-Briefly, this function operates on a space of two part representations for a given string $x$, consisting of some finite set $S$ (containing $x$) and the index of $x$ in $S$.
+Here, he proposed what has become known as the \emph{Kolmogorov structure function}.Minimizes over two part representations for a given string $x$, consisting of some finite set $S$ (containing $x$) and the index of $x$ in $S$.
 The \emph{minimal sufficient statistic} is the set with the lowest complexity for which the length of the two-part representation reaches the prefix-free Kolmogorov complexity.
@@
 Koppel \cite{koppelSoph1988,koppel1991almost} built on this notion, coining the phrase \emph{sophistication}.
-He changed the class of `models' from finite sets to total functions and used monotone functions, instead of prefix-free Kolmogorov complexity.
+He changed the class of models from finite sets to total functions and used monotone functions, instead of prefix-free Kolmogorov complexity.
 We expect that this change was made to extend the theory to infinite strings, an approach which has been abandoned by subsequent authors.

### 2015-01-05 PhD Thesis/src/thesis.tex

--- 2014-01-16 PhD Thesis/src/thesis.tex
@@
 \usepackage{hyperref}
+\usepackage{makeidx}
 
@@
 
-\title{Statistics for complex graphs: an algorithmic approach}
+\title{Single sample statistics}
 \date{\today}
@@
 
+\makeindex
+
 \begin{document}
+
+\maketitle
+
+\tableofcontents
 
@@
 
-% Start with a story
-% Mention RDF: the future of data storage
+Imagine a mysterious signal arriving from outer space: for a brief period, radio-astronomers receive a sequence of long and short pulses from a largely empty section of the sky.
+Then, the sky goes quiet again.
 
-\section{The complex graph}
+This scenario represents the ultimate challenge to statistics.
+We have a finite amount of data, and nothing more.
+We cannot take another sample, and no assumptions are justified.
+We don't know the source distribution, we don't know whether the data can be cut into meaningful parts.
+All we have is one finite piece of data.
 
-% Complex graphs are not just funny things to look at, the are the fundamental unit of science in the 21st century.
+Yet, even in this bleakest situation, we are not totally lost.
+If the signal is entirely without structure, we feel we can dismiss it as white noise.
+If the signal is highly regular, like a periodic pulse, some natural phenomenon might be responsible for it, perhaps a planet in close orbit of a star.
+If the signal contains many complex structures, and yet remains unpredictable, like a binary encoding of a novel, we would suspect an intelligent origin.
 
-\section{Computer programs as statistical models}
+This story describes the setting we will analyse in this dissertation.
+We, the experimenter, are presented with a single, finite sample of data, described in some discrete manner, perhaps as a string of bits.
+We are given no extra samples, and no basis for any further assumptions.
+The data may represent a sequence of iid.
+samples from some distribution, or it may describe a single, complex object.
+If we wish to see it as such, we must find our motivation for doing so within the data.
 
-\chapter{Statistical analyses of complex graphs}
+The thesis behind this dissertation is that this settings, which we'll call \emph{single sample statistics}, is the most generic view of statistical practice that still allows us to make useful inference.
+All other settings, repeated IID samples, time-series analysis, can be seen as  version of this setting with additional assumptions or constraints.
+If we want to establish a simple, comprehensive framework for statistical practice from first principles, this setting should be our foundation.
+A single, finite datum, with no further assumptions or restrictions.
 
-\begin{summary}We review the literature on complex graphs, and the approaches that have been used so far to provide statistical analyses on them.
+\textbf{We will show that statistical inference is feasible in the single sample setting setting}.
+We can determine the likelihood of the signal in a general sense, and we can compare different explanations for the data.
+However, there are also common, uncontroversial practices that are unlikely to be possible in the setting of single sample statistics without further assumptions and restrictions.
+Model selection, for instance, seems to be a lost cause.
+It's not just that we can't find the data's inherent model, we cannot define what it should be in an unambiguous, objective manner.
+Both are interesting results.
+By studying what works and what doesn't for single sample statistics we show both the value and the danger of additional constraints.
+Assuming that our data is sampled IID from a Bernoulli distribution will allow us to perform model selection, but if this assumption is unjustified
+
+\chapter{Kolmogorov complexity: a universal probability}
+
+\begin{summary}If we make no assumptions on the source of our data, we are, in essence taking the model class of all computable probability distributions: if our data is produced by a model outside of this class, we will never be able to discover that fact anyway.
+Following this line of reasoning, we end up naturally with Algorithmic complexity, and Algorithmic statistics.
+In this chapter we make this reasoning explicit.
+While Kolmogorov complexity is itself incomputable, the surrounding framework does allow us to perform practical statistical inference, even in the single sample setting.
+We also explore what further practical approaches are possible if we allow a model assumption.
 \end{summary}
 
-% Algorithmic statistics is how you stud graphs. Nothing else makes sense.
+\pb{A safe approximation for Kolmogorov Complexity}
 
-\chapter{Algorithmic Statistics: Bridging the gap between Kolmogorov complexity and MDL}
+\chapter{Model selection and sophistication}
 
-\begin{summary}MDL and Kolmogorov complexity are two offshoots of the same basic idea: that good compression makes good statistics.
-MDL takes this as an axiom and builds on it, while Kolmogorov complexity makes only the assumption that the data is produced by a computational process.
-The two methods are not incompatible, but they have produced very different results and techniques.
-In order to have access to the best of both worlds, we start with the rigorous theoretical foundations of Kolmogorov complexity and build a platform that will allow us to fold in the techniques of MDL.
-We will revisit the principle of two-part coding, considered to be a usable, but cumbersome technique in MDL and show that when combined with Kolmogorov complexity it has some deep advantages over other methods.
+\begin{summary}In this chapter we investigate the limits of the single sample setting.
+We show that in a  straightforward approach to Algorithmic Statistics, model selection cannot be performed in an objective manner.
+This tells us that in the single sample setting, the data does not have an inherent model.
+It may have inherent patterns, as witnessed by the Kolmogorov complexity, but an objective separation into model and noise is not feasible.
 \end{summary}
 
-\chapter{What makes data interesting: two-part coding in action}
+\pb{A critical review of Algorithmic Model Selection and Sophistication}
 
-\begin{summary}The question of what makes data complex is has been decisively answered.
-First for IID data in, as maximal Entropy, and then for any data as maximal Kolmogorov Complexity.
-But there's some dissonance in calling this data information-rich.
-We are not very interested in watching TV broadcasts of static, or recordings of white noise.
-This may contain maximal information to a computer, but to us, it's not very interesting.
-In fact the most interesting data seems to be somewhere in between the zero complexity signal (s tv which is turned off) and the maximum complexity one (analog noise).
-The question of whether we can capture this notion of `interestingness' in a metric analogous to information entropy and Kolmogorov complexity has occupied research for over half a century, and has produced some surprisingly deep results.
-This chapter provides a review.
+\chapter{Graph modeling}
+
+\begin{summary} Analysis of graphs is a perfect exemplar of the single sample setting.
+Take the world wide web, for instance.
+It is a single datum, and we cannot sample another.
+Yet it is large enough that it approaches the size of datasets that we imagine when talking about algorithmic statistics.
+It is certainly highly structures, and highly unpredictable.
+It cannot be partitioned into logical units, but it can be sub-sampled.
 \end{summary}
 
-\section{The question of interestingness}
+\section{Complex graphs, an MDL perspective}
+\pb{MDL models for graphs}
 
-\chapter{Graph modeling: extracting motifs}
+\subsection{Motif extraction}
 
-\begin{summary}Modeling complex objects starts with finding out what the components are.
-If we know the letters, we can find the words.
-If we know the words, we can find the sentences.
-Graphs give us nodes and links for free.
-How do they combine into functional, recurring units?
-How can we build up to the global structure?
+\pb{Motifs}
+
+\section{Inference on the nodes of a graph}
+
+\pb{Simplifying RDF data for graph-based machine learning, Machine Learning on Linked Data, a Position Paper}
+
+\chapter{Exploiting self-similarity}
+\begin{summary}
+In the single sample setting, the only thing we have to work with is the internal structure of the data.
+A particularly useful structure to exploit is that of self-similarity.
+In this chapter, we investigate the possibilities.
 \end{summary}
 
-\chapter{Graph modeling with rewrite grammars}
+If all we have is the data, we must find compressible structure.
+The most common form of structure is symmetry: transforming the data in some way, leaves it in some way invariant.
+For example: if we take a text, and chuck it into words, we have a kind of translational symmetry.
+Shifting the data one chunk left or right exposes a different word, but from the same distribution.
+Of course, this assumes that the words are sampled IID, which is far from true, but this translational symmetry alone is enough to compress the data.
 
-\begin{summary}In last chapter we induced motifs for various types of complex networks.
-We ended by applying the procedure recursively, collapsing upwards across several levels to compress the whole structure.
-\end{summary}
+All such symmetries, are commonly exploited in many different statistical settings.
+Most practitioners are unlikely to take this perspective, of course: a machine learning with a table of instances is unlikely to think of his dataset as a single datum with internal translational symmetry, but from the single sample perspective, that is what is happening.\index{Symmetry}
 
-\chapter{Graph modeling with deep neural networks}
+While it may be awkward to think of classification from the single sample perspective (sometimes we must allow ourselves to forget our first principles), the notion does suggest a new kind of symmetry to analyse: self similarity.
 
-\begin{summary}To model graphs in an algorithmic setting we can use any family of computational structures that can turn randomness into graph structures.
-One recently popular model that has been shown to be effective at producing random examples of any given model is the Deep Belief Network.
-We investigate their potential to model graph structure, and fit them in to the framework of algorithmic statistics.
-\end{summary}
+Self similarity if symmetry of scale: when the data is shrunk down, it remains the same.
+The most exact and literal form of self similarity can be found in fractal figures, like the Sierpinski gasket.
+In this chapter we investigate whether we can model self similarity: we will attempt to tackle what is known in fractal geometry as \emph{the inverse problem}.
 
-\chapter{Machine learning on graphs}
+\pb{An Expectation Maximization Algorithm for the Fractal Inverse Problem}
 
-% This does not fit the line too well, refocus?
+\chapter{Single sample statistics}
 
-\begin{summary}The classic view of offline machine learning starts with a list of instances from which to learn.
-These instances can then be represented in some way---preferably a vector of numeric features---and analyzed statistically.
-In the past decade a situation has emerged and re-emerged that is an ill fit to this paradigm.
-We are faced with information in a graph: a social network, a web graph, an RDF dataset.
-We may know our instances, if they correspond to specific nodes in the graph, but the things we know about them are spread throughout the graph.
-We investigate the problem and provide a bridge to traditional machine learning.
-\end{summary}
-
-\section{A pipeline}
-\subsection{Instance extraction}
-\subsection{Feature extraction}
-\section{Methods}
-\subsection{Classification}
-\subsection{Clustering}
-\subsection{Regression}
-
-\chapter{Citation networks, a use case}
-
-\begin{summary}Throughout the previous chapters, we have performed many quick tests on small datasets, to illustrate and validate our points.
-In this chapter we take one use case and investigate more thoroughly, in order to show the full potential of our methods.
-We choose the domain of citation networks, and investigate various questions.
-\end{summary}
-
-\begin{itemize}
-  \item Who are the hubs?
-  \item H-index: can we do better?
-  \item Multi-resolution analysis
-  \item Visualization
-\end{itemize}
-
-\chapter{Statistics for complex graphs}
-
-\begin{summary}\ldots
-\end{summary}
-
-Conclusatory chapter.
+\pb{Review and Conclusion}
 
@@
 \chapter{Tutorials}
-\section{Kolmogorov Complexity}
-\section{Minimum Description Length}
-\section{Graphs and RDF}
+\begin{summary}
+This appendix provides brief introductions to the subjects that are required to understand this thesis.
+They illustrate the use of our notation and provide a concise basis for readers who are well-versed in mathematics and computer science, but lack knowledge of these specific subjects.
+\end{summary}
+
+\section{Kolmogorov complexity}
+\section{Recommended literature}
+
+\chapter{Symbols and Terms}
+
+\begin{description}
+ \item[$K(\cdot)$]
+ \item[$k(\cdot)$]
+ \item[Shannon Entropy]
+\end{description}
+
 \chapter{Algorithms and techniques}
 \chapter{Proofs}
+\chapter{index}
+\printindex
 

### 2015-01-06 Facticity/Facticity/paper.tex
...

@@
 
-\section{Preliminaries, definitions and desiderata}
+\subsection{Preliminaries, definitions and desiderata}
 
@@
 
-\section{Measures of meaningful information}
+\section{Inefficient indices}
 
-\subsection{Koppel's Sophistication}
+We start from a simple intuition: the definition of Kolmogorov complexity already uses a description of the data in two parts; a computable function and and input.
+Can we assume that such a description, that is as efficient as the Kolmogorov complexity, creates a good separation into the structural part of the data and the noise?
 
-\begin{itemize}
-\item Model class is $\T$
-\item Threshold function $\tau(x)=L^\M(x)+c$
-\item Description method for models is required to be prefix-free, but
-  can otherwise be chosen arbitrarily.
-\item Koppel uses monotone complexity, and monotone functions and discusses 
-  infinite strings, TODO figure out if and how this affects our results.
-  \item Koppel has a `machine-independence result', but not one that does what we want.
-\end{itemize}
+The first problem that arises, is that allowing any acceptable numbering can lead to inefficient model representations.
 
-We start with sophistication as originally introduced by Koppel and Atlan \cite{koppelSoph1988,koppel1991almost}.
-Their sophistication is built on a two-argument universal Turing Machine $U(i, y) = \phi_p(y)$.
-They consider a pair $(i, y)$ a representation of a string $x$ if (a) $\phi_i$ is total, and (b) $\phi_i(y)$ has $x$ as a prefix.\footnotemark
+\pb{Switch naar two-part perspective}
+\begin{lemma}
+Let $\phi$ be some acceptable numbering, and $\s_\phi(x)$ be any function which gives the length of some index $i$ such that $\phi_i(y) = x$ for some $y$.
 
-\footnotetext{The original sophistication also allows infinite binary strings in the range of the computable functions considered.
-We ignore this aspect for the sake of simplicity, although our conclusions still hold with such functions included.}
-
-This construction shows one of the first pitfalls of sophistication: using an inefficient representation for the model. We show that unless 
-\begin{lemma}
-Koppel and Atlan's sophistication is not both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
+No such function can be $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
 \end{lemma}
 \begin{proof}
-Assume that the sophistication is $\gamma$-invariant, for some $\gamma(n) \in o(n)$. We will show that this implies that it is bounded.
+Choose some $\gamma(n) \in o(n)$. Assume towards contradiction that the sophistication is $\gamma$-invariant and unbounded.
 
-Let $\num{\phi}$ be the canonical acceptable numbering.
-Choose $\alpha < 1$ so that $\gamma(i) < \alpha i$ for sufficiently large $i$.
-Let $s: \N \to \B$ be a computable function such that $|s(i)| = \left\lceil(1-\alpha)^{-i}\right\rceil$.
+Let $\num{\phi}$ be the canonical acceptable numbering. Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.
 
@@
 \begin{cases}
-	\phi_i(x) \;\text{if}\; j = s(2i) \\
+	\phi_i(x) \;\text{if}\; j = s_{2i}\\
 	\infty\;\text{otherwise}
@@
 \begin{cases}
-	\phi_i(x) \;\text{if}\; j = s(2i+1) \\
+	\phi_i(x) \;\text{if}\; j = s_{2i+1} \\
 	\infty \;\text{otherwise}\p
@@
 that $\s_\psi(x) < \s_\xi(x)$.
-By construction of the numberings $\psi$ and $\xi$, we have $\s_\psi(x)/ (1-\alpha)  \leq \s_\xi(x) + 1$, which gives us:
+By construction of the numberings $\psi$ and $\xi$, we have $2\;\s_\psi(x)  \leq \s_\xi(x)$, which gives us:
 \begin{align*}
-	\s_\xi(x) - \s_\psi(x) &\geq \alpha\, \s_\xi(x) - 1\\
-	& \in \Omega(\s_\xi(x)) - 1\\
+	\s_\xi(x) - \s_\psi(x) &\geq \frac{1}{2}\, \s_\xi(x)  \in \Omega(\s_\xi(x))\p 
 \end{align*}
-while, by our original assumption $\s_\xi(x) - \s_\psi(x) \in O(\gamma(\s_\xi(x))) \subset o(\s_\xi(x))$, which contradicts the bound above.
+Since, by assumption, $\s_\xi$ can be arbitrarily large and $\s_\xi(x) - \s_\psi(x) \in O(\gamma(\s_\xi(x))) \subset o(\s_\xi(x))$, we have a contradiction.
 \end{proof}
 
-The main trick we use in this proof is that we can inflate the indexes arbitrarily.
-In short, the representation of the models is highly inefficient.
-For the definition of Kolmogorov complexity, this is not a problem, since we can switch to another universal machine at only a constant cost, but for the definition of sophistication, we must make sure that both parts of our two-part code are efficient.
+In this proof, we deliberately use ridiculously inefficient index functions.
+Obviously one could simply pick a more efficient index; the point is that many definitions of sophistication to not mention any such restriction.
+As we shall see in the next, any inefficiency in the index will lead to a lack of invariance in sophistication.
 
-We define the complexity of a given model as its shortest effective description: a description that can be used to compute the function for a given input:
+The treatments that fall prey to this problem are Koppel and Atlan's \cite{koppelSoph1988,koppel1991almost},  and several of its variants (\cite{antunes2009sophistication,antunes2013sophistication}).
+
+Interestingly, using an inefficient index is no problem in the definition of Kolmogorov complexity, since we can select as a model a UTM with an efficient index at only a constant penalty (by the Invariance Theorem \cite{li1993introduction}).
+But this robustness property does \emph{not} carry over to definitions of sophistication, if we are interested in separating the model and its input, we must represent the model efficiently.
+
+Another perspective ion this problem is that the length of a model in an ordinary description does not represent its complexity, much like the length of a compressible string does not reflect its information content.
+In order to obtain a more robust measure of model complexity, we describe a natural extension of Kolmogorov complexity to functions, and show that this retains invariance.
 
@@
   \begin{align*}
-    C(f) &:= \min\{C(i):\phi_i=f\}\\
-    K(f) &:= \min\{K(i):\phi_i=f\}.
+    C_\phi(f) &:= \min\{C(i):\phi_i=f\}\\
+    K_\phi(f) &:= \min\{K(i):\phi_i=f\}.
   \end{align*} \label{definition:model-complexity}
 \end{definition}
-These definitions also appear in \cite{} and \cite{}, but are not discussed in depth.
-We will show that these properties are still invariant up to a constant (ie.
-complexity is a property of a function):
+These definitions also appear in \cite{grunwald2004shannon} and \cite{vitanyi2004meaningful}.
+We will show that these values are invariant up to a constant (ie.
+complexity is an objective property of a function):
 \begin{lemma}[Invariance]
@@
 \end{align*}
-We can reverse $\phi$ and $\psi$ without loss of generality to achieve the same inequality the other way around, completing the proof.
+We can reverse $\phi$ and $\psi$ to achieve the same inequality the other way around, completing the proof.
 \end{proof}
+This lemma shows that the complexity of a function depends on the choice of
+enumeration by no more than a constant term.
 
-This lemma shows that the complexity of a function depends on the choice of
-enumeration by no more than a constant term. Note that this implies
-that for any enumeration $\psi_1,\psi_2,\ldots$, the complexity $K(\psi_i)$
-is less than the literal description length of the index
-$|\bar\imath|$ up to a constant. For some numberings, there is a gap
-between these two description methods; in those cases, the index is an
-inefficient representation of the corresponding function. However it
-is possible to construct numberings that are \emph{faithful} in the
-sense that no such gap exists:
+There are two ways to use this notion of model complexity for more robust attempts to define sophistication, that differ in the way the two-part code is conceptualized.
+Both approaches are used in the literature, which is therefore sometimes confusing.
+\begin{enumerate}
+  \item On the one hand one can stick with Koppel's approach, but repair it by imposing the additional requirement that the used length of the indices used is equal to $C(\phi_i)$, and that the numbering is still acceptable.
+We construct such an index below.
+This approach is taken by Adriaans \cite{adriaans2012facticity}.
+In this case, the total code length is simply $C(x)$, and the sophistication is the length of the model index within this code.
+This approach is the most direct formalization of the intuition that the two part code used within the definition of Kolmogorov complexity separates structure from noise.
+  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use Kolmogorov complexity as a black box to measure the complexity of the model, which is then the first part of a two-part code describing the data.
+With this approach, the total code length is $K(f)+|y|$, where $y$ is an input to the model $f$ such that $f(y)=x$.
+The difference is that in this approach, we can construct everything on the basis of a single, unconstrained, acceptable numbering, without using additional contraints.
+It is also more natural in this setting to limit the model class of allowed functions.
+This approach is used for Kolmogorov's structure function \cite{cover1985kolmogorov}, Algorithmic Statistics \cite{gacs2001algorithmic}, Vi\'anyi's sophistication \cite{vitanyi2004meaningful} and Effective complexity \cite{gellmann1996information}.
+\end{enumerate}
+
+To show that the two approaches are equivalent with the correct restriction on the index, we must show that there exists a numbering for which the length of the index of a model is equal to its complexity, and that such a numbering is acceptable.
 
@@
 \end{definition}
-
+This definition is a refinement of the one used in \cite{adriaans2012facticity}.
+That publication defines an \emph{index function}, rather than a numbering, and notes that it is not computable.
+Since we use numberings, we only need to show that they are acceptable, ie.
+we only need to be able to unpakc the efficient description, so the incomputability of the index function is no issue:
 \begin{lemma}
@@
 $\bar\imath p$, the 
-$s^n_m$-theorem (see~\cite{TODO}) states that we can compute the $h$
+$s^n_m$-theorem \cite{kleene193notation} states that we can compute the $h$
 such that $\phi_h(y)=\phi_v(\bar\imath p,y)$. Let $h(\bar\imath p)$
@@
 
-\subsection{Antunes's Coarse Sophistication}
-
-\begin{itemize}
-\item Model class is $\T$
-\item Description method $L(m)$ for models is unclear, but appears to be
-  taken from Koppel, i.e. arbitrary but self-delimiting.
-\item Sophistication is defined slightly differently: $S^\T(x)=L(m^*)$ where $m^*=\argmin_m L(m)+L^m(x)-C(x)$. Thus, there is no slack.
-\end{itemize}
-
-\subsection{Adriaans' Facticity}
-
-\begin{itemize}
-\item Model class is $\C$
-\item Description method for models is $L(m)=|\bar\imath|$ where
-  $\phi_i=m$. The index function is assumed to be \emph{faithful}, see
-  Definition~\ref{def:faithful}
-\item Threshold function is $\tau(x)=L^\M(x)$, the constant in the candidate set is $0$.
-\item Facticity is actually defined as index length without prefix
-  encoding:  $S^\C_0(x):=\min\{|i|:\phi_i\in C^\C_0(x)\}$. But this
-  difference is not important for the discussion.
-\end{itemize}
+Unfortunately, even with a faithful index, we can fail, if the prefix-function used to separate the model from its input is inefficient:
 
 \begin{lemma}
-  If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then facticity is bounded.
+Let $L(\phi) = |\bar\imath|$ If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then facticity is bounded.
 \end{lemma}
@@
 \end{proof}
+
 Note that the condition for this lemma applies to all known partial recursive prefix functions:
@@
 
-\subsection{Effective complexity}
+\section{Underfitting and overfitting}
 
-\begin{itemize}
-\item Model class is \emph{ensembles}.
-Equivalent to $\K$.
-Subsequent literature suggests that ensembles have finite support, which would mean that effective complexity reduces to the structure function.
-\item Description method for models is $L(m) = K(m)$ (the minimization is not mentioned explicitly, but we'll give them the benefit of the doubt).
-\item The threshold function is equal to the minimum two-part code: $\tau(x)=L^m(x)$.
-\item Gell-mann and Lloyd acknowledge that representations exist with small and large models.
-The choice from the candidate set is made by applying a time-bound to the computation from the string to the model, but not to the decoding of the model itself.
-The idea is that this forces `deep' information into the model, and shallow information into the data.
-Presumably, the representation with the smallest model code under this constraint is chosen.
-\end{itemize}
+\pb{explain title}
 
-\pb{\cite{ay2010effective} provides a more formal definition.}
-\begin{theorem}
-There exists a UTM so that for all $x$, $\s_\text{EC}(x) = K(x)$, up to a constant. 
-\end{theorem}
-
-Subsequent publications suggest that all ensemebke
-
-\begin{theorem}[Reduction of Effective Complexity to Structure
-  Function]
-  There is a constant $c$ such that, for every representation of $x$
-  consisting of an ensemble $E$ with $x\in E$ with entropy $H$,
-  yielding a total information of $K(E)+H$, there is a finite set $S$
-  with $x\in S$ such that $K(S)\le K(E)+c$ and $\log|S|\le H$.
-\end{theorem}
-\begin{proof}
-  First note that although the ensemble $E$ may itself be infinite, it
-  is ``typical'', and therefore a valid candidate representation, for
-  only a finite number of its elements, namely the set $S=\{y\in E:-\log
-  P(y)\le H\} = \{y:P(y)\ge 2^{-H}\}$. Since the sum of probabilities must
-  converge, this set cannot be infinite: since $1\ge\sum_{y\in S} P(y)\ge
-  \sum_{y\in S}2^{-H} = |S|2^{-H}$ so $|S|\le 2^H$.
-
-\end{proof}
-
-\subsection{Roll your own}
-
-\begin{itemize}
-\item Model class is either $\C$ or $\K$
-\item Description method for models is $L(m)=K(m)$.
-\item Threshold function is $\tau(x)=L^m(x)+c$, which matches $C(x)$
-  resp. $K(x)$ up to a constant.
-\end{itemize}
+\subsection{Underfitting: the universal model}
 
@@
 
-\paragraph{TODO: instantiate for UTM}
-
-\begin{itemize}
-\item Model class is $\K$
-\item Description method for models is $L(m)=K(m)$.
-\item Threshold function is $\tau(x)=L^m(x)+c$, which matches $K(x)$
-  up to a constant.
-\end{itemize}
-
 \begin{lemma}
@@
 
-\subsection{Kolmogorov's Structure Function}
+\pb{What happens if we just look at the witnesses to Komogorov complexity.}
+\pb{
+stelling: Bounded door universeel model \\
+literatuur: Gell-man, Vitanyi noemen univ model.
+oplossing: beperk de model klasse
+}
 
-\begin{itemize}
-\item Model class is $\F$
-\item Description method for models $L(m)=K(S)$ where $S$ is the range
-  of $m$. There is some ambiguity as to the definition of prefix
-  complexity for finite sets; discuss and refer to Shen \cite{TODO}.
-\end{itemize}
+\subsection{Overfitting: the singletons}
+
+\pb{
+stelling: singletons kunnen domineren in structure function.
+lit: Kolmogorov, Vitanyi, Antunes, Gell-Man
+oplossing: Paul's meaningful information?
+}
+
+\section{The problem of depth}
 
@@
 \end{proof}
-
-\subsection{Vit\'anyi's Sophistication / Meaningful Information}
-
-\begin{itemize}
-\item Model class is $\T$
-\item Description method for models is $L(m)=K(m)$
-\item Candidate set is defined differently: $C_c=\{m\in\T:|L^m(x)-K(x)|\le
-  c\}$. There appears to be a problem with this definition, because
-  the two-part descriptions of $x$ may be shorter than $K(x)$ by
-  exploiting the fact that the models do not need to be prefix functions.
-\end{itemize}
-
-% stru fu: model class is all finite sets
-% code length L(m)+L_m(x)
-% acceptability criterion acc(m) = 2L(m)+L_m(x)-C(x)
-%
 
@@
 Note that we favoured simplicity of the proof over the strength of the bounds: the multiplicative constant $1/6$ can be improved with a more detailed treatment.
+
+\pb{
+stelling: voor alle niet superdiepe strings is Paul's sophistication bounded.
+}
+
 \section{A world without sophistication}
@@
 \pb{Total functions. They ruin efficiency and push sophistication into a land of unprovable claims.}
+\pb{reference Antunes instability}
 
@@
 
-\section{Stuff to move up or delete}
+\section{=======Stuff to move up or delete=======}
+
+\section{Old structure}
+
+\subsection{Koppel and Atlan}
+
+We start with sophistication as originally introduced by Koppel and Atlan \cite{koppelSoph1988,koppel1991almost}.
+Their sophistication is built on a two-argument universal Turing Machine $U(i, y) = \phi_p(y)$.
+They consider a pair $(i, y)$ a representation of a string $x$ if (a) $\phi_i$ is total, and (b) $\phi_i(y)$ has $x$ as a prefix.
+The length of a representation is measured as $|i| + |y|$.
+Note that this ignores the cost of delimiting the model $i$ from the input $y$, ie.
+this is not the length of a single unambiguous code word for $x$.
+Candidates are those models with representations $c$ bits from the minimum length, where $c$ is a pre-set `significance' parameter.
+
+\footnotemark
+
+Koppel and Atlan actually provide a machine-independence result, but only in the setting of infinite strings.
+
+\footnotetext{The original sophistication also allows infinite binary strings in the range of the computable functions considered.
+We ignore this aspect for the sake of simplicity, although our conclusions still hold with such functions included.}
+
+This construction shows one of the first pitfalls of sophistication: using an inefficient representation for the model:
+
+\subsection{Antunes's Coarse Sophistication}
+
+\begin{itemize}
+\item Model class is $\T$
+\item Description method $L(m)$ for models is unclear, but appears to be
+  taken from Koppel, i.e. arbitrary but self-delimiting.
+\item Sophistication is defined slightly differently: $S^\T(x)=L(m^*)$ where $m^*=\argmin_m L(m)+L^m(x)-C(x)$. Thus, there is no slack.
+\end{itemize}
+
+\subsection{Adriaans' Facticity}
+
+\begin{itemize}
+\item Model class is $\C$
+\item Description method for models is $L(m)=|\bar\imath|$ where
+  $\phi_i=m$. The index function is assumed to be \emph{faithful}, see
+  Definition~\ref{def:faithful}
+\item Threshold function is $\tau(x)=L^\M(x)$, the constant in the candidate set is $0$.
+\item Facticity is actually defined as index length without prefix
+  encoding:  $S^\C_0(x):=\min\{|i|:\phi_i\in C^\C_0(x)\}$. But this
+  difference is not important for the discussion.
+\end{itemize}
+
+\subsection{Effective complexity}
+
+\begin{itemize}
+\item Model class is \emph{ensembles}.
+Equivalent to $\K$.
+Subsequent literature suggests that ensembles have finite support, which would mean that effective complexity reduces to the structure function.
+\item Description method for models is $L(m) = K(m)$ (the minimization is not mentioned explicitly, but we'll give them the benefit of the doubt).
+\item The threshold function is equal to the minimum two-part code: $\tau(x)=L^m(x)$.
+\item Gell-mann and Lloyd acknowledge that representations exist with small and large models.
+The choice from the candidate set is made by applying a time-bound to the computation from the string to the model, but not to the decoding of the model itself.
+The idea is that this forces `deep' information into the model, and shallow information into the data.
+Presumably, the representation with the smallest model code under this constraint is chosen.
+\end{itemize}
+
+\pb{\cite{ay2010effective} provides a more formal definition.}
+\begin{theorem}
+There exists a UTM so that for all $x$, $\s_\text{EC}(x) = K(x)$, up to a constant. 
+\end{theorem}
+
+Subsequent publications suggest that all ensemeble\ldots.
+
+\begin{theorem}[Reduction of Effective Complexity to Structure
+  Function]
+  There is a constant $c$ such that, for every representation of $x$
+  consisting of an ensemble $E$ with $x\in E$ with entropy $H$,
+  yielding a total information of $K(E)+H$, there is a finite set $S$
+  with $x\in S$ such that $K(S)\le K(E)+c$ and $\log|S|\le H$.
+\end{theorem}
+\begin{proof}
+  First note that although the ensemble $E$ may itself be infinite, it
+  is ``typical'', and therefore a valid candidate representation, for
+  only a finite number of its elements, namely the set $S=\{y\in E:-\log
+  P(y)\le H\} = \{y:P(y)\ge 2^{-H}\}$. Since the sum of probabilities must
+  converge, this set cannot be infinite: since $1\ge\sum_{y\in S} P(y)\ge
+  \sum_{y\in S}2^{-H} = |S|2^{-H}$ so $|S|\le 2^H$.
+\end{proof}
+
+\subsection{Roll your own}
+
+\begin{itemize}
+\item Model class is either $\C$ or $\K$
+\item Description method for models is $L(m)=K(m)$.
+\item Threshold function is $\tau(x)=L^m(x)+c$, which matches $C(x)$
+  resp. $K(x)$ up to a constant.
+\end{itemize}
+
+\paragraph{TODO: instantiate for UTM}
+
+\begin{itemize}
+\item Model class is $\K$
+\item Description method for models is $L(m)=K(m)$.
+\item Threshold function is $\tau(x)=L^m(x)+c$, which matches $K(x)$
+  up to a constant.
+\end{itemize}
+
+\subsection{Kolmogorov's Structure Function}
+
+\begin{itemize}
+\item Model class is $\F$
+\item Description method for models $L(m)=K(S)$ where $S$ is the range
+  of $m$. There is some ambiguity as to the definition of prefix
+  complexity for finite sets; discuss and refer to Shen \cite{TODO}.
+\end{itemize}
+
+\subsection{Vit\'anyi's Sophistication / Meaningful Information}
+
+\begin{itemize}
+\item Model class is $\T$
+\item Description method for models is $L(m)=K(m)$
+\item Candidate set is defined differently: $C_c=\{m\in\T:|L^m(x)-K(x)|\le
+  c\}$. There appears to be a problem with this definition, because
+  the two-part descriptions of $x$ may be shorter than $K(x)$ by
+  exploiting the fact that the models do not need to be prefix functions.
+\end{itemize}
+
+% stru fu: model class is all finite sets
+% code length L(m)+L_m(x)
+% acceptability criterion acc(m) = 2L(m)+L_m(x)-C(x)
+%
 

### 2015-01-06 Facticity/Facticity/paper.tex
added playground

@@
 \title{A Critical Review of Sophistication and Algorithmic Model Selection}
+
 \author{Peter Bloem and Steven de Rooij}
@@
 \end{abstract}
+
+\section{Steven's playground}
+
+Potential alternative title: ``Four problems with all current attempts to define meaningful information''
+
+Requirements for sophistication:
+
+(1) Being an information measure, the value should count the number of bits required for an effective description of the meaningful properties of a binary object.
+
+(2) The Invariance Theorem for Kolmogorov complexity states that a change in reference UTM can only influence the complexity up to an additive constant term.
+Analogously, if sophistication is to be interpreted as an objective property of the data, then there should be strict limits on the degree to which a change in the UTM can affect the sophistication.
+
+(3) Sophistication should not be $\gamma$-equivalent to 0.
+While a valid information measure, the function 0 is clearly at odds with the intuition that there is no universal cap on the amount of meaningful information in the world.
+
+(4) Sophistication should not be $\gamma$-equivalent to the Kolmgorov complexity, because then there is no point in defining a new word for the same concept.
+Moreover, Kolmogorov complexity is at odds with the intuition that there is no meaningful information in a random string.
 

### 2015-01-07 Facticity/Facticity/paper.tex
...

--- 2015-01-06 Facticity/Facticity/paper.tex
@@
 \begin{lemma}
-Let $L(\phi) = |\bar\imath|$ If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then facticity is bounded.
+Let $L(\phi) = |\bar\imath|$ for a faithful index, and some prefix encoding function, and let $\s^\C$ be defined on this $L$ with the default threshold function.
+
+If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then $s^\C$ is bounded for all numberings $\phi$
 \end{lemma}
@@
 \end{proof}
+This result applies to Adriaans' facticity\cite{adriaans2012facticity}.
+In order to circumvent the problem, the UTM should should encode the model $f$ in $K(f)$ bits, relying on $K$'s own prefix encoding to delimit the model from its input.
 
-Note that the condition for this lemma applies to all known partial recursive prefix functions:
+Note that the condition for this lemma applies to all known partial computable prefix functions. We conjecture that it applies to all:
 
 \begin{conjecture}
-For any partial recursive prefix code length function $L$, the function
+For any partial computable prefix code length function $L$, the function
 \[
@@
 
+This result shows that even if we have a faithful numbering, we can only use it in settings where we don't count the cost of delimiting the model from the input (eg.
+\cite{koppelSoph1988}).
+In the cases where the data representations are taken as a single code word, the model-part must represent the model as efficient as $K(f)$ does.
+Instead of a numbering, we require a self-delimiting code for functions rather than a numbering.
+
+In the rest of this paper, we will take the second perspective: we will consider the representations as two-part descriptions $K(f)$, and $|y|$, such that $f(y)=x$.
+This description can be unpacked by a very specific universal Turing machine, but contrary to the construction of the Kolmogorov complexity, we are not free to choose any UTM.
+
 \section{Underfitting and overfitting}
 
-\pb{explain title}
+In the previous section, we saw a first example of underfitting: a UTM with a more efficient model representation became a model for all strings.
+Even if we chose our model-description carefully, to avoid the issues of the previous section, however, universal Turing machines are still candidate model for the data.
+These candidates represent the data efficiently (they reach the Kolmogorov complexity up to a constant), but they are extremely generic: taking a second sample from the model can produce a string that has nothing to do with the data.
+Nevertheless, these models perform well, in terms of compression, and there is no justification for excluding them from the candidate set.
+
+The opposite problem is overfitting.
+A model that can produce only $x$ can be described in as many bits as it takes to describe $x$, plus some constant amount.
+Again, such models perform as well as any other, but they are not very informative.
+We call this \emph{overfitting}.
+
+As we shall see, most variants of sophistication are constructed specifically to prevent over- and underfitting.
 
 \subsection{Underfitting: the universal model}
+
+Any sophistication function built on a model class that includes singletons, or universal models is in danger of over or underfitting.
+In that case, we can construct a numbering, so that a particular model, or class of models, represents all data more efficiently than any other model by an arbitrary constant.
+This constant boost is enough to force all other models out of the candidate set.
+
+The following lemma generalizes the principle. We will instantiate it for different variants of sophistication:
 

### 2015-01-07 Facticity/Facticity/paper.tex
...

@@
 
-\title{A Critical Review of Sophistication and Algorithmic Model Selection}
+\title{Four Problems for Sophistication}
 
@@
 
-\section{Steven's playground}
+\section*{Steven's playground}
 
@@
 
-The first problem that arises, is that allowing any acceptable numbering can lead to inefficient model representations.
-
-\pb{Switch naar two-part perspective}
+The first problem that arises is that allowing any acceptable numbering can lead to inefficient model representations.
+
 \begin{lemma}
@@
 
-This result shows that even if we have a faithful numbering, we can only use it in settings where we don't count the cost of delimiting the model from the input (eg.
-\cite{koppelSoph1988}).
+This result shows that even if we have a faithful numbering, we can only use it in settings where we don't count the cost of delimiting the model from the input (such as \cite{koppelSoph1988}).
 In the cases where the data representations are taken as a single code word, the model-part must represent the model as efficient as $K(f)$ does.
@@
 
-\section{Underfitting and overfitting}
-
-In the previous section, we saw a first example of underfitting: a UTM with a more efficient model representation became a model for all strings.
+\section{Underfitting: the universal model}
+
+In the previous section, we saw a first example of \emph{underfitting}: a second UTM with a more efficient model representation became a model for all strings.
 Even if we chose our model-description carefully, to avoid the issues of the previous section, however, universal Turing machines are still candidate model for the data.
@@
 
-The opposite problem is overfitting.
+The opposite problem is \emph{overfitting}.
 A model that can produce only $x$ can be described in as many bits as it takes to describe $x$, plus some constant amount.
-Again, such models perform as well as any other, but they are not very informative.
+Again, such models perform optimally, but they are not very informative.
 We call this \emph{overfitting}.
 
-As we shall see, most variants of sophistication are constructed specifically to prevent over- and underfitting.
-
-\subsection{Underfitting: the universal model}
+As we shall see, most variants of sophistication are constructed specifically to prevent the problems of over- and underfitting.
 
@@
 
-\begin{lemma}\label{lem:thecoolone}
+\begin{lemma}\label{lemma:thecoolone}
   Let $\psi_1,\psi_2,\ldots$ be any acceptable enumeration of the partial recursive functions.
@@
 \end{enumerate}
-Then is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
+Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
 \end{lemma}
@@
 
-\begin{lemma}
+This is a very technical and opaque lemma, due to the double effect of the choice of numbering.
+The choice of numbering affects both the model complexity and the data complexity by a constant amount.
+The idea behind the lemma, however, is simple.
+Given a subset $\M'$ of the model class, we can choose the numbering so that representations use a model in $\M'$ perform an arbitrary constant better than all other models.
+This lets us effectively `push' all other models out of the candidate set, so that the sophistication is determined by a model in $\M'$.
+
+The simplest instantiation shows a the most well known problem with sophistication: the universal model (ie.
+a UTM) is a candidate model too.
+We can manipulate the numbering so that the universal model always determines the sophistication.
+That is, the sophistication measure \emph{underfits}.
+
+\begin{lemma}[underfitting]
+Let $\M$ be a model class containing model $u$, called a \emph{universal model} with the following property
+\[
+\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p
+\]  
+Then any $\s^\M$ with the default threshold function is either non-invariant or unbounded.
+\end{lemma}
+\begin{proof}
+We will construct an acceptable numbering for which $\s^\M$ is bounded. This means that if $\s^\M$
+ is either bounded for all acceptable numberings, or not invariant.
+
+Let $D$ be a prefix function as in Lemma~\ref{lemma:thecoolone} such that it returns the index of $u$ for the argument $\epsilon$ and $\infty$ for any other argument.
+That is, $\M' = \{u\}$.
+The assumptions 1 and 2 from Lemma~\ref{lemma:thecoolone} follow directly from this construction, so that we can instantiate the Lemma.
+This tells us that there exists an acceptable numbering for which $\s_\M(x) = \s_{\M'}(x) + c$.
+Since $\M'$ contains only a single model, $\s_{\M'}$ must be constant, completing the proof.
+\end{proof}
+
+The underfitting problem is the reason that almost all sophistication measures limit their model class, usually to either finite sets ($\F$) or total functions ($\T$ or $\K$).
+This eliminates the universal model, but as we shall see in Section~\ref{section:depth}, this does not solve the problem for most data.
+The variants that are sucseptible to the problem of underfitting are facticity \cite{adriaans2012facticity} and some versions of effective complexity \ref{gellmann1996information}.
+The latter aims to bypass the problem by selecting carefully from the candidate set, but as we have shown, for some numberings the candidate set contains only the universal model.
+
+\section{Overfitting: the singletons}
+
+On the other end of the spectrum, we find the problem of optimally performing, highly-specific models.
+Most poigniantly, the singletons models.
+Sets which only contain $x$, probability distributions with $p(x) = 1$, or functions which produce $x$ for all arguments.
+Lemma~\ref{lemma:thecoolone} allows us to give these models a constant boost as well:
+
+\begin{lemma}[overfitting]
 Let $\M$ be a prefix model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$.
@@
 
-\pb{What happens if we just look at the witnesses to Komogorov complexity.}
-\pb{
-stelling: Bounded door universeel model \\
-literatuur: Gell-man, Vitanyi noemen univ model.
-oplossing: beperk de model klasse
-}
-
-\subsection{Overfitting: the singletons}
-
-\pb{
-stelling: singletons kunnen domineren in structure function.
-lit: Kolmogorov, Vitanyi, Antunes, Gell-Man
-oplossing: Paul's meaningful information?
-}
+In short, this lemma tells us that there are numberings for which the singleton models always determine the sophistication.
+This means that for any dataset, the sophistication is equal to the Kolmogorov complexity.
+
+Many variants in the literature are susceptible to this problem, including the structure function \cite{cover1985kolmogorov,gacs2001algorithmic}.
+\cite{vitanyi2004meaningful} and \cite{adriaans2012facticity} circumvent this problem by using non-self-delimiting representations.
+Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically less efficient than representations placing some information in the input.
+While this side-steps the issue of the singletons, there is no suggestion that this makes the sophistication invariant.
+
+Another solution would be to make the constant in the threshold function dependent on the choice of enumeration.
+We are not aware of any approach to this effect, and there does not seem to be a simple way to determine the thresholkd based on the properties of the numbering.
 
 \section{The problem of depth}
+
+We have seen that some variants of sophistication sidestep the simplest examples of underfitting by limiting the model class to exclude universal models.
+While this eliminates the simplest proof of boundedness, it does not follow immediately that the sophistication based on a limited model class must be bounded.
+In fact, a well-known result in algorithmic statistics shows that there exist strings for which no finite set is as good a model as the singleton set.
+We will show the consequences for sophistication:
 
@@
 We continue from the proof of Theorem~\ref{theorem:structure-function-is-not-bounded}.
-Suppose towards contradiction that either $C^\T_K(x_i)$ or $C^T_{\tau^\circ}(x_i)$ contains a candidate $f$ with $K(f)\le i/5$.
+Suppose towards contradiction that either $C^\T_K(x_i)$ or $C^\T_{\tau^\circ}(x_i)$ contains a candidate $f$ with $K(f)\le i/5$.
 Then by Lemma~\ref{lemma:total-to-sets}, there is a set $S$ with $\log|S|\le|d|\le\tau^\circ(x)\le K(x)$, so (a) in the proof of Theorem~\ref{theorem:structure-function-is-not-bounded} above cannot be the case, and $K(S)\le K(f)+K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) in the same proof for large enough $i$.
@@
 
-\pb{
-stelling: voor alle niet superdiepe strings is Paul's sophistication bounded.
-}
+This tells us that for the right strings, a sophistication so defined can truly grow to a significant part of the total description length.
+While this shows that sophistication, if carefully defined, is not entirely vacuous, we can shed some light on exactly which strings end up with high sophistication, using the principle of depth.
+We will use the following definition:
+
+\begin{definition}[depth]
+Let $U$ be some universal Turing machine, so that $U(\bar\imath y) = \phi_i(y)$.
+Let $U^t$ be a simulation of this machine, which is allowed to run for at most $t$ steps, and returns $0$ if it has not yet finished at that point.
+Let $C^t_\M(x) = \min{|\bar\imath y| : U^t(\bar\imath y) = x, \phi_i \in \M}$.
+
+The \emph{$c$-depth} $d^\M_c(x)$ of a string is defined as:
+\[
+	d^\M_c(x) = \min \left\{t : C^\M_t(x) - C^\M(x) \leq c \right\}
+\] 
+\end{definition}
+
+Thus, deep strings are those that can only be optimally compressed with a great investment of time.
+The notion of depth was first proposed by Bennett \cite{} and later refined to variants of this form \cite{}.
+
+\begin{theorem}
+Let $A(n)$ be the single-argument Ackermann function and $c$ some arbitrary constant.
+For all strings with depth $d^\C_c(x) \leq A(|x|)$, $\s^\T(x)$, $s^\P(x)$ and $\s^\F$ are bounded for specific numberings.
+\end{theorem}
+\begin{proof}
+\end{proof}
+This shows that while high-sophistication strings exist for variants with limited model class, the mechanics do not follow the basic intuition of sophistication.
+Unless we encounter data that would take longer than the lifetime of the universe to produce (and then some), sophistication behaves exactly as it would without the restricted model class.
 
@@
 
-\pb{Procedural view: we cannot between dominating models.
+\pb{Procedural view: we cannot choose between dominating models.
 This may not be the aim of algorithmic statistics, but it suggests that it's kind of unreasonable to ask for model selection without repeated samples, or assumptions on the model class.}

### 2015-01-08 Facticity/Facticity/paper.tex
...

@@
 Let $A(n)$ be the single-argument Ackermann function and $c$ some arbitrary constant.
-For all strings with depth $d^\C_c(x) \leq A(|x|)$, $\s^\T(x)$, $s^\P(x)$ and $\s^\F$ are bounded for specific numberings.
+There are numberings such that for all strings with depth $d^\C_c(x) \leq A(C(x))$, $\s^\T(x)$, $s^\P(x)$ and $\s^\F$ are bounded.
 \end{theorem}
 \begin{proof}
-\end{proof}
+Let $U(\bar\imath y)$ be some (non-prefix) universal Turing machine, and let $U^A(\bar\imath y)$ be a simulation of that machine which outputs $0$ if the number of steps taken exceeds $A(|\bar\imath y|)$.
+Let $u$ be the index of the function $U^A$ in the standard enumeration.
+
+Let $D$ be a prefix function with $D(\epsilon) = u$.
+We can instantiate Lemma~\ref{lemma:thecoolone} with $D$, $\M' = \{\phi_u\}$ and $X = \{x : d^\C_c(x) \leq A(C(x))\}$.
+This tells us that there exists a numbering for which $\s^\T(x) = \s^{\M'}(x) + |\bar0| \leq c$ for all $x \in X$.
+\end{proof}
+
 This shows that while high-sophistication strings exist for variants with limited model class, the mechanics do not follow the basic intuition of sophistication.
-Unless we encounter data that would take longer than the lifetime of the universe to produce (and then some), sophistication behaves exactly as it would without the restricted model class.
+Unless we encounter data that would take longer than the lifetime of the universe (and then some) to produce, sophistication behaves exactly as it would without the restricted model class: a solution to a superexponential problem would have no greater sophistication that a string sampled from a simple Bernoulli model.
 

### 2015-01-08 Facticity/Facticity/paper.tex
abstract

@@
 \begin{abstract}
-Kolmogorov complexity is a sound expression of the information content of a binary string, but it clashes with our intuitive notions of information: the strings with maximal information content are not the ones we would regard as interesting, or meaningful.
-One attempt to capture this idea of `useful' information is \emph{sophistication}.
-It is a simple idea with many attempted formalizations: describe the string efficiently with a two-part coding approach, and take the complexity of the model as a measure of the amount of interesting information.
-Kolmogorov complexity is invariant to the ad-hoc choices in its construction: change them, and the result will not change more than a constant.
-We review the history of sophistication critically, asking whether invariance exists for the proposed sophistication functions as well.
-In most cases, we can show that it doesn't.
-We conjecture that no formalization of sophistication can be invariant, and we briefly review the consequences of this conjecture.
+Kolmogorov complexity is a sound measure of the amount information in a binary string, but it does not distinguish between the information that expresses the structural properties of the data, and noise.
+Kolmogorov's definition of the \emph{structure function} in his lecture Tallinn in 1973 was the first attempt to separate meaningful information from noise, but many alternative definitions of the \emph{sophistication}, or the amount of meaningful information in the data, have subsequently been proposed.
+This indicates that the idea is considered worth pursuing by multiple distinguished scholars, but at the same time it shows that the issue has never been resolved completely satisfactorily.
+
+In this paper we describe four fundamental problems that arise with all existing proposals, showing each of them to be \emph{unsound}.
+Based on these results we we lay down the view that this unsoundness is fundamental: while it is certainly possible to make nontrivial and useful statements about the structural properties of binary strings, it is \emph{not} possible to objectively quantify its amount of meaningful information.
 \end{abstract}
@@
 \section*{Steven's playground}
-
-Potential alternative title: ``Four problems with all current attempts to define meaningful information''
 

### 2015-01-08 Facticity/Facticity/paper.tex
...

--- 2015-01-08 Facticity/Facticity/paper.tex
@@
 
-\section{A world without sophistication}
-
-We have conjectured that sophistication is a lost cause, and that the notion is impossible to formalize in a satisfying manner.
-
-\pb{Procedural view: we cannot choose between dominating models.
-This may not be the aim of algorithmic statistics, but it suggests that it's kind of unreasonable to ask for model selection without repeated samples, or assumptions on the model class.}
-\pb{The problem with making a fixed choice.}
-\pb{The existence of balanced strings.}
-\pb{Total functions. They ruin efficiency and push sophistication into a land of unprovable claims.}
-\pb{reference Antunes instability}
-
-However, even if sophistication must be abandoned, that does not mean that the efforts surrounding it are lost. The are many elements of algorithmic statistics that do not rely 
-\pb{Non-stochastic strings are in some sense invariant}
-\pb{We can eliminate almost all models. Complexity is countevidence.No difficult business with total functions.}
-
-Finally, the question remains how we can solve the matter of meaningful information without the ability to perform model selection. \pb{All candidates likely Take the finite set case. }
+\section{Conclusion}
+
+As we have seen, the idea behind sophistication has a rich history.
+It has been proposed many times, by many different authors: sometimes building on the work of others, sometimes proposing it independently.
+This tells us two things.
+Firstly, there is a strong intuition behind the idea of sophistication, a sense that it can be both meaningful and properly defined.
+Secondly, it shows that it is not an easy intuition to pin down in an unambiguous manner.
+
+So why should we consider this intuition reasonable at all?
+The starting point behind sophistication is that high structure and high randomness are objectively `uninteresting'.
+A more sophisticated signal, such as a television broadcast must contain, in some objective sense, more meaningful information.
+Whether this intuition seems reasonable likely differs per person, but sophistication goes one step further: it also posits an unambiguous separation into structure and randomness.
+Not for nothing is the sophistication also known as the \emph{algorithmic minimal sufficient statistic}.
+
+Let us consider whether an unambiguous separation of structure and noise is valid.
+Imagine a prefix-free universal Turing machine, to which we feed random bits whenever it reads from its input tape.
+We continue doing so until it halts (or indefinitely as the case may be).
+In other words, we are sampling from $m(x)$, the universal distribution.
+What can we say about the resulting data?
+If the universal Turing is constructed in the conventional manner, it will first sample random bits to select a non-universal Turing machine, and then sample an output from that.
+What's the characteristic model of this data?
+Certainly, we cannot exclude the possibility of a universal model.
+Also, we cannot exclude the possibility that the model actually selected was a singleton model.
+Ultimately, selecting that model that was actually used in generating the data comes down to guessing where the model ended and the input began.
+In an unbounded scenario, this is an entirely arbitrary choice.
+
+It must be said that such probabilistic views are strongly discouraged by the proponents of algorithmic statistics, mostly due to the sentiment expressed earlier in the words of Kolmogorov.
+Information theory, should precede probability, not the other way around.
+Nevertheless, this sampling scenario somewhat undercuts the intuition behind sophistication: there is no reason to assume that a string with high sophistication was sampled from a complex model.
+Repeated samples offer an unambiguous golden standard, seeing the same string a million times, strongly suggests a singleton model,
+
+For this reason, we advocate a skeptical view of sophistication.
+In particular the notion that it can be made invariant must be mentioned and investigated skeptically.
+We offer no generic proofs to the contrary, be we do offer the following insight.
+As we have seen, candidates within a constant of the optimal representation can occur all over the spectrum from small to large models.
+A small change in construction, such as a different acceptable numbering, or a change in threshold function, can arbitrarily push models into or out of the candidate set.
+Whereas in Kolmogorov complexity constant changes have constant effects on the outcome, here constant changes can cause very large changes in the outcome.
+
+The most common solution, restricting the models to be total, is problematic.
+We have show that the behavior of the resulting function does not correspond with the intuition behind it.
+Additionaly, we must not that this restriction seems arbitrary.
+What makes the total functions better than any other model class?
+If it is an arbitrary choice, then the sophistication should be invariant to changing it to some other value.
+If it isn't, which model classes are permissable and why?
+Finally, we note that the use of total functions has a very subtle effect on the discussion surrounding sophistication.
+Since the total functions are not enumerable, the value of the sophistication relies strongly on the difference between the Komogorov complexity and the
 

### 2015-01-08 Facticity/Facticity/paper.tex
abstract

--- 2015-01-08 Facticity/Facticity/paper.tex
@@
 Kolmogorov complexity is a sound measure of the amount information in a binary string, but it does not distinguish between the information that expresses the structural properties of the data, and noise.
-Kolmogorov's definition of the \emph{structure function} in his lecture Tallinn in 1973 was the first attempt to separate meaningful information from noise, but many alternative definitions of the \emph{sophistication}, or the amount of meaningful information in the data, have subsequently been proposed.
+Kolmogorov's definition of the \emph{structure function} in his lecture in Tallinn in 1973 was the first attempt to separate meaningful information from noise, but many alternative definitions of the \emph{sophistication}, or the amount of meaningful information in the data, have subsequently been proposed.
 This indicates that the idea is considered worth pursuing by multiple distinguished scholars, but at the same time it shows that the issue has never been resolved completely satisfactorily.

### 2015-01-08 Facticity/Facticity/paper.tex
abstract

@@
 In this paper we describe four fundamental problems that arise with all existing proposals, showing each of them to be \emph{unsound}.
-Based on these results we we lay down the view that this unsoundness is fundamental: while it is certainly possible to make nontrivial and useful statements about the structural properties of binary strings, it is \emph{not} possible to objectively quantify its amount of meaningful information.
+Based on these results we we lay down the view that this unsoundness is fundamental: while many nontrivial and useful structural properties of binary strings can be detected without a priori assumptions, it is \emph{not} possible to objectively quantify its sophistication.
 \end{abstract}

### 2015-01-15 Facticity/Facticity/paper.tex
...

--- 2015-01-08 Facticity/Facticity/paper.tex
@@
 
-\title{Four Problems for Sophistication}
+\title{Four Problems with Sophistication}
 
@@
 Kolmogorov complexity is a sound measure of the amount information in a binary string, but it does not distinguish between the information that expresses the structural properties of the data, and noise.
-Kolmogorov's definition of the \emph{structure function} in his lecture in Tallinn in 1973 was the first attempt to separate meaningful information from noise, but many alternative definitions of the \emph{sophistication}, or the amount of meaningful information in the data, have subsequently been proposed.
-This indicates that the idea is considered worth pursuing by multiple distinguished scholars, but at the same time it shows that the issue has never been resolved completely satisfactorily.
+Kolmogorov's definition of the \emph{structure function} in his lecture Tallinn in 1973 was the first attempt to separate meaningful information from noise, but many alternative definitions of the \emph{sophistication}, or the amount of meaningful information in the data, have subsequently been proposed.
+This indicates that the idea is considered worth pursuing by multiple distinguished scholars, but at the same time it shows that the issue has never been completely satisfactorily resolved.
 
@@
 
+\section{Overfitting: the singletons}
+
+On the other end of the spectrum, we find the problem of optimally performing, highly-specific models.
+Most poigniantly, the singletons models.
+Sets which only contain $x$, probability distributions with $p(x) = 1$, or functions which produce $x$ for all arguments.
+Lemma~\ref{lemma:thecoolone} allows us to give these models a constant boost as well:
+
+\begin{lemma}[overfitting]
+Let $\M$ be a prefix model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$.
+Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the prefix partial recursive functions, and a constant $c$, such that
+\[
+K(x)-S^{\M,\phi}(x)\le c
+\]
+for all $x\in\X$.
+\end{lemma}
+\begin{proof}
+Let $\psi_1,\psi_2,\ldots$ be any default enumeration of the partial recursive prefix functions.
+Note that since $f$ is a prefix function, if $f$ is defined for input $\epsilon$ then it cannot be defined for any other input.
+Pick $f,x$ with $f(\epsilon)=x$.
+Note that $x$ can be computed from $f$ and a fixed program, so there is a $c$ such that $K(x)\le K(f)+c$.
+Vice versa, given any $x$ we can construct an index of $f$, since $\psi$ is an acceptable numbering.
+Therefore $|K(f)-K(x)|\le c$.
+
+We now define a computable function $D$ by $D(\bar\imath p)=j$ where $\psi_j(\epsilon) = \psi_i(p)$.
+We will show that the two conditions of Lemma~\ref{lem:thecoolone} hold for the prefix function $D$.
+
+\begin{enumerate}
+\item Let $f$ be any function in the range of $D$, and $x$ its output. Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\psi_i(q)=x\}=K(x)\le K(f)+c$.
+\item On the one hand $L^{\M',\psi}(x)\le K(f)+|\epsilon|\le K(x)+c$. On the other hand, $L^{\M,\psi}(x)$ is an effective description of $x$, so $K(x)$ is at most a constant larger.
+Together, these inequalities establish the second condition.
+\end{enumerate}
+
+Then by Lemma~\ref{lem:thecoolone} there is an enumeration $\phi_1,\phi_2,\ldots$ such that $S^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
+We observed that $|K(f)-K(x)|\le c$ for any $f\in\M'$, so $S^{\M',\psi}(x)\ge K(x)+c$.
+This proves the lemma.
+\end{proof}
+
+In short, this lemma tells us that there are numberings for which the singleton models always determine the sophistication.
+This means that for any dataset, the sophistication is equal to the Kolmogorov complexity.
+
+Many variants in the literature are susceptible to this problem, including the structure function \cite{cover1985kolmogorov,gacs2001algorithmic}.
+\cite{vitanyi2004meaningful} and \cite{adriaans2012facticity} circumvent this problem by using non-self-delimiting representations.
+Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically less efficient than representations placing some information in the input.
+While this side-steps the issue of the singletons, there is no suggestion that this makes the sophistication invariant.
+
+Another solution would be to make the constant in the threshold function dependent on the choice of enumeration.
+We are not aware of any approach to this effect, and there does not seem to be a simple way to determine the thresholkd based on the properties of the numbering.
+
 \section{Underfitting: the universal model}
@@
 
-The simplest instantiation shows a the most well known problem with sophistication: the universal model (ie.
+The simplest instantiation shows the most well known problem with sophistication: the universal model (ie.
 a UTM) is a candidate model too.
@@
 The latter aims to bypass the problem by selecting carefully from the candidate set, but as we have shown, for some numberings the candidate set contains only the universal model.
-
-\section{Overfitting: the singletons}
-
-On the other end of the spectrum, we find the problem of optimally performing, highly-specific models.
-Most poigniantly, the singletons models.
-Sets which only contain $x$, probability distributions with $p(x) = 1$, or functions which produce $x$ for all arguments.
-Lemma~\ref{lemma:thecoolone} allows us to give these models a constant boost as well:
-
-\begin{lemma}[overfitting]
-Let $\M$ be a prefix model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$.
-Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the prefix partial recursive functions, and a constant $c$, such that
-\[
-K(x)-S^{\M,\phi}(x)\le c
-\]
-for all $x\in\X$.
-\end{lemma}
-\begin{proof}
-Let $\psi_1,\psi_2,\ldots$ be any default enumeration of the partial recursive prefix functions.
-Note that since $f$ is a prefix function, if $f$ is defined for input $\epsilon$ then it cannot be defined for any other input.
-Pick $f,x$ with $f(\epsilon)=x$.
-Note that $x$ can be computed from $f$ and a fixed program, so there is a $c$ such that $K(x)\le K(f)+c$.
-Vice versa, given any $x$ we can construct an index of $f$, since $\psi$ is an acceptable numbering.
-Therefore $|K(f)-K(x)|\le c$.
-
-We now define a computable function $D$ by $D(\bar\imath p)=j$ where $\psi_j(\epsilon) = \psi_i(p)$.
-We will show that the two conditions of Lemma~\ref{lem:thecoolone} hold for the prefix function $D$.
-
-\begin{enumerate}
-\item Let $f$ be any function in the range of $D$, and $x$ its output. Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\psi_i(q)=x\}=K(x)\le K(f)+c$.
-\item On the one hand $L^{\M',\psi}(x)\le K(f)+|\epsilon|\le K(x)+c$. On the other hand, $L^{\M,\psi}(x)$ is an effective description of $x$, so $K(x)$ is at most a constant larger.
-Together, these inequalities establish the second condition.
-\end{enumerate}
-
-Then by Lemma~\ref{lem:thecoolone} there is an enumeration $\phi_1,\phi_2,\ldots$ such that $S^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
-We observed that $|K(f)-K(x)|\le c$ for any $f\in\M'$, so $S^{\M',\psi}(x)\ge K(x)+c$.
-This proves the lemma.
-\end{proof}
-
-In short, this lemma tells us that there are numberings for which the singleton models always determine the sophistication.
-This means that for any dataset, the sophistication is equal to the Kolmogorov complexity.
-
-Many variants in the literature are susceptible to this problem, including the structure function \cite{cover1985kolmogorov,gacs2001algorithmic}.
-\cite{vitanyi2004meaningful} and \cite{adriaans2012facticity} circumvent this problem by using non-self-delimiting representations.
-Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically less efficient than representations placing some information in the input.
-While this side-steps the issue of the singletons, there is no suggestion that this makes the sophistication invariant.
-
-Another solution would be to make the constant in the threshold function dependent on the choice of enumeration.
-We are not aware of any approach to this effect, and there does not seem to be a simple way to determine the thresholkd based on the properties of the numbering.
 
@@
 Unless we encounter data that would take longer than the lifetime of the universe (and then some) to produce, sophistication behaves exactly as it would without the restricted model class: a solution to a superexponential problem would have no greater sophistication that a string sampled from a simple Bernoulli model.
+This result also suggests that the `non-stochastic' property of strings with high sophistication\cite{shen1983concept,vereshchagin2004kolmogorov} says more about depth and totality than it does about stochasticity.
+
+\pb{We may need to cite Antunes for similar results.}
 
@@
 As we have seen, the idea behind sophistication has a rich history.
-It has been proposed many times, by many different authors: sometimes building on the work of others, sometimes proposing it independently.
+It has been proposed many times, by many different authors.
 This tells us two things.
-Firstly, there is a strong intuition behind the idea of sophistication, a sense that it can be both meaningful and properly defined.
-Secondly, it shows that it is not an easy intuition to pin down in an unambiguous manner.
+Firstly, there is a strong intuition behind sophistication, shared by many respected authors, and second, it is not easy to get the definition exactly right.
 
 So why should we consider this intuition reasonable at all?
-The starting point behind sophistication is that high structure and high randomness are objectively `uninteresting'.
+The first assumption of sophistication is that high structure and high randomness are objectively `uninteresting'.
 A more sophisticated signal, such as a television broadcast must contain, in some objective sense, more meaningful information.
-Whether this intuition seems reasonable likely differs per person, but sophistication goes one step further: it also posits an unambiguous separation into structure and randomness.
-Not for nothing is the sophistication also known as the \emph{algorithmic minimal sufficient statistic}.
-
-Let us consider whether an unambiguous separation of structure and noise is valid.
+Whether this intuition seems reasonable likely differs per person, but sophistication goes one step further: it also posits an objective separation of a string into structure and randomness.
+
+Let us consider whether this assumption is valid.
 Imagine a prefix-free universal Turing machine, to which we feed random bits whenever it reads from its input tape.
@@
 What can we say about the resulting data?
-If the universal Turing is constructed in the conventional manner, it will first sample random bits to select a non-universal Turing machine, and then sample an output from that.
+If the universal Turing machine is constructed in the conventional manner, it will first sample random bits to select a non-universal Turing machine, and then sample an output from that.
 What's the characteristic model of this data?
@@
 Ultimately, selecting that model that was actually used in generating the data comes down to guessing where the model ended and the input began.
-In an unbounded scenario, this is an entirely arbitrary choice.
-
-It must be said that such probabilistic views are strongly discouraged by the proponents of algorithmic statistics, mostly due to the sentiment expressed earlier in the words of Kolmogorov.
-Information theory, should precede probability, not the other way around.
-Nevertheless, this sampling scenario somewhat undercuts the intuition behind sophistication: there is no reason to assume that a string with high sophistication was sampled from a complex model.
-Repeated samples offer an unambiguous golden standard, seeing the same string a million times, strongly suggests a singleton model,
-
-For this reason, we advocate a skeptical view of sophistication.
-In particular the notion that it can be made invariant must be mentioned and investigated skeptically.
-We offer no generic proofs to the contrary, be we do offer the following insight.
+In an unbounded scenario, an entirely arbitrary choice, since the singleton and the universal model are equally likely explanations for the same data.
+
+Such a probabilistic view is strongly discouraged by the proponents of algorithmic statistics, as expressed earlier in the words of Kolmogorov: information theory should precede probability theory.
+Nevertheless, this sampling scenario undercuts the intuition behind sophistication: there is no reason to assume that a string with high sophistication was sampled from a complex model, so how can the complex model be intrinsic to the data, moreso than any other model that reaches the Kolmogorov complexity.
+
+For this reason, we take a skeptical view of sophistication.
+In particular the question whether it can be defined in an invariant way must be asked, and carefully investigated.
+It is exceedingly difficult to prove that this is impossible, but we can offer the following insight.
 As we have seen, candidates within a constant of the optimal representation can occur all over the spectrum from small to large models.
@@
 Whereas in Kolmogorov complexity constant changes have constant effects on the outcome, here constant changes can cause very large changes in the outcome.
-
-The most common solution, restricting the models to be total, is problematic.
-We have show that the behavior of the resulting function does not correspond with the intuition behind it.
-Additionaly, we must not that this restriction seems arbitrary.
+While no authors have touched on the issue of invariance explicitly, it is strongly related to the issue of instability, discussed in \cite{antunes2013sophistication,vereshchagin2013algorithmic}.
+
+The most common approach, restricting the models to be total, is problematic.
+We have shown that the behavior of the resulting function does not correspond with the intuition behind it.
+Additionally, we must note that this restriction seems arbitrary.
 What makes the total functions better than any other model class?
-If it is an arbitrary choice, then the sophistication should be invariant to changing it to some other value.
+If it is an arbitrary choice, then the sophistication should be invariant to changing it to some other reasonable class.
 If it isn't, which model classes are permissable and why?
 Finally, we note that the use of total functions has a very subtle effect on the discussion surrounding sophistication.
-Since the total functions are not enumerable, the value of the sophistication relies strongly on the difference between the Komogorov complexity and the
+Since the total functions are not enumerable, the value of the sophistication relies strongly on the difference between the Kolmogorov complexity and the Kolmogorov complexity bounded to total functions.
+This gap is exceedingly difficult to capture, because the total functions cannot be enumerated.
+
+This is dangerous: if such definitions cannot be proved wrong or right, they will stand as the ad-hoc official view, while the restrictions used to elude proofs of incorrectness make them very difficult to build on.
+This creates an artificial dead-end for what is at heart a useful and valuable topic of research.
+
+\subsection{A world without sophistication}
+
+It is too early to say that sophistication is an impossible ideal, but the points we have raised certainly show that it is riddled with subtle pitfalls, and inspired by questionable intuitions.
+We will conclude by asking what it would mean if we were to abandon this ideal entirely.
+
+Of course the definition of Kolmogorov complexity still stands, and the space of two-part descriptions, ordered by strength of compression is still meaningful.
+If we give up on the ability to distinguish between representations that are close to the minimum, and consider them equivalent, we lose the ability to select a model, but we can still eliminate all those representation that are far from the Kolmogorov complexity.
+This gives us a kind of hypothesis test for a model.
+The probability of the model forming a good hypothesis is $2^|\bar\imath y|$, where $\bar\imath y$ is the representation.
+We can exclude any model if we know that that there exists a more probable representation.
+We might worry whether, in this scenario, the Kolmogorov complexity might compress better than the original model.
+
+Consider the following metaphor.
+We are given a datastring which can be read as a high-resolution bitmap image of the painting\emph{Impression of a Sunrise}.
+The candidate set allows us to say that this data represents an object, an image, a painting, an impressionist painting, a Monet and that it is specifically `Impression of a sunrise'.
+Each of these descriptions are models, which require further specification of the exact data, and each of these descriptions may end up in the candidate set.
+Which model takes precedence is a subjective choice.We cannot  say that it is in any way \emph{more} of an image than it is a painting, or vice versa.
+
+It does allow us to say, however, that it is unlikely to be a Jackson Pollock painting, or a piece of music.
+And, if we were to get a second sample from the same source, we \emph{could} answer the question: we would receive either the same painting again, another Monet, the same painting again, or simply another `thing', and we would gain information about the model.
+
+The question of meaningful information can be resolved by other means.
+There is firstly, the notion of depth in its different variants, which tells us that the complexity of a representation can be measured by its use of computational resources.
+Additionally, we might look at the patterns found  by models in the candidate set.
+In both views, the candidate set must be relatively homogenous.
+For instance, all programs must exceed the depth of $x$.
+WE know of no framework that would allow this kind of investigation, but it is conceivable that all one-part codes that approximate the Kolmogorov complexity exploit the same patterns.
+It is only the separation into random and non-random structure that is in question.
+
+We note that these model-selection problems are not specific to Algorithmic statistics.
+They can largely be directly translated to the MDL and Bayesian frameworks.
+While proponents in these fields have never made explicit claims of asymptotically objective model selection, it is a sobering thought that when the model class is extended to all computable distributions, model selection on a single sample becomes arbitrary.
+(Among the candidate set)
+
+Our conclusion, is thus that in giving up sophistication, we do not lose Algorithmic Statistics, but rather clean up some its more unmanageable components.
+Letting go of the minimal sufficient statistic will allow us to return to the simplicity of $\K$ as model classes.
+This gives us a simple framework, with a one-to-one relation to Bayesian statistics and MDL.
+Within this framework, we can simply study the consequences of assuming only that the source of our data is a computable process with access to a source of randomness.
 
@@
 
-This publication was supported by the Dutch national program COMMIT and by the the Netherlands eScience center.
+This publication was supported by the Dutch national program COMMIT and by  the Netherlands eScience center.
 

### 2015-01-15 Facticity/Facticity/paper.tex
intro

--- 2015-01-08 Facticity/Facticity/paper.tex
@@
 
+\newcommand{\phiv}{\mathbf\phi}
+\newcommand{\psiv}{\mathbf\psi}
+
 \newtheorem{conj}{Conjecture}
@@
 
-\title{Four Problems for Sophistication}
+\title{Three Problems for Sophistication}
 
@@
 
-In this paper we describe four fundamental problems that arise with all existing proposals, showing each of them to be \emph{unsound}.
-Based on these results we we lay down the view that this unsoundness is fundamental: while many nontrivial and useful structural properties of binary strings can be detected without a priori assumptions, it is \emph{not} possible to objectively quantify its sophistication.
+In this paper we describe three fundamental problems that arise with all existing proposals, showing each of them to be \emph{unsound}.
+Based on these results we we put forward the view that the problem is fundamental: while many nontrivial and useful structural properties of binary strings can be detected without a priori assumptions, it is \emph{not} possible to objectively quantify its sophistication.
 \end{abstract}
 
-\section*{Steven's playground}
-
-Requirements for sophistication:
-
-(1) Being an information measure, the value should count the number of bits required for an effective description of the meaningful properties of a binary object.
-
-(2) The Invariance Theorem for Kolmogorov complexity states that a change in reference UTM can only influence the complexity up to an additive constant term.
-Analogously, if sophistication is to be interpreted as an objective property of the data, then there should be strict limits on the degree to which a change in the UTM can affect the sophistication.
-
-(3) Sophistication should not be $\gamma$-equivalent to 0.
-While a valid information measure, the function 0 is clearly at odds with the intuition that there is no universal cap on the amount of meaningful information in the world.
-
-(4) Sophistication should not be $\gamma$-equivalent to the Kolmgorov complexity, because then there is no point in defining a new word for the same concept.
-Moreover, Kolmogorov complexity is at odds with the intuition that there is no meaningful information in a random string.
-
 \section{Introduction}
+
+Kolmogorov complexity breaks away from the tradition in statistics and information theory by allowing the analysis of \emph{individual} binary sequences rather than studying probability distributions.
+This is important because for many binary strings, such as a binary encoding of a novel, or a holiday snapshot, or the fossil record, it is impossible to determine what distribution generates such objects.
+There is, however, a nagging discrepancy between Kolmogorov complexity and what most people whould intuitively call ``complex''.
+For example, a sequence of a million coin flips is exceedingly unlikely to have a Kolmogorov complexity of much less than one million bits, even though there is nothing complex about flipping a lot of coins --- it's just hard work.
+For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data.
+This tradition started with a lecture by Kolmogorov in Talinn, Estonia in 1973, but many alternative definitions of meaningful information, or \emph{sophistication}, were proposed, see Section~\ref{sec:related_work} for an overview.
+In this paper, we investigate three problems that have plagued these definitions, and show that existing definitions of sophistication are unsatisfactory.
+We believe the problem to be fundamental, and provide several intuitive arguments why these attempts to define sophistication must fail.
+
+The Kolmogorov complexity of a binary string $x$ is, roughly, the length of the shortest computer program that will output $x$.
+This length depends on the chosen computer language.
+However, the Invariance Theorem states that the choice between any two Turing complete languages affects the Kolmogorov complexity by at most a constant term: if $K^L(x)$ denotes Kolmogorov complexity of $x$ with respect to computer language $L$, then there $|K^L(x)-K^M(x)|<c$ for some constant $c$ that depends on $L$ and $M$, but \emph{not on $x$}.
+This means that for sufficiently complex objects, the choice of computer language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings.
+The invariance theorem allows us to fix a single reference computer language and omit it from the superscript.
+
+A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees if it is to be interpreted as an objective property of a binary string.
+We are interested in a measure satisfying the following requirements:
+
+\begin{enumerate}
+\item Being an information measure, the sophistication should count the number of bits required for an effective description of the structural properties of a binary string.
+\item An analogue of the Invariance Theorem should hold for sophistication: at the very least, there should be strict limits on the degree to which a change in the used computer language can affect the sophistication.
+\item There should be no constant $c$ such that $S(x)\le c$ for every input sequence $x$.
+If sophistication is bounded, then its value can change entirely when another computer language is adopted, even for inputs of very high Kolmogorov complexity.
+Moreover, a bounded sophistication would be at odds with the intuition that there is no limit to the amount structure that a binary string can exhibit.
+\item Similarly, there should be no constant $c$ such that the $K(x)-S(x)\le c$, because then sophistication would be equivalent to Kolmogorov complexity, and the sophistication would incorporate the amount of noise in the data just like Kolmogorov complexity does.
+\end{enumerate}
+
+For most of proposed definitions of sophistication, we can prove that they fail one or more of the attributes above.
+Vit\'anyi's definition \cite{TODO} is the one exception that we cannot \emph{prove} to be at odds with our requirements, but we can show that only extremely deep strings (i.e., strings that require an enormous amount of processing to construct) yield a large sophistication, which is at odds with the intuition that there can be objects that there can be objects that contain a lot of structure that can be easily described.
+Moreover in our discussion section we provide arguments that suggest that it should be possible to generalise our negative results, so that they show that Vit\'anyi's sophistication is not invariant.
+
+A valid definition of sophistication must contend with the following issues.
+First, while all definitions are based on a two-part encoding of the data sequence, some researchers use Kolmogorov complexity as a building block of their proposed code, while others look for the two parts of the code \emph{inside} the Kolmgorov complexity: the definition of Kolmgorov complexity is opened up, and is observed to consist of two parts, one of which is then interpreted as sophistication.
+In Section~\ref{sec:nickname}, we show how this second approach sets us up to suffer from to the so-called \emph{nickname problem}: the sophistication almost trivially becomes highly dependent on the chosen enumeration of partial recursive functions.
+This immediately breaks invariance of the sophistication, unless additional requirements are made on the enumeration.
+We also show that there do exist optimally efficient enumerations of the partial recursive functions that address this issue.
+
+The second issue is that of \emph{overfitting}, which we consider in Section~\ref{sec:overfitting}.
+Overfitting is a common problem in the statistical literature, that refers to the tendency of some model selection procedures to select a model that provides a very good fit to the observed data, but does not generalise well to unseen data.
+For example, in a polynomial regression setting, a higher order polynomial can always provide better fit to a data set.
+Therefore, even if the data are in fact drawn from a second order polynomial with gaussian errors, naive model selection methods will select a higher order polynomial that provides a worse description of the trend in the data.
+The same can occur for the sophistication.
+We somehow need to make sure that the two-part representation that determines the sophistication does not store noise in the ``structure'' part of the code.
+In statistics, this is typically addressed by penalising complex models, e.g.
+using the BIC criterion \cite{TODO}.
+However, such penalisations tend to break the delicate balance between structural information and noise, and lead to the opposite problem: underfitting.
+
+Underfitting is the subject of Section~\ref{sec:underfitting}.
+It is an especially serious problem for sophistication because the considered models for sophistication are so powerful.
+In particular, in any numbering of the partial recursive functions, there are universal elements that can mimic any other partial recursive function.
+Such universal elements are a good models for \emph{any} data sequence, and if complex models are penalized to avoid overfitting, then the problem becomes to make sure that they are not \emph{always} preferred for complex data.
+This problem is widely recognised, and is sometimes referred to as the ``collapse'' of the structure function.
+The usual workaround is to restrict the set of considered models such that universal elements are no longer allowed.
+For example, the model class may be restricted to the set of all total recursive functions rather than all partial recursive functions.
+While this patches the ``collapse'' problem to an extent, it is very questionable whether it adequately solves the problem of underfitting in general.
+
+Finally, in the discussion in Section~\ref{sec:discussion} we argue that while two-part coding can yield useful insights into the structure of the data and allows us to identify some models as unlikely to capture all structure in the data, it is probably not possible to identify a \emph{single} model as ``best'': in general many models, of vastly different complexities, may be reasonable candidates for describing the data generating mechanism.
+Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that sometimes the data allows for multiple very different interpretations, and there is no such thing as sophistication.
+
+\subsection{Related work}
+This originated with Kolmogorov's lecture about his \emph{structure function}, in Tallinn, Estonia in 1973.
+This work was never published, but summarized by Cover \cite{cover1985kolmogorov} and later discussed more extensively, and generalised, by Vereshchagin, G\'acs and Vit\'anyi \cite{vereshchagin2004kolmogorov,gacs2001algorithmic}).
+Koppel \cite{koppelSoph1988,koppel1991almost} built on this notion to define a measure of the meaningful information in a binary string, called \emph{sophistication}, which is the term we will adopt in this paper.
+A similar notion called ``effective complexity'' was described by Gell-Mann and Lloyd \cite{gellmann1996information}.
+Antunes and Fortnow, worried about a certain instability with Koppel's sophistication, tweaked the definition to obtain ``coarse sophistication'' \cite{antunes2009sophistication}.
+Adriaans is the first to address the p a different problem in Koppel's definition gain very similar is Adriaans' \emph{facticity}\cite{adriaans2012facticity}.
+
+\subsection{Old stuff}
 
@@
 
-Koppel \cite{koppelSoph1988,koppel1991almost} built on this notion, coining the phrase \emph{sophistication}.
 He changed the class of models from finite sets to total functions and used monotone functions, instead of prefix-free Kolmogorov complexity.

### 2015-01-15 Facticity/Facticity/paper.tex
intro v1

--- 2015-01-15 Facticity/Facticity/paper.tex
@@
 
-\title{Four Problems with Sophistication}
+\title{Three Problems for Sophistication}
 
@@
 
-In this paper we describe four fundamental problems that arise with all existing proposals, showing each of them to be \emph{unsound}.
-Based on these results we we lay down the view that this unsoundness is fundamental: while many nontrivial and useful structural properties of binary strings can be detected without a priori assumptions, it is \emph{not} possible to objectively quantify its sophistication.
+In this paper we describe three fundamental problems that arise with all existing proposals, showing each of them to be \emph{unsound}.
+Based on these results we we put forward the view that the problem is fundamental: while many nontrivial and useful structural properties of binary strings can be detected without a priori assumptions, it may be impossible to objectively quantify its sophistication.
 \end{abstract}
 
-\section*{Steven's playground}
-
-Requirements for sophistication:
-
-(1) Being an information measure, the value should count the number of bits required for an effective description of the meaningful properties of a binary object.
-
-(2) The Invariance Theorem for Kolmogorov complexity states that a change in reference UTM can only influence the complexity up to an additive constant term.
-Analogously, if sophistication is to be interpreted as an objective property of the data, then there should be strict limits on the degree to which a change in the UTM can affect the sophistication.
-
-(3) Sophistication should not be $\gamma$-equivalent to 0.
-While a valid information measure, the function 0 is clearly at odds with the intuition that there is no universal cap on the amount of meaningful information in the world.
-
-(4) Sophistication should not be $\gamma$-equivalent to the Kolmgorov complexity, because then there is no point in defining a new word for the same concept.
-Moreover, Kolmogorov complexity is at odds with the intuition that there is no meaningful information in a random string.
-
 \section{Introduction}
 
-\noindent Kolmogorov complexity is an objective measure of the information content of an object.
-Although its definition contains several arbitrary choices, these only affect its value up to a constant additive term.
-This property is crucial for the measure's success: it allows us to view Kolmogorov complexity as a \emph{property of the data} rather than an arbitrary function computed on it.
-This, in turn, allows us to define the intrinsic entropy, or randomness, inherent in a single object expressed as a binary string, without any assumptions on the process that produced it.
-However, this notion of complexity does not in all respects match our intuitive idea of which objects are complex, and which objects are not.
-
-Imagine an old-fashioned, analog television set which can be tuned to three frequencies.
-The first produces a blank screen, the second an ordinary TV show, and the third white noise.
-Judging by their Kolmogorov complexity, these  broadcasts have increasing information content.
-The first is easiest to describe, the second is harder but still contains a great deal of regularity, and the third cannot be compressed.
-However, intuitively speaking the second signal clearly conveys more meaningful information than the other two.
-This notion has proved far harder to capture in formal terms than the simple notion of `raw' information content.
-
-Over the years one solution in particular has been proposed in many different forms.
-We select a description for the data by two-part coding; describing a model first, and then the data in terms of the model.
-The size in bits, used to describe the model is taken as a measure of the useful information in our string.
-This idea has been proposed many times in many different guises, which shows that it is both a persistent intuition but also that it is a difficult idea to formalize in a meaningful way.
-For the purposes of this paper, we will refer to this concept as \emph{sophistication}\footnotemark.
-
-\footnotetext{As the historical overview shows, sophistication is only one name for this idea, and not the first.
-The alternatives, however, are more awkward, and sophistication has been used by different authors for different realizations of the same principle.}
-
-The intuitive idea of sophistication makes one crucial demand: we can find an objectively value for the model complexity of a string, either by having a single objectively optimal two-part description, or a set of descriptions with equivalent model complexity.
-In the framework of Kolmogorov complexity, two observers cannot disagree by more than a constant about the information content of a dataset.
-In formal terms: the Kolmogorov complexity of a dataset $x$ is defined as the shortest effective description for $x$ in some Turing complete description language.
-This description language is usually expressed as a universal Turing machine $U(\bar\imath y) = \phi_i(y)$ where $\phi_i$ is the $i$'th partial computable function in some effective enumeration, and the $\bar\imath$ is a self-delimiting encoding of $i$.
-If two people disagree about the shortest effective description of $x$, we can interpret this as them using different description languages, ie.
-different universal Turing machines.
-The crucial aspect of Kolmogorov complexity, the starting theorem that makes it interesting as a function at all, is \emph{invariance}: two people using different description languages will get the same value for a string's Kolmogorov complexity, albeit up to a fixed constant.
-Without invariance, Kolmogorov complexity would be an arbitrary function.
-With invariance, we can be sure that we are talking about an intrinsic property of the data.
-The same consideration must be given to each proposal for formalising sophistication: is it invariant?
-
-There are many approaches to sophistication, but the following pattern is common to all of them.
-We start with a space of \emph{models}.
-There can be finite sets, computable probability distributions or simple functions from strings to strings.
-The function perspective is the most generic, all other model classes can be translated to it.
-The approach to sophistication is to find a representation for a dataset $x$ in two parts: a computable function $f$ and and input $y$, such that $f(y) = x$.
-As with Kolmogorov complexity, the shorter a description is, the better.
-We take the shortest description and those close to it, and consider them \emph{candidates}.
-How close a description must be to the optimum differs between variants, one popular choice is to use a pre-chosen constant value.
-We then choose a single description from the candidate set and return the description length of the model as the sophistication.
-Usually, the candidate with the smallest model determines the sophistication.
-
-The implicit assumption behind sophistication is that this approach will separate the description of the string into a structural part, representing its regularities, and a stochastic part representing its non-deterministic properties.
-In other words, the string is cast as a random choice from a model for which it is `typical'.
-
-In this article we will show that most proposals are not invariant.
-\sdr{Ik wil dit nog steeds graag sterker gesteld hebben.} For one remaining case, we can offer no such proof, but no proof of invariance exists either.
-This is likely due to the complex nature of the model class chosen, and the difficulty of resolving the matter of invariance shows how complex such a sophistication measure would be to build on, even if it did prove to be invariant.
-
-\subsection{A brief historical overview}
-
-We first provide a brief chronological overview of the existing variants of sophistication. We will return to each in-depth in the body of the paper.
-
-The first suggestion that a representation of a string can be separated cleanly into an encoding of its regular and its stochastic proeperties came from Kolmogorov himself, in a talk given at a conference in Tallinn in 1973 (first committed to paper by Cover \cite{cover1985kolmogorov} and later discussed more extensively discussed by Vereshchagin and Vit\'anyi \cite{vereshchagin2004kolmogorov}).
-Here, he proposed what has become known as the \emph{Kolmogorov structure function}.Minimizes over two part representations for a given string $x$, consisting of some finite set $S$ (containing $x$) and the index of $x$ in $S$.
-The \emph{minimal sufficient statistic} is the set with the lowest complexity for which the length of the two-part representation reaches the prefix-free Kolmogorov complexity.
-It was first shown by Shen \cite{shen1983concept} that there are strings for which the best model cannot be small.
-
-Koppel \cite{koppelSoph1988,koppel1991almost} built on this notion, coining the phrase \emph{sophistication}.
-He changed the class of models from finite sets to total functions and used monotone functions, instead of prefix-free Kolmogorov complexity.
-We expect that this change was made to extend the theory to infinite strings, an approach which has been abandoned by subsequent authors.
+Kolmogorov complexity breaks away from the tradition in statistics and information theory by allowing the analysis of \emph{individual} binary sequences rather than studying probability distributions.
+This is important because for many binary strings, such as a binary encoding of a novel, or a holiday snapshot, or the fossil record, it is impossible to determine what distribution generates such objects.
+There is, however, a nagging discrepancy between Kolmogorov complexity and what most people whould intuitively call ``complex''.
+For example, a sequence of a million coin flips is exceedingly unlikely to have a Kolmogorov complexity of much less than one million bits, even though there is nothing complex about flipping a lot of coins---it's just hard work.
+For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data.
+This tradition started with a lecture by Kolmogorov in Talinn, Estonia in 1973, but many alternative definitions of meaningful information, or \emph{sophistication}, were proposed, see the related work section below.
+In this paper, we investigate three problems with these definitions, and show that they are unsatisfactory.
+We believe the problem to be fundamental, and provide several intuitive arguments why these attempts to define sophistication must fail.
+
+The Kolmogorov complexity of a binary string $x$ is, roughly, the length of the shortest computer program that will output $x$.
+This length depends on the used programming language.
+However, the Invariance Theorem states that the choice between any two Turing complete languages affects the Kolmogorov complexity by at most a constant term: if $K^L(x)$ denotes Kolmogorov complexity of $x$ with respect to language $L$, then $|K^L(x)-K^M(x)|<c$ for some constant $c$ that depends on $L$ and $M$, but \emph{not on $x$}.
+This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings.
+The invariance theorem allows us to fix a single reference programming language and omit it from the superscript.
+
+A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees if it is to be interpreted as an objective property of a binary string.
+We are interested in a measure satisfying the following requirements:
+
+\begin{enumerate}
+\item Being an information measure, the sophistication should count the number of bits required for an effective description of the structural properties of a binary string.
+\item An analogue of the Invariance Theorem should hold for sophistication: at the very least, there should be strict limits on the degree to which a change in the used programming language can affect the sophistication.
+\item There should be no constant $c$ such that $S(x)\le c$ for every input sequence $x$.
+If sophistication is bounded, then knowing its value under one programming language provides no constraints on its value under another language, even for inputs of very high Kolmogorov complexity.
+Moreover, a bounded sophistication would be at odds with the intuition that there is no limit to the amount structure that a binary string can exhibit.
+\item Similarly, there should be no constant $c$ such that the $K(x)-S(x)\le c$ for all $x$, because then sophistication would be equivalent to Kolmogorov complexity, and the sophistication would incorporate the amount of noise in the data just like Kolmogorov complexity does.
+\end{enumerate}
+
+For most proposed definitions of sophistication, we can prove that they fail one or more of the attributes above.
+Vit\'anyi's definition \cite{TODO} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we can show that only extremely deep strings (i.e., strings that require an enormous amount of processing to construct) yield a large sophistication, while intuitively it seems there must be objects that contain a lot of structure all of which can be described without much calculation.
+Moreover in our discussion section we provide arguments that suggest that it should be possible to generalise our negative results to show that Vit\'anyi's sophistication is not invariant.
+
+A valid definition of sophistication must contend with the following issues.
+First, while all definitions are based on a two-part encoding of the data sequence, some researchers use Kolmogorov complexity as a \emph{building block} of their proposed code, while others look for the two parts of the code \emph{inside} the Kolmgorov complexity: when the definition of Kolmogorov complexity is opened up, it turns out to consist of two parts itself.
+In Section~\ref{sec:nickname}, we show how this second approach sets us up to suffer from to the so-called \emph{nickname problem}: the sophistication becomes highly dependent on the chosen programming language.
+This immediately breaks invariance, unless additional requirements are made on the language.
+We also show that there do exist languages (efficient enumerations of the partial recursive functions) that address this issue.
+
+The second issue is that of \emph{overfitting}, which we consider in Section~\ref{sec:overfitting}.
+Overfitting is a common problem in the statistical literature, that refers to the tendency of some model selection procedures to select a very complex model that provides a very good fit to the observed data, but does not generalise well to unseen data.
+For example, in a polynomial regression setting, a higher order polynomial can always provide better fit to a data set.
+Therefore, even if the data are in fact noisy samples from a parabola, naive model selection methods may select a much higher order polynomial that provides a worse description of the trend in the data.
+The same can occur for the sophistication.
+We somehow need to make sure that the two-part representation that determines the sophistication does not store noise in the ``structure'' part of the code.
+In statistics, this is typically addressed by penalising complex models, e.g.
+using the BIC criterion \cite{TODO}.
+However, such penalisations tend to break the delicate balance between structural information and noise, and lead to the opposite problem: underfitting.
+
+Underfitting (Section~\ref{sec:underfitting}) is the opposite problem, where the selected model is simple, but fails to represent part of the structure in the data.
+It is an especially serious problem for sophistication because the considered models for sophistication are so powerful.
+In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language.
+Such interpreters can generate \emph{any} data sequence with only a short input, and if complex models are penalized to avoid overfitting, then the problem becomes to make sure that interpreters are not \emph{always} preferred for complex data.
+This problem is widely recognised, and is sometimes referred to as the ``collapse'' of the structure function \cite{}.
+The usual workaround is to restrict the set of considered models such that interpreters are no longer allowed.
+For example, the model class may be restricted to the set of all total recursive functions (only programs that never hang) rather than all partial recursive functions (all programs).
+While this patches the ``collapse'' problem to an extent, it is questionable whether it adequately solves the problem of underfitting in general.
+
+Finally, in the discussion in Section~\ref{sec:discussion} we argue that while two-part coding can yield useful insights into the structure of the data and allows us to identify some models as unlikely to capture all structure in the data, it is probably not possible to uniquely separate structure from noise and identify a \emph{single} model as ``best'': in general many models, of vastly different complexities, may be reasonable models of the data generating mechanism.
+Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that sometimes the data allows for multiple very different interpretations, and that there is no such thing as sophistication.
+
+\subsection{Related work}
+The idea of separating structure from noise originated with Kolmogorov's lecture in Tallinn, Estonia in 1973 about the \emph{structure function},.
+This work was never published, but summarized by Cover \cite{cover1985kolmogorov}.
+In this theory, the two-part description of the data consists of a program that generates a finite set, and an index of the original object in the set.
+The length of the program then measures the structure in the data.
+This approach was discussed more extensively, and generalised, by Vereshchagin, G\'acs and Vit\'anyi \cite{vereshchagin2004kolmogorov,gacs2001algorithmic}).
+
+Koppel \cite{koppelSoph1988,koppel1991almost} also built on Kolmogorov and Cover's work to define a measure of the meaningful information in a binary string, called \emph{sophistication}, the term we adopt in this paper.
+He changed the class of models from finite sets to total functions and used monotone functions, instead of prefix-free Kolmogorov complexity, to be able to extend the theory to infinite strings, an approach which has been abandoned by subsequent authors.
 We will only consider the finite version (based on regular functions) in this paper.
-
-While Koppel expands the model class from finite sets to functions,he does require that the functions be total.
-This is a recurring theme: if the model class is expanded to all computable functions, the universal Turing machine $U$ itself becomes a model.
-If we assume that $U$ computes the function $\phi_u$ in the standard enumeration of computable functions, and that $\bar\imath y$ is a shortest description for $x$ on $U$ (ie.
-$U(\bar\imath y) = x)$), then $\br{u}\br{\imath}y$ is also a description for $x$ on $\bar u$.
-This description has a constant model size ($|\br{u}|$), ie.
-every $x$ has constant sophistication.
-This `collapse' to a universal model is the reason that most treatments of sophistication constrain the model class to exclude the universal Turing machine.
-This makes it more expensive to move information out of the model.
-As we shall see, defining sophistication quickly becomes a balancing act, forcing information one way or the other.
 
@@
 Working from the perspective of statistical mechanics, they arrived (largely independently, it seems) at a notion that fits the mold of sophistication.
-Effective complexity describes a binary string in terms of a probability distribution over finite strings (an `ensemble' in the terminology of Gell-Mann and Lloyd) and the entropy of the distribution.
-The reasoning is that, if the string is typical for the distribution, the entropy is sufficient to describe the string given the distribution, so that the length of the description of the distribution and the size of its entropy suffices as a two-part description.
-The two-part description which determines the effective complexity is chosen by placing a time bound (taken from the singleton description) on the computation of the string from the distribution.
-The hope is that this will move deep (ie.
-hard to compute) information to the model part, and shallow information to the data.
-In a later treatment\cite{gell2004nonextensive}, Gell-Mann and Lloyd select the decisive two-part representation from the candidates based on ``the average quantities judged to be important'', which seems to be a conscious choice to make the measure subjective and context-dependant (sacrificing invariance).
-
-G\'acs, Tromp and Vit\'anyi \cite{gacs2001algorithmic}, built on the idea of the structure function to establish a framework of \emph{algorithmic statistics}, as an attempt to found a statistical theory not on probabilistic notions, but on combinatorial ones.
-This echoes Kolmogorov's own hope for the theory of information:
-\begin{quotation}
-\noindent Information theory must precede probability theory and not be based on it.
-By the very essence of this discipline, the foundations of information theory must have a finite combinatorial character.
-\cite{kolmogorov1983combinatorial}
-\end{quotation}
-Algorithmic statistics studies binary strings and their possible two-part representations.
-In a later paper \cite{gacs2001algorithmic}, Vit\'anyi specifically studies the role of the minimal sufficient statistic as an expression of the `meaningful' information of a string (taking the name sophistication from Koppel).
-Like Koppel, Vit\'anyi restricts the model class to total functions, eliminating the universal model.
 
 Antunes et al.
-\cite{antunes2009sophistication} note that the constant cutoff used to separate candidate two-part descriptions from those deemed to be too inefficient, is a source of instability.
-Small changes to this constant could shift the result sophistication by arbitrary amounts.
-They introduce a more elaborate criterion for the cutoff, calling this variant \emph{coarse sophistication}.
+\cite{antunes2009sophistication} note that the constant cutoff used by Koppel to separate candidate two-part descriptions from those deemed to be too inefficient, is a source of instability, a result in the spirit of this paper.
+They propose a different definition called \emph{coarse sophistication}.
 As we shall see, this version behaves very differently from regular sophistication, and is closer to computational depth.
@@
 Adriaans' \emph{Facticity}, introduced in 2012\cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions.
-By constructing a very specific universal Turing machine, and taking constant terms into account, the facticity is expected to become robust against the collapse to a universal model.
-
-\pb{Vereshchagin's strongly sufficient statistic}
-
-\pb{Antunes' instability result}
-
-\pb{Diagrammetje}
+Adriaans also identifies the nickname problem discussed in Section~\ref{sec:nickname}.
 

### 2015-01-16 Facticity/Facticity/paper.tex
...

@@
 
+% --- DELETE BEFORE SUBMUSSIONS ---
+\pagestyle{headings}
+
 \title{Three Problems for Sophistication}
@@
 \begin{abstract}
-Kolmogorov complexity is a sound measure of the amount information in a binary string, but it does not distinguish between the information that expresses the structural properties of the data, and noise.
+Kolmogorov complexity is a sound measure of the amount of information in a binary string, but it does not distinguish between the information that expresses the structural properties of the data, and noise.
 Kolmogorov's definition of the \emph{structure function} in his lecture Tallinn in 1973 was the first attempt to separate meaningful information from noise, but many alternative definitions of the \emph{sophistication}, or the amount of meaningful information in the data, have subsequently been proposed.
@@
 
-Before we start, we will define a framework into which we can fit all current variants of sophistication.
-The principle is always the same: we start with a class of models, we consider the space of two-part descriptions using these models, we select the set of representations with the shortest code-lengths and we select, by some method, the best model from this set.
+We will generalizes the main proposals for sophistication into a single, consistent notation.
+The unifying principle is this: we start with a class of models, we consider the space of two-part descriptions using these models, we select the set of representations with the shortest code-lengths and we select, by some method, the best model from this set.
 
-Let $\B = {0,1}^*$.
+Let $\B = \{0,1\}^*$.
 We deal with partial functions $\phi: \B \to \B$, which we will also call \emph{models}.
@@
 
-\footnotetext{Completeness is analogous to totality, in that a prefix-free Turing machine that does not halt corresponds to a complete prefix-free function.}
+\footnotetext{Completeness is analogous to totality: a total function corresponds to a Turing machine that halts for all its inputs, while a complete function corresponds to a Turing machine that always halts, so long as there are enough bits on the input tape.}
 
-\begin{definition}[Models and model classes]
-  A \emph{model} is a function $\phi:\B\to\B$. A
-  \emph{model class} is a set of models. We distinguish four important
-  model classes:
+As noted in the introduction, the Kolmogorov complexity is invariant to the choice of description language. Formally, our description language creates a numbering of the partial computable functions.
+
+\begin{definition}[Numberings]
+A numbering is an enumeration of $\C$, denoted as $\phi_1, \phi_2, \ldots$ or by a function $g: \N \to \C$.
+We fix one canonical numbering $g^\circ$, which works as an effective description, that is for any $i \in \N, y\in \B$, we can effectively compute $(g^\circ(i))(y)$.
+We call a numbering $g$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ as that for all $i$ $g^\circ(a(i)) = h(i)$ and  $g^\circ(i) = h(b(i))$.
+\end{definition}
+
+When we say that Kolmogorov complexity is invariant, we mean that changing the canonical acceptable numbering changes the Kolmogorov complexity by at most a constant.
+
+\begin{definition}[Prefix encoding function]
+  A prefix encoding function $f:\B\to\B$ is an injective
+  function that maps the binary strings to a prefix-free set. There
+  are prefix encoding functions $g$ with the property that
+  $|g(x)|\le|x|+2\log(|x|+2)|$ \cite{li1993introduction}. We fix such a function and
+  denote it by $\br{x}=g(x)$.
+\end{definition}
+
+\begin{definition}[Complexity]
+Let $\phi_1,\phi_2,\ldots$ be the canonical acceptable enumeration of $\C$.
+Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ corresponfs to the set of all computable prefix functions.
+Now define Kolmogorov complexity by
+\begin{align*}
+C(x)&=\min\{|\bar\imath y|:\phi_i(y)=x\},\\
+K(x)&=\min\{|\bar\imath y|:\phi_i(y)=x, i\in\Pi\}.\\
+\end{align*}
+\end{definition}
+This definition of $K$ is uncommon.
+It is equivalent to the standard definition as used in \cite{li1993introduction}, up to a constant, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).
+
+We now have the basic tools with which we set up our generic sophistication function. We start with the class of allowed models:
+
+\begin{definition}[Models and model classes] A \emph{model class} is a set of models. We distinguish four important model classes:
   \begin{itemize}
@@
 \end{definition}
-\begin{definition}[Numberings]
-A numbering is an enumeration of $\C$, denoted as $\phi_1, \phi_2, \ldots$ or by a function $g: \N \to \C$.
-We fix one canonical numbering $g^\circ$, which works as an effective description, that is for any $i \in \N, y\in \B$, we can effectively compute $(g^\circ(i))(y)$.
-We call a numbering $g$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ as that for all $i$ $g^\circ(a(i)) = h(i)$ and  $g^\circ(i) = h(b(i))$.
-\end{definition}
 
-\begin{definition}[Prefix encoding function]
-  A prefix encoding function $f:\B\to\B$ is an injective
-  function that maps the binary strings to a prefix-free set. There
-  are prefix encoding functions $g$ with the property that
-  $|g(x)|\le|x|+2\log(|x|+2)|$ \cite{li1993introduction}. We fix such a function and
-  denote it by $\br{x}=g(x)$.
-\end{definition}
-
-\begin{definition}[Complexity]
-Let $\phi_1,\phi_2,\ldots$ be the canonical acceptable enumeration of $\C$.
-Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\} = \K$.
-Now define Kolmogorov complexity by
-\begin{align*}
-C(x)&=\min\{|\bar\imath| y:\phi_i(y)=x\},\\
-K(x)&=\min\{|\bar\imath| y:\phi_i(y)=x, i\in\Pi\}.\\
-\end{align*}
-\end{definition}
-This definition of $K$ is uncommon.
-It is functionally equivalent to the standard definition as used in \cite{li1993introduction}, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).
-
-We now have the basic tools with which we set up our generic sophistication function.
-We first define the space of two-part representations for a given string $x$, and the codelength that each representation achieves:
+We note that the model classes $\T$ and $\P$ are unusual in that they are not enumerable.
 
@@
 \]
-with $c$ a constant independent of $x$ which may be chosen arbitrarily. The default threshold function is $\tau^\circ(x)=L^\M(x)$.
+with $c$ a constant independent of $x$ which may be chosen arbitrarily.
+The default threshold function is $\tau^\circ(x)=L^\M(x)$.
+We will often omit the subscript if the default threshold function is used.
 \end{definition}
@@
 \end{definition}
-For a function like sophistication to be meaningful, it must satisfy certain properties. We will check all currently published variants against the following minimal requirements:
 
-\begin{description}
-  \item[Invariance] Like the Kolmogorov complexity, the construction of a sophistication function makes certain ad-hoc choices, which are captured in the choice of an arbitrary universal Turing machine $U$ or an arbitrary numbering of $\C$.
-If we change what choice we make, the Kolmogorov complexity remains invariant up to a constant.
-This is the sole reason we allow ourselves to interpret the Komogorov complexioty as a property of the data, rather than simply the result of some computation on it.
-A similar invariance must hold for any sophistication function.
-   \item[Unboundedness] Both the sophistication and the residual information (roughly the complexity minus the sophistication) must not be bounded by a constant.
-If the sophistication is bekow a constant for all strings it cannot have much meaning.
-Also, if it is always withing a constant of the Kolmogorov complexity, it cannot correspond to the notion we hope to capture with it.
-   \item[Efficiency] The two-part representation $(m, y)$ that witnesses the sophistication (that is the one that reaches the minimum described by $\s^\M_\tau$) must reach the Kolmogorov complexity (or at least the difference should be insignificant).
-If there is an effective one-part representation that is much more efficient than the two-part representation that witnesses the sophistication, we cannot, in all honesty, claim that $(m, y)$ captures all the information in $x$.
-\end{description}
+Finally, we must formalize what we mean when we say that we desire sophistication to be invariant.
+For the Kolmogorov complexity, we know that a change in acceptable numbering causes no more than a constant change.
+It may be that a less stringent invariance (for instance a $\log$ term) can still yield a worthwhile definition of sophistication.
+To allow such distinctions, we generalize the notion of invariance:
 
-Given the philosophical underpinnings of sophistication, these three requirements are the very least that we could expect. We will show that no current proposal satisfies all three.
+\begin{definition}
+Let $\gamma:\N\to \N$ be some nondecreasing function. Two functions $f$ and $g$ are $\gamma$-equivalent if $|f(x) - g(x)| \in O(\gamma(\max(f(x), g(x)))$.
+\end{definition}
 
-For all variants but one, we show that all measures that are invariant are bounded.
-This leads us to the following conjecture.
-Let $\gamma(n)$ be a function representing a \emph{granularity}: the extent to which we care about differences between functions.
-That is, we will call two functions $f$ and $g$ $\gamma$-equal if $|f(x) - g(x)| \in O\left(\gamma\left(f(x)\right)\right)$.
-
-\begin{conj}
-Let $\s_g$ be a sophistication function constructed as in Definition~\ref{definition:sophistication}, built on some acceptable numbering $g$, and some fixed model class $\M$ and threshold function $\tau$ (omitted from the sub/superscript for the sake of clarity).
-
-For any sophistication function so constructed, and any granularity $\gamma$ we have that for any two acceptable numberings $g$ and $h$, either:
-\begin{itemize}
-  \item there exist $\s_g$ and $\s_h$ which are not are $\gamma$-equal (ie $\s$ is not invariant),
-  \item $s_g$ is $\gamma$-equal to $0$ or to the Kolmogorov complexity.
-\end{itemize}
-\end{conj}
-
-In these terms, all instantiations of the Kolmogorov complexity, using different Universal Turing machines are $1$-equal.
-It may be unreasonable to require the sophistication to attain the same level of invariance (a log-invariant sophistication could still be useful).
-This conjecture allows such relaxations: if it is true, sophistication must be considered fundamentally flawed.
-We hope that this conjecture will prompt authors to address the matter of invariance critically, and to consider the possibility that sophistication and its ilk are an untenable proposition.
-
-In the rest of this paper, we will analyse the existing variants one by one.
-We will conclude with the informal reasoning supporting this conjecture and consider its the consequences.
-What does this mean for the problem of meaningful information, what does it mean for algorithmic statistics and most importantly, what are the consequences for the idea of model selection in general?
+We say that a formalization of sophistication $s_\phi$ built on acceptable numbering $\phi$ is $\gamma$-invariant to the choice of numbering is $s_\phi$ and $s_\psi$ are $\gamma$-equal for any two acceptable numberings $\phi$ and $\psi$.
+Thus, by this definition, Kolmogorov complexity is $1$-invariant.
 
@@
 We start from a simple intuition: the definition of Kolmogorov complexity already uses a description of the data in two parts; a computable function and and input.
-Can we assume that such a description, that is as efficient as the Kolmogorov complexity, creates a good separation into the structural part of the data and the noise?
+Can we assume that when such a description is as short as the Kolmogorov complexity, it automatically creates a good separation into the structural part of the data and the noise?
+As we will show, the representation of the structural part plays a fundamental part in answering this question.
+If it is in any way inefficient, the resulting sophistication function does not behave as expected.
 
-The first problem that arises is that allowing any acceptable numbering can lead to inefficient model representations.
+The issues can be very subtle and require a careful attention that is not always afforded in the literature. We will first illustrate what happens if any acceptable numbering is allowed:
 
@@
 
-No such function can be $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
+No such function can be both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
 \end{lemma}
@@
 
-In this proof, we deliberately use ridiculously inefficient index functions.
-Obviously one could simply pick a more efficient index; the point is that many definitions of sophistication to not mention any such restriction.
-As we shall see in the next, any inefficiency in the index will lead to a lack of invariance in sophistication.
+In this proof, we deliberately used ridiculously inefficient index functions.
+Obviously, one could simply pick a more efficient index; the point is that many definitions of sophistication do not mention any such restriction.
+As we shall see, any inefficiency in the index will lead to a lack of invariance in sophistication.
 
@@
 
-Interestingly, using an inefficient index is no problem in the definition of Kolmogorov complexity, since we can select as a model a UTM with an efficient index at only a constant penalty (by the Invariance Theorem \cite{li1993introduction}).
-But this robustness property does \emph{not} carry over to definitions of sophistication, if we are interested in separating the model and its input, we must represent the model efficiently.
+Interestingly, using an inefficient index is no problem in the definition of Kolmogorov complexity, since we can select as a model a universal Turing machine with an efficient index at only a constant penalty (by the Invariance Theorem \cite{li1993introduction}).
+But this robustness property does \emph{not} carry over to definitions of sophistication.
+If we are interested in separating the model and its input, we must represent the model efficiently.
 
-Another perspective ion this problem is that the length of a model in an ordinary description does not represent its complexity, much like the length of a compressible string does not reflect its information content.
+Another perspective on this problem is that the length of a model in an ordinary description does not represent its complexity, much like the length of a compressible string does not reflect its information content.
 In order to obtain a more robust measure of model complexity, we describe a natural extension of Kolmogorov complexity to functions, and show that this retains invariance.
@@
 \end{proof}
-This lemma shows that the complexity of a function depends on the choice of
-enumeration by no more than a constant term.
+This lemma shows that the complexity of a function does not depend on the choice of
+enumeration by more than a constant term.
 
-There are two ways to use this notion of model complexity for more robust attempts to define sophistication, that differ in the way the two-part code is conceptualized.
-Both approaches are used in the literature, which is therefore sometimes confusing.
+There are two ways to use this notion of model complexity for more robust attempts to define sophistication. Confusingly, both approaches are used in the literature,
 \begin{enumerate}
-  \item On the one hand one can stick with Koppel's approach, but repair it by imposing the additional requirement that the used length of the indices used is equal to $C(\phi_i)$, and that the numbering is still acceptable.
+  \item One can stick with the approach of using the model used by the representation that witnesses the Kolmogorov complexity, but repair it by imposing the additional constraint that an efficient numbering is used.
+Specifically, that the length of the indices satisfies $|i| = C(\phi_i)$, and that the numbering is still acceptable.
 We construct such an index below.
@@
 This approach is the most direct formalization of the intuition that the two part code used within the definition of Kolmogorov complexity separates structure from noise.
-  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use Kolmogorov complexity as a black box to measure the complexity of the model, which is then the first part of a two-part code describing the data.
+  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use Kolmogorov complexity as a black box to measure the complexity of the \emph{model}, which is then the first part of a two-part code describing the data.
 With this approach, the total code length is $K(f)+|y|$, where $y$ is an input to the model $f$ such that $f(y)=x$.
-The difference is that in this approach, we can construct everything on the basis of a single, unconstrained, acceptable numbering, without using additional contraints.
+The difference is that in this approach, we can construct everything on the basis of a single acceptable numbering, without using additional contraints.
 It is also more natural in this setting to limit the model class of allowed functions.
@@
 
-To show that the two approaches are equivalent with the correct restriction on the index, we must show that there exists a numbering for which the length of the index of a model is equal to its complexity, and that such a numbering is acceptable.
+To show that the two approaches are equivalent with the correct restriction on the index, we must show that there exists an acceptable numbering for which the length of the index of a model is equal to its complexity.
 
@@
 That publication defines an \emph{index function}, rather than a numbering, and notes that it is not computable.
-Since we use numberings, we only need to show that they are acceptable, ie.
-we only need to be able to unpakc the efficient description, so the incomputability of the index function is no issue:
+Since a numbering is all that is needed, we only need to show acceptability.
+The incomputability of the index function is no issue:
 \begin{lemma}
@@
 \begin{proof}
-Let $i_\tn{div}(y)$ be an index such that $\phi_{i_\tn{div}}(y)=\infty$ for all $y$. Define
+Let $\tn{div} \in \N$ be an index such that $\phi_{\tn{div}}(y)=\infty$ for all $y$. Define
   \[\psi_q=\begin{cases}
@@
 \begin{lemma}
-Let $L(\phi) = |\bar\imath|$ for a faithful index, and some prefix encoding function, and let $\s^\C$ be defined on this $L$ with the default threshold function.
+Let $L(\phi_i) = |\bar\imath|$ for a faithful index, and some prefix encoding function, and let $\s^\C$ be defined on this $L$ with the default threshold function.
 
-If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then $s^\C$ is bounded for all numberings $\phi$
+If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then $\s^\C$ is bounded for all numberings $\phi$.\label{lemma:prefix-inefficiency}
 \end{lemma}
 \begin{proof}
-Let $\bar\imath p$ be any two-part representation for the data $x$, i.e.
-$\phi_i(p)=x$.
-Then construct an alternative two-part representation $\bar vi^* p$, where $i^*$ is the shortest $\psi_u$-program for $i$ such that $\phi_v(i^* p)=\phi_{\psi_u(i^*)}(p) = \phi_i(p)=x$.
+Let $\bar\imath y$ be any two-part representation for the data $x$, i.e.
+$\phi_i(y)=x$.
+Then construct an alternative two-part representation $\bar vi^* y$, where $i^*$ is the shortest $\psi_u$-program for $i$ such that $\phi_v(i^* y)=\phi_{\psi_u(i^*)}(y) = \phi_i(y)=x$.
 We compare the lengths of these two representations.
@@
 \[
-|\bar\imath p|-|\bar v i^* p| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-K(i).
+|\bar\imath p|-|\bar v i^* y| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-K(i).
 \]
 By assumption there must be an $i_0$ such that the above expression is positive for all $i>i_0$.
-From this $i$ onwards, the second representation (using $v$) will have a shorter code length, so $\bar\imath p$ cannot achieve the minimum in the definition of the facticity.
-Consequently, the facticity is bounded by $F^\phi(x)<|\overline{\imath_0}|$ for all $x$.
+From this $i_0$ onwards, the second representation (using $v$) will have a shorter code length, so $\bar\imath p$ cannot achieve the minimum in the definition of the sophistication.
+Consequently, the sophistication is bounded by $\s(x)<|\overline{\imath_0}|$ for all $x$.
 \end{proof}
@@
 
-Note that the condition for this lemma applies to all known partial computable prefix functions. We conjecture that it applies to all:
+Note that the condition for this lemma applies to all commonly used prefix encoding functions. We conjecture that it applies to all possible computable prefix encoding functions:
 
 \begin{conjecture}
-For any partial computable prefix code length function $L$, the function
+For any partial computable prefix codelength function $L$, the function
 \[
@@
 
-This result shows that even if we have a faithful numbering, we can only use it in settings where we don't count the cost of delimiting the model from the input (such as \cite{koppelSoph1988}).
-In the cases where the data representations are taken as a single code word, the model-part must represent the model as efficient as $K(f)$ does.
-Instead of a numbering, we require a self-delimiting code for functions rather than a numbering.
+Lemma~\ref{lemma:prefix-inefficiency} shows that even if we have a faithful numbering, we can only use it in settings where we don't count the cost of delimiting the model from the input (such as \cite{koppelSoph1988}).
+In the cases where the data representations are taken as a single code word, the model-part, including delimiting information, must represent the model as efficiently as $K(f)$ does.
 
-In the rest of this paper, we will take the second perspective: we will consider the representations as two-part descriptions $K(f)$, and $|y|$, such that $f(y)=x$.
-This description can be unpacked by a very specific universal Turing machine, but contrary to the construction of the Kolmogorov complexity, we are not free to choose any UTM.
+In the rest of this paper, we will avoid these issues and take the second perspective: we will consider the representations as two-part descriptions $(f, y)$ with codelength $K(f) + |y|$, such that $f(y)=x$.
 
@@
 
-On the other end of the spectrum, we find the problem of optimally performing, highly-specific models.
-Most poigniantly, the singletons models.
-Sets which only contain $x$, probability distributions with $p(x) = 1$, or functions which produce $x$ for all arguments.
-Lemma~\ref{lemma:thecoolone} allows us to give these models a constant boost as well:
+We now come to the main issue plaguing sophistication.
+In the naive setting, where we allow all prefix-free functions as models, and use the default threshold function, the candidate set always contains two representations that clearly do not provide a good separation into structure and content.
+The first is a universal model: a universal Turing machine which simply runs its shortest program for $x$.
+The second is a \emph{singleton model} a model which returns $x$ for any input.
+In the first representation all information about $x$ is stored in the second part of the two-part description.
+In the second representation all information is stored in the model.
+The model complexity suggested by these representations is either constant, or equal to the Kolmogorov complexity.
 
-\begin{lemma}[overfitting]
-Let $\M$ be a prefix model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$.
-Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the prefix partial recursive functions, and a constant $c$, such that
-\[
-K(x)-S^{\M,\phi}(x)\le c
-\]
-for all $x\in\X$.
-\end{lemma}
-\begin{proof}
-Let $\psi_1,\psi_2,\ldots$ be any default enumeration of the partial recursive prefix functions.
-Note that since $f$ is a prefix function, if $f$ is defined for input $\epsilon$ then it cannot be defined for any other input.
-Pick $f,x$ with $f(\epsilon)=x$.
-Note that $x$ can be computed from $f$ and a fixed program, so there is a $c$ such that $K(x)\le K(f)+c$.
-Vice versa, given any $x$ we can construct an index of $f$, since $\psi$ is an acceptable numbering.
-Therefore $|K(f)-K(x)|\le c$.
-
-We now define a computable function $D$ by $D(\bar\imath p)=j$ where $\psi_j(\epsilon) = \psi_i(p)$.
-We will show that the two conditions of Lemma~\ref{lem:thecoolone} hold for the prefix function $D$.
-
-\begin{enumerate}
-\item Let $f$ be any function in the range of $D$, and $x$ its output. Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\psi_i(q)=x\}=K(x)\le K(f)+c$.
-\item On the one hand $L^{\M',\psi}(x)\le K(f)+|\epsilon|\le K(x)+c$. On the other hand, $L^{\M,\psi}(x)$ is an effective description of $x$, so $K(x)$ is at most a constant larger.
-Together, these inequalities establish the second condition.
-\end{enumerate}
-
-Then by Lemma~\ref{lem:thecoolone} there is an enumeration $\phi_1,\phi_2,\ldots$ such that $S^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
-We observed that $|K(f)-K(x)|\le c$ for any $f\in\M'$, so $S^{\M',\psi}(x)\ge K(x)+c$.
-This proves the lemma.
-\end{proof}
-
-In short, this lemma tells us that there are numberings for which the singleton models always determine the sophistication.
-This means that for any dataset, the sophistication is equal to the Kolmogorov complexity.
-
-Many variants in the literature are susceptible to this problem, including the structure function \cite{cover1985kolmogorov,gacs2001algorithmic}.
-\cite{vitanyi2004meaningful} and \cite{adriaans2012facticity} circumvent this problem by using non-self-delimiting representations.
-Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically less efficient than representations placing some information in the input.
-While this side-steps the issue of the singletons, there is no suggestion that this makes the sophistication invariant.
-
-Another solution would be to make the constant in the threshold function dependent on the choice of enumeration.
-We are not aware of any approach to this effect, and there does not seem to be a simple way to determine the thresholkd based on the properties of the numbering.
-
-\section{Underfitting: the universal model}
-
-In the previous section, we saw a first example of \emph{underfitting}: a second UTM with a more efficient model representation became a model for all strings.
-Even if we chose our model-description carefully, to avoid the issues of the previous section, however, universal Turing machines are still candidate model for the data.
-These candidates represent the data efficiently (they reach the Kolmogorov complexity up to a constant), but they are extremely generic: taking a second sample from the model can produce a string that has nothing to do with the data.
-Nevertheless, these models perform well, in terms of compression, and there is no justification for excluding them from the candidate set.
-
-The opposite problem is \emph{overfitting}.
-A model that can produce only $x$ can be described in as many bits as it takes to describe $x$, plus some constant amount.
-Again, such models perform optimally, but they are not very informative.
-We call this \emph{overfitting}.
-
-As we shall see, most variants of sophistication are constructed specifically to prevent the problems of over- and underfitting.
-
-Any sophistication function built on a model class that includes singletons, or universal models is in danger of over or underfitting.
-In that case, we can construct a numbering, so that a particular model, or class of models, represents all data more efficiently than any other model by an arbitrary constant.
-This constant boost is enough to force all other models out of the candidate set.
-
-The following lemma generalizes the principle. We will instantiate it for different variants of sophistication:
+The key problem is that both can be present in the candidate set, and the formalization of sophistication must ensure that neither are selected.
+As we will show, in many settings, we can ensure that they are the \emph{only} models in the candidate set.
+The principle we use is that the threshold on the candidate set is determined by a constant, and a change in numbering affects model size by a constant.
+Thus, We can choose our numbering  so that all but a certain class of models take a constant penalty, pushing them out of the candidate set.
+Since we use this trick repeatedly we have generalized it into an (admittedly complex) lemma.
 
@@
 
-This is a very technical and opaque lemma, due to the double effect of the choice of numbering.
-The choice of numbering affects both the model complexity and the data complexity by a constant amount.
-The idea behind the lemma, however, is simple.
+This is a technical and opaque lemma and proof, due to the double effect of the numbering, which occurs in both the definition of model complexity and the definition of data complexity.
+The \emph{idea} behind the lemma, however, is simple.
 Given a subset $\M'$ of the model class, we can choose the numbering so that representations use a model in $\M'$ perform an arbitrary constant better than all other models.
@@
 
-The simplest instantiation shows the most well known problem with sophistication: the universal model (ie.
-a UTM) is a candidate model too.
-We can manipulate the numbering so that the universal model always determines the sophistication.
-That is, the sophistication measure \emph{underfits}.
+We can now show that there exist numberings for which the sophistication always selects the singleton models:
+
+\begin{lemma}[overfitting]
+Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$.
+Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the prefix partial recursive functions, and a constant $c$, such that
+\[
+K(x)-S^{\M,\phi}(x)\le c
+\]
+for all $x\in\X$.
+\end{lemma}
+\begin{proof}
+Let $\psi_1,\psi_2,\ldots$ be any default enumeration of the partial recursive prefix functions.
+Note that since $f$ is a prefix function, if $f$ is defined for input $\epsilon$ then it cannot be defined for any other input.
+Pick $f,x$ with $f(\epsilon)=x$.
+Note that $x$ can be computed from $f$ and a fixed program, so there is a $c$ such that $K(x)\le K(f)+c$.
+Vice versa, given any $x$ we can construct an index of $f$, since $\psi$ is an acceptable numbering.
+Therefore $|K(f)-K(x)|\le c$.
+
+We now define a computable function $D$ by $D(\bar\imath p)=j$ where $\psi_j(\epsilon) = \psi_i(p)$.
+We will show that the two conditions of Lemma~\ref{lemma:thecoolone} hold for the prefix function $D$.
+
+\begin{enumerate}
+\item Let $f$ be any function in the range of $D$, and $x$ its output. Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\psi_i(q)=x\}=K(x)\le K(f)+c$.
+\item On the one hand $L^{\M',\psi}(x)\le K(f)+|\epsilon|\le K(x)+c$. On the other hand, $L^{\M,\psi}(x)$ is an effective description of $x$, so $K(x)$ is at most a constant larger.
+Together, these inequalities establish the second condition.
+\end{enumerate}
+
+Then by Lemma~\ref{lemma:thecoolone} there is an enumeration $\phi_1,\phi_2,\ldots$ such that $S^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
+We observed that $|K(f)-K(x)|\le c$ for any $f\in\M'$, so $S^{\M',\psi}(x)\ge K(x)+c$.
+This proves the lemma.
+\end{proof}
+
+In short, this lemma tells us that there are numberings for which the singleton models always determine the sophistication.
+This means that for any dataset, the sophistication is equal to the Kolmogorov complexity.
+
+Many variants in the literature are susceptible to this problem, including the structure function \cite{cover1985kolmogorov,gacs2001algorithmic}.
+\cite{vitanyi2004meaningful} and \cite{adriaans2012facticity} void this issue by using non-self-delimiting representations.
+Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically less efficient than representations placing some information in the input.
+While this side-steps the issue of the singletons, there is no suggestion that this makes the sophistication invariant.
+
+Another approach may be to make the constant in the threshold function dependent on the choice of enumeration.
+We are not aware of any approach to this effect, and there does not seem to be a simple way to determine the threshold based on the properties of the numbering.
+We are not aware of such considerations being mentioned anywhere in the literature.
+
+\section{Underfitting: the universal model}
+
+On the other end of the spectrum, there exist candidate representations which place all information in the `noise part', using a universal function as the model.
+We can easily show that if such models exist, we can choose our numbering so that they are always chosen.
 
@@
 
-The underfitting problem is the reason that almost all sophistication measures limit their model class, usually to either finite sets ($\F$) or total functions ($\T$ or $\K$).
-This eliminates the universal model, but as we shall see in Section~\ref{section:depth}, this does not solve the problem for most data.
-The variants that are sucseptible to the problem of underfitting are facticity \cite{adriaans2012facticity} and some versions of effective complexity \ref{gellmann1996information}.
+Most sophistication measures avoid the underfitting problem by limiting their model class, usually to either finite sets ($\F$) or total functions ($\T$ or $\K$).
+This eliminates the universal model, but as we shall see in the next section, this does not solve the problem for most data.
+The variants that are susceptible to the problem of underfitting are facticity \cite{adriaans2012facticity} and some versions of effective complexity \ref{gellmann1996information}.
 The latter aims to bypass the problem by selecting carefully from the candidate set, but as we have shown, for some numberings the candidate set contains only the universal model.
 
-\section{The problem of depth}
+\subsection{The problem of depth}
 
@@
 This tells us that for the right strings, a sophistication so defined can truly grow to a significant part of the total description length.
-While this shows that sophistication, if carefully defined, is not entirely vacuous, we can shed some light on exactly which strings end up with high sophistication, using the principle of depth.
+While this shows that sophistication, if defined with particular care, is not entirely vacuous, we can shed some light on exactly which strings end up with high sophistication, using the principle of depth.
 We will use the following definition:
@@
 Thus, deep strings are those that can only be optimally compressed with a great investment of time.
-The notion of depth was first proposed by Bennett \cite{} and later refined to variants of this form \cite{}.
+The notion of depth was first proposed by Bennett \cite{bennett1988logical} and later refined to variants of this form \cite{antunes2006computational}.
+It is important to note that it is exceedingly unlikely to sample deep string from a non-deep distribution.
+We note that it is exceedingly unlikely that a deep string is sampled from a shallow distribution \cite{bloem2014safe,bennett1988logical}.
 
@@
 This shows that while high-sophistication strings exist for variants with limited model class, the mechanics do not follow the basic intuition of sophistication.
-Unless we encounter data that would take longer than the lifetime of the universe (and then some) to produce, sophistication behaves exactly as it would without the restricted model class: a solution to a superexponential problem would have no greater sophistication that a string sampled from a simple Bernoulli model.
-This result also suggests that the `non-stochastic' property of strings with high sophistication\cite{shen1983concept,vereshchagin2004kolmogorov} says more about depth and totality than it does about stochasticity.
+Unless we encounter data that would take longer than the lifetime of the universe (and then some) to produce, sophistication behaves exactly as it would without the restricted model class: a solution to a super-exponential problem would have no greater sophistication than a string sampled from a simple Bernoulli model.
+This result also suggests that the `non-stochastic' property of strings with high sophistication\cite{shen1983concept,vereshchagin2004kolmogorov} says more about depth and totality than it does about the lack of randomness in its information content.
 
-\pb{We may need to cite Antunes for similar results.}
+The relation between sophistication, depth and absolutely non-stochastic strings was also investigated in \cite{antunes2013sophistication}.
 
@@
 A more sophisticated signal, such as a television broadcast must contain, in some objective sense, more meaningful information.
-Whether this intuition seems reasonable likely differs per person, but sophistication goes one step further: it also posits an objective separation of a string into structure and randomness.
+Whether this intuition seems reasonable likely differs per person, but sophistication goes one step further: it also suggests that we can objectively separate the information in a string into random and structured.
 
-Let us consider whether this assumption is valid.
-Imagine a prefix-free universal Turing machine, to which we feed random bits whenever it reads from its input tape.
-We continue doing so until it halts (or indefinitely as the case may be).
+To see whether this assumption is valid; imagine a prefix-free universal Turing machine, to which we feed random bits whenever it reads from its input tape.
+We continue doing so until it halts (or indefinitely if it does not).
 In other words, we are sampling from $m(x)$, the universal distribution.
 What can we say about the resulting data?
-If the universal Turing machine is constructed in the conventional manner, it will first sample random bits to select a non-universal Turing machine, and then sample an output from that.
+If the universal Turing machine is constructed in the conventional manner, the initial bits will determine the index of a partial computable function $i$, and the rest will be the input to that machine $y$.
 What's the characteristic model of this data?
-Certainly, we cannot exclude the possibility of a universal model.
-Also, we cannot exclude the possibility that the model actually selected was a singleton model.
-Ultimately, selecting that model that was actually used in generating the data comes down to guessing where the model ended and the input began.
-In an unbounded scenario, an entirely arbitrary choice, since the singleton and the universal model are equally likely explanations for the same data.
+Certainly, we cannot exclude the possibility that $\phi_i$ is also a universal model.
+Also, we cannot exclude the possibility that $\phi_i$ is a singleton model for $x$.
+Ultimately, choosing the specific model that was actually used in generating the data comes down to guessing where the model ended and the input began.
 
-Such a probabilistic view is strongly discouraged by the proponents of algorithmic statistics, as expressed earlier in the words of Kolmogorov: information theory should precede probability theory.
-Nevertheless, this sampling scenario undercuts the intuition behind sophistication: there is no reason to assume that a string with high sophistication was sampled from a complex model, so how can the complex model be intrinsic to the data, moreso than any other model that reaches the Kolmogorov complexity.
+This kind of perspective, with a probabilistic source for the data is strongly discouraged by the proponents of algorithmic statistics: they often argue that information theory should precede probability theory \cite{kolmogorov1983combinatorial}.
+Nevertheless, this sampling scenario undercuts the intuition behind sophistication: there is no reason to assume that a string with high sophistication was sampled from a complex model, so how can the complex model be intrinsic to the data, moreso than any other model that reaches the Kolmogorov complexity?
 
 For this reason, we take a skeptical view of sophistication.
-In particular the question whether it can be defined in an invariant way must be asked, and carefully investigated.
-It is exceedingly difficult to prove that this is impossible, but we can offer the following insight.
-As we have seen, candidates within a constant of the optimal representation can occur all over the spectrum from small to large models.
+The matter of its invariance, in particular cannot be glossed over.
+It should be the first question asked of any proposal.
+We can offer the following insight, to back up our skepticism.
+As we have seen, candidates within a constant of the optimal representation can occur all over the spectrum, from small to large models.
 A small change in construction, such as a different acceptable numbering, or a change in threshold function, can arbitrarily push models into or out of the candidate set.
@@
 We have shown that the behavior of the resulting function does not correspond with the intuition behind it.
-Additionally, we must note that this restriction seems arbitrary.
+We can also ask what justification there is for this restriction?
 What makes the total functions better than any other model class?
@@
 If it isn't, which model classes are permissable and why?
+
+As we have seen, the restriction to a model class $\M$ tends to assign high sophistication to those strings that can be compressed significantly better by models outside $\M$ than by models within it.
+For the class $\T$, this results in extremely deep strings getting high sophistication.
+While this is an interesting property, which is worth investigating, the interpretation of an objective separation into model and noise is spurious at best.
+Instead of interpreting these strings as ``non-stochastic'', we consider it more natural to consider them \emph{non-typical for $\M$}.
+
 Finally, we note that the use of total functions has a very subtle effect on the discussion surrounding sophistication.
@@
 If we give up on the ability to distinguish between representations that are close to the minimum, and consider them equivalent, we lose the ability to select a model, but we can still eliminate all those representation that are far from the Kolmogorov complexity.
-This gives us a kind of hypothesis test for a model.
-The probability of the model forming a good hypothesis is $2^|\bar\imath y|$, where $\bar\imath y$ is the representation.
-We can exclude any model if we know that that there exists a more probable representation.
-We might worry whether, in this scenario, the Kolmogorov complexity might compress better than the original model.
+This is analogous to hypothesis testing in classical statistics.
+We can eliminate models but not confirm them.
+Even if we knew the Kolmogorov complexity of our data, there would still be equivalent representations we could not objectively  choose between.
+Given our choice of numbering we may hav an absolute favorite, but objectively speaking, we must accept all explanations that are `close enough' to the Kolmogorov complexity.
 
 Consider the following metaphor.
-We are given a datastring which can be read as a high-resolution bitmap image of the painting\emph{Impression of a Sunrise}.
-The candidate set allows us to say that this data represents an object, an image, a painting, an impressionist painting, a Monet and that it is specifically `Impression of a sunrise'.
-Each of these descriptions are models, which require further specification of the exact data, and each of these descriptions may end up in the candidate set.
-Which model takes precedence is a subjective choice.We cannot  say that it is in any way \emph{more} of an image than it is a painting, or vice versa.
+We are given a bitstring which can be read as a high-resolution bitmap image of the painting \emph{Impression of a Sunrise}.
+The candidate set for this painting contains various models from very generic to very specific.
+The theory behind sophistication suggests that we can choose one of these as the objective, intrinsic model of the data.
+Different models say different things: the universal models says that it is a `thing'.
+A more specfic model might say that it is an image.
+Even more specific would be a painting, a Monet, or specifically the painting Impression of a sunrise.
+These are the models we imagine we may find in the candidate set.
+Can we say that the data is intrinsically more of an image than a painting?
+More of a Monet than a painting?
+We see no intuitive reason why such a distinction should be possible.
 
-It does allow us to say, however, that it is unlikely to be a Jackson Pollock painting, or a piece of music.
+Of course, Kolmogorov complexity does allow us to say that it is unlikely to be a Jackson Pollock painting, or a piece of music.
 And, if we were to get a second sample from the same source, we \emph{could} answer the question: we would receive either the same painting again, another Monet, the same painting again, or simply another `thing', and we would gain information about the model.
-
-The question of meaningful information can be resolved by other means.
-There is firstly, the notion of depth in its different variants, which tells us that the complexity of a representation can be measured by its use of computational resources.
-Additionally, we might look at the patterns found  by models in the candidate set.
-In both views, the candidate set must be relatively homogenous.
-For instance, all programs must exceed the depth of $x$.
-WE know of no framework that would allow this kind of investigation, but it is conceivable that all one-part codes that approximate the Kolmogorov complexity exploit the same patterns.
-It is only the separation into random and non-random structure that is in question.
+But a single sample by itself should give us no reason to choose an intrinsic model.
+Ultimately, this is just a metaphor, but since the notion of sophistication is driven by intuition, we feel that an intuition to the contrary should serve to sharpen the debate.
 
@@
 
-Our conclusion, is thus that in giving up sophistication, we do not lose Algorithmic Statistics, but rather clean up some its more unmanageable components.
+Our conclusion is that in giving up sophistication, we do not lose algorithmic statistics, but rather clean up some of its more unmanageable components.
 Letting go of the minimal sufficient statistic will allow us to return to the simplicity of $\K$ as model classes.
 This gives us a simple framework, with a one-to-one relation to Bayesian statistics and MDL.
-Within this framework, we can simply study the consequences of assuming only that the source of our data is a computable process with access to a source of randomness.
+
+We are left with the original question that sophistication was designed to answer.
+How can we distinguish pure randomness from intuitively interesting information.
+We note that the idea of depth is built on a more solid fundament.
+For instance, contrary to sophistication, encountering a string with high sophistication does imply, with high probability, that the source was of high depth also.
+Unfortunaly, depth is not measured in bits, and so moves us away from the idea that the useful complexity of a stirng should be measured in bits.
+It remains to be seen whether this is a fundamental issue, or whether the interesting information in a string can be objectively expressed in the same vein as it plain information content.
 
@@
 
-\section{=======Stuff to move up or delete=======}
-
-\section{Old structure}
-
-\subsection{Koppel and Atlan}
-
-We start with sophistication as originally introduced by Koppel and Atlan \cite{koppelSoph1988,koppel1991almost}.
-Their sophistication is built on a two-argument universal Turing Machine $U(i, y) = \phi_p(y)$.
-They consider a pair $(i, y)$ a representation of a string $x$ if (a) $\phi_i$ is total, and (b) $\phi_i(y)$ has $x$ as a prefix.
-The length of a representation is measured as $|i| + |y|$.
-Note that this ignores the cost of delimiting the model $i$ from the input $y$, ie.
-this is not the length of a single unambiguous code word for $x$.
-Candidates are those models with representations $c$ bits from the minimum length, where $c$ is a pre-set `significance' parameter.
-
-\footnotemark
-
-Koppel and Atlan actually provide a machine-independence result, but only in the setting of infinite strings.
-
-\footnotetext{The original sophistication also allows infinite binary strings in the range of the computable functions considered.
-We ignore this aspect for the sake of simplicity, although our conclusions still hold with such functions included.}
-
-This construction shows one of the first pitfalls of sophistication: using an inefficient representation for the model:
-
-\subsection{Antunes's Coarse Sophistication}
-
-\begin{itemize}
-\item Model class is $\T$
-\item Description method $L(m)$ for models is unclear, but appears to be
-  taken from Koppel, i.e. arbitrary but self-delimiting.
-\item Sophistication is defined slightly differently: $S^\T(x)=L(m^*)$ where $m^*=\argmin_m L(m)+L^m(x)-C(x)$. Thus, there is no slack.
-\end{itemize}
-
-\subsection{Adriaans' Facticity}
-
-\begin{itemize}
-\item Model class is $\C$
-\item Description method for models is $L(m)=|\bar\imath|$ where
-  $\phi_i=m$. The index function is assumed to be \emph{faithful}, see
-  Definition~\ref{def:faithful}
-\item Threshold function is $\tau(x)=L^\M(x)$, the constant in the candidate set is $0$.
-\item Facticity is actually defined as index length without prefix
-  encoding:  $S^\C_0(x):=\min\{|i|:\phi_i\in C^\C_0(x)\}$. But this
-  difference is not important for the discussion.
-\end{itemize}
-
-\subsection{Effective complexity}
-
-\begin{itemize}
-\item Model class is \emph{ensembles}.
-Equivalent to $\K$.
-Subsequent literature suggests that ensembles have finite support, which would mean that effective complexity reduces to the structure function.
-\item Description method for models is $L(m) = K(m)$ (the minimization is not mentioned explicitly, but we'll give them the benefit of the doubt).
-\item The threshold function is equal to the minimum two-part code: $\tau(x)=L^m(x)$.
-\item Gell-mann and Lloyd acknowledge that representations exist with small and large models.
-The choice from the candidate set is made by applying a time-bound to the computation from the string to the model, but not to the decoding of the model itself.
-The idea is that this forces `deep' information into the model, and shallow information into the data.
-Presumably, the representation with the smallest model code under this constraint is chosen.
-\end{itemize}
-
-\pb{\cite{ay2010effective} provides a more formal definition.}
-\begin{theorem}
-There exists a UTM so that for all $x$, $\s_\text{EC}(x) = K(x)$, up to a constant. 
-\end{theorem}
-
-Subsequent publications suggest that all ensemeble\ldots.
-
-\begin{theorem}[Reduction of Effective Complexity to Structure
-  Function]
-  There is a constant $c$ such that, for every representation of $x$
-  consisting of an ensemble $E$ with $x\in E$ with entropy $H$,
-  yielding a total information of $K(E)+H$, there is a finite set $S$
-  with $x\in S$ such that $K(S)\le K(E)+c$ and $\log|S|\le H$.
-\end{theorem}
-\begin{proof}
-  First note that although the ensemble $E$ may itself be infinite, it
-  is ``typical'', and therefore a valid candidate representation, for
-  only a finite number of its elements, namely the set $S=\{y\in E:-\log
-  P(y)\le H\} = \{y:P(y)\ge 2^{-H}\}$. Since the sum of probabilities must
-  converge, this set cannot be infinite: since $1\ge\sum_{y\in S} P(y)\ge
-  \sum_{y\in S}2^{-H} = |S|2^{-H}$ so $|S|\le 2^H$.
-\end{proof}
-
-\subsection{Roll your own}
-
-\begin{itemize}
-\item Model class is either $\C$ or $\K$
-\item Description method for models is $L(m)=K(m)$.
-\item Threshold function is $\tau(x)=L^m(x)+c$, which matches $C(x)$
-  resp. $K(x)$ up to a constant.
-\end{itemize}
-
-\paragraph{TODO: instantiate for UTM}
-
-\begin{itemize}
-\item Model class is $\K$
-\item Description method for models is $L(m)=K(m)$.
-\item Threshold function is $\tau(x)=L^m(x)+c$, which matches $K(x)$
-  up to a constant.
-\end{itemize}
-
-\subsection{Kolmogorov's Structure Function}
-
-\begin{itemize}
-\item Model class is $\F$
-\item Description method for models $L(m)=K(S)$ where $S$ is the range
-  of $m$. There is some ambiguity as to the definition of prefix
-  complexity for finite sets; discuss and refer to Shen \cite{TODO}.
-\end{itemize}
-
-\subsection{Vit\'anyi's Sophistication / Meaningful Information}
-
-\begin{itemize}
-\item Model class is $\T$
-\item Description method for models is $L(m)=K(m)$
-\item Candidate set is defined differently: $C_c=\{m\in\T:|L^m(x)-K(x)|\le
-  c\}$. There appears to be a problem with this definition, because
-  the two-part descriptions of $x$ may be shorter than $K(x)$ by
-  exploiting the fact that the models do not need to be prefix functions.
-\end{itemize}
-
-% stru fu: model class is all finite sets
-% code length L(m)+L_m(x)
-% acceptability criterion acc(m) = 2L(m)+L_m(x)-C(x)
-%
-
-\hide{
-
-\subsection{Plan}
-
-Perhaps surprisingly, it is also reinforced by Vit\'anyi, when he writes : (From Vereshchagin and Vit\'anyi, ``Kolmogorov’s Structure Functions and Model Selection'':)
-\begin{quotation}
-  This expression emphasizes the two-part code nature of Kolmogorov complexity. In the example
-\[x=10101010101010101010101010\]
-we can encode $x$ by a small Turing machine printing a specified number of copies of the pattern `10' which computes $x$ from the program `13'.” This way, $K(x)$ is viewed as the shortest length of a two-part code for $x$, one part describing a Turing machine, or model, for the regular aspects of $x$, and the second part describing the irregular aspects of $x$ in the form of a program to be interpreted by $T$.
-The regular, or “valuable,” information in  is constituted by the bits in the “model” while the random or ``useless'' information of $x$ constitutes the remainder.
-\end{quotation}
-
-\subsection{Conclusion}
-Stel onze conjecture is waar, is het dan helemaal nutteloos?
-Nee, want lengte van two-part representations is nog steeds een redelijke maat voor counterevidence against a model.
-No-hypercompression, $p$-values, randomness deficiency.
-Probleem zit in harde cutoffs zoals minimalisatie en een vaste constante waarboven representaties niet meer meedoen.
-Daar gaat invariantie van kapot.
-
-Onze mening: als onze conjecture waar is, dan moet je constraints weglaten: maak het makkelijk en doe prefix functies en alle partial recursive functions, en accepteer de collapse gewoon.
-Maar dat betekent weer dat je een hele wolk van two-part optimale representaties hebt waar je niet tussen moet willen kiezen.
-
-Er is een gat tussen $K(f)+|y|-K(x)$ en $\delta(x\mid f)$. De laatste is de juiste maat voor de counterevidence tegen $f$.
-
-\subsection{Thoughts and questions}
-
-\begin{itemize}
-\item Paul argues bleurg that the randomness deficiency of a model is a measure of the amount of structure left in the noise part.
-I feel that it is also a measure of the amount of counterevidence against a model in the statistical sense, as follows: the randomness deficiency of $x$ given model $M$ is $\delta(x|M)=L_M(x)-K(x|M)$.
-Suppose that model $M$ is true.
-We can then use the no hypercompression inequality to rewrite:
-%
-\[
-P_{x\sim M}(\delta(x|M)\ge c) = P_{x\sim M}(L_M(x) - K(x|M) \ge c)\le 2^{-c}
-\]
-Thus, if $M$ is true the probability of a large randomness deficiency is exponentially small!
-This amount of evidence is robust in the sense that changing the UTM around will change the amount of evidence by at most a constant.
-If the structure function is going to give us a \emph{cloud} of candidate models instead of just a single one, I think this is a really nice interpretation of the meaning of that cloud: we simply measure the available evidence against all individual models.
-
-\item Shen explains the collapse of the structure function if you use enumerating complexity rather than listing complexity.
-How can we most cleanly explain that this collapse does not vanish for weaker model classes?
-
-\item Properties of representations on $K(M)$ vs $L_M(x)$ graph: diagonals identify representations with the same two-part codelength.
-The randomness deficiency is the distance to the minimum, which is the $K(x)$ diagonal:
-\[L(x;M)=K(M)+L_M(x) =K(M)+K(x|M)+L_M(x)-K(x|M)  = K(x)+\delta(x|M).\]
- Models move horizontally by a constant when the ordering of TMs is changed.
-Thus the randomness deficiency and two part code are also changed by at most a constant.
-In contrast the \emph{minimum} can now be achieved by a very different representation.
-
-\item ``Power and peril of MDL'': bespreken ook de ``collapse of the structure function'' als de model class te sterk is, dwz sets worden te efficient gecodeerd (wsch equivalent met Shen's enumerating complexity).
-
-\end{itemize}
-
-\subsection{Dead darlings}
-
-\begin{definition}
-Given an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions, and an enumeration $\psi_1,\psi_2,\ldots$ of the partial recursive prefix functions.
-Let $\psi_u$ be a universal element with property $\psi_u(\bar\imath p)=\psi_i(p)$.
-Then define $v$ by
-\[
-v(a p)=\begin{cases}\phi_{\psi_u(a)}(p)&\tn{if $a$ is in the domain of $\psi_u$,}\\
-\infty&\tn{otherwise}.\end{cases}
-\]
-\end{definition}
-\begin{lemma}
-The function $v=\phi_v$ as defined above is partial recursive.
-\end{lemma}
-\begin{proof}
-Let $r$ be the input of $v$.
-Simulate $\psi_u$ using a TM with a one-way read-only input tape.
-If the TM does not halt, then enter an infinite loop.
-Otherwise, let $r=ap$ where $a$ is the part of the input tape that has been read when the TM halts, and let $i$ be the output of the TM.
-Now simulate $\phi_i$ on input $p$.
-\end{proof}
-
-\begin{lemma}
-Assume the prefix function $\bar\cdot$ used in the definition of complexity is such that $\bar x-x$ is monotonically increasing.
-Let $\psi_1,\psi_2,\ldots$ be any acceptable numbering.
-Then for every $c$ there is another acceptable numbering $\phi_1,\phi_2,\ldots$ such that for all partial recursive functions $f$ we have
-\[
-K^\phi(f)-K^\psi(f) \ge c.
-\] \label{lemma:building-block}
-\end{lemma}
-\begin{proof}
-First note that there is a $c'$ such that 
-\begin{equation}
-	K(1^r 0  x) + c' \geq K(x) \label{eq:prelim1}
-\end{equation}
-for all $x$, since we can create a program that first runs the shortest program for $1^r 0 x$ and then strips the prefix $1^r 0$ from the result.
-
-Define $\phi_{1^r0i}(x) = \psi_{i}(x)$, with $\phi_{j}(x) = \infty$ if $j$ does not contain a zero.
-Intuitively, any $\phi$-program is $r+1$ bits longer than the corresponding $\psi$-program.
-However, the used prefix function complicates this slightly.
-But all we need is the bound
-\begin{equation}\label{eq:prelim2}\begin{split}
-K^\phi(x) &= \min \{ |\overline{1^r 0 \imath}| y : \phi_{1^r 0 i}(y) = x \}\\
-&  \ge \min\{|1^r0\bar\imath y| : \phi_{1^r 0 i}(y) = x \}\qquad\tn{(using assumption on $\bar\cdot$)}\\
-& = r+1 +\min\{|\bar\imath y| : \psi_i(y) = x \}\\
-& = r+1+K^\psi(x).
-\end{split}\end{equation}
-
-We can now relate the complexity of $f$ under the two enumerations as follows:
-\begin{align*}
-K^\phi(f) &= \min\left\{ K^\phi(j) : \phi_j = f \right\}& \\
-       &= \min\left\{ K^\phi(1^r0i) : \phi_{1^r0i} = f \right\}& \\
-       &\ge  \min\left\{ K^\phi(i) : \phi_{1^r0i} = f \right\}-c'&\text{by~\eqref{eq:prelim1}}\\
-       &\geq \min\left\{ K^\psi(i)  : \phi_{1^r0i} = f \right\}+r+1-c'& \text{by \eqref{eq:prelim2}}\\
-       &= \min\left\{ K^\psi(i) : \psi_{i} = f \right\} + r+1 -c'& \\
-       &= K^\psi(f) + r + 1 - c'.
-\end{align*}
-The proof is completed by choosing $r=c+c'-1$.
-\end{proof}
-
-\begin{definition}
-  An enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive
-  functions is called \emph{compact} if for all $i,y$ there is a $j$
-  such that $|j|\le|i y|$ and $\phi_i(y z)=\phi_j(z)$ for all $z$.
-\end{definition}
-
-\begin{lemma}
-There is a compact acceptable numbering.
-\end{lemma}
-\begin{proof}
-  Let $\phi_1,\phi_2,\ldots$ be any acceptable numbering.
-Let $\psi_1,\psi_2,\ldots$ be a new acceptable numbering defined by $\psi_j(x)=\phi_i(y x)$ if $j$ is a pair satisfying $j=\bar\imath y$, and $\psi_j(x)=\infty$ otherwise.
-We will show that $\psi$ is both $0$-compact and acceptable.
-
-To show compactness, pick any index $i$ and binary string $y$.
-First assume $i$ is a valid pair $i=\bar a b$.
-Then  $\psi_i(y x) = \psi_{\bar a b}(y x)=\phi_a(b y x)$, but also $\psi_{\bar a b y}(x)=\phi_a(b y x)$, so $\psi_i(y x) = \psi_j(x)$ for $j=\bar a b y$; so $|j|=|iy|+0$.
-On the other hand, if $i$ is not a valid pair, we can simply use $j=i$ so that $\psi_i(y x)=\psi_j(x)=\infty$ for all $x$.
-
-To show that $\psi_1,\psi_2,\ldots$ is acceptable, we define total
-computable functions $f$ and $g$ such that $\psi_i(x)=\phi_{f(i)}(x)$
-and $\phi_j(x)=\psi_{g(j)}(x)$. The easy direction is
-$g(j)=\bar\jmath\epsilon$. The definition of the mapping in the other
-direction $f(i)$ depends on whether $i$ is a valid pair. If it is not,
-then $f(i)=h$ for some $h$ such that $\phi_h(x)=\infty$ for all $x$.
-On the other hand if $i=\bar\jmath y$, then $f(i)$ is the index of a partial recursive function $\phi_{f(i)}(x)$ defined by a Turing machine that simulates the machine for $\phi_j$ but with $yx$ on the input tape.
-\end{proof}
-
-\begin{lemma}
-Compact numberings are sort of faithful.
-\end{lemma}
-\begin{proof}
-First, let $v$ be the index of the partial recursive function given by $\phi_v(\bar k p x)=\phi_{\phi_k(p)}(x)$.
-Let $\bar k p$ be a shortest program for $i$, that is $|\bar k p| = K(i)$ and $\phi_k(p)=i$.
-Then $\phi_i(x)=\phi_v(\bar k \bar p x)$.
-We now use compactness to find an index $j$ with $\phi_v(\bar k \bar p x)=\phi_j(x)$ and length
-\[|j|\le |v\bar k\bar p|\le|v|+|\bar k p|+2\log|\bar k p|=K(i)+|v|+2\log K(i),\]
-as required.
-\end{proof}
-}
-
 \end{document}

### 2015-01-18 Facticity/Facticity/paper.tex
...

@@
 \newcommand{\br}[1]{\overline{#1}}
-\newcommand{\s}{\tn{soph}}
+\newcommand{\s}{S}
 \newcommand{\num}[1]{#1_1, #1_2, \ldots}
@@
 \begin{abstract}
-Kolmogorov complexity is a sound measure of the amount of information in a binary string, but it does not distinguish between the information that expresses the structural properties of the data, and noise.
-Kolmogorov's definition of the \emph{structure function} in his lecture Tallinn in 1973 was the first attempt to separate meaningful information from noise, but many alternative definitions of the \emph{sophistication}, or the amount of meaningful information in the data, have subsequently been proposed.
-This indicates that the idea is considered worth pursuing by multiple distinguished scholars, but at the same time it shows that the issue has never been completely satisfactorily resolved.
-
-In this paper we describe three fundamental problems that arise with all existing proposals, showing each of them to be \emph{unsound}.
-Based on these results we we put forward the view that the problem is fundamental: while many nontrivial and useful structural properties of binary strings can be detected without a priori assumptions, it may be impossible to objectively quantify its sophistication.
+Kolmogorov complexity is a sound measure of the amount of information in data, but it does not distinguish structural information from random noise.
+Kolmogorov's definition of the \emph{structure function} was the first attempt to capture this separation.
+Since then many variations of this idea have been proposed, for which we use \emph{sophistication} as an umbrella term.
+The multitude of treatments indicates that the idea is considered worth pursuing by multiple distinguished scholars, but at the same time it shows that the issue has never been satisfactorily resolved.
+We describe three fundamental problems with all existing proposals, showing each of them to be \emph{unsound}.
+Based on these results we we put forward the view that the problem is fundamental: while many nontrivial and useful structural properties of binary strings can be detected without a priori assumptions, it may be impossible to objectively quantify the sophistication.
 \end{abstract}
@@
 
-Kolmogorov complexity breaks away from the tradition in statistics and information theory by allowing the analysis of \emph{individual} binary sequences rather than studying probability distributions.
+Kolmogorov complexity breaks away from the tradition in statistics and information theory by allowing the analysis of \emph{individual} binary sequences rather than studying large samples from probability distributions in an assumed class.
 This is important because for many binary strings, such as a binary encoding of a novel, or a holiday snapshot, or the fossil record, it is impossible to determine what distribution generates such objects.
-There is, however, a nagging discrepancy between Kolmogorov complexity and what most people whould intuitively call ``complex''.
+Kolmogorov complexity is often seen as the first step to a statistics without assumptions.
+
+There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would call ``complex''.
 For example, a sequence of a million coin flips is exceedingly unlikely to have a Kolmogorov complexity of much less than one million bits, even though there is nothing complex about flipping a lot of coins---it's just hard work.
 For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data.
-This tradition started with a lecture by Kolmogorov in Talinn, Estonia in 1973, but many alternative definitions of meaningful information, or \emph{sophistication}, were proposed, see the related work section below.
+This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication},\footnotemark were proposed, see the related work section below.
 In this paper, we investigate three problems with these definitions, and show that they are unsatisfactory.
-We believe the problem to be fundamental, and provide several intuitive arguments why these attempts to define sophistication must fail.
+We believe the problem is fundamental, and provide several intuitive arguments why these attempts to define sophistication must fail.
+
+\footnotetext{Most proposals coin their own name along with a specific treatment.
+For the sake of clarity, we have chosen to use \emph{sophistication} as an umbrella term to refer to all ideas that fit this mold.}
 
 The Kolmogorov complexity of a binary string $x$ is, roughly, the length of the shortest computer program that will output $x$.
-This length depends on the used programming language.
-However, the Invariance Theorem states that the choice between any two Turing complete languages affects the Kolmogorov complexity by at most a constant term: if $K^L(x)$ denotes Kolmogorov complexity of $x$ with respect to language $L$, then $|K^L(x)-K^M(x)|<c$ for some constant $c$ that depends on $L$ and $M$, but \emph{not on $x$}.
+This length depends on which programming language is used, but only by a constant, independent of $x$.
 This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings.
-The invariance theorem allows us to fix a single reference programming language and omit it from the superscript.
+We can see it as \emph{a property of the data}.
 
@@
 \begin{enumerate}
-\item Being an information measure, the sophistication should count the number of bits required for an effective description of the structural properties of a binary string.
-\item An analogue of the Invariance Theorem should hold for sophistication: at the very least, there should be strict limits on the degree to which a change in the used programming language can affect the sophistication.
-\item There should be no constant $c$ such that $S(x)\le c$ for every input sequence $x$.
+\item Since it measures information, the sophistication should count the number of bits required for an effective description of the structural properties of a binary string.
+\item An analogue of the Invariance Theorem should hold for sophistication: at the very least, there should be strict limits on the degree to which a change in the programming language can affect the sophistication.
+\item There should be no constant $c$ such that $S(x)\le c$ for every input $x$.
 If sophistication is bounded, then knowing its value under one programming language provides no constraints on its value under another language, even for inputs of very high Kolmogorov complexity.
-Moreover, a bounded sophistication would be at odds with the intuition that there is no limit to the amount structure that a binary string can exhibit.
-\item Similarly, there should be no constant $c$ such that the $K(x)-S(x)\le c$ for all $x$, because then sophistication would be equivalent to Kolmogorov complexity, and the sophistication would incorporate the amount of noise in the data just like Kolmogorov complexity does.
+Moreover, a bounded sophistication would be at odds with the intuition that there is no limit to the amount of structure that a binary string can exhibit.
+\item Similarly, there should be no constant $c$ such that $K(x)-S(x)\le c$ for all $x$, because then sophistication would be equivalent to Kolmogorov complexity. 
 \end{enumerate}
 
-For most proposed definitions of sophistication, we can prove that they fail one or more of the attributes above.
-Vit\'anyi's definition \cite{TODO} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we can show that only extremely deep strings (i.e., strings that require an enormous amount of processing to construct) yield a large sophistication, while intuitively it seems there must be objects that contain a lot of structure all of which can be described without much calculation.
+For most proposed definitions of sophistication, we can prove that they fail one or more of the conditions above.
+Vit\'anyi's definition \cite{vitanyi2004meaningful} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we can show that only extremely deep strings (i.e., strings that require an enormous amount of processing to construct) yield a large sophistication, while intuitively it seems there must be objects that contain a lot of structure all of which can be described without much calculation.
 Moreover in our discussion section we provide arguments that suggest that it should be possible to generalise our negative results to show that Vit\'anyi's sophistication is not invariant.
@@
 A valid definition of sophistication must contend with the following issues.
-First, while all definitions are based on a two-part encoding of the data sequence, some researchers use Kolmogorov complexity as a \emph{building block} of their proposed code, while others look for the two parts of the code \emph{inside} the Kolmgorov complexity: when the definition of Kolmogorov complexity is opened up, it turns out to consist of two parts itself.
-In Section~\ref{sec:nickname}, we show how this second approach sets us up to suffer from to the so-called \emph{nickname problem}: the sophistication becomes highly dependent on the chosen programming language.
+First, while all definitions are based on a two-part encoding of the data sequence, some researchers use Kolmogorov complexity as a \emph{building block} of their proposed code, while others look for the two parts of the code \emph{inside} the Kolmogorov complexity: when the definition of Kolmogorov complexity is opened up, it turns out to consist of two parts itself.
+In Section~\ref{section:indices}, we show how this second approach sets us up to suffer from to the so-called \emph{nickname problem}: the sophistication becomes highly dependent on the chosen programming language.
 This immediately breaks invariance, unless additional requirements are made on the language.
-We also show that there do exist languages (efficient enumerations of the partial recursive functions) that address this issue.
+We also show that there do exist languages (efficient enumerations of the partial recursive functions) that avoid this issue.
 
-The second issue is that of \emph{overfitting}, which we consider in Section~\ref{sec:overfitting}.
+The second issue is that of \emph{overfitting}, which we consider in Section~\ref{section:overfitting}.
 Overfitting is a common problem in the statistical literature, that refers to the tendency of some model selection procedures to select a very complex model that provides a very good fit to the observed data, but does not generalise well to unseen data.
 For example, in a polynomial regression setting, a higher order polynomial can always provide better fit to a data set.
-Therefore, even if the data are in fact noisy samples from a parabola, naive model selection methods may select a much higher order polynomial that provides a worse description of the trend in the data.
+Therefore, even if the data are in fact noisy samples from a parabola, naive model selection methods may select a much higher order polynomial that provides a worse description of the trend in the data, encoding all noise in the model parameters.
 The same can occur for the sophistication.
 We somehow need to make sure that the two-part representation that determines the sophistication does not store noise in the ``structure'' part of the code.
-In statistics, this is typically addressed by penalising complex models, e.g.
-using the BIC criterion \cite{TODO}.
-However, such penalisations tend to break the delicate balance between structural information and noise, and lead to the opposite problem: underfitting.
+In statistics, this is typically addressed by penalising complex models.
+However, in the sophistication setting, such penalties tend to break the delicate balance between structural information and noise, and lead to the opposite problem: underfitting.
 
-Underfitting (Section~\ref{sec:underfitting}) is the opposite problem, where the selected model is simple, but fails to represent part of the structure in the data.
-It is an especially serious problem for sophistication because the considered models for sophistication are so powerful.
+Underfitting (Section~\ref{section:underfitting}) is the opposite problem, where the selected model is simple, but fails to represent part of the structure in the data.
+It is an especially serious problem for sophistication because the models for used in sophistication are so powerful.
 In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language.
-Such interpreters can generate \emph{any} data sequence with only a short input, and if complex models are penalized to avoid overfitting, then the problem becomes to make sure that interpreters are not \emph{always} preferred for complex data.
-This problem is widely recognised, and is sometimes referred to as the ``collapse'' of the structure function \cite{}.
-The usual workaround is to restrict the set of considered models such that interpreters are no longer allowed.
-For example, the model class may be restricted to the set of all total recursive functions (only programs that never hang) rather than all partial recursive functions (all programs).
-While this patches the ``collapse'' problem to an extent, it is questionable whether it adequately solves the problem of underfitting in general.
+Such interpreters, or \emph{universal models}, can be described in a constant amount of bits, and can generate the data with from a program the length of the Kolmogorov complexity, essentially encoding all information as noise.
+If complex models are penalized to avoid overfitting, then the problem becomes to make sure that universal models are not \emph{always} preferred for complex data.
+The usual workaround is to restrict the set of allowed models to exclude the universal ones.
+For example, the model class may be restricted to the set of all total recursive functions (only programs that halt for all inputs) rather than all partial recursive functions (all programs).
+While this patches the problem of universal models to an extent, it is questionable whether it adequately solves the problem of underfitting in general.
 
-Finally, in the discussion in Section~\ref{sec:discussion} we argue that while two-part coding can yield useful insights into the structure of the data and allows us to identify some models as unlikely to capture all structure in the data, it is probably not possible to uniquely separate structure from noise and identify a \emph{single} model as ``best'': in general many models, of vastly different complexities, may be reasonable models of the data generating mechanism.
-Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that sometimes the data allows for multiple very different interpretations, and that there is no such thing as sophistication.
+Finally, in the discussion in Section~\ref{section:conclusion} we argue that while two-part coding can yield useful insights into the structure of the data and allows us to identify some models as unlikely to capture all structure in the data, it is probably not possible to uniquely separate structure from noise and identify a \emph{single} model as ``best'': in general many models, of vastly different complexities, may be reasonable models of the data generating mechanism.
+Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that sometimes the data allows for multiple equivalent interpretations, and that there is no such thing as sophistication.
 
 \subsection{Related work}
-The idea of separating structure from noise originated with Kolmogorov's lecture in Tallinn, Estonia in 1973 about the \emph{structure function},.
-This work was never published, but summarized by Cover \cite{cover1985kolmogorov}.
+The idea of separating structure from noise originated with Kolmogorov's lecture in Tallinn, Estonia in 1973 introducing the \emph{structure function} \cite{cover1985kolmogorov}.
 In this theory, the two-part description of the data consists of a program that generates a finite set, and an index of the original object in the set.
 The length of the program then measures the structure in the data.
-This approach was discussed more extensively, and generalised, by Vereshchagin, G\'acs and Vit\'anyi \cite{vereshchagin2004kolmogorov,gacs2001algorithmic}).
+This approach was discussed more extensively, and generalised, by Vereshchagin, G\'acs and Vit\'anyi \cite{vereshchagin2004kolmogorov,gacs2001algorithmic}.
 
@@
 
-In 1996 Gell-Mann and Lloyd introduced `Effective complexity' \cite{gellmann1996information}.
+In 1996 Gell-Mann and Lloyd introduced \emph{effective complexity} \cite{gellmann1996information}.
 Working from the perspective of statistical mechanics, they arrived (largely independently, it seems) at a notion that fits the mold of sophistication.
@@
 As we shall see, this version behaves very differently from regular sophistication, and is closer to computational depth.
+Vereshchagin \cite{vereshchagin2013algorithmic} proposes an alternative solution to the instability issue.
 
 Adriaans' \emph{Facticity}, introduced in 2012\cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions.
-Adriaans also identifies the nickname problem discussed in Section~\ref{sec:nickname}.
+Adriaans also identifies the nickname problem discussed in Section~\ref{section:indices}.
 
 \subsection{Preliminaries, definitions and desiderata}
-
-We will generalizes the main proposals for sophistication into a single, consistent notation.
-The unifying principle is this: we start with a class of models, we consider the space of two-part descriptions using these models, we select the set of representations with the shortest code-lengths and we select, by some method, the best model from this set.
 
@@
 
-As noted in the introduction, the Kolmogorov complexity is invariant to the choice of description language. Formally, our description language creates a numbering of the partial computable functions.
-
 \begin{definition}[Numberings]
-A numbering is an enumeration of $\C$, denoted as $\phi_1, \phi_2, \ldots$ or by a function $g: \N \to \C$.
-We fix one canonical numbering $g^\circ$, which works as an effective description, that is for any $i \in \N, y\in \B$, we can effectively compute $(g^\circ(i))(y)$.
-We call a numbering $g$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ as that for all $i$ $g^\circ(a(i)) = h(i)$ and  $g^\circ(i) = h(b(i))$.
+A numbering is an enumeration of of the partial computable functions, denoted as $\phi_1, \phi_2, \ldots$.
+We fix one canonical, which we know to be an effective description, that is for any $i \in \N, y\in \B$, we can effectively compute $\phi_i(y)$, where the computation diverges if $\phi_i(y) = \infty$.
+We call a numbering $\num{\psi}$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ such that for all $i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{b(i)}$.
+We use the greek letters without a subscript (eg.
+$\phi$ and $\psi$) to refer to the numbering as a whole.
 \end{definition}
-
-When we say that Kolmogorov complexity is invariant, we mean that changing the canonical acceptable numbering changes the Kolmogorov complexity by at most a constant.
 
@@
 
-\begin{definition}[Complexity]
+\begin{definition}[Complexity]\belowdisplayskip=-12pt
 Let $\phi_1,\phi_2,\ldots$ be the canonical acceptable enumeration of $\C$.
-Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ corresponfs to the set of all computable prefix functions.
-Now define Kolmogorov complexity by
+Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ corresponds to the set of all computable prefix functions.
+We define the plain and prefix Kolmogorov complexity as
 \begin{align*}
@@
 K(x)&=\min\{|\bar\imath y|:\phi_i(y)=x, i\in\Pi\}.\\
-\end{align*}
+\end{align*}\label{definition:complexity}\vspace{-18pt}
 \end{definition}
 This definition of $K$ is uncommon.
-It is equivalent to the standard definition as used in \cite{li1993introduction}, up to a constant, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).
+It is equivalent to the standard definition as used in \cite{li1993introduction}, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).
 
-We now have the basic tools with which we set up our generic sophistication function. We start with the class of allowed models:
+We will generalize the main proposals for sophistication into a single notation.
+We start with a class of models, we consider all two-part descriptions of the data using these models, we select the set of representations with the shortest code-lengths and we select the smallest model from this set.
+The size of the smallest model represents the sophistication of the data.
 
@@
     surjective function mapping the binary sequences of length
-    $\lceil\log S\rceil$ onto $S$. In other words, this is the model class of finite sets, with uniform distributions.
+    $\lceil\log S\rceil$ onto $S$. In other words, this is the model class of finite sets.
   \end{itemize}
 \end{definition}
-
-We note that the model classes $\T$ and $\P$ are unusual in that they are not enumerable.
-
+We note that classes $\T$ and $\P$ are unusual in that they are not enumerable.
 \begin{definition}[Representations and code-lengths]
 A pair $(m,y)$ is a \emph{representation} for $x$ if $m$ is a model and $m(y)=x$.
-
 Define the following code-length functions for the representations:
@@
 \item $L(m)$ is the length of a self-delimiting description of
-  the model $m$, i.e. it is the length function of a prefix code for models. Each variant of sophistication must supply such a function.
+  the model $m$, i.e. it is the length function of a prefix code for models. 
 \item $L^m(x)=L(m)+\min\{|y|:m(y)=x\}$ measures the number of bits required to describe both the model $m$ and the data $x$ using the model.
@@
 
-\begin{definition}[Sophistication]
+\begin{definition}[Sophistication]\belowdisplayskip=-12pt
   The model complexity, meaningful information, effective complexity, facticity or \emph{sophistication},  of a string $x$ given a model class $\M$ and threshold function $\tau$ is
@@
 \end{definition}
-
-Finally, we must formalize what we mean when we say that we desire sophistication to be invariant.
-For the Kolmogorov complexity, we know that a change in acceptable numbering causes no more than a constant change.
-It may be that a less stringent invariance (for instance a $\log$ term) can still yield a worthwhile definition of sophistication.
-To allow such distinctions, we generalize the notion of invariance:
-
+Finally, we must formalize what we mean when we say that we expect sophistication to be invariant.
+It may be that a less strict invariance than we have for the Kolmogorov complexity, still yields a useful sophistication.
 \begin{definition}
@@
 \end{definition}
-
-We say that a formalization of sophistication $s_\phi$ built on acceptable numbering $\phi$ is $\gamma$-invariant to the choice of numbering is $s_\phi$ and $s_\psi$ are $\gamma$-equal for any two acceptable numberings $\phi$ and $\psi$.
-Thus, by this definition, Kolmogorov complexity is $1$-invariant.
+We say that a formalization of sophistication $\s_\phi$ built on acceptable numbering $\phi$ is $\gamma$-invariant to the choice of numbering is $s_\phi$ and $s_\psi$ are $\gamma$-equal for any two acceptable numberings $\phi$ and $\psi$.
+\begin{lemma}
+If a function $\S_\phi(x)$ is bounded for a given numbering, it is either not $\gamma$-invariant or it is  $\gamma$-equal to a bounded function.
+\end{lemma}
+The proof follows directly from the definitions.
+Most lemmas in the rest of the paper will simply prove that a given variant is bounded for a specific numbering, implicitly using this lemma to show that the variant is unsound.
 
 \section{Inefficient indices}
+\label{section:indices}
+We start from a simple intuition: the definition of Kolmogorov complexity already uses a description of the data in two parts; a computable function and an input.
+Can we assume that, when such a description is as short as the Kolmogorov complexity, it automatically creates a good separation into the structural part of the data and the noise?
+As we will show, the representation of function plays a fundamental part in answering this question.
 
-We start from a simple intuition: the definition of Kolmogorov complexity already uses a description of the data in two parts; a computable function and and input.
-Can we assume that when such a description is as short as the Kolmogorov complexity, it automatically creates a good separation into the structural part of the data and the noise?
-As we will show, the representation of the structural part plays a fundamental part in answering this question.
-If it is in any way inefficient, the resulting sophistication function does not behave as expected.
-
-The issues can be very subtle and require a careful attention that is not always afforded in the literature. We will first illustrate what happens if any acceptable numbering is allowed:
+The issues can be very subtle and require a careful attention that is rarely given in the literature. We will first illustrate what happens if any acceptable numbering is allowed:
 
@@
 \[
-\psi_j(x) = 
-\begin{cases}
+\psi_j(x) = \begin{cases}
 	\phi_i(x) \;\text{if}\; j = s_{2i}\\
 	\infty\;\text{otherwise}
-\end{cases}
-\]
-and 
-\[
+\end{cases}\,\,\,
 \xi_j(x) = 
@@
 \end{proof}
-
 In this proof, we deliberately used ridiculously inefficient index functions.
@@
 
-Interestingly, using an inefficient index is no problem in the definition of Kolmogorov complexity, since we can select as a model a universal Turing machine with an efficient index at only a constant penalty (by the Invariance Theorem \cite{li1993introduction}).
-But this robustness property does \emph{not} carry over to definitions of sophistication.
+Interestingly, using an inefficient index is no problem in the definition of Kolmogorov complexity, since we can select as a model a universal Turing machine with an efficient index at only a constant penalty.
+But this robustness does \emph{not} carry over to definitions of sophistication.
 If we are interested in separating the model and its input, we must represent the model efficiently.
 
-Another perspective on this problem is that the length of a model in an ordinary description does not represent its complexity, much like the length of a compressible string does not reflect its information content.
+Simply put, the length of a model in an ordinary description does not represent its complexity, much like the length of a compressible string does not reflect its information content.
 In order to obtain a more robust measure of model complexity, we describe a natural extension of Kolmogorov complexity to functions, and show that this retains invariance.
 
-\begin{definition}
+\begin{definition}\belowdisplayskip=-12pt
   The complexity of a partial recursive function is defined by
   \begin{align*}
-    C_\phi(f) &:= \min\{C(i):\phi_i=f\}\\
-    K_\phi(f) &:= \min\{K(i):\phi_i=f\}.
+    C_\phi(f) &:= \min\{C(i):\phi_i=f\}\;\tn{,}\\
+    K_\phi(f) &:= \min\{K(i):\phi_i=f\}\p
   \end{align*} \label{definition:model-complexity}
@@
 \begin{lemma}[Invariance]
-Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be two acceptable numberings.
-There exists a constant $c$ such that $\left| K^\phi(f) - K^\psi(f)\right | \leq c$ for all $f$.
+Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be any two acceptable numberings.
+There exists a constant $c$ such that $\left| K_\phi(f) - K_\psi(f)\right | \leq c$ for all $f$.
 \label{lemma:invariance}
@@
 \begin{proof}
-Let $g(i)$ be the function which which translates from $\psi$ to $\phi$.
+Let $g(i)$ be the function such that $\psi_i$ to $\phi_{g(i)}$.
 \begin{align*}
-K^\phi(f) &= \min\left\{ K^\phi(i) : \phi_i= f\right\} \\
-&\geq \min\left\{ K^\psi(i) : \phi_i= f\right\} - c\\
-&= \min\left\{ K^\psi(i) : \psi_{g(i)}= f\right\} - c\\
-&\geq \min\left\{ K^\psi(i) : \psi_i= f\right\} - c\\
-&= K^\psi(f).
+K_\phi(f) &= \min\left\{ K_\phi(i) : \phi_i= f\right\} 
+\geq \min\left\{ K_\psi(i) : \phi_i= f\right\} - c\\
+&= \min\left\{ K_\psi(i) : \psi_{g(i)}= f\right\} - c
+\geq \min\left\{ K_\psi(i) : \psi_i= f\right\} - c = K_\psi(f).
 \end{align*}
-We can reverse $\phi$ and $\psi$ to achieve the same inequality the other way around, completing the proof.
+The opposite inequality can be achieved by reversing $\phi$ and $\psi$. 
 \end{proof}
@@
 \begin{enumerate}
-  \item One can stick with the approach of using the model used by the representation that witnesses the Kolmogorov complexity, but repair it by imposing the additional constraint that an efficient numbering is used.
-Specifically, that the length of the indices satisfies $|i| = C(\phi_i)$, and that the numbering is still acceptable.
+  \item We maintain that the representation witnessing the Kolmogorov complexity determines the sophistication, but impose the additional constraint that an efficient, acceptable numbering is used.
+Specifically, that the length of the indices satisfies $|i| = C(\phi_i)$.
 We construct such an index below.
 This approach is taken by Adriaans \cite{adriaans2012facticity}.
-In this case, the total code length is simply $C(x)$, and the sophistication is the length of the model index within this code.
-This approach is the most direct formalization of the intuition that the two part code used within the definition of Kolmogorov complexity separates structure from noise.
-  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use Kolmogorov complexity as a black box to measure the complexity of the \emph{model}, which is then the first part of a two-part code describing the data.
-With this approach, the total code length is $K(f)+|y|$, where $y$ is an input to the model $f$ such that $f(y)=x$.
-The difference is that in this approach, we can construct everything on the basis of a single acceptable numbering, without using additional contraints.
-It is also more natural in this setting to limit the model class of allowed functions.
-This approach is used for Kolmogorov's structure function \cite{cover1985kolmogorov}, Algorithmic Statistics \cite{gacs2001algorithmic}, Vi\'anyi's sophistication \cite{vitanyi2004meaningful} and Effective complexity \cite{gellmann1996information}.
+  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use Kolmogorov complexity to measure the complexity of the \emph{model} (as in Definition~\ref{definition:model-complexity}), which is then the first part of a two-part code describing the data.
+With this approach, the total code length is $K(\phi_i)+|y|$, where $y$ is an input to the model $\phi_i$ such that $\phi_i(y)=x$.
+The difference is that in this approach, we can construct everything on the basis of a single acceptable numbering, without additional constraints.
+This approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information}.
 \end{enumerate}
-
-To show that the two approaches are equivalent with the correct restriction on the index, we must show that there exists an acceptable numbering for which the length of the index of a model is equal to its complexity.
+To show that the two perspective are equivalent, we must show that there exists an acceptable numbering which is efficient enough.
 
@@
 \end{definition}
-This definition is a refinement of the one used in \cite{adriaans2012facticity}.
-That publication defines an \emph{index function}, rather than a numbering, and notes that it is not computable.
+This definition is a refinement of the one used in \cite{adriaans2012facticity}. \footnotemark
+
+\footnotetext{That publication defines a faithful \emph{index function}, rather than a numbering, and notes that it is not computable.
 Since a numbering is all that is needed, we only need to show acceptability.
-The incomputability of the index function is no issue:
+The incomputability of the index function is no issue.}
 \begin{lemma}
@@
 \[\begin{split}
-C(f)&=\min\{C(i):\phi_i=f\}\\
-&=\min\{\min\{|\bar a b|:\phi_a(b)=i\}:\phi_i=f\}\\
-&=\min\{|\bar a b|:\phi_{\phi_a(b)}=f\}\\
-&=\min\{|\bar a b|:\psi_{\bar a b}=f\}.
+C(f)&=\min\{C(i):\phi_i=f\} =\min\{\min\{|\bar a b|:\phi_a(b)=i\}:\phi_i=f\} \\
+& =\min\{|\bar a b|:\phi_{\phi_a(b)}=f\}
+ =\min\{|\bar a b|:\psi_{\bar a b}=f\}.
 \end{split}\]
@@
 denote this index as a function of the program; further define
-$h(q)=i_\tn{div}$ if $q$ cannot be expressed as $\bar\imath p$. By
+$h(q)=\tn{div}$ if $q$ cannot be expressed as $\bar\imath p$. By
 construction $h$ is total and computable. To check that the mapping
@@
 \end{proof}
-
-Unfortunately, even with a faithful index, we can fail, if the prefix-function used to separate the model from its input is inefficient:
-
+Unfortunately, even with a faithful index, we can fail if the prefix-function used to separate the model from its input is inefficient:
 \begin{lemma}
@@
 
-If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then $\s^\C$ is bounded for all numberings $\phi$.\label{lemma:prefix-inefficiency}
+If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then $\s^\C$ is bounded for all numberings.\label{lemma:prefix-inefficiency}
 \end{lemma}
@@
 This result applies to Adriaans' facticity\cite{adriaans2012facticity}.
-In order to circumvent the problem, the UTM should should encode the model $f$ in $K(f)$ bits, relying on $K$'s own prefix encoding to delimit the model from its input.
+In order to circumvent the problem, the representation should should encode the model $f$ in $K(f)$ bits, relying on $K$'s own prefix encoding to delimit the model from its input.
 
@@
 \begin{conjecture}
-For any partial computable prefix codelength function $L$, the function
-\[
-\min\{L(i)-K(i):i\ge i_0\}
-\]
+For any partial computable prefix codelength function $L$, the function $\min\{L(i)-K(i):i\ge i_0\}$
 is unbounded in $i_0$.
@@
 
-In the rest of this paper, we will avoid these issues and take the second perspective: we will consider the representations as two-part descriptions $(f, y)$ with codelength $K(f) + |y|$, such that $f(y)=x$.
+In the rest of this paper, we will avoid these issues and take the second perspective: we will consider the representations as two-part descriptions $(i, y)$ with codelength $K(\phi_i) + |y|$, such that $\phi_i(y)=x$.
 
 \section{Overfitting: the singletons}
-
+\label{section:overfitting}
 We now come to the main issue plaguing sophistication.
 In the naive setting, where we allow all prefix-free functions as models, and use the default threshold function, the candidate set always contains two representations that clearly do not provide a good separation into structure and content.
-The first is a universal model: a universal Turing machine which simply runs its shortest program for $x$.
+The first is a \emph{universal model}: a universal Turing machine which simply runs its shortest program for $x$.
 The second is a \emph{singleton model} a model which returns $x$ for any input.
-In the first representation all information about $x$ is stored in the second part of the two-part description.
-In the second representation all information is stored in the model.
-The model complexity suggested by these representations is either constant, or equal to the Kolmogorov complexity.
 
@@
 As we will show, in many settings, we can ensure that they are the \emph{only} models in the candidate set.
-The principle we use is that the threshold on the candidate set is determined by a constant, and a change in numbering affects model size by a constant.
-Thus, We can choose our numbering  so that all but a certain class of models take a constant penalty, pushing them out of the candidate set.
-Since we use this trick repeatedly we have generalized it into an (admittedly complex) lemma.
+The following lemma generalizes the principle.
 
@@
 First suppose $f\in\M'$. Then
-\[\begin{split}
-K^\phi(f)&=\min\{K^\phi(i):\phi_i=f\}\\
-&=\min\{|\bar\jmath q|:\phi_j(q)=i, \phi_i=f\}\\
-&=\min\{|\bar\jmath q|:\phi_{\phi_j(q)}=f\}\\
-&\le\min\{|\bar0 q|:\phi_{\phi_0(q)}=f\}~\text{(using $j=0$)}\\
-&=\min\{|\bar0 q|:\phi_{1^r0 D(q)}=f\}\\
-&=|\bar 0|+\min\{|q|:\psi_{D(q)}=f\}\\
-&\le K^\psi(f)+c+|\bar 0|,
-\end{split}
-\]
+\begin{align*}
+K^\phi(f) &=\min\{|\bar\jmath q|:\phi_{\phi_j(q)}=f\}
+\leq\min\{|\bar0 q|:\phi_{\phi_0(q)}=f\} \\ 
+&=\min\{|\bar0 q|:\phi_{1^r0 D(q)}=f\}  
+=|\bar 0|+\min\{|q|:\psi_{D(q)}=f\}
+\leq K^\psi(f)+c+|\bar 0|,
+\end{align*}
 where the last inequality uses the first assumption.
@@
 Now assume that the best model $f$ for $x$ is not in $\M'$. 
-Let $i$ be the index of $f$ with the shortest description, i.e. it achieves the minimum in
-\[
-K^\phi(f)=\min\{K^\phi(i):\phi_i=f\}.
-\]
+Let $i$ be the index of $f$ with the shortest description, i.e. it achieves the minimum in $K^\phi(f)=\min\{K^\phi(i):\phi_i=f\}$.
 There are two possibilities. Either $i=0$, in which case we have $K^\phi(f)\ge r$ because $\phi_0$ cannot output zero and all other $\phi$-programs are at least $r$ bits long.
@@
 Otherwise, we can bound
-\[\begin{split}
+\begin{align*}
 K^\phi(f)&=K^\phi(i)=K^\phi(1^r0j)\\
-&\ge K^\phi(j)-c'\\
-&\ge K^\psi(j)+r+1-c'\\
-&\ge\min\{K^\psi(j):\phi_j=f\}+r+1-c'\\
+&\ge K^\phi(j)-c' \ge K^\psi(j)+r+1-c'\\
 &=K^\psi(f)+r+1-c'.
-\end{split}
-\]
+\end{align*}
 Now choose $r=c''+\max\{K^\psi(\phi_0),c'-1\}$. Then substitution yields, for both cases, 
@@
 K^\phi(g)+C^g(x) &\ge K^\psi(g)+C^g(x)+c''\\
-&\ge K^\psi(f)+C^f(x)-c+c''\\
 &\ge K^\phi(f)+C^f(x) -2c+c'' -|\bar0|.
@@
 
-This is a technical and opaque lemma and proof, due to the double effect of the numbering, which occurs in both the definition of model complexity and the definition of data complexity.
+This is a technical lemma, due to the double effect of the numbering, which occurs in both the definition of model complexity and the definition of data complexity.
 The \emph{idea} behind the lemma, however, is simple.
@@
 
-\begin{enumerate}
-\item Let $f$ be any function in the range of $D$, and $x$ its output. Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\psi_i(q)=x\}=K(x)\le K(f)+c$.
-\item On the one hand $L^{\M',\psi}(x)\le K(f)+|\epsilon|\le K(x)+c$. On the other hand, $L^{\M,\psi}(x)$ is an effective description of $x$, so $K(x)$ is at most a constant larger.
+(1) Let $f$ be any function in the range of $D$, and $x$ its output.
+Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\psi_i(q)=x\}=K(x)\le K(f)+c$.
+(2) On the one hand $L^{\M',\psi}(x)\le K(f)+|\epsilon|\le K(x)+c$.
+On the other hand, $L^{\M,\psi}(x)$ is an effective description of $x$, so $K(x)$ is at most a constant larger.
 Together, these inequalities establish the second condition.
-\end{enumerate}
 
-Then by Lemma~\ref{lemma:thecoolone} there is an enumeration $\phi_1,\phi_2,\ldots$ such that $S^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
+Now, by Lemma~\ref{lemma:thecoolone} there is an enumeration $\phi_1,\phi_2,\ldots$ such that $S^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
 We observed that $|K(f)-K(x)|\le c$ for any $f\in\M'$, so $S^{\M',\psi}(x)\ge K(x)+c$.
@@
 Many variants in the literature are susceptible to this problem, including the structure function \cite{cover1985kolmogorov,gacs2001algorithmic}.
-\cite{vitanyi2004meaningful} and \cite{adriaans2012facticity} void this issue by using non-self-delimiting representations.
+\cite{vitanyi2004meaningful,adriaans2012facticity} avoid this issue by not using self-delimiting representations.
 Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically less efficient than representations placing some information in the input.
@@
 We are not aware of any approach to this effect, and there does not seem to be a simple way to determine the threshold based on the properties of the numbering.
-We are not aware of such considerations being mentioned anywhere in the literature.
 
 \section{Underfitting: the universal model}
-
+\label{section:underfitting}
 On the other end of the spectrum, there exist candidate representations which place all information in the `noise part', using a universal function as the model.
@@
 \begin{lemma}[underfitting]
-Let $\M$ be a model class containing model $u$, called a \emph{universal model} with the following property
-\[
-\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p
-\]  
-Then any $\s^\M$ with the default threshold function is either non-invariant or unbounded.
+Let $\M$ be a model class containing model $u$, called a \emph{universal model} with $\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p$.
+Then, there exist numberings such that $\s^\M$ is bounded.
 \end{lemma}
 \begin{proof}
-We will construct an acceptable numbering for which $\s^\M$ is bounded. This means that if $\s^\M$
- is either bounded for all acceptable numberings, or not invariant.
-
 Let $D$ be a prefix function as in Lemma~\ref{lemma:thecoolone} such that it returns the index of $u$ for the argument $\epsilon$ and $\infty$ for any other argument.
@@
 \end{proof}
-
-Most sophistication measures avoid the underfitting problem by limiting their model class, usually to either finite sets ($\F$) or total functions ($\T$ or $\K$).
+Most sophistication measures avoid the underfitting problem by limiting their model class, usually to either finite sets ($\F$) or total functions ($\T$ or $\P$).
 This eliminates the universal model, but as we shall see in the next section, this does not solve the problem for most data.
-The variants that are susceptible to the problem of underfitting are facticity \cite{adriaans2012facticity} and some versions of effective complexity \ref{gellmann1996information}.
+The variants that are susceptible to the this type of underfitting are facticity \cite{adriaans2012facticity} and some versions of effective complexity \cite{gellmann1996information}.
 The latter aims to bypass the problem by selecting carefully from the candidate set, but as we have shown, for some numberings the candidate set contains only the universal model.
@@
 In fact, a well-known result in algorithmic statistics shows that there exist strings for which no finite set is as good a model as the singleton set.
-We will show the consequences for sophistication:
+In this section we will focus on the variants $S^\T$ and $S^T_K$ \cite{vitanyi2004meaningful}, as those are the variants most resistant to proofs of non-invariance.
 
-\begin{theorem}
-There are infinitely many $x$ such that $S^\F_K(x) \geq K(x)/5$. \label{theorem:structure-function-is-not-bounded}
-\end{theorem}
+\begin{lemma}
+There are infinitely many $x$ such that $S^\F_K(x) \geq K(x)/5$. \label{lemma:structure-function-is-not-bounded}
+\end{lemma}
 \begin{proof}
@@
 
-We use this lemma to transport the result of Theorem~\ref{theorem:structure-function-is-not-bounded} to the model class of total functions:
+We use this lemma to transport the result of Theorem~\ref{lemma:structure-function-is-not-bounded} to the model class of total functions:
 
@@
   \item $S^\T_K(x) \geq K(x)/6$, {and}\label{eq:poezenvoer}
-  \item $S^\T_{\tau^\circ}(x) \geq K(x)/6$. \label{eq:hondevoer}
+  \item $S^\T(x) \geq K(x)/6$. \label{eq:hondevoer}
 \end{enumerate}
@@
 \begin{proof}
-We continue from the proof of Theorem~\ref{theorem:structure-function-is-not-bounded}.
+We continue from the proof of Theorem~\ref{lemma:structure-function-is-not-bounded}.
 Suppose towards contradiction that either $C^\T_K(x_i)$ or $C^\T_{\tau^\circ}(x_i)$ contains a candidate $f$ with $K(f)\le i/5$.
-Then by Lemma~\ref{lemma:total-to-sets}, there is a set $S$ with $\log|S|\le|d|\le\tau^\circ(x)\le K(x)$, so (a) in the proof of Theorem~\ref{theorem:structure-function-is-not-bounded} above cannot be the case, and $K(S)\le K(f)+K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) in the same proof for large enough $i$.
+Then by Lemma~\ref{lemma:total-to-sets}, there is a set $S$ with $\log|S|\le|d|\le\tau^\circ(x)\le K(x)$, so (a) in the proof of Theorem~\ref{lemma:structure-function-is-not-bounded} above cannot be the case, and $K(S)\le K(f)+K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) in the same proof for large enough $i$.
 Therefore such candidates $f$ cannot exist and all candidates must satisfy $K(f)>i/5>K(x_i)/6$.
@@
 This tells us that for the right strings, a sophistication so defined can truly grow to a significant part of the total description length.
-While this shows that sophistication, if defined with particular care, is not entirely vacuous, we can shed some light on exactly which strings end up with high sophistication, using the principle of depth.
+While this shows that sophistication, if defined with particular care, is not entirely vacuous, we can shed some light on exactly which strings end up with high sophistication, using the principle of \emph{depth}.
 We will use the following definition:
 
-\begin{definition}[depth]
+\begin{definition}[depth]\belowdisplayskip=-12pt
 Let $U$ be some universal Turing machine, so that $U(\bar\imath y) = \phi_i(y)$.
 Let $U^t$ be a simulation of this machine, which is allowed to run for at most $t$ steps, and returns $0$ if it has not yet finished at that point.
-Let $C^t_\M(x) = \min{|\bar\imath y| : U^t(\bar\imath y) = x, \phi_i \in \M}$.
+Let $C^t_\M(x) = \min\{|\bar\imath y| : U^t(\bar\imath y) = x, \phi_i \in \M\}$.
 
@@
 \end{definition}
-
-Thus, deep strings are those that can only be optimally compressed with a great investment of time.
+Deep strings are those that can only be optimally compressed with a great investment of time.
 The notion of depth was first proposed by Bennett \cite{bennett1988logical} and later refined to variants of this form \cite{antunes2006computational}.
-It is important to note that it is exceedingly unlikely to sample deep string from a non-deep distribution.
 We note that it is exceedingly unlikely that a deep string is sampled from a shallow distribution \cite{bloem2014safe,bennett1988logical}.
@@
 Unless we encounter data that would take longer than the lifetime of the universe (and then some) to produce, sophistication behaves exactly as it would without the restricted model class: a solution to a super-exponential problem would have no greater sophistication than a string sampled from a simple Bernoulli model.
-This result also suggests that the `non-stochastic' property of strings with high sophistication\cite{shen1983concept,vereshchagin2004kolmogorov} says more about depth and totality than it does about the lack of randomness in its information content.
+This result also suggests that the `non-stochastic' property of strings with high sophistication \cite{shen1983concept,vereshchagin2004kolmogorov} says more about depth and totality than it does about the lack of randomness in its information content.
 
@@
 \section{Conclusion}
-
+\label{section:conclusion}
 As we have seen, the idea behind sophistication has a rich history.
@@
 So why should we consider this intuition reasonable at all?
-The first assumption of sophistication is that high structure and high randomness are objectively `uninteresting'.
+The first assumption of sophistication is that high compressibility and pure randomness are objectively `uninteresting'.
 A more sophisticated signal, such as a television broadcast must contain, in some objective sense, more meaningful information.
@@
 
-To see whether this assumption is valid; imagine a prefix-free universal Turing machine, to which we feed random bits whenever it reads from its input tape.
-We continue doing so until it halts (or indefinitely if it does not).
-In other words, we are sampling from $m(x)$, the universal distribution.
-What can we say about the resulting data?
-If the universal Turing machine is constructed in the conventional manner, the initial bits will determine the index of a partial computable function $i$, and the rest will be the input to that machine $y$.
+To see whether this intuition is valid; imagine that we sample from the universal distribution by feeding random bits to a universal Turing machine.
+If this UTM is constructed in the conventional manner, the initial bits will determine the index of a partial computable function $i$, and the rest will be the input to that machine $y$.
 What's the characteristic model of this data?
@@
 Also, we cannot exclude the possibility that $\phi_i$ is a singleton model for $x$.
-Ultimately, choosing the specific model that was actually used in generating the data comes down to guessing where the model ended and the input began.
 
 This kind of perspective, with a probabilistic source for the data is strongly discouraged by the proponents of algorithmic statistics: they often argue that information theory should precede probability theory \cite{kolmogorov1983combinatorial}.
-Nevertheless, this sampling scenario undercuts the intuition behind sophistication: there is no reason to assume that a string with high sophistication was sampled from a complex model, so how can the complex model be intrinsic to the data, moreso than any other model that reaches the Kolmogorov complexity?
+Nevertheless, this sampling scenario undercuts the intuition behind sophistication: there is no reason to assume that a string with high sophistication was sampled from a complex model, so how can the complex model be intrinsic to the data, moreso than any other model that reaches the Kolmogorov complexity?\footnotemark
+
+\footnotetext{Note that computational depth and Kolmogorov complexity do not have this problem.
+The properties of a distribution allow us to put bounds on both the depth and the complexity we can expect from strings produced by it.}
 
@@
 A small change in construction, such as a different acceptable numbering, or a change in threshold function, can arbitrarily push models into or out of the candidate set.
-Whereas in Kolmogorov complexity constant changes have constant effects on the outcome, here constant changes can cause very large changes in the outcome.
+Whereas in Kolmogorov complexity, constant changes have constant effects on the outcome, here constant changes can cause very large changes in the outcome.
 While no authors have touched on the issue of invariance explicitly, it is strongly related to the issue of instability, discussed in \cite{antunes2013sophistication,vereshchagin2013algorithmic}.
@@
 The most common approach, restricting the models to be total, is problematic.
-We have shown that the behavior of the resulting function does not correspond with the intuition behind it.
-We can also ask what justification there is for this restriction?
-What makes the total functions better than any other model class?
-If it is an arbitrary choice, then the sophistication should be invariant to changing it to some other reasonable class.
-If it isn't, which model classes are permissable and why?
-
 As we have seen, the restriction to a model class $\M$ tends to assign high sophistication to those strings that can be compressed significantly better by models outside $\M$ than by models within it.
-For the class $\T$, this results in extremely deep strings getting high sophistication.
+For the class $\T$, this results in only extremely deep strings getting high sophistication.
 While this is an interesting property, which is worth investigating, the interpretation of an objective separation into model and noise is spurious at best.
@@
 This is dangerous: if such definitions cannot be proved wrong or right, they will stand as the ad-hoc official view, while the restrictions used to elude proofs of incorrectness make them very difficult to build on.
-This creates an artificial dead-end for what is at heart a useful and valuable topic of research.
+This creates an artificial dead-end for a very valuable area of research.
 
@@
 
-Of course the definition of Kolmogorov complexity still stands, and the space of two-part descriptions, ordered by strength of compression is still meaningful.
-If we give up on the ability to distinguish between representations that are close to the minimum, and consider them equivalent, we lose the ability to select a model, but we can still eliminate all those representation that are far from the Kolmogorov complexity.
-This is analogous to hypothesis testing in classical statistics.
-We can eliminate models but not confirm them.
-Even if we knew the Kolmogorov complexity of our data, there would still be equivalent representations we could not objectively  choose between.
-Given our choice of numbering we may hav an absolute favorite, but objectively speaking, we must accept all explanations that are `close enough' to the Kolmogorov complexity.
-
 Consider the following metaphor.
-We are given a bitstring which can be read as a high-resolution bitmap image of the painting \emph{Impression of a Sunrise}.
-The candidate set for this painting contains various models from very generic to very specific.
+We are given a bitstring which can be read as a bitmap image of the painting \emph{Impression of a Sunrise}.
+The candidate set for this string contains various models from very generic to very specific.
 The theory behind sophistication suggests that we can choose one of these as the objective, intrinsic model of the data.
 Different models say different things: the universal models says that it is a `thing'.
-A more specfic model might say that it is an image.
+A more specific model might say that it is an image.
 Even more specific would be a painting, a Monet, or specifically the painting Impression of a sunrise.
@@
 Of course, Kolmogorov complexity does allow us to say that it is unlikely to be a Jackson Pollock painting, or a piece of music.
-And, if we were to get a second sample from the same source, we \emph{could} answer the question: we would receive either the same painting again, another Monet, the same painting again, or simply another `thing', and we would gain information about the model.
+And, if we were to get a second sample from the same source, we \emph{could} answer the question: we would receive either the same painting again, or perhaps another Monet, and we would gain information about the model.
 But a single sample by itself should give us no reason to choose an intrinsic model.
 Ultimately, this is just a metaphor, but since the notion of sophistication is driven by intuition, we feel that an intuition to the contrary should serve to sharpen the debate.
-
-We note that these model-selection problems are not specific to Algorithmic statistics.
-They can largely be directly translated to the MDL and Bayesian frameworks.
-While proponents in these fields have never made explicit claims of asymptotically objective model selection, it is a sobering thought that when the model class is extended to all computable distributions, model selection on a single sample becomes arbitrary.
-(Among the candidate set)
 
@@
 This gives us a simple framework, with a one-to-one relation to Bayesian statistics and MDL.
-
-We are left with the original question that sophistication was designed to answer.
-How can we distinguish pure randomness from intuitively interesting information.
-We note that the idea of depth is built on a more solid fundament.
-For instance, contrary to sophistication, encountering a string with high sophistication does imply, with high probability, that the source was of high depth also.
-Unfortunaly, depth is not measured in bits, and so moves us away from the idea that the useful complexity of a stirng should be measured in bits.
-It remains to be seen whether this is a fundamental issue, or whether the interesting information in a string can be objectively expressed in the same vein as it plain information content.
 

### 2015-01-19 Facticity/Facticity/paper.tex
10 pagina's

@@
 \usepackage[mathscr]{eucal}
+\usepackage{thmtools}
 
@@
 \begin{abstract}
-Kolmogorov complexity is a sound measure of the amount of information in data, but it does not distinguish structural information from random noise.
-Kolmogorov's definition of the \emph{structure function} was the first attempt to capture this separation.
-Since then many variations of this idea have been proposed, for which we use \emph{sophistication} as an umbrella term.
-The multitude of treatments indicates that the idea is considered worth pursuing by multiple distinguished scholars, but at the same time it shows that the issue has never been satisfactorily resolved.
-We describe three fundamental problems with all existing proposals, showing each of them to be \emph{unsound}.
-Based on these results we we put forward the view that the problem is fundamental: while many nontrivial and useful structural properties of binary strings can be detected without a priori assumptions, it may be impossible to objectively quantify the sophistication.
+Kolmogorov complexity measures the amount of information in data, but does not distinguish structure from noise.
+Kolmogorov's definition of the \emph{structure function} was the first attempt to measure only the structural information in data, by measuring the complexity of an objectively best model under two-part coding.
+Since then, many variations of this idea have been proposed, for which we use \emph{sophistication} as an umbrella term.
+We describe three fundamental problems with all existing proposals, showing each  to be \emph{unsound}.
+Consequently, we put forward the view that the problem is fundamental, it may be impossible to objectively quantify the sophistication.
 \end{abstract}
-
 \section{Introduction}
 
-Kolmogorov complexity breaks away from the tradition in statistics and information theory by allowing the analysis of \emph{individual} binary sequences rather than studying large samples from probability distributions in an assumed class.
-This is important because for many binary strings, such as a binary encoding of a novel, or a holiday snapshot, or the fossil record, it is impossible to determine what distribution generates such objects.
-Kolmogorov complexity is often seen as the first step to a statistics without assumptions.
-
-There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would call ``complex''.
-For example, a sequence of a million coin flips is exceedingly unlikely to have a Kolmogorov complexity of much less than one million bits, even though there is nothing complex about flipping a lot of coins---it's just hard work.
+Komogorov complexity gives us a sound definition of the amount of information contained in a bitstring.
+There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would consider complexity.
+For example, a sequence of a million coin flips will almost certainly have maximal Kolmogoroc complexity, even though there is nothing complex about flipping a lot of coins---it's just hard work.
 For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data.
-This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication},\footnotemark were proposed, see the related work section below.
-In this paper, we investigate three problems with these definitions, and show that they are unsatisfactory.
-We believe the problem is fundamental, and provide several intuitive arguments why these attempts to define sophistication must fail.
-
-\footnotetext{Most proposals coin their own name along with a specific treatment.
-For the sake of clarity, we have chosen to use \emph{sophistication} as an umbrella term to refer to all ideas that fit this mold.}
+This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication},\footnotemark were proposed.
+In this paper, we investigate three serious problems with these definitions.
+We believe that ultimately, some of these issues are insurmountable, and provide several intuitive arguments why any attempt to define sophistication must fail.
 
 The Kolmogorov complexity of a binary string $x$ is, roughly, the length of the shortest computer program that will output $x$.
-This length depends on which programming language is used, but only by a constant, independent of $x$.
-This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings.
-We can see it as \emph{a property of the data}.
+This length depends on which programming language is used, but only by a constant independent of $x$.
+This property is known as \emph{invariance}.
+This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and we can see the Kolmogorov complexity  as \emph{a property of the data}.
 
-A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees if it is to be interpreted as an objective property of a binary string.
-We are interested in a measure satisfying the following requirements:
+A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees. We are interested in a measure satisfying the following requirements:
 
 \begin{enumerate}
-\item Since it measures information, the sophistication should count the number of bits required for an effective description of the structural properties of a binary string.
-\item An analogue of the Invariance Theorem should hold for sophistication: at the very least, there should be strict limits on the degree to which a change in the programming language can affect the sophistication.
+\item Since it measures information, $S(x)$ should count the bits required for an effective description of the structural properties of a binary string.
+\item An analogue of invariance should hold for sophistication: at the very least, there should be strict limits on the degree to which a change in the programming language can affect the sophistication.
 \item There should be no constant $c$ such that $S(x)\le c$ for every input $x$.
-If sophistication is bounded, then knowing its value under one programming language provides no constraints on its value under another language, even for inputs of very high Kolmogorov complexity.
+If sophistication is bounded, then knowing its value under one programming language provides no constraints on its value under another language.
 Moreover, a bounded sophistication would be at odds with the intuition that there is no limit to the amount of structure that a binary string can exhibit.
@@
 \end{enumerate}
+For most proposed definitions of sophistication, we can prove that they fail one or more of the conditions above.
+Vit\'anyi's definition \cite{vitanyi2004meaningful} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we can show that only strings that require an enormous amount of processing to construct, yield a large sophistication.
 
-For most proposed definitions of sophistication, we can prove that they fail one or more of the conditions above.
-Vit\'anyi's definition \cite{vitanyi2004meaningful} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we can show that only extremely deep strings (i.e., strings that require an enormous amount of processing to construct) yield a large sophistication, while intuitively it seems there must be objects that contain a lot of structure all of which can be described without much calculation.
-Moreover in our discussion section we provide arguments that suggest that it should be possible to generalise our negative results to show that Vit\'anyi's sophistication is not invariant.
-
-A valid definition of sophistication must contend with the following issues.
-First, while all definitions are based on a two-part encoding of the data sequence, some researchers use Kolmogorov complexity as a \emph{building block} of their proposed code, while others look for the two parts of the code \emph{inside} the Kolmogorov complexity: when the definition of Kolmogorov complexity is opened up, it turns out to consist of two parts itself.
-In Section~\ref{section:indices}, we show how this second approach sets us up to suffer from to the so-called \emph{nickname problem}: the sophistication becomes highly dependent on the chosen programming language.
-This immediately breaks invariance, unless additional requirements are made on the language.
-We also show that there do exist languages (efficient enumerations of the partial recursive functions) that avoid this issue.
+A valid definition of $S$ must contend with the following issues.
+First, the representation of the model is crucial.
+This is particularly important to those treatments that `open up' the definition of $K$, to consider the representation that witnesses the Kolmogorov complexity.
+In Section~\ref{section:indices}, we discuss the \emph{nickname problem}: the sophistication becomes highly dependent on the chosen programming language, unless it can represent functions with maximum efficiency.
 
 The second issue is that of \emph{overfitting}, which we consider in Section~\ref{section:overfitting}.
-Overfitting is a common problem in the statistical literature, that refers to the tendency of some model selection procedures to select a very complex model that provides a very good fit to the observed data, but does not generalise well to unseen data.
-For example, in a polynomial regression setting, a higher order polynomial can always provide better fit to a data set.
-Therefore, even if the data are in fact noisy samples from a parabola, naive model selection methods may select a much higher order polynomial that provides a worse description of the trend in the data, encoding all noise in the model parameters.
+Overfitting is a common problem in the statistical literature, that refers to the tendency of model selection to choose a very complex model that provides a very good fit to the observed data, but does not generalise well to unseen data.
+For example, in a polynomial regression setting, a polynomial of sufficiently high order can always cover all data points.
+However, if the data was sampled from a straight line, with a small amount of noise, such a polynomial would be far too complex: it would encode all noise inside the model.
 The same can occur for the sophistication.
@@
 In statistics, this is typically addressed by penalising complex models.
-However, in the sophistication setting, such penalties tend to break the delicate balance between structural information and noise, and lead to the opposite problem: underfitting.
+In sophistication, however, such penalties tend to break the balance between structural information and noise, and lead to the opposite problem: underfitting.
 
-Underfitting (Section~\ref{section:underfitting}) is the opposite problem, where the selected model is simple, but fails to represent part of the structure in the data.
-It is an especially serious problem for sophistication because the models for used in sophistication are so powerful.
+Underfitting (Section~\ref{section:underfitting}) occurs when the selected model is simple, but fails to represent some of the structure present in the data.
 In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language.
-Such interpreters, or \emph{universal models}, can be described in a constant amount of bits, and can generate the data with from a program the length of the Kolmogorov complexity, essentially encoding all information as noise.
-If complex models are penalized to avoid overfitting, then the problem becomes to make sure that universal models are not \emph{always} preferred for complex data.
+Such \emph{universal models} can be described in a constant amount of bits, and  represent the data with a program the length of the Kolmogorov complexity, essentially encoding all information as noise.
+If complex models are penalized, then the problem becomes to make sure that universal models are not \emph{always} preferred for complex data.
 The usual workaround is to restrict the set of allowed models to exclude the universal ones.
-For example, the model class may be restricted to the set of all total recursive functions (only programs that halt for all inputs) rather than all partial recursive functions (all programs).
-While this patches the problem of universal models to an extent, it is questionable whether it adequately solves the problem of underfitting in general.
+For example, the model class may be restricted to the set of all total recursive functions.
+While excludes universal models, it is questionable whether it adequately solves the problem of underfitting in general.
 
-Finally, in the discussion in Section~\ref{section:conclusion} we argue that while two-part coding can yield useful insights into the structure of the data and allows us to identify some models as unlikely to capture all structure in the data, it is probably not possible to uniquely separate structure from noise and identify a \emph{single} model as ``best'': in general many models, of vastly different complexities, may be reasonable models of the data generating mechanism.
-Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that sometimes the data allows for multiple equivalent interpretations, and that there is no such thing as sophistication.
+Finally, in the discussion in Section~\ref{section:conclusion} we argue that while two-part coding can yield useful insights into the structure of the data and allows us to identify some models as poor representations, it is probably not possible to uniquely separate structure from noise and identify a \emph{single} model as ``best'': in general many models, of vastly different complexities, may be reasonable representations.
+Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that the data allows for multiple equivalent interpretations, and that there is no such thing as sophistication.
 
 \subsection{Related work}
-The idea of separating structure from noise originated with Kolmogorov's lecture in Tallinn, Estonia in 1973 introducing the \emph{structure function} \cite{cover1985kolmogorov}.
-In this theory, the two-part description of the data consists of a program that generates a finite set, and an index of the original object in the set.
-The length of the program then measures the structure in the data.
-This approach was discussed more extensively, and generalised, by Vereshchagin, G\'acs and Vit\'anyi \cite{vereshchagin2004kolmogorov,gacs2001algorithmic}.
+The idea of separating structure from noise originated with Kolmogorov's \emph{structure function} \cite{cover1985kolmogorov}.
+In this theory, the data is described as finite set together with an index in the set.
+The complexity of the set determines the sophistication.
+This approach was discussed more extensively, and generalised in \cite{vereshchagin2004kolmogorov,gacs2001algorithmic}.
 
-Koppel \cite{koppelSoph1988,koppel1991almost} also built on Kolmogorov and Cover's work to define a measure of the meaningful information in a binary string, called \emph{sophistication}, the term we adopt in this paper.
-He changed the class of models from finite sets to total functions and used monotone functions, instead of prefix-free Kolmogorov complexity, to be able to extend the theory to infinite strings, an approach which has been abandoned by subsequent authors.
-We will only consider the finite version (based on regular functions) in this paper.
+Koppel \cite{koppelSoph1988,koppel1991almost} also built on the structure function work to define a measure of the meaningful information in a binary string, coining \emph{sophistication}, the name we adopt in this paper.
 
 In 1996 Gell-Mann and Lloyd introduced \emph{effective complexity} \cite{gellmann1996information}.
-Working from the perspective of statistical mechanics, they arrived (largely independently, it seems) at a notion that fits the mold of sophistication.
+Working from the perspective of statistical mechanics, they arrived (independently, it seems) at a notion that fits the mold of sophistication.
 
 Antunes et al.
-\cite{antunes2009sophistication} note that the constant cutoff used by Koppel to separate candidate two-part descriptions from those deemed to be too inefficient, is a source of instability, a result in the spirit of this paper.
+\cite{antunes2009sophistication} note that the constant cutoff used by Koppel is a source of instability, a result in the spirit of this paper.
 They propose a different definition called \emph{coarse sophistication}.
@@
 Vereshchagin \cite{vereshchagin2013algorithmic} proposes an alternative solution to the instability issue.
-
 Adriaans' \emph{Facticity}, introduced in 2012\cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions.
@@
 
-\subsection{Preliminaries, definitions and desiderata}
+\subsection{Notation}
 
 Let $\B = \{0,1\}^*$.
-We deal with partial functions $\phi: \B \to \B$, which we will also call \emph{models}.
-Such functions are \emph{computable} if they can be computed by a Turing machine as described in \cite[Definition~1.7.1]{li1993introduction}.
-A function is prefix-free if its domain, $D_f = \{y : f(y) \neq \infty\}$ is a prefix set.
+We deal with computable partial functions $\phi: \B \to \B$, which we will call \emph{models}.
+A function is prefix-free if its domain, $D_f = \{y : f(y) \neq \infty\}$ is a prefix free set.
 A function $f$ is \emph{total} if $D_f = \B$.
 A prefix-free function is \emph{complete} if every infinite binary string has a member of $D_f$ as a prefix.
-\footnotemark
-
-\footnotetext{Completeness is analogous to totality: a total function corresponds to a Turing machine that halts for all its inputs, while a complete function corresponds to a Turing machine that always halts, so long as there are enough bits on the input tape.}
+We fix a prefix code with $|\bar{x}| \le|x|+2\log(|x|+2)|$ (eg.
+\cite[Section~1.4]{li1993introduction}).
 
 \begin{definition}[Numberings]
-A numbering is an enumeration of of the partial computable functions, denoted as $\phi_1, \phi_2, \ldots$.
-We fix one canonical, which we know to be an effective description, that is for any $i \in \N, y\in \B$, we can effectively compute $\phi_i(y)$, where the computation diverges if $\phi_i(y) = \infty$.
-We call a numbering $\num{\psi}$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ such that for all $i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{b(i)}$.
-We use the greek letters without a subscript (eg.
-$\phi$ and $\psi$) to refer to the numbering as a whole.
+A numbering is an enumeration of of the partial computable functions, denoted as $\phi_1, \phi_2, \ldots$ or simply $\phi$.
+We fix one effective numbering $\phi$.
+We call a numbering $\num{\psi}$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ with $\forall: i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{b(i)}$.
 \end{definition}
 
-\begin{definition}[Prefix encoding function]
-  A prefix encoding function $f:\B\to\B$ is an injective
-  function that maps the binary strings to a prefix-free set. There
-  are prefix encoding functions $g$ with the property that
-  $|g(x)|\le|x|+2\log(|x|+2)|$ \cite{li1993introduction}. We fix such a function and
-  denote it by $\br{x}=g(x)$.
+\begin{definition}[Complexity]
+Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ corresponds to the set of all computable prefix functions.
+We define the plain and prefix Kolmogorov complexity as $C(x)=\min\{|\bar\imath y|:\phi_i(y)=x\}$ and $K(x)=\min\{|\bar\imath y|:\phi_i(y)=x, i\in\Pi\}$ respectively.
+\label{definition:complexity}
 \end{definition}
 
-\begin{definition}[Complexity]\belowdisplayskip=-12pt
-Let $\phi_1,\phi_2,\ldots$ be the canonical acceptable enumeration of $\C$.
-Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ corresponds to the set of all computable prefix functions.
-We define the plain and prefix Kolmogorov complexity as
-\begin{align*}
-C(x)&=\min\{|\bar\imath y|:\phi_i(y)=x\},\\
-K(x)&=\min\{|\bar\imath y|:\phi_i(y)=x, i\in\Pi\}.\\
-\end{align*}\label{definition:complexity}\vspace{-18pt}
-\end{definition}
-This definition of $K$ is uncommon.
-It is equivalent to the standard definition as used in \cite{li1993introduction}, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).
-
-We will generalize the main proposals for sophistication into a single notation.
-We start with a class of models, we consider all two-part descriptions of the data using these models, we select the set of representations with the shortest code-lengths and we select the smallest model from this set.
-The size of the smallest model represents the sophistication of the data.
-
-\begin{definition}[Models and model classes] A \emph{model class} is a set of models. We distinguish four important model classes:
-  \begin{itemize}
-  \item $\C$ is the set of all partial computable functions,
-  \item $\K\subset\C$ is the set of all prefix-free
-    functions,
-  \item $\T\subset\C$ is the set of all total functions,
-  \item $\P\subset\K$ is the set of all complete, prefix-free functions,
-  \item $\F\subset\P$ contains, for every finite set $S$, a computable
-    surjective function mapping the binary sequences of length
-    $\lceil\log S\rceil$ onto $S$. In other words, this is the model class of finite sets.
-  \end{itemize}
-\end{definition}
-We note that classes $\T$ and $\P$ are unusual in that they are not enumerable.
-\begin{definition}[Representations and code-lengths]
-A pair $(m,y)$ is a \emph{representation} for $x$ if $m$ is a model and $m(y)=x$.
-Define the following code-length functions for the representations:
-\begin{itemize}
-\item $L(m)$ is the length of a self-delimiting description of
-  the model $m$, i.e. it is the length function of a prefix code for models. 
-\item $L^m(x)=L(m)+\min\{|y|:m(y)=x\}$ measures the number of bits required to describe both the model $m$ and the data $x$ using the model.
-\item $L^\M(x)=\min\{L^m(x):m\in\M\}$ measures the number of bits
-  required to describe $x$ using the best model in the model class.
-\end{itemize}
-\end{definition}
-We can now define which of these representations express the data most efficiently:
-\begin{definition}[Candidate set]
-The \emph{candidate models} for $x$ from model class $\M$ and threshold function $\tau$ are
-\[
-  C^\M_\tau(x):=\{m\in\M:L^m(x)\le \tau(x) + c\}
-\]
-with $c$ a constant independent of $x$ which may be chosen arbitrarily.
-The default threshold function is $\tau^\circ(x)=L^\M(x)$.
-We will often omit the subscript if the default threshold function is used.
+\begin{definition}[Model classes] A \emph{model class} is a set of models.
+We distinguish four important model classes: the partial (computable) functions ($\C$), the prefix functions ($\K \subset \C$), the total functions ($\T \subset \C$), the complete functions ($\P \subset \K$) and the finite sets ($\F \subset \P$, encoded as a prefix-free function mapping all strings of length $\lceil\log S\rceil$ to members of $S$).
 \end{definition}
 
-\begin{definition}[Sophistication]\belowdisplayskip=-12pt
-  The model complexity, meaningful information, effective complexity, facticity or \emph{sophistication},  of a string $x$ given a model class $\M$ and threshold function $\tau$ is
-  \[
-  \s_\tau^{\M}(x):=\min\{L(m):m\in C^\M_\tau(x)\}\p
-  \]\label{definition:sophistication}
+\begin{definition}[Representations]
+A pair $(i,y)$ is a \emph{representation} for $x$ if $\phi_i(y)=x$.
+Define the following code-length functions for the representations:\\
+\-\hspace{1cm}$L(m)$ is the length of a self-delimiting description of the model $m$.\\ 
+\-\hspace{1cm}$L^m(x)=L(m)+\min\{|y|:m(y)=x\}$\\
+\-\hspace{1cm}$L^\M(x)=\min\{L^m(x):m\in\M\}$
 \end{definition}
-Finally, we must formalize what we mean when we say that we expect sophistication to be invariant.
-It may be that a less strict invariance than we have for the Kolmogorov complexity, still yields a useful sophistication.
-\begin{definition}
-Let $\gamma:\N\to \N$ be some nondecreasing function. Two functions $f$ and $g$ are $\gamma$-equivalent if $|f(x) - g(x)| \in O(\gamma(\max(f(x), g(x)))$.
+We can now define sophistication generically:
+\begin{definition}[Sophistication]
+The \emph{candidate models} for $x$ from model class $\M$ and threshold function $\tau$ are
+$C^\M_\tau(x):=\{m\in\M:L^m(x)\le \tau(x) + c\}$
+with $c$ a constant independent of $x$ which may be chosen arbitrarily. The default threshold function is $L^\M(x)$. The \emph{sophistication} is \\ 
+\-\hspace{1cm}$\s_\tau^{\M}(x):=\min\{L(m):m\in C^\M_\tau(x)\}$\label{definition:sophistication}
 \end{definition}
-We say that a formalization of sophistication $\s_\phi$ built on acceptable numbering $\phi$ is $\gamma$-invariant to the choice of numbering is $s_\phi$ and $s_\psi$ are $\gamma$-equal for any two acceptable numberings $\phi$ and $\psi$.
+Finally, we generalize the notion of invariance to allow for less strict versions. 
+\begin{definition}[$\gamma-invariance$]
+Let $\gamma:\N\to \N$ be some nondecreasing function.
+Two functions $f$ and $g$ are $\gamma$-equivalent if $|f(x) - g(x)| \in O(\gamma(\max(f(x), g(x)))$.
+We say that a function $\s_\phi(x)$ is \emph{$\gamma$-invariant} to the choice of numbering if $s_\phi$ and $s_\psi$ are $\gamma$-equal for any two acceptable numberings $\phi$ and $\psi$.
+\end{definition}
+
 \begin{lemma}
-If a function $\S_\phi(x)$ is bounded for a given numbering, it is either not $\gamma$-invariant or it is  $\gamma$-equal to a bounded function.
+If a function $\s_\psi(x)$ is bounded for some $\psi$, it is either not $\gamma$-invariant or it is  $\gamma$-equivalent to a bounded function.
 \end{lemma}
-The proof follows directly from the definitions.
-Most lemmas in the rest of the paper will simply prove that a given variant is bounded for a specific numbering, implicitly using this lemma to show that the variant is unsound.
 
@@
 \label{section:indices}
-We start from a simple intuition: the definition of Kolmogorov complexity already uses a description of the data in two parts; a computable function and an input.
-Can we assume that, when such a description is as short as the Kolmogorov complexity, it automatically creates a good separation into the structural part of the data and the noise?
-As we will show, the representation of function plays a fundamental part in answering this question.
+We start from a simple intuition: the definition of $K$ already uses a description of the data in two parts; a computable function and an input.
+Can we assume that, when such a description is optimal, it automatically separates the structural and random information?
+As we will show, the representation of functions plays crucial fundamental.
 
-The issues can be very subtle and require a careful attention that is rarely given in the literature. We will first illustrate what happens if any acceptable numbering is allowed:
-
-\begin{lemma}
+We will first illustrate what happens if any acceptable numbering is allowed:
+\begin{restatable}{lemma}{ineffindices}
 Let $\phi$ be some acceptable numbering, and $\s_\phi(x)$ be any function which gives the length of some index $i$ such that $\phi_i(y) = x$ for some $y$.
@@
 No such function can be both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
-\end{lemma}
-\begin{proof}
-Choose some $\gamma(n) \in o(n)$. Assume towards contradiction that the sophistication is $\gamma$-invariant and unbounded.
-
-Let $\num{\phi}$ be the canonical acceptable numbering. Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.
-
-We now define two new acceptable numberings $\num{\psi}$ and $\num{\xi}$:
-\[
-\psi_j(x) = \begin{cases}
-	\phi_i(x) \;\text{if}\; j = s_{2i}\\
-	\infty\;\text{otherwise}
-\end{cases}\,\,\,
-\xi_j(x) = 
-\begin{cases}
-	\phi_i(x) \;\text{if}\; j = s_{2i+1} \\
-	\infty \;\text{otherwise}\p
-\end{cases}
-\]
-Let $\s_\psi$ and $\s_\xi$ be the sophistication functions built on these enumerations.
-Choose any $x$ and assume w.l.o.g.
-that $\s_\psi(x) < \s_\xi(x)$.
-By construction of the numberings $\psi$ and $\xi$, we have $2\;\s_\psi(x)  \leq \s_\xi(x)$, which gives us:
-\begin{align*}
-	\s_\xi(x) - \s_\psi(x) &\geq \frac{1}{2}\, \s_\xi(x)  \in \Omega(\s_\xi(x))\p 
-\end{align*}
-Since, by assumption, $\s_\xi$ can be arbitrarily large and $\s_\xi(x) - \s_\psi(x) \in O(\gamma(\s_\xi(x))) \subset o(\s_\xi(x))$, we have a contradiction.
-\end{proof}
-In this proof, we deliberately used ridiculously inefficient index functions.
-Obviously, one could simply pick a more efficient index; the point is that many definitions of sophistication do not mention any such restriction.
+\end{restatable}
+\noindent In the proof of this lemma (Section~\ref{section:appendix-inefficient-indices}), we deliberately used ridiculously inefficient index functions.
+We could use a more efficient index, but most definitions of sophistication do not mention any such restrictions.
 As we shall see, any inefficiency in the index will lead to a lack of invariance in sophistication.
 
-The treatments that fall prey to this problem are Koppel and Atlan's \cite{koppelSoph1988,koppel1991almost},  and several of its variants (\cite{antunes2009sophistication,antunes2013sophistication}).
+The treatments that fall prey to this problem are Koppel and Atlan's \cite{koppelSoph1988,koppel1991almost},  and several of its variants \cite{antunes2009sophistication,antunes2013sophistication}.
 
-Interestingly, using an inefficient index is no problem in the definition of Kolmogorov complexity, since we can select as a model a universal Turing machine with an efficient index at only a constant penalty.
-But this robustness does \emph{not} carry over to definitions of sophistication.
-If we are interested in separating the model and its input, we must represent the model efficiently.
+Interestingly, inefficient indices are no problem in the definition of Kolmogorov complexity.
+However, this robustness does \emph{not} carry over to definitions of sophistication.
+Simply put, the length of a model in an ordinary description does not represent its complexity, much like the length of a compressible string does not reflect its information content.
+In order to obtain a robust model complexity, we use an extension of Kolmogorov complexity:
 
-Simply put, the length of a model in an ordinary description does not represent its complexity, much like the length of a compressible string does not reflect its information content.
-In order to obtain a more robust measure of model complexity, we describe a natural extension of Kolmogorov complexity to functions, and show that this retains invariance.
+\begin{definition}[\cite{grunwald2004shannon,vitanyi2004meaningful}]
+The complexity of a partial recursive function is defined by $C_\phi(f) = \min\{C(i):\phi_i=f\}$ and $K_\phi(f) = \min\{K(i):\phi_i=f\}$.\label{definition:model-complexity}
+\end{definition}
+We show in the appendix (Lemma~\ref{lemma:invariance}) that these are invariant, ie.
+complexity is a property of a function.
+There are two ways to use this notion of model complexity for more robust attempts to define sophistication.
+Confusingly, both approaches are used in the literature.
+\begin{enumerate}
+  \item We maintain that the representation witnessing the Kolmogorov complexity determines the sophistication, but impose the additional constraint that an efficient, acceptable numbering is used.
+Specifically, that the length of the indices satisfies $|i| = C(\phi_i)$.
+This approach is taken by Adriaans \cite{adriaans2012facticity}.
+  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use $K$ to measure the complexity of the \emph{model} (as in Definition~\ref{definition:model-complexity}), which is then the first part of a two-part code describing the data.
+With this approach, the total code length is $K(\phi_i)+|y|$, with $\phi_i(y)=x$.
+In this approach, we can construct everything on the basis of a single acceptable numbering, without additional constraints.
+This approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information}.
+\end{enumerate}
+To show that the two perspectives are equivalent, we must show that there exists an acceptable numbering which is efficient enough:
+\begin{definition}[Faithful Numberings]\label{def:faithful}
+  A numbering $\num{\psi}$ is \emph{faithful} if
+  for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and
+  $|j|\le C(\psi_j)+c$, for some constant $c$.
+\end{definition}
+This definition is a refinement of \cite[Definition~10]{adriaans2012facticity}.
+We show in the appendix (Lemma~\ref{lemma:faithful-numberings}) that such numberings exist.
+Unfortunately, even with a faithful numbering, we can fail:
+\begin{restatable}{lemma}{ineffprefix}
+Let $\phi$ be faithful and let $L(\phi_i) = |\bar\imath|$, for some prefix encoding function.
+If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then $\s^\C$ is bounded for all numberings.\label{lemma:prefix-inefficiency}
+\end{restatable}
+\noindent This result applies to \cite{adriaans2012facticity}.
+To avoid the problem, the model $\phi_i$ should be encoded in $K(\phi_i)$ bits, relying on $K$'s own prefix encoding to delimit the model.
 
-\begin{definition}\belowdisplayskip=-12pt
-  The complexity of a partial recursive function is defined by
-  \begin{align*}
-    C_\phi(f) &:= \min\{C(i):\phi_i=f\}\;\tn{,}\\
-    K_\phi(f) &:= \min\{K(i):\phi_i=f\}\p
-  \end{align*} \label{definition:model-complexity}
+Note that the condition for this lemma likely applies to all possible computable  prefix encoding functions (see Section~\ref{section:prefix-inefficiency}).
+Thus, even if we have a faithful numbering, we can only use it if we don't count the cost of delimiting the model from the input (as in \cite{koppelSoph1988}).
+In the following, we will take the second perspective: the representations are two-part descriptions $(i, y)$ such that $\phi_i(y)=x$, with codelength $K(\phi_i) + |y|$.
+\section{Overfitting: the singletons}
+\label{section:overfitting}
+We now come to the main issue plaguing sophistication.
+If we use model class $\K$, the candidate set always contains two representations that do not provide a good separation.
+The first is a \emph{universal model}: a basic universal Turing machine.
+The second is a \emph{singleton model} a model which returns $x$ for any input.
+As we will show, in many settings, we can ensure that one of them is the \emph{only} model in the candidate set.
+The following lemma generalizes the principle:
+
+\begin{restatable}{lemma}{coolone}
+\label{lemma:thecoolone}
+  Let $\psi_1,\psi_2,\ldots$ be any acceptable enumeration of the partial recursive functions.
+  Let $\M$ be any model class, let $\X$ be any set of binary sequences and let $D:\{0,1\}^*\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e.
+if $f=\psi_{D(p)}$ then $p$ is a $D$-description of $f$.
+Let $\M'=\{\psi_i:i\in\tn{range}(D)\}$.
+Further assume there is a constant $c$ such that:\\
+\-\hspace{1cm}(1)$\forall_{f\in\M'}:\min\{|p|:\psi_{D(p)}=f\}\le K^\psi(f)+c$\\
+\-\hspace{1cm}(2)$\forall_{x\in\X}:L^{\M',\psi}(x)-L^{\M,\psi}(x)\le c$.\\
+Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
+\end{restatable}
+\noindent We can now show numberings for which the singletons are always selected:
+
+\begin{restatable}[Overfitting]{lemma}{overfitting}
+Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$.
+Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the prefix partial recursive functions, and a constant $c$, such that
+\[
+K(x)-S^{\M,\phi}(x)\le c
+\]
+for all $x\in\X$.
+\end{restatable}
+\noindent In short, this lemma tells us that there are numberings for which the singleton models always determine the sophistication.
+This means that for any dataset, the sophistication is equal to the Kolmogorov complexity.
+
+Many variants in the literature (including the structure function) are susceptible to this problem \cite{cover1985kolmogorov,gacs2001algorithmic,vereshchagin2013algorithmic}.
+\cite{vitanyi2004meaningful,adriaans2012facticity} avoid it by not using self-delimiting representations.
+Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically not optimally efficient.
+While this dodges the issue of the singletons, it does not show that $S$ is invariant.
+
+Another approach may be to make the constant threshold dependent on the choice of enumeration, although we are not aware of any published efforts.
+
+\section{Underfitting: the universal model}
+\label{section:underfitting}
+On the other end of the spectrum, there exist representations with all information in the noise, using a universal model.
+We can easily show that if such models exist, we can choose our numbering so that they are always chosen.
+\begin{restatable}[underfitting]{lemma}{underfitting}
+Let $\M$ be a model class containing model $u$, called a \emph{universal model} with $\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p$.
+Then, there exist numberings such that $\s^\M$ is bounded.
+\end{restatable}
+\noindent Most sophistication measures avoid the underfitting problem by limiting their model class, usually to either finite sets ($\F$) or total functions ($\T$ or $\P$).
+This eliminates the universal model, but as we shall see, it does not solve the problem for most data.
+The variants that are susceptible to the this type of underfitting are facticity \cite{adriaans2012facticity} and some versions of effective complexity \cite{gellmann1996information}.
+The latter aims to bypass the problem by selecting carefully from the candidate set, but as we have shown, for some numberings the candidate set contains only the universal model.
+
+\subsection{The problem of depth}
+
+We have seen that some variants of sophistication sidestep the simplest examples of underfitting by using total models.
+Does this ensure that these variants are not bounded?
+A well-known result in algorithmic statistics shows that there exist strings for which no finite set is as good a model as the singleton set.
+In this section we will focus on the variants $S^\T$ and $S^T_K$ \cite{vitanyi2004meaningful}, as those are the variants most resistant to proofs of non-invariance.
+
+\begin{restatable}{lemma}{nonstochastic}
+There are infinitely  $x$ such that $S^\F_K(x) \geq K(x)/5$. \label{lemma:structure-function-is-not-bounded}
+\end{restatable}
+
+\begin{restatable}{lemma}{vitanyi}
+Let $f$ be a total, computable function, such that for some $d$, $f(d) = x$ and let $k = K(f) + |d|$.
+Then there exists a finite set $S$ containing $x$ and a constant $c$, such that $K(S) \leq K(f) + K(|d|) + c$ and $\log |S| \leq |d|$.\label{lemma:total-to-sets}
+\end{restatable}
+\noindent This lemma is a variation on \cite[Lemma~7.2]{vitanyi2004meaningful}.
+We use this lemma to transport the result of Theorem~\ref{lemma:structure-function-is-not-bounded} to the model class of total functions:
+
+\begin{restatable}{theorem}{dogfood}
+There are infinitely many $x$ with $S^\T_K(x) \geq K(x)/6$ and $S^\T(x) \geq K(x)/6$. 
+\end{restatable} 
+\noindent This tells us that $S^\T_K(x)$ and $S^\T(x)$ can truly grow to a significant part of the total description length.
+While this shows that sophistication, if defined with particular care, is not entirely vacuous, we can shed some light on exactly which strings end up with high sophistication, using the principle of \emph{depth}.
+We will use the following definition:
+
+\begin{definition}[Depth\cite{bennett1988logical,antunes2006computational}]\belowdisplayskip=-12pt
+Let $U$ be some universal Turing machine, so that $U(\bar\imath y) = \phi_i(y)$.
+Let $U^t$ be a simulation of this machine, which is allowed to run for at most $t$ steps, and returns $0$ if it has not yet finished at that point.
+Let $C^t_\M(x) = \min\{|\bar\imath y| : U^t(\bar\imath y) = x, \phi_i \in \M\}$.
+
+The \emph{$c$-depth} $d^\M_c(x)$ of a string is $d^\M_c(x) = \min \left\{t : C^\M_t(x) - C^\M(x) \leq c \right\}$.
 \end{definition}
-These definitions also appear in \cite{grunwald2004shannon} and \cite{vitanyi2004meaningful}.
-We will show that these values are invariant up to a constant (ie.
-complexity is an objective property of a function):
-\begin{lemma}[Invariance]
+Deep strings are those that can only be optimally compressed with a great investment of time.
+We note that it is exceedingly unlikely that a deep string is sampled from a shallow distribution \cite{bloem2014safe,bennett1988logical}.
+
+\begin{restatable}{theorem}{depth}
+Let $A(n)$ be the single-argument Ackermann function and $c$ some arbitrary constant.
+There are numberings such that for all strings with depth $d^\C_c(x) \leq A(C(x))$, $\s^\T(x)$, $s^\P(x)$ and $\s^\F$ are bounded.
+\end{restatable}
+\noindent This shows that while high-sophistication strings exist, the mechanics do not follow the basic intuition of sophistication.
+Unless we encounter data that would take longer than the lifetime of the universe (and then some) to produce, sophistication behaves exactly as it would without the restricted model class: a solution to a super-exponential problem would have no greater sophistication than a string sampled from a simple Bernoulli model.
+This result also suggests that the `non-stochastic' property of strings with high sophistication\cite{shen1983concept,vereshchagin2004kolmogorov} says more about depth and totality than it does aboutrandomness and structure.
+
+The relation between sophistication, depth and absolutely non-stochastic strings was also investigated in \cite{antunes2013sophistication}.
+
+\section{Conclusion}
+\label{section:conclusion}
+As we have seen, the idea behind sophistication has a rich history.
+It has been proposed many times, by many different authors.
+This tells us two things.
+Firstly, there is a strong intuition behind sophistication, shared by many respected authors, and second, it is not easy to get the definition exactly right.
+
+So why should we consider this intuition reasonable at all?
+The first assumption of sophistication is that high compressibility and pure randomness are objectively `uninteresting'.
+A more sophisticated signal, such as a television broadcast must contain, in some objective sense, more meaningful information.
+Whether this intuition seems reasonable likely differs per person, but sophistication goes one step further: it also suggests that we can objectively separate the information in a string into random and structured.
+
+To see whether this intuition is valid; imagine that we sample from the universal distribution by feeding random bits to a universal Turing machine.
+If this UTM is constructed in the conventional manner, the initial bits will determine the index of a partial computable function $i$, and the rest will be the input to that machine $y$.
+What's the characteristic model of this data?
+Certainly, we cannot exclude the possibility that $\phi_i$ is also a universal model.
+Also, we cannot exclude the possibility that $\phi_i$ is a singleton model for $x$.
+
+Although sampling perspectives are discouraged by proponents of algorithmic statistics, this example does clash with the intuition behind sophistication: there is no reason to assume that a string with high sophistication was sampled from a complex model, so how can a complex model be intrinsic to the data?
+\footnotemark
+
+\footnotetext{Note that computational depth does not have this problem. Only deep sources produce deep strings.}
+
+Consider the following metaphor.
+We are given a bitstring which can be read as a bitmap image of the painting \emph{Impression of a Sunrise}.
+The candidate set for this string contains various models from very generic to very specific.
+The theory behind sophistication suggests that we can choose one of these as the objective, intrinsic model of the data.
+Different models say different things: the universal models says that it is a `thing'.
+A more specific model might say that it is an image.
+Even more specific would be a painting, a Monet, or specifically the painting Impression of a sunrise.
+These are the models we imagine we may find in the candidate set.
+Can we say that the data is intrinsically more of an image than a painting?
+More of a Monet than a painting?
+We see no intuitive reason why such a distinction should be possible.
+
+Of course, Kolmogorov complexity allows us to say that it is unlikely to be a Jackson Pollock, or a piece of music.
+And, if we were to get a second sample from the same source, we \emph{could} answer the question: we would receive either the same painting again, or perhaps another Monet.
+But, a single sample does not give us enough evidence to choose a single model.
+Ultimately, this is just a metaphor, but since the notion of sophistication is driven by intuition, we feel that an intuition to the contrary should serve to sharpen the debate.
+
+For this reason, we take a skeptical view of sophistication.
+We can offer the following insight, to back up our skepticism.
+As we have seen, candidates within a constant of the optimal representation can occur all over the spectrum, from small to large models.
+A small change in construction, such as a different acceptable numbering, or a change in threshold function, can arbitrarily push models into or out of the candidate set.
+Whereas in Kolmogorov complexity, constant changes have constant effects on the outcome, here constant changes can cause large changes in the outcome.
+While no authors have touched on the issue of invariance, it is related to the issue of instability \cite{antunes2013sophistication,vereshchagin2013algorithmic}.
+We consider invariance the first criterion by which any proposal must be judged, with the burden of proof resting with the proposer, rather than the critic.
+
+Finally, we note that the use of total functions has a very subtle effect on the discussion surrounding sophistication.
+Since the total functions are not enumerable, the value of the sophistication relies strongly on the difference between the Kolmogorov complexity and the Kolmogorov complexity bounded to total functions.
+This gap is exceedingly difficult to capture, because the total functions cannot be enumerated.
+This is dangerous: if such definitions cannot be proved wrong or right, they will stand as the ad-hoc official view, while the restrictions used to elude proofs of incorrectness make them very difficult to build on.
+This creates an artificial dead end for a very valuable area of research.
+
+\subsubsection*{\ackname}
+
+This publication was supported by the Dutch national program COMMIT and by  the Netherlands eScience center.
+
+\bibliographystyle{plain}
+\bibliography{facticity}
+
+\appendix
+
+\section{Proofs and further discussion}
+
+\subsection{Invariance of function complexity}
+
+\begin{lemma}[Invariance of function complexity]
 Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be any two acceptable numberings.
@@
 
-There are two ways to use this notion of model complexity for more robust attempts to define sophistication. Confusingly, both approaches are used in the literature,
-\begin{enumerate}
-  \item We maintain that the representation witnessing the Kolmogorov complexity determines the sophistication, but impose the additional constraint that an efficient, acceptable numbering is used.
-Specifically, that the length of the indices satisfies $|i| = C(\phi_i)$.
-We construct such an index below.
-This approach is taken by Adriaans \cite{adriaans2012facticity}.
-  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use Kolmogorov complexity to measure the complexity of the \emph{model} (as in Definition~\ref{definition:model-complexity}), which is then the first part of a two-part code describing the data.
-With this approach, the total code length is $K(\phi_i)+|y|$, where $y$ is an input to the model $\phi_i$ such that $\phi_i(y)=x$.
-The difference is that in this approach, we can construct everything on the basis of a single acceptable numbering, without additional constraints.
-This approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information}.
-\end{enumerate}
-To show that the two perspective are equivalent, we must show that there exists an acceptable numbering which is efficient enough.
+\subsection{Existence of faithful numberings}
 
-\begin{definition}[Faithful Numbering]\label{def:faithful}
-  A numbering $\psi_1,\psi_2,\ldots$ of the partial recursive
-  functions is \emph{faithful} if there is a constant $c$ such that
-  for all indices $i$ there is a $j$ such that $\psi_i=\psi_j$ and
-  $|j|\le C(\psi_j)+c$.
-\end{definition}
-This definition is a refinement of the one used in \cite{adriaans2012facticity}. \footnotemark
-
-\footnotetext{That publication defines a faithful \emph{index function}, rather than a numbering, and notes that it is not computable.
-Since a numbering is all that is needed, we only need to show acceptability.
-The incomputability of the index function is no issue.}
 \begin{lemma}
-  There is a faithful acceptable numbering.
+  There is a faithful acceptable numbering.\label{lemma:faithful-numberings}
 \end{lemma}
@@
 \end{proof}
-Unfortunately, even with a faithful index, we can fail if the prefix-function used to separate the model from its input is inefficient:
-\begin{lemma}
-Let $L(\phi_i) = |\bar\imath|$ for a faithful index, and some prefix encoding function, and let $\s^\C$ be defined on this $L$ with the default threshold function.
 
-If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then $\s^\C$ is bounded for all numberings.\label{lemma:prefix-inefficiency}
-\end{lemma}
+\subsection{Inefficient indices}
+\label{section:appendix-inefficient-indices}
+
+\ineffindices*
+
+\begin{proof}
+Choose some $\gamma(n) \in o(n)$. Assume towards contradiction that the sophistication is $\gamma$-invariant and unbounded.
+
+Let $\num{\phi}$ be the canonical acceptable numbering. Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.
+
+We now define two new acceptable numberings $\num{\psi}$ and $\num{\xi}$:
+\[
+\psi_j(x) = \begin{cases}
+	\phi_i(x) \;\text{if}\; j = s_{2i}\\
+	\infty\;\text{otherwise}
+\end{cases}\,\,\,
+\xi_j(x) = 
+\begin{cases}
+	\phi_i(x) \;\text{if}\; j = s_{2i+1} \\
+	\infty \;\text{otherwise}\p
+\end{cases}
+\]
+Let $\s_\psi$ and $\s_\xi$ be the sophistication functions built on these enumerations.
+Choose any $x$ and assume w.l.o.g.
+that $\s_\psi(x) < \s_\xi(x)$.
+By construction of the numberings $\psi$ and $\xi$, we have $2\;\s_\psi(x)  \leq \s_\xi(x)$, which gives us:
+\begin{align*}
+	\s_\xi(x) - \s_\psi(x) &\geq \frac{1}{2}\, \s_\xi(x)  \in \Omega(\s_\xi(x))\p 
+\end{align*}
+Since, by assumption, $\s_\xi$ can be arbitrarily large and $\s_\xi(x) - \s_\psi(x) \in O(\gamma(\s_\xi(x))) \subset o(\s_\xi(x))$, we have a contradiction.
+\end{proof}
+
+\ineffprefix*
 \begin{proof}
@@
 \end{proof}
-This result applies to Adriaans' facticity\cite{adriaans2012facticity}.
-In order to circumvent the problem, the representation should should encode the model $f$ in $K(f)$ bits, relying on $K$'s own prefix encoding to delimit the model from its input.
 
-Note that the condition for this lemma applies to all commonly used prefix encoding functions. We conjecture that it applies to all possible computable prefix encoding functions:
+\subsection{Inefficiency of computable prefix codes}
+\label{section:prefix-inefficiency}
+Lemma~\ref{lemma:prefix-inefficiency} only holds if the prefix function $L(i)$ has the property that $\min\{L(i)-K(i):i\ge i_0\}$ is unbounded.
+We conjecture that this is the case for any computable function.
 
+We conjecture that it applies to all possible computable prefix encoding functions:
 \begin{conjecture}
@@
 
-Lemma~\ref{lemma:prefix-inefficiency} shows that even if we have a faithful numbering, we can only use it in settings where we don't count the cost of delimiting the model from the input (such as \cite{koppelSoph1988}).
-In the cases where the data representations are taken as a single code word, the model-part, including delimiting information, must represent the model as efficiently as $K(f)$ does.
+\subsection{Boosting a model class by manipulating the numbering}
 
-In the rest of this paper, we will avoid these issues and take the second perspective: we will consider the representations as two-part descriptions $(i, y)$ with codelength $K(\phi_i) + |y|$, such that $\phi_i(y)=x$.
-
-\section{Overfitting: the singletons}
-\label{section:overfitting}
-We now come to the main issue plaguing sophistication.
-In the naive setting, where we allow all prefix-free functions as models, and use the default threshold function, the candidate set always contains two representations that clearly do not provide a good separation into structure and content.
-The first is a \emph{universal model}: a universal Turing machine which simply runs its shortest program for $x$.
-The second is a \emph{singleton model} a model which returns $x$ for any input.
-
-The key problem is that both can be present in the candidate set, and the formalization of sophistication must ensure that neither are selected.
-As we will show, in many settings, we can ensure that they are the \emph{only} models in the candidate set.
-The following lemma generalizes the principle.
-
-\begin{lemma}\label{lemma:thecoolone}
-  Let $\psi_1,\psi_2,\ldots$ be any acceptable enumeration of the partial recursive functions.
-  Let $\M$ be any model class, let $\X$ be any set of binary sequences and let $D:\{0,1\}^*\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e.
-if $f=\psi_{D(p)}$ then $p$ is a $D$-description of $f$.
-Let $\M'=\{\psi_i:i\in\tn{range}(D)\}$.
-Further assume there is a constant $c$ such that
-\begin{enumerate}
-  \item $\forall_{f\in\M'}:\min\{|p|:\psi_{D(p)}=f\}\le K^\psi(f)+c$
-  \item $\forall_{x\in\X}:L^{\M',\psi}(x)-L^{\M,\psi}(x)\le c$.
-\end{enumerate}
-Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
-\end{lemma}
+\coolone*
 \begin{proof}
@@
 
-We can now show that there exist numberings for which the sophistication always selects the singleton models:
+\subsection{Overfitting}
 
-\begin{lemma}[overfitting]
-Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$.
-Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the prefix partial recursive functions, and a constant $c$, such that
-\[
-K(x)-S^{\M,\phi}(x)\le c
-\]
-for all $x\in\X$.
-\end{lemma}
+\overfitting*
 \begin{proof}
@@
 
-In short, this lemma tells us that there are numberings for which the singleton models always determine the sophistication.
-This means that for any dataset, the sophistication is equal to the Kolmogorov complexity.
+\subsection{Underfitting}
 
-Many variants in the literature are susceptible to this problem, including the structure function \cite{cover1985kolmogorov,gacs2001algorithmic}.
-\cite{vitanyi2004meaningful,adriaans2012facticity} avoid this issue by not using self-delimiting representations.
-Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically less efficient than representations placing some information in the input.
-While this side-steps the issue of the singletons, there is no suggestion that this makes the sophistication invariant.
-
-Another approach may be to make the constant in the threshold function dependent on the choice of enumeration.
-We are not aware of any approach to this effect, and there does not seem to be a simple way to determine the threshold based on the properties of the numbering.
-
-\section{Underfitting: the universal model}
-\label{section:underfitting}
-On the other end of the spectrum, there exist candidate representations which place all information in the `noise part', using a universal function as the model.
-We can easily show that if such models exist, we can choose our numbering so that they are always chosen.
-
-\begin{lemma}[underfitting]
-Let $\M$ be a model class containing model $u$, called a \emph{universal model} with $\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p$.
-Then, there exist numberings such that $\s^\M$ is bounded.
-\end{lemma}
+\underfitting*
 \begin{proof}
@@
 \end{proof}
-Most sophistication measures avoid the underfitting problem by limiting their model class, usually to either finite sets ($\F$) or total functions ($\T$ or $\P$).
-This eliminates the universal model, but as we shall see in the next section, this does not solve the problem for most data.
-The variants that are susceptible to the this type of underfitting are facticity \cite{adriaans2012facticity} and some versions of effective complexity \cite{gellmann1996information}.
-The latter aims to bypass the problem by selecting carefully from the candidate set, but as we have shown, for some numberings the candidate set contains only the universal model.
 
@@
 
-We have seen that some variants of sophistication sidestep the simplest examples of underfitting by limiting the model class to exclude universal models.
-While this eliminates the simplest proof of boundedness, it does not follow immediately that the sophistication based on a limited model class must be bounded.
-In fact, a well-known result in algorithmic statistics shows that there exist strings for which no finite set is as good a model as the singleton set.
-In this section we will focus on the variants $S^\T$ and $S^T_K$ \cite{vitanyi2004meaningful}, as those are the variants most resistant to proofs of non-invariance.
-
-\begin{lemma}
-There are infinitely many $x$ such that $S^\F_K(x) \geq K(x)/5$. \label{lemma:structure-function-is-not-bounded}
-\end{lemma}
+\nonstochastic*
 \begin{proof}
@@
 
-\begin{lemma}
-Let $f$ be a total, computable function, such that for some $d$, $f(d) = x$ and let $k = K(f) + |d|$.
-Then there exists a finite set $S$ containing $x$ and a constant $c$, such that $K(S) \leq K(f) + K(|d|) + c$ and $\log |S| \leq |d|$.\label{lemma:total-to-sets}
-\end{lemma}
+\vitanyi*
 \begin{proof}
@@
 \end{proof}
-This lemma is a variation on \cite[Lemma~7.2]{vitanyi2004meaningful}.
 
-We use this lemma to transport the result of Theorem~\ref{lemma:structure-function-is-not-bounded} to the model class of total functions:
-
-\begin{theorem}
-There are infinitely many $x$ with 
-\begin{enumerate}
-  \item $S^\T_K(x) \geq K(x)/6$, {and}\label{eq:poezenvoer}
-  \item $S^\T(x) \geq K(x)/6$. \label{eq:hondevoer}
-\end{enumerate}
-\end{theorem}
+\dogfood*
 \begin{proof}
@@
 \end{proof}
+
 Note that we favoured simplicity of the proof over the strength of the bounds: the multiplicative constant $1/6$ can be improved with a more detailed treatment.
 
-This tells us that for the right strings, a sophistication so defined can truly grow to a significant part of the total description length.
-While this shows that sophistication, if defined with particular care, is not entirely vacuous, we can shed some light on exactly which strings end up with high sophistication, using the principle of \emph{depth}.
-We will use the following definition:
-
-\begin{definition}[depth]\belowdisplayskip=-12pt
-Let $U$ be some universal Turing machine, so that $U(\bar\imath y) = \phi_i(y)$.
-Let $U^t$ be a simulation of this machine, which is allowed to run for at most $t$ steps, and returns $0$ if it has not yet finished at that point.
-Let $C^t_\M(x) = \min\{|\bar\imath y| : U^t(\bar\imath y) = x, \phi_i \in \M\}$.
-
-The \emph{$c$-depth} $d^\M_c(x)$ of a string is defined as:
-\[
-	d^\M_c(x) = \min \left\{t : C^\M_t(x) - C^\M(x) \leq c \right\}
-\] 
-\end{definition}
-Deep strings are those that can only be optimally compressed with a great investment of time.
-The notion of depth was first proposed by Bennett \cite{bennett1988logical} and later refined to variants of this form \cite{antunes2006computational}.
-We note that it is exceedingly unlikely that a deep string is sampled from a shallow distribution \cite{bloem2014safe,bennett1988logical}.
-
-\begin{theorem}
-Let $A(n)$ be the single-argument Ackermann function and $c$ some arbitrary constant.
-There are numberings such that for all strings with depth $d^\C_c(x) \leq A(C(x))$, $\s^\T(x)$, $s^\P(x)$ and $\s^\F$ are bounded.
-\end{theorem}
+\depth*
 \begin{proof}
@@
 
-This shows that while high-sophistication strings exist for variants with limited model class, the mechanics do not follow the basic intuition of sophistication.
-Unless we encounter data that would take longer than the lifetime of the universe (and then some) to produce, sophistication behaves exactly as it would without the restricted model class: a solution to a super-exponential problem would have no greater sophistication than a string sampled from a simple Bernoulli model.
-This result also suggests that the `non-stochastic' property of strings with high sophistication \cite{shen1983concept,vereshchagin2004kolmogorov} says more about depth and totality than it does about the lack of randomness in its information content.
+\section{Coarse sophistication}
 
-The relation between sophistication, depth and absolutely non-stochastic strings was also investigated in \cite{antunes2013sophistication}.
-
-\section{Conclusion}
-\label{section:conclusion}
-As we have seen, the idea behind sophistication has a rich history.
-It has been proposed many times, by many different authors.
-This tells us two things.
-Firstly, there is a strong intuition behind sophistication, shared by many respected authors, and second, it is not easy to get the definition exactly right.
-
-So why should we consider this intuition reasonable at all?
-The first assumption of sophistication is that high compressibility and pure randomness are objectively `uninteresting'.
-A more sophisticated signal, such as a television broadcast must contain, in some objective sense, more meaningful information.
-Whether this intuition seems reasonable likely differs per person, but sophistication goes one step further: it also suggests that we can objectively separate the information in a string into random and structured.
-
-To see whether this intuition is valid; imagine that we sample from the universal distribution by feeding random bits to a universal Turing machine.
-If this UTM is constructed in the conventional manner, the initial bits will determine the index of a partial computable function $i$, and the rest will be the input to that machine $y$.
-What's the characteristic model of this data?
-Certainly, we cannot exclude the possibility that $\phi_i$ is also a universal model.
-Also, we cannot exclude the possibility that $\phi_i$ is a singleton model for $x$.
-
-This kind of perspective, with a probabilistic source for the data is strongly discouraged by the proponents of algorithmic statistics: they often argue that information theory should precede probability theory \cite{kolmogorov1983combinatorial}.
-Nevertheless, this sampling scenario undercuts the intuition behind sophistication: there is no reason to assume that a string with high sophistication was sampled from a complex model, so how can the complex model be intrinsic to the data, moreso than any other model that reaches the Kolmogorov complexity?\footnotemark
-
-\footnotetext{Note that computational depth and Kolmogorov complexity do not have this problem.
-The properties of a distribution allow us to put bounds on both the depth and the complexity we can expect from strings produced by it.}
-
-For this reason, we take a skeptical view of sophistication.
-The matter of its invariance, in particular cannot be glossed over.
-It should be the first question asked of any proposal.
-We can offer the following insight, to back up our skepticism.
-As we have seen, candidates within a constant of the optimal representation can occur all over the spectrum, from small to large models.
-A small change in construction, such as a different acceptable numbering, or a change in threshold function, can arbitrarily push models into or out of the candidate set.
-Whereas in Kolmogorov complexity, constant changes have constant effects on the outcome, here constant changes can cause very large changes in the outcome.
-While no authors have touched on the issue of invariance explicitly, it is strongly related to the issue of instability, discussed in \cite{antunes2013sophistication,vereshchagin2013algorithmic}.
-
-The most common approach, restricting the models to be total, is problematic.
-As we have seen, the restriction to a model class $\M$ tends to assign high sophistication to those strings that can be compressed significantly better by models outside $\M$ than by models within it.
-For the class $\T$, this results in only extremely deep strings getting high sophistication.
-While this is an interesting property, which is worth investigating, the interpretation of an objective separation into model and noise is spurious at best.
-Instead of interpreting these strings as ``non-stochastic'', we consider it more natural to consider them \emph{non-typical for $\M$}.
-
-Finally, we note that the use of total functions has a very subtle effect on the discussion surrounding sophistication.
-Since the total functions are not enumerable, the value of the sophistication relies strongly on the difference between the Kolmogorov complexity and the Kolmogorov complexity bounded to total functions.
-This gap is exceedingly difficult to capture, because the total functions cannot be enumerated.
-
-This is dangerous: if such definitions cannot be proved wrong or right, they will stand as the ad-hoc official view, while the restrictions used to elude proofs of incorrectness make them very difficult to build on.
-This creates an artificial dead-end for a very valuable area of research.
-
-\subsection{A world without sophistication}
-
-It is too early to say that sophistication is an impossible ideal, but the points we have raised certainly show that it is riddled with subtle pitfalls, and inspired by questionable intuitions.
-We will conclude by asking what it would mean if we were to abandon this ideal entirely.
-
-Consider the following metaphor.
-We are given a bitstring which can be read as a bitmap image of the painting \emph{Impression of a Sunrise}.
-The candidate set for this string contains various models from very generic to very specific.
-The theory behind sophistication suggests that we can choose one of these as the objective, intrinsic model of the data.
-Different models say different things: the universal models says that it is a `thing'.
-A more specific model might say that it is an image.
-Even more specific would be a painting, a Monet, or specifically the painting Impression of a sunrise.
-These are the models we imagine we may find in the candidate set.
-Can we say that the data is intrinsically more of an image than a painting?
-More of a Monet than a painting?
-We see no intuitive reason why such a distinction should be possible.
-
-Of course, Kolmogorov complexity does allow us to say that it is unlikely to be a Jackson Pollock painting, or a piece of music.
-And, if we were to get a second sample from the same source, we \emph{could} answer the question: we would receive either the same painting again, or perhaps another Monet, and we would gain information about the model.
-But a single sample by itself should give us no reason to choose an intrinsic model.
-Ultimately, this is just a metaphor, but since the notion of sophistication is driven by intuition, we feel that an intuition to the contrary should serve to sharpen the debate.
-
-Our conclusion is that in giving up sophistication, we do not lose algorithmic statistics, but rather clean up some of its more unmanageable components.
-Letting go of the minimal sufficient statistic will allow us to return to the simplicity of $\K$ as model classes.
-This gives us a simple framework, with a one-to-one relation to Bayesian statistics and MDL.
-
-\subsubsection*{\ackname}
-
-This publication was supported by the Dutch national program COMMIT and by  the Netherlands eScience center.
-
-\bibliographystyle{plain}
-\bibliography{facticity}
-
-\appendix
+\pb{TODO}
 

### 2015-01-19 Facticity/Facticity/paper.tex
aant lezen

--- 2015-01-18 Facticity/Facticity/paper.tex
@@
 \newcommand{\s}{S}
-\newcommand{\num}[1]{#1_1, #1_2, \ldots}
-
-\newcommand{\phiv}{\mathbf\phi}
-\newcommand{\psiv}{\mathbf\psi}
+\newcommand{\dom}[1]{\mathop{\tn{dom}(#1)}}
 
@@
 
-% --- DELETE BEFORE SUBMUSSIONS ---
+% --- DELETE BEFORE SUBMISSIONS ---
 \pagestyle{headings}
@@
 
+\sdr{Course sophistication handled sufficiently?}
+\sdr{Do we really need the $\gamma$-invariance? Or can $\gamma$ be 1?}
+\sdr{We need the definitions! But maybe we can simplify, do we really need the candidate set?}
+
 Kolmogorov complexity breaks away from the tradition in statistics and information theory by allowing the analysis of \emph{individual} binary sequences rather than studying large samples from probability distributions in an assumed class.
@@
 In this paper, we investigate three problems with these definitions, and show that they are unsatisfactory.
-We believe the problem is fundamental, and provide several intuitive arguments why these attempts to define sophistication must fail.
+We believe the problem is fundamental, and provide arguments why such attempts to define sophistication must fail.
 
@@
 
-The Kolmogorov complexity of a binary string $x$ is, roughly, the length of the shortest computer program that will output $x$.
-This length depends on which programming language is used, but only by a constant, independent of $x$.
+The Kolmogorov complexity of a binary string $x$, denoted $K(x)$, is (roughly) the length of the shortest computer program that will output $x$.
+This length depends on which programming language is used, but, as shown by a core result known as the Invariance Theorem, the language affects the result only by a constant, independent of $x$.
 This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings.
@@
 In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language.
-Such interpreters, or \emph{universal models}, can be described in a constant amount of bits, and can generate the data with from a program the length of the Kolmogorov complexity, essentially encoding all information as noise.
-If complex models are penalized to avoid overfitting, then the problem becomes to make sure that universal models are not \emph{always} preferred for complex data.
+Such interpreters, or \emph{universal models}, can be described in a constant amount of bits, and can generate the data with an input of the  the length of the Kolmogorov complexity, essentially encoding all information as noise.
+If complex models are penalized to avoid overfitting, then the problem becomes to make sure that simple universal models are not \emph{always} preferred for complex data.
 The usual workaround is to restrict the set of allowed models to exclude the universal ones.
 For example, the model class may be restricted to the set of all total recursive functions (only programs that halt for all inputs) rather than all partial recursive functions (all programs).
-While this patches the problem of universal models to an extent, it is questionable whether it adequately solves the problem of underfitting in general.
+While this patches the problem with universal models, it is questionable whether it adequately solves the problem of underfitting in general.
 
 Finally, in the discussion in Section~\ref{section:conclusion} we argue that while two-part coding can yield useful insights into the structure of the data and allows us to identify some models as unlikely to capture all structure in the data, it is probably not possible to uniquely separate structure from noise and identify a \emph{single} model as ``best'': in general many models, of vastly different complexities, may be reasonable models of the data generating mechanism.
-Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that sometimes the data allows for multiple equivalent interpretations, and that there is no such thing as sophistication.
+Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that sometimes the data allows for multiple interpretations, and that there is no such thing as sophistication.
 
@@
 Koppel \cite{koppelSoph1988,koppel1991almost} also built on Kolmogorov and Cover's work to define a measure of the meaningful information in a binary string, called \emph{sophistication}, the term we adopt in this paper.
-He changed the class of models from finite sets to total functions and used monotone functions, instead of prefix-free Kolmogorov complexity, to be able to extend the theory to infinite strings, an approach which has been abandoned by subsequent authors.
+Koppel changed the class of models from finite sets to total functions and used monotone functions, instead of prefix-free Kolmogorov complexity, to be able to extend the theory to infinite strings, an approach which has been abandoned by subsequent authors.
 We will only consider the finite version (based on regular functions) in this paper.
@@
 Vereshchagin \cite{vereshchagin2013algorithmic} proposes an alternative solution to the instability issue.
+\sdr{But if his stuff works, this paper is nonsense, so we should say more about it!}
 
@@
 Such functions are \emph{computable} if they can be computed by a Turing machine as described in \cite[Definition~1.7.1]{li1993introduction}.
-A function is prefix-free if its domain, $D_f = \{y : f(y) \neq \infty\}$ is a prefix set.
-A function $f$ is \emph{total} if $D_f = \B$.
-A prefix-free function is \emph{complete} if every infinite binary string has a member of $D_f$ as a prefix.
-\footnotemark
+A function is prefix-free if its domain, $\dom{f} = \{y : f(y) \neq \infty\}$ is a prefix free set.
+A function $f$ is \emph{total} if $\dom{f} = \B$.
+A prefix-free function is \emph{complete} if every infinite binary string has a member of $\dom{f}$ as a prefix.\footnotemark
+
+\sdr{I think this footnote should not be a footnote and better explained. It's important}
 
@@
 \begin{definition}[Numberings]
-A numbering is an enumeration of of the partial computable functions, denoted as $\phi_1, \phi_2, \ldots$.
-We fix one canonical, which we know to be an effective description, that is for any $i \in \N, y\in \B$, we can effectively compute $\phi_i(y)$, where the computation diverges if $\phi_i(y) = \infty$.
-We call a numbering $\num{\psi}$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ such that for all $i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{b(i)}$.
-We use the greek letters without a subscript (eg.
-$\phi$ and $\psi$) to refer to the numbering as a whole.
+A numbering $\psi_1,\psi_2,\ldots$ is an enumeration of of the partial computable functions.
+We fix one canonical numbering $\phi$ that we know to be effective, that is for any $i \in \N, y\in \B$, we can effectively compute $\phi_i(y)$, where the computation diverges if $\phi_i(y) = \infty$.
+We call a numbering $\psi$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ such that for all $i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{a(i)}$.
+We abbreviate the sequence by its symbol without the index: $\phi=\phi_1,\phi_2,\ldots$
 \end{definition}
@@
   $|g(x)|\le|x|+2\log(|x|+2)|$ \cite{li1993introduction}. We fix such a function and
-  denote it by $\br{x}=g(x)$.
+  denote it using a bar, $\br{x}=g(x)$.
 \end{definition}
 
+\sdr{Clearly explain what it means if a function function is total, complete, prefix.
+Do we need to fuss about equating the numbers and binary strings?
+Perhaps it's better to do the conversions explicitly in this paper}
+
 \begin{definition}[Complexity]\belowdisplayskip=-12pt
-Let $\phi_1,\phi_2,\ldots$ be the canonical acceptable enumeration of $\C$.
-Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ corresponds to the set of all computable prefix functions.
-We define the plain and prefix Kolmogorov complexity as
+Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ is the set of all computable prefix functions.
+\sdr{Cite textbook that this is possible.} We define the plain and prefix Kolmogorov complexity as
 \begin{align*}
@@
 This definition of $K$ is uncommon.
-It is equivalent to the standard definition as used in \cite{li1993introduction}, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).
+It is equivalent to the standard definition in \cite{li1993introduction}, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).
 
@@
 
-\begin{definition}[Models and model classes] A \emph{model class} is a set of models. We distinguish four important model classes:
+\begin{definition}[Models and model classes] A \emph{model class} is a set of models. We distinguish five important model classes:
   \begin{itemize}
@@
 We say that a formalization of sophistication $\s_\phi$ built on acceptable numbering $\phi$ is $\gamma$-invariant to the choice of numbering is $s_\phi$ and $s_\psi$ are $\gamma$-equal for any two acceptable numberings $\phi$ and $\psi$.
-\begin{lemma}
-If a function $\S_\phi(x)$ is bounded for a given numbering, it is either not $\gamma$-invariant or it is  $\gamma$-equal to a bounded function.
+\begin{lemma}\label{lem:broken}
+Given a slack function $\gamma$, if $\s_\phi(x)$ is either $\gamma$-equivalent to $0$ or to $K(x)$ for \emph{some} numbering $\phi$, then either this equivalence exists for all other numberings as well, or $\s_\phi(x)$ is not $\gamma$-invariant.
 \end{lemma}
@@
 \label{section:indices}
-We start from a simple intuition: the definition of Kolmogorov complexity already uses a description of the data in two parts; a computable function and an input.
+We start from a simple intuition: the definition of Kolmogorov complexity already uses a description of the data in two parts; the index of a computable function and an input to that function.
 Can we assume that, when such a description is as short as the Kolmogorov complexity, it automatically creates a good separation into the structural part of the data and the noise?
-As we will show, the representation of function plays a fundamental part in answering this question.
+As we will show, this depends in part on the used enumeration $\phi$.
 
-The issues can be very subtle and require a careful attention that is rarely given in the literature. We will first illustrate what happens if any acceptable numbering is allowed:
+The issue is subtle and require a careful attention that is rarely given in the literature. We will first illustrate what happens if any acceptable numbering is allowed:
 
 \begin{lemma}
-Let $\phi$ be some acceptable numbering, and $\s_\phi(x)$ be any function which gives the length of some index $i$ such that $\phi_i(y) = x$ for some $y$.
-
-No such function can be both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
+Any sophistication $\s_\phi(x)$ that gives the length of some index $i$ such that $\phi_i(y) = x$ for some $y$ cannot be both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
 \end{lemma}
@@
 
-Let $\num{\phi}$ be the canonical acceptable numbering. Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.
+Let $\phi$ be the canonical acceptable numbering. Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.
 
-We now define two new acceptable numberings $\num{\psi}$ and $\num{\xi}$:
+We now define two new acceptable numberings $\psi$ and $\chi$:
 \[
@@
 \]
-Let $\s_\psi$ and $\s_\xi$ be the sophistication functions built on these enumerations.
-Choose any $x$ and assume w.l.o.g.
-that $\s_\psi(x) < \s_\xi(x)$.
-By construction of the numberings $\psi$ and $\xi$, we have $2\;\s_\psi(x)  \leq \s_\xi(x)$, which gives us:
+Now choose any $x$ and assume w.l.o.g. that $\s_\psi(x) < \s_\xi(x)$. By construction of the numberings $\psi$ and $\xi$, we have $2\;\s_\psi(x)  \leq \s_\xi(x)$, which gives us:
 \begin{align*}
@@
 
-The treatments that fall prey to this problem are Koppel and Atlan's \cite{koppelSoph1988,koppel1991almost},  and several of its variants (\cite{antunes2009sophistication,antunes2013sophistication}).
+This is a serious problem with several papers, in particular
+Koppel and Atlan's \cite{koppelSoph1988,koppel1991almost} and several of its variants \cite{antunes2009sophistication,antunes2013sophistication}.
 
@@
 \begin{lemma}[Invariance]
-Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be any two acceptable numberings.
-There exists a constant $c$ such that $\left| K_\phi(f) - K_\psi(f)\right | \leq c$ for all $f$.
-\label{lemma:invariance}
+Let $\phi$ and $\psi$ be any two acceptable numberings. There exists a constant $c$ such that $\left| K_\phi(f) - K_\psi(f)\right | \leq c$ for all $f$. \label{lemma:invariance}
 \end{lemma}
 \begin{proof}
-Let $g(i)$ be the function such that $\psi_i$ to $\phi_{g(i)}$.
+Let $g(i)$ be the function such that $\psi_i=\phi_{g(i)}$.
 \begin{align*}
@@
 &= \min\left\{ K_\psi(i) : \psi_{g(i)}= f\right\} - c
-\geq \min\left\{ K_\psi(i) : \psi_i= f\right\} - c = K_\psi(f).
+\geq \min\left\{ K_\psi(i) : \psi_i= f\right\} - c' = K_\psi(f).
 \end{align*}
@@
   \item We maintain that the representation witnessing the Kolmogorov complexity determines the sophistication, but impose the additional constraint that an efficient, acceptable numbering is used.
-Specifically, that the length of the indices satisfies $|i| = C(\phi_i)$.
+Specifically, that for every function $f$ there is an $i$ with $\phi_i=f$ and $|i|=C(f)$.
 We construct such an index below.
@@
   \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use Kolmogorov complexity to measure the complexity of the \emph{model} (as in Definition~\ref{definition:model-complexity}), which is then the first part of a two-part code describing the data.
-With this approach, the total code length is $K(\phi_i)+|y|$, where $y$ is an input to the model $\phi_i$ such that $\phi_i(y)=x$.
-The difference is that in this approach, we can construct everything on the basis of a single acceptable numbering, without additional constraints.
+With this approach, the total code length is $K(f)+|y|$, where $y$ is an input to the model $f$ such that $f(y)=x$.
+The difference is that in this approach, we do not need any constraints on the default numbering $\phi$.
 This approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information}.
 \end{enumerate}
-To show that the two perspective are equivalent, we must show that there exists an acceptable numbering which is efficient enough.
+To show that the two perspectives are equivalent, we must show that there exists an acceptable numbering which is efficient enough.
 
 \begin{definition}[Faithful Numbering]\label{def:faithful}
-  A numbering $\psi_1,\psi_2,\ldots$ of the partial recursive
+  A numbering $\psi$ of the partial recursive
   functions is \emph{faithful} if there is a constant $c$ such that
@@
 \end{definition}
-This definition is a refinement of the one used in \cite{adriaans2012facticity}. \footnotemark
+This definition is a refinement of the one used in \cite{adriaans2012facticity}.\footnotemark
 
@@
 \end{proof}
-Unfortunately, even with a faithful index, we can fail if the prefix-function used to separate the model from its input is inefficient:
+\sdr{Explain following better: point out that it provides a slight imbalance in favour of putting stuff in the data part, which, if the model class is $\C$, already to underfitting.
+Refer to the underfitting section.}
+Even with a faithful index, we can fail if the prefix-function used to separate the model from its input is inefficient:
 \begin{lemma}
 Let $L(\phi_i) = |\bar\imath|$ for a faithful index, and some prefix encoding function, and let $\s^\C$ be defined on this $L$ with the default threshold function.
+\sdr{Dit wringt als een gek om t zo te zeggen, moet eleganter.
+Volgens mij gebruiken we maar twee $L$ functies: de index met een bar erop, en $K$.}
 
@@
 This result applies to Adriaans' facticity\cite{adriaans2012facticity}.
-In order to circumvent the problem, the representation should should encode the model $f$ in $K(f)$ bits, relying on $K$'s own prefix encoding to delimit the model from its input.
+\sdr{I think we should not talk about circumventing this problem, it is too small an imbalance.
+Rather we should point out from the start that the difference in cost for the model $|\bar\imath|$ instead of $K(i)$ is unbounded, so even a faithful model index is still not an optimal description, and the gap is a direct consequence of the definition of Kolmogorov complexity.
+And that this slight imbalance can already lead to underfitting.
+We can then say that the model does not occur if we use the other two-part coding approach, which is why we'll adopt that other perspective in the rest of the paper.} In order to circumvent the problem, the representation should encode the model $f$ in $K(f)$ bits, relying on $K$'s own prefix encoding to delimit the model from its input.
 
@@
 We now come to the main issue plaguing sophistication.
-In the naive setting, where we allow all prefix-free functions as models, and use the default threshold function, the candidate set always contains two representations that clearly do not provide a good separation into structure and content.
+\sdr{Beetje bombastisch.
+En hierna zou ik zeggen, we first consider the simplest, most natural setting, \dots} In the naive setting, where we allow all prefix-free functions as models, and use the default threshold function, the candidate set always contains two representations that clearly do not provide a good separation into structure and content \sdr{noise?}.
 The first is a \emph{universal model}: a universal Turing machine which simply runs its shortest program for $x$.
@@
 
-The key problem is that both can be present in the candidate set, and the formalization of sophistication must ensure that neither are selected.
+The key problem is that both can be present in the candidate set, and the formalization of sophistication must ensure that neither are selected \sdr{well occasionally they may be selected\dots should be more precise here}.
 As we will show, in many settings, we can ensure that they are the \emph{only} models in the candidate set.
@@
 \begin{lemma}\label{lemma:thecoolone}
-  Let $\psi_1,\psi_2,\ldots$ be any acceptable enumeration of the partial recursive functions.
-  Let $\M$ be any model class, let $\X$ be any set of binary sequences and let $D:\{0,1\}^*\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e.
+  Let $\psi$ be an acceptable enumeration of the partial recursive functions.
+  Let $\M$ be any model class, let $\X$ be any set of binary sequences and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e.
 if $f=\psi_{D(p)}$ then $p$ is a $D$-description of $f$.
@@
 \end{enumerate}
-Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
+Then there is an enumeration $\phi$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
 \end{lemma}
@@
 
+\sdr{Move this proof to appendix. Explain the lemma above it rather than below it.}
 This is a technical lemma, due to the double effect of the numbering, which occurs in both the definition of model complexity and the definition of data complexity.
@@
 
-\begin{lemma}[overfitting]
+\begin{lemma}[Overfitting]
 Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$.
-Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the prefix partial recursive functions, and a constant $c$, such that
+Then there is an enumeration $\phi$ of the prefix partial recursive functions, and a constant $c$, such that
 \[
@@
 \begin{proof}
-Let $\psi_1,\psi_2,\ldots$ be any default enumeration of the partial recursive prefix functions.
+Let $\psi$ \sdr{This should probably be $\phi$} be any default enumeration of the partial recursive prefix functions.
 Note that since $f$ is a prefix function, if $f$ is defined for input $\epsilon$ then it cannot be defined for any other input.
@@
 In short, this lemma tells us that there are numberings for which the singleton models always determine the sophistication.
-This means that for any dataset, the sophistication is equal to the Kolmogorov complexity.
+This means that for any dataset, there is a numbering $\phi$ such that the sophistication is equivalent to Kolmogorov complexity, which implies that this sophistication is unsound (by Lemma~\ref{lem:broken}).
 
-Many variants in the literature are susceptible to this problem, including the structure function \cite{cover1985kolmogorov,gacs2001algorithmic}.
-\cite{vitanyi2004meaningful,adriaans2012facticity} avoid this issue by not using self-delimiting representations.
-Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically less efficient than representations placing some information in the input.
-While this side-steps the issue of the singletons, there is no suggestion that this makes the sophistication invariant.
+Many variants in the literature are susceptible to this problem, \sdr{Don't say ``many'', just list explicitly} including the structure function \sdr{Type error, the structure function is not a sophistication} \cite{cover1985kolmogorov,gacs2001algorithmic}.
+In \cite{vitanyi2004meaningful,adriaans2012facticity} this issue is avoided by not using self-delimiting representations.
+\sdr{Prep the balance idea so that this is understandable.
+We need it} Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically less efficient than representations placing some information in the input.
+While this side-steps the issue of the singletons and may be a sufficient measure against overfitting, it is not clear that it helps make the sophistication invariant.
 
-Another approach may be to make the constant in the threshold function dependent on the choice of enumeration.
+Another approach may be to make the constant in the threshold function \sdr{Nice point but very hard to understand the way its done now since we didn't really rub the framework in the reader's face} dependent on the choice of enumeration.
 We are not aware of any approach to this effect, and there does not seem to be a simple way to determine the threshold based on the properties of the numbering.
@@
 On the other end of the spectrum, there exist candidate representations which place all information in the `noise part', using a universal function as the model.
-We can easily show that if such models exist, we can choose our numbering so that they are always chosen.
+If such models are allowed, we can choose our numbering so that they always determine the sophistication.
 
-\begin{lemma}[underfitting]
-Let $\M$ be a model class containing model $u$, called a \emph{universal model} with $\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p$.
+\begin{lemma}[Underfitting]
+Let $\M$ be a model class containing model a universal model $u$, with the property that $\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p$.
 Then, there exist numberings such that $\s^\M$ is bounded.
@@
 That is, $\M' = \{u\}$.
-The assumptions 1 and 2 from Lemma~\ref{lemma:thecoolone} follow directly from this construction, so that we can instantiate the Lemma.
-This tells us that there exists an acceptable numbering for which $\s_\M(x) = \s_{\M'}(x) + c$.
-Since $\M'$ contains only a single model, $\s_{\M'}$ must be constant, completing the proof.
+This construction satisfies the conditions 1 and 2 from Lemma~\ref{lemma:thecoolone}; invoking it we find that there exists an acceptable numbering for which $\s_\M(x) = \s_{\M'}(x) + c$.
+Since $\M'$ contains only a single model, $\s_{\M'}$ is constant.
 \end{proof}
+The variants that are susceptible to the this type of underfitting are facticity \cite{adriaans2012facticity} and some versions of effective complexity \cite{gellmann1996information}.
+The latter aims to bypass the problem by selecting carefully from the candidate set, but as we have shown, for some numberings the candidate set contains only the universal model.
+
+\sdr{Explain that this is actually a widely recognised problem}
+
 Most sophistication measures avoid the underfitting problem by limiting their model class, usually to either finite sets ($\F$) or total functions ($\T$ or $\P$).
 This eliminates the universal model, but as we shall see in the next section, this does not solve the problem for most data.
-The variants that are susceptible to the this type of underfitting are facticity \cite{adriaans2012facticity} and some versions of effective complexity \cite{gellmann1996information}.
-The latter aims to bypass the problem by selecting carefully from the candidate set, but as we have shown, for some numberings the candidate set contains only the universal model.
 
@@
 We have seen that some variants of sophistication sidestep the simplest examples of underfitting by limiting the model class to exclude universal models.
-While this eliminates the simplest proof of boundedness, it does not follow immediately that the sophistication based on a limited model class must be bounded.
-In fact, a well-known result in algorithmic statistics shows that there exist strings for which no finite set is as good a model as the singleton set.
+While this eliminates the simplest proof of boundedness, it does not follow immediately that the sophistication based on a restricted model class must be unbounded.
+However, a well-known result in algorithmic statistics shows that there are strings for which no finite set is as good a model as the singleton set.
 In this section we will focus on the variants $S^\T$ and $S^T_K$ \cite{vitanyi2004meaningful}, as those are the variants most resistant to proofs of non-invariance.
+\sdr{Do I know this notation?}
 

### 2015-01-19 Facticity/Facticity/paper.tex

--- 2015-01-19 Facticity/Facticity/paper.tex
@@
 
+\sdr{Course sophistication handled sufficiently?}
+\sdr{Do we really need the $\gamma$-invariance? Or can $\gamma$ be 1?}
+\sdr{We need the definitions! But maybe we can simplify, do we really need the candidate set?}
+
 Komogorov complexity gives us a sound definition of the amount of information contained in a bitstring.
@@
 This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication},\footnotemark were proposed.
-In this paper, we investigate three serious problems with these definitions.
-We believe that ultimately, some of these issues are insurmountable, and provide several intuitive arguments why any attempt to define sophistication must fail.
+In this paper, we investigate three serious problems with these definitions, and show why they are unsatisfactory.
+We believe that the problems are fundamental, and provide arguments why such attempts to define sophistication must fail.
 
-The Kolmogorov complexity of a binary string $x$ is, roughly, the length of the shortest computer program that will output $x$.
-This length depends on which programming language is used, but only by a constant independent of $x$.
-This property is known as \emph{invariance}.
-This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and we can see the Kolmogorov complexity  as \emph{a property of the data}.
+The Kolmogorov complexity of a binary string $x$, denoted $K(x)$, is (roughly) the length of the shortest computer program that will output $x$.
+This length depends on which programming language is used, but, as shown by a core result known as the Invariance Theorem, the language affects the result only by a constant, independent of $x$.
+This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings.
+We can see it as \emph{a property of the data}.
 
-A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees. We are interested in a measure satisfying the following requirements:
+A robust definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees. We are interested in a measure satisfying the following requirements:
 
@@
 Underfitting (Section~\ref{section:underfitting}) occurs when the selected model is simple, but fails to represent some of the structure present in the data.
+This is a problem for sophistication because the considered models can be so powerful.
 In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language.
-Such \emph{universal models} can be described in a constant amount of bits, and  represent the data with a program the length of the Kolmogorov complexity, essentially encoding all information as noise.
+Such \emph{universal models} can be described in a constant amount of bits, and  represent the data with an input of the length of the Kolmogorov complexity, essentially encoding all information as noise.
 If complex models are penalized, then the problem becomes to make sure that universal models are not \emph{always} preferred for complex data.
@@
 
-\subsection{Notation}
-
+\section{Preliminaries}
 Let $\B = \{0,1\}^*$.
@@
 
+\sdr{I think this completeness vs totalness should be explained much better, its important because without it the differences between the settings cannot be explained}
+
 \begin{definition}[Numberings]
-A numbering is an enumeration of of the partial computable functions, denoted as $\phi_1, \phi_2, \ldots$ or simply $\phi$.
-We fix one effective numbering $\phi$.
-We call a numbering $\num{\psi}$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ with $\forall: i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{b(i)}$.
+
+\begin{definition}[Prefix encoding function]
+  A prefix encoding function $f:\B\to\B$ is an injective
+  function that maps the binary strings to a prefix-free set. There
+  are prefix encoding functions $g$ with the property that
+  $|g(x)|\le|x|+2\log(|x|+2)|$ \cite{li1993introduction}. We fix such a function and
+  denote it using a bar, $\br{x}=g(x)$.
+\end{definition}
+
+\sdr{Clearly explain what it means if a function function is total, complete, prefix.
+Do we need to fuss about equating the numbers and binary strings?
+Perhaps it's better to do the conversions explicitly in this paper}
+
+We will generalize the main proposals for sophistication into a single notation.
+We start with a class of models, we consider all two-part descriptions of the data using these models, we select the set of representations with the shortest code-lengths and we select the smallest model from this set.
+The size of the smallest model represents the sophistication of the data.
+
+A numbering is an enumeration of of the partial computable functions, denoted as $\psi_1, \psi_2, \ldots$ or simply $\psi$.
+We fix one canonical effective numbering $\phi$.
+We call a numbering $\psi$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ with $\forall: i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{a(i)}$.
 \end{definition}
@@
 \begin{definition}[Complexity]
-Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ corresponds to the set of all computable prefix functions.
-We define the plain and prefix Kolmogorov complexity as $C(x)=\min\{|\bar\imath y|:\phi_i(y)=x\}$ and $K(x)=\min\{|\bar\imath y|:\phi_i(y)=x, i\in\Pi\}$ respectively.
+Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ corresponds to the set of all computable prefix functions.\sdr{Cite textbook that this is possible} We define the plain and prefix Kolmogorov complexity as $C(x)=\min\{|\bar\imath y|:\phi_i(y)=x\}$ and $K(x)=\min\{|\bar\imath y|:\phi_i(y)=x, i\in\Pi\}$ respectively.
 \label{definition:complexity}
 \end{definition}
+\sdr{Added back, liked it} This definition of $K$ is uncommon.
+It is equivalent to the standard definition in \cite{li1993introduction}, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).
 
 \begin{definition}[Model classes] A \emph{model class} is a set of models.
-We distinguish four important model classes: the partial (computable) functions ($\C$), the prefix functions ($\K \subset \C$), the total functions ($\T \subset \C$), the complete functions ($\P \subset \K$) and the finite sets ($\F \subset \P$, encoded as a prefix-free function mapping all strings of length $\lceil\log S\rceil$ to members of $S$).
+We distinguish five important model classes: the partial (computable) functions ($\C$), the prefix functions ($\K \subset \C$), the total functions ($\T \subset \C$), the complete functions ($\P \subset \K$) and the finite sets ($\F \subset \P$, encoded as a prefix-free function mapping all strings of length $\lceil\log S\rceil$ to members of $S$).
 \end{definition}
+
+\sdr{We used to define: A pair $(m,y)$ is a \emph{representation} for $x$ if $m$ is a model and $m(y)=x$.
+I don't think we can change this.
+Otherwise an element of $\F$ an index is not a representation.}
 
@@
 \-\hspace{1cm}$L^\M(x)=\min\{L^m(x):m\in\M\}$
+>>>>>>> 3ebd420d297744b1c9bdacf2ebd8051f62dcfd13
 \end{definition}
@@
 \end{definition}
-
-\begin{lemma}
-If a function $\s_\psi(x)$ is bounded for some $\psi$, it is either not $\gamma$-invariant or it is  $\gamma$-equivalent to a bounded function.
+\begin{lemma}\label{lem:broken}
+If $\s_\phi(x)$ is $\gamma$-equivalent to either $0$ or to $K(x)$ for \emph{some} numbering $\phi$, then either this equivalence exists for all other numberings as well, or $\s_\phi(x)$ is not $\gamma$-invariant.
 \end{lemma}
@@
 \label{section:indices}
+
 We start from a simple intuition: the definition of $K$ already uses a description of the data in two parts; a computable function and an input.
@@
 \begin{restatable}{lemma}{ineffindices}
-Let $\phi$ be some acceptable numbering, and $\s_\phi(x)$ be any function which gives the length of some index $i$ such that $\phi_i(y) = x$ for some $y$.
-
-No such function can be both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
+Any sophistication $\s_\phi(x)$ that returns the length of some index $i$ such that $\phi_i(y) = x$ for some $y$, cannot be both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
  \end{restatable}
+\sdr{Perhaps this proof should be here? It is informative}
 \noindent In the proof of this lemma (Section~\ref{section:appendix-inefficient-indices}), we deliberately used ridiculously inefficient index functions.
@@
   \item We maintain that the representation witnessing the Kolmogorov complexity determines the sophistication, but impose the additional constraint that an efficient, acceptable numbering is used.
-Specifically, that the length of the indices satisfies $|i| = C(\phi_i)$.
+Specifically, that for every function $f$ there is an $i$ with $\phi_i=f$ and $|i|=C(f)$.
+We construct such an index below.
 This approach is taken by Adriaans \cite{adriaans2012facticity}.
@@
 To show that the two perspectives are equivalent, we must show that there exists an acceptable numbering which is efficient enough:
-\begin{definition}[Faithful Numberings]\label{def:faithful}
-  A numbering $\num{\psi}$ is \emph{faithful} if
+\begin{definition}[Faithful Numbering]\label{def:faithful}
+  A numbering $\psi$ is \emph{faithful} if
   for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and
@@
 To avoid the problem, the model $\phi_i$ should be encoded in $K(\phi_i)$ bits, relying on $K$'s own prefix encoding to delimit the model.
-
+\sdr{The above is incomprehensible.
+Explain better why this can't be fixed; point out that it provides a slight imbalance in favour of putting stuff in the data part, which, if the model class is $\C$, already to underfitting.
+Prime the reader to think in terms of balance.
+Refer to the underfitting section.}
+\sdr{We should point out from the start that the difference in cost for the model $|\bar\imath|$ instead of $K(i)$ is unbounded, so even a faithful model index is still not an optimal description, and the gap is a direct consequence of the definition of Kolmogorov complexity.
+And that this slight imbalance can already lead to underfitting.
+We can then say that the model does not occur if we use the other two-part coding approach, which is why we'll adopt that other perspective in the rest of the paper.} In order to circumvent the problem, the representation should encode the model $f$ in $K(f)$ bits, relying on $K$'s own prefix encoding to delimit the model from its input.
 Note that the condition for this lemma likely applies to all possible computable  prefix encoding functions (see Section~\ref{section:prefix-inefficiency}).
@@
 In the following, we will take the second perspective: the representations are two-part descriptions $(i, y)$ such that $\phi_i(y)=x$, with codelength $K(\phi_i) + |y|$.
+
+\sdr{Moved this back here, the conjecture needs to be in the main text}
+Lemma~\ref{lemma:prefix-inefficiency} only holds if the prefix function $L(i)$ has the property that $\min\{L(i)-K(i):i\ge i_0\}$ is unbounded.
+We conjecture that this is the case for any computable function.
+
+We conjecture that it applies to all possible computable prefix encoding functions:
+\begin{conjecture}
+For any partial computable prefix codelength function $L$, the function $\min\{L(i)-K(i):i\ge i_0\}$
+is unbounded in $i_0$.
+\end{conjecture}
+
 \section{Overfitting: the singletons}
@@
 We now come to the main issue plaguing sophistication.
-If we use model class $\K$, the candidate set always contains two representations that do not provide a good separation.
+\sdr{Beetje bombastisch.
+En hierna zou ik zeggen, we first consider the simplest, most natural setting, \dots} If we use model class $\K$, the candidate set always contains two representations that do not provide a good separation.
 The first is a \emph{universal model}: a basic universal Turing machine.
@@
 \label{lemma:thecoolone}
-  Let $\psi_1,\psi_2,\ldots$ be any acceptable enumeration of the partial recursive functions.
-  Let $\M$ be any model class, let $\X$ be any set of binary sequences and let $D:\{0,1\}^*\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e.
+  Let $\psi$ be any acceptable enumeration of the partial recursive functions.
+  Let $\M$ be any model class, let $\X$ be any set of binary sequences and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e.
 if $f=\psi_{D(p)}$ then $p$ is a $D$-description of $f$.
@@
 \-\hspace{1cm}(2)$\forall_{x\in\X}:L^{\M',\psi}(x)-L^{\M,\psi}(x)\le c$.\\
-Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
+Then there is an enumeration $\phi$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
 \end{restatable}
@@
 Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$.
-Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the prefix partial recursive functions, and a constant $c$, such that
+Then there is an enumeration $\phi$ of the prefix partial recursive functions, and a constant $c$, such that
 \[
@@
 \noindent In short, this lemma tells us that there are numberings for which the singleton models always determine the sophistication.
-This means that for any dataset, the sophistication is equal to the Kolmogorov complexity.
+This means that there is a numbering $\phi$ such that the sophistication is equivalent to Kolmogorov complexity, which implies that this sophistication is unsound (by Lemma~\ref{lem:broken}).
 
-Many variants in the literature (including the structure function) are susceptible to this problem \cite{cover1985kolmogorov,gacs2001algorithmic,vereshchagin2013algorithmic}.
-\cite{vitanyi2004meaningful,adriaans2012facticity} avoid it by not using self-delimiting representations.
-Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically not optimally efficient.
+Many variants in the literature (including the structure function) are susceptible to this problem  \sdr{Don't say ``many'', just list explicitly} \cite{cover1985kolmogorov,gacs2001algorithmic,vereshchagin2013algorithmic}.
+In \cite{vitanyi2004meaningful,adriaans2012facticity} this issue is avoided by not using self-delimiting representations.
+\sdr{Prep the balance idea so that this is understandable.
+We need it}  Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically not optimally efficient.
 While this dodges the issue of the singletons, it does not show that $S$ is invariant.
@@
 \begin{restatable}[underfitting]{lemma}{underfitting}
-Let $\M$ be a model class containing model $u$, called a \emph{universal model} with $\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p$.
-Then, there exist numberings such that $\s^\M$ is bounded.
+Let $\M$ be a model class containing a universal model $u$, with the property that $\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p$. Then, there exist numberings such that $\s^\M$ is bounded.
 \end{restatable}
@@
 The latter aims to bypass the problem by selecting carefully from the candidate set, but as we have shown, for some numberings the candidate set contains only the universal model.
+\sdr{Explain this is actually a widely recognised problem}
 
@@
 In this section we will focus on the variants $S^\T$ and $S^T_K$ \cite{vitanyi2004meaningful}, as those are the variants most resistant to proofs of non-invariance.
-
+\sdr{Do I know this notation? Also, I wanna emphasize that the structure function underfits in this way.}
+\sdr{Combine this lemma with the one after next}
 \begin{restatable}{lemma}{nonstochastic}
@@
 
+\sdr{Move this lemma to the appendix altogether}
 \begin{restatable}{lemma}{vitanyi}
@@
 \end{definition}
+
 Deep strings are those that can only be optimally compressed with a great investment of time.
@@
 \begin{lemma}[Invariance of function complexity]
-Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be any two acceptable numberings.
-There exists a constant $c$ such that $\left| K_\phi(f) - K_\psi(f)\right | \leq c$ for all $f$.
-\label{lemma:invariance}
+Let $\phi$ and $\psi$ be any two acceptable numberings. There exists a constant $c$ such that $\left| K_\phi(f) - K_\psi(f)\right | \leq c$ for all $f$. \label{lemma:invariance}
 \end{lemma}
 \begin{proof}
-Let $g(i)$ be the function such that $\psi_i$ to $\phi_{g(i)}$.
+Let $g(i)$ be the function such that $\psi_i=\phi_{g(i)}$.
 \begin{align*}
@@
 enumeration by more than a constant term.
-
-\subsection{Existence of faithful numberings}
 
@@
 
-Let $\num{\phi}$ be the canonical acceptable numbering. Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.
+Let $\phi$ be the canonical acceptable numbering. Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.
 
-We now define two new acceptable numberings $\num{\psi}$ and $\num{\xi}$:
+We now define two new acceptable numberings $\psi$ and $\xi$:
 \[
@@
 
-\subsection{Inefficiency of computable prefix codes}
-\label{section:prefix-inefficiency}
-Lemma~\ref{lemma:prefix-inefficiency} only holds if the prefix function $L(i)$ has the property that $\min\{L(i)-K(i):i\ge i_0\}$ is unbounded.
-We conjecture that this is the case for any computable function.
-
-We conjecture that it applies to all possible computable prefix encoding functions:
-\begin{conjecture}
-For any partial computable prefix codelength function $L$, the function $\min\{L(i)-K(i):i\ge i_0\}$
-is unbounded in $i_0$.
-\end{conjecture}
-
 \subsection{Boosting a model class by manipulating the numbering}
@@
 \coolone*
+
 \begin{proof}
@@
 \overfitting*
+
 \begin{proof}
-Let $\psi_1,\psi_2,\ldots$ be any default enumeration of the partial recursive prefix functions.
+Let $\psi$ \sdr{This should probably be $\phi$} be any default enumeration of the partial recursive prefix functions.
 Note that since $f$ is a prefix function, if $f$ is defined for input $\epsilon$ then it cannot be defined for any other input.
@@
 \end{proof}
-
 \subsection{Underfitting}
@@
 \underfitting*
+
 \begin{proof}
@@
 That is, $\M' = \{u\}$.
-The assumptions 1 and 2 from Lemma~\ref{lemma:thecoolone} follow directly from this construction, so that we can instantiate the Lemma.
-This tells us that there exists an acceptable numbering for which $\s_\M(x) = \s_{\M'}(x) + c$.
-Since $\M'$ contains only a single model, $\s_{\M'}$ must be constant, completing the proof.
+This construction satisfies the conditions 1 and 2 from Lemma~\ref{lemma:thecoolone}; invoking it we find that there exists an acceptable numbering for which $\s_\M(x) = \s_{\M'}(x) + c$.
+Since $\M'$ contains only a single model, $\s_{\M'}$ is constant.
 \end{proof}
@@
 \nonstochastic*
+
 \begin{proof}

### 2015-01-19 Facticity/Facticity/paper.tex
nog wat verder gelezen

@@
 \noindent This tells us that $S^\T_K(x)$ and $S^\T(x)$ can truly grow to a significant part of the total description length.
-While this shows that sophistication, if defined with particular care, is not entirely vacuous, we can shed some light on exactly which strings end up with high sophistication, using the principle of \emph{depth}.
+While this shows that sophistication, if defined with particular care, is not entirely vacuous, \sdr{hold on, it might still be completely meaningless and/or not invariant} we can shed some light on exactly which strings end up with high sophistication, using the principle of \emph{depth}.
 We will use the following definition:
@@
 \noindent This shows that while high-sophistication strings exist, the mechanics do not follow the basic intuition of sophistication.
-Unless we encounter data that would take longer than the lifetime of the universe (and then some) to produce, sophistication behaves exactly as it would without the restricted model class: a solution to a super-exponential problem would have no greater sophistication than a string sampled from a simple Bernoulli model.
-This result also suggests that the `non-stochastic' property of strings with high sophistication\cite{shen1983concept,vereshchagin2004kolmogorov} says more about depth and totality than it does aboutrandomness and structure.
+Unless we encounter data that would take longer than the lifetime of the universe (and then some) to produce \sdr{Not sure about this philosophically, but lets discuss}, sophistication behaves exactly as it would without the restricted model class: a solution to a super-exponential problem \sdr{Bad example, use something that clearly does not require much computation but has a lot of structure.
+A midpoint-displaced heightmap comes to mind} would have no greater sophistication than a string sampled from a simple Bernoulli model.
+This result also suggests that the `non-stochastic' property of strings with high sophistication\cite{shen1983concept,vereshchagin2004kolmogorov} says more about depth and totality than it does about structure and noise.
 
@@
 
-\section{Conclusion}
+\section{Discussion and Conclusion}
 \label{section:conclusion}
@@
 This tells us two things.
-Firstly, there is a strong intuition behind sophistication, shared by many respected authors, and second, it is not easy to get the definition exactly right.
+First, there is a strong intuition behind sophistication, shared by many respected authors, and second, it is not easy to get the definition exactly right.
 
@@
 A more sophisticated signal, such as a television broadcast must contain, in some objective sense, more meaningful information.
-Whether this intuition seems reasonable likely differs per person, but sophistication goes one step further: it also suggests that we can objectively separate the information in a string into random and structured.
+Whether this intuition seems reasonable likely differs per person \sdr{weak}, but sophistication goes one step further: it also suggests that we can objectively separate the information in a string into random and structured.
+\sdr{I think here we should sell the intuition better before we tear it down.}
 
-To see whether this intuition is valid; imagine that we sample from the universal distribution by feeding random bits to a universal Turing machine.
-If this UTM is constructed in the conventional manner, the initial bits will determine the index of a partial computable function $i$, and the rest will be the input to that machine $y$.
-What's the characteristic model of this data?
-Certainly, we cannot exclude the possibility that $\phi_i$ is also a universal model.
-Also, we cannot exclude the possibility that $\phi_i$ is a singleton model for $x$.
+To see whether this intuition is valid; imagine that we sample from the universal distribution by feeding random bits to a universal Turing machine, machine $u$ in the $\phi$-enumeration.
+If this UTM is constructed in the conventional manner, the initial bits will determine the index of a partial computable function $i$, and the rest will be the input $y$ to that machine.
+What is the characteristic model of the resulting data?
+Certainly, if we have to judge based only on the data we cannot exclude the possibility that the data was sampled from $\phi_u$: after all, it was.
+On the other hand, neither can we exclude the possibility that it was generated by $\phi_i$, as again, it was!
 
-Although sampling perspectives are discouraged by proponents of algorithmic statistics, this example does clash with the intuition behind sophistication: there is no reason to assume that a string with high sophistication was sampled from a complex model, so how can a complex model be intrinsic to the data?
-\footnotemark
+Although sampling perspectives are discouraged by proponents of algorithmic statistics, this statistical sanity check illustrates why it seems unjustified to assume that, based on only the data, all but one model can always be disqualified.
+\sdr{Hier stond: There is no reason to assume that a string with high sophistication was sampled from a complex model, so how can a complex model be intrinsic to the data?
+\footnotemark}
 
@@
 The theory behind sophistication suggests that we can choose one of these as the objective, intrinsic model of the data.
-Different models say different things: the universal models says that it is a `thing'.
-A more specific model might say that it is an image.
+Different models say different things: the universal models says that it is `something computable'.
+A more specific model might say that it is `an image'.
 Even more specific would be a painting, a Monet, or specifically the painting Impression of a sunrise.
@@
 More of a Monet than a painting?
-We see no intuitive reason why such a distinction should be possible.
+We see no intuitive reason why such a distinction should be possible without making any further assumptions.
 
-Of course, Kolmogorov complexity allows us to say that it is unlikely to be a Jackson Pollock, or a piece of music.
+Of course, \sdr{Actually this is not trivial at all} Kolmogorov complexity allows us to say that it is unlikely to be a Jackson Pollock, or a piece of music.
 And, if we were to get a second sample from the same source, we \emph{could} answer the question: we would receive either the same painting again, or perhaps another Monet.
 But, a single sample does not give us enough evidence to choose a single model.
-Ultimately, this is just a metaphor, but since the notion of sophistication is driven by intuition, we feel that an intuition to the contrary should serve to sharpen the debate.
+\sdr{Volgende zin weer zwak, de metafoor wordt backed up door ons voorbeeld hierboven dus is behoorlijk hard} Ultimately, this is just a metaphor, but since the notion of sophistication is driven by intuition, we feel that an intuition to the contrary should serve to sharpen the debate.
 
@@
 Whereas in Kolmogorov complexity, constant changes have constant effects on the outcome, here constant changes can cause large changes in the outcome.
-While no authors have touched on the issue of invariance, it is related to the issue of instability \cite{antunes2013sophistication,vereshchagin2013algorithmic}.
+\sdr{Not sufficiently clear} While no authors have touched on the issue of invariance, it is related to the issue of instability \cite{antunes2013sophistication,vereshchagin2013algorithmic}.
 We consider invariance the first criterion by which any proposal must be judged, with the burden of proof resting with the proposer, rather than the critic.
@@
 Finally, we note that the use of total functions has a very subtle effect on the discussion surrounding sophistication.
-Since the total functions are not enumerable, the value of the sophistication relies strongly on the difference between the Kolmogorov complexity and the Kolmogorov complexity bounded to total functions.
+Since the total functions are not enumerable, the value of the sophistication relies strongly on the difference between the Kolmogorov complexity and the Kolmogorov complexity restricted to total functions.
 This gap is exceedingly difficult to capture, because the total functions cannot be enumerated.
@@
 This creates an artificial dead end for a very valuable area of research.
+
+\sdr{Something about what is useful about the theory}
+\sdr{A picture of the structure function (perhaps sooner as well)}
+\sdr{This has become a discussion section. It does not recap. That's a bit unusual.}
 

### 2015-01-20 Facticity/Facticity/paper.tex
boe

@@
 \end{abstract}
+
 \section{Introduction}
-
-\sdr{Course sophistication handled sufficiently?}
-\sdr{Do we really need the $\gamma$-invariance? Or can $\gamma$ be 1?}
-\sdr{We need the definitions! But maybe we can simplify, do we really need the candidate set?}
-
 Komogorov complexity gives us a sound definition of the amount of information contained in a bitstring.
@@
 
+\sdr{Discuss Vereshchagin's solution}
+
 \section{Preliminaries}
 Let $\B = \{0,1\}^*$.
-We deal with computable partial functions $\phi: \B \to \B$, which we will call \emph{models}.
-A function is prefix-free if its domain, $D_f = \{y : f(y) \neq \infty\}$ is a prefix free set.
-A function $f$ is \emph{total} if $D_f = \B$.
-A prefix-free function is \emph{complete} if every infinite binary string has a member of $D_f$ as a prefix.
-We fix a prefix code with $|\bar{x}| \le|x|+2\log(|x|+2)|$ (eg.
+We deal with partial computable functions $\phi: \B \to \B$, which we will also call \emph{models}.
+A function is called \emph{prefix} if its domain, $\dom{f} = \{y : f(y) \neq \infty\}$ is a prefix free set, i.e.
+no string in $\dom{f}$ is a prefix of another.
+A function $f$ is \emph{total} if $\dom{f} = \B$.
+
+A \emph{numbering} is an enumeration of of the partial computable functions, denoted as $\psi_1, \psi_2, \ldots$ or simply $\psi$.
+We fix one canonical effective numbering $\phi$.
+We call a numbering $\psi$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ with $\forall: i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{a(i)}$.
+
+A \emph{model class} is a set of model indices. We distinguish four important model classes: 
+\begin{enumerate}
+\item $\C=\N$, the indices of all partial computable functions
+\item $\T=\{i:\tn{$\phi_i$ is total}\}$. Note that $\T$ is not enumerable.
+\item $\K$ is an enumerable set such that $\{\phi_i:i\in\K\}$ is the set of all partial computable prefix functions
+\item $\F$ is an enumerable set such that $\{\phi_i:i\in\F\}$ contains a uniform code for every finite set.
+A uniform code for a finite set $S$ is a total computable surjective function $f:\{0,1\}^{\lceil|S|\rceil}\to S$.
+\end{enumerate}
+
+We sometimes need to map binary strings to a prefix-free set.
+We denote by $\bar x$ the prefix-encoded representation for $x$.
+We require that the mapping satisfies $|\bar{x}| = |x|+O(\log(|x|)$ (see eg.
 \cite[Section~1.4]{li1993introduction}).
 
-\sdr{I think this completeness vs totalness should be explained much better, its important because without it the differences between the settings cannot be explained}
+\sdr{Define mapping between numbers and binary strings}
 
-\begin{definition}[Numberings]
+Now define the Kolmogorov complexity of a binary string $x$, with respect to a model class $\M$ and a numbering $\psi$, as follows:
+\[
+C^{\M,\psi}(x)=\min\{\bar\imath y:\psi_i(y)=x,i\in\M\}.
+\]
+The Kolmogorov complexity is invariant for the model classes $\C$ and $\K$; for those model classes the default numbering is used if it is omitted in the superscript.
 
-\begin{definition}[Prefix encoding function]
-  A prefix encoding function $f:\B\to\B$ is an injective
-  function that maps the binary strings to a prefix-free set. There
-  are prefix encoding functions $g$ with the property that
-  $|g(x)|\le|x|+2\log(|x|+2)|$ \cite{li1993introduction}. We fix such a function and
-  denote it using a bar, $\br{x}=g(x)$.
-\end{definition}
+A pair $(i,y)$ is a \emph{representation} for $x$ if $\phi_i(y)=x$.
 
-\sdr{Clearly explain what it means if a function function is total, complete, prefix.
-Do we need to fuss about equating the numbers and binary strings?
-Perhaps it's better to do the conversions explicitly in this paper}
-
-We will generalize the main proposals for sophistication into a single notation.
-We start with a class of models, we consider all two-part descriptions of the data using these models, we select the set of representations with the shortest code-lengths and we select the smallest model from this set.
-The size of the smallest model represents the sophistication of the data.
-
-A numbering is an enumeration of of the partial computable functions, denoted as $\psi_1, \psi_2, \ldots$ or simply $\psi$.
-We fix one canonical effective numbering $\phi$.
-We call a numbering $\psi$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ with $\forall: i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{a(i)}$.
-\end{definition}
-
-\begin{definition}[Complexity]
-Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ corresponds to the set of all computable prefix functions.\sdr{Cite textbook that this is possible} We define the plain and prefix Kolmogorov complexity as $C(x)=\min\{|\bar\imath y|:\phi_i(y)=x\}$ and $K(x)=\min\{|\bar\imath y|:\phi_i(y)=x, i\in\Pi\}$ respectively.
-\label{definition:complexity}
-\end{definition}
-\sdr{Added back, liked it} This definition of $K$ is uncommon.
-It is equivalent to the standard definition in \cite{li1993introduction}, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).
-
-\begin{definition}[Model classes] A \emph{model class} is a set of models.
-We distinguish five important model classes: the partial (computable) functions ($\C$), the prefix functions ($\K \subset \C$), the total functions ($\T \subset \C$), the complete functions ($\P \subset \K$) and the finite sets ($\F \subset \P$, encoded as a prefix-free function mapping all strings of length $\lceil\log S\rceil$ to members of $S$).
-\end{definition}
-
-\sdr{We used to define: A pair $(m,y)$ is a \emph{representation} for $x$ if $m$ is a model and $m(y)=x$.
-I don't think we can change this.
-Otherwise an element of $\F$ an index is not a representation.}
-
-\begin{definition}[Representations]
-A pair $(i,y)$ is a \emph{representation} for $x$ if $\phi_i(y)=x$.
-Define the following code-length functions for the representations:\\
-\-\hspace{1cm}$L(m)$ is the length of a self-delimiting description of the model $m$.\\ 
-\-\hspace{1cm}$L^m(x)=L(m)+\min\{|y|:m(y)=x\}$\\
-\-\hspace{1cm}$L^\M(x)=\min\{L^m(x):m\in\M\}$
->>>>>>> 3ebd420d297744b1c9bdacf2ebd8051f62dcfd13
-\end{definition}
-We can now define sophistication generically:
-\begin{definition}[Sophistication]
-The \emph{candidate models} for $x$ from model class $\M$ and threshold function $\tau$ are
-$C^\M_\tau(x):=\{m\in\M:L^m(x)\le \tau(x) + c\}$
-with $c$ a constant independent of $x$ which may be chosen arbitrarily. The default threshold function is $L^\M(x)$. The \emph{sophistication} is \\ 
-\-\hspace{1cm}$\s_\tau^{\M}(x):=\min\{L(m):m\in C^\M_\tau(x)\}$\label{definition:sophistication}
-\end{definition}
-Finally, we generalize the notion of invariance to allow for less strict versions. 
-\begin{definition}[$\gamma-invariance$]
-Let $\gamma:\N\to \N$ be some nondecreasing function.
-Two functions $f$ and $g$ are $\gamma$-equivalent if $|f(x) - g(x)| \in O(\gamma(\max(f(x), g(x)))$.
-We say that a function $\s_\phi(x)$ is \emph{$\gamma$-invariant} to the choice of numbering if $s_\phi$ and $s_\psi$ are $\gamma$-equal for any two acceptable numberings $\phi$ and $\psi$.
-\end{definition}
-\begin{lemma}\label{lem:broken}
-If $\s_\phi(x)$ is $\gamma$-equivalent to either $0$ or to $K(x)$ for \emph{some} numbering $\phi$, then either this equivalence exists for all other numberings as well, or $\s_\phi(x)$ is not $\gamma$-invariant.
-\end{lemma}
+The \emph{sophistication} of $x$ is 
+\begin{equation}\label{eq:soph}
+S^{\M,\psi,c}(x)=\min\{C^\K(\phi_i):i\in\M,C^{\{i\},\psi}(x)\le C^{\M,\psi}(x)+c\}.
+\end{equation}
 
@@
 \label{section:indices}
+As mentioned in the introduction, the definition of sophistication is sometimes obtained by opening up the definition of Kolmogorov complexity, and measuring the length of the index part of that two-part code.
+We call such a sophistication \emph{index sophistication}.
+Examples are Koppel and Atlan's works \cite{koppelSoph1988,koppel1991almost},  and several of its variants \cite{antunes2009sophistication,antunes2013sophistication}.
 
-We start from a simple intuition: the definition of $K$ already uses a description of the data in two parts; a computable function and an input.
-Can we assume that, when such a description is optimal, it automatically separates the structural and random information?
-As we will show, the representation of functions plays crucial fundamental.
+In these works, no constraints are placed on the used numbering.
+This makes these approaches vulnerable to the so-called \emph{nickname problem}: the number of bits required to specify a model becomes very dependent on the numbering.
+This problem is severe: any index sophistication without requirements on the numbering is either bounded, or not invariant:
 
-We will first illustrate what happens if any acceptable numbering is allowed:
-\begin{restatable}{lemma}{ineffindices}
-Any sophistication $\s_\phi(x)$ that returns the length of some index $i$ such that $\phi_i(y) = x$ for some $y$, cannot be both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
- \end{restatable}
-\sdr{Perhaps this proof should be here? It is informative}
-\noindent In the proof of this lemma (Section~\ref{section:appendix-inefficient-indices}), we deliberately used ridiculously inefficient index functions.
-We could use a more efficient index, but most definitions of sophistication do not mention any such restrictions.
-As we shall see, any inefficiency in the index will lead to a lack of invariance in sophistication.
+\begin{lemma}
+Let $s^\psi$ denote any index sophistication with respect to numbering $\psi$.
+There are acceptable numberings $\psi$ and $\xi$ such that $|s^\xi(x) - s^\psi(x)|\geq \frac{1}{2}\min\{s^\psi(x),s^\xi(x)\}$ for all $x$.
+\end{lemma}
 
-The treatments that fall prey to this problem are Koppel and Atlan's \cite{koppelSoph1988,koppel1991almost},  and several of its variants \cite{antunes2009sophistication,antunes2013sophistication}.
+To prove this, we use numberings that are deliberately extremely inefficient:
 
-Interestingly, inefficient indices are no problem in the definition of Kolmogorov complexity.
-However, this robustness does \emph{not} carry over to definitions of sophistication.
-Simply put, the length of a model in an ordinary description does not represent its complexity, much like the length of a compressible string does not reflect its information content.
-In order to obtain a robust model complexity, we use an extension of Kolmogorov complexity:
+\begin{proof}
+Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.
+
+Define $\psi$ and $\xi$ by
+\[
+\psi_j(x) = \begin{cases}
+	\phi_i(x) \;\text{if}\; j = s_{2i}\\
+	\infty\;\text{otherwise}
+\end{cases}\,\,\,
+\xi_j(x) = 
+\begin{cases}
+	\phi_i(x) \;\text{if}\; j = s_{2i+1} \\
+	\infty \;\text{otherwise}\p
+\end{cases}
+\]
+Choose any $x$ and assume w.l.o.g. that $s^\psi(x)\le s^\xi(x)$. By construction of $\psi$ and $\xi$ we have $2s_\psi(x)  \leq s_\xi(x)$ from which the result follows.
+\end{proof}
+
+While the invariance theorem expresses that inefficient indices do not affect Kolmogorov complexity too much, this robustness property does \emph{not} carry over to index sophistication.
+In other words, the length of a model index does \emph{not} represent its complexity.
+
+In order to obtain a reasonable and robust measure of model complexity, we  extend the definition of Kolmogorov complexity to functions:
 
 \begin{definition}[\cite{grunwald2004shannon,vitanyi2004meaningful}]
-The complexity of a partial recursive function is defined by $C_\phi(f) = \min\{C(i):\phi_i=f\}$ and $K_\phi(f) = \min\{K(i):\phi_i=f\}$.\label{definition:model-complexity}
+The complexity of a computable function is defined by $C^{\M,\psi}(f) = \min\{C^{\M,\psi}(i):\phi_i=f\}$.\label{definition:model-complexity}
 \end{definition}
-We show in the appendix (Lemma~\ref{lemma:invariance}) that these are invariant, ie.
-complexity is a property of a function.
+
+We show in the appendix (Lemma~\ref{lemma:invariance}) that this definition is invariant, so as the Kolmogorov complexity is an objective property of a function, as it is of a binary string.
 There are two ways to use this notion of model complexity for more robust attempts to define sophistication.
@@
 \begin{enumerate}
-  \item We maintain that the representation witnessing the Kolmogorov complexity determines the sophistication, but impose the additional constraint that an efficient, acceptable numbering is used.
-Specifically, that for every function $f$ there is an $i$ with $\phi_i=f$ and $|i|=C(f)$.
-We construct such an index below.
-This approach is taken by Adriaans \cite{adriaans2012facticity}.
+  \item We maintain that the representation witnessing the Kolmogorov complexity determines the sophistication, but impose the additional constraint that an efficient acceptable numbering is used.
+We construct such an efficient index, called a \emph{faithful} index, below.
+This is the approach taken by Adriaans \cite{adriaans2012facticity}.
   \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use $K$ to measure the complexity of the \emph{model} (as in Definition~\ref{definition:model-complexity}), which is then the first part of a two-part code describing the data.
@@
 \end{enumerate}
-To show that the two perspectives are equivalent, we must show that there exists an acceptable numbering which is efficient enough:
-\begin{definition}[Faithful Numbering]\label{def:faithful}
-  A numbering $\psi$ is \emph{faithful} if
-  for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and
-  $|j|\le C(\psi_j)+c$, for some constant $c$.
-\end{definition}
+
+We first construct an efficient numbering.
+We will call a numbering $\psi$ \emph{faithful} if for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and $|j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
 This definition is a refinement of \cite[Definition~10]{adriaans2012facticity}.
-We show in the appendix (Lemma~\ref{lemma:faithful-numberings}) that such numberings exist.
+
+\begin{lemma}
+There exist faithful acceptable numberings.
+\end{lemma}
+
 Unfortunately, even with a faithful numbering, we can fail:
 \begin{restatable}{lemma}{ineffprefix}
-Let $\phi$ be faithful and let $L(\phi_i) = |\bar\imath|$, for some prefix encoding function.
-If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then $\s^\C$ is bounded for all numberings.\label{lemma:prefix-inefficiency}
+Assume $\phi$ is faithful and define the index sophistication $s^c(x)=\min\{|\bar\imath|:C^{\{i\}}(x)\le C^{\C}(x)+c\}$.
+If $\min\{i\ge i_0:|\bar\imath|-C^\K(i)\}$ is an unbounded function of $i_0$, then $s^c$ is bounded.\label{lemma:prefix-inefficiency}
 \end{restatable}
@@
 And that this slight imbalance can already lead to underfitting.
-We can then say that the model does not occur if we use the other two-part coding approach, which is why we'll adopt that other perspective in the rest of the paper.} In order to circumvent the problem, the representation should encode the model $f$ in $K(f)$ bits, relying on $K$'s own prefix encoding to delimit the model from its input.
+[cue for lemma] We can then say that the problem does not occur if we use the other two-part coding approach, which is why we'll adopt that other perspective in the rest of the paper.} In order to circumvent the problem, the representation should encode the model $f$ in $K(f)$ bits, relying on $K$'s own prefix encoding to delimit the model from its input.
 Note that the condition for this lemma likely applies to all possible computable  prefix encoding functions (see Section~\ref{section:prefix-inefficiency}).
@@
 \begin{conjecture}
-For any partial computable prefix codelength function $L$, the function $\min\{L(i)-K(i):i\ge i_0\}$
+For any partial computable prefix codelength function $L$, the function $\min\{L(i)-C^\K(i):i\ge i_0\}$
 is unbounded in $i_0$.
@@
 
-\section{Overfitting: the singletons}
-\label{section:overfitting}
+\section{Balancing under- and overfitting}
+
+- define a default sophistication. PROBLEMS, PROBLEMS!
+- ingewikkeld lemma, laat zien:
+- can make it equal to C^K
+- can make it bounded
+- it's certainly not invariant
+- any tweak that upsets balance -> can still go wrong the other way
+- overfitting: one solution is to consider all representations that are within a constant of optimal, and then select the one with the smallest model.
+To effectively do this, the constant must depend on the numbering, but no-one knows how.
+In statistics overfitting is often addressed by imposing a penality on complex models.
+For sophistication a natural way of doing this is to allow non-prefix functions as models, which makes it comparatively cheap to put stuff in the data part.
+If you do this, overfitting cannot really happen anymore (?), but obviously only makes underfitting problem worse
+- underfitting: most common solution is to restrict the model class. This helps (theorem), but obviously makes overfitting problem only worse
+- only solution: upset balance both ways at the same time.
+There are approaches that do this and that therefore do not provably overfit or underfit, but (1) not clear if these methods are invariant (2) not clear if this sophistication still corresponds to original intuition: depth.
+See discussion
+
+Overfitting: the singletons}
+\label{section:balance}
 We now come to the main issue plaguing sophistication.
@@
 
+\subsection(Overfitting}\label{section:overfitting}
+
 \begin{restatable}[Overfitting]{lemma}{overfitting}
@@
 
-\section{Underfitting: the universal model}
+\subsection{Underfitting: the universal model}
 \label{section:underfitting}
@@
 
-\subsection{The problem of depth}
+\subsection{The problem of depth}\label{section:depth}
 
@@
 
-\section{Discussion and Conclusion}
+\section{Discussion}
 \label{section:conclusion}
-As we have seen, the idea behind sophistication has a rich history.
-It has been proposed many times, by many different authors.
-This tells us two things.
-First, there is a strong intuition behind sophistication, shared by many respected authors, and second, it is not easy to get the definition exactly right.
 
@@
 \sdr{This has become a discussion section. It does not recap. That's a bit unusual.}
+
+\section{Conclusion}
+As we have seen, the idea behind sophistication has a rich history.
+It has been proposed many times, by many different authors.
+This tells us two things.
+First, there is a strong intuition behind sophistication, shared by many respected authors, and second, it is not easy to get the definition exactly right.
 
@@
 \label{section:appendix-inefficient-indices}
-
-\ineffindices*
-
-\begin{proof}
-Choose some $\gamma(n) \in o(n)$. Assume towards contradiction that the sophistication is $\gamma$-invariant and unbounded.
-
-Let $\phi$ be the canonical acceptable numbering. Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.
-
-We now define two new acceptable numberings $\psi$ and $\xi$:
-\[
-\psi_j(x) = \begin{cases}
-	\phi_i(x) \;\text{if}\; j = s_{2i}\\
-	\infty\;\text{otherwise}
-\end{cases}\,\,\,
-\xi_j(x) = 
-\begin{cases}
-	\phi_i(x) \;\text{if}\; j = s_{2i+1} \\
-	\infty \;\text{otherwise}\p
-\end{cases}
-\]
-Let $\s_\psi$ and $\s_\xi$ be the sophistication functions built on these enumerations.
-Choose any $x$ and assume w.l.o.g.
-that $\s_\psi(x) < \s_\xi(x)$.
-By construction of the numberings $\psi$ and $\xi$, we have $2\;\s_\psi(x)  \leq \s_\xi(x)$, which gives us:
-\begin{align*}
-	\s_\xi(x) - \s_\psi(x) &\geq \frac{1}{2}\, \s_\xi(x)  \in \Omega(\s_\xi(x))\p 
-\end{align*}
-Since, by assumption, $\s_\xi$ can be arbitrarily large and $\s_\xi(x) - \s_\psi(x) \in O(\gamma(\s_\xi(x))) \subset o(\s_\xi(x))$, we have a contradiction.
-\end{proof}
 

### 2015-01-20 Facticity/Facticity/paper.tex
...

@@
 \newcommand{\hide}[1]{}
+\newcommand{\old}[1]{}
 
@@
 \section{Balancing under- and overfitting}
+\label{section:balance}
 
 - define a default sophistication. PROBLEMS, PROBLEMS!
-- ingewikkeld lemma, laat zien:
-- can make it equal to C^K
-- can make it bounded
-- it's certainly not invariant
-- any tweak that upsets balance -> can still go wrong the other way
-- overfitting: one solution is to consider all representations that are within a constant of optimal, and then select the one with the smallest model.
-To effectively do this, the constant must depend on the numbering, but no-one knows how.
-In statistics overfitting is often addressed by imposing a penality on complex models.
-For sophistication a natural way of doing this is to allow non-prefix functions as models, which makes it comparatively cheap to put stuff in the data part.
-If you do this, overfitting cannot really happen anymore (?), but obviously only makes underfitting problem worse
-- underfitting: most common solution is to restrict the model class. This helps (theorem), but obviously makes overfitting problem only worse
-- only solution: upset balance both ways at the same time.
-There are approaches that do this and that therefore do not provably overfit or underfit, but (1) not clear if these methods are invariant (2) not clear if this sophistication still corresponds to original intuition: depth.
-See discussion
-
-Overfitting: the singletons}
-\label{section:balance}
-We now come to the main issue plaguing sophistication.
-\sdr{Beetje bombastisch.
-En hierna zou ik zeggen, we first consider the simplest, most natural setting, \dots} If we use model class $\K$, the candidate set always contains two representations that do not provide a good separation.
-The first is a \emph{universal model}: a basic universal Turing machine.
-The second is a \emph{singleton model} a model which returns $x$ for any input.
-As we will show, in many settings, we can ensure that one of them is the \emph{only} model in the candidate set.
-The following lemma generalizes the principle:
 
@@
 \end{restatable}
-\noindent We can now show numberings for which the singletons are always selected:
 
-\subsection(Overfitting}\label{section:overfitting}
-
+- ingewikkeld lemma, laat zien:\\
+- can make it equal to $C^K$
 \begin{restatable}[Overfitting]{lemma}{overfitting}
@@
 \end{restatable}
-\noindent In short, this lemma tells us that there are numberings for which the singleton models always determine the sophistication.
-This means that there is a numbering $\phi$ such that the sophistication is equivalent to Kolmogorov complexity, which implies that this sophistication is unsound (by Lemma~\ref{lem:broken}).
 
-Many variants in the literature (including the structure function) are susceptible to this problem  \sdr{Don't say ``many'', just list explicitly} \cite{cover1985kolmogorov,gacs2001algorithmic,vereshchagin2013algorithmic}.
-In \cite{vitanyi2004meaningful,adriaans2012facticity} this issue is avoided by not using self-delimiting representations.
-\sdr{Prep the balance idea so that this is understandable.
-We need it}  Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically not optimally efficient.
-While this dodges the issue of the singletons, it does not show that $S$ is invariant.
-
-Another approach may be to make the constant threshold dependent on the choice of enumeration, although we are not aware of any published efforts.
-
-\subsection{Underfitting: the universal model}
-\label{section:underfitting}
-On the other end of the spectrum, there exist representations with all information in the noise, using a universal model.
-We can easily show that if such models exist, we can choose our numbering so that they are always chosen.
+- can make it bounded
 \begin{restatable}[underfitting]{lemma}{underfitting}
@@
 \end{restatable}
-\noindent Most sophistication measures avoid the underfitting problem by limiting their model class, usually to either finite sets ($\F$) or total functions ($\T$ or $\P$).
-This eliminates the universal model, but as we shall see, it does not solve the problem for most data.
-The variants that are susceptible to the this type of underfitting are facticity \cite{adriaans2012facticity} and some versions of effective complexity \cite{gellmann1996information}.
-The latter aims to bypass the problem by selecting carefully from the candidate set, but as we have shown, for some numberings the candidate set contains only the universal model.
-\sdr{Explain this is actually a widely recognised problem}
 
-\subsection{The problem of depth}\label{section:depth}
+- it's certainly not invariant \\
+- any tweak that upsets balance -> can still go wrong the other way
 
-We have seen that some variants of sophistication sidestep the simplest examples of underfitting by using total models.
-Does this ensure that these variants are not bounded?
-A well-known result in algorithmic statistics shows that there exist strings for which no finite set is as good a model as the singleton set.
-In this section we will focus on the variants $S^\T$ and $S^T_K$ \cite{vitanyi2004meaningful}, as those are the variants most resistant to proofs of non-invariance.
-\sdr{Do I know this notation? Also, I wanna emphasize that the structure function underfits in this way.}
-\sdr{Combine this lemma with the one after next}
-\begin{restatable}{lemma}{nonstochastic}
-There are infinitely  $x$ such that $S^\F_K(x) \geq K(x)/5$. \label{lemma:structure-function-is-not-bounded}
+\subsection{Overfitting}
+
+- overfitting: one solution is to consider all representations that are within a constant of optimal, and then select the one with the smallest model.
+To effectively do this, the constant must depend on the numbering, but no-one knows how.
+\\
+- In statistics overfitting is often addressed by imposing a penalty on complex models.
+For sophistication a natural way of doing this is to allow non-prefix functions as models, which makes it comparatively cheap to put stuff in the data part.
+If you do this, overfitting cannot really happen anymore (?), but obviously only makes underfitting problem worse
+
+\subsection{Underfitting}
+- underfitting: most common solution is to restrict the model class. This helps
+
+\begin{restatable}{theorem}{dogfood}
+There are infinitely many $x$ with $\s^{\F,\phi,c}(x) \geq C^{\K}(x)/5$ and $\s^{\F,\phi,c}(x) \geq C^{\K}(x)/6$. 
 \end{restatable}
 
-\sdr{Move this lemma to the appendix altogether}
-\begin{restatable}{lemma}{vitanyi}
-Let $f$ be a total, computable function, such that for some $d$, $f(d) = x$ and let $k = K(f) + |d|$.
-Then there exists a finite set $S$ containing $x$ and a constant $c$, such that $K(S) \leq K(f) + K(|d|) + c$ and $\log |S| \leq |d|$.\label{lemma:total-to-sets}
-\end{restatable}
-\noindent This lemma is a variation on \cite[Lemma~7.2]{vitanyi2004meaningful}.
-We use this lemma to transport the result of Theorem~\ref{lemma:structure-function-is-not-bounded} to the model class of total functions:
+(theorem), but obviously makes overfitting problem only worse
 
-\begin{restatable}{theorem}{dogfood}
-There are infinitely many $x$ with $S^\T_K(x) \geq K(x)/6$ and $S^\T(x) \geq K(x)/6$. 
-\end{restatable} 
-\noindent This tells us that $S^\T_K(x)$ and $S^\T(x)$ can truly grow to a significant part of the total description length.
-While this shows that sophistication, if defined with particular care, is not entirely vacuous, \sdr{hold on, it might still be completely meaningless and/or not invariant} we can shed some light on exactly which strings end up with high sophistication, using the principle of \emph{depth}.
-We will use the following definition:
+- only solution: upset balance both ways at the same time.
+There are approaches that do this and that therefore do not provably overfit or underfit, but (1) not clear if these methods are invariant (2) not clear if this sophistication still corresponds to original intuition: depth.
+See discussion
 
@@
 The relation between sophistication, depth and absolutely non-stochastic strings was also investigated in \cite{antunes2013sophistication}.
+
+\old{
+
+We now come to the main issue plaguing sophistication.
+\sdr{Beetje bombastisch.
+En hierna zou ik zeggen, we first consider the simplest, most natural setting, \dots} If we use model class $\K$, the candidate set always contains two representations that do not provide a good separation.
+The first is a \emph{universal model}: a basic universal Turing machine.
+The second is a \emph{singleton model} a model which returns $x$ for any input.
+As we will show, in many settings, we can ensure that one of them is the \emph{only} model in the candidate set.
+The following lemma generalizes the principle:
+
+\noindent We can now show numberings for which the singletons are always selected:
+
+\noindent In short, this lemma tells us that there are numberings for which the singleton models always determine the sophistication.
+This means that there is a numbering $\phi$ such that the sophistication is equivalent to Kolmogorov complexity, which implies that this sophistication is unsound (by Lemma~\ref{lem:broken}).
+
+Many variants in the literature (including the structure function) are susceptible to this problem  \sdr{Don't say ``many'', just list explicitly} \cite{cover1985kolmogorov,gacs2001algorithmic,vereshchagin2013algorithmic}.
+In \cite{vitanyi2004meaningful,adriaans2012facticity} this issue is avoided by not using self-delimiting representations.
+\sdr{Prep the balance idea so that this is understandable.
+We need it}  Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically not optimally efficient.
+While this dodges the issue of the singletons, it does not show that $S$ is invariant.
+
+Another approach may be to make the constant threshold dependent on the choice of enumeration, although we are not aware of any published efforts.
+
+On the other end of the spectrum, there exist representations with all information in the noise, using a universal model.
+We can easily show that if such models exist, we can choose our numbering so that they are always chosen.
+
+\noindent Most sophistication measures avoid the underfitting problem by limiting their model class, usually to either finite sets ($\F$) or total functions ($\T$ or $\P$).
+This eliminates the universal model, but as we shall see, it does not solve the problem for most data.
+The variants that are susceptible to the this type of underfitting are facticity \cite{adriaans2012facticity} and some versions of effective complexity \cite{gellmann1996information}.
+The latter aims to bypass the problem by selecting carefully from the candidate set, but as we have shown, for some numberings the candidate set contains only the universal model.
+\sdr{Explain this is actually a widely recognised problem}
+
+\noindent This tells us that $S^\T_K(x)$ and $S^\T(x)$ can truly grow to a significant part of the total description length.
+While this shows that sophistication, if defined with particular care, is not entirely vacuous, \sdr{hold on, it might still be completely meaningless and/or not invariant} we can shed some light on exactly which strings end up with high sophistication, using the principle of \emph{depth}.
+We will use the following definition:
+
+}
 
@@
 
-\nonstochastic*
+We first show that representions using $\T$ can be reduced to representations in $\F$ with a small overhead.
+\pb{Definitie nonstochastic strings uit \cite[Proposition~I.3 (b)]{gacs2001algorithmic}}
 
-\begin{proof}
-From \cite[Proposition~I.3 (b)]{gacs2001algorithmic} we know that for sufficiently large $i$ there are strings of length less than $i$ that are not $(i/4, i/4)$-stochastic.
-For each $i$, let $x_i$ be the smallest binary string that is not $(i/4, i/4)$-stochastic.
-Since every string is $(\alpha,\alpha)$-stochastic for sufficiently large $\alpha$, this sequence contains infinitely many distinct binary strings.
-Now pick any $i$ that is large enough that (a) $|x_i|<i$ and (b) $K(x_i)/5 < i/4$.
-By definition of $(\alpha,\beta)$-stochasticity, if a string is not $(i/4,i/4)$-stochastic, then for every finite set $S$ containing $x$, either (a) $\log|S|\ge K(x_i)+i/4$, in which case $S$ cannot be a candidate representation for $x$ in $C^\F_K$, or (b) $K(S)>i/4>K(x_i)/5$.
-This proves the theorem.
-\end{proof}
-
-\vitanyi*
+\begin{restatable}{lemma}{vitanyi}
+Let $f$ be a total, computable function, such that for some $d$, $f(d) = x$ and let $k = K(f) + |d|$.
+Then there exists a finite set $S$ containing $x$ and a constant $c$, such that $K(S) \leq K(f) + K(|d|) + c$ and $\log |S| \leq |d|$.\label{lemma:total-to-sets}
+\end{restatable}
 \begin{proof}
@@
 \end{proof}
+\noindent This lemma is a variation on \cite[Lemma~7.2]{vitanyi2004meaningful}.
 
+We use this to show that sophistications with a restricted model class are not bounded:
 \dogfood*
 \begin{proof}
-We continue from the proof of Theorem~\ref{lemma:structure-function-is-not-bounded}.
-Suppose towards contradiction that either $C^\T_K(x_i)$ or $C^\T_{\tau^\circ}(x_i)$ contains a candidate $f$ with $K(f)\le i/5$.
-Then by Lemma~\ref{lemma:total-to-sets}, there is a set $S$ with $\log|S|\le|d|\le\tau^\circ(x)\le K(x)$, so (a) in the proof of Theorem~\ref{lemma:structure-function-is-not-bounded} above cannot be the case, and $K(S)\le K(f)+K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) in the same proof for large enough $i$.
+From \cite[Proposition~I.3 (b)]{gacs2001algorithmic} we know that for sufficiently large $i$ there are strings of length less than $i$ that are not $(i/4, i/4)$-stochastic.
+For each $i$, let $x_i$ be the smallest binary string that is not $(i/4, i/4)$-stochastic.
+Since every string is $(\alpha,\alpha)$-stochastic for sufficiently large $\alpha$, this sequence contains infinitely many distinct binary strings.
+Now pick any $i$ that is large enough that (a) $|x_i|<i$ and (b) $C^\K(x_i)/5 < i/4$.
+By definition of $(\alpha,\beta)$-stochasticity, if a string is not $(i/4,i/4)$-stochastic, then for every finite set $S$ containing $x$, either (a) $\log|S|\ge C^\K(x_i)+i/4$, in which case $S$ cannot be a candidate representation for $x$ in \pb{candidate set}, or (b) $C^\K(S)>i/4>C^\K(x_i)/5$.
+This proves the first inequality.
+
+Now, suppose towards contradiction that \pb{$\T$ candidate set} contains a candidate $f$ with $C^\K(f)\le i/5$.
+Then by Lemma~\ref{lemma:total-to-sets}, there is a set $S$ with $\log|S|\le|d|\le\tau^\circ(x)\le C^\K(x)$, so (a) above cannot be the case, and $K(S)\le K(f)+K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) for large enough $i$.
 Therefore such candidates $f$ cannot exist and all candidates must satisfy $K(f)>i/5>K(x_i)/6$.

### 2015-01-20 Facticity/Facticity/paper.tex
depth hopefully fixed

@@
 \newcommand{\dom}[1]{\mathop{\tn{dom}(#1)}}
+\newcommand{\range}[1]{\mathop{\tn{range}(#1)}}
 
@@
 \begin{proof}
-From \cite[Proposition~I.3 (b)]{gacs2001algorithmic} we know that for sufficiently large $i$ there are strings of length less than $i$ that are not $(i/4, i/4)$-stochastic.
-For each $i$, let $x_i$ be the smallest binary string that is not $(i/4, i/4)$-stochastic.
+For all $i$, let $x_i$ be the smallest binary string that is not $(i/4, i/4)$-stochastic.
 Since every string is $(\alpha,\alpha)$-stochastic for sufficiently large $\alpha$, this sequence contains infinitely many distinct binary strings.
+From \cite[Proposition~I.3 (b)]{gacs2001algorithmic} we know that for all sufficiently large $i$ there are strings of length less than $i$ that are not $(i/4, i/4)$-stochastic.
 Now pick any $i$ that is large enough that (a) $|x_i|<i$ and (b) $C^\K(x_i)/5 < i/4$.
-By definition of $(\alpha,\beta)$-stochasticity, if a string is not $(i/4,i/4)$-stochastic, then for every finite set $S$ containing $x$, either (a) $\log|S|\ge C^\K(x_i)+i/4$, in which case $S$ cannot be a candidate representation for $x$ in \pb{candidate set}, or (b) $C^\K(S)>i/4>C^\K(x_i)/5$.
-This proves the first inequality.
+By definition of $(\alpha,\beta)$-stochasticity, if a string is not $(i/4,i/4)$-stochastic, then for every finite set $S$ containing $x$, either (a) $\log|S|\ge C^\K(x_i)+i/4\ge 5C^\K(x_i)/4$, or (b) $C^\K(S)>i/4>C^\K(x_i)/5$.
+In case (a), any two-part description using a uniform code for $S$ is substantially larger than the Kolmogorov complexity, for large enough $i$, so it does not determine the sophistication.
+Case (b) provides the required lower bound on the complexity of the model.
 
-Now, suppose towards contradiction that \pb{$\T$ candidate set} contains a candidate $f$ with $C^\K(f)\le i/5$.
-Then by Lemma~\ref{lemma:total-to-sets}, there is a set $S$ with $\log|S|\le|d|\le\tau^\circ(x)\le C^\K(x)$, so (a) above cannot be the case, and $K(S)\le K(f)+K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) for large enough $i$.
-Therefore such candidates $f$ cannot exist and all candidates must satisfy $K(f)>i/5>K(x_i)/6$.
+Now, take $\T$ as model class and suppose towards contradiction that the sophistication is determined by a total function $f=\phi_j,j\in\T$ with $f(d)=x$ and $C^\K(f)\le i/5$.
+Then by Lemma~\ref{lemma:total-to-sets}, for some fixed constant $c$ there is a finite set $S$ with $\log|S|\le|d|\le C^\K(x_i)$, so (a) above cannot be the case, and $C^\K(S)\le C^\K(f)+C^\K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) for large enough $i$.
+Therefore such candidates $f$ cannot exist and all candidates must satisfy $C^\K(f)>i/5>C^\K(x_i)/6$.
 \end{proof}
 
-Note that we favoured simplicity of the proof over the strength of the bounds: the multiplicative constant $1/6$ can be improved with a more detailed treatment.
+Note that we favoured brevity of the proof over the strength of the bounds: the multiplicative constant $1/6$ can be improved with a more detailed treatment.
 

### 2015-01-20 Facticity/Facticity/paper.tex
...

--- 2015-01-20 Facticity/Facticity/paper.tex
@@
 Let $s^\psi$ denote any index sophistication with respect to numbering $\psi$.
-There are acceptable numberings $\psi$ and $\xi$ such that $|s^\xi(x) - s^\psi(x)|\geq \frac{1}{2}\min\{s^\psi(x),s^\xi(x)\}$ for all $x$.
+There are acceptable numberings $\psi$ and $\xi$ such that for all $x$:
+\[|s^\xi(x) - s^\psi(x)|\geq \frac{1}{2}\min\{s^\psi(x),s^\xi(x)\}\p\]
 \end{lemma}
@@
 \end{proof}
-
 While the invariance theorem expresses that inefficient indices do not affect Kolmogorov complexity too much, this robustness property does \emph{not} carry over to index sophistication.
@@
 \begin{definition}[\cite{grunwald2004shannon,vitanyi2004meaningful}]
-The complexity of a computable function is defined by $C^{\M,\psi}(f) = \min\{C^{\M,\psi}(i):\phi_i=f\}$.\label{definition:model-complexity}
+The complexity of a computable function is defined by \label{definition:model-complexity}
+\[
+C^{\M,\psi}(f) = \min\{C^{\M,\psi}(i):\phi_i=f\}\p
+\]
 \end{definition}
@@
 \end{enumerate}
-
-We first construct an efficient numbering.
-We will call a numbering $\psi$ \emph{faithful} if for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and $|j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
+We first construct an efficient numbering: we will call a numbering $\psi$ \emph{faithful} if for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and $|j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
 This definition is a refinement of \cite[Definition~10]{adriaans2012facticity}.
@@
 
-Unfortunately, even with a faithful numbering, we can fail:
+Unfortunately, even with a faithful numbering, we can fail.
+The Kolmogorov complexity uses representations of the form $\bar\imath y$, where the bar denotes some straightforward prefix encoding to delimit the model description $i$ from its input $y$.
+If we define a second prefix encoding $\tilde{\imath}$, which is always better than our default by some slowly growing term, such as $O(\log\log\log i)$, we can define a second representation $\bar u \tilde \imath y$, which takes a constant amount to switch to a different encoding of the numbering, but gains more than a constant, for sufficiently complex strings.
+Ultimately, $u$ would always be chosen as a model for sufficiently complex $x$.
+
+Can we make our prefix encoding so efficient that no other encoding can undercut it by more than a constant amount?
+One self-delimiting description of $\phi_i$ is teh one implied by $C^\K(\phi_i)$.
+By definition, this is the smallest effective, self-delimiting description of $\phi_i$.
+Let $i^*$ be a self-delimiting program for $\phi_i$, witnessing $C^\K(\phi_i)$.
+If our prefix function is less efficient than $C^\K(x)$ by an unbounded amount, the effective representation $i^*y$ will always beat $\bar\imath y$ for complex enough data.
+We conjecture that no computable prefix function can avoid this:
+\begin{conjecture}
+For any partial computable prefix codelength function $L$, the function $\min\{L(i)-C^\K(i):i\ge i_0\}$
+is unbounded in $i_0$.
+\end{conjecture}
+The following lemma shows that for all prefix functions that are inefficient in this way, the index sophistication fails. This icludes \cite{adriaans2012facticity}.
 \begin{restatable}{lemma}{ineffprefix}
@@
 \end{restatable}
-\noindent This result applies to \cite{adriaans2012facticity}.
-To avoid the problem, the model $\phi_i$ should be encoded in $K(\phi_i)$ bits, relying on $K$'s own prefix encoding to delimit the model.
-\sdr{The above is incomprehensible.
-Explain better why this can't be fixed; point out that it provides a slight imbalance in favour of putting stuff in the data part, which, if the model class is $\C$, already to underfitting.
-Prime the reader to think in terms of balance.
-Refer to the underfitting section.}
-\sdr{We should point out from the start that the difference in cost for the model $|\bar\imath|$ instead of $K(i)$ is unbounded, so even a faithful model index is still not an optimal description, and the gap is a direct consequence of the definition of Kolmogorov complexity.
-And that this slight imbalance can already lead to underfitting.
-[cue for lemma] We can then say that the problem does not occur if we use the other two-part coding approach, which is why we'll adopt that other perspective in the rest of the paper.} In order to circumvent the problem, the representation should encode the model $f$ in $K(f)$ bits, relying on $K$'s own prefix encoding to delimit the model from its input.
-Note that the condition for this lemma likely applies to all possible computable  prefix encoding functions (see Section~\ref{section:prefix-inefficiency}).
-Thus, even if we have a faithful numbering, we can only use it if we don't count the cost of delimiting the model from the input (as in \cite{koppelSoph1988}).
-In the following, we will take the second perspective: the representations are two-part descriptions $(i, y)$ such that $\phi_i(y)=x$, with codelength $K(\phi_i) + |y|$.
 
-\sdr{Moved this back here, the conjecture needs to be in the main text}
-Lemma~\ref{lemma:prefix-inefficiency} only holds if the prefix function $L(i)$ has the property that $\min\{L(i)-K(i):i\ge i_0\}$ is unbounded.
-We conjecture that this is the case for any computable function.
+If our conjecture is correct, the index sophistication is always bounded.
+The only way to avoid the problem is to use a the self-delimiting program of length $C^\K(x)$ to describe the model.
+Any inefficieny in the model description, and a second represention using a universal model with a more efficient model description will win, causing a bounded sophistication.
 
-We conjecture that it applies to all possible computable prefix encoding functions:
-\begin{conjecture}
-For any partial computable prefix codelength function $L$, the function $\min\{L(i)-C^\K(i):i\ge i_0\}$
-is unbounded in $i_0$.
-\end{conjecture}
+This leads us to the follwong definition of sophistication:
+\begin{definition}
+The sophistication, for a model class $\M$, any acceptable numbering $\psi$ and constant $c$ is:
+\end{definition}
 
@@
 
-- define a default sophistication. PROBLEMS, PROBLEMS!
+In the previous section, we saw the first glimpse of how delicate the internal balance in sophistication is.
+If the description of the model $m$ is less efficient than $C^\K(m)$ by anything more than a constant, representations with a bounded model size will always win out in the limit.
 
@@
 \begin{restatable}{theorem}{dogfood}
-There are infinitely many $x$ with $\s^{\F,\phi,c}(x) \geq C^{\K}(x)/5$ and $\s^{\F,\phi,c}(x) \geq C^{\K}(x)/6$. 
+There are infinitely many $x$ with $\s^{\F,\phi,c}(x) \geq C^{\K}(x)/5$ and $\s^{\T,\phi,c}(x) \geq C^{\K}(x)/6$. 
 \end{restatable}

### 2015-01-20 Facticity/Facticity/paper.tex
...

@@
 As we shall see, this version behaves very differently from regular sophistication, and is closer to computational depth.
-Vereshchagin \cite{vereshchagin2013algorithmic} proposes an alternative solution to the instability issue.
+Vereshchagin \cite{vereshchagin2013algorithmic}, working in the finite set setting,  suggests additional restrictions on models that are allowed to determine the sophistication, in hopes of curbing the instability issue.
 Adriaans' \emph{Facticity}, introduced in 2012\cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions.
 Adriaans also identifies the nickname problem discussed in Section~\ref{section:indices}.
-
-\sdr{Discuss Vereshchagin's solution}
 
@@
 \cite[Section~1.4]{li1993introduction}).
-
-\sdr{Define mapping between numbers and binary strings}
+To simplify notation, we will sometimes conflate natural numbers and binary strings, implicitly using the ordering $(0, \epsilon)$, $(1, 0)$, $(2, 1)$, $(3, 00)$, $(4, 01)$, \ldots
 
@@
 \]
-The Kolmogorov complexity is invariant for the model classes $\C$ and $\K$; for those model classes the default numbering is used if it is omitted in the superscript.
+The model class $\C$ gives us the plain Kolmogorov complexity and the $\K$ gives us the prefix-free Kolmogorov complexity.
+The complexity is invariant to the numbering for these model classes, so we may omit it from the superscript.
 
 A pair $(i,y)$ is a \emph{representation} for $x$ if $\phi_i(y)=x$.
-
-The \emph{sophistication} of $x$ is 
-\begin{equation}\label{eq:soph}
-S^{\M,\psi,c}(x)=\min\{C^\K(\phi_i):i\in\M,C^{\{i\},\psi}(x)\le C^{\M,\psi}(x)+c\}.
-\end{equation}
 
@@
 \end{definition}
-
-We show in the appendix (Lemma~\ref{lemma:invariance}) that this definition is invariant, so as the Kolmogorov complexity is an objective property of a function, as it is of a binary string.
+We show in the appendix (Lemma~\ref{lemma:invariance}) that this definition is invariant, so that the Kolmogorov complexity is an objective property of a function, as it is of a binary string.
 There are two ways to use this notion of model complexity for more robust attempts to define sophistication.
@@
 \end{lemma}
-
 Unfortunately, even with a faithful numbering, we can fail.
 The Kolmogorov complexity uses representations of the form $\bar\imath y$, where the bar denotes some straightforward prefix encoding to delimit the model description $i$ from its input $y$.
-If we define a second prefix encoding $\tilde{\imath}$, which is always better than our default by some slowly growing term, such as $O(\log\log\log i)$, we can define a second representation $\bar u \tilde \imath y$, which takes a constant amount to switch to a different encoding of the numbering, but gains more than a constant, for sufficiently complex strings.
+If we define a second prefix encoding $\tilde{\imath}$, which is always better than our default by some slowly growing term, such as $O(\log\log\log i)$, we can define a second representation $\bar u \tilde \imath y$, which takes a constant amount of bits $|\bar u|$ to switch to a different universal model, but gains more than a constant, for sufficiently complex strings, because a different prefix function is used.
 Ultimately, $u$ would always be chosen as a model for sufficiently complex $x$.
@@
 Can we make our prefix encoding so efficient that no other encoding can undercut it by more than a constant amount?
-One self-delimiting description of $\phi_i$ is teh one implied by $C^\K(\phi_i)$.
+One self-delimiting description of $\phi_i$ is the one used in $C^\K(\phi_i)$.
 By definition, this is the smallest effective, self-delimiting description of $\phi_i$.
-Let $i^*$ be a self-delimiting program for $\phi_i$, witnessing $C^\K(\phi_i)$.
+Let $i^*$ be a the shortest self-delimiting program for $\phi_i$.
 If our prefix function is less efficient than $C^\K(x)$ by an unbounded amount, the effective representation $i^*y$ will always beat $\bar\imath y$ for complex enough data.
-We conjecture that no computable prefix function can avoid this:
+We conjecture that no computable prefix encoding function can avoid this issue:
 \begin{conjecture}
@@
 \end{conjecture}
-The following lemma shows that for all prefix functions that are inefficient in this way, the index sophistication fails. This icludes \cite{adriaans2012facticity}.
+The following lemma shows that for all index sophistications using an inefficient prefix encoding will fail. This includes \cite{adriaans2012facticity}.
 \begin{restatable}{lemma}{ineffprefix}
@@
 
-This leads us to the follwong definition of sophistication:
-\begin{definition}
-The sophistication, for a model class $\M$, any acceptable numbering $\psi$ and constant $c$ is:
-\end{definition}
+For this reason we will using the following definition of sophistication in the remainder of this article: 
+\begin{equation}\label{eq:soph}
+S^{\M,\psi,c}(x)=\min\left\{C^\K(\phi_i):i\in\M,C^{\{i\},\psi}(x)\le C^{\M,\psi}(x)+c\right\}.
+\end{equation}
 
@@
 If the description of the model $m$ is less efficient than $C^\K(m)$ by anything more than a constant, representations with a bounded model size will always win out in the limit.
-
+In light of this balance issues, the most natural definition of sophistication is $S^{\K,\psi,c}$.
+In this setting, we know that there always exist candidate models---those models $i$ with $i \in \K$ and $C^{i}(x) \leq C^\K(x) + c$---for which al information except a constant amount is stored in the data, and candidates for which all information is stored in the model.
+This tells us that $S^\K$ is \emph{balanced}: there is no inefficiency in the representation forcing information in or out of the model.
+The flipside of having a balanced sophistication is that invariance is easy to show.
+We will show that there are numberings for which the universal models always determine the $S^\K$ and numberings for which the singletons always determine $S^\K$.
+	The following lemma generalizes the principle: 
 \begin{restatable}{lemma}{coolone}
@@
 
-- ingewikkeld lemma, laat zien:\\
-- can make it equal to $C^K$
+This is a technical lemma, due to the double effect of the numbering, which occurs in both the definition of model complexity and the definition of data complexity.
+The \emph{idea} behind the lemma, however, is simple.
+Given a subset $\M'$ of the model class, we can choose the numbering so that representations use a model in $\M'$ perform an arbitrary constant better than all other models.
+This lets us effectively `push' all other models out of the candidate set, so that the sophistication is determined by a model in $\M'$.- can make it equal to $C^K$
+
+This allows us to ensure that for some numberings, the unversal model always determines the sophistication, making is bounded:
+
+\begin{restatable}[underfitting]{lemma}{underfitting}
+Let $\M$ be a model class containing a universal model $u$, with the property that $\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p$. Then, there exist numberings such that $\s^\M$ is bounded.
+\end{restatable}
+For other numberings, the singletons determine the sophistication, making it equal to the complexity:
 \begin{restatable}[Overfitting]{lemma}{overfitting}
@@
 \end{restatable}
-
-- can make it bounded
-\begin{restatable}[underfitting]{lemma}{underfitting}
-Let $\M$ be a model class containing a universal model $u$, with the property that $\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p$. Then, there exist numberings such that $\s^\M$ is bounded.
-\end{restatable}
-
-- it's certainly not invariant \\
-- any tweak that upsets balance -> can still go wrong the other way
+This shows that the price we pay for balance is absolute invariance.
+All information can be seen as structure of noise, depending on the numbering.
+To avoid these issues, existing proposals upset the balance to exclude, or penalize at least the singleton models and the universal models.
 
@@
 
-- overfitting: one solution is to consider all representations that are within a constant of optimal, and then select the one with the smallest model.
-To effectively do this, the constant must depend on the numbering, but no-one knows how.
-\\
-- In statistics overfitting is often addressed by imposing a penalty on complex models.
-For sophistication a natural way of doing this is to allow non-prefix functions as models, which makes it comparatively cheap to put stuff in the data part.
-If you do this, overfitting cannot really happen anymore (?), but obviously only makes underfitting problem worse
+The existing proposals that do not avoid overfitting as the structure function \cite{cover1985kolmogorov}, the sophistication as defined in \cite{mota2013sophistication}, the miinimal sufficient statistic \cite{gacs2001algorithmic} and Vereshchagin's strongly  sufficient statistic \cite{vereshchagin2013algorithmic}.
+It may be argued that the constant determining the candidates should be allowed to depend on the numbering, but this dependence has not been mentioned in the literature and there is no clear way to determine what such a constant should be, on the basis of a given numbering.
+
+In traditional statistics, overfitting is often addressed by imposing a penalty on complex models.
+As we have seen, a strong penalty, such as the one imposed by an inefficient prefix encoding of the model, will immediately cause underfitting.
+A more subtle approach is to use $C$ as a model class, allowing representations that are not self-delimiting.
+This eliminates at least the singletons, as placing all information in the model automatically creates a self delimiting representation.
+Self delimiting representations are longer than plain representations, and this gap is unbounded \cite[Section~4.5.5]{vitanyi2004meaningful}, which tells us that singletons will be candidates for only finitely many strings.
+This approach is taken by Vit\'anyi \cite{vitanyi2004meaningful} and by Adriaans \cite{adriaans2012facticity}.
 
 \subsection{Underfitting}
-- underfitting: most common solution is to restrict the model class. This helps
+
+Underfitting is a widely acknowledged problem, and almost all proposals avoid it by limiting the class of allowed models, to exclude universal models.
+Does this make the sophistication unbounded?
+The following theorem shows that it does.
 
@@
 
-(theorem), but obviously makes overfitting problem only worse
+But clearly, this only exacerbates the overfitting problem.
+Only one model class eliminates both the singletons and the unversal model: $S^\T(x)$.
+The only proposal that uses an efficient model representation, excludes the universal models and excludes the singletons is Vit\'anyi's \cite{vitanyi2004meaningful}.
+While this invalidates straightforward proofs of boundedness, there is no suggestion that $S^\T(x)$ is actually invariant.
 
-- only solution: upset balance both ways at the same time.
-There are approaches that do this and that therefore do not provably overfit or underfit, but (1) not clear if these methods are invariant (2) not clear if this sophistication still corresponds to original intuition: depth.
-See discussion
+While under $S^{\T, \phi, c}(x)$ there are clearly string with high sophistication, these may not conform to the original intuition that motivated sophistication.
+To show this, we require the concept of depth:
 
@@
 \end{restatable}
-\noindent This shows that while high-sophistication strings exist, the mechanics do not follow the basic intuition of sophistication.
-Unless we encounter data that would take longer than the lifetime of the universe (and then some) to produce \sdr{Not sure about this philosophically, but lets discuss}, sophistication behaves exactly as it would without the restricted model class: a solution to a super-exponential problem \sdr{Bad example, use something that clearly does not require much computation but has a lot of structure.
-A midpoint-displaced heightmap comes to mind} would have no greater sophistication than a string sampled from a simple Bernoulli model.
-This result also suggests that the `non-stochastic' property of strings with high sophistication\cite{shen1983concept,vereshchagin2004kolmogorov} says more about depth and totality than it does about structure and noise.
+\noindent This shows that while high-sophistication strings exist, they do not behave as expected.
+Consider a string that is typical for a low-depth model, say some high-complexity Markov chain.
+Under sophistication, no matter how complex the markov chain, the sophistication would stay the same.
+Any structure that is simple enough to be exploited within the time bound of the Ackermann function (or some other fast-growing, simply stated bound) will be counted as noise.
+Only structure that is so deep that if we used it to compress $x$ we could never hope to decompress it within the lifetime of the universe, would count towards sophistication.
+This clearly contradicts the motivating principle of sophistication, which is to capture the number of bits required to express \emph{all} regularity in the data. 
+What makes a string high-sophistication is not the fact that it has a great deal of structure, but the fact that it can be compressed far better with partial functions than with total.
+That is, it is non-typical for the model class $\T$.
+This suggests that the `non-stochastic' property of strings with high sophistication \cite{shen1983concept,vereshchagin2004kolmogorov} says more about depth and totality than it does about structure and noise.
 
 The relation between sophistication, depth and absolutely non-stochastic strings was also investigated in \cite{antunes2013sophistication}.
-
-\old{
-
-We now come to the main issue plaguing sophistication.
-\sdr{Beetje bombastisch.
-En hierna zou ik zeggen, we first consider the simplest, most natural setting, \dots} If we use model class $\K$, the candidate set always contains two representations that do not provide a good separation.
-The first is a \emph{universal model}: a basic universal Turing machine.
-The second is a \emph{singleton model} a model which returns $x$ for any input.
-As we will show, in many settings, we can ensure that one of them is the \emph{only} model in the candidate set.
-The following lemma generalizes the principle:
-
-\noindent We can now show numberings for which the singletons are always selected:
-
-\noindent In short, this lemma tells us that there are numberings for which the singleton models always determine the sophistication.
-This means that there is a numbering $\phi$ such that the sophistication is equivalent to Kolmogorov complexity, which implies that this sophistication is unsound (by Lemma~\ref{lem:broken}).
-
-Many variants in the literature (including the structure function) are susceptible to this problem  \sdr{Don't say ``many'', just list explicitly} \cite{cover1985kolmogorov,gacs2001algorithmic,vereshchagin2013algorithmic}.
-In \cite{vitanyi2004meaningful,adriaans2012facticity} this issue is avoided by not using self-delimiting representations.
-\sdr{Prep the balance idea so that this is understandable.
-We need it}  Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically not optimally efficient.
-While this dodges the issue of the singletons, it does not show that $S$ is invariant.
-
-Another approach may be to make the constant threshold dependent on the choice of enumeration, although we are not aware of any published efforts.
-
-On the other end of the spectrum, there exist representations with all information in the noise, using a universal model.
-We can easily show that if such models exist, we can choose our numbering so that they are always chosen.
-
-\noindent Most sophistication measures avoid the underfitting problem by limiting their model class, usually to either finite sets ($\F$) or total functions ($\T$ or $\P$).
-This eliminates the universal model, but as we shall see, it does not solve the problem for most data.
-The variants that are susceptible to the this type of underfitting are facticity \cite{adriaans2012facticity} and some versions of effective complexity \cite{gellmann1996information}.
-The latter aims to bypass the problem by selecting carefully from the candidate set, but as we have shown, for some numberings the candidate set contains only the universal model.
-\sdr{Explain this is actually a widely recognised problem}
-
-\noindent This tells us that $S^\T_K(x)$ and $S^\T(x)$ can truly grow to a significant part of the total description length.
-While this shows that sophistication, if defined with particular care, is not entirely vacuous, \sdr{hold on, it might still be completely meaningless and/or not invariant} we can shed some light on exactly which strings end up with high sophistication, using the principle of \emph{depth}.
-We will use the following definition:
-
-}
 
@@
 
-\sdr{Move this proof to appendix. Explain the lemma above it rather than below it.}
-This is a technical lemma, due to the double effect of the numbering, which occurs in both the definition of model complexity and the definition of data complexity.
-The \emph{idea} behind the lemma, however, is simple.
-Given a subset $\M'$ of the model class, we can choose the numbering so that representations use a model in $\M'$ perform an arbitrary constant better than all other models.
-This lets us effectively `push' all other models out of the candidate set, so that the sophistication is determined by a model in $\M'$.
-
 \subsection{Overfitting}

### 2015-01-20 Facticity/Facticity/paper.tex
intro tweaked

@@
 
-\title{Three Problems for Sophistication}
+\title{Two Problems for Sophistication}
 
@@
 \end{enumerate}
-For most proposed definitions of sophistication, we can prove that they fail one or more of the conditions above.
+There have been many related proposals for such a measure, that are all based on a \emph{two-part code}: rather than encoding $x$ by means of a single computer program, we encode a \emph{model} for $x$ in the first part of the code, that is interpreted as a representation of $x$'s structural properties.
+The model does not fully specify $x$, but when combined with the second part of the code, which specifies the noise, the original string becomes fully determined.
+Thus, the model can be interpreted as a function mapping the noise to the original object.
+
+For any string $x$, there may be many different two-part representations.
+The total code length can never be less than the Kolmogorov complexity, but it can come close.
+The balance between the two parts may vary; Figure~\ref{fig:todo} provides an illustration.
+The various proposals then define sophistication in terms of the length of the structure part of this two-part code for one of the representations that (almost) achieves the Kolmogorov complexity.
+However, For most of these definitions, we can prove that they fail one or more of the conditions above.
 Vit\'anyi's definition \cite{vitanyi2004meaningful} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we can show that only strings that require an enormous amount of processing to construct, yield a large sophistication.
 
-A valid definition of $S$ must contend with the following issues.
-First, the representation of the model is crucial.
-This is particularly important to those treatments that `open up' the definition of $K$, to consider the representation that witnesses the Kolmogorov complexity.
-In Section~\ref{section:indices}, we discuss the \emph{nickname problem}: the sophistication becomes highly dependent on the chosen programming language, unless it can represent functions with maximum efficiency.
+A valid definition of $S$ must contend with two important issues.
+First, the details of the way the model is encoded are important.
+There are two technically distinct approaches; in one of these one has to deal with the so-called ``nickname problem'' that strangely remains unresolved in several publications.
+These definitions yield a sophistication that is highly dependent on the chosen programming language, unless special care is taken, as discussed in Section~\ref{section:indices}.
 
-The second issue is that of \emph{overfitting}, which we consider in Section~\ref{section:overfitting}.
+The second issue is that of striking the right balance between underfitting and overfitting, which we consider in Section~\ref{section:balance}.
 Overfitting is a common problem in the statistical literature, that refers to the tendency of model selection to choose a very complex model that provides a very good fit to the observed data, but does not generalise well to unseen data.
-For example, in a polynomial regression setting, a polynomial of sufficiently high order can always cover all data points.
-However, if the data was sampled from a straight line, with a small amount of noise, such a polynomial would be far too complex: it would encode all noise inside the model.
-The same can occur for the sophistication.
-We somehow need to make sure that the two-part representation that determines the sophistication does not store noise in the ``structure'' part of the code.
-In statistics, this is typically addressed by penalising complex models.
+In the case of sophistication, overfitting occurs if the model that determines the sophistication contains much or even all of the noise.
+This is a real danger for sophistication, because the models are encoded so efficiently.
+In statistics, overfitting is often addressed by penalising complex models.
 In sophistication, however, such penalties tend to break the balance between structural information and noise, and lead to the opposite problem: underfitting.
@@
 Underfitting (Section~\ref{section:underfitting}) occurs when the selected model is simple, but fails to represent some of the structure present in the data.
-This is a problem for sophistication because the considered models can be so powerful.
+This is also a problem for sophistication because the considered models are so powerful.
 In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language.
-Such \emph{universal models} can be described in a constant amount of bits, and  represent the data with an input of the length of the Kolmogorov complexity, essentially encoding all information as noise.
+Such \emph{universal models} can be described in a constant number of bits, and are able to reconstruct the data with an input of the length of the Kolmogorov complexity.
+Such a two-part representation essentially encodes all information as noise.
 If complex models are penalized, then the problem becomes to make sure that universal models are not \emph{always} preferred for complex data.
@@
 For example, the model class may be restricted to the set of all total recursive functions.
-While excludes universal models, it is questionable whether it adequately solves the problem of underfitting in general.
+While this excludes universal models, it is questionable whether it adequately solves the problem of underfitting in general.
 

### 2015-01-20 Facticity/Facticity/paper.tex
plaatjes

--- 2015-01-20 Facticity/Facticity/paper.tex
@@
 \usepackage{thmtools}
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 A pair $(i,y)$ is a \emph{representation} for $x$ if $\phi_i(y)=x$.
+
+\begin{figure*}[tb]
+  \hspace{-0.9cm}
+  \begin{subfigure}[b]{0.55\textwidth}    
+     \includegraphics[width=\textwidth]{./img/sophistication.pdf}
+      \label{figure:sophistication}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.55\textwidth} 
+     \includegraphics[width=\textwidth]{./img/sophistication-jump.pdf}
+      \label{figure:sophistication-jump}
+  \end{subfigure}
+
+  \caption{\small (left) A schematic representation of the idea of sophistication.
+We can represent the data $x$ in two parts: by describing first, a model, and then the residual information required to specify the data given the model.
+The smallest possible effective description, denoted by a black diagonal, is $C^\M(x)$.
+We accept all representation that come within a given \emph{slack} of this optimum, represented by a dashed line.
+The size of the smallest model within the slack is the sophistication of the data.
+(right) The same image, after a constant perturbation  in te model complexity caused by a change in numbering.}
+  \label{fig:diagram}
+\end{figure*}
 

### 2015-01-21 Facticity/Facticity/paper.tex
plaatjes in intro

--- 2015-01-20 Facticity/Facticity/paper.tex
@@
 \usepackage{thmtools}
+\usepackage{graphicx}
+\usepackage{caption}
 
@@
 
+\begin{figure*}[tb]
+  \hspace{-2.8em}
+  \begin{minipage}{0.55\textwidth}
+     \includegraphics[width=\textwidth]{./img/sophistication.pdf}
+  \end{minipage}
+  \begin{minipage}{0.55\textwidth}
+     \includegraphics[width=\textwidth]{./img/sophistication-jump.pdf}
+  \end{minipage}
+  \caption{\small (left) Two-part representations of $x$.
+The horizontal axis measures the bits used to describe the model, the vertical axis measures the number of bits used to describe the noise.
+The smallest possible effective description, denoted by a black diagonal, is $C^\M(x)$.
+We accept all representations that are close to this optimum, the threshold being represented by a dashed line.
+The size of the smallest model within the slack is the sophistication of the data.
+(right) The same image, after a constant perturbation in te model complexity caused by a change in numbering.}
+  \label{fig:diagram}
+\end{figure*}
+
 For any string $x$, there may be many different two-part representations.
 The total code length can never be less than the Kolmogorov complexity, but it can come close.
-The balance between the two parts may vary; Figure~\ref{fig:todo} provides an illustration.
+The balance between the two parts may vary; Figure~\ref{fig:diagram} provides an illustration.
 The various proposals then define sophistication in terms of the length of the structure part of this two-part code for one of the representations that (almost) achieves the Kolmogorov complexity.

### 2015-01-21 Facticity/Facticity/paper.tex
bewijsje nonstochastic beetje geedit en definitie toegevoegd

@@
 We first show that representions using $\T$ can be reduced to representations in $\F$ with a small overhead.
-\pb{Definitie nonstochastic strings uit \cite[Proposition~I.3 (b)]{gacs2001algorithmic}}
+
+\begin{definition}[Stochastic strings, Definition~1.1 from~\cite{gacs2001algorithmic}]
+  Let $\alpha$ and $\beta$ be natural numbers. A finite binary string $x$ is called \emph{$(\alpha,\beta)$-stochastic} if there exists a finite set $S\subseteq\B$ such that 
+  \[
+  x\in S,~C^\K(S)\le\alpha,~C^\K(x)\ge\log|A|-\beta.
+\]
+\end{definition}
 
@@
 Let $f$ be a total, computable function, such that for some $d$, $f(d) = x$ and let $k = K(f) + |d|$.
-Then there exists a finite set $S$ containing $x$ and a constant $c$, such that $K(S) \leq K(f) + K(|d|) + c$ and $\log |S| \leq |d|$.\label{lemma:total-to-sets}
+Then there exists a finite set $A$ containing $x$ and a constant $c$, such that $K(A) \leq K(f) + K(|d|) + c$ and $\log |A| \leq |d|$.\label{lemma:total-to-sets}
 \end{restatable}
@@
 Let $S = f\left(\{0,1\}^{|d|}\right)$.
-Since $f$ is total, this set can be explicity computed from a description of $f$ and a description of $|d|$, which tells us that there is a constant $c$ such that $K(S) \leq K(f) + K(|d|) + c$.
+Since $f$ is total, this set can be explicity computed from a description of $f$ and a description of $|d|$, which tells us that there is a constant $c$ such that $C^\K(A) \leq C^\K(f) + C^\K(|d|) + c$.
 
-Since $S$ is the image of a set of size $2^{|d|}$ under $f$, we have $\log |S| \leq |d|$.
+Since $A$ is the image of a set of size $2^{|d|}$ under $f$, we have $\log |A| \leq |d|$.
 \end{proof}
@@
 \begin{proof}
+In this proof, for convenience we identify the models indexed by $\F$ with the sets that they encode.
+Thus the two-part code consists of the description of a set $A$, and then an index in this set using $\log|A|$ bits.
 For all $i$, let $x_i$ be the smallest binary string that is not $(i/4, i/4)$-stochastic.
-Since every string is $(\alpha,\alpha)$-stochastic for sufficiently large $\alpha$, this sequence contains infinitely many distinct binary strings.
+Since every string is $(\alpha,\alpha)$-stochastic for sufficiently large $\alpha$, this is a sequence of infinitely many distinct binary strings of increasing length.
 From \cite[Proposition~I.3 (b)]{gacs2001algorithmic} we know that for all sufficiently large $i$ there are strings of length less than $i$ that are not $(i/4, i/4)$-stochastic.
 Now pick any $i$ that is large enough that (a) $|x_i|<i$ and (b) $C^\K(x_i)/5 < i/4$.
-By definition of $(\alpha,\beta)$-stochasticity, if a string is not $(i/4,i/4)$-stochastic, then for every finite set $S$ containing $x$, either (a) $\log|S|\ge C^\K(x_i)+i/4\ge 5C^\K(x_i)/4$, or (b) $C^\K(S)>i/4>C^\K(x_i)/5$.
-In case (a), any two-part description using a uniform code for $S$ is substantially larger than the Kolmogorov complexity, for large enough $i$, so it does not determine the sophistication.
+By definition of $(\alpha,\beta)$-stochasticity, if a string is not $(i/4,i/4)$-stochastic, then for every finite set $A$ containing $x$, either (a) $\log|A|\ge C^\K(x_i)+i/4\ge 5C^\K(x_i)/4$, or (b) $C^\K(A)>i/4>C^\K(x_i)/5$.
+In case (a), any two-part description using $A$ is substantially larger than the Kolmogorov complexity, for large enough $i$, so it does not determine the sophistication.
 Case (b) provides the required lower bound on the complexity of the model.
@@
 Now, take $\T$ as model class and suppose towards contradiction that the sophistication is determined by a total function $f=\phi_j,j\in\T$ with $f(d)=x$ and $C^\K(f)\le i/5$.
-Then by Lemma~\ref{lemma:total-to-sets}, for some fixed constant $c$ there is a finite set $S$ with $\log|S|\le|d|\le C^\K(x_i)$, so (a) above cannot be the case, and $C^\K(S)\le C^\K(f)+C^\K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) for large enough $i$.
+Then by Lemma~\ref{lemma:total-to-sets}, for some fixed constant $c$ there is a finite set $A$ with $\log|A|\le|d|\le C^\K(x_i)$, so (a) above cannot be the case, and $C^\K(A)\le C^\K(f)+C^\K(|d|)+c\le i/5+C^\K(|d|)+c$, which contradicts (b) for large enough $i$.
 Therefore such candidates $f$ cannot exist and all candidates must satisfy $C^\K(f)>i/5>C^\K(x_i)/6$.

### 2015-01-21 Facticity/Facticity/paper.tex
discussion

@@
 The horizontal axis measures the bits used to describe the model, the vertical axis measures the number of bits used to describe the noise.
-The smallest possible effective description, denoted by a black diagonal, is $C^\M(x)$.
+The Kolmogorov complexity $C^\M(x)$, appearing as a black diagonal, provides a lower bound.
 We accept all representations that are close to this optimum, the threshold being represented by a dashed line.
@@
 For any string $x$, there may be many different two-part representations.
-The total code length can never be less than the Kolmogorov complexity, but it can come close.
+The total codelength can never be less than the Kolmogorov complexity, but it can come close.
 The balance between the two parts may vary; Figure~\ref{fig:diagram} provides an illustration.
@@
 \label{section:indices}
-As mentioned in the introduction, the definition of sophistication is sometimes obtained by opening up the definition of Kolmogorov complexity, and measuring the length of the index part of that two-part code.
+In some works, the definition of sophistication is obtained by opening up the definition of Kolmogorov complexity, and measuring the length of the index part of that two-part code.
 We call such a sophistication \emph{index sophistication}.
@@
 This makes these approaches vulnerable to the so-called \emph{nickname problem}: the number of bits required to specify a model becomes very dependent on the numbering.
-This problem is severe: any index sophistication without requirements on the numbering is either bounded, or not invariant:
+This problem is severe: \emph{any} index sophistication without requirements on the numbering is either bounded, or not invariant:
 
@@
   \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use $K$ to measure the complexity of the \emph{model} (as in Definition~\ref{definition:model-complexity}), which is then the first part of a two-part code describing the data.
-With this approach, the total code length is $K(\phi_i)+|y|$, with $\phi_i(y)=x$.
+With this approach, the total codelength is $K(\phi_i)+|y|$, with $\phi_i(y)=x$.
 In this approach, we can construct everything on the basis of a single acceptable numbering, without additional constraints.
@@
 
-So why should we consider this intuition reasonable at all?
-The first assumption of sophistication is that high compressibility and pure randomness are objectively `uninteresting'.
-A more sophisticated signal, such as a television broadcast must contain, in some objective sense, more meaningful information.
-Whether this intuition seems reasonable likely differs per person \sdr{weak}, but sophistication goes one step further: it also suggests that we can objectively separate the information in a string into random and structured.
-\sdr{I think here we should sell the intuition better before we tear it down.}
+We have criticized existing measures of the sophistication of a binary string.
+Our results show that there are technical problems with each of them.
+But that does not in itself mean that it should be impossible to come up with a sound measure.
+The intuition throughout the literature, starting with Kolmogorov's proposal of the structure function, appears to have been that it should be possible to identify noise, and distinguish it from structural information, by its property of being incompressible.
 
-To see whether this intuition is valid; imagine that we sample from the universal distribution by feeding random bits to a universal Turing machine, machine $u$ in the $\phi$-enumeration.
-If this UTM is constructed in the conventional manner, the initial bits will determine the index of a partial computable function $i$, and the rest will be the input $y$ to that machine.
-What is the characteristic model of the resulting data?
-Certainly, if we have to judge based only on the data we cannot exclude the possibility that the data was sampled from $\phi_u$: after all, it was.
+This idea is bold, but not necessarily unreasonable.
+Nevertheless, in this section we offer the opinion that such a clean-cut separation \emph{cannot} be made to work.
+We provide two arguments to support this belief.
+
+For the first argument, we take a generative perspective.
+If data are generated by a computable probability distribution, then a sophistication could be called \emph{consistent} if, given a large enough sample, it accurately reflects the complexity of that distribution.
+While in general we do not wish to make any assumptions about the origin of the data, it can make sense to do so as a sanity check, as we will do here.
+Suppose that the data are generated by a universal distribution, e.g.
+by feeding random bits to a universal partial computable function.
+Let $u$ be the index of the function in the $\phi$-enumeration with the property that $\phi_u(\bar\imath y)=\phi_i(y)$.
+Then the initial random bits will determine the prefix encoded index $\bar\imath$ of the function $\phi_i$ that $\phi_u$ will subsequently emulate, and the remaining bits are used as inputs to $\phi_i$.
+We now ask, what should be the sophistication of data that is generated in this fashion?
+
+Certainly, if we have to judge based only on the data, we cannot exclude the possibility that the data was sampled from $\phi_u$: after all, it was.
 On the other hand, neither can we exclude the possibility that it was generated by $\phi_i$, as again, it was!
-
-Although sampling perspectives are discouraged by proponents of algorithmic statistics, this statistical sanity check illustrates why it seems unjustified to assume that, based on only the data, all but one model can always be disqualified.
-\sdr{Hier stond: There is no reason to assume that a string with high sophistication was sampled from a complex model, so how can a complex model be intrinsic to the data?
-\footnotemark}
-
-\footnotetext{Note that computational depth does not have this problem. Only deep sources produce deep strings.}
+For this reason it seems unjustified to assume that, based on only the data, all but one model can always be disqualified.
 
@@
 We are given a bitstring which can be read as a bitmap image of the painting \emph{Impression of a Sunrise}.
-The candidate set for this string contains various models from very generic to very specific.
+There are many good models for this string, ranging from very generic to very specific.
 The theory behind sophistication suggests that we can choose one of these as the objective, intrinsic model of the data.
@@
 A more specific model might say that it is `an image'.
-Even more specific would be a painting, a Monet, or specifically the painting Impression of a sunrise.
-These are the models we imagine we may find in the candidate set.
-Can we say that the data is intrinsically more of an image than a painting?
+Even more specific would be a painting, a Monet, or specifically the painting Impression of a Sunrise.
+A sound measure of Sophistication should be able to select one of these models as the proper representation of structure in the data, and disqualify the others as either over- or underfitting.
+But how should we be able to say that the data is intrinsically more of an image than a painting?
 More of a Monet than a painting?
-We see no intuitive reason why such a distinction should be possible without making any further assumptions.
+Intuitively such distinctions require further assumptions.
 
-Of course, \sdr{Actually this is not trivial at all} Kolmogorov complexity allows us to say that it is unlikely to be a Jackson Pollock, or a piece of music.
-And, if we were to get a second sample from the same source, we \emph{could} answer the question: we would receive either the same painting again, or perhaps another Monet.
-But, a single sample does not give us enough evidence to choose a single model.
-\sdr{Volgende zin weer zwak, de metafoor wordt backed up door ons voorbeeld hierboven dus is behoorlijk hard} Ultimately, this is just a metaphor, but since the notion of sophistication is driven by intuition, we feel that an intuition to the contrary should serve to sharpen the debate.
+The second argument why we believe sophistication will elude definition is more technical.
+Given an input $x$ and some model class, say $\C$, consider the set of all possible two-part representations of $x$.
+Figure~\ref{fig:diagram} provides an illustration.
+Now, when the numbering of partial computable functions is changed, the codelength of the model part of all these representations will change.
+This is illustrated in the right graph in the figure.
+However, the Invariance Theorem expresses that this change is limited by a constant term that does not depend on $x$.
+So for sufficiently large $x$, the horizontal shift of all the representations will be relatively small.
+\emph{However}, even this small shift can push some representations out of the acceptable region (indicated by the dashed line), and pull others in.
+This may lead to a different representation determining the sophistication, one whose \emph{total} codelength is close to what it was before, but whose \emph{model} codelength can be anywhere between $0$ and $C^\C(x)$.
+If such jumps can occur, the sophistication is not invariant.
+And while we cannot \emph{prove} in general that such jumps do not occur, there seems to be no reason to believe that they don't.
 
 For this reason, we take a skeptical view of sophistication.
-We can offer the following insight, to back up our skepticism.
-As we have seen, candidates within a constant of the optimal representation can occur all over the spectrum, from small to large models.
-A small change in construction, such as a different acceptable numbering, or a change in threshold function, can arbitrarily push models into or out of the candidate set.
-Whereas in Kolmogorov complexity, constant changes have constant effects on the outcome, here constant changes can cause large changes in the outcome.
-\sdr{Not sufficiently clear} While no authors have touched on the issue of invariance, it is related to the issue of instability \cite{antunes2013sophistication,vereshchagin2013algorithmic}.
-We consider invariance the first criterion by which any proposal must be judged, with the burden of proof resting with the proposer, rather than the critic.
+Note that \emph{part} of the theory is fine: there is nothing wrong with evaluating models for the data by comparing their two-part code lengths.
+In fact, the randomness deficiency $\delta(\phi_i,x)=\min\{|y|:\phi_i(y)=x\}-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data --- if $\phi_i$ truly generated the data, then the probability that the random deficiency is more than $k$ bits is less than $2^{-k}$ (proof omitted).
+But fundamental problems arise as soon as hard cut-offs are introduced.
+We also wish to point out that a lot of measures taken in the literature to address problems with sophistication, such as restricting the model class or introducing model penalties, are no longer necessary as soon as such hard cut-offs are avoided.
+What remains is a simple and elegant theory that can disprove most models, leaving not one, but a limited number of interesting candidates.
+
+\section{Conclusion}
+As we have seen, the idea behind sophistication has a rich history.
+It has been proposed many times, by many different authors.
+This tells us two things.
+First, there is a strong intuition behind sophistication, shared by many respected authors, and second, it is not easy to get the definition exactly right.
+ While no authors have touched on the issue of invariance, it is related to the issue of instability \cite{antunes2013sophistication,vereshchagin2013algorithmic}.
+We consider invariance a first criterion that any proposal must satisfy, , with the burden of proof resting with the proposer, rather than the critic.
 
@@
 This creates an artificial dead end for a very valuable area of research.
-
-\sdr{Something about what is useful about the theory}
-\sdr{A picture of the structure function (perhaps sooner as well)}
-\sdr{This has become a discussion section. It does not recap. That's a bit unusual.}
-
-\section{Conclusion}
-As we have seen, the idea behind sophistication has a rich history.
-It has been proposed many times, by many different authors.
-This tells us two things.
-First, there is a strong intuition behind sophistication, shared by many respected authors, and second, it is not easy to get the definition exactly right.
 
@@
 By assumption there must be an $i_0$ such that the above expression is positive for all $i>i_0$.
-From this $i_0$ onwards, the second representation (using $v$) will have a shorter code length, so $\bar\imath p$ cannot achieve the minimum in the definition of the sophistication.
+From this $i_0$ onwards, the second representation (using $v$) will have a shorter codelength, so $\bar\imath p$ cannot achieve the minimum in the definition of the sophistication.
 Consequently, the sophistication is bounded by $\s(x)<|\overline{\imath_0}|$ for all $x$.

### 2015-01-21 Facticity/Facticity/paper.tex
discussion

@@
 
-\section{Discussion}
+\section{Discussion and conclusion}
 \label{section:conclusion}
@@
 But that does not in itself mean that it should be impossible to come up with a sound measure.
-The intuition throughout the literature, starting with Kolmogorov's proposal of the structure function, appears to have been that it should be possible to identify noise, and distinguish it from structural information, by its property of being incompressible.
+The intuition throughout the literature, starting with Kolmogorov's proposal of the structure function, appears to be that it should be possible to identify noise, and distinguish it from structural information, by its property of being incompressible.
 
@@
 If data are generated by a computable probability distribution, then a sophistication could be called \emph{consistent} if, given a large enough sample, it accurately reflects the complexity of that distribution.
-While in general we do not wish to make any assumptions about the origin of the data, it can make sense to do so as a sanity check, as we will do here.
 Suppose that the data are generated by a universal distribution, e.g.
 by feeding random bits to a universal partial computable function.
-Let $u$ be the index of the function in the $\phi$-enumeration with the property that $\phi_u(\bar\imath y)=\phi_i(y)$.
+(While in general we do not wish to make any assumptions about the origin of the data, it can make sense to do so as a sanity check, as we will do here.) Let $u$ be the index of the function in the $\phi$-enumeration with the property that $\phi_u(\bar\imath y)=\phi_i(y)$.
 Then the initial random bits will determine the prefix encoded index $\bar\imath$ of the function $\phi_i$ that $\phi_u$ will subsequently emulate, and the remaining bits are used as inputs to $\phi_i$.
@@
 The theory behind sophistication suggests that we can choose one of these as the objective, intrinsic model of the data.
-Different models say different things: the universal models says that it is `something computable'.
+Different models say different things: the universal model says that it is `something computable'.
 A more specific model might say that it is `an image'.
@@
 If such jumps can occur, the sophistication is not invariant.
-And while we cannot \emph{prove} in general that such jumps do not occur, there seems to be no reason to believe that they don't.
+And while we cannot \emph{prove} in general that such jumps can alwys occur, there seems to be no reason to believe that they don't.
 
@@
 In fact, the randomness deficiency $\delta(\phi_i,x)=\min\{|y|:\phi_i(y)=x\}-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data --- if $\phi_i$ truly generated the data, then the probability that the random deficiency is more than $k$ bits is less than $2^{-k}$ (proof omitted).
+In the Monet example above, this will allow us to disqualify the model expressing that the data is actually, say, a recording of Jazz music.
+
 But fundamental problems arise as soon as hard cut-offs are introduced.
-We also wish to point out that a lot of measures taken in the literature to address problems with sophistication, such as restricting the model class or introducing model penalties, are no longer necessary as soon as such hard cut-offs are avoided.
-What remains is a simple and elegant theory that can disprove most models, leaving not one, but a limited number of interesting candidates.
-
-\section{Conclusion}
-As we have seen, the idea behind sophistication has a rich history.
-It has been proposed many times, by many different authors.
-This tells us two things.
-First, there is a strong intuition behind sophistication, shared by many respected authors, and second, it is not easy to get the definition exactly right.
- While no authors have touched on the issue of invariance, it is related to the issue of instability \cite{antunes2013sophistication,vereshchagin2013algorithmic}.
-We consider invariance a first criterion that any proposal must satisfy, , with the burden of proof resting with the proposer, rather than the critic.
-
-Finally, we note that the use of total functions has a very subtle effect on the discussion surrounding sophistication.
-Since the total functions are not enumerable, the value of the sophistication relies strongly on the difference between the Kolmogorov complexity and the Kolmogorov complexity restricted to total functions.
-This gap is exceedingly difficult to capture, because the total functions cannot be enumerated.
-This is dangerous: if such definitions cannot be proved wrong or right, they will stand as the ad-hoc official view, while the restrictions used to elude proofs of incorrectness make them very difficult to build on.
-This creates an artificial dead end for a very valuable area of research.
+In our opinion, a lot of measures taken in the literature to address problems with sophistication, such as restricting the model class or introducing model penalties, complicate the method and make problems harder to analyse, without actually addressing the fundamental issue.
+This is dangerous: if such ad-hoc ``fixes'' result in a theory that is hard to prove either wrong or right, then this can create an artificial dead end for a valuable area of research.
+When hard cut-offs are avoided on the other hand, all such measures are suddenly no longer necessary.
+What remains is a simple and elegant theory that can be used to sift through all possible models, disproving most while retaining a select number of interesting candidates for our further consideration.
 

### 2015-01-21 Facticity/Facticity/paper.tex
...

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 We describe three fundamental problems with all existing proposals, showing each  to be \emph{unsound}.
-Consequently, we put forward the view that the problem is fundamental, it may be impossible to objectively quantify the sophistication.
+Consequently, we put forward the view that the problem is fundamental: it may be impossible to objectively quantify the sophistication.
 \end{abstract}
@@
 There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would consider complexity.
-For example, a sequence of a million coin flips will almost certainly have maximal Kolmogoroc complexity, even though there is nothing complex about flipping a lot of coins---it's just hard work.
+For example, a sequence of a million coin flips will almost certainly have maximal Kolmogorov complexity, even though there is nothing complex about flipping a lot of coins---it's just a lot of work.
 For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data.
@@
 
-The Kolmogorov complexity of a binary string $x$, denoted $K(x)$, is (roughly) the length of the shortest computer program that will output $x$.
-This length depends on which programming language is used, but, as shown by a core result known as the Invariance Theorem, the language affects the result only by a constant, independent of $x$.
+The Kolmogorov complexity of a binary string $x$, is (roughly) the length of the shortest computer program that will output $x$.
+This length depends on which programming language is used, but, as shown by a central result, known as the Invariance Theorem\cite[Section~2.1]{li1993introduction}, the language affects the result only by a constant, independent of $x$.
 This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings.

### 2015-01-21 Facticity/Facticity/paper.tex
../

@@
 
-A robust definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees. We are interested in a measure satisfying the following requirements:
+A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees. We are interested in a measure satisfying the following requirements:
 
@@
   \caption{\small (left) Two-part representations of $x$.
-The horizontal axis measures the bits used to describe the model, the vertical axis measures the number of bits used to describe the noise.
+The horizontal axis measures the bits used to describe the model, the vertical axis measures the bits required to described the data given the model.
 The Kolmogorov complexity $C^\M(x)$, appearing as a black diagonal, provides a lower bound.
-We accept all representations that are close to this optimum, the threshold being represented by a dashed line.
-The size of the smallest model within the slack is the sophistication of the data.
-(right) The same image, after a constant perturbation in te model complexity caused by a change in numbering.}
+We accept all representations that are close to this optimum, with the threshold represented by a dashed line.
+The size of the smallest model below the threshold is the sophistication of the data.
+(right) The same image, after a constant perturbation in the model complexity caused by a change in numbering.}
   \label{fig:diagram}
@@
 The second issue is that of striking the right balance between underfitting and overfitting, which we consider in Section~\ref{section:balance}.
-Overfitting is a common problem in the statistical literature, that refers to the tendency of model selection to choose a very complex model that provides a very good fit to the observed data, but does not generalise well to unseen data.
+Overfitting is a common problem in the statistical literature, that refers to the tendency of model selection to choose a complex model that provides a very good fit to the observed data, but does not generalise well to unseen data.
 In the case of sophistication, overfitting occurs if the model that determines the sophistication contains much or even all of the noise.
@@
 
-Underfitting (Section~\ref{section:underfitting}) occurs when the selected model is simple, but fails to represent some of the structure present in the data.
+Underfitting occurs when the selected model is simple, but fails to represent some of the structure present in the data.
 This is also a problem for sophistication because the considered models are so powerful.
 In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language.
-Such \emph{universal models} can be described in a constant number of bits, and are able to reconstruct the data with an input of the length of the Kolmogorov complexity.
+Such \emph{universal models} are \emph{simple} models in the language of sophistication, since they can be described with a small number of bits, yet are able to represent the data within a codelength a constant away from the Kolmogorov complexity.
 Such a two-part representation essentially encodes all information as noise.
@@
 Finally, in the discussion in Section~\ref{section:conclusion} we argue that while two-part coding can yield useful insights into the structure of the data and allows us to identify some models as poor representations, it is probably not possible to uniquely separate structure from noise and identify a \emph{single} model as ``best'': in general many models, of vastly different complexities, may be reasonable representations.
-Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that the data allows for multiple equivalent interpretations, and that there is no such thing as sophistication.
+Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that the data allows for multiple equivalent interpretations of which information is structured, and which is random, and that there is no such thing as sophistication.
 
@@
 As we shall see, this version behaves very differently from regular sophistication, and is closer to computational depth.
-Vereshchagin \cite{vereshchagin2013algorithmic}, working in the finite set setting,  suggests additional restrictions on models that are allowed to determine the sophistication, in hopes of curbing the instability issue.
+Vereshchagin \cite{vereshchagin2013algorithmic} suggests additional restrictions on models that are allowed to determine the sophistication, in hopes of curbing the instability issue.
 Adriaans' \emph{Facticity}, introduced in 2012\cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions.

### 2015-01-21 Facticity/Facticity/paper.tex
...

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 
-\title{Two Problems for Sophistication}
+\title{Three Problems for Sophistication}
 
@@
 There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would consider complexity.
-For example, a sequence of a million coin flips will almost certainly have maximal Kolmogorov complexity, even though there is nothing complex about flipping a lot of coins---it's just a lot of work.
+For example, a sequence of a million coin flips will almost certainly have maximal Kolmogoroc complexity, even though there is nothing complex about flipping a lot of coins---it's just hard work.
 For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data.
@@
 
-The Kolmogorov complexity of a binary string $x$, is (roughly) the length of the shortest computer program that will output $x$.
-This length depends on which programming language is used, but, as shown by a central result, known as the Invariance Theorem\cite[Section~2.1]{li1993introduction}, the language affects the result only by a constant, independent of $x$.
+The Kolmogorov complexity of a binary string $x$, denoted $K(x)$, is (roughly) the length of the shortest computer program that will output $x$.
+This length depends on which programming language is used, but, as shown by a core result known as the Invariance Theorem, the language affects the result only by a constant, independent of $x$.
 This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings.
@@
 In our opinion, a lot of measures taken in the literature to address problems with sophistication, such as restricting the model class or introducing model penalties, complicate the method and make problems harder to analyse, without actually addressing the fundamental issue.
-This is dangerous: if such ad-hoc ``fixes'' result in a theory that is hard to prove either wrong or right, then this can create an artificial dead end for a valuable area of research.
+This is dangerous: if such ad-hoc fixes result in a theory that is hard to prove either wrong or right, then this can create an artificial dead end for a valuable area of research.
 When hard cut-offs are avoided on the other hand, all such measures are suddenly no longer necessary.

### 2015-01-21 Facticity/Facticity/paper.tex
...

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 \item $\C=\N$, the indices of all partial computable functions
-\item $\T=\{i:\tn{$\phi_i$ is total}\}$. Note that $\T$ is not enumerable.
+\item $\T=\{i:\tn{$\phi_i$ is total}\}$. Note that $\T$ is not computably enumerable.
 \item $\K$ is an enumerable set such that $\{\phi_i:i\in\K\}$ is the set of all partial computable prefix functions
@@
 \end{enumerate}
-
 We sometimes need to map binary strings to a prefix-free set.
-We denote by $\bar x$ the prefix-encoded representation for $x$.
-We require that the mapping satisfies $|\bar{x}| = |x|+O(\log(|x|)$ (see eg.
+We denote by $\br{x}$ the prefix-encoded representation for $x$.
+We require that the mapping satisfies $|\br{x}| = |x|+O(\log(|x|)$ (see eg.
 \cite[Section~1.4]{li1993introduction}).
@@
 \]
-The model class $\C$ gives us the plain Kolmogorov complexity and the $\K$ gives us the prefix-free Kolmogorov complexity.
-The complexity is invariant to the numbering for these model classes, so we may omit it from the superscript.
+The model class $\C$ gives us the plain Kolmogorov complexity and $\K$ gives us the prefix-free Kolmogorov complexity.
+For $\C$ and $\K$, the complexity is invariant to the numbering, so we may omit it from the superscript.
 
@@
 
-In these works, no constraints are placed on the used numbering.
+In these works, no constraints are placed on the numbering.
 This makes these approaches vulnerable to the so-called \emph{nickname problem}: the number of bits required to specify a model becomes very dependent on the numbering.

### 2015-01-21 Facticity/Facticity/paper.tex
...

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 \section{Introduction}
-Komogorov complexity gives us a sound definition of the amount of information contained in a bitstring.
+Komogorov complexity gives us a sound definition of the amount of information contained in a binary string.
 There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would consider complexity.
@@
 For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data.
-This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication},\footnotemark were proposed.
+This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication}, were proposed.
 In this paper, we investigate three serious problems with these definitions, and show why they are unsatisfactory.
@@
 
-The Kolmogorov complexity of a binary string $x$, is (roughly) the length of the shortest computer program that will output $x$.
-This length depends on which programming language is used, but, as shown by a central result, known as the Invariance Theorem\cite[Section~2.1]{li1993introduction}, the language affects the result only by a constant, independent of $x$.
+The Kolmogorov complexity $C(x)$ of a binary string $x$, is (roughly) the length of the shortest computer program that will output $x$.
+This length depends on which programming language is used, but, as shown by a central result known as the Invariance Theorem\cite[Section~2.1]{li1993introduction}, the language affects the value only by a constant, independent of $x$.
 This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings.
@@
 \begin{enumerate}
-\item Since it measures information, $S(x)$ should count the bits required for an effective description of the structural properties of a binary string.
+\item $S(x)$ should count the bits required for an effective description of the structural properties of a binary string.
 \item An analogue of invariance should hold for sophistication: at the very least, there should be strict limits on the degree to which a change in the programming language can affect the sophistication.
 \item There should be no constant $c$ such that $S(x)\le c$ for every input $x$.
-If sophistication is bounded, then knowing its value under one programming language provides no constraints on its value under another language.
+If sophistication is bounded, then knowing its value under one programming language provides no constraints on its value under another language (except that it's also bounded).
 Moreover, a bounded sophistication would be at odds with the intuition that there is no limit to the amount of structure that a binary string can exhibit.
-\item Similarly, there should be no constant $c$ such that $K(x)-S(x)\le c$ for all $x$, because then sophistication would be equivalent to Kolmogorov complexity. 
+\item Similarly, there should be no constant $c$ such that $C(x)-S(x)\le c$ for all $x$, because then sophistication would be equivalent to Kolmogorov complexity. 
 \end{enumerate}
@@
 The various proposals then define sophistication in terms of the length of the structure part of this two-part code for one of the representations that (almost) achieves the Kolmogorov complexity.
-However, For most of these definitions, we can prove that they fail one or more of the conditions above.
-Vit\'anyi's definition \cite{vitanyi2004meaningful} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we can show that only strings that require an enormous amount of processing to construct, yield a large sophistication.
+However, for most of these definitions, we can prove that they fail one or more of the conditions above.
+Vereshchagin and Vit\'anyi's definition \cite{vitanyi2004meaningful} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we can show that only strings that require an enormous amount of processing to construct, yield a large sophistication.
 
@@
 In the case of sophistication, overfitting occurs if the model that determines the sophistication contains much or even all of the noise.
-This is a real danger for sophistication, because the models are encoded so efficiently.
+This is a real danger, because the models are encoded so efficiently.
 In statistics, overfitting is often addressed by penalising complex models.
@@
 In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language.
-Such \emph{universal models} are \emph{simple} models in the language of sophistication, since they can be described with a small number of bits, yet are able to represent the data within a codelength a constant away from the Kolmogorov complexity.
+Such \emph{universal models} are \emph{simple} in the language of sophistication, since they can be described with a relatively small number of bits, yet are able to represent the data within a codelength a constant away from the Kolmogorov complexity.
 Such a two-part representation essentially encodes all information as noise.
@@
   \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use $K$ to measure the complexity of the \emph{model} (as in Definition~\ref{definition:model-complexity}), which is then the first part of a two-part code describing the data.
-With this approach, the total codelength is $K(\phi_i)+|y|$, with $\phi_i(y)=x$.
+With this approach, the total codelength is $C^\K(\phi_i)+|y|$, with $\phi_i(y)=x$.
 In this approach, we can construct everything on the basis of a single acceptable numbering, without additional constraints.

### 2015-01-21 Facticity/Facticity/paper.tex
section 3

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 There are acceptable numberings $\psi$ and $\xi$ such that for all $x$:
-\[|s^\xi(x) - s^\psi(x)|\geq \frac{1}{2}\min\{s^\psi(x),s^\xi(x)\}\p\]
+\[|s^\psi(x) - s^\xi(x)|\geq \frac{1}{2}\min\{s^\psi(x),s^\xi(x)\}\p\]
 \end{lemma}
@@
 \end{proof}
-While the invariance theorem expresses that inefficient indices do not affect Kolmogorov complexity too much, this robustness property does \emph{not} carry over to index sophistication.
-In other words, the length of a model index does \emph{not} represent its complexity.
+The Kolmogorov complexity does not suffer from inefficient indices.
+A switch to to a more efficient index will only cost a constant amount of bits in the complexity.
+However, as this lemma shows, it creates a massive change in the size of the model that witnesses the complexity.
+The plain length of the index in a given acceptable numbering may be a very poor indicator of model complexity.
 
@@
 The Kolmogorov complexity uses representations of the form $\bar\imath y$, where the bar denotes some straightforward prefix encoding to delimit the model description $i$ from its input $y$.
-If we define a second prefix encoding $\tilde{\imath}$, which is always better than our default by some slowly growing term, such as $O(\log\log\log i)$, we can define a second representation $\bar u \tilde \imath y$, which takes a constant amount of bits $|\bar u|$ to switch to a different universal model, but gains more than a constant, for sufficiently complex strings, because a different prefix function is used.
-Ultimately, $u$ would always be chosen as a model for sufficiently complex $x$.
+If we define a second prefix encoding $\tilde{\imath}$, with $|\bar\imath|-|\tilde\imath|$ unbounded, we can define a second representation $\bar u \tilde \imath y$, at a constant overhead $|\br{u}|$, and gain more than $|\br{u}|$ for sufficiently complex strings.
+Ultimately, $u$ will always be chosen as a model for sufficiently complex $x$, and the sophistication becomes bounded.
 
-Can we make our prefix encoding so efficient that no other encoding can undercut it by more than a constant amount?
-One self-delimiting description of $\phi_i$ is the one used in $C^\K(\phi_i)$.
-By definition, this is the smallest effective, self-delimiting description of $\phi_i$.
-Let $i^*$ be a the shortest self-delimiting program for $\phi_i$.
-If our prefix function is less efficient than $C^\K(x)$ by an unbounded amount, the effective representation $i^*y$ will always beat $\bar\imath y$ for complex enough data.
+Can we make our prefix encoding so efficient that no other encoding can undercut it?
+By definition, $C^{\M}(m)$ this is the smallest effective, self-delimiting description of $\phi_i$.
+We cannot compute the prefix code from the data, but this is not necessary to define a representation.
+Let $i^*$ be the shortest self-delimiting program for $\phi_i$, then $i^*y$ is a decodable representation for $x$.
+If our prefix function is less efficient than $C^\K(m)$ by an unbounded amount, $i^*y$ will always beat $\bar\imath y$ for complex enough data.
 We conjecture that no computable prefix encoding function can avoid this issue:
@@
 \end{conjecture}
-The following lemma shows that for all index sophistications using an inefficient prefix encoding will fail. This includes \cite{adriaans2012facticity}.
+The following lemma shows that for index sophistications using a faithful index, but an inefficient prefix encoding are bounded. This includes \cite{adriaans2012facticity}.
 \begin{restatable}{lemma}{ineffprefix}
@@
 \end{restatable}
-
 If our conjecture is correct, the index sophistication is always bounded.
-The only way to avoid the problem is to use a the self-delimiting program of length $C^\K(x)$ to describe the model.
+The only way to avoid the problem is to use a self-delimiting program of length $C^\K(m)$ to describe the model.
 Any inefficieny in the model description, and a second represention using a universal model with a more efficient model description will win, causing a bounded sophistication.
 
-For this reason we will using the following definition of sophistication in the remainder of this article: 
+For this reason we will use the following definition of sophistication in the remainder of this article: 
 \begin{equation}\label{eq:soph}

### 2015-01-21 Facticity/Facticity/paper.tex
intro squeezed

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 Since then, many variations of this idea have been proposed, for which we use \emph{sophistication} as an umbrella term.
-We describe three fundamental problems with all existing proposals, showing each  to be \emph{unsound}.
+We describe two fundamental problems with all existing proposals, showing each  to be \emph{unsound}.
 Consequently, we put forward the view that the problem is fundamental: it may be impossible to objectively quantify the sophistication.
@@
 
-\section{Introduction}
+\section{Introduction}\enlargethispage{4ex}
 Komogorov complexity gives us a sound definition of the amount of information contained in a binary string.
-There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would consider complexity.
+It is, however, not what most people would consider complexity.
 For example, a sequence of a million coin flips will almost certainly have maximal Kolmogorov complexity, even though there is nothing complex about flipping a lot of coins---it's just a lot of work.
-For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data.
-This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication}, were proposed.
-In this paper, we investigate three serious problems with these definitions, and show why they are unsatisfactory.
-We believe that the problems are fundamental, and provide arguments why such attempts to define sophistication must fail.
+Many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data: its \emph{sophistication}.
+In this paper, we investigate two serious problems with sophistication measures, and show why they are unsatisfactory.
+We believe that the problems are fundamental, and provide arguments why attempts to define sophistication must fail.
 
@@
 This length depends on which programming language is used, but, as shown by a central result known as the Invariance Theorem\cite[Section~2.1]{li1993introduction}, the language affects the value only by a constant, independent of $x$.
-This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings.
-We can see it as \emph{a property of the data}.
-
-A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees. We are interested in a measure satisfying the following requirements:
+This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings, a \emph{property} of the data.
+A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees:
 
@@
 \item $S(x)$ should count the bits required for an effective description of the structural properties of a binary string.
-\item An analogue of invariance should hold for sophistication: at the very least, there should be strict limits on the degree to which a change in the programming language can affect the sophistication.
+\item An analogue of invariance should hold: there must be strict limits on how much sophistication can be affected by a change in programming language.
 \item There should be no constant $c$ such that $S(x)\le c$ for every input $x$.
@@
 
+For any string $x$, there may be many different two-part representations.
+The total codelength can never be less than the Kolmogorov complexity, but it can come close.
+The balance between the two parts may vary; Figure~\ref{fig:diagram} provides an illustration.
+The various proposals then define sophistication in terms of the length of the structure part of this two-part code for one of the representations that (almost) achieves the Kolmogorov complexity.
+However, for most of these definitions, we can prove that they fail one or more of the conditions above.
+Vereshchagin and Vit\'anyi's definition \cite{vitanyi2004meaningful} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we show that their method only assigns substantial sophistication to strings that require an enormous amount of processing to construct.
+
 \begin{figure*}[tb]
@@
   \end{minipage}
-  \caption{\small (left) Two-part representations of $x$.
-The horizontal axis measures the bits used to describe the model, the vertical axis measures the bits required to described the data given the model.
-The Kolmogorov complexity $C^\M(x)$, appearing as a black diagonal, provides a lower bound.
-We accept all representations that are close to this optimum, with the threshold represented by a dashed line.
+  \caption{\small (left) Two-part representations of $x$ by the two components of their codelength.
+The Kolmogorov complexity $C^\M(x)$, appearing as a black diagonal, provides a lower bound on the total codelength.
+We consider only representations that are close to this optimum, with the threshold represented by a dashed line.
 The size of the smallest model below the threshold is the sophistication of the data.
@@
 \end{figure*}
-
-For any string $x$, there may be many different two-part representations.
-The total codelength can never be less than the Kolmogorov complexity, but it can come close.
-The balance between the two parts may vary; Figure~\ref{fig:diagram} provides an illustration.
-The various proposals then define sophistication in terms of the length of the structure part of this two-part code for one of the representations that (almost) achieves the Kolmogorov complexity.
-However, for most of these definitions, we can prove that they fail one or more of the conditions above.
-Vereshchagin and Vit\'anyi's definition \cite{vitanyi2004meaningful} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we can show that only strings that require an enormous amount of processing to construct, yield a large sophistication.
 
@@
 The idea of separating structure from noise originated with Kolmogorov's \emph{structure function} \cite{cover1985kolmogorov}.
-In this theory, the data is described as finite set together with an index in the set.
+In this theory, the data $x$ is described as finite set $A$ that contains $x$, together with the index of $x$ in $A$.
 The complexity of the set determines the sophistication.
@@
 The model class $\C$ gives us the plain Kolmogorov complexity and $\K$ gives us the prefix-free Kolmogorov complexity.
-For $\C$ and $\K$, the complexity is invariant to the numbering, so we may omit it from the superscript.
+For $\C$ and $\K$, the complexity is invariant under a change of numbering, which may therefore be omitted from the superscript.
 
-A pair $(i,y)$ is a \emph{representation} for $x$ if $\phi_i(y)=x$.
+A pair $(i,y)$ is a \emph{representation} for $x$ if $\phi_i(y)=x$. \sdr{Gebruiken we dit?}
 

### 2015-01-21 Facticity/Facticity/paper.tex
...

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 If the description of the model $m$ is less efficient than $C^\K(m)$ by anything more than a constant, representations with a bounded model size will always win out in the limit.
-In light of this balance issues, the most natural definition of sophistication is $S^{\K,\psi,c}$.
-In this setting, we know that there always exist candidate models---those models $i$ with $i \in \K$ and $C^{i}(x) \leq C^\K(x) + c$---for which al information except a constant amount is stored in the data, and candidates for which all information is stored in the model.
-This tells us that $S^\K$ is \emph{balanced}: there is no inefficiency in the representation forcing information in or out of the model.
-The flipside of having a balanced sophistication is that invariance is easy to show.
-We will show that there are numberings for which the universal models always determine the $S^\K$ and numberings for which the singletons always determine $S^\K$.
+In light of such balance issues, the most natural definition of sophistication is $S^{\K,\psi,c}$.
+In this setting, we know that there always exist candidate models---those models $i$ with $i \in \K$ and $C^{\{i\}}(x) \leq C^\K(x) + c$---for which all but a constant amount of information is stored in the data, and candidates for which all information is stored in the model.
+This tells us that $S^\K$ is \emph{balanced}: there is no inefficiency in the description of either the data or the model, forcing information one way or the other.
+The flipside of having a balanced sophistication is that invariance is easy to show: there are numberings for which the universal models always determine the $S^\K$ and numberings for which the singletons always determine $S^\K$.
 	The following lemma generalizes the principle: 
@@
 \label{lemma:thecoolone}
-  Let $\psi$ be any acceptable enumeration of the partial recursive functions.
-  Let $\M$ be any model class, let $\X$ be any set of binary sequences and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e.
-if $f=\psi_{D(p)}$ then $p$ is a $D$-description of $f$.
-Let $\M'=\{\psi_i:i\in\tn{range}(D)\}$.
+  Let $\M$ be any model class, let $\X$ be any set of binary strings and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\phi$, i.e.
+if $m=\phi_{D(p)}$ then $p$ is a $D$-description of $m$.
+Let $\M'=\{\phi_i:i\in\tn{range}(D)\}$.
 Further assume there is a constant $c$ such that:\\
-\-\hspace{1cm}(1)$\forall_{f\in\M'}:\min\{|p|:\psi_{D(p)}=f\}\le K^\psi(f)+c$\\
-\-\hspace{1cm}(2)$\forall_{x\in\X}:L^{\M',\psi}(x)-L^{\M,\psi}(x)\le c$.\\
-Then there is an enumeration $\phi$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
+\-\hspace{1cm}(1) $\forall_{m\in\M'}:\min\{|p|:\phi_{D(p)}=m\}\le C^{\K,\phi}(m)+c$\\
+\-\hspace{1cm}(2) $\forall_{x\in\X}:C^{\M',\phi}(x)-C^{\M,\phi}(x)\le c$.\\
+Then there is an enumeration $\psi$ of the partial recursive functions such that $S^{\M,\psi}(x) = |\br{0}|+S^{\M',\phi}(x)$ for all $x\in\X$.
 \end{restatable}
+This is a technical lemma, but the \emph{idea} behind the it is simple: given a subset $\M'$ of the model class, we can choose the numbering so that representations use a model in $\M'$ perform an arbitrary constant better than all other models.
+This lets us effectively `push' all other models out of the candidate set, so that the sophistication is determined by a model in $\M'$.
 
-This is a technical lemma, due to the double effect of the numbering, which occurs in both the definition of model complexity and the definition of data complexity.
-The \emph{idea} behind the lemma, however, is simple.
-Given a subset $\M'$ of the model class, we can choose the numbering so that representations use a model in $\M'$ perform an arbitrary constant better than all other models.
-This lets us effectively `push' all other models out of the candidate set, so that the sophistication is determined by a model in $\M'$.- can make it equal to $C^K$
-
-This allows us to ensure that for some numberings, the unversal model always determines the sophistication, making is bounded:
+This allows us to ensure that for some numberings, a universal model always determines the sophistication, making it bounded:
 
 \begin{restatable}[underfitting]{lemma}{underfitting}
-Let $\M$ be a model class containing a universal model $u$, with the property that $\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p$. Then, there exist numberings such that $\s^\M$ is bounded.
+Let $\M$ be a model class containing a universal model $\phi_u$, with the property that $\exists c \forall i \in \M : C^{\{u\}}(x) \leq C^{\{i\}}(x) + c \p$.
+Then, there exist numberings such that $\s^\M$ is bounded.
 \end{restatable}
-For other numberings, the singletons determine the sophistication, making it equal to the complexity:
+\noindent For other numberings, the singletons determine the sophistication, making it equal to the complexity:
 \begin{restatable}[Overfitting]{lemma}{overfitting}
-Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$.
-Then there is an enumeration $\phi$ of the prefix partial recursive functions, and a constant $c$, such that
+Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $i\in\M$ with $\phi_i(\epsilon)=x$.
+Then there is a numbering $\psi$, and a constant $c$, such that for all $x\in\X$
 \[
-K(x)-S^{\M,\phi}(x)\le c
-\]
-for all $x\in\X$.
+C^\M(x)-S^{\M,\psi}(x)\leq c
+\]\label{lemma:overfitting}
 \end{restatable}
-This shows that the price we pay for balance is absolute invariance.
+\noindent This shows that the price we pay for balance is absolute invariance.
 All information can be seen as structure of noise, depending on the numbering.
-To avoid these issues, existing proposals upset the balance to exclude, or penalize at least the singleton models and the universal models.
+To avoid these issues, existing proposals upset the balance to exclude or penalize at least the singleton models and the universal models.
 
@@
 
-The existing proposals that do not avoid overfitting as the structure function \cite{cover1985kolmogorov}, the sophistication as defined in \cite{mota2013sophistication}, the miinimal sufficient statistic \cite{gacs2001algorithmic} and Vereshchagin's strongly  sufficient statistic \cite{vereshchagin2013algorithmic}.
+The existing proposals that do not avoid overfitting are the structure function \cite{cover1985kolmogorov}, the sophistication as defined in \cite{mota2013sophistication}, the minimal sufficient algorithmic statistic \cite{gacs2001algorithmic} and Vereshchagin's strongly sufficient statistic \cite{vereshchagin2013algorithmic}.
+It is widely recognized that singleton models are always candidates in these settings, but not, it seems that they may be the only candidates for any string, as Lemma~\ref{lemma:overfitting} shows.
+
 It may be argued that the constant determining the candidates should be allowed to depend on the numbering, but this dependence has not been mentioned in the literature and there is no clear way to determine what such a constant should be, on the basis of a given numbering.
@@
 As we have seen, a strong penalty, such as the one imposed by an inefficient prefix encoding of the model, will immediately cause underfitting.
-A more subtle approach is to use $C$ as a model class, allowing representations that are not self-delimiting.
+A more subtle approach is to use $\C$ as a model class, allowing representations that are not self-delimiting.
 This eliminates at least the singletons, as placing all information in the model automatically creates a self delimiting representation.
@@
 This approach is taken by Vit\'anyi \cite{vitanyi2004meaningful} and by Adriaans \cite{adriaans2012facticity}.
+The price we pay is that the models can no longer be equated with probability measures, removing the link to traditional statistics and probability theory.
 
@@
 \begin{restatable}{theorem}{dogfood}
-There are infinitely many $x$ with $\s^{\F,\phi,c}(x) \geq C^{\K}(x)/5$ and $\s^{\T,\phi,c}(x) \geq C^{\K}(x)/6$. 
+There are infinitely many $x$ with 
+\[
+	\s^{\F,\phi,c}(x) \geq C^{\K}(x)/5  \text{\;\;\;and\;\;\;} \s^{\T,\phi,c}(x) \geq C^{\K}(x)/6\p
+\] 
 \end{restatable}
@@
 But clearly, this only exacerbates the overfitting problem.
-Only one model class eliminates both the singletons and the unversal model: $S^\T(x)$.
-The only proposal that uses an efficient model representation, excludes the universal models and excludes the singletons is Vit\'anyi's \cite{vitanyi2004meaningful}.
-While this invalidates straightforward proofs of boundedness, there is no suggestion that $S^\T(x)$ is actually invariant.
+Only one model class eliminates both the singletons and the universal model: $\T$.
+The only proposal that uses an efficient model representation \emph{and} excludes the universal models \emph{and} excludes the singletons is Vit\'anyi's \cite{vitanyi2004meaningful}.
+While this invalidates straightforward proofs of boundedness, there is no suggestion that $S^{\T, \phi, c}(x)$ is actually invariant.
 
-While under $S^{\T, \phi, c}(x)$ there are clearly string with high sophistication, these may not conform to the original intuition that motivated sophistication.
+While under $S^{\T, \phi, c}(x)$ there are strings with high sophistication, these may not conform to the original intuition that motivated sophistication.
 To show this, we require the concept of depth:
@@
 Let $U^t$ be a simulation of this machine, which is allowed to run for at most $t$ steps, and returns $0$ if it has not yet finished at that point.
-Let $C^t_\M(x) = \min\{|\bar\imath y| : U^t(\bar\imath y) = x, \phi_i \in \M\}$.
+Let $C^\M_t(x) = \min\{|\bar\imath y| : U^t(\bar\imath y) = x, \phi_i \in \M\}$.
 
-The \emph{$c$-depth} $d^\M_c(x)$ of a string is $d^\M_c(x) = \min \left\{t : C^\M_t(x) - C^\M(x) \leq c \right\}$.
+The \emph{$c$-depth} is $d^{\M,c}(x) = \min \left\{t : C^\M_t(x) - C^\M(x) \leq c \right\}$.
 \end{definition}
-
 Deep strings are those that can only be optimally compressed with a great investment of time.
@@
 Let $A(n)$ be the single-argument Ackermann function and $c$ some arbitrary constant.
-There are numberings such that for all strings with depth $d^\C_c(x) \leq A(C(x))$, $\s^\T(x)$, $s^\P(x)$ and $\s^\F$ are bounded.
+There are numberings such that for all strings with depth $d^{\C,c}(x) \leq A(C^\C(x))$, $\s^{\T, \psi, c}$ is bounded.
+The same is true for $\s^{\F}$
 \end{restatable}
@@
 Consider a string that is typical for a low-depth model, say some high-complexity Markov chain.
-Under sophistication, no matter how complex the markov chain, the sophistication would stay the same.
+Under sophistication, no matter how complex the Markov chain, the sophistication would stay the same.
 Any structure that is simple enough to be exploited within the time bound of the Ackermann function (or some other fast-growing, simply stated bound) will be counted as noise.

### 2015-01-21 Facticity/Facticity/paper.tex

@@
 
-\section{Introduction}
-Komogorov complexity gives us a sound definition of the amount of information contained in a bitstring.
-There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would consider complexity.
+\section{Introduction}\enlargethispage{4ex}
+Komogorov complexity gives us a sound definition of the amount of information contained in a binary string.
+It is, however, not what most people would consider complexity.
 For example, a sequence of a million coin flips will almost certainly have maximal Kolmogorov complexity, even though there is nothing complex about flipping a lot of coins---it's just a lot of work.
-For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data.
-This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication},\footnotemark were proposed.
-In this paper, we investigate three serious problems with these definitions, and show why they are unsatisfactory.
-We believe that the problems are fundamental, and provide arguments why such attempts to define sophistication must fail.
+Many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data: its \emph{sophistication}.
+In this paper, we investigate two serious problems with sophistication measures, and show why they are unsatisfactory.
+We believe that the problems are fundamental, and provide arguments why attempts to define sophistication must fail.
 
-The Kolmogorov complexity of a binary string $x$, is (roughly) the length of the shortest computer program that will output $x$.
-This length depends on which programming language is used, but, as shown by a central result, known as the Invariance Theorem\cite[Section~2.1]{li1993introduction}, the language affects the result only by a constant, independent of $x$.
-This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings.
-We can see it as \emph{a property of the data}.
-
-A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees. We are interested in a measure satisfying the following requirements:
+The Kolmogorov complexity $C(x)$ of a binary string $x$, is (roughly) the length of the shortest computer program that will output $x$.
+This length depends on which programming language is used, but, as shown by a central result known as the Invariance Theorem\cite[Section~2.1]{li1993introduction}, the language affects the value only by a constant, independent of $x$.
+This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings, a \emph{property} of the data.
+A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees:
 
 \begin{enumerate}
-\item Since it measures information, $S(x)$ should count the bits required for an effective description of the structural properties of a binary string.
-\item An analogue of invariance should hold for sophistication: at the very least, there should be strict limits on the degree to which a change in the programming language can affect the sophistication.
+\item $S(x)$ should count the bits required for an effective description of the structural properties of a binary string.
+\item An analogue of invariance should hold: there must be strict limits on how much sophistication can be affected by a change in programming language.
 \item There should be no constant $c$ such that $S(x)\le c$ for every input $x$.
-If sophistication is bounded, then knowing its value under one programming language provides no constraints on its value under another language.
+If sophistication is bounded, then knowing its value under one programming language provides no constraints on its value under another language (except that it's also bounded).
 Moreover, a bounded sophistication would be at odds with the intuition that there is no limit to the amount of structure that a binary string can exhibit.
-\item Similarly, there should be no constant $c$ such that $K(x)-S(x)\le c$ for all $x$, because then sophistication would be equivalent to Kolmogorov complexity. 
+\item Similarly, there should be no constant $c$ such that $C(x)-S(x)\le c$ for all $x$, because then sophistication would be equivalent to Kolmogorov complexity. 
 \end{enumerate}
@@
 Thus, the model can be interpreted as a function mapping the noise to the original object.
+
+For any string $x$, there may be many different two-part representations.
+The total codelength can never be less than the Kolmogorov complexity, but it can come close.
+The balance between the two parts may vary; Figure~\ref{fig:diagram} provides an illustration.
+The various proposals then define sophistication in terms of the length of the structure part of this two-part code for one of the representations that (almost) achieves the Kolmogorov complexity.
+However, for most of these definitions, we can prove that they fail one or more of the conditions above.
+Vereshchagin and Vit\'anyi's definition \cite{vitanyi2004meaningful} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we show that their method only assigns substantial sophistication to strings that require an enormous amount of processing to construct.
 
@@
 \end{figure*}
-
-For any string $x$, there may be many different two-part representations.
-The total codelength can never be less than the Kolmogorov complexity, but it can come close.
-The balance between the two parts may vary; Figure~\ref{fig:diagram} provides an illustration.
-The various proposals then define sophistication in terms of the length of the structure part of this two-part code for one of the representations that (almost) achieves the Kolmogorov complexity.
-However, For most of these definitions, we can prove that they fail one or more of the conditions above.
-Vit\'anyi's definition \cite{vitanyi2004meaningful} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we can show that only strings that require an enormous amount of processing to construct, yield a large sophistication.
 

### 2015-01-21 Facticity/Facticity/paper.tex
sec 3 processed

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 This makes these approaches vulnerable to the so-called \emph{nickname problem}: the number of bits required to specify a model becomes very dependent on the numbering.
-This problem is severe: \emph{any} index sophistication without requirements on the numbering is either bounded, or not invariant:
+This problem is severe: \emph{any} index sophistication without requirements on the numbering is either bounded, or nowhere near invariant:
 
@@
 
-To prove this, we use numberings that are deliberately extremely inefficient:
+To prove this, we simply use extremely inefficient numberings:
 
@@
 However, as this lemma shows, it creates a massive change in the size of the model that witnesses the complexity.
-The plain length of the index in a given acceptable numbering may be a very poor indicator of model complexity.
+The plain length of the index in a given acceptable numbering is a very poor indicator of model complexity.
 
@@
 
-\begin{definition}[\cite{grunwald2004shannon,vitanyi2004meaningful}]
-The complexity of a computable function is defined by \label{definition:model-complexity}
+\begin{definition}[\cite{grunwald2004shannon,vitanyi2004meaningful}]\label{definition:model-complexity}
+The complexity of a computable function $f$ is given by 
 \[
@@
 We show in the appendix (Lemma~\ref{lemma:invariance}) that this definition is invariant, so that the Kolmogorov complexity is an objective property of a function, as it is of a binary string.
+
 There are two ways to use this notion of model complexity for more robust attempts to define sophistication.
 Confusingly, both approaches are used in the literature.
-\begin{enumerate}
-  \item We maintain that the representation witnessing the Kolmogorov complexity determines the sophistication, but impose the additional constraint that an efficient acceptable numbering is used.
+First, we can stick to using the length of the index as the measure of sophistication, but impose the additional constraint that an efficient acceptable numbering is used.
 We construct such an efficient index, called a \emph{faithful} index, below.
 This is the approach taken by Adriaans \cite{adriaans2012facticity}.
-  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use $K$ to measure the complexity of the \emph{model} (as in Definition~\ref{definition:model-complexity}), which is then the first part of a two-part code describing the data.
+Second, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use $C^\K$ to measure the complexity of the \emph{model} (as in Definition~\ref{definition:model-complexity}), which is then the first part of a two-part code describing the data.
 With this approach, the total codelength is $C^\K(\phi_i)+|y|$, with $\phi_i(y)=x$.
@@
 This approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information}.
-\end{enumerate}
+
 We first construct an efficient numbering: we will call a numbering $\psi$ \emph{faithful} if for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and $|j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
 This definition is a refinement of \cite[Definition~10]{adriaans2012facticity}.
+The following Lemma is proved in the appendix:
 
@@
 Can we make our prefix encoding so efficient that no other encoding can undercut it?
-By definition, $C^{\M}(m)$ this is the smallest effective, self-delimiting description of $\phi_i$.
+By definition, $C^{\K}(\phi_i)$ is the smallest effective, self-delimiting description of $\phi_i$.
 We cannot compute the prefix code from the data, but this is not necessary to define a representation.
@@
 \end{conjecture}
-The following lemma shows that for index sophistications using a faithful index, but an inefficient prefix encoding are bounded. This includes \cite{adriaans2012facticity}.
+The following lemma shows that even index sophistications using a faithful index are bounded if the prefix encoding is inefficient. This includes \cite{adriaans2012facticity}.
 \begin{restatable}{lemma}{ineffprefix}
 Assume $\phi$ is faithful and define the index sophistication $s^c(x)=\min\{|\bar\imath|:C^{\{i\}}(x)\le C^{\C}(x)+c\}$.
-If $\min\{i\ge i_0:|\bar\imath|-C^\K(i)\}$ is an unbounded function of $i_0$, then $s^c$ is bounded.\label{lemma:prefix-inefficiency}
+If $\min\{|\bar\imath|-C^\K(i):i\ge i_0\}$ is an unbounded function of $i_0$, then $s^c$ is bounded.\label{lemma:prefix-inefficiency}
 \end{restatable}
-If our conjecture is correct, the index sophistication is always bounded.
-The only way to avoid the problem is to use a self-delimiting program of length $C^\K(m)$ to describe the model.
+If our conjecture is correct, the index sophistication is always bounded for model class $\C$.
 Any inefficieny in the model description, and a second represention using a universal model with a more efficient model description will win, causing a bounded sophistication.
 
-For this reason we will use the following definition of sophistication in the remainder of this article: 
+Because of these issues, rather than index sophistication we will use the following definition of sophistication in the remainder of this article: 
 \begin{equation}\label{eq:soph}
@@
 
-In the previous section, we saw the first glimpse of how delicate the internal balance in sophistication is.
-If the description of the model $m$ is less efficient than $C^\K(m)$ by anything more than a constant, representations with a bounded model size will always win out in the limit.
-In light of this balance issues, the most natural definition of sophistication is $S^{\K,\psi,c}$.
-In this setting, we know that there always exist candidate models---those models $i$ with $i \in \K$ and $C^{i}(x) \leq C^\K(x) + c$---for which al information except a constant amount is stored in the data, and candidates for which all information is stored in the model.
+In the previous section, we saw the first glimpse of the delicate balance between the two code components in sophistication.
+If the description of the model $m$ is less efficient than $C^\K(m)$ by anything more than a constant, the sophistication becomes bounded.
+We will investigate this balance, starting with the sophistication $S^{\K,\psi,c}$.
+In this setting, we know that, on the one hand, there always exist candidate models---those models $i$ with $i \in \K$ and $C^{i}(x) \leq C^\K(x) + c$---for which al information except a constant amount is stored in the data, and on the other hand, there are also candidate representations for which all information is stored in the model.
 This tells us that $S^\K$ is \emph{balanced}: there is no inefficiency in the representation forcing information in or out of the model.
-The flipside of having a balanced sophistication is that invariance is easy to show.
+The flip side of having a balanced sophistication is that it becomes easy to show invariance.
 We will show that there are numberings for which the universal models always determine the $S^\K$ and numberings for which the singletons always determine $S^\K$.
-	The following lemma generalizes the principle: 
+We use the following lemma to do so:
 \begin{restatable}{lemma}{coolone}
@@
 Further assume there is a constant $c$ such that:\\
-\-\hspace{1cm}(1)$\forall_{f\in\M'}:\min\{|p|:\psi_{D(p)}=f\}\le K^\psi(f)+c$\\
-\-\hspace{1cm}(2)$\forall_{x\in\X}:L^{\M',\psi}(x)-L^{\M,\psi}(x)\le c$.\\
+\-\hspace{1cm}(1)~$\forall_{f\in\M'}:\min\{|p|:\psi_{D(p)}=f\}\le C^\K^\psi(f)+c$\\
+\-\hspace{1cm}(2)~$\forall_{x\in\X}:L^{\M',\psi}(x)-L^{\M,\psi}(x)\le c$.\\
 Then there is an enumeration $\phi$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.

### 2015-01-21 Facticity/Facticity/paper.tex

@@
 \label{lemma:thecoolone}
-  Let $\psi$ be any acceptable enumeration of the partial recursive functions.
-  Let $\M$ be any model class, let $\X$ be any set of binary sequences and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e.
-if $f=\psi_{D(p)}$ then $p$ is a $D$-description of $f$.
-Let $\M'=\{\psi_i:i\in\tn{range}(D)\}$.
+  Let $\M$ be any model class, let $\X$ be any set of binary strings and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\phi$, i.e.
+if $m=\phi_{D(p)}$ then $p$ is a $D$-description of $m$.
+Let $\M'=\{\phi_i:i\in\tn{range}(D)\}$.
 Further assume there is a constant $c$ such that:\\
-\-\hspace{1cm}(1)~$\forall_{f\in\M'}:\min\{|p|:\psi_{D(p)}=f\}\le C^\K^\psi(f)+c$\\
-\-\hspace{1cm}(2)~$\forall_{x\in\X}:L^{\M',\psi}(x)-L^{\M,\psi}(x)\le c$.\\
-Then there is an enumeration $\phi$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
+\-\hspace{1cm}(1) $\forall_{m\in\M'}:\min\{|p|:\phi_{D(p)}=m\}\le C^{\K,\phi}(m)+c$\\
+\-\hspace{1cm}(2) $\forall_{x\in\X}:C^{\M',\phi}(x)-C^{\M,\phi}(x)\le c$.\\
+Then there is an enumeration $\psi$ of the partial recursive functions such that $S^{\M,\psi}(x) = |\br{0}|+S^{\M',\phi}(x)$ for all $x\in\X$.
 \end{restatable}
+This is a technical lemma, but the idea behind it is simple: given a subset $\M'$ of the model class, we can choose the numbering so that representations use a model in $\M'$ perform an arbitrary constant better than all other models, so that a model in $\M'$ ends up determining the sophistication.
 
-This is a technical lemma, due to the double effect of the numbering, which occurs in both the definition of model complexity and the definition of data complexity.
-The \emph{idea} behind the lemma, however, is simple.
-Given a subset $\M'$ of the model class, we can choose the numbering so that representations use a model in $\M'$ perform an arbitrary constant better than all other models.
-This lets us effectively `push' all other models out of the candidate set, so that the sophistication is determined by a model in $\M'$.- can make it equal to $C^K$
-
-This allows us to ensure that for some numberings, the unversal model always determines the sophistication, making is bounded:
-
+This means that for model classes with a universal model, there is a numbering such that sophistication is bounded:
 \begin{restatable}[underfitting]{lemma}{underfitting}
-Let $\M$ be a model class containing a universal model $u$, with the property that $\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p$. Then, there exist numberings such that $\s^\M$ is bounded.
+Let $\M$ be a model class containing a universal model $\phi_u$, with the property that $\exists c \forall i \in \M : C^{\{u\}}(x) \leq C^{\{i\}}(x) + c \p$.
+Then, there exist numberings such that $\s^\M$ is bounded.
 \end{restatable}
-For other numberings, the singletons determine the sophistication, making it equal to the complexity:
+\noindent For other numberings, the singletons determine the sophistication, making it equal to the complexity:
 \begin{restatable}[Overfitting]{lemma}{overfitting}
-Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$.
-Then there is an enumeration $\phi$ of the prefix partial recursive functions, and a constant $c$, such that
+Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $i\in\M$ with $\phi_i(\epsilon)=x$.
+Then there is a numbering $\psi$, and a constant $c$, such that for all $x\in\X$
 \[
-K(x)-S^{\M,\phi}(x)\le c
-\]
-for all $x\in\X$.
+C^\M(x)-S^{\M,\psi}(x)\leq c
+\]\label{lemma:overfitting}
 \end{restatable}
-This shows that the price we pay for balance is absolute invariance.
-All information can be seen as structure of noise, depending on the numbering.
-To avoid these issues, existing proposals upset the balance to exclude, or penalize at least the singleton models and the universal models.
+\noindent Thus, in this balanced sophistication, there is no invariance: all information can be seen as structure, or as noise, depending on the numbering.
+To avoid these issues, existing proposals upset the balance to exclude or penalize at least the singleton models and the universal models.
 
@@
 
-The existing proposals that do not avoid overfitting as the structure function \cite{cover1985kolmogorov}, the sophistication as defined in \cite{mota2013sophistication}, the miinimal sufficient statistic \cite{gacs2001algorithmic} and Vereshchagin's strongly  sufficient statistic \cite{vereshchagin2013algorithmic}.
-It may be argued that the constant determining the candidates should be allowed to depend on the numbering, but this dependence has not been mentioned in the literature and there is no clear way to determine what such a constant should be, on the basis of a given numbering.
+The existing proposals that do not avoid overfitting are the minimal sufficient statistic in the structure function \cite{cover1985kolmogorov}, the sophistication as defined in \cite{mota2013sophistication}, the minimal sufficient algorithmic statistic \cite{gacs2001algorithmic} and Vereshchagin's strongly sufficient statistic \cite{vereshchagin2013algorithmic}.
+It is widely recognized that singleton models are always candidates in these settings, but not, it seems that they may be the only candidates for any string, as Lemma~\ref{lemma:overfitting} shows.
+
+It may be argued that the slack parameter $c$ in the sophistication, that determines the allowed gap between a candidate representation and the complexity, should be allowed to depend on the numbering, but this dependence has not been mentioned in the literature and there is no clear way to determine how this constant should be determined on the basis of a given numbering.
 
@@
 As we have seen, a strong penalty, such as the one imposed by an inefficient prefix encoding of the model, will immediately cause underfitting.
-A more subtle approach is to use $C$ as a model class, allowing representations that are not self-delimiting.
-This eliminates at least the singletons, as placing all information in the model automatically creates a self delimiting representation.
-Self delimiting representations are longer than plain representations, and this gap is unbounded \cite[Section~4.5.5]{vitanyi2004meaningful}, which tells us that singletons will be candidates for only finitely many strings.
+A more subtle approach is to use $\C$ as a model class rather than $\K$, thus allowing representations that are not self-delimiting.
+The gap between self-delimiting and non self-delimiting descriptions grows without bound
+\cite[Section~4.5.5]{vitanyi2004meaningful}, so that for sufficiently complex inputs all information will have to be placed in the noise part of the code, eliminating the singletons as viable candidates.
 This approach is taken by Vit\'anyi \cite{vitanyi2004meaningful} and by Adriaans \cite{adriaans2012facticity}.
+Such measures thus help reduce the overfitting problem, but they only increase the tendency to underfit.
+We also pay the price that the models can no longer be equated with probability measures, removing the link to traditional statistics and probability theory.
 
@@
 
-Underfitting is a widely acknowledged problem, and almost all proposals avoid it by limiting the class of allowed models, to exclude universal models.
+Underfitting is a widely acknowledged problem for sophistication \sdr{provide refs}, and most proposals avoid it by limiting the class of allowed models, to exclude universal models.
 Does this make the sophistication unbounded?
@@
 \begin{restatable}{theorem}{dogfood}
-There are infinitely many $x$ with $\s^{\F,\phi,c}(x) \geq C^{\K}(x)/5$ and $\s^{\T,\phi,c}(x) \geq C^{\K}(x)/6$. 
+There are infinitely many $x$ with 
+\[
+	\s^{\F,\phi,c}(x) \geq C^{\K}(x)/5  \text{\;\;\;and\;\;\;} \s^{\T,\phi,c}(x) \geq C^{\K}(x)/6\p
+\] 
 \end{restatable}
 
-But clearly, this only exacerbates the overfitting problem.
-Only one model class eliminates both the singletons and the unversal model: $S^\T(x)$.
-The only proposal that uses an efficient model representation, excludes the universal models and excludes the singletons is Vit\'anyi's \cite{vitanyi2004meaningful}.
-While this invalidates straightforward proofs of boundedness, there is no suggestion that $S^\T(x)$ is actually invariant.
+Thus, there are no longer any numberings for which sophistication is bounded, so the underfitting problem is at least partly solved.
+However, as we show below, the resulting sophistication is not an intuitive measure of the amount of structure in the data.
+Moreover, we push the balance back in the direction of overfitting, which is only exacerbated by this change.
+Only one model class eliminates both the singletons and the universal model: $\T$.
+The only proposal that uses an efficient model representation \emph{and} excludes the universal models \emph{and} excludes the singletons is Vereshchagin and Vit\'anyi's \cite{vitanyi2004meaningful}.
+While this invalidates straightforward proofs of boundedness, there is no suggestion that $S^{\T, \phi, c}(x)$ is actually invariant.
 
-While under $S^{\T, \phi, c}(x)$ there are clearly string with high sophistication, these may not conform to the original intuition that motivated sophistication.
-To show this, we require the concept of depth:
+While under $S^{\T, \phi, c}(x)$ there are strings with high sophistication, these may not conform to the original intuition that motivated sophistication. To show this, we need the concept of depth:
 
@@
 Let $A(n)$ be the single-argument Ackermann function and $c$ some arbitrary constant.
-There are numberings such that for all strings with depth $d^\C_c(x) \leq A(C(x))$, $\s^\T(x)$, $s^\P(x)$ and $\s^\F$ are bounded.
+There are numberings such that for all strings with depth $d^{\C,c}(x) \leq A(C^\C(x))$ the sophistication $\s^{\T, \psi, c}$ is bounded.
+The same is true for $\s^{\F}$ .
 \end{restatable}
-\noindent This shows that while high-sophistication strings exist, they do not behave as expected.
-Consider a string that is typical for a low-depth model, say some high-complexity Markov chain.
-Under sophistication, no matter how complex the markov chain, the sophistication would stay the same.
+\noindent This shows that while high-sophistication strings do exist, they do not behave as expected.
+Consider a string that is typical for a shallow model, say some elaborate Markov chain.
+Under $\s^{\T,\psi,c}$, no matter how long an effective description of the Markov chain would have to be, the sophistication is bounded.
 Any structure that is simple enough to be exploited within the time bound of the Ackermann function (or some other fast-growing, simply stated bound) will be counted as noise.
 Only structure that is so deep that if we used it to compress $x$ we could never hope to decompress it within the lifetime of the universe, would count towards sophistication.
-This clearly contradicts the motivating principle of sophistication, which is to capture the number of bits required to express \emph{all} regularity in the data. 
-What makes a string high-sophistication is not the fact that it has a great deal of structure, but the fact that it can be compressed far better with partial functions than with total.
+This clearly contradicts the motivating principle of sophistication, which is to capture the number of bits required to express \emph{all} regularity in the data, not just the regularity that is difficult to describe.
+In this setting, a string high-sophistication not only has a lot of structure, but it also has the property that it can be compressed far better with partial functions than with total.
 That is, it is non-typical for the model class $\T$.

### 2015-01-21 Facticity/Facticity/paper.tex
..

@@
 We have criticized existing measures of the sophistication of a binary string.
-Our results show that there are technical problems with each of them.
+Our results show that there are technical problems with all of them.
 But that does not in itself mean that it should be impossible to come up with a sound measure.
-The intuition throughout the literature, starting with Kolmogorov's proposal of the structure function, appears to be that it should be possible to identify noise, and distinguish it from structural information, by its property of being incompressible.
+The intuition throughout the literature, starting with the structure function, appears to be that the crucial property is whether a string is typical for a model.
+If so, another random choice from that mode selects a string with the same structure: a string of the same `type'.
 
@@
 For the first argument, we take a generative perspective.
-If data are generated by a computable probability distribution, then a sophistication could be called \emph{consistent} if, given a large enough sample, it accurately reflects the complexity of that distribution.
-Suppose that the data are generated by a universal distribution, e.g.
-by feeding random bits to a universal partial computable function.
-(While in general we do not wish to make any assumptions about the origin of the data, it can make sense to do so as a sanity check, as we will do here.) Let $u$ be the index of the function in the $\phi$-enumeration with the property that $\phi_u(\bar\imath y)=\phi_i(y)$.
+We can generate data from a model in $i \in \K$ by feeding it random bits until it produces an argument, we will call the resulting probability distribution for model $p_i(x) = \sum_{y:\phi_i(y) = x} 2^{-|y|}$.
+In this scenario, a sophistication could be called \emph{consistent} if, for data of sufficiently high complexity, it accurately reflects the complexity of that distribution from which it originated.\footnotemark Suppose that the data issampled from  universal distribution: let $\phi_u(\bar\imath y)=\phi_i(y)$ and sample from $p_u$ as described above.
 Then the initial random bits will determine the prefix encoded index $\bar\imath$ of the function $\phi_i$ that $\phi_u$ will subsequently emulate, and the remaining bits are used as inputs to $\phi_i$.
-We now ask, what should be the sophistication of data that is generated in this fashion?
+We now ask, what should be the sophistication of the resulting data?
 
-Certainly, if we have to judge based only on the data, we cannot exclude the possibility that the data was sampled from $\phi_u$: after all, it was.
-On the other hand, neither can we exclude the possibility that it was generated by $\phi_i$, as again, it was!
-For this reason it seems unjustified to assume that, based on only the data, all but one model can always be disqualified.
+\footnotetext{While in general we do not wish to make any assumptions about the origin of the data, in some cases, it can make sense to do so as a sanity check.}
+
+Certainly, if we have to judge based only on the data, we cannot exclude the possibility that the data was sampled from $p_u$: after all, it was.
+On the other hand, neither can we exclude the possibility that it came from $p_i$, as again, it did!
+It seems unjustified to assume that, based on only the data, all but one model can always be disqualified.
 
@@
 The theory behind sophistication suggests that we can choose one of these as the objective, intrinsic model of the data.
-Different models say different things: the universal model says that it is `something computable'.
+Different models say different things: the universal model says that it is `some finite object'.
 A more specific model might say that it is `an image'.
@@
 A sound measure of Sophistication should be able to select one of these models as the proper representation of structure in the data, and disqualify the others as either over- or underfitting.
-But how should we be able to say that the data is intrinsically more of an image than a painting?
+But how should we be able to say that the data is intrinsically more of a painting than an image?
 More of a Monet than a painting?
-Intuitively such distinctions require further assumptions.
+Intuitively, such distinctions require further assumptions, or a second sample from the same distribution.
 
-The second argument why we believe sophistication will elude definition is more technical.
+The second reason we believe sophistication will elude definition is more technical.
 Given an input $x$ and some model class, say $\C$, consider the set of all possible two-part representations of $x$.
@@
 If such jumps can occur, the sophistication is not invariant.
-And while we cannot \emph{prove} in general that such jumps can alwys occur, there seems to be no reason to believe that they don't.
+And while we cannot \emph{prove} in general that such jumps can always occur, there seems to be no reason to believe that they don't.
 
@@
 Note that \emph{part} of the theory is fine: there is nothing wrong with evaluating models for the data by comparing their two-part code lengths.
-In fact, the randomness deficiency $\delta(\phi_i,x)=\min\{|y|:\phi_i(y)=x\}-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data --- if $\phi_i$ truly generated the data, then the probability that the random deficiency is more than $k$ bits is less than $2^{-k}$ (proof omitted).
+In fact, the randomness deficiency $\delta(\phi_i,x)=-\log p_i(x)-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data---if $\phi_i$ truly generated the data, then the probability that the randomness deficiency is more than $k$ bits is less than $2^{-k}$ \cite{bloem2014safe}[Lemma~6].
 In the Monet example above, this will allow us to disqualify the model expressing that the data is actually, say, a recording of Jazz music.
@@
 In our opinion, a lot of measures taken in the literature to address problems with sophistication, such as restricting the model class or introducing model penalties, complicate the method and make problems harder to analyse, without actually addressing the fundamental issue.
-This is dangerous: if such ad-hoc fixes result in a theory that is hard to prove either wrong or right, then this can create an artificial dead end for a valuable area of research.
-When hard cut-offs are avoided on the other hand, all such measures are suddenly no longer necessary.
+This is dangerous: if such ad-hoc fixes result in a theory that is hard to prove either wrong or right, it creates an artificial dead end for a valuable area of research.
+When hard cut-offs are avoided, on the other hand, all such measures are suddenly no longer necessary.
 What remains is a simple and elegant theory that can be used to sift through all possible models, disproving most while retaining a select number of interesting candidates for our further consideration.

### 2015-01-21 Facticity/Facticity/paper.tex
related work

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 
+Adriaans' \emph{Facticity}, introduced in 2012\cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions.
+Adriaans also identifies the nickname problem discussed in Section~\ref{section:indices}.
+
+Other authors have pointed out problems with sophistication.
 Antunes et al.
@@
 They propose a different definition called \emph{coarse sophistication}.
-As we shall see, this version behaves very differently from regular sophistication, and is closer to computational depth.
-Vereshchagin \cite{vereshchagin2013algorithmic} suggests additional restrictions on models that are allowed to determine the sophistication, in hopes of curbing the instability issue.
-Adriaans' \emph{Facticity}, introduced in 2012\cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions.
-Adriaans also identifies the nickname problem discussed in Section~\ref{section:indices}.
-
+This version behaves very differently from regular sophistication, and is closer to computational depth.
+Vereshchagin \cite{vereshchagin2013algorithmic} also points out that the Kolmogorov minimal sufficient statistic is not always well-defined and suggests requiring that the finite set model should described using a \emph{total} function rather than any computable function.
+This yields a more robust sophistication that is nonetheless still susceptible to the issues discussed in this paper.
 \section{Preliminaries}
@@
 Note that \emph{part} of the theory is fine: there is nothing wrong with evaluating models for the data by comparing their two-part code lengths.
-In fact, the randomness deficiency $\delta(\phi_i,x)=\min\{|y|:\phi_i(y)=x\}-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data --- if $\phi_i$ truly generated the data, then the probability that the random deficiency is more than $k$ bits is less than $2^{-k}$ (proof omitted).
+In fact, the randomness deficiency $\delta(\phi_i,x)=\min\{|y|:\phi_i(y)=x\}-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data---if $\phi_i$ truly generated the data, then the probability that the random deficiency is more than $k$ bits is less than $2^{-k}$ (proof omitted).
 In the Monet example above, this will allow us to disqualify the model expressing that the data is actually, say, a recording of Jazz music.

### 2015-01-21 Facticity/Facticity/paper.tex

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 For the first argument, we take a generative perspective.
-We can generate data from a model in $i \in \K$ by feeding it random bits until it produces an argument, we will call the resulting probability distribution for model $p_i(x) = \sum_{y:\phi_i(y) = x} 2^{-|y|}$.
-In this scenario, a sophistication could be called \emph{consistent} if, for data of sufficiently high complexity, it accurately reflects the complexity of that distribution from which it originated.\footnotemark Suppose that the data issampled from  universal distribution: let $\phi_u(\bar\imath y)=\phi_i(y)$ and sample from $p_u$ as described above.
+We can generate data from a model in $\phi_i, i \in \K$ by feeding it random bits until it produces an output.
+We will call the resulting probability distribution $p_i(x) = \sum_{y:\phi_i(y) = x} 2^{-|y|}$.
+In this scenario, a sophistication could be called \emph{consistent} if, for data of sufficiently high complexity, it accurately reflects the complexity of that distribution from which it originated.
+Suppose that the data issampled from  universal distribution: let $\phi_u(\bar\imath y)=\phi_i(y)$ and sample from $p_u$ as described above.
 Then the initial random bits will determine the prefix encoded index $\bar\imath$ of the function $\phi_i$ that $\phi_u$ will subsequently emulate, and the remaining bits are used as inputs to $\phi_i$.
 We now ask, what should be the sophistication of the resulting data?
-
-\footnotetext{While in general we do not wish to make any assumptions about the origin of the data, in some cases, it can make sense to do so as a sanity check.}
 
@@
 Given an input $x$ and some model class, say $\C$, consider the set of all possible two-part representations of $x$.
-Figure~\ref{fig:diagram} provides an illustration.
 Now, when the numbering of partial computable functions is changed, the codelength of the model part of all these representations will change.
-This is illustrated in the right graph in the figure.
+This is illustrated in the right graph in Figure~\ref{fig:diagram}.
 However, the Invariance Theorem expresses that this change is limited by a constant term that does not depend on $x$.
 So for sufficiently large $x$, the horizontal shift of all the representations will be relatively small.
-\emph{However}, even this small shift can push some representations out of the acceptable region (indicated by the dashed line), and pull others in.
+However, even this small shift can push some representations out of the acceptable region (indicated by the dashed line), and pull others in.
 This may lead to a different representation determining the sophistication, one whose \emph{total} codelength is close to what it was before, but whose \emph{model} codelength can be anywhere between $0$ and $C^\C(x)$.
@@
 For this reason, we take a skeptical view of sophistication.
-Note that \emph{part} of the theory is fine: there is nothing wrong with evaluating models for the data by comparing their two-part code lengths.
-In fact, the randomness deficiency $\delta(\phi_i,x)=-\log p_i(x)-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data---if $\phi_i$ truly generated the data, then the probability that the randomness deficiency is more than $k$ bits is less than $2^{-k}$ \cite{bloem2014safe}[Lemma~6].
+Note that part of the theory is fine: there is nothing wrong with evaluating models for the data by comparing their two-part code lengths.
+In fact, the randomness deficiency $\delta(\phi_i,x)=-\log p_i(x)-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data---if $p_i$ truly generated the data, then the probability that the randomness deficiency is more than $k$ bits is less than $2^{-k}$ \cite{bloem2014safe}[Lemma~6].
 In the Monet example above, this will allow us to disqualify the model expressing that the data is actually, say, a recording of Jazz music.

### 2015-01-21 Facticity/Facticity/paper.tex
...

@@
 \item $\F$ is an enumerable set such that $\{\phi_i:i\in\F\}$ contains a uniform code for every finite set.
-A uniform code for a finite set $S$ is a total computable surjective function $f:\{0,1\}^{\lceil|S|\rceil}\to S$.
+A uniform code for a finite set $S$ is a total computable surjective function $f:\{0,1\}^{\lceil\log|S|\rceil}\to S$.
 \end{enumerate}
@@
 
-A pair $(i,y)$ is a \emph{representation} for $x$ if $\phi_i(y)=x$. \sdr{Gebruiken we dit?}
+A pair $(i,y)$ is a \emph{representation} for $x$ if $\phi_i(y)=x$.
 
@@
 Let $s^\psi$ denote any index sophistication with respect to numbering $\psi$.
-There are acceptable numberings $\psi$ and $\xi$ such that for all $x$:
-\[|s^\psi(x) - s^\xi(x)|\geq \frac{1}{2}\min\{s^\psi(x),s^\xi(x)\}\p\]
+There are acceptable numberings $\psi$ and $\xi$ such that for all $x$
+\[|s^\psi(x) - s^\xi(x)|\geq \tfrac{1}{2}\min\{s^\psi(x),s^\xi(x)\}\p\]
 \end{lemma}
@@
 \end{proof}
-The Kolmogorov complexity does not suffer from inefficient indices.
-A switch to to a more efficient index will only cost a constant amount of bits in the complexity.
-However, as this lemma shows, it creates a massive change in the size of the model that witnesses the complexity.
-The plain length of the index in a given acceptable numbering is a very poor indicator of model complexity.
+Thus, even though the Kolmogorov complexity changes by only a constant number of bits under a switch to a different numbering, the plain length of the index in a given acceptable numbering is a very poor indicator of model complexity.
 
-In order to obtain a reasonable and robust measure of model complexity, we extend the definition of Kolmogorov complexity to functions:
-
-\begin{definition}[\cite{grunwald2004shannon,vitanyi2004meaningful}]\label{definition:model-complexity}
-The complexity of a computable function $f$ is given by 
-\[
-C^{\M,\psi}(f) = \min\{C^{\M,\psi}(i):\phi_i=f\}\p
-\]
-\end{definition}
-We show in the appendix (Lemma~\ref{lemma:invariance}) that this definition is invariant, so that the Kolmogorov complexity is an objective property of a function, as it is of a binary string.
+In order to obtain a reasonable and robust measure of model complexity, we define the Kolmogorov complexity of functions as in \cite{grunwald2004shannon,vitanyi2004meaningful}] by $C^{\M,\psi}(f) = \min\{C^{\M,\psi}(i):\phi_i=f\}$.
+Lemma~\ref{lemma:invariance} in the appendix shows that this definition is invariant, so that the Kolmogorov complexity is an objective property of a function.
 
@@
 This is the approach taken by Adriaans \cite{adriaans2012facticity}.
-Second, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use $C^\K$ to measure the complexity of the \emph{model} (as in Definition~\ref{definition:model-complexity}), which is then the first part of a two-part code describing the data.
+Second, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use $C^\K$ to measure the complexity of the \emph{model} (as defined above), which is then the first part of a two-part code describing the data.
 With this approach, the total codelength is $C^\K(\phi_i)+|y|$, with $\phi_i(y)=x$.
@@
 Because of these issues, rather than index sophistication we will use the following definition of sophistication in the remainder of this article: 
-\begin{equation}\label{eq:soph}
+\[
 S^{\M,\psi,c}(x)=\min\left\{C^\K(\phi_i):i\in\M,C^{\{i\},\psi}(x)\le C^{\M,\psi}(x)+c\right\}.
-\end{equation}
+\]
 
@@
 We will show that there are numberings for which the universal models always determine the $S^\K$ and numberings for which the singletons always determine $S^\K$.
-We use the following lemma to do so:
+The following lemma is based on the idea that, given a subset $\M'$ of the model class, we can choose the numbering so that representations use a model in $\M'$ perform an arbitrary constant better than all other models, so that a model in $\M'$ always ends up determining the sophistication.
 \begin{restatable}{lemma}{coolone}
@@
 \end{restatable}
-This is a technical lemma, but the idea behind it is simple: given a subset $\M'$ of the model class, we can choose the numbering so that representations use a model in $\M'$ perform an arbitrary constant better than all other models, so that a model in $\M'$ ends up determining the sophistication.
-
 This means that for model classes with a universal model, there is a numbering such that sophistication is bounded:
@@
 Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $i\in\M$ with $\phi_i(\epsilon)=x$.
-Then there is a numbering $\psi$, and a constant $c$, such that for all $x\in\X$
-\[
-C^\M(x)-S^{\M,\psi}(x)\leq c
-\]\label{lemma:overfitting}
+Then there is a numbering $\psi$, and a constant $c$, such that for all $x\in\X$ we have $C^\M(x)-S^{\M,\psi}(x)\leq c$.\label{lemma:overfitting}
 \end{restatable}
@@
 
-The existing proposals that do not avoid overfitting are the minimal sufficient statistic in the structure function \cite{cover1985kolmogorov}, the sophistication as defined in \cite{mota2013sophistication}, the minimal sufficient algorithmic statistic \cite{gacs2001algorithmic} and Vereshchagin's strongly sufficient statistic \cite{vereshchagin2013algorithmic}.
-It is widely recognized that singleton models are always candidates in these settings, but not, it seems that they may be the only candidates for any string, as Lemma~\ref{lemma:overfitting} shows.
+The existing proposals that can overfit are the minimal sufficient statistic in the structure function \cite{cover1985kolmogorov}, the sophistication as defined in \cite{mota2013sophistication}, the minimal sufficient algorithmic statistic \cite{gacs2001algorithmic} and Vereshchagin's strongly sufficient statistic \cite{vereshchagin2013algorithmic}.
+It is widely recognized that singleton models are always candidates in these settings, but not, it seems, that they may be the only candidates for any string, as Lemma~\ref{lemma:overfitting} shows.
 
@@
 
-Underfitting is a widely acknowledged problem for sophistication \sdr{provide refs}, and most proposals avoid it by limiting the class of allowed models, to exclude universal models.
-Does this make the sophistication unbounded?
-The following theorem shows that it does.
+As mentioned in the related work, underfitting is a widely acknowledged problem for sophistication, and most proposals avoid it by limiting the class of allowed models, to exclude universal models.
+The following theorem shows that this does indeed yield an unbounded sophistication.
 
 \begin{restatable}{theorem}{dogfood}
-There are infinitely many $x$ with 
-\[
-	\s^{\F,\phi,c}(x) \geq C^{\K}(x)/5  \text{\;\;\;and\;\;\;} \s^{\T,\phi,c}(x) \geq C^{\K}(x)/6\p
-\] 
+There are infinitely many $x$ with $\s^{\F,\phi,c}(x) \geq C^{\K}(x)/5$, and also infinitely many $x$ with $\s^{\T,\phi,c}(x) \geq C^{\K}(x)/6$.
 \end{restatable}
@@
 Let $C^\M_t(x) = \min\{|\bar\imath y| : U^t(\bar\imath y) = x, \phi_i \in \M\}$.
-
 The \emph{$c$-depth} is $d^{\M,c}(x) = \min \left\{t : C^\M_t(x) - C^\M(x) \leq c \right\}$.
@@
 Let $A(n)$ be the single-argument Ackermann function and $c$ some arbitrary constant.
-There are numberings such that for all strings with depth $d^{\C,c}(x) \leq A(C^\C(x))$ the sophistication $\s^{\T, \psi, c}$ is bounded.
+There are numberings such that for strings with depth $d^{\C,c}(x) \leq A(C^\C(x))$ the sophistication $\s^{\T, \psi, c}$ is bounded.
 The same is true for $\s^{\F}$ .
@@
 
-\section{Coarse sophistication}
-
-\pb{TODO}
-
 \end{document}

### 2015-01-21 Facticity/Facticity/paper.tex
...

@@
 
-In order to obtain a reasonable and robust measure of model complexity, we define the Kolmogorov complexity of functions as in \cite{grunwald2004shannon,vitanyi2004meaningful}] by $C^{\M,\psi}(f) = \min\{C^{\M,\psi}(i):\phi_i=f\}$.
-Lemma~\ref{lemma:invariance} in the appendix shows that this definition is invariant, so that the Kolmogorov complexity is an objective property of a function.
+In order to obtain a reasonable and robust measure of model complexity, we define the Kolmogorov complexity of functions as in \cite{grunwald2004shannon,vitanyi2004meaningful} by $C^{\M,\psi}(f) = \min\{C^{\M,\psi}(i):\phi_i=f\}$.
+Lemma~\ref{lemma:invariance} in the appendix shows that this definition is invariant: the Kolmogorov complexity is an objective property of a function.
 
-There are two ways to use this notion of model complexity for more robust attempts to define sophistication.
-Confusingly, both approaches are used in the literature.
-First, we can stick to using the length of the index as the measure of sophistication, but impose the additional constraint that an efficient acceptable numbering is used.
-We construct such an efficient index, called a \emph{faithful} index, below.
+There are two ways to use this notion of model complexity for more robust attempts to define sophistication. Confusingly, both approaches are used in the literature.
+First, we can measure the complexity of the \emph{model} using $C^{\M,\psi}(\phi_i)$, which is then the first part of a two-part code describing the data.
+This approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information} and in most of the present paper.
+
+Second, we can stick to using the length of the index as the measure of sophistication, but impose the additional constraint that an efficient acceptable numbering is used.
 This is the approach taken by Adriaans \cite{adriaans2012facticity}.
-Second, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use $C^\K$ to measure the complexity of the \emph{model} (as defined above), which is then the first part of a two-part code describing the data.
-With this approach, the total codelength is $C^\K(\phi_i)+|y|$, with $\phi_i(y)=x$.
-In this approach, we can construct everything on the basis of a single acceptable numbering, without additional constraints.
-This approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information}.
+Before switching to the other approach we construct such an efficient index.
 
-We first construct an efficient numbering: we will call a numbering $\psi$ \emph{faithful} if for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and $|j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
+We call a numbering $\psi$ \emph{faithful} if for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and $|j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
 This definition is a refinement of \cite[Definition~10]{adriaans2012facticity}.

### 2015-01-21 Facticity/Facticity/paper.tex
...

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 
+Antunes et al.
+\cite{antunes2009sophistication} note that the constant cutoff used by Koppel is a source of instability, a result in the spirit of this paper.
+They propose a different definition called \emph{coarse sophistication}.
+As we shall see, this version behaves very differently from regular sophistication, and is closer to computational depth.
+Vereshchagin \cite{vereshchagin2013algorithmic} suggests additional restrictions on models that are allowed to determine the sophistication, in hopes of curbing the instability issue.
 Adriaans' \emph{Facticity}, introduced in 2012\cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions.
@@
 
-Other authors have pointed out problems with sophistication.
-Antunes et al.
-\cite{antunes2009sophistication} note that the constant cutoff used by Koppel is a source of instability, a result in the spirit of this paper.
-They propose a different definition called \emph{coarse sophistication}.
-This version behaves very differently from regular sophistication, and is closer to computational depth.
-Vereshchagin \cite{vereshchagin2013algorithmic} also points out that the Kolmogorov minimal sufficient statistic is not always well-defined and suggests requiring that the finite set model should described using a \emph{total} function rather than any computable function.
-This yields a more robust sophistication that is nonetheless still susceptible to the issues discussed in this paper.
 \section{Preliminaries}
@@
 For the first argument, we take a generative perspective.
-We can generate data from a model in $\phi_i, i \in \K$ by feeding it random bits until it produces an output.
-We will call the resulting probability distribution $p_i(x) = \sum_{y:\phi_i(y) = x} 2^{-|y|}$.
-In this scenario, a sophistication could be called \emph{consistent} if, for data of sufficiently high complexity, it accurately reflects the complexity of that distribution from which it originated.
-Suppose that the data issampled from  universal distribution: let $\phi_u(\bar\imath y)=\phi_i(y)$ and sample from $p_u$ as described above.
+We can generate data from a model in $i \in \K$ by feeding it random bits until it produces an argument, we will call the resulting probability distribution for model $p_i(x) = \sum_{y:\phi_i(y) = x} 2^{-|y|}$.
+In this scenario, a sophistication could be called \emph{consistent} if, for data of sufficiently high complexity, it accurately reflects the complexity of that distribution from which it originated.\footnotemark Suppose that the data issampled from  universal distribution: let $\phi_u(\bar\imath y)=\phi_i(y)$ and sample from $p_u$ as described above.
 Then the initial random bits will determine the prefix encoded index $\bar\imath$ of the function $\phi_i$ that $\phi_u$ will subsequently emulate, and the remaining bits are used as inputs to $\phi_i$.
 We now ask, what should be the sophistication of the resulting data?
+
+\footnotetext{While in general we do not wish to make any assumptions about the origin of the data, in some cases, it can make sense to do so as a sanity check.}
 
@@
 Given an input $x$ and some model class, say $\C$, consider the set of all possible two-part representations of $x$.
+Figure~\ref{fig:diagram} provides an illustration.
 Now, when the numbering of partial computable functions is changed, the codelength of the model part of all these representations will change.
-This is illustrated in the right graph in Figure~\ref{fig:diagram}.
+This is illustrated in the right graph in the figure.
 However, the Invariance Theorem expresses that this change is limited by a constant term that does not depend on $x$.
 So for sufficiently large $x$, the horizontal shift of all the representations will be relatively small.
-However, even this small shift can push some representations out of the acceptable region (indicated by the dashed line), and pull others in.
+\emph{However}, even this small shift can push some representations out of the acceptable region (indicated by the dashed line), and pull others in.
 This may lead to a different representation determining the sophistication, one whose \emph{total} codelength is close to what it was before, but whose \emph{model} codelength can be anywhere between $0$ and $C^\C(x)$.
@@
 For this reason, we take a skeptical view of sophistication.
-Note that part of the theory is fine: there is nothing wrong with evaluating models for the data by comparing their two-part code lengths.
-In fact, the randomness deficiency $\delta(\phi_i,x)=-\log p_i(x)-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data---if $p_i$ truly generated the data, then the probability that the randomness deficiency is more than $k$ bits is less than $2^{-k}$ \cite{bloem2014safe}[Lemma~6].
+Note that \emph{part} of the theory is fine: there is nothing wrong with evaluating models for the data by comparing their two-part code lengths.
+In fact, the randomness deficiency $\delta(\phi_i,x)=-\log p_i(x)-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data---if $\phi_i$ truly generated the data, then the probability that the randomness deficiency is more than $k$ bits is less than $2^{-k}$ \cite{bloem2014safe}[Lemma~6].
 In the Monet example above, this will allow us to disqualify the model expressing that the data is actually, say, a recording of Jazz music.
@@
 
-\section{Proofs and further discussion}
+\section{Proofs}
 
@@
 \begin{lemma}[Invariance of function complexity]
-Let $\phi$ and $\psi$ be any two acceptable numberings. There exists a constant $c$ such that $\left| K_\phi(f) - K_\psi(f)\right | \leq c$ for all $f$. \label{lemma:invariance}
+Let $\psi$ and $\eta$ be any two acceptable numberings.
+There exists a constant $c$ such that $\left| C^{\K,\psi}(f) - C^{\K, \eta}(f)\right | \leq c$ for all $f \in \C$ and $\left| C^{\C,\psi}(f) - C^{\C, \eta}(f)\right | \leq c$ for all $f \in \C$.
+\label{lemma:invariance}
 \end{lemma}
 \begin{proof}
-Let $g(i)$ be the function such that $\psi_i=\phi_{g(i)}$.
+Let $g(i)$ be the function such that $\psi_i=\eta_{g(i)}$.
 \begin{align*}
-K_\phi(f) &= \min\left\{ K_\phi(i) : \phi_i= f\right\} 
-\geq \min\left\{ K_\psi(i) : \phi_i= f\right\} - c\\
-&= \min\left\{ K_\psi(i) : \psi_{g(i)}= f\right\} - c
-\geq \min\left\{ K_\psi(i) : \psi_i= f\right\} - c' = K_\psi(f).
+C^{\C,\psi}(f) &= \min\left\{ C^{\C,\psi}(i) : \psi_i= f\right\} 
+\geq \min\left\{ C^{\C, \eta}(i) : \psi_i= f\right\} - c\\
+&= \min\left\{ C^{\C, \eta}(i) : \eta_{g(i)}= f\right\} - c
+\geq \min\left\{ C^{\C, \eta}(i) : \eta_i= f\right\} - c' \\
+&= C^{\C, \eta}(f).
 \end{align*}
-The opposite inequality can be achieved by reversing $\phi$ and $\psi$. 
+The opposite inequality can be achieved by reversing $\psi$ and $\eta$. The same proof holds with model class $\K$.
 \end{proof}
@@
 \begin{proof}
-Let $\tn{div} \in \N$ be an index such that $\phi_{\tn{div}}(y)=\infty$ for all $y$. Define
+Let $d \in \N$ be an index such that $\phi_d(y)=\infty$ for all $y$. Define
   \[\psi_q=\begin{cases}
     \phi_{\phi_i(p)}&\tn{if $q$ can be written as $\bar\imath p$ and $\phi_i(p)<\infty$,}\\
-    \phi_{i_\tn{div}}&\tn{otherwise.}\end{cases}
+    \phi_d&\tn{otherwise.}\end{cases}
   \]
@@
 \[\begin{split}
-C(f)&=\min\{C(i):\phi_i=f\} =\min\{\min\{|\bar a b|:\phi_a(b)=i\}:\phi_i=f\} \\
+C^{\C, \phi}(f)&=\min\{C^{\C, \phi}(i):\phi_i=f\} =\min\{\min\{|\bar a b|:\phi_a(b)=i\}:\phi_i=f\} \\
 & =\min\{|\bar a b|:\phi_{\phi_a(b)}=f\}
@@
 denote this index as a function of the program; further define
-$h(q)=\tn{div}$ if $q$ cannot be expressed as $\bar\imath p$. By
+$h(q)=d$ if $q$ cannot be expressed as $\bar\imath p$. By
 construction $h$ is total and computable. To check that the mapping
@@
 We compare the lengths of these two representations.
-Note that $|i^*|=K^\psi(i)$.
+Note that $|i^*|=C^{\K,\phi}(i)$.
 Therefore,
 \[
-|\bar\imath p|-|\bar v i^* y| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-K(i).
+|\bar\imath y|-|\bar v i^* y| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-C^{\K}(i).
 \]
@@
 From this $i_0$ onwards, the second representation (using $v$) will have a shorter codelength, so $\bar\imath p$ cannot achieve the minimum in the definition of the sophistication.
-Consequently, the sophistication is bounded by $\s(x)<|\overline{\imath_0}|$ for all $x$.
+Consequently, the sophistication is bounded by $s(x)<|\overline{\imath_0}|$ for all $x$.
 \end{proof}
@@
 \begin{proof}
-We define the numbering $\phi$ as follows:
+We define the numbering $\psi$ as follows:
 \[\begin{cases}
-\phi_0(p) = 1^r 0 D(p) \\
-\phi_{1^r0i}(p) = \psi_i(p) \\
-\phi_j(\cdot) = \infty &\text{if $j$ contains no zeroes.}
+\psi_0(p) = 1^r 0 D(p) \\
+\psi_{1^r0i}(p) = \phi_i(p) \\
+\psi_j(\cdot) = \infty &\text{if $j$ contains no zeroes.}
 \end{cases}\]
-We will show that under the $\phi$-enumeration, the best representation for $x$ using a model $f\in\M'$ is always better than the best representation using some $f\not\in\M'$.
+We will show that under the $\psi$-enumeration, the best representation for $x$ using a model $m\in\M'$ is always better than the best representation using some $m\notin\M'$.
 
-First suppose $f\in\M'$. Then
+First suppose $m\in\M'$. Then
 \begin{align*}
-K^\phi(f) &=\min\{|\bar\jmath q|:\phi_{\phi_j(q)}=f\}
-\leq\min\{|\bar0 q|:\phi_{\phi_0(q)}=f\} \\ 
-&=\min\{|\bar0 q|:\phi_{1^r0 D(q)}=f\}  
-=|\bar 0|+\min\{|q|:\psi_{D(q)}=f\}
-\leq K^\psi(f)+c+|\bar 0|,
+C^{\K,\psi}(m) &=\min\{|\bar\jmath q|:\psi_{\psi_j(q)}=m\}
+\leq\min\{|\bar0 q|:\psi_{\psi_0(q)}=m\} \\ 
+&=\min\{|\bar0 q|:\psi_{1^r0 D(q)}=m\}  
+=\min\{|q|:\phi_{D(q)}=m\} + |\bar 0|\\
+&\leq C^{\K,\phi}(m)+c+|\br{0}|,
 \end{align*}
-where the last inequality uses the first assumption.
+where the last inequality uses assumption (1).
 
-Now assume that the best model $f$ for $x$ is not in $\M'$. 
-Let $i$ be the index of $f$ with the shortest description, i.e. it achieves the minimum in $K^\phi(f)=\min\{K^\phi(i):\phi_i=f\}$.
-There are two possibilities. Either $i=0$, in which case we have $K^\phi(f)\ge r$ because $\phi_0$ cannot output zero and all other $\phi$-programs are at least $r$ bits long.
+Now assume that the best model $m$ for $x$ is not in $\M'$. 
+Let $i$ be the index of $m$ with the shortest description, i.e. it achieves the minimum in $C^{\K, \psi}(m)=\min\left\{C^{\K,\psi}(i):\psi_i=m\right\}$.
+There are two possibilities. Either $i = 0$, in which case we have $C^{\K,\psi}(m)\ge r$ because $\phi_0$ cannot output zero and all other $\psi$-programs are at least $r$ bits long.
 
@@
 \begin{align*}
-K^\phi(f)&=K^\phi(i)=K^\phi(1^r0j)\\
+C^{\K,\psi}(m)&=C^{\K\psi}(i)=K^\phi(1^r0j)\\
 &\ge K^\phi(j)-c' \ge K^\psi(j)+r+1-c'\\

### 2015-01-21 Facticity/Facticity/paper.tex
...

@@
 
-\begin{lemma}
-There exist faithful acceptable numberings.
-\end{lemma}
+\begin{restatable}{lemma}{faithful}
+  There exist faithful, acceptable numberings.\label{lemma:faithful-numberings}
+\end{restatable}
+
 Unfortunately, even with a faithful numbering, we can fail.
@@
 
-\begin{lemma}
-  There is a faithful acceptable numbering.\label{lemma:faithful-numberings}
-\end{lemma}
+\faithful*
+
 \begin{proof}

### 2015-01-21 Facticity/Facticity/paper.tex
...

@@
 This definition is a refinement of \cite[Definition~10]{adriaans2012facticity}.
-The following Lemma is proved in the appendix:
-
-\begin{restatable}{lemma}{faithful}
-  There exist faithful, acceptable numberings.\label{lemma:faithful-numberings}
-\end{restatable}
-
-Unfortunately, even with a faithful numbering, we can fail.
+We prove in the appendix that there actually do exist faithful, acceptable numberings (Lemma~\ref{lemma:faithful-numberings}).
+However even choosing a faithful index is not quite enough.
 The Kolmogorov complexity uses representations of the form $\bar\imath y$, where the bar denotes some straightforward prefix encoding to delimit the model description $i$ from its input $y$.
 If we define a second prefix encoding $\tilde{\imath}$, with $|\bar\imath|-|\tilde\imath|$ unbounded, we can define a second representation $\bar u \tilde \imath y$, at a constant overhead $|\br{u}|$, and gain more than $|\br{u}|$ for sufficiently complex strings.
-Ultimately, $u$ will always be chosen as a model for sufficiently complex $x$, and the sophistication becomes bounded.
+As shown in Lemma~\ref{lemma:prefix-inefficiency} in the appendix, this results in a bounded sophistication.
+This is in fact an issue with \cite{adriaans2012facticity}. Moreover, we conjecture that \emph{no} computable prefix function $\bar\cdot$ is sufficiently efficient to avoid this.
 
-Can we make our prefix encoding so efficient that no other encoding can undercut it?
-By definition, $C^{\K}(\phi_i)$ is the smallest effective, self-delimiting description of $\phi_i$.
-We cannot compute the prefix code from the data, but this is not necessary to define a representation.
-Let $i^*$ be the shortest self-delimiting program for $\phi_i$, then $i^*y$ is a decodable representation for $x$.
-If our prefix function is less efficient than $C^\K(m)$ by an unbounded amount, $i^*y$ will always beat $\bar\imath y$ for complex enough data.
-We conjecture that no computable prefix encoding function can avoid this issue:
-\begin{conjecture}
-For any partial computable prefix codelength function $L$, the function $\min\{L(i)-C^\K(i):i\ge i_0\}$
-is unbounded in $i_0$.
-\end{conjecture}
-The following lemma shows that even index sophistications using a faithful index are bounded if the prefix encoding is inefficient. This includes \cite{adriaans2012facticity}.
-\begin{restatable}{lemma}{ineffprefix}
-Assume $\phi$ is faithful and define the index sophistication $s^c(x)=\min\{|\bar\imath|:C^{\{i\}}(x)\le C^{\C}(x)+c\}$.
-If $\min\{|\bar\imath|-C^\K(i):i\ge i_0\}$ is an unbounded function of $i_0$, then $s^c$ is bounded.\label{lemma:prefix-inefficiency}
-\end{restatable}
-If our conjecture is correct, the index sophistication is always bounded for model class $\C$.
-Any inefficieny in the model description, and a second represention using a universal model with a more efficient model description will win, causing a bounded sophistication.
-
-Because of these issues, rather than index sophistication we will use the following definition of sophistication in the remainder of this article: 
+Because of these issues, rather than index sophistication we will use the following definition of sophistication in the remainder of this article.
+It is parameterised by a model class, a numbering, and a slack parameter $c$ that determines the maximal allowed slack between the two-part code length using a representation and the Kolmogorov complexity.
 \[
@@
 
-\faithful*
+\begin{restatable}{lemma}{faithful}\label{lemma:faithful-numberings}
+  There are faithful acceptable numberings.
+\end{restatable}
 
@@
 
-\ineffprefix*
+Can we make our prefix encoding $\bar\cdot$ so efficient that no other encoding can undercut it?
+By definition, $C^{\K}(\phi_i)$ is the smallest effective, self-delimiting description of $\phi_i$.
+We cannot compute the prefix code from the data, but this is not necessary to define a representation.
+Let $i^*$ be the shortest self-delimiting program for $\phi_i$, then $i^*y$ is a decodable representation for $x$.
+If our prefix function is less efficient than $C^\K(m)$ by an unbounded amount, $i^*y$ will always beat $\bar\imath y$ for complex enough data.
+We conjecture that no computable prefix encoding function can avoid this issue:
+\begin{conjecture}
+For any partial computable prefix codelength function $L$, the function $\min\{L(i)-C^\K(i):i\ge i_0\}$
+is unbounded in $i_0$.
+\end{conjecture}
+The following lemma shows that even index sophistications using a faithful index are bounded if the prefix encoding is inefficient.
+
+\begin{restatable}{lemma}{ineffprefix}
+Assume $\phi$ is faithful and define the index sophistication $s^c(x)=\min\{|\bar\imath|:C^{\{i\}}(x)\le C^{\C}(x)+c\}$.
+If $\min\{|\bar\imath|-C^\K(i):i\ge i_0\}$ is an unbounded function of $i_0$, then $s^c$ is bounded.\label{lemma:prefix-inefficiency}
+\end{restatable}
+If our conjecture is correct, the index sophistication is always bounded for model class $\C$.
+Any inefficieny in the model description, and a second represention using a universal model with a more efficient model description will win, causing a bounded sophistication.
+
 \begin{proof}

### 2015-01-21 Facticity/Facticity/paper.tex
...

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 
-\subsection{Boosting a model class by manipulating the numbering}
+\subsection{Boosting submodel by manipulating the numbering}
 
@@
 \begin{align*}
-C^{\K,\psi}(m)&=C^{\K\psi}(i)=K^\phi(1^r0j)\\
-&\ge K^\phi(j)-c' \ge K^\psi(j)+r+1-c'\\
-&=K^\psi(f)+r+1-c'.
+C^{\K,\psi}(m)&=C^{\K,\psi}(i)=C^{\K,\psi}(1^r0j)\\
+&\ge C^{\K,\psi}(j)-c' \ge C^{\K,\phi}(j)+r+1-c'\\
+&=C^{\K,\phi}(m)+r+1-c'.
 \end{align*}
-Now choose $r=c''+\max\{K^\psi(\phi_0),c'-1\}$. Then substitution yields, for both cases, 
-$K^\phi(f)\ge c''+K^\psi(f)$.
+Now choose $r=c''+\max\left\{C^{\K,\psi}(\phi_0),c'-1\right\}$. Then substitution yields, for both cases, 
+$C^{\K,\phi}(m)\ge c''+C^{\K,\psi}(m)$.
 
@@
 \[\begin{split}
-K^\phi(g)+C^g(x) &\ge K^\psi(g)+C^g(x)+c''\\
-&\ge K^\phi(f)+C^f(x) -2c+c'' -|\bar0|.
+C^{\K,\psi}(g)+\min \{|y| : g(y) = x\} &\ge C^{\K,^\phi}(g)+\min\{|y| : g(y) = x\}+c''\\
+&\ge C^{\K,\psi}(f)+\min \{|y| : f(y) = x\} -2c+c'' -|\bar0|.
 \end{split}\]
-
 By choosing $c''$ sufficiently large, we can ensure that the best representation is in $\M'$ for all $x\in\X$, which completes the proof.
@@
 \begin{proof}
-Let $\psi$ \sdr{This should probably be $\phi$} be any default enumeration of the partial recursive prefix functions.
-Note that since $f$ is a prefix function, if $f$ is defined for input $\epsilon$ then it cannot be defined for any other input.
-Pick $f,x$ with $f(\epsilon)=x$.
-Note that $x$ can be computed from $f$ and a fixed program, so there is a $c$ such that $K(x)\le K(f)+c$.
+Note that since $\phi_i$ is a prefix function, if $\phi_i$ is defined for input $\epsilon$ then it cannot be defined for any other input.
+Pick $i,x$ with $\phi_i(\epsilon)=x$.
+Note that $x$ can be computed from $i$ and a fixed program, so there is a $c$ such that $C^{\K, \phi}(x)\le C^{\K, \phi}(\phi_i)+c$.
 Vice versa, given any $x$ we can construct an index of $f$, since $\psi$ is an acceptable numbering.

### 2015-01-21 Facticity/Facticity/paper.tex
...

@@
 \begin{proof}
+Let $\phi_i$ be a singleton for $x$ as described.
 Note that since $\phi_i$ is a prefix function, if $\phi_i$ is defined for input $\epsilon$ then it cannot be defined for any other input.
@@
 Note that $x$ can be computed from $i$ and a fixed program, so there is a $c$ such that $C^{\K, \phi}(x)\le C^{\K, \phi}(\phi_i)+c$.
-Vice versa, given any $x$ we can construct an index of $f$, since $\psi$ is an acceptable numbering.
-Therefore $|K(f)-K(x)|\le c$.
+Vice versa, given any $x$ we can construct an index of $\phi_i$, since $\phi$ is an acceptable numbering.
+Therefore $|C^{\K,\phi}(\phi_i)-C^{\K, \phi}(x)|\le c$.
 
-We now define a computable function $D$ by $D(\bar\imath p)=j$ where $\psi_j(\epsilon) = \psi_i(p)$.
+We now define a computable function $D$ by $D(\bar\imath y)=j$ where $\psi_j(\epsilon) = \psi_i(y)$.
 We will show that the two conditions of Lemma~\ref{lemma:thecoolone} hold for the prefix function $D$.
@@
 
-We first show that representions using $\T$ can be reduced to representations in $\F$ with a small overhead.
+We first show that representations using $\T$ can be reduced to representations in $\F$ with a small overhead.
 

### 2015-01-21 Facticity/Facticity/paper.tex
...

@@
 (1) Let $f$ be any function in the range of $D$, and $x$ its output.
-Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\psi_i(q)=x\}=K(x)\le K(f)+c$.
-(2) On the one hand $L^{\M',\psi}(x)\le K(f)+|\epsilon|\le K(x)+c$.
-On the other hand, $L^{\M,\psi}(x)$ is an effective description of $x$, so $K(x)$ is at most a constant larger.
+Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\phi_i(q)=x\}=C^{\K}(x) \le C^{\K}(f)+c$.
+(2) On the one hand $C^{\M',\psi}(x)\le C^\K(f)+|\epsilon|\le C^\K(x)+c$.
+On the other hand, $C^{\M,\psi}(x)$ is an effective description of $x$, so $C^{\K}(x)$ is at most a constant larger.
 Together, these inequalities establish the second condition.
 
-Now, by Lemma~\ref{lemma:thecoolone} there is an enumeration $\phi_1,\phi_2,\ldots$ such that $S^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
-We observed that $|K(f)-K(x)|\le c$ for any $f\in\M'$, so $S^{\M',\psi}(x)\ge K(x)+c$.
+Now, by Lemma~\ref{lemma:thecoolone} there is an numbering $\psi_1,\psi_2,\ldots$ such that $\s^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
+We observed that $|C^{\K}(f)-C^{\K}(x)|\le c$ for any $f\in\M'$, so $\s^{\M',\psi}(x)\ge C^{\K}(x)-c$.
 This proves the lemma.
@@
 That is, $\M' = \{u\}$.
-This construction satisfies the conditions 1 and 2 from Lemma~\ref{lemma:thecoolone}; invoking it we find that there exists an acceptable numbering for which $\s_\M(x) = \s_{\M'}(x) + c$.
-Since $\M'$ contains only a single model, $\s_{\M'}$ is constant.
+This construction satisfies the conditions 1 and 2 from Lemma~\ref{lemma:thecoolone}; invoking it we find that there exists an acceptable numbering $\psi$ for which $\s^{\M,\psi}(x) = \s^{\M', phi}(x) + c$.
+Since $\M'$ contains only a single model, $\s_{\M',\phi}(x)$ is constant.
 \end{proof}
@@
 
-Note that we favoured brevity of the proof over the strength of the bounds: the multiplicative constant $1/6$ can be improved with a more detailed treatment.
+Note that we favored brevity of the proof over the strength of the bounds: the multiplicative constant $1/6$ can be improved with a more detailed treatment.
 
@@
 Let $D$ be a prefix function with $D(\epsilon) = u$.
-We can instantiate Lemma~\ref{lemma:thecoolone} with $D$, $\M' = \{\phi_u\}$ and $X = \{x : d^\C_c(x) \leq A(C(x))\}$.
+We can instantiate Lemma~\ref{lemma:thecoolone} with $D$, $\M' = \{\phi_u\}$ and $X = \{x : d^{\C,c}(x) \leq A(C(x))\}$.
 This tells us that there exists a numbering for which $\s^\T(x) = \s^{\M'}(x) + |\bar0| \leq c$ for all $x \in X$.

### 2015-01-21 Facticity/Facticity/paper.tex
...

@@
 \begin{figure*}[tb]
-  \hspace{-2.8em}
-  \begin{minipage}{0.55\textwidth}
+  \centering
+  \begin{minipage}{0.45\textwidth}
      \includegraphics[width=\textwidth]{./img/sophistication.pdf}
   \end{minipage}
-  \begin{minipage}{0.55\textwidth}
+  \begin{minipage}{0.45\textwidth}
      \includegraphics[width=\textwidth]{./img/sophistication-jump.pdf}
@@
 
-\subsection{Related work}
+\paragraph{Related work}
 The idea of separating structure from noise originated with Kolmogorov's \emph{structure function} \cite{cover1985kolmogorov}.
@@
 Working from the perspective of statistical mechanics, they arrived (independently, it seems) at a notion that fits the mold of sophistication.
-
-Adriaans' \emph{Facticity}, introduced in 2012\cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions.
+\emph{Facticity} \cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions.
 Adriaans also identifies the nickname problem discussed in Section~\ref{section:indices}.
@@
 
-A \emph{model class} is a set of model indices. We distinguish four important model classes: 
-\begin{enumerate}
-\item $\C=\N$, the indices of all partial computable functions
-\item $\T=\{i:\tn{$\phi_i$ is total}\}$. Note that $\T$ is not computably enumerable.
-\item $\K$ is an enumerable set such that $\{\phi_i:i\in\K\}$ is the set of all partial computable prefix functions
-\item $\F$ is an enumerable set such that $\{\phi_i:i\in\F\}$ contains a uniform code for every finite set.
+A \emph{model class} is a set of model indices.
+We distinguish four important model classes: (1) The indices of the partial computable functions $\C=\N$.
+(2) The total functions $\T=\{i:\tn{$\phi_i$ is total}\}$.
+Note that $\T$ is not computably enumerable.
+(3) $\K$ is an enumerable set such that $\{\phi_i:i\in\K\}$ is the set of all partial computable prefix functions.
+(4) The finite sets: $\F$ is an enumerable set such that $\{\phi_i:i\in\F\}$ contains a uniform code for every finite set.
 A uniform code for a finite set $S$ is a total computable surjective function $f:\{0,1\}^{\lceil\log|S|\rceil}\to S$.
-\end{enumerate}
+
 We sometimes need to map binary strings to a prefix-free set.
@@
 To simplify notation, we will sometimes conflate natural numbers and binary strings, implicitly using the ordering $(0, \epsilon)$, $(1, 0)$, $(2, 1)$, $(3, 00)$, $(4, 01)$, \ldots
-
-Now define the Kolmogorov complexity of a binary string $x$, with respect to a model class $\M$ and a numbering $\psi$, as follows:
-\[
-C^{\M,\psi}(x)=\min\{\bar\imath y:\psi_i(y)=x,i\in\M\}.
-\]
-The model class $\C$ gives us the plain Kolmogorov complexity and $\K$ gives us the prefix-free Kolmogorov complexity.
-For $\C$ and $\K$, the complexity is invariant under a change of numbering, which may therefore be omitted from the superscript.
-
-A pair $(i,y)$ is a \emph{representation} for $x$ if $\phi_i(y)=x$.
+Now define the Kolmogorov complexity of a binary string $x$, with respect to a model class $\M$ and a numbering $\psi$, as follows $C^{\M,\psi}(x)=\min\{\bar\imath y:\psi_i(y)=x,i\in\M\}$.
+The model classes $\C$ and $\K$ give us the plain and prefix-free Kolmogorov complexity repectively.
 
@@
 Let $s^\psi$ denote any index sophistication with respect to numbering $\psi$.
-There are acceptable numberings $\psi$ and $\xi$ such that for all $x$
-\[|s^\psi(x) - s^\xi(x)|\geq \tfrac{1}{2}\min\{s^\psi(x),s^\xi(x)\}\p\]
+There are acceptable numberings $\psi$ and $\xi$ such that for all $x$:\\
+\-\hspace{2.8em} $|s^\psi(x) - s^\xi(x)|\geq \tfrac{1}{2}\min\{s^\psi(x),s^\xi(x)\}$.
 \end{lemma}
-
-To prove this, we simply use extremely inefficient numberings:
-
 \begin{proof}
 Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.
-
-Define $\psi$ and $\xi$ by
-\[
-\psi_j(x) = \begin{cases}
-	\phi_i(x) \;\text{if}\; j = s_{2i}\\
-	\infty\;\text{otherwise}
-\end{cases}\,\,\,
-\xi_j(x) = 
-\begin{cases}
-	\phi_i(x) \;\text{if}\; j = s_{2i+1} \\
-	\infty \;\text{otherwise}\p
-\end{cases}
-\]
-Choose any $x$ and assume w.l.o.g. that $s^\psi(x)\le s^\xi(x)$. By construction of $\psi$ and $\xi$ we have $2s_\psi(x)  \leq s_\xi(x)$ from which the result follows.
+Define $\phi, \xi$ such that $\psi_j(x) = \phi_i(x)$ for $j = s_{2i}$ and $\xi_j(x) = \phi_i(x)$ for $j = s_{2i+1}$, with all other functions returning $\infty$.
+Choose any $x$ and assume w.l.o.g.
+that $s^\psi(x)\le s^\xi(x)$.
+By construction of $\psi$ and $\xi$ we have $2s_\psi(x)  \leq s_\xi(x)$ from which the result follows.
 \end{proof}
@@
 \end{restatable}
-\noindent For other numberings, the singletons determine the sophistication, making it equal to the complexity:
 \begin{restatable}[Overfitting]{lemma}{overfitting}

### 2015-01-21 Facticity/Facticity/paper.tex
...

@@
 \end{restatable}
-This means that for model classes with a universal model, there is a numbering such that sophistication is bounded:
+This means that for model classes with a universal model, we can choose our numbering to make $S^\K$ bounded, or equal to the complexity.
 \begin{restatable}[underfitting]{lemma}{underfitting}
@@
 
-Thus, there are no longer any numberings for which sophistication is bounded, so the underfitting problem is at least partly solved.
+Thus, there are no longer any numberings for which sophistication is bounded.
 However, as we show below, the resulting sophistication is not an intuitive measure of the amount of structure in the data.
-Moreover, we push the balance back in the direction of overfitting, which is only exacerbated by this change.
+Moreover, the balance is pushed back in the direction of overfitting, which is only exacerbated by this change.
 Only one model class eliminates both the singletons and the universal model: $\T$.

### 2015-01-21 Facticity/Facticity/paper.tex
bib first pass

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 We will investigate this balance, starting with the sophistication $S^{\K,\psi,c}$.
-In this setting, we know that, on the one hand, there always exist candidate models---those models $i$ with $i \in \K$ and $C^{i}(x) \leq C^\K(x) + c$---for which al information except a constant amount is stored in the data, and on the other hand, there are also candidate representations for which all information is stored in the model.
+In this setting, we know that, on the one hand, there always exist candidate models---those models $i$ with $i \in \K$ and $C^{\{i\}}(x) \leq C^\K(x) + c$---for which al information except a constant amount is stored in the data, and on the other hand, there are also candidate representations for which all information is stored in the model.
 This tells us that $S^\K$ is \emph{balanced}: there is no inefficiency in the representation forcing information in or out of the model.
-The flip side of having a balanced sophistication is that it becomes easy to show invariance.
-We will show that there are numberings for which the universal models always determine the $S^\K$ and numberings for which the singletons always determine $S^\K$.
-The following lemma is based on the idea that, given a subset $\M'$ of the model class, we can choose the numbering so that representations use a model in $\M'$ perform an arbitrary constant better than all other models, so that a model in $\M'$ always ends up determining the sophistication.
-\begin{restatable}{lemma}{coolone}
-\label{lemma:thecoolone}
-  Let $\M$ be any model class, let $\X$ be any set of binary strings and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\phi$, i.e.
-if $m=\phi_{D(p)}$ then $p$ is a $D$-description of $m$.
-Let $\M'=\{\phi_i:i\in\tn{range}(D)\}$.
-Further assume there is a constant $c$ such that:\\
-\-\hspace{1cm}(1) $\forall_{m\in\M'}:\min\{|p|:\phi_{D(p)}=m\}\le C^{\K,\phi}(m)+c$\\
-\-\hspace{1cm}(2) $\forall_{x\in\X}:C^{\M',\phi}(x)-C^{\M,\phi}(x)\le c$.\\
-Then there is an enumeration $\psi$ of the partial recursive functions such that $S^{\M,\psi}(x) = |\br{0}|+S^{\M',\phi}(x)$ for all $x\in\X$.
-\end{restatable}
+The flip side of having a balanced sophistication is that it becomes easy to show invariance, by tweaking the numbering in such a way that models in a specific subset $\M'\subset\M$ of the model class becomes an arbitrary constant number of bits cheaper to represent than all other models, so that a model in $\M'$ always ends up determining the sophistication.
+
 This means that for model classes with a universal model, there is a numbering such that sophistication is bounded:
-\begin{restatable}[underfitting]{lemma}{underfitting}
+\begin{restatable}[Underfitting]{lemma}{underfitting}
 Let $\M$ be a model class containing a universal model $\phi_u$, with the property that $\exists c \forall i \in \M : C^{\{u\}}(x) \leq C^{\{i\}}(x) + c \p$.
@@
 
-\coolone*
+\begin{restatable}{lemma}{coolone}
+\label{lemma:thecoolone}
+  Let $\M$ be any model class, let $\X$ be any set of binary strings and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\phi$, i.e.
+if $m=\phi_{D(p)}$ then $p$ is a $D$-description of $m$.
+Let $\M'=\{\phi_i:i\in\tn{range}(D)\}$.
+Further assume there is a constant $c$ such that:\\
+\-\hspace{1cm}(1) $\forall_{m\in\M'}:\min\{|p|:\phi_{D(p)}=m\}\le C^{\K,\phi}(m)+c$\\
+\-\hspace{1cm}(2) $\forall_{x\in\X}:C^{\M',\phi}(x)-C^{\M,\phi}(x)\le c$.\\
+Then there is an enumeration $\psi$ of the partial recursive functions such that $S^{\M,\psi}(x) = |\br{0}|+S^{\M',\phi}(x)$ for all $x\in\X$.
+\end{restatable}
 
@@
 (1) Let $f$ be any function in the range of $D$, and $x$ its output.
-Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\phi_i(q)=x\}=C^{\K}(x) \le C^{\K}(f)+c$.
-(2) On the one hand $C^{\M',\psi}(x)\le C^\K(f)+|\epsilon|\le C^\K(x)+c$.
-On the other hand, $C^{\M,\psi}(x)$ is an effective description of $x$, so $C^{\K}(x)$ is at most a constant larger.
+Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\psi_i(q)=x\}=K(x)\le K(f)+c$.
+(2) On the one hand $L^{\M',\psi}(x)\le K(f)+|\epsilon|\le K(x)+c$.
+On the other hand, $L^{\M,\psi}(x)$ is an effective description of $x$, so $K(x)$ is at most a constant larger.
 Together, these inequalities establish the second condition.
 
-Now, by Lemma~\ref{lemma:thecoolone} there is an numbering $\psi_1,\psi_2,\ldots$ such that $\s^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
-We observed that $|C^{\K}(f)-C^{\K}(x)|\le c$ for any $f\in\M'$, so $\s^{\M',\psi}(x)\ge C^{\K}(x)-c$.
+Now, by Lemma~\ref{lemma:thecoolone} there is an enumeration $\phi_1,\phi_2,\ldots$ such that $S^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
+We observed that $|K(f)-K(x)|\le c$ for any $f\in\M'$, so $S^{\M',\psi}(x)\ge K(x)+c$.
 This proves the lemma.
@@
 That is, $\M' = \{u\}$.
-This construction satisfies the conditions 1 and 2 from Lemma~\ref{lemma:thecoolone}; invoking it we find that there exists an acceptable numbering $\psi$ for which $\s^{\M,\psi}(x) = \s^{\M', phi}(x) + c$.
-Since $\M'$ contains only a single model, $\s_{\M',\phi}(x)$ is constant.
+This construction satisfies the conditions 1 and 2 from Lemma~\ref{lemma:thecoolone}; invoking it we find that there exists an acceptable numbering for which $\s_\M(x) = \s_{\M'}(x) + c$.
+Since $\M'$ contains only a single model, $\s_{\M'}$ is constant.
 \end{proof}
@@
 
-Note that we favored brevity of the proof over the strength of the bounds: the multiplicative constant $1/6$ can be improved with a more detailed treatment.
+Note that we favoured brevity of the proof over the strength of the bounds: the multiplicative constant $1/6$ can be improved with a more detailed treatment.
 
@@
 Let $D$ be a prefix function with $D(\epsilon) = u$.
-We can instantiate Lemma~\ref{lemma:thecoolone} with $D$, $\M' = \{\phi_u\}$ and $X = \{x : d^{\C,c}(x) \leq A(C(x))\}$.
+We can instantiate Lemma~\ref{lemma:thecoolone} with $D$, $\M' = \{\phi_u\}$ and $X = \{x : d^\C_c(x) \leq A(C(x))\}$.
 This tells us that there exists a numbering for which $\s^\T(x) = \s^{\M'}(x) + |\bar0| \leq c$ for all $x \in X$.

### 2015-01-21 Facticity/Facticity/paper.tex

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 We will investigate this balance, starting with the sophistication $S^{\K,\psi,c}$.
-In this setting, we know that, on the one hand, there always exist candidate models---those models $i$ with $i \in \K$ and $C^{i}(x) \leq C^\K(x) + c$---for which al information except a constant amount is stored in the data, and on the other hand, there are also candidate representations for which all information is stored in the model.
+In this setting, we know that, on the one hand, there always exist candidate models---those models $i$ with $i \in \K$ and $C^{\{i\}}(x) \leq C^\K(x) + c$---for which al information except a constant amount is stored in the data, and on the other hand, there are also candidate representations for which all information is stored in the model.
 This tells us that $S^\K$ is \emph{balanced}: there is no inefficiency in the representation forcing information in or out of the model.
-The flip side of having a balanced sophistication is that it becomes easy to show invariance.
-We will show that there are numberings for which the universal models always determine the $S^\K$ and numberings for which the singletons always determine $S^\K$.
-The following lemma is based on the idea that, given a subset $\M'$ of the model class, we can choose the numbering so that representations use a model in $\M'$ perform an arbitrary constant better than all other models, so that a model in $\M'$ always ends up determining the sophistication.
-\begin{restatable}{lemma}{coolone}
-\label{lemma:thecoolone}
-  Let $\M$ be any model class, let $\X$ be any set of binary strings and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\phi$, i.e.
-if $m=\phi_{D(p)}$ then $p$ is a $D$-description of $m$.
-Let $\M'=\{\phi_i:i\in\tn{range}(D)\}$.
-Further assume there is a constant $c$ such that:\\
-\-\hspace{1cm}(1) $\forall_{m\in\M'}:\min\{|p|:\phi_{D(p)}=m\}\le C^{\K,\phi}(m)+c$\\
-\-\hspace{1cm}(2) $\forall_{x\in\X}:C^{\M',\phi}(x)-C^{\M,\phi}(x)\le c$.\\
-Then there is an enumeration $\psi$ of the partial recursive functions such that $S^{\M,\psi}(x) = |\br{0}|+S^{\M',\phi}(x)$ for all $x\in\X$.
-\end{restatable}
-This means that for model classes with a universal model, we can choose our numbering to make $S^\K$ bounded, or equal to the complexity.
+The flip side of having a balanced sophistication is that it becomes easy to show invariance, by tweaking the numbering in such a way that models in a specific subset $\M'\subset\M$ of the model class becomes an arbitrary constant number of bits cheaper to represent than all other models, so that a model in $\M'$ always ends up determining the sophistication.
+
+This means that for model classes with a universal model, we can choose our numbering to make $S^\K$ bounded, or equal to the complexity:
 \begin{restatable}[underfitting]{lemma}{underfitting}

### 2015-01-21 Facticity/Facticity/paper.tex
...

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 To simplify notation, we will sometimes conflate natural numbers and binary strings, implicitly using the ordering $(0, \epsilon)$, $(1, 0)$, $(2, 1)$, $(3, 00)$, $(4, 01)$, \ldots
-Now define the Kolmogorov complexity of a binary string $x$, with respect to a model class $\M$ and a numbering $\psi$, as follows $C^{\M,\psi}(x)=\min\{\bar\imath y:\psi_i(y)=x,i\in\M\}$.
-The model classes $\C$ and $\K$ give us the plain and prefix-free Kolmogorov complexity repectively.
+We define the $\M$-Kolmogorov complexity of a binary string $x$, with respect to a model class $\M$ and a numbering $\psi$, as follows $C^{\M,\psi}(x)=\min\{\bar\imath y:\psi_i(y)=x,i\in\M\}$.
+The model classes $\C$ and $\K$ give us the plain and prefix-free Kolmogorov complexity respectively.
 
@@
 \label{section:indices}
-In some works, the definition of sophistication is obtained by opening up the definition of Kolmogorov complexity, and measuring the length of the index part of that two-part code.
+In some works, the definition of sophistication is obtained by opening up the definition of Kolmogorov complexity, and measuring the length of the index in that two-part code.
 We call such a sophistication \emph{index sophistication}.

### 2015-01-21 Facticity/Facticity/paper.tex
...

@@
 Only one model class eliminates both the singletons and the universal model: $\T$.
-The only proposal that uses an efficient model representation \emph{and} excludes the universal models \emph{and} excludes the singletons is Vereshchagin and Vit\'anyi's \cite{vitanyi2004meaningful}.
+The only proposal that uses an efficient model representation \emph{and} excludes the universal models \emph{and} excludes the singletons is \cite{vitanyi2004meaningful}.
 While this invalidates straightforward proofs of boundedness, there is no suggestion that $S^{\T, \phi, c}(x)$ is actually invariant.
@@
 We note that it is exceedingly unlikely that a deep string is sampled from a shallow distribution \cite{bloem2014safe,bennett1988logical}.
+\cite{antunes2013sophistication} also investigates the relation between $S(x)$ and $d(x)$.
 
@@
 \end{restatable}
-\noindent This shows that while high-sophistication strings do exist, they do not behave as expected.
+\noindent This shows that while high-sophistication strings exist, they do not behave as expected.
 Consider a string that is typical for a shallow model, say some elaborate Markov chain.
-Under $\s^{\T,\psi,c}$, no matter how long an effective description of the Markov chain would have to be, the sophistication is bounded.
+Under $\s^{\T,\psi,c}$, no matter how high the complexity of the Markov chain, the sophistication is bounded.
 Any structure that is simple enough to be exploited within the time bound of the Ackermann function (or some other fast-growing, simply stated bound) will be counted as noise.
-Only structure that is so deep that if we used it to compress $x$ we could never hope to decompress it within the lifetime of the universe, would count towards sophistication.
+Only structure so deep that it would take beyond the lifetime of the universe to decompress would count towards sophistication.
+
 This clearly contradicts the motivating principle of sophistication, which is to capture the number of bits required to express \emph{all} regularity in the data, not just the regularity that is difficult to describe.
@@
 
-The relation between sophistication, depth and absolutely non-stochastic strings was also investigated in \cite{antunes2013sophistication}.
-
 \section{Discussion and conclusion}
@@
 
-We have criticized existing measures of the sophistication of a binary string.
-Our results show that there are technical problems with all of them.
+We have criticized existing measures of sophistication and shown technical problems with all of them.
 But that does not in itself mean that it should be impossible to come up with a sound measure.

### 2015-03-23 Facticity/Facticity/cie-abstract.tex
First draft CiE abstract

--- 2015-01-21 Facticity/Facticity/paper.tex
@@
 % --- DELETE BEFORE SUBMISSIONS ---
-\pagestyle{headings}
+% \pagestyle{headings}
 
-\title{Two Problems for Sophistication}
+\title{The problem of Sophistication}
 
@@
 \begin{abstract}
-Kolmogorov complexity measures the amount of information in data, but does not distinguish structure from noise.
-Kolmogorov's definition of the \emph{structure function} was the first attempt to measure only the structural information in data, by measuring the complexity of an objectively best model under two-part coding.
-Since then, many variations of this idea have been proposed, for which we use \emph{sophistication} as an umbrella term.
-We describe two fundamental problems with all existing proposals, showing each  to be \emph{unsound}.
-Consequently, we put forward the view that the problem is fundamental: it may be impossible to objectively quantify the sophistication.
+Kolmogorov complexity is a sound, unambiguous formalization of the information content of an object.
+However, researchers have long been plagued by its inability to express a more intuitive notion of complexity: in terms of Kolmogorov complexity, a completely random bitstring is the most complex object possible.
+This is not what we mean when we we use the word complexity informally: a television broadcast is complex when we can see structures, faces and landscapes, not when we can only see white noise.
+
+\hspace{0.05\textwidth} The desire to capture this notion of complexity in formal terms has led to many proposals for additional measures.
+The largest family of proposals is built on the following idea:  an object consists of \emph{structural} and \emph{stochastic} information, and the information content of the structural information determines the complexity of the object.
+We will use the name \emph{sophistication} for such proposals.
+
+\hspace{0.05\textwidth} In essence, sophistication performs model selection using the class of all computable models.
+This means that we are dealing not just with the problem of formally defining complexity, but also with a limiting case of statistics: what happens if we extend our model class to the entirety of conceivable models?
+Do we end up with an ultimate statistics machine, or does the whole framework collapse, and does model selection only work with carefully constrained model classes?
+
+\hspace{0.05\textwidth} While sophistication is a simple enough idea, any attempt to formalize it quickly runs into problems.
+We highlight the two major issues, and show how they invalidate each proposal.
+We are forced to conclude that no current proposal truly solves the problem of sophistication, and we offer some informal arguments which suggest that no such formalization may be possible at all.
+
 \end{abstract}
 
-\section{Introduction}\enlargethispage{4ex}
-Komogorov complexity gives us a sound definition of the amount of information contained in a binary string.
-It is, however, not what most people would consider complexity.
-For example, a sequence of a million coin flips will almost certainly have maximal Kolmogorov complexity, even though there is nothing complex about flipping a lot of coins---it's just a lot of work.
-Many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data: its \emph{sophistication}.
-In this paper, we investigate two serious problems with sophistication measures, and show why they are unsatisfactory.
-We believe that the problems are fundamental, and provide arguments why attempts to define sophistication must fail.
-
-The Kolmogorov complexity $C(x)$ of a binary string $x$, is (roughly) the length of the shortest computer program that will output $x$.
-This length depends on which programming language is used, but, as shown by a central result known as the Invariance Theorem\cite[Section~2.1]{li1993introduction}, the language affects the value only by a constant, independent of $x$.
-This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings, a \emph{property} of the data.
-A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees:
-
-\begin{enumerate}
-\item $S(x)$ should count the bits required for an effective description of the structural properties of a binary string.
-\item An analogue of invariance should hold: there must be strict limits on how much sophistication can be affected by a change in programming language.
-\item There should be no constant $c$ such that $S(x)\le c$ for every input $x$.
-If sophistication is bounded, then knowing its value under one programming language provides no constraints on its value under another language (except that it's also bounded).
-Moreover, a bounded sophistication would be at odds with the intuition that there is no limit to the amount of structure that a binary string can exhibit.
-\item Similarly, there should be no constant $c$ such that $C(x)-S(x)\le c$ for all $x$, because then sophistication would be equivalent to Kolmogorov complexity. 
-\end{enumerate}
-There have been many related proposals for such a measure, that are all based on a \emph{two-part code}: rather than encoding $x$ by means of a single computer program, we encode a \emph{model} for $x$ in the first part of the code, that is interpreted as a representation of $x$'s structural properties.
-The model does not fully specify $x$, but when combined with the second part of the code, which specifies the noise, the original string becomes fully determined.
-Thus, the model can be interpreted as a function mapping the noise to the original object.
-
-For any string $x$, there may be many different two-part representations.
-The total codelength can never be less than the Kolmogorov complexity, but it can come close.
-The balance between the two parts may vary; Figure~\ref{fig:diagram} provides an illustration.
-The various proposals then define sophistication in terms of the length of the structure part of this two-part code for one of the representations that (almost) achieves the Kolmogorov complexity.
-However, for most of these definitions, we can prove that they fail one or more of the conditions above.
-Vereshchagin and Vit\'anyi's definition \cite{vitanyi2004meaningful} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we show that their method only assigns substantial sophistication to strings that require an enormous amount of processing to construct.
-
-\begin{figure*}[tb]
-  \centering
-  \begin{minipage}{0.45\textwidth}
-     \includegraphics[width=\textwidth]{./img/sophistication.pdf}
-  \end{minipage}
-  \begin{minipage}{0.45\textwidth}
-     \includegraphics[width=\textwidth]{./img/sophistication-jump.pdf}
-  \end{minipage}
-  \caption{\small (left) Two-part representations of $x$ by the two components of their codelength.
-The Kolmogorov complexity $C^\M(x)$, appearing as a black diagonal, provides a lower bound on the total codelength.
-We consider only representations that are close to this optimum, with the threshold represented by a dashed line.
-The size of the smallest model below the threshold is the sophistication of the data.
-(right) The same image, after a constant perturbation in the model complexity caused by a change in numbering.}
-  \label{fig:diagram}
-\end{figure*}
-
-A valid definition of $S$ must contend with two important issues.
-First, the details of the way the model is encoded are important.
-There are two technically distinct approaches; in one of these one has to deal with the so-called ``nickname problem'' that strangely remains unresolved in several publications.
-These definitions yield a sophistication that is highly dependent on the chosen programming language, unless special care is taken, as discussed in Section~\ref{section:indices}.
-
-The second issue is that of striking the right balance between underfitting and overfitting, which we consider in Section~\ref{section:balance}.
-Overfitting is a common problem in the statistical literature, that refers to the tendency of model selection to choose a complex model that provides a very good fit to the observed data, but does not generalise well to unseen data.
-In the case of sophistication, overfitting occurs if the model that determines the sophistication contains much or even all of the noise.
-This is a real danger, because the models are encoded so efficiently.
-In statistics, overfitting is often addressed by penalising complex models.
-In sophistication, however, such penalties tend to break the balance between structural information and noise, and lead to the opposite problem: underfitting.
-
-Underfitting occurs when the selected model is simple, but fails to represent some of the structure present in the data.
-This is also a problem for sophistication because the considered models are so powerful.
-In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language.
-Such \emph{universal models} are \emph{simple} in the language of sophistication, since they can be described with a relatively small number of bits, yet are able to represent the data within a codelength a constant away from the Kolmogorov complexity.
-Such a two-part representation essentially encodes all information as noise.
-If complex models are penalized, then the problem becomes to make sure that universal models are not \emph{always} preferred for complex data.
-The usual workaround is to restrict the set of allowed models to exclude the universal ones.
-For example, the model class may be restricted to the set of all total recursive functions.
-While this excludes universal models, it is questionable whether it adequately solves the problem of underfitting in general.
-
-Finally, in the discussion in Section~\ref{section:conclusion} we argue that while two-part coding can yield useful insights into the structure of the data and allows us to identify some models as poor representations, it is probably not possible to uniquely separate structure from noise and identify a \emph{single} model as ``best'': in general many models, of vastly different complexities, may be reasonable representations.
-Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that the data allows for multiple equivalent interpretations of which information is structured, and which is random, and that there is no such thing as sophistication.
-
-\paragraph{Related work}
-The idea of separating structure from noise originated with Kolmogorov's \emph{structure function} \cite{cover1985kolmogorov}.
-In this theory, the data $x$ is described as finite set $A$ that contains $x$, together with the index of $x$ in $A$.
-The complexity of the set determines the sophistication.
-This approach was discussed more extensively, and generalised in \cite{vereshchagin2004kolmogorov,gacs2001algorithmic}.
-
-Koppel \cite{koppelSoph1988,koppel1991almost} also built on the structure function work to define a measure of the meaningful information in a binary string, coining \emph{sophistication}, the name we adopt in this paper.
-
-In 1996 Gell-Mann and Lloyd introduced \emph{effective complexity} \cite{gellmann1996information}.
-Working from the perspective of statistical mechanics, they arrived (independently, it seems) at a notion that fits the mold of sophistication.
-\emph{Facticity} \cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions.
-Adriaans also identifies the nickname problem discussed in Section~\ref{section:indices}.
-
-Other authors have pointed out problems with sophistication.
-Antunes et al.
-\cite{antunes2009sophistication} note that the constant cutoff used by Koppel is a source of instability, a result in the spirit of this paper.
-They propose a different definition called \emph{coarse sophistication}.
-This version behaves very differently from regular sophistication, and is closer to computational depth.
-Vereshchagin \cite{vereshchagin2013algorithmic} also points out that the Kolmogorov minimal sufficient statistic is not always well-defined and suggests requiring that the finite set model should described using a \emph{total} function rather than any computable function.
-This yields a more robust sophistication that is nonetheless still susceptible to the issues discussed in this paper.
-\section{Preliminaries}
-Let $\B = \{0,1\}^*$.
-We deal with partial computable functions $\phi: \B \to \B$, which we will also call \emph{models}.
-A function is called \emph{prefix} if its domain, $\dom{f} = \{y : f(y) \neq \infty\}$ is a prefix free set, i.e.
-no string in $\dom{f}$ is a prefix of another.
-A function $f$ is \emph{total} if $\dom{f} = \B$.
-
-A \emph{numbering} is an enumeration of of the partial computable functions, denoted as $\psi_1, \psi_2, \ldots$ or simply $\psi$.
-We fix one canonical effective numbering $\phi$.
-We call a numbering $\psi$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ with $\forall: i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{a(i)}$.
-
-A \emph{model class} is a set of model indices.
-We distinguish four important model classes: (1) The indices of the partial computable functions $\C=\N$.
-(2) The total functions $\T=\{i:\tn{$\phi_i$ is total}\}$.
-Note that $\T$ is not computably enumerable.
-(3) $\K$ is an enumerable set such that $\{\phi_i:i\in\K\}$ is the set of all partial computable prefix functions.
-(4) The finite sets: $\F$ is an enumerable set such that $\{\phi_i:i\in\F\}$ contains a uniform code for every finite set.
-A uniform code for a finite set $S$ is a total computable surjective function $f:\{0,1\}^{\lceil\log|S|\rceil}\to S$.
-
-We sometimes need to map binary strings to a prefix-free set.
-We denote by $\br{x}$ the prefix-encoded representation for $x$.
-We require that the mapping satisfies $|\br{x}| = |x|+O(\log(|x|)$ (see eg.
-\cite[Section~1.4]{li1993introduction}).
-To simplify notation, we will sometimes conflate natural numbers and binary strings, implicitly using the ordering $(0, \epsilon)$, $(1, 0)$, $(2, 1)$, $(3, 00)$, $(4, 01)$, \ldots
-We define the $\M$-Kolmogorov complexity of a binary string $x$, with respect to a model class $\M$ and a numbering $\psi$, as follows $C^{\M,\psi}(x)=\min\{\bar\imath y:\psi_i(y)=x,i\in\M\}$.
-The model classes $\C$ and $\K$ give us the plain and prefix-free Kolmogorov complexity respectively.
-
-\section{Inefficient indices}
-\label{section:indices}
-In some works, the definition of sophistication is obtained by opening up the definition of Kolmogorov complexity, and measuring the length of the index in that two-part code.
-We call such a sophistication \emph{index sophistication}.
-Examples are Koppel and Atlan's works \cite{koppelSoph1988,koppel1991almost},  and several of its variants \cite{antunes2009sophistication,antunes2013sophistication}.
-
-In these works, no constraints are placed on the numbering.
-This makes these approaches vulnerable to the so-called \emph{nickname problem}: the number of bits required to specify a model becomes very dependent on the numbering.
-This problem is severe: \emph{any} index sophistication without requirements on the numbering is either bounded, or nowhere near invariant:
-
-\begin{lemma}
-Let $s^\psi$ denote any index sophistication with respect to numbering $\psi$.
-There are acceptable numberings $\psi$ and $\xi$ such that for all $x$:\\
-\-\hspace{2.8em} $|s^\psi(x) - s^\xi(x)|\geq \tfrac{1}{2}\min\{s^\psi(x),s^\xi(x)\}$.
-\end{lemma}
-\begin{proof}
-Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.
-Define $\phi, \xi$ such that $\psi_j(x) = \phi_i(x)$ for $j = s_{2i}$ and $\xi_j(x) = \phi_i(x)$ for $j = s_{2i+1}$, with all other functions returning $\infty$.
-Choose any $x$ and assume w.l.o.g.
-that $s^\psi(x)\le s^\xi(x)$.
-By construction of $\psi$ and $\xi$ we have $2s_\psi(x)  \leq s_\xi(x)$ from which the result follows.
-\end{proof}
-Thus, even though the Kolmogorov complexity changes by only a constant number of bits under a switch to a different numbering, the plain length of the index in a given acceptable numbering is a very poor indicator of model complexity.
-
-In order to obtain a reasonable and robust measure of model complexity, we define the Kolmogorov complexity of functions as in \cite{grunwald2004shannon,vitanyi2004meaningful} by $C^{\M,\psi}(f) = \min\{C^{\M,\psi}(i):\phi_i=f\}$.
-Lemma~\ref{lemma:invariance} in the appendix shows that this definition is invariant: the Kolmogorov complexity is an objective property of a function.
-
-There are two ways to use this notion of model complexity for more robust attempts to define sophistication. Confusingly, both approaches are used in the literature.
-First, we can measure the complexity of the \emph{model} using $C^{\M,\psi}(\phi_i)$, which is then the first part of a two-part code describing the data.
-This approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information} and in most of the present paper.
-
-Second, we can stick to using the length of the index as the measure of sophistication, but impose the additional constraint that an efficient acceptable numbering is used.
-This is the approach taken by Adriaans \cite{adriaans2012facticity}.
-Before switching to the other approach we construct such an efficient index.
-
-We call a numbering $\psi$ \emph{faithful} if for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and $|j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
-This definition is a refinement of \cite[Definition~10]{adriaans2012facticity}.
-We prove in the appendix that there actually do exist faithful, acceptable numberings (Lemma~\ref{lemma:faithful-numberings}).
-However even choosing a faithful index is not quite enough.
-The Kolmogorov complexity uses representations of the form $\bar\imath y$, where the bar denotes some straightforward prefix encoding to delimit the model description $i$ from its input $y$.
-If we define a second prefix encoding $\tilde{\imath}$, with $|\bar\imath|-|\tilde\imath|$ unbounded, we can define a second representation $\bar u \tilde \imath y$, at a constant overhead $|\br{u}|$, and gain more than $|\br{u}|$ for sufficiently complex strings.
-As shown in Lemma~\ref{lemma:prefix-inefficiency} in the appendix, this results in a bounded sophistication.
-This is in fact an issue with \cite{adriaans2012facticity}. Moreover, we conjecture that \emph{no} computable prefix function $\bar\cdot$ is sufficiently efficient to avoid this.
-
-Because of these issues, rather than index sophistication we will use the following definition of sophistication in the remainder of this article.
-It is parameterised by a model class, a numbering, and a slack parameter $c$ that determines the maximal allowed slack between the two-part code length using a representation and the Kolmogorov complexity.
-\[
-S^{\M,\psi,c}(x)=\min\left\{C^\K(\phi_i):i\in\M,C^{\{i\},\psi}(x)\le C^{\M,\psi}(x)+c\right\}.
-\]
-
-\section{Balancing under- and overfitting}
-\label{section:balance}
-
-In the previous section, we saw the first glimpse of the delicate balance between the two code components in sophistication.
-If the description of the model $m$ is less efficient than $C^\K(m)$ by anything more than a constant, the sophistication becomes bounded.
-We will investigate this balance, starting with the sophistication $S^{\K,\psi,c}$.
-In this setting, we know that, on the one hand, there always exist candidate models---those models $i$ with $i \in \K$ and $C^{\{i\}}(x) \leq C^\K(x) + c$---for which al information except a constant amount is stored in the data, and on the other hand, there are also candidate representations for which all information is stored in the model.
-This tells us that $S^\K$ is \emph{balanced}: there is no inefficiency in the representation forcing information in or out of the model.
-The flip side of having a balanced sophistication is that it becomes easy to show invariance, by tweaking the numbering in such a way that models in a specific subset $\M'\subset\M$ of the model class becomes an arbitrary constant number of bits cheaper to represent than all other models, so that a model in $\M'$ always ends up determining the sophistication.
-
-This means that for model classes with a universal model, we can choose our numbering to make $S^\K$ bounded, or equal to the complexity:
-\begin{restatable}[underfitting]{lemma}{underfitting}
-Let $\M$ be a model class containing a universal model $\phi_u$, with the property that $\exists c \forall i \in \M : C^{\{u\}}(x) \leq C^{\{i\}}(x) + c \p$.
-Then, there exist numberings such that $\s^\M$ is bounded.
-\end{restatable}
-\begin{restatable}[Overfitting]{lemma}{overfitting}
-Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $i\in\M$ with $\phi_i(\epsilon)=x$.
-Then there is a numbering $\psi$, and a constant $c$, such that for all $x\in\X$ we have $C^\M(x)-S^{\M,\psi}(x)\leq c$.\label{lemma:overfitting}
-\end{restatable}
-\noindent Thus, in this balanced sophistication, there is no invariance: all information can be seen as structure, or as noise, depending on the numbering.
-To avoid these issues, existing proposals upset the balance to exclude or penalize at least the singleton models and the universal models.
-
-\subsection{Overfitting}
-
-The existing proposals that can overfit are the minimal sufficient statistic in the structure function \cite{cover1985kolmogorov}, the sophistication as defined in \cite{mota2013sophistication}, the minimal sufficient algorithmic statistic \cite{gacs2001algorithmic} and Vereshchagin's strongly sufficient statistic \cite{vereshchagin2013algorithmic}.
-It is widely recognized that singleton models are always candidates in these settings, but not, it seems, that they may be the only candidates for any string, as Lemma~\ref{lemma:overfitting} shows.
-
-It may be argued that the slack parameter $c$ in the sophistication, that determines the allowed gap between a candidate representation and the complexity, should be allowed to depend on the numbering, but this dependence has not been mentioned in the literature and there is no clear way to determine how this constant should be determined on the basis of a given numbering.
-
-In traditional statistics, overfitting is often addressed by imposing a penalty on complex models.
-As we have seen, a strong penalty, such as the one imposed by an inefficient prefix encoding of the model, will immediately cause underfitting.
-A more subtle approach is to use $\C$ as a model class rather than $\K$, thus allowing representations that are not self-delimiting.
-The gap between self-delimiting and non self-delimiting descriptions grows without bound
-\cite[Section~4.5.5]{vitanyi2004meaningful}, so that for sufficiently complex inputs all information will have to be placed in the noise part of the code, eliminating the singletons as viable candidates.
-This approach is taken by Vit\'anyi \cite{vitanyi2004meaningful} and by Adriaans \cite{adriaans2012facticity}.
-Such measures thus help reduce the overfitting problem, but they only increase the tendency to underfit.
-We also pay the price that the models can no longer be equated with probability measures, removing the link to traditional statistics and probability theory.
-
-\subsection{Underfitting}
-
-As mentioned in the related work, underfitting is a widely acknowledged problem for sophistication, and most proposals avoid it by limiting the class of allowed models, to exclude universal models.
-The following theorem shows that this does indeed yield an unbounded sophistication.
-
-\begin{restatable}{theorem}{dogfood}
-There are infinitely many $x$ with $\s^{\F,\phi,c}(x) \geq C^{\K}(x)/5$, and also infinitely many $x$ with $\s^{\T,\phi,c}(x) \geq C^{\K}(x)/6$.
-\end{restatable}
-
-Thus, there are no longer any numberings for which sophistication is bounded.
-However, as we show below, the resulting sophistication is not an intuitive measure of the amount of structure in the data.
-Moreover, the balance is pushed back in the direction of overfitting, which is only exacerbated by this change.
-Only one model class eliminates both the singletons and the universal model: $\T$.
-The only proposal that uses an efficient model representation \emph{and} excludes the universal models \emph{and} excludes the singletons is \cite{vitanyi2004meaningful}.
-While this invalidates straightforward proofs of boundedness, there is no suggestion that $S^{\T, \phi, c}(x)$ is actually invariant.
-
-While under $S^{\T, \phi, c}(x)$ there are strings with high sophistication, these may not conform to the original intuition that motivated sophistication. To show this, we need the concept of depth:
-
-\begin{definition}[Depth\cite{bennett1988logical,antunes2006computational}]\belowdisplayskip=-12pt
-Let $U$ be some universal Turing machine, so that $U(\bar\imath y) = \phi_i(y)$.
-Let $U^t$ be a simulation of this machine, which is allowed to run for at most $t$ steps, and returns $0$ if it has not yet finished at that point.
-Let $C^\M_t(x) = \min\{|\bar\imath y| : U^t(\bar\imath y) = x, \phi_i \in \M\}$.
-The \emph{$c$-depth} is $d^{\M,c}(x) = \min \left\{t : C^\M_t(x) - C^\M(x) \leq c \right\}$.
-\end{definition}
-Deep strings are those that can only be optimally compressed with a great investment of time.
-We note that it is exceedingly unlikely that a deep string is sampled from a shallow distribution \cite{bloem2014safe,bennett1988logical}.
-\cite{antunes2013sophistication} also investigates the relation between $S(x)$ and $d(x)$.
-
-\begin{restatable}{theorem}{depth}
-Let $A(n)$ be the single-argument Ackermann function and $c$ some arbitrary constant.
-There are numberings such that for strings with depth $d^{\C,c}(x) \leq A(C^\C(x))$ the sophistication $\s^{\T, \psi, c}$ is bounded.
-The same is true for $\s^{\F}$ .
-\end{restatable}
-\noindent This shows that while high-sophistication strings exist, they do not behave as expected.
-Consider a string that is typical for a shallow model, say some elaborate Markov chain.
-Under $\s^{\T,\psi,c}$, no matter how high the complexity of the Markov chain, the sophistication is bounded.
-Any structure that is simple enough to be exploited within the time bound of the Ackermann function (or some other fast-growing, simply stated bound) will be counted as noise.
-Only structure so deep that it would take beyond the lifetime of the universe to decompress would count towards sophistication.
-
-This clearly contradicts the motivating principle of sophistication, which is to capture the number of bits required to express \emph{all} regularity in the data, not just the regularity that is difficult to describe.
-In this setting, a string high-sophistication not only has a lot of structure, but it also has the property that it can be compressed far better with partial functions than with total.
-That is, it is non-typical for the model class $\T$.
-This suggests that the `non-stochastic' property of strings with high sophistication \cite{shen1983concept,vereshchagin2004kolmogorov} says more about depth and totality than it does about structure and noise.
-
-\section{Discussion and conclusion}
-\label{section:conclusion}
-
-We have criticized existing measures of sophistication and shown technical problems with all of them.
-But that does not in itself mean that it should be impossible to come up with a sound measure.
-The intuition throughout the literature, starting with the structure function, appears to be that the crucial property is whether a string is typical for a model.
-If so, another random choice from that mode selects a string with the same structure: a string of the same `type'.
-
-This idea is bold, but not necessarily unreasonable.
-Nevertheless, in this section we offer the opinion that such a clean-cut separation \emph{cannot} be made to work.
-We provide two arguments to support this belief.
-
-For the first argument, we take a generative perspective.
-We can generate data from a model in $\phi_i, i \in \K$ by feeding it random bits until it produces an output.
-We will call the resulting probability distribution $p_i(x) = \sum_{y:\phi_i(y) = x} 2^{-|y|}$.
-In this scenario, a sophistication could be called \emph{consistent} if, for data of sufficiently high complexity, it accurately reflects the complexity of that distribution from which it originated.
-Suppose that the data issampled from  universal distribution: let $\phi_u(\bar\imath y)=\phi_i(y)$ and sample from $p_u$ as described above.
-Then the initial random bits will determine the prefix encoded index $\bar\imath$ of the function $\phi_i$ that $\phi_u$ will subsequently emulate, and the remaining bits are used as inputs to $\phi_i$.
-We now ask, what should be the sophistication of the resulting data?
-
-Certainly, if we have to judge based only on the data, we cannot exclude the possibility that the data was sampled from $p_u$: after all, it was.
-On the other hand, neither can we exclude the possibility that it came from $p_i$, as again, it did!
-It seems unjustified to assume that, based on only the data, all but one model can always be disqualified.
-
-Consider the following metaphor.
-We are given a bitstring which can be read as a bitmap image of the painting \emph{Impression of a Sunrise}.
-There are many good models for this string, ranging from very generic to very specific.
-The theory behind sophistication suggests that we can choose one of these as the objective, intrinsic model of the data.
-Different models say different things: the universal model says that it is `some finite object'.
-A more specific model might say that it is `an image'.
-Even more specific would be a painting, a Monet, or specifically the painting Impression of a Sunrise.
-A sound measure of Sophistication should be able to select one of these models as the proper representation of structure in the data, and disqualify the others as either over- or underfitting.
-But how should we be able to say that the data is intrinsically more of a painting than an image?
-More of a Monet than a painting?
-Intuitively, such distinctions require further assumptions, or a second sample from the same distribution.
-
-The second reason we believe sophistication will elude definition is more technical.
-Given an input $x$ and some model class, say $\C$, consider the set of all possible two-part representations of $x$.
-Now, when the numbering of partial computable functions is changed, the codelength of the model part of all these representations will change.
-This is illustrated in the right graph in Figure~\ref{fig:diagram}.
-However, the Invariance Theorem expresses that this change is limited by a constant term that does not depend on $x$.
-So for sufficiently large $x$, the horizontal shift of all the representations will be relatively small.
-However, even this small shift can push some representations out of the acceptable region (indicated by the dashed line), and pull others in.
-This may lead to a different representation determining the sophistication, one whose \emph{total} codelength is close to what it was before, but whose \emph{model} codelength can be anywhere between $0$ and $C^\C(x)$.
-If such jumps can occur, the sophistication is not invariant.
-And while we cannot \emph{prove} in general that such jumps can always occur, there seems to be no reason to believe that they don't.
-
-For this reason, we take a skeptical view of sophistication.
-Note that part of the theory is fine: there is nothing wrong with evaluating models for the data by comparing their two-part code lengths.
-In fact, the randomness deficiency $\delta(\phi_i,x)=-\log p_i(x)-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data---if $p_i$ truly generated the data, then the probability that the randomness deficiency is more than $k$ bits is less than $2^{-k}$ \cite{bloem2014safe}[Lemma~6].
-In the Monet example above, this will allow us to disqualify the model expressing that the data is actually, say, a recording of Jazz music.
-
-But fundamental problems arise as soon as hard cut-offs are introduced.
-In our opinion, a lot of measures taken in the literature to address problems with sophistication, such as restricting the model class or introducing model penalties, complicate the method and make problems harder to analyse, without actually addressing the fundamental issue.
-This is dangerous: if such ad-hoc fixes result in a theory that is hard to prove either wrong or right, it creates an artificial dead end for a valuable area of research.
-When hard cut-offs are avoided, on the other hand, all such measures are suddenly no longer necessary.
-What remains is a simple and elegant theory that can be used to sift through all possible models, disproving most while retaining a select number of interesting candidates for our further consideration.
-
-\subsubsection*{\ackname}
-
-This publication was supported by the Dutch national program COMMIT and by  the Netherlands eScience center.
-
-\bibliographystyle{plain}
-\bibliography{facticity}
-
-\appendix
-
-\section{Proofs}
-
-\subsection{Invariance of function complexity}
-
-\begin{lemma}[Invariance of function complexity]
-Let $\psi$ and $\eta$ be any two acceptable numberings.
-There exists a constant $c$ such that $\left| C^{\K,\psi}(f) - C^{\K, \eta}(f)\right | \leq c$ for all $f \in \C$ and $\left| C^{\C,\psi}(f) - C^{\C, \eta}(f)\right | \leq c$ for all $f \in \C$.
-\label{lemma:invariance}
-\end{lemma}
-\begin{proof}
-Let $g(i)$ be the function such that $\psi_i=\eta_{g(i)}$.
-\begin{align*}
-C^{\C,\psi}(f) &= \min\left\{ C^{\C,\psi}(i) : \psi_i= f\right\} 
-\geq \min\left\{ C^{\C, \eta}(i) : \psi_i= f\right\} - c\\
-&= \min\left\{ C^{\C, \eta}(i) : \eta_{g(i)}= f\right\} - c
-\geq \min\left\{ C^{\C, \eta}(i) : \eta_i= f\right\} - c' \\
-&= C^{\C, \eta}(f).
-\end{align*}
-The opposite inequality can be achieved by reversing $\psi$ and $\eta$. The same proof holds with model class $\K$.
-\end{proof}
-This lemma shows that the complexity of a function does not depend on the choice of
-enumeration by more than a constant term.
-
-\begin{restatable}{lemma}{faithful}\label{lemma:faithful-numberings}
-  There are faithful acceptable numberings.
-\end{restatable}
-
-\begin{proof}
-Let $d \in \N$ be an index such that $\phi_d(y)=\infty$ for all $y$. Define
-  \[\psi_q=\begin{cases}
-    \phi_{\phi_i(p)}&\tn{if $q$ can be written as $\bar\imath p$ and $\phi_i(p)<\infty$,}\\
-    \phi_d&\tn{otherwise.}\end{cases}
-  \]
-  To show that $\psi$ is faithful, pick any function $f$. Then
-\[\begin{split}
-C^{\C, \phi}(f)&=\min\{C^{\C, \phi}(i):\phi_i=f\} =\min\{\min\{|\bar a b|:\phi_a(b)=i\}:\phi_i=f\} \\
-& =\min\{|\bar a b|:\phi_{\phi_a(b)}=f\}
- =\min\{|\bar a b|:\psi_{\bar a b}=f\}.
-\end{split}\]
-This shows there is a sufficiently small $\psi$ index.
-
-To show that $\psi$ is acceptable, let $\phi_j$ denote the identity
-function. Then a $\phi$-index $i$ can be mapped to a $\psi$-index
-using the computable function $r(i)=\bar\jmath i$, so that
-$\psi_{r(i)}(y)=\psi_{\bar\jmath i}(y)=\phi_i(y)$. For the reverse,
-define $\phi_v(\bar\imath p, y)=\phi_{\phi_i(p)}(y)$. For fixed
-$\bar\imath p$, the 
-$s^n_m$-theorem \cite{kleene193notation} states that we can compute the $h$
-such that $\phi_h(y)=\phi_v(\bar\imath p,y)$. Let $h(\bar\imath p)$
-denote this index as a function of the program; further define
-$h(q)=d$ if $q$ cannot be expressed as $\bar\imath p$. By
-construction $h$ is total and computable. To check that the mapping
-returns the correct function, rewrite $\phi_{h(\bar\imath
-  p)}(y)=\phi_v(\bar\imath p,y)=\phi_{\phi_i(p)}(y)=\psi_{\bar\imath p}(y)$.
-\end{proof}
-
-\subsection{Inefficient indices}
-\label{section:appendix-inefficient-indices}
-
-Can we make our prefix encoding $\bar\cdot$ so efficient that no other encoding can undercut it?
-By definition, $C^{\K}(\phi_i)$ is the smallest effective, self-delimiting description of $\phi_i$.
-We cannot compute the prefix code from the data, but this is not necessary to define a representation.
-Let $i^*$ be the shortest self-delimiting program for $\phi_i$, then $i^*y$ is a decodable representation for $x$.
-If our prefix function is less efficient than $C^\K(m)$ by an unbounded amount, $i^*y$ will always beat $\bar\imath y$ for complex enough data.
-We conjecture that no computable prefix encoding function can avoid this issue:
-\begin{conjecture}
-For any partial computable prefix codelength function $L$, the function $\min\{L(i)-C^\K(i):i\ge i_0\}$
-is unbounded in $i_0$.
-\end{conjecture}
-The following lemma shows that even index sophistications using a faithful index are bounded if the prefix encoding is inefficient.
-
-\begin{restatable}{lemma}{ineffprefix}
-Assume $\phi$ is faithful and define the index sophistication $s^c(x)=\min\{|\bar\imath|:C^{\{i\}}(x)\le C^{\C}(x)+c\}$.
-If $\min\{|\bar\imath|-C^\K(i):i\ge i_0\}$ is an unbounded function of $i_0$, then $s^c$ is bounded.\label{lemma:prefix-inefficiency}
-\end{restatable}
-If our conjecture is correct, the index sophistication is always bounded for model class $\C$.
-Any inefficieny in the model description, and a second represention using a universal model with a more efficient model description will win, causing a bounded sophistication.
-
-\begin{proof}
-Let $\bar\imath y$ be any two-part representation for the data $x$, i.e.
-$\phi_i(y)=x$.
-Then construct an alternative two-part representation $\bar vi^* y$, where $i^*$ is the shortest $\psi_u$-program for $i$ such that $\phi_v(i^* y)=\phi_{\psi_u(i^*)}(y) = \phi_i(y)=x$.
-We compare the lengths of these two representations.
-Note that $|i^*|=C^{\K,\phi}(i)$.
-Therefore,
-\[
-|\bar\imath y|-|\bar v i^* y| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-C^{\K}(i).
-\]
-By assumption there must be an $i_0$ such that the above expression is positive for all $i>i_0$.
-From this $i_0$ onwards, the second representation (using $v$) will have a shorter codelength, so $\bar\imath p$ cannot achieve the minimum in the definition of the sophistication.
-Consequently, the sophistication is bounded by $s(x)<|\overline{\imath_0}|$ for all $x$.
-\end{proof}
-
-\subsection{Boosting submodel by manipulating the numbering}
-
-\begin{restatable}{lemma}{coolone}
-\label{lemma:thecoolone}
-  Let $\M$ be any model class, let $\X$ be any set of binary strings and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\phi$, i.e.
-if $m=\phi_{D(p)}$ then $p$ is a $D$-description of $m$.
-Let $\M'=\{\phi_i:i\in\tn{range}(D)\}$.
-Further assume there is a constant $c$ such that:\\
-\-\hspace{1cm}(1) $\forall_{m\in\M'}:\min\{|p|:\phi_{D(p)}=m\}\le C^{\K,\phi}(m)+c$\\
-\-\hspace{1cm}(2) $\forall_{x\in\X}:C^{\M',\phi}(x)-C^{\M,\phi}(x)\le c$.\\
-Then there is an enumeration $\psi$ of the partial recursive functions such that $S^{\M,\psi}(x) = |\br{0}|+S^{\M',\phi}(x)$ for all $x\in\X$.
-\end{restatable}
-
-\begin{proof}
-We define the numbering $\psi$ as follows:
-\[\begin{cases}
-\psi_0(p) = 1^r 0 D(p) \\
-\psi_{1^r0i}(p) = \phi_i(p) \\
-\psi_j(\cdot) = \infty &\text{if $j$ contains no zeroes.}
-\end{cases}\]
-We will show that under the $\psi$-enumeration, the best representation for $x$ using a model $m\in\M'$ is always better than the best representation using some $m\notin\M'$.
-
-First suppose $m\in\M'$. Then
-\begin{align*}
-C^{\K,\psi}(m) &=\min\{|\bar\jmath q|:\psi_{\psi_j(q)}=m\}
-\leq\min\{|\bar0 q|:\psi_{\psi_0(q)}=m\} \\ 
-&=\min\{|\bar0 q|:\psi_{1^r0 D(q)}=m\}  
-=\min\{|q|:\phi_{D(q)}=m\} + |\bar 0|\\
-&\leq C^{\K,\phi}(m)+c+|\br{0}|,
-\end{align*}
-where the last inequality uses assumption (1).
-
-Now assume that the best model $m$ for $x$ is not in $\M'$. 
-Let $i$ be the index of $m$ with the shortest description, i.e. it achieves the minimum in $C^{\K, \psi}(m)=\min\left\{C^{\K,\psi}(i):\psi_i=m\right\}$.
-There are two possibilities. Either $i = 0$, in which case we have $C^{\K,\psi}(m)\ge r$ because $\phi_0$ cannot output zero and all other $\psi$-programs are at least $r$ bits long.
-
-Otherwise, we can bound
-\begin{align*}
-C^{\K,\psi}(m)&=C^{\K,\psi}(i)=C^{\K,\psi}(1^r0j)\\
-&\ge C^{\K,\psi}(j)-c' \ge C^{\K,\phi}(j)+r+1-c'\\
-&=C^{\K,\phi}(m)+r+1-c'.
-\end{align*}
-Now choose $r=c''+\max\left\{C^{\K,\psi}(\phi_0),c'-1\right\}$. Then substitution yields, for both cases, 
-$C^{\K,\phi}(m)\ge c''+C^{\K,\psi}(m)$.
-
-Combining the inequalities above, for any model $g\not\in\M'$, there is a model $f\in\M$ such that
-\[\begin{split}
-C^{\K,\psi}(g)+\min \{|y| : g(y) = x\} &\ge C^{\K,^\phi}(g)+\min\{|y| : g(y) = x\}+c''\\
-&\ge C^{\K,\psi}(f)+\min \{|y| : f(y) = x\} -2c+c'' -|\bar0|.
-\end{split}\]
-By choosing $c''$ sufficiently large, we can ensure that the best representation is in $\M'$ for all $x\in\X$, which completes the proof.
-\end{proof}
-
-\subsection{Overfitting}
-
-\overfitting*
-
-\begin{proof}
-Let $\phi_i$ be a singleton for $x$ as described.
-Note that since $\phi_i$ is a prefix function, if $\phi_i$ is defined for input $\epsilon$ then it cannot be defined for any other input.
-Pick $i,x$ with $\phi_i(\epsilon)=x$.
-Note that $x$ can be computed from $i$ and a fixed program, so there is a $c$ such that $C^{\K, \phi}(x)\le C^{\K, \phi}(\phi_i)+c$.
-Vice versa, given any $x$ we can construct an index of $\phi_i$, since $\phi$ is an acceptable numbering.
-Therefore $|C^{\K,\phi}(\phi_i)-C^{\K, \phi}(x)|\le c$.
-
-We now define a computable function $D$ by $D(\bar\imath y)=j$ where $\psi_j(\epsilon) = \psi_i(y)$.
-We will show that the two conditions of Lemma~\ref{lemma:thecoolone} hold for the prefix function $D$.
-
-(1) Let $f$ be any function in the range of $D$, and $x$ its output.
-Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\phi_i(q)=x\}=C^{\K}(x) \le C^{\K}(f)+c$.
-(2) On the one hand $C^{\M',\psi}(x)\le C^\K(f)+|\epsilon|\le C^\K(x)+c$.
-On the other hand, $C^{\M,\psi}(x)$ is an effective description of $x$, so $C^{\K}(x)$ is at most a constant larger.
-Together, these inequalities establish the second condition.
-
-Now, by Lemma~\ref{lemma:thecoolone} there is an numbering $\psi_1,\psi_2,\ldots$ such that $\s^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
-We observed that $|C^{\K}(f)-C^{\K}(x)|\le c$ for any $f\in\M'$, so $\s^{\M',\psi}(x)\ge C^{\K}(x)-c$.
-This proves the lemma.
-\end{proof}
-\subsection{Underfitting}
-
-\underfitting*
-
-\begin{proof}
-Let $D$ be a prefix function as in Lemma~\ref{lemma:thecoolone} such that it returns the index of $u$ for the argument $\epsilon$ and $\infty$ for any other argument.
-That is, $\M' = \{u\}$.
-This construction satisfies the conditions 1 and 2 from Lemma~\ref{lemma:thecoolone}; invoking it we find that there exists an acceptable numbering $\psi$ for which $\s^{\M,\psi}(x) = \s^{\M', phi}(x) + c$.
-Since $\M'$ contains only a single model, $\s_{\M',\phi}(x)$ is constant.
-\end{proof}
-
-\subsection{The problem of depth}
-
-We first show that representations using $\T$ can be reduced to representations in $\F$ with a small overhead.
-
-\begin{definition}[Stochastic strings, Definition~1.1 from~\cite{gacs2001algorithmic}]
-  Let $\alpha$ and $\beta$ be natural numbers. A finite binary string $x$ is called \emph{$(\alpha,\beta)$-stochastic} if there exists a finite set $S\subseteq\B$ such that 
-  \[
-  x\in S,~C^\K(S)\le\alpha,~C^\K(x)\ge\log|A|-\beta.
-\]
-\end{definition}
-
-\begin{restatable}{lemma}{vitanyi}
-Let $f$ be a total, computable function, such that for some $d$, $f(d) = x$ and let $k = K(f) + |d|$.
-Then there exists a finite set $A$ containing $x$ and a constant $c$, such that $K(A) \leq K(f) + K(|d|) + c$ and $\log |A| \leq |d|$.\label{lemma:total-to-sets}
-\end{restatable}
-\begin{proof}
-Let $S = f\left(\{0,1\}^{|d|}\right)$.
-Since $f$ is total, this set can be explicity computed from a description of $f$ and a description of $|d|$, which tells us that there is a constant $c$ such that $C^\K(A) \leq C^\K(f) + C^\K(|d|) + c$.
-
-Since $A$ is the image of a set of size $2^{|d|}$ under $f$, we have $\log |A| \leq |d|$.
-\end{proof}
-\noindent This lemma is a variation on \cite[Lemma~7.2]{vitanyi2004meaningful}.
-
-We use this to show that sophistications with a restricted model class are not bounded:
-\dogfood*
-\begin{proof}
-In this proof, for convenience we identify the models indexed by $\F$ with the sets that they encode.
-Thus the two-part code consists of the description of a set $A$, and then an index in this set using $\log|A|$ bits.
-For all $i$, let $x_i$ be the smallest binary string that is not $(i/4, i/4)$-stochastic.
-Since every string is $(\alpha,\alpha)$-stochastic for sufficiently large $\alpha$, this is a sequence of infinitely many distinct binary strings of increasing length.
-From \cite[Proposition~I.3 (b)]{gacs2001algorithmic} we know that for all sufficiently large $i$ there are strings of length less than $i$ that are not $(i/4, i/4)$-stochastic.
-Now pick any $i$ that is large enough that (a) $|x_i|<i$ and (b) $C^\K(x_i)/5 < i/4$.
-By definition of $(\alpha,\beta)$-stochasticity, if a string is not $(i/4,i/4)$-stochastic, then for every finite set $A$ containing $x$, either (a) $\log|A|\ge C^\K(x_i)+i/4\ge 5C^\K(x_i)/4$, or (b) $C^\K(A)>i/4>C^\K(x_i)/5$.
-In case (a), any two-part description using $A$ is substantially larger than the Kolmogorov complexity, for large enough $i$, so it does not determine the sophistication.
-Case (b) provides the required lower bound on the complexity of the model.
-
-Now, take $\T$ as model class and suppose towards contradiction that the sophistication is determined by a total function $f=\phi_j,j\in\T$ with $f(d)=x$ and $C^\K(f)\le i/5$.
-Then by Lemma~\ref{lemma:total-to-sets}, for some fixed constant $c$ there is a finite set $A$ with $\log|A|\le|d|\le C^\K(x_i)$, so (a) above cannot be the case, and $C^\K(A)\le C^\K(f)+C^\K(|d|)+c\le i/5+C^\K(|d|)+c$, which contradicts (b) for large enough $i$.
-Therefore such candidates $f$ cannot exist and all candidates must satisfy $C^\K(f)>i/5>C^\K(x_i)/6$.
-\end{proof}
-
-Note that we favored brevity of the proof over the strength of the bounds: the multiplicative constant $1/6$ can be improved with a more detailed treatment.
-
-\depth*
-\begin{proof}
-Let $U(\bar\imath y)$ be some (non-prefix) universal Turing machine, and let $U^A(\bar\imath y)$ be a simulation of that machine which outputs $0$ if the number of steps taken exceeds $A(|\bar\imath y|)$.
-Let $u$ be the index of the function $U^A$ in the standard enumeration.
-
-Let $D$ be a prefix function with $D(\epsilon) = u$.
-We can instantiate Lemma~\ref{lemma:thecoolone} with $D$, $\M' = \{\phi_u\}$ and $X = \{x : d^{\C,c}(x) \leq A(C(x))\}$.
-This tells us that there exists a numbering for which $\s^\T(x) = \s^{\M'}(x) + |\bar0| \leq c$ for all $x \in X$.
-\end{proof}
-
 \end{document}

### 2015-03-23 Facticity/Facticity/paper.tex
First draft CiE abstract

@@
 % --- DELETE BEFORE SUBMISSIONS ---
-\pagestyle{headings}
+% \pagestyle{headings}
 
@@
 \section{Introduction}\enlargethispage{4ex}
-Komogorov complexity gives us a sound definition of the amount of information contained in a binary string.
-It is, however, not what most people would consider complexity.
+Kolmogorov complexity gives us a sound definition of the amount of information contained in a binary string.
+It does not, however, capture what most people would consider complexity.
 For example, a sequence of a million coin flips will almost certainly have maximal Kolmogorov complexity, even though there is nothing complex about flipping a lot of coins---it's just a lot of work.
@@
 However, for most of these definitions, we can prove that they fail one or more of the conditions above.
-Vereshchagin and Vit\'anyi's definition \cite{vitanyi2004meaningful} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we show that their method only assigns substantial sophistication to strings that require an enormous amount of processing to construct.
+Vit\'anyi's definition \cite{vitanyi2004meaningful} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we show that this method only assigns substantial sophistication to strings that require an enormous amount of processing to construct.
 
@@
 The idea of separating structure from noise originated with Kolmogorov's \emph{structure function} \cite{cover1985kolmogorov}.
-In this theory, the data $x$ is described as finite set $A$ that contains $x$, together with the index of $x$ in $A$.
+In this theory, the data $x$ is described as a finite set $A$ that contains $x$, together with the index of $x$ in $A$.
 The complexity of the set determines the sophistication.
@@
 \emph{Facticity} \cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions.
-Adriaans also identifies the nickname problem discussed in Section~\ref{section:indices}.
+This publication also identifies the nickname problem discussed in Section~\ref{section:indices}.
 
@@
 
-\section{Inefficient indices}
+\section{Inefficient indices}\enlargethispage{4ex}
 \label{section:indices}
@@
 In these works, no constraints are placed on the numbering.
-This makes these approaches vulnerable to the so-called \emph{nickname problem}: the number of bits required to specify a model becomes very dependent on the numbering.
-This problem is severe: \emph{any} index sophistication without requirements on the numbering is either bounded, or nowhere near invariant:
+This makes these approaches vulnerable to the \emph{nickname problem} \cite{adriaans2012facticity}: the bits required to specify a model strongly depends on the numbering: \emph{any} index sophistication without requirements on the numbering is either bounded, or nowhere near invariant:
 
@@
 \begin{proof}
-Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.
+Let $s_i$ be the string consisting of $2^{i}$ zeroes followed by a one.
 Define $\phi, \xi$ such that $\psi_j(x) = \phi_i(x)$ for $j = s_{2i}$ and $\xi_j(x) = \phi_i(x)$ for $j = s_{2i+1}$, with all other functions returning $\infty$.
@@
 that $s^\psi(x)\le s^\xi(x)$.
-By construction of $\psi$ and $\xi$ we have $2s_\psi(x)  \leq s_\xi(x)$ from which the result follows.
+By construction, we have $2s_\psi(x)  \leq s_\xi(x)$ from which the lemma follows.
 \end{proof}
-Thus, even though the Kolmogorov complexity changes by only a constant number of bits under a switch to a different numbering, the plain length of the index in a given acceptable numbering is a very poor indicator of model complexity.
-
-In order to obtain a reasonable and robust measure of model complexity, we define the Kolmogorov complexity of functions as in \cite{grunwald2004shannon,vitanyi2004meaningful} by $C^{\M,\psi}(f) = \min\{C^{\M,\psi}(i):\phi_i=f\}$.
-Lemma~\ref{lemma:invariance} in the appendix shows that this definition is invariant: the Kolmogorov complexity is an objective property of a function.
+Thus, the length of the index is a very poor indicator of model complexity.
+In order to obtain a robust measure, define the complexity of functions as in \cite{grunwald2004shannon,vitanyi2004meaningful} by $C^{\M,\psi}(f) = \min\{C^{\M,\psi}(i):\phi_i=f\}$.
+Lemma~\ref{lemma:invariance} in the appendix shows that this definition is invariant: it is an objective property of a function.
 
@@
 
-Second, we can stick to using the length of the index as the measure of sophistication, but impose the additional constraint that an efficient acceptable numbering is used.
-This is the approach taken by Adriaans \cite{adriaans2012facticity}.
-Before switching to the other approach we construct such an efficient index.
-
-We call a numbering $\psi$ \emph{faithful} if for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and $|j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
-This definition is a refinement of \cite[Definition~10]{adriaans2012facticity}.
+Second, we can stick to using the length of the index as the measure of sophistication, but require that numbering is \emph{faithful}: for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and $|j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
+This approach is taken is in \cite{adriaans2012facticity}.
 We prove in the appendix that there actually do exist faithful, acceptable numberings (Lemma~\ref{lemma:faithful-numberings}).
@@
 In the previous section, we saw the first glimpse of the delicate balance between the two code components in sophistication.
-If the description of the model $m$ is less efficient than $C^\K(m)$ by anything more than a constant, the sophistication becomes bounded.
 We will investigate this balance, starting with the sophistication $S^{\K,\psi,c}$.
@@
 This means that for model classes with a universal model, we can choose our numbering to make $S^\K$ bounded, or equal to the complexity:
-\begin{restatable}[underfitting]{lemma}{underfitting}
+\begin{restatable}[Underfitting]{lemma}{underfitting}
 Let $\M$ be a model class containing a universal model $\phi_u$, with the property that $\exists c \forall i \in \M : C^{\{u\}}(x) \leq C^{\{i\}}(x) + c \p$.
@@
 \end{restatable} 
-
-Thus, there are no longer any numberings for which sophistication is bounded.
+\noindent Thus, there are no longer any numberings for which sophistication is bounded.
 However, as we show below, the resulting sophistication is not an intuitive measure of the amount of structure in the data.
@@
 But that does not in itself mean that it should be impossible to come up with a sound measure.
-The intuition throughout the literature, starting with the structure function, appears to be that the crucial property is whether a string is typical for a model.
-If so, another random choice from that mode selects a string with the same structure: a string of the same `type'.
+The intuition throughout the literature, starting with the structure function, appears to be that the crucial property is whether a string is typical for a model, and that this typicality can be tested: another random choice from that model should select a string with the same structure
 
@@
 In this scenario, a sophistication could be called \emph{consistent} if, for data of sufficiently high complexity, it accurately reflects the complexity of that distribution from which it originated.
-Suppose that the data issampled from  universal distribution: let $\phi_u(\bar\imath y)=\phi_i(y)$ and sample from $p_u$ as described above.
+Suppose that the data is sampled from the universal distribution: let $\phi_u(\bar\imath y)=\phi_i(y)$ and sample from $p_u$ as described above.
 Then the initial random bits will determine the prefix encoded index $\bar\imath$ of the function $\phi_i$ that $\phi_u$ will subsequently emulate, and the remaining bits are used as inputs to $\phi_i$.
@@
 Note that part of the theory is fine: there is nothing wrong with evaluating models for the data by comparing their two-part code lengths.
-In fact, the randomness deficiency $\delta(\phi_i,x)=-\log p_i(x)-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data---if $p_i$ truly generated the data, then the probability that the randomness deficiency is more than $k$ bits is less than $2^{-k}$ \cite{bloem2014safe}[Lemma~6].
+In fact, the randomness deficiency $\delta(\phi_i,x)=-\log p_i(x)-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data---if $p_i$ truly generated the data, then the probability that the randomness deficiency is more than $k$ bits is less than $2^{-k}$ \cite[Lemma~6]{bloem2014safe}.
 In the Monet example above, this will allow us to disqualify the model expressing that the data is actually, say, a recording of Jazz music.
@@
 
-This publication was supported by the Dutch national program COMMIT and by  the Netherlands eScience center.
+This publication was supported by the Dutch national program COMMIT and by  the Netherlands eScience center. We thank Tom Sterkenburg for interesting discussions.
 
@@
 
-\subsection{Boosting submodel by manipulating the numbering}
+\subsection{Boosting a submodel by manipulating the numbering}
 

### 2015-05-03 Facticity/Facticity/cie-abstract.tex
First attempt

@@
 \documentclass{style/llncs}
-
-\usepackage{amsmath,amsfonts}
-\usepackage{color}
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-\usepackage[mathscr]{eucal}
-\usepackage{thmtools}
-\usepackage{graphicx}
-\usepackage{caption}
-
-\newcommand{\M}{\mathscr M}
-\newcommand{\C}{\mathscr C}
-\newcommand{\T}{\mathscr T}
-\newcommand{\F}{\mathscr F}
-\renewcommand{\P}{\mathscr P}
-\newcommand{\K}{\mathscr K}
-\newcommand{\X}{\mathscr X}
-\newcommand{\B}{\mathbb B}
-\newcommand{\D}{\Delta}
-\newcommand{\N}{\mathbb N}
-\newcommand{\tn}[1]{\textnormal{#1}}
-\newcommand{\pair}[1]{\left\langle{#1}\right\rangle}
-\newcommand{\concat}{\oplus}
-\newcommand{\symb}[1]{\texttt{#1}}
-\newcommand{\br}[1]{\overline{#1}}
-\newcommand{\s}{S}
-\newcommand{\dom}[1]{\mathop{\tn{dom}(#1)}}
-\newcommand{\range}[1]{\mathop{\tn{range}(#1)}}
-
-\newtheorem{conj}{Conjecture}
-
-\let\doendproof\endproof
-\renewcommand\endproof{~\hfill\qed\doendproof}
-
-\newcommand{\p}{\,\text{.}}
-
-\newcommand{\tuple}[1]{\left\langle{#1}\right\rangle}
-
-\newcommand{\hide}[1]{}
-\newcommand{\old}[1]{}
-
-\newcommand{\sdr}[1]{\textcolor{blue}{\small #1\textsuperscript{[Steven]} }}
-\newcommand{\pb}[1]{\textcolor{OliveGreen}{\small #1 \textsuperscript{[Peter]} }}
-
-\newcommand{\argmin}{\mathop{\arg\min}}
-
-% --- DELETE BEFORE SUBMISSIONS ---
-% \pagestyle{headings}
 
@@
 \begin{abstract}
-Kolmogorov complexity is a sound, unambiguous formalization of the information content of an object.
-However, researchers have long been plagued by its inability to express a more intuitive notion of complexity: in terms of Kolmogorov complexity, a completely random bitstring is the most complex object possible.
-This is not what we mean when we we use the word complexity informally: a television broadcast is complex when we can see structures, faces and landscapes, not when we can only see white noise.
+Kolmogorov complexity formalizes the idea of information content, but clashes with intuitive ideas of complexity: data seems most complex to us when it contains structure, like language or interesting visual information, not when it contains only random noise.
 
-\hspace{0.05\textwidth} The desire to capture this notion of complexity in formal terms has led to many proposals for additional measures.
-The largest family of proposals is built on the following idea:  an object consists of \emph{structural} and \emph{stochastic} information, and the information content of the structural information determines the complexity of the object.
-We will use the name \emph{sophistication} for such proposals.
+\hspace{0.05\textwidth}A popular approach is to assume that an object consists of \emph{structural} and \emph{stochastic} information, and that only the former determines the complexity.
+A dataset $x$ is separated into structural and stochastic parts by representing it as a two-part code: a computable function $\phi$, the \emph{model}, and a string $y$, the \emph{input}, such that $\phi(y) = x$.
+The aim is that for efficient representations, $\phi$ contains all structure, and $y$ contains all stochastic information.
+Many proposals, including the structure function, the minimal sufficient algorithmic statistic and sophistication fit this mold.
+We use \emph{sophistication} as a catch-all term.
 
-\hspace{0.05\textwidth} In essence, sophistication performs model selection using the class of all computable models.
-This means that we are dealing not just with the problem of formally defining complexity, but also with a limiting case of statistics: what happens if we extend our model class to the entirety of conceivable models?
-Do we end up with an ultimate statistics machine, or does the whole framework collapse, and does model selection only work with carefully constrained model classes?
+\hspace{0.05\textwidth} We highlight two issues.
+First, it is crucial to correctly measure the information content of the model.
+All treatments rely on a \emph{numbering}; a canonical description of the computable functions.
+Kolmogorov complexity is not affected by the choice of numbering by more than a constant, but for sophistication the impact can be dramatic when the length of the plain description of $\phi$ is used to measure information content.
 
-\hspace{0.05\textwidth} While sophistication is a simple enough idea, any attempt to formalize it quickly runs into problems.
-We highlight the two major issues, and show how they invalidate each proposal.
-We are forced to conclude that no current proposal truly solves the problem of sophistication, and we offer some informal arguments which suggest that no such formalization may be possible at all.
+\hspace{0.05\textwidth} Second, we discuss the problems of over- and underfitting.
+To give an extreme example of underfitting: we can use a universal Turing machine as a model.
+This representation has all information in the input, yet is maximally efficient, up to a constant.
+We show that for some numberings, such models are always preferred, leading to a bounded sophistication.
 
+\hspace{0.05\textwidth} Underfitting is a known issue which many treatments seek to circumvent by restricting the class of models to total functions.
+This leads to the opposite problem: for some numberings, all information ends up inside the model.
+Only one formalization escapes this problem by allowing non-self-delimiting descriptions.
+While the question whether this form is invariant remains open, we show that numberings exist that cause it to be bounded for all but extremely deep strings.
+Thus, it certainly does not fit the original idea behind sophistication.
+
+\hspace{0.05\textwidth} We draw three conclusions.
+First, sophistication is an interesting problem that leads to diverse questions about statistics, computability and compression.
+Second, we show that all currently published approaches are either incorrect, or do not behave as desired.
+Finally we propose that a sound definition of sophistication, capturing the right intuition may be impossible.
+We ask what we can truly discover about individual objects if our only assumption is that the model should be computable.
 \end{abstract}

### 2015-05-03 Facticity/Facticity/paper.tex
First attempt

@@
 \newcommand{\s}{S}
+\newcommand{\bi}{\bar\imath}
 \newcommand{\dom}[1]{\mathop{\tn{dom}(#1)}}
@@
 Kolmogorov complexity measures the amount of information in data, but does not distinguish structure from noise.
-Kolmogorov's definition of the \emph{structure function} was the first attempt to measure only the structural information in data, by measuring the complexity of an objectively best model under two-part coding.
+Kolmogorov's definition of the \emph{structure function} was the first attempt to measure only the structural information in data, by measuring the complexity of the smallest model allowing for optimal compression of the data.
 Since then, many variations of this idea have been proposed, for which we use \emph{sophistication} as an umbrella term.
@@
 
-\section{Introduction}\enlargethispage{4ex}
+\section{Introduction}
+
 Kolmogorov complexity gives us a sound definition of the amount of information contained in a binary string.
 It does not, however, capture what most people would consider complexity.
-For example, a sequence of a million coin flips will almost certainly have maximal Kolmogorov complexity, even though there is nothing complex about flipping a lot of coins---it's just a lot of work.
-Many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data: its \emph{sophistication}.
-In this paper, we investigate two serious problems with sophistication measures, and show why they are unsatisfactory.
-We believe that the problems are fundamental, and provide arguments why attempts to define sophistication must fail.
+For example, a sequence of a million coin flips will almost certainly have maximal Kolmogorov complexity, even though there is nothing complex about flipping a coin repeatedly.
+Many scholars have defined additional complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information.
+While this concept has been given many names, we use \emph{sophistication} as an umbrella term.
+In this paper, we investigate two serious problems with sophistication measures.
+We conclude with three arguments that suggest the problems are fundamental, explaining our belief that sophistication cannot be defined in a satisfactory manner.
 
 The Kolmogorov complexity $C(x)$ of a binary string $x$, is (roughly) the length of the shortest computer program that will output $x$.
-This length depends on which programming language is used, but, as shown by a central result known as the Invariance Theorem\cite[Section~2.1]{li1993introduction}, the language affects the value only by a constant, independent of $x$.
-This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings, a \emph{property} of the data.
-A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees:
+This length depends on which programming language is used, but, as shown by a central result known as the Invariance Theorem \cite[Section~2.1]{li1993introduction}, the language affects the value only by a constant, independent of $x$.
+This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings, a property \emph{of the data.} A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees:
 
@@
 \end{enumerate}
-There have been many related proposals for such a measure, that are all based on a \emph{two-part code}: rather than encoding $x$ by means of a single computer program, we encode a \emph{model} for $x$ in the first part of the code, that is interpreted as a representation of $x$'s structural properties.
+There have been many related proposals for such a measure, all based on a \emph{two-part code}: rather than encoding $x$ by means of a single computer program, we encode a \emph{model} for $x$ in the first part of the code, that is interpreted as a representation of $x$'s structural properties.
 The model does not fully specify $x$, but when combined with the second part of the code, which specifies the noise, the original string becomes fully determined.
@@
 Underfitting occurs when the selected model is simple, but fails to represent some of the structure present in the data.
-This is also a problem for sophistication because the considered models are so powerful.
+This is also a problem for sophistication because the models under consideration are so powerful.
 In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language.
@@
 
-\paragraph{Related work}
-The idea of separating structure from noise originated with Kolmogorov's \emph{structure function} \cite{cover1985kolmogorov}.
-In this theory, the data $x$ is described as a finite set $A$ that contains $x$, together with the index of $x$ in $A$.
-The complexity of the set determines the sophistication.
-This approach was discussed more extensively, and generalised in \cite{vereshchagin2004kolmogorov,gacs2001algorithmic}.
+\section{Preliminaries}
+One of the main issues in dealing with sophistication in a general sense, is that one must unify many different approaches, all founded on slightly different principles.
+We have devised a general framework which, while it may not represent the idiosyncracies of all methods, allows us to rephrase them in a uniform way that illustrates the properties we discuss.
+We hope the reader will trust that these properties remain also in the original statements of the measures.
 
-Koppel \cite{koppelSoph1988,koppel1991almost} also built on the structure function work to define a measure of the meaningful information in a binary string, coining \emph{sophistication}, the name we adopt in this paper.
-
-In 1996 Gell-Mann and Lloyd introduced \emph{effective complexity} \cite{gellmann1996information}.
-Working from the perspective of statistical mechanics, they arrived (independently, it seems) at a notion that fits the mold of sophistication.
-\emph{Facticity} \cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions.
-This publication also identifies the nickname problem discussed in Section~\ref{section:indices}.
-
-Other authors have pointed out problems with sophistication.
-Antunes et al.
-\cite{antunes2009sophistication} note that the constant cutoff used by Koppel is a source of instability, a result in the spirit of this paper.
-They propose a different definition called \emph{coarse sophistication}.
-This version behaves very differently from regular sophistication, and is closer to computational depth.
-Vereshchagin \cite{vereshchagin2013algorithmic} also points out that the Kolmogorov minimal sufficient statistic is not always well-defined and suggests requiring that the finite set model should described using a \emph{total} function rather than any computable function.
-This yields a more robust sophistication that is nonetheless still susceptible to the issues discussed in this paper.
-\section{Preliminaries}
 Let $\B = \{0,1\}^*$.
@@
 To simplify notation, we will sometimes conflate natural numbers and binary strings, implicitly using the ordering $(0, \epsilon)$, $(1, 0)$, $(2, 1)$, $(3, 00)$, $(4, 01)$, \ldots
-We define the $\M$-Kolmogorov complexity of a binary string $x$, with respect to a model class $\M$ and a numbering $\psi$, as follows $C^{\M,\psi}(x)=\min\{\bar\imath y:\psi_i(y)=x,i\in\M\}$.
-The model classes $\C$ and $\K$ give us the plain and prefix-free Kolmogorov complexity respectively.
 
-\section{Inefficient indices}\enlargethispage{4ex}
+For technical reasons, we deviate slightly from the traditional notation of Kolmogorov complexity: we define the $\M$-Kolmogorov complexity of a binary string $x$, with respect to a model class $\M$ and a numbering $\psi$, as follows $C^{\M,\psi}(x)=\min\{\bar\imath y:\psi_i(y)=x,i\in\M\}$.
+This means that $C^\C(x)$ corresponds to the plain Kolmogorov complexity, traditionally denoted $C(x)$, and $C^\K(x)$ corresponds to the prefix-free Kolmogorov complexity $K$.
+This notation also allows us to represent the smallest possible description of $x$ using a single model $\phi_i$ as $C^{\{i\}}(x)$.
+Note that this includes the description of $i$ as well as its shortest program for $x$.
+
+Finally we not that the choice of numbering, combined with a choice of prefix function, serves the same purpose that is expressed in the choice of a universal Turing machine.
+Requiring an acceptable numbering and a computable prefix function is equivalent to choosing a universal Turing machine: this requirement suffices to make the Kolmogorov complexity invariant.
+As we shall see, however simply choosing a universal Turing machine is not enough for a proper treatment of sophistication: we must choose our numbering carefully.
+Thus, we phrase our requirement for an invariant sophistication with respect to the \emph{numbering} chosen.
+There should be strict guarantees on how much the value of $S(x)$ can changes if the numbering is varied.
+
+\section{Inefficient indices}
 \label{section:indices}
-In some works, the definition of sophistication is obtained by opening up the definition of Kolmogorov complexity, and measuring the length of the index in that two-part code.
-We call such a sophistication \emph{index sophistication}.
-Examples are Koppel and Atlan's works \cite{koppelSoph1988,koppel1991almost},  and several of its variants \cite{antunes2009sophistication,antunes2013sophistication}.
 
-In these works, no constraints are placed on the numbering.
-This makes these approaches vulnerable to the \emph{nickname problem} \cite{adriaans2012facticity}: the bits required to specify a model strongly depends on the numbering: \emph{any} index sophistication without requirements on the numbering is either bounded, or nowhere near invariant:
+The simplest approach to sophistication would simply be to `open up' the Kolmogorov complexity and to see which program on the universal Turing machine achieves the smallest description length: the program that \emph{witnesses} the Kolmogorov complexity.
+Since any UTM is defined in terms of a two-part coding, such programs necessarily consist of a model and an input.
+While this approach is not literally used anywhere in the literature, it is an illustrative point to start our investigation.
+As is common, we will allow all witnesses within a constant slack term $c$ from the Kolmogorov complexity, and choose the one with the smallest model.
+
+\begin{definition}[Index sophistication]
+\[
+s_{\text{index}}^{\M_0,\M_1,\psi,c}(X) = \min\left\{ |i| : \i in \M_0, C^{\{i\}, \psi}(x) \leq C^{\M_1, \psi}(x)+c   \right\}
+\]
+\end{definition}
+
+Koppel and Atlan's treatment \cite{koppelSoph1988,koppel1991almost}, where the name \emph{sophistication} originates, follows this basic logic, although it contains many idiosyncracies like the use ot monotonic models, and an inherent extension to infinite strings.
+As the subsequent history of sophistication has discarded these, we will not discuss them here.
+
+In \cite{antunes2009sophistication} Koppel's principle is first recast to finite strings.
+The definition used is similar to $s_\text{index}^{\T, \C,\psi,c}$, with one exception: the total complexity of a witness $(i,y)$ is measured as $|i|+|y|$ without the cost of delimiting the two.
+This differences are not relevant to the point we make in this section.
+The restriction to total classes is a common technique to avoid underfitting, which we will discuss in the next section.
 
@@
 
-There are two ways to use this notion of model complexity for more robust attempts to define sophistication. Confusingly, both approaches are used in the literature.
+Note that such perverse numberings are not an issue for Kolmogorov complexity.
+If our universal Turing machine uses an inefficient numbering, we can switch to a moe efficient one at only a constant cost.
+This gives us an insight into what happens inside Kolmogorov complexity.
+We may define the program $\bi y$ using a model $i$ and an input $y$, but if there's any inefficiency in the description of $i$, the shortest program will be achieved by using $i$ to define a more efficient model encoding, and encoding both the model and its input in $y$.
+For the Kolmogorov complexity, such considerations are academic, since we only look at the total length of the program.
+For sophistication, however, they are crucial.
+
+There are two ways to use the model complexity $C^\M(\phi_i)$ for more robust attempts to define sophistication.
+Confusingly, both are used in the literature.
 First, we can measure the complexity of the \emph{model} using $C^{\M,\psi}(\phi_i)$, which is then the first part of a two-part code describing the data.
@@
 
-Second, we can stick to using the length of the index as the measure of sophistication, but require that numbering is \emph{faithful}: for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and $|j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
+Second, we can stick to using the length of the index as the measure of sophistication, but restrict the allowed numberings.
 This approach is taken is in \cite{adriaans2012facticity}.
-We prove in the appendix that there actually do exist faithful, acceptable numberings (Lemma~\ref{lemma:faithful-numberings}).
+Adriaans  defines \emph{facticity} as $s_\text{index}^{\C,\C,\psi,0}$, but only allows numberings that are \emph{faithful}: for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and $|j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
+
+We prove in the appendix that---contrary to Adriaans' suggestion---there actually do exist faithful, acceptable numberings (Lemma~\ref{lemma:faithful-numberings}).
 However even choosing a faithful index is not quite enough.
@@
 If we define a second prefix encoding $\tilde{\imath}$, with $|\bar\imath|-|\tilde\imath|$ unbounded, we can define a second representation $\bar u \tilde \imath y$, at a constant overhead $|\br{u}|$, and gain more than $|\br{u}|$ for sufficiently complex strings.
-As shown in Lemma~\ref{lemma:prefix-inefficiency} in the appendix, this results in a bounded sophistication.
-This is in fact an issue with \cite{adriaans2012facticity}. Moreover, we conjecture that \emph{no} computable prefix function $\bar\cdot$ is sufficiently efficient to avoid this.
+As shown in Lemma~\ref{lemma:prefix-inefficiency} in the appendix, this results in a bounded sophistication (thus the facticity is bounded).
 
@@
 In the previous section, we saw the first glimpse of the delicate balance between the two code components in sophistication.
-We will investigate this balance, starting with the sophistication $S^{\K,\psi,c}$.
-In this setting, we know that, on the one hand, there always exist candidate models---those models $i$ with $i \in \K$ and $C^{\{i\}}(x) \leq C^\K(x) + c$---for which al information except a constant amount is stored in the data, and on the other hand, there are also candidate representations for which all information is stored in the model.
+We will investigate this balance, starting with the variant $S^{\K,\psi,c}$.
+For reasons which will become clear, this variant is not used anywhere in the literature, but it helps to illustrate the issues we wich to discuss.
+
+In this setting, we know that, on the one hand, there always exist witnesses for which all information except a constant amount is stored in the input, and on the other hand, there are also candidate representations for which all information is stored in the model.
 This tells us that $S^\K$ is \emph{balanced}: there is no inefficiency in the representation forcing information in or out of the model.
-The flip side of having a balanced sophistication is that it becomes easy to show invariance, by tweaking the numbering in such a way that models in a specific subset $\M'\subset\M$ of the model class becomes an arbitrary constant number of bits cheaper to represent than all other models, so that a model in $\M'$ always ends up determining the sophistication.
+The flip side of having a balanced sophistication is that it becomes easy to show a lack of invariance, by tweaking the numbering in such a way that models in a specific subset $\M'\subset\M$ of the model class becomes an arbitrary constant number of bits cheaper to represent than all other models, so that a model in $\M'$ always ends up determining the sophistication.
 
 This means that for model classes with a universal model, we can choose our numbering to make $S^\K$ bounded, or equal to the complexity:
+
 \begin{restatable}[Underfitting]{lemma}{underfitting}
@@
 \end{restatable}
+
+For this reason, many treatments restrict the model class.
+What is perhaps less well known is that the same property holds in the other direction: there are numberings for which the singleton models always determine the sophistication.
+
 \begin{restatable}[Overfitting]{lemma}{overfitting}
@@
 \end{restatable}
+
+The proof of both these lemmas relies on a simple principle: for certain subclasses $\M' \subset \M$, there exist numberings which, when used, have the effect of penalizing the codelength $C^\K(\phi_i)$ for any model outside $\M'$ by an arbitrary constant amount.
+We can use this penatly to effectively `push' these models outside of the range of considered models, ensuring that, under this numbering, a model in $\M'$ always determines the sophistication.
+
+For the lemmas above this means setting $\M' = {u}$, with $\phi_u$ a universal model, or letting $\M'$ be the set of all singletons.
+The requirements for $\M'$ are strict, and slightly complex.
+The following lemma illustrates the principle.
+
+\begin{restatable}{lemma}{coolone}
+\label{lemma:thecoolone}
+  Let $\M$ be any model class, let $\X$ be any set of binary strings and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\phi$, i.e.
+if $m=\phi_{D(p)}$ then $p$ is a $D$-description of $m$.
+Let $\M'=\{\phi_i:i\in\tn{range}(D)\}$.
+Further assume there is a constant $c$ such that:\\
+\-\hspace{1cm}(1) $\forall_{m\in\M'}:\min\{|p|:\phi_{D(p)}=m\}\le C^{\K,\phi}(m)+c$\\
+\-\hspace{1cm}(2) $\forall_{x\in\X}:C^{\M',\phi}(x)-C^{\M,\phi}(x)\le c$.\\
+Then there is an enumeration $\psi$ of the partial recursive functions such that $S^{\M,\psi}(x) = |\br{0}|+S^{\M',\phi}(x)$ for all $x\in\X$.
+\end{restatable}
+
+\begin{proof}
+We define the numbering $\psi$ as follows:
+\[\begin{cases}
+\psi_0(p) = 1^r 0 D(p) \\
+\psi_{1^r0i}(p) = \phi_i(p) \\
+\psi_j(\cdot) = \infty &\text{if $j$ contains no zeroes.}
+\end{cases}\]
+We will show that under the $\psi$-enumeration, the best representation for $x$ using a model $m\in\M'$ is always better than the best representation using some $m\notin\M'$.
+
+First suppose $m\in\M'$. Then
+\begin{align*}
+C^{\K,\psi}(m) &=\min\{|\bar\jmath q|:\psi_{\psi_j(q)}=m\}
+\leq\min\{|\bar0 q|:\psi_{\psi_0(q)}=m\} \\ 
+&=\min\{|\bar0 q|:\psi_{1^r0 D(q)}=m\}  
+=\min\{|q|:\phi_{D(q)}=m\} + |\bar 0|\\
+&\leq C^{\K,\phi}(m)+c+|\br{0}|,
+\end{align*}
+where the last inequality uses assumption (1).
+
+Now assume that the best model $m$ for $x$ is not in $\M'$. 
+Let $i$ be the index of $m$ with the shortest description, i.e. it achieves the minimum in $C^{\K, \psi}(m)=\min\left\{C^{\K,\psi}(i):\psi_i=m\right\}$.
+There are two possibilities. Either $i = 0$, in which case we have $C^{\K,\psi}(m)\ge r$ because $\phi_0$ cannot output zero and all other $\psi$-programs are at least $r$ bits long.
+
+Otherwise, we can bound
+\begin{align*}
+C^{\K,\psi}(m)&=C^{\K,\psi}(i)=C^{\K,\psi}(1^r0j)\\
+&\ge C^{\K,\psi}(j)-c' \ge C^{\K,\phi}(j)+r+1-c'\\
+&=C^{\K,\phi}(m)+r+1-c'.
+\end{align*}
+Now choose $r=c''+\max\left\{C^{\K,\psi}(\phi_0),c'-1\right\}$. Then substitution yields, for both cases, 
+$C^{\K,\phi}(m)\ge c''+C^{\K,\psi}(m)$.
+
+Combining the inequalities above, for any model $g\not\in\M'$, there is a model $f\in\M$ such that
+\[\begin{split}
+C^{\K,\psi}(g)+\min \{|y| : g(y) = x\} &\ge C^{\K,^\phi}(g)+\min\{|y| : g(y) = x\}+c''\\
+&\ge C^{\K,\psi}(f)+\min \{|y| : f(y) = x\} -2c+c'' -|\bar0|.
+\end{split}\]
+By choosing $c''$ sufficiently large, we can ensure that the best representation is in $\M'$ for all $x\in\X$, which completes the proof.
+\end{proof}
+
+Lemmas~\ref{lemma:underfitting} and \ref{lemma:overfitting} follow as corollaries. For Lemma~\ref{lemma:underfitting}:
+\begin{proof}
+Let $D$ be a prefix function as in Lemma~\ref{lemma:thecoolone} such that it returns the index of $u$ for the argument $\epsilon$ and $\infty$ for any other argument.
+That is, $\M' = \{u\}$.
+This construction satisfies the conditions 1 and 2 from Lemma~\ref{lemma:thecoolone}; invoking it we find that there exists an acceptable numbering $\psi$ for which $\s^{\M,\psi}(x) = \s^{\M', phi}(x) + c$.
+Since $\M'$ contains only a single model, $\s_{\M',\phi}(x)$ is constant.
+\end{proof}
+
+And for Lemma~\ref{lemma:overfitting}
+
+\begin{proof}
+Let $\phi_i$ be a singleton for $x$ as described.
+Note that since $\phi_i$ is a prefix function, if $\phi_i$ is defined for input $\epsilon$ then it cannot be defined for any other input.
+Pick $i,x$ with $\phi_i(\epsilon)=x$.
+Note that $x$ can be computed from $i$ and a fixed program, so there is a $c$ such that $C^{\K, \phi}(x)\le C^{\K, \phi}(\phi_i)+c$.
+Vice versa, given any $x$ we can construct an index of $\phi_i$, since $\phi$ is an acceptable numbering.
+Therefore $|C^{\K,\phi}(\phi_i)-C^{\K, \phi}(x)|\le c$.
+
+We now define a computable function $D$ by $D(\bar\imath y)=j$ where $\psi_j(\epsilon) = \psi_i(y)$.
+We will show that the two conditions of Lemma~\ref{lemma:thecoolone} hold for the prefix function $D$.
+
+(1) Let $f$ be any function in the range of $D$, and $x$ its output.
+Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\phi_i(q)=x\}=C^{\K}(x) \le C^{\K}(f)+c$.
+(2) On the one hand $C^{\M',\psi}(x)\le C^\K(f)+|\epsilon|\le C^\K(x)+c$.
+On the other hand, $C^{\M,\psi}(x)$ is an effective description of $x$, so $C^{\K}(x)$ is at most a constant larger.
+Together, these inequalities establish the second condition.
+
+Now, by Lemma~\ref{lemma:thecoolone} there is an numbering $\psi_1,\psi_2,\ldots$ such that $\s^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
+We observed that $|C^{\K}(f)-C^{\K}(x)|\le c$ for any $f\in\M'$, so $\s^{\M',\psi}(x)\ge C^{\K}(x)-c$.
+This proves the lemma.
+\end{proof}
+
 \noindent Thus, in this balanced sophistication, there is no invariance: all information can be seen as structure, or as noise, depending on the numbering.
@@
 
-The existing proposals that can overfit are the minimal sufficient statistic in the structure function \cite{cover1985kolmogorov}, the sophistication as defined in \cite{mota2013sophistication}, the minimal sufficient algorithmic statistic \cite{gacs2001algorithmic} and Vereshchagin's strongly sufficient statistic \cite{vereshchagin2013algorithmic}.
+We will now review the treatments in the literature that show overfitting.
+The first is the structure function, proposed by Kolmogorov himself.
+Most likely the first attempt at separating structure from noise in an object manner.
+Kolmogorov defined the function
+\[
+h_x(\alpha) = \min \left \{\log |S| : S \ni x, C^\K(s) \leq \alpha \right \} 
+\]
+and suggested that the smallest set for which $K(S) + \log|S| \leq C^\K(S) + c$ for some pre-chosen constant$c$ can be seen as capturing all the structural information in $x$.
+\footnote{\pb{Discuss how much of this Kolmogorov believed himself and note that this was unpublished work.}} This is equivalent to the sophistication $\s^{\F,\K,\psi,c}$.
+Lemma~\ref{lemma:underfitting} shows us that given $c$, there are numberings for which this sophistication is always equal, up to a constant, to $C^\K(x)$.
+Thus, while the general shape of the structure function might tell an interesting story, its sophistication function does not satisfy our requirements.
+
+$\s^{\F,\K,\psi,c}$ is also used in \cite{antunes2013sophistication}. \pb{This version doesn't work at all\ldots}
+
+Also the regular and naive versions from \cite{mota2013sophistication} \ldots
+
+the structure function \cite{cover1985kolmogorov}
+
+the sophistication as defined in \cite{mota2013sophistication} (and the naive soph?)
+
+the minimal sufficient algorithmic statistic \cite{gacs2001algorithmic}
+
+Vereshchagin's strongly sufficient statistic \cite{vereshchagin2013algorithmic}. \pb{fix citation}
+
 It is widely recognized that singleton models are always candidates in these settings, but not, it seems, that they may be the only candidates for any string, as Lemma~\ref{lemma:overfitting} shows.
@@
 
-\subsection{Boosting a submodel by manipulating the numbering}
-
-\begin{restatable}{lemma}{coolone}
-\label{lemma:thecoolone}
-  Let $\M$ be any model class, let $\X$ be any set of binary strings and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\phi$, i.e.
-if $m=\phi_{D(p)}$ then $p$ is a $D$-description of $m$.
-Let $\M'=\{\phi_i:i\in\tn{range}(D)\}$.
-Further assume there is a constant $c$ such that:\\
-\-\hspace{1cm}(1) $\forall_{m\in\M'}:\min\{|p|:\phi_{D(p)}=m\}\le C^{\K,\phi}(m)+c$\\
-\-\hspace{1cm}(2) $\forall_{x\in\X}:C^{\M',\phi}(x)-C^{\M,\phi}(x)\le c$.\\
-Then there is an enumeration $\psi$ of the partial recursive functions such that $S^{\M,\psi}(x) = |\br{0}|+S^{\M',\phi}(x)$ for all $x\in\X$.
-\end{restatable}
-
-\begin{proof}
-We define the numbering $\psi$ as follows:
-\[\begin{cases}
-\psi_0(p) = 1^r 0 D(p) \\
-\psi_{1^r0i}(p) = \phi_i(p) \\
-\psi_j(\cdot) = \infty &\text{if $j$ contains no zeroes.}
-\end{cases}\]
-We will show that under the $\psi$-enumeration, the best representation for $x$ using a model $m\in\M'$ is always better than the best representation using some $m\notin\M'$.
-
-First suppose $m\in\M'$. Then
-\begin{align*}
-C^{\K,\psi}(m) &=\min\{|\bar\jmath q|:\psi_{\psi_j(q)}=m\}
-\leq\min\{|\bar0 q|:\psi_{\psi_0(q)}=m\} \\ 
-&=\min\{|\bar0 q|:\psi_{1^r0 D(q)}=m\}  
-=\min\{|q|:\phi_{D(q)}=m\} + |\bar 0|\\
-&\leq C^{\K,\phi}(m)+c+|\br{0}|,
-\end{align*}
-where the last inequality uses assumption (1).
-
-Now assume that the best model $m$ for $x$ is not in $\M'$. 
-Let $i$ be the index of $m$ with the shortest description, i.e. it achieves the minimum in $C^{\K, \psi}(m)=\min\left\{C^{\K,\psi}(i):\psi_i=m\right\}$.
-There are two possibilities. Either $i = 0$, in which case we have $C^{\K,\psi}(m)\ge r$ because $\phi_0$ cannot output zero and all other $\psi$-programs are at least $r$ bits long.
-
-Otherwise, we can bound
-\begin{align*}
-C^{\K,\psi}(m)&=C^{\K,\psi}(i)=C^{\K,\psi}(1^r0j)\\
-&\ge C^{\K,\psi}(j)-c' \ge C^{\K,\phi}(j)+r+1-c'\\
-&=C^{\K,\phi}(m)+r+1-c'.
-\end{align*}
-Now choose $r=c''+\max\left\{C^{\K,\psi}(\phi_0),c'-1\right\}$. Then substitution yields, for both cases, 
-$C^{\K,\phi}(m)\ge c''+C^{\K,\psi}(m)$.
-
-Combining the inequalities above, for any model $g\not\in\M'$, there is a model $f\in\M$ such that
-\[\begin{split}
-C^{\K,\psi}(g)+\min \{|y| : g(y) = x\} &\ge C^{\K,^\phi}(g)+\min\{|y| : g(y) = x\}+c''\\
-&\ge C^{\K,\psi}(f)+\min \{|y| : f(y) = x\} -2c+c'' -|\bar0|.
-\end{split}\]
-By choosing $c''$ sufficiently large, we can ensure that the best representation is in $\M'$ for all $x\in\X$, which completes the proof.
-\end{proof}
-
-\subsection{Overfitting}
-
-\overfitting*
-
-\begin{proof}
-Let $\phi_i$ be a singleton for $x$ as described.
-Note that since $\phi_i$ is a prefix function, if $\phi_i$ is defined for input $\epsilon$ then it cannot be defined for any other input.
-Pick $i,x$ with $\phi_i(\epsilon)=x$.
-Note that $x$ can be computed from $i$ and a fixed program, so there is a $c$ such that $C^{\K, \phi}(x)\le C^{\K, \phi}(\phi_i)+c$.
-Vice versa, given any $x$ we can construct an index of $\phi_i$, since $\phi$ is an acceptable numbering.
-Therefore $|C^{\K,\phi}(\phi_i)-C^{\K, \phi}(x)|\le c$.
-
-We now define a computable function $D$ by $D(\bar\imath y)=j$ where $\psi_j(\epsilon) = \psi_i(y)$.
-We will show that the two conditions of Lemma~\ref{lemma:thecoolone} hold for the prefix function $D$.
-
-(1) Let $f$ be any function in the range of $D$, and $x$ its output.
-Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\phi_i(q)=x\}=C^{\K}(x) \le C^{\K}(f)+c$.
-(2) On the one hand $C^{\M',\psi}(x)\le C^\K(f)+|\epsilon|\le C^\K(x)+c$.
-On the other hand, $C^{\M,\psi}(x)$ is an effective description of $x$, so $C^{\K}(x)$ is at most a constant larger.
-Together, these inequalities establish the second condition.
-
-Now, by Lemma~\ref{lemma:thecoolone} there is an numbering $\psi_1,\psi_2,\ldots$ such that $\s^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
-We observed that $|C^{\K}(f)-C^{\K}(x)|\le c$ for any $f\in\M'$, so $\s^{\M',\psi}(x)\ge C^{\K}(x)-c$.
-This proves the lemma.
-\end{proof}
-\subsection{Underfitting}
-
-\underfitting*
-
-\begin{proof}
-Let $D$ be a prefix function as in Lemma~\ref{lemma:thecoolone} such that it returns the index of $u$ for the argument $\epsilon$ and $\infty$ for any other argument.
-That is, $\M' = \{u\}$.
-This construction satisfies the conditions 1 and 2 from Lemma~\ref{lemma:thecoolone}; invoking it we find that there exists an acceptable numbering $\psi$ for which $\s^{\M,\psi}(x) = \s^{\M', phi}(x) + c$.
-Since $\M'$ contains only a single model, $\s_{\M',\phi}(x)$ is constant.
-\end{proof}
-
 \subsection{The problem of depth}

### 2015-05-08 Facticity/Facticity/paper.tex
changes...

@@
 
-% --- DELETE BEFORE SUBMISSIONS ---
-% \pagestyle{headings}
-
 \title{Two Problems for Sophistication}
@@
 Since then, many variations of this idea have been proposed, for which we use \emph{sophistication} as an umbrella term.
-We describe two fundamental problems with all existing proposals, showing each  to be \emph{unsound}.
+We describe two fundamental problems with existing proposals, showing many of them to be unsound.
 Consequently, we put forward the view that the problem is fundamental: it may be impossible to objectively quantify the sophistication.
@@
 The Kolmogorov complexity $C(x)$ of a binary string $x$, is (roughly) the length of the shortest computer program that will output $x$.
-This length depends on which programming language is used, but, as shown by a central result known as the Invariance Theorem \cite[Section~2.1]{li1993introduction}, the language affects the value only by a constant, independent of $x$.
+This length depends on which programming language is used, but, as shown by a central result known as the invariance theorem \cite[Section~2.1]{li1993introduction}, the language affects the value only by a constant, independent of $x$.
 This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings, a property \emph{of the data.} A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees:
@@
 \end{enumerate}
-There have been many related proposals for such a measure, all based on a \emph{two-part code}: rather than encoding $x$ by means of a single computer program, we encode a \emph{model} for $x$ in the first part of the code, that is interpreted as a representation of $x$'s structural properties.
+There have been many related proposals for such a measure, all based on a \emph{two-part code}: rather than encoding $x$ by means of a single computer program, we encode a \emph{model} for $x$ in the first part of the code, which is interpreted as a representation of $x$'s structural properties.
 The model does not fully specify $x$, but when combined with the second part of the code, which specifies the noise, the original string becomes fully determined.
@@
 Let $\B = \{0,1\}^*$.
-We deal with partial computable functions $\phi: \B \to \B$, which we will also call \emph{models}.
-A function is called \emph{prefix} if its domain, $\dom{f} = \{y : f(y) \neq \infty\}$ is a prefix free set, i.e.
+We deal with partial computable functions $f: \B \times \B \to \B$, which we will also call \emph{models}.
+A function is called \emph{prefix} if its domain, $\dom{f} = \{y : f(y; \cdot) \neq \infty\}$ is a prefix free set, i.e.
 no string in $\dom{f}$ is a prefix of another.
 A function $f$ is \emph{total} if $\dom{f} = \B$.
+In most cases, we do not use the second argument, and we adopt the convention that $f(x) = f(x, \epsilon)$.
 
@@
 
-A \emph{model class} is a set of model indices.
-We distinguish four important model classes: (1) The indices of the partial computable functions $\C=\N$.
-(2) The total functions $\T=\{i:\tn{$\phi_i$ is total}\}$.
-Note that $\T$ is not computably enumerable.
-(3) $\K$ is an enumerable set such that $\{\phi_i:i\in\K\}$ is the set of all partial computable prefix functions.
-(4) The finite sets: $\F$ is an enumerable set such that $\{\phi_i:i\in\F\}$ contains a uniform code for every finite set.
+A \emph{model class} is a set of model indices. We distinguish four important model classes:
+\begin{itemize} 
+  \item The indices of the partial computable functions $\C=\N$.
+  \item The total functions $\T=\{i:\tn{$\phi_i$ is total}\}$. Note that $\T$ is not computably enumerable.
+  \item $\K$ is an enumerable set such that $\{\phi_i:i\in\K\}$ is the set of all partial computable prefix functions.
+  \item The finite sets: $\F$ is an enumerable set such that $\{\phi_i:i\in\F\}$ contains a uniform code for every finite set.
 A uniform code for a finite set $S$ is a total computable surjective function $f:\{0,1\}^{\lceil\log|S|\rceil}\to S$.
+\end{itemize}
 
@@
 
-For technical reasons, we deviate slightly from the traditional notation of Kolmogorov complexity: we define the $\M$-Kolmogorov complexity of a binary string $x$, with respect to a model class $\M$ and a numbering $\psi$, as follows $C^{\M,\psi}(x)=\min\{\bar\imath y:\psi_i(y)=x,i\in\M\}$.
+For technical reasons, we deviate slightly from the traditional notation of Kolmogorov complexity: we define the $\M$-Kolmogorov complexity of a binary string $x$, with respect to a model class $\M$ and a numbering $\psi$, as follows $C^{\M,\psi}(x\mid z)=\min\{\bar\imath y:\psi_i(y; z)=x,i\in\M\}$, with $C^\M(x) = C^\M(x\mid \epsilon)$.
 This means that $C^\C(x)$ corresponds to the plain Kolmogorov complexity, traditionally denoted $C(x)$, and $C^\K(x)$ corresponds to the prefix-free Kolmogorov complexity $K$.
@@
 
-Finally we not that the choice of numbering, combined with a choice of prefix function, serves the same purpose that is expressed in the choice of a universal Turing machine.
+Finally we note that the choice of numbering, combined with a choice of prefix function, serves the same purpose that is normally expressed in the choice of a Turing machine.
 Requiring an acceptable numbering and a computable prefix function is equivalent to choosing a universal Turing machine: this requirement suffices to make the Kolmogorov complexity invariant.
-As we shall see, however simply choosing a universal Turing machine is not enough for a proper treatment of sophistication: we must choose our numbering carefully.
+As we shall see, however, this is not enough for a proper treatment of sophistication: we must choose our numbering with care.
 Thus, we phrase our requirement for an invariant sophistication with respect to the \emph{numbering} chosen.
-There should be strict guarantees on how much the value of $S(x)$ can changes if the numbering is varied.
+There should be strict guarantees on how much the value of $S(x)$ can change if the numbering is varied.
 
@@
 Since any UTM is defined in terms of a two-part coding, such programs necessarily consist of a model and an input.
-While this approach is not literally used anywhere in the literature, it is an illustrative point to start our investigation.
 As is common, we will allow all witnesses within a constant slack term $c$ from the Kolmogorov complexity, and choose the one with the smallest model.
@@
 \[
-s_{\text{index}}^{\M_0,\M_1,\psi,c}(X) = \min\left\{ |i| : \i in \M_0, C^{\{i\}, \psi}(x) \leq C^{\M_1, \psi}(x)+c   \right\}
+\s_{\text{index}}^{\M_1,\M_2,\psi,c}(X) = \min\left\{ |i| : \i \in \M_1, C^{\{i\}, \psi}(x) \leq C^{\M_2, \psi}(x)+c   \right\}
 \]
+When $\M_1 = \M_2$, we will abbreviate with $\s^{\M_1,\psi,c}$
 \end{definition}
 
-Koppel and Atlan's treatment \cite{koppelSoph1988,koppel1991almost}, where the name \emph{sophistication} originates, follows this basic logic, although it contains many idiosyncracies like the use ot monotonic models, and an inherent extension to infinite strings.
+Koppel and Atlan's treatment \cite{koppelSoph1988,koppel1991almost}, where the name \emph{sophistication} originates, follows this basic logic, although it contains many idiosyncracies like the use of monotonic models, and an inherent extension to infinite strings.
 As the subsequent history of sophistication has discarded these, we will not discuss them here.
 
-In \cite{antunes2009sophistication} Koppel's principle is first recast to finite strings.
+In \cite{antunes2009sophistication,antunes2013sophistication} Koppel's principle is recast to finite strings.
 The definition used is similar to $s_\text{index}^{\T, \C,\psi,c}$, with one exception: the total complexity of a witness $(i,y)$ is measured as $|i|+|y|$ without the cost of delimiting the two.
-This differences are not relevant to the point we make in this section.
+This difference is not relevant to the point we make in this section.
 The restriction to total classes is a common technique to avoid underfitting, which we will discuss in the next section.
@@
 
-Note that such perverse numberings are not an issue for Kolmogorov complexity.
-If our universal Turing machine uses an inefficient numbering, we can switch to a moe efficient one at only a constant cost.
+Note that such perversely inefficient numberings are not an issue for Kolmogorov complexity.
+If our universal Turing machine uses an inefficient numbering, we can switch to a more efficient one at only a constant cost.
 This gives us an insight into what happens inside Kolmogorov complexity.
@@
 This approach is taken is in \cite{adriaans2012facticity}.
-Adriaans  defines \emph{facticity} as $s_\text{index}^{\C,\C,\psi,0}$, but only allows numberings that are \emph{faithful}: for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and $|j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
+Adriaans  defines \emph{facticity} as $\s_\text{index}^{\C,\psi,0}$, but only allows numberings that are \emph{faithful}: for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and $|j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
 
@@
 \[
-S^{\M,\psi,c}(x)=\min\left\{C^\K(\phi_i):i\in\M,C^{\{i\},\psi}(x)\le C^{\M,\psi}(x)+c\right\}.
+\s^{\M_1,\M_2,\psi,c}(x)=\min\left\{C^\K(\phi_i):i\in\M_1,C^{\{i\},\psi}(x)\le C^{\M_2,\psi}(x)+c\right\}.
 \]
@@
 In the previous section, we saw the first glimpse of the delicate balance between the two code components in sophistication.
-We will investigate this balance, starting with the variant $S^{\K,\psi,c}$.
-For reasons which will become clear, this variant is not used anywhere in the literature, but it helps to illustrate the issues we wich to discuss.
+We will investigate this balance, starting with the variant $\s^{\K,\psi,c}$.
+For reasons which will become clear, this variant is not used anywhere in the literature, but it helps to illustrate the issues we wish to discuss.
 
@@
 Let $\M$ be a model class containing a universal model $\phi_u$, with the property that $\exists c \forall i \in \M : C^{\{u\}}(x) \leq C^{\{i\}}(x) + c \p$.
-Then, there exist numberings such that $\s^\M$ is bounded.
+Then, there exist numberings such that $\s^\M$ is bounded.\label{lemma:underfitting}
 \end{restatable}
@@
 The proof of both these lemmas relies on a simple principle: for certain subclasses $\M' \subset \M$, there exist numberings which, when used, have the effect of penalizing the codelength $C^\K(\phi_i)$ for any model outside $\M'$ by an arbitrary constant amount.
-We can use this penatly to effectively `push' these models outside of the range of considered models, ensuring that, under this numbering, a model in $\M'$ always determines the sophistication.
+We can use this penalty to effectively `push' these models outside of the range of considered models, ensuring that, under this numbering, a model in $\M'$ always determines the sophistication.
 
-For the lemmas above this means setting $\M' = {u}$, with $\phi_u$ a universal model, or letting $\M'$ be the set of all singletons.
+For the lemmas above, this means setting $\M' = {u}$, with $\phi_u$ a universal model, or letting $\M'$ be the set of all singletons.
 The requirements for $\M'$ are strict, and slightly complex.
-The following lemma illustrates the principle.
+The following lemma gives a set of sufficient conditions.
 
@@
 
-Now, by Lemma~\ref{lemma:thecoolone} there is an numbering $\psi_1,\psi_2,\ldots$ such that $\s^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
+Now, by Lemma~\ref{lemma:thecoolone} there is a numbering $\psi_1,\psi_2,\ldots$ such that $\s^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
 We observed that $|C^{\K}(f)-C^{\K}(x)|\le c$ for any $f\in\M'$, so $\s^{\M',\psi}(x)\ge C^{\K}(x)-c$.
@@
 \noindent Thus, in this balanced sophistication, there is no invariance: all information can be seen as structure, or as noise, depending on the numbering.
-To avoid these issues, existing proposals upset the balance to exclude or penalize at least the singleton models and the universal models.
+To avoid these issues, existing proposals upset the balance to exclude or penalize the universal models, and possibly the singleton models.
 
@@
 We will now review the treatments in the literature that show overfitting.
-The first is the structure function, proposed by Kolmogorov himself.
-Most likely the first attempt at separating structure from noise in an object manner.
+The first is the structure function, proposed by Kolmogorov: most likely the first attempt at separating structure from noise in an objective manner.
 Kolmogorov defined the function
@@
 \]
-and suggested that the smallest set for which $K(S) + \log|S| \leq C^\K(S) + c$ for some pre-chosen constant$c$ can be seen as capturing all the structural information in $x$.
-\footnote{\pb{Discuss how much of this Kolmogorov believed himself and note that this was unpublished work.}} This is equivalent to the sophistication $\s^{\F,\K,\psi,c}$.
+and suggested that the smallest set for which $K(S) + \log|S| \leq C^\K(S) + c$ for some pre-chosen constant $c$ can be seen as capturing all the structural information in $x$ \cite{cover1985kolmogorov}.
+This is equivalent to the sophistication $\s^{\F,\K,\psi,c}$.
 Lemma~\ref{lemma:underfitting} shows us that given $c$, there are numberings for which this sophistication is always equal, up to a constant, to $C^\K(x)$.
-Thus, while the general shape of the structure function might tell an interesting story, its sophistication function does not satisfy our requirements.
+Thus, while the general shape of the structure function might tell an interesting story, its sophistication function is either not invariant, or vacuous.
 
-$\s^{\F,\K,\psi,c}$ is also used in \cite{antunes2013sophistication}. \pb{This version doesn't work at all\ldots}
-
-Also the regular and naive versions from \cite{mota2013sophistication} \ldots
-
-the structure function \cite{cover1985kolmogorov}
-
-the sophistication as defined in \cite{mota2013sophistication} (and the naive soph?)
-
-the minimal sufficient algorithmic statistic \cite{gacs2001algorithmic}
-
-Vereshchagin's strongly sufficient statistic \cite{vereshchagin2013algorithmic}. \pb{fix citation}
+\pb{Discuss the minimal sufficient algorithmic statistic \cite{gacs2001algorithmic}}
 
@@
 
-It may be argued that the slack parameter $c$ in the sophistication, that determines the allowed gap between a candidate representation and the complexity, should be allowed to depend on the numbering, but this dependence has not been mentioned in the literature and there is no clear way to determine how this constant should be determined on the basis of a given numbering.
+It may be argued that the slack parameter $c$ in the sophistication, that determines the allowed gap between a candidate representation and the complexity, should be allowed to depend on the numbering, but this dependence has not been mentioned in the literature and there is no clear way to determine how this constant should be chosen on the basis of a given numbering.
 
@@
 The gap between self-delimiting and non self-delimiting descriptions grows without bound
-\cite[Section~4.5.5]{vitanyi2004meaningful}, so that for sufficiently complex inputs all information will have to be placed in the noise part of the code, eliminating the singletons as viable candidates.
+\cite[Section~4.5.5]{li1993introduction}, so that for sufficiently complex inputs, some information will have to be placed in the noise part of the code, eliminating the singletons as viable candidates.
 This approach is taken by Vit\'anyi \cite{vitanyi2004meaningful} and by Adriaans \cite{adriaans2012facticity}.
@@
 \subsection{Underfitting}
+\label{section:underfitting}
 
-As mentioned in the related work, underfitting is a widely acknowledged problem for sophistication, and most proposals avoid it by limiting the class of allowed models, to exclude universal models.
+Underfitting is a widely acknowledged problem for sophistication, and most proposals avoid it by limiting the class of allowed models to exclude universal models.
 The following theorem shows that this does indeed yield an unbounded sophistication.
 
+\pb{Recheck what was known already}.
 \begin{restatable}{theorem}{dogfood}
@@
 We note that it is exceedingly unlikely that a deep string is sampled from a shallow distribution \cite{bloem2014safe,bennett1988logical}.
-\cite{antunes2013sophistication} also investigates the relation between $S(x)$ and $d(x)$.
 
@@
 
-This clearly contradicts the motivating principle of sophistication, which is to capture the number of bits required to express \emph{all} regularity in the data, not just the regularity that is difficult to describe.
-In this setting, a string high-sophistication not only has a lot of structure, but it also has the property that it can be compressed far better with partial functions than with total.
-That is, it is non-typical for the model class $\T$.
+The relation between $S(x)$ and $d(x)$ is also investigated in \cite{antunes2013sophistication}, where it is shown that within a logarithmic error term on the sophistication and the slack, they are identical.
+Our point here is not that the two functions are similar, but that for all practical strings, the sophistication is bounded.This contradicts the intuition that sophistication measures structure, as it seems to suggest that all strings we can possibly hope to encounter contain no structure, save a constant amount.
+The alternative is that under other numberings these strings \emph{would} have structure, but then the sophistication would no longer be invariant.
+
+As for the strings with high sophistication, they have the property that they can be compressed far better with partial functions than with total: they are non-typical for the model class $\T$.
 This suggests that the `non-stochastic' property of strings with high sophistication \cite{shen1983concept,vereshchagin2004kolmogorov} says more about depth and totality than it does about structure and noise.
+
+\pb{Vereshchagin's strongly sufficient statistic \cite{vereshchagin2013algorithmic}}
+
+\subsection{Other variants}
+
+By moving away from the idea of two-part coding, the mechanics of lemma~\ref{lemmathecoolone} can be avoided. In \cite{mota2013sophistication}, the \emph{naive sophistication} is introduced:
+\[
+\s_\text{naive}^{\psi, c}(x) = \min\left\{C^\K(x) : \log |S| - C^\K(x|s) \leq c \right\} \p
+\] 
+The condition is that the \emph{randomness deficiency} of $S$, the extent to which $x$ is typical for $S$, is constant.
+This is equivalent to the algorithmic minimal sufficient statistic based on the $\beta_x$ function from \cite{vereshchagin2004kolmogorov}.
+While the logic of Lemma~\ref{lemma:overfitting} does not work for this form, it is still an open question if it is invariant.
+
+Another approach is the \emph{coarse sophistication} \cite{antunes2009sophistication}. The simplest definition comes from \cite{mota2013sophistication}:
+\[
+\s_\text{coarse}(x) = \min_c\left\{\s^{\F,\K,\psi,c}(x) + c \right\} \p
+\]
+Again, this variant avoids the pitfalls of Lemma~\ref{lemma:overfitting}.
+If a set of $x$s has witnesses $S_x$ that are as good as the singleton set, up to a constant, but with smaller $K(S)$ by more than a constant, the added penalty $c$ will eventually be much less than the gain for the simpler witness, and the singletons will not determine the coarse sophistication.
+However, whether this form is invariant is again an open question.
+We also note that when extended from $\F$ to $\T$, it runs into the same problem as discussed in Section~\ref{section:underfitting}
+
+Finally we note that coarse sophistication moves further away from the original philosophy of sophistication than may at first be apparent\ldots \pb{Show, with some rewriting, how far this really deviates form the original sophistication.}
+
+Finally, we must mention \emph{effective complexity} \cite{gellmann1996information}.
+This measure is unique, in that it was formulated from the perspective of physics.
+Nevertheless, it fits the mold of sophistication very neatly.
+The models used are computable probability distributions on finite sets, and all candidates that come within a constant of the Kolmogorov complexity are considered.
+The complexity of the model is correctly measured by its Kolmogorov complexity.
+We note that this means that Lemmas \ref{lemma:underfitting} and \ref{lemma:overfitting} both apply.
+Unlike other sophistication measures, here it is not the smallest model which is chosen, but the one which reproduces the data with in the shortest time.
+Thus, if there are multiple candidates, this approach would likely favor the singleton sets.
+In \cite{gell2004nonextensive}, the authors abandon this approach, and note the choice from the set of candidates is a subjective one, which depends on context.
 
@@
 Nevertheless, in this section we offer the opinion that such a clean-cut separation \emph{cannot} be made to work.
-We provide two arguments to support this belief.
+We provide three arguments to support this belief.
 
 For the first argument, we take a generative perspective.
-We can generate data from a model in $\phi_i, i \in \K$ by feeding it random bits until it produces an output.
+We can generate data from a model in $\phi_i, i \in \K$, by feeding it random bits until it produces an output.
 We will call the resulting probability distribution $p_i(x) = \sum_{y:\phi_i(y) = x} 2^{-|y|}$.
-In this scenario, a sophistication could be called \emph{consistent} if, for data of sufficiently high complexity, it accurately reflects the complexity of that distribution from which it originated.
+In this scenario, a sophistication could be called \emph{consistent} if, for data of sufficiently high Kolmogorov complexity, it accurately reflects the complexity of that distribution from which it originated.
 Suppose that the data is sampled from the universal distribution: let $\phi_u(\bar\imath y)=\phi_i(y)$ and sample from $p_u$ as described above.
@@
 Given an input $x$ and some model class, say $\C$, consider the set of all possible two-part representations of $x$.
-Now, when the numbering of partial computable functions is changed, the codelength of the model part of all these representations will change.
+Now, when the numbering is changed, the codelength of the model part of all these representations will change.
 This is illustrated in the right graph in Figure~\ref{fig:diagram}.
-However, the Invariance Theorem expresses that this change is limited by a constant term that does not depend on $x$.
+The invariance theorem expresses that this change is limited by a constant term that does not depend on $x$.
 So for sufficiently large $x$, the horizontal shift of all the representations will be relatively small.
@@
 And while we cannot \emph{prove} in general that such jumps can always occur, there seems to be no reason to believe that they don't.
+Indeed, in \cite{antunes2013sophistication} it is shown that logarithmic changes in the slack parameter can already cause these effects.
 
-For this reason, we take a skeptical view of sophistication.
+Our third insight uses the opposite perspective from the one that guides sophistication: the task of \emph{randomness extraction}.
+This is a common task in computer science, where one is faced with data containing some randomness and some structure, and one wishes to extract a true random bitstring from it.
+This is easily cast in terms of Kolmogorov complexity: we are given a string $x$, and are required to extract the maximum amount of random bits from it in a deterministic fashion.
+Instead of finding the structure and discarding the noise, we are attempting the opposite.
+Consider all two part codes for $x$, using any model in $\C$.
+If we take the approach of discarding the model and returning the input as randomness, which gives us the best result?
+Clearly the answer is the universal model.
+And this should come as no suprise: if we have a shortest program for $x$, a witness to the Kolmogorov complexity, it is random in its entirety, model and input.
+The most randomness we can extract is the whole of the program, not just the input to the model.
+This emphasises a final time that in the model class of all computable functions, what is noise and what is model is a subjective matter.
+If we want maximal noise we can get, and if we want maximal structure, we can get that too.
+\pb{Does this make any sense?}
+
+We contrast this view with statements of the following kind (from \cite{vereshchagin2004kolmogorov}), commonly found in the literature:
+\begin{quote}
+This expression emphasizes the two-part code nature of Kolmogorov complexity. In the example
+\[
+x = 01010101010101010101010101
+\]
+we can encode $x$ by a small Turing machine printing a specified number of copies of the pattern ``01'' which computes $x$ from the program ``13.'' This way, $K(x)$ is viewed as the shortest length of a two-part code for $x$, one part describing a Turing machine, or model, for the regular aspects of $x$, and the second part describing the irregular aspects of $x$ in the form of a program to be interpreted by $T$.
+The regular, or ``valuable,'' information in $x$ is constituted by the bits in the ``model'' while the random or ``useless'' information of $x$ constitutes the remainder.
+\end{quote}
+Sophistication may yet yield a function which can be proved invariant and non-vacuous, and its investigation has yielded many interesting results.
+However, this view of and objective separation of structure and noise does not seem to coincide with the reality of how sophistication functions actually behave.
+
+So we take a skeptical view of sophistication.
 Note that part of the theory is fine: there is nothing wrong with evaluating models for the data by comparing their two-part code lengths.
 In fact, the randomness deficiency $\delta(\phi_i,x)=-\log p_i(x)-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data---if $p_i$ truly generated the data, then the probability that the randomness deficiency is more than $k$ bits is less than $2^{-k}$ \cite[Lemma~6]{bloem2014safe}.
-In the Monet example above, this will allow us to disqualify the model expressing that the data is actually, say, a recording of Jazz music.
+In the Monet example above, this will allow us to disqualify the model expressing that the data is actually, say, a recording of jazz music.
 
-But fundamental problems arise as soon as hard cut-offs are introduced.
+But fundamental problems arise as soon as a hard cut-off is introduced on how far we are allowed to deviate from the minimum determined by the Kolmogorov complexity.
 In our opinion, a lot of measures taken in the literature to address problems with sophistication, such as restricting the model class or introducing model penalties, complicate the method and make problems harder to analyse, without actually addressing the fundamental issue.
@@
 \appendix
-
-\section{Proofs}
-
-\subsection{Invariance of function complexity}
-
+\section{Appendix}
 \begin{lemma}[Invariance of function complexity]
@@
 If our prefix function is less efficient than $C^\K(m)$ by an unbounded amount, $i^*y$ will always beat $\bar\imath y$ for complex enough data.
-We conjecture that no computable prefix encoding function can avoid this issue:
-\begin{conjecture}
-For any partial computable prefix codelength function $L$, the function $\min\{L(i)-C^\K(i):i\ge i_0\}$
-is unbounded in $i_0$.
-\end{conjecture}
 The following lemma shows that even index sophistications using a faithful index are bounded if the prefix encoding is inefficient.
@@
 \end{restatable}
-If our conjecture is correct, the index sophistication is always bounded for model class $\C$.
-Any inefficieny in the model description, and a second represention using a universal model with a more efficient model description will win, causing a bounded sophistication.
-
 \begin{proof}
@@
 \end{proof}
+This particular consequence can be avoided if the prefix function is a \emph{Solovay function} \cite[Section~???]{downey2010algorithmic}, but the resulting sophistication would still be far from well-behaved.
 

### 2015-05-12 Facticity/Facticity/paper.tex
Ugh...

@@
 \newcommand{\N}{\mathbb N}
+\newcommand{\Nm}{\mathscr N}
 \newcommand{\tn}[1]{\textnormal{#1}}
@@
 Kolmogorov complexity measures the amount of information in data, but does not distinguish structure from noise.
-Kolmogorov's definition of the \emph{structure function} was the first attempt to measure only the structural information in data, by measuring the complexity of the smallest model allowing for optimal compression of the data.
+Kolmogorov's definition of the \emph{structure function} was the first attempt to measure only the structural information in data, by measuring the complexity of the smallest model that allows for optimal compression of the data.
 Since then, many variations of this idea have been proposed, for which we use \emph{sophistication} as an umbrella term.
@@
 \section{Introduction}
+\enlargethispage{2\baselineskip}
+\footnotetext{Peter Bloem is a student author (PhD student).}
 
@@
 For example, a sequence of a million coin flips will almost certainly have maximal Kolmogorov complexity, even though there is nothing complex about flipping a coin repeatedly.
-Many scholars have defined additional complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information.
+Many scholars have defined additional measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful}.
 While this concept has been given many names, we use \emph{sophistication} as an umbrella term.
-In this paper, we investigate two serious problems with sophistication measures.
-We conclude with three arguments that suggest the problems are fundamental, explaining our belief that sophistication cannot be defined in a satisfactory manner.
+In this paper, we investigate two serious problems with sophistication.
+We conclude with two arguments suggesting the problems are fundamental, explaining our belief that sophistication cannot be defined in a satisfactory manner.
 
-The Kolmogorov complexity $C(x)$ of a binary string $x$, is (roughly) the length of the shortest computer program that will output $x$.
-This length depends on which programming language is used, but, as shown by a central result known as the invariance theorem \cite[Section~2.1]{li1993introduction}, the language affects the value only by a constant, independent of $x$.
-This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings, a property \emph{of the data.} A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees:
+The Kolmogorov complexity $C(x)$ of a binary string $x$ is, informally, the length of the shortest computer program to print $x$.
+This length depends on the choice of programming language, but, by the invariance theorem \cite[Section~2.1]{li1993introduction}, this choice affects the value only by a constant, independent of $x$.
+For sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure.
+A definition of sophistication $S(x)$ in the spirit of $C(x)$ should have similar guarantees:
 
@@
 \item There should be no constant $c$ such that $S(x)\le c$ for every input $x$.
-If sophistication is bounded, then knowing its value under one programming language provides no constraints on its value under another language (except that it's also bounded).
-Moreover, a bounded sophistication would be at odds with the intuition that there is no limit to the amount of structure that a binary string can exhibit.
+If sophistication is bounded, then knowing its value under one programming language provides no constraints on its value under another language (except it is also bounded).
 \item Similarly, there should be no constant $c$ such that $C(x)-S(x)\le c$ for all $x$, because then sophistication would be equivalent to Kolmogorov complexity. 
 \end{enumerate}
-There have been many related proposals for such a measure, all based on a \emph{two-part code}: rather than encoding $x$ by means of a single computer program, we encode a \emph{model} for $x$ in the first part of the code, which is interpreted as a representation of $x$'s structural properties.
-The model does not fully specify $x$, but when combined with the second part of the code, which specifies the noise, the original string becomes fully determined.
-Thus, the model can be interpreted as a function mapping the noise to the original object.
+There have been many proposals for such a measure, all based on a \emph{two-part code}: we encode a \emph{model} in the first part of the code, which is interpreted as a representation of $x$'s structural properties.
+The model does not fully specify $x$, but when combined with the second part of the code, which specifies the noise, the original string becomes fully determined.\footnotemark
 
-For any string $x$, there may be many different two-part representations.
-The total codelength can never be less than the Kolmogorov complexity, but it can come close.
-The balance between the two parts may vary; Figure~\ref{fig:diagram} provides an illustration.
-The various proposals then define sophistication in terms of the length of the structure part of this two-part code for one of the representations that (almost) achieves the Kolmogorov complexity.
-However, for most of these definitions, we can prove that they fail one or more of the conditions above.
-Vit\'anyi's definition \cite{vitanyi2004meaningful} is the one exception that we cannot \emph{prove} to conflict with our requirements, but we show that this method only assigns substantial sophistication to strings that require an enormous amount of processing to construct.
+\footnotetext{Some variants deviate from the two-part coding format, see Section~\ref{section:other}.}
+
+For any string $x$, there may be many different two-part codes.
+The total length can never be less than the Kolmogorov complexity, but it can come close.
+Figure~\ref{fig:diagram} illustrates the principle.
+The various proposals define sophistication as the size of the model for one of the representations that approaches the Kolmogorov complexity.
+However, for most definitions, we can prove  they fail one of the conditions above.
+For certain variants, we cannot \emph{prove} they conflict with our requirements, but we show these methods only assign substantial sophistication to strings that require an enormous amount of processing to construct.
 
@@
   \centering
-  \begin{minipage}{0.45\textwidth}
+  \begin{minipage}{0.40\textwidth}
      \includegraphics[width=\textwidth]{./img/sophistication.pdf}
   \end{minipage}
-  \begin{minipage}{0.45\textwidth}
+  \begin{minipage}{0.40\textwidth}
      \includegraphics[width=\textwidth]{./img/sophistication-jump.pdf}
@@
 
-The second issue is that of striking the right balance between underfitting and overfitting, which we consider in Section~\ref{section:balance}.
-Overfitting is a common problem in the statistical literature, that refers to the tendency of model selection to choose a complex model that provides a very good fit to the observed data, but does not generalise well to unseen data.
+The second issue is that of striking the right balance between under- and overfitting, which we consider in Section~\ref{section:balance}.
+Overfitting is a common problem in statistics, that refers to the tendency to choose a complex model that provides a very good fit to the observed data, but does not generalise well to unseen data.
 In the case of sophistication, overfitting occurs if the model that determines the sophistication contains much or even all of the noise.
-This is a real danger, because the models are encoded so efficiently.
 In statistics, overfitting is often addressed by penalising complex models.
@@
 In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language.
-Such \emph{universal models} are \emph{simple} in the language of sophistication, since they can be described with a relatively small number of bits, yet are able to represent the data within a codelength a constant away from the Kolmogorov complexity.
+Such \emph{universal models} are \emph{simple}, since they can be described with a relatively small number of bits, yet are able to represent the data using a code within a constant from the Kolmogorov complexity.
 Such a two-part representation essentially encodes all information as noise.
 If complex models are penalized, then the problem becomes to make sure that universal models are not \emph{always} preferred for complex data.
-The usual workaround is to restrict the set of allowed models to exclude the universal ones.
-For example, the model class may be restricted to the set of all total recursive functions.
+The usual workaround is to restrict the set of allowed models, for instance to total functions.
 While this excludes universal models, it is questionable whether it adequately solves the problem of underfitting in general.
 
-Finally, in the discussion in Section~\ref{section:conclusion} we argue that while two-part coding can yield useful insights into the structure of the data and allows us to identify some models as poor representations, it is probably not possible to uniquely separate structure from noise and identify a \emph{single} model as ``best'': in general many models, of vastly different complexities, may be reasonable representations.
+Finally, In the discussion in Section~\ref{section:conclusion} we argue that while two-part coding can yield useful insights into the structure of the data and allows us to identify some models as poor representations, it is probably not possible to uniquely separate structure from noise and identify a \emph{single} model as ``best'': in general many models, of vastly different complexities, may be reasonable representations.
 Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that the data allows for multiple equivalent interpretations of which information is structured, and which is random, and that there is no such thing as sophistication.
 
-\section{Preliminaries}
-One of the main issues in dealing with sophistication in a general sense, is that one must unify many different approaches, all founded on slightly different principles.
-We have devised a general framework which, while it may not represent the idiosyncracies of all methods, allows us to rephrase them in a uniform way that illustrates the properties we discuss.
-We hope the reader will trust that these properties remain also in the original statements of the measures.
+\enlargethispage{3\baselineskip} 
+\section{Notation}
+The following notation allows us to generalize across all definitions and variants, save the occasional exception which we will highlight individually.
 
@@
 We deal with partial computable functions $f: \B \times \B \to \B$, which we will also call \emph{models}.
-A function is called \emph{prefix} if its domain, $\dom{f} = \{y : f(y; \cdot) \neq \infty\}$ is a prefix free set, i.e.
-no string in $\dom{f}$ is a prefix of another.
-A function $f$ is \emph{total} if $\dom{f} = \B$.
+A function is called \emph{prefix} if its domain, $\text{dom}_z(f)~=~\{y~:~f(y,~z)~\neq \infty\}$, is a prefix free set for all $z$, i.e.
+no string in $\text{dom}_z(f)$ is a prefix of another.
+A function $f$ is \emph{total} if $\forall_z \text{dom}_z(f) = \B$.
 In most cases, we do not use the second argument, and we adopt the convention that $f(x) = f(x, \epsilon)$.
@@
 
-A \emph{model class} is a set of model indices. We distinguish four important model classes:
+A \emph{model class} is a set of model indices. We distinguish four classes:
 \begin{itemize} 
@@
   \item $\K$ is an enumerable set such that $\{\phi_i:i\in\K\}$ is the set of all partial computable prefix functions.
-  \item The finite sets: $\F$ is an enumerable set such that $\{\phi_i:i\in\F\}$ contains a uniform code for every finite set.
-A uniform code for a finite set $S$ is a total computable surjective function $f:\{0,1\}^{\lceil\log|S|\rceil}\to S$.
+  \item The finite sets: $\F$ is an enumerable set such that $\{\phi_i:i\in\F\}$ contains a uniform code for every finite set.\footnote{A uniform code for a set $S$ is a surjective prefix function $f:\{0,1\}^{\lceil\log|S|\rceil}\to S$.}
 \end{itemize}
+A model class is thus defined in reference to a specific numbering, indicated by the context.
 
@@
 
-For technical reasons, we deviate slightly from the traditional notation of Kolmogorov complexity: we define the $\M$-Kolmogorov complexity of a binary string $x$, with respect to a model class $\M$ and a numbering $\psi$, as follows $C^{\M,\psi}(x\mid z)=\min\{\bar\imath y:\psi_i(y; z)=x,i\in\M\}$, with $C^\M(x) = C^\M(x\mid \epsilon)$.
-This means that $C^\C(x)$ corresponds to the plain Kolmogorov complexity, traditionally denoted $C(x)$, and $C^\K(x)$ corresponds to the prefix-free Kolmogorov complexity $K$.
-This notation also allows us to represent the smallest possible description of $x$ using a single model $\phi_i$ as $C^{\{i\}}(x)$.
-Note that this includes the description of $i$ as well as its shortest program for $x$.
+For technical reasons, we deviate slightly from the traditional notation of Kolmogorov complexity: we define the $\M$-Kolmogorov complexity of a binary string $x$, with respect to a model class $\M$ and a numbering $\psi$, as follows $C^{\M,\psi}(x\mid z)=\min\{\bar\imath y:\psi_i(y; z)=x,i\in\M\}$, with $C^{\M,\psi}(x) = C^{\M,\psi}(x\mid \epsilon)$.
+We may drop the numbering parameter when the distinction isn't relevant.
+Under this notation $C^\C(x)$ corresponds to the plain Kolmogorov complexity $C(x)$ and $C^\K(x)$ corresponds to the prefix-free version $K(x)$.
+We can also represent the smallest two-part description of $x$ using a given model $\phi_i$ as $C^{\{i\}}(x)$.
+The complexity of a finite set is defined an the complexity of an explicit enumeration.
 
-Finally we note that the choice of numbering, combined with a choice of prefix function, serves the same purpose that is normally expressed in the choice of a Turing machine.
-Requiring an acceptable numbering and a computable prefix function is equivalent to choosing a universal Turing machine: this requirement suffices to make the Kolmogorov complexity invariant.
-As we shall see, however, this is not enough for a proper treatment of sophistication: we must choose our numbering with care.
-Thus, we phrase our requirement for an invariant sophistication with respect to the \emph{numbering} chosen.
-There should be strict guarantees on how much the value of $S(x)$ can change if the numbering is varied.
+We use a choice of numbering for the purpose that is normally served by the universal Turing machine.
+The choice is a matter of taste: a UTM determines a numbering, and vice versa.
+The reason that we prefer to work with numberings is that it highlights an important issue: while Kolmogorov complexity is invariant to the choice of numbering this property does not immediately carry over to sophistication: for some treatments, the result is highly dependent on the chosen numbering (as shown in the next section).
 
@@
 
-The simplest approach to sophistication would simply be to `open up' the Kolmogorov complexity and to see which program on the universal Turing machine achieves the smallest description length: the program that \emph{witnesses} the Kolmogorov complexity.
-Since any UTM is defined in terms of a two-part coding, such programs necessarily consist of a model and an input.
-As is common, we will allow all witnesses within a constant slack term $c$ from the Kolmogorov complexity, and choose the one with the smallest model.
+The simplest approach to sophistication would be to `open up' the Kolmogorov complexity and to see which program on the universal Turing machine achieves the smallest description length: the program that \emph{witnesses} the Kolmogorov complexity.
+Since any UTM uses a two-part coding, its programs consist of a model and an input:
 
 \begin{definition}[Index sophistication]
+Let $\psi$ be an acceptable numbering.
+Let $\M$ be the model class from which candidates are chosen, and let $\Nm$ be the model class that determines the minimum achievable complexity.
+Let $c$ be a fixed constant.
+The \emph{index sophistication} is:
 \[
-\s_{\text{index}}^{\M_1,\M_2,\psi,c}(X) = \min\left\{ |i| : \i \in \M_1, C^{\{i\}, \psi}(x) \leq C^{\M_2, \psi}(x)+c   \right\}
+\s_{\text{index}}^{\M,\Nm,\psi,c}(x) = \min\left\{ |i| : i \in \M, C^{\{i\}, \psi}(x) \leq C^{\Nm, \psi}(x)+c   \right\} \p
 \]
-When $\M_1 = \M_2$, we will abbreviate with $\s^{\M_1,\psi,c}$
+When $\M = \Nm$, we will abbreviate with $\s^{\Nm,\psi,c}$. \label{definition:index}
 \end{definition}
-
-Koppel and Atlan's treatment \cite{koppelSoph1988,koppel1991almost}, where the name \emph{sophistication} originates, follows this basic logic, although it contains many idiosyncracies like the use of monotonic models, and an inherent extension to infinite strings.
+Koppel and Atlan's treatment \cite{koppelSoph1988,koppel1991almost}, where the name \emph{sophistication} originates, follows this basic logic, although it contains idiosyncracies like the use of monotonic models, and an extension to infinite strings.
 As the subsequent history of sophistication has discarded these, we will not discuss them here.
 
-In \cite{antunes2009sophistication,antunes2013sophistication} Koppel's principle is recast to finite strings.
+In \cite{antunes2009sophistication,antunes2013sophistication} Koppel's principle is limited to finite strings, with the total functions as a model class.
 The definition used is similar to $s_\text{index}^{\T, \C,\psi,c}$, with one exception: the total complexity of a witness $(i,y)$ is measured as $|i|+|y|$ without the cost of delimiting the two.
 This difference is not relevant to the point we make in this section.
-The restriction to total classes is a common technique to avoid underfitting, which we will discuss in the next section.
+The restriction to total classes is a common approach, perhaps used because it helps to avoid underfitting, which we will discuss in the next section.
 
 \begin{lemma}
-Let $s^\psi$ denote any index sophistication with respect to numbering $\psi$.
+Let $\s^\psi$ denote any index sophistication with respect to numbering $\psi$.
 There are acceptable numberings $\psi$ and $\xi$ such that for all $x$:\\
-\-\hspace{2.8em} $|s^\psi(x) - s^\xi(x)|\geq \tfrac{1}{2}\min\{s^\psi(x),s^\xi(x)\}$.
+\-\hspace{2.8em} $|\s^\psi(x) - \s^\xi(x)|\geq \tfrac{1}{2}\min\{\s^\psi(x),\s^\xi(x)\}$.
 \end{lemma}
@@
 Let $s_i$ be the string consisting of $2^{i}$ zeroes followed by a one.
-Define $\phi, \xi$ such that $\psi_j(x) = \phi_i(x)$ for $j = s_{2i}$ and $\xi_j(x) = \phi_i(x)$ for $j = s_{2i+1}$, with all other functions returning $\infty$.
+Define $\psi, \xi$ such that $\psi_j(x) = \phi_i(x)$ for $j = s_{2i}$ and $\xi_j(x) = \phi_i(x)$ for $j = s_{2i+1}$, with all other functions returning $\infty$ for all inputs.
 Choose any $x$ and assume w.l.o.g.
-that $s^\psi(x)\le s^\xi(x)$.
-By construction, we have $2s_\psi(x)  \leq s_\xi(x)$ from which the lemma follows.
+that $\s^\psi(x)\le s^\xi(x)$.
+By construction, we have $2\s^\psi(x)  \leq \s^\xi(x)$.
 \end{proof}
 Thus, the length of the index is a very poor indicator of model complexity.
-In order to obtain a robust measure, define the complexity of functions as in \cite{grunwald2004shannon,vitanyi2004meaningful} by $C^{\M,\psi}(f) = \min\{C^{\M,\psi}(i):\phi_i=f\}$.
-Lemma~\ref{lemma:invariance} in the appendix shows that this definition is invariant: it is an objective property of a function.
+For a robust measure, define the complexity of functions as in \cite{grunwald2004shannon,vitanyi2004meaningful} by $C^{\M,\psi}(f) = \min\{C^{\M,\psi}(i):\phi_i=f\}$.
+Lemma~\ref{lemma:invariance} in the appendix shows that $C^\C(f)$ and $C^\K(f)$ are invariant: they are objective properties of a function.
 
 Note that such perversely inefficient numberings are not an issue for Kolmogorov complexity.
-If our universal Turing machine uses an inefficient numbering, we can switch to a more efficient one at only a constant cost.
-This gives us an insight into what happens inside Kolmogorov complexity.
-We may define the program $\bi y$ using a model $i$ and an input $y$, but if there's any inefficiency in the description of $i$, the shortest program will be achieved by using $i$ to define a more efficient model encoding, and encoding both the model and its input in $y$.
-For the Kolmogorov complexity, such considerations are academic, since we only look at the total length of the program.
-For sophistication, however, they are crucial.
+If our UTM uses an inefficient numbering, we can switch to a more efficient one at only a constant cost, keeping the complexity the same, up to a constant.
+For sophistication, however the choice of numbering is crucial.
 
@@
 Confusingly, both are used in the literature.
-First, we can measure the complexity of the \emph{model} using $C^{\M,\psi}(\phi_i)$, which is then the first part of a two-part code describing the data.
-This approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information} and in most of the present paper.
+First, we can measure the complexity of the \emph{model} using $C^{\M,\psi}(\phi_i)$, which is then the size of the first part of a two-part code describing the data.
+For instance, this approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information} and in most of the present paper.
 
 Second, we can stick to using the length of the index as the measure of sophistication, but restrict the allowed numberings.
-This approach is taken is in \cite{adriaans2012facticity}.
-Adriaans  defines \emph{facticity} as $\s_\text{index}^{\C,\psi,0}$, but only allows numberings that are \emph{faithful}: for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and $|j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
+This approach is taken by Adriaans in \cite{adriaans2012facticity}, which defines \emph{facticity} as $\s_\text{index}^{\C,\psi,0}$, but only allows numberings that are \emph{faithful}: $\forall i \exists j : \psi_i=\psi_j, |j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
 
@@
 The Kolmogorov complexity uses representations of the form $\bar\imath y$, where the bar denotes some straightforward prefix encoding to delimit the model description $i$ from its input $y$.
-If we define a second prefix encoding $\tilde{\imath}$, with $|\bar\imath|-|\tilde\imath|$ unbounded, we can define a second representation $\bar u \tilde \imath y$, at a constant overhead $|\br{u}|$, and gain more than $|\br{u}|$ for sufficiently complex strings.
-As shown in Lemma~\ref{lemma:prefix-inefficiency} in the appendix, this results in a bounded sophistication (thus the facticity is bounded).
+If we define a second prefix encoding $\tilde{\imath}$, with $|\bar\imath|-|\tilde\imath|$ unbounded, we can define a second representation $\bar u \tilde \imath y$, at a constant overhead $|\br{u}|$, and gain more than $|\br{u}|$ for sufficiently complex strings, resulting in a bounded sophistication.
 
-Because of these issues, rather than index sophistication we will use the following definition of sophistication in the remainder of this article.
-It is parameterised by a model class, a numbering, and a slack parameter $c$ that determines the maximal allowed slack between the two-part code length using a representation and the Kolmogorov complexity.
+We continue with a sophistication that avoids these issues. Let $C'^{\M, \psi} = C^\K(\psi_i) + min\{|y| : \psi_i(y) = x\}$. We define sophistication as:
+
+\begin{definition}[Sophistication]
 \[
-\s^{\M_1,\M_2,\psi,c}(x)=\min\left\{C^\K(\phi_i):i\in\M_1,C^{\{i\},\psi}(x)\le C^{\M_2,\psi}(x)+c\right\}.
+\s^{\M,\Nm,\psi,c}(x)=\min\left\{C^\K(\phi_i):i\in\M,C'^{\{i\},\psi}(x)\le C^{\Nm,\psi}(x)+c\right\}.
 \]
+\end{definition}
 
@@
 We will investigate this balance, starting with the variant $\s^{\K,\psi,c}$.
-For reasons which will become clear, this variant is not used anywhere in the literature, but it helps to illustrate the issues we wish to discuss.
+This variant itself is not used anywhere in the literature, but it helps to illustrate the issues we wish to discuss.
 
-In this setting, we know that, on the one hand, there always exist witnesses for which all information except a constant amount is stored in the input, and on the other hand, there are also candidate representations for which all information is stored in the model.
-This tells us that $S^\K$ is \emph{balanced}: there is no inefficiency in the representation forcing information in or out of the model.
-The flip side of having a balanced sophistication is that it becomes easy to show a lack of invariance, by tweaking the numbering in such a way that models in a specific subset $\M'\subset\M$ of the model class becomes an arbitrary constant number of bits cheaper to represent than all other models, so that a model in $\M'$ always ends up determining the sophistication.
+$\K$ has representations with all information except a constant in the input and representations with all information in the model.
+The downside to this balance is that it becomes easy to show a lack of invariance: by tweaking the numbering so that models in a specific subset $\M'\subset\M$ become cheaper to represent by an arbitrary amount relative to others, so that a model in $\M'$ always determines the sophistication.
 
-This means that for model classes with a universal model, we can choose our numbering to make $S^\K$ bounded, or equal to the complexity:
+This means that for model classes with a universal model, we can choose our numbering to make $S^\K$ either bounded, or equal to the complexity:
 
-\begin{restatable}[Underfitting]{lemma}{underfitting}
-Let $\M$ be a model class containing a universal model $\phi_u$, with the property that $\exists c \forall i \in \M : C^{\{u\}}(x) \leq C^{\{i\}}(x) + c \p$.
-Then, there exist numberings such that $\s^\M$ is bounded.\label{lemma:underfitting}
-\end{restatable}
+\begin{theorem}[Underfitting]
+Let $\M$ be a model class containing a universal model $\phi_u$, with the property that $\exists c \forall i \in \M, x : C'^{\{u\}}(x) \leq C'^{\{i\}}(x) + c$.
+Then, for some numberings $\s^{\M,\psi,c}$ is bounded.
+\label{theorem:underfitting}
+\end{theorem}
 
-For this reason, many treatments restrict the model class.
-What is perhaps less well known is that the same property holds in the other direction: there are numberings for which the singleton models always determine the sophistication.
+This result is well known and many treatments avoid it by restricting the model class.
+Less well known is perhaps that the same holds in the other direction: for some numberings the singletons always determine the sophistication.
 
-\begin{restatable}[Overfitting]{lemma}{overfitting}
+\begin{theorem}[Overfitting]
 Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $i\in\M$ with $\phi_i(\epsilon)=x$.
-Then there is a numbering $\psi$, and a constant $c$, such that for all $x\in\X$ we have $C^\M(x)-S^{\M,\psi}(x)\leq c$.\label{lemma:overfitting}
-\end{restatable}
+Then there is a numbering $\psi$, and a constant $c$, such that for all $x\in\X$ we have $C^\M(x)-S^{\M,\psi,c}(x)\leq c$.\label{theorem:overfitting}
+\end{theorem}
+The proofs of both theorems rely on a simple principle: for certain $\M' \subset \M$, there exist numberings which have the effect of penalizing $C^\K(\phi_i)$ for any model outside $\M'$ by an arbitrary constant amount.
+We can use this to effectively `push' these models outside of the range of candidates, ensuring that, under this numbering, a model in $\M'$ always determines the sophistication.
 
-The proof of both these lemmas relies on a simple principle: for certain subclasses $\M' \subset \M$, there exist numberings which, when used, have the effect of penalizing the codelength $C^\K(\phi_i)$ for any model outside $\M'$ by an arbitrary constant amount.
-We can use this penalty to effectively `push' these models outside of the range of considered models, ensuring that, under this numbering, a model in $\M'$ always determines the sophistication.
-
-For the lemmas above, this means setting $\M' = {u}$, with $\phi_u$ a universal model, or letting $\M'$ be the set of all singletons.
+For these theorems, this means setting $\M' = \{u\}$, with $\phi_u$ a universal model, or letting $\M'$ be the singletons.
 The requirements for $\M'$ are strict, and slightly complex.
@@
 \label{lemma:thecoolone}
-  Let $\M$ be any model class, let $\X$ be any set of binary strings and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\phi$, i.e.
-if $m=\phi_{D(p)}$ then $p$ is a $D$-description of $m$.
-Let $\M'=\{\phi_i:i\in\tn{range}(D)\}$.
+  Let $\M$ be any model class, let $\X$ be any set of binary strings and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\phi$.
+Let $\M'= \tn{range}(D)$.
 Further assume there is a constant $c$ such that:\\
-\-\hspace{1cm}(1) $\forall_{m\in\M'}:\min\{|p|:\phi_{D(p)}=m\}\le C^{\K,\phi}(m)+c$\\
-\-\hspace{1cm}(2) $\forall_{x\in\X}:C^{\M',\phi}(x)-C^{\M,\phi}(x)\le c$.\\
-Then there is an enumeration $\psi$ of the partial recursive functions such that $S^{\M,\psi}(x) = |\br{0}|+S^{\M',\phi}(x)$ for all $x\in\X$.
+\-\hspace{1cm}(1) $\forall_{m\in\M'}:\min\{|p|:\phi_{D(p)}=\phi_m\}\le C^{\K,\phi}(\phi_m)+c$\\
+\-\hspace{1cm}(2) $\forall_{x\in\X}:C'^{\M',\phi}(x)-C'^{\M,\phi}(x)\le c$.\\
+Then there is an enumeration $\psi$ of the partial recursive functions such that $S^{\M,\psi,c}(x) = |\br{0}|+S^{\M',\phi,c}(x)$ for all $x\in\X$.
 \end{restatable}
@@
 We define the numbering $\psi$ as follows:
-\[\begin{cases}
-\psi_0(p) = 1^r 0 D(p) \\
-\psi_{1^r0i}(p) = \phi_i(p) \\
-\psi_j(\cdot) = \infty &\text{if $j$ contains no zeroes.}
-\end{cases}\]
-We will show that under the $\psi$-enumeration, the best representation for $x$ using a model $m\in\M'$ is always better than the best representation using some $m\notin\M'$.
-
-First suppose $m\in\M'$. Then
+\[
+\psi_0(p) = 1^r 0 D(p), \;\;\;
+\psi_{1^r0i}(p) = \phi_i(p), \;\;\;
+\psi_j(\cdot) = \infty \;\text{if $j$ contains no zeroes.}
+\]
+We will show that under the $\psi$, the best representation for $x$ with $m\in\M'$ is always better than the best using some $m\notin\M'$. First suppose $m\in\M'$. Then
 \begin{align*}
-C^{\K,\psi}(m) &=\min\{|\bar\jmath q|:\psi_{\psi_j(q)}=m\}
-\leq\min\{|\bar0 q|:\psi_{\psi_0(q)}=m\} \\ 
-&=\min\{|\bar0 q|:\psi_{1^r0 D(q)}=m\}  
-=\min\{|q|:\phi_{D(q)}=m\} + |\bar 0|\\
-&\leq C^{\K,\phi}(m)+c+|\br{0}|,
+   C^{\K,\psi}(\phi_m) &=\min\{|\bar\jmath q|:\psi_{\psi_j(q)}=\phi_m\} \leq\min\{|\bar0 q|:\psi_{\psi_0(q)}=\phi_mm\} \\ 
+   &=\min\{|\bar0 q|:\psi_{1^r0 D(q)}=\phi_m\} =\min\{|q|:\phi_{D(q)}=\phi_m\} + |\bar 0|\\
+   &\leq C^{\K,\phi}(\phi_m)+c+|\br{0}|,
 \end{align*}
@@
 Now assume that the best model $m$ for $x$ is not in $\M'$. 
-Let $i$ be the index of $m$ with the shortest description, i.e. it achieves the minimum in $C^{\K, \psi}(m)=\min\left\{C^{\K,\psi}(i):\psi_i=m\right\}$.
-There are two possibilities. Either $i = 0$, in which case we have $C^{\K,\psi}(m)\ge r$ because $\phi_0$ cannot output zero and all other $\psi$-programs are at least $r$ bits long.
-
+Let $i$ be the index of $m$ with the shortest description, i.e. it achieves the minimum in $C^{\K, \psi}(\phi_m)=\min\left\{C^{\K,\psi}(i):\psi_i=\phi_m\right\}$.
+There are two possibilities.
+Either $i = 0$, in which case we have $C^{\K,\psi}(m)\ge r$ because $\phi_0$ cannot output zero and all other $\psi$-programs are at least $r$ bits long.
 Otherwise, we can bound
 \begin{align*}
-C^{\K,\psi}(m)&=C^{\K,\psi}(i)=C^{\K,\psi}(1^r0j)\\
+C^{\K,\psi}(\phi_m)&=C^{\K,\psi}(m)=C^{\K,\psi}(1^r0j)\\
 &\ge C^{\K,\psi}(j)-c' \ge C^{\K,\phi}(j)+r+1-c'\\
@@
 Now choose $r=c''+\max\left\{C^{\K,\psi}(\phi_0),c'-1\right\}$. Then substitution yields, for both cases, 
-$C^{\K,\phi}(m)\ge c''+C^{\K,\psi}(m)$.
-
-Combining the inequalities above, for any model $g\not\in\M'$, there is a model $f\in\M$ such that
+$C^{\K,\phi}(m)\ge c''+C^{\K,\psi}(m)$. Combining the inequalities above, for any model $\phi_g$ with $g\not\in\M'$, there is a model $\phi_f$ with$f\in\M$ such that
 \[\begin{split}
-C^{\K,\psi}(g)+\min \{|y| : g(y) = x\} &\ge C^{\K,^\phi}(g)+\min\{|y| : g(y) = x\}+c''\\
-&\ge C^{\K,\psi}(f)+\min \{|y| : f(y) = x\} -2c+c'' -|\bar0|.
+C^{\K,\psi}(\phi_g)+&\min \{|y| : \phi_g(y) = x\} \ge C^{\K,^\phi}(\phi_g)+\min\{|y| : \phi_g(y) = x\}+c''\\
+&\ge C^{\K,\psi}(\phi_f)+\min \{|y| : \phi_f(y) = x\} -2c+c'' -|\bar0|.
 \end{split}\]
-By choosing $c''$ sufficiently large, we can ensure that the best representation is in $\M'$ for all $x\in\X$, which completes the proof.
+A sufficiently large $c''$ ensures the best representation is in $\M'$ for all $x\in\X$.
 \end{proof}
 
-Lemmas~\ref{lemma:underfitting} and \ref{lemma:overfitting} follow as corollaries. For Lemma~\ref{lemma:underfitting}:
+Theorems~\ref{theorem:underfitting} and \ref{theorem:overfitting} follow as corollaries. For Theorem~\ref{theorem:underfitting}:
 \begin{proof}
@@
 That is, $\M' = \{u\}$.
-This construction satisfies the conditions 1 and 2 from Lemma~\ref{lemma:thecoolone}; invoking it we find that there exists an acceptable numbering $\psi$ for which $\s^{\M,\psi}(x) = \s^{\M', phi}(x) + c$.
-Since $\M'$ contains only a single model, $\s_{\M',\phi}(x)$ is constant.
+This construction satisfies the conditions 1 and 2 from Lemma~\ref{lemma:thecoolone}; invoking it we find that there exists an acceptable numbering $\psi$ for which $\s^{\M,\psi}(x) = \s^{\M', \phi, c}(x) + c$.
+Since $\M'$ contains only a single model, $\s^{\M',\phi, c}(x)$ is constant.
 \end{proof}
 
-And for Lemma~\ref{lemma:overfitting}
+And for Theorem~\ref{theorem:overfitting}:
 
@@
 Let $\phi_i$ be a singleton for $x$ as described.
-Note that since $\phi_i$ is a prefix function, if $\phi_i$ is defined for input $\epsilon$ then it cannot be defined for any other input.
+Note that since $\phi_i$ is a prefix function, since $\phi_i$ is defined for input $\epsilon$ then it cannot be defined for any other input.
 Pick $i,x$ with $\phi_i(\epsilon)=x$.
@@
 
-We now define a computable function $D$ by $D(\bar\imath y)=j$ where $\psi_j(\epsilon) = \psi_i(y)$.
+We now define a computable function $D$ by $D(\bar\imath y)=j$ where $\psi_j(\epsilon) = \phi_i(y)$.
 We will show that the two conditions of Lemma~\ref{lemma:thecoolone} hold for the prefix function $D$.
@@
 Now, by Lemma~\ref{lemma:thecoolone} there is a numbering $\psi_1,\psi_2,\ldots$ such that $\s^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
-We observed that $|C^{\K}(f)-C^{\K}(x)|\le c$ for any $f\in\M'$, so $\s^{\M',\psi}(x)\ge C^{\K}(x)-c$.
+We observed that $|C^{\K}(f)-C^{\K}(x)|\le c_0$ for any $f\in\M'$, so $\s^{\M',\psi, c}(x)\ge C^{\K}(x)-c_0$.
 This proves the lemma.
@@
 \]
-and suggested that the smallest set for which $K(S) + \log|S| \leq C^\K(S) + c$ for some pre-chosen constant $c$ can be seen as capturing all the structural information in $x$ \cite{cover1985kolmogorov}.
+and suggested that the smallest set for which $C^\K(S) + \log|S| \leq C^\K(S) + c$ for some pre-chosen constant $c$ can be seen as capturing all the structure in $x$ \cite{cover1985kolmogorov}.
 This is equivalent to the sophistication $\s^{\F,\K,\psi,c}$.
-Lemma~\ref{lemma:underfitting} shows us that given $c$, there are numberings for which this sophistication is always equal, up to a constant, to $C^\K(x)$.
-Thus, while the general shape of the structure function might tell an interesting story, its sophistication function is either not invariant, or vacuous.
+Theorem~\ref{theorem:overfitting} shows  there are numberings for which this sophistication is always equal to $C^\K(x)$.
+Thus, either this is true for all numberings, or this sophistication is not invariant.
 
-\pb{Discuss the minimal sufficient algorithmic statistic \cite{gacs2001algorithmic}}
+In \cite{gacs2001algorithmic} the idea of the structure function is extended to the notion of an \emph{algorithmic sufficient statistic}.
+This is essentially the witness to the sophistication $\s^{\F,\K,\psi,c}(x)$.
+Thus, by Lemma~\ref{theorem:overfitting}, there exist numberings such that for all $x$, $\{x\}$ is the minimal sufficient statistic.
 
-It is widely recognized that singleton models are always candidates in these settings, but not, it seems, that they may be the only candidates for any string, as Lemma~\ref{lemma:overfitting} shows.
-
-It may be argued that the slack parameter $c$ in the sophistication, that determines the allowed gap between a candidate representation and the complexity, should be allowed to depend on the numbering, but this dependence has not been mentioned in the literature and there is no clear way to determine how this constant should be chosen on the basis of a given numbering.
+It may be argued that the slack parameter $c$ in the sophistication, which determines the allowed gap between a candidate representation and the complexity, should be allowed to depend on the numbering, but this dependence has not been mentioned in the literature and there is no clear way to determine how this constant should be chosen on the basis of a given numbering.
 
@@
 As we have seen, a strong penalty, such as the one imposed by an inefficient prefix encoding of the model, will immediately cause underfitting.
-A more subtle approach is to use $\C$ as a model class rather than $\K$, thus allowing representations that are not self-delimiting.
+A more subtle approach is to allow descriptions that are not self delimiting.
 The gap between self-delimiting and non self-delimiting descriptions grows without bound
-\cite[Section~4.5.5]{li1993introduction}, so that for sufficiently complex inputs, some information will have to be placed in the noise part of the code, eliminating the singletons as viable candidates.
+\cite[Section~4.5.5]{li1993introduction}, so that for sufficiently complex inputs, some information will have to be placed in the noise part of the code, since placing all information in the model results in a self-delimiting description.
+This eliminates the singletons as viable candidates.
 This approach is taken by Vit\'anyi \cite{vitanyi2004meaningful} and by Adriaans \cite{adriaans2012facticity}.
 Such measures thus help reduce the overfitting problem, but they only increase the tendency to underfit.
-We also pay the price that the models can no longer be equated with probability measures, removing the link to traditional statistics and probability theory.
+We also pay the price that the models can no longer be equated with probability measures, weakening the link to traditional statistics and probability theory.
 
@@
 
-Underfitting is a widely acknowledged problem for sophistication, and most proposals avoid it by limiting the class of allowed models to exclude universal models.
-The following theorem shows that this does indeed yield an unbounded sophistication.
+The existence of universal models is a widely acknowledged problem for sophistication, and most proposals avoid it by limiting the class of allowed models to exclude them.
+It is known that there are strings for which $\s^{F, K, \psi, c}$, $\s^{T, \psi, c}$, $\s^{T, \K, \psi, c}$ are close to the length of the string (up to a logarithmic term).
+These are the so-called absolutely non-stochastic strings \cite{shen1983concept}.
+Proofs for the sophistications mentioned can be found in \cite{gacs2001algorithmic}, \cite{antunes2009sophistication} and \cite{vitanyi2004meaningful} respectively.
+The existence of these strings is independent of the numbering chosen.
 
-\pb{Recheck what was known already}.
-\begin{restatable}{theorem}{dogfood}
-There are infinitely many $x$ with $\s^{\F,\phi,c}(x) \geq C^{\K}(x)/5$, and also infinitely many $x$ with $\s^{\T,\phi,c}(x) \geq C^{\K}(x)/6$.
-\end{restatable} 
-\noindent Thus, there are no longer any numberings for which sophistication is bounded.
-However, as we show below, the resulting sophistication is not an intuitive measure of the amount of structure in the data.
-Moreover, the balance is pushed back in the direction of overfitting, which is only exacerbated by this change.
+However, the problem of the singletons remains.
 Only one model class eliminates both the singletons and the universal model: $\T$.
-The only proposal that uses an efficient model representation \emph{and} excludes the universal models \emph{and} excludes the singletons is \cite{vitanyi2004meaningful}.
-While this invalidates straightforward proofs of boundedness, there is no suggestion that $S^{\T, \phi, c}(x)$ is actually invariant.
+The only proposal we are aware of that uses an efficient model representation \emph{and} excludes the universal models \emph{and} excludes the singletons is \cite{vitanyi2004meaningful}, where $\s^{\T,\K,\psi,c}$ is used.
+While this invalidates straightforward proofs of boundedness, there is no evidence that $S^{\T,\K,\psi, c}(x)$ is actually invariant either.
 
-While under $S^{\T, \phi, c}(x)$ there are strings with high sophistication, these may not conform to the original intuition that motivated sophistication. To show this, we need the concept of depth:
+While under $S^{\T,\K,\psi, c}(x)$ there are strings with high sophistication, these may not conform to sophistication's motivating intuition. To show this, we need the concept of depth:
 
@@
 
-\begin{restatable}{theorem}{depth}
+\begin{theorem}
 Let $A(n)$ be the single-argument Ackermann function and $c$ some arbitrary constant.
-There are numberings such that for strings with depth $d^{\C,c}(x) \leq A(C^\C(x))$ the sophistication $\s^{\T, \psi, c}$ is bounded.
-The same is true for $\s^{\F}$ .
-\end{restatable}
+There are numberings such that for strings with depth $d^{\C,c}(x) \leq A(C^\C(x))$ the sophistication $\s^{\T, \K, \psi, c}$ is bounded.\label{thm:depth}
+\end{theorem}
+
+\begin{proof}
+Let $U(\bar\imath y)$ be some universal Turing machine, and let $U^A(\bar\imath y)$ be a simulation of that machine which outputs $0$ if the number of steps taken exceeds $A(|\bar\imath y|)$.
+Let $u$ be the index of the function $U^A$ in the standard enumeration.
+
+Let $D$ be a prefix function with $D(\epsilon) = u$.
+We can instantiate Lemma~\ref{lemma:thecoolone} with $D$, $\M' = \{\phi_u\}$ and $X = \{x : d^{\C,c}(x) \leq A(C(x))\}$.
+This tells us that there exists a numbering for which $\s^{\T, \K,\psi,c}(x) = \s^{\M', \psi, c}(x) + |\bar0| \leq c$ for all $x \in X$.
+\end{proof}
+
 \noindent This shows that while high-sophistication strings exist, they do not behave as expected.
-Consider a string that is typical for a shallow model, say some elaborate Markov chain.
-Under $\s^{\T,\psi,c}$, no matter how high the complexity of the Markov chain, the sophistication is bounded.
-Any structure that is simple enough to be exploited within the time bound of the Ackermann function (or some other fast-growing, simply stated bound) will be counted as noise.
+Consider a string that is typical for a shallow model, say some elaborate probabilistic automaton.
+Under $\s^{\T,\K, \psi,c}$, no matter how high the complexity of the automaton, the sophistication is bounded.
+We could encode the collected works of Shakespeare in its transition graph, and this information would be counted as noise by the sophistication.
+Any structure simple enough to be exploited within the time bound of the Ackermann function will not be seen as `meaningful information'.
 Only structure so deep that it would take beyond the lifetime of the universe to decompress would count towards sophistication.
+In the remainder we will refer to strings $x$ with $d^{\C,c}(x) \leq A(C(x))$ as \emph{non-deep} strings.
 
 The relation between $S(x)$ and $d(x)$ is also investigated in \cite{antunes2013sophistication}, where it is shown that within a logarithmic error term on the sophistication and the slack, they are identical.
-Our point here is not that the two functions are similar, but that for all practical strings, the sophistication is bounded.This contradicts the intuition that sophistication measures structure, as it seems to suggest that all strings we can possibly hope to encounter contain no structure, save a constant amount.
+Our point is not the similarity between the two, but that for all practical strings, the sophistication is bounded.
+This contradicts the intuition that sophistication measures structure, as it seems to suggest that all strings we can possibly hope to understand or generate contain no structure, save a constant amount.
 The alternative is that under other numberings these strings \emph{would} have structure, but then the sophistication would no longer be invariant.
@@
 
-\pb{Vereshchagin's strongly sufficient statistic \cite{vereshchagin2013algorithmic}}
+\subsection{Other variants}
+\label{section:other}
 
-\subsection{Other variants}
+By moving away from the idea of two-part coding, the mechanics of lemma~\ref{lemma:thecoolone} can be avoided.
+In \cite{mota2013sophistication}, the \emph{naive sophistication} is introduced.
+We will define a generic version, parametrized by model class.
 
-By moving away from the idea of two-part coding, the mechanics of lemma~\ref{lemmathecoolone} can be avoided. In \cite{mota2013sophistication}, the \emph{naive sophistication} is introduced:
+\begin{definition}[Naive sophistication]
+Let $C_i(x) = \min\left\{|y| : \phi_i(y) = x\right\}$. Then we define the \emph{naive sophistication} as:
 \[
-\s_\text{naive}^{\psi, c}(x) = \min\left\{C^\K(x) : \log |S| - C^\K(x|s) \leq c \right\} \p
+\s_\text{naive}^{\M, \psi, c}(x) = \min\left\{C^\K(\psi_i) : i \in \M, C_i(x) - C^\K(x\mid s) \leq c \right\} \p
 \]
-The condition is that the \emph{randomness deficiency} of $S$, the extent to which $x$ is typical for $S$, is constant.
-This is equivalent to the algorithmic minimal sufficient statistic based on the $\beta_x$ function from \cite{vereshchagin2004kolmogorov}.
-While the logic of Lemma~\ref{lemma:overfitting} does not work for this form, it is still an open question if it is invariant.
+\end{definition}
 
+The condition now is not that the two-part code length corresponds to the Kolmogorov complexity, as in the regular sophistication, but that the \emph{randomness deficiency} is less than a constant.
+$\s_\text{naive}^{\F,\psi,c}(x)$ corresponds to the version in \cite{mota2013sophistication}.
+The switch to the randomness deficiency avoids Theorem~\ref{theorem:overfitting}, but we end up with the same problem as Vit\'anyi's sophistication (Theorem~\ref{thm:depth}): for non-deep strings $s^{\T, \psi, c}$ is defined by the model $U^A$, and thus bounded.
+
+We cannot show that $s^{\F, \psi, c}(x)$ is bounded for non-deep strings, but this is only a consequence of the use of sets, not of the switch to randomness deficiency as a condition.
+\emph{Any} set sophistication is necessarily lower-bounded by the function $\text{set}(x) = \min\left\{C^\K(S) : x \in S\right\}$ and if this function were bounded, it would suggest that a finite amount of sets contained all strings.
+\begin{theorem}
+Let $x$ be a non-deep string. Then $\s_\text{naive}^{\T,\K,\psi,c}(x)$ is bounded and
+\[
+\text{set}(x) \leq \s_\text{naive}^{\F,\K,\psi,c}(x) \leq C^\K(C^\K(x)) \p
+\]\label{theorem:naive}
+\end{theorem}
+\begin{proof}
+For the first part, we take the model $U^A$ as defined in Section~\ref{section:underfitting}.For some numbering $\psi$, we have $C_i(x) - C^\K(x\mid S) \leq 0$, thus $\s_\text{naive}^{\T,\psi,c}(x) \leq C^K(U^A)$.
+For the second part, let $C^\K(x) = k$ and $S^A_k = \{x \mid \exists p : U^A(p) = x, |p| \leq k\}$, and let $S_k = \{x : \exists p : U(p) = x, |p| \leq k\}$.
+We will first show that $C^\K(x) = C^\K(S_k^A) + \log |S_k^A|$ up to a constant., ie $S^A_k$ is \emph{optimal} for $x$.
+It's known that  $\log |S^A_k| \leq \log |S_k| = k - K(k) + O(1)$ \cite{gacs2001algorithmic}.
+We also have $C^\K(S_k^A) \leq C^\K(k) + O(1)$.
+Thus
+\begin{align*} 
+C^\K(x) &= k + C^\K(k) -C^\K(k) \\
+&\geq C^\K(S^k) + \log |S^A_k| \geq C^\K(x) \\  
+\end{align*}
+Where all (in)equalities are up to a constant.
+The last inequality follows from the fact that $C^\K(x)$ is the smallest possible self-delimiting description.
+By Lemma III.9 in \cite{gacs2001algorithmic}, optimality implies typicality, so the randomness deficiency of $x$ in $S_k^A$ is bounded.
+Thus, for large enough $c$, $\s_\text{naive}^{\F,\psi,c}(x) \leq C^\K(C^\K(k))$.
+\end{proof}
 Another approach is the \emph{coarse sophistication} \cite{antunes2009sophistication}. The simplest definition comes from \cite{mota2013sophistication}:
 \[
-\s_\text{coarse}(x) = \min_c\left\{\s^{\F,\K,\psi,c}(x) + c \right\} \p
+\s^{\M,\Nm,\psi, c}_\text{coarse}(x) = \min_c\left\{\s^{\M,\Nm,\psi,c}(x) + c \right\} \p
 \]
-Again, this variant avoids the pitfalls of Lemma~\ref{lemma:overfitting}.
-If a set of $x$s has witnesses $S_x$ that are as good as the singleton set, up to a constant, but with smaller $K(S)$ by more than a constant, the added penalty $c$ will eventually be much less than the gain for the simpler witness, and the singletons will not determine the coarse sophistication.
-However, whether this form is invariant is again an open question.
-We also note that when extended from $\F$ to $\T$, it runs into the same problem as discussed in Section~\ref{section:underfitting}
+Again, this variant avoids the pitfalls of Theorem~\ref{theorem:overfitting}.
+If a set of $x$s has witnesses $S_x$ that are as good as the singleton set, up to a constant, but with smaller $C^\K(S)$ by more than a constant, the added penalty $c$ will eventually be much less than the gain for the simpler witness, and the singletons will not determine the coarse sophistication.
+It was noted in \cite{antunes2009sophistication} that the coarse sophistication is within a logarithmic term of the busy beaver depth, again relating sophistication to depth.
+For our purposes we can show that, as with the naive sophistication the total function version is bounded and the set version grows very slowly:
+\begin{theorem}
+Let $c_0$ be some constant. Then for all non-deep strings $x$ we have that $\s_\text{coarse}^{\T,\K,\psi,c}(x)$ is bounded and
+\[
+\text{set}(x) \leq s_\text{coarse}^{\F,\K,\psi,c}(x) \leq C^\K(C^\K(x)) + c_0
+\].\label{theorem:coarse}
+\end{theorem}
+\begin{proof}
+For $\s_\text{coarse}^{\T,\K,\psi,c}(x)$, again take $U^A$ as a model.
+By Lemma~\ref{lemma:thecoolone}, there exists a numbering $\psi$ so that for all shallow $x$, a representation with $U^A$ witnesses the Kolmogorov complexity exactly.
+Thus $\s_\text{coarse}^{\T,\K,\psi,c}(x) \leq C^\K(U^A)$ for all non-deep strings.
+For the set case, we re-use $S_k^A$ as defined in the proof of Theorem~\ref{theorem:naive}.
+If we set $c$ to $C^\K(C^\K(x))$ in the regular sophistication, the upper bound follows.
+The lower bound follows as established earlier.
+\end{proof}
+In \cite{vereshchagin2013algorithmic}, Vereshchagin proposes a strongly algorithmic sufficient statistic.
+Whereas the regular algorithmic sufficient statistic $S$ from \cite{gacs2001algorithmic} has as one of its requirement that $C^\K(S\mid x)$ is constant, the strong variant imposes the requirement that $C^\T(S\mid x)$ is also constant.
+This reduces the problems of underfitting discussed in this section, but since $C^\T(\{x\}\mid x)$ is bounded, by Theorem~\ref{theorem:overfitting}, overfitting remains a problem: for a given constant deviation allowed from the Kolmogorov complexity, there exist numberings so that only the singletons come within that constant.
 
-Finally we note that coarse sophistication moves further away from the original philosophy of sophistication than may at first be apparent\ldots \pb{Show, with some rewriting, how far this really deviates form the original sophistication.}
-
-Finally, we must mention \emph{effective complexity} \cite{gellmann1996information}.
+Finally, we must mention \emph{effective complexity} \cite{gellmann1996information}, proposed by Gell-Man and Lloyd.
 This measure is unique, in that it was formulated from the perspective of physics.
@@
 The complexity of the model is correctly measured by its Kolmogorov complexity.
-We note that this means that Lemmas \ref{lemma:underfitting} and \ref{lemma:overfitting} both apply.
+We note that this means that Lemma~\ref{theorem:overfitting} applies.
 Unlike other sophistication measures, here it is not the smallest model which is chosen, but the one which reproduces the data with in the shortest time.
@@
 But that does not in itself mean that it should be impossible to come up with a sound measure.
-The intuition throughout the literature, starting with the structure function, appears to be that the crucial property is whether a string is typical for a model, and that this typicality can be tested: another random choice from that model should select a string with the same structure
-
+The intuition throughout the literature, starting with the structure function, appears to be that the crucial property is whether a string is typical for a model, and that this typicality can be tested: another random choice from that model should select a string with the same structure.
 This idea is bold, but not necessarily unreasonable.
 Nevertheless, in this section we offer the opinion that such a clean-cut separation \emph{cannot} be made to work.
-We provide three arguments to support this belief.
+We provide two arguments to support this belief.
 
 For the first argument, we take a generative perspective.
-We can generate data from a model in $\phi_i, i \in \K$, by feeding it random bits until it produces an output.
+We can generate data from a model $\phi_i, i \in \K$, by feeding it random bits until it produces an output.
 We will call the resulting probability distribution $p_i(x) = \sum_{y:\phi_i(y) = x} 2^{-|y|}$.
-In this scenario, a sophistication could be called \emph{consistent} if, for data of sufficiently high Kolmogorov complexity, it accurately reflects the complexity of that distribution from which it originated.
-Suppose that the data is sampled from the universal distribution: let $\phi_u(\bar\imath y)=\phi_i(y)$ and sample from $p_u$ as described above.
-Then the initial random bits will determine the prefix encoded index $\bar\imath$ of the function $\phi_i$ that $\phi_u$ will subsequently emulate, and the remaining bits are used as inputs to $\phi_i$.
+In this setting, a sophistication could be called \emph{consistent} if, for data of sufficiently high Kolmogorov complexity, it reflects the complexity of that distribution from which it originated.
+Suppose the data is sampled from the universal distribution: let $\phi_u(\bar\imath y)=\phi_i(y)$ and sample from $p_u$ as described above.
+Then the initial bits will determine the prefix encoded index $\bar\imath$ of the function $\phi_i$ that $\phi_u$ will subsequently emulate, and the remaining bits are used as inputs to $\phi_i$.
 We now ask, what should be the sophistication of the resulting data?
+\enlargethispage{3\baselineskip}
 
@@
 Consider the following metaphor.
-We are given a bitstring which can be read as a bitmap image of the painting \emph{Impression of a Sunrise}.
-There are many good models for this string, ranging from very generic to very specific.
-The theory behind sophistication suggests that we can choose one of these as the objective, intrinsic model of the data.
-Different models say different things: the universal model says that it is `some finite object'.
-A more specific model might say that it is `an image'.
-Even more specific would be a painting, a Monet, or specifically the painting Impression of a Sunrise.
-A sound measure of Sophistication should be able to select one of these models as the proper representation of structure in the data, and disqualify the others as either over- or underfitting.
+We are given a a bitmap image of the painting \emph{Impression of a Sunrise}.
+There are many good models for this string, from very generic to very specific.
+Sophistication suggests that we can choose one of these as the objective, intrinsic model of the data.
+The universal model says that it is `some compressible, finite object'.
+Another might say that it is `an image'.
+Even more specific would be `a painting', `a Monet', or specifically `the painting \emph{Impression of a Sunrise}'.
+A sound sophistication should be able to select one of these as the proper representation of structure in the data, and disqualify the others as over- or underfitting.
 But how should we be able to say that the data is intrinsically more of a painting than an image?
@@
 If such jumps can occur, the sophistication is not invariant.
-And while we cannot \emph{prove} in general that such jumps can always occur, there seems to be no reason to believe that they don't.
-Indeed, in \cite{antunes2013sophistication} it is shown that logarithmic changes in the slack parameter can already cause these effects.
-
-Our third insight uses the opposite perspective from the one that guides sophistication: the task of \emph{randomness extraction}.
-This is a common task in computer science, where one is faced with data containing some randomness and some structure, and one wishes to extract a true random bitstring from it.
-This is easily cast in terms of Kolmogorov complexity: we are given a string $x$, and are required to extract the maximum amount of random bits from it in a deterministic fashion.
-Instead of finding the structure and discarding the noise, we are attempting the opposite.
-Consider all two part codes for $x$, using any model in $\C$.
-If we take the approach of discarding the model and returning the input as randomness, which gives us the best result?
-Clearly the answer is the universal model.
-And this should come as no suprise: if we have a shortest program for $x$, a witness to the Kolmogorov complexity, it is random in its entirety, model and input.
-The most randomness we can extract is the whole of the program, not just the input to the model.
-This emphasises a final time that in the model class of all computable functions, what is noise and what is model is a subjective matter.
-If we want maximal noise we can get, and if we want maximal structure, we can get that too.
-\pb{Does this make any sense?}
-
-We contrast this view with statements of the following kind (from \cite{vereshchagin2004kolmogorov}), commonly found in the literature:
-\begin{quote}
-This expression emphasizes the two-part code nature of Kolmogorov complexity. In the example
-\[
-x = 01010101010101010101010101
-\]
-we can encode $x$ by a small Turing machine printing a specified number of copies of the pattern ``01'' which computes $x$ from the program ``13.'' This way, $K(x)$ is viewed as the shortest length of a two-part code for $x$, one part describing a Turing machine, or model, for the regular aspects of $x$, and the second part describing the irregular aspects of $x$ in the form of a program to be interpreted by $T$.
-The regular, or ``valuable,'' information in $x$ is constituted by the bits in the ``model'' while the random or ``useless'' information of $x$ constitutes the remainder.
-\end{quote}
-Sophistication may yet yield a function which can be proved invariant and non-vacuous, and its investigation has yielded many interesting results.
-However, this view of and objective separation of structure and noise does not seem to coincide with the reality of how sophistication functions actually behave.
+And while we cannot \emph{prove} in general that such jumps can always occur, there seems to be no reason to believe that they do not.
+Indeed, in \cite{antunes2013sophistication} it is shown that logarithmic changes in the slack parameter can cause these effects.
 
@@
 This is dangerous: if such ad-hoc fixes result in a theory that is hard to prove either wrong or right, it creates an artificial dead end for a valuable area of research.
-When hard cut-offs are avoided, on the other hand, all such measures are suddenly no longer necessary.
+When the hard cut-off on which models are considered is avoided, on the other hand, all such measures are suddenly no longer necessary.
 What remains is a simple and elegant theory that can be used to sift through all possible models, disproving most while retaining a select number of interesting candidates for our further consideration.
@@
 \begin{align*}
-C^{\C,\psi}(f) &= \min\left\{ C^{\C,\psi}(i) : \psi_i= f\right\} 
+C^{\C,\psi}&(f) = \min\left\{ C^{\C,\psi}(i) : \psi_i= f\right\} 
 \geq \min\left\{ C^{\C, \eta}(i) : \psi_i= f\right\} - c\\
 &= \min\left\{ C^{\C, \eta}(i) : \eta_{g(i)}= f\right\} - c
-\geq \min\left\{ C^{\C, \eta}(i) : \eta_i= f\right\} - c' \\
-&= C^{\C, \eta}(f).
+\geq \min\left\{ C^{\C, \eta}(i) : \eta_i= f\right\} - c' = C^{\C, \eta}(f).
 \end{align*}
-The opposite inequality can be achieved by reversing $\psi$ and $\eta$. The same proof holds with model class $\K$.
+Reversing $\psi$ and $\eta$ for the opposite inequality. The same proof holds for $\K$.
 \end{proof}
@@
 enumeration by more than a constant term.
+%\enlargethispage{2\baselineskip}
 
@@
 
-\subsection{Inefficient indices}
-\label{section:appendix-inefficient-indices}
-
-Can we make our prefix encoding $\bar\cdot$ so efficient that no other encoding can undercut it?
-By definition, $C^{\K}(\phi_i)$ is the smallest effective, self-delimiting description of $\phi_i$.
-We cannot compute the prefix code from the data, but this is not necessary to define a representation.
-Let $i^*$ be the shortest self-delimiting program for $\phi_i$, then $i^*y$ is a decodable representation for $x$.
-If our prefix function is less efficient than $C^\K(m)$ by an unbounded amount, $i^*y$ will always beat $\bar\imath y$ for complex enough data.
-The following lemma shows that even index sophistications using a faithful index are bounded if the prefix encoding is inefficient.
-
-\begin{restatable}{lemma}{ineffprefix}
-Assume $\phi$ is faithful and define the index sophistication $s^c(x)=\min\{|\bar\imath|:C^{\{i\}}(x)\le C^{\C}(x)+c\}$.
-If $\min\{|\bar\imath|-C^\K(i):i\ge i_0\}$ is an unbounded function of $i_0$, then $s^c$ is bounded.\label{lemma:prefix-inefficiency}
-\end{restatable}
-\begin{proof}
-Let $\bar\imath y$ be any two-part representation for the data $x$, i.e.
-$\phi_i(y)=x$.
-Then construct an alternative two-part representation $\bar vi^* y$, where $i^*$ is the shortest $\psi_u$-program for $i$ such that $\phi_v(i^* y)=\phi_{\psi_u(i^*)}(y) = \phi_i(y)=x$.
-We compare the lengths of these two representations.
-Note that $|i^*|=C^{\K,\phi}(i)$.
-Therefore,
-\[
-|\bar\imath y|-|\bar v i^* y| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-C^{\K}(i).
-\]
-By assumption there must be an $i_0$ such that the above expression is positive for all $i>i_0$.
-From this $i_0$ onwards, the second representation (using $v$) will have a shorter codelength, so $\bar\imath p$ cannot achieve the minimum in the definition of the sophistication.
-Consequently, the sophistication is bounded by $s(x)<|\overline{\imath_0}|$ for all $x$.
-\end{proof}
-This particular consequence can be avoided if the prefix function is a \emph{Solovay function} \cite[Section~???]{downey2010algorithmic}, but the resulting sophistication would still be far from well-behaved.
-
-\subsection{The problem of depth}
-
-We first show that representations using $\T$ can be reduced to representations in $\F$ with a small overhead.
-
-\begin{definition}[Stochastic strings, Definition~1.1 from~\cite{gacs2001algorithmic}]
-  Let $\alpha$ and $\beta$ be natural numbers. A finite binary string $x$ is called \emph{$(\alpha,\beta)$-stochastic} if there exists a finite set $S\subseteq\B$ such that 
-  \[
-  x\in S,~C^\K(S)\le\alpha,~C^\K(x)\ge\log|A|-\beta.
-\]
-\end{definition}
-
-\begin{restatable}{lemma}{vitanyi}
-Let $f$ be a total, computable function, such that for some $d$, $f(d) = x$ and let $k = K(f) + |d|$.
-Then there exists a finite set $A$ containing $x$ and a constant $c$, such that $K(A) \leq K(f) + K(|d|) + c$ and $\log |A| \leq |d|$.\label{lemma:total-to-sets}
-\end{restatable}
-\begin{proof}
-Let $S = f\left(\{0,1\}^{|d|}\right)$.
-Since $f$ is total, this set can be explicity computed from a description of $f$ and a description of $|d|$, which tells us that there is a constant $c$ such that $C^\K(A) \leq C^\K(f) + C^\K(|d|) + c$.
-
-Since $A$ is the image of a set of size $2^{|d|}$ under $f$, we have $\log |A| \leq |d|$.
-\end{proof}
-\noindent This lemma is a variation on \cite[Lemma~7.2]{vitanyi2004meaningful}.
-
-We use this to show that sophistications with a restricted model class are not bounded:
-\dogfood*
-\begin{proof}
-In this proof, for convenience we identify the models indexed by $\F$ with the sets that they encode.
-Thus the two-part code consists of the description of a set $A$, and then an index in this set using $\log|A|$ bits.
-For all $i$, let $x_i$ be the smallest binary string that is not $(i/4, i/4)$-stochastic.
-Since every string is $(\alpha,\alpha)$-stochastic for sufficiently large $\alpha$, this is a sequence of infinitely many distinct binary strings of increasing length.
-From \cite[Proposition~I.3 (b)]{gacs2001algorithmic} we know that for all sufficiently large $i$ there are strings of length less than $i$ that are not $(i/4, i/4)$-stochastic.
-Now pick any $i$ that is large enough that (a) $|x_i|<i$ and (b) $C^\K(x_i)/5 < i/4$.
-By definition of $(\alpha,\beta)$-stochasticity, if a string is not $(i/4,i/4)$-stochastic, then for every finite set $A$ containing $x$, either (a) $\log|A|\ge C^\K(x_i)+i/4\ge 5C^\K(x_i)/4$, or (b) $C^\K(A)>i/4>C^\K(x_i)/5$.
-In case (a), any two-part description using $A$ is substantially larger than the Kolmogorov complexity, for large enough $i$, so it does not determine the sophistication.
-Case (b) provides the required lower bound on the complexity of the model.
-
-Now, take $\T$ as model class and suppose towards contradiction that the sophistication is determined by a total function $f=\phi_j,j\in\T$ with $f(d)=x$ and $C^\K(f)\le i/5$.
-Then by Lemma~\ref{lemma:total-to-sets}, for some fixed constant $c$ there is a finite set $A$ with $\log|A|\le|d|\le C^\K(x_i)$, so (a) above cannot be the case, and $C^\K(A)\le C^\K(f)+C^\K(|d|)+c\le i/5+C^\K(|d|)+c$, which contradicts (b) for large enough $i$.
-Therefore such candidates $f$ cannot exist and all candidates must satisfy $C^\K(f)>i/5>C^\K(x_i)/6$.
-\end{proof}
-
-Note that we favored brevity of the proof over the strength of the bounds: the multiplicative constant $1/6$ can be improved with a more detailed treatment.
-
-\depth*
-\begin{proof}
-Let $U(\bar\imath y)$ be some (non-prefix) universal Turing machine, and let $U^A(\bar\imath y)$ be a simulation of that machine which outputs $0$ if the number of steps taken exceeds $A(|\bar\imath y|)$.
-Let $u$ be the index of the function $U^A$ in the standard enumeration.
-
-Let $D$ be a prefix function with $D(\epsilon) = u$.
-We can instantiate Lemma~\ref{lemma:thecoolone} with $D$, $\M' = \{\phi_u\}$ and $X = \{x : d^{\C,c}(x) \leq A(C(x))\}$.
-This tells us that there exists a numbering for which $\s^\T(x) = \s^{\M'}(x) + |\bar0| \leq c$ for all $x \in X$.
-\end{proof}
-
 \end{document}

### 2015-07-09 Facticity/Facticity/paper.tex
...

@@
 
-\author{Peter Bloem and Steven de Rooij}
+\author{\ldots}
 
-\institute{
-  System and Network Engineering Group, \\University of Amsterdam, the Netherlands\\
-  \email{uva@peterbloem.nl, steven.de.rooij@gmail.com}
-}
+%institute{
+%  System and Network Engineering Group, \\University of Amsterdam, the Netherlands\\
+%  \email{uva@peterbloem.nl, steven.de.rooij@gmail.com}
+%}
 
@@
 \enlargethispage{2\baselineskip}
-\footnotetext{Peter Bloem is a student author (PhD student).}
 
@@
 For technical reasons, we deviate slightly from the traditional notation of Kolmogorov complexity: we define the $\M$-Kolmogorov complexity of a binary string $x$, with respect to a model class $\M$ and a numbering $\psi$, as follows $C^{\M,\psi}(x\mid z)=\min\{\bar\imath y:\psi_i(y; z)=x,i\in\M\}$, with $C^{\M,\psi}(x) = C^{\M,\psi}(x\mid \epsilon)$.
-We may drop the numbering parameter when the distinction isn't relevant.
+We may drop the numbering parameter when the distinction is not relevant.
 Under this notation $C^\C(x)$ corresponds to the plain Kolmogorov complexity $C(x)$ and $C^\K(x)$ corresponds to the prefix-free version $K(x)$.
@@
 \subsubsection*{\ackname}
-
-This publication was supported by the Dutch national program COMMIT and by  the Netherlands eScience center. We thank Tom Sterkenburg for interesting discussions.
+\ldots
+\hide{This publication was supported by the Dutch national program COMMIT and by  the Netherlands eScience center. We thank Tom Sterkenburg for interesting discussions.}
 

### 2015-07-31 Facticity/Facticity/paper.tex
...

--- 2015-05-12 Facticity/Facticity/paper.tex
@@
 \newcommand{\range}[1]{\mathop{\tn{range}(#1)}}
+\newcommand{\Cc}{C_\tn{comp}}
 
@@
 
-\author{Peter Bloem and Steven de Rooij}
+\author{Peter Bloem, Steven de Rooij and Pieter Adriaans}
 
@@
   System and Network Engineering Group, \\University of Amsterdam, the Netherlands\\
-  \email{uva@peterbloem.nl, steven.de.rooij@gmail.com}
+  \email{uva@peterbloem.nl, steven.de.rooij@gmail.com, p.w.adriaans@uva.nl}
 }
@@
 
-\section{Introduction}
-\enlargethispage{2\baselineskip}
-\footnotetext{Peter Bloem is a student author (PhD student).}
+\section{Introduction}\enlargethispage{3\baselineskip}
 
@@
 The Kolmogorov complexity $C(x)$ of a binary string $x$ is, informally, the length of the shortest computer program to print $x$.
-This length depends on the choice of programming language, but, by the invariance theorem \cite[Section~2.1]{li1993introduction}, this choice affects the value only by a constant, independent of $x$.
+This length depends on the choice of programming language, but, by the invariance theorem \cite[Section~2.1]{li1993introduction}, only by a constant, independent of $x$.
 For sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure.
@@
 If sophistication is bounded, then knowing its value under one programming language provides no constraints on its value under another language (except it is also bounded).
-\item Similarly, there should be no constant $c$ such that $C(x)-S(x)\le c$ for all $x$, because then sophistication would be equivalent to Kolmogorov complexity. 
+\item Similarly, there should be no constant $c$ such that $\left | C(x)-S(x)\right |\le c$ for all $x$, because then sophistication would be equivalent to Kolmogorov complexity. 
 \end{enumerate}
@@
 Figure~\ref{fig:diagram} illustrates the principle.
-The various proposals define sophistication as the size of the model for one of the representations that approaches the Kolmogorov complexity.
-However, for most definitions, we can prove  they fail one of the conditions above.
-For certain variants, we cannot \emph{prove} they conflict with our requirements, but we show these methods only assign substantial sophistication to strings that require an enormous amount of processing to construct.
+The key to sophistication is to take the representations that come close to the Kolmogorov complexity, the \emph{candidates}, and define the sophistication as the size of the smallest model in this set.
+However, for most definitions, we can prove that they fail one of the conditions above.
+For others, we cannot \emph{prove} they conflict with our requirements, but we show these methods only assign substantial sophistication to strings that require an enormous amount of processing to construct.
 
 \begin{figure*}[tb]
+  \vspace{-2\baselineskip}
   \centering
@@
   \end{minipage}
-  \caption{\small (left) Two-part representations of $x$ by the two components of their codelength.
-The Kolmogorov complexity $C^\M(x)$, appearing as a black diagonal, provides a lower bound on the total codelength.
-We consider only representations that are close to this optimum, with the threshold represented by a dashed line.
+  \caption{\small (left) Two-part representations of $x$ by the two components of their code.
+The Kolmogorov complexity $C(x)$, appearing as a black diagonal, provides a lower bound on the total codelength.
+We consider only representations that are close to this optimum---called \emph{candidates}---with the threshold represented by a dashed line.
 The size of the smallest model below the threshold is the sophistication of the data.
@@
   \label{fig:diagram}
-\end{figure*}
+\end{figure*}\enlargethispage{3\baselineskip}
 
-A valid definition of $S$ must contend with two important issues.
+A valid definition of $S(x)$ must contend with two important issues.
 First, the details of the way the model is encoded are important.
@@
 
-Underfitting occurs when the selected model is simple, but fails to represent some of the structure present in the data.
+Underfitting occurs when the selected model is simple, but fails to capture all structure in the data.
 This is also a problem for sophistication because the models under consideration are so powerful.
 In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language.
-Such \emph{universal models} are \emph{simple}, since they can be described with a relatively small number of bits, yet are able to represent the data using a code within a constant from the Kolmogorov complexity.
+Such \emph{universal models} are \emph{simple}, since they can be described with a relatively small number of bits, yet are able to represent any data using a code within a constant from the Kolmogorov complexity.
 Such a two-part representation essentially encodes all information as noise.
@@
 
-Finally, In the discussion in Section~\ref{section:conclusion} we argue that while two-part coding can yield useful insights into the structure of the data and allows us to identify some models as poor representations, it is probably not possible to uniquely separate structure from noise and identify a \emph{single} model as ``best'': in general many models, of vastly different complexities, may be reasonable representations.
-Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that the data allows for multiple equivalent interpretations of which information is structured, and which is random, and that there is no such thing as sophistication.
+Finally, in Section~\ref{section:conclusion} we argue that while two-part coding can yield useful insights into the structure of the data and identifies some models as poor representations, it is probably not possible to objectively separate structure from noise and identify a \emph{single} model as ``best'': many models of different complexities may be reasonable representations.
+Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that the data allows for multiple, equivalent interpretations of which information is structured, and which is random, and that there is no such thing as sophistication.
 
-\enlargethispage{3\baselineskip} 
-\section{Notation}
+\section{Notation}\enlargethispage{3\baselineskip}
 The following notation allows us to generalize across all definitions and variants, save the occasional exception which we will highlight individually.
 
-Let $\B = \{0,1\}^*$.
-We deal with partial computable functions $f: \B \times \B \to \B$, which we will also call \emph{models}.
-A function is called \emph{prefix} if its domain, $\text{dom}_z(f)~=~\{y~:~f(y,~z)~\neq \infty\}$, is a prefix free set for all $z$, i.e.
+Let $\B = \{0,1\}^*$. We deal with partial computable functions $f: \B \times \B \to \B$, which we
+also call \emph{models}.
+$f$ is called \emph{prefix} if $\text{dom}_z(f)~=~\{y~:~f(y,~z)~\neq \infty\}$, is a prefix free set for all $z$, i.e.
 no string in $\text{dom}_z(f)$ is a prefix of another.
 A function $f$ is \emph{total} if $\forall_z \text{dom}_z(f) = \B$.
-In most cases, we do not use the second argument, and we adopt the convention that $f(x) = f(x, \epsilon)$.
+In most cases, we do not use the second argument, and let $f(x) = f(x, \epsilon)$.
 
-A \emph{numbering} is an enumeration of of the partial computable functions, denoted as $\psi_1, \psi_2, \ldots$ or simply $\psi$.
-We fix one canonical effective numbering $\phi$.
+A \emph{numbering} is an enumeration of the partial computable functions, denoted by $\psi_1, \psi_2, \ldots$ or simply $\psi$.
+We fix one canonical numbering $\phi$, chosen to be \emph{effective}: ie.
+given $i$ and $y$, we can effectively compute $\phi_i(y)$.
 We call a numbering $\psi$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ with $\forall: i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{a(i)}$.
 
-A \emph{model class} is a set of model indices. We distinguish four classes:
+A \emph{model class} is a set of indices in a numbering $\psi$. We define four classes:
 \begin{itemize} 
   \item The indices of the partial computable functions $\C=\N$.
-  \item The total functions $\T=\{i:\tn{$\phi_i$ is total}\}$. Note that $\T$ is not computably enumerable.
-  \item $\K$ is an enumerable set such that $\{\phi_i:i\in\K\}$ is the set of all partial computable prefix functions.
-  \item The finite sets: $\F$ is an enumerable set such that $\{\phi_i:i\in\F\}$ contains a uniform code for every finite set.\footnote{A uniform code for a set $S$ is a surjective prefix function $f:\{0,1\}^{\lceil\log|S|\rceil}\to S$.}
+  \item The total functions $\T=\{i:\tn{$\psi_i$ is total}\}$. Note that $\T$ is not computably enumerable.
+  \item $\K$ is an enumerable set such that $\{\psi_i:i\in\K\}$ is the set of all partial computable prefix functions.
+  \item The finite sets: $\F$ is an enumerable set such that $\{\psi_i:i\in\F\}$ is the set of uniform codes for all finite sets.\footnote{A uniform code for a set $F$ is a surjective prefix function $f:\{0,1\}^{\lceil\log|F|\rceil}\to F$.}
 \end{itemize}
-A model class is thus defined in reference to a specific numbering, indicated by the context.
 
-We sometimes need to map binary strings to a prefix-free set.
-We denote by $\br{x}$ the prefix-encoded representation for $x$.
-We require that the mapping satisfies $|\br{x}| = |x|+O(\log(|x|)$ (see eg.
+Let $\br{x}$ denote the prefix-encoded representation for $x$.
+We require that the mapping satisfies $|\br{x}| = |x|+O(\log|x|)$ (see eg.
 \cite[Section~1.4]{li1993introduction}).
@@
 
-For technical reasons, we deviate slightly from the traditional notation of Kolmogorov complexity: we define the $\M$-Kolmogorov complexity of a binary string $x$, with respect to a model class $\M$ and a numbering $\psi$, as follows $C^{\M,\psi}(x\mid z)=\min\{\bar\imath y:\psi_i(y; z)=x,i\in\M\}$, with $C^{\M,\psi}(x) = C^{\M,\psi}(x\mid \epsilon)$.
-We may drop the numbering parameter when the distinction isn't relevant.
-Under this notation $C^\C(x)$ corresponds to the plain Kolmogorov complexity $C(x)$ and $C^\K(x)$ corresponds to the prefix-free version $K(x)$.
-We can also represent the smallest two-part description of $x$ using a given model $\phi_i$ as $C^{\{i\}}(x)$.
-The complexity of a finite set is defined an the complexity of an explicit enumeration.
+For technical reasons, we deviate slightly from the traditional notation of Kolmogorov complexity: let $\M$ be a model class and $\psi$ an acceptable numbering, then let $C^{\M,\psi}(x\mid z)=\min\{|\bar\imath y|:\psi_i(y, z)=x,i\in\M\}$, with $C^{\M,\psi}(x) = C^{\M,\psi}(x\mid \epsilon)$.
+We omit the numbering when the distinction is not relevant.
+$C^\C(x)$ corresponds to the plain Kolmogorov complexity $C(x)$ and $C^\K(x)$ corresponds to the prefix-free version $K(x)$.
+Note that the notation $C^{\{i\}, \psi}(x)$ represents the smallest two-part description of $x$ using model $\psi_i$.
 
-We use a choice of numbering for the purpose that is normally served by the universal Turing machine.
-The choice is a matter of taste: a UTM determines a numbering, and vice versa.
-The reason that we prefer to work with numberings is that it highlights an important issue: while Kolmogorov complexity is invariant to the choice of numbering this property does not immediately carry over to sophistication: for some treatments, the result is highly dependent on the chosen numbering (as shown in the next section).
+We use the principle of a numbering for the purpose normally served by the universal Turing machine.
+We prefer to work with numberings as it highlights an important issue: while Kolmogorov complexity is invariant to the choice of numbering this property does not immediately carry over to sophistication: for some treatments, the result is highly dependent on the chosen numbering, as we will see in the next section.
 
@@
 
-The simplest approach to sophistication would be to `open up' the Kolmogorov complexity and to see which program on the universal Turing machine achieves the smallest description length: the program that \emph{witnesses} the Kolmogorov complexity.
-Since any UTM uses a two-part coding, its programs consist of a model and an input:
+The simplest approach to sophistication would be to `open up' the Kolmogorov complexity and to see which program achieves the smallest description length: the program that \emph{witnesses} the Kolmogorov complexity.
+This witness is a two-part coding; it consists of a model and an input.
 
@@
 \[
-\s_{\text{index}}^{\M,\Nm,\psi,c}(x) = \min\left\{ |i| : i \in \M, C^{\{i\}, \psi}(x) \leq C^{\Nm, \psi}(x)+c   \right\} \p
+\s_{\tn{index}}^{\M,\Nm,\psi,c}(x) = \min\left\{ |i| : C^{\{i\}, \psi}(x) \leq C^{\Nm, \psi}(x)+c,\;i \in \M\right\} \p
 \]
-When $\M = \Nm$, we will abbreviate with $\s^{\Nm,\psi,c}$. \label{definition:index}
+When $\M = \Nm$, we will use $\s_\tn{index}^{\M,\psi,c}$. If the set over which the minimum is taken is empty, the sophistication is undefined.\label{definition:index}
 \end{definition}
@@
 
-In \cite{antunes2009sophistication,antunes2013sophistication} Koppel's principle is limited to finite strings, with the total functions as a model class.
-The definition used is similar to $s_\text{index}^{\T, \C,\psi,c}$, with one exception: the total complexity of a witness $(i,y)$ is measured as $|i|+|y|$ without the cost of delimiting the two.
-This difference is not relevant to the point we make in this section.
-The restriction to total classes is a common approach, perhaps used because it helps to avoid underfitting, which we will discuss in the next section.
+In \cite{antunes2009sophistication,antunes2013sophistication} Koppel's principle is limited to finite strings, with $\T$ as a model class.
+The definition is similar to $S_\tn{index}^{\T, \C,\psi,c}$, except the total complexity of a witness $(i,y)$ is measured as $|i|+|y|$ without the cost of delimiting the two.
+This difference is not relevant to the current discussion.
+The restriction to $\T$ is a common approach, which avoids underfitting, as discussed in the next section.
 
-\begin{lemma}
-Let $\s^\psi$ denote any index sophistication with respect to numbering $\psi$.
-There are acceptable numberings $\psi$ and $\xi$ such that for all $x$:\\
-\-\hspace{2.8em} $|\s^\psi(x) - \s^\xi(x)|\geq \tfrac{1}{2}\min\{\s^\psi(x),\s^\xi(x)\}$.
+\begin{lemma}\label{lemma:perverse}
+Let $\s_\tn{index}^\psi$ denote any index sophistication with respect to numbering $\psi$ (with any choice for $\M$, $\Nm$ and $c$).
+There are acceptable numberings $\psi$ and $\xi$ such that for all $x$: $|\s_\tn{index}^\psi(x) - \s_\tn{index}^\xi(x)|\geq \tfrac{1}{2}\min\{\s_\tn{index}^\psi(x),\s_\tn{index}^\xi(x)\}$.
 \end{lemma}
 \begin{proof}
-Let $s_i$ be the string consisting of $2^{i}$ zeroes followed by a one.
-Define $\psi, \xi$ such that $\psi_j(x) = \phi_i(x)$ for $j = s_{2i}$ and $\xi_j(x) = \phi_i(x)$ for $j = s_{2i+1}$, with all other functions returning $\infty$ for all inputs.
+Let $z_i \in \B$ consist of $2^{i}-1$ zeroes followed by a one.
+Define $\psi, \xi$ such that $\psi_j(x) = \phi_i(x)$ for $j = z_{2i}$ and $\xi_j(x) = \phi_i(x)$ for $j = z_{2i+1}$, with all other functions returning $\infty$ for all inputs.
 Choose any $x$ and assume w.l.o.g.
-that $\s^\psi(x)\le s^\xi(x)$.
-By construction, we have $2\s^\psi(x)  \leq \s^\xi(x)$.
-\end{proof}
-Thus, the length of the index is a very poor indicator of model complexity.
-For a robust measure, define the complexity of functions as in \cite{grunwald2004shannon,vitanyi2004meaningful} by $C^{\M,\psi}(f) = \min\{C^{\M,\psi}(i):\phi_i=f\}$.
-Lemma~\ref{lemma:invariance} in the appendix shows that $C^\C(f)$ and $C^\K(f)$ are invariant: they are objective properties of a function.
+that $\s_\text{index}^\psi(x)\le \s_\text{index}^\xi(x)$.
+By construction, we have $2\s_\text{index}^\psi(x)  \leq \s_\text{index}^\xi(x)$.
+\end{proof}\enlargethispage{3\baselineskip}
+Thus, the length of the index is a very poor indicator of model complexity. For a robust measure, we define the complexity of a function $f$ as in \cite{grunwald2004shannon,vitanyi2004meaningful} by 
+\begin{align}
+C^{\M,\psi}(f) = \min\{C^{\M,\psi}(i):\psi_i=f\} \p \label{equation:function-complexity}
+\end{align}
+Lemma~\ref{lemma:invariance} in the appendix shows that $C^\C(f)$ and $C^\K(f)$ are invariant.
+Note that the perversely inefficient numberings of Lemma~\ref{lemma:perverse} are no issue for Kolmogorov complexity: we can use a UTM with a more efficient numbering as a model at only a constant penalty.
+For sophistication, however, the numbering is crucial.
 
-Note that such perversely inefficient numberings are not an issue for Kolmogorov complexity.
-If our UTM uses an inefficient numbering, we can switch to a more efficient one at only a constant cost, keeping the complexity the same, up to a constant.
-For sophistication, however the choice of numbering is crucial.
+There are two ways to use $C^\M(f)$ for more robust attempts to define sophistication.
+Confusingly, both are used in the literature.
+First, we can measure the complexity of the \emph{model} $\phi_i$ as $C^\K(\phi_i)$, which is then the size of the first part of a two-part code describing the data.
+This approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information}.
 
-There are two ways to use the model complexity $C^\M(\phi_i)$ for more robust attempts to define sophistication.
-Confusingly, both are used in the literature.
-First, we can measure the complexity of the \emph{model} using $C^{\M,\psi}(\phi_i)$, which is then the size of the first part of a two-part code describing the data.
-For instance, this approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information} and in most of the present paper.
+Second, we can stick to using the length of the index as the measure of sophistication, but restrict the allowed numberings to those that can represent a given function \emph{efficiently}.
+This approach is taken by Adriaans in \cite{adriaans2012facticity}, who defines \emph{facticity} as $\s_\text{index}^{\C,\psi,0}$, but only allows \emph{faithful} numberings.
+Formally, a faithful numbering has the property that $\forall i \exists j : \psi_i=\psi_j, |j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
+Essentially, this means that a faithful numbering can represent a function $f$ with an index the same length as the Kolmogorov complexity $C^\C(f)$.
 
-Second, we can stick to using the length of the index as the measure of sophistication, but restrict the allowed numberings.
-This approach is taken by Adriaans in \cite{adriaans2012facticity}, which defines \emph{facticity} as $\s_\text{index}^{\C,\psi,0}$, but only allows numberings that are \emph{faithful}: $\forall i \exists j : \psi_i=\psi_j, |j|\le C^{\C}(\psi_j)+c$, for some constant $c$.
+We prove in the appendix that---contrary to Adriaans' suggestion---there do exist faithful, acceptable numberings (Lemma~\ref{lemma:faithful-numberings}).
+However, even choosing a faithful numbering is not enough.
+The Kolmogorov complexity uses representations of the form $\bar\imath y$, with $\psi_i(y) = x$,  where the bar denotes some straightforward prefix encoding to delimit the model description $i$ from its input $y$.
+If we define a second prefix encoding $\tilde{\imath}$, with $|\bar\imath|-|\tilde\imath|$ unbounded, we can define a second representation $\bar u \tilde \imath y$, with $\psi_u(\tilde{\imath} y) = \psi_i(y)$, at a constant overhead $|\br{u}|$, and gain more than $|\br{u}|$ for sufficiently complex strings, resulting again in a bounded sophistication.
 
-We prove in the appendix that---contrary to Adriaans' suggestion---there actually do exist faithful, acceptable numberings (Lemma~\ref{lemma:faithful-numberings}).
-However even choosing a faithful index is not quite enough.
-The Kolmogorov complexity uses representations of the form $\bar\imath y$, where the bar denotes some straightforward prefix encoding to delimit the model description $i$ from its input $y$.
-If we define a second prefix encoding $\tilde{\imath}$, with $|\bar\imath|-|\tilde\imath|$ unbounded, we can define a second representation $\bar u \tilde \imath y$, at a constant overhead $|\br{u}|$, and gain more than $|\br{u}|$ for sufficiently complex strings, resulting in a bounded sophistication.
-
-We continue with a sophistication that avoids these issues. Let $C'^{\M, \psi} = C^\K(\psi_i) + min\{|y| : \psi_i(y) = x\}$. We define sophistication as:
+We continue with a sophistication that avoids the issues of inefficient indices and of inefficient prefix encodings.
+We change the definition of index sophistication so that its two-part representations use $C^\K(\phi_i)$ bits for the representation of the model.
+We first introduce the following notation for the $\M$-Kolmogorov complexity using such compact two-part representations:
+\[
+\Cc^{\M, \psi}(x) = \min\{ C^{\K, \psi}(\psi_i) + |y| :  \psi_i(y) = x, i \in \M\}\p
+\]
+For model classes $\K$ and $\C$ this is equivalent to the existing definition and invariant to the numbering.
+Note that again, we use $\Cc^{\{i\}, \psi}(x)$ to represent the smallest two-part code using model $\psi_i$.
 
@@
 \[
-\s^{\M,\Nm,\psi,c}(x)=\min\left\{C^\K(\phi_i):i\in\M,C'^{\{i\},\psi}(x)\le C^{\Nm,\psi}(x)+c\right\}.
+S^{\M,\Nm,\psi,c}(x)=\min\left\{C^\K(\phi_i): \Cc^{\{i\}, \psi} \le \Cc^{\Nm,\psi}(x)+c, i\in\M\right\}.
 \]
@@
 
-In the previous section, we saw the first glimpse of the delicate balance between the two code components in sophistication.
-We will investigate this balance, starting with the variant $\s^{\K,\psi,c}$.
-This variant itself is not used anywhere in the literature, but it helps to illustrate the issues we wish to discuss.
+In the last section, we began to see the delicate balance between the two code components.
+We will study this balance, starting with the variant $S^{\K,\psi,c}$, which is not used in the literature, but helps to illustrate the issues we wish to discuss.
 
-$\K$ has representations with all information except a constant in the input and representations with all information in the model.
-The downside to this balance is that it becomes easy to show a lack of invariance: by tweaking the numbering so that models in a specific subset $\M'\subset\M$ become cheaper to represent by an arbitrary amount relative to others, so that a model in $\M'$ always determines the sophistication.
-
-This means that for model classes with a universal model, we can choose our numbering to make $S^\K$ either bounded, or equal to the complexity:
+$\K$ has optimal representations with all but a constant part of the information in the input and it has optimal representations with all information in the model.
+The downside to this balance is that it becomes easy to show a lack of invariance.
+We can tweak the numbering so that models in a specific subset $\M'\subset\K$ become cheaper to represent by an arbitrary amount relative to others: we can ensure that a model in $\M'$ always determines the sophistication.
+For instance, if we let $\M'$ contain only a universal model we get a bounded sophistication.
 
 \begin{theorem}[Underfitting]
-Let $\M$ be a model class containing a universal model $\phi_u$, with the property that $\exists c \forall i \in \M, x : C'^{\{u\}}(x) \leq C'^{\{i\}}(x) + c$.
-Then, for some numberings $\s^{\M,\psi,c}$ is bounded.
+Let $\M, \Nm$ be model classes with $\M \subseteq \Nm$ and let $\M$ contain a universal model $\phi_u$, with the property that $\exists c \forall i \in \Nm, x \in \B : \Cc^{\{u\},\phi}(x) \leq \Cc^{\Nm, \phi}(x)+ c$.
+Then, for some numbering $\psi$, $S^{\M,\Nm, \psi,c}$ is bounded.
 \label{theorem:underfitting}
 \end{theorem}
-
-This result is well known and many treatments avoid it by restricting the model class.
-Less well known is perhaps that the same holds in the other direction: for some numberings the singletons always determine the sophistication.
+This problem is well known and many treatments avoid it by restricting the model class.
+Less well known, perhaps, is that the same holds in the other direction:  if $\M'$ is the set of \emph{singleton} models---those models that output a single $x$ for an empty input---we get a sophistication equal to the Kolmogorov complexity.
 
 \begin{theorem}[Overfitting]
-Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $i\in\M$ with $\phi_i(\epsilon)=x$.
-Then there is a numbering $\psi$, and a constant $c$, such that for all $x\in\X$ we have $C^\M(x)-S^{\M,\psi,c}(x)\leq c$.\label{theorem:overfitting}
+Let $\X \subseteq \B$.
+Let $\M \subseteq \Nm \subseteq \K$ be model classes where for every $x\in\X$ there is a singleton model $i\in\M$ with $\phi_i(\epsilon)=x$.
+Then there is a numbering $\psi$, and a constant $c$, such that for all $x\in\X$ we have $C^\K(x)-S^{\M,\Nm,\psi,c}(x) \leq c$.\label{theorem:overfitting}
 \end{theorem}
-The proofs of both theorems rely on a simple principle: for certain $\M' \subset \M$, there exist numberings which have the effect of penalizing $C^\K(\phi_i)$ for any model outside $\M'$ by an arbitrary constant amount.
+The proofs of both theorems rely on a simple principle: there exist numberings which have the effect of penalizing $C^\K(\phi_i)$ for any model outside $\M'$ by an arbitrary constant amount.
 We can use this to effectively `push' these models outside of the range of candidates, ensuring that, under this numbering, a model in $\M'$ always determines the sophistication.
-
-For these theorems, this means setting $\M' = \{u\}$, with $\phi_u$ a universal model, or letting $\M'$ be the singletons.
-The requirements for $\M'$ are strict, and slightly complex.
+The requirements for $\M'$ are somewhat complex.
 The following lemma gives a set of sufficient conditions.
 
-\begin{restatable}{lemma}{coolone}
+\begin{lemma}
 \label{lemma:thecoolone}
-  Let $\M$ be any model class, let $\X$ be any set of binary strings and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\phi$.
+Let $\M$ and $\Nm$ be any model class, let $\X$ be any set of binary strings and let $D:\B\to\N$ be a partial computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\phi$.
 Let $\M'= \tn{range}(D)$.
 Further assume there is a constant $c$ such that:\\
-\-\hspace{1cm}(1) $\forall_{m\in\M'}:\min\{|p|:\phi_{D(p)}=\phi_m\}\le C^{\K,\phi}(\phi_m)+c$\\
-\-\hspace{1cm}(2) $\forall_{x\in\X}:C'^{\M',\phi}(x)-C'^{\M,\phi}(x)\le c$.\\
-Then there is an enumeration $\psi$ of the partial recursive functions such that $S^{\M,\psi,c}(x) = |\br{0}|+S^{\M',\phi,c}(x)$ for all $x\in\X$.
-\end{restatable}
+\-\hspace{1cm}(a) $\forall_{m\in\M'}:\min\{|p|:\phi_{D(p)}=\phi_m\}\le C^{\K,\phi}(\phi_m)+c$\\
+\-\hspace{1cm}(b) $\forall_{x\in\X}: \Cc^{\M', \phi}(x) - \Cc^{\Nm, \phi}(x) \le c$.\\
+Then, there is a $\psi$ such that if $S^{\M,\Nm,\psi,k}(x)$ is defined, then $S^{\M,\Nm,\psi,k}(x) = S^{\M',\Nm,\psi,k}(x)$ up to a constant. 
+\end{lemma}
+\begin{proof}\enlargethispage{2\baselineskip}
+Pick any $x \in \X$.
+Let $f$ and $g$ be $\phi$-indices such that $f \in \M'$ and $g \notin \M'$ nor is $\phi_g$ equivalent to any function indexed by $\M'$.
+Furthermore let $\Cc^{\{f\}, \phi}(x)$ and $\Cc^{\{g\}, \phi}(x)$ both be within a constant $q$ of $\Cc^{\Nm,\phi}(x)$.
+Assumption (b) ensures that $\M'$ always provides such an $f$.
 
-\begin{proof}
-We define the numbering $\psi$ as follows:
+We will show that for every integer $r$, there is a numbering $\psi$ such that $\Cc^{\{g'\}, \psi}(x) - \Cc^{\{f'\}, \psi}(x) \geq r$ for all $x \in \X$, where $f'$ and $g'$ are the $\psi$-indices equivalent to $f$ and $g$.
+Thus, for large enough $r$, $\phi_g$ is eliminated as a candidate model, while $\phi_f$ remains in place.
+Thus, under $\psi$, a member of $\M'$ determines the sophistication, or the sophistication is undefined.
+
+Let $d$ be a positive constant. We will show later how to choose it to achieve the required result. We define $\psi$ as follows:
 \[
-\psi_0(p) = 1^r 0 D(p), \;\;\;
-\psi_{1^r0i}(p) = \phi_i(p), \;\;\;
-\psi_j(\cdot) = \infty \;\text{if $j$ contains no zeroes.}
+\psi_0(p) = 0^d 1 D(p), \;\;\;
+\psi_{0^d1i}(p) = \phi_i(p), \;\;\;
+\psi_j(\cdot) = \infty \;\text{if $j\neq 0$ and $j \neq 0^d1\ldots$}
 \]
-We will show that under the $\psi$, the best representation for $x$ with $m\in\M'$ is always better than the best using some $m\notin\M'$. First suppose $m\in\M'$. Then
+The key to the proof is the way that the function complexity $C^\K(\cdot)$ changes when we change the numbering from $\phi$ to $\psi$.
+For $f$, the value increases by no more than a fixed constant, but for $g$, it increases by a constant that we can arbitrarily increase by increasing $d$.
+
+We will first show that for $f$, the value does not increase by more than a constant $c_f$. Assume w.l.o.g. that $0 \in \K$.
 \begin{align*}
-   C^{\K,\psi}(\phi_m) &=\min\{|\bar\jmath q|:\psi_{\psi_j(q)}=\phi_m\} \leq\min\{|\bar0 q|:\psi_{\psi_0(q)}=\phi_mm\} \\ 
-   &=\min\{|\bar0 q|:\psi_{1^r0 D(q)}=\phi_m\} =\min\{|q|:\phi_{D(q)}=\phi_m\} + |\bar 0|\\
-   &\leq C^{\K,\phi}(\phi_m)+c+|\br{0}|,
+C^{\K, \psi}(\phi_f) &= \min\left\{|\bar \jmath q| : \psi_{\psi_j(q)} = \phi_f, j \in \K \right\} & \text{rewriting (\ref{equation:function-complexity})} \\
+&\leq \min \left\{ |\overline{0} q| : \psi_{\psi_0(q)} = \phi_f \right\}& \text{choose $j=0$} \\
+&= \min \left\{ |q| : \phi_{D(q)} = \phi_f \right\} + |\overline{0}| \\
+&\leq C^{\K, \phi}(\phi_f) + c_f & \text{by assumption (a).}
 \end{align*}
-where the last inequality uses assumption (1).
+In order to show that for $g$, we can increase the difference by an arbitrary constant, we first show that, for any $z$ not in the range of $\psi_0$, the Kolmogorov complexity itself increases by at least $d$ when we switch from $\phi$ to $\psi$:
+\begin{align}
+C^{\K, \psi}(z) &= \min\left\{|\bi y| : \psi_i(y) = z, i \in \K \right\} & \text{by definition}\nonumber\\
+&= \min\left\{|\overline{0^d1j} y| : \psi_{0^d1j}(y) = z \right\} & \text{since $z\notin \text{range}(\psi_0)$} \nonumber\\
+&\geq \min\left\{|\bar \jmath y| : \phi_{j}(y) = z \right\} + d &\nonumber\\
+&= C^{\K, \phi}(z) + d\p & \label{line:kolmogorov-switch}
+\end{align}
+We now show the increase in model complexity for $g$. First, assume $\phi_g \neq \psi_0$:
+\begin{align*}
+C^{\K,\psi}(\phi_g) &= \min \left\{C^{\K, \psi}(i) : \psi_i = \phi_g \right\} \\
+&= \min \left\{C^{\K, \psi}(0^d1j) : \phi_j = \phi_g \right\} \\
+&= C^{\K, \psi}(0^d1j) \\
+&\geq C^{\K,\phi}(0^d1j) + d & \text{by (\ref{line:kolmogorov-switch})} \\
+&\geq C^{\K,\phi}(j) - c_g + d &\text{since $C^\K(j) \leq C^\K(0^d1j) + c_0$}\\
+&\geq C^{\K, \phi}(\phi_g) - c_g + d \p
+\end{align*}
+Now assume $\phi_g = \psi_0$.
+We have $C^{\K,\psi}(\phi_g) = \min\left\{C^{\K, \psi}(i) : \psi_i = \psi_0 \right\} \geq d$.
+This follows from the fact that the minimum is achieved either at $i = 0$ or at $i = 0^d1m$ with $m \notin \M'$.
+Neither have a representation using a function with a $\psi$-index without the $0^d1$ prefix.
 
-Now assume that the best model $m$ for $x$ is not in $\M'$. 
-Let $i$ be the index of $m$ with the shortest description, i.e. it achieves the minimum in $C^{\K, \psi}(\phi_m)=\min\left\{C^{\K,\psi}(i):\psi_i=\phi_m\right\}$.
-There are two possibilities.
-Either $i = 0$, in which case we have $C^{\K,\psi}(m)\ge r$ because $\phi_0$ cannot output zero and all other $\psi$-programs are at least $r$ bits long.
-Otherwise, we can bound
+Choosing $d \geq r + \max \left\{C^{\K,\phi}(\psi_0), c_g \right\} + c_f + 2q$ ensures that for both cases, we have $C^{\K, \psi}(\phi_g) \geq C^{\K, \phi}(\phi_g) + r + c_f + 2q$.
+While $C^\K(\psi_0)$ depends on the choice of $d$, we have $C^\K(\psi_0) \leq C^\K(d) + C^\K(D)$, up to a constant, which is in $O(\log d)$, so we can choose $d$ to satisfy the inequality.
+
+Finally, we can show the result: \belowdisplayskip=-26pt
 \begin{align*}
-C^{\K,\psi}(\phi_m)&=C^{\K,\psi}(m)=C^{\K,\psi}(1^r0j)\\
-&\ge C^{\K,\psi}(j)-c' \ge C^{\K,\phi}(j)+r+1-c'\\
-&=C^{\K,\phi}(m)+r+1-c'.
+&\Cc^{\{g'\}, \psi}(x) - \Cc^{\{f'\}, \psi}(x)\\
+&=C^{\K, \psi}(\phi_g) + \min\{|y| : \phi_g(y) = x\} - C^{\K, \psi}(\phi_f) - \min\{|y| : \phi_f(y) = x\}\\
+&\geq C^{\K, \phi}(\phi_g) + r + c_f + 2q + \min\{|y| : \phi_g(y) = x\} \\
+&\quad\quad- C^{\K, \phi}(\phi_f) - c_f - \min\{|y| : \phi_f(y) = x\} \\
+&= \Cc^{\{g\}, \phi}(x) - \Cc^{\{f\}, \phi}(x) + r + 2q\geq r \p\\
 \end{align*}
-Now choose $r=c''+\max\left\{C^{\K,\psi}(\phi_0),c'-1\right\}$. Then substitution yields, for both cases, 
-$C^{\K,\phi}(m)\ge c''+C^{\K,\psi}(m)$. Combining the inequalities above, for any model $\phi_g$ with $g\not\in\M'$, there is a model $\phi_f$ with$f\in\M$ such that
-\[\begin{split}
-C^{\K,\psi}(\phi_g)+&\min \{|y| : \phi_g(y) = x\} \ge C^{\K,^\phi}(\phi_g)+\min\{|y| : \phi_g(y) = x\}+c''\\
-&\ge C^{\K,\psi}(\phi_f)+\min \{|y| : \phi_f(y) = x\} -2c+c'' -|\bar0|.
-\end{split}\]
-A sufficiently large $c''$ ensures the best representation is in $\M'$ for all $x\in\X$.
 \end{proof}
-
 Theorems~\ref{theorem:underfitting} and \ref{theorem:overfitting} follow as corollaries. For Theorem~\ref{theorem:underfitting}:
 \begin{proof}
-Let $D$ be a prefix function as in Lemma~\ref{lemma:thecoolone} such that it returns the index of $u$ for the argument $\epsilon$ and $\infty$ for any other argument.
+Let $D$ be a prefix function as in Lemma~\ref{lemma:thecoolone} that returns the index of $u$ for the argument $\epsilon$ and $\infty$ for any other argument.
 That is, $\M' = \{u\}$.
-This construction satisfies the conditions 1 and 2 from Lemma~\ref{lemma:thecoolone}; invoking it we find that there exists an acceptable numbering $\psi$ for which $\s^{\M,\psi}(x) = \s^{\M', \phi, c}(x) + c$.
-Since $\M'$ contains only a single model, $\s^{\M',\phi, c}(x)$ is constant.
+This construction satisfies the conditions 1 and 2 from Lemma~\ref{lemma:thecoolone}.
+Invoking it, we find that there exists an acceptable numbering $\psi$ for which $S^{\M,\Nm,\psi,k}(x) = S^{\M',\Nm,\phi,k}(x) + c$.
+Since $\M'$ contains only a single model, $S^{\M',\Nm,\phi, c}(x)$ is constant.
 \end{proof}
-
 And for Theorem~\ref{theorem:overfitting}:
@@
 \begin{proof}
-Let $\phi_i$ be a singleton for $x$ as described.
-Note that since $\phi_i$ is a prefix function, since $\phi_i$ is defined for input $\epsilon$ then it cannot be defined for any other input.
-Pick $i,x$ with $\phi_i(\epsilon)=x$.
-Note that $x$ can be computed from $i$ and a fixed program, so there is a $c$ such that $C^{\K, \phi}(x)\le C^{\K, \phi}(\phi_i)+c$.
-Vice versa, given any $x$ we can construct an index of $\phi_i$, since $\phi$ is an acceptable numbering.
-Therefore $|C^{\K,\phi}(\phi_i)-C^{\K, \phi}(x)|\le c$.
+Let $x$ be any string.
+Given a description of $x$, we construct some index $i$ such that $\phi_i(\epsilon) = x$ (a singleton for $x$).
+Thus, $C^{\K, \psi}(\phi_i) \leq C^{\K,\psi}(x)$ up to a constant.
+Likewise, given $\phi$ we can produce $x$, so that $|C^{\K,\phi}(\phi_i)-C^{\K, \phi}(x)|\le c$ for some constant $c$.
 
-We now define a computable function $D$ by $D(\bar\imath y)=j$ where $\psi_j(\epsilon) = \phi_i(y)$.
+We now define a computable function $D$ by $D(\bar\imath y)=j$ where $\phi_j(\epsilon) = \phi_i(y)$ and $i \in \K$, and let $\M'$ be its range.
 We will show that the two conditions of Lemma~\ref{lemma:thecoolone} hold for the prefix function $D$.
 
-(1) Let $f$ be any function in the range of $D$, and $x$ its output.
-Then $\min\{|p|:\psi_{D(p)}=f\}=\min\{|\bar\imath q|:\phi_i(q)=x\}=C^{\K}(x) \le C^{\K}(f)+c$.
-(2) On the one hand $C^{\M',\psi}(x)\le C^\K(f)+|\epsilon|\le C^\K(x)+c$.
-On the other hand, $C^{\M,\psi}(x)$ is an effective description of $x$, so $C^{\K}(x)$ is at most a constant larger.
-Together, these inequalities establish the second condition.
+(a) Let $f \in \M'$ with $\phi_f(\epsilon) = x$.
+Then $\min\{|p|:\psi_{D(p)}=\phi_f\}=\min\{|\bar\imath q|:\phi_i(q)=x\}=C^{\K}(x) \le C^{\K}(\phi_f)+c$.
+(b) On the one hand $\Cc^{\M',\psi}(x)\le C^\K(\phi_f)+|\epsilon|\le C^\K(x)+c$.
+On the other hand, the witness to $\Cc^{\M,\psi}(x)$ is an effective description of $x$, so $C^{\K}(x)$ is at most a constant larger.
 
-Now, by Lemma~\ref{lemma:thecoolone} there is a numbering $\psi_1,\psi_2,\ldots$ such that $\s^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$.
-We observed that $|C^{\K}(f)-C^{\K}(x)|\le c_0$ for any $f\in\M'$, so $\s^{\M',\psi, c}(x)\ge C^{\K}(x)-c_0$.
-This proves the lemma.
+Now, by Lemma~\ref{lemma:thecoolone} there is a numbering $\psi$ such that we have $S^{\M,\Nm,\psi,k}(x) =S^{\M',\Nm,\psi,k}(x) + c$.
+We observed that $|C^{\K}(\phi_i)-C^{\K}(x)|\le c_0$ for all singletons, so $S^{\M',\Nm,\psi, k}(x)\ge C^{\K,\psi}(x)-c_0$.
+This proves the theorem.
 \end{proof}
@@
 We will now review the treatments in the literature that show overfitting.
-The first is the structure function, proposed by Kolmogorov: most likely the first attempt at separating structure from noise in an objective manner.
-Kolmogorov defined the function
+The first is the structure function, proposed by Kolmogorov, most likely the first attempt at separating structure from noise in an objective manner.
+Kolmogorov defined the following function, using the finite sets $\F$ as models:
 \[
-h_x(\alpha) = \min \left \{\log |S| : S \ni x, C^\K(s) \leq \alpha \right \} 
+h_x(\alpha) = \min \left \{\log |F| : x \in F, C^\K(F) \leq \alpha \right \} 
 \]
-and suggested that the smallest set for which $C^\K(S) + \log|S| \leq C^\K(S) + c$ for some pre-chosen constant $c$ can be seen as capturing all the structure in $x$ \cite{cover1985kolmogorov}.
-This is equivalent to the sophistication $\s^{\F,\K,\psi,c}$.
+and suggested that the smallest set for which $C^\K(F) + \log|F| \leq C^\K(x) + c$ holds for some pre-chosen constant $c$, can be seen as capturing all the structure in $x$ \cite{cover1985kolmogorov}.
+This is equivalent to the sophistication $S^{\F,\K,\psi,c}(x)$.
 Theorem~\ref{theorem:overfitting} shows there are numberings for which this sophistication is always equal to $C^\K(x)$.
@@
 
-In \cite{gacs2001algorithmic} the idea of the structure function is extended to the notion of an \emph{algorithmic sufficient statistic}.
-This is essentially the witness to the sophistication $\s^{\F,\K,\psi,c}(x)$.
-Thus, by Lemma~\ref{theorem:overfitting}, there exist numberings such that for all $x$, $\{x\}$ is the minimal sufficient statistic.
+In \cite{gacs2001algorithmic} the structure function is extended to an \emph{algorithmic sufficient statistic}.
+This is, again, essentially the witness to the sophistication $S^{\F,\K,\psi,c}(x)$.
+A probabilistic version is also introduced, which uses the model class $\mathcal P$, which indexes the set of functions that compute computable probability semimeasures up to a multiplicative constant error, yielding $S^{{\mathcal P}, \K, \psi, c}(x)$.
+For both, Lemma~\ref{theorem:overfitting} gives us a numbering such that the singleton is always the minimal sufficient statistic.
 
-It may be argued that the slack parameter $c$ in the sophistication, which determines the allowed gap between a candidate representation and the complexity, should be allowed to depend on the numbering, but this dependence has not been mentioned in the literature and there is no clear way to determine how this constant should be chosen on the basis of a given numbering.
+It may be argued that the slack parameter $c$ in the sophistication, which determines the allowed gap between a candidate representation and the complexity, should depend on the numbering, but this dependence has not been mentioned in the literature and there is no obvious method to choose this constant for a given numbering.
 
-In traditional statistics, overfitting is often addressed by imposing a penalty on complex models.
-As we have seen, a strong penalty, such as the one imposed by an inefficient prefix encoding of the model, will immediately cause underfitting.
-A more subtle approach is to allow descriptions that are not self delimiting.
-The gap between self-delimiting and non self-delimiting descriptions grows without bound
-\cite[Section~4.5.5]{li1993introduction}, so that for sufficiently complex inputs, some information will have to be placed in the noise part of the code, since placing all information in the model results in a self-delimiting description.
+In traditional statistics, overfitting is often addressed by a penalty on complex models.
+As we have seen, a strong penalty, such as the one imposed by an inefficient prefix encoding of the model, will cause underfitting.
+A more subtle approach is to allow descriptions that are not self-delimiting.
+The gap between the smallest self-delimiting description and the smallest non self-delimiting description grows without bound
+\cite[Section~4.5.5]{li1993introduction}, so that some information ends up in the noise, since placing all information in the model results in a self-delimiting, and thus non-optimal description.
 This eliminates the singletons as viable candidates.
 This approach is taken by Vit\'anyi \cite{vitanyi2004meaningful} and by Adriaans \cite{adriaans2012facticity}.
-Such measures thus help reduce the overfitting problem, but they only increase the tendency to underfit.
-We also pay the price that the models can no longer be equated with probability measures, weakening the link to traditional statistics and probability theory.
+Such measures reduce the overfitting problem, but they only increase the tendency to underfit.
+We also pay the price that the models can no longer be equated with probability measures, weakening the link to traditional statistics.
 
@@
 
-The existence of universal models is a widely acknowledged problem for sophistication, and most proposals avoid it by limiting the class of allowed models to exclude them.
-It is known that there are strings for which $\s^{F, K, \psi, c}$, $\s^{T, \psi, c}$, $\s^{T, \K, \psi, c}$ are close to the length of the string (up to a logarithmic term).
-These are the so-called absolutely non-stochastic strings \cite{shen1983concept}.
-Proofs for the sophistications mentioned can be found in \cite{gacs2001algorithmic}, \cite{antunes2009sophistication} and \cite{vitanyi2004meaningful} respectively.
-The existence of these strings is independent of the numbering chosen.
+Universal models are a widely acknowledged problem for sophistication, and most proposals avoid them by limiting the allowed models to exclude them.
+It is known that there are strings $x$ for which $S^{\F, \K, \psi, c}(x)$, $S^{\T, \psi, c}(x)$ and $S^{\T, \K, \psi, c}(x)$ are close to $|x|$ (up to a logarithmic term).
+Proofs can be found in \cite{gacs2001algorithmic}, \cite{antunes2009sophistication} and \cite{vitanyi2004meaningful} respectively.
+These are the \emph{absolutely non-stochastic strings} \cite{shen1983concept}.
+The existence of these strings is independent of the numbering.
 
@@
 Only one model class eliminates both the singletons and the universal model: $\T$.
-The only proposal we are aware of that uses an efficient model representation \emph{and} excludes the universal models \emph{and} excludes the singletons is \cite{vitanyi2004meaningful}, where $\s^{\T,\K,\psi,c}$ is used.
-While this invalidates straightforward proofs of boundedness, there is no evidence that $S^{\T,\K,\psi, c}(x)$ is actually invariant either.
+The only proposal we are aware of that uses an efficient model representation \emph{and} excludes the universal models \emph{and} excludes the singletons is: $S^{\T,\K,\psi,c}$, from \cite{vitanyi2004meaningful}.
+While this avoids our proofs of boundedness, there is no evidence that $S^{\T,\K,\psi, c}$ is actually invariant.
 
-While under $S^{\T,\K,\psi, c}(x)$ there are strings with high sophistication, these may not conform to sophistication's motivating intuition. To show this, we need the concept of depth:
+While high sophistication strings exist for $S^{\T,\K,\psi, c}$, they may not conform to sophistication's motivating intuition. To show this, we use the concept of depth:
 
@@
 \begin{theorem}
-Let $A(n)$ be the single-argument Ackermann function and $c$ some arbitrary constant.
-There are numberings such that for strings with depth $d^{\C,c}(x) \leq A(C^\C(x))$ the sophistication $\s^{\T, \K, \psi, c}$ is bounded.\label{thm:depth}
+Let $A(n)$ be the single-argument Ackermann function and $c_d$ some constant.
+For all $k$, there is a numbering $\psi$ such that for all strings with depth $d^{\C,c_d}(x) \leq A(C^\C(x))$ the sophistication $S^{\T, \K, \psi, k}(x)$ is bounded.\label{thm:depth}
 \end{theorem}
@@
 
-Let $D$ be a prefix function with $D(\epsilon) = u$.
-We can instantiate Lemma~\ref{lemma:thecoolone} with $D$, $\M' = \{\phi_u\}$ and $X = \{x : d^{\C,c}(x) \leq A(C(x))\}$.
-This tells us that there exists a numbering for which $\s^{\T, \K,\psi,c}(x) = \s^{\M', \psi, c}(x) + |\bar0| \leq c$ for all $x \in X$.
+Let $D(\epsilon) = u$.
+We can instantiate Lemma~\ref{lemma:thecoolone} with $D$, $\M' = \{u\}$ and $\X = \{x : d^{\C,c_d}(x) \leq A(C^\C(x))\}$.
+This tells us that there exists a numbering $\psi$ for which $S^{\T, \K,\psi,k}(x) = S^{\M', \K, \psi, k}(x) + |\bar0| \leq c$ for all $x \in \X$.
 \end{proof}
@@
 Consider a string that is typical for a shallow model, say some elaborate probabilistic automaton.
-Under $\s^{\T,\K, \psi,c}$, no matter how high the complexity of the automaton, the sophistication is bounded.
-We could encode the collected works of Shakespeare in its transition graph, and this information would be counted as noise by the sophistication.
+Under $S^{\T,\K, \psi,c}$, no matter how high the complexity of the automaton, the sophistication is bounded.
+We could encode the collected works of Shakespeare in its transition graph, and this information would be counted as noise.
 Any structure simple enough to be exploited within the time bound of the Ackermann function will not be seen as `meaningful information'.
 Only structure so deep that it would take beyond the lifetime of the universe to decompress would count towards sophistication.
-In the remainder we will refer to strings $x$ with $d^{\C,c}(x) \leq A(C(x))$ as \emph{non-deep} strings.
+In the remainder we will refer to strings $x$ with $d^{\C,c}(x) \leq A(C^\C(x))$ as \emph{shallow} strings.
+Note that any string whose shortest program can be run in any time bound represented by a primitive recursive function is shallow.
 
@@
 This contradicts the intuition that sophistication measures structure, as it seems to suggest that all strings we can possibly hope to understand or generate contain no structure, save a constant amount.
-The alternative is that under other numberings these strings \emph{would} have structure, but then the sophistication would no longer be invariant.
+The alternative is that under other numberings these strings \emph{do} have structure, but then the sophistication is not invariant.
 
@@
 We will define a generic version, parametrized by model class.
+Let $C_{\psi_i}(x) = \min\left\{|y| : \psi_i(y) = x\right\}$.
+Then we define the \emph{naive sophistication} as:
+\[
+\s_\tn{naive}^{\M, \psi, c}(x) = \min\left\{C^{\K, \psi}(\psi_i) : C_{\psi_i}(x) - C^{\K,\psi}(x\mid i) \leq c, i \in \M \right\} \p
+\]
+\noindent The condition now is not that the two-part code length is minimal, but that the \emph{randomness deficiency} $C_{\psi_i}(x) - C^{\K,\psi}(x\mid i)$ is less than a constant.
+$S_\tn{naive}^{\F,\psi,c}(x)$ corresponds to the version in \cite{mota2013sophistication}.
+The switch to the randomness deficiency avoids Theorem~\ref{theorem:overfitting}, but we end up with the same problem as in Theorem~\ref{thm:depth}: for shallow strings $S_\tn{naive}^{\T, \psi, c}$ is defined by the model $U^A$, and thus bounded.
 
-\begin{definition}[Naive sophistication]
-Let $C_i(x) = \min\left\{|y| : \phi_i(y) = x\right\}$. Then we define the \emph{naive sophistication} as:
+We cannot show that $S_\tn{naive}^{\F, \psi, c}(x)$ is bounded for shallow strings, but this is only a consequence of the use of sets, not of the switch to randomness deficiency as a condition.
+\emph{Any} set sophistication is necessarily lower-bounded by the function $\tn{set}(x) = \min\left\{C^\K(F) : x \in F\right\}$ and if this function were bounded, it would suggest that a finite amount of finite sets contained all strings.
+\begin{theorem}
+Let $\psi$ be \emph{any} acceptable numbering. Then for all shallow $x$ and large enough $c$, $S_\tn{naive}^{\T,\K,\psi,c}(x)$ is bounded and for come constant $c_F$, we have:\belowdisplayskip=-12pt
 \[
-\s_\text{naive}^{\M, \psi, c}(x) = \min\left\{C^\K(\psi_i) : i \in \M, C_i(x) - C^\K(x\mid s) \leq c \right\} \p
-\]
-\end{definition}
-
-The condition now is not that the two-part code length corresponds to the Kolmogorov complexity, as in the regular sophistication, but that the \emph{randomness deficiency} is less than a constant.
-$\s_\text{naive}^{\F,\psi,c}(x)$ corresponds to the version in \cite{mota2013sophistication}.
-The switch to the randomness deficiency avoids Theorem~\ref{theorem:overfitting}, but we end up with the same problem as Vit\'anyi's sophistication (Theorem~\ref{thm:depth}): for non-deep strings $s^{\T, \psi, c}$ is defined by the model $U^A$, and thus bounded.
-
-We cannot show that $s^{\F, \psi, c}(x)$ is bounded for non-deep strings, but this is only a consequence of the use of sets, not of the switch to randomness deficiency as a condition.
-\emph{Any} set sophistication is necessarily lower-bounded by the function $\text{set}(x) = \min\left\{C^\K(S) : x \in S\right\}$ and if this function were bounded, it would suggest that a finite amount of sets contained all strings.
-\begin{theorem}
-Let $x$ be a non-deep string. Then $\s_\text{naive}^{\T,\K,\psi,c}(x)$ is bounded and
-\[
-\text{set}(x) \leq \s_\text{naive}^{\F,\K,\psi,c}(x) \leq C^\K(C^\K(x)) \p
+\tn{set}(x) \leq S_\tn{naive}^{\F,\K,\psi,c}(x) \leq C^{\K,\psi}(C^{\K, \psi}(x)) + c_F \p
 \]\label{theorem:naive}
@@
 \begin{proof}
-For the first part, we take the model $U^A$ as defined in Section~\ref{section:underfitting}.For some numbering $\psi$, we have $C_i(x) - C^\K(x\mid S) \leq 0$, thus $\s_\text{naive}^{\T,\psi,c}(x) \leq C^K(U^A)$.
-For the second part, let $C^\K(x) = k$ and $S^A_k = \{x \mid \exists p : U^A(p) = x, |p| \leq k\}$, and let $S_k = \{x : \exists p : U(p) = x, |p| \leq k\}$.
-We will first show that $C^\K(x) = C^\K(S_k^A) + \log |S_k^A|$ up to a constant., ie $S^A_k$ is \emph{optimal} for $x$.
-It's known that  $\log |S^A_k| \leq \log |S_k| = k - K(k) + O(1)$ \cite{gacs2001algorithmic}.
-We also have $C^\K(S_k^A) \leq C^\K(k) + O(1)$.
-Thus
-\begin{align*} 
-C^\K(x) &= k + C^\K(k) -C^\K(k) \\
-&\geq C^\K(S^k) + \log |S^A_k| \geq C^\K(x) \\  
-\end{align*}
-Where all (in)equalities are up to a constant.
-The last inequality follows from the fact that $C^\K(x)$ is the smallest possible self-delimiting description.
-By Lemma III.9 in \cite{gacs2001algorithmic}, optimality implies typicality, so the randomness deficiency of $x$ in $S_k^A$ is bounded.
-Thus, for large enough $c$, $\s_\text{naive}^{\F,\psi,c}(x) \leq C^\K(C^\K(k))$.
+Let $\psi$ be any acceptable numbering.
+Let the Turing machine $U$ be defined as $U(\bi y) = \psi_i(y)$ if $i \in \K$, and $U(\bi y) = \infty$ otherwise.
+Let $U^A$ be derived from $U$ as in section in Section~\ref{section:underfitting} and let $\phi_u$ compute $U^A$.
+
+For the first part we have $C_{\psi_u}(x) - C^{\K, \psi}(x\mid u) \leq c_0$ for some $c_0$, thus for large enough $c$, $\s_\text{naive}^{\T,\psi,c}(x) \leq C^K(\psi_u)$.
+For the second part, let $C^{\K,\psi}(x) = k$ and $F^A_k = \{x \mid \exists p : U^A(p) = x, |p| = k\}$.
+$|F^A_k| \leq |\{p : |p| = k\}|$, so that $\log |F^A_k| \leq k$, which gives us $\log |F^A_k| - C^{K, \psi}(x \mid F^A_k) \leq c_1$.
+Thus, for large enough $c$, $\s_\text{naive}^{\F,\psi,c}(x) \leq C^K(F^A_k)$.
+From a description of $k$, we can compute $F^A_k$ with a finite program, so that $C^{\K,\psi}(F^A_k) \leq C^{\K,\psi}(k) + c_F$, which completes the proof.
 \end{proof}
-Another approach is the \emph{coarse sophistication} \cite{antunes2009sophistication}. The simplest definition comes from \cite{mota2013sophistication}:
+Note that the constant $c$ only needs to be large enough to ensure that $C^{\K,\psi}(x) - C^{\K,\psi}(x\mid u) \leq c$ and $C^{\K,\psi}(x) - C^{\K,\psi}(x\mid F^A_k)\leq c$.
+Since $u$ and $F^A_k$ are generally of no value in computing $x$, $c$ is likely very small.
+
+Another approach is the \emph{coarse sophistication} \cite{antunes2009sophistication}, defined in \cite{mota2013sophistication} as:
 \[
-\s^{\M,\Nm,\psi, c}_\text{coarse}(x) = \min_c\left\{\s^{\M,\Nm,\psi,c}(x) + c \right\} \p
+S^{\M,\Nm,\psi}_\tn{coarse}(x) = \min_c\left\{S^{\M,\Nm,\psi,c}(x) + c \right\} \p
 \]
 Again, this variant avoids the pitfalls of Theorem~\ref{theorem:overfitting}.
-If a set of $x$s has witnesses $S_x$ that are as good as the singleton set, up to a constant, but with smaller $C^\K(S)$ by more than a constant, the added penalty $c$ will eventually be much less than the gain for the simpler witness, and the singletons will not determine the coarse sophistication.
-It was noted in \cite{antunes2009sophistication} that the coarse sophistication is within a logarithmic term of the busy beaver depth, again relating sophistication to depth.
-For our purposes we can show that, as with the naive sophistication the total function version is bounded and the set version grows very slowly:
+If there are candidates that are as good as the singletons but with smaller size by more than a constant, the constant penalty $c$ will eventually be much less than the gain for the simpler witness, and the singletons will not determine the coarse sophistication.
+The coarse sophistication is within a logarithmic term of the busy beaver depth \cite{antunes2009sophistication}.
+As with the naive sophistication, we can show that for shallow strings, the total function version is bounded, and the set version grows very slowly:
 \begin{theorem}
-Let $c_0$ be some constant. Then for all non-deep strings $x$ we have that $\s_\text{coarse}^{\T,\K,\psi,c}(x)$ is bounded and
+Let $\psi$ be \emph{any} acceptable numbering. For all shallow $x$, $\s_\tn{coarse}^{\T,\K,\psi}(x)$ is bounded and there is a constant $c_F$ such that:\belowdisplayskip=-12pt
 \[
-\text{set}(x) \leq s_\text{coarse}^{\F,\K,\psi,c}(x) \leq C^\K(C^\K(x)) + c_0
-\].\label{theorem:coarse}
+\text{set}(x) \leq s_\tn{coarse}^{\F,\K,\psi}(x) \leq 2 C^{\K,\psi}(C^{\K,\psi}(x)) + c_F \p
+\]\label{theorem:coarse}
 \end{theorem}
 \begin{proof}
-For $\s_\text{coarse}^{\T,\K,\psi,c}(x)$, again take $U^A$ as a model.
-By Lemma~\ref{lemma:thecoolone}, there exists a numbering $\psi$ so that for all shallow $x$, a representation with $U^A$ witnesses the Kolmogorov complexity exactly.
-Thus $\s_\text{coarse}^{\T,\K,\psi,c}(x) \leq C^\K(U^A)$ for all non-deep strings.
-For the set case, we re-use $S_k^A$ as defined in the proof of Theorem~\ref{theorem:naive}.
-If we set $c$ to $C^\K(C^\K(x))$ in the regular sophistication, the upper bound follows.
-The lower bound follows as established earlier.
+Let $\psi$ be any acceptable numbering and define $U^A$, $\phi_u$ and $F^A_k$ as in the proof of Theorem~\ref{theorem:naive}.
+For the first part, we know that for some constant $c_0$, $\Cc^{\{u\}, \psi}(x) \leq \Cc^{\K,\psi}(x) + c_0$ so that $S^{\T, \K,\psi,c_0}(x) \leq C^{\K,\psi}(\phi_u)$, thus $S_\text{coarse}^{\T,\K,\psi}(x) \leq C^{\K,\psi}(\phi_u) + c_0$.
+For the second part, we know that for some $c_1$, $C^{\K,\psi}(F^A_k) + \log|F^A_k| \leq C^{\K,\psi}(k) + C^{\K,\psi}(x)+ c_1$, so that $S^{\F,\K,\psi,C^\K(k)+c_1}(x) \leq C^{\K,\psi}(F^A_k)$.
+Thus, $S_\tn{coarse}^{\F,\K,\psi}(x) \leq C^{\K,\psi}(F^A_k) + C^{\K,\psi}(k) + c_1 = 2 C^{\K,\psi}(C^{\K,\psi}(x)) + c_F$.
 \end{proof}
-In \cite{vereshchagin2013algorithmic}, Vereshchagin proposes a strongly algorithmic sufficient statistic.
-Whereas the regular algorithmic sufficient statistic $S$ from \cite{gacs2001algorithmic} has as one of its requirement that $C^\K(S\mid x)$ is constant, the strong variant imposes the requirement that $C^\T(S\mid x)$ is also constant.
-This reduces the problems of underfitting discussed in this section, but since $C^\T(\{x\}\mid x)$ is bounded, by Theorem~\ref{theorem:overfitting}, overfitting remains a problem: for a given constant deviation allowed from the Kolmogorov complexity, there exist numberings so that only the singletons come within that constant.
+In \cite{vereshchagin2013algorithmic}, Vereshchagin proposes a \emph{strongly algorithmic sufficient statistic}.
+Where the regular algorithmic sufficient statistic $F$ from \cite{gacs2001algorithmic} has $C^\K(F\mid x)$ constant, the strong variant imposes the stronger requirement that $C^\T(F\mid x)$ is also constant.
+This reduces the problems of underfitting discussed in this section, but since $C^\T(\{x\}\mid x)$ is bounded, by Theorem~\ref{theorem:overfitting}, overfitting remains a problem: there exist numberings under which the singletons are the only candidates.
 
-Finally, we must mention \emph{effective complexity} \cite{gellmann1996information}, proposed by Gell-Man and Lloyd.
-This measure is unique, in that it was formulated from the perspective of physics.
-Nevertheless, it fits the mold of sophistication very neatly.
-The models used are computable probability distributions on finite sets, and all candidates that come within a constant of the Kolmogorov complexity are considered.
-The complexity of the model is correctly measured by its Kolmogorov complexity.
-We note that this means that Lemma~\ref{theorem:overfitting} applies.
-Unlike other sophistication measures, here it is not the smallest model which is chosen, but the one which reproduces the data with in the shortest time.
-Thus, if there are multiple candidates, this approach would likely favor the singleton sets.
-In \cite{gell2004nonextensive}, the authors abandon this approach, and note the choice from the set of candidates is a subjective one, which depends on context.
+Finally, \emph{effective complexity} \cite{gellmann1996information}, proposed by Gell-Man and Lloyd, was formulated from the perspective of physics, but fits the mold of sophistication.
+The model class consists of all computable probability distributions on finite sets.
+The complexity of the model is measured by its Kolmogorov complexity, avoiding the problems of Section~\ref{section:indices}.
+Theorem~\ref{theorem:overfitting}, however, still applies to effective complexity.
+Unlike other sophistication measures, it is not the candidate with the smallest model which is chosen, but the one which reproduces the data within the shortest time.
+Thus, if there are multiple candidates, this approach would likely favor the singletons.
+In \cite{gell2004nonextensive}, the authors abandon this approach, and note that the choice from the set of candidates is a subjective one, which depends on context, which is in line with the view we express in the next section.
 
@@
 But that does not in itself mean that it should be impossible to come up with a sound measure.
-The intuition throughout the literature, starting with the structure function, appears to be that the crucial property is whether a string is typical for a model, and that this typicality can be tested: another random choice from that model should select a string with the same structure.
-This idea is bold, but not necessarily unreasonable.
-Nevertheless, in this section we offer the opinion that such a clean-cut separation \emph{cannot} be made to work.
-We provide two arguments to support this belief.
+The common intuition, starting with the structure function, appears to be that the crucial property is whether a string is typical for a model, and that this typicality can be tested: another random choice from that model should select a string with the same structure.
+This idea is bold, but not  unreasonable.
+Nevertheless, we offer the opinion that such a clean-cut separation \emph{cannot} be made to work.
+We provide two arguments.
 
@@
 We can generate data from a model $\phi_i, i \in \K$, by feeding it random bits until it produces an output.
-We will call the resulting probability distribution $p_i(x) = \sum_{y:\phi_i(y) = x} 2^{-|y|}$.
-In this setting, a sophistication could be called \emph{consistent} if, for data of sufficiently high Kolmogorov complexity, it reflects the complexity of that distribution from which it originated.
-Suppose the data is sampled from the universal distribution: let $\phi_u(\bar\imath y)=\phi_i(y)$ and sample from $p_u$ as described above.
+We will call the resulting probability distribution $p_i$.
+Call a sophistication \emph{consistent} if, for sufficiently large data, it reflects the complexity of the source of the data.
+Now, let $\phi_u(\bar\imath y)=\phi_i(y)$ and sample from $p_u$.
 Then the initial bits will determine the prefix encoded index $\bar\imath$ of the function $\phi_i$ that $\phi_u$ will subsequently emulate, and the remaining bits are used as inputs to $\phi_i$.
 We now ask, what should be the sophistication of the resulting data?
-\enlargethispage{3\baselineskip}
 
 Certainly, if we have to judge based only on the data, we cannot exclude the possibility that the data was sampled from $p_u$: after all, it was.
-On the other hand, neither can we exclude the possibility that it came from $p_i$, as again, it did!
-It seems unjustified to assume that, based on only the data, all but one model can always be disqualified.
+Yet, neither can we deny that it may have came from $p_i$, as again, it did!
+Eliminating the universal models does not solve this problem: the same argument holds if $\phi_u$ indexes, for instance, only those models computable by finite automata.
+\emph{Any} model that dominates a set of other models creates this kind of ambiguity.
 
@@
 
-The second reason we believe sophistication will elude definition is more technical.
-Given an input $x$ and some model class, say $\C$, consider the set of all possible two-part representations of $x$.
-Now, when the numbering is changed, the codelength of the model part of all these representations will change.
-This is illustrated in the right graph in Figure~\ref{fig:diagram}.
-The invariance theorem expresses that this change is limited by a constant term that does not depend on $x$.
-So for sufficiently large $x$, the horizontal shift of all the representations will be relatively small.
+The second reason we doubt sophistication is more technical.
+Consider the set of all possible two-part representations of $x$.
+When the numbering is changed, the codelength of the model part of all these representations will change.
+This is illustrated in the second diagram in Figure~\ref{fig:diagram}.
+The invariance theorem expresses that this change is limited by a constant term.
 However, even this small shift can push some representations out of the acceptable region (indicated by the dashed line), and pull others in.
@@
 And while we cannot \emph{prove} in general that such jumps can always occur, there seems to be no reason to believe that they do not.
-Indeed, in \cite{antunes2013sophistication} it is shown that logarithmic changes in the slack parameter can cause these effects.
+Indeed, in \cite{antunes2013sophistication} it is shown that logarithmic changes in the slack parameter can already cause these effects.
 
@@
 Note that part of the theory is fine: there is nothing wrong with evaluating models for the data by comparing their two-part code lengths.
-In fact, the randomness deficiency $\delta(\phi_i,x)=-\log p_i(x)-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data---if $p_i$ truly generated the data, then the probability that the randomness deficiency is more than $k$ bits is less than $2^{-k}$ \cite[Lemma~6]{bloem2014safe}.
+In fact, the randomness deficiency $-\log p_i(x)-C^\K(x\mid i)$ has a direct statistical interpretation as a measure of counterevidence---under $p_i$, the probability of a randomness deficiency above $k$ is less than $2^{-k}$ \cite[Lemma~6]{bloem2014safe}.
 In the Monet example above, this will allow us to disqualify the model expressing that the data is actually, say, a recording of jazz music.
@@
 But fundamental problems arise as soon as a hard cut-off is introduced on how far we are allowed to deviate from the minimum determined by the Kolmogorov complexity.
-In our opinion, a lot of measures taken in the literature to address problems with sophistication, such as restricting the model class or introducing model penalties, complicate the method and make problems harder to analyse, without actually addressing the fundamental issue.
+In our opinion, a lot of measures taken in the literature, such as restricting the model class or introducing model penalties, complicate the method and make problems harder to analyse, without actually addressing the fundamental issue.
 This is dangerous: if such ad-hoc fixes result in a theory that is hard to prove either wrong or right, it creates an artificial dead end for a valuable area of research.
-When the hard cut-off on which models are considered is avoided, on the other hand, all such measures are suddenly no longer necessary.
-What remains is a simple and elegant theory that can be used to sift through all possible models, disproving most while retaining a select number of interesting candidates for our further consideration.
+When the hard cut-off on candidates is avoided, however, all such measures are no longer necessary.
+What remains is an elegant theory that can be used to sift through all possible models, disproving most while retaining a select number of interesting candidates for our further consideration.
 
 \subsubsection*{\ackname}
-
 This publication was supported by the Dutch national program COMMIT and by  the Netherlands eScience center. We thank Tom Sterkenburg for interesting discussions.
@@
 \begin{lemma}[Invariance of function complexity]
-Let $\psi$ and $\eta$ be any two acceptable numberings.
-There exists a constant $c$ such that $\left| C^{\K,\psi}(f) - C^{\K, \eta}(f)\right | \leq c$ for all $f \in \C$ and $\left| C^{\C,\psi}(f) - C^{\C, \eta}(f)\right | \leq c$ for all $f \in \C$.
-\label{lemma:invariance}
+Let $\psi$ and $\eta$ be any two acceptable numberings Let $f$ be any partial computable function. There exists a constant $c$ independent of $f$ such that \belowdisplayskip=-12pt
+\[
+\left| C^{\K,\psi}(f) - C^{\K, \eta}(f)\right | \leq c \;\text{and}\; \left| C^{\C,\psi}(f) - C^{\C, \eta}(f)\right | \leq c \p
+\] \label{lemma:invariance}
 \end{lemma}
@@
 &= \min\left\{ C^{\C, \eta}(i) : \eta_{g(i)}= f\right\} - c
-\geq \min\left\{ C^{\C, \eta}(i) : \eta_i= f\right\} - c' = C^{\C, \eta}(f).
+= \min\left\{ C^{\C, \eta}(g(i)) : \eta_{g(i)}= f\right\} - c'\\
+&\geq \min\left\{ C^{\C, \eta}(j) : \eta_j= f\right\} - c' 
+= C^{\C, \eta}(f).
 \end{align*}
-Reversing $\psi$ and $\eta$ for the opposite inequality. The same proof holds for $\K$.
+Reverse $\psi$ and $\eta$ for the opposite inequality. The same proof holds for $C^\K$.
 \end{proof}
-This lemma shows that the complexity of a function does not depend on the choice of
-enumeration by more than a constant term.
-%\enlargethispage{2\baselineskip}
 
-\begin{restatable}{lemma}{faithful}\label{lemma:faithful-numberings}
+\begin{lemma}\label{lemma:faithful-numberings}
   There are faithful acceptable numberings.
-\end{restatable}
+\end{lemma}
 
@@
   \]
+It may seem that the second line requires a test whether $\phi_i(p)$ halts, for $\psi$ to be acceptable, but as we will show below, this is not the case.
+
 To show that $\psi$ is faithful, pick any function $f$. Then
@@
 
-To show that $\psi$ is acceptable, let $\phi_j$ denote the identity
-function. Then a $\phi$-index $i$ can be mapped to a $\psi$-index
-using the computable function $r(i)=\bar\jmath i$, so that
+To show that $\psi$ is acceptable, let $\phi_j(z)=z$. Then a $\phi$-index $i$ can be mapped to a $\psi$-index
+with $r(i)=\bar\jmath i$, so that
 $\psi_{r(i)}(y)=\psi_{\bar\jmath i}(y)=\phi_i(y)$. For the reverse,
@@
   p)}(y)=\phi_v(\bar\imath p,y)=\phi_{\phi_i(p)}(y)=\psi_{\bar\imath p}(y)$.
+Note that if $q$ can be written as $\bar \imath p$, but $\phi_i(p)$ diverges, $h(q)$ will still return a function, but one which doesn't halt, making it equivalent to $\phi_d$ as required. 
 \end{proof}

### 2015-10-12 Deliverable 13Q2/src/priors.tex
..

\documentclass[a4paper]{article}
\usepackage{amsmath,amsthm}

\theoremstyle{example}
\newtheorem{example}{Example}[section]
\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]

\newcommand{\del}[1]{\left(#1\right)}
\newcommand{\dif}{\,\textnormal{d}}

\newcommand{\pim}{\ensuremath{\pi_\textsc{\tiny{m}}}}
\newcommand{\pit}{\ensuremath{\pi_\textsc{\tiny{t}}}}
\newcommand{\pik}{\ensuremath{\pi_\textsc{\tiny{k}}}}

\title{Priors in the Switch Distribution}
\author{Steven de Rooij}

\begin{document}

\maketitle

\section{Introduction}
In the original switch distribution, we use a geometric prior $\pim$
on the number of switches, an arbitrary prior $\pit$ on the switch
indices and an arbitrary prior $\pik$ on the expert to switch to. In
this document we disregard both $\pim$ and $\pik$. The aim is to
analyse an interesting prior $\pit$, defined via the conditional
probabilities $\alpha_n:=\pit(Z>n|Z\ge n)$. Thus $\alpha_n$ denotes the
probability that we do not switch at sample size $n$. Defining priors
in terms of these conditional probabilities is useful because those
are the numbers that are actually used in the algorithm, and some
sequences $\alpha_n$ turn out to correspond to interesting $\pit$ that
would be hard to come up with otherwise.

\section{Loss Bound}
Here we develop a loss bound on the cost of encoding switch points
using the prior specified by the sequence $\alpha_1,\ldots,\alpha_n$.
We compute the $-\ln$ prior probability that the switches that occur
within the first $n$ samples are exactly $t_1,\ldots,t_m$:
\begin{equation}\label{eq:bnd}
\begin{split}
  &-\ln\pit(t^m\text{, no switches between $t_m$ and $n$})\\
&=-\ln\del{\del{\prod_{j=1}^m\del{\prod_{i=t_{m-1}+1}^{t_m-1}\alpha_i}(1-\alpha_{t_m})}\prod_{i=t_m+1}^n\alpha_i}\\
  &= \sum_{i=1}^n-\ln\alpha_i+\sum_{j=1}^m-\ln\del{\frac{1}{\alpha_{t_j}}-1}.
\end{split}
\end{equation}
The last expression can be read as follows: the first sum denotes the
cost of not switching during the first $n$ outcomes, and the second
sum denotes the correction for the switches that actually did occur.

Now set
\begin{equation}
  \alpha_n:=e^{-c\cdot w(n)},
\end{equation}
for some $c>0$ and some mass function $w$ on the positive integers. We
can now bound
\begin{equation}\label{eq:bnd2}
-\ln\pit(t^m)=c\sum_{i=1}^n w(n)+\sum_{j=1}^m -\ln\del{e^{c\cdot w(t_j)}-1}\le
c-m\ln c+\sum_{j=1}^m -\ln\del{w(t_j)},
\end{equation}
where we used that $w$ is a mass function summing to one, and the rule
$e^x\ge x+1$.

\begin{remark}
  Note that the loss does not depend on $n$, even though we do not
  allow ``stabilising'' by means of $\pim$. This is possible because
  distributions $\pit$ that are indirectly defined through their
  conditional probabilities and $w$ as above, are defective. Namely,
  the prior probability of never switching is
  $\prod_{i=1}^\infty\alpha_i=e^{-\sum w(i)}=e^{-1}$.
\end{remark}
We now choose $w(n)$ to be the universal prior for the integers,
$w(n)=1/(n(\log n)^2)$, which allows us to derive
\begin{equation}
  -\ln\pit(t^m)\le c-m\ln c+\sum_{j=1}^m\ln t_j+2\ln\ln t_j,
\end{equation}
a bound that is comparable to the one we obtain for the original
switch code (there we have $-\ln\pit(t^m)\le
m\ln2+\sum_{j=1}^m\ln t_j+\ln(t_m+1)$). However, we now have the following
two advantages:
\begin{itemize}
\item[(a)] We no longer need the stable states, which yields a much
  simpler algorithm with only half the running time of the previous
  version.
\item[(b)] There is a direct relationship between $w(n)$ and the
  obtained loss bound. The relationship is even easier than it used to
  be between $\pit$ and the loss bound in the original switch
  distribution.
\end{itemize}

\noindent The next step is to work out the relationship between $\pit$
and $w$. Assuming $c=1$, we have
\begin{equation}
  \pit(Z>n|Z\ge n)=\alpha_n=e^{-w(n)}\approx 1-w(n).
\end{equation}
A last question is: to which value should we set $c$. Optimisation
yields
\begin{equation}
0=\frac{\dif}{\dif c}\del{c-m\ln c}=1-\frac{m}{c}\quad\Rightarrow\quad c=m.
\end{equation}
Thus, if we know a priori that the number of switches will be large,
it can pay to set $c$ to a value larger than one. However, using a
conservative value for $c$ is better in the worst case.

\newpage
\section{A Method for Constructing $w(n)$}
Here we describe a convenient way to construct a discrete,
non-defective, easy to use prior $w(n)$ on the positive integers such
that it asymptotically behaves like a given continuous function
$f(x)$. This function should be positive and decreasing, and
$\int_1^\infty f(x)\dif x$ should converge. For example, $f(x)$ may
itself be a density on $[1,\infty)$, but we will not worry about
normalisation until after discretisation. As an example, we will
consider $f(x)=1/((x+o)(\ln(x+o))^2)$ for some $o\ge1$, since we want
to obtain something resembling the universal prior on the
integers. Now define
\begin{equation}
  g(x):=-\int f(x)\dif x.
\end{equation}
Note that $g(x)$ is decreasing and convex. In the example we obtain
$g(x)=1/\ln(x+o)$. We use $g$ to define the prior mass function
\begin{equation}
  w(n):=\frac{g(n)-g(n+1)}{g(1)}.
\end{equation}
The probability under $w$ of a range of numbers can be calculated
using the telescoping sum
\[
  \frac{1}{g(1)}\cdot \sum_{i=1}^n w(i) =
  g(1)-g(2)+g(2)-g(3)+\ldots+g(n)-g(n+1)=g(1)-g(n+1),
\]
so that
\begin{equation}
  \sum_{i=1}^n w(i)=1-\frac{g(n+1)}{g(1)}.
\end{equation}
In particular, by summing over all $i=1,2,\ldots$, we find that the
distribution is normalised. We will now compare $w$ to the function
$f$ we started out with: we have
\begin{equation}
  f(n+1)\le w(n)g(1)\le f(n)
\end{equation}
To see why this is true, note that $w(n)g(1)$ approximates the
derivative of $g$, which is $f$. Since $g$ is convex, $w(n)g(1)$
underestimates the derivative at $n$ and overestimates it at $n+1$.

We will treat two more examples. Let $f(x)=x^{-\alpha}$ for
$\alpha>1$. This can easily be normalised (it integrates to
$1/(\alpha-1)$), but if we would just set $w(n)=n^{-\alpha}$, the
normalising sum can no longer be computed. Therefore, we follow the
above procedure instead. We obtain $g(x)=(x^{1-a})/(\alpha-1)$ and
$w(n)=n^{1-\alpha}-(n+1)^{1-\alpha}$. We now have a distribution that is easy to
compute, with the bound
\[
(\alpha-1)(n+1)^{-\alpha}\le w(n)\le (\alpha-1)n^{-\alpha}.
\]
In the special case where $\alpha=2$ this reduces to $g(x)=1/n$ and
the familiar prior $w(n)=1/n-1/(n+1)=1/(n(n+1))$.

\end{document}

### 2015-10-12 fractal paper/src/fractal.old2.tex
..

--- 2013-02-08 fractal paper/src/fractal.old.tex
@@
 
-\title{Fractal Modeling for Machine Learning}
+\title{EM-algorithms for learning Discrete and Random Iterated Function Systems}
 \date{\today}
@@
 \begin{abstract}
-\noindent We present an algorithm for fitting IFS fractals to data and we investigate the use of fractal models in machine learning.
-Fractals are mathematical objects that display great detail and complexity despite having simple descriptions.
-It has been suggested that fractals model many natural and other real-world structures better than the shapes of classical geometry \cite{mandelbrot1982fractal}.
+\noindent We present an algorithm for solving the \emph{inverse problem}, the task of finding a fractal model for a given data set.
+The algorithm is inspired by the famous Expectation Maximization algorithm for fitting models with hidden variables.
+We build on a connection between Mixture-of-Gaussian models and Iterated function Systems (IFSs) to design an EM-style algorithm for IFS models.
+We also extend our algorithm to fit Random Iterated Function Systems (RIFSs) to data.
 \end{abstract}
 
-\begin{figure}[b]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \includegraphics[width=\textwidth]{../img/sierpinski.pdf}
-    \caption{}
-    \label{fig:sierpinski}
-  \end{subfigure}  
-  \hspace{0.03\textwidth}
-  \begin{subfigure}[b]{0.3\textwidth}
-   \includegraphics[width=\textwidth]{../img/sierpinski-evo.pdf}
-    \caption{}
-  \end{subfigure}
-  \hspace{0.03\textwidth}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \includegraphics[width=0.45\textwidth]{../img/sierpinski-off.png}
-    \includegraphics[width=0.45\textwidth]{../img/random013.png} \\
-    \includegraphics[width=0.45\textwidth]{../img/random010.png}
-    \includegraphics[width=0.45\textwidth]{../img/mvn1k.png}
-    \caption{}
-    \label{fig:measures}
+\section*{Definitions}
 
-  \end{subfigure}
-  \caption{\small (a) The Sierpinski triangle, a simple example fractal. (b) Constructing \textit{(a)} from an arbitrary initial image. (c) Examples of IFS probability distributions.}
-  \label{fig:examples}
-\end{figure}
+We will briefly review the established definitions and basic properties of Iterated Function Systems and their random variant.
+As the properties are well established, we will forego their proofs.
+For a more extensive introduction, including proofs and detailed references, we recommend \cite{hutchinson2000deterministic}.
 
-Fractals are mathematical objects with certain remarkable properties. What constitutes a fractal is not precisely defined\footnotemark, but the following properties are common:
+\begin{dfn}{Iterated Function System (Set)}
+An iterated function system $S$ is a set of $k$ contractive functions $S_i:{\mathbb R}^d \rightarrow {\mathbb R}^d$. The individual functions $S_i$ are called the IFS's \emph{components}.
+\end{dfn}
 
-\begin{description}
-  \item[Self similarity] A part is a scaled down copy of the whole.
-See for instance the Sierpinski triangle (fig.
-\ref{fig:sierpinski}).
-It consists of three triangular shapes which are scaled down copies of itself.
-  \item[Infinitely small detail] 'Zooming in' reveals ever finer detail.
-In the case of the Sierpinski triangle we will see the same shape recurring again and again, but fractals like the Mandelbrot set reveal a great variety of shapes.
-  \item[Non-integer dimension] The Sierpinski triangle, for example, has a dimension of $1.58$.
-This is a counter-intuitive notion, that is not simple to explain.
-We refer the reader to \cite{schroeder2009fractals} for an intuitive explanation.
-\end{description}
+\begin{thm}
+Given an iterated function system $S = \{S_i\}$, there exists a unique set $Q_S$ such that the following holds:
+\[
+Q_S = \bigcup_i S_i(Q_S)
+\]
+\end{thm}
 
-The earliest examples of fractals were described around the beginning of the twentieth century as counterexamples to conjectures in planar geometry and analysis and mostly considered pathological.
-It was not until the 1970s that they began to be seen as a single and useful family of mathematical objects that could accurately describe many physical phenomena.
-Mandelbrot put it as follows in the \emph{The Fractal Geometry of Nature} \cite{mandelbrot1982fractal}:
+$Q_S$ is called the \emph{limit set} of $S$.
 
-\begin{quotation}
-\small
-\noindent Clouds are not spheres, mountains are not cones, coastlines are not circles, and bark is not smooth, nor does lightning travel in a straight line.
-\end{quotation}
+\begin{thm}
+Given an iterated function system $S = \{S_i\}$, the function $f_S:{\mathbb R}^d \rightarrow {\mathbb R^d}$, defined as 
+\[
+f_S(X) = \bigcup_i S_i(X)
+\]
+and iterated as
+\[
+f_S^{n+1}(X) = f_S(f_S^n(X)), \; f_S^{1}(X) = f_S(X)
+\]
+will converge to $Q_S$ with $n$ for any nonempty, compact $X \subset {\mathbb R}^d$. That is
+\[
+\lim_{n \rightarrow \infty} f_S^{n}(X) = Q_S
+\]
+\end{thm}
 
-Fractal geometry has been used in many fields, including physics \cite{mandelbrot1984fractals}, geology \cite{cheng1997multifractal}, biology \cite{goldberger1992fractal} and economics \cite{turiel2003multifractal}.
-The first connection between fractals and the natural world was in the form of tangible objects such as coast lines, trees or bronchia, but the more hidden aspects of reality were soon shown to have fractal properties as well.
-The path traced out by the Brownian motion of a particle, the price chart for a commodity or the level of floods in a river delta have all been studied as fractals to great effect.
+Since, in statistics, probability measures are of greater value than point sets, we can adapt our definition and their results to that domain.
+First, we must define what it means to apply a transformation to a (probability) measure.
+Let $p$ be a probability measure on ${\mathbb R}^d$ and $f$ a function $f:{\mathbb R}^d \rightarrow {\mathbb R}^d$.
+We then define $f(p)$ as
+\[
+	\left (f(p) \right ) (x) = p(f^{-1}(x))
+\]
 
-The main problem with fractal analysis has always been the difficulty of finding a fractal model for a given set of data.
-The assumption that data contains fractal properties by itself already allows us to calculate certain statistics (such as its dimension), but the various families of fractals that have been proposed are difficult to fit to data.
-This is called the \emph{inverse problem}.
-Instead of generating an image or dataset from a given mathematical description of a fractal, we are asked to consider an image or dataset and find a good fractal model for it.
+That is, the density of $x$ under the transformed measure is equal to the density of the inversely transformed $x$ under the original measure.
+We will call the space of all probability distributions on ${\mathbb R}^d$ with non-empty, compact support ${\mathbb P}^d$
 
-This problem was investigated thoroughly in the context of image compression (see for instance \cite{hart1996fractal}), but has received relatively little attention in the area of statistics.
-In particular the case of data in arbitrary dimensions has been left largely uninvestigated.
+The transition to measures also allows us to give each component a different weight. This makes the IFS measure a \emph{multi-fractal} [@! cite] and creates a variety of densities.
 
-\footnotetext{Mandelbrot originally defined it as a set whose topological dimension differs from its Hausdorff dimension, but retracted this definition, stating that he preferred the word to be not precisely defined.}
+\begin{dfn}[Iterated Function System (Measure)]
+An IFS measure consists of $k$ contractive functions $S_i:{\mathbb R}^d \rightarrow {\mathbb R}^d$, and $k$ associated nonnegative \emph{priors} (or \emph{weights}) $w_i \in {\mathbb R}$ such that $\sum_i w_i = 1$.
+\end{dfn}
 
-\subsection*{Previous research}
+\begin{thm}
+Given an IFS measure $S = \{(S_i, w_i)\}$, there exists a unique measure $q_S$ such that the following holds:
+\[
+q_S(x) = \sum_i w_i \left(S_i(q_S)\right)(x)
+\]
+\end{thm}
 
-The most fruitful algorithms for the induction of fractal models so far have been partitioned iterated function systems and evolutionary algorithms.
+$q_S$ is called the \emph{limit measure} of $S$.
 
-Partitioned iterated function systems \cite{hart1996fractal} are a subset of iterated function systems (described in the next section) designed for fast image compression by fractal models.
-The compression algorithm segments an image in to two tilings of different size, and then attempts to map the larger tiles to the smaller tiles.
-The resulting set of contraction mappings can be iterated to approximate the original image.
-The approach works well for two dimensional data stored in binned form, but does not translate well to higher dimensions.
+\begin{thm}
+Given an IFS measure $S = \{(S_i, w_i)\}$, the function $f_S:{\mathbb R}^d \rightarrow {\mathbb R^d}$, defined as 
+\[
+f_S(p) = \sum_i w_i S_i(p)
+\]
 
-Another approach is to represent a fractal model as a point in some space (eg.
-a vector or a bitstring) and to search the space for a model which approximates the data by some target function using generic search algorithms.
-Fractal models tend to be difficult to optimize for and only specific methods show promising results.
-For instance, the use of evolutionary strategies combined with Hausdorff distance as a target function has been shown to be able to reconstruct fractals from data, and to fit models to common datasets.
-These approaches can be found in \cite{bloem2010fractal,nettleton1994evolutionary} and many others.
+will converge to $q_S$ with $n$ for $p \in {\mathbb P}^d$. That is
+\[
+\lim_{n \rightarrow \infty} f_S^{n}(p)(x) = q_S(x)
+\]
+\end{thm}
 
-The generic nature of these search methods means that they do not exploit any knowledge of the fractal models.
-We expect that the induction of fractal models can be sped up and made more accurate if we tailor the search algorithm to the model.
+As shown in the examples section, the family of IFS measures is rich and diverse, and contains some surprisingly organic forms.
+However, the structures in nature that have traditionally been considered fractals (eg.
+coast lines, trees and lightning strikes) also display a great degree of randomness, not captured by these discrete IFSs.
+To model them succesfully, we require the framework of Random Iterated Function Systems (RIFSs).
 
-The investigation of iterated function systems has yielded some results that are promising, but have not resulted in an algorithm for the induction of fractal models in arbitrary dimensions.
-Examples include the method of moments \cite{rinaldo1994inverse}, similarity hashing \cite{hart1997similarity} and wavelet analysis \cite{struzik1996coastline}.
+The key thing to keep in, when dealing with RIFS models is that they do not describe a singular set, or measure on ${\mathbb R}^d$, but a probability distribution of such sets.
+For the sake of brevity we will skip the RIFS sets and move directly to RIFS measures.
 
-\section*{Methods}
+To define RIFS models we will need to consider the set of all possible deterministic IFS measures $S = {(S_i, w_i)}$. We will call this set ${\mathbb S}$
+
+\begin{dfn}[Random Iterated Function System (Measure)]
+A RIFS measure ${\cal S}$consists of a probability distribution over ${\mathbb S}$
+\end{dfn}
+
+This is a broad definition.
+The simplest version, and the subset that we will be focusing on is the Discrete RIFS, which consists of a finite set of $h$ deterministic IFS, each with a nonnegative prior $\rho_i$ such that $\sum_i \rho_i = 1$ for each.
+
+There is are analog of the limit set theorems for RIFSs, but we will not go into them here.
+
+\begin{dfn}[The inverse problem for IFSs]
+Given a finite set of points $X \subset {\mathbb R}^d$, find an IFS measure $S$, such that $S$ `fits the data' by some criterion. 
+\end{dfn}
+
+The obvious criterion for fitting the data in this case would be maximum likelihood (although others exist).
+Unfortunately, as explained below, Likelihood poses some problems for IFS fitting (specifically for models near the maximum).
+Since our algorithm is only an approximation to the maximum likelihood fit, we will leave the specifics of the criterion open for now.
+
+\begin{dfn}[The inverse problem for RIFSs]
+Given a set ${\cal X}$ of sets of points (both finite), find a discrete RIFS measure ${\cal S}$ such that each $X \in {\cal X}$ can be said to fit a (probable) instance of ${\cal S}$.
+\end{dfn}
+
+While we believe this to be a very relevant problem in the field of fractal modeling, the task of fitting a model to sets of sets of points is not a common one in statistics.
+Again, we will only show emprically that the algorithm has the possibility to converge to the correct model in some cases, and we will not give any results on which criterion (if any) is optimized, either locally or globally.
+
+\subsection*{Examples}
+
+\section*{Methods for deterministic IFS models}
 
@@
 
-\subsection*{Iterated Function Systems}
-
-If we are to fit fractal models to data then we require a family of fractals which is easy to parametrize, and likely to encompass a large number of organic and useful patterns.
-We will use the formalism of \emph{Iterated Function Systems} (IFSs) for this purpose.
-
-Since the aim of this paper is to present our induction algorithm and to introduce IFS models to a machine learning context, we will provide only an informal treatment of the properties of IFSs that are already well established.
-For a more rigorous treatment, we recommend \cite{hutchinson2000deterministic} and the references therein.
-
-Iterated Function Systems are a simple generalization of many classic fractals.
-Consider the Sierpinski Gasket (fig.
-\ref{fig:sierpinski}).
-As noted before, it consists of three scaled down versions of itself.
-In mathematical terms, if we view the Sierpinski gasket as a set in ${\mathbb R}^2$ and take the union of three specific transformations applied to it, we will get the Sierpinski gasket again.
-The key result in the theory of iterated function systems is that for these three transformations there is only one set that has this property.
-The three transformations together form a scaling law, and we say the Sierpinski gasket satisfies the scaling law.
-
-We will use the following notation.
-Let $S_1$, $S_2$, \ldots, $S_k$ be a set of contractive transformations on ${\mathbb R}^d$.
-(We will later limit the family of allowable functions, but from a mathematical point of view we only require them to be contractive).
-There exists a single set $Q$ that satisfies the following property:
-\begin{align}
-Q = S_1(Q) \cup S_2(Q) \cup \ldots \cup S_k(Q) \label{scaling law}
-\end{align}
-
-We call $\{S_i\}$ a \textit{scaling law}, and we say that $Q$ satisfies it.
-
-If we apply the transformations $S_i$ to a set $X$ that doesn't satisfy the scaling law (and take the union of the results), we get a different set $X'$.
-Call this function $S$: $S(X) = \bigcup S_i(X)$.
-Another important result about IFSs is that when we iterate $S$ as $S^{n+1}(X) = S(S^n(X))$ we get
-\[
-\lim_{n \rightarrow \infty} S^n(X_0) = Q
-\]
-with $X_0$ any non-empty set and $Q$ the unique set that satisfies the scaling law $\{S_i\}$.
-
-Of course, in statistics, we are interested in finding probability distributions rather than sets.
-The IFS formalism can be extended to measures in a natural way.
-First, we must extend the concept of a map to measures.
-Let $p$ be some probability measure over ${\mathbb R}^d$ and $F:{\mathbb R}^d \rightarrow {\mathbb R}^d$ be some map on ${\mathbb R}^d$.
-Then we define the mapping of $p$ by $F$ as:
-\[
-(F(p))(X) = v(F^{-1}(X))
-\]
-So if $p$ is a probability measure, then $F(p)$ is also a probability measure and $F(p)$ assigns $X$ the same probability as $p$ would assign to $F^{-1}(X)$, where $F^{-1}$ is the inverse of $F$.
-
-Our IFS over probability measures consists, as before, of $k$ transformations, but here we also introduce $k$ weights $w_i$ with the constraint that they sum to one.
-Let $q$ be a probability distribution and $S = \{(S_i, w_i)\}$ be a scaling law.
-We say that $q$ satifies $S$ iff:
-\[
-q(X) = \sum_{i\in[1, k]} w_i S_i(q)(X)
-\]
-
-As before q is uniquely defined by $S$. If we apply the transformations and weights to a different probability distribution iteratively, we get a sequence which converges to $q$:
-\begin{align*}
-p_{n+1} &= \sum w_i S_i(p_n) \\
-\lim_{n \rightarrow \infty} p_n &= q 
-\end{align*}
-
-This is the type of IFS model we will attempt to learn from data. See figure \ref{fig:measures} for some examples IFS measures.
+For the random case, our data ${\cal X}$ consists of $m$ distinct datasets ${X_i}$ where each $X_i$ is defined as in the previous paragraph.
 
@@
 
-The method described below can easily be extended to other families of transformations such as affine transformations or neural networks.
-The only requirement is that we must have some way of finding a transformation $F$ within this family given two ordered point sets $\{x_i\}, \{y_i\} \subseteq {\mathbb R}^d$ so that the error is minimized.
-For similarities there is a direct and optimal method.
-\cite{umeyama1991least}
+This means, of course that our method is not suited to learn all IFSs, but only those based on similitudes.
+However the skeleton of the algorithm does not depend strongly on the choice of map type.
+Any type of map for which there exists a method to fit a map to an existing set of point pairs $\{(x_i, y_i)\}$ can potentially be used with this algorithm.
+We chose the $similitudes$ as a family, because there is an efficient, optimal method for solving this sub-problem.
 
@@
 
-\subsection*{Convergence}
+\section*{Methods for Random IFS models}
+
+\subsection*{Choice trees}
+
+In the case of fitting deterministic fractals, our main problem turned out to be the existence of a complex hidden variable, in the form of a code for each data point.
+
+In the random case, we are faced with further unobservable information.
+Assuming that a single dataset is drawn from a given instance of a random fractal ${\cal S}$, we must know how this instance is built up from ${\cal S}$'s deterministic components.
+To simplify the problem, we will assume that ${\cal S}$'s components all have the same number $k$ of component maps, and that ${\cal S}$ has $m$ components.
+
+The full representation of a single dataset generated from ${\cal S}$ then consists of the generated points (the obeervables), the code tree (with the points at its leaves and a nuber representing the IFS) and at each node of the code tree a number from $\{1, \ldots, m\}$ to denote the choice of component IFS at that level.
+
+[@! Picture]
+
+Analogous to the soft responsibilities of the classical EM algorithm, we will use choice trees that hold at their nodes not a single choice of component IFS, but a probability distribution over all components.
+We'll call these stochastic choice trees.
+
+We now have three elements that together generate a dataset: a discrete RIFS model, codes for each point in each dataset, a choice tree per dataset).
+Each of these three can be found in a straightforward manner, if the other two are known.
+Our algorithm consists of an iteration of all three steps.
+
+When we speak of coding, we must now distinguis between the \emph{map coding} of a point, which consists of a string of integers from $(1,k)$ denoting the choice of map of the component IFS at each level and the \emph{complete coding} of a point, which consists of a string of pairs of two integers the frist from $(1, m)$ and the second from $(1,k)$.
+This code give both the choice IFS component and its map at each level.
+
+[@! algorithm]
+
+\subsection*{Finding codes given a model and a set of choice trees}
+
+\subsection*{Finding a choice tree given a model and a coding}
+
+As we know the coding we can, for each node in the code tree, find the points $X_\sigma$ that have the code prefix $\sigma$.
+We can also, for every code $\tau$ that extends $\sigma$ by a single symbols, fond the points with that code.
+
+\subsection*{Finding a model given a coding}
+
+Given a complete code for each point, we can repeat the principle used for the
 
@@
 
+\subsection*{Discrete}
 We perform a number of experiments to give an impression of the behavior of IFSs as statistical models, and the EM algorithm as a learning method.
@@
 
+\subsection*{Rnadom}
+
 \section*{Conclusions}
@@
 We introduced an EM-style algorithm for fitting fractal models to data.
-Compared to search algorithms for other types of models the algorithm is relatively expensive, but compared to other approaches for fractal modeling the algorithm represents a significant improvement @!.
+Compared to search algorithms for other types of models the algorithm is relatively expensive, but compared to other approaches for fractal modeling the algorithm represents a significant improvement [@!].
 
-Further, we have given some indication of how the study of fractal modeling can open up interesting new prospects in machine learning.
-We expect that these methods will allow many avenues of future research.
-One option would be to investigate the use of more expressive maps.
-The similitudes used here could be replaced by neural networks, for instance.
-The EM search algorithm is fairly simple, and could be extended to try various branches and backtrack if it deviates too far from a good model.
-
-\subsubsection*{Other fractals}
-
-It is important to note that there are many other families of fractals besides iterated function systems.
-They are not covered by our method, and cannot be properly represented by Iterated Function Systems.
-Examples include Julia sets and dynamical systems.
-
-There is one family of fractals in particular that would be very interesting to explore in light of this research, and that is the family of random fractals.
-When we think of natural phenomena that seem to resemble iterated function systems---coast lines, trees, the veins of a leaf---there is always a combination of fractal scaling and some random component.
-The framework most likely to capture these phenomena is that of random Iterated Function Systems, as described in \cite{hutchinson2000deterministic}.
-
-The random IFS model is an extension of the Iterated Function System which produces not one single set but a distribution over all sets.
-The dataset for which to learn a random fractal would not consist of one list of vectors, but of a list of lists of vectors.
-For instance, a collection of timeseries preceding a critical event, or a collection of images of snowflakes.
-In \cite{bloem2010fractal} we showed a proof-of-concept, indicating that such models can be learned by evolutionary algorithm in principle, but we did not advance beyond the simplest conceivable model.
-If the EM algorithm described here could be extended to random fractals, it would open up a new type of machine learning task, where instances ae represented by point patterns, and the model to be learned is a distribution over distributions.
+with these tools in place, we believe the door is opened to new opportunities for statistical modeling with fractals.
+Rather than simply proving that fractal structure exists in a dataset (for instance by measuring its fractal dimension), these algorithms will help researchers to expose that structure.
+Which datasets and modeling problems are susceptible to fractal analysis of this kind is a question for future research.
 

### 2015-10-12 fractal paper/src/fractal.tex
..

--- 2012-12-04 fractal paper/src/fractal.tex
@@
 
-\usepackage{charter}
-\usepackage{eulervm}
 \usepackage{amsmath, amsthm, amssymb}
@@
 
-\title{Fractal Modeling for Machine Learning}
+\newcommand{\p}{\mbox{\,.}}
+\newcommand{\bD}{\boldsymbol\Lambda}
+\newcommand{\bm}{\boldsymbol\mu}
+\newcommand{\cN}{{\cal N}}
+\newcommand{\bs}[1]{{\boldsymbol {#1} }}
+
+\title{A variational algorithm for the Fractal Inverse Problem.}
 \date{\today}
@@
 
+\newcommand{\pb}[1]{\textcolor{OliveGreen}{\small #1 \textsuperscript{[pb]} }}
+
 \begin{document}
@@
 \begin{abstract}
-\noindent We present an algorithm for fitting IFS fractals to data and we investigate the use of fractal models in machine learning.
-Fractals are mathematical objects that display great detail and complexity despite having simple descriptions.
-It has been suggested that fractals model many natural and other real-world structures better than the shapes of classical geometry \cite{mandelbrot1982fractal}.
+\noindent We present an algorithm for solving the \emph{inverse problem} for iterated function systems (IFS) and random iterated function systems (RIFS).
+The central idea is to cast the IFS as an generalization of a mixture of of Gaussians model, and to treat the particular sequence of components that leads to a particular point as a latent variable.
 \end{abstract}
+\section{Introduction}
+\pb{Briefly: importance of fractals, inverse problem.}
 
-\begin{figure}[b]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \includegraphics[width=\textwidth]{../img/sierpinski.pdf}
-    \caption{}
-    \label{fig:sierpinski}
-  \end{subfigure}  
-  \hspace{0.03\textwidth}
-  \begin{subfigure}[b]{0.3\textwidth}
-   \includegraphics[width=\textwidth]{../img/sierpinski-evo.pdf}
-    \caption{}
-  \end{subfigure}
-  \hspace{0.03\textwidth}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \includegraphics[width=0.45\textwidth]{../img/sierpinski-off.png}
-    \includegraphics[width=0.45\textwidth]{../img/random013.png} \\
-    \includegraphics[width=0.45\textwidth]{../img/random010.png}
-    \includegraphics[width=0.45\textwidth]{../img/mvn1k.png}
-    \caption{}
-    \label{fig:measures}
+\pb{Iterated Function Systems.}
 
-  \end{subfigure}
-  \caption{\small (a) The Sierpinski triangle, a simple example fractal. (b) Constructing \textit{(a)} from an arbitrary initial image. (c) Examples of IFS probability distributions.}
-  \label{fig:examples}
-\end{figure}
+\pb{Idea behind our approach. EM principle, using variational Bayes.}
 
-Fractals are mathematical objects with certain remarkable properties. What constitutes a fractal is not precisely defined\footnotemark, but the following properties are common:
+\subsection{Preliminaries}
+We will deal with datasets which are a set of points $\{x_1, \ldots, x_n\}$ with $x_i \in \mathbb{R}^D$.
 
+Let $F_{t, R}(y) = t + Rx$ be an affine transformation represented by a vector $t$ and a matrix $R$.
+If $X ~ \cN(\mu, \Sigma)$, then $F_{t, R}(Y) ~ \cN(t + R\mu, R\SigmaR^T)$.
+Specifically this means that if we take the standard multivariate normal $\cN_0 = \cN(0, I)$ as  canonical source, we can represent every multivariate normal $\langle \mu, \Sigma \rangle$ by an affine transformation $\langle t, R \rangle$ with $t = \mu$, $\Sigma = RR^T$.
+It also tells us that sampling from $\cN_0$ and applying a series of affine transformations is equivalent to sampling from a particular MVN.
+
+\paragraph{Variational Bayes} Let $x$ be the observed data of a problem, for which we have a model, and let $z$ be the latent variables of the model.
+If the model has any parameters, we place priors on them so that we can consider them latent variables, and absorb them into $z$ The variational Bayes approach attempts to approximate the distribution $p(z \mid x)$ using a \emph{variational distribution} $q(z)$.
+We assume that $z$ can be split into subsets $z_1, \ldots, z_m$, for which $q$ factorizes:
+\[
+q(z) = \prod_i q_i(z_i)
+\]
+A central result in variational Bayes is that the optimum for each factor is given by:
+\begin{equation}
+q^*_i(z_i) = \langle \ln p(x, z)\rangle_{/i} + \text {const} \p \label{line:var-bayes}
+\end{equation}
+Where $\langle \ln p(x, z) \rangle_{/i} = \sum \prod_{j = {1, \ldots, m}/i} q_j(z_j) ln p(x, z)$ with the sum over all values of all $z_j$ except $z_i$.
+That is, it represents the expectation with respect to all latent variable factors except $z_i$.
+
+Usually, the values of these optimal factors have cyclical dependencies, which means that we cannot compute $q^*$ directly.
+We can, however, search for a local optimum by iteration: we start with some initial value, and use (\ref{line:var-bayes}) to compute new values for each factor based on those of the previous generation.
+
+\section{The IFS model}
+
+\begin{figure*}[t]
+  \includegraphics[width=\textwidth]{../img/factor-graph.pdf}
+  \caption{A graphical model, illustrating the components of the IFS model.
+The gray box is a plate, representing a repetition of the nodes inside the box for each datapoint.
+The black node represents the observed data.}
+  \label{figure:ifs-diagram}
+\end{figure*}
+
+The complete IFS model is illustrated in Figure~\ref{figure:ifs-diagram}. If contains the following components:
 \begin{description}
-  \item[Self similarity] A part is a scaled down copy of the whole.
-See for instance the Sierpinski triangle (fig.
-\ref{fig:sierpinski}).
-It consists of three triangular shapes which are scaled down copies of itself.
-  \item[Infinitely small detail] 'Zooming in' reveals ever finer detail.
-In the case of the Sierpinski triangle we will see the same shape recurring again and again, but fractals like the Mandelbrot set reveal a great variety of shapes.
-  \item[Non-integer dimension] The Sierpinski triangle, for example, has a dimension of $1.58$.
-This is a counter-intuitive notion, that is not simple to explain.
-We refer the reader to \cite{schroeder2009fractals} for an intuitive explanation.
+\item[the component weights $\pi$] This is a vector of $k$ nonnegative real values, with $\sum_i \pi_i = 1$.
+Each weight determines the probability of its associated component being chosen at each iteration of the IFS.
+\item[the depth $d$] The number of times the IFS is iterated.
+\item[the components $\bm = \{\mu_1, \ldots, \mu_k\}$, $\bD = \{\Lambda_1, \ldots, \Lambda_k\}$] The means and precision matrices for the $k$ components. 
+\item[The post-transformation $\mu_\text{post}$, $\Lambda_\text{post}$] The parameters for a single post-transformation.
+To see why this is necessary, consider the Koch curve in Ficgure~\ref{figure{fractals}}.
+While a dataset sampled from this measure can be prefectly represented by an IFS, it is slighty off-center.
+If this were real dataset, it might be mean-centered instead, in which case no IFS would fit.
+Unlike the mixture of Gaussians model, an IFS cannot be translated by a straightforward change of the parameters,a nd as the Koch curve shows, simple solutions like mean-centering the data do not always solve the problem.
+For this reason, we add a single post-transformation into the model, to allow for any simple, affine transformations to the final limit measure.
+This factor has the additional advantage that at $d = 0$, the model collapses to a single MVN distribution.
+\item[The transformation sequence $\sigma_i$] For each data point $x_i$, we have a sequence of length $d$ representing the sequence of component-transformations we need to apply to a point sampled from $\cN_\text{pre}$ to get the final point sampled from the limit measure of the IFS.
+\item[The data $x_i$] This is the point after the post-transformation is applied: ie. the observed data.
 \end{description}
 
-The earliest examples of fractals were described around the beginning of the twentieth century as counterexamples to conjectures in planar geometry and analysis and mostly considered pathological.
-It was not until the 1970s that they began to be seen as a single and useful family of mathematical objects that could accurately describe many physical phenomena.
-Mandelbrot put it as follows in the \emph{The Fractal Geometry of Nature} \cite{mandelbrot1982fractal}:
+The sampling procedure modeled
 
-\begin{quotation}
-\small
-\noindent Clouds are not spheres, mountains are not cones, coastlines are not circles, and bark is not smooth, nor does lightning travel in a straight line.
-\end{quotation}
-
-Fractal geometry has been used in many fields, including physics \cite{mandelbrot1984fractals}, geology \cite{cheng1997multifractal}, biology \cite{goldberger1992fractal} and economics \cite{turiel2003multifractal}.
-The first connection between fractals and the natural world was in the form of tangible objects such as coast lines, trees or bronchia, but the more hidden aspects of reality were soon shown to have fractal properties as well.
-The path traced out by the Brownian motion of a particle, the price chart for a commodity or the level of floods in a river delta have all been studied as fractals to great effect.
-
-The main problem with fractal analysis has always been the difficulty of finding a fractal model for a given set of data.
-The assumption that data contains fractal properties by itself already allows us to calculate certain statistics (such as its dimension), but the various families of fractals that have been proposed are difficult to fit to data.
-This is called the \emph{inverse problem}.
-Instead of generating an image or dataset from a given mathematical description of a fractal, we are asked to consider an image or dataset and find a good fractal model for it.
-
-This problem was investigated thoroughly in the context of image compression (see for instance \cite{hart1996fractal}), but has received relatively little attention in the area of statistics.
-In particular the case of data in arbitrary dimensions has been left largely uninvestigated.
-
-\footnotetext{Mandelbrot originally defined it as a set whose topological dimension differs from its Hausdorff dimension, but retracted this definition, stating that he preferred the word to be not precisely defined.}
-
-\subsection*{Previous research}
-
-The most fruitful algorithms for the induction of fractal models so far have been partitioned iterated function systems and evolutionary algorithms.
-
-Partitioned iterated function systems \cite{hart1996fractal} are a subset of iterated function systems (described in the next section) designed for fast image compression by fractal models.
-The compression algorithm segments an image in to two tilings of different size, and then attempts to map the larger tiles to the smaller tiles.
-The resulting set of contraction mappings can be iterated to approximate the original image.
-The approach works well for two dimensional data stored in binned form, but does not translate well to higher dimensions.
-
-Another approach is to represent a fractal model as a point in some space (eg.
-a vector or a bitstring) and to search the space for a model which approximates the data by some target function using generic search algorithms.
-Fractal models tend to be difficult to optimize for and only specific methods show promising results.
-For instance, the use of evolutionary strategies combined with Hausdorff distance as a target function has been shown to be able to reconstruct fractals from data, and to fit models to common datasets.
-These approaches can be found in \cite{bloem2010fractal,nettleton1994evolutionary} and many others.
-
-The generic nature of these search methods means that they do not exploit any knowledge of the fractal models.
-We expect that the induction of fractal models can be sped up and made more accurate if we tailor the search algorithm to the model.
-
-The investigation of iterated function systems has yielded some results that are promising, but have not resulted in an algorithm for the induction of fractal models in arbitrary dimensions.
-Examples include the method of moments \cite{rinaldo1994inverse}, similarity hashing \cite{hart1997similarity} and wavelet analysis \cite{struzik1996coastline}.
-
-\section*{Methods}
-
-In this paper we focus on numerical data in an arbitrary number of dimensions.
-Each instance in our dataset is represented by a fixed number of numerical features.
-Call our dataset $X$, our instance $x \in X$ and $x \in {\mathbb R}^d$.
-We will assume that our dataset represents iid samples from a probability distribution over ${\mathbb R}^d$ which we will attempt to model.
-
-\subsection*{Iterated Function Systems}
-
-If we are to fit fractal models to data then we require a family of fractals which is easy to parametrize, and likely to encompass a large number of organic and useful patterns.
-We will use the formalism of \emph{Iterated Function Systems} (IFSs) for this purpose.
-
-Since the aim of this paper is to present our induction algorithm and to introduce IFS models to a machine learning context, we will provide only an informal treatment of the properties of IFSs that are already well established.
-For a more rigorous treatment, we recommend \cite{hutchinson2000deterministic} and the references therein.
-
-Iterated Function Systems are a simple generalization of many classic fractals.
-Consider the Sierpinski Gasket (fig.
-\ref{fig:sierpinski}).
-As noted before, it consists of three scaled down versions of itself.
-In mathematical terms, if we view the Sierpinski gasket as a set in ${\mathbb R}^2$ and take the union of three specific transformations applied to it, we will get the Sierpinski gasket again.
-The key result in the theory of iterated function systems is that for these three transformations there is only one set that has this property.
-The three transformations together form a scaling law, and we say the Sierpinski gasket satisfies the scaling law.
-
-We will use the following notation.
-Let $S_1$, $S_2$, \ldots, $S_k$ be a set of contractive transformations on ${\mathbb R}^d$.
-(We will later limit the family of allowable functions, but from a mathematical point of view we only require them to be contractive).
-There exists a single set $Q$ that satisfies the following property:
+This gives us the joint distribution $p(\bs{x}, \bs{\sigma}, \mu_\text{pre}, \Lambda_\text{pre}, \bm, \bD, \mu_\text{post}, \Lambda_\text{post}, \pi, d)$ which factorizes as follows:
 \begin{align}
-Q = S_1(Q) \cup S_2(Q) \cup \ldots \cup S_k(Q) \label{scaling law}
+p(&\bs{x}, \bs{x'}, \bs{\sigma}, \mu_\text{pre}, \Lambda_\text{pre}, \bm, \bD, \mu_\text{post}, \Lambda_\text{post}, \pi, d) \notag\\ 
+=\;& p(\mathbf{x} \mid \mathbf{x'}, \mu_\text{post}, \Lambda_\text{post}) \; p(\mu_\text{post}, \Lambda_\text{post}) \notag\\
+& p(\mathbf{x'} \mid \bs{\sigma}, \bm, \bD, \mu_\text{pre}, \Lambda_\text{pre}) \notag\\
+& p(\mathbf{\sigma} \mid \pi, d) \; p(\pi) \; p(d) \notag\\
+& \prod_i p(\mu_i, \Lambda_i) \p \label{line:factorization}\\
 \end{align}
 
-We call $\{S_i\}$ a \textit{scaling law}, and we say that $Q$ satisfies it.
-
-If we apply the transformations $S_i$ to a set $X$ that doesn't satisfy the scaling law (and take the union of the results), we get a different set $X'$.
-Call this function $S$: $S(X) = \bigcup S_i(X)$.
-Another important result about IFSs is that when we iterate $S$ as $S^{n+1}(X) = S(S^n(X))$ we get
-\[
-\lim_{n \rightarrow \infty} S^n(X_0) = Q
-\]
-with $X_0$ any non-empty set and $Q$ the unique set that satisfies the scaling law $\{S_i\}$.
-
-Of course, in statistics, we are interested in finding probability distributions rather than sets.
-The IFS formalism can be extended to measures in a natural way.
-First, we must extend the concept of a map to measures.
-Let $p$ be some probability measure over ${\mathbb R}^d$ and $F:{\mathbb R}^d \rightarrow {\mathbb R}^d$ be some map on ${\mathbb R}^d$.
-Then we define the mapping of $p$ by $F$ as:
-\[
-(F(p))(X) = v(F^{-1}(X))
-\]
-So if $p$ is a probability measure, then $F(p)$ is also a probability measure and $F(p)$ assigns $X$ the same probability as $p$ would assign to $F^{-1}(X)$, where $F^{-1}$ is the inverse of $F$.
-
-Our IFS over probability measures consists, as before, of $k$ transformations, but here we also introduce $k$ weights $w_i$ with the constraint that they sum to one.
-Let $q$ be a probability distribution and $S = \{(S_i, w_i)\}$ be a scaling law.
-We say that $q$ satifies $S$ iff:
-\[
-q(X) = \sum_{i\in[1, k]} w_i S_i(q)(X)
-\]
-
-As before q is uniquely defined by $S$. If we apply the transformations and weights to a different probability distribution iteratively, we get a sequence which converges to $q$:
+We use the following priors:
 \begin{align*}
-p_{n+1} &= \sum w_i S_i(p_n) \\
-\lim_{n \rightarrow \infty} p_n &= q 
+p(\pi) &= \text{Dir}(\bs{\alpha}_0) \\
+p(d) &= {\cal U}(0, d_\text{max}) \\
+p(\mu, \Lambda) &= p(\mu \mid \Lambda) p(\Lambda) = \cN(\mu \mid \bs{m}_0, (\beta_0\Lambda)^{-1}) {\cal W}(\Lambda \mid \bs{W}_0, \nu_0) \p
 \end{align*}
 
-This type of IFS is what we will use as a formalism for modeling datasets with fractals. See figure\ref{fig:measures} for some examples IFS measures.
+Where the last line, a Gaussian-Wishart prior, applies to each pair of $\mu$ and $\Lambda$.
 
-\subsection*{Representing IFSs}
+\subsection{The Inference algorithm for IFS models}
+We assume the following factorization for our variational distribution to approximate  $p(\bs{z}|\bs{x})$:
+\begin{align*}
+q(& \mu_\text{pre}, \Lambda_\text{pre}, \bm, \bD, \bs{\sigma}, \pi, d) \\ 
+=\;&  q_{\mu_\text{pre}, \Lambda_\text{pre}}(\mu_\text{pre}, \Lambda_\text{pre}) \; q_{\bm, \bD}(\bm, \bD)\; q_{\bs{\sigma}}(\bs{\sigma}) \; q_{\pi, d}(\pi, d) \p \\
+\end{align*}
+For the sake of clarity, we will omit the subscripts from the factors in the following.
 
-As discussed, an IFS is completely determined by its component transformations and their weights.
-The definition does not specify any kind of transformation, so to make fitting IFSs to data feasible we must fix a family of transformations which we can search.
-For the purposes of this paper we will use \emph{similarities} (also known as \emph{similitudes}): maps on ${\mathbb R}^d$ consisting of a translation, a rotation and a uniform scaling.
+\subsection{Update rules}
 
-The method described below can easily be extended to other families of transformations such as affine transformations or neural networks.
-The only requirement is that we must have some way of finding a transformation $F$ within this family given two ordered point sets $\{x_i\}, \{y_i\} \subseteq {\mathbb R}^d$ so that the error is minimized.
-For similarities there is a direct and optimal method.
-\cite{umeyama1991least}
+\paragraph{Responsibilities}
 
-\subsection*{Drawing points from an IFS distribution}
+The derivation of the update rule for the responsibilities, ie.
+$q_{bs{\sigma}}^{n}(\bs{\sigma})$ takes the same form as the case for the mixture of Gaussians model \cite{}, since our model is essentially a mixture of $k^d$ Gaussians.
+Note that we can visualize $q(\bs{sigma})$ as a large matrix with $n$ rows and $2^d$ columns.
 
-For various reasons, it can be useful to generate a set of points drawn from an IFS probability distribution.
-The most common method to achieve this is known as the \emph{chaos game}.
-The algorithm is simple: we start with some initial point $x_0$ we choose a random component $S_i$ from the IFS (so that $S_i$ is chosen with probability $w_i$) and and let $x_1 = S_i(x_0)$.
-We iterate this process and discard the first 50 or so points.
-The rest of the sequence $x_n$ can be used as a set of points drawn from the IFS.
+We have:
+\begin{align*}
+q^n(\bs{\sigma}) &= \langle \ln p(x, \bm, \bD, \mu', \Lambda', \bs{\sigma}, \pi, d) \rangle_{\bm, \bD, \mu', \Lambda', \pi, d} + const \p
+\end{align*}
 
-It should be mentioned that the sequence of points generated this way does not strictly represent a sequence of independent draws.
-We can only say that the natural measure of this process coincides with the IFS.
-However, any two successive points in the sequence will satisfy the relation $x_{n+1} = S_i(x_{n})$ for some $i$.
-This is not true for two independently drawn points.
-For our experiments we generate a sequence and then shuffle the data to remove this dependency where it is undesirable.
+We factorize $\ln p(\ldots)$ according to (\ref{line:factorization}), and absorb any terms that don't depend on $\bs{\sigma}$ into the constant:   
+\begin{align*}
+q^n(\bs{\sigma}) &= \langle \ln p(x \mid \bm, \bD, \mu', \Lambda', \bs{\sigma}) \rangle_{\bm, \bD, \mu', \Lambda'} +  \langle \ln p(\bs{\sigma} \mid \pi, d) \rangle_{\pi, d} + const \p\\
+& = 
+\end{align*}
 
-A more correct (and more expensive) way to generate random data from an IFS is to start with a draw from some initial distribution (eg.
-a multivariate normal) and apply $n$ transformations randomly chosen from $S$.
-The distribution of the resulting random variable will converge to $S$ with $n$.
-For multiple points the process can be repeated.
-This process adds a factor of $n$ to the time complexity of the chaos game, so it is not commonly used.
+\paragraph{Components}
 
-\subsection*{Code space}
-\label{sec:code-space}
+\paragraph{Component weights}
 
-\begin{figure}[b!]
-  \centering
-  \begin{subfigure}[b]{0.48\textwidth}
-    \includegraphics[width=\textwidth]{../img/sierpinski-codes.pdf}
-    \caption{}
-    \label{fig:sierpinski-codes}
-  \end{subfigure}  
-  \hspace{0.015\textwidth}
-  \begin{subfigure}[b]{0.48\textwidth}
-    \includegraphics[width=\textwidth]{../img/code-construction.pdf}
-    \caption{}
-    \label{fig:code-construction}
-  \end{subfigure}
+\paragraph{Pre-transformation}
 
-  \caption{\small (a) Codes of length three on the Sierpinski triangle and the subsets they code for. (b) The construction of a subset from its code.}
-  \label{fig:codes}
-\end{figure}
+\paragraph{Depth}
 
-Let $(\sigma_i)$ be a sequence with $\sigma_i \in [1, n]$ for some IFS $S$ with $n$ components.
-We can compose the components of the IFS acording to the elements of $\sigma$.
-Call this function $S_\sigma$:
-\[
-S_\sigma = S_{\sigma_1} \circ S_{\sigma_2} \circ \ldots \circ S_{\sigma_m}
-\]
+\subsection{Results}
 
-Since each $S_i$ is contractive, we can let the length of $\sigma$ go to infinity and get 
-\[
-S_\sigma(X) = x_\sigma  
-\] 
-That is, the sequence of functions maps any non-empty $X$ to a unique point $x_\sigma$.
-In the case of probability measures, every probability measure is mapped to a probability measure that gives the single point $x_\sigma$ full probability mass.
-Figure \ref{fig:codes} illustrates the concept: an Iterated Function System assigns a code to each point on its support.
-Finite codes map to subsets of the support.
-Since the size of the subsets decreases exponentially we can talk about the endpoint of a finite code without losing much accuracy.
-We will take the point to which the origin is mapped by $S_\sigma$ as the canonical choice of the endpoint of finite $\sigma$.
+\section{The RIFS model}
+\paragraph{Random Iterated Function Systems}
 
-Finding the endpoint for given code is straightforward, but we will also require the inverse process: given a point $x$ on or near the support of some IFS, we wish to find $\sigma$.
-To this end we start with the standard  multivariate normal distribution and apply all maps $S_\sigma$ for all $\sigma$ of a given length $l$ giving us a set of $k^l$ normal distributions (since an affine transformation of a normal distribution is also a normal distribution).
-The $\sigma$ which assigns $x$ the greatest probability density is chosen.
+\subsection{The Inference algorithm for RIFS models}
+\subsection{Results}
 
-Since the variance of the distribution drops off exponentially as the length $l$ is increased, this method will fail if points are too far off the support of the IFS (even if logarithmic probabilities are used).
-To increase the range of the method, we fall back on the distance from the point to the mean of the distribution.
-If the first method cannot distinguish between codes, we choose the code whose probability distribution has its mean closest to $x$.
+\section{Discussion}
 
-A word of caution on codes is appropriate: the ordering is often the reverse of what our intuition suggests.
-If we number the three components of the Sierpinski triangle clockwise from the top 0, 1 and 2, we might number a small section of the triangle by first moving into the top subtriangle, then into the left subtriangle (2) of that and then into the right subtriangle (1) of that.
-Intuitively, this is the most natural way of coding sections of the Sierpinski triangle.
+\pb{Recap.}
 
-If we wish to map the whole triangle to this subset, we must reverse the codes: the transformation that maps to the smallest subsection must be applied first.
-The sequence of transformations which will map the whole triangle into the required subspace is:
-\[
-S_1(S_2(S_0(x)))
-\]
-so its code $\sigma$ is $(1, 2, 0)$. If we wish to `travel into' the Sierpinski gasket by this code, we must read it right-to-left.
-
-Finally, we wish to note that a line segment, the square, the cube and higher dimensional analogues can be represented as simple IFSs of $2^d$ components.
-For instance, for the unit line, we map the whole onto its lower and upper half with two components.
-The codes of this IFS are binary strings and the subsets they code are the diadic intervals.
-For the square, we can map the whole onto each one of its quadrants and get quaternary codes.
-Using this mechanism we can always map any dataset into the bi-unit hypercube and find a straightforward coding for the instances in $2^d$-ary codes.
-What fractal fitting offers us is a potential for a subspace that fits the data more compactly (ie.
-using fewer components) and gives us a good code of lower arity.
-
-\subsection*{Evaluating models}
-
-The standard way of evaluating the fit of a model to a dataset is the likelihood of the model given the data.
-This is often used in optimization algorithms for fitting probability models to data.
-In the case of fractal models, the likelihood function has an important drawback.
-
-Consider a dataset drawn from the Sierpinski gasket with a minute amount of two dimensional noise added.
-Since the Sierpinski gasket has dimension less than two, the probability that the point will still fall within the Sierpinski gasket will be zero.
-Thus all points will have probability zero under the Sierpinski gasket and its likelihood will be zero.
-Intuitively however, the Sierpinski gasket is still a good model for this data.
-
-For this reason we will often use the Hausdorff distance to evaluate our models. The Hausdorff distance  $H(X, Y)$ between two sets $X$ and $Y$ is defined as follows:
-\[ 
-H'(X, Y) = \max_{x\in X} \, \min_{y\in Y} \, d(x, y) 
-\]
-\[
-H(X, Y) = \max\left (H'(X, Y), H'(Y, X)\right )
-\]
-where $d(\cdot, \cdot)$ is the Euclidean distance. For infinite sets the $\max$ and $\min$ functions should be replaced with $\sup$ and $\inf$.
-
-The most intuitive way of understanding the Hausdorff distance is as a game where the first player must get from one set to the other in as small a distance as possible, and the malevolant second player will choose his starting point to maximize that distance.
-
-Hausdorff distance is a common tool in dealing with the fractal inverse problem \cite{nettleton1994evolutionary}.
-We will use it to evaluate an IFS model by generating a number of points from the model and calculating the Hausdorff distance of that set to the model.
-Since the Hausdorff distance has not been substantiated as a statistical measure, we can only use it as an intermediate indication of model quality.
-To prove the utility of IFS models we will test them on classification and dimensionality reduction tasks.
-
-\subsection*{Dimension}
-
-The \textit{curse of dimensionality} is a blanket term for many effects related the analysis of high dimensional data.
-The main cause of these problems is that as dimensionality increases, the volume of a subspace (such as the unit ball) increases exponentially with it.
-This means, for instance, that vast amounts of samples are required to maintain the same level of coverage in a high dimensional space as compared to a low dimensional one.
-
-The solution to many of these problems is that while the data has a high-dimensional representation, it has a low intrinsic dimension.
-That is, it can be mapped to a low dimensional space without any loss of information.
-A simple example is that of all points lying on a line in some high dimensional space.
-
-IFSs have a particular advantage over classical models like Gaussian mixtures: their intrinsic dimension varies over the whole range of positive reals.
-A search through the space of IFSs can return models of any dimension.
-
-We measure the dimension of datasets and data generated from models using Takens' estimator \cite{takens1985numerical}:
-\[
-D(d_{\mbox{max}}) = -\frac{|A|}{\sum_{a \in A} \ln a} 
-\]
-Where $A$ is the set of distances between pairs of data points lower than parameter $d_{\mbox{max}}$.
-We choose $d_{\mbox{max}}$ as the value that produces the lowest Kolmogorov-Smirnov statistic with the probability distribution underlying the Takens estimator.
-This method is based on the work for general power laws in \cite{clauset2009power} and will be detailed in a later publication.
-
-We estimate the uncertainty of this estimate as follows: bootstrap a version of the dataset by sampling to the same size with replacement, and estimate the dimension of that dataset.
-We repeat the process 50 times and take the sample standard deviation as an estimate of the uncertainty.
-
-\subsection*{Analogous model: Mixture of Gaussians}
-
-To study the IFS as a statistical model, it helps to have a commonly used analogous model to compare it with.
-We use the Mixture of Gaussians (MoG) model.
-This model consists of $n$ multivariate Gaussians over ${\mathbb R}^d$ each represented by a mean $\mu \in {\mathbb R}^d$ and a $d \times d$ covariance matrix $\Sigma$.
-Each component also has a prior $w_i$.
-The complete model describes the following probability distribution:
-\[
-p(x) = \sum_{i \in [1,n]} w_i N(x|\mu_i, \Sigma)
-\]
-Since every multivariate Gaussian can also be described as an affine transformation of the standard normal distribution (mean at the origin and $\Sigma$ the identity matrix), we can also represent the mixture of Gaussians as $n$ affine transformations and $n$ weights, just as the IFS:
-\[
-p(x) = \sum_{i \in [1,n]} w_i S_i(N(0, I))(x)
-\]
-This gives us a classical model with a similar representation.
-Note that the representation would be equal if we used similarities for the MoG model.
-However since an MVN transformed with uniform scaling is rotationally symmetric that would make all the rotational parameters useless.
-For this reason we will allow the MoG model full affine transformations as parameters and the IFS model only similarities.
-Even with this discrepancy, we still consider the MoG to be a reasonable analog.
-
-Unlike the single MVN distribution, there is no closed form solution for the maximum likelihood fit of a MoG model to data.
-We must approximate it.
-The most common method is by the well-known EM algorithm.
-Since our solution to the fractal inverse problem takes inspiration from this algorithm, we describe it here briefly.
-The key intuition behind the EM algorithm is that if we knew which component of the MoG model `generated' each datapoint, we would have a closed form solution for the parameters of the component, and if we knew the parameters of the components we could find out which component was most likely to have generated the data point.
-We turn this circular dependency into an iterative algorithm.
-Starting with some initial MoG model we iterate the following steps
-
-\begin{description}
-  \item[Expectation] Assign each data point a probability distribution over the components of the MoG model proportional to the likelihood of the model given the datapoint.
-In other words, each component takes a certain amount of `responsibility' for the data point.
-  \item[Maximization] Estimate the parameters for each component based on the probability weights calculated in the previous step.
-\end{description}
-
-\section*{Fitting IFS models to data}
-
-We can now describe the central algorithm of this paper.
-To make the explanation as clear as possible we will first describe a skeletal version of the algorithm that works in principle and then describe various alterations to this skeleton that are required to improve the performance of the algorithm on real data with finite approximations of the IFS measures.
-
-The results described are always for the full algorithm with all improvements.
-An implementation of only the basic algorithm should be enough to retrieve the Sierpinski gasket (or a crude approximation) from data.
-
-\subsection*{The basic algorithm}
-
-As with the EM algorithm for MoG models, the main barrier to inducing an IFS model from the data directly is a dependency on latent variables.
-In the MoG case each data point needed to be assigned to a component, in the IFS case we need to know, for each data point the code $\sigma$ that represents the sequence of components.
-
-Given these codes it's a relatively simple matter to find which points maps to which under the data's self-similarity.
-And given two ordered lists of points that are supposed to map into one another we can easily find a map to minimize the error.
-Thus our algorithm works as follows.
-\footnote{the names `expectation' and `maximization' of these steps are only by analogy with the EM algorithm for MoG models. We do not claim any true expectation or maximization procedures.}
-
-\begin{itemize}
-\item We start with in initial IFS $S$, chosen by one of the methods described below. 
-\item (Expectation) We assign each datapoint a code $\sigma$ of length $d$ (where $d$ is a parameter of the algorithm). The codes are provided by the model $S$ as described above. 
-\item (Maximization) We know that if a point $x_\sigma$ is mapped using one of our components we have $S_i(x_\sigma) = x_{i\sigma}$.
-That is, the symbol for the mapping component is prepended to the code.
-For each $x_\sigma$ in our dataset we search the data for a point with a code $\tau = i\sigma_{1:k-1}$, the symbol $i$ followed by the elements of $\sigma$ save for the last.
-For each $i$ we maintain ordered lists of the matches found.
-\item (Maximization) For each component $S_i$ we now have two lists of points such that $S_i$ should map one into the other with minimal squared error.
-We used the method described in \cite{umeyama1991least} to find $S_i$.
-\end{itemize}
-
-We will often use a subsample of the data, to allow us to increase the depth or number of components. In such cases, we reample at each EM iteration.
-
-\subsection*{Improvements}
-
-\subsubsection*{Centering and scaling the data}
-
-Iterated function systems are not very good at approximating data that is weighted far from the origin.
-For this reason we find the affine transformation which maps the data to the range $[0,1]$ in each dimension.
-
-We store this transformation with the model to map the model back to the data space as needed.
-
-\subsubsection*{Using all depths}
-
-The algorithm as described here can work at all depths $d$.
-Since each depth may contain information about self similar scaling in the data, we also want to consider all depths less than $d$ rather than just $d$.
-This is easily done by considering codes of all length up to and including $k$.
-We build a code tree of the codes generated from the dataset and search it systematically for the required matches at all levels.
-
-% This means that for each $i$ we have two sets of points at each depth.
-If the components are equally weighted, we will have a set of about $1/n$ of the data, mapping into a set of about $1/{n^2}$ of the data.
-At depth two the ratios become $1/{n^2}$ and $1/{n^3}$ and so on.
-For large depths, only specific codes will have matching points.
-
-\subsubsection*{Weights}
-
-The algorithm as described above does not give us a way to retrieve the component weights.
-For the coding step we already take the weights into account, so the only thing we must do is retrieve them from the set of codes with which we start the maximization step.
-
-If we have a collection of points $X_\sigma$ with code $\sigma$ and a collection of points $X_\tau$ with code $\tau$ such that $S_i$ should satisfy $S_i(x_\sigma) = x_\tau$, we know that if we generate the IFS model, $w_i$ of the probability mass at stage $t$ always gets distributed to component $S_i$ for stage $t + 1$.
-Thus we can expect the relative size of these sets to follow $w_i$: $|X_\tau| = w_i |X_\sigma|$.
-Over all such sets $X_\sigma^j$ and $X_\tau^j$ we must minimize the squared error:
-\[
-\overline{w_i} = \argmin_{w_i} \sum_{j} \left( |X_\tau^j| - w_i |X_\sigma^j| \right)^2
-\]
-This works out as 
-\[
-\overline{w_i} = \frac{\sum_{j}|X_\sigma^j||X_\tau^j|}{\sum_{j}|X_\sigma^j|^2}
-\]
-
-\subsubsection*{Multiple matches ($\sigma$)}
-
-It is perfectly possible that we get multiple points per code.
-Indeed, at low depths, it is almost certain.
-In this case we have two sets of points $X_\sigma$ and $X_\tau$ such that the points in the former should map onto the latter.
-But our algorithm for finding the map $S_i$ requires the individual points to be matched to each other.
-If the number of points is low we can simply take the mean of both sets, but if the number of points is high, we risk throwing away a lot of information (in particular about rotation).
-Theoretically each point in the first set should map onto every point in the other but this creates an explosion of possiblities.
-If we pair the points up randomly we introduce artifacts in the induced rotation.
-
-We tackle this problem by finding the MVN for both sets.
-Ideally we would then extend our method for finding maps to paired MVNs rather than paired points, but this is a highly complex problem.
-Instead, we make use of the correspondence between MVNs and affine maps.
-We find for each MVN the transformation that transforms the standard distribution ($N(0, I)$) into that MVN.
-We then generate $|X_\tau|$ `spanning points' from some initial distribution and map these to both sets.
-We add these point pairs to the list of matched points from which the maps will be calculated.
-
-To generate the spanning points we use a spherical MVN with variance $\sigma$.
-This value becomes an important parameter in the algorithm.
-If the spanning points have a high variance, then the rotation becomes a more important part of the found affine map.
-The limiting case with $\sigma \rightarrow 0$ is essentially equal to just using the mean of all points with a given code and getting no rotation information from the mapping between single code pairs.
-
-The behaviour of the algorithm under varying $\sigma$ is somewhat mysterious, but trial-and-error testing suggests that for high values the algorithm is more likely to converge.
-This comes at the cost of some solutions, as with $\sigma = 1.0$, the algorithm is unable to find the two-component Koch curve.
-For all experiments reported in this paper we use $\sigma = 0.1$.
-
-\subsubsection*{Edge cases}
-
-As with standard EM algorithms, some undesired behavior can emerge when the initial model is far from the optimum.
-Consider for instance the case where the Sierpinski triangle is the current model and all data is to the top right of the triangle.
-The symbol 2 would not occur in any of the codes and we would have no basis to find $S_2$ in the maximization step.
-If this occurs, we will `split' one of the components that could be found  (distributing the weight equally) and perturb each component slightly to ensure that the components differ.
-The idea behind this approach is simple.
-For every place in every code that used to be assigned to the original component there is now around a fifty/fifty probability of the place being assigned to the other code.
-In the next step roughly half of the matches of the component will have been `donated' to the other component.
-
-The perturbation is done by flattening the model into a vector of values in ${\mathbb R}$ and adding a small random value $N(0, \pi)$ to each. In our experiments $\pi$ was set to 0.3.
-
-\subsubsection*{Use the data as a basis}
-
-When we generate a sample from the model (to a fixed depth, rather than with the chaos game), and when we compute the probability density to a fixed point, we make use of some initial normal distribution.
-As the depth increases the influence of the choice of initial distribution increases exponentially.
-For low depths, however, our choice can affect the quality of our model.
-
-To determine our what our initial model should be, we examine the situation for depth one.
-In this case, our data points get a code of one symbol, and the transformations induced are those that optimally map the whole dataset to these subsets.
-The equivalent MOG model fits a normal distribution to each subset.
-For these normal distributions to be the result of a depth one IFS model, we must choose as its initial distribution the MVN fit to the data.
-With this choice, the MOG model and the MVN model are equivalent (so long as they use the same class of transformations).
-
-To summarize, we choose as our initial distribution an MVN fit for the data restricted to our transformation class.
-Since we are using Similitudes, we fit a symmetrical MVN to our data, and use that as the initial distribution.
-
-% \subsubsection{Branching}
-
-% [@! Test if this helps]
-
-%As an optimization algorithm, EM is strictly linear in the sense that it only considers a single model at each iteration.
-To avoid local minima it can often help to evaluate multiple models per iteration.
-To test whether our algorithm benefits from such additions we introduce a branching mode which works as follows.
-
-%During the iteration of the algorithm we maintain a buffer of models.
-After each iteration, we generate $b$ versions of the current best model perturbed by a parameter $\sigma$ as with the edge cases.
-We add these models to the buffer and sort it by Hausdorff distance with the data (as described in section @!).
-We discard all but the top $h$ models.
-At the next iteration, we continue with the top model from the list and repeat the procedure.
-
-%This method is analogous with beam search.
-Its advantage is that we perform the expensive EM step the same number of times as the non-branching model, and the extra cost lies purely in testing evaluating the extra models.
-(This is not inexpensive in itself, but any branching algorithm will need to test candidate models anyway).
-
-\subsection*{Choosing the initial model}
-
-The choice of initial model is crucial for achieving good performance
-
-\subsection*{Convergence}
-
-\section*{Results}
-
-We perform a number of experiments to give an impression of the behavior of IFSs as statistical models, and the EM algorithm as a learning method.
-
-First, we show a selection of results on 2D datasets to give a visual impression of the behavior of the algorithm.
-These images represent the value of the EM algorithm over earlier approaches, as methods based on evolutionary algorithms commonly fail to induce fractals from a fractal dataset with this speed or accuracy.
-
-We then investigate the use of IFS models on higher dimensional data.
-To attempt to visualize the way an iterated function system models a high dimensional dataset, we use a four component model so that we can map the data back to two dimensions and plot it.
-
-\subsection*{Datasets}
-
-\begin{description}
-  \item[forest]
-  \item[census]
-  \item[abalone]
-\end{description}
-
-\subsection*{Model fitting and dimension scaling}
-
-\begin{table}[p]
-\begin{tabular}{c c c c c}
-\hline
-data & initial & & & best model \\
-\hline
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/data.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/generation0000.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/generation0001.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/generation0010.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinski/generation0032.png}
-\\
- & & 6 & 48 & 83 \\
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/data.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/generation0000.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/generation0003.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/generation0020.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/sierpinskioff/generation0079.png}
-\\    
- &  & 3 & 20 & 79 \\
-	\includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/data.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/generation0000.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/generation0008.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/generation0022.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/koch2/generation0078.png}
-\\    
- &  & 8 & 22 & 78 \\
-	\includegraphics[width=0.2\textwidth]{../img/runs2d/ball/data.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/ball/generation0000.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/ball/generation0003.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/ball/generation0024.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/ball/generation0072.png}
-\\   
- &   & 3  & 24 & 72 \\
-	\includegraphics[width=0.2\textwidth]{../img/runs2d/coast/data.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/coast/generation0000.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/coast/generation0008.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/coast/generation0020.png}
-& 
-    \includegraphics[width=0.2\textwidth]{../img/runs2d/coast/generation0069.png}
-\\
- &  & 8 & 20 & 69 \\
-\hline
-
-\end{tabular}
-\caption{Selected training steps of the algorithm for various data sets.
-All models were trained with a dataset of 100000 points, sampling 300 per iteration for 100 iterations.
-Model depth was 6 for all runs.
-1000 points were sampled from each model and from the dataset to estimate the Hausdorff distance (and determine the `best model').
-The blue squares indicate the transformations (of the red square) that make up the model.
-The green bar indicates component weight.
-By row: (1) The sierpinski triangle, $\sigma = 0.1$ (2) The sierpinski triangle with unequal weights.
-$\sigma = 0.1$ (3) The 2-component koch curve.
-$\sigma = 0.001$ (4) A uniform distribution over the bi-unit ball (3 components, $\sigma = 0.1$) (5) Points sampled from a small stretch of coast line (3 components, $sigma = 0.1$).
-}
-\label{tablelabel}
-\end{table}
-
-[@!
-EXPERIMENT for various datasets, for various depths and component sizes, scatter plot the difference between the model dimension and the data dimension against the model accuracy (by hausdorff distance) show uncertainty with error bars]
-
-First, we fit IFS models to our datasets (ignoring the class information).
-We report the Haussdorff distance of a sample of 1000 points from the data to 1000 points generated by the model (we consider the iris dataset to be too small for this purpose).
-We withhold a random 25\% of the data for testing, repeat the experiment 20 times and report the sample standard deviation and the mean of the resulting values for symmetric error.
-
-We are also interested in the dimensional scaling of the model.
-In this case we are interested in the training behavior of the model, so we do not withhold test data.
-We measure the dimension of the dataset on a random sample of 300 points (repeating 25 times to measure the uncertainty) and generate 300 points from the best model and measure the dimension in the same way.
-We expect that the difference between the model dimension and the data dimension depends on the quality of the model.
-Figure @!
-plots the data dimension against the model dimension.
-
-In all cases we train the model for 300 generations and select the best model by Hausdorff distance.
-
-\begin{figure}[htb]
-\centering
-
-  \begin{subfigure}[b]{0.66\textwidth}
-	\includegraphics[width=\textwidth]{../img/dimension/out-table.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.33\textwidth}
-    \begin{tabular}{c c }
-    data & model \\
-	\hline
-		\includegraphics[width=0.5\textwidth]{../img/dimension/data-henon.png} & 
-		\includegraphics[width=0.5\textwidth]{../img/dimension/model-henon.png} \\
-		\includegraphics[width=0.5\textwidth]{../img/dimension/data-ikeda.png} & 
-		\includegraphics[width=0.5\textwidth]{../img/dimension/model-henon.png} \\
-		\includegraphics[width=0.5\textwidth]{../img/dimension/data-henon.png} & 
-		\includegraphics[width=0.5\textwidth]{../img/dimension/model-henon.png} \\
-	\end{tabular}
-  \end{subfigure}
-  \caption{Dimensional scaling. The marker size and color both indicate the quality of the model fit by Hausdorff distance. }
-  \label{fig:dim_scale}
-\end{figure}
-
-\subsection*{Classification}
-
-We explore two methods for using IFS models for classification.
-In the first, we train a single model for each class and combine in a straightforward manner using Bayes' rule.
-In the second, we learn a single model for the whole set of points, and use the codes generated by the model to perform classification.
-
-\subsubsection*{Bayesian}
-
-Given some data set $X = \{x_i\} \subseteq {\mathbb R}^d$ with associated classes $c_i \in \{1, 2, \ldots, m\}$, we split the data by class: $X_c = \{ x \in X \mid c_i = c\}$.
-We train a probability model $p_c$ for each $X_c$ and classify each new point with
-\[
-C(x) = \argmax_{c} p_c(x) p(c)
-\]
-The class prior $p(c)$ is estimated as $\frac{|X_c|}{|X|}$.
-
-\subsubsection*{Single model}
-
-An alternative way to construct a classifier with IFS modelling is to learn a single model for the whole dataset, ignoring class labels.
-For some datasets, the set of all points may present a more coherent and simple distribution than the sets split by class label.
-Another advantage of this approach is that we only have to learn one model, which is cheaper both in training and in evaluation.
-
-The model provides us with a mapping from the data space (${\mathbb R}^d$) to code space ($\{1, \ldots, k\}^d$).
-To turn this model into a classifier we proceed as follows.
-We build a code tree such that each possible code represents a path from the root to the leaf node.
-We then traverse this tree for each point in the training set, maintaining a frequency count for each class observed at each node.
-Thus, the root node gives us the frequency distribution over the classes for the whole dataset and each node below it refines that to a distribution over a section of the dataset whose code starts with a given prefix.
-This turns each node into a very basic classifier, which simply returns the class with the highest frequency count.
-The root node will give us the classifier that uses nothing more than class frequencies and each node below it will use more information about the neighbourhood of the point.
-
-For each point to be classified we take the probabilities at each node in the path and average them evenly.
-
-\begin{table}[h]
-\begin{tabular}{| l | r l r l | r l r l  r l r l |}
-\hline
-  & ifs bayes & & ifs single & & mog & & ann & & knn & & tree & \\
-\hline
-census  & & & & & & & 0.054 &\color{gray}(0) & 0.06 & \color{gray} (0.002) & 0.05 & \color{gray} (0.001) \\
-forest  & & & & & & & 0.163 &\color{gray}(0.001) & 0.138 &\color{gray}(0.001) & 0.12&\color{gray}(0.003) \\
-abalone & & & & & & & 0.47 & \color{gray}(0.008) & 0.47 & \color{gray}(0.006) & 0.48 & \color{gray}(0.01) \\
-mnist   & & & & & & & 0.09 & & 0.05 & & 0.20 & \\
-\hline
-
-\end{tabular}
-\caption{Hausdorff distances for models trained against various datasets.}
-\label{classification-results}
-\end{table}
-\subsection{Time performance}
-
-An informal analysis of the algorithm shows that the expectation step costs $O(k^dn)$, with the maximization step taking roughly $O(n^2)$.
-This falsely suggests that the expectation step, being linear in the number of data points, is negligable compared to the maximization step, which is quadratic.
-Of course, when viewed in terms of the parameters the expectation step is exponential and this is what causes the expectation step to become the bottleneck for anything but very low values of $k$ and $d$.
-
-\section*{Conclusions}
-
-We have presented a framework for a fractal approach to the statistics of point patterns.
-We introduced an EM-style algorithm for fitting fractal models to data.
-Compared to search algorithms for other types of models the algorithm is relatively expensive, but compared to other approaches for fractal modeling the algorithm represents a significant improvement @!.
-
-Further, we have given some indication of how the study of fractal modeling can open up interesting new prospects in machine learning.
-We expect that these methods will allow many avenues of future research.
-One option would be to investigate the use of more expressive maps.
-The similitudes used here could be replaced by neural networks, for instance.
-The EM search algorithm is fairly simple, and could be extended to try various branches and backtrack if it deviates too far from a good model.
-
-\subsubsection*{Other fractals}
-
-It is important to note that there are many other families of fractals besides iterated function systems.
-They are not covered by our method, and cannot be properly represented by Iterated Function Systems.
-Examples include Julia sets and dynamical systems.
-
-There is one family of fractals in particular that would be very interesting to explore in light of this research, and that is the family of random fractals.
-When we think of natural phenomena that seem to resemble iterated function systems---coast lines, trees, the veins of a leaf---there is always a combination of fractal scaling and some random component.
-The framework most likely to capture these phenomena is that of random Iterated Function Systems, as described in \cite{hutchinson2000deterministic}.
-
-The random IFS model is an extension of the Iterated Function System which produces not one single set but a distribution over all sets.
-The dataset for which to learn a random fractal would not consist of one list of vectors, but of a list of lists of vectors.
-For instance, a collection of timeseries preceding a critical event, or a collection of images of snowflakes.
-In \cite{bloem2010fractal} we showed a proof-of-concept, indicating that such models can be learned by evolutionary algorithm in principle, but we did not advance beyond the simplest conceivable model.
-If the EM algorithm described here could be extended to random fractals, it would open up a new type of machine learning task, where instances ae represented by point patterns, and the model to be learned is a distribution over distributions.
+\pb{}
 

### 2015-11-06 fractal paper/src/fractal.tex
...

@@
 \usepackage{subcaption}
+\usepackage{hyperref}
+\usepackage{float}
+
+\floatstyle{ruled}
+\newfloat{pseudo}{h}{lop}
+\floatname{pseudo}{Algorithm}
+
+\usepackage{eulervm}
+\usepackage{charter}
 
@@
 \newcommand{\bD}{\boldsymbol\Lambda}
+\newcommand{\R}{\mathbb{R}}
 \newcommand{\bm}{\boldsymbol\mu}
@@
 \newcommand{\bs}[1]{{\boldsymbol {#1} }}
+\newcommand{\vc}{\textnormal{vec}}
+\newcommand{\tab}{\hspace*{5mm}}
+\newcommand{\ol}[1]{\overline{#1}}
 
-\title{A variational algorithm for the Fractal Inverse Problem.}
+\theoremstyle{definition}
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}
+\newtheorem{corollary}{Corollary}
+\newtheorem{lemma}{Lemma}
+\newtheorem{example}{Example}
+
+\title{An Expectation-Maximization Algorithm for the Fractal Inverse Problem}
 \date{\today}
@@
 \begin{abstract}
-\noindent We present an algorithm for solving the \emph{inverse problem} for iterated function systems (IFS) and random iterated function systems (RIFS).
-The central idea is to cast the IFS as an generalization of a mixture of of Gaussians model, and to treat the particular sequence of components that leads to a particular point as a latent variable.
+\noindent We present a expectation-maximization algorithm to solve the \emph{inverse problem} for iterated function systems (IFS): the task of fitting a fractal model to a given dataset.
+An IFS is defined by a small number of maps on $\mathbb{R}^H$, the IFS's \emph{components}.
+A point is sampled from the IFS by starting with some initial point and applying a sequence of randomly chosen components.
+The central idea of our algorithm is to treat this sequence as a latent variable.
+If we are given such a \emph{code} for each point in the data, we know which points should map to one another under each transformation of the IFS, allowing us to reconstruct the components.
+Given the components, we can easily assign codes each point.
+Iterating these two steps provides us with a basic algorithm to fit an IFS model to data.
+This idea can be cast as an instance of the Expectation-Maximization (EM) algorithm.
+Our algorithm  quickly and accurately determines the IFS behind a given dataset.
+The model reduces to a simple mixture of Gaussians when the data contains no self-similarity.
+In this treatment, all components are similitudes, but extensions to more general families are simple.
 \end{abstract}
+
 \section{Introduction}
-\pb{Briefly: importance of fractals, inverse problem.}
+What constitutes a fractal is not precisely defined\footnotemark. The following properties, however, are common to all examples.
 
-\pb{Iterated Function Systems.}
+First, there is \textbf{self similarity}: a part is a scaled down copy of the whole.
+See for instance the Sierpinski triangle (fig.
+\ref{figure:codes}).
+It consists of three triangular shapes which are scaled down copies of itself.
+Second, fractals tend to posess \textbf{infinitely fine structure}.
+'Zooming in' reveals ever finer detail.
+In the case of the Sierpinski triangle, we will see the same shape recurring again and again, but other fractals like the Mandelbrot set reveal a great variety of shapes.
+Finally, fractals tend to have \textbf{non-integer dimension}. The Sierpinski triangle, for example, has a dimension of $1.58$.
 
-\pb{Idea behind our approach. EM principle, using variational Bayes.}
+\begin{figure}[b!]
+  \centering
+  \begin{subfigure}[b]{0.48\textwidth}
+    \includegraphics[width=\textwidth]{../img/sierpinski-codes.pdf}
+    \caption{}
+    \label{fig:sierpinski-codes}
+  \end{subfigure}  
+  \hspace{0.015\textwidth}
+  \begin{subfigure}[b]{0.48\textwidth}
+    \includegraphics[width=\textwidth]{../img/code-construction.pdf}
+    \caption{}
+    \label{fig:code-construction}
+  \end{subfigure}
+
+  \caption{\small (a) Codes of length three on the Sierpinski triangle and the subsets they code for. (b) The construction of a subset from its code.}
+  \label{figure:codes}
+\end{figure}
+
+Since the name fractal was coined in the 1970s, fractals have been seen as a potential model for many natural phenomena.
+Mandelbrot put it as follows in the \emph{The Fractal Geometry of Nature} \cite{mandelbrot1982fractal}:
+
+\begin{quotation}
+\small
+\noindent Clouds are not spheres, mountains are not cones, coastlines are not circles, and bark is not smooth, nor does lightning travel in a straight line.
+\end{quotation}
+
+Fractal geometry has been used in many fields, including physics \cite{mandelbrot1984fractals}, geology \cite{cheng1997multifractal}, biology \cite{goldberger1992fractal} and economics \cite{turiel2003multifractal}.
+
+The greatest problem with fractal analysis has always been the difficulty of finding a fractal model for a given set of data.
+It may be visually clear that a cloud or a coastline `looks' fractal, and we may be able to determine that it has a non-integer dimension, but how do we get from a dataset to a model?
+How do we determine the parameters of a fractal-generating model that lead to this particular dataset?
+Current approaches tend to rely on evolutionary algorithms \cite{deliu1991genetic}.
+Such techniques take a long time, and it can be difficult to get it to converge to precise parameter values, even if the dataset itself is sampled from a fractal model.
+Other approaches are highly domain-specific approaches such as the fractal image compression \cite{hart1996fractal}.
+Some interesting results have been derived from the method of moments \cite{rinaldo1994inverse} and sampling random transformations from the data \cite{hart1997similarity}, but so far these have not led to a practical algorithm.
+
+\footnotetext{Mandelbrot originally defined it as a set whose topological dimension differs from its Hausdorff dimension, but retracted this definition, stating that he preferred the word to be not precisely defined.}
+
+A very popular fractal model is the iterated function system (IFS).
+Many fractals, like the Sierpinski gasket, the Koch curve and the Menger sponge can be seen as IFSs.
+We will use the following definition:
+\begin{definition}
+An \emph{Iterated Function System} of order $K$ and dimension $H$ is a pair $(\{f_k\}, \{w_k\})$ of $K$ \emph{components} $f_k$ with $K$ associated\emph{weights} $w_k$.
+Each component is is a function $f_k: \R^H \to \R^H$, and each weight is a nonnegative scalar, with $\sum_i w_i = 1$.
+\end{definition}
+In this paper, all components are similitudes, ie. $f_k$ is defined by a scalar $s_k$, a rotation matrix $R_k$ and a translation vector $t_k$:
+\[
+f_k(x) = s_kR_k x + t_k \p
+\]
+The definition can be extended to generic functions and generic metric spaces \cite{hutchinson2000deterministic}, but this formulation is sufficient for our purposes.
+
+The IFS determines a probability distribution on $\R_H$ in the following way.
+Let $p_0$ be some initial distribution on $\R^H$ and let $D$ be some nonnegative integer.
+Sample a point $x_0$ from $p_0$ and pick a random component $f_k$ (according to probability $p(f_k) = w_k$).
+Let $x_1 = f_k(x_0)$.
+Sample another random $f_k$ and let $x_2 = f_k(x_1)$, and so on until $x_D$.
+
+There are two basic properties to of IFSs.
+First, as $D$ grows, the distribution on $x_D$ converges.
+We call the distribution it converges to the \emph{limit distribution}.
+Second, the limit distribution is independent of the choice of $p_0$.
+For formal statements of these properties and their proofs we refer the reader to \cite{hutchinson2000deterministic}.
+
+We can now frame the fractal inverse problem as a problem of statistical parameter estimation: we would like to find the IFS for which the probability of of the data under its limit distribution is maximal.
+The algorithm we describe is based on the following ideas:
+\begin{itemize}
+  \item Since the initial distribution doesn't matter, we can choose what we like.
+We choose the standard multivariate Normal (MVN) distribution $\cN_0$.
+The benefit of this choice is that every affine transformation of an MVN is itself an MVN.
+Thus, no matter how high we set $D$, the resulting distribution will always be a mixture of MVNs.
+  \item We assume that each point in the dataset was the result of a length $D$ sequence of applications of the components of the model.
+We cast this sequence---called a \emph{code} hereafter---as a latent variable in our model.
+Each code $c$determines an \emph{endpoint distribution}, and each endpoint distribution is an MVN $\cN_c$.
+  \item  Given the IFS that generated our data, it is a simple matter to find the most likely code for each point $x$.
+Conversely, if we know the most likely code for each point, we can determine which points in the dataset should map to one another, under a particular component $f_k$.
+This correspondence allows us to reconstruct $f_k$.
+\end{itemize}
+
+We make the correspondences in the last point probabilistic: instead of assigning each point its most likely code, we compute a \emph{responsibility} that each possible component takes for the point.
+This is the posterior probability of the component given the point: $p(c \mid x) \propto p(x \mid c) p(c)$, where $p(x\mid c) = N_c(x)$ and $p(x) = \prod_{s \in c} w_s$.
+This is the \emph{expectation} step of the algorithm.
+The \emph{maximization} step, optimizing the IFS model given these probabilities is detailed in the body of the paper.
+
+Note that finding a code corresponding to a particular endpoint can be counter-intuitive, as the first component applied determines the location at the smallest scale, and the last component applied determines the location at the largest scale.
+Figure~\ref{figure:codes} provides an illustration.
+
+We show in Section~\ref{section:experiments} that our algorithm quickly reconstructs many known fractals, such as the Sierpinski gasket and the Koch curve.
+We also apply the algorithm to some datasets sampled from images, showing that, in principle, it can find approximations of naturally occurring fractals, that are not specifically IFSs.
 
 \subsection{Preliminaries}
-We will deal with datasets which are a set of points $\{x_1, \ldots, x_n\}$ with $x_i \in \mathbb{R}^D$.
+\paragraph{Measures and transformations} Let $\{x_i\}_{i\in[1,N]}$, $x_i\in\R^H$ be our dataset. Let $X$ be the $N\times H$ matrix with $x_i$ as its columns. All vectors are column vectors.
 
-Let $F_{t, R}(y) = t + Rx$ be an affine transformation represented by a vector $t$ and a matrix $R$.
-If $X ~ \cN(\mu, \Sigma)$, then $F_{t, R}(Y) ~ \cN(t + R\mu, R\SigmaR^T)$.
-Specifically this means that if we take the standard multivariate normal $\cN_0 = \cN(0, I)$ as  canonical source, we can represent every multivariate normal $\langle \mu, \Sigma \rangle$ by an affine transformation $\langle t, R \rangle$ with $t = \mu$, $\Sigma = RR^T$.
-It also tells us that sampling from $\cN_0$ and applying a series of affine transformations is equivalent to sampling from a particular MVN.
+Let $1_L$ be the length-$L$ vector with $1$ at all indices.
+Such vectors are useful tools in dealing with matrices, for instance, the sum of the elements of the $K \times L$ matrix $M$ can be expressed as ${1_K}^T M 1_L$, while its marginal vectors are $({1_K}^T M)^T$ and $M 1_L$.
+Let $0_L^j$ be the length-$M$ vector with element $j$ $1$ and all others $1$.
+This vector functions as a row- or column-selector, eg.
+$X 0^j_N = x_i$.
 
-\paragraph{Variational Bayes} Let $x$ be the observed data of a problem, for which we have a model, and let $z$ be the latent variables of the model.
-If the model has any parameters, we place priors on them so that we can consider them latent variables, and absorb them into $z$ The variational Bayes approach attempts to approximate the distribution $p(z \mid x)$ using a \emph{variational distribution} $q(z)$.
-We assume that $z$ can be split into subsets $z_1, \ldots, z_m$, for which $q$ factorizes:
+Let $v(X)$ be a probability distribution with $X \subseteq \R^H$.
+Let $f_{t, A}(x) = Ax + t$ be an invertible affine transformation represented by a vector $t$ and a matrix $A$.
+Then the transformation of $v$ by $f_{t, A}$ is defined by the relation $f_{t, A}(v)(X) = v({f_{t, A}}^{-1}(X))$.
+Let $v(x)$ be the density function of  the measure $v$ (with respect to the Lebesgue measure).
+Then the density function of the measure $f_{t, A}(v)$ is $f_{t, A}(v)(x) = |A^{-1}| v({f_{t, A}}^{-1}(x))$.
+
+A multivariate normal distribution (MVN) on $\R^H$ is determined by a mean $\mu \in \R_H$ and a covariance matrix $\Sigma \in \R^{H\times H}$. Its probability density function is:
 \[
-q(z) = \prod_i q_i(z_i)
+\cN(x \mid \mu, \Sigma) = (2\pi)^{-\frac{H}{2}}|\Sigma|^{-\frac{1}{2}} \exp \left [-\frac{1}{2} \left \| (x-\mu)^T\Sigma^{-1}(x-\mu) \right\|^2\right] 
 \]
-A central result in variational Bayes is that the optimum for each factor is given by:
-\begin{equation}
-q^*_i(z_i) = \langle \ln p(x, z)\rangle_{/i} + \text {const} \p \label{line:var-bayes}
-\end{equation}
-Where $\langle \ln p(x, z) \rangle_{/i} = \sum \prod_{j = {1, \ldots, m}/i} q_j(z_j) ln p(x, z)$ with the sum over all values of all $z_j$ except $z_i$.
-That is, it represents the expectation with respect to all latent variable factors except $z_i$.
+We will call the MVN with $\mu = 0$, $\Sigma = I_H$ the standard multivariate normal distribution, or $\cN_0$.
+A \emph{spherical MVN} has $\Sigma = sI$ for some scalar $s$.
+Let $x$ be random variable with $x \sim \cN(\mu, \Sigma)$, then $f_{t, A}(x) \sim \cN(t + A\mu, A\Sigma A^T)$.
 
-Usually, the values of these optimal factors have cyclical dependencies, which means that we cannot compute $q^*$ directly.
-We can, however, search for a local optimum by iteration: we start with some initial value, and use (\ref{line:var-bayes}) to compute new values for each factor based on those of the previous generation.
+A \emph{rotation matrix} $R$ is a matrix with $R_R = I$ and $|R| = 1$.
+A \emph{similitude} (or rigid transformation) on $\R^H$ is a transformation defined by a rotation matrix $R \in \R^{H\times H}$, a translation vector $t \in \R^H$ and uniform scaling $s \in R$: $f_{t, R, s}(x) = sRx + t$.
+The inverse of a similitude is also a similitude: ${f_{s, R, t}}^{-1}(x) = f_{\frac{1}{s}, R^T, -\frac{1}{s}R^Tt}(x)$.
+
+$f_{s, R, t}(\cN_0)$ is a spherical MVN with mean $t$ and covariance $s^2I$. Transforming a generic spherical MVN by a similitude can be cast as the transformation of $\cN_0$ by two similitudes:
+\begin{align}
+f_{t, R, s}(\cN_{t_0, {s_0}^2 I})(x) &= f_{t, R, s}(f_{t_0, R_0, s_0}(\cN_0))(x) \notag \\
+&=  (2\pi)^{-\frac{H}{2}} \frac{1}{ss_0} \exp \left[- \frac{1}{2{s_0}^2s^2}\left\| x - (s R t_0 + t)\right \|^2\right] \label{line:inverse}
+\end{align}
+\paragraph{The EM Algorithm} \cite{dempster1977maximum}
+Let $p(X\mid Z, \theta)$ be probability distribution with parameters $\theta$ and latent variables $z$.
+The most common example is a mixture of MVNs: in this case $\theta$ contains the parameters of $K$ MVNs together with a weight $w_k$ for each, and $Z$ is a matrix whole columns $z_i$ determine which component is responsible for $x_i$: ie.
+if component $k$ is responsible, then $z_ki$ is 1 and all other $z_ki$ are 0.
+
+The maximum likelihood parameters for a dataset $X$ are determined by 
+\[
+\argmax_{\theta} \ln p(X|\theta) = \argmax_{\theta} \ln \sum_Z p(X|Z, \theta)p(Z) 
+\]
+where the sum is over all possible values of $Z$.
+The value in side the sum is the \emph{complete-data likelihood} $p(X, Z\mid \theta)$.
+In the mixture-of-MVNs example, there are $K^N$ possibilities for $Z$, making this sum intractable for practical data.
+
+The intuition behind EM is to maximize $\theta$ with respect to our best guess for $Z$ and vice versa.
+Let $\theta^\text{old}$ be our current best guess for the parameters.
+We first optimize $p(Z\mid X, \theta^\text{old})$, the posterior on the latent variables.
+Note that, in the MVN example, we can represent this as an $N\times K$ matrix $Z$ with $Z_{ij} = p(j\mid x,\theta) = N^\text{old}_j(x) w_j$.
+This is known as the \emph{expectation} step.
+
+We then determine the expectation of the logarithmic complete-data likelihood under  our posterior $p(Z|X, \theta^\text{old})$ as a function of $\theta$.
+\[
+Q(\theta) = \sum_Z p(X|Z, \theta)p(Z) \ln p(X, Z \mid \theta)
+\]  
+Note that the complete-data likelihood is now inside the logarithm.
+We optimize this $Q$-function, with respect to $\theta$ to get $\theta^\text{new}$, our new best guess for the parameters.
+This is known as the \emph{maximization} step.
+The EM algorithm iterates the two steps until the algorithm converges.
 
@@
 \begin{figure*}[t]
-  \includegraphics[width=\textwidth]{../img/factor-graph.pdf}
+  \center
+  \includegraphics[width=0.7\textwidth]{../img/factor-graph.pdf}
   \caption{A graphical model, illustrating the components of the IFS model.
@@
 
-The complete IFS model is illustrated in Figure~\ref{figure:ifs-diagram}. If contains the following components:
+The complete IFS model is illustrated in Figure~\ref{figure:ifs-diagram}. It consists of:
 \begin{description}
-\item[the component weights $\pi$] This is a vector of $k$ nonnegative real values, with $\sum_i \pi_i = 1$.
+\item[the component weights $w$] This is a vector of $K$ non-negative real values, with $\sum_k w_k = 1$.
 Each weight determines the probability of its associated component being chosen at each iteration of the IFS.
-\item[the depth $d$] The number of times the IFS is iterated.
-\item[the components $\bm = \{\mu_1, \ldots, \mu_k\}$, $\bD = \{\Lambda_1, \ldots, \Lambda_k\}$] The means and precision matrices for the $k$ components. 
-\item[The post-transformation $\mu_\text{post}$, $\Lambda_\text{post}$] The parameters for a single post-transformation.
-To see why this is necessary, consider the Koch curve in Ficgure~\ref{figure{fractals}}.
-While a dataset sampled from this measure can be prefectly represented by an IFS, it is slighty off-center.
-If this were real dataset, it might be mean-centered instead, in which case no IFS would fit.
-Unlike the mixture of Gaussians model, an IFS cannot be translated by a straightforward change of the parameters,a nd as the Koch curve shows, simple solutions like mean-centering the data do not always solve the problem.
-For this reason, we add a single post-transformation into the model, to allow for any simple, affine transformations to the final limit measure.
-This factor has the additional advantage that at $d = 0$, the model collapses to a single MVN distribution.
-\item[The transformation sequence $\sigma_i$] For each data point $x_i$, we have a sequence of length $d$ representing the sequence of component-transformations we need to apply to a point sampled from $\cN_\text{pre}$ to get the final point sampled from the limit measure of the IFS.
+\item[the depth  priors $v$] The number of times the IFS is iterated for each point.
+$v$ is a vector representing a categorical distribution on $[0,D]$, from which the length of the code $c$ is sampled.
+\item[the components $f_k = \{s_k, R_k, t_k\}$] $K$ similitudes. 
+\item[The post-transformation $s_p, R_p, t_p$] The parameters a single \emph{post-transformation}. This transformation is applied once to all sampled points.
+\item[The code $c_i$] An ordered sequence $c_i = \langle c_{i1}, \ldots, c_{id} \rangle$. Each element of the code is an integer in $[1,K]$ representing a component. 
 \item[The data $x_i$] This is the point after the post-transformation is applied: ie. the observed data.
@@
 
-The sampling procedure modeled
+To sample a point $x_i$ from this model we first sample a depth $d$ with $p(d=a) = v_a$.
+We then sample a sequence $c_i$ of $d$ components $c_{i1}, \ldots, c_{dj}$ with each $c_ij \in [1,K]$ chosen independently with $p(c_ij = a) = w_a$.
+We then create the composite function $f = f_p \circ f_{c_{i1}} \circ \ldots \circ f_{c_{id}}$ and sample $x$ from $f_(N_0)$.
 
-This gives us the joint distribution $p(\bs{x}, \bs{\sigma}, \mu_\text{pre}, \Lambda_\text{pre}, \bm, \bD, \mu_\text{post}, \Lambda_\text{post}, \pi, d)$ which factorizes as follows:
-\begin{align}
-p(&\bs{x}, \bs{x'}, \bs{\sigma}, \mu_\text{pre}, \Lambda_\text{pre}, \bm, \bD, \mu_\text{post}, \Lambda_\text{post}, \pi, d) \notag\\ 
-=\;& p(\mathbf{x} \mid \mathbf{x'}, \mu_\text{post}, \Lambda_\text{post}) \; p(\mu_\text{post}, \Lambda_\text{post}) \notag\\
-& p(\mathbf{x'} \mid \bs{\sigma}, \bm, \bD, \mu_\text{pre}, \Lambda_\text{pre}) \notag\\
-& p(\mathbf{\sigma} \mid \pi, d) \; p(\pi) \; p(d) \notag\\
-& \prod_i p(\mu_i, \Lambda_i) \p \label{line:factorization}\\
-\end{align}
+Giving the model a variable depth has several advantages.
+First, it makes the model a generalization of good fall-back models.
+with $v_0 = 1$, the model becomes a spherical multivariate Gaussian.
+With $v_1 = 1$, the model becomes a mixture of spherical Gaussians.\footnotemark Since each point has its own depth, the full model is a mixture over all these models, and the deeper IFS models.
+If the data is not self similar, or only partially self similar (like an IFS model with ourliers), the lower depth parts of the mixture can capture this.
 
-We use the following priors:
+This also allows the EM search a `gentle start'.
+In the limit, most IFSs have a support with lower dimension than the embedding dimension of the data: ie.
+almost all of $\R^H$ has probability zero.
+This means that even if the model fits the source of the data perfectly, the slightest addition of noise will make the entire likelihood zero.
+It also means that a minute change in parameters can mean the difference between the maximum likelihood, and likelihood zero.
+In effect, for high values of $d$, the fitness landscape becomes very jagged.
+
+The variable depth provides an automatic defense: if such a drop occurs, the lower models, whose fitness landscape is smoother, automatically get a higher posterior, smoothing out the fitness landscape for the complete mixture.
+As the search converges to the correct model, the depth prior becomes more weighted towards the higher depths.
+
+However, the variable depth also creates a problem: while any transformation of an IFS by a similitude is also an IFS, how the components of the original IFS are transformed depends on the parameter $D$.
+Imagine a dataset sampled from a Sierpinski gasket translated away form the origin.
+We have good models at every depth for this data, but their components are different.
+Only if the data is centered do the components coincide for all depths, but how the data should be centered depends on the parameters of the IFS.
+For this reason, we introduce the\emph{post-transform}:we learn a centered IFS, and transform the full limit distribution to fit the data.
+
+\footnotetext{To generalize to non-spherical MVNs and mixtures over non-spherical MVNs, the component family should be extended to positive definite affine functions.
+This extension is discussed in the conclusion.}
+
+\section{The EM algorithm for IFS models}
+
+We will use the notation $[1,K]^d$ to refer to the set of all length-$d$ codes.
+Let $[1,K]^{[a,b]} = \bigcup_{d \in [a,b]} [1,K]^d$.
+Thus, the set of all codes, including the empty code, is $[1, K]^{[0, D]}$.
+For every code $c_j$ in this set, we have
 \begin{align*} 
-p(\pi) &= \text{Dir}(\bs{\alpha}_0) \\
-p(d) &= {\cal U}(0, d_\text{max}) \\
-p(\mu, \Lambda) &= p(\mu \mid \Lambda) p(\Lambda) = \cN(\mu \mid \bs{m}_0, (\beta_0\Lambda)^{-1}) {\cal W}(\Lambda \mid \bs{W}_0, \nu_0) \p
+p(x \mid c_j) &= \cN(x \mid \mu_j, \Sigma_j) \\
+p(c_j) &= p(d) \prod_i p(c_j) = v_{|c_j|}\prod_{i \in c_j} w_{i}
 \end{align*}
 
-Where the last line, a Gaussian-Wishart prior, applies to each pair of $\mu$ and $\Lambda$.
+Where $\mu_j$ is the mean corresponding to $f_j(\cN_0)$ with $f_j = f_p \circ f_{c_{j1}} \circ f_{c_{j2}} \circ \ldots \circ f_{c_{jd}}$, and $\Sigma_j$ is the covariance matrix $f_j(\cN_0)$.
 
-\subsection{The Inference algorithm for IFS models}
-We assume the following factorization for our variational distribution to approximate  $p(\bs{z}|\bs{x})$:
+\subsection{The latent variables}
+
+The responsibilities for an MVN mixture model are well-known. Let $M = |[1, K]^{[0, D]}|$ and let $P$ be an $N\times M$ matrix with: 
+\[
+P_{ij} = p(c_j \mid x_i) = \frac{p(x_i\mid c_j) p(c_j) }{\sum_{a \in [1,K]^{[0, D]}} p(x_i\mid c_a) p(c_a)} \p 
+\]
+We also create a specific matrix for each $f_k$ for purposes that will become clear in the next section. Let $M = |[1, K]^{[0, D-1}]|$, and let $P^k$ be an $N \times M$ matrix with: 
+\[
+P_{ij}^k = p(k.c_j \mid x_i) = \frac{p(x_i\mid k.c_j) p(k.c_j) }{\sum_{a \in [1,K]^{[0, D]}} p(x_i\mid c_a) p(c_a)} 
+\]
+where $k.c_j$ is the code $\langle k, c_{j1}, \ldots, c_{j|c|}\rangle$. Note that this is a sub-matrix of $P$.
+
+Note that the size of these matrices will grow very fast with $D$, and most of its values will be practically indistinguishable from $0$ when normalized.
+For this reason, it may be advisable to set all but the largest values to $0$, and store the matrix in a sparse datastructure.
+For our experiments, such optimizations were not necessary.
+
+\subsection{The parameters}
+
+Let $F^\text{old}$ be the model we used to compute the responsibilities. Our $Q$-function is:
 \begin{align*}
-q(& \mu_\text{pre}, \Lambda_\text{pre}, \bm, \bD, \bs{\sigma}, \pi, d) \\ 
-=\;&  q_{\mu_\text{pre}, \Lambda_\text{pre}}(\mu_\text{pre}, \Lambda_\text{pre}) \; q_{\bm, \bD}(\bm, \bD)\; q_{\bs{\sigma}}(\bs{\sigma}) \; q_{\pi, d}(\pi, d) \p \\
-\end{align*}
-For the sake of clarity, we will omit the subscripts from the factors in the following.
-
-\subsection{Update rules}
-
-\paragraph{Responsibilities}
-
-The derivation of the update rule for the responsibilities, ie.
-$q_{bs{\sigma}}^{n}(\bs{\sigma})$ takes the same form as the case for the mixture of Gaussians model \cite{}, since our model is essentially a mixture of $k^d$ Gaussians.
-Note that we can visualize $q(\bs{sigma})$ as a large matrix with $n$ rows and $2^d$ columns.
-
-We have:
-\begin{align*}
-q^n(\bs{\sigma}) &= \langle \ln p(x, \bm, \bD, \mu', \Lambda', \bs{\sigma}, \pi, d) \rangle_{\bm, \bD, \mu', \Lambda', \pi, d} + const \p
+Q(F) &= \sum_{i} p(z_i \mid x_i, F^\text{old}) \ln p(x_i \mid z_i, F) p (z_i \mid F) \\
+     &= \sum_{i=1,j=1}^{N, M} P_{ij} \ln \cN_j(x) \;v_{|c_j|}\; \prod_{a\in c_j}w_a \,\,\, \text{with } M = |[1,K]^{[0,D]}|
 \end{align*}
 
-We factorize $\ln p(\ldots)$ according to (\ref{line:factorization}), and absorb any terms that don't depend on $\bs{\sigma}$ into the constant:   
+For clarity of notation, when optimizing for a certain subset of the parameters $q$ (such as the depths, weights or parts of the components) we will write $Q(q)$, and omit any terms that are constant with respect to $q$.
+We may also omit any constant multiplier of the whole function.
+
+To find the optimal depth we solve $\partial Q(v)/\partial v$, using a Lagrange multiplier to enforce that $v$ sums to one, which gives us:
+\[
+\hat v_d = \frac{p_d}{\sum_i p_i} \,\,\text{with } p_d = \sum_{j: |c_j| = d} P_{ij}
+\]
+
+\paragraph{The components $f_k$ and weights $w_k$}
+
+Even with the simplifications of the EM algorithm $Q$ is very difficult to optimize for $f_k$ and $f_p$.
+We must find $K$ maps and weights, and a post-transformation $f_p$, such that all the $M$ endpoint distributions, provide maximal likelihood to their assigned points.
+The problem is that each term in $Q$ is a complicated mix of multiple components.
+
+To make the optimization of $Q$ practical, we simplify the task in two ways.
+First, we optimize $f_p$ and $(\{f_k\}, \{w_k\})$ separately, taking the parameters not being optimized from $F^\text{old}$.
+Let $Y = {f_p^\text{old}}^{-1}(X)$.
+Second, we simplify the $Q$ function as follows:
+
 \begin{align*}
-q^n(\bs{\sigma}) &= \langle \ln p(x \mid \bm, \bD, \mu', \Lambda', \bs{\sigma}) \rangle_{\bm, \bD, \mu', \Lambda'} +  \langle \ln p(\bs{\sigma} \mid \pi, d) \rangle_{\pi, d} + const \p\\
-& = 
+Q(\{f_k\}, \{w_k\}) = \sum_k \sum_{i=1,j=1}^{N, M} & p(k.c_j \mid x, F^\text{old}) \ln f_p f_k(\cN_{j})(x_i) p^{\text{old}}(c_j) p(k)\\ & \hfill \text{with } M = |[1,K]^{[0, D-1]}|  \\
+= \frac{1}{s^\text{old}_p}\sum_k \sum_{i, j} &P^k_{ij} \ln \left[ f_k(\cN_{j})(y_i) \; v_{|c_j|+1} p(c_j) \; w_k \right]\p
+\end{align*}
+We then \emph{approximate} all elements indexed by $j$, so that of all the transformations in each code, only the last one functions as an argument of $Q$.
+We approximate $\cN_j$ by fitting a maximum likelihood spherical MVN to the data weighted by $P^k1_M$: that is, the part of the dataset that $\cN_j$ was responsible for under the old model.
+This gives us $\cN_j = f_{s_j, I, t_j}(\cN_0)$ with
+\[
+t_j = \frac{YP^k0_M^j}{P^k0_M^j 1_M} \;\;\;\;\; s_j = \sqrt{\frac{\sum_i \|P^k0_M^j - t_j\|^2 }{({1_M}^TP^k)_j}} \p
+\]
+That is, $t_j$ and $s_j$ are the weighted mean and weighted standard deviation respectively. Arranging the different $t_j$ as the columns of a matrix $T^k$, we get 
+\begin{align}
+T^k = Y(P^k / (1_N{1_M}^TP^k)) \label{line:tk}
+\end{align}
+where the division is element-wise. $p(c_j)$ becomes a term independent of any parameters, so we need not approximate it. This gives us:
+\begin{align*}
+Q(f_k, w_k) &= \sum_{i, j} P^k_{ij} \ln f_k(\cN_j)(y_i) + \sum_{i, j} P^k_{ij} \ln w_k  
+\end{align*}
+Finding $w_k$ follows the same principle as the depths. We find:
+\[
+\hat w_k = p^k / \sum_{i} p^i \p
+\]
+with $p^k = {1_N}^TP^k1_M$, the sum of the elements of $P^k$.
+
+If we isolate $f_k$, the problem is very similar to the one solved to construct the Coherent Point Drift (CPD) algorithm \cite{myronenko2010point}: transform a set of MVNs to maximize the likelihood of a dataset, with responsibilities of components for datapoints given.
+The main difference is that in our situation each component $N_j$ has its own variance.
+We follow the same approach, and as we shall see, the solutions for our problem are very similar to those of the CPD algorithm.
+
+Using Equation~\ref{line:inverse} from the preliminaries, we can rewrite $Q(f_k)$ as a mixture of transformations of $\cN_0$:
+\begin{align*} 
+Q(s_k, R_k, t_k) &= - p^k\ln s_k - \sum_{i, j} P^k_{ij} \frac{1}{2{s_j}^2{s_k}^2} \left\|y_i-t_k - s_k R_k t_j \right\|^2 
+\end{align*}
+where $s_j$, $R_j$ and $t_j$ are the parameters of the similitude $f_p \circ c_{j1} \circ \ldots \circ c_{jd}$.
+
+To find the optimal translation $t_k$, we solve $\partial Q(t_k)/\partial t_k = 0$, which yields, in matrix notation:
+\begin{align*}
+\hat t_k &= \frac{1}{p^k_z}YP^kZ1_M - \frac{1}{p^k_z}s_kR_kT^kZ{P^k}^T1_N = y^k - s_kR_k t^k
+\end{align*} 
+where $T^k$ is the matrix with $t_j$ as its columns, $Z = \text{diag}({s_1}^{-2}, \ldots, {s_M}^{-2})$ and ${p^k_z} = {1_N}^TP^kZ1_M$, the sum of the elements of the matrix $P^kZ$.
+Thus $\hat t_k$ is the difference between a weighted mean of the data $y^k$ and a weighted mean of the endpoint means $t_j$, scaled and rotated, where in both cases, the matrix  $P^kZ$, normalized to sum to one, determines the weights.
+Note that this is not a complete solution, since it still depends on $s_k$ and $R_k$.
+We can, however, plug $\hat t_k$ back into the $Q$-function to optimize for $R_k$.
+
+Finding the optimal rotation matrix $R_k$ is more complex than simply finding the derivative and setting it equal to zero, since we have the constraint that $R_k$ is orthogonal and has determinant $1$.
+Luckily, a similar problem was solved in \cite{umeyama1991least} and \cite{myronenko2010point}.
+To use this solution, the objective function must be rewritten to the form $\text{tr}(A^TR_k)$ for some $A$.
+We fill in $\hat t_k$, and reduce to
+\[ 
+Q(R_k) = \text{tr}\left(\left[\ol{Y}^kP^kZ{\ol{T}^k}^T\right]^TR_k\right) 
+\]
+with 
+\begin{align*}
+\ol{Y}^k &= Y - y^k {1_N}^T\\
+\ol{T}^k &= T^k - t^k {1_M}^T\\
+\end{align*}
+By \cite[Lemma~1]{myronenko2010point}, the optimal $R_k$ can be derived from the singular value decomposition (SVD) of $A=Y^kP^kZ{T^k}^T$.
+If $A = USV^T$, with $U$, $S$ and $V$ defined as normal for the SVD then we have $\hat R_k = U \;\text{diag} (1, \ldots, 1, |UV^T|)\; V^T$.
+
+Finally, we derive the scaling parameter $s_k$ by filling in $\hat t_k$ and solving $\partial Q(S_k)/\partial s_k$. We get:
+
+\begin{align*}
+0 &= {s_k}^{-2}\;\text{tr}({\ol{Y}^k}^T\text{diag}(P^kZ1_M)\ol{Y}^k) + {s_k}^{-1}\;\text{tr}(\ol{T}^kZ{P^k}^T{\ol{Y}^k}^TR_k) -p^k \p
+\end{align*}
+This is a quadratic polynomial in ${s_k}^{-1}$, which we can solve and invert to find $\hat s_k$.
+
+\subsection{The post transform: $s_p$, $R_p$, $t_p$} Let $M = \left |[1,K]^{[0,D]}\right|$ The $Q$-function becomes:
+\begin{align*}
+Q(s_p, R_p, t_p) &= \sum_{i=1,j=1}^{N,M} p(c_j\mid x_i) \ln f_p(N_j)(x_i) p(c_j) \\
+&= - p\ln s_p  - \frac{1}{2 {s_p}^2} \sum_{i, j} {s_j}^{-2} P_{ij} \| x - t_p -s_pR_pt_j \|^2 \\
+\end{align*}
+where $p = {1_N}^T P 1^M$ and $s_j$ and $t_j$ can be derived from the optimal components, as determined above. Note that $j$ now iterates over a larger set of codes.
+
+The form of the $Q$ function is the same as the ones we used to derive the optimal components $s_k, R_k, t_k$. We follow the same derivations and get:
+
+\begin{align*}
+\hat t_p &= \overline{x} - s_pR_p\overline{t}\\
+\hat R_p &= U\;\text{diag}(1, \ldots, 1, |UV^T|)\;V^T & \text{with } USV = \text{svd}(A), A = \ol{X} P Z {\ol{T}}^T \\
+\end{align*}
+with 
+\begin{align*}
+\ol{x} &= ({1_N}^TPZ1_M)^{-1} XPZ1_M & & \ol{t} = ({1_N}^TPZ1_M)^{-1}TZP^T1_M\\
+\ol{X} &= X - \overline{x}{1_N}^T & &\ol{T} = T - \overline{t} {1_M}^T\\
+Z &= \text{diag}({s_1}^{-2}, \ldots, {s_M}^{-2})\\
 \end{align*}
 
-\paragraph{Components}
+Finally, for $s_p$, we solve
+\[
+0 = {s_p}^{-2}\text{tr}(\ol{X}^T\text{d}(PZ1_M)\ol{X}) + {s_p}^{-1} \text{tr}(\ol{T}ZP^T\ol{X}^TR_p) - p
+\]
 
-\paragraph{Component weights}
+\begin{pseudo}[th]
+\caption{One iteration of the EM algorithm for fitting a fractal model.}
+{
+Given: a dataset $X$, a number of components $K$, a maximum depth $D$. \\
 
-\paragraph{Pre-transformation}
+$C \leftarrow [0,K]^{[0, D-1]}$, $M \leftarrow |C|$ \emph{\# the set of all codes of length up to $D$} \\
+\emph{\# Expectation step}\\
+$P_{ij} \leftarrow N_j(x_i) p(d) \prod_{c \in [1,K]^d} w_c$  \\ 
+Normalize $P$ so that $P1_M = 1_N$ \\
+\textbf{for each} $k \in [1,K]$: \\
+\tab Let $P^k$ be the submatrix of $P$'s columns $j$ such that $c_{j1} = k$\\
+\\
+\emph{\# Maximization step} \\
+$Y \leftarrow {f^\text{old}_p}^{-1}(X)$\\
+for all $d$, $v_d \propto \sum_{j : |c_j| = d} P_{ij} $\\
+\textbf{for each} $k \in [1,K]$: \\
+\tab $w_k \leftarrow {1_N}^T P^k 1_M / \sum_i {{1_N}^T P^i 1_M}$\\
+\tab $y^k \leftarrow {p^k}^{-1}YP^kZ1_M, \,\,\, t^k \leftarrow {p^k}^{-1} T^kZ{P^k}^T 1_N$\\
+\tab $\ol{Y}^k \leftarrow Y + y^k{1_N}^T$, \,\,\, $\ol{T}^k \leftarrow T^k + t^k{1_M}^T$\\
+\tab $Z \leftarrow \text{diag}({s_1}^{-2}, \ldots, {s_M}^{-2})$\\
+\tab $U,S,V^T \leftarrow \text{svd}(\ol{Y}^kP^kZ{\ol{T}^k}^T)$\\ 
+\tab $R_k \leftarrow U\;\text{diag}(1,\ldots,1,|UV^T|) V^T$ \\
+\tab $s_k$: solve ${s_k}^{-2}\; \text{tr}({\ol{Y}^k}^T\text{d}(P^kZ1_M)\ol{Y}^k) + {s_k}^{-1}\;\text{tr}(\ol{T}^kZ{P^k}^T{\ol{Y}^k}^TR_k) = 0$ \\
+\tab $t_k \leftarrow y^k - s_k R_k t^k$ \\
 
-\paragraph{Depth}
+$C \leftarrow [0,K]^{[0, D]}$, $M \leftarrow |C|$\\
+$X' \leftarrow X - \overline{x}{1_N}^T$, $T' \leftarrow T + \overline{t}{1_M}^T$\\
+\tab $Z \leftarrow \text{diag}({s_1}^{-2}, \ldots, {s_M}^{-2})$\\
+$U,S,V^T \leftarrow \text{svd}(X'PZ{T'}^T)$\\ 
+$R_p \leftarrow U\;\text{diag}(1,\ldots,1,|UV^T|) V^T$ \\
+$s_p$: solve ${s_p}^{-2}\text{tr}(X'PZ1_NX'+ T'Z{P}^T1_MT') - s_p^{-1} \text{tr}(ZP^T{x'}^TT') -2p = 0$ \\
+$t_p \leftarrow \overline{x} - s_pR_p\overline{t}$ \\
+}
+\end{pseudo}
 
-\subsection{Results}
+\subsection{Dealing with singularities}
 
-\section{The RIFS model}
-\paragraph{Random Iterated Function Systems}
+In the EM algorithm for MVN mixtures, there is a danger that the algorithm becomes stuck in a situation where one or more of the components do not have responsibility for any part of the data, or for only a single point.
+In this case, the covariance matrix for such components cannot be computed, and the algorithm must be reset in some way.
 
-\subsection{The Inference algorithm for RIFS models}
-\subsection{Results}
+In the IFS algorithm a similar thing can happen, causing one or more of the mtrices $P^k$ to become low-rank.
+In this case, the SVD decomposition required to find $R_k$ cannot be computed.
+When this happens, we use the following strategy: we remove the undetermined component, and for each one we \emph{split} one of the well-determined components.
+Let $f_a = (s_a, R_a, t_a)$ with weight be the well-determined component and $f_b$ be the singleton component.
+We resolve the situation by setting:
+\begin{align*}
+f_a \leftarrow (s_a, R_a, t_a + \epsilon_a)\\
+f_b \leftarrow (s_a, R_a, t_a + \epsilon_b) 
+\end{align*}
+where $\epsilon_a, \epsilon_b$ are vectors with elements randomly drawn from $\cN(0, \sigma)$, where we use $\sigma = 0.001$ in all experiments.
+The weight $w_a$ is distributed equally over the components $f_a$ and $f_b$ and the weight vector is re-normalized.
+
+The rationale is easiest to understand if we take $v_1=1$ and view the model as a mixture of Guassians.
+In that case, the well-determined components cover all the data.
+By splitting $f_a$, and adding some small noise, the points formerly `claimed' by $f_a$ will now be distributed approximately evenly between $f_a$ and $f_b$.
+
+\section{Results}
+\pb{subsampling}
+\pb{initial models}
+
+\label{section:experiments}
+
+\subsection{Synthetic distributions}
+
+\pb{sierpinski, koch-2, ball}
+
+\subsection{Images}
+
+\pb{Sunflower}
+
+\subsection{Higher-dimensional data}
+
+\pb{spherical-MOG, full MOG, IFS, three datasets}
 
@@
 
-\pb{Recap.}
+We have introduced a new algorithm for the induction of fractal models.
+To our knowledge this is the first such algorithm that does not use a general-purpose optimization technique like genetic algorithms.
 
-\pb{}
+In the introduction we mentioned the intuitive motivation for the algorithm that given the codes for each point, we can consbtruct paired lists of points for each component such that the first list should map onto the second, reducing the inverse problem to the absolute orientation problem, as solved in \cite{umeyama1991least}.
+The reader may wonder if we have strayed from this original idea in our EM treatment, since our $Q$-function do not seem to map points onto one, but rather endpoint distributions onto points.
+However, since our simplification of the $Q$-function in Equation~\ref{} estimates these endpoint distributions from a re-weighted dataset, we can substitue these, and show that there we are in effect building an $N \times N$ matrix ${\cal P}^k$ and solving a weighted version of the absolute orientation problem, mapping every point $x_i$ in the data onto every other point $x_j$ with weight ${\cal P}^k_{ij}$.
+This is proved in the appendix.
+
+Similitudes were chosen as a nice balance between expressiveness and parameter complexity.
+More general function classes are certainly possible, although deriving the maximization step analytically may not be possible for these.
+For instance, if we allow generic, affine  functions, defined by a translation $t$ and a matrix $A$, we find a derivative of the form $\frac{\partial Q}{\partial A} |A^{-1}| + cA$.
+This is difficult, if not impossible to solve analytically, although if we restrict $A$ to be positive definite, we can see that the problem is convex, so that we can easily solve it numerically.
+
+An alternative approach would be to use a variational Bayesian algorithm instead of an EM algorithm.
+If a variational algorithm exists for the desired function class, it should be straightforward to plug these into a general variational algorithm for iterated function systems.
+Such an approach would also avoid the problem of singularities, and it would allow some tuning of the algorithm through the selection of priors.
+We consider this a promising direction for future research.
+
+Many natural fractal phenomena are not precisely captured by iterated functions systems.
+Coastlines trees and clouds are self-similar, but in a much more random manner than IFSs can describe.
+One solution takes the form of an extension to random iterated function sytems, as described in \cite{hart1996fractal}.
+Here, at each application the IFS is itself chosen at random from a distribution on IFSs.
+Another option would be to manually add domain knowledge about the data into the model, for instance specific knowledge about the development of clouds or the growth of trees.
+In both cases, we believe the EM algorithm provides a basic template for a solution: the main issue is that of finding the specific choices made inside the model to arrive at each point in the dataset.
+By casting this sequence of choices as a latent variable, and applying Bayes rule (either using EM or a variational approach), we divide and conquer: we cannot solve the problem as a whole, but we can find the latent variables given the parameters, and the parameters given the latent variables.
 
@@
 
+\appendix
+
+\section{Derivations}
+
+Transforming a generic spherical MVN by a similitude can be cast as the transformation of $\cN_0$ by two similitudes:
+\begin{align*}
+f_{t, R, s}(\cN_{t_0, {s_0}^2 I})(x) &= f_{t, R, s}(f_{t_0, R_0, s_0}(\cN_0))(x) = f_{ sRt_0+t, RR_0,ss_0}(\cN_0)(x) \\
+&= \frac{1}{ss_0} \cN_0\left(\frac{1}{ss_0} {R_0}^T R^T x - \frac{s}{ss_0} {R_0}^Tt_0 - \frac{1}{ss_0} {R_0}^TR^Tt \right) \\
+&= \frac{1}{ss_0} \cN_0\left(\frac{1}{ss_0} x - \frac{1}{s_0} {R}t_0 - \frac{1}{ss_0} t \right) \\
+&= \frac{1}{ss_0} \cN_0\left(\frac{1}{ss_0} {R_0}^T R^T x - \frac{s}{ss_0} {R_0}^Tt_0 - \frac{1}{ss_0} {R_0}^TR^Tt \right) \\
+&=  (2\pi)^{-\frac{H}{2}} \frac{1}{ss_0} \exp \left[- \frac{1}{2{s_0}^2s^2}\left\| x- t - s R t_0 \right \|^2\right]
+\end{align*}
+
+\paragraph{The depth $v$ and weights $w$} For the depth priors $v$, the $Q$-function, with constant terms omitted, reduces to:
+\begin{align*}
+Q(v) &= \sum_{i,j} P_{ij} \ln v_{|c_j|} \\ 
+&= \sum_{d \in [0,D]} \sum_{j: |c_j| = d} P_{ij} \ln v_d = \sum_{d \in [0,D]} \left[\sum_{c: |c| = d} P_{x, c}\right] \ln v_d\\
+&= \sum_{d \in [0,D]} p_d \ln v_d \;\;\text{with $p_d = \sum_{c: |c| = d} P_{x, c} $}  \\
+\end{align*}
+For $v$, we have the additional constraint that $\sum_d v_d = 1$, which we can take into consideration with lagrange multipliers, giving us the objective function ${\cal L}(v, \lambda) = p_d/v_d - \lambda\sum_d v_d - \lambda$.
+We differentiate and set equal to zero, giving us the equations $p_d/v_d - \lambda = 0$ amd $\sum_i v_i -1 =0$.
+Solving these gives us:
+\[
+\hat v_d = \frac{p_d}{\sum_i p_i}
+\]
+
+Finding $w_k$ follows the same principle as the depths. The $Q$-function reduces to:
+\[
+Q(w) = \sum_{k} p_k \ln w_k
+\]
+
+Using Lagrange multipliers to incorporate the constraint that $\sum_i w_i = 1$, we find
+\[
+\hat w_k = p^k / \sum_{i} p^i \p
+\]
+
+\paragraph{The translations $t_k$} The optimal translation can be found by straightforward differentiation:
+\begin{align*}
+\frac{\partial Q(t_k)}{\partial t_k} &= 
+- \sum_{i,j} P_{ij}^k \frac{1}{{s_j}^2{s_k}^2}\frac{\left \|  y_i - t_k - s_kR_kt_j \right\|^2}{\partial t_k}
+\\
+&= \sum_{i,j} P_{ij}^k \frac{1}{{s_j}^2{s_k}^2} \left(y_i - t_k - s_kR_kt_j\right)^T \\
+\end{align*}
+
+This sum represents a row vector, which we transpose to get a column vector, and set equal to zero to get $\hat t_k$:
+\begin{align*}
+0 &= \frac{1} {{s_k}^2} \left (\sum_{i,j} P^k_{ij} {s_j}^{-2} y_i  - \sum_{i,j} P^k_{ij} {s_j}^{-2} t_k - s_kR_k \sum_{i,j} P^k_{ij} {s_j}^{-2} t_j\right) \\
+\end{align*}
+
+Let $Z = \text{diag}({s_1}^{-2}, \ldots, {s_M}^{-2})$, we can then rewrite to matrix notation:
+\begin{align*}
+\hat t_k &= \frac{YP^kZ1_M - s_kR_k\;T_j(P^kZ)^T1_N}{{1_N}^TP^kZ1_M} 
+\end{align*}
+
+\paragraph{The rotations $R_k$}
+\begin{align*} 
+Q(R_k) &= - \frac{1}{2}\sum_{i,j} P^k_{ij} \frac{1}{{s_j}^2{s_k}^2} \left\| y_i - t_k - s_kR_k t_j \right\|^2\\
+&= - \frac{1}{2}\sum_{i,j} P^k_{ij} {s_j}^{-2} \left\| \left(y_i - y^k\right) - s_kR_k \left(t_j-t^k\right) \right\|^2 \p 
+\end{align*}
+Let ${y^k_i} = y_i - y^k$ and ${t^k_j} = t_j - t^k$, ie.
+mean-centered versions of the data points and the component means.
+Let $Y^k$ and ${T^k}$ be the corresponding matrices with these vectors as columns.
+We multiply out the inner product to get
+\begin{align*}
+Q(R_k) &= - \frac{1}{2}\sum_{i,j} P^k_{ij} {s_j}^{-2} \left[ {y_i^k}^T{y_i^k} - 2 s_k {y_i^k}^TR_k{t_j^k} + {t_j^k}^T{t_j^k} \right] \\
+\end{align*}
+from which we remove the terms and global factors that are independent of $R_k$:
+\begin{align*}
+Q(R_k)&= \sum_{i,j} P^k_{ij}\; {s_j}^{-2}\; {y_i^k}^TR_k{t_j^k}
+\end{align*}
+Which in matrix notation becomes:
+\begin{align*} 
+Q(R_k) &= \text{tr}((P^k Z)^T {Y^k}^TR_kT^k) = \text{tr}(T^k Z {P^k}^T{Y^k}^TR_k) = \text{tr}([Y^kP^kZ{T^k}^T]^TR_k) \p
+\end{align*}
+Where we use the fact that $\text{tr}(A^TB) = \sum_{i,j}(A\circ B)_{ij}$ and the fact that the trace is invariant under cyclic permutations.
+
+\paragraph{The scaling $s_k$}
+We derive the scaling by filling in $\hat t_k$ and solving for $\partial Q(S_k)/\partial s_k$.
+\begin{align*}
+\frac{\partial Q(s_k)}{\partial s_k} &=  \frac{- p^k \ln s_k - \sum_{i,j} P^k_{ij} \frac{1}{2{s_j}^2} \left\|y_i - t_k - s_k R_k t_j \right\|^2}{\partial s_k}\\
+\frac{\partial Q(s_k)}{\partial s_k} &= - p^k {s_k}^{-1} - \sum_{i,j} P^k_{ij} \frac{1}{2{s_j}^2} \left(-{s_k}^{-3} {y^k_i}^T y^k_i + {s_k^{-2}} {y^k_i}^TR_kt_j^k + {s_k}^{-3}{t_j^k}^T t_j^k\right) \\
+&= - p^k {s_k}^{-1} + {s_k}^{-3} \frac{1}{2}\sum_{i,j} P^k_{ij} 2{s_j}^{-2} {y^k_i}^T y^k_i \\ 
+& - {s_k}^{-2} \frac{1}{2}\sum_{i,j} P^k_{ij} 2{s_j}^{-2} {y^k_i}^TR_kt_j^k - {s_k}^{-3} \frac{1}{2}\sum_{i,j} P^k_{ij} 2{s_j}^{-2} {t_j^k}^T t_j^k)
+\end{align*}
+In matrix notation:
+\begin{align*}
+0 &= {s_k}^{-2} \text{tr}({Y^k}^T\text{d}(P^kZ1_M)Y^k) + {s_k}^{-1} \text{tr}(T^kZ{P^k}^T{Y^k}^TR_k) -p^k \p
+\end{align*}
+
+\section{Relation to the absolute orientation problem}
+
+In the introduction we described the intuitive motivation for the algorithm: given the codes for each point, we can construct paired lists of points for each component such that the first list should map onto the second, reducing the inverse problem to the absolute orientation problem, as solved in \cite{umeyama1991least}.
+The reader may wonder if we have strayed from this original idea in our EM treatment, since our update rules do not seem to map points onto one another.
+
+\ldots
+
 \end{document}

### 2015-11-06 fractal paper/src/fractal.tex
....

@@
 \[
-t_j = \frac{YP^k0_M^j}{P^k0_M^j 1_M} \;\;\;\;\; s_j = \sqrt{\frac{\sum_i \|P^k0_M^j - t_j\|^2 }{({1_M}^TP^k)_j}} \p
+t_j = \frac{YP^k0_M^j}{P^k0_M^j 1_M} \;\;\;\;\; s_j = \sqrt{\frac{\sum_i P^k0_M^j\| y_i - t_j\|^2 }{({1_M}^TP^k)_j}} \p
 \]
@@
 \begin{align}
-T^k = Y(P^k / (1_N{1_M}^TP^k)) \label{line:tk}
+T^k = (YP^k) / (1_H({1_N}^T)P^k) \label{line:tk}
 \end{align}
-where the division is element-wise. $p(c_j)$ becomes a term independent of any parameters, so we need not approximate it. This gives us:
+where the division is pointwise. $p(c_j)$ becomes a term independent of any parameters, so we need not approximate it. This gives us:
 \begin{align*}
@@
 \end{align*}
-By \cite[Lemma~1]{myronenko2010point}, the optimal $R_k$ can be derived from the singular value decomposition (SVD) of $A=Y^kP^kZ{T^k}^T$.
+By \cite[Lemma~1]{myronenko2010point}, the optimal $R_k$ can be derived from the singular value decomposition (SVD) of $A=\ol{Y}^kP^kZ{\ol{T}^k}^T$.
 If $A = USV^T$, with $U$, $S$ and $V$ defined as normal for the SVD then we have $\hat R_k = U \;\text{diag} (1, \ldots, 1, |UV^T|)\; V^T$.
@@
 \begin{align*}
-\frac{\partial Q(s_k)}{\partial s_k} &=  \frac{- p^k \ln s_k - \sum_{i,j} P^k_{ij} \frac{1}{2{s_j}^2} \left\|y_i - t_k - s_k R_k t_j \right\|^2}{\partial s_k}\\
-\frac{\partial Q(s_k)}{\partial s_k} &= - p^k {s_k}^{-1} - \sum_{i,j} P^k_{ij} \frac{1}{2{s_j}^2} \left(-{s_k}^{-3} {y^k_i}^T y^k_i + {s_k^{-2}} {y^k_i}^TR_kt_j^k + {s_k}^{-3}{t_j^k}^T t_j^k\right) \\
-&= - p^k {s_k}^{-1} + {s_k}^{-3} \frac{1}{2}\sum_{i,j} P^k_{ij} 2{s_j}^{-2} {y^k_i}^T y^k_i \\ 
-& - {s_k}^{-2} \frac{1}{2}\sum_{i,j} P^k_{ij} 2{s_j}^{-2} {y^k_i}^TR_kt_j^k - {s_k}^{-3} \frac{1}{2}\sum_{i,j} P^k_{ij} 2{s_j}^{-2} {t_j^k}^T t_j^k)
+\frac{\partial Q(s_k)}{\partial s_k} &=  \frac{- p^k \ln s_k - \frac{1}{2}\sum_{i,j} (P^kZ)_{ij} {s_k}^{-2} \left\|y_i - t_k - s_k R_k t_j \right\|^2}{\partial s_k}\\
+\frac{\partial Q(s_k)}{\partial s_k} &= - p^k {s_k}^{-1} - \frac{1}{2}\sum_{i,j} (P^kZ)_{ij} \left(-2{s_k}^{-3} {y^k_i}^T y^k_i + 2{s_k^{-2}} {y^k_i}^TR_kt_j^k \right) \\
+&= -p^k {s_k}^{-1} + {s_k}^{-3} \sum_{i, j} (P^kZ)_{ij}{s_k}^{-3} {y^k_i}^T y^k_i - s_k^{-1} \sum_{i, j}(P^kZ)_{ij}{s_k^{-2}} {y^k_i}^TR_kt_j^k
 \end{align*}
@@
 \begin{align*}
-0 &= {s_k}^{-2} \text{tr}({Y^k}^T\text{d}(P^kZ1_M)Y^k) + {s_k}^{-1} \text{tr}(T^kZ{P^k}^T{Y^k}^TR_k) -p^k \p
+0 &= {s_k}^{-2} \text{tr}({\ol{Y}^k}^T\text{d}(P^kZ1_M)\ol{Y}^k) + {s_k}^{-1} \text{tr}(\ol{T}^kZ{P^k}^T{\ol{Y}^k}^TR_k) -p^k \p
 \end{align*}

### 2015-11-07 fractal paper/src/fractal.tex
...

@@
 We then \emph{approximate} all elements indexed by $j$, so that of all the transformations in each code, only the last one functions as an argument of $Q$.
-We approximate $\cN_j$ by fitting a maximum likelihood spherical MVN to the data weighted by $P^k1_M$: that is, the part of the dataset that $\cN_j$ was responsible for under the old model.
+We approximate $\cN_j$ by fitting a maximum likelihood spherical MVN to the data weighted by $P1_M$: that is, the part of the dataset that $\cN_j$ was responsible for under the old model.
+
+Let $P'$ consist of all but the last $K^D$ columns of $P$ (ie., those corresponding to values of $j$ in the sum above.
+Note that this is different from any $P^k$: both $P^k$ and $P'$ are a submatrix of $P$, but column $j$ in $P'$ contains $p(k.c_j\mid x)$, while column $j$ in $P'$ contains $p(c_j \mid x)$.
+
 This gives us $\cN_j = f_{s_j, I, t_j}(\cN_0)$ with
 \[
-t_j = \frac{YP^k0_M^j}{P^k0_M^j 1_M} \;\;\;\;\; s_j = \sqrt{\frac{\sum_i P^k0_M^j\| y_i - t_j\|^2 }{({1_M}^TP^k)_j}} \p
+t_j = \frac{YP'0_M^j}{P'0_M^j 1_M} \;\;\;\;\; s_j = \sqrt{\frac{\sum_i P'0_M^j\| y_i - t_j\|^2 }{({1_M}^TP')_j}} \p
 \]
@@
 \begin{align}
-T^k = (YP^k) / (1_H({1_N}^T)P^k) \label{line:tk}
+T = (YP') / (1_H({1_N}^T)P') \label{line:tk}
 \end{align}
-where the division is pointwise. $p(c_j)$ becomes a term independent of any parameters, so we need not approximate it. This gives us:
+where the division is pointwise. We remove the last $K^D$ columns from
+
+$p(c_j)$ becomes a term independent of any parameters, so we need not approximate it. This gives us:
 \begin{align*}

### 2017-06-09 fractal-plr/src/fractal.tex
...

\documentclass[reprint,amsmath,amssymb,aps,prl]{revtex4-1}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{float}

\floatstyle{ruled}
\newfloat{pseudo}{h}{lop}
\floatname{pseudo}{Algorithm}

\usepackage{color}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\p}{\mbox{\,.}}
\newcommand{\bD}{\boldsymbol\Lambda}
\newcommand{\R}{\mathbb{R}}
\newcommand{\cN}{{\cal N}}
\newcommand{\bs}[1]{{\boldsymbol {#1} }}
\newcommand{\vc}{\textnormal{vec}}
\newcommand{\tab}{\hspace*{5mm}}
\newcommand{\ovl}[1]{\overline{#1}}
\newcommand{\run}[1]{\includegraphics[width=0.15\linewidth]{../img/runs/{#1}.png}}

\newcommand{\bx}{\boldsymbol x}
\newcommand{\by}{\boldsymbol y}
\newcommand{\bX}{\boldsymbol X}
\newcommand{\bY}{\boldsymbol Y}
\newcommand{\bo}{\boldsymbol 1}
\newcommand{\be}{\boldsymbol e}
\newcommand{\bt}{\boldsymbol t}
\newcommand{\bA}{\boldsymbol A}
\newcommand{\bmu}{\boldsymbol \mu}
\newcommand{\bSig}{\boldsymbol \Sigma}
\newcommand{\bI}{\boldsymbol I}
\newcommand{\bR}{\boldsymbol R}
\newcommand{\bZ}{\boldsymbol Z}
\newcommand{\bz}{\boldsymbol z}
\newcommand{\bth}{\boldsymbol \theta}
\newcommand{\bw}{\boldsymbol w}
\newcommand{\bv}{\boldsymbol v}
\newcommand{\bP}{\boldsymbol P}
\newcommand{\bT}{\boldsymbol T}
\newcommand{\bU}{\boldsymbol U}
\newcommand{\bS}{\boldsymbol S}
\newcommand{\bV}{\boldsymbol V}

\begin{document}

\preprint{APS/123-QED}

\title{An Expectation-Maximization Algorithm for the Fractal Inverse Problem}

\author{Peter Bloem}
\email{vu@peterbloem.nl}
\affiliation{VU University Amsterdam}

\author{Steven de Rooij}
\email{s.derooij@gmail.com}
\affiliation{Leiden University}

\date{\today}

\begin{abstract}
\noindent We present an Expectation-Maximization algorithm for the \emph{fractal inverse problem}: the problem of fitting a fractal model to data. In our setting the fractals are Iterated Function Systems (IFS), with \emph{similitudes} as the family of transformations. The data is a point cloud in ${\mathbb R}^H$ with arbitrary dimension $H$. Each IFS defines a probability distribution on ${\mathbb R}^H$, so that the fractal inverse problem can be cast as a problem of parameter estimation. We show that the algorithm reconstructs well-known fractals from data, with the model converging to high precision parameters. We also show the utility of the model as an approximation for for datasources outside the IFS model class. 
\end{abstract}

\maketitle

\noindent Fractals are mathematical objects; commonly formalised as a class of sets or of probability distributions. What, in general, constitutes a fractal is not precisely defined \footnotemark[1], but certain properties are common to most examples; for instance self-similarity, infinitely fine structure, and a non-integer dimension.

\footnotetext[1]{Mandelbrot originally defined a fractal as a set whose topological dimension differs from its Hausdorff dimension, but later retracted this definition, stating that he preferred the word to be not precisely defined.}

\begin{figure}[b]
  \includegraphics[width=\linewidth]{../img/copies.png}
  \caption{An example of an IFS.}
\end{figure}

Since they were named in the 1970s, fractals have increasingly been seen as a potential model for many natural phenomena. Mandelbrot put it as follows in \emph{The Fractal Geometry of Nature} \cite{mandelbrot1982fractal}: 
\begin{quote}
\noindent Clouds are not spheres, mountains are not cones, coastlines are not circles, and bark is not smooth, nor does lightning travel in a straight line.
\end{quote}
Fractals have been used in many fields, including physics \cite{mandelbrot1984fractals}, geology \cite{cheng1997multifractal}, biology \cite{goldberger1992fractal} and economics \cite{turiel2003multifractal}.

One of the greatest problems with fractal analysis has always been the difficulty of finding a fractal model, given some observations. It may be visually clear that a cloud or a coastline `looks' fractal, and we may be able to determine that it has a non-integer dimension \cite{theiler1990estimating}, but how do we get from a dataset to a model? How do we determine the parameters of the fractal-generating model that led to this particular image or dataset? This is called the \emph{fractal inverse problem}.

Current approaches tend to rely on evolutionary algorithms \cite{deliu1991genetic,collet2000polar,nettleton1994evolutionary}. Such algorithms are expensive, and it can be difficult to get them to converge to precise models, even if the dataset itself is sampled from a fractal model. Other approaches are highly domain-specific, such as fractal image compression \cite{hart1996fractal}, which does not generalize well to data of arbitrary dimension. Some interesting results have been also been derived from the method of moments \cite{rinaldo1994inverse} and sampling random transformations from the data \cite{hart1997similarity}, but apart from the evolutionary methods, we are not aware of any other generic, practical methods tailored to fractal induction.

We focus on the family of \emph{Iterated Function Systems} (IFSs) \cite{barnsley2014fractals,hutchinson2000deterministic}, a broad class of fractals, capturing many well-known examples. Figure~\ref{figure:examples} shows the basic principle: we start with some initial image and apply a small number, $K$, of contracting transformations to it, resulting in a second image consisting of $K$ scaled down copies of the initial image. We then apply the transformations again, to this second image, and so on. As we iterate, the image converges to a fractal. Which fractal emerges is entirely determined by the chosen transformations. By fixing a family of transformations $\cal F$, we define a family of fractals, each determined by a set of $K$ transformations chosen from $\cal F$: its \emph{components}. Here, we choose \emph{similitudes}, a subset of affine maps, as our family of transformations. Similitudes offer a good trade-off between expressiveness, and ease of optimization.

The IFS concept generalizes naturally to \emph{probability distributions}: we choose the components of the IFS, and instead of applying them to an initial image, we apply them to an initial \emph{distribution}, combining these into a mixture of scaled-down copies of the original distribution. The iteration now results in a fractal probability distribution. This allows us to frame the fractal inverse problem as a problem of parameter estimation: for a set of points $\{x\} \subset \R^H$, find a set of $K$ similitudes, such that the likelihood of $\{x\}$ under the resulting IFS distribution is maximal.

We present an Expectation-Maximization (EM) algorithm \cite{dempster1977maximum} to find this model. Since the specific sequence of components that ``generated'' a point $x$ in the dataset is unknown, we cast this information as a \emph{latent variable} and iterate between optimizing the latent variables given some model, and optimizing the model given the latent variables.

We show that our algorithm quickly reconstructs known fractals, such as the Sierpinski triangle and the Koch curve. We also apply the algorithm to some datasets sampled from images, and some natural data of higher dimension, showing that, while there are no IFSs to perfectly capture these images, the self similarity in the model does often allow a good fit, comparable to that of a mixture-of-Gaussians model. An open source implementation is available \footnote{\url{https://github.com/pbloem/ifsem}}.

\paragraph{Notation}

Let $\{\bx_i\}_{i\in[1,N]} \subset \R^H$ be our dataset. Let $\bX$ be the $N\times H$ matrix with $\bx_i$ as its columns. All vectors are column vectors.

Let $\bo$ be vector with $1$ at all indices and let $e^i$ be the one-hot vector for $i$, with length implied by context.

For $S \subseteq \R^H$, let $V(S)$ be a probability distribution on $\R^H$ with p.d.f. $v(\bx)$. Let $f_{\bt, \bA}(\bx) = \bA\bx + \bt$ be an invertible affine transformation. Then the transformation of $V$ by $f_{\bt, \bA}$ is defined by the relation $f_{\bt, \bA}(V)(S) = V({f_{\bt, \bA}}^{-1}(S))$. The density function of $f_{\bt, \bA}(V)$ is $f_{\bt, \bA}(V)(\bx) = |\bA^{-1}| v({f_{\bt, \bA}}^{-1}(\bx))$.

A \emph{Gaussian}, on $\R^H$ is determined by a mean $\bmu \in \R_H$ and a covariance matrix $\bSig \in \R^{H\times H}$. We call $\cN_0 = (\boldsymbol 0$, $\bSig = \bI^H)$ the \emph{standard} Gaussian. A \emph{spherical} or \emph{isometric} Guassian has $\bSig = s\bI$ for some scalar $s$. Let $\bx$ be a random variable with $\bx \sim \cN(\bmu, \bSig)$, then $f_{\bt, \bA}(\bx) \sim \cN(\bt + \bA\bmu, \bA\bSig \bA^T)$.
A \emph{similitude} $f:\R^H \to \R^H$ is defined by a rotation matrix $\bR$, vector $\bt$ and scalar $s$ as follows $f_{\bt, \bR, s}(\bx) = s\bR \bx + \bt$.

Transforming a spherical MVN by a similitude yields
\begin{align}
&f_{\bt, \bR, s}\left(\cN(\bt_0, {s_0}^2 \bI)\right)(\bx) = \notag \\
&(2\pi)^{-\frac{H}{2}} ({ss_0})^{-H} \exp \left[- \frac{1}{2{s_0}^2s^2}\left\| \bx - (s \bR \bt_0 + \bt)\right \|^2\right] \label{line:inverse}
\end{align}

\section{The IFS Model}

We define an \emph{Iterated Function System} of order $K$ and dimension $H$ as a pair $(\{f_k\}, \{w_k\})$ of $K$ similitude \emph{components} $f_k: \R^H \to \R^H$ with $K$ associated \emph{weights} $w_k$, nonnegative scalars, with $\sum_i w_i = 1$.

Let $\bx_0 \sim p_0$, with $p_0$ any distribution on $\R^H$ with compact support. Define $\bx_{n+1}$ as the random variable with a mixture distribution over the $K$ random variables $f_k(\bx_n)$, with mixture weights $w_k$. We note two important properties of IFSs. First, the distribution on $\bx_D$ converges with $D$: the KL divergence between successive distributions goes to zero. We call the distribution it converges to the \emph{limit distribution}. Second, the limit distribution is independent of the choice of $p_0$. Thus, we can see the weights and components as \emph{parameters} of the limit distribution. For proofs and generalizations, see \cite{hutchinson2000deterministic}.

\begin{figure}[t]
  \center
  \includegraphics[width=\linewidth]{../img/factor-graph.pdf}
  \caption{A graphical model, illustrating the components of the IFS model. The gray plate represents a repetition of the nodes inside for each datapoint.}
  \label{figure:ifs-diagram}
\end{figure}

To make the model easier to fit to natural data, we extend the IFS model with some additional factors, as illustrated in Figure~\ref{figure:ifs-diagram}. It contains:
\begin{description}
\item[the components $\{(s_k, \bR_k, \bt_k)\}_k$] The $K$ similitudes that make up the model: $f_k(x) = s_k\bR_k\bx + \bt_k$.
\item[the weights $\bw$] A length-$K$ probability vector. 
\item[the depth weights $\bv$] A length-$D$ probability vector.
\item[the post-transform $s_p, \bR_p, \bt_p$] A single similitude.
\item[the code $c_i$] A sequence of integers $c_i = \langle c_{i1}, \ldots, c_{id} \rangle$.
\item[the data $\bx_i$] A single point in $\R^H$.
\end{description}
The first four items form the \emph{parameters} $\theta$ of the model. The rest are \emph{observed} variables $\bx_i$ and the \emph{latent} variables $c_i$.

These variables combine into a model as follows. Let $\{c_j\} = [1, K]^{[0, D]}$, the set of all sequences of integers from $[1, K]$ of length up to $D$, including the empty sequence. We define the function $f_j:\R^H \to \R^H$ as the composition of the post-transform and the components indicated by the code $c_j$:$f_j = f_p \circ f_{c_{j1}} \circ \ldots \circ f_{c_{jd}}$. We can now write the p.d.f. of the model as:
\begin{align*}
p(x\mid \theta) &= \sum_{j=1}^{|[1, K]^{[0, D]}|} \bv_{|c_j|} \left[\prod_i \bw_{c_{ji}}\right ] f_j(\cN_0)(x) \p
\end{align*}
That is, a mixture of $|[1, K]^{[0, D]}|$ Gaussians, where the $j$-th Gaussian has the weight $\bv_{|c_j|} \prod_i \bw_{c_{ji}}$. Equivalently, we evaluate the IFS to all depths from $0$ to $D$ and mix these with the weights in $\bv$.

Mixing different depths makes the model a generalization of good fall-back models: with $\bv_0 = 1$, the model becomes a spherical Gaussian. With $\bv_1 = 1$, the model becomes a mixture of K Gaussians \footnotemark[2]. Mixing depths also allows the EM search a ``gentle start''. Most IFSs have a support with lower dimension than the embedding dimension of the data: i.e. almost all of $\R^H$ has probability zero. Thus, the slightest change in parameters means the difference between maximum likelihood, and likelihood zero. The variable depth allows the search algorithm to start with low-fitness models with a smooth error surface, and slowly converge to the more challenging, deeper models.

The post-transform allows us to deal with data that is not properly centered. While we can make off-center IFSs by tuning the components, such models will not overlap properly at different depths. The post-transform allows us to learn a centered IFS in place, and map it to the data.

\footnotetext[2]{To generalize to generic MOG models, the components should be extended to positive definite affine functions.}

\begin{figure}[tb]
  \centering
  \begin{subfigure}[b]{0.47\linewidth}
    \includegraphics[width=\textwidth]{../img/sierpinski-codes.pdf}
    \caption{}
    \label{fig:sierpinski-codes}
  \end{subfigure}  
  \hspace{0.015\textwidth}
  \begin{subfigure}[b]{0.47\linewidth}
    \includegraphics[width=\textwidth]{../img/code-construction.pdf}
    \caption{}
    \label{fig:code-construction}
  \end{subfigure}

  \caption{\small (a) Codes of length three on the Sierpinski triangle and the subsets they code for. (b) The construction of a subset from its code.}
  \label{figure:codes}
\end{figure}

\section{The EM Algorithm}
\label{section:algorithm}

Our task is to select the parameters $\theta$ such that he probability of our data $\bX$ under the model $\theta$ is maximized. Unfortunately, we do not know the code $c_i$ corresponding to each point $\bx_i$. A common way to deal with such \emph{latent variables} is the EM algorithm. In brief, it starts with an initial model, computes the most likely values of the latent variables for each datapoint, and then re-computes the most likely model given these values of the latent variables, iterating until convergence.

The most concise way to describe the EM algorithm is in terms of a \emph{Q-function}:
\[
Q(\theta) = \sum_{\bZ} p(\bZ \mid \bX, \theta^\text{old}) \ln p(\bX, \bZ \mid \theta) \label{line:q-function}
\]
Where the sum iterates over all possible values of all latent variables $\bZ$. Note that the variable $\theta^\text{old}$, our existing estimate for the model parameters, is a constant. The \emph{expectation} step computes $p(\bZ \mid \bX, \theta^\text{old})$, and the \emph{maximization} step computes the $\theta$ that maximizes $Q(\theta)$.

We first rewrite to:
\begin{align*}
Q(\theta) &= \sum_{i} p(\bz_i \mid \bx_i, \theta^\text{old}) \ln p(\bx_i \mid \bz_i, \theta) p(\bz_i \mid \theta) \\
     &= \sum_{i=1,j=1}^{N, M} \bP_{ij} \ln \cN_j(x) \;v_{|c_j|}\; \prod_{k\in c_j}w_k 
\end{align*}
with $M = |[1, K]^{[0, D]}|$ and $\bP$ an $N\times M$ matrix with 
\[
\bP_{ij} = p(c_j \mid x_i) = \frac{p(x_i\mid c_j) p(c_j) }{\sum_{a \in [1,K]^{[0, D]}} p(x_i\mid c_a) p(c_a)} \p 
\]
The computation of $\bP$ is the expectation step of the algorithm: $\bP_{ij}$ represents the \emph{responsibility} that the $j$-th Gaussian takes for the $i$-th datapoint. For the maximization step we must optimize $Q$ with respect to the various elements of $\theta$, taking the partial derivative and setting it equal to zero. For clarity of notation, when optimizing for a certain subset $q$ of the parameters $\theta$ (such as the depths, weights or parts of the components) we will write $Q(q)$, and omit any terms that are constant with respect to $q$. We may also omit any constant multiplier of the whole function.

The optimization of the components and the post-transform $f_k$ and $f_p$ is difficult. We must find $K$ maps and weights, and a post-transformation $f_p$, such that all the $M$ endpoint distributions provide maximal likelihood to their assigned points. The problem is that each term in $Q$ is a complicated mix of multiple components.

To make the optimization of $Q$ practical, we simplify the task in two ways. First, we optimize $f_p$ and $(\{f_k\}, \{w_k\})$ separately, taking the parameters not being optimized from $\theta^\text{old}$. Second, we simplify the $Q$-function. We first rewrite it as follows: Let $kc_j$ be the code $\langle k, c_{j1}, \ldots\rangle$ and let $M^- = |[1, K]^{[0, D-1}]|$. Let $\bY = {f_p^\text{old}}^{-1}(\bX)$. Then:
\begin{align*}
Q&(\{f_k\}, \{w_k\})  \\ 
&= \frac{1}{s_p}\sum_k \sum_{i, j} &\bP^k_{ij} \ln \left[ f_k(\cN_{j})(y_i) \; v_{|c_j|+1} p(c_j) \; w_k \right]\p
\end{align*}
Where we have defined a submatrix $\bP^k$ of $\bP$, which contains all columns corresponding to codes beginning with the symbol $k$: $\bP_{ij}^k = p(kc_j \mid \bx_i)$.

We have now written $Q$ in a form that isolates only the first component in the code. Of course, each $\cN_j$ in this sum still depends on all the components in the code $c_j$, but this is where we simplify the function: we take $\cN_j$ to be a \emph{constant}, computed from $\theta^\text{old}$, and optimize only for the first component in the code. This gives us, for the $k$-th component: $Q(f_k) = \sum_{i, j} \bP^k_{ij} \ln f_k(\cN_j)(y_i)$.

Using (\ref{line:inverse}), we can rewrite $Q(f_k)$ as a mixture of transformations of $\cN_0$:
\begin{align*} 
Q(s_k, \bR_k, \bt_k) =& - p^k H \ln s_k - \\ 
&\sum_{i, j} \bP^k_{ij} \frac{1}{2{s_j}^2{s_k}^2} \left\|y_i-\bt_k - s_k \bR_k \bt_j \right\|^2 
\end{align*}
where $s_j$, $\bR_j$ and $\bt_j$ are the parameters of the similitude $f_p \circ c_{j1} \circ \ldots \circ c_{jd}$.

To find the optimal translation $\hat\bt_k$, we solve $\partial Q(\bt_k)/\partial \bt_k = 0$, which yields:
\begin{align*}
\hat \bt_k &= \frac{1}{p^k_z}\bY\bP^k\bZ\bo - \frac{1}{p^k_z}s_kR_k\bT\bZ{\bP^k}^T\bo = \by^k - s_k\bR_k \bt^k
\end{align*} 
where $\bT$ is the matrix with $\bt_j$ as its columns, $\bZ = \text{diag}({s_1}^{-2}, \ldots, {s_M}^{-2})$ and ${p^k_z} = \bo^TP^kZ\bo$. Thus $\hat \bt_k$ is the difference between a weighted mean of the data $\by^k$ and a weighted mean of the MVN means $\bt^k$, scaled and rotated, where in both cases, the matrix  $\bP^k\bZ$, normalized to sum to one, determines the weights.

Finding the optimal rotation matrix $\bR_k$ is complex, since $\bR_k$ must be a \emph{rotation} matrix. We use the technique described in \cite{myronenko2009closed}. We first rewrite the objective function to the form $\text{tr}(\bA^T\bR_k)$, for some $\bA$:
\[ 
Q(\bR_k) = \text{tr}\left(\left[\bY^k\bP^k\bZ{\bT^k}^T\right]^T\bR_k\right) \text{,}\;\;
\begin{aligned}
\bY^k &= Y - \by^k {1}^T\\
\bT^k &= \bT - \bt^k {1}^T
\end{aligned}
\]
The optimal $\bR_k$ can then be derived from the singular value decomposition (SVD) of $\bA=\bY^k\bP^k\bZ{\bT^k}^T$: if $\bA = \bU\bS\bV^T$, with $\bU$, $\bS$ and $\bV$ defined as normal for the SVD then we have $\hat \bR_k = \bU \;\text{diag} (1, \ldots, 1, |\bU\bV^T|)\; \bV^T$.

Finally, we derive the scaling parameter $s_k$ by filling in $\hat \bt_k$ and solving $\partial Q(s_k)/\partial s_k = 0$.

The optimization of the post transform follows the same pattern. The optimal values for the weights $\bv$ and the depths $\bv$ follow from straightforward differentiation under constraints. The complete update rules are provided in Figure~\ref{figure:algorithm}. For detailed derivations, see \cite{bloem2016single} Parts of this method are inspired by the Coherent Point Drift (CPD) algorithm \cite{myronenko2010point}.

\begin{figure}[tbh]
{
\hrule
\vspace{3mm}
\raggedright
% Given: a dataset $\bX$, a number of components $K$, a maximum depth $D$. \\
%\tab \\
$\bP_{ij} \leftarrow \cN_j(\bx_i) \bv_{|c_j|} \prod_{a \in c_j} \bw_a$ \hfill \emph{\# Expectation step}\\ 
Normalize $\bP$ so that $\bP\bo = \bo$ \\
\textbf{for each} $k \in [1,K]$: \\
\tab $\bP^k \leftarrow$ submatrix of $\bP$'s columns $j$ for which $c_{j1} = k$\\
\tab \\
$\bY \leftarrow {f^\text{old}_p}^{-1}(\bX)$ \hfill \emph{\# Maximization step}\\
for all $d$, $\hat\bv_d \propto \sum_{j : |c_j| = d} \bP_{ij} $\\
\textbf{for each} $k \in [1,K]$: \\
\tab $\hat\bw_k \leftarrow \bo^T \bP^k \bo / \sum_i {\bo^T \bP^i \bo}$\\
\tab $\by^k \leftarrow {p^k}^{-1}\bY\bP^k\bZ\bo, \,\,\, \bt^k \leftarrow {p^k}^{-1} \bT \bZ{\bP^k}^T \bo$\\
\tab $\bY^k \leftarrow \bY + \by^k\bo^T$, \,\,\, $\bT^k \leftarrow \bT^k + \bt^k\bo^T$\\
\tab $\bZ \leftarrow \text{diag}({s_1}^{-2}, \ldots, {s_{M^-}}^{-2})$\\
\tab $\bU,\bS,\bV^T \leftarrow \text{svd}(\bY^k\bP^k\bZ{\bT^k}^T)$\\ 
\tab $\hat\bR_k \leftarrow \bU\;\text{diag}(1,\ldots,1,|\bU\bV^T|) \bV^T$ \\
\tab $\hat s_k$: solve ${s_k}^{-2}\; \text{tr}({\bY^k}^T\text{d}(\bP^k\bZ\bo)\bY^k)$ \\
\tab \tab $+ {s_k}^{-1}\;\text{tr}(\bT^k \bZ{\bP^k}^T{\bY^k}^T\bR_k) - Hp^k= 0$ \\
\tab $\hat\bt_k \leftarrow \by^k - s_k \bR_k \bt^k$ \\
\tab \\
$\bX^p \leftarrow X - \bx^p\bo^T$, $\bT^p \leftarrow \bT + \bt^p\bo^T$\\
$\bZ \leftarrow \text{diag}({s_1}^{-2}, \ldots, {s_M}^{-2})$\\
$\bU,\bS,\bV^T \leftarrow \text{svd}(\bX^p\bP\bZ{\bT^p}^T)$\\ 
$\hat\bR_p \leftarrow \bU\;\text{diag}(1,\ldots,1,|UV^T|) V^T$ \\
$\hat s_p$: solve \\
\tab ${s_p}^{-2}\text{tr}(\bX^T\text{d}(\bP\bZ\bo)\bX) + {s_p}^{-1} \text{tr}(\bT \bZ\bP^T\bX^T\bR_p) - H p = 0$ \\
$\hat\bt_p \leftarrow \bx^p - s_p\bR_p\bt^p$ \\
\vspace{3mm}
\hrule
}
\caption{One iteration of the IFS-EM algorithm.}
\label{figure:algorithm}
\end{figure}

\section{Results}
\label{section:results}

\begin{figure*}[tbh]
  \centering
   \begin{subfigure}{0.5\linewidth}
	\begin{tabular}{c c c c c c}
		 data  & it 1 & it 20 & it 300 & full \\
		 \run{sierpinski/data}  & \run{sierpinski/best.000000} & \run{sierpinski/best.000019} &  \run{sierpinski/best.000299} & \run{sierpinski/best.000299.deep} \\
		 \run{sierpinski-off/data}  & \run{sierpinski-off/best.000000} & \run{sierpinski-off/best.000019} & \run{sierpinski-off/best.000299} & \run{sierpinski-off/best.000299.deep} \\
		\run{koch2/data} & \run{koch2/best.000000} & \run{koch2/best.000019} & \run{koch2/best.000299} & \run{koch2/best.000299.deep} \\
		 \run{koch4/data}  & \run{koch4/best.000000} & \run{koch4/best.000019} & \run{koch4/best.000299} & \run{koch4/best.000299.deep} \\
		 \run{square/data}  & \run{square/best.000000} & \run{square/best.000019} &  \run{square/best.000299} & \run{square/best.000299.deep}  \\
		 \run{coast/data}& \run{coast/best.000000} & \run{coast/best.000019} & \run{coast/best.000299} & \run{coast/best.000299.deep}  \\
		 \run{broccoli/data} & \run{broccoli/best.000000} & \run{broccoli/best.000019} & \run{broccoli/best.000299} & \run{broccoli/best.000299.deep}  \\
		 \run{sphere/data} & \run{sphere/best.000000} & \run{sphere/best.000019} &  \run{sphere/best.000299} & \run{sphere/best.000299.deep} \\
	\end{tabular}   
	\caption{}
	\label{figure:twod}
  \end{subfigure}  
  \begin{subfigure}{0.49\linewidth}
  	\begin{subfigure}{\linewidth}  
	  	\begin{tabular}{ r r r r r r r}
	  		data & $c$ & 0 & 10 & 50 & 100 & 500 \\
			\hline
			sierpinski & 3 & 0.05 & 0.10 & 0.21 & 0.21 &  0.63 \\
			sierp. (nu) & 3 & 0.04 & 0.68 & 0.95 & 1.00 & \\
			koch  & 2 & 0.16 & 0.27 & 0.30 & 0.33 & 0.48 \\
			koch  & 4 & 0.26 & 0.72 & 0.54 & 0.37 & 0.35 \\
			square & 4 & 0 & 0.32 & 0.42 & 0.60 & 0.61 \\
			\hline
		\end{tabular}
		\caption{}
		\label{figure:convergence}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\begin{subfigure}{0.31\linewidth}  
			\includegraphics[width=\linewidth]{../img/currency/likelihoods.pdf}\\
			currency
		\end{subfigure} 
		\begin{subfigure}{0.31\linewidth}  
			\includegraphics[width=\linewidth]{../img/gait/likelihoods.pdf}\\
			gait
		\end{subfigure} 		
		\begin{subfigure}{0.31\linewidth}  
			\includegraphics[width=\linewidth]{../img/galaxies/likelihoods.pdf}\\
			galaxies
		\end{subfigure} 
		\caption{}		
	\end{subfigure} 
	\begin{subfigure}{\linewidth}
			\begin{subfigure}{0.31\linewidth}  
				\includegraphics[width=\linewidth]{../img/currency/{data.matrix}.png}\\
				data
			\end{subfigure} 
			\begin{subfigure}{0.31\linewidth} 
				\includegraphics[width=\linewidth]{../img/currency/{mog.matrix}.png}\\
				mog
			\end{subfigure}
			\begin{subfigure}{0.31\linewidth}  
				\includegraphics[width=\linewidth]{../img/currency/{pre.matrix}.png}\\
				pre
			\end{subfigure}
			\caption{} 
			\label{figure:scatter}
	\end{subfigure} 
  \end{subfigure}
  \caption{\small (a) Runs on several 2D datasets: the Sierpinski triangle with uniform and non-uniform weights, the Koch curve with 2 and 4 components, a uniform distribution over a square, a section of coast line, shadows on a head of romanesco broccoli and a circle. The red frame represents the bi-unit square before the post transform is applied. Each component maps the red frame onto one of the blue frames. The last column shows a sample from the final model at infinite depth. (b) The proportion of 128 runs that convergence for a given dataset, for different numbers of initial candidates. (c) Likelihoods of the final model on three 3-dimensional datasets compared against two mixture-of-Gaussian models, one constrained to spherical Gaussians (iso), one with unconstraind Gaussians (mog). We report results for a run with a simple initial model (ifs), and one with candidate pre-selection (pre). Inner error bars show the standard error, outer error bars show the range. (3) Scatter matrices for currency experiment.}
  \label{figure:results}
\end{figure*}

When fitting to data from an IFS model, it can be difficult to establish whether the algorithm has converged to the correct model. The same model can be built in many equivalent ways, and the likelihood under the true model grows unbounded with the depth. As a proxy for convergence, we use the model dpeth $\bv$. If at least 95\% of the weight is assigned to the deepest model, we consider the model to have converged. Deep models are highly unstable, so only the true model will converge to higher depths.

For performance reasons, we subsample a minibatch of size $N'$ for each iteration.

The model is very sensitive to the choice of initial parameters. We use two methods to choose the initial model. The first generates $K$ points from a uniform distribution on the bi-unit $H$-sphere, and then constructs the $K$ components with those points as fixed points. Each component has $s = \frac{1}{2}$ and $R$ chosen from the uniform distribution over rotation matrices. The second initialization strategy, called \emph{pre-selection} uses a pool of candidate models (each trained at depth 3 for 100 iterations, with $N' = 500$). We select the model with the highest mean depth ($\sum_i iv_i$) as the initial model.

First, we run the model on several 2D datasets (Figure~\ref{figure:twod}). In all cases, we use preselection with 500 candidates. We then search for 300 iterations, with $N' = 500$and $D=6$. For each dataset, we show the run that resulted in the model with the highest likelihood (on withheld data), out of 8 repeats. For all experiments we set $K=3$, except for the koch curve ($K=2$ and $K=4$) and the square ($K=4$).

The second experiment (Figure~\ref{figure:convergence}) shows the influence of the candidate pool. We repeat the experiments described above 128 times with different sized pools of candidates. We report the percentage of runs that converged. We see that in general more candidates leads to higher convergence, with the exception of the 4-component Koch curve. This may be because mean-depth is not a good pre-selection criterion in this case.

In our final experiment, we apply the model to 3D data and compare it to basic MOG models. Specifically, we test a mixture of isometric Gaussians and an unconstrained MOG. We report the result of a basic IFS and a model with pre-selection on 100 candidates. We repeat each experiment 20 times.

The datasets are: a timeseries of currency conversion rates for three currencies (\cite{franses2000non,hyndman2010time} 4773 points), human motion data from an accelerometer worn around the trunk  (\cite{bachlin2010wearable}\footnote{We select only class-0 motions (i.e. natural motion, recorded in between the experiments).} 777052 points) and the large scale distribution of galaxy clusters in the observable universe (\cite{skrutskie2006two}\footnotemark 5641 points). Each model was trained by Expectation-Maximization. For the MOG models, the whole dataset was used each iteration, for the IFS models, one minibatch of 10 000 points was used, resampling at each iteration.

\footnotetext{Preprocessed data downloaded from \url{https://www.cfa.harvard.edu/~dfabricant/huchra/seminar/lsc/}.}

We note that the IFS model outperforms the isometric MOG (which has similar parameter complexity). The unconstrained MOG model, which has more degrees of freedom than the IFS models, outperforms it, or stays on par with it. These results show that there is no reason to assume that the IFS model will replace the MOG model as a general-purpose probability density model. However, there may be a niche of datasets for which the IFS models is suitable. The scatter matrices shown in Figure~\ref{figure:scatter} show the difference in strategies between the two models.

\section{Discussion}
\label{section:discussion}

We have introduced a new algorithm for the induction of fractal models. To our knowledge, this is the first such algorithm that does not use a general-purpose optimization technique like genetic algorithms.

The model easily generalizes to function classes more general than similitudes, although convex optimization or gradient descent may then be required for the maximization step. Possibilities for future work include developing a variational algorithm \cite{beal2003variational} to increase stability, and the extension of the algorithm to random IFSs \cite{hart1996fractal} to model natural, non-deterministic fractals.

\paragraph{Acknowledgements} We thank Pieter Adriaans for valuable discussions in the early stages of this project. This publication was supported by the Dutch national program COMMIT, by the Netherlands eScience center, and by the Amsterdam Academic Alliance Data Science (AAA-DS) Program Award to the UvA and VU Universities.

\bibliography{fractal}

\end{document}

### 2017-06-09 fractal-plr/img/runs/broccoli/broccoli.sh

@@ -1,0 +1,12 @@
+# shell for the job:
+#PBS -S /bin/bash
+# use one node with eight cores:
+#PBS -lnodes=1:cores16:cpu3:mem64gb
+#PBS -lwalltime=12:00:00
+module load java/1.8u40
+
+cd $HOME/experiments/fractals/runs/broccoli
+
+# call the programs
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --from-image $HOME/datasets/tabular/romanesco/romanesco.png --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+wait          # wait until programs are finished

### 2017-06-09 fractal-plr/img/runs/coast/coast.sh

@@ -1,0 +1,12 @@
+# shell for the job:
+#PBS -S /bin/bash
+# use one node with eight cores:
+#PBS -lnodes=1:cores16:cpu3:mem64gb
+#PBS -lwalltime=12:00:00
+module load java/1.8u40
+
+cd $HOME/experiments/fractals/runs/coast
+
+# call the programs
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --from-image $HOME/datasets/tabular/coastlines/density/australia0.png --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+wait          # wait until programs are finished

### 2017-06-09 fractal-plr/img/runs/koch2/koch2.sh

@@ -1,0 +1,12 @@
+# shell for the job:
+#PBS -S /bin/bash
+# use one node with eight cores:
+#PBS -lnodes=1:cores16:cpu3:mem64gb
+#PBS -lwalltime=12:00:00
+module load java/1.8u40
+
+cd $HOME/experiments/fractals/runs/koch2
+
+# call the programs
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 2 --repeats 8 --samples 500 --iterations 300 --data-source koch --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+wait          # wait until programs are finished

### 2017-06-09 fractal-plr/img/runs/koch4/koch4.sh

@@ -1,0 +1,12 @@
+# shell for the job:
+#PBS -S /bin/bash
+# use one node with eight cores:
+#PBS -lnodes=1:cores16:cpu3:mem64gb
+#PBS -lwalltime=24:00:00
+module load java/1.8u40
+
+cd $HOME/experiments/fractals/runs/koch4
+
+# call the programs
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 4 --repeats 8 --samples 500 --iterations 300 --data-source koch --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+wait          # wait until programs are finished

### 2017-06-09 fractal-plr/img/runs/sierpinski-off/sierpinski-off.sh

@@ -1,0 +1,12 @@
+# shell for the job:
+#PBS -S /bin/bash
+# use one node with eight cores:
+#PBS -lnodes=1:cores16:cpu3:mem64gb
+#PBS -lwalltime=12:00:00
+module load java/1.8u40
+
+cd $HOME/experiments/fractals/runs/sierpinski-off
+
+# call the programs
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --data-source sierpinski-off --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+wait          # wait until programs are finished

### 2017-06-09 fractal-plr/img/runs/sierpinski/sierpinski.sh

@@ -1,0 +1,12 @@
+# shell for the job:
+#PBS -S /bin/bash
+# use one node with eight cores:
+#PBS -lnodes=1:cores16:cpu3:mem64gb
+#PBS -lwalltime=12:00:00
+module load java/1.8u40
+
+cd $HOME/experiments/fractals/runs/sierpinski
+
+# call the programs
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --data-source sierpinski --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+wait          # wait until programs are finished

### 2017-06-09 fractal-plr/img/runs/sphere/sphere.sh

@@ -1,0 +1,12 @@
+# shell for the job:
+#PBS -S /bin/bash
+# use one node with eight cores:
+#PBS -lnodes=1:cores16:cpu3:mem64gb
+#PBS -lwalltime=12:00:00
+module load java/1.8u40
+
+cd $HOME/experiments/fractals/runs/sphere
+
+# call the programs
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --data-source sphere --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+wait          # wait until programs are finished

### 2017-06-09 fractal-plr/img/runs/square/square.sh

@@ -1,0 +1,12 @@
+# shell for the job:
+#PBS -S /bin/bash
+# use one node with eight cores:
+#PBS -lnodes=1:cores16:cpu3:mem64gb
+#PBS -lwalltime=24:00:00
+module load java/1.8u40
+
+cd $HOME/experiments/fractals/runs/square
+
+# call the programs
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 4 --repeats 8 --samples 500 --iterations 300 --data-source square --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+wait          # wait until programs are finished

### 2017-06-09 fractal-plr/img/runs/sub.sh

@@ -1,0 +1,32 @@
+cd broccoli  
+qsub broccoli.sh 
+cd .. 
+
+cd coast  
+qsub coast.sh  
+cd .. 
+
+cd koch2  
+qsub koch2.sh	
+cd .. 
+
+cd koch4  
+qsub koch4.sh  
+cd .. 
+
+cd sierpinski  
+qsub sierpinski.sh  
+cd .. 
+
+cd sierpinski-off  
+qsub sierpinski-off.sh
+cd .. 
+
+cd sphere  
+qsub sphere.sh  
+cd .. 
+
+cd square
+qsub square.sh
+cd .. 
+

### 2017-07-03 fractal-plr/src/fractal.tex
...

@@
 \email{vu@peterbloem.nl}
-\affiliation{VU University Amsterdam}
+\affiliation{Knowledge Representation and Reasoning Group\\
+        VU University Amsterdam \\
+		De Boelelaan 1105, 1081 HV Amsterdam, NL}
 
@@
 \email{s.derooij@gmail.com}
-\affiliation{Leiden University}
+\affiliation{Mathematical Institute \\
+        University of Leiden\\
+        Niels Bohrweg 1, 2333 CA Leiden, NL}
 
@@
 We show that the algorithm reconstructs well-known fractals from data, with the model converging to high precision parameters.
-We also show the utility of the model as an approximation for for datasources outside the IFS model class.
+We also show the utility of the model as an approximation for datasources outside the IFS model class.
 \end{abstract}
@@
 
-\noindent Fractals are mathematical objects; commonly formalised as a class of sets or of probability distributions.
-What, in general, constitutes a fractal is not precisely defined \footnotemark[1], but certain properties are common to most examples; for instance self-similarity, infinitely fine structure, and a non-integer dimension.
-
-\footnotetext[1]{Mandelbrot originally defined a fractal as a set whose topological dimension differs from its Hausdorff dimension, but later retracted this definition, stating that he preferred the word to be not precisely defined.}
+\noindent The word \emph{fractal} is not precisely defined\footnote{Mandelbrot originally defined fractals using Haussdorf dimension, but later stated that he preferred the word to be not precisely defined.}, but typically refers to objects with self-similarity, a non-integer Hausdorff dimension, and an infinitely fine structure.
 
@@
   \caption{An example of an IFS.}
+  \label{figure:example}
 \end{figure}
 
-Since they were named in the 1970s, fractals have increasingly been seen as a potential model for many natural phenomena.
-Mandelbrot put it as follows in \emph{The Fractal Geometry of Nature} \cite{mandelbrot1982fractal}:
+Since they were named in the 1970s, fractals have increasingly been seen as a potential model for many natural phenomena. Mandelbrot put it as follows \cite{mandelbrot1982fractal}: 
 \begin{quote}
@@
 It may be visually clear that a cloud or a coastline `looks' fractal, and we may be able to determine that it has a non-integer dimension \cite{theiler1990estimating}, but how do we get from a dataset to a model?
-How do we determine the parameters of the fractal-generating model that led to this particular image or dataset?
 This is called the \emph{fractal inverse problem}.
@@
 Other approaches are highly domain-specific, such as fractal image compression \cite{hart1996fractal}, which does not generalize well to data of arbitrary dimension.
-Some interesting results have been also been derived from the method of moments \cite{rinaldo1994inverse} and sampling random transformations from the data \cite{hart1997similarity}, but apart from the evolutionary methods, we are not aware of any other generic, practical methods tailored to fractal induction.
+Some interesting results have been also been derived from the method of moments \cite{rinaldo1994inverse} and sampling random transformations from the data \cite{hart1997similarity}, but apart from the evolutionary methods, we are not aware of any other generic, practical methods.
 
 We focus on the family of \emph{Iterated Function Systems} (IFSs) \cite{barnsley2014fractals,hutchinson2000deterministic}, a broad class of fractals, capturing many well-known examples.
-Figure~\ref{figure:examples} shows the basic principle: we start with some initial image and apply a small number, $K$, of contracting transformations to it, resulting in a second image consisting of $K$ scaled down copies of the initial image.
-We then apply the transformations again, to this second image, and so on.
-As we iterate, the image converges to a fractal.
-Which fractal emerges is entirely determined by the chosen transformations.
+Figure~\ref{figure:example} shows the basic principle, (in the domain of images, for simplicity): we start with some initial image and apply a small number, $K$, of contracting transformations to it, resulting in $K$ scaled down copies of the initial image.
+As we iterate this process, the image converges to a fractal.
+Which fractal emerges is entirely determined by the chosen transformations, and entirely independent of the chosen starting image.
 By fixing a family of transformations $\cal F$, we define a family of fractals, each determined by a set of $K$ transformations chosen from $\cal F$: its \emph{components}.
-Here, we choose \emph{similitudes}, a subset of affine maps, as our family of transformations.
-Similitudes offer a good trade-off between expressiveness, and ease of optimization.
+Here, we choose \emph{similitudes}, a subset of affine maps.
+Similitudes offer a good trade-off between expressiveness, and efficient optimization.
 
@@
 The iteration now results in a fractal probability distribution.
-This allows us to frame the fractal inverse problem as a problem of parameter estimation: for a set of points $\{x\} \subset \R^H$, find a set of $K$ similitudes, such that the likelihood of $\{x\}$ under the resulting IFS distribution is maximal.
+This allows us to frame the fractal inverse problem as a problem of parameter estimation: for a set of points $X \subset \R^H$, find the set of $K$ similitudes that maximizes the likelihood  of the resulting IFS distribution given $X$.
 
 We present an Expectation-Maximization (EM) algorithm \cite{dempster1977maximum} to find this model.
-Since the specific sequence of components that ``generated'' a point $x$ in the dataset is unknown, we cast this information as a \emph{latent variable} and iterate between optimizing the latent variables given some model, and optimizing the model given the latent variables.
+Since the specific sequence of components that ``generated'' a point $x$ in the dataset is unknown, we cast this information as a \emph{latent variable} and iterate between optimizing the latent variables given the model, and optimizing the model given the latent variables.
 
-We show that our algorithm quickly reconstructs known fractals, such as the Sierpinski triangle and the Koch curve.
-We also apply the algorithm to some datasets sampled from images, and some natural data of higher dimension, showing that, while there are no IFSs to perfectly capture these images, the self similarity in the model does often allow a good fit, comparable to that of a mixture-of-Gaussians model.
-An open source implementation is available \footnote{\url{https://github.com/pbloem/ifsem}}.
+We show that our algorithm reconstructs known fractals.
+We also apply the algorithm to some datasets sampled from images, and some natural data of higher dimension, showing that, while there are no IFSs to perfectly capture these images, the self similarity in the model does often allow a good fit, comparable in likelihood to that of a Gaussian mixture model (GMM).
+An open-source implementation is available \footnote{\url{https://github.com/pbloem/ifsem}}.
 
@@
 
-Let $\{\bx_i\}_{i\in[1,N]} \subset \R^H$ be our dataset. Let $\bX$ be the $N\times H$ matrix with $\bx_i$ as its columns. All vectors are column vectors.
-
-Let $\bo$ be vector with $1$ at all indices and let $e^i$ be the one-hot vector for $i$, with length implied by context.
+Let $\{\bx_i\}_{i\in[1,N]} \subset \R^H$ be our dataset. Let $\bX$ be the $H\times N$ matrix with $\bx_i$ as its columns.
 
@@
 
-A \emph{Gaussian}, on $\R^H$ is determined by a mean $\bmu \in \R_H$ and a covariance matrix $\bSig \in \R^{H\times H}$.
-We call $\cN_0 = (\boldsymbol 0$, $\bSig = \bI^H)$ the \emph{standard} Gaussian.
-A \emph{spherical} or \emph{isometric} Guassian has $\bSig = s\bI$ for some scalar $s$.
+A \emph{Gaussian}, on $\R^H$ is determined by a mean $\bmu \in \R_H$ and a covariance matrix $\bSig \in \R^{H\times H}$, with $\cN_0 = (\boldsymbol 0, \bI^H)$.
+A \emph{spherical} or \emph{isometric} Gaussian has $\bSig = s\bI$ for some scalar $s$.
 Let $\bx$ be a random variable with $\bx \sim \cN(\bmu, \bSig)$, then $f_{\bt, \bA}(\bx) \sim \cN(\bt + \bA\bmu, \bA\bSig \bA^T)$.
@@
 
-Transforming a spherical MVN by a similitude yields
+Transforming a spherical Gaussian by a similitude yields
 \begin{align}
@@
 \end{align}
+Let $[a ..
+b]$ be the integer interval from $a$ to $b$, including both.
+The \emph{Cartesian power} $[a ..
+b]^n$ is the set of all length $n$ sequences containing only elements from $[a ..
+b]$ (with repetition allowed).
+We extend this definition as follows $[a ..
+b]^{[c ..
+d]} = \bigcup_{i \in [c ..
+d]} [a ..
+b]^i$.
+Define $[a ..
+b]^0$ as the set containing the empty sequence.
 
-\section{The IFS Model}
+\paragraph{The IFS Model}
+
+\begin{figure}[bt]
+  \center
+  \includegraphics[width=\linewidth]{../img/factor-graph.pdf}
+  \caption{A graphical model, illustrating the components of the IFS model.}
+  \label{figure:ifs-diagram}
+\end{figure}
 
@@
 We note two important properties of IFSs.
-First, the distribution on $\bx_D$ converges with $D$: the KL divergence between successive distributions goes to zero.
+First, the distribution on $\bx_D$ converges with $D$.
 We call the distribution it converges to the \emph{limit distribution}.
@@
 Thus, we can see the weights and components as \emph{parameters} of the limit distribution.
-For proofs and generalizations, see \cite{hutchinson2000deterministic}.
+For formal proofs and generalizations, see \cite{hutchinson2000deterministic}.
 
-\begin{figure}[t]
-  \center
-  \includegraphics[width=\linewidth]{../img/factor-graph.pdf}
-  \caption{A graphical model, illustrating the components of the IFS model. The gray plate represents a repetition of the nodes inside for each datapoint.}
-  \label{figure:ifs-diagram}
-\end{figure}
-
-To make the model easier to fit to natural data, we extend the IFS model with some additional factors, as illustrated in Figure~\ref{figure:ifs-diagram}. It contains:
+To make the model easier to fit to natural data, we extend the IFS model with some additional factors, as illustrated in Figure~\ref{figure:ifs-diagram}.
 \begin{description}
@@
 \end{description}
-The first four items form the \emph{parameters} $\theta$ of the model. The rest are \emph{observed} variables $\bx_i$ and the \emph{latent} variables $c_i$.
+The first four form the \emph{parameters} $\theta$ of the model. The rest are \emph{observed} variables $\bx_i$ and the \emph{latent} variables $c_i$.
 
-These variables combine into a model as follows.
-Let $\{c_j\} = [1, K]^{[0, D]}$, the set of all sequences of integers from $[1, K]$ of length up to $D$, including the empty sequence.
+These combine into a model as follows.
+Let $\{c_j\} = [1 ..
+K]^{[0 ..
+D]}$, the set of all sequences containing integers from $[1 ..
+K]$ of length up to and including $D$, including the empty sequence.
 We define the function $f_j:\R^H \to \R^H$ as the composition of the post-transform and the components indicated by the code $c_j$:$f_j = f_p \circ f_{c_{j1}} \circ \ldots \circ f_{c_{jd}}$.
 We can now write the p.d.f.
-of the model as:
+of the model as
 \begin{align*}
-p(x\mid \theta) &= \sum_{j=1}^{|[1, K]^{[0, D]}|} \bv_{|c_j|} \left[\prod_i \bw_{c_{ji}}\right ] f_j(\cN_0)(x) \p
+p(x\mid \theta) &= \sum_{j=1}^{|[1 .. K]^{[0 .. D]}|} f_j(\cN_0)(x)\; \bv_{|c_j|} \prod\nolimits_i \bw_{c_{ji}} \p
 \end{align*}
-That is, a mixture of $|[1, K]^{[0, D]}|$ Gaussians, where the $j$-th Gaussian has the weight $\bv_{|c_j|} \prod_i \bw_{c_{ji}}$.
-Equivalently, we evaluate the IFS to all depths from $0$ to $D$ and mix these with the weights in $\bv$.
+That is, a mixture of $|[1 ..
+K]^{[0 ..
+D]}|$ Gaussians, where the $j$-th Gaussian has the weight $\bv_{|c_j|} \prod_i \bw_{c_{ji}}$.
+Equivalently: we evaluate the IFS to all depths from $0$ to $D$ and mix these with the weights in $\bv$.
 
 Mixing different depths makes the model a generalization of good fall-back models: with $\bv_0 = 1$, the model becomes a spherical Gaussian.
-With $\bv_1 = 1$, the model becomes a mixture of K Gaussians \footnotemark[2].
-Mixing depths also allows the EM search a ``gentle start''.
-Most IFSs have a support with lower dimension than the embedding dimension of the data: i.e.
+With $\bv_1 = 1$, the model becomes a mixture of K Gaussians \footnote{To generalize to generic GMM models, the components should be extended to positive definite affine functions.}.
+Mixing depths also provides a ``gentle start.'' Most IFSs have a support with low dimension: i.e.
 almost all of $\R^H$ has probability zero.
-Thus, the slightest change in parameters means the difference between maximum likelihood, and likelihood zero.
-The variable depth allows the search algorithm to start with low-fitness models with a smooth error surface, and slowly converge to the more challenging, deeper models.
+Thus, the slightest change in parameters means the difference between maximal and zero likelihood.
+The variable depth allows the search algorithm to start with low-fitness models with more a more stable response to perturbations (i.e.
+a smoother error surface), and slowly converge to the more challenging, deeper models.
 
-The post-transform allows us to deal with data that is not properly centered.
+The post-transform caters for data that is not properly centered.
 While we can make off-center IFSs by tuning the components, such models will not overlap properly at different depths.
 The post-transform allows us to learn a centered IFS in place, and map it to the data.
-
-\footnotetext[2]{To generalize to generic MOG models, the components should be extended to positive definite affine functions.}
 
@@
 
-\section{The EM Algorithm}
+\paragraph{The EM Algorithm}
 \label{section:algorithm}
@@
 \end{align*}
-with $M = |[1, K]^{[0, D]}|$ and $\bP$ an $N\times M$ matrix with 
+with $M = |[1 .. K]^{[0 .. D]}|$ and $\bP$ an $N\times M$ matrix with 
 \[
-\bP_{ij} = p(c_j \mid x_i) = \frac{p(x_i\mid c_j) p(c_j) }{\sum_{a \in [1,K]^{[0, D]}} p(x_i\mid c_a) p(c_a)} \p 
+\bP_{ij} = p(c_j \mid x_i) = \frac{p(x_i\mid c_j) p(c_j) }{\sum_{a \in [1 .. K]^{[0 .. D]}} p(x_i\mid c_a) p(c_a)} \p 
 \]
 The computation of $\bP$ is the expectation step of the algorithm: $\bP_{ij}$ represents the \emph{responsibility} that the $j$-th Gaussian takes for the $i$-th datapoint.
-For the maximization step we must optimize $Q$ with respect to the various elements of $\theta$, taking the partial derivative and setting it equal to zero.
-For clarity of notation, when optimizing for a certain subset $q$ of the parameters $\theta$ (such as the depths, weights or parts of the components) we will write $Q(q)$, and omit any terms that are constant with respect to $q$.
-We may also omit any constant multiplier of the whole function.
+For the maximization step we optimize $Q$ with respect to the various elements of $\theta$, taking the partial derivative and setting it equal to zero.
+For clarity of notation, when optimizing for a certain subset $q$ of the parameters $\theta$ we write $Q(q)$, and omit any terms any multipliers constant in $q$.
 
-The optimization of the components and the post-transform $f_k$ and $f_p$ is difficult.
-We must find $K$ maps and weights, and a post-transformation $f_p$, such that all the $M$ endpoint distributions provide maximal likelihood to their assigned points.
+To optimize $f_k$ and $f_p$, we must find $K$ maps and weights, and a post-transformation $f_p$, such that all the $M$ endpoint distributions provide maximal likelihood to their assigned points.
 The problem is that each term in $Q$ is a complicated mix of multiple components.
@@
 Second, we simplify the $Q$-function.
-We first rewrite it as follows: Let $kc_j$ be the code $\langle k, c_{j1}, \ldots\rangle$ and let $M^- = |[1, K]^{[0, D-1}]|$.
+We first rewrite it as follows: Let $kc_j$ be the code $\langle k, c_{j1}, \ldots\rangle$ and let $M^- = |[1 ..
+K]^{[0 ..
+D-1}]|$.
 Let $\bY = {f_p^\text{old}}^{-1}(\bX)$.
@@
 Q&(\{f_k\}, \{w_k\})  \\ 
-&= \frac{1}{s_p}\sum_k \sum_{i, j} &\bP^k_{ij} \ln \left[ f_k(\cN_{j})(y_i) \; v_{|c_j|+1} p(c_j) \; w_k \right]\p
+&= \frac{1}{s_p}\sum_k \sum_{i, j} &\bP^k_{ij} \ln \left[ f_k(\cN_{j})(y_i) \; v_{|c_j|+1} p(c_j) \; w_k \right]
 \end{align*}
-Where we have defined a submatrix $\bP^k$ of $\bP$, which contains all columns corresponding to codes beginning with the symbol $k$: $\bP_{ij}^k = p(kc_j \mid \bx_i)$.
+with $\bP_{ij}^k = p(kc_j \mid \bx_i)$.
 
@@
 
-Using (\ref{line:inverse}), we can rewrite $Q(f_k)$ as a mixture of transformations of $\cN_0$:
+Using (\ref{line:inverse}), we rewrite $Q(f_k)$ as a mixture of transformations of $\cN_0$:
 \begin{align*} 
@@
 \end{align*}
-where $s_j$, $\bR_j$ and $\bt_j$ are the parameters of the similitude $f_p \circ c_{j1} \circ \ldots \circ c_{jd}$.
+where $s_j$, $\bR_j$ and $\bt_j$ are the parameters of the similitude $f_j = f_p \circ f_{c_{j1}} \circ \ldots \circ f_{c_{jd}}$.
 
-To find the optimal translation $\hat\bt_k$, we solve $\partial Q(\bt_k)/\partial \bt_k = 0$, which yields:
+To find $\hat\bt_k$, we solve $\partial Q(\bt_k)/\partial \bt_k = 0$:
 \begin{align*}
@@
 where $\bT$ is the matrix with $\bt_j$ as its columns, $\bZ = \text{diag}({s_1}^{-2}, \ldots, {s_M}^{-2})$ and ${p^k_z} = \bo^TP^kZ\bo$.
-Thus $\hat \bt_k$ is the difference between a weighted mean of the data $\by^k$ and a weighted mean of the MVN means $\bt^k$, scaled and rotated, where in both cases, the matrix  $\bP^k\bZ$, normalized to sum to one, determines the weights.
 
-Finding the optimal rotation matrix $\bR_k$ is complex, since $\bR_k$ must be a \emph{rotation} matrix.
-We use the technique described in \cite{myronenko2009closed}.
-We first rewrite the objective function to the form $\text{tr}(\bA^T\bR_k)$, for some $\bA$:
+Finding the optimal rotation matrix $\hat\bR_k$ is complex, since it must be a \emph{rotation} matrix. We first rewrite the objective function to the form $\text{tr}(\bA^T\bR_k)$, for some $\bA$:
 \[ 
@@
 \begin{aligned}
-\bY^k &= Y - \by^k {1}^T\\
+\bY^k &= \bY - \by^k {1}^T\\
 \bT^k &= \bT - \bt^k {1}^T
@@
 \]
-The optimal $\bR_k$ can then be derived from the singular value decomposition (SVD) of $\bA=\bY^k\bP^k\bZ{\bT^k}^T$: if $\bA = \bU\bS\bV^T$, with $\bU$, $\bS$ and $\bV$ defined as normal for the SVD then we have $\hat \bR_k = \bU \;\text{diag} (1, \ldots, 1, |\bU\bV^T|)\; \bV^T$.
+The optimal $\bR_k$ can then be derived from the singular value decomposition (SVD) of $\bA=\bY^k\bP^k\bZ{\bT^k}^T$: if $\bA = \bU\bS\bV^T$, with $\bU$, $\bS$ and $\bV$ defined as normal for the SVD then we have $\hat \bR_k = \bU \;\text{diag} (1, \ldots, 1, |\bU\bV^T|)\; \bV^T$ \cite{myronenko2009closed}.
 
@@
 
-The optimization of the post transform follows the same pattern.
-The optimal values for the weights $\bv$ and the depths $\bv$ follow from straightforward differentiation under constraints.
+The optimization of the post-transform follows the same pattern.
+The optimal values for the weights $\bw$ and the depths $\bv$ follow from straightforward differentiation under constraints.
 The complete update rules are provided in Figure~\ref{figure:algorithm}.
-For detailed derivations, see \cite{bloem2016single} Parts of this method are inspired by the Coherent Point Drift (CPD) algorithm \cite{myronenko2010point}.
+For detailed derivations, see \cite{bloem2016single}.
+The derivations of $\hat\bt_k$ and $\hat\bR_k$ are inspired by \cite{myronenko2010point}.
 
-\begin{figure}[tbh]
+\begin{figure}[bt]
 {
@@
 Normalize $\bP$ so that $\bP\bo = \bo$ \\
-\textbf{for each} $k \in [1,K]$: \\
+\textbf{for each} $k \in [1 .. K]$: \\
 \tab $\bP^k \leftarrow$ submatrix of $\bP$'s columns $j$ for which $c_{j1} = k$\\
@@
 
-\section{Results}
+\paragraph{Results}
 \label{section:results}
@@
 			sierpinski & 3 & 0.05 & 0.10 & 0.21 & 0.21 &  0.63 \\
-			sierp. (nu) & 3 & 0.04 & 0.68 & 0.95 & 1.00 & \\
+			sierp. (nu) & 3 & 0.04 & 0.68 & 0.95 & 0.99 & 1.00 \\
 			koch  & 2 & 0.16 & 0.27 & 0.30 & 0.33 & 0.48 \\
@@
 				\includegraphics[width=\linewidth]{../img/currency/{mog.matrix}.png}\\
-				mog
+				gmm
 			\end{subfigure}
@@
 (b) The proportion of 128 runs that convergence for a given dataset, for different numbers of initial candidates.
-(c) Likelihoods of the final model on three 3-dimensional datasets compared against two mixture-of-Gaussian models, one constrained to spherical Gaussians (iso), one with unconstraind Gaussians (mog).
+(c) Likelihoods of the final model on three 3-dimensional datasets compared against two mixture-of-Gaussian models, one constrained to spherical Gaussians (iso), one with unconstrained Gaussians (gmm).
 We report results for a run with a simple initial model (ifs), and one with candidate pre-selection (pre).
 Inner error bars show the standard error, outer error bars show the range.
-(3) Scatter matrices for currency experiment.}
+(3) Scatter matrices for the currency experiment.}
   \label{figure:results}
@@
 
-When fitting to data from an IFS model, it can be difficult to establish whether the algorithm has converged to the correct model.
+When fitting to data from an IFS model, it can be difficult to establish whether the algorithm has converged to the target.
 The same model can be built in many equivalent ways, and the likelihood under the true model grows unbounded with the depth.
-As a proxy for convergence, we use the model dpeth $\bv$.
+As a proxy for convergence, we use the model depth $\bv$.
 If at least 95\% of the weight is assigned to the deepest model, we consider the model to have converged.
-Deep models are highly unstable, so only the true model will converge to higher depths.
+Deep models are highly unstable, so only an accurate model will converge to higher depths.
 
@@
 We use two methods to choose the initial model.
-The first generates $K$ points from a uniform distribution on the bi-unit $H$-sphere, and then constructs the $K$ components with those points as fixed points.
-Each component has $s = \frac{1}{2}$ and $R$ chosen from the uniform distribution over rotation matrices.
-The second initialization strategy, called \emph{pre-selection} uses a pool of candidate models (each trained at depth 3 for 100 iterations, with $N' = 500$).
-We select the model with the highest mean depth ($\sum_i iv_i$) as the initial model.
+The first generates $K$ points from a uniform distribution on the bi-unit $H$-sphere, and then constructs the $K$ components with these as fixed points.
+Each component has $s_k = \frac{1}{2}$ and $\bR_k$ chosen from the uniform distribution over rotation matrices.
+With this strategy we see visually accurate models emerging for between 20\% and 70\% of runs, depending on the dataset \cite{bloem2016single}.
+The convergence rate, however, even for models that visually seem like good approximations, is low.
+The second initialization strategy, called \emph{pre-selection} remedies this.
+We use a pool of candidate models, initialized as described above, each trained at depth 3 for 100 iterations with $N' = 500$.
+We select the model with the highest mean depth ($\sum_i iv_i$) as the initial model \footnote{This principle could be extended further, iteratively using the best model at depth $D$, as the initial model for depth $D+1$.
+For our experiments, however, we wish to emphasize the performance of a single-model search.}.
 
 First, we run the model on several 2D datasets (Figure~\ref{figure:twod}).
-In all cases, we use preselection with 500 candidates.
-We then search for 300 iterations, with $N' = 500$and $D=6$.
+In all cases, we use pre-selection with 10 candidates.
+We then search for 300 iterations, with $N' = 500$ and $D=6$.
 For each dataset, we show the run that resulted in the model with the highest likelihood (on withheld data), out of 8 repeats.
@@
 The second experiment (Figure~\ref{figure:convergence}) shows the influence of the candidate pool.
-We repeat the experiments described above 128 times with different sized pools of candidates.
-We report the percentage of runs that converged.
-We see that in general more candidates leads to higher convergence, with the exception of the 4-component Koch curve.
+We repeat the experiments described above 128 times with different sized pools.
+We report the proportion of runs that converged.
+In general, larger pools lead to higher convergence, with the exception of the 4-component Koch curve.
 This may be because mean-depth is not a good pre-selection criterion in this case.
 
-In our final experiment, we apply the model to 3D data and compare it to basic MOG models.
-Specifically, we test a mixture of isometric Gaussians and an unconstrained MOG.
+In our final experiment, we apply the model to 3D data and compare it to basic GMMs (Gausian mixture models).
+Specifically, we test a mixture of isometric Gaussians and an unconstrained GMM.
 We report the result of a basic IFS and a model with pre-selection on 100 candidates.
@@
 The datasets are: a timeseries of currency conversion rates for three currencies (\cite{franses2000non,hyndman2010time} 4773 points), human motion data from an accelerometer worn around the trunk  (\cite{bachlin2010wearable}\footnote{We select only class-0 motions (i.e.
-natural motion, recorded in between the experiments).} 777052 points) and the large scale distribution of galaxy clusters in the observable universe (\cite{skrutskie2006two}\footnotemark 5641 points).
-Each model was trained by Expectation-Maximization.
-For the MOG models, the whole dataset was used each iteration, for the IFS models, one minibatch of 10 000 points was used, resampling at each iteration.
+natural motion, recorded in between the experiments).} 777052 points) and the large scale distribution of galaxy clusters in the observable universe (\cite{skrutskie2006two}\footnote{Preprocessed data downloaded from \url{https://www.cfa.harvard.edu/~dfabricant/huchra/seminar/lsc/}.} 5641 points).
+For the GMMs, the whole dataset was used each iteration.
+For the IFS models we used $N' = 10000$.
+In all cases, we trained for 100 iterations, with $K=4$ and $D=5$.
 
-\footnotetext{Preprocessed data downloaded from \url{https://www.cfa.harvard.edu/~dfabricant/huchra/seminar/lsc/}.}
+The IFS model outperforms the isometric GMM (which has similar parameter complexity).
+The unconstrained GMM, which has more degrees of freedom than the IFS models, tends to outperform the IFS model.
+Clearly, the IFS model will not soon replace the GMM as a general-purpose probability density model.
+However, there may be a niche of domains for which the IFS model is suitable.
+Figure~\ref{figure:scatter} shows the difference in strategies between the two models.
 
-We note that the IFS model outperforms the isometric MOG (which has similar parameter complexity).
-The unconstrained MOG model, which has more degrees of freedom than the IFS models, outperforms it, or stays on par with it.
-These results show that there is no reason to assume that the IFS model will replace the MOG model as a general-purpose probability density model.
-However, there may be a niche of datasets for which the IFS models is suitable.
-The scatter matrices shown in Figure~\ref{figure:scatter} show the difference in strategies between the two models.
-
-\section{Discussion}
+\paragraph{Discussion}
 \label{section:discussion}
 
-We have introduced a new algorithm for the induction of fractal models.
-To our knowledge, this is the first such algorithm that does not use a general-purpose optimization technique like genetic algorithms.
+To our knowledge, this is the first solution to the fractal inverse problem that does not use a general-purpose optimization technique like genetic algorithms.
 
 The model easily generalizes to function classes more general than similitudes, although convex optimization or gradient descent may then be required for the maximization step.
-Possibilities for future work include developing a variational algorithm \cite{beal2003variational} to increase stability, and the extension of the algorithm to random IFSs \cite{hart1996fractal} to model natural, non-deterministic fractals.
+Future work includes developing a variational algorithm \cite{beal2003variational} to increase stability, and the extension of the algorithm to random IFSs \cite{hart1996fractal} to model non-deterministic fractals.
 

### 2017-07-03 fractal-plr/img/runs/broccoli/broccoli.sh

@@ -8,5 +8,5 @@ module load java/1.8u40
 cd $HOME/experiments/fractals/runs/broccoli
 
 # call the programs
-java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --from-image $HOME/datasets/tabular/romanesco/romanesco.png --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --from-image $HOME/datasets/tabular/romanesco/romanesco.png --data-size 50000 --with-init --num-candidates 10 > sys.log 2>&1
 wait          # wait until programs are finished

### 2017-07-03 fractal-plr/img/runs/coast/coast.sh

@@ -8,5 +8,5 @@ module load java/1.8u40
 cd $HOME/experiments/fractals/runs/coast
 
 # call the programs
-java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --from-image $HOME/datasets/tabular/coastlines/density/australia0.png --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --from-image $HOME/datasets/tabular/coastlines/density/australia0.png --data-size 50000 --with-init --num-candidates 10 > sys.log 2>&1
 wait          # wait until programs are finished

### 2017-07-03 fractal-plr/img/runs/koch2/koch2.sh

@@ -8,5 +8,5 @@ module load java/1.8u40
 cd $HOME/experiments/fractals/runs/koch2
 
 # call the programs
-java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 2 --repeats 8 --samples 500 --iterations 300 --data-source koch --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 2 --repeats 8 --samples 500 --iterations 300 --data-source koch --data-size 50000 --with-init --num-candidates 10 > sys.log 2>&1
 wait          # wait until programs are finished

### 2017-07-03 fractal-plr/img/runs/koch4/koch4.sh

@@ -8,5 +8,5 @@ module load java/1.8u40
 cd $HOME/experiments/fractals/runs/koch4
 
 # call the programs
-java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 4 --repeats 8 --samples 500 --iterations 300 --data-source koch --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 4 --repeats 8 --samples 500 --iterations 300 --data-source koch --data-size 50000 --with-init --num-candidates 10 > sys.log 2>&1
 wait          # wait until programs are finished

### 2017-07-03 fractal-plr/img/runs/sierpinski-off/sierpinski-off.sh

@@ -8,5 +8,5 @@ module load java/1.8u40
 cd $HOME/experiments/fractals/runs/sierpinski-off
 
 # call the programs
-java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --data-source sierpinski-off --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --data-source sierpinski-off --data-size 50000 --with-init --num-candidates 10 > sys.log 2>&1
 wait          # wait until programs are finished

### 2017-07-03 fractal-plr/img/runs/sierpinski/sierpinski.sh

@@ -8,5 +8,5 @@ module load java/1.8u40
 cd $HOME/experiments/fractals/runs/sierpinski
 
 # call the programs
-java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --data-source sierpinski --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --data-source sierpinski --data-size 50000 --with-init --num-candidates 10 > sys.log 2>&1
 wait          # wait until programs are finished

### 2017-07-03 fractal-plr/img/runs/sphere/sphere.sh

@@ -8,5 +8,5 @@ module load java/1.8u40
 cd $HOME/experiments/fractals/runs/sphere
 
 # call the programs
-java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --data-source sphere --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 3 --repeats 8 --samples 500 --iterations 300 --data-source sphere --data-size 50000 --with-init --num-candidates 10 > sys.log 2>&1
 wait          # wait until programs are finished

### 2017-07-03 fractal-plr/img/runs/square/square.sh

@@ -8,5 +8,5 @@ module load java/1.8u40
 cd $HOME/experiments/fractals/runs/square
 
 # call the programs
-java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 4 --repeats 8 --samples 500 --iterations 300 --data-source square --data-size 50000 --with-init --num-candidates 500 > sys.log 2>&1
+java -Xmx56g -jar $HOME/lilian/ifsem.jar --ifs.hq --type ifs --components 4 --repeats 8 --samples 500 --iterations 300 --data-source square --data-size 50000 --with-init --num-candidates 10 > sys.log 2>&1
 wait          # wait until programs are finished

### 2017-07-05 kgmotifs/kgmotifs.tex
...

\documentclass[11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}

\newcommand{\N}{{\mathbb N}}

\title{Detecting Network Motifs in Knowledge Graphs using Compression}

\begin{document}

\maketitle

\begin{abstract}
We introduce a method to detect \emph{network motifs} in knowledge graphs. Network motifs are useful patterns or meaningful subunits of the graph that recur frequently. We introduce a scalable approach for detecting such motifs in large knowledge graphs, inspired by recent work for simple graphs, and show that the motifs returned reflect the basic structure of the graph. Specifically, we show that common motifs reflect graph patterns used in common queries, basic schematic units of the graph and meaningful functional subunits, for various knowledge graphs.
\end{abstract}

The Linked, Open Data cloud contains a wealth of knowledge graphs, and is growing quickly. At the time of writing the average knowledge graph contains overt 65 000 edges, with many datasets containing more than a million \footnotemark. For such large graphs, it can be difficult to see the forest for the trees: how is the graph structured at the lowest level? What kind of things can I ask of what types of objects? What are small recurring patterns that might represent a novel insight into the data?

\footnotetext{\url{http://stats.lod2.eu/stats}}

In the domain of plain (unlabelled graphs), \emph{network motifs} were recently introduced as a tool to provide users of large graphs insight into their data. Network motifs are small subgraphs whose frequency is unexpected with respect to a null model. That is, for a given graph dataset $g$ and small graph $m$, we count the frequency $F(m, g)$: how often $m$ occurs in $g$ as a subgraph. We define a probability distribution over graphs $p^\text{null}(g)$, and estimate the probability that a graph sampled from $p^\text{null}$ contains more instances of $m$ than observed in our data: $p^\text{null}(F(m, G) \geq F(m, g))$, where $G$ is a random variable representing a graph. If this probability is low (commonly, below 0.05), we consider $m$ a motif. \footnotemark

\footnotetext{This procedure has the structure of a hypothesis test, but it is important not to interpret it as statistical evidence for the meaningfulness of the motif. The only thing it proves (in a frequentist statistical sense) is that $p^\text{null}$ is not the true source of the data. This is usually not a surprise: we are rarely able to model all aspects of a realistic data-generating process in a single distribution. The $p$-values used in motif analysis should be interpreted strictly as \emph{heuristics}. See also \cite{bloem2017large}.}

In \cite{bloem2017large}, an alternative method is presented that uses compression as a heuristic for motif value: the better a motif compresses the data, the more likely it is to be meaningful. It is shown that this principle can be implemented in a similar sort of hypothesis test for a null model, allowing a comparable workflow to classical motif analysis. 
In this paper, we extend the compression-based motif analysis to Knowledge Graphs. For the purposes of this research we define Knowledge graphs as labeled multigraphs. Nodes are labeled with entity names, and links are labeled with relations. We extend the definition of a motif to that of a basic graph pattern: that is, a motif is a small graph with some of its nodes and links labeled. The motif occurs in the graph where the graph structure matches, and the labels. The unlabeled nodes and links are free to contain whatever label when the motif is mapped to the data.
To maintain the connection to graph pattern queries, we do not require the motifs to be\emph{induced} subgraphs: that is, when a motif is mapped to a graph, the graph can contain additional links that are not specified by the motif. 
We perform several experiments to show that our method returns meaningful subgraphs. First, we look at several graphs for which schema information is provided. That is, we know the basic properties available for each entity. We show that the motifs found by our method correspond well to the schema, and compare performance against several basic benchmarks. Second, we run motif analysis on several datasets for which a list of real SPARQL queries entered by users is available. From these, we extract the basic graph patterns, and show that (a) these correspond to good motifs using our criterion and (b) our method can return such motifs with good precision and recall.
All code and benchmark datasets used in the paper are available, under open licenses.

% TODO: Make illustration

\subsection{Related Work}

\subsection{Preliminaries}

\section{Method}

We will first describe our method under the assumption that a good null model $p^\text{null}$ is available, and that $- \log p^\text{null}(G)$ can be efficiently computed. In Section~\ref{section:null-model}, we explain which null model we use for our experiments, and how it is implemented.
We will start with some formal definitions of our basic ingredients. To help with notation, we make the simplifying assumption that all names (for entities and relations) have been mapped to the natural numbers in some arbitrary way.  We can then define a \textbf{knowledge graph} $G$ as a finite set of triples of integers: $G \subseteq \N \times\N \times \N$. If $(s, p, r) \in G$ with $s, p, r \in \N$, then the entities mapped to integers $s$ and $o$ are related by relation $r$.
This representation means that when we used codes for all integers to represent these triples, that those nodes and relations assigned to small natural numbers will receive smaller codes. In practice, however we will always remap these integers to other symbols based on their frequency in the data, so this mapping will not affect performance. We may also worry that the names contain structure \emph{internally} that could be used for compression (since the labels of knowledge graphs are usually URLs), as was shown to be true in \cite{}. For our purposes, however we would like our analysis to use only the graph structure itself, and not the internal structure, so this simplified definition is a good way to ensure that we are blind to the internal structure of the labels.

\subsection{Motif search}

\subsection{Relevance test}

\subsection{Motif code}

\section{Null model}

\label{section:null-model}

\section{Experiments}

\subsection{Schema reconstruction}

\subsection{Query induction}

\subsection{Scale}

To show the scale of our method, we show its performance on some very large graphs, culminating with adump of the complete LOD cloud, from \cite{}.

\subsection{Various}

% Note: only if there's time. If not, save for later.

In this section, we show some experiments that are not intended to substantiate any claims about our method, but rather to illustrate the variety of uses for succesfull motif detection.

\subsubsection{Analogical reasoning}

\subsubsection{Visualization}

\subsubsection{Feature extraction}

\subsubsection{Node embedding}

\bibliography{kgmotifs}
\end{document}

### 2017-07-05 kgpsr/kgpsr.tex
...

\documentclass[11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}

\title{A Variational Approach to Knowledge Graph Node Embeddings}

\begin{document}

\newcommand{\R}{{\mathbb R}}
\newcommand{\fw}[1]{\text{\texttt{#1}}}

\maketitle

\begin{abstract}

	We present a Variational approach to to the problem of embedding knowledge graph nodes.
\end{abstract}

\section{Introduction}
	Knowledge graph node embeddings are vector representations of the nodes in a knowledge graph. Effective embeddings are an important first step in many pipelines for knowledge graphs visualisation, completion, classification and other tasks. Many embedding algorithms model the relations in the knowledge graph as maps in the embedding space: either translations, or more complex transformations. If for instance, we have the triple $(\fw{John}, \fw{parentof}, \fw{Mary})$, then we learn embeddings $v_\fw{John}, v_\fw{Mary} \in \R^k$  (for some chosen embedding dimension $k$) and a map $f_\fw{parentof} : \R^k \to \R^k$ for the relation \texttt{parentof} such that $f_\fw{parentof}(v_\fw{John})$ should be as close to $v_\fw{Mary}$ as possible.
	This scheme works well, but has trouble representing multiple relations. If John has more than one child, what should $f_\fw{parentof}(v_\fw{John})$ be? Many approaches have been introduced, often requiring an advanced mathematical structure for the embedding space, such as complex vectors \cite{trouillon2016complex}, or holographic projections \cite{nickel2016holographic}.
	We introduce a variational approach. We model each knowledge graph node not as a point in the embedding space, but as a spherical Gaussian of fixed variance. Transforming this Gaussian by the relational map should then result in a new Gaussian which maximizes the likelihood of the nodes connected to the original by the given relation.\footnotemark
	\footnotetext{This approach is inspired by the Coherent Point Drift algorithm \cite{}, which uses a similar approach for the problem of point set registration.}
	The variational approach has several advantages:
	\begin{itemize}
		\item The embedding space is a simple Euclidean space, making no special demands on downstream architectures.
		\item The multiple-relation issue is solved in an intuitive manner: mapping a parent node to its children creates a probability distribution that covers the children. 
		\item For linear maps, we can decompose the loss function into factors which can be solved analytically. Thus, the problem can be solved using an iterative algorithm which is guaranteed to converge to a local optimum.
		\item Negative sampling is not required. The basic framework constrains the problem sufficiently to lead to models that generalize well.
	\end{itemize}

	We test our embeddings on the tasks of knowledge base completion and node classification. We outperform the state-of the art in each case. We also compare a general-purpose stochastic gradient descent approach for our model, to the variational search, showing that the latter converges faster, in fewer iterations, and to better solutions.

\section{Model}

\section{Experiments}

\section{Conclusion}

\bibliography{kgpsr}

\end{document}

### 2017-07-05 sparse-hyper/Sparse-hyper.tex
...

\documentclass[11pt]{article}

\title{Sparse, Adaptive Hypernetworks}

\begin{document}

\maketitle

\begin{abstract}
Many neural network architectures benefit from sparse connections between layers This sparseness can be designed (as in convolutional neural networks), but resent experiments have shown that it can also be learned, even as part of the end-to-end learning by gradient descent. Unfortunately, such sturcture learning requires all potential weights to be stored in memory (usually in a dense matrix structure): few weights will be nonzero, but we do not know which ones, so all zero values must be stored explicitly. To remedy this problem, we introduce sparse, adaptive hypernetworks. We use a second, dense network to store the weights of the sparse layer---a hypernetwork---and construct the sparse layer dynamically. Moreover, we provide the data as input to the hypernetwork so the structure of the sparse layer can change adaptively. We show that such sparse, adaptive layers can be used to build models that achieve state-of-the-art performance in several domains where large, sparse structures and transformations are important, with minimal memory footprint.
\end{abstract}

\section{Introduction}

Many forms of data in modern machine learning are highly structured: vector images, (knowledge) graphs, natural language, etc. Such data is usually best stored in sparse format: usually using one-hot coding for the individual elements, and something akin to an adjacency matrix for the relations holding them together. Consequently, the transformations we want to learn for such instances are also highly sparse.

Consider, for instance the example of learning a graph-to-graph transformation for directed graphs. We will allow self-loops, but not multiple connections, so that each graph can be represented by a square, binary adjacency matrix. Let us also assume that we want to learn a simple graph transformation (possibly as part of a large network). The most straightforward approach would be to use a single, fully-connected layer from one adjacency matrix to another. Let $x$ be the input adjacency matrix, $W$ the wight matrix, and $y$ the output.

In this scenario, our input is likely very sparse: for large, natural graphs, most elements of the adjacency will likely be zero. For good transformations, most weights will likely also be zero: an effective transformation should exploit the locality of the data as suggested by the graph structure. Unfortunately, the structure of our data is not regular. When processing images, we know in advance that each instance will consist of a grid of pixels, and we can wire our network accordingly for the whole dataset. In our case, each instance contains a new structure, which is encoded in the input itself. We must therefore learn the sparse structure of our transformation adaptively.

In a naive implementation, this would still require a dense matrix of weights, even though the matrix we store will be sparse. We don't know if a particular weight will be zero or not for any given instance, so all zero weights must be stored explicitly. Assuming that the output of our transformation is as large as the input, the number of weights of our layer grows as the square of the size of the input. Since the size of the input is already the square of the number of nodes in the network, this is clearly an unfeasible approach for large graphs.

We solve this problem with \emph{sparse, adaptive hypernetworks}. Hypernetworks \cite{ha2016hypernetworks} operate by using a secondary, small network to produce the weights of  the primary network layer. We make this approach adaptive, by supplying the input of the layer to the hypernetwork as well, allowing the hypernetwork to change the wiring of the main network on the fly. We apply a regulariser to steer the optimisation towards sparse $W$, combined with a clipping of all but the highest $k$ weights produced by the hypernetwork.

We test the performance of hypernetworks in three domains. First, we show that an image autoencoder with sparse adaptive layers can match the performance of one with a similar number of parameters and weights. Additionally, we show that the adaptive wiring acts as a kind of attention mechanism, allowing the network to devote more parameters to areas that require more attention. Second, we test the hypothesis that multilayer sparse networks can be used to learn complex invariant representations of graphs. We design a graph classification task which cannot be solved without successfully detecting isomorphic graphs as an intermediate step. We train a model end-to-end and show that after training, the embeddings learning for graphs in the same isomorphism class show strong clustering in the embedding space. Finally, we show that sparse, adaptive layers can be used in \emph{generative models}. We use a variational autoencoder \cite{kingma2013auto} consisting of sparse, adaptive layers to model a collection of MIDI files stored as knowledge graphs (using the midi-ld standard). We show that the knowledge graphs sampled from the resulting distribution adhere to the required standard (i.e. they can be parsed into music) and result in generated music that is judged to be more likely to have been produced by a human being than several baseline music generators.

\section{Model}

\section{Experiments}

\section{Conclusion}

\bibliography{sparse-hyper}

\end{document}

### 2017-07-10 kgmotifs/kgmotifs.tex
...

@@
 \usepackage{amssymb}
+\usepackage{url}
 
 \newcommand{\N}{{\mathbb N}}
+\newcommand{\G}{{\cal G}}
 
@@
 
-The Linked, Open Data cloud contains a wealth of knowledge graphs, and is growing quickly.
-At the time of writing the average knowledge graph contains overt 65 000 edges, with many datasets containing more than a million \footnotemark.
-For such large graphs, it can be difficult to see the forest for the trees: how is the graph structured at the lowest level?
+\noindent The Linked, Open Data cloud contains a wealth of knowledge graphs, and is growing quickly.
+At the time of writing the average knowledge graph contains over 65 000 edges, with many datasets containing more than a million.\footnotemark~For such large graphs, it can be difficult to see the forest for the trees: how is the graph structured at the lowest level?
 What kind of things can I ask of what types of objects?
-What are small recurring patterns that might represent a novel insight into the data?
+What are small, recurring patterns that might represent a novel insight into the data?
 
@@
 
-In the domain of plain (unlabelled graphs), \emph{network motifs} were recently introduced as a tool to provide users of large graphs insight into their data.
+In the domain of plain (unlabelled) graphs, \emph{network motifs} \cite{milo} were recently introduced as a tool to provide users of large graphs insight into their data.
 Network motifs are small subgraphs whose frequency is unexpected with respect to a null model.
@@
 
+The most common null model for classical motif analysis is the degree-sequence model (also known a the configuration model \cite{}): a uniform distribution over all graphs that share the same in and out degrees of the data for every node.
+We extend this to knowledge graphs by also including the frequency with which each relation occurs.
+Let a \emph{degree sequence} $D$ of length $n$ be a triple of three integer sequences: $(D^\text{in}, D^\text{rel}, D^\text{out})$.
+If $D$ is the degree sequence of a graph, then node $i$ has $D^\text{in}_i$ incoming links,  $D^\text{out}_i$ outgoing links and for each relation $r$, there are $D^\text{rel}_r$ links.
+
+Let $\G_D$ be the set of all graphs with degree sequence $D$. Then the configuration model can be expressed simply as
+\[
+p(G) = \frac{1}{|\G_D|}
+\]
+for any $G$ that satisfies $D$ and $p(G) = 0$ otherwise.
+Unfortunately, there is no efficient way to compute $|\G_D|$ and even approximations tend to be costly for large graphs.
+Following the approach in \cite{bloem2017large}, we define an approximation.
+We can define a knowledge graph by three length-$m$ integer sequences: $S$, $P$, $O$, with ${(S_j, P_j, O_j)}_j$ the graph's tripleset.
+If the graph satisfies degree sequence $D$, then we know that $S$ should contain node $j$ $D^\text{out}_j$ times, $P$ should contain relation $r$ $D^\text{rel}_r$ times and $O$ should contain node $j$ $D^\text{in}_j$ times.
+Let ${\cal S}_D$ be the set of all such triples of integer sequences satisfying $D$.
+We have
+\[
+|{\cal S}_D| =
+ {m \choose {D_1^\text{out}, \ldots, D_n^\text{out}} }
+ {m \choose {D_1^\text{rel}, \ldots, D_n^\text{rel}} }
+ {m \choose {D_1^\text{in}, \ldots, D_n^\text{in}} } \text{.}
+\]
+
+While every member of ${\cal S}_D$ contains a valid graph satisfying $D$, many graphs are represented multiple times.
+Firstly, many elements of ${S}_D$ contain the same link multiple times, which means they are not is ${\cal G}_D$.
+We call the set without these elements ${\cal S'}_D \subset {\cal S}_D$.
+Secondly the links of the graph are listed in arbitrary order; if we apply the same permutation to all three lists $S$, $P$ and $O$, we get an encoding of the same graph.
+Since we know that any element in ${\cal S}'_D$ contains only unique triples, we know that each graph is present exactly $m!$ times.
+Thus, we have
+\[
+|\G_D| = |{\cal S}'_D| \frac{1}{m!} \leq  |{\cal S}_D| \frac{1}{m!} \text{.}
+\]
+
+We can thus use 
+\[
+p^\text{EL}_D(G) =  \frac{m!}{{m \choose {D_1^\text{out}, \ldots, D_n^\text{out}} }
+ {m \choose {D_1^\text{rel}, \ldots, D_n^\text{rel}} }
+ {m \choose {D_1^\text{in}, \ldots, D_n^\text{in}} }}
+\]
+as an approximation for the DS model.
+We call this the edge-list (EL) model.
+It always lower-bounds the true degree sequence model, since it affords some probability mass to graphs that cannot exist.
+\footnotemark Experiments in the classical motif setting have shown that the ES model is an acceptable proxy \cite{bloem2017large}, especially considering the extra scalability it affords.
+
+\footnotetext{Note that we cannot simply think of $p^\text{ES}$ as a uniform model for graphs containing multiple triples, since in we can only divide by $m!$ is we know that all triples are unique.}
+
+ \paragraph{Encoding D} In order to encode a graph with $L^\text{EL}_D$, we must first encode $D$.
+\footnote{Or, equivalently, to make $p^\text{EL}$ a complete distribution on all graphs, we must provide it with a prior on $D$.} For each of the three sequences in $D$ we use the following model:
+\[
+ p(D) = \prod_i p^{\N}(D_i) \\
+ L(D) = - \sum_i \log p^{\N}(D_i)
+\]
+
+where $p^{\N}$ is a distribution on the natural numbers.
+This is an optimal encoding for $D$ assuming that its members are independently drawn from $p^{\N}$.
+When we use $p^\text{DS}$ as the null model, we use the data distribution for $p^{\N}$ to ensure that we have a lower bound to the optimal code-length.
+When we use $p^\text{DS}$ as part of the motif code, we must use a fair encoding, so we use a Dirichlet-Multinomial model to store $D$.
+
 \section{Experiments}
@@
 
-To show the scale of our method, we show its performance on some very large graphs, culminating with adump of the complete LOD cloud, from \cite{}.
+To show the scale of our method, we show its performance on some very large graphs, culminating with a dump of the complete LOD cloud, containing 38 billion triples, from \cite{}.
 

### 2017-07-10 kgmotifs/kgmotifs.tex
...

@@
 
-In the domain of plain (unlabelled) graphs, \emph{network motifs} \cite{milo} were recently introduced as a tool to provide users of large graphs insight into their data.
+In the domain of plain (unlabelled) graphs, \emph{network motifs} \cite{milo2002network} were recently introduced as a tool to provide users of large graphs insight into their data.
 Network motifs are small subgraphs whose frequency is unexpected with respect to a null model.
@@
 It is shown that this principle can be implemented in a similar sort of hypothesis test for a null model, allowing a comparable workflow to classical motif analysis.
+
 In this paper, we extend the compression-based motif analysis to Knowledge Graphs.
@@
 The unlabeled nodes and links are free to contain whatever label when the motif is mapped to the data.
-To maintain the connection to graph pattern queries, we do not require the motifs to be\emph{induced} subgraphs: that is, when a motif is mapped to a graph, the graph can contain additional links that are not specified by the motif.
+
+To maintain the connection to graph pattern queries, we do not require the motifs to be \emph{induced} subgraphs: that is, when a motif is mapped to a graph, the graph can contain additional links that are not specified by the motif.
+
 We perform several experiments to show that our method returns meaningful subgraphs.
@@
 From these, we extract the basic graph patterns, and show that (a) these correspond to good motifs using our criterion and (b) our method can return such motifs with good precision and recall.
-All code and benchmark datasets used in the paper are available, under open licenses.
+
+All code and benchmark datasets used in the paper are available, under open licenses.\footnotemark
+
+\footnotetext{\url{...}}
 
@@
 In Section~\ref{section:null-model}, we explain which null model we use for our experiments, and how it is implemented.
+
 We will start with some formal definitions of our basic ingredients.
-To help with notation, we make the simplifying assumption that all names (for entities and relations) have been mapped to the natural numbers in some arbitrary way.
-We can then define a \textbf{knowledge graph} $G$ as a finite set of triples of integers: $G \subseteq \N \times\N \times \N$.
-If $(s, p, r) \in G$ with $s, p, r \in \N$, then the entities mapped to integers $s$ and $o$ are related by relation $r$.
-This representation means that when we used codes for all integers to represent these triples, that those nodes and relations assigned to small natural numbers will receive smaller codes.
-In practice, however we will always remap these integers to other symbols based on their frequency in the data, so this mapping will not affect performance.
-We may also worry that the names contain structure \emph{internally} that could be used for compression (since the labels of knowledge graphs are usually URLs), as was shown to be true in \cite{}.
-For our purposes, however we would like our analysis to use only the graph structure itself, and not the internal structure, so this simplified definition is a good way to ensure that we are blind to the internal structure of the labels.
+To comfortably define probabilities and codes on graphs, we will slightly deviate from conventional definitions.
+Let $V$ be a countably infinite set containing all possible entities, and let $R$ be a countably infinite set containing all possible relations between them.
+Let $v: V \to \N$ and $r: R \to \N$ be arbitrary mappings of these sets to the natural numbers.\footnotemark
+
+\footnotetext{For instance, let $V$ and $E$ be the set of finite strings, and let $v$ and $e$ represent length-lexicographical order.}
+
+A \emph{knowledge graph} $G$, hereafter known simply as a \emph{graph} is a, is tuple $G = (V_G, E_G)$ with finite sets $V_G \subset V$ and $E_G \subset V \times R \times V$.
+We define the tripleset of $G$ as $G^T = \{(s, p, o) \in \N^3 : \left (v(s), r(p), v(o) \right ) \in E_G\}$.
+In this paper we will \emph{identify} the graph with the triple list.
+That is when we speak of encoding the graph, we mean encoding the triple list, and when we speak of a probability distribution over graphs, we mean a probability distribution over triple lists.
+
+This means that we are purposely \emph{not} modelling certain aspects of the data.
+There might, for instance bet a similarity between two node labels $v(a)$ and $v(b)$ (such as a shared IRI prefix).
+We are not interested in exploiting such structure, so we eliminate it from our notation.
+Additionally, is we use a simple code to store the integers in the triple-list we will end up using fewer bits for those nodes and triples with labels who were arbitrarily assigned short integers.
+Equivalently, when sampling from a simple distribution over the integers, small numbers will be more likely to be generated.
+In practice, however, we will always learn a distribution over the integers from the data, so that the choice of $e$ and $r$ will not affect the code.
+
+As shown in Figure~\ref{figure:diagram}, the motifs that we aim to find are partially labeled: some nodes and edges are labeled with the marker "?" indicating that when we fit the motif to the graph, those edges can contain any label.
+In other words, the motif is a \emph{basic graph pattern}, and the edges and nodes marked with "?" are variables.
+To represent this, we add a single variable symbol $?$.
+Thus, a motif $M$ is defined by the triple set $M \subset (?
+\cup \N) ^3$, with integer values in triples interpreted as in complete knowledge graphs.
 
@@
 
+Given a graph $G$, our aim is to find motifs $M$ that represent meaningful building blocks, or that are otherwise characteristic for $G$.
+In the design of our method, we will constantly aim to find a trade-off between completeness and efficiency that allows the method to scale to very large graphs.
+Specifically, when we economize, we will only do so in a way that makes the hypothesis test described in Section~\ref{section:relevance-test} \emph{more conservative}.
+For instance, we will use a very fast search for candidate motifs, which returns only a limit number of instances.
+A more expensive search might yield more motifs, but never less.
+That is, for those motifs that we find, we can be sure that they would also be found by a more expensive method.
+Since pattern mining of this kind often results in too many potential patterns rather than too few, it seems efficient to start with the motifs that can be found with fast, simple methods, and only extend the search once these have been exhausted.
+
+Specifically, we generate candidate motifs by \emph{sampling} them from the data.
+This method, inspired by \cite{kashtan2004efficient} is based on the simple insight that when we sample a pattern from the data, we are more likely to sample a frequently occurring pattern than a infrequent one.
+In we sample $K$ times, and rank the results by relative frequency within the sample, we are likely toe end up with a good indication of the most frequent patterns in the data.
+For unlabeled graphs, as little as 500 samples can be sufficient to find the most frequent subgraph.
+Theoretically, this allows us to generate promising motifs with time complexity indepenedent of the size of the graph. Our sampling algorithm is described in Figure~\ref{algorithm:sampling}
+
 \subsection{Relevance test}
+
+\label{section:relevance-test}
 
@@
 \bibliography{kgmotifs}
+
 \end{document}

### 2017-07-28 kgmotifs/kgmotifs.tex
...

@@
 \newcommand{\G}{{\cal G}}
+\newcommand{\I}{{\cal I}}
+\newcommand{\tab}{\hspace*{5mm}}
+\newcommand{\from}{\leftarrow}
 
@@
 
+Unfortunately, counting all occurrences of a motif in a network is an expensive operation.
+One that we need to perform not only on the data, but also on many samples from the null model, in order to estimate the probability described above.
+
 \footnotetext{This procedure has the structure of a hypothesis test, but it is important not to interpret it as statistical evidence for the meaningfulness of the motif.
@@
 It is shown that this principle can be implemented in a similar sort of hypothesis test for a null model, allowing a comparable workflow to classical motif analysis.
+Crucially, this method scales to graphs with billions of edges, suggesting that it might be suitable for the scale of knowledge graphs we encounter in the LOD cloud.
 
@@
 Nodes are labeled with entity names, and links are labeled with relations.
-We extend the definition of a motif to that of a basic graph pattern: that is, a motif is a small graph with some of its nodes and links labeled.
-The motif occurs in the graph where the graph structure matches, and the labels.
+We extend the definition of a motif to that of a basic graph pattern: that is, a motif is a small graph with \emph{some} of its nodes and links labeled.
+The motif \emph{occurs} in a graph where both the graph structure and the labels match.
 The unlabeled nodes and links are free to contain whatever label when the motif is mapped to the data.
@@
 
-We will first describe our method under the assumption that a good null model $p^\text{null}$ is available, and that $- \log p^\text{null}(G)$ can be efficiently computed.
-In Section~\ref{section:null-model}, we explain which null model we use for our experiments, and how it is implemented.
-
 We will start with some formal definitions of our basic ingredients.
@@
 
+% TODO: Clarify that we aren't storing node labels
+
 A \emph{knowledge graph} $G$, hereafter known simply as a \emph{graph} is a, is tuple $G = (V_G, E_G)$ with finite sets $V_G \subset V$ and $E_G \subset V \times R \times V$.
-We define the tripleset of $G$ as $G^T = \{(s, p, o) \in \N^3 : \left (v(s), r(p), v(o) \right ) \in E_G\}$.
+We define the triple list of $G$ as $G^T = \{(s, p, o) \in \N^3 : \left (v(s), r(p), v(o) \right ) \in E_G\}$.
 In this paper we will \emph{identify} the graph with the triple list.
 That is when we speak of encoding the graph, we mean encoding the triple list, and when we speak of a probability distribution over graphs, we mean a probability distribution over triple lists.
+Let $R_G \cub \N$ be the set of relations used in the $G$.
 
 This means that we are purposely \emph{not} modelling certain aspects of the data.
-There might, for instance bet a similarity between two node labels $v(a)$ and $v(b)$ (such as a shared IRI prefix).
+There might, for instance be a similarity between two node labels $v(a)$ and $v(b)$ (such as a shared IRI prefix).
 We are not interested in exploiting such structure, so we eliminate it from our notation.
-Additionally, is we use a simple code to store the integers in the triple-list we will end up using fewer bits for those nodes and triples with labels who were arbitrarily assigned short integers.
+Additionally, we introduce some bias with our choice of $e$ and $r$.
+If we were to use a simple code to store the integers in the triple-list we will end up using fewer bits for those nodes and triples with labels who were assigned short integers.
 Equivalently, when sampling from a simple distribution over the integers, small numbers will be more likely to be generated.
-In practice, however, we will always learn a distribution over the integers from the data, so that the choice of $e$ and $r$ will not affect the code.
+In practice, however, we will always \emph{learn} a distribution over the integers from the data, so that $e$ and $r$ are purely a means to simplify the notation.
+Which functions we choose will not affect our model.
 
@@
 Thus, a motif $M$ is defined by the triple set $M \subset (?
-\cup \N) ^3$, with integer values in triples interpreted as in complete knowledge graphs.
+\cup \N) ^3$, with integer values in triples interpreted as they are in complete knowledge graphs.
 
-\subsection{Motif search}
-
-Given a graph $G$, our aim is to find motifs $M$ that represent meaningful building blocks, or that are otherwise characteristic for $G$.
-In the design of our method, we will constantly aim to find a trade-off between completeness and efficiency that allows the method to scale to very large graphs.
-Specifically, when we economize, we will only do so in a way that makes the hypothesis test described in Section~\ref{section:relevance-test} \emph{more conservative}.
-For instance, we will use a very fast search for candidate motifs, which returns only a limit number of instances.
-A more expensive search might yield more motifs, but never less.
-That is, for those motifs that we find, we can be sure that they would also be found by a more expensive method.
-Since pattern mining of this kind often results in too many potential patterns rather than too few, it seems efficient to start with the motifs that can be found with fast, simple methods, and only extend the search once these have been exhausted.
-
-Specifically, we generate candidate motifs by \emph{sampling} them from the data.
-This method, inspired by \cite{kashtan2004efficient} is based on the simple insight that when we sample a pattern from the data, we are more likely to sample a frequently occurring pattern than a infrequent one.
-In we sample $K$ times, and rank the results by relative frequency within the sample, we are likely toe end up with a good indication of the most frequent patterns in the data.
-For unlabeled graphs, as little as 500 samples can be sufficient to find the most frequent subgraph.
-Theoretically, this allows us to generate promising motifs with time complexity indepenedent of the size of the graph. Our sampling algorithm is described in Figure~\ref{algorithm:sampling}
+We will first assume that a target pattern $M$ is given for the data $G$ and that we have a set of \emph{instances} $\cal I'$ where $I \in \cal I'$ is a subset of $E_G$ which matches the pattern.
+Moverover, we require that all instances in $\cal I$ are mutually disjoint.
+Thus, we cannot simply return all matches of the pattern, we require some selection to ensure that none of them overlap.
+In most settings we do not have a given $M$ and $I$ and we will need to search for them.
+We describe a simple and fast search algorithm in Section~\ref{section:search}.
 
@@
 
+We will first describe our method under the assumption that a good null model $p^\text{null}$ is available, and that $- \log p^\text{null}(G)$ can be efficiently computed.
+In Section~\ref{section:null-model}, we explain which null model we use for our experiments, and how it is implemented.
+
+% TODO
+
+In the design of our method, we will constantly aim to find a trade-off between completeness and efficiency that allows the method to scale to very large graphs.
+Specifically, when we economize, we will only do so in a way that makes the hypothesis test described in Section~\ref{section:relevance-test} \emph{more conservative}.
+
 \subsection{Motif code}
+
+As described above, we can perform our relevance test with any compression method which exploits the pattern $M$, and its instances $I$ to store the graph efficiently.
+The better our method, the more motifs we will find.
+Note that there is no need for our code to be optimal in any sense.
+So long as we accept that we won't find all motifs that exist, we are free to trade off compression performance against efficiency of computation.
+
+We store the graph by encoding various aspects, one after the other.
+The information in all of these together is sufficient to reconstruct the graph.
+Note that everything is stored using prefix-free codes, so that we can simply sum the codelengths we get for each aspect to get the codelength for the complete graph.
+
+We also assume that we are given a code $L^\text{base}$ for generic knowledge graphs.
+
+We store, in order:
+
+\begin{description}
+ \item[the graph dimensions] We first store $|V_G|$, $|E_G|$ and $|R_G|$ using the generic code $L^{\N}(\cdot)$.
+ \item[the pattern] We store the pattern $M$ using $L^\text{base}$.
+We increment each of the pattern's existing link labels by 1, and use the label 0 to represent unlabeled links.
+Since $L^\text{base}$ does not store node labels ($V_M$ is always the set of the first $|V_M|$ natural numbers) we need to store separately to which nodes in $G$ the labeled nodes of $M$ refer.
+We store a sequence of integers ${g_i}$ where $g_i = 0$ indicates that the $i$-th node of $M$ is unlabeled, and $g_i \neq 0$ indicates that $g_i$ should map to node $g_i - 1$ in $G$.
+ \item[the template] This is the graph, minus all links occurring in  instances.
+Let $E_G'$ be $E_G$ minus any link occurring in any member of $\cal I$.
+We then store $(V_G, E_G')$ using $L^\text{base}$.
+ \item[the instances] In order to reconstruct the graph from the template, we need to know which nodes correspond to instances of the motif, and in what order.
+For each unlabeled node in M, we store one list of length $\cal I$ of the nodes in $G$ to which it maps for each instance.
+We store these lists using the DM code, since it is likely that a particular subset of the nodes of $G$ is frequent for a particular unlabeled node of a pattern.
+ \item[the edge labels] We now have enough information to reconstruct the graph structure.
+All that remains to fully reconstruct $G$ is to label the unlabeled link of the motif instances.
+As for the nodes, we create one list of length $\cal I$ for each unlabeled  link in the pattern which we store using the DM code.
+\end{description}
+
+We can note a few aspects of this code.
+Firstly, it achieves its compression from not having to store the edges of the motif instances.
+For each additional instance, it pays the cost of storing the instance, and any unlabeled nodes and links, but it saves $E_M$ edges.
+The precise trade-off depends on the choice of base code, and on the structure of the rest of the graph (some edges are cheaper to store than others).
+
+Nevertheless, we can make a few broad observations: patterns with many unlabeled nodes and links are more costly to store than fully labeled patterns.
+On the other hand, a pattern with many unlabeled nodes and links is more likely to have many instances, so that \emph{if} it results in a small amount compression per instance, the large number of instances may cause it to compress a lot.
+For the unlabeled nodes and links, we use a DM model for each specific node and link.
+This means that the more structure there is to the label given to a particular part of the pattern, the better we compress.
+For instance, suppose a particular node of the pattern is always labeled with either a node corresponding to YES or NO, in equal proportion, we will require only a bit per instance to record those labels.
+If all nodes in the graph are equally likely to appear in that position, we will pay approximately $\log |V_G|$ bits per instance.
+If only a single node appears in the position, we pay only a a constant number of bits to record the entire sequence of instances.
+\footnotemark
+
+\footnote{In fact, we could simplify the model by keeping the pattern entirely unlabeled, and trusting that the DM code will store everything efficiently.
+However, for our search algorithm, and for the interpretation of the motif as a graph pattern, we prefer to distinguish between labeled and unlabeled elements explicitly.}
+
+\subsection{Motif search}
+
+\label{section:search}
+
+In most use cases, we do not start with a given motif and a set of instances.
+We have only the graph $G$, and we wish to find any motifs that might be interesting.
+In order to do this efficiently, we generate candidate patterns by \emph{sampling} them from the data.
+This method, inspired by \cite{kashtan2004efficient} is based on the simple insight that when we sample a pattern from the data, we are more likely to sample a frequently occurring pattern than a infrequent one.
+In we sample $K$ times, and rank the results by relative frequency within the sample, we are likely to end up with a good indication of the most frequent patterns in the data.
+For unlabeled graphs, as little as 500 samples can be sufficient to find the most frequent subgraph.
+
+Theoretically, this allows us to generate promising motifs with time complexity independent of the size of the graph.
+Our algorithm samples a pattern in three steps.
+First, it samples a weakly connected subset of $n$ nodes, and extracts the subgraph induced by these nodes.
+Second, it converts this subgraph to a non-induced pattern by randomly removing labels from some of the nodes and links and randomly removing some links entirely (as detailed below).
+Finally, we sort the nodes of this pattern into a canonical ordering, using the Nauty algorithm \cite{mckay1981practical}.
+
+For a subgraph with $n$ nodes and $e$ links, we can generate $n^2e^3$ patterns.
+A uniform random choice among these would make patterns with many labels very unlikely.
+Instead, we sample $n' ~ U([0..
+n])$, $e' ~ U([0..
+e])$ and $e^*~ U([0..e'])$: the number of nodes to leave labeled, the number of nodes links to leave, and the number of links to leave labeled respectively.
+If the resulting pattern is disconnected, we reject the sample.
+
+We sample $K$ such patterns, recording a dictionary mapping canonical patterns $M$ to a list of observed instances $\I'_M$ of $M$ in $G$.
+Note that we must store the \emph{link} of the occurrence in $\I'_M$, since the same pattern may emerge from a single induced subgraph in multiple ways.
+
+Before we can can pass each pattern to the relevance test to see if it is a motif, we must first ensure that $\I'_M$ contains only link-disjoint instances.
+Assuming that all instances contribute equally to the compression, we would like to select the largest subset of $\I'_M$ consisting entirely of mutually disjoint sets.
+This is an instance of the Maximum Independent Set Problem, which is NP-Hard to solve optimally.
+Since we are aiming for scalability, we use a simple greedy algorithm to create our set $\I_M$ of disjoint instances: we include the first instance $I$ in $\I'_M$ in $\I_M$ and exclude any other instances that overlap with $I$.
+We then move to the next instance that has not already been included or excluded, and repeat the procedure.
+We continue until all instances have been included or excluded.
+
+We then pass each $M$ and $\I_M$ to the relevance test to see if they are motifs. For experiments reported here, we check only the 100 patterns with the largest number of instances.\footnote
+
+\footnotemark{In \cite{bloem2017large}, the number of edges between nodes inside the instance and outside it (the exdegree of the instance), was a strong predictor of whether the instances would benefit compression.
+Thus, the algorithm was strongly helped by pruning the list of instances using a ternary search.
+For the current code, the exdegree does not affect compression (since the instance nodes are not removed from the template graph).
+Therefore, we pass the entire instance list $\I_M$ to the relevance test wihtout pruning it.}
+
+Our sampling algorithm is described more precisely in Figure~\ref{figure:sampling}.
+
+\begin{figure}[bt]
+{
+\hrule
+\vspace{3mm}
+\raggedright
+Given: a graph $G$, \\
+\tab a minimum motif size $s_\text{min}$ and a maximum motif size $s_\text{max}$ \\
+\tab a target number of graphs sampled\\
+\tab\\
+Initialize $d$, a dictionary from motif $M$ to a list of instances $d[M] \in 2^{2^N}$ \\ 
+  \\
+Choose $s$ uniformly from $[s_\text{min} .. s_\text{max}]$ \\
+Choose a starting node $n$ uniformly from $G$ \\
+$N \from \{n\}$ \\
+\textbf{while} $|N| < s$:\\
+\tab Choose $n$ uniformly from $N$
+\tab Choose $n'$ uniformly from the neighbors of $n$
+\tab $N \from N \cup n'$ \\
+\tab After 500 cycles, sample a different starting node\\
+
+$M \from G[N]$, the induced subgraph of $N$ \\
+$M \from \text{canonical}(M)$ using the Nauty canonization algorithm, ignoring label \\
+\vspace{3mm}
+\hrule
+}
+\caption{The motif sampling algorithm.}
+\label{figure:sampling}
+\end{figure}
+
+The sampling algorithm produces full labeled, induced subgraphs.
+
+For instance, we will use a very fast search for candidate motifs, which returns only a limited number of instances.
+A more expensive search might yield more instances, but never fewer.
+That is, for those motifs that we find, we can be sure that they would also be found by a more expensive method.
+Since pattern mining of this kind often results in too many potential patterns rather than too few, it seems efficient to start with the motifs that can be found with fast, simple methods, and only extend the search once these have been exhausted.
 

### 2017-07-28 kgpsr/kgpsr.tex
..

@@
 \usepackage{amssymb}
+
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 Knowledge graph node embeddings are vector representations of the nodes in a knowledge graph.
-Effective embeddings are an important first step in many pipelines for knowledge graphs visualisation, completion, classification and other tasks.
+For a knowledge graph with nodes $N$, an embedding is a function $v: N \to \R^k$ such that $v$ organizes the nodes meaningfully in the spacer $\R^k$ according to some criterion.
+
+Effective embeddings are an important first step in many pipelines for knowledge graph visualisation, completion, classification and other tasks.
 Many embedding algorithms model the relations in the knowledge graph as maps in the embedding space: either translations, or more complex transformations.
-If for instance, we have the triple $(\fw{John}, \fw{parentof}, \fw{Mary})$, then we learn embeddings $v_\fw{John}, v_\fw{Mary} \in \R^k$  (for some chosen embedding dimension $k$) and a map $f_\fw{parentof} : \R^k \to \R^k$ for the relation \texttt{parentof} such that $f_\fw{parentof}(v_\fw{John})$ should be as close to $v_\fw{Mary}$ as possible.
+If for instance, we have the triple $(\fw{Mary}, \fw{childof}, \fw{John})$, then we learn embeddings $v_\fw{Mary}, v_\fw{John} \in \R^k$  (for some chosen embedding dimension $k$) and a map $f_\fw{childof} : \R^k \to \R^k$ for the relation \texttt{childof} such that $f_\fw{childof}(v_\fw{John})$ should be as close to $v_\fw{Mary}$ as possible (for some chosen distance metric).
+
 This scheme works well, but has trouble representing multiple relations.
-If John has more than one child, what should $f_\fw{parentof}(v_\fw{John})$ be?
+If John has more than one child, what should $f_\fw{childof}(v_\fw{John})$ be?
 Many approaches have been introduced, often requiring an advanced mathematical structure for the embedding space, such as complex vectors \cite{trouillon2016complex}, or holographic projections \cite{nickel2016holographic}.
-	We introduce a variational approach.
-We model each knowledge graph node not as a point in the embedding space, but as a spherical Gaussian of fixed variance.
+
+Our approach is based on a simple intuition.
+For each source node, we place a spherical Gaussian distribution over the embedding.
 Transforming this Gaussian by the relational map should then result in a new Gaussian which maximizes the likelihood of the nodes connected to the original by the given relation.\footnotemark
-	\footnotetext{This approach is inspired by the Coherent Point Drift algorithm \cite{}, which uses a similar approach for the problem of point set registration.}
-	The variational approach has several advantages:
+\footnotetext{This approach is inspired by the Coherent Point Drift algorithm \cite{}, which uses a similar approach for the problem of point set registration.} In other words, transforming a spherical Gaussian centered on $v_\fw{John}$ by $f_\fw{childof}$ should result in a distribution that maximizes the probability of John's children.
+
+Intuitively, this is a simple latent variable problem, that can be solved by an alternating optimization algorithm in the vein of expectation-maximization: given a choice for the node embeddings, we can find the optimal functions to represent the relations, and given a choice for the relation functions, we compute the optimal embeddings.
+Alternating these two optimizations is likely to converge to a good solution.
+To formalize this intuition in an elegant manner, take a Bayesian approach, and find the optimum using variational inference.
+Thge preicsae model is detailed in Section~\ref{section:model}.
+
+Our approach has several advantages over the alternatives:
 \begin{itemize}
-		\item The embedding space is a simple Euclidean space, making no special demands on downstream architectures.
+	\item The nodes are embedded in a simple Euclidean space, making no special demands on downstream architectures.
 	\item The multiple-relation issue is solved in an intuitive manner: mapping a parent node to its children creates a probability distribution that covers the children. 
+	\item The method is agnostic to the family of maps used to represent relations.
+We choose affine maps as a good tradeoff between ease of optimization, rigidity to protect against overfitting, and expressiveness.
+The method easily generalizes to other function classes.
 	\item For linear maps, we can decompose the loss function into factors which can be solved analytically.
@@
 	\item Negative sampling is not required. The basic framework constrains the problem sufficiently to lead to models that generalize well.
+	\item The variational framework protects against convergence to solutions with singularities, and allows us to incorporate hyperparameters like the dimension of the embedding space into the optimization process.
 \end{itemize}
@@
 \section{Model}
+\label{section:model}
+
+Define a knowledge graph $G$ as a triple $G = (V_G, R_G, E_G)$ with $V_G$ representing a set of labeled nodes, $R_G$ representing a set of available relations, and $E_G \subset V_G \times R_G \times V_G$ the set of \emph{triples}---edges labeled with relations.
+Since we are not interested in modeling the internal structure of labels, we will assume that $V$ is always the set of the first $|V|$ natural numbers, and likewise for $R$.
+
+To simplify notation, we will assume that the graph being modeled is clear from context, and drop the subscript $G$.
+
+Let $x_i \in \R^k$ be the embedding chosen for node $i \in V$.
+Let $f_r$ be a parametrized function from $\cal F$, representing relation $r \in R$.
+For our method, we will let $\cal F$ be the set of affine functions, each parametrized by a transformation matrix $A_r \in \R^{k\times k}$, and a translation vector $t_r\in \R^k$.
+Thus our model contains $|N|$ vectors in $\R^k$, and $|R|$ functions each defined by a matrix $A_r$ and vector $t_r$.
+
+To formalize the intuitions expressed in the introduction, we first express the constraints in a graphical model.
+Let $E^r_\text{subject} = \{s \mid \exists o, (s, r, o) \in E\}$ and $E^{r, o}_\text{object} = \{ s \mid (s, r, o) \in E\}$.
+Figure~\ref{figure:graphical-model} shows the relations between the different random variables.
+Interestingly, this model contains \emph{no observed variables}.
+The data is used purely to inform the topology of the graphical model.\footnotemark The model also gives us the priors: the prior probability for the embeddings, for which we will use the standard normal distribution: $\mu = 0$ and $\Lambda = \bf{1}$ and the prior $W_t$ on the
+
+\footnote{Modeling the data as observations would lead to tensor-factorization approaches like ReSCAL \cite{}.}
+
+% TODO Read up on Bayes and conjugate priors. GEt a better sense of what priors I should use.
+
+\[
+\argmin_{\{x_i\}, \{f_r\}, \sigma} \prod_{(s,r,o) \in E_G} f_r(N_{x_s,\sigma})(x_o)
+\]
 

### 2017-08-10 kgmotifs/kgmotifs.tex
...

@@
 \begin{abstract}
-We introduce a method to detect \emph{network motifs} in knowledge graphs.
+\noindent We introduce a method to detect \emph{network motifs} in knowledge graphs.
 Network motifs are useful patterns or meaningful subunits of the graph that recur frequently.
@@
 
-\noindent The Linked, Open Data cloud contains a wealth of knowledge graphs, and is growing quickly.
-At the time of writing the average knowledge graph contains over 65 000 edges, with many datasets containing more than a million.\footnotemark~For such large graphs, it can be difficult to see the forest for the trees: how is the graph structured at the lowest level?
+\noindent The Linked Open Data cloud contains a wealth of knowledge graphs, and is growing quickly.
+At the time of writing the average knowledge graph contains over 65\,000 edges, with many datasets containing more than a million.\footnotemark~For such large graphs, it can be difficult to see the forest for the trees: how is the graph structured at the lowest level?
 What kind of things can I ask of what types of objects?
@@
 
-In \cite{bloem2017large}, an alternative method is presented that uses compression as a heuristic for motif value: the better a motif compresses the data, the more likely it is to be meaningful.
+In \cite{bloem2017large}, an alternative method is presented that uses \emph{compression} as a heuristic for motif value: the better a motif compresses the data, the more likely it is to be meaningful.
 It is shown that this principle can be implemented in a similar sort of hypothesis test for a null model, allowing a comparable workflow to classical motif analysis.
@@
 
+% TODO: General purpose codes: natural numbers, integers, sequences.
+
 \section{Method}
@@
 A \emph{knowledge graph} $G$, hereafter known simply as a \emph{graph} is a, is tuple $G = (V_G, E_G)$ with finite sets $V_G \subset V$ and $E_G \subset V \times R \times V$.
-We define the triple list of $G$ as $G^T = \{(s, p, o) \in \N^3 : \left (v(s), r(p), v(o) \right ) \in E_G\}$.
-In this paper we will \emph{identify} the graph with the triple list.
-That is when we speak of encoding the graph, we mean encoding the triple list, and when we speak of a probability distribution over graphs, we mean a probability distribution over triple lists.
-Let $R_G \cub \N$ be the set of relations used in the $G$.
+We define the \emph{tripleset} of $G$ as $G^T = \{(s, p, o) \in \N^3 : \left (v(s), r(p), v(o) \right ) \in E_G\}$.
+In this paper we will \emph{identify} the graph with the tripleset.
+That is, when we speak of encoding the graph, we mean encoding the triple list, and when we speak of a probability distribution over graphs, we mean a probability distribution over triplesets.
+Let $R_G \subset \N$ be the set of relations used in $G$.
 
 This means that we are purposely \emph{not} modelling certain aspects of the data.
-There might, for instance be a similarity between two node labels $v(a)$ and $v(b)$ (such as a shared IRI prefix).
+There might, for instance be a similarity between two node labels $a$ and $b$ (such as a shared IRI prefix).
 We are not interested in exploiting such structure, so we eliminate it from our notation.
-Additionally, we introduce some bias with our choice of $e$ and $r$.
-If we were to use a simple code to store the integers in the triple-list we will end up using fewer bits for those nodes and triples with labels who were assigned short integers.
-Equivalently, when sampling from a simple distribution over the integers, small numbers will be more likely to be generated.
+We also introduce some bias with our choice of $e$ and $r$.
+Entities and relations that are assigned small integers may end up receiving short codes or high probability mass, if a simple code or disgribution is used.
 In practice, however, we will always \emph{learn} a distribution over the integers from the data, so that $e$ and $r$ are purely a means to simplify the notation.
@@
 In other words, the motif is a \emph{basic graph pattern}, and the edges and nodes marked with "?" are variables.
-To represent this, we add a single variable symbol $?$.
-Thus, a motif $M$ is defined by the triple set $M \subset (?
-\cup \N) ^3$, with integer values in triples interpreted as they are in complete knowledge graphs.
+To represent this, we require additional symbols: one for variable relations, and one for each node in the motif marked with a variable.
+We will define a pattern by an \emph{integer} triple list: $M \subset \Z^3$.
+Natural numbers mean the same as they do in a graph, and negative integers represent variable nodes.
+Variable links marked with the same negative integer, must map to the same relation in order for the pattern to match, but variable links labeled with different negative integers \emph{may} map to the same relation.
+For the current paper, we will only search for patterns where each variable link contains a unique label.
 
 We will first assume that a target pattern $M$ is given for the data $G$ and that we have a set of \emph{instances} $\cal I'$ where $I \in \cal I'$ is a subset of $E_G$ which matches the pattern.
-Moverover, we require that all instances in $\cal I$ are mutually disjoint.
+Moreover, we require that all instances in $\cal I$ are mutually disjoint.
 Thus, we cannot simply return all matches of the pattern, we require some selection to ensure that none of them overlap.
@@
 We describe a simple and fast search algorithm in Section~\ref{section:search}.
+
+Note that unlike the graph, the nodes of $M$ \emph{are} explicitly labeled. The positive-valued nodes of $M$ are not simply the first $|V_M|$ integers, but refer specifically to the nodes of $G$.
 
@@
 In Section~\ref{section:null-model}, we explain which null model we use for our experiments, and how it is implemented.
-
-% TODO
 
@@
  \item[the graph dimensions] We first store $|V_G|$, $|E_G|$ and $|R_G|$ using the generic code $L^{\N}(\cdot)$.
- \item[the pattern] We store the pattern $M$ using $L^\text{base}$.
-We increment each of the pattern's existing link labels by 1, and use the label 0 to represent unlabeled links.
-Since $L^\text{base}$ does not store node labels ($V_M$ is always the set of the first $|V_M|$ natural numbers) we need to store separately to which nodes in $G$ the labeled nodes of $M$ refer.
-We store a sequence of integers ${g_i}$ where $g_i = 0$ indicates that the $i$-th node of $M$ is unlabeled, and $g_i \neq 0$ indicates that $g_i$ should map to node $g_i - 1$ in $G$.
+ \item[the pattern] Let $M'$ be the unlabeled graph representation of $M$: that is, $M'$ is a plain graph (not a pattern), $V_{M'}$ contains only the first $V_{M'}$ natural numbers and similarly for $R_{M'}$.
+Otherwise, $M$ is isomorphic to $M'$: the node and edge labels of $M$ can be mapped to those of $M'$ so that $M$ only contains an edge $E$ if and only if $M'$ contains the edge resulting from mapping $E$'s symbols.
  \item[the template] This is the graph, minus all links occurring in  instances.
@@
  \item[the instances] In order to reconstruct the graph from the template, we need to know which nodes correspond to instances of the motif, and in what order.
-For each unlabeled node in M, we store one list of length $\cal I$ of the nodes in $G$ to which it maps for each instance.
-We store these lists using the DM code, since it is likely that a particular subset of the nodes of $G$ is frequent for a particular unlabeled node of a pattern.
+For each node in $M'$, we create a list of length $\cal I$ of the nodes in $G$ to which it maps for each instance.
+For the non-variable list, this will be a a list simply repeating one node, and for the variable nodes it will be a list containing some subset of the ndoes in $G$.
+ We store each list using the DM code.
+It may seem wasteful to store a list repeating a single element, but under the DM code such lists will be stored efficiently.
+The more varied the list, the more bits it will require to store.
  \item[the edge labels] We now have enough information to reconstruct the graph structure.
-All that remains to fully reconstruct $G$ is to label the unlabeled link of the motif instances.
-As for the nodes, we create one list of length $\cal I$ for each unlabeled  link in the pattern which we store using the DM code.
+All that remains to fully reconstruct $G$ is to label the links of the motif instances.
+As for the nodes, we create one list of length $\cal I$ for each type of link in $M'$, which we store using the DM code.
 \end{description}
 
-We can note a few aspects of this code.
-Firstly, it achieves its compression from not having to store the edges of the motif instances.
-For each additional instance, it pays the cost of storing the instance, and any unlabeled nodes and links, but it saves $E_M$ edges.
+The precise computation of the codelength is given in Algorithm~\ref{algorithm:codelength}.
+
+We note a few aspects of this code.
+Firstly, it achieves its compression by not storing the edges of the motif instances.
+For each additional instance, it pays the cost of storing the instance, and any variable nodes and links, but it saves $|E_M|$ edges in storing the template.
 The precise trade-off depends on the choice of base code, and on the structure of the rest of the graph (some edges are cheaper to store than others).
 
-Nevertheless, we can make a few broad observations: patterns with many unlabeled nodes and links are more costly to store than fully labeled patterns.
-On the other hand, a pattern with many unlabeled nodes and links is more likely to have many instances, so that \emph{if} it results in a small amount compression per instance, the large number of instances may cause it to compress a lot.
-For the unlabeled nodes and links, we use a DM model for each specific node and link.
+Nevertheless, we can make a few broad observations: patterns with many variable nodes and links are more costly to store than fully labeled patterns.
+On the other hand, a pattern with many variable nodes and links is more likely to have many instances, so that \emph{if} it results in a small amount compression per instance, the large number of instances may cause it to compress a lot.
+For the variable nodes and links, we use a DM model for each specific node and link.
 This means that the more structure there is to the label given to a particular part of the pattern, the better we compress.
 For instance, suppose a particular node of the pattern is always labeled with either a node corresponding to YES or NO, in equal proportion, we will require only a bit per instance to record those labels.
-If all nodes in the graph are equally likely to appear in that position, we will pay approximately $\log |V_G|$ bits per instance.
+If, however, all nodes in the graph are equally likely to appear in the position of the variable, we will pay approximately $\log |V_G|$ bits per instance.
 If only a single node appears in the position, we pay only a a constant number of bits to record the entire sequence of instances.
-\footnotemark
-
-\footnote{In fact, we could simplify the model by keeping the pattern entirely unlabeled, and trusting that the DM code will store everything efficiently.
-However, for our search algorithm, and for the interpretation of the motif as a graph pattern, we prefer to distinguish between labeled and unlabeled elements explicitly.}
+Thus, to our code, non-variable nodes are just a highly regular type of variable node.
+The distinction is only meaningful in the search algorithm.
 
@@
 This method, inspired by \cite{kashtan2004efficient} is based on the simple insight that when we sample a pattern from the data, we are more likely to sample a frequently occurring pattern than a infrequent one.
-In we sample $K$ times, and rank the results by relative frequency within the sample, we are likely to end up with a good indication of the most frequent patterns in the data.
-For unlabeled graphs, as little as 500 samples can be sufficient to find the most frequent subgraph.
+Ik we sample $K$ times, and rank the results by relative frequency within the sample, we are likely to end up with a good indication of the most frequent patterns in the data.
+For unlabeled graphs, as few as 500 samples can be sufficient to find the most frequent subgraph.
 
@@
 Second, it converts this subgraph to a non-induced pattern by randomly removing labels from some of the nodes and links and randomly removing some links entirely (as detailed below).
-Finally, we sort the nodes of this pattern into a canonical ordering, using the Nauty algorithm \cite{mckay1981practical}.
+Finally, we sort the nodes of this pattern into a canonical ordering, using the Nauty algorithm \cite{mckay1981practical}, this allows us to record the instances for each subgraph without counting isomorphic subgraphs separately.
 
-For a subgraph with $n$ nodes and $e$ links, we can generate $n^2e^3$ patterns.
+For an induced subgraph with $n$ nodes and $e$ links, we can generate $n^2e^3$ patterns.
 A uniform random choice among these would make patterns with many labels very unlikely.
@@
 n])$, $e' ~ U([0..
-e])$ and $e^*~ U([0..e'])$: the number of nodes to leave labeled, the number of nodes links to leave, and the number of links to leave labeled respectively.
+e])$ and $e^*~ U([0..e'])$, with $n'$ the number of nodes to leave labeled, $e'$ the number of links to leave, and $e^*$ the number of links to leave labeled, respectively.
 If the resulting pattern is disconnected, we reject the sample.
@@
 For the current code, the exdegree does not affect compression (since the instance nodes are not removed from the template graph).
-Therefore, we pass the entire instance list $\I_M$ to the relevance test wihtout pruning it.}
+Therefore, we pass the entire instance list $\I_M$ to the relevance test without pruning it.}
 
@@
 Initialize $d$, a dictionary from motif $M$ to a list of instances $d[M] \in 2^{2^N}$ \\ 
-  \\
+\tab\\
 Choose $s$ uniformly from $[s_\text{min} .. s_\text{max}]$ \\

### 2017-08-12 kgmotifs/kgmotifs.tex
...

@@
 \usepackage{url}
+\usepackage{float}
 
 \newcommand{\N}{{\mathbb N}}
+\newcommand{\R}{{\mathbb R}}
+\newcommand{\Z}{{\mathbb Z}}
+
 \newcommand{\G}{{\cal G}}
@@
 
-\title{Detecting Network Motifs in Knowledge Graphs using Compression}
+\floatstyle{ruled}
+\newfloat{pseudo}{h}{lop}
+\floatname{pseudo}{Algorithm}
+
+\title{Detecting Motifs in Knowledge Graphs using Compression}
 
@@
 At the time of writing the average knowledge graph contains over 65\,000 edges, with many datasets containing more than a million.\footnotemark~For such large graphs, it can be difficult to see the forest for the trees: how is the graph structured at the lowest level?
-What kind of things can I ask of what types of objects?
+What kind of things can I ask of what types of entities?
 What are small, recurring patterns that might represent a novel insight into the data?
@@
 
-In \cite{bloem2017large}, an alternative method is presented that uses \emph{compression} as a heuristic for motif value: the better a motif compresses the data, the more likely it is to be meaningful.
+In \cite{bloem2017large}, an alternative method is presented that uses \emph{compression} as a heuristic for motif relevance: the better a motif compresses the data, the more likely it is to be meaningful.
 It is shown that this principle can be implemented in a similar sort of hypothesis test for a null model, allowing a comparable workflow to classical motif analysis.
@@
 
-\subsection{Related Work}
+\paragraph{Related Work}
 
-\subsection{Preliminaries}
+\paragraph{Preliminaries}
 
@@
 To comfortably define probabilities and codes on graphs, we will slightly deviate from conventional definitions.
-Let $V$ be a countably infinite set containing all possible entities, and let $R$ be a countably infinite set containing all possible relations between them.
-Let $v: V \to \N$ and $r: R \to \N$ be arbitrary mappings of these sets to the natural numbers.\footnotemark
+Specifically, we will analyze the \emph{structure} of knowledge graphs only, ignoring any meaning they have outside the graph.
+For instance, we will distinguish between links with different labels, but we will ignore any internal structure of the label itself (such as shared IRI prefixes).
+Under this assumption we can assume that the node and link labels of our knowledge graphs are simply natural numbers.
+This leads to the following definition.
 
-\footnotetext{For instance, let $V$ and $E$ be the set of finite strings, and let $v$ and $e$ represent length-lexicographical order.}
+A \emph{knowledge graph} $G$, hereafter known simply as a \emph{graph} is a, is tuple $G = (v_G, r_G, E_G)$.
+$v_G \in \N$ is the number of nodes in the graph, and $r_G$ is the number of relations.
+We will define the nodeset as $V_G = \{0, \ldots, v_G-1\}$ and the relation-set as $R_G = \{0, \ldots, r_G\}$.
+The \emph{tripleset} $E_G \subset V_G \times R \times V_G$ defines the links of the graph and their labels.
+\footnotemark
 
-% TODO: Clarify that we aren't storing node labels
+\footnotetext{Commonly, the elements of the nodeset of a graph are defined to be simple atoms from some larger unspecified universe.
+However, since we are discussing sampling graphs, we need to specify where these atoms come from.
+This definition specifies that graphs are essentially unlabeled, but nodes do have a definite ordering (i.e.
+two graphs can be isomorphic without being the same graph).}
 
-A \emph{knowledge graph} $G$, hereafter known simply as a \emph{graph} is a, is tuple $G = (V_G, E_G)$ with finite sets $V_G \subset V$ and $E_G \subset V \times R \times V$.
-We define the \emph{tripleset} of $G$ as $G^T = \{(s, p, o) \in \N^3 : \left (v(s), r(p), v(o) \right ) \in E_G\}$.
-In this paper we will \emph{identify} the graph with the tripleset.
-That is, when we speak of encoding the graph, we mean encoding the triple list, and when we speak of a probability distribution over graphs, we mean a probability distribution over triplesets.
-Let $R_G \subset \N$ be the set of relations used in $G$.
+As shown in Figure~\ref{figure:diagram}, the patterns that we aim to find are partially labeled: some nodes and edges are labeled with negative integers.
+In other words, the motif is a \emph{basic graph pattern}, and the edges and nodes marked with negative integers are variables.
+We will define a pattern in the same way we define a graph, but we will extend its nodeset en relationset.
+A pattern $M$ for graph $G$ is a tuple $(v'_M, r'_M, G, E_M)$.
+$v'_M$ and $r'_M$ indicate the number of variable nodes and links respectively.
+$M$'s nodeset is defined as $V_M = \{-v'_M, \ldots, v_G-1\}$ and the relationset as $R_M = \{-r'_M, \ldots, r_G-1\}$, with $E_M \subset V_M \times R_M \times V_M$, as for graphs.
+In short, nodes can be labeled with nonnegative integers referring to $G$'s nodes or with negative integers representing a variable node, and similar for relations.
 
-This means that we are purposely \emph{not} modelling certain aspects of the data.
-There might, for instance be a similarity between two node labels $a$ and $b$ (such as a shared IRI prefix).
-We are not interested in exploiting such structure, so we eliminate it from our notation.
-We also introduce some bias with our choice of $e$ and $r$.
-Entities and relations that are assigned small integers may end up receiving short codes or high probability mass, if a simple code or disgribution is used.
-In practice, however, we will always \emph{learn} a distribution over the integers from the data, so that $e$ and $r$ are purely a means to simplify the notation.
-Which functions we choose will not affect our model.
+An \emph{instance} for $M$ in $G$ is a pair of sequences of integers: $I = (I^n, I^r)$.
+$I^n$ is a sequence of distinct integers of length $v'_M$.
+$I^r$ is a sequence of non-distinct integers of length $r'_M$.
+For each link $(s, p, o) \in E_M$ containing a negative $s$, $p$ or $o$, there is a corresponding link in $E_G$ with a negative $s$ replaced by $I^n_{-s}$, a negative $o$ replaced by $I^n_{-o}$, and a negative $p$ replaced by $I^r_{-p}$.
+Put simply: for a pattern to match, variable links marked with the same negative integer, must map to the same relation in order for the pattern to match, but variable links labeled with different negative integers \emph{may} map to the same relation.
+Variable nodes are always labeled distinctly and may never map to the same node in $G$.
+An instance describes a subgraph of $G$ that \emph{matches} the pattern $M$.
 
-As shown in Figure~\ref{figure:diagram}, the motifs that we aim to find are partially labeled: some nodes and edges are labeled with the marker "?" indicating that when we fit the motif to the graph, those edges can contain any label.
-In other words, the motif is a \emph{basic graph pattern}, and the edges and nodes marked with "?" are variables.
-To represent this, we require additional symbols: one for variable relations, and one for each node in the motif marked with a variable.
-We will define a pattern by an \emph{integer} triple list: $M \subset \Z^3$.
-Natural numbers mean the same as they do in a graph, and negative integers represent variable nodes.
-Variable links marked with the same negative integer, must map to the same relation in order for the pattern to match, but variable links labeled with different negative integers \emph{may} map to the same relation.
-For the current paper, we will only search for patterns where each variable link contains a unique label.
-
-We will first assume that a target pattern $M$ is given for the data $G$ and that we have a set of \emph{instances} $\cal I'$ where $I \in \cal I'$ is a subset of $E_G$ which matches the pattern.
-Moreover, we require that all instances in $\cal I$ are mutually disjoint.
-Thus, we cannot simply return all matches of the pattern, we require some selection to ensure that none of them overlap.
-In most settings we do not have a given $M$ and $I$ and we will need to search for them.
+We will first assume that a target pattern $M$ is given for the data $G$ and that we have a set of instances $\I$.
+Moreover, we require that all instances in $\I$ are mutually disjoint: no two subgraphs defined by member of $\I$ may share an edge, but nodes may be shared.
+In most settings we do not actually have a given $M$ and $\I$ and we will need to search for them.
 We describe a simple and fast search algorithm in Section~\ref{section:search}.
-
-Note that unlike the graph, the nodes of $M$ \emph{are} explicitly labeled. The positive-valued nodes of $M$ are not simply the first $|V_M|$ integers, but refer specifically to the nodes of $G$.
 
@@
 \begin{description}
- \item[the graph dimensions] We first store $|V_G|$, $|E_G|$ and $|R_G|$ using the generic code $L^{\N}(\cdot)$.
+ \item[the graph dimensions] We first store $|V_G|$, $|E_G|$ and $|R_G|$ using the generic code $L^{\N}(\cdot)$. We also store the size of $\I$.
  \item[the pattern] Let $M'$ be the unlabeled graph representation of $M$: that is, $M'$ is a plain graph (not a pattern), $V_{M'}$ contains only the first $V_{M'}$ natural numbers and similarly for $R_{M'}$.
@@
 Let $E_G'$ be $E_G$ minus any link occurring in any member of $\cal I$.
-We then store $(V_G, E_G')$ using $L^\text{base}$.
+We then store $(V_G, E_G')$ using $L^\text{base}(\cdot)$.
  \item[the instances] In order to reconstruct the graph from the template, we need to know which nodes correspond to instances of the motif, and in what order.
-For each node in $M'$, we create a list of length $\cal I$ of the nodes in $G$ to which it maps for each instance.
-For the non-variable list, this will be a a list simply repeating one node, and for the variable nodes it will be a list containing some subset of the ndoes in $G$.
+For each node in $M'$, we create a list of length $\cal |I|$ of the nodes in $G$ to which it maps for each instance.
+For the non-variable nodes of the pattern, this will be a list simply repeating one node node of $G$.
+For the variable nodes, it will be a list containing some subset of the nodes in $G$.
  We store each list using the DM code.
-It may seem wasteful to store a list repeating a single element, but under the DM code such lists will be stored efficiently.
-The more varied the list, the more bits it will require to store.
+It may seem wasteful to store a list repeating a single element (instead of storing a labeled version of the pattern), but under the DM code such lists will be stored efficiently, using only a small, constant number of bits.
+The more varied the list, the more bits will be required to store it.
  \item[the edge labels] We now have enough information to reconstruct the graph structure.
@@
 
-The precise computation of the codelength is given in Algorithm~\ref{algorithm:codelength}.
+The precise computation of the codelength is given in Algorithm~\ref{algorithm:motif-code}.
 
 We note a few aspects of this code.
-Firstly, it achieves its compression by not storing the edges of the motif instances.
-For each additional instance, it pays the cost of storing the instance, and any variable nodes and links, but it saves $|E_M|$ edges in storing the template.
-The precise trade-off depends on the choice of base code, and on the structure of the rest of the graph (some edges are cheaper to store than others).
+Firstly, it achieves its compression by reducing the number of links it has to store.
+For each additional instance, it pays the cost of storing the instance, and any variable nodes and links, but it saves $|E_M|$ links in storing the template.
+The precise trade-off depends on the choice of base code, and on the structure of the rest of the graph (some edges are cheaper to store than others, for some codes).
 
 Nevertheless, we can make a few broad observations: patterns with many variable nodes and links are more costly to store than fully labeled patterns.
-On the other hand, a pattern with many variable nodes and links is more likely to have many instances, so that \emph{if} it results in a small amount compression per instance, the large number of instances may cause it to compress a lot.
+On the other hand, a pattern with many variable nodes and links is more likely to have many instances, so that \emph{if} it results in a small amount of compression per instance, the large number of instances may cause it to compress a lot.
 For the variable nodes and links, we use a DM model for each specific node and link.
@@
 
+\begin{pseudo}[tb]
+\caption{The motif code $L^\text{motif}(G ; M, {\cal I}, L^\text{base})$. Note that the nodes and relations of the graph are integers.}
+\label{algorithm:motif-code}
+{ 
+\textbf{function} $\text{codelength}(G; M, {\cal I}, L^\text{base})$:\\
+\tab\tab a graph $G$, a pattern $M$\\ 
+\tab\tab a list $\cal I$ of instances of $M$ in $G$, a code $L^\text{base}$ on knowledge graphs.\\
+\\
+\tab $b_\text{dim} \leftarrow L^\N(|V_G|) + L^\N(|R_G|) + L^\N(|E_G|) + L^\N(|\I|)$ \\
+
+\tab \emph{// Turn the pattern into a normal knowledge graph}\\
+\tab $E_{M'} = \left\{(\text{i}(s, V_M),  \text{i}(p, R_M), \text{i}(o, V_M)) \mid (s, p, o) \in E_M\right\}$\\
+\tab $M' = ([0 ..  |V_M|-1], [0 .. |R_M|-1], E_{M'})$\\
+\tab $b_\text{pattern} \leftarrow L^\text{base}(M')$ \\
+
+\tab $E'_G = E_G - \cup_{\I \in {\cal I}} I$ \hfill \emph{// Store the graph, with instances removed}\\
+\tab $b_\text{template} = L_\text{base}((V_G, R_G, E'_G))$\\
+
+\tab $b_\text{labels} \leftarrow 0$ \hfill \emph{// Store the node labels} \\
+\tab \textbf{for} $n$ in $V_M$\\
+\tab\tab $\I_n = \langle v(n, I_1), v(n, I_2), ... \rangle$\\
+\tab\tab\tab with $v(n, I)$ the label (from $V_G$) of node $n$ in instance $I$\\
+\tab\tab $b_\text{labels} \leftarrow b_\text{labels} + L^{DM}(\I_n)$\\
+
+\tab \textbf{for} $e$ in $E_M$ \hfill \emph{// Store the edge labels}\\
+\tab\tab $\I_e = \langle r(e, I_1), r(n, I_2), ... \rangle$\\
+\tab\tab\tab with $r(e, I)$ the relation (from $R_G$) of edge $e$ in instance $I$\\
+\tab\tab $b_\text{labels} \leftarrow b_\text{labels} + L^{DM}(\I_e)$\\
+
+\tab \textbf{return} $b_\text{dim} + b_\text{pattern} + b_\text{template} + b_\text{labels}$\\
+}
+
+\textbf{function} $\text{i}(s, S)$:\\
+\tab \textbf{return} index of $s$ in $\text{sorted}(S)$\\
+
+\end{pseudo}
+
 \subsection{Motif search}
@@
 
-In most use cases, we do not start with a given motif and a set of instances.
-We have only the graph $G$, and we wish to find any motifs that might be interesting.
+Usually, we do not start with a given pattern $M$ and a set of instances $\I$.
+We have only the graph $G$, and we wish to find \emph{any} pattern that might be interesting.
 In order to do this efficiently, we generate candidate patterns by \emph{sampling} them from the data.
 This method, inspired by \cite{kashtan2004efficient} is based on the simple insight that when we sample a pattern from the data, we are more likely to sample a frequently occurring pattern than a infrequent one.
-Ik we sample $K$ times, and rank the results by relative frequency within the sample, we are likely to end up with a good indication of the most frequent patterns in the data.
+If we sample $K$ times, and rank the results by relative frequency within the sample, we are likely to end up with a good indication of the most frequent patterns in the data.
 For unlabeled graphs, as few as 500 samples can be sufficient to find the most frequent subgraph.
 
-Theoretically, this allows us to generate promising motifs with time complexity independent of the size of the graph.
+Theoretically, this allows us to generate promising candidates with time complexity independent of the size of the graph.
 Our algorithm samples a pattern in three steps.
-First, it samples a weakly connected subset of $n$ nodes, and extracts the subgraph induced by these nodes.
-Second, it converts this subgraph to a non-induced pattern by randomly removing labels from some of the nodes and links and randomly removing some links entirely (as detailed below).
-Finally, we sort the nodes of this pattern into a canonical ordering, using the Nauty algorithm \cite{mckay1981practical}, this allows us to record the instances for each subgraph without counting isomorphic subgraphs separately.
+First, it samples a weakly connected subset of $n$ nodes, and extracts the \emph{induced subgraph} for these nodes.
+Second, it converts this induced subgraph to a non-induced pattern by randomly removing labels from some of the nodes and links and randomly removing some links entirely (as detailed below).
+Finally, we sort the nodes of this pattern into a canonical ordering, using the Nauty algorithm \cite{mckay1981practical} (so that we record isomorphic samples as the same pattern).
 
-For an induced subgraph with $n$ nodes and $e$ links, we can generate $n^2e^3$ patterns.
+For an induced subgraph with $n$ nodes and $e$ links, we can generate up to $n^2e^3$ different patterns: each node can be left labeled or made a variable node, each link can be removed, left unlabeled or be made variable.
 A uniform random choice among these would make patterns with many labels very unlikely.
@@
 
-We sample $K$ such patterns, recording a dictionary mapping canonical patterns $M$ to a list of observed instances $\I'_M$ of $M$ in $G$.
-Note that we must store the \emph{link} of the occurrence in $\I'_M$, since the same pattern may emerge from a single induced subgraph in multiple ways.
+We sample $K$ such patterns, recording a dictionary which maps canonical patterns $M$ to a list of observed instances $\I'_M$ of $M$ in $G$.
+Note that we must store the \emph{links} of the occurrence in $\I'_M$, since the same pattern may emerge from a single induced subgraph in multiple ways.
 
@@
 
-Our sampling algorithm is described more precisely in Figure~\ref{figure:sampling}.
+Our sampling algorithm is described more precisely in Algorithm~\ref{algorithm:sampling}.
 
-\begin{figure}[bt]
-{
-\hrule
-\vspace{3mm}
-\raggedright
-Given: a graph $G$, \\
-\tab a minimum motif size $s_\text{min}$ and a maximum motif size $s_\text{max}$ \\
-\tab a target number of graphs sampled\\
-\tab\\
-Initialize $d$, a dictionary from motif $M$ to a list of instances $d[M] \in 2^{2^N}$ \\ 
-\tab\\
-Choose $s$ uniformly from $[s_\text{min} .. s_\text{max}]$ \\
-Choose a starting node $n$ uniformly from $G$ \\
-$N \from \{n\}$ \\
-\textbf{while} $|N| < s$:\\
-\tab Choose $n$ uniformly from $N$
-\tab Choose $n'$ uniformly from the neighbors of $n$
-\tab $N \from N \cup n'$ \\
-\tab After 500 cycles, sample a different starting node\\
+\begin{pseudo}[tb]
+\caption{The motif sampling algorithm.}
 
-$M \from G[N]$, the induced subgraph of $N$ \\
-$M \from \text{canonical}(M)$ using the Nauty canonization algorithm, ignoring label \\
-\vspace{3mm}
-\hrule
-}
-\caption{The motif sampling algorithm.}
-\label{figure:sampling}
-\end{figure}
+\textbf{function} $\text{sample}(G, s_\text{min}, s_\text{max})$:\\
+\tab\tab a graph $G$, a minimum and maximum motif size $s_\text{min}$, $s_\text{max}$ \\
 
-The sampling algorithm produces full labeled, induced subgraphs.
+\tab Choose $s$ uniformly from $[s_\text{min} .. s_\text{max}]$ \\
+\tab Choose a starting node $n$ uniformly from $V_G$ \\
+\tab $N \from \{n\}$ \\
+\tab \textbf{while} $|N| < s$:\\
+\tab \tab Choose $n$ uniformly from $N$\\
+\tab \tab Choose $n'$ uniformly from the neighbors of $n$\\
+\tab \tab $N \from N \cup n'$ \\
+\tab \tab After 500 cycles, sample a different starting node\\
 
-For instance, we will use a very fast search for candidate motifs, which returns only a limited number of instances.
-A more expensive search might yield more instances, but never fewer.
-That is, for those motifs that we find, we can be sure that they would also be found by a more expensive method.
-Since pattern mining of this kind often results in too many potential patterns rather than too few, it seems efficient to start with the motifs that can be found with fast, simple methods, and only extend the search once these have been exhausted.
+\tab $M \from G[N]$, the induced subgraph of $N$ \\
+\tab  Sample $n' ~ U([0.. n])$, $e' ~ U([0.. e])$ and $e^*~ U([0..e'])$\\
+\tab Randomly remove $e^*$ links, $e'$ linklabels and $n'$ nodelabels from $M$.\\
+\tab\tab (removed labels replaced by the next available negative integer)\\
+
+\tab \textbf{return} $\text{canonical}(M, N)$ \\
+
+\textbf{function} $\text{canonical}(G, I)$:\\
+\tab\tab a graph $G$, list of integers $I$\\
+\tab \textbf{return} $G$ with nodes in canonical order, $I$ permuted the same way
+\label{algorithm:sampling}
+\end{pseudo}
 
@@
 We extend this to knowledge graphs by also including the frequency with which each relation occurs.
-Let a \emph{degree sequence} $D$ of length $n$ be a triple of three integer sequences: $(D^\text{in}, D^\text{rel}, D^\text{out})$.
+Let a \emph{degree sequence} $D$ of length $n$ be a triple of three length-$n$ integer sequences: $(D^\text{in}, D^\text{rel}, D^\text{out})$.
 If $D$ is the degree sequence of a graph, then node $i$ has $D^\text{in}_i$ incoming links,  $D^\text{out}_i$ outgoing links and for each relation $r$, there are $D^\text{rel}_r$ links.
@@
 Unfortunately, there is no efficient way to compute $|\G_D|$ and even approximations tend to be costly for large graphs.
-Following the approach in \cite{bloem2017large}, we define an approximation.
-We can define a knowledge graph by three length-$m$ integer sequences: $S$, $P$, $O$, with ${(S_j, P_j, O_j)}_j$ the graph's tripleset.
+Following the approach in \cite{bloem2017large}, we define a fast approximation to the configuration model, which works well in practice for motif detection.
+
+We can define a knowledge graph by three length-$m$ integer sequences: $S$, $P$, $O$, with $\{(S_j, P_j, O_j)\}_j$ the graph's tripleset.
 If the graph satisfies degree sequence $D$, then we know that $S$ should contain node $j$ $D^\text{out}_j$ times, $P$ should contain relation $r$ $D^\text{rel}_r$ times and $O$ should contain node $j$ $D^\text{in}_j$ times.
@@
 
-While every member of ${\cal S}_D$ contains a valid graph satisfying $D$, many graphs are represented multiple times.
-Firstly, many elements of ${S}_D$ contain the same link multiple times, which means they are not is ${\cal G}_D$.
+While every member of ${\cal S}_D$ represents a valid graph satisfying $D$, many graphs are represented multiple times.
+Firstly, many elements of ${S}_D$ contain the same link multiple times, which means they are not in ${\cal G}_D$.
 We call the set without these elements ${\cal S'}_D \subset {\cal S}_D$.
-Secondly the links of the graph are listed in arbitrary order; if we apply the same permutation to all three lists $S$, $P$ and $O$, we get an encoding of the same graph.
+Secondly the links of the graph are listed in arbitrary order; if we apply the same permutation to all three lists $S$, $P$ and $O$, we get a new representation of the same graph.
 Since we know that any element in ${\cal S}'_D$ contains only unique triples, we know that each graph is present exactly $m!$ times.
@@
 We call this the edge-list (EL) model.
-It always lower-bounds the true degree sequence model, since it affords some probability mass to graphs that cannot exist.
-\footnotemark Experiments in the classical motif setting have shown that the ES model is an acceptable proxy \cite{bloem2017large}, especially considering the extra scalability it affords.
+It always lower-bounds the true degree sequence model, since it affords some probability mass to graphs that cannot exist.\footnotemark ~Experiments in the classical motif setting have shown that the ES model is an acceptable proxy \cite{bloem2017large}, especially considering the extra scalability it affords.
 
-\footnotetext{Note that we cannot simply think of $p^\text{ES}$ as a uniform model for graphs containing multiple triples, since in we can only divide by $m!$ is we know that all triples are unique.}
+\footnotetext{Note that we cannot simply think of $p^\text{EL}$ as a uniform model for graphs containing multiple links, since in we can only divide by $m!$ if we know that all triples are unique.}
 
@@
 \footnote{Or, equivalently, to make $p^\text{EL}$ a complete distribution on all graphs, we must provide it with a prior on $D$.} For each of the three sequences in $D$ we use the following model:
-\[
- p(D) = \prod_i p^{\N}(D_i) \\
- L(D) = - \sum_i \log p^{\N}(D_i)
-\]
+\begin{align*}
+ p(D) &= \prod_i p^{\N}(D_i) \\
+ L(D) &= - \sum_i \log p^{\N}(D_i)
+\end{align*}
 
@@
 This is an optimal encoding for $D$ assuming that its members are independently drawn from $p^{\N}$.
-When we use $p^\text{DS}$ as the null model, we use the data distribution for $p^{\N}$ to ensure that we have a lower bound to the optimal code-length.
-When we use $p^\text{DS}$ as part of the motif code, we must use a fair encoding, so we use a Dirichlet-Multinomial model to store $D$.
+When we use $p^\text{EL}$ as the null model, we use the data distribution for $p^{\N}$ to ensure that we have a lower bound to the optimal code-length.
+When we use $p^\text{EL}$ as part of the motif code, we must use a fair encoding, so we use a Dirichlet-Multinomial model to store each sequence in $D$.
 

### 2017-08-12 kgpsr/kgpsr.tex
...

@@
 
-\title{A Variational Approach to Knowledge Graph Node Embeddings}
+\title{A Bayesian Approach to Knowledge Graph Node Embeddings}
 
@@
 
-	We present a Variational approach to to the problem of embedding knowledge graph nodes.
+	We present a variational approach to to the problem of embedding knowledge graph nodes.
 \end{abstract}
@@
 Knowledge graph node embeddings are vector representations of the nodes in a knowledge graph.
-For a knowledge graph with nodes $N$, an embedding is a function $v: N \to \R^k$ such that $v$ organizes the nodes meaningfully in the spacer $\R^k$ according to some criterion.
+For a knowledge graph with nodeset $N$, an embedding is a function $v: N \to \R^k$ such that $v$ organizes the nodes meaningfully in the space $\R^k$ according to some criterion.
 
@@
 Many embedding algorithms model the relations in the knowledge graph as maps in the embedding space: either translations, or more complex transformations.
-If for instance, we have the triple $(\fw{Mary}, \fw{childof}, \fw{John})$, then we learn embeddings $v_\fw{Mary}, v_\fw{John} \in \R^k$  (for some chosen embedding dimension $k$) and a map $f_\fw{childof} : \R^k \to \R^k$ for the relation \texttt{childof} such that $f_\fw{childof}(v_\fw{John})$ should be as close to $v_\fw{Mary}$ as possible (for some chosen distance metric).
+If for instance, we have the triple $(\fw{Mary}, \fw{childof}, \fw{John})$, then we learn embeddings $v_\fw{Mary}, v_\fw{John} \in \R^k$  (for some chosen embedding dimension $k$) and a map $f_\fw{childof} : \R^k \to \R^k$ for the relation \texttt{childof} such that $f_\fw{childof}(v_\fw{John})$ should be as close to $v_\fw{Mary}$ as possible (for some chosen notion of distance).
 
@@
 
-Intuitively, this is a simple latent variable problem, that can be solved by an alternating optimization algorithm in the vein of expectation-maximization: given a choice for the node embeddings, we can find the optimal functions to represent the relations, and given a choice for the relation functions, we compute the optimal embeddings.
+The task, then, given a knowledge graph, is to find an embedding $v_i$ for each node $i$, and a function $f_r$ for each relations, such that the resultant Gaussians optimally cover the required nodes.
+Intuitively, this problem can be solved by an alternating optimization algorithm in the vein of expectation-maximization: given a choice for the node embeddings, we can find the optimal functions to represent the relations, and given a choice for the relation functions, we can compute the optimal embeddings.
 Alternating these two optimizations is likely to converge to a good solution.
-To formalize this intuition in an elegant manner, take a Bayesian approach, and find the optimum using variational inference.
-Thge preicsae model is detailed in Section~\ref{section:model}.
+To formalize this intuition in an elegant manner, we take a Bayesian approach, and find a solution using variational inference.
+The precise model is detailed in Section~\ref{section:model}.
 
@@
 \begin{itemize}
-	\item The nodes are embedded in a simple Euclidean space, making no special demands on downstream architectures.
+	\item The nodes are embedded in a simple Euclidean space, making no special demands on downstream architectures. %TODO Check if ComplEx and HolE do actually do this. Explain better
 	\item The multiple-relation issue is solved in an intuitive manner: mapping a parent node to its children creates a probability distribution that covers the children. 
@@
 For our method, we will let $\cal F$ be the set of affine functions, each parametrized by a transformation matrix $A_r \in \R^{k\times k}$, and a translation vector $t_r\in \R^k$.
-Thus our model contains $|N|$ vectors in $\R^k$, and $|R|$ functions each defined by a matrix $A_r$ and vector $t_r$.
+Thus our model contains $|V|$ vectors in $\R^k$, and $|R|$ functions each defined by a matrix $A_r$ and vector $t_r$.
 
@@
 
-% TODO Read up on Bayes and conjugate priors. GEt a better sense of what priors I should use.
+% TODO Read up on Bayes and conjugate priors. Get a better sense of what priors I should use.
 
@@
 \]
+
+If we have $(s, p, o) \in G$, then we have:
+\begin{align*}
+p(X_o \mid F_p = f_p) = f_p(p(X_s))(X_o) \\
+p(X_s \mid F_p = f_p) = f_p^{-1}(p(X_o))(X_s) 
+\end{align*}
+
+Where $X_s$ are random variables over $\R^k$ representing the embedding of $s$, and $F_p$ is a random variable over $\cal F$ representing the function chosen for relation $p$.
+What this states, is that for a given choice of function embedding $f_p$, the probability distribution over $X_o$ can be found by taking the probability distribution over $X_s$, and transforming it by $f_p$.
+Starting with the joint distribution, and marginalizing out $F_p$:
+
+\begin{align*}
+p(X_o) = \int p(F_p = f_p) f_p(p(X_s))(X_o) df_p\\
+p(X_s) = \int p(F_p = f_p) f_p(p(X_o))(X_s) df_p\\
+\end{align*}
+
+Which is a recursive, cyclic definition. Use KL divergence instead, to measure error.
 

### 2017-08-12 newton/newton.tex
...

\documentclass[11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}

\newcommand{\N}{{\mathbb N}}
\newcommand{\R}{{\mathbb R}}
\newcommand{\W}{{\boldsymbol W}}
\newcommand{\G}{{\cal G}}
\newcommand{\I}{{\cal I}}
\newcommand{\tab}{\hspace*{5mm}}
\newcommand{\from}{\leftarrow}

\title{Subspace search: exact Newton's method, restricted to a linear subspace}

\begin{document}

\maketitle

\begin{abstract}
\noindent It is well known that Newton's method can lead to far better optimization performance than plain gradient descent. Unfortunately the memory and time required to compute and invert the Hessian, make Newton's method infeasable for the optimization of functions with many parameters, such as those used in deep learning applications. Instead of approximating the inverse Hessian, as is done is quasi-newton methods, we present a method which first reduces the parameter space of the function, and then applies Newton's method without approximation. We compare our method to several quasi-newton methods and other optimizers, on several common models from the deep learning literature. We show that our method converges faster and to lower loss than the alternatives, at the price of minimal overhead.
\end{abstract}

\emph{Line search} is a commonly used algorithm to choose a step size in iterative optimization. Broadly it works as follows. Let $f(x)$ be a function over $\R^d$ we aim to minimize, and let $x^n$ be our current candidate. Compute a direction in which we expect the minimium to lie (using either the negative gradient, or Newton's method). Minimize the function along this direction only, using ternary search, and let $x^{n+1}$ be the location of this minimum.

They key insight is that reducing $f(x)$ to a function of one parameter allows us to use a much more powerful search method to find the optimum. It may only be the optimum in a one-dimensional subspace, but as a next step in the iteration, it makes a much better candidate than a step of arbitrary size in our chosen direction.

 Our method is a generalization of this idea. Let $f(x)$ be a function over a very high-dimensional space (e.g. the loss of a ConvNet with millions of parameters). We can reasonably compute its gradient, and at most a diagonal approximation of the Hessian, but not much more. Instead of approximating the Hessian, we first reduce the function to a lower parameter space. Let $\W \in \R^{d\times k}$ be a matrix with $k$ basis vectors as its columns (with $k$ many orders of magnitude smaller than $d$). The function $g(y) = f(\W y)$  with $y \in \R^k$ is the function $f$ restricted to the linear subspace spanned by $\W$.

We then apply the exact Newton's method to $g(z)$. This requires inversion of the Hessian of $g(z)$, but this is only a $k\times k$ matrix, so if we choose $k$ to be small enough, the computation required will be small.

The question remains how we should choose $\W$. In line search based on the gradient, the direction chosen is often characterized as the direction of steepest ascent. For our purposes, it is more instructive to take a slightly different perspective. The gradient gives us the direction of the optimum of the graph, if we approximate it with a tangent linear function. The direction provided by Newton's method, likewise, gives us the direction of the optimum if we approximate the graph locally with a quadratic function. Thus, we are looking for the $k$-dimensional linear subspace most likely to contain the optimum. Using only the gradient, there is not much more we can say than that the chosen subspace should contain the gradient direction. Since we will be applying Newton's method in the second step, we would prefer the subspace for which the second derivatives are largest.

\paragraph{Related work}

\begin{itemize}
\item Deep Learning via Hessian-Free Optimization
\item Krylov subspace
\item subspace optimization
\end{itemize}}

\section{Method}

\begin{itemize}
\item Compute the diagonal of the Hessian $h$
\item Select the $k$ standard basis vectors (i.e. one-hot-vectors) for which the corresponding value $h_i$ is largest, plus the gradient. (ie. we restrict $f(x)$ to those dimensions in which the second derivative is greatest).  \item 
\end{itemize}

\section{Results}

\section{Discussion}

\bibliography{newton}

\end{document}

### 2017-08-16 fractal-plr/src/fractal.prl-final.tex
...

--- 2017-07-03 fractal-plr/src/fractal.tex
@@
 
-\noindent The word \emph{fractal} is not precisely defined\footnote{Mandelbrot originally defined fractals using Haussdorf dimension, but later stated that he preferred the word to be not precisely defined.}, but typically refers to objects with self-similarity, a non-integer Hausdorff dimension, and an infinitely fine structure.
+\noindent The word \emph{fractal} is not precisely defined \footnote{Mandelbrot originally defined fractals using Haussdorf dimension, but later stated that he preferred the word to be not precisely defined.}, but typically refers to objects with self-similarity, a non-integer Hausdorff dimension, and an infinitely fine structure.
 

### 2017-08-16 kgmotifs/kgmotifs.tex
...

@@
 \usepackage{float}
+
+\usepackage{easylist}
+\usepackage{soul}
 
@@
 \begin{document}
+
+\section{TODO}
+
+\paragraph{Code}
+
+\begin{easylist}[itemize]
+§ Create fast KGraph implements DTGraph$<$Integer, Integer$>$.
+§§ \st{memory-back implementation}
+§§ \st{Load RDF data (from HDT)}
+§§ \st{Copy all tests from DTGraph, test}
+§§ Create Disk-backed version.
+§ Implement null model.
+§ Implement sampling code. 
+§§ $\to$ Implement Nauty for knowledge graphs
+§§§ Test \emph{really} thoroughly.
+§§ Implement sampling.
+§§ Test throughly. 
+§ Implement \textbf{motif code}
+§§ Tests:
+§§§ Implement random sampler. Check that random samples do not have significant motifs.
+§§§ Sum probabilities for large sample of small graphs. Check that they do not sum to more than one.
+§§§ ???
+\end{easylist}
+
+\paragraph{Paper}
+
+\begin{easylist}[itemize]
+§ Write preliminaries
+§§ Code and MDL
+§§ DM code, $\N$ code
+§ Write hypothesis test method
+§ Experiments
+§ Conclusion
+\end{easylist}
 
@@
 % TODO: General purpose codes: natural numbers, integers, sequences.
+
+\paragraph{Common codes}
+
+Extended DM code: test if necessary
 
@@
 For the variable nodes, it will be a list containing some subset of the nodes in $G$.
- We store each list using the DM code.
-It may seem wasteful to store a list repeating a single element (instead of storing a labeled version of the pattern), but under the DM code such lists will be stored efficiently, using only a small, constant number of bits.
+ We store each list using the extended DM code.
+It may seem wasteful to store a list repeating a single element (instead of storing a labeled version of the pattern), but under the extended DM code such lists will be stored efficiently, using only a small, constant number of bits.
 The more varied the list, the more bits will be required to store it.
@@
  {m \choose {D_1^\text{out}, \ldots, D_n^\text{out}} }
- {m \choose {D_1^\text{rel}, \ldots, D_n^\text{rel}} }
+ {m \choose {D_1^\text{rel}, \ldots, D_{|R_G|}^\text{rel}} }
  {m \choose {D_1^\text{in}, \ldots, D_n^\text{in}} } \text{.}
@@
 p^\text{EL}_D(G) =  \frac{m!}{{m \choose {D_1^\text{out}, \ldots, D_n^\text{out}} }
- {m \choose {D_1^\text{rel}, \ldots, D_n^\text{rel}} }
+ {m \choose {D_1^\text{rel}, \ldots, D_{|R_G|}^\text{rel}} }
  {m \choose {D_1^\text{in}, \ldots, D_n^\text{in}} }}
@@
 We call this the edge-list (EL) model.
-It always lower-bounds the true degree sequence model, since it affords some probability mass to graphs that cannot exist.\footnotemark ~Experiments in the classical motif setting have shown that the ES model is an acceptable proxy \cite{bloem2017large}, especially considering the extra scalability it affords.
+It always lower-bounds the true degree sequence model, since it affords some probability mass to graphs that cannot exist.\footnotemark ~Experiments in the classical motif setting have shown that the EL model is an acceptable proxy for the DS model \cite{bloem2017large}, especially considering the extra scalability it affords.
 
-\footnotetext{Note that we cannot simply think of $p^\text{EL}$ as a uniform model for graphs containing multiple links, since in we can only divide by $m!$ if we know that all triples are unique.}
+\footnotetext{Note that we cannot simply think of $p^\text{EL}$ as a uniform model for graphs containing multiple links, since we can only divide by $m!$ if we know that all triples are unique.}
 
@@
 
-In this section, we show some experiments that are not intended to substantiate any claims about our method, but rather to illustrate the variety of uses for succesfull motif detection.
+In this section, we show some experiments that are not intended to substantiate any claims about our method, but rather to illustrate the variety of uses for succesful motif detection.
 
@@
 
+\appendix
+
+\section{Appendix}
+
+\subsection{The canonical isomorphism algorithm for knowledge graphs}
+
+During sampling, we want to re-order the nodes in a sampled pattern to some canonical ordering.
+This way, when we sample isomorphic patterns, we recognize it, and group their instances together.
+The \emph{nauty} algorithm is the standard solution for findings such \emph{canonical isomorphs}.
+As we are sampling partially labeled, directed graphs, we have a lot of information that can speed up canonicalization.
+Since this steop is usually the bottleneck in sampling, it's important to perform it as efficiently as possible.
+Below, we provide a very succinct description of how nauty works, when applied to our patterns.
+
+Let $V'_M$ and $R'_M$ represent the labels of the variable nodes and links in $M$.
+Pattern $M$ is an isomorphism of pattern $M'$ if there exist bijections $\sigma_v : V'_M \times V'M$ and $\sigma_r: \R'_M \times \R'_M$ such that applying these to the variable elements of $M$ yields $M'$ (under graph equivalence).
+The set of all patterns isomorphic to $G$ is the isomorphism class $[G]$.
+A function $C(P)$ is a canonicalization function if it produces for all any pattern $M$ in $[M]$ the same, canonical member of $M$.
+
+Note that a very simple canonicalization algorithm would be to pass the unlabeled graph to the nauty algorithm, and use its canonical ordering on the labeled graph.
+However, the labels can make the canonicalization procedure significantly faster.
+
+The key principle behind the nauty algorithm is the use of a \emph{colouring}: a partition of the nodes of a graph.
+A \emph{safe colouring} is defined as a partition $\pi$ of the nodes of the graph such that any permutation that maps nodes from different partitions to one another is guaranteed \emph{not} to be an automorphism.
+No guarantees are made about nodes in the same cell.
+For instance, the partition that puts all nodes in the same cell is always safe.
+If two nodes have different degrees, we can put them in different cells and be sure that we have a safe colouring.
+Given one safe colouring $\pi$, if node $a$ from partition $i$ has a neighbor in partition $j$ and node $b$ from partition $j$ has no such node, we can \emph{refine} the partition $\pi$ by separating $i$ into a new partition (one contain $a$ and one containing $b$).
+
+The nauty algorithm operates on a graph with a given colouring (with the unit colouring used by default).
+It first refines the colouring as much as possible, using degree information, and propagating the distinguishing features through the graph.
+Once the most refined colouring has been achieved, it searches through all autorphisms of the graph to produce a canonical example (using various optimizations to prune the search tree).
+
+In our case, we use the colouring to indicate what our labeling tells us about the colouring of the nodes.
+Note the following:
+\begin{easylist}
+§ Labeled nodes in the pattern can be placed in a discrete cell of the colouring.
+§ Variable nodes should be placed into a single cell in the partition.
+§ We define a partition over nodes \emph{and} relation labels. 
+joh\end{easylist}
+
 \end{document}

### 2017-08-16 kgpsr/kgpsr.tex
...

@@
 
+\usepackage{easylist}
+\usepackage{soul}
+
 \DeclareMathOperator*{\argmin}{arg\,min}
+
+\newcommand{\R}{{\mathbb R}}
+\newcommand{\cN}{{\mathcal N}}
+\newcommand{\fw}[1]{\text{\texttt{#1}}}
+\newcommand{\p}{\text{.}}
 
@@
 
-\newcommand{\R}{{\mathbb R}}
-\newcommand{\fw}[1]{\text{\texttt{#1}}}
+\begin{easylist}
+§ Strict Bayesian approach doesn't seem to work: no data, no posterior.
+But, posterior over latent variables.
+Latent variables are embeddings and tranformations, \emph{observed are the memberships/responsibilities }.
+So we \emph{do} have observed variables?
+§§ Optimize for the posterior over the connections... This would tie in better with the link prediction use case.  
+§§ If the structure of the grpah is the latent variable, we should be bale to sample a structure from the model.
+§ Take inspiration from the fractal model start with $X_0 ~ N(0, I)$ as embedding for all
+nodes and iterate $X_1$ is $X_0$ transformed by all translations, for all points.
+§ It doesn't look like the constraints of just a probability model are sufficient to determine the problem.
+It seems most elegant to assume that the embeddings are (isometric) Gaussians, and to optimize for minimal KL divergence to the transformed ones.
+
+\end{easylist}
 
@@
 
-Let $x_i \in \R^k$ be the embedding chosen for node $i \in V$.
-Let $f_r$ be a parametrized function from $\cal F$, representing relation $r \in R$.
-For our method, we will let $\cal F$ be the set of affine functions, each parametrized by a transformation matrix $A_r \in \R^{k\times k}$, and a translation vector $t_r\in \R^k$.
-Thus our model contains $|V|$ vectors in $\R^k$, and $|R|$ functions each defined by a matrix $A_r$ and vector $t_r$.
+Let $T_{spo}$ be a random bolean variable indicating whether the triple $\langle s, p, o\rangle$ is true.
+Let $T^u$ be the set of all such variables (for all combination of $s$, $p$, $o$) and let $T$ be the set of such variables for the observed triples.
+We will assume that the observed triples we samples from the set of all true triples by some unspecified distribution $p(G| G^u)$.
+This graph is fully specified by a set of node-embeddings $x_i$ and relation-maps $f_p$: we have $f_p(x_s) = x_o$ is and only if the triple $s, p, o$ is true.
+Thus, the embeddings specify the graph.
+We assume that the observed triples $T$ constitute a sequence of i.i.d.
+draws from the set of all triples satisfying the embeddings.
 
-To formalize the intuitions expressed in the introduction, we first express the constraints in a graphical model.
-Let $E^r_\text{subject} = \{s \mid \exists o, (s, r, o) \in E\}$ and $E^{r, o}_\text{object} = \{ s \mid (s, r, o) \in E\}$.
-Figure~\ref{figure:graphical-model} shows the relations between the different random variables.
-Interestingly, this model contains \emph{no observed variables}.
-The data is used purely to inform the topology of the graphical model.\footnotemark The model also gives us the priors: the prior probability for the embeddings, for which we will use the standard normal distribution: $\mu = 0$ and $\Lambda = \bf{1}$ and the prior $W_t$ on the
-
-\footnote{Modeling the data as observations would lead to tensor-factorization approaches like ReSCAL \cite{}.}
-
-% TODO Read up on Bayes and conjugate priors. Get a better sense of what priors I should use.
-
+To develop our model, we will first assume we know a value $x_i \in \R^k$ for each node $i$ in the knowledge graph, and a function $f_p$ for each predicate $p$.
+We will choose affine functions as our family, so that we can say (with some abuse of notation) $f_p \in \R^{k\times k} \times \R^k$.
+For such a fully known model, we will say that a triple is true if and only if $f_p$ is mapped exactly onto $x_s$:
 \[
-\argmin_{\{x_i\}, \{f_r\}, \sigma} \prod_{(s,r,o) \in E_G} f_r(N_{x_s,\sigma})(x_o)
+p(T_{spo} = \text{true} \mid f_p, x_s, x_o) = \begin{cases} 1 & \text{if} f_p(x_s) = x_o \\ 0 & \text{otherwise} \end{cases} \p
 \]
 
-If we have $(s, p, o) \in G$, then we have:
+This model suffers from the same problem as TransE: we cannot represent multiple relations.
+Interestingly, the fact that we do not observer $x_s$, $x_o$ and $f_p$ directly allows us to solve this problem.
+We express our uncertainty about these values in probability.
+If $s$ is mapped to several points by $p$, we are thus \emph{uncertain} about which point $f_p$ should map $x_s$ to.
+Expressing this uncertainty as probability as well allows as to retrieve a probability distribution over the model parameters that let $x_s$ map to deveral different $x_o$'s under $f_p$.
+
+We start dy removing the assumption that we know the model parameters.
+We express our uncertainty as probabilities: $p(F_p = f_p)$, $p(X_s = x_s)$, $p(X_o = x_o)$ and marginalize them out.
+The probability of a triple becomes:
 \begin{align*}
-p(X_o \mid F_p = f_p) = f_p(p(X_s))(X_o) \\
-p(X_s \mid F_p = f_p) = f_p^{-1}(p(X_o))(X_s) 
+p(T_{spo} = \text{true}) &= \int_{x_s, f_p, x_o \in S_{spo}} p(X_s = x_s) p(F_p = f_p) p(X_s = x_s) \,d x_s df_p dx_o\\
+\text{with} \,\,S_{spo} &= \{(x_s, f_p, x_o) \mid f_p(x_s) = x_o\}
 \end{align*}
 
-Where $X_s$ are random variables over $\R^k$ representing the embedding of $s$, and $F_p$ is a random variable over $\cal F$ representing the function chosen for relation $p$.
-What this states, is that for a given choice of function embedding $f_p$, the probability distribution over $X_o$ can be found by taking the probability distribution over $X_s$, and transforming it by $f_p$.
-Starting with the joint distribution, and marginalizing out $F_p$:
-
 \begin{align*}
-p(X_o) = \int p(F_p = f_p) f_p(p(X_s))(X_o) df_p\\
-p(X_s) = \int p(F_p = f_p) f_p(p(X_o))(X_s) df_p\\
+p(\{X_i\}, \{F_p\}\mid T) = \frac{p(T\mid \{X_i\}, \{F_p\}) \prod_i p(X_i)\prod_p p(F_p)}{P(T)}
 \end{align*}
-
-Which is a recursive, cyclic definition. Use KL divergence instead, to measure error.
 
@@
 
+It is worth noting that we started from an impossible definition of the structure of our graph.
+The presumed unobserved true embeddings cannot exist for any graph contain multiple children or parents.
+Moreover, the system may be overdetermined: even with unique relations for every triple, there may still be no way to satisfy the whole knowledge graph in a single set of embeddings.
+Nevertheless, the Bayesian approach allows us to satisfy the resulting contradictions, by not choosing a single truth, but but providing a probability distribution over configurations that are each close to a correct one.
+In other words, the uncertainty about how the contradictions should be resolved is simply modeled as another probability, and folded into the model, combined with our other uncertainties.
+
 \bibliography{kgpsr}

### 2017-11-14 motifs/motifs.tex
Ready for second resubmission

\documentclass[twoside,11pt]{article}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{jmlr2e}

\usepackage{float}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}

\usepackage{amsmath}
\usepackage{threeparttable}

\usepackage{booktabs}

%% Include all macros below

\newcommand{\p}{\mbox{\,.}}
\newcommand{\fl}[1]{\left \lfloor #1 \right \rfloor}
\newcommand{\g}[1]{\color{gray} #1 \color{black}}
\newcommand{\B}{{\mathbb B}}
% \newcommand{\G}{{\mathbb G}}
\newcommand{\cG}{{\mathcal G}}
\newcommand{\cS}{{\mathcal S}}
\newcommand{\Rbb}{{\mathbb R}}
\newcommand{\N}{{\mathbb N}}
\newcommand{\C}{{\mathscr C}}
\newcommand{\Co}{{\overline{\mathscr C}}}
\newcommand{\R}{{\mathscr R}}
\newcommand{\m}{{\overline{m}}}
\newcommand{\ok}{{\overline{\kappa}}}
\newcommand{\s}[1]{{\overline{#1}}}
\newcommand{\mmid}{\;\middle|\;}
\newcommand{\hide}[1]{}
\newcommand{\tab}{\hspace*{5mm}}
\newcommand{\stack}[2]{\genfrac{}{}{0pt}{}{#1}{#2}}

\floatstyle{ruled}
\newfloat{pseudo}{h}{lop}
\floatname{pseudo}{Algorithm}

\newcommand{\pb}[1]{\textcolor{OliveGreen}{\small #1 \textsuperscript{[pb]} }}

% Heading arguments are {volume}{year}{pages}{submitted}{published}{author-full-names}

\jmlrheading{1}{2000}{1-48}{12/16}{10/00}{Peter Bloem and Steven de Rooij}

% Short headings should be running head and authors last names

\ShortHeadings{Finding Network Motifs in Large Graphs}{Bloem and de Rooij}
\firstpageno{1}

\begin{document}

\title{Finding Network Motifs in Large Graphs using Compression as a Measure of Relevance}

\author{\name Peter Bloem \email vu@peterbloem.nl \\
       \addr Knowledge Representation and Reasoning Group\\
        VU University Amsterdam \\
		De Boelelaan 1105, 1081 HV Amsterdam, NL \AND
       \name Steven de Rooij \email steven.de.rooij@gmail.com \\
       \addr Mathematical Institute \\
        University of Leiden\\
        Niels Bohrweg 1, 2333 CA Leiden, NL}

\editor{\ldots}

\maketitle

\begin{abstract}%
We introduce a new method for finding \emph{network motifs}: interesting or informative subgraph patterns in a network. Subgraphs are motifs when their frequency in the data is high compared to the expected frequency under a \emph{null model}. To compute this expectation, a full or approximate count of the occurrences of a motif is normally repeated on as many as 1000 random graphs sampled from the null model; a prohibitively expensive step. We use ideas from the Minimum Description Length (MDL) literature to define a new measure of motif relevance. With our method, samples from the null model are not required. Instead we compute the probability of the data under the null model and compare this to the probability under a specially designed alternative model. With this new relevance test, we can search for motifs by random sampling, rather than requiring an accurate count of all instances of a motif. This allows motif analysis to scale to networks with billions of links, while still resulting in informative motifs. 
\end{abstract}

 \begin{keywords} graphs, minimum description length, unsupervised learning, pattern mining, network motifs \end{keywords}

\section{Introduction}

Graphlets are small, induced subgraphs a large network. \emph{Network motifs} \citep{milo2002network} are those graphlets that occur more frequently in the data than expected. Graphlets and network motifs are used for many purposes, for instance: as features in machine learning \citep{shervashidze2009efficient,wu2011characterizing}, as candidates for ``functional units'' of the system represented by the graph \citep{milo2002network}, as a method to reduce the complexity of the graph \citep{koutra2015summarizing} or as a building block in community detection \citep{arenas2008motif}. To be able to conclude that such frequent subgraphs really represent meaningful aspects of the data, we must first show that they are not simply a product of chance. That is, a subgraph may simply be a frequent subgraph in \emph{any} random graph: a subgraph is only a \emph{motif} if its frequency is higher \emph{than expected}.

This expectation is defined in reference to a \emph{null model}: a probability distribution over graphs. We determine what the expected frequency of the subgraph is under the null model, and if the observed frequency is substantially higher than this expectation, the subgraph is a motif.\footnotemark

\footnotetext{Commonly, all possible graphlets of a given size are checked. Depending on the task, the full list of motif scores is used as a feature vector representing the graph, or the top-$k$ motifs are used by a domain expert as candidates for further analysis. Since each motif is judged by a hypothesis test, the reader may wonder whether this approach constitutes multiple testing. However, as discussed in Section~\ref{section:multiple-testing}, motif analysis uses the significance test merely as a pattern-mining heuristic, not as a true statistical test.}

Unfortunately, there is usually no efficient way to compute the expected frequency of a subgraph under a null model. The most common approach is to generate a large number of random graphs, say 1000, from the null model and compare the frequencies of the subgraph in the graphs in this sample to its frequency in the data \citep{milo2002network}. This means that any resources invested in extracting the motifs from the data must be invested again 1000 times to find out which subgraphs are motifs.

We introduce an alternative method that does not require such sampling from the null model. To achieve this, we use two probability distributions on graphs: the null model $p^\text{null}$, and a distribution $p^\text{motif}$ under which graphs with one or more frequent subgraphs have high probability. If a subgraph $M$ of a given graph $G$ allows us to show that $p^\text{motif}(G)$ is larger than $p^\text{null}(G)$, then $M$ is a motif.

To design $p^\text{motif}$, we use the Minimum Description Length (MDL) Principle \citep{rissanen1978modeling,grunwald2007minimum}. We design a description method for graphs, a \emph{code}, which uses the frequent occurrence of a potential motif $M$ to create a compressed description of the graph. Our approach is analogous to compressing text by giving a frequent word a brief codeword: we describe $M$ once and refer back to this description wherever it occurs.

By a commonly used correspondence between codes and probability distributions (detailed in the preliminaries), we derive $p^\text{motif}$ from this code, a distribution that assigns high probability to graphs containing motifs. Section~\ref{section:model-selection} explains the principle and its theoretical justification.

This approach removes the need to take samples from the null model: we only need to compute $p^\text{null}(G)$ and $p^\text{motif}(G)$. It also removes the need for accurate subgraph counts. For a potential motif $M$, we only need to find enough occurrences in  $G$ to achieve the required level of compression. We use a simple random sampling algorithm, which operates in constant time with respect to the size of the graph. These two improvements lead to a considerable increase in the scale of graphs which can be analysed.

Specifically, we show the following:
\begin{itemize}
  \item On commodity hardware, our method can be used to analyze graphs with millions of links in minutes. We can analyze graphs with billions of links in under 9 hours on a basic compute node, and in about 45 hours on hardware equivalent to a single laptop (Section~\ref{section:large}). Contrast this with the current scale of motif analysis, which is limited to graphs in the order of 1000 to 10 000 links.\footnote{\citet{hovcevar2014combinatorial} report a complete subgraph \emph{count} on a graph with 137 775 links, the largest we have found, but the comparison to a null model required for motif analysis is not performed.}
  \item The method can be used for motifs with at least 10 nodes, although space and time complexity do increase for larger motif sizes (Section~\ref{section:large}).  
  \item Our method can retrieve motifs that have been injected into the data, even at relatively low quantities (Section~\ref{section:recovering}).
  \item In real data, the judgments on whether a subgraph is a motif or not produced by our method are as informative in characterizing the graph as those returned by the traditional method. We evaluate this by using the motif judgments as binary features in a graph classification task (Section~\ref{section:classification}). 
\end{itemize}

The paper is structured as follows. The next two subsections discuss related work and preliminary information on graphs and MDL. Section~\ref{section:model-selection} provides an in-depth explanation of the method we use to perform hypothesis testing through MDL. In Section~\ref{section:motif-code}, we detail our motif model. In Section~\ref{section:null-models} we detail the three null-models we will use to test its performance. In Section \ref{section:experiments}, we detail briefly how we search for motifs, and select candidates, and then report the results of four experiments to validate the claims made above.

All software used in this paper is available open source. \footnote{See \url{https://github.com/pbloem/motive} and \url{https://github.com/pbloem/motive-cls}.}

\subsection{Related Work}
Many different algorithms, techniques and tools have been proposed for the detection of motifs, all based on a common framework, consisting of three basic steps:

\begin{enumerate}
  \item Obtain a count $f_M$ of the frequency of subgraph $M$ in graph $G$. \label{step1}
  \item Obtain or approximate the probability distribution over the number of instances $F_M$ given that $G$ came from a particular null model $p^\text{null}$.  \label{step2}
  \item If $p^\text{null}(F_M \leq f_M) \leq \alpha$ (usually with $\alpha = 0.05$), we consider $M$ a motif. \label{step3}
\end{enumerate}

\noindent This was the approach proposed by \cite{milo2002network}, in the paper that where the phrase \emph{network motif} was coined. One problem with this approach is that it is very expensive to perform naively. Step \ref{step1} requires a full \emph{graph census},\footnote{A graph census for size $k$ counts, for each connected graph $M$ of size $k$, the number of times the graph occurs as an induced subgraph of the data.} and since the probability in step \ref{step3} cannot usually be computed analytically, we are required to perform the census again on thousands of graphs sampled from the null model in order to approximate it.

Most subsequent approaches have attempted to improve efficiency by focusing on step~\ref{step1}: either by designing algorithms to get exact counts more efficiently \citep{koskas2011nemo,li2012netmode,khakabimamaghani2013quatexelero,meira2014acc}, or by approximating the exact count. The most extreme example of the latter is \citet{kashtan2004efficient}, which simply counts randomly sampled subgraphs. The complexity of this algorithm is \emph{independent of the size of the data}, suggesting an exceptionally scalable approach to motif detection. Unfortunately, while the resulting ranking of motifs by frequency is usually accurate, the estimate of their total frequency is not \citep{wernicke2005faster}, which makes it difficult to build on this approach in steps \ref{step2} and \ref{step3}. Other algorithms provide more accurate and unbiased estimates \citep{wernicke2005faster,ribeiro2010g,bhuiyan2012guise,jha2015path,slota2014complex,paredes2015rand}, but they do not maintain the extreme scalability of the sampling approach.

We take an alternative route: instead of improving the sampling, we change the measure of motif relevance: we define a new hypothesis test as an alternative to steps \ref{step2} and \ref{step3}, which does not require an accurate estimate of the number of instances of the motif. All that is required is a set of some instances; as many as can be found with the resources available. This means that the highly scalable sampling approach from \cite{kashtan2004efficient} can be maintained.

The frequency of a motif is not the only statistic we can use. We might, for instance, compare the level of clustering among the instances with the value expected under the null model. This is relevant, because our compression method can be seen as a combination of several statistics. Specifically, properties that allow a motif to reject the null hypothesis are: a high frequency of \emph{non-overlapping}\footnotemark~instances, a large number of instances with few links to nodes outside the instance and a high number of internal links. While this deviates from the more common statistic of pure frequency, we believe these are all properties that fit the intuitive idea of a motif. In the graph mining literature, different notions of frequency of a subgraph are often called \emph{support measures} \citep{wang2012efficiently}.

\footnotetext{Using non-overlapping instances is known as \emph{frequency concept 3} in \citep{schreiber2004towards}.}

\cite{picard2008assessing} provide an analytical solution, which can perform the hypothesis test without generating a random sample of networks. The constraint here is that the null model should be \emph{exchangeable}. For the commonly used degree-sequence model, a sampling based-approximation is used.\footnotemark~The complexity of the hypothesis test is independent of the the size of the graph, but it is is $O(k!^2)$ in the size $k$ of the motif. While this method could be used for very scalable motif detection (for small $k$, with exchangeable null models), we are not aware of any practical efforts to this effect.

\footnotetext{The degree sequence model always seems to require approximate solutions. FANMOD \citep{wernicke2005faster} requires the computation of the number of graphs with a given degree sequence and a specific subgraph at a particular location. They use the approximation from \citet{bender1978asymptotic}, which is asymptotically correct, but provides no bouds for the case of a finite graph. For our algorithm, we require a similar estimate (the number of graphs with a particular degree sequence). We use the sampling-based methods from \cite{blitzstein2011sequential,charo2010efficient}. These are slower, but they provide a clearer bound on the accuracy of the estimate, as described in Section~\ref{section:null models} and the supplementary materials.}

The idea that compression can be used as a heuristic for subgraph discovery was also used in the SUBDUE algorithm by \citet{cook1994substructure}. Our approach uses a more refined compression method and we connect it explicitly to the framework of motif analysis and the use of hypothesis tests as a heuristic for pattern mining. We also exploit the possibility that the MDL approach offers, for a very scalable sampling algorithm, to replace the more restrictive beamsearch used in SUBDUE. In itemset mining, the use of MDL has a long and successful history, as detailed by \citet[Chapter~8]{aggarwal2014frequent}.

The literature behind graph mining, graphlets and network motifs seems to have developed largely in parallel, repeating much of the same ground. While graph mining tends to focus on datasets consisting of many small graphs, rather than one large graph, there are efforts focusing on the latter. For a good overview of recent work, we refer the reader to \citet[Chapter~13]{aggarwal2014frequent}. Research on graphlets likewise tends to focus on datasets where each instance is represented by a separate graph, usually in a classification setting \cite{shervashidze2009efficient}. However, recently, a great deal of machine learning research has emerged that deals with the case where the dataset consists of a single graph, and the instances to be classified are a subset of its nodes \citep{pham2016column}. Where graphlet approaches tend to focus on the raw frequencies of subgraphs in the data, network motif research focuses on the frequency relative to the expectation under a null model.

\subsection{Preliminaries: Graphs and Codes}
\label{section:preliminaries}

\paragraph{Graphs} A graph $G$ of size $n$ is a tuple $(V_G, E_G)$ containing a set of \emph{nodes} (or \emph{vertices}) $V_G$ and a set of \emph{links} (or \emph{edges}) $E_G$. For convenience in defining probability distributions on graphs, $V_G$ is always the set of the first $n$ natural numbers. $E_G$ contains pairs of elements from $V_G$. For the dimensions of the graph, we use the functions $n(G) = |V_G|$ and $m(G) = |E_G|$. If a graph $G$ is \emph{directed}, the pairs in $E_G$ are ordered, if it is \emph{undirected}, they are unordered. A \emph{multigraph} has the same definition as a graph, but with $E_G$ a multiset, i.e. the same link can occur more than once.

There are many types of graphs and tailoring a method to each one is a laborious task. Here, we limit ourselves to datasets that are (directed or undirected) \emph{simple graphs}: i.e. no link connects a node to itself and the same link cannot occur more than once. Probability distributions on the set of simple graphs are usually the most complex, so that we can trust that a method that works for simple graphs is easily translated to other types.

Two graphs $G$ and $H$ are \emph{isomorphic} if there exists a bijection $f: N_G \to N_H$ on the nodes of $G$ such that two nodes $a$ and $b$ are adjacent in $G$ if and only if $f(a)$ and $f(b)$ are adjacent in $H$. If two graphs $G$ and $H$ are isomorphic, we say that they belong to the same \emph{isomorphism class} $[G]$.

\paragraph{Codes and MDL} The intuition behind MDL is that compressing data and learning from it are much the same process: in both cases, we are analyzing data to find patterns that are characteristic for the data, or patterns that represent the data well. This is not just an intuition, there exists a very precise correspondence between optimizing for probability and optimizing for description length. We will detail the basic principle below, and give an intuitive example of how we apply the method in the next section. For more details, we refer the reader to \cite{grunwald2007minimum}.

Let $\B$ be the set of all finite-length binary strings. We use $|b|$ to represent the length of $b \in \B$. Let $\log(x) = \log_2(x)$. A \emph{code} for a set of objects $\cal X$ (usually graphs, in our case) is an injective function $f: {\cal X} \to \B$, mapping objects to binary code words. All codes in this paper are \emph{prefix-free}: no code word is the prefix of another. We will denote a \emph{codelength function} with the letter $L$, ie. $L(x) = |f(x)|$. It is common practice to compute $L$ directly, without explicitly computing the codewords. In fact, we will adopt the convention of referring to $L$ itself as a code.

A well known result in information theory is the association between codes and probability distributions, implied by the  \emph{Kraft inequality}: for each probability distribution $p$ on $\cal X$, there exists a prefix-free code $L$ such that for all $x \in \cal X$: $- \log p(x) \leq L(x) < -\log p(x) + 1$. Inversely, for every prefix-free code $L$ for $\cal X$, there exists a probability distribution $p$ such that for all $x \in \cal X$: $p(x) = 2^{-L(x)}$. For proofs, see \cite[Section~3.2.1]{grunwald2007minimum} or \cite[Theorem~5.2.1]{cover2006elements}. To explain the intuition, note that we can easily transform a code $L$ into a sampling algorithm for $p$ by feeding the decoding function random bits until it produces an output. To transform a probability distribution to a code, techniques like arithmetic coding \citep{rissanen1979arithmetic} can be used.

As explained by \citet[page 96]{grunwald2007minimum}, the fact that $-\log p^*(x)$ is real-valued and $L^*(x)$ is integer-valued can be safely ignored and we may \emph{identify} codes with probability distributions, allowing codes to take non-integer values.

When we need to encode a single choice from a finite set $S$ of options, we will often use the code with length $\log |S|$, corresponding to a uniform distribution on $S$. In some cases, we can allow codes with multiple codewords for a single object. If a code $L$ has multiple codewords for some object  $x$, we may indicate the choice for a particular codeword by a parameter $a$ as $L(x; a)$.

\section{Model Rejection by Codelength}

\label{section:model-selection}

In this section we provide an intuitive discussion of the principle behind our method. To simplify the explanation, we will illustrate the idea on a simpler problem than motif analysis: determining whether a single large clique in a graph is an unusual feature, or to be expected under a particular null model. For certain null models, such as the Erd\H{o}s-Renyi (ER) model, the expected size of the largest clique is well understood \citep{bollobas1976cliques}. For other models, less so. Without resorting to sampling a large number of graphs from the model, how can we determine that a particular size of clique is unusually large under a given null model? We will define two codes, $L^\text{null}$ and $L^\text{clique}$, and show that if $L^\text{clique}$ compresses the data better than $L^\text{null}$ by enough bits, it is very unlikely that the data came from the null model, and we cannot dismiss the clique as a consequence of random chance.

Consider an undirected graph $G$ containing a large clique: all nodes in some subset $V_C \subseteq V_G$ are directly connected to one another. We can describe the graph by first describing the set $V_C$, and then describing $G$ in a canonical manner. Since every node in $V_C$ is connected to every other node in the clique, we can omit these links from the second part of our description, shortening the total description length, if $V_C$ is large enough. This gives us a code $L^\text{clique}$ with short codelengths for graphs containing a single large clique, and---by the correspondence mentioned in the preliminaries---also a probability distribution $p^\text{clique}$ with high probabilities for graphs with a large clique. The principle is illustrated in Figure~\ref{figure:clique-example}.

It is important to be aware that this is a code with \emph{multiple codewords for the same graph}. To convert this to a probability distribution, we sum the probabilities of all codewords: $p^\text{clique}(G) = \sum_c 2^{-L^\text{clique}(G;c)}$, where each $c$ identifies a clique in $G$.\footnote{We can convert $p^\text{clique}$ back to a another code by taking the negative logarithm: $\bar L^\text{clique}(G) = -\log(G)$. $\bar L^\text{clique}$ is a superior code (it always lowerbounds $L^\text{clique}$), but inefficient to compute.} As we shall see, we can use such codes without problems, \emph{so long as they are not used as the null model}. For null models we require a code with a single codeword per graph.

Of course, there is no guarantee that of all the distributions with a bias towards large cliques, $p^\text{clique}$ is the precise distribution from which we sampled our data. Luckily, it does not need to be: as long as we can show that our clique-based model encodes the data more efficiently than the null model, the presence of the clique disproves the hypothesis that the data came from the null model. We do not know where it came from, but it did not come from the null model. Under the hypothesis that the null model was the source of the data, we can show that the probability that any other model compresses the data better by $k$ bits or more, decays exponentially in $k$. This is known as  the \emph{no-hypercompression inequality} \cite[p103]{grunwald2007minimum}. More precisely, let $p^\text{null}(x)$ be any probability distribution, with $L^\text{null}(x) = - \log p^\text{null}(x)$ and let $L(x)$ be \emph{any} code, then we have
\[
p^\text{null}\left(L^\text{null}(x) - L(x) \geq k\right) \leq 2^{-k} \p
\]

To provide some intuition for why this should be so, imagine that $L^\text{null}$ and $L$ are both codes from uniform distributions: $L^\text{null}$ provides each graph in $X^\text{null}$ with a codeword of constant length $n$, and $L$ provides each graph in $X$ with a codeword of constant length $n-k$. Then, $L^\text{null}$ provides enough codewords for $2^n$ graphs and, $L$ provides enough codewords for $2^{n-k}$ graphs. Thus, if a graph can be represented by both (and thus compressed by $k$ bits), its probability under the null model is at most $2^{-k}$. This is the basic principle allowing us to use compression for statistical purposes: \emph{there are exponentially fewer short codes than long ones, so compressible data is rare}.

To give a sense of scale; under the null model, the probability that $L^\text{clique}$ will compress the data better than the null model by 10 bits or more is less than one in one-thousand. For twenty bits, we get one in a million, for thirty bits, one in a billion, and so on.

We can interpret this procedure as a hypothesis test: the difference in compression $\Delta$ between the null model and the alternative model is a statistic \cite[Example~14.2]{grunwald2007minimum}. The no-hypercompression inequality gives us a bound on the probability $p^\text{null}(D \geq k)$. To reject the null model with significance level $\alpha$, we must find some code on the set of all graphs and show that it compresses the data better than the null model by $k$ bits, with $2^{-k} \leq \alpha$. Any code will do, as long as it was chosen before seeing the data. The better we design our code, the greater the power of the test.

Note that $\Delta$ is also the logarithm of the likelihood ratio between the null model and $L$, so we can see this as a likelihood ratio test. We can also interpret the difference in codelength between two models $p_a$ and $p_b$  as the logarithm of the \emph{Bayes factor} $p_a(x)/p_b(x)$ \cite[Section~14.2.3]{grunwald2007minimum}.

Now, while our test  \emph{rejects} the null model, it does not necessarily \emph{confirm} anything about the pattern we used to compress the data. We would like to make it as likely as possible that it was the pattern we are interested in (e.g. the clique) that allowed us to reject the null model, and not some other aspect of the alternative model. This is a complex problem and, it is difficult to provide any guarantees. If the alternative model stores, say, the linkset of a graph more efficiently than the null model does, it may well be that the alternative model always beats the null model, regardless of the pattern used. To rule out most such artifacts, we must re-use the same model we used for the null model within the alternative model to store every part of the graph except the pattern. There is usually an intuitive way to do this: for instance, in the alternative model for the clique, we store the graph in two parts: first the nodes belonging to the clique, and then the full graph, minus any links in the clique. If we use, say, the Erd\H{o}s-Renyi model as our null model, we also use it to store the graph in this second part of the code. In this way, we know that any compression achieved must be due to the clique.\footnotemark

\footnotetext{The lack of guarantees may be a cause for skepticism. Consider, however, that the hypothesis test by itself  is \emph{never} proof that the pattern found is meaningful, only that the null hypothesis is incorrect. See also Section~\ref{section:multiple-testing}.}

\begin{figure}[tb]
  \includegraphics[width=\textwidth]{./images/mdlexplanation.pdf}
  \caption{A simple example of graph analysis by MDL: finding cliques. (a) The data. (b) A representation of the data that exploits the existence of a large clique. We first store a set indicating which nodes belong to the clique. We then store the graph, omitting those links that belong to the clique. (c) An illustration of the principle of a bound. For a complete code on graphs, using $L^{\text{uniform}}$, we need to first store the size of the graph using some arbitrary code on the integers $L^\N$. To make sure that it does not matter which code we choose for $L^\N$, we require our code ($L^{\text{clique}}$) to beat the bound $B(G)$ (by at least $k$ bits). }
   \label{figure:clique-example}
\end{figure}

To ensure this, we aim to have the alternative model exploit only the pattern, and nothing else: we reuse the same model we used as the null model to store all of the graph, save for the pattern. For instance, in the example above, the clique model must store the graph minus the links of the clique. If we use the null model for this, we know that the only change between the null model and the alternative is the use of the clique, so that must be what made the difference.

\subsection{Rejecting Multiple Null Models}

\label{section:multiplerejection}

A final benefit of this method is that we can reject multiple null models with a single test. In many situations we will have a function $B(G)$ that lowerbounds any code in some set $\cal L$. If our alternative model provides a codelength below $B(G) - k_\alpha$ with $k_\alpha$ the number of bits required for our chosen $\alpha$, we can reject all of $\cal L$.

As an example, Let $\cG_n$ be the set of all undirected graphs of size $n$. For our null model, we define a uniform code on such graphs: $L^\text{uniform}_n = \log |\cG_n|$. This null model captures the idea that the size of the graph is the only informative statistic: given the size, all graphs  are equally likely. However, it is \emph{parametrized}. It is currently not a code on \emph{all} graphs, just those of size $n$. To turn it into a code that can represent all graphs, we need to encode the parameter $n$ as well, with some code $L^\N$ over the natural numbers
\[
L^\text{complete}(G) = L^\N(n(G)) + L^\text{uniform}_{n(G)}(G) \p
\]
This is called \emph{two-part coding}, we encode the parameters of a model first, and then the data given the parameters. For some parametrized model $L_\theta$, we can choose any code for $\theta$ to make it complete. We will call the set of all such complete codes the \emph{two-part codes on $L_\theta$}. Note that we can simply concatenate the two codewords, since all codes are prefix-free.

Which code we choose for the parameter is arbitrary. We may be able to reject the uniform code for one choice of $L^\N$, or several, but how can we prove that $L^\text{complete}$ will be rejected whatever $L^\N$ we choose? Instead of choosing an arbitrary code for the size, we can use the \emph{bound} $B(G) = L^\text{uniform}_{n(G)}(G)$ as our null model. This is not a code, but it \emph{is} a lower bound for any two-part code on $L^\text{uniform}_n$. If $L^\text{clique}(G)$ is shorter than $B(G)$, it is also shorter than $L^\text{complete}(G)$ whatever the choice of $L^\N$.\footnotemark

\footnotetext{
In probabilistic terms, the code on the parameter corresponds to a prior on the parameter. The two-part codes correspond to maximum likelihood posterior probabilities: $p(\hat\theta) p(x \mid\hat\theta)$. Our bound corresponds to the maximal likelihood of the data: $p(x \mid \hat \theta)$. This shows us that the bound applies not only to the two-part codes, but also to the full Bayesian mixture: $\sum_\theta p(x \mid \theta) p(\theta) \leq \sum_\theta p(x \mid \hat\theta) p(\theta) = p(x \mid\hat\theta)$
}

% Contrast this with the traditional approach, where we would define a statistic on $G$, like the size of the largest clique, and compare the observed value of the statistic with the expectation under the null model. In this case the models would have to be rejected with separate tests. If a large clique is unlikely in a sample from $p^\text{uniform}_n$, we have no guarantee that it will also be unlikely in a sample from $p^\text{complete}$.

Note that when we store the rest of the graph within $L^\text{clique}$ we \emph{cannot} use $B(G)$ in place of $L^\text{complete}(G)$. We want a \emph{conservative} hypothesis test: the probability of incorrectly rejecting a null model may be lower than $\alpha$ but never higher. By this principle, bounds chosen in place of either model should always decrease $\Delta$. The code corresponding to the null model must always be lowerbounded, and the optimal code for the alternative model must always be upperbounded.

This is also the reason that we allow the alternative code to have multiple codewords for one graph. Replacing such a code with the complete equivalent model, found by summing over all codewords, would increase $\Delta$, strengthening the hypothesis test. By evaluating only one codeword, we are effectively using an upper bound on this complete model, trading off power for efficiency of computation.

% So, to apply this method to techniques, we need a code $L^\text{motif}$ which exploits the presence of motifs in the data to describe it succinctly. This code is explained in the next section. We will contract this with three null models described in the section \emph{Null Models}. As with the clique example, we will re-use the null model in $L^\text{motif}$ to describe any part of the graph that is not a motif.

\section{Encoding with Motifs}

\label{section:motif-code}

We will now use the principles described above to define the alternative model for the detection of a motif. The idea is that we will use a given motif, and a list of its occurrences in the data to try to find an efficient description of the data.

Let $S = \langle S_1, \ldots, S_k \rangle$ be a sequence of nodes from $V_G$. The \emph{induced subgraph} $G[S]$ is a graph  with $k$ nodes, containing a link  $(i, j)$ if and only if $G$ has a link $(S_i, S_j)$. That is, the induced subgraph extracts all links existing between members of $S$.

Assume that we are given a graph $G$, a potential motif $M$, and a list ${\cal I}^\text{raw} = \langle I_1, \ldots, I_k\rangle$ of \emph{instances} of $M$ in $G$. That is, each sequence $I\in {\cal I}^\text{raw}$ consists  of nodes in $N_G$, such that the induced subgraph $G[I]$ is equal to $M$. Note that that ${\cal I}^\text{raw}$ need not contain all instances of $M$ in the data. Additionally, sequences in ${\cal I}^\text{raw}$ may overlap, i.e. two instances may share one or more nodes. We are also provided with a generic graph code $L^\text{base}(G)$ on the simple graphs.

The basic principle behind our code is illustrated in Figure~\ref{figure:motif-code}: we want to store the motif only once, remove as many instances of the motif from the data as we can, and replace them with references to the stored motif. The two graphs combined contain enough information to recover the data, but we have only had to describe the motif once. Algorithm~\ref{algorithm:motif-code} describes the exact process.

As noted above, this is a ``best effort'' approach. We do not necessarily require an optimal encoding of the graph using the candidate motif. Indeed, depending on how the problem is framed, such an encoding would be either incomputable or prohibitively expensive to compute. We are free to trade off the efficiency of the encoding against the number of motifs revealed.

\begin{figure}[htb]
  \includegraphics[width=\textwidth]{./images/illustration.pdf}
  \caption{An illustration of the motif code. We store $M$ once, and remove its instances from $G$, replacing them with a single, special node. The links to special nodes are annotated with `rewiring' information, which tells us how to rewire the subgraph back into $H$. Storing only $H$ and $M$ is enough to reconstruct the data.}
   \label{figure:motif-code}
\end{figure}

\paragraph{Removing overlaps} The first thing we need is a subset $\cal I$ of ${\cal I}^\text{raw}$ such that the instances contained within it do not overlap: i.e. for each $I_a$ and $I_b$ in $\cal I$, we have $I_a \cap I_b = \emptyset$. To maximize the scalability of the algorithm, we will use a greedy approach to find a reasonable subset, rather than attempting to approximate an optimal solution.

As we will see later, the most important factor for compression is the number of links an instance has to nodes outside the instance. We call this the \emph{exdegree}.\footnote{Unlike the in- and outdegree, the exdegree is not a property of a node, but of a subgraph.} In order to find a subset of instances with low exdegree, we first sort ${\cal I}^\text{raw}$ by exdegree in ascending order. We then remove the first $I$, add it to our subset $\cal I$ and remove all other instances that overlap with it. We continue removing the first remaining instance until $\cal I$ is empty.

\paragraph{Encoding integers} In the following, we will often need to encode single natural numbers, or a sequence of natural number from a finite range. For single numbers, we will use the code corresponding to the probability distribution $p^\N(n) = 1/ (n(n+1))$, and denote it $L^\N(n)$.\footnotemark

\footnotetext{Note that this code is only used for storing single integers like graph dimensions, or the parameters of other codes. For long sequences of integers we use the DM code. Therefore, the specific choice of $p^\N(n)$ does not affect the codelength very much.}

For sequences of elements from a finite set, we use the code corresponding to a \emph{Dirichlet-Multinomial} (DM) distribution. Let $S$ be a sequence of length $k$ of elements from some alphabet $\Sigma$. Conceptually, the DM distribution models the following sampling process: we sample a probability vector $p$ on $[0, |\Sigma|]$ from a Dirichlet distribution with parameter vector $\alpha$, and then sample $k$ symbols from the categorical distribution represented by $p$. The probability mass function corresponding to this process can be expressed as
\begin{align*}
&p^\text{DirM}_\alpha(S\mid k, \Sigma) = \prod_{i \in [1,k]} \text{DirM}_\alpha(S_i\mid S_{1:i-1}, k, \Sigma) \\
&\text{DirM}_\alpha(S_i \mid S', k, \Sigma) = \frac{f(S_i, S') + \alpha_i}{|S'| + \sum_i \alpha_i}
\end{align*}
where $f(x, X)$ denotes the frequency of $x$ in $X$. We use $\alpha_i = 1/2$ for all $i$. Let $L^\text{DirM}_{k,\Sigma} (S) = -\log p^\text{DirM}(S \mid k, \Sigma)$. The DM model can be seen as encoding each element from $S_i$, using the smoothed relative frequency of $S_i$ in the subsequence $S_{1:i=1}$ preceding it. Thus the probability of a given symbol changes at each point in the sequence, based on how often it has been observed up to that point.

Note that this code is parametrized with $k$ and $\Sigma$. If these cannot be deduced from information already stored, they need to be encoded separately. When encoding natural numbers, we will have $\Sigma = [0, n_\text{max}]$, and we only need to encode $n_\text{max}$. A useful property of the DM code is that it is \emph{exchangeable}: if we re-arrange the elements of $S$, the codelength remains the same.

Note that, since we use $L^\N(n)$ and $L^\text{DirM}(n)$ only in the motif code, there is no need for them to be optimal. The better they compress, the more motifs we will find, but we do not require optimal results for the algorithm to be valid.

\paragraph{The motif code} We can now define the full motif code. It stores the following elements. Since each corresponds to a prefix-free code, we can simply concatenate their codewords for a prefix-free codeword for the data.
\begin{description}
\item[subgraph] First, we store the subgraph $M$ using $L^\text{base}(M)$ bits.
\item[template] We then create the \emph{template graph} $H$ by removing the nodes of each instance $I \in \cal I$, except for the first, which becomes a specially marked node, called an \emph{instance node}. The internal links of $I$---those incident to two nodes both in $I$---are removed from the graph. Any link connecting a node outside of $I$ to a node inside of $I$ is kept, and rewired to the instance node.
\item[instance nodes] $L^\text{base}$ does not record which nodes of $H$ are instance nodes, so we must record this separately. Once we have recorded how many instance nodes there are, there are $n(H) \choose |{\cal I}|$ possible placements, so we can encode this information in $L^\N (|{\cal I}|) + \log {n(H) \choose |{\cal I}|}$ bits. 
\item[rewiring] For each side of a link in $H$ incident to an instance node, we need to know which node in the motif it originally connected to. Let there be some agreed-upon order in which to enumerate the links of any given graph. Given this order, we only need to encode the sequence $W$ of integers $w_i \in [1,\ldots, n(M)]$. We do so using the DM model described above. The maximum symbol and length of $W$ can be deduced from parts already encoded. Note that this code is invariant to the ordering of $W$, so the particulars of the canonical node ordering do not need to be specified.
\item[multiple edges] Since $L^\text{base}$ can only encode simple graphs, we cannot use it to store $H$ directly, since collapsing the instances into single nodes may have created multiple edges. We remove all multiple edges and encode them separately. We assume a canonical ordering over the links and record for each link incident to an instance node, how many copies of it were removed. This gives us a sequence $R$ of natural numbers $R_i \in [0, r_\text{max}]$ which we store by first recording the maximum value in $L^\N(\max(R))$ bits, and then recording $R$ with the DM model.
\item[insertions] Finally, while $H$ and $M$ give us enough information to recover a graph isomorphic to $G$, we cannot yet reconstruct where each node of a motif instance belongs in the node ordering of $G$. Note that the first node in the instance became the instance node, so we only need to record where to insert the rest of the nodes of the motif. This means that we perform $|{\cal I}| (n(M)-1)$ such insertions. Each insertion requires $\log (t+1)$ bits to describe, where $t$ is the size of the graph before the insertion. We require $\sum_{t=n(H)}^{n(G)-1} \log (t+1) = \log (n(G)!) - \log (n(H)!)$ bits to record the correct insertions.
\end{description}

\begin{pseudo}[th]
\caption{The motif code $L^\text{motif}(G ; M, {\cal I}, L^\text{base})$. Note that the nodes of the graph are integers.}
\label{algorithm:motif-code}
{ 
Given:\\ 
\tab a graph $G$, a subgraph $M$,\\ 
\tab a list $\cal I$ of instances of $M$ in $G$, a code $L^\text{base}$ on the simple graphs.\\
\\
$b_\text{subgraph} \leftarrow L^\text{base}(M)$\hfill \textbf{subgraph} \\
\\
\emph{\# replace each instance with a single node} \\
$H \leftarrow \text{copy}(G)$, $W = [] $\hfill \textbf{template} \\
\textbf{for each } $I = \{ m_1, \ldots m_{n(M)}\}$ \textbf{in} $\cal I$:\\
\tab \emph{\# We use $m_1$ (the $m_1$-th node in $G$) as the instance node}\\
\tab \textbf{for each} link $l$ between a node $n_\text{out}$ not in $I$ and a node $m_j$ in $I$:\\
\tab \tab \textbf{if} $j \neq 1$: add a link between $n_\text{out}$ and $m_j$\\
\tab \tab $W$.append$(j)$\\
\tab remove all nodes $m_i$ except $m_1$, and all incident links\\
$b_\text{rewiring} \leftarrow  L^\text{DirM}_{|W|, n(M)}(W)\hfill\textbf{rewiring}$\\
\\
\emph{\#Remove multiple edges from $H$  and record the duplicates in $R$}\\
$R, H' \leftarrow \text{simple}(H)$ \\
$b_\text{template} \leftarrow L^\text{base}(H')$\\ 
$b_\text{multi-edges} \leftarrow L^\N(\max(R)) + L^\text{DirM}_{|R|, \max(R)}(R)$ \hfill \textbf{multiple edges}\\
\\
$b_\text{instances} \leftarrow L^\N (|{\cal I}|) \log {n(H) \choose |{\cal I}|}$ \hfill \textbf{instance nodes}\\
$b_\text{insertions} \leftarrow \log (n(G))! - \log (n(H))!$ \hfill \textbf{insertions}\\    
\\
\textbf{return} $b_\text{subgraph} + b_\text{template} + b_\text{rewiring} + b_\text{multi-edges} + b_\text{instances} + b_\text{insertions}$\\
}
\end{pseudo}

\paragraph{Pruning the list of instances} Since our code accepts any list of motif instances, we are free to take the list $\cal I$ and remove instances before passing it to the motif code, effectively discounting instances of the motif. This can often improve compression, as storing the rewiring information for instances with high exdegrees may cost more than we gain from removing them from the graph. We sort  $\cal I$ by exdegree and search for the value $c$ for which compressing the graph with only the first $c$ elements of $\cal I$ gives the lowest codelength.

The codelength $L^\text{motif}$ as a function of $c$ is roughly unimodal, which means that a ternary search should give us a good value of $c$ while reducing the number of times we have to compute the full codelength. We use a \emph{Fibonacci search} \citep{kiefer1953sequential}, an elegant variation on ternary search requiring only one sample per recursion.

\paragraph{Implementation} The \textbf{template} part of the code can be time and memory intensive to compute for large graphs, as it involves creating a copy of the data. For any given $L^\text{base}$, we can create a specific implementation which computes the codelength required for storing the template graph without constructing $H$ explicitly. This will speed up the computation of the code at the expense of creating a new implementation for each new null model. We use such specific implementations for our three null models.

More precisely, since we would like to operate efficiently on very large graphs, with a relatively low number of motif instances, we would like to avoid, where possible, operations requiring a full pass over the entire graph. We do this by storing the parameters of the null model (either the degree sequence or the size and number of links) for the complete graph, and computing how they change when the template graph is created. We can do this with a loop over the motif instances, computing directly how the model parameters change. Since the parameters determine the codelength, this is all we need to determine the number of bits required to store the template graph. In this computation we must make sure also to compute which links of the template graph become multiple links, and how many external links connect to each node within the motif.

\section{Null Models}

\label{section:null-models}
We will define three null models. For each model we follow the same pattern: we first describe a parametrized model (which does not represent a code on all graphs). We then use this to derive a bound as described in Section~\ref{section:multiplerejection}, so that we can reject a set of null models, and finally we describe how to turn the parametrized model into a complete model to store graphs within the motif code.

Specifically, let $L^\text{name}_\theta(G)$ be a parametrized model with parameter $\theta$. Let $\hat\theta(G)$ be the value of $\theta$ that minimizes $L^\text{name}_\theta(G)$ (the maximum likelihood parameter). From this we derive a bound $B^\text{name}(G)$---usually using $B^\text{name}(G) = L^\text{name}_{\hat\theta(G)}(G)$---which we will use in place of the null model. Finally, we create the complete model by two-part coding: $L^\text{name}(G) = L^{\theta}(\hat\theta(G)) + L^\text{name}_{\hat\theta(G)}(G)$.

\subsection{The Erd\H{o}s-Renyi Model}
\label{section:null models}

The Erd\H{o}s-Renyi (ER) model is probably the best known probability distribution on graphs \citep{renyi1959random,gilbert1959random}. It takes a number of nodes $n$ and a number of links $m$ as parameters, and assigns equal probability to all graphs with these attributes, and zero probability to all others. This gives us 
\[
L^\text{ER}_{n, m}(G) = \log{(n^2-n)/2 \choose m} 
\] 
for undirected graphs, and 
\[
L^\text{ER}_{n, m}(G) = \log{n^2-n \choose m} 
\] 
for directed graphs. We use the bound $B^\text{ER}(G) = L^\text{ER}_{n(G), m(G)}(G)$.

For a complete code on simple graphs, we encode $n$ with $L^\N$. For $m$ we know that the value is at most $m_\text{max}=(n^2-n)/2$ in the undirected case, and at most $m_\text{max}=n^2-n$ in the directed case, and we can encode such a value in $\log (m_\text{max} + 1)$ bits ($+1$ because $0$ is also a possiblity). This gives us:
\[
L^\text{ER}(G) = L^\N(n(G)) + \log (m_\text{max} + 1) + L^\text{ER}_{\theta}(G)\;\text{with}\;\theta=(n(G),m(G))\p
\]

\subsection{The Degree-Sequence Model}
\label{section:degree-sequence-model}

The most common null model in motif analysis is the \emph{degree-sequence model}, also known as the \emph{configuration model} \citep{newman2010networks}. 
For undirected graphs, we define the degree sequence of graph $G$ as the sequence $D(G)$ of length $n(G)$ such that $D_i$ is the number of links incident to node $i$ in $G$. For directed graphs, the degree sequence is a pair of such sequences $D(G) = (D^\text{in}, D^\text{out})$, such that $D^\text{in}_i$ is the number of incoming links of node $i$, and $D^\text{out}_i$ is the number of outgoing links.

\paragraph{The parametrized model $L^\textnormal{DS}_D(G)$} The degree-sequence model $L^\text{DS}_D(G)$ takes a degree sequence $D$ as a parameter and assigns equal probability to all graphs with that degree sequence. Assuming that $G$ matches the degree sequence, we have $L^\text{DS}_D(G) = \log |\cG_D|$ where $\cG_D$ is the set of simple graphs with degree sequence $D$. There is no known efficient way to compute this value for either directed or undirected graphs, but various estimation procedures exist. We use an importance sampling algorithm discovered independently by \cite{blitzstein2011sequential} and \cite{charo2010efficient}.\footnotemark~This algorithm is guaranteed to produce any graph matching $D$ with some nonzero probability. Crucially, the algorithm does not backtrack or reject candidates, which means that if we multiply the probability of each random choice made in sampling, we get the probability of the sample under our sampling procedure. That is, the algorithm produces, along with a sample $G \in \cG_D$, the probability $q^\text{DS}_D(G)$ of the algorithm producing $G$. While the samples are not uniform, we do have
\begin{equation}
{E} \left [\frac{1}{q^\text{DS}_D(G)}\right] = |\cG_D| \label{line:expectation}
\end{equation}
where $G$ is a random variable representing a sample from the algorithm. Thus, we can sample a number of graphs and take the mean of their inverse probability under $q^\text{DS}_D$ to estimate $p^\text{DS}_D(G)$. This is a form of \emph{importance sampling}.

\footnotetext{Specifically, our implementation uses the algorithms described in \cite{charo2010efficient} and \cite{kim2012constructing}. However the non-uniform sampling from the candidate set, discussed in \cite[p10, step 5]{blitzstein2011sequential} is crucial to achieving a low variance in the sampling distribution, and thus a fast convergence.}

Unfortunately, even with the highly optimized implementations described in \cite{charo2010efficient} and \cite{kim2012constructing} sampling can be slow for large graphs. Luckily, we are only interested in an estimate of the codelength accurate to around the level of single bits, which means that we only need to sample until we have a rough estimate of the order of magnitude of $|\cG_D|$. For instance, if we accept a margin of error of only 15 bits (of the potentially $10^6$ bits required to store the graph), we can underestimate the number of graphs by 4 orders of magnitude and still end up within the margin. All we need is a reliable confidence interval for our estimate, so that we can choose a suitably conservative bound. Our method of obtaining such a confidence interval is described in the appendix. In all cases, we use a one-sided confidence interval: when computing the codelength under the null model, we use a lower bound for the true value, and when computing the codelength for the motif code, we use an upper bound. Thus, the difference in codelength is a lower bound for the true value.

\paragraph{The bound $B^\textnormal{DS}(G)$} To get a bound for all two-part codes on $L^\text{DS}_D$, we could use ${B'}^\text{DS}(G) = L^\text{DS}_{D(G)}(G)$. Beating such a bound would tell us that no property of the degree sequence could explain the motif we had found. Unfortunately, the degree sequence forms a large part of the code, and a lot of evidence is required to compress better than ${B'}^\text{DS}(G)$ with a complete code.

Instead, we make the assumption that the degrees are sampled independently from a single distribution $p^\text{deg}(n)$ on the the natural numbers. This corresponds to a code $\sum_{D_i \in D}L^\text{deg}(D_i)$ on the entire degree sequence. Let $f(s, D)$ be the frequency of symbol $s$ in sequence $D$. It can be shown that $B^\text{deg}(D) = \sum_{D_i \in D} f(D_i, D)/|D|$ is a lower bound for any such code on the degree sequence. This gives us the bound $B^\text{DS}(G) = B^\text{deg}(D(G))) + L^\text{DS}_{D(G)}(G)$. For directed graphs, we use $B^\text{DS}(G) = B^\text{deg}(D^\text{in}(G))) + B^\text{deg}(D^\text{out}(G))) + L^\text{DS}_{D(G)}(G)$.

Note that the \emph{scale-free} property of many complex graphs is captured by this model. Compressing better than $B^\textnormal{DS}(G)$ with a certain motif $M$, indicates that the occurrences of $M$ are not explained by the scale-freeness of the graph.

\paragraph{The complete model $L^\textnormal{DS}(G)$} For the alternative model we need a complete code. First, we store $n(G)$ with $L^\N$. We then store the maximum degree and encode the degree sequence with the DM model. For undirected graphs we get: 
\[
L^\text{DS}(G) = L^\N(n(G)) + L^\N(d) + L^\text{DirM}_{\theta}(D) + L^\text{DS}_{D(G)}(G) \;\text{with}\; \theta = \left(n(G), max(D)\right)
\]
and for directed graphs
\begin{align*}
L^\text{DS}(G) = &L^\N(n(G)) \\ 
 + &L^\N(\max(D^\text{in})) + L^\text{DirM}_{\theta}(D^\text{in})\\
 + &L^\N(\max(D^\text{out})) + L^\text{DirM}_{\phi}(D^\text{out}) + L^\text{DS}_{D(G)}(G) \\
 \text{with}&\; \theta = \left(n(G), \max(D^\text{in})\right), \phi = \left(n(G), \max(D^\text{out})\right) 
\end{align*}

Note that in the computation of $L^\text{motif}$ with $L^\text{DS}$ as a base model, we estimate $|\cG_D|$ for both the template graph and the motif. It is important to combine the confidence intervals over these two estimates carefully, so that we end up with a correct confidence interval over the total codelength. This is discussed in the supporting materials. For $L^\text{motif}$, we compute a one-sided confidence interval to get an \emph{upper}bound, so that with 95\% confidence we are \emph{over}estimating the size of the motif code.

\subsection{The Edgelist Model}

While estimating $|\cG_D|$ can be costly, we can compute an upper bound efficiently. Assume that we have a directed graph $G$ with $n$ nodes, $m$ links and a pair of degree sequences $D = (D^\text{in}, D^\text{out})$. To describe $G$, we might write down the links as a pair of sequences $(F, T)$ of nodes: with  $F_i$ the node from which link $i$ originates, and $T_i$ the node to which it points. Let $\cS_d$ be the set of all pairs of such sequences satisfying $D$. We have $ v$ possibilities for the first sequence, and $m \choose D_1^\text{out}, \ldots, D_n^\text{out}$ for the second. This gives us $|\cS_D| = {m \choose D_1^\text{in}, \ldots, D_n^\text{in}}{m \choose D_1^\text{out}, \ldots, D_n^\text{out}} = m!m! / \prod_{i=1}^n D^\text{in}_i ! D^\text{out}_i !$. We have $|\cS_D| > |\cG_D|$ for two reasons. First, many of the graphs represented by such a sequence pair contain multiple links and self-loops, which means they are not in $\cG_D$. Second, the link order is arbitrary: we can interchange any two different links, and we would get a different pair of sequences, representing the same graph, so that for a graph with no multiple edges, there are $m!$ different sequence-pairs to represent them.

To refine this upper bound, let $\cS'_D \subset \cS_D$ be the set of sequence pairs representing simple graphs. Since all links in such graphs are distinct, we have $|\cG_D| = |\cS'_D|/m!$. Since $|\cS'_D| \leq |\cS_D|$, we have \footnotemark
\[
|\cG_D| \leq \frac{m!}{\prod_{i=1}^n D^\text{in}_i ! D^\text{out}_i !} \p
\]

\footnotetext{This value was previously used in \cite{bezakova2006graph} as a precise value for the number of graphs with multiple edges. This is incorrect, as we can only divide by $m!$ if we know that no graphs have multiple edges.}

In the undirected case, we can imagine a single, long list of nodes of length $2m$. We construct a graph from this by connecting the node at index $i$ in this list to the node at index $m+i$ for all $i \in [1, m]$. In this list, node $a$ should occur $D_a$ times. We define $\cS_D$ as the set of all lists such that the resulting graph satisfies $D$. There are $(2m)! \choose D_1, \ldots, D_n$ such lists.
We now have an additional reason why $|\cS_D| > |\cG_D|$: each pair of nodes describing a link can be swapped around to give us the exact same graph. This gives us:
\[
|\cG_D| \leq |\cS'_D| / (2^m m!) = \frac{(2m)!}{2^m m! \prod_{i=1}^n D_i!} \p
\]

In both cases, the fact that we have an upperbound gives us a code: while the code as described assigns some probability mass to non-simple graphs, we can easily assume that this is assigned instead to some null-element, since we are only interested in the codelengths and probabilites of simple graphs. This gives us the following parametrized code for directed graphs:
\[
L^\text{EL}_D(G) = \log m! - \sum_{i=0}^n \log D_i^\text{in}! - \sum_{i=0}^n \log D_i^\text{out}!   
\]
where $(D^\text{in}, D^\text{out})$ are the degree sequences of $G$, and for the undirected case:
\[
L^\text{EL}_D(G) = \log (2m)! - \log m! - m - \sum_{i=0}^n \log D_i! \p   
\]

For the bound and the complete model, we follow the same strategy we used for the degree-sequence model: $B^\text{EL}(G) = B^\text{deg}(G) + L^\text{EL}_{D(G)}(G)$ and, 
\[
L^\text{EL}(G) = L^\N(n(G)) + L^\N(\max(D)) + L^\text{DirM}_{\theta}(D) + L^\text{EL}_{D(G)}(G) \;\text{with}\; \theta=(n(G), \max(D))
\] for undirected graphs and 
\begin{align*}
L^\text{EL}(G) = &L^\N(n(G)) \\ 
 + &L^\N(\max(D^\text{in})) + L^\text{DirM}_{\theta}(D^\text{in})\\
 + &L^\N(\max(D^\text{out})) + L^\text{DirM}_{\phi}(D^\text{out}) + L^\text{EL}_{D(G)}(G)\\
 \text{with}&\; \theta=(n(G), \max(D^\text{in})), \phi=(n(G), \max(D^\text{out})) \
\end{align*} for directed graphs.

\section{Experiments}

\label{section:experiments}

To validate and illustrate our method, we will perform four experiments. First, we will construct a graph by injecting instances of a single motif into a random network. The method should then recover only this motif as significant. Second, we will run the method on data sets from four different domains, and show the results for the most frequent subgraphs, using the three null models we have described. Third, we use the motifs identified by our method, and by the traditional approach, as features in a classification task, showing comparable performance, well above chance. This tells us that, at least for the purposes of classification, the motifs returned by our method are no less informative than those found by the traditional method. Finally, to show the scalability of the method with fast null models, we will run the analysis on several large graphs, first using purely memory based methods, and then with a disk-backed datastructure to store the graph.

In all experiments we search for motifs by sampling, based on the method described by \citet{kashtan2004efficient}. Note that we have no particular need for a sampling algorithm which provides an accurate approximation of the actual frequencies present in the graph, so long as it can provide us with a large selection of non-overlapping instances with low exdegree. For this reason, we adapt the algorithm to improve its speed: we start with a set $N'$ containing a single random node drawn uniformly. We then add to $N'$ a random neighbour of a random member of $N'$, and repeat until $N'$ has the required size. We extract and return $G[N']$. In the case of a directed graph, nodes reachable by incoming and outgoing links are both considered neighbours.

The size $n(M)$ of the subgraph is chosen before each sample from a uniform distribution over the interval $[n_\text{min}, n_\text{max}]$. This distribution over sizes is biased towards small motifs: since there are fewer connected graphs for small sizes, small graphs are more likely to be sampled. As the results show, however, the method still finds motifs with many nodes, so we opt for this simple, ad-hoc method.

We re-order the nodes of the extracted graph to a canonical ordering for its isomorphism class, using the Nauty algorithm \citep{mckay1981practical}. We maintain a map from each  subgraph in canonical form to a list of instances found for the subgraph. After sampling is completed, we end up with a set of potential motifs and a list of instances for each, to pass to the motif code described in Section~\ref{section:motif-code}.

In all experiments, we report the \emph{log-factor}: $B^\text{null}(G) - L^\text{motif}(G; M, {\cal I}, L^\text{null})$. That is, we use the bound in place of the null model, and the complete code of the same null model is used in the motif code (to store the template graph and the motif). If the log-factor is larger than 10 bits, we can interpret it, as described in Section~\ref{section:model-selection}, as a successful hypothesis test, allowing us to reject the null model at $\alpha=0.001$. In all cases, a negative log-factor means that we do not have sufficient evidence to reject the null model, but a different experiment might yet achieve a positive log-factor. This could be achieved by sampling more subgraphs, using a different algorithm to find motif instances or taking more samples from the degree-sequence estimator.

All experiments in this paper were run on a single machine with a 2.60 Ghz Intel Xeon processor (E5-2650 v2) with 64 Gigabytes of memory and 8 physical cores. The memory and cores available to the program differ per experiment and are reported where relevant.

\subsection{Recovering Motifs from Generated Data}

\label{section:recovering}

We use the following procedure to sample an undirected graph with $5000$ nodes and $10000$ links, containing $k$ injected instances of a particular motif $M$ with $n'$ nodes and $m'$ links. Let $M$ be given (in our experiment $M$ is always the graph indicated in red in Figure~\ref{figure:plot-synthetic}).

\begin{enumerate}
  \item Let $n = 5000 - (n'-1)k$ and $m = 10000 - m'k$ and sample a graph $H$ from the uniform distribution over all graphs with $n$ nodes and $m$ links.   
  \item Label $k$ random nodes, with degree 5 or less, as instance nodes.
  \item Let $p^\text{cat}$ be a categorical distribution on $\{1, \ldots, 5\}$, chosen randomly from the uniform distribution over all such distributions.
  \item Label every connection between an instance node and a link with a random value from $p^\text{cat}$. Links incident to two instance nodes, will thus get \emph{two} values.
  \item Reconstruct the graph $G$ from $M$ and $H$.
\end{enumerate}

\begin{figure*}[tb]
  \hspace{-0.1\textwidth}
  \includegraphics[width=1.2\textwidth]{./images/synthetic-plot.pdf}
  \caption{\small The results of the experiment on generated data. The bottom row shows all 21 simple connected graphs with 5 nodes (up to isomorphism). The middle row shows the number of non-overlapping instances found by the sampling algorithm for $k = 0$, $k=10$ and $k=100$ from left to right, for each motif. The bars show the average value over 10 randomly sampled graphs, with the same subgraph (shown in red) injected each time. The top row shows the difference between  the code length under the null model (the ER model) and under the motif code. Error bars represent the \emph{range}, i.e. they are drawn from the smallest to the largest observation.}
  \label{figure:plot-synthetic}
\end{figure*}

This is roughly similar to sampling from our motif code. In this graph, $M$ should be the only significant motif, with the exception of motifs that can be explained from the prevalence of $M$, ie. subgraphs and supergraphs of $M$, or graphs that contain part of $M$. However, these should have a markedly lower log-factor than $M$. For our experiment, we will only extract subgraphs of size 5, to rule out the first two cases.

On this sampled graph, we run our motif analysis. We run the experiment multiple times, with $k = 0$, $k = 10$ and $k = 100$, using the same subgraph $M$ over all runs, but sampling a different $H$ each time. For each value of $k$, we repeat the experiment 10 times. Per run we sample 5000 motifs. This value is chosen to show that even a very \emph{low} sample size is sufficient to recover the motif. The null model in all cases is the ER model, as that corresponds to the source of the data.

Figure~\ref{figure:plot-synthetic} shows the results for the 21 possible connected simple graphs of size 5. This result shows that, on such generated data, the method behaves as expected in the following ways:
\begin{itemize}
	\item If no motifs are injected---that is, the graph is sampled from the null model---no subgraphs are motifs.
	\item For a small number of injected motifs ($k = 10$), the correct motif is given a positive log-factor. Other subgraphs are shown to have very high frequencies, but the method correctly gives these a negative log factor. In other words, these are shown to be natural consequences of the null model.
	\item If the number of motifs is high ($k = 100$), the resulting significance increases. Note that in the traditional method, a such high significances require many more samples from the null model.
\end{itemize}

We can also see that once we insert 100 instances of the motif, two other subgraphs ``become motifs'': in both cases, these share a part of the inserted motif (a rectangle and a triangle). This is an important lesson for motif analysis: not every significant subgraph represents a meaningful result, some may be a byproduct of other motifs. This effect is not unique to our method, but occurs in all motif analysis.

\subsection{Motifs from Real-World Data}
\label{section:various}

Next, we show how our our approach operates on a selection of data sets across domains. Our main aim with this experiment is to show how the three null models influence the results. Specifically, to ascertain whether the edgelist model provides a reasonable approximation for the degree-sequence model.

We use the following data sets:
\begin{description}
\item[kingjames (undirected, $n=1773, m=9131$)] Co-occurrences of nouns in the text of the King James Bible \citep{konect:2014:moreno_names,konect:harrison}. Nodes represent nouns (places and names) and links represent whether these occur together in one or more verses. The full motif analysis took 23 hours and 16 minutes. 
\item[yeast (undirected, $n=1528, m=2844$)] A network of the protein interactions in yeast, based on a literature review \citep{reguly2006comprehensive}. Nodes are proteins, and links are reported interactions between proteins. We removed 81 self-loops. The full motif analysis took 3 hours and 19 minutes. 
\item[physicians (directed, $n=241, m=1098$)] Nodes are physicians in Illinois \citep{konect:2015:moreno_innovation,konect:coleman1957}. Links indicate that one physician turns to the other for advice. The full motif analysis took 31 minutes. 
\item[citations (directed, $n=1769, m=4222$)] The arXiv citation network in the category of theoretical astrophysics, as created for the 2003 KDD Cup \citep{gehrke2003overview}. To create a workable graph, we follow the procedure outlined in \citet{carstens2013motifs}: we include only papers published before 1994, remove citations to papers published after the citing paper, and select the largest connected component. The full motif analysis took 8 hours and 53 minutes. 
\end{description}
All data sets are simple (no multiple edges, no self-loops). In each case we take $5 \cdot 10^6$ samples with $n_\text{min} = 3$ and $n_\text{max} = 6$. We test the 100 motifs with the highest number of instances (after overlap removal), and report the log-factor for each null model. For the edgelist and ER models we use a Fibonacci search at full depth, for the degree-sequence model we restrict the search depth to $3$. For the degree-sequence estimator, we use $40$ samples and $\alpha=0.05$ to determine our confidence interval. We use the same set of instances for each null model.

\begin{figure*}[p]
  \includegraphics[width=\textwidth]{./images/kingjames/compare-plot.pdf}\\
  \includegraphics[width=\textwidth]{./images/yeast/compare-plot.pdf}\\
  \caption{The results of the motif extraction on the 2 undirected networks.}
  \label{figure:plot-und}
\end{figure*}

\begin{figure*}[p]
  \includegraphics[width=\textwidth]{./images/physicians/compare-plot.pdf}\\
  \includegraphics[width=\textwidth]{./images/citations/compare-plot.pdf}\\
  \caption{The results of the motif extraction on the 2 directed networks.}
  \label{figure:plot-dir}
\end{figure*}

The negative log-factors in some of the results may appear surprising at first, since the motif code should never do worse than the null model by more than a few bits: in the worst case, it can simply prune all motif instances, and use the null model for the entire graph. The negative values shown here are due to the use of a \emph{bound} for the parameters: the motif code must store the degree sequence explicitly, while the null model uses a lower bound for that part of the code.

Our first observation is that for the physician data set, there are no motifs under the degree-sequence null model. This likely because the physicians network is too small: the use of a bound for the null model means that the alternative model requires a certain amount of data before the differences become significant. Note, however, that if we were to compare against a complete model (instead of the bound), a constant term would be added to all compression lengths under the null model. In other words, the ordering of the motifs by relevance would remain the same.

In both the kingjames and the yeast graphs, many motifs contain cliques or near-cliques. This suggests that the data contains local communities of highly interconnected nodes which the null model cannot explain.

Finally, we would like to emphasize the scale of the null-hypothesis rejections. In the kingjames test, the most significant motif under the DS model has a log-factor of over 3500 bits. This corresponds to a rejection of the null model at significance $\alpha = 2^{-3500} \approx 10^{-1050}$. Note that accurate estimates of strong significances are a common problem in the traditional approach, since this requires very high numbers of samples from the null model \citep{picard2008assessing}.

For the experiments in this section, the maximum Java heap space was set to 2 Gigabytes. The computation of the log-factor of each motif was done in parallel, as was the sampling for the degree sequence model, with at most 16 threads runnning concurrently, taking advantage of the 8 available (physical) cores.

Now, clearly, these analyses took relatively long, for data sets that are of medium size. The bottleneck here is the computation of the degree-sequence model. If we eliminate that, as we do in Section~\ref{section:large}, we see that we can run the same analysis in minutes on graphs that are many orders of magnitude larger than these. Moreover, the plots show a reasonable degree of agreement between the EL model and the DS model, suggesting that the former might make an acceptable proxy. But are the motifs returned still, in some sense, informative? We investigate this in the next section, comparing the traditional method of motif detection to our method, using the EL model.

\subsection{Comparison with the traditional method}

\label{section:classification}

\begin{figure}[tbh]
{
  \includegraphics[width=\textwidth]{./images/experimentexplanation.pdf}	
}

  \caption{A schematic illustration of the classification experiment. (a) We start with a classification task on simple, undirected graphs. (b) These graphs are reduced to 29-dimensional binary feature vectors. Each feature corresponds to an undirected, connected mini-graph of size 3, 4, or 5. If the motif algorithm considers the mini-graph a motif for the current instance, the feature is 1, otherwise it is 0. This reduces the entire graph to just 29 bits of information, using only the judgements of the motif algorithm to represent the graph. If these judgements do indeed contain some information that characterises the graph, these feature vectors should be sufficient to perform better than a majority class classifier. (c) We apply a simple, linear SVM in the 29-dimensional feature space to perform the classification task on just the motif judgments.}
  \label{figure:experiment-explanation}
\end{figure}

Pattern mining on graph data is a very difficult task to evaluate. Not only is it unsupervised, we also have no ground truth data for what people might consider interesting subgraphs. Often, we do not even know what such a ground truth would look like: even a domain expert may not recognize a ``correct'' motif at first sight. Nevertheless, we have rather stretched the definition of a motif---choosing a different statistic, and an approximation to the default null model---so we need some assurance that after all this, the motifs found by our method are not completely meaningless.

The definition of what constitutes a good motif is exceedingly vague: papers variously describe a motif as a ``functional unit'', a ``characteristic pattern'' or a ``statistically significant subgraph.'' To be able to operationalize the definition into something that we can test empirically, we will define a network motif as a subgraph that is \emph{characteristic} for the full graph. That is, in some manner, the information that $M$ is a motif for $G$ should \emph{characterize} $G$: it distinguishes $G$ from the graphs for which $M$ is not a motif, and that distinction should be meaningful in the domain of the data.

Using this definition, we can test how effective a motif detection method is, using a graph classification task. We start with a set of undirected simple graphs, with associated classes. We then translate each graph into a binary feature vector using \emph{only the motif judgements} of the algorithm under evaluation. We test all connected subgraphs of size 3, 4, and 5, giving us 29 binary features. As we established in the previous paragraph, a good motif characterizes a graph in a way that is relevant to its domain. In the case of a graph classification task we are provided with a domain-relevant distinction in the form of the classes provided. If a simple classifier (in our case a linear SVM) can classify the graphs purely on the basis of these 29 motif judgments, the algorithm has succeeded in characterizing the graph. This approach---quantifying unsupervised pattern extraction through classification---was also used by \citet{leeuwen2006compression}.

Our main aim is to establish that the resulting classifier performs better than chance. For this purpose we compare it to a majority class baseline. Our secondary aim is to show that we do not perform much worse than the traditional method. Note that our method allows motif analysis to scale up by at least four orders of magnitude, so we consider a small reduction in classification accuracy acceptable, so long as the method still significantly outperforms the majority class baseline.

Finding a graph classification task that fits our requirement is not easy: the graphs must be large enough to provide our method with enough data,\footnotemark~but small enough that the traditional method works without approximation. Moreover, the graphs must be undirected, so that we can use the highly efficient ORCA algorithm \citep{hovcevar2014combinatorial} to get exact subgraph counts for the traditional method. In order to tune the graph classification tasks to our needs, we adapt them from classification tasks on knowledge graphs, described in \cite{ristoski2016collection}. In these tasks, the data consists of a single labeled, directed multigraph, and the task is to predict classes for a specific subset of nodes (the \emph{instance nodes}). We translate the graph to an unlabeled simple graph by using the same nodes (ignoring their labels) and connecting them with a single undirected edge only if there are one or more directed edges between them in the original knowledge graph.

\footnotetext{For smaller graphs our method can be adapted by testing against a single model, instead of using a bound to reject multiple null models as described in Section~\ref{section:model-selection}. However, for this test, we would like to evaluate the method as we expect it to be used.}

This gives us a classification task on the nodes of a single, undirected simple graph. We turn this into a classification task on \emph{separate} graphs by extracting the 3-neighborhood around each instance node. To control the size of the extracted neighborhoods, we remove the $h$ nodes with the highest degrees from the data before extracting the neighborhoods. $h$ was chosen by trial-and-error, before seeing the classification performance, to achieve neighborhoods with around 1000--2000 nodes.

We now have a graph classification task from which we can create feature vectors as described above. For our method, we sample 100 000 subgraphs, with size 3, 4, 5 having equal probability and test the compression levels under the edgelist model. We judge a subgraph to be a motif if it beats the EL bound by more than $- \log \alpha$ bits with $\alpha = 0.05$.

Many methods for motif analysis have been published, but most are approximations or more efficient counting algorithms. Therefore, a single algorithm based on exact counts can act as a representation for most of the existing approaches. We perform exact subgraph counts on both the data and 1000 samples from the DS model. The samples from the null model are taken using the Curveball algorithm \citep{strona2014fast,carstens2016curveball}. We estimated the mixing time to be around 10 000 steps, and set the run-in accordingly. The subgraph counts were performed using the ORCA method.\footnote{We created a Java implementation, available at \url{https://github.com/pbloem/orca}.} We mark a subgraph as a motif if fewer than 5\% of the graphs generated from the DS model have more instances of the subgraph than the data does.\footnotemark~

\footnotetext{Note that the commonly used z-score method is seriously flawed, as discussed by \citet{picard2008assessing}, so we do not use it here.}

For performance reasons, we use only 100 randomly chosen instances from the classification task. On these 100 instances, we perform five-fold cross-validation.
We repeat the complete experiment, from sampling instances to cross-validation, 10 times. The classifier is a linear SVM ($C=1$). For tasks with more than 2 classes, the one-against-one approach \citep{knerr1990single} is used.

Note that the accuracy should not be compared to that achieved by others on these benchmarks, as we have thrown away almost all information in the original knowledge graph. We are merely using classification accuracy as a proxy for the quality of motif analysis.

\begin{figure}[tbh]
{
  \begin{center}
  \includegraphics[width=0.8\textwidth]{./images/bars.pdf}
	\begin{tabular}{ r | r r r r r r}
		data & \# nodes & \# links & $h$ & $K$ & $\overline{n}$ & $\overline{m}$ \\
		\hline
		AIFB & 8 275 & 17 911 & 10 & 4 & 1877.11 & 7141.48 \\
		AM   & 1 495 566 & 2 393 604 & 3 000 & 11 & 2506.07 & 4392.06 \\
		BGS  & 333 613 & 362 627 & 250 & 2 & 3097.47 & 4404.49 \\
		\hline
	\end{tabular}
	\end{center}	
}

  \caption{The results of the classification experiment for the traditional method (`counting') and ours (`motive'). The bars show the mean accuracy of ten runs (with five-fold cross-validation performed within each run. The error bars show a 95\% confidence interval (assuming normality). The baseline shows the performance of a majority-vote classifier which ignores the features.  The table shows the size of the data, the average size and number of links ($\overline{n}$, $\overline{m}$) of the instance graphs, the number of classes $K$ and the number $h$ of hubs removed.}
  \label{figure:classification}
\end{figure}

The results are shown in Figure~\ref{figure:classification}. For one data set, our method is significantly better, for another, the traditional approach is significantly better, and for one, the difference is not significant. While the performance of neither method is stellar, the fact that both beat the baseline significantly, shows that at the very least, the motifs contain \emph{some} information about the class labels of the instance represented by the graph from which the motifs were taken.

\subsection{Large-Scale Motif Extraction}
\label{section:large}

\begin{threeparttable}[tbhp]
\begin{tabular}{ l  r r r r r r r r r }
data & disk & $n$ & $m$ & $|\cal I|$ & mem. & $t$ & preload & search & motifs \\ 
\hline
wiki-nl\tnote{a} & & 1 M & 13 M & 3--6 & 16 Gb & 16 & & 7m & 8 \\
		&  &  &  & 3--6 & 5 Gb & 16 & & 13m & 8  \\
		&  &  &  & 3--6 & 2 Gb & 1 & & 25m & 8  \\
		&  &  &  & 10 & 11 Gb & 1 & & 2h 41m & 0 \\
		& \checkmark  &  &  & 3--6 & 1 Gb & 1 & 8m & 1h 30m &  8 \\
\hline
wiki-en\tnote{b} & \checkmark & 12 M & 378 M & 3--6 & 2 Gb & 1 & 4h 58m & 6h 6m & 10 \\
 & \checkmark & & & 8 & 8 Gb & 1 & & 6h 5m & 23 \\
 \hline
twitter\tnote{c} & \checkmark  & 53 M & 1 963 M & 3--6 & 6 Gb & 1 & 17h 12m & 33h 19m & 0\\
 & \checkmark  &  &  & 7 & 8 Gb & 1 & & 54h 26m & 0 \\
\hline
friendster\tnote{d} & \checkmark  & 68 M & 2 586 M & 3--6 & 6 Gb & 1 & 42h 37m & 45h 2m & 68 \\ 
&\checkmark  & & & 3--6 & 56 Gb & 9 & & 8h 38m & 68 \\ 
&\checkmark  & & & 10 & 7 Gb & 1 & & 35h 7m & 57 \\ 
\hline
\end{tabular}
\begin{tablenotes}
\item[a]\citet{konect:2016:link-dynamic-nlwiki,konect:unlink}, multiple links were removed.
\item[b]\citet{konect:2016:dbpedia-link,konect:dbpedia2}, self-loops were removed.
\item[c]\citet{konect:2016:twitter,konect:twitter1}
\item[d]\citet{konect:2016:friendster}
\end{tablenotes}

\caption{The results of various runs of the algorithm on large data sets. Sizes are rounded to the nearest million. The second column indicates whether the graph was stored on disk, or in memory. The `size' column indicates the sizes of motifs that were sampled. The $t$ column shows the number of threads allowed to run concurrently. The last column indicates how many significant motifs were returned (under the EL model). Only the 100 subgraphs with the highest number of instances after sampling were tested. All runtimes are rounded to the nearest minute. The memory column indicates the maximum heapspace allowed for the Java Virtual Machine. Preloading was always done with the lowest amount of memory indicated for that data set, and can be sped up if more memory is available.}
\label{table:various}
\end{threeparttable}

\noindent Section~\ref{section:classification} showed that our method can, in principle, return characteristic motifs, even when used with the edgelist null-model. Since the codelength under the EL model can be computed very efficiently, this configuration should be extremely scalable. To test its limits, we run several experiments on large data sets ranging from a million to a billion links.

In all experiments, we sample $10^6$ motifs in total (if multiple motif sizes are used, each size is equally likely to be sampled). We take the 100 most frequent motifs in this sample and compute their log-factors under the ER and EL models. We report the number of significant motifs found under the EL model. All experiments were executed on a single, dedicated machine, with 64 Gb of memory, and a single 2.60 Ghz Intel Xeon processor (E5-2650 v2).

Table~\ref{table:various} shows the results. The largest data set that we can analyse stored in-memory with commodity hardware is the wiki-nl data set. For larger data sets, we store the graph on disk. The graph is stored in two lists, as it is in the in-memory version. The first, the \emph{forward list}, contains at index $i$ a sorted list of integers $j$ for all links $(n_i,n_j)$ that exist: i.e. a list of outgoing neighbors of $n_i$. The second, the \emph{backward list}, contains lists of incoming neighbors for each node. The data is stored on disk so that random access can be performed efficiently (using the MapDB database engine\footnotemark).

\footnotetext{\url{http://www.mapdb.org/}}

For large graphs, converting a file from the common edgelist encoding (a file with a line for each link, encoded as a pair of integers) to this format can take considerable time, but this only needs to be done once, so we show the preloading and analysis times separately. Loading the graph is done by performing a disk-based sort of the edgelist-encoded file, on the first element of each pair, loading the forward list, sorting again by the second element, and loading the backward list. This minimizes random access as both lists can be filled sequentially in one pass.

We only require one pass over the whole data, to compute the model parameters (eg. the degree sequence). For the sampling and the computation of the log factors only relatively small amounts of random access are required. Since a graph can, in principle, be compressed with only a very small number of instances of a given motif, this gives us a very scalable method to find motifs in large data.

For disk-based experiments, we also limit the total number of rewritten links in the template graph to 500 000, to limit memory use. If the motif with a given list of instances results in more rewritten links, we do not consider it. Note that, since we search for a good pruning of the instance list, the motif will still be considered with a more heavily pruned instance list. A large number of rewritten links suggest that there are many instances with high ex-degree, so we likely do not lose much by this heuristic.

The main conclusion from this experiment is that we can perform motif analysis on data in the scale of billions of edges with very modest hardware. The problem is `embarrassingly parallel', and indeed, a decent speedup is achieved for multithreaded execution. We also observe that the amount of motifs found can vary wildly between data sets. The twitter and friendster data sets are from similar domains (large social networks), and yet for twitter, no significant motifs are found, by a wide margin,\footnote{The EL model compressed better than the motif model by millions of bits in all cases.} whereas for the friendster data the majority of the 100 most frequent motifs in the sample are significant. What exactly causes the difference in these data sets is a matter of future research; most likely requiring domain experts to investigate the motifs and their instances.

With large data, using the full parallelism available is not always the best option. There is a trade-off between maximizing concurrency and avoiding garbage collection. The second line for the friendster data shows the fastest runtime (using the maximum stable heapspace) which used 9 concurrently running threads (with 16 logical cores available).

We also show that our method can scale to larger motifs, often with a modest increase in resources. This is highly dependent on the data, however. On the twitter data, sampling motifs larger than 7 did not finish within 48 hours. This may be due to an incomplete implementation of the Nauty algorithm: the data may contain subgraphs that take a long time to convert to their canonical ordering. A more efficient canonization algorithm (like the complete Nauty) could improve performance. However, as the results show, some data allows for fast analysis on larger motifs.

Preloading can be prohibitively expensive, in the same order as the analysis itself. However, the graph in database format does not take up considerably more space than it does in raw edgelist-encoding. Preloading times could therefore be eliminated by distributing graph data in a suitable indexed binary format. For example, in the domain of knowledge graphs the HDT format by \cite{fernandez2013binary} provides both compression and indexing over the links of a graph.

\section{Discussion}

We have introduced a new method of testing motif relevance. This test allows for a new method of motif analysis scalable to graphs with billions of nodes, even on commodity hardware.

\subsection{Multiple Testing and Significance as Motif Evidence}
\label{section:multiple-testing}

The issue of multiple testing deserves some closer scrutiny. In most of our experiments, we presented significance scores of up to a hundred subgraphs. Should the reader take into account multiple testing when interpreting these results? Does the probability that one or more of these motifs is a ``false positive'' increase with the number of significances computed?

This question highlights an important issue in the use of hypothesis testing in motif analysis: the question we are trying to answer---is $M$ a motif?---is not the question we are testing for. Strictly interpreted, the hypothesis test investigates the question \emph{whether the full data came from the null model}. Thus, while we test multiple motifs, we only test them on a \emph{single} null-hypothesis. To perform a valid single null-hypothesis test, we can simply select the best compressing motif, and use the log-factor of that motif as the significance with which we reject the hypothesis that the null model generated the data. This is a valid code: we can check as many motifs as we like before choosing the one we will use to compress the data with. Since it is a valid code, it is a valid, single hypothesis test, and there is no need to correct for multiple testing.

Of course, in general practice, we are not actually interested in whether or not the null model is true. What we are investigating is whether a given \emph{motif} is ``true.'' We are using the framework of hypothesis testing merely as a \emph{heuristic} to guide unsupervised pattern mining. Does the general intuition of multiple hypothesis testing still apply? Does testing for more than one motif raise the probability of false positives? As shown in Section~\ref{section:recovering}, even without multiple testing, we should expect false positives, due to overlap with the true motif. However, increasing the number of motifs checked does not raise the amount of such false positives. This is because we analyze multiple motifs on a \emph{single dataset}. The process generating the data is stochastic, but once the data is known, the log-factors of all motifs are determined. We use stochastic methods to approximate these log factors from above, but higher sampling rates, or more motifs checked will only ever get us closer to the true value, it will not put us in danger of an increased number of false positives.

In summary, while the log-factor is technically a significance in a hypothesis test, this hypothesis test does not state anything directly about the motif found. Thus, the significance should not be regarded as proof or evidence of the motif itself. This is regardless of multiple testing, even if a single motif is tested, returning a high log-factor, there is no guarantee that the motif is in anyway meaningful. The best we can say is that the ranking of motifs provided by the log-factor provides a reasonable heuristic for practitioners to decide which motifs to investigate further.

Note also that these considerations hold for all analysis of network motifs using significance tests; they are not specific to our method.

\subsection{Limitations and Extensions}

Our current model does not allow for the detection of \emph{anti-motifs}, subgraphs that occur significantly less than expected. For that purpose, another model would be required; one which exploits the property that a subgraph has a \emph{lower} frequency than expected to compress the data. In theory, this is certainly possible: any such non-randomness can be exploited for the purposes of compression. We leave this as a matter for future research.

While our method allows fast motif search with the ER and EL null models, for other models, it can be as slow as the traditional method. To be scalable, our method requires a null model for which the probability of a graph under the null model can be quickly computed or approximated. The traditional method in contrast, requires a null model that allows efficient sampling. Currently, motif analysis has not matured to the point where different null models are routinely tested.

For such experiments, the size of the data determines the available options. If the data contains more than 1 million links (and hardware is limited to one compute node) our method is the only practical approach we are aware of. For smaller sizes, the traditional method becomes available, and with it, any null model from which samples can be taken efficiently.

Note also that the constraints on the null model are stronger than those on the alternative model. In our motif code, we technically use the length of one of many codewords, rather than the true probability (which would require a sum over all codewords). In the null model, such approximations are not appropriate, and each graph must have a single codeword. For the DS model, this makes the codelength so expensive to compute that we gain little over the traditional method. We did however, manage to define a similar model, which could be computed efficiently, and served as an acceptable proxy. This is the main advantage of our approach: there is no guarantee that every null model can be efficiently computed, but often, we can make approximations that allow us to trade off accuracy for scalability.

\subsection{Conclusion}

\label{section:conclusion}

We have presented a novel method for finding network motifs. Our method has several advantages:
\begin{itemize}
  \item The search for motif instances only needs to be run once: on the data $G$. Where the traditional approach requires a graph census to be repeated on samples from the null model, we only need to compute $p^\text{null}(G)$ and compare it to $p^\text{motif}(G)$ 
  \item The search does not need to find all instances of a motif; we do not even require an accurate estimate of the total number of instances. We only require as many instances as can be found with the resources available. For large graphs, a relatively small number of instances may suffice to prove some motifs significant.
  \item This also allows us to retain a list of exactly those instances that made the subgraph a relevant motif. These can then be inspected by a domain-expert to establish whether the motif truly represents a ``functional unit''. 
  \item Given sufficiently strong evidence, a single test can be used to eliminate multiple null models. 
  \item The resulting measure of relevance can be  used to compare the significance of motifs of different sizes in a meaningful way.
\end{itemize}

\noindent Motif detection is hampered by the difficulty of properly evaluating the motifs found. Even showing the results to a domain expert might not suffice, as a given motif might not be obviously good or bad at first sight. Depending on the domain, confirming a motif as meaningful might require considerable further experimentation. If the method were extended to knowledge graphs (labeled graphs expressing relational data), the resulting motifs will likely be easier to evaluate. Such an extension, however, is not trivial, and we leave this as a matter for future research.

Finally, we hope that our approach is illustrative of the general benefit of MDL techniques in the analysis of complex graphs. In conventional graph analysis a researcher often starts with a structural property that is observed in a graph, and then attempts to construct a process which generates graphs with that structural property. A case in point is the property of scale-freeness and the preferential attachment algorithm that was introduced to explain it \citep{albert2002statistical}. The correspondence between codes and probability distributions allows us instead to build models based on a description method for graphs. The trick then becomes to find a code that describes such graphs  with the desired property efficiently, instead of finding a process that is likely to generate such graphs. For many properties, such as cliques, motifs or specific degree sequences, such a code readily suggests itself.

\acks We thank Pieter Adriaans for valuable discussions. This publication was supported by the Dutch national program COMMIT, by the Netherlands eScience center, and by the Amsterdam Academic Alliance Data Science (AAA-DS) Program Award to the UvA and VU Universities. We thank the anonymous referees for valuable comments.

\section{Appendix}

\label{section:appendix}
\paragraph{Confidence Intervals for the Degree-Sequence Model}
\label{section:confidence-intervals}

In \cite{blitzstein2011sequential}, the sample mean $1/n \sum_i 1/q^\text{DS}_D(G_i)$ was used as an estimator for the expectation in Equation~\ref{line:expectation} in this paper. However, as shown in \cite{charo2010efficient}, the distribution of $1/q^\text{DS}_D(G)$ tends to be very close to log-normal. This means that the sample mean will converge \emph{very} slowly to the correct value. Specifically, the standard deviation of this estimate after $n$ samples is $\frac{1}{\sqrt{n}} \sqrt{(e^{\sigma^2}-1)e^{2\mu +\sigma^2}}$, which for a distribution with $\mu=200$ and $\sigma=10$, leads to a standard deviation of approximately $\frac{1}{\sqrt{n}} e^{300}$.

For this reason, we use the \emph{maximum-likelihood estimator} for the log-normal distribution instead. Let $Q_i = 1/q^\text{DS}_{D}(G_i)$. We assume $Q_i$ is log-normally distributed, so that $Y_i = \log Q_i$ is normally distributed. Let $\overline{Y} = \frac{1}{n}\sum_i Y_i$ and $S_Y = 1/n \sum_i (\overline{Y} - Y_i)^2$; then the maximum-likelihood estimator of $EQ$ is $\exp\left(\overline{Y} + \frac{1}{2}S_Y\right)$. Thus, the codelength under the degree sequence model can be estimated as $\left(\overline{Y} + \frac{1}{2}S_Y\right)\log_2(e)$.

As mentioned in the body of the text, even with the highly optimized implementations described by \cite{charo2010efficient} and \cite{kim2012constructing}, sampling can be slow for large graphs. In our implementation, a modern day laptop can take several minutes to produce a single sample for a random graph with $10^4$ nodes and $10^5$ links. However, we are not interested in precision beyond several orders of magnitude, so if we have a reliable method for determining confidence intervals, we can use those to provide us with safe bounds.

Since we are dealing with a log-normal source, we cannot simply use twice the standard error of the mean to approximate our error bars. We will use the parametric bootstrap procedure provided in \citep{angus1994bootstrap,zhou1997confidence}. To substantiate this method, we test the coverage of the two-sided symmetric confidence interval on three data sets. We proceed as follows: first we estimate the true value of $|\cG_D|$ with the ML estimator, using $10^6$ samples. Call this value $g$. We use this as our gold standard. We then sample a small number (5, 10, 20) of graphs and their associated probabilities. Using the bootstrap method we construct a two-sided symmetric confidence interval with $\alpha = 0.05$ on this sample. We repeat the procedure of sampling data and constructing an interval 5000 times and report the proportion of times $g$ was inside the confidence interval. If the bootstrap method is accurate, the resulting value should be close to 0.95. We use the following data sets:
  \begin{description}
  \item[cattle] Observed dominance behaviors between cows. A directed graph with 28 nodes, and 217 links. \cite{schein1955social,konect:2015:moreno_cattle}
  \item[revolution] Affiliations of 136 people to 5 organizations encoded as a bipartite graph. 141 nodes and 160 links. \cite{konect:2015:brunson_revolution}
  \item[random] A simple undirected random graph of 50 nodes,with  each pair of distinct nodes connected with probability $0.5$.
  \end{description}
As Table~\ref{table:coverage-experiment} shows, this method becomes relatively reliable at around 10 samples, although the intervals are quite large at that sample size.

\begin{table}
  \begin{center}
  \begin{tabular}{r r r r r r r }
  \hline
   &  \multicolumn{2}{|c|}{5 samples} & \multicolumn{2}{c|}{10 samples}  & \multicolumn{2}{c}{20 samples}  \\
   \hline
  cattle  & 0.93 & 511 (379) & 0.94 & 194 (94.6) & 0.94 & 107 (35.3) \\
  revolution   & 0.89 & 147 (120) & 0.92 & 59.3 (29.2) & 0.93 & 33.5 (11.3) \\
  random  & 0.91 & 108 (97.0) & 0.94 & 44.7 (24.5) & 0.94 & 25.3 (9.44) \\
  \hline
  \end{tabular}
  \end{center}

  \caption{Coverage and interval length at $\alpha=0.05$ for three small data sets. The coverage is the number of times the true value $g$ was contained in the interval, over 5000 trials. The length is the mean length of these intervals in bits. The standard deviation is given in parentheses. } 
  \label{table:coverage-experiment}
\end{table}

Now, when we use $L^\text{DS}$ as a base model in $L^\text{motif}$, the intractable value $|\cG_D|$ occurs in two places: the encoding of the subgraph, and the encoding of the template graph. Since we use our estimator for both, we must be careful to end up with a correct confidence interval for the resulting motif code. Let $D'$ be the degree sequence of the subgraph, and $D$ be the degree sequence of the template. Then, we can split the total code length into three components: $\log |\cG_{D'}|$, $\log |\cG_D|$ and $R$. $R$ is the sum of all parts of the code that we can compute exactly, including the sizes and sequences of the motif and template graph (ie. everything but $\log |\cG_{D'}|$ and $\log |\cG_D|$). The total codelength is described by $\log |\cG_{D'}| + \log |\cG_D| + R$, where the first two terms require the use of the estimator. Let $Q_m$ and $Q_h$ be random variables representing the inverse probability of graphs sampled from the importance sampling algorithm, for the degree sequence of the motif and the template graph respectively. In other words, the true codelength for the motif code is
\begin{align*}
& \log E Q_m + \log  E Q_h + R \\ 
= & \log E Q_m E Q_h + R \\
= & \log E [Q_m Q_h] + R
\end{align*}
where the last line follows from the fact $Q_m$ and $Q_h$ are independent. So to get a correct confidence interval, we can take the same number of samples of both $Q_m$ and $Q_h$, multiply their probabilities, and perform the bootstrap analysis on the list of these multiplied probabilities (since we are summing the logarithms of log-normally distributed variables, the result is log-normally distributed as well).

\bibliography{motifs}

\end{document}

### 2018-09-05 kgmotifs/kgmotifs.tex
deadline approaching

--- 2017-08-12 kgmotifs/kgmotifs.tex
@@
-\documentclass[11pt]{article}
+\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
+
+\usepackage{aaai19}    % Required
+\usepackage{times}     % Required
+\usepackage{helvet}    % Required
+\usepackage{courier}   % Required
+\usepackage{url}       % Required
+\usepackage{graphicx}  % Required
+\frenchspacing         % Required
+\setlength{\pdfpagewidth}{8.5in}  % Required
+\setlength{\pdfpageheight}{11in}  % Required
+
+% PDF Info Is Required:
+\pdfinfo{
+/Title (Detecting Motifs in Knowledge Graphs using Compression)
+/Author (...)}
+
+\setcounter{secnumdepth}{0}
 
@@
 
+\usepackage{easylist}
+\usepackage{soul}
+
 \newcommand{\N}{{\mathbb N}}
@@
 
+\newcommand{\X}{{\cal X}}
+\newcommand{\B}{{\mathbb B}}
 \newcommand{\G}{{\cal G}}
+\newcommand{\cS}{{\cal S}}
+
 \newcommand{\I}{{\cal I}}
@@
 \title{Detecting Motifs in Knowledge Graphs using Compression}
+\author{
+Author names withheld
+% Peter Bloem\\
+% Knowledge Representation \& Reasoning Group\\
+% VU University\\
+% vu@peterbloem.nl\\
+}
 
@@
 Network motifs are useful patterns or meaningful subunits of the graph that recur frequently.
-We introduce a scalable approach for detecting such motifs in large knowledge graphs, inspired by recent work for simple graphs, and show that the motifs returned reflect the basic structure of the graph.
-Specifically, we show that common motifs reflect graph patterns used in common queries, basic schematic units of the graph and meaningful functional subunits, for various knowledge graphs.
+We extend the common definition of a network motif to coincide with a \emph{basic graph pattern}, commonly used in the knowledge graph literature.
+We introduce an approach, inspired by recent work for simple graphs, for inducing these for a given knowledge knowledge graph, and show that the motifs returned reflect the basic structure of the graph.
+Specifically we show that in random graphs, no motifs are found, and that when we insert a motif artificially, it can be detected, even if as few as 10 instances are inserted.
+Finally, we show the results of motif induction on three real-world knowledge graphs, and show that our method provides good suggestions for motifs that reflect the structure of the graph in an intuitive way.
 \end{abstract}
 
-\noindent The Linked Open Data cloud contains a wealth of knowledge graphs, and is growing quickly.
-At the time of writing the average knowledge graph contains over 65\,000 edges, with many datasets containing more than a million.\footnotemark~For such large graphs, it can be difficult to see the forest for the trees: how is the graph structured at the lowest level?
+\noindent \emph{Knowledge graphs} are an extremely versatile way of storing knowledge.
+They allow knowledge to be encoded without a predefined schema, they allow different amounts of data to be encoded for each known entity and they allow different modalities to be integrated seamlessly \cite{wilcke2017knowledge}.
+This versatility comes at a price.
+For a given knowledge graph, it can be difficult to see the forest for the trees: how is the graph structured at the lowest level?
 What kind of things can I ask of what types of entities?
@@
 
-\footnotetext{\url{http://stats.lod2.eu/stats}}
-
-In the domain of plain (unlabelled) graphs, \emph{network motifs} \cite{milo2002network} were recently introduced as a tool to provide users of large graphs insight into their data.
-Network motifs are small subgraphs whose frequency is unexpected with respect to a null model.
-That is, for a given graph dataset $g$ and small graph $m$, we count the frequency $F(m, g)$: how often $m$ occurs in $g$ as a subgraph.
-We define a probability distribution over graphs $p^\text{null}(g)$, and estimate the probability that a graph sampled from $p^\text{null}$ contains more instances of $m$ than observed in our data: $p^\text{null}(F(m, G) \geq F(m, g))$, where $G$ is a random variable representing a graph.
-If this probability is low (commonly, below 0.05), we consider $m$ a motif.
+In the domain of unlabeled simple graphs, \emph{network motifs} \cite{milo2002network} were introduced as a tool to provide insight into local graph structure.
+Network motifs are small subgraphs whose frequency in the graph is unexpected with respect to a \emph{null model}.
+That is, for a given graph dataset $G$ and small graph $M$, we count the frequency $f(M, G)$: how often $M$ occurs in $G$ as a subgraph.
+We define a probability distribution over graphs $p^\text{null}(\G)$, and estimate the probability that a graph sampled from $p^\text{null}$ contains more instances of $M$ than observed in our data: $p^\text{null}(f(M, \G) \geq f(M, G))$, where $\G$ is a random variable representing a graph.
+If this probability is very low (commonly, below 0.05), we consider $m$ a motif.
 \footnotemark
 
-Unfortunately, counting all occurrences of a motif in a network is an expensive operation.
-One that we need to perform not only on the data, but also on many samples from the null model, in order to estimate the probability described above.
+\begin{figure*}[bth]
+  \centering
+    \includegraphics[width=\textwidth]{example.pdf}
+    \caption{An example of the principle behind our motif code.
+a) A basic knowledge graph.
+Note that in our analysis, we consider only the indices of the nodes and relations, not their labels.
+b) A motif that occurs frequently.
+c) A compressed representation; we remove all edges that are part of an occurrence of the motif.
+We then store separately which nodes match the motif.
+The motif, together with the template graph and the matches can be used to reconstruct the graph.
+We take this as a heuristic: the better the motif compresses, the more characteristic it is likely to be for the graph.}
+    \label{figure:example}
+  \label{figure:codes}
+\end{figure*}
 
-\footnotetext{This procedure has the structure of a hypothesis test, but it is important not to interpret it as statistical evidence for the meaningfulness of the motif.
-The only thing it proves (in a frequentist statistical sense) is that $p^\text{null}$ is not the true source of the data.
+Unfortunately, estimating this probability usually requires repeating the subgraph count on many samples from the null model.
+To avoid this costly operation, \cite{bloem2017large} introduces  an alternative method, using \emph{compression} as a heuristic for motif relevance: the better a motif compresses the data, the more likely it is to be meaningful.
+It is shown that this principle can be implemented in a similar sort of hypothesis test for a null model, allowing a comparable workflow to classical motif analysis.
+
+\footnotetext{This procedure has the structure of a hypothesis test, but it is important not to interpret it as \emph{statistical evidence} for the meaningfulness of the motif.
+The only thing it shows (in a frequentist statistical sense) is that $p^\text{null}$ is not the true source of the data.
 This is usually not a surprise: we are rarely able to model all aspects of a realistic data-generating process in a single distribution.
@@
 
-In \cite{bloem2017large}, an alternative method is presented that uses \emph{compression} as a heuristic for motif relevance: the better a motif compresses the data, the more likely it is to be meaningful.
-It is shown that this principle can be implemented in a similar sort of hypothesis test for a null model, allowing a comparable workflow to classical motif analysis.
-Crucially, this method scales to graphs with billions of edges, suggesting that it might be suitable for the scale of knowledge graphs we encounter in the LOD cloud.
-
-In this paper, we extend the compression-based motif analysis to Knowledge Graphs.
-For the purposes of this research we define Knowledge graphs as labeled multigraphs.
-Nodes are labeled with entity names, and links are labeled with relations.
-We extend the definition of a motif to that of a basic graph pattern: that is, a motif is a small graph with \emph{some} of its nodes and links labeled.
-The motif \emph{occurs} in a graph where both the graph structure and the labels match.
-The unlabeled nodes and links are free to contain whatever label when the motif is mapped to the data.
-
-To maintain the connection to graph pattern queries, we do not require the motifs to be \emph{induced} subgraphs: that is, when a motif is mapped to a graph, the graph can contain additional links that are not specified by the motif.
+In this paper, we extend the compression-based motif analysis to knowledge graphs.
+For the purposes of this research we define knowledge graphs as labeled, directed multigraphs.
+Nodes are uniquely labeled with entity names, and links are non-uniquely labeled with relations.
+We extend the definition of a motif to that of a \emph{basic graph pattern}: that is, a motif is a small graph with \emph{some} of its nodes and links labeled with entities and relations as they occur in the graph.
+Other nodes and links are labeled as \emph{variables}.
+A pattern matches at a particular location in the graph, if the variables can be replaced with values from the graph, and the motif becomes a subgraph as a result.
+The intuition behind our method is that the more compactly a pattern compresses the graph, the better the motif.
+Figure~\ref{figure:example} shows an illustration of the basic principle.
+In the preliminaries section below, we justify this intuition more formally.
 
 We perform several experiments to show that our method returns meaningful subgraphs.
-First, we look at several graphs for which schema information is provided.
-That is, we know the basic properties available for each entity.
-We show that the motifs found by our method correspond well to the schema, and compare performance against several basic benchmarks.
-Second, we run motif analysis on several datasets for which a list of real SPARQL queries entered by users is available.
-From these, we extract the basic graph patterns, and show that (a) these correspond to good motifs using our criterion and (b) our method can return such motifs with good precision and recall.
+First we test the intuition that a random graph shoudl contain no motifs.
+We also show that when we artifically insert motifs into a random graph, we can then detect these as motifs, sometimes with as few as 10 instances inserted into the graph.
+Finally, we show the results of motif analysis on three real-world knowledge graphs.
 
-All code and benchmark datasets used in the paper are available, under open licenses.\footnotemark
+All code and datasets used in this paper are available, under open licenses.\footnotemark
 
-\footnotetext{\url{...}}
-
-% TODO: Make illustration
+\footnotetext{URL withheld for purposes of blind reviewing.}
 
@@
 
-\paragraph{Preliminaries}
+Network motifs for unlabeled simple graphs were introduced in \cite{milo2002network}.
+A more comprehensive overview of the related literature can be found in \cite[Section~1.1]{bloem2017large}.
+In \cite{bloem2017large}, the principle of Minimum Description Length (MDL) was first connected to motif analysis.
+However, the idea had earlier been exploited for detecting meaningful subgraphs in the SUBDUE algorithm \cite{cook1994substructure}.
 
-% TODO: General purpose codes: natural numbers, integers, sequences.
+A few other methods have been proposed for inducing the structure of a given knowledge graph.
+In \cite{pham2015deriving}, the authors use the principle of characteristic sets to characterize a knowledge graph in terms of the star patterns it contains.
+In \cite{pham2016exploiting}, they show that the majority of the LOD cloud can be efficiently described using such principles, showing the highly tabular structure of many knowledge graphs.
+In \cite{volker2011statistical}, association rules mining is used to induce basic patterns in the graph.
+
+To the best of our knowledge, ours is the first method presented that can potentially induce any basic graph pattern.
+
+\subsection{Preliminaries}
+
+\paragraph{Minimum Description Length}
+Our method is based on the MDL principle: that we should favour models that compress the data.
+We will show briefly how this intuition can be made mathematically precise.
+For more details, we refer the reader to \cite{grunwald2007minimum} for MDL in general, and to \cite{bloem2017large}, for a more extensive discussion these principles in the domain of graph analysis.
+
+Let $\B$ be the set of all finite-length binary strings.
+We use $|b|$ to represent the length of $b \in \B$.
+Let $\log(x) = \log_2(x)$.
+A \emph{code} for a set of objects $\cal X$ (such as graphs) is an injective function $f: {\cal X} \to \B$, mapping objects to binary code words.
+All codes in this paper are \emph{prefix-free}: no code word is the prefix of another.
+We will denote a \emph{codelength function} with the letter $L$, ie.
+$L(x) = |f(x)|$.
+We commonly compute $L(x)$ directly, without first computing $f(x)$ and we refer to $L(x)$ as a code.
+
+A well known result in information theory is the association between codes and probability distributions, implied by the \emph{Kraft inequality}: for each probability distribution $p$ on $\cal X$, there exists a prefix-free code $L$ such that for all $x \in \cal X$: $- \log p(x) \leq L(x) < -\log p(x) + 1$.
+Inversely, for every prefix-free code $L$ for $\cal X$, there exists a probability distribution $p$ such that for all $x \in \cal X$: $p(x) = 2^{-L(x)}$.
+For proofs, see \cite[Section~3.2.1]{grunwald2007minimum} or \cite[Theorem~5.2.1]{cover2006elements}.
+To explain the intuition, note that we can easily transform a code $L$ into a sampling algorithm for $p$ by feeding the decoding function random bits until it produces an output.
+To transform a probability distribution to a code, techniques like arithmetic coding \cite{rissanen1979arithmetic} can be used.\footnotemark
+\footnotetext{
+As explained in \cite[page 96]{grunwald2007minimum}, the fact that $-\log p^*(x)$ is real-valued and $L^*(x)$ is integer-valued can be safely ignored and we may \emph{identify} codes with probability distributions, allowing codes to take non-integer values.
+}
+
+When we need to encode a single choice from a finite set $S$ of options, we will often use the code with length $\log |S|$, corresponding to a uniform distribution on $S$.
+In some cases, we can allow codes with multiple codewords for a single object.
+If a code $L$ has multiple codewords for some object  $x$, we may indicate the choice for a particular codeword by a parameter $a$ as $L(x; a)$.
+
+\paragraph{Relevance testing} We will use the MDL principle to perform a hypothesis test.
+Assume we have some data $x$ and a null hypothesis that it was sampled from distribution $p^\text{null}$.
+Let $L^\text{null}$ be the corresponding codelength.
+A simple but crucial result, known as the \emph{no-hypercompression inequality} \cite[p103]{grunwald2007minimum} tells us that the probability of sampling data $x$ from $p^\text{null}$ that can be described shorter than $L^\text{null}(x)$ by $k$ or more bits, \emph{using any code} is less than $2^{-k}$.
+Thus, a standard MDL hypothesis test would consist of stating the null hypothesis in terms of a code $L^\text{null}$, designing a code (before seeing the data) which compresses the data better than $L^\text{null}$ by, say, 10 bits and rejecting the null hypothesis with confidence $2^{-10}$.
+For a longer, more intuitive explanation of this principle in pattern induction, we refer the reader to \cite[Section~2]{bloem2017large}.
+
+We must note again, that when we use this procedure to find motifs, we are not providing evidence for the hypothesis that the motif is ``correct.'' The only thing we are statistically proving is that the null-model is incorrect.
+Nevertheless, hypothesis testing has shown to be a valuable heuristic in motif analysis.
+
+\paragraph{Common codes}
+In the construction of our codes for graphs, we require some simpler codes as building blocks.
+First, when we store any positive integer $n$, we do so with the code corresponding to the distribution $p^\N(n) = 1/(n(n+1))$, and denote it $L^\N(n)$.
+For nonnegative numbers we add 1 to the argument.
+for the full range of integers, we add an extra bit for the sign, and then use the first code for negative integers and the second for positive ones.
+
+We will often need to encode sequences of integers.
+These will be highly skewed, with only a subset of integers occurring frequently, and others occurring infrequently or not at all.
+As noted in \cite{de2016names} a code based on the Pitman-Yor model \cite{pitman1997two} is very effective in such situations.
+Let $S = \langle S_1, ..., S_n\rangle$ be a sequence of integers of length $n$.
+We first store its members $m(s)$ in the order in which they first occur: we first store $n$ and the first member using the codes described above.
+The rest of the members are stored sequentially using the distance from the current to the next member.
+
+Let $f(A, B)$ be the frequency of symbol $A$ in sequence $B$.We then store the complete sequence using the code corresponding to the following distribution:
+\begin{align*}
+&p^\text{PY}_{\alpha,d}(S) = \prod_{i \in [1,k]} \text{PY}_{\alpha,d}(S_i\mid S_{1:i-1}) \\
+&\text{PY}_{\alpha,d}(S_i \mid S') = 
+\begin{cases}
+\frac{\alpha - d |m(S)|}{|m(S)| + \alpha} & \text{if} f(S_i, S') = 0 \\ 
+\frac{f(S_i, S') - d}{|m(S)| + \alpha} & \text{otherwise}
+\end{cases}
+\end{align*}
+See \cite{de2016names} for a more intuitive explanation. In all experiments we use $\alpha = 0.5$, $d=0.1$. We will refer to the total for a sequence  codelength under this procedure as $L^PY(S)$
 
@@
 To comfortably define probabilities and codes on graphs, we will slightly deviate from conventional definitions.
-Specifically, we will analyze the \emph{structure} of knowledge graphs only, ignoring any meaning they have outside the graph.
+Specifically, we will analyse the \emph{structure} of knowledge graphs only, ignoring any meaning they have outside the graph.
 For instance, we will distinguish between links with different labels, but we will ignore any internal structure of the label itself (such as shared IRI prefixes).
-Under this assumption we can assume that the node and link labels of our knowledge graphs are simply natural numbers.
-This leads to the following definition.
+Under this assumption we can assume that the node and link labels of our knowledge graphs are simply natural numbers.\footnotemark This leads to the following definition.
 
-A \emph{knowledge graph} $G$, hereafter known simply as a \emph{graph} is a, is tuple $G = (v_G, r_G, E_G)$.
-$v_G \in \N$ is the number of nodes in the graph, and $r_G$ is the number of relations.
-We will define the nodeset as $V_G = \{0, \ldots, v_G-1\}$ and the relation-set as $R_G = \{0, \ldots, r_G\}$.
-The \emph{tripleset} $E_G \subset V_G \times R \times V_G$ defines the links of the graph and their labels.
-\footnotemark
+\footnotetext{For practitioners this restriction is not noticeable, as the indices can simply be mapped back to the original strings when the found motifs are presented.}
 
-\footnotetext{Commonly, the elements of the nodeset of a graph are defined to be simple atoms from some larger unspecified universe.
-However, since we are discussing sampling graphs, we need to specify where these atoms come from.
-This definition specifies that graphs are essentially unlabeled, but nodes do have a definite ordering (i.e.
-two graphs can be isomorphic without being the same graph).}
+A \emph{knowledge graph} $G$, hereafter simply a \emph{graph}, is a tuple $G = (v_G, r_G, E_G)$.
+$v_G \in \N$ is the number of nodes in the graph, and $r_G \in \N$ is the number of relations.
+We define the nodeset as $V_G = \{0, \ldots, v_G-1\}$ and the relation-set as $R_G = \{0, \ldots, r_G\}$.
+The \emph{tripleset} $E_G \subset V_G \times R \times V_G$ determines the edges of the graph and their labels.
 
-As shown in Figure~\ref{figure:diagram}, the patterns that we aim to find are partially labeled: some nodes and edges are labeled with negative integers.
-In other words, the motif is a \emph{basic graph pattern}, and the edges and nodes marked with negative integers are variables.
-We will define a pattern in the same way we define a graph, but we will extend its nodeset en relationset.
-A pattern $M$ for graph $G$ is a tuple $(v'_M, r'_M, G, E_M)$.
-$v'_M$ and $r'_M$ indicate the number of variable nodes and links respectively.
-$M$'s nodeset is defined as $V_M = \{-v'_M, \ldots, v_G-1\}$ and the relationset as $R_M = \{-r'_M, \ldots, r_G-1\}$, with $E_M \subset V_M \times R_M \times V_M$, as for graphs.
-In short, nodes can be labeled with nonnegative integers referring to $G$'s nodes or with negative integers representing a variable node, and similar for relations.
+As shown in Figure~\ref{figure:example}, the patterns that we aim to find are partially labeled: some nodes and edges are labeled with as variables.
+We will represent these with negative integers.
+A pattern $M$ for graph $G$ is a tuple $(V_M, R_M, G, E_M)$.
+Let $v_M$ and $r_M$ indicate the number of variable nodes and variable links in $M$ respectively, then $V_M \subseteq \{-v'_M, \ldots, v_G-1\}$ and $R_M \subseteq \{-(r'_M+v'M), \ldots,-v'M, 0,\ldots, r_G-1\}$, with $E_M \subset V_M \times R_M \times V_M$, as for regular graphs.
+That is; nodes can be labeled either with nonnegative integers referring to $G$'s nodes or with negative integers representing a variable node, and similar for relations.
+The negative integers are always contiguous within a single graph, with the highest representing the node labels and the lowest representing the edge labels.
 
@@
 $I^r$ is a sequence of non-distinct integers of length $r'_M$.
-For each link $(s, p, o) \in E_M$ containing a negative $s$, $p$ or $o$, there is a corresponding link in $E_G$ with a negative $s$ replaced by $I^n_{-s}$, a negative $o$ replaced by $I^n_{-o}$, and a negative $p$ replaced by $I^r_{-p}$.
-Put simply: for a pattern to match, variable links marked with the same negative integer, must map to the same relation in order for the pattern to match, but variable links labeled with different negative integers \emph{may} map to the same relation.
+For each edge $(s, p, o) \in E_M$ containing a negative $s$, $p$ or $o$, there is a corresponding link in $E_G$ with a negative $s$ replaced by $I^n_{-s}$, a negative $o$ replaced by $I^n_{-o}$, and a negative $p$ replaced by $I^r_{-p - v'_M}$.
+Put simply: for a pattern to match, variable edges marked with the same negative integer, must map to the same relation in order for the pattern to match, but variable links labeled with different negative integers \emph{may} map to the same relation.
 Variable nodes are always labeled distinctly and may never map to the same node in $G$.
 An instance describes a subgraph of $G$ that \emph{matches} the pattern $M$.
+Each edge in the motif may only match one edge in the graph.
+In other words, the occurrence of the motif in the graph must have as many edges as the motif itself.
 
@@
 In most settings we do not actually have a given $M$ and $\I$ and we will need to search for them.
-We describe a simple and fast search algorithm in Section~\ref{section:search}.
-
-\subsection{Relevance test}
+We describe a simple and search algorithm in Section~\ref{section:search}.
 
@@
 
-We will first describe our method under the assumption that a good null model $p^\text{null}$ is available, and that $- \log p^\text{null}(G)$ can be efficiently computed.
-In Section~\ref{section:null-model}, we explain which null model we use for our experiments, and how it is implemented.
+\section{Null model}
+
+\label{section:null-model}
+
+The most common null model for classical motif analysis is the degree-sequence model (also known a the configuration model \cite{newman2010networks}): a uniform distribution over all graphs with a particular degree sequence.
+We extend this to knowledge graphs by also including the degree of each relation: that is, the frequency with which it occurs in the tripleset.
+Let a \emph{degree sequence} $D$ of length $n$ be a triple of three integer sequences: $(D^\text{in}, D^\text{rel}, D^\text{out})$.
+If $D$ is the degree sequence of a graph, then node $i$ has $D^\text{in}_i$ incoming links,  $D^\text{out}_i$ outgoing links and for each relation $r$, there are $D^\text{rel}_r$ triples.
+
+Let $\G_D$ be the set of all graphs with degree sequence $D$. Then the configuration model can be expressed simply as
+\[
+p^C(G) = \frac{1}{|\G_D|}
+\]
+for any $G$ that satisfies $D$ and $p(G) = 0$ otherwise.
+Unfortunately, there is no efficient way to compute $|\G_D|$ and even approximations tend to be costly for large graphs.
+Following the approach in \cite{bloem2017large}, we define a fast approximation to the configuration model, which works well in practice for motif detection.
+
+We can describe a knowledge graph by three length-$m$ integer sequences: $S$, $P$, $O$, such that $\{(S_j, P_j, O_j)\}_j$ is the graph's tripleset.
+If the graph satisfies degree sequence $D$, then we know that $S$ should contain node $j$ $D^\text{out}_j$ times, $P$ should contain relation $r$ $D^\text{rel}_r$ times and $O$ should contain node $j$ $D^\text{in}_j$ times.
+Let ${\cal S}_D$ be the set of all such triples of integer sequences satisfying $D$.
+We have
+\[
+|{\cal S}_D| =
+ {m \choose {D_1^\text{out}, \ldots, D_n^\text{out}} }
+ {m \choose {D_1^\text{rel}, \ldots, D_{|R_G|}^\text{rel}} }
+ {m \choose {D_1^\text{in}, \ldots, D_n^\text{in}} } \text{.}
+\]
+
+While every member of ${\cal S}_D$ represents a valid graph satisfying $D$, many graphs are represented multiple times.
+Firstly, many elements of ${S}_D$ contain the same link multiple times.
+We call the set without these elements ${\cal S'}_D \subset {\cal S}_D$.
+Secondly the links of the graph are listed in arbitrary order; if we apply the same permutation to all three lists $S$, $P$ and $O$, we get a new representation of the same graph.
+Since we know that any element in ${\cal S}'_D$ contains only unique triples, we know that each graph is present exactly $m!$ times.
+This gives us
+\[
+|\G_D| = |{\cal S}'_D| \frac{1}{m!} \leq  |{\cal S}_D| \frac{1}{m!} \text{.}
+\]
+
+We can thus use 
+\[
+p^\text{EL}_D(G) =  \frac{m!}{{m \choose {D_1^\text{out}, \ldots, D_n^\text{out}} }
+ {m \choose {D_1^\text{rel}, \ldots, D_{|R_G|}^\text{rel}} }
+ {m \choose {D_1^\text{in}, \ldots, D_n^\text{in}} }} \leq p^C(G) \text{.}
+\]
+
+Filling in the definition of the multinomial coefficient, and rewriting, we get a codelength of:
+\begin{align*}
+- &\log p^\text{EL}_D(G) = 2 \log(m!) \\
+&- \sum_i \log(D_i^\text{in}!)- \sum_i \log(D_i^\text{rel}!)- \sum_i \log(D_i^\text{out}!)
+\end{align*}
+
+as an approximation for the DS model.
+We call this the edgelist (EL) model.
+It gives a probability that always lower-bounds the configuration model, since it affords some probability mass to graphs that cannot exist.\footnotemark ~Experiments in the classical motif setting have shown that the EL model is an acceptable proxy for the DS model \cite{bloem2017large}, especially considering the extra scalability it affords.
+
+\footnotetext{Note that we cannot simply think of $p^\text{EL}$ as a uniform model for graphs containing multiple links, since we can only divide by $m!$ if we know that all triples are unique.}
+
+ \paragraph{Encoding D} In order to encode a graph with $L^\text{EL}_D$, we must first encode $D$.
+\footnote{Or, equivalently, to make $p^\text{EL}$ a complete distribution on all graphs, we must provide it with a prior on $D$.} For each of the three sequences in $D$ we use the following model:
+\begin{align*}
+ p(D) &= \prod_i q^{\N}(D_i) \\
+ L(D) &= - \sum_i \log q^{\N}(D_i)
+\end{align*}
+
+where $1^{\N}$ is any distribution on the natural numbers.
+This is an optimal encoding for $D$ assuming that its members are independently drawn from $q^{\N}$.
+When we use $p^\text{EL}$ as the null model, we use the data distribution for $q^{\N}$ to ensure that we have a lower bound to the optimal code-lengthn (in essence, we cheat, giving the null model a slightly lower than optimal codelength).
+When we use $p^\text{EL}$ as part of the motif code, we must use a fair encoding, so we use the Pitman-Yor code to store each sequence in $D$.
 
 In the design of our method, we will constantly aim to find a trade-off between completeness and efficiency that allows the method to scale to very large graphs.
-Specifically, when we economize, we will only do so in a way that makes the hypothesis test described in Section~\ref{section:relevance-test} \emph{more conservative}.
+Specifically, when we economize, we will only do so in a way that makes the hypothesis test \emph{more conservative}.
 
@@
 
-As described above, we can perform our relevance test with any compression method which exploits the pattern $M$, and its instances $I$ to store the graph efficiently.
+As described above, we can perform our relevance test with any compression method which exploits the pattern $M$, and its instances $\I$ to store the graph efficiently.
 The better our method, the more motifs we will find.
@@
 The information in all of these together is sufficient to reconstruct the graph.
-Note that everything is stored using prefix-free codes, so that we can simply sum the codelengths we get for each aspect to get the codelength for the complete graph.
+Note that everything is stored using prefix-free codes, so that we can simply concatenate the codewords (or rather, sum the codelengths we get for each aspect).
 
-We also assume that we are given a code $L^\text{base}$ for generic knowledge graphs.
+We also assume that we are given a code $L^\text{base}$ for generic knowledge graphs (in practice, this will be the null model, although the motif code is valid for any base code).
 
@@
 \begin{description}
- \item[the graph dimensions] We first store $|V_G|$, $|E_G|$ and $|R_G|$ using the generic code $L^{\N}(\cdot)$. We also store the size of $\I$.
- \item[the pattern] Let $M'$ be the unlabeled graph representation of $M$: that is, $M'$ is a plain graph (not a pattern), $V_{M'}$ contains only the first $V_{M'}$ natural numbers and similarly for $R_{M'}$.
-Otherwise, $M$ is isomorphic to $M'$: the node and edge labels of $M$ can be mapped to those of $M'$ so that $M$ only contains an edge $E$ if and only if $M'$ contains the edge resulting from mapping $E$'s symbols.
- \item[the template] This is the graph, minus all links occurring in  instances.
+ \item[the graph dimensions] We first store $|V_G|$, $|E_G|$ and $|R_G|$ using the generic code $L^{\N}(\cdot)$. 
+ \item[the pattern] We store the structure of the pattern using the base code, and its labels as a sequence using the Pitman-Yor code.
+ \item[the template] This is the graph, minus all links occurring in instances of $M$.
 Let $E_G'$ be $E_G$ minus any link occurring in any member of $\cal I$.
 We then store $(V_G, E_G')$ using $L^\text{base}(\cdot)$.
- \item[the instances] In order to reconstruct the graph from the template, we need to know which nodes correspond to instances of the motif, and in what order.
-For each node in $M'$, we create a list of length $\cal |I|$ of the nodes in $G$ to which it maps for each instance.
-For the non-variable nodes of the pattern, this will be a list simply repeating one node node of $G$.
-For the variable nodes, it will be a list containing some subset of the nodes in $G$.
- We store each list using the DM code.
-It may seem wasteful to store a list repeating a single element (instead of storing a labeled version of the pattern), but under the DM code such lists will be stored efficiently, using only a small, constant number of bits.
-The more varied the list, the more bits will be required to store it.
- \item[the edge labels] We now have enough information to reconstruct the graph structure.
-All that remains to fully reconstruct $G$ is to label the links of the motif instances.
-As for the nodes, we create one list of length $\cal I$ for each type of link in $M'$, which we store using the DM code.
+ \item[the instances] To store the instances, we view the connections between the nodes made by motifs as a hypergraph, and we extend the EL code to store it. The details are given below.
 \end{description}
@@
 \tab\tab a graph $G$, a pattern $M$\\ 
-\tab\tab a list $\cal I$ of instances of $M$ in $G$, a code $L^\text{base}$ on knowledge graphs.\\
+\tab\tab instances $\cal I$ of $M$ in $G$, a code $L^\text{base}$.\\
 \\
-\tab $b_\text{dim} \leftarrow L^\N(|V_G|) + L^\N(|R_G|) + L^\N(|E_G|) + L^\N(|\I|)$ \\
+$b_\text{dim} \leftarrow L^\N(|V_G|) + L^\N(|R_G|) + L^\N(|E_G|)$ \\
 
-\tab \emph{// Turn the pattern into a normal knowledge graph}\\
-\tab $E_{M'} = \left\{(\text{i}(s, V_M),  \text{i}(p, R_M), \text{i}(o, V_M)) \mid (s, p, o) \in E_M\right\}$\\
-\tab $M' = ([0 ..  |V_M|-1], [0 .. |R_M|-1], E_{M'})$\\
-\tab $b_\text{pattern} \leftarrow L^\text{base}(M')$ \\
+\emph{\# Turn the pattern into a normal knowledge graph}\\
+$E_{M'} \leftarrow$ the edges of $M$ with contiguous integer labels \\
+$M' \leftarrow (|V_M|, |R_M|, E_{M'})$ \\
+$S_M \leftarrow$ the labels of M in canonical order \\
+$b_\text{pattern} \leftarrow L^\text{base}(M') + L^{PY}(S_M) $\\
 
-\tab $E'_G = E_G - \cup_{\I \in {\cal I}} I$ \hfill \emph{// Store the graph, with instances removed}\\
-\tab $b_\text{template} = L_\text{base}((V_G, R_G, E'_G))$\\
+\emph{\# Store the template graph}\\
+$E'_G \leftarrow E_G - \cup_{\I \in {\cal I}} \text{triples}(I)$ \\
+$b_\text{template} \leftarrow L_\text{base}((V_G, R_G, E'_G))$\\
 
-\tab $b_\text{labels} \leftarrow 0$ \hfill \emph{// Store the node labels} \\
-\tab \textbf{for} $n$ in $V_M$\\
-\tab\tab $\I_n = \langle v(n, I_1), v(n, I_2), ... \rangle$\\
-\tab\tab\tab with $v(n, I)$ the label (from $V_G$) of node $n$ in instance $I$\\
-\tab\tab $b_\text{labels} \leftarrow b_\text{labels} + L^{DM}(\I_n)$\\
+$b_\text{instances} \leftarrow -\log p_M(\I) + \sum_{D \in D^{\I}} L^{PY}(D)$\\
 
-\tab \textbf{for} $e$ in $E_M$ \hfill \emph{// Store the edge labels}\\
-\tab\tab $\I_e = \langle r(e, I_1), r(n, I_2), ... \rangle$\\
-\tab\tab\tab with $r(e, I)$ the relation (from $R_G$) of edge $e$ in instance $I$\\
-\tab\tab $b_\text{labels} \leftarrow b_\text{labels} + L^{DM}(\I_e)$\\
+\textbf{return} $b_\text{dim} + b_\text{pattern} + b_\text{template} + b_\text{instances}$\\
+}
+\end{pseudo}
 
-\tab \textbf{return} $b_\text{dim} + b_\text{pattern} + b_\text{template} + b_\text{labels}$\\
-}
+\subsection{Encoding motif instances}
 
-\textbf{function} $\text{i}(s, S)$:\\
-\tab \textbf{return} index of $s$ in $\text{sorted}(S)$\\
+To encode a list of instances $\I$ of a given pattern $M$, we generalize  the idea of the edgelist model described above.
+To understand the following ,it may be instructive to think of a single triple as an instance for the pattern
+\[
+\texttt{?n1 ?rel ?n2}.
+\]
+ We can now think of the edgelist model as storing a sequence of instances for this pattern, which together make up the entire knowledge graph.
 
-\end{pseudo}
+ To generalize this notion to arbitrary patterns, to be defined for a given template graph, we define the \emph{degree constraint} $D^\I$ of a list of instamces for a given pattern as follows: for each variable node $i$ in the pattern, the degree sequence provides an integer sequence $D^i$ of length $v_G$, indicating how often each node in the completed knowledge graph takes that position in the pattern.
+Similarly, for each variable link $j$ in the pattern, the degree sequence provides an integer sequence $C^j$ of length $r_G$ indicating for each relation how often it takes that position in the pattern.
+
+ We store these sequences in the same manner as the degree sequence of the template graph, using the Pitman-Yor code for each.
+
+Given this information, all we need to do is describe which of the possible sequences of matches for this pattern satisfying the given degree sequence we are encoding.
+As with the configuration model, the ideal is a uniform code over all possible configurations, for which we will define an approximation.
+Given $w$ variable nodes in a pattern, and $l$ variable links, we can define such a collection of instances using $w+l$ integer sequences: $N^1, \ldots, N^n, L^1, \ldots, L^l$, with the $t$-th instance defined by the integer tuple $(N^1_t, \ldots, N^n_t, L^1_t, \ldots, L^l_t)$.
+If this set of sequences satisfies the degree constraint, we know that node $q$ must occur $D^i_q$ times in sequence $N^i$, and similarly for the variable links.
+Let $\cS_\I$ be the set of all such integer sequences satisfying the degree constraint.
+We will follow the same logic as for the EdgeList model.
+Let $k$ be the number of matches of the pattern.
+We have:
+\begin{align*}
+|\cS| = &{k \choose D^1_1, \ldots, D^1_v}\times \ldots \times{k \choose D^w_1, \ldots, D^w_v}\times \\
+&{k \choose C^1_1, \ldots, C^1_r}\times \ldots \times{k \choose C^l_1, \ldots, C^l_r}
+\end{align*}
+
+As before, this set is larger than the set we are interested in.
+First, each set of pattern matches is contained multiple times (once for each permutation) and second, not all elements are valid pattern matches (in some, a single triple may be represented by multiple instances).
+Let $\cS_\I'$ be the subset representing only valid matches, and let $\G_\I$ be the set of valid instances with permutations removed.
+As before, we have
+\[
+|\G_\I| = |\cS'_\I| \frac{1}{k!} \leq  |\cS_\I| \frac{1}{k!} \text{.}
+\]
+Which gives us the following distribution
+\[
+p_M(G) =  \frac{k!}{|\cS_D|} < \frac{1}{\G_D} \text{,}
+\]
+ with $-\log p_M(\I)$ as a code to store the instances. Rewriting as before, gives us a codelength of 
+ \begin{align*}
+ - \log p_M(G) &= (w+l-1)\log(k!) \\
+  -&\sum_{j\in [1,w], i} \log(D^j_i!) -\sum_{j\in [1,l], i} \log(C^j_i!)
+ \end{align*}
+
+Note that if we store a graph with the pattern $\texttt{?n1 ?rel ?n2.}$, we obtain an empty template graph, and the motif code achieves the same codelength as the edgelist model, up to a small constant amount for storing the pattern.
+
+For a given graph and pattern, we can simply find the complete list of instances using a graph patter search.
+Since we require a slightly different semantics than standard graph pattern matchers, we adapt the DualIso algorithm \cite{saltz2014dualiso} for knowledge graph matching.
+Before computing the motif code, we prune the list of instances provided by this search iterating over the instances and removing and instance that produces a triple also produced by an earlier instance.
+
+Since graph matching encompasses subgraph isomorphism, it is known to be NP-Complete.
+To guard against rare patterns that produce long-running searches we terminate all searches after 5 seconds, returning  only those matches that were found within the time limit.
+
+We express the strength of a motif by its log-factor: $-\log p^\text{null}(G) + \log p^\text{motif}(G; M, \I, L^\text{base})$.
+If this value is positive, the motif code compresses the graph better than the null model.
+If the log-factor is greater than 10 bits, it corresponds to a rejection of the null model at $p < 0.001$.
 
@@
 
-Usually, we do not start with a given pattern $M$ and a set of instances $\I$.
-We have only the graph $G$, and we wish to find \emph{any} pattern that might be interesting.
-In order to do this efficiently, we generate candidate patterns by \emph{sampling} them from the data.
-This method, inspired by \cite{kashtan2004efficient} is based on the simple insight that when we sample a pattern from the data, we are more likely to sample a frequently occurring pattern than a infrequent one.
-If we sample $K$ times, and rank the results by relative frequency within the sample, we are likely to end up with a good indication of the most frequent patterns in the data.
-For unlabeled graphs, as few as 500 samples can be sufficient to find the most frequent subgraph.
+Ultimately, we want to find any motifs that have a high log-factor for a given graph $G$.
+Since we can readily compute the log-factor, any black-box optimization algorithm can be used to search the space of all possible motifs.
+For the sake of simplicity, we will use basic simulated annealing here.
+The starting pattern is always a single random triple from the graph, with its relation made a variable.
+We define seven possible transition from one pattern to another:
+\begin{description}
+	\item[Extend (0.1)] Choose an instance of the pattern and an adjacent triple not part of the instance. Add the triple to the pattern.
+	\item[Make a node a variable (3)] Choose a random constant node, and turn it in to a variable node.
+	\item[Make an edge a variable (3)] Choose a random constant edge label, and turn it in to a variable (always introducing a new variable).
+	\item[Make a variable node constant (2)] Choose a random variable node and turn it into a constant. Take the value from a random instance.
+	\item[Make a variable edge constant (2)] Choose a random variable edge and turn it into a constant. Take the value from a random instance.
+	\item[Remove edge (3)] Remove a random edge from the pattern, ensuring that it stays connected.
+	\item[Couple (1)] Take two distinct edge variables, which for at least one instance hold the same value and turn them into a single variable.
+\end{description}
 
-Theoretically, this allows us to generate promising candidates with time complexity independent of the size of the graph.
-Our algorithm samples a pattern in three steps.
-First, it samples a weakly connected subset of $n$ nodes, and extracts the \emph{induced subgraph} for these nodes.
-Second, it converts this induced subgraph to a non-induced pattern by randomly removing labels from some of the nodes and links and randomly removing some links entirely (as detailed below).
-Finally, we sort the nodes of this pattern into a canonical ordering, using the Nauty algorithm \cite{mckay1981practical} (so that we record isomorphic samples as the same pattern).
+The values in brackets are the relative weights of each transition type (chosen by trial and error to keep the algorithm stable).
+We choose a transition at random and compute a new pattern.
+If the transition cannot be made (for instance, there are no constant node to make variable) or if the resultant pattern is in some way invalid, we sample a new transition.
 
-For an induced subgraph with $n$ nodes and $e$ links, we can generate up to $n^2e^3$ different patterns: each node can be left labeled or made a variable node, each link can be removed, left unlabeled or be made variable.
-A uniform random choice among these would make patterns with many labels very unlikely.
-Instead, we sample $n' ~ U([0..
-n])$, $e' ~ U([0..
-e])$ and $e^*~ U([0..e'])$, with $n'$ the number of nodes to leave labeled, $e'$ the number of links to leave, and $e^*$ the number of links to leave labeled, respectively.
-If the resulting pattern is disconnected, we reject the sample.
+Once a new pattern has been sampled, we compare its codelength under the motif model to that of the previous sample.
+If the codelength is lower, we continue with the new pattern.
+If the codelentgh is longer, we continue with the new sample with probability $\alpha$ or return to the previous pattern otherwise.
+We use $\alpha = 0.5$ in all experiments.
 
-We sample $K$ such patterns, recording a dictionary which maps canonical patterns $M$ to a list of observed instances $\I'_M$ of $M$ in $G$.
-Note that we must store the \emph{links} of the occurrence in $\I'_M$, since the same pattern may emerge from a single induced subgraph in multiple ways.
+We store all patterns seen and their scores.
+In order to exploit all available processor cores, we run several searches in parallel.
+After each the searches are finished, we tajke the top 1000 patterns from each and sort them by motif codelength.
+If two searches return different codelengths\footnotemark for a motif we take the lowest value.
 
-Before we can can pass each pattern to the relevance test to see if it is a motif, we must first ensure that $\I'_M$ contains only link-disjoint instances.
-Assuming that all instances contribute equally to the compression, we would like to select the largest subset of $\I'_M$ consisting entirely of mutually disjoint sets.
-This is an instance of the Maximum Independent Set Problem, which is NP-Hard to solve optimally.
-Since we are aiming for scalability, we use a simple greedy algorithm to create our set $\I_M$ of disjoint instances: we include the first instance $I$ in $\I'_M$ in $\I_M$ and exclude any other instances that overlap with $I$.
-We then move to the next instance that has not already been included or excluded, and repeat the procedure.
-We continue until all instances have been included or excluded.
-
-We then pass each $M$ and $\I_M$ to the relevance test to see if they are motifs. For experiments reported here, we check only the 100 patterns with the largest number of instances.\footnote
-
-\footnotemark{In \cite{bloem2017large}, the number of edges between nodes inside the instance and outside it (the exdegree of the instance), was a strong predictor of whether the instances would benefit compression.
-Thus, the algorithm was strongly helped by pruning the list of instances using a ternary search.
-For the current code, the exdegree does not affect compression (since the instance nodes are not removed from the template graph).
-Therefore, we pass the entire instance list $\I_M$ to the relevance test without pruning it.}
-
-Our sampling algorithm is described more precisely in Algorithm~\ref{algorithm:sampling}.
-
-\begin{pseudo}[tb]
-\caption{The motif sampling algorithm.}
-
-\textbf{function} $\text{sample}(G, s_\text{min}, s_\text{max})$:\\
-\tab\tab a graph $G$, a minimum and maximum motif size $s_\text{min}$, $s_\text{max}$ \\
-
-\tab Choose $s$ uniformly from $[s_\text{min} .. s_\text{max}]$ \\
-\tab Choose a starting node $n$ uniformly from $V_G$ \\
-\tab $N \from \{n\}$ \\
-\tab \textbf{while} $|N| < s$:\\
-\tab \tab Choose $n$ uniformly from $N$\\
-\tab \tab Choose $n'$ uniformly from the neighbors of $n$\\
-\tab \tab $N \from N \cup n'$ \\
-\tab \tab After 500 cycles, sample a different starting node\\
-
-\tab $M \from G[N]$, the induced subgraph of $N$ \\
-\tab  Sample $n' ~ U([0.. n])$, $e' ~ U([0.. e])$ and $e^*~ U([0..e'])$\\
-\tab Randomly remove $e^*$ links, $e'$ linklabels and $n'$ nodelabels from $M$.\\
-\tab\tab (removed labels replaced by the next available negative integer)\\
-
-\tab \textbf{return} $\text{canonical}(M, N)$ \\
-
-\textbf{function} $\text{canonical}(G, I)$:\\
-\tab\tab a graph $G$, list of integers $I$\\
-\tab \textbf{return} $G$ with nodes in canonical order, $I$ permuted the same way
-\label{algorithm:sampling}
-\end{pseudo}
-
-\section{Null model}
-
-\label{section:null-model}
-
-The most common null model for classical motif analysis is the degree-sequence model (also known a the configuration model \cite{}): a uniform distribution over all graphs that share the same in and out degrees of the data for every node.
-We extend this to knowledge graphs by also including the frequency with which each relation occurs.
-Let a \emph{degree sequence} $D$ of length $n$ be a triple of three length-$n$ integer sequences: $(D^\text{in}, D^\text{rel}, D^\text{out})$.
-If $D$ is the degree sequence of a graph, then node $i$ has $D^\text{in}_i$ incoming links,  $D^\text{out}_i$ outgoing links and for each relation $r$, there are $D^\text{rel}_r$ links.
-
-Let $\G_D$ be the set of all graphs with degree sequence $D$. Then the configuration model can be expressed simply as
-\[
-p(G) = \frac{1}{|\G_D|}
-\]
-for any $G$ that satisfies $D$ and $p(G) = 0$ otherwise.
-Unfortunately, there is no efficient way to compute $|\G_D|$ and even approximations tend to be costly for large graphs.
-Following the approach in \cite{bloem2017large}, we define a fast approximation to the configuration model, which works well in practice for motif detection.
-
-We can define a knowledge graph by three length-$m$ integer sequences: $S$, $P$, $O$, with $\{(S_j, P_j, O_j)\}_j$ the graph's tripleset.
-If the graph satisfies degree sequence $D$, then we know that $S$ should contain node $j$ $D^\text{out}_j$ times, $P$ should contain relation $r$ $D^\text{rel}_r$ times and $O$ should contain node $j$ $D^\text{in}_j$ times.
-Let ${\cal S}_D$ be the set of all such triples of integer sequences satisfying $D$.
-We have
-\[
-|{\cal S}_D| =
- {m \choose {D_1^\text{out}, \ldots, D_n^\text{out}} }
- {m \choose {D_1^\text{rel}, \ldots, D_n^\text{rel}} }
- {m \choose {D_1^\text{in}, \ldots, D_n^\text{in}} } \text{.}
-\]
-
-While every member of ${\cal S}_D$ represents a valid graph satisfying $D$, many graphs are represented multiple times.
-Firstly, many elements of ${S}_D$ contain the same link multiple times, which means they are not in ${\cal G}_D$.
-We call the set without these elements ${\cal S'}_D \subset {\cal S}_D$.
-Secondly the links of the graph are listed in arbitrary order; if we apply the same permutation to all three lists $S$, $P$ and $O$, we get a new representation of the same graph.
-Since we know that any element in ${\cal S}'_D$ contains only unique triples, we know that each graph is present exactly $m!$ times.
-Thus, we have
-\[
-|\G_D| = |{\cal S}'_D| \frac{1}{m!} \leq  |{\cal S}_D| \frac{1}{m!} \text{.}
-\]
-
-We can thus use 
-\[
-p^\text{EL}_D(G) =  \frac{m!}{{m \choose {D_1^\text{out}, \ldots, D_n^\text{out}} }
- {m \choose {D_1^\text{rel}, \ldots, D_n^\text{rel}} }
- {m \choose {D_1^\text{in}, \ldots, D_n^\text{in}} }}
-\]
-as an approximation for the DS model.
-We call this the edge-list (EL) model.
-It always lower-bounds the true degree sequence model, since it affords some probability mass to graphs that cannot exist.\footnotemark ~Experiments in the classical motif setting have shown that the ES model is an acceptable proxy \cite{bloem2017large}, especially considering the extra scalability it affords.
-
-\footnotetext{Note that we cannot simply think of $p^\text{EL}$ as a uniform model for graphs containing multiple links, since in we can only divide by $m!$ if we know that all triples are unique.}
-
- \paragraph{Encoding D} In order to encode a graph with $L^\text{EL}_D$, we must first encode $D$.
-\footnote{Or, equivalently, to make $p^\text{EL}$ a complete distribution on all graphs, we must provide it with a prior on $D$.} For each of the three sequences in $D$ we use the following model:
-\begin{align*}
- p(D) &= \prod_i p^{\N}(D_i) \\
- L(D) &= - \sum_i \log p^{\N}(D_i)
-\end{align*}
-
-where $p^{\N}$ is a distribution on the natural numbers.
-This is an optimal encoding for $D$ assuming that its members are independently drawn from $p^{\N}$.
-When we use $p^\text{EL}$ as the null model, we use the data distribution for $p^{\N}$ to ensure that we have a lower bound to the optimal code-length.
-When we use $p^\text{EL}$ as part of the motif code, we must use a fair encoding, so we use a Dirichlet-Multinomial model to store each sequence in $D$.
+\footnotemark{This is possible if the pattern search is cut off due to the 5 second time limit.}
 
@@
 
-\subsection{Schema reconstruction}
+To validate the method, we first test it on random graphs.
+We sample a directed graph with a given number of nodes $n$ and edges $m$, with no self-connections and multiple edges (that is, we sample from from the $G(n, m)$ Erd\H{o}s-Renyi model).
+We then label the nodes uniformly at random with one of the relations in $0, \ldots, r$.
+To make the dimensions realistic we base them on thos of the Mutag dataset used later: $n=23644$, $m74567$, $r=24$.
 
-\subsection{Query induction}
+We then take one randomly chosen pattern, and insert $k$ instances of the pattern into the graph. We run the experiment for $k=0$, $k=10$ and $k=100$.
 
-\subsection{Scale}
+\section{Acknowledgments}
 
-To show the scale of our method, we show its performance on some very large graphs, culminating with a dump of the complete LOD cloud, containing 38 billion triples, from \cite{}.
+Withheld for purposes of blind reviewing.
+% This publication was supported by the Amsterdam Academic Alliance Data Science (AAA-DS) Program Award to the UvA and VU Universities.
 
-\subsection{Various}
-
-% Note: only if there's time. If not, save for later.
-
-In this section, we show some experiments that are not intended to substantiate any claims about our method, but rather to illustrate the variety of uses for succesfull motif detection.
-
-\subsubsection{Analogical reasoning}
-
-\subsubsection{Visualization}
-
-\subsubsection{Feature extraction}
-
-\subsubsection{Node embedding}
-
+% References and End of Paper
 \bibliography{kgmotifs}
+\bibliographystyle{aaai}
 

### 2018-09-05 kgmotifs/supplement.tex
deadline approaching

\documentclass[letterpaper]{article} %DO NOT CHANGE THIS

\usepackage{pdflscape}
\usepackage{times}     % Required
\usepackage{helvet}    % Required
\usepackage{courier}   % Required
\usepackage{url}       % Required
\usepackage{graphicx}  % Required
\frenchspacing         % Required
\setlength{\pdfpagewidth}{8.5in}  % Required
\setlength{\pdfpageheight}{11in}  % Required
\usepackage{longtable}
\usepackage{makecell}

% PDF Info Is Required:
\pdfinfo{
/Title (Supplement: Detecting Motifs in Knowledge Graphs using Compression)
/Author (...)}

\setcounter{secnumdepth}{0}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}
\usepackage{float}

\usepackage{easylist}
\usepackage{soul}

\newcommand{\N}{{\mathbb N}}
\newcommand{\R}{{\mathbb R}}
\newcommand{\Z}{{\mathbb Z}}

\newcommand{\X}{{\cal X}}
\newcommand{\B}{{\mathbb B}}
\newcommand{\G}{{\cal G}}
\newcommand{\cS}{{\cal S}}

\newcommand{\I}{{\cal I}}
\newcommand{\tab}{\hspace*{5mm}}
\newcommand{\from}{\leftarrow}

\renewcommand\cellalign{tl}

\floatstyle{ruled}
\newfloat{pseudo}{h}{lop}
\floatname{pseudo}{Algorithm}

\title{Supplement: Detecting Motifs in Knowledge Graphs using Compression}

\begin{document}
\maketitle 
\section{The canonical isomorphism algorithm for patterns}

During search, we want to re-order the nodes and links in an encountered pattern to some canonical ordering. This way, when we later return an isomorphic pattern, we recognize it, and will not need to compute its logfactor again. 
The \emph{nauty} algorithm is the standard solution for findings such \emph{canonical isomorphs}. As we are sampling partially labeled, directed graphs, we have a lot of information that can speed up canonicalization. Since this step is usually the bottleneck in sampling, it's important to perform it as efficiently as possible. Below, we provide a very succinct description of how nauty works, when applied to our patterns.

Let $V'_M$ and $R'_M$ represent the labels of the variable nodes and links in $M$. Pattern $M$ is an isomorphism of pattern $M'$ if there exist bijections $\sigma_v : V'_M \times V'M$ and $\sigma_r: \R'_M \times \R'_M$ such that applying these to the variable elements of $M$ yields $M'$ (under graph equivalence). The set of all patterns isomorphic to $G$ is the isomorphism class $[G]$. A function $C(P)$ is a canonicalization function if it produces for all any pattern $M$ in $[M]$ the same, canonical member of $M$.

Note that a very simple canonicalization algorithm would be to pass the unlabeled graph to the Nauty algorithm, and use its canonical ordering on the labeled graph. However, the labels can make the canonicalization procedure significantly faster.

The key principle behind the Nauty algorithm is the use of a \emph{colouring}: a partition of the nodes of a graph. A \emph{safe colouring} is defined as a partition $\pi$ of the nodes of the graph such that any permutation that maps nodes from different partitions to one another is guaranteed \emph{not} to be an automorphism. No guarantees are made about nodes in the same cell. For instance, the partition that puts all nodes in the same cell is always safe. If two nodes have different degrees, we can put them in different cells and be sure that we have a safe colouring. Given one safe colouring $\pi$, if node $a$ from partition $i$ has a neighbor in partition $j$ and node $b$ from partition $j$ has no such node, we can \emph{refine} the partition $\pi$ by separating $i$ into a new partition (one contain $a$ and one containing $b$).

The nauty algorithm operates on a graph with a given colouring (with the unit colouring used by default). It first refines the colouring as much as possible, using degree information, and propagating the distinguishing features through the graph. Once the most refined colouring has been achieved, it searches through all autorphisms of the graph to produce a canonical example (using various optimizations to prune the search tree).

In our case, we use the colouring to indicate what our labeling tells us about the colouring of the nodes.
Note the following:
\begin{easylist}
§ Labeled nodes in the pattern can be placed in a discrete cell of the colouring.
§ Variable nodes should be placed into a single cell in the partition.
§ We define a partition over nodes \emph{and} relation labels. 
\end{easylist}

\begin{landscape}

\section*{Motifs found for the three datasets}

\subsection{Dogfood, top 100 by log factor}

\begin{longtable}{ r r p{18cm} }
 log factor & frequency & \\
\hline\endhead
 28232.1 & 5807 & \makecell{\texttt{?n1 swc:heldBy ?n2.} } \\ 
 5125.5 & 2538 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
 3997.0 & 2180 & \makecell{\texttt{?n1 swc:isPartOf ?n2.} } \\ 
 1755.1 & 566 & \parbox{2cm}{\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_tzid ?n2.} } \\ 
 378.1 & 16651 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 rdf:\_1 ?n3.} } \\ 
 201.2 & 539 & \makecell{\texttt{?n1 swc:hasRole ?n2.} \\\texttt{?n2 rdf:type swc:Presenter.} } \\ 
 -2723.1 & 466 & \makecell{\texttt{?n1 rdf:type swc:Presenter.} } \\ 
 -2994.2 & 2031 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
 -3191.3 & 30 & \makecell{dsw:conference/eswc/2006/sessions/poster-session swc:hasRole ?n1. \\\texttt{?n1 rdf:type swc:Presenter.} } \\ 
 -3227.7 & 154 & \makecell{\texttt{dsw:conference/iswc/2008/research/pcmember swc:heldBy ?n1.} } \\ 
 -3271.9 & 156 & \makecell{\texttt{dsw:conference/iswc/2008/research/pcmember ?p2 ?n1.} } \\ 
 -3415.6 & 49 & \texttt{dsw:conference/dc/2010/programme-committee-member swc:heldBy ?n1.} \\ 
 -3514.0 & 138 & \makecell{\texttt{?n1 ?p2 dsw:person/eero-hyvoenen.} } \\ 
 -3596.0 & 29 & \makecell{\texttt{dsw:conference/iswc/2008/in\_use/pcmember swc:heldBy ?n1.} } \\ 
 -3605.8 & 33 & \makecell{\texttt{?n1 ?p2 dsw:person/heiko-paulheim.} } \\ 
 -3608.1 & 31 & \makecell{\texttt{dsw:conference/eswc/2012/poster-track-committee-member swc:heldBy ?n1.} } \\ 
 -3608.5 & 96 & \makecell{\texttt{\_:genid8 ?p2 ?n1.} } \\ 
 -3619.3 & 81 & \makecell{\texttt{\_:genid14 ?p2 ?n1.} } \\ 
 -3637.2 & 55 & \makecell{\texttt{\_:genid37 ?p2 ?n1.} } \\ 
 -3638.1 & 60 & \makecell{\texttt{\_:genid17 ?p2 ?n1.} } \\ 
 -3642.2 & 36 & \makecell{\texttt{dsw:conference/iswc/2010/dc/pcmember swc:heldBy ?n1.} } \\ 
 -3643.8 & 47 & \makecell{\texttt{?n1 ?p2 dsw:person/kim-viljanen.} } \\ 
 -3644.5 & 46 & \makecell{\texttt{?n1 ?p2 dsw:person/bernhard-schandl.} } \\ 
 -3656.0 & 1 & \makecell{\texttt{dsw:conference/eswc/2010/paper/services/3/authorlist rdf:\_1 dsw:person/heiko-paulheim.} } \\ 
 -3658.5 & 1 & \makecell{\texttt{\_:genid12 rdf:\_10 dsw:person/michael-pool.} } \\ 
 -3660.9 & 1 & \makecell{\texttt{\_:genid14 rdf:\_1 dsw:person/siegfried-handschuh.} } \\ 
 -3660.9 & 19 & \makecell{\texttt{?n1 ?p2 dsw:person/xi-bai.} } \\ 
 -3662.3 & 1 & \makecell{dsw:person/doroteo-t-toledano foaf:name "Doroteo T. Toledano". } \\ 
 -3662.8 & 34 & \makecell{\texttt{dsw:conference/eswc/2012/poster-track-committee-member ?p2 ?n1.} } \\ 
 -3662.9 & 1 & \makecell{\texttt{dsw:person/doroteo-t-toledano rdfs:label "Doroteo T. Toledano".} } \\ 
 -3663.6 & 1 & \makecell{\texttt{dsw:conference/www/2012/demo/demo-chair swc:heldBy dsw:person/bebo-white.} } \\ 
 -3664.9 & 1 & \makecell{\texttt{\_:genid17 rdf:\_1 dsw:person/john-barton.} } \\ 
 -3665.5 & 1 & \makecell{dsw:organization/czech-technical-university-in-prague foaf:member dsw:person/petr-pollak. } \\ 
 -3665.9 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/431/authorlist rdf:\_1 dsw:person/alexandre-passant.} } \\ 
 -3666.2 & 1 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 dc:creator dsw:person/violeta-damjanovic.} } \\ 
 -3666.4 & 1 & \makecell{\texttt{\_:genid37 rdf:\_2 dsw:person/stephen-larson.} } \\ 
 -3666.7 & 1 & \makecell{\texttt{\_:genid117 rdf:\_1 dsw:person/tomas-vitvar.} } \\ 
 -3666.8 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember rdf:type swc:OrganisingCommitteeMember.} } \\ 
 -3666.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/468/authorlist rdf:\_1 dsw:person/jerome-euzenat.} } \\ 
 -3666.8 & 1 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 dc:creator dsw:person/raluca-zaharia.} } \\ 
 -3666.8 & 1 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 dc:creator dsw:person/laurentiu-vasiliu.} } \\ 
 -3667.2 & 1 & \makecell{\texttt{dsw:conference/iswc/2008/paper/poster\_demo/43 foaf:maker dsw:person/malte-kiesel.} } \\ 
 -3667.2 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember swc:heldBy dsw:person/achille-fokoue.} } \\ 
 -3667.6 & 3 & \makecell{\texttt{dsw:conference/lrec/2008/papers/421 ?p1 dsw:person/petr-pollak.} } \\ 
 -3667.6 & 1 & \makecell{\texttt{dsw:conference/lrec/2008/papers/421 dc:creator dsw:person/petr-pollak.} } \\ 
 -3667.6 & 1 & \makecell{\texttt{dsw:workshop/iwod/2008/main/chair/organisingcommitteemember rdf:type swc:OrganisingCommitteeMember.} } \\ 
 -3667.7 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/poster-track-committee-member swc:heldBy dsw:person/tomas-vitvar.} } \\ 
 -3667.8 & 1 & \makecell{\texttt{\_:genid39 rdf:\_1 dsw:person/tomas-vitvar.} } \\ 
 -3667.9 & 3 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 ?p1 dsw:person/raluca-zaharia.} } \\ 
 -3667.9 & 3 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 ?p1 dsw:person/laurentiu-vasiliu.} } \\ 
 -3668.0 & 1 & \makecell{\texttt{dsw:workshop/swsc/2008/main/pcmember swc:heldBy dsw:person/tomas-vitvar.} } \\ 
 -3668.1 & 59 & \makecell{\texttt{?n1 ?p2 dsw:person/achille-fokoue.} } \\ 
 -3668.3 & 1 & \makecell{\texttt{\_:genid38 rdf:\_1 dsw:person/nicola-fanizzi.} } \\ 
 -3668.3 & 1 & \makecell{\texttt{dsw:conference/lrec/2008/papers/301 swrs:author dsw:person/hiroyuki-shinnou.} } \\ 
 -3669.0 & 1 & \makecell{\texttt{\_:genid37 rdf:\_2 dsw:person/hanmin-jung.} } \\ 
 -3670.3 & 1 & \makecell{\texttt{dsw:conference/iswc/2008/research/pcmember swc:heldBy dsw:person/chiara-ghidini.} } \\ 
 -3670.6 & 13 & \makecell{\texttt{?n1 rdf:type swc:OrganisingCommitteeMember.} } \\ 
 -3670.9 & 80 & \makecell{\texttt{\_:genid12 ?p2 ?n1.} } \\ 
 -3671.1 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/dc/pcmember swc:heldBy dsw:person/andreas-harth.} } \\ 
 -3671.2 & 18 & \makecell{\texttt{\_:genid17 rdf:\_1 ?n1.} } \\ 
 -3673.4 & 3 & \makecell{\texttt{dsw:conference/iswc/2008/paper/poster\_demo/43 ?p1 dsw:person/malte-kiesel.} } \\ 
 -3673.7 & 13 & \makecell{\texttt{?n1 ?p2 dsw:person/hiroyuki-shinnou.} } \\ 
 -3674.2 & 3 & \makecell{\texttt{dsw:conference/lrec/2008/papers/301 ?p1 dsw:person/hiroyuki-shinnou.} } \\ 
 -3674.7 & 69 & \makecell{\texttt{?n1 ?p2 dsw:person/andreas-harth.} } \\ 
 -3675.6 & 20 & \makecell{\texttt{?n1 ?p2 dsw:person/jianfeng-du.} } \\ 
 -3676.1 & 1 & \makecell{\texttt{dsw:person/akira-mori rdf:type foaf:Person.} } \\ 
 -3677.1 & 1 & \makecell{\texttt{dsw:person/alberto-musetti rdf:type foaf:Person.} } \\ 
 -3677.9 & 1 & \makecell{\texttt{dsw:person/alejandro-jaimes rdf:type foaf:Person.} } \\ 
 -3678.1 & 1 & \makecell{\texttt{dsw:conference/eswc/2010/paper/services/3/authorlist ?p1 dsw:person/heiko-paulheim.} } \\ 
 -3678.7 & 13 & \makecell{\texttt{\_:genid38 rdf:\_1 ?n1.} } \\ 
 -3678.8 & 1 & \makecell{\texttt{dsw:person/georg-lausen rdf:type foaf:Person.} } \\ 
 -3679.2 & 2 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 ?p1 dsw:person/violeta-damjanovic.} } \\ 
 -3679.5 & 10 & \makecell{\texttt{\_:genid8 rdf:\_4 ?n1.} } \\ 
 -3679.7 & 10 & \makecell{\texttt{\_:genid59 rdf:\_2 ?n1.} } \\ 
 -3680.2 & 2 & \makecell{\texttt{dsw:person/doroteo-t-toledano ?p1 "Doroteo T. Toledano".} } \\ 
 -3680.3 & 13 & \makecell{\texttt{\_:genid39 rdf:\_1 ?n1.} } \\ 
 -3680.5 & 1 & \makecell{\texttt{\_:genid12 ?p1 dsw:person/michael-pool.} } \\ 
 -3680.7 & 3 & \makecell{\texttt{dsw:conference/iswc/2009/paper/research/347 ?p1 dsw:person/bernhard-schandl.} } \\ 
 -3682.2 & 33 & \makecell{\texttt{\_:genid38 ?p2 ?n1.} } \\ 
 -3682.9 & 1 & \makecell{\texttt{\_:genid14 ?p1 dsw:person/siegfried-handschuh.} } \\ 
 -3683.8 & 13 & \makecell{\texttt{\_:genid37 rdf:\_2 ?n1.} } \\ 
 -3685.6 & 1 & \makecell{\texttt{dsw:conference/www/2012/demo/demo-chair ?p1 dsw:person/bebo-white.} } \\ 
 -3685.7 & 1 & \makecell{\texttt{\_:genid150 ?p1 dsw:person/bettina-bauer-messmer.} } \\ 
 -3686.3 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/semantic-web-in-use/60/authorlist ?p1 dsw:person/xi-bai.} } \\ 
 -3686.8 & 169 & \makecell{\texttt{\_:genid38 rdf:\_1 ?n2.} \\\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid38.} } \\ 
 -3687.0 & 43 & \makecell{\texttt{\_:genid42 ?p2 ?n1.} } \\ 
 -3687.1 & 1 & \makecell{\texttt{dsw:conference/eswc/2010/social\_web/pcmember ?p1 dsw:person/eero-hyvoenen.} } \\ 
 -3687.2 & 3 & \makecell{\texttt{dsw:conference/eswc/2008/paper/54 ?p1 dsw:person/eero-hyvoenen.} } \\ 
 -3687.5 & 1 & \makecell{\texttt{dsw:organization/czech-technical-university-in-prague ?p1 dsw:person/petr-pollak.} } \\ 
 -3688.1 & 1 & \makecell{\texttt{dsw:conference/rr/2008/full/4/authors ?p1 dsw:person/jianfeng-du.} } \\ 
 -3688.4 & 1 & \makecell{\texttt{\_:genid37 ?p1 dsw:person/stephen-larson.} } \\ 
 -3688.7 & 1 & \makecell{\texttt{\_:genid117 ?p1 dsw:person/tomas-vitvar.} } \\ 
 -3688.7 & 2 & \makecell{\texttt{\_:genid8 ?p1 dsw:person/irene-celino.} } \\ 
 -3688.8 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember ?p1 swc:OrganisingCommitteeMember.} } \\ 
 -3688.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/468/authorlist ?p1 dsw:person/jerome-euzenat.} } \\ 
 -3689.0 & 13 & \makecell{\texttt{\_:genid37 rdf:\_1 ?n1.} } \\ 
 -3689.1 & 1 & \makecell{\texttt{\_:genid36 ?p1 dsw:person/benjamin-coe.} } \\ 
 -3689.2 & 2 & \makecell{\texttt{dsw:conference/eswc/2011/paper/in-use/52 ?p1 dsw:person/eero-hyvoenen.} } \\ 
 -3689.3 & 1 & \makecell{\texttt{dsw:conference/eswc/2006/demo-papers/FD01-hitzler ?p1 dsw:person/malte-kiesel.} } \\ 
 -3689.3 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember ?p1 dsw:person/achille-fokoue.} } \\ 
 -3689.7 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/poster-track-committee-member ?p1 dsw:person/tomas-vitvar.} } \\ 
 -3689.8 & 8 & \makecell{\texttt{\_:genid96 rdf:\_1 ?n1.} } \\ 
 -3690.1 & 1 & \makecell{\texttt{dsw:workshop/swsc/2008/main/pcmember ?p1 dsw:person/tomas-vitvar.} } \\ 
 -3690.3 & 1 & \makecell{\texttt{\_:genid38 ?p1 dsw:person/nicola-fanizzi.} } \\ 
 -3691.0 & 1 & \makecell{\texttt{\_:genid37 ?p1 dsw:person/hanmin-jung.} } \\ 
 -3692.3 & 3 & \makecell{\texttt{?n1 swc:heldBy dsw:person/bebo-white.} } \\ 
 -3692.9 & 47 & \makecell{\texttt{\_:genid36 ?p2 ?n1.} } \\ 
 -3693.1 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/dc/pcmember ?p1 dsw:person/andreas-harth.} } \\ 
 -3694.6 & 1 & \makecell{\texttt{?n1 rdf:\_10 dsw:person/michael-pool.} } \\ 
 -3694.6 & 1 & \makecell{\texttt{\_:genid12 rdf:\_10 ?n1.} } \\ 
 -3695.0 & 2 & \makecell{\texttt{?n1 swc:heldBy dsw:person/marco-neumann.} } \\ 
 -3695.4 & 13 & \makecell{\texttt{\_:genid38 rdf:\_1 ?n1.} \\\texttt{dsw:conference/iswc/2008/paper/poster\_demo/18 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid38.} } \\ 
 -3695.9 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/hiroyuki-shinnou.} } \\ 
 -3696.4 & 39 & \makecell{\texttt{\_:genid39 ?p2 ?n1.} } \\ 
 -3696.7 & 6 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/45 foaf:maker ?n1.} } \\ 
 -3697.1 & 1 & \makecell{\texttt{dsw:conference/www/2012/dev/16/talk swc:hasRole ?n1.} \\\texttt{?n1 rdf:type swc:Presenter.} } \\ 
 -3697.1 & 2 & \makecell{\texttt{dsw:person/doroteo-t-toledano foaf:made ?n1.} } \\ 
 -3697.6 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/bruno-alves.} } \\ 
 -3698.0 & 1 & \makecell{\texttt{dsw:conference/dc/2010/tutorial/2/tutorial-presenter rdf:type swc:Presenter.} \\\texttt{?n1 swc:hasRole dsw:conference/dc/2010/tutorial/2/tutorial-presenter.} } \\ 
 -3698.1 & 1 & \makecell{\texttt{dsw:person/akira-mori ?p1 foaf:Person.} } \\ 
 -3698.5 & 1 & \makecell{\texttt{?n1 foaf:name "Doroteo T. Toledano".} } \\ 
 -3698.8 & 1 & \makecell{\texttt{dsw:person/michalis-foukarakis ?p1 foaf:Person.} } \\ 
 -3699.0 & 1 & \makecell{\texttt{?n1 rdfs:label "Doroteo T. Toledano".} } \\ 
 -3699.0 & 1 & \makecell{\texttt{dsw:person/doroteo-t-toledano rdfs:label ?n1.} } \\ 
 -3699.2 & 1 & \makecell{\texttt{dsw:person/alberto-musetti ?p1 foaf:Person.} } \\ 
 -3699.2 & 2 & \makecell{\texttt{\_:genid152 rdf:\_1 ?n1.} } \\ 
 -3699.3 & 1 & \makecell{\texttt{dsw:person/michalis-foukarakis foaf:name ?n1.} } \\ 
 -3699.7 & 1 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 rdfs:label ?n1.} } \\ 
 -3699.8 & 1 & \makecell{\texttt{dsw:person/michalis-foukarakis rdfs:label ?n1.} } \\ 
 -3699.9 & 1 & \makecell{\texttt{dsw:person/alejandro-jaimes ?p1 foaf:Person.} } \\ 
 -3699.9 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/paper/in-use-industrial/138/authorlist rdf:\_1 ?n1.} } \\ 
 -3700.1 & 1 & \makecell{\texttt{dsw:person/alain-leger ?p1 foaf:Person.} } \\ 
 -3700.2 & 4 & \makecell{\texttt{\_:genid117 rdf:\_1 ?n1.} } \\ 
 -3700.4 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/nikos-loutas.} } \\ 
 -3700.5 & 4 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember swc:heldBy ?n1.} } \\ 
 -3700.8 & 1 & \makecell{\texttt{dsw:person/georg-lausen ?p1 foaf:Person.} } \\ 
 -3700.9 & 1 & \makecell{\texttt{?n1 rdfs:label "Graph Kernels for RDF data".} } \\ 
 -3700.9 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/petr-pollak.} } \\ 
 -3701.1 & 2 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/aidan-boran.} } \\ 
 -3701.8 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-633 rdfs:label ?n1.} } \\ 
 -3702.1 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/431/authorlist rdf:\_1 ?n1.} } \\ 
 -3702.1 & 2 & \makecell{\texttt{?n1 foaf:maker dsw:person/bettina-bauer-messmer.} } \\ 
 -3702.1 & 2 & \makecell{\texttt{dsw:conference/www/2012/eu/10 swc:isPartOf ?n1.} } \\ 
 -3702.3 & 1 & \makecell{\texttt{?n1 dc:creator dsw:person/violeta-damjanovic.} } \\ 
 -3702.3 & 1 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 dc:creator ?n1.} } \\ 
 -3702.5 & 37 & \makecell{\texttt{?n1 ?p2 dsw:person/chiara-ghidini.} } \\ 
 -3702.5 & 1 & \makecell{\texttt{?n1 rdf:\_2 dsw:person/stephen-larson.} } \\ 
 -3702.8 & 2 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/siegfried-handschuh.} } \\ 
 -3702.9 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember rdf:type ?n1.} } \\ 
 -3702.9 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/468/authorlist rdf:\_1 ?n1.} } \\ 
 -3703.0 & 1 & \makecell{\texttt{?n1 dc:creator dsw:person/raluca-zaharia.} } \\ 
 -3703.4 & 4 & \makecell{\texttt{?n1 foaf:maker dsw:person/malte-kiesel.} } \\ 
 -3703.5 & 1 & \makecell{\texttt{?n1 dc:creator dsw:person/naimul-basher.} } \\ 
 -3703.6 & 37 & \makecell{\texttt{\_:genid94 ?p2 ?n1.} } \\ 
 -3703.8 & 4 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 dc:creator ?n1.} } \\ 
 -3704.0 & 1 & \makecell{\texttt{?n1 dc:creator dsw:person/vinhtuan-thai.} } \\ 
 -3704.3 & 2 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/jerome-euzenat.} } \\ 
 -3704.7 & 39 & \makecell{\texttt{\_:genid62 ?p2 ?n1.} } \\ 
 -3705.4 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/tomas-vitvar.} } \\ 
 -3705.7 & 3 & \makecell{\texttt{?n1 swc:heldBy dsw:person/tomas-vitvar.} } \\ 
 -3706.6 & 2 & \makecell{\texttt{?n1 foaf:maker dsw:person/aidan-boran.} } \\ 
 -3706.9 & 3 & \makecell{\texttt{?n1 swrs:author dsw:person/hiroyuki-shinnou.} } \\ 
 -3708.8 & 2 & \makecell{\texttt{\_:genid65 http://www.w3.org/2002/12/cal/ical\_tzid ?n1.} } \\ 
 -3709.0 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-766/authors rdf:type ?n1.} } \\ 
 -3710.4 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/achille-fokoue.} } \\ 
 -3710.8 & 2 & \makecell{\texttt{?n1 swc:heldBy dsw:person/aidan-boran.} } \\ 
 -3711.7 & 6 & \makecell{\texttt{?n1 ?p2 dsw:person/bruno-alves.} } \\ 
 -3714.2 & 1 & \makecell{\texttt{dsw:conference/eswc/2010/paper/services/3/authorlist ?p2 ?n1.} } \\ 
 -3714.3 & 5 & \makecell{\texttt{?n1 swc:heldBy dsw:person/chiara-ghidini.} } \\ 
 -3714.9 & 1 & \makecell{\texttt{dsw:person/georg-lausen rdf:type ?n1.} } \\ 
 -3715.9 & 12 & \makecell{\texttt{?n1 rdf:\_3 dsw:person/eero-hyvoenen.} } \\ 
 -3716.7 & 5 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/andreas-harth.} } \\ 
 -3717.0 & 9 & \makecell{\texttt{dsw:person/georg-lausen swc:holdsRole ?n1.} } \\ 
 -3717.4 & 13 & \makecell{\texttt{\_:genid38 rdf:\_1 ?n1.} \\\texttt{dsw:conference/iswc/2008/paper/poster\_demo/18 ?p2 \_:genid38.} } \\ 
 -3718.4 & 8 & \makecell{\texttt{?n1 swc:heldBy dsw:person/achille-fokoue.} } \\ 
 -3719.7 & 37 & \makecell{\texttt{\_:genid59 ?p2 ?n1.} } \\ 
 -3719.9 & 13 & \makecell{\texttt{?n1 ?p2 dsw:person/aidan-boran.} } \\ 
 -3720.5 & 20 & \makecell{\texttt{?n1 ?p2 dsw:person/tomas-vitvar.} } \\ 
 -3721.7 & 4 & \makecell{\texttt{dsw:conference/eswc/2012/paper/in-use-industrial/138/authorlist ?p2 ?n1.} } \\ 
 -3723.0 & 9 & \makecell{\texttt{?n1 dc:creator dsw:person/andreas-harth.} } \\ 
 -3724.0 & 12 & \makecell{\texttt{?n1 swc:heldBy dsw:person/bernhard-schandl.} } \\ 
 -3726.0 & 58 & \makecell{\texttt{dsw:conference/iswc/2009/paper/inuse/143 dc:creator dsw:person/achille-fokoue.} \\\texttt{?n1 ?p2 dsw:person/achille-fokoue.} } \\ 
 -3726.2 & 26 & \makecell{\texttt{?n1 swrs:author dsw:person/eero-hyvoenen.} } \\ 
 -3726.4 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/petr-pollak.} } \\ 
 -3727.7 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/benjamin-coe.} } \\ 
 -3728.1 & 24 & \makecell{\texttt{?n1 swc:heldBy dsw:person/mathieu-daquin.} } \\ 
 -3728.9 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/naimul-basher.} } \\ 
 -3729.6 & 0 & \makecell{\texttt{?n1 ?p3 ?n2.} \\\texttt{?n2 rdf:type swc:Presenter.} } \\ 
 -3730.0 & 2 & \makecell{\texttt{dsw:conference/iswc/2010/paper/445/authorlist ?p2 ?n1.} } \\ 
 -3731.5 & 30 & \makecell{\texttt{\_:genid96 ?p2 ?n1.} } \\ 
 -3731.5 & 4 & \makecell{\texttt{?n1 ?p2 dsw:person/michael-pool.} } \\ 
 -3731.7 & 2 & \makecell{\texttt{dsw:conference/iswc/2010/paper/431/authorlist ?p2 ?n1.} } \\ 
 -3731.9 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/raluca-zaharia.} } \\ 
 -3732.4 & 13 & \makecell{\texttt{dsw:conference/iswc/2008/paper/poster\_demo/18 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n1.} \\\texttt{?n1 rdf:\_1 ?n2.} } \\ 
 -3732.6 & 6 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember ?p2 ?n1.} } \\ 
 -3733.0 & 3 & \makecell{\texttt{?n2 rdfs:label "Graph Kernels for RDF data".} \\\texttt{?n2 dc:creator ?n1.} } \\ 
 -3735.3 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/julien-gaugaz.} } \\ 
 -3738.0 & 45 & \makecell{\texttt{http://www.eswc2007.org/pdf/eswc07-dellavalle.pdf rdf:type ?n2.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -3738.4 & 2 & \makecell{\texttt{?n1 ?p2 dsw:person/sukwon-chung.} } \\ 
 -3739.9 & 5 & \makecell{\texttt{\_:bNode14 ?p2 ?n1.} } \\ 
 -3740.6 & 7 & \makecell{\texttt{?n1 ?p2 dsw:person/bettina-bauer-messmer.} } \\ 
 -3741.2 & 5 & \makecell{\texttt{dsw:conference/iswc/2010/paper/489/authorlist ?p2 ?n1.} } \\ 
 -3741.4 & 6 & \makecell{\texttt{dsw:conference/iswc/2010/paper/261/authorlist ?p2 ?n1.} } \\ 
 -3742.5 & 36 & \makecell{\texttt{\_:genid99 ?p2 ?n1.} } \\ 
 -3743.8 & 16 & \makecell{\texttt{\_:genid121 ?p2 ?n1.} } \\ 
 -3744.2 & 0 & \makecell{\texttt{?n1 ?p4 ?n2.} \\\texttt{?n2 rdf:\_1 ?n3.} } \\ 
 -3746.9 & 4 & \makecell{\texttt{?n1 ?p2 dsw:person/stephen-larson.} } \\ 
 -3749.0 & 7 & \makecell{\texttt{dsw:conference/www/2012/demo/demo-chair ?p2 ?n1.} } \\ 
 -3749.7 & 20 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-633 ?p2 ?n1.} } \\ 
 -3750.6 & 13 & \makecell{\texttt{\_:genid152 ?p2 ?n1.} } \\ 
 -3751.7 & 29 & \makecell{\texttt{\_:genid65 ?p2 ?n1.} } \\ 
 -3752.6 & 12 & \makecell{\texttt{?n1 ?p2 dsw:person/vinhtuan-thai.} } \\ 
 -3753.1 & 6 & \makecell{\texttt{dsw:person/doroteo-t-toledano ?p2 ?n1.} } \\ 
 -3755.1 & 59 & \makecell{\texttt{dsw:conference/iswc-aswc/2007/tracks/posters-demos/papers/375 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
 -3755.4 & 9 & \makecell{\texttt{?n1 ?p2 dsw:person/violeta-damjanovic.} } \\ 
 -3759.6 & 34 & \makecell{\texttt{\_:genid69 ?p2 ?n1.} } \\ 
 -3761.7 & 9 & \makecell{\texttt{dsw:person/alberto-musetti ?p2 ?n1.} } \\ 
 -3763.9 & 7 & \makecell{\texttt{dsw:person/michalis-foukarakis ?p2 ?n1.} } \\ 
 -3764.3 & 17 & \makecell{\texttt{\_:genid150 ?p2 ?n1.} } \\ 
 -3769.0 & 15 & \makecell{\texttt{?n1 ?p2 dsw:person/malte-kiesel.} } \\ 
 -3770.1 & 10 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 ?p2 ?n1.} } \\ 
 -3772.9 & 45 & \makecell{\texttt{?n1 ?p2 dsw:person/elena-paslaru-bontas-simperl.} } \\ 
 -3780.1 & 18 & \makecell{\texttt{dsw:conference/iswc/2008/paper/poster\_demo/43 ?p2 ?n1.} } \\ 
 -3780.2 & 18 & \makecell{\texttt{dsw:conference/eswc/2006/demo-papers/FD01-hitzler ?p2 ?n1.} } \\ 
 -3784.0 & 1290 & \makecell{\texttt{?n1 rdf:type ?n3.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
 -3797.8 & 12 & \makecell{\texttt{dsw:person/alejandro-jaimes ?p2 ?n1.} } \\ 
 -3800.4 & 26 & \makecell{\texttt{dsw:person/georg-lausen ?p2 ?n1.} } \\ 
 -3833.3 & 233 & \makecell{\texttt{?n1 ?p5 ?n3.} \\\texttt{?n3 ?p4 ?n2.} } \\ 
 -3902.5 & 47 & \makecell{\texttt{?n1 rdf:\_10 ?n2.} } \\ 
 -3993.4 & 405 & \makecell{\texttt{?n1 ?p4 ?n2.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
 -4021.2 & 2102 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
 -4106.8 & 156 & \makecell{\texttt{?n1 rdfs:label ?n2.} } \\ 
 -4204.5 & 373 & \makecell{\texttt{?n1 ?p4 ?n3.} \\\texttt{?n2 ?p4 ?n3.} } \\ 
 -4750.2 & 2914 & \makecell{\texttt{?n1 rdf:\_1 ?n2.} } \\ 
 -4761.6 & 248 & \makecell{\texttt{?n1 ?p2 foaf:Person.} } \\ 
 -4802.7 & 293 & \makecell{\texttt{?n1 rdf:type foaf:Person.} } \\ 
 -5632.4 & 301 & \makecell{\texttt{?n1 foaf:name ?n2.} } \\ 
 -6987.6 & 3047 & \makecell{\texttt{?n1 rdf:\_2 ?n2.} } \\ 
 -9429.5 & 2385 & \makecell{\texttt{?n1 foaf:maker ?n2.} } \\ 
 -10493.1 & 2528 & \makecell{\texttt{?n1 dc:creator ?n2.} } \\ 
\hline
\end{longtable}

\subsection{Dogfood, top 500 by frequency}

\begin{longtable}{ r r p{19cm} }
 log factor & frequency & \\
\hline\endhead
 -5052.5 & 16582 & \makecell{\texttt{?n1 rdf:\_1 ?n3.} \\\texttt{?n1 rdf:\_1 ?n2.} } \\ 
 -4021.9 & 6804 & \makecell{\texttt{\_:genid9 ?p4 ?n1.} \\\texttt{\_:genid9 ?p3 ?n2.} } \\ 
 28232.1 & 5807 & \makecell{\texttt{?n1 swc:heldBy ?n2.} } \\ 
 122602.4 & 5709 & \makecell{\texttt{?n1 swc:heldBy ?n2.} \\\texttt{?n2 swc:holdsRole ?n1.} } \\ 
 -3801.0 & 3734 & \makecell{\texttt{?n1 ?p3 dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{?n2 ?p3 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -4178.4 & 3157 & \makecell{\texttt{?n2 ?p4 ?n1.} \\\texttt{?n2 rdf:\_4 ?n3.} } \\ 
 -7299.7 & 2799 & \makecell{\texttt{?n1 rdf:\_2 ?n2.} } \\ 
 -4442.9 & 2449 & \makecell{\texttt{?n1 rdf:\_1 ?n2.} } \\ 
 -4599.6 & 2393 & \makecell{\texttt{?n1 rdf:\_1 ?n4.} \\\texttt{?n2 rdf:\_1 ?n4.} \\\texttt{?n2 rdf:\_1 ?n3.} } \\ 
 -9551.9 & 2193 & \makecell{\texttt{?n1 foaf:maker ?n2.} } \\ 
 6218.0 & 1939 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
 -3804.8 & 1938 & \makecell{\texttt{dsw:conference/iswc-aswc/2007/chair\_12 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n2.} } \\ 
 -1305.2 & 1713 & \makecell{\texttt{?n1 http://purl.org/dc/terms/creator ?n2.} } \\ 
 -15966.3 & 1672 & \makecell{\texttt{?n1 http://purl.org/ontology/bibo/authorList ?n2.} } \\ 
 3849.2 & 1663 & \makecell{\texttt{?n1 swc:holdsRole ?n2.} } \\ 
 -1446.4 & 1587 & \makecell{\texttt{?n1 rdf:type rdf:Seq.} } \\ 
 271.9 & 1587 & \makecell{\texttt{?n1 ?p2 rdf:Seq.} } \\ 
 -2977.3 & 1541 & \makecell{\texttt{?n2 rdf:\_3 ?n3.} \\\texttt{?n2 http://www.w3.org/2002/12/cal/ical\_date ?n1.} } \\ 
 -197.2 & 1466 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} } \\ 
 -1595.5 & 1463 & \makecell{\texttt{?n2 rdf:type ?n3.} \\\texttt{?n2 rdf:\_4 ?n1.} } \\ 
 -3974.4 & 1292 & \makecell{\texttt{?n1 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n2.} } \\ 
 -3829.0 & 1216 & \makecell{\texttt{\_:genid9 ?p3 ?n2.} \\\texttt{\_:genid9 ?p3 ?n1.} } \\ 
 -3828.3 & 1209 & \makecell{\texttt{dsw:conference/iswc-aswc/2007/chair\_12 ?p5 ?n3.} \\\texttt{?n1 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n2.} } \\ 
 -3082.3 & 1044 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
 -4012.3 & 670 & \makecell{\texttt{?n1 ?p5 ?n4.} \\\texttt{?n2 ?p5 ?n4.} \\\texttt{?n2 ?p5 ?n3.} } \\ 
 -4179.5 & 613 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 rdf:\_7 ?n3.} } \\ 
 -3790.2 & 479 & \makecell{\texttt{?n1 ?p3 ?n2.} \\\texttt{?n1 ?p3 dsw:person/jason-jung.} } \\ 
 -3067.7 & 421 & \makecell{\texttt{?n1 rdf:type ?n3.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -3808.7 & 398 & \makecell{\texttt{dsw:person/rayid-ghani ?p3 ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 -1939.1 & 396 & \makecell{\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date ?n2.} } \\ 
 -3745.9 & 359 & \makecell{\texttt{dsw:person/rayid-ghani swc:holdsRole ?n2.} \\\texttt{?n1 swc:holdsRole ?n2.} } \\ 
 -2779.2 & 335 & \makecell{\texttt{?n1 ?p2 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -2727.6 & 334 & \makecell{\texttt{?n1 swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 1626.1 & 334 & \makecell{\texttt{dsw:conference/eswc/2011/programme-committee-member swc:heldBy ?n1.} \\\texttt{?n1 swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3707.5 & 333 & \makecell{\texttt{dsw:person/rayid-ghani swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{?n1 swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3729.5 & 333 & \makecell{\texttt{dsw:person/rayid-ghani ?p2 dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{?n1 ?p2 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3171.2 & 309 & \makecell{\texttt{?n1 rdf:type geo:Point.} } \\ 
 -2738.5 & 277 & \makecell{\texttt{dsw:conference/eswc/2012/research-track-committee-member ?p2 ?n1.} } \\ 
 -2678.0 & 274 & \makecell{\texttt{dsw:conference/eswc/2012/research-track-committee-member swc:heldBy ?n1.} } \\ 
 -3887.4 & 271 & \makecell{\texttt{\_:genid95 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n2.} } \\ 
 -3754.4 & 164 & \makecell{\texttt{\_:genid9 ?p3 ?n1.} \\\texttt{\_:genid9 http://www.w3.org/2002/12/cal/ical\_date ?n2.} } \\ 
 -3817.4 & 147 & \makecell{\texttt{\_:genid95 rdf:type ?n3.} \\\texttt{?n1 rdf:type ?n3.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -3708.6 & 126 & \makecell{\texttt{?n2 ?p3 ?n1.} \\\texttt{?n2 http://www.w3.org/2002/12/cal/ical\_date "2009-05-31T18:00:00".} } \\ 
 -3550.9 & 97 & \makecell{\texttt{\_:genid10 ?p2 ?n1.} } \\ 
 -4294.2 & 95 & \makecell{\texttt{?n1 ?p2 foaf:Person.} } \\ 
 -3625.0 & 89 & \makecell{\texttt{\_:genid5 ?p2 ?n1.} } \\ 
 -3505.4 & 89 & \makecell{\texttt{dsw:organization/yahoo-inc ?p2 ?n1.} } \\ 
 -3698.7 & 86 & \makecell{\texttt{\_:genid6 ?p2 ?n1.} } \\ 
 -3633.7 & 86 & \makecell{\texttt{\_:genid11 ?p2 ?n1.} } \\ 
 -3757.9 & 82 & \makecell{\texttt{?n2 ?p3 ?n1.} \\\texttt{?n2 rdf:\_2 dsw:person/jan-hettenhausen.} } \\ 
 -3468.0 & 73 & \makecell{\texttt{dsw:organization/yahoo-inc foaf:member ?n1.} } \\ 
 -3815.9 & 72 & \makecell{\texttt{\_:genid11 rdf:\_1 ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 -3755.1 & 70 & \makecell{\texttt{dsw:workshop/om/2008/paper/main/21 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
 -3596.9 & 67 & \makecell{\texttt{?n1 ?p2 dsw:person/haofen-wang.} } \\ 
 -3660.9 & 67 & \makecell{\texttt{\_:genid18 ?p2 ?n1.} } \\ 
 -3721.2 & 66 & \makecell{\texttt{\_:genid18 ?p2 ?n1.} \\\texttt{\_:genid18 rdf:\_1 dsw:person/gary-wills.} } \\ 
 -3576.6 & 65 & \makecell{\texttt{?n1 ?p2 dsw:person/deborah-mcguinness.} } \\ 
 -3719.8 & 65 & \makecell{\texttt{\_:genid1 ?p2 ?n1.} } \\ 
 -3970.0 & 64 & \makecell{\texttt{?n1 rdf:\_9 ?n2.} } \\ 
 -3695.6 & 59 & \makecell{\texttt{\_:genid27 ?p2 ?n1.} } \\ 
 -3696.5 & 59 & \makecell{\texttt{\_:genid50 ?p2 ?n1.} } \\ 
 -3688.2 & 52 & \makecell{\texttt{\_:genid24 ?p2 ?n1.} } \\ 
 -3740.6 & 52 & \makecell{\texttt{?n1 ?p2 dsw:person/denny-vrandecic.} } \\ 
 -3737.8 & 51 & \makecell{\texttt{\_:genid11 rdf:\_1 ?n2.} \\\texttt{?n1 swrs:author ?n2.} } \\ 
 -3679.5 & 49 & \makecell{\texttt{\_:genid23 ?p2 ?n1.} } \\ 
 -3706.1 & 48 & \makecell{\texttt{\_:genid48 ?p2 ?n1.} } \\ 
 -3716.6 & 48 & \makecell{\texttt{\_:genid48 ?p2 ?n1.} \\\texttt{dsw:conference/www/2008/paper/431 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid48.} } \\ 
 -3661.4 & 47 & \makecell{\texttt{?n1 ?p2 dsw:person/maria-esther-vidal.} } \\ 
 -3681.1 & 47 & \makecell{\texttt{\_:genid46 ?p2 ?n1.} } \\ 
 -3692.9 & 47 & \makecell{\texttt{\_:genid36 ?p2 ?n1.} } \\ 
 -3609.9 & 43 & \makecell{\texttt{dsw:workshop/swcs/2012/programme-committee-member ?p2 ?n1.} } \\ 
 -3729.1 & 43 & \makecell{\texttt{\_:genid56 ?p2 ?n1.} } \\ 
 -3677.4 & 41 & \makecell{\texttt{\_:genid45 ?p2 ?n1.} } \\ 
 -3564.6 & 40 & \makecell{\texttt{dsw:workshop/swcs/2012/programme-committee-member swc:heldBy ?n1.} } \\ 
 -3611.2 & 39 & \makecell{\texttt{?n1 ?p2 dsw:person/gong-cheng.} } \\ 
 -3698.4 & 39 & \makecell{\texttt{?n1 ?p2 dsw:person/matthew-rowe.} } \\ 
 -3743.1 & 38 & \makecell{\texttt{?n1 ?p2 dsw:person/giorgos-stoilos.} } \\ 
 -3725.6 & 37 & \makecell{\texttt{\_:genid82 ?p2 ?n1.} } \\ 
 -3597.5 & 37 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26 ?p2 ?n1.} } \\ 
 -3716.3 & 36 & \makecell{\texttt{?n1 ?p2 dsw:person/katharina-siorpaes.} } \\ 
 -3758.7 & 36 & \makecell{\texttt{\_:genid66 ?p2 ?n1.} } \\ 
 -3730.4 & 36 & \makecell{\texttt{\_:genid78 ?p2 ?n1.} } \\ 
 -3649.8 & 36 & \makecell{\texttt{?n1 ?p2 dsw:person/vanja-josifovski.} } \\ 
 -3759.6 & 34 & \makecell{\texttt{\_:genid69 ?p2 ?n1.} } \\ 
 -3731.8 & 33 & \makecell{\texttt{\_:genid72 ?p2 ?n1.} } \\ 
 -3693.2 & 32 & \makecell{\texttt{dsw:workshop/sdow/2011/pcmember ?p2 ?n1.} } \\ 
 -3745.8 & 31 & \makecell{\texttt{\_:genid90 ?p2 ?n1.} } \\ 
 -3732.2 & 30 & \makecell{\texttt{\_:genid9 rdf:\_3 ?n2.} \\\texttt{\_:genid9 http://www.w3.org/2002/12/cal/ical\_date ?n1.} } \\ 
 -3730.5 & 30 & \makecell{\texttt{?n1 rdf:\_1 ?n2.} \\\texttt{?n1 rdf:\_1 dsw:person/jason-jung.} } \\ 
 -3751.7 & 29 & \makecell{\texttt{\_:genid65 ?p2 ?n1.} } \\ 
 -3737.3 & 26 & \makecell{\texttt{?n1 ?p2 dsw:person/karl-aberer.} } \\ 
 -3665.6 & 24 & \makecell{\texttt{\_:genid5 rdf:\_1 ?n1.} } \\ 
 -3737.1 & 23 & \makecell{\texttt{?n1 ?p2 dsw:person/gang-wang.} } \\ 
 -3703.3 & 21 & \makecell{\texttt{?n1 rdf:\_3 ?n2.} \\\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date "2009-05-31T18:00:00".} } \\ 
 -3728.1 & 21 & \makecell{\texttt{?n1 ?p3 "2009-05-31T18:00:00".} \\\texttt{?n1 rdf:\_3 ?n2.} } \\ 
 -3769.4 & 20 & \makecell{\texttt{\_:genid135 ?p2 ?n1.} } \\ 
 -3670.9 & 19 & \makecell{\texttt{\_:genid15 rdf:\_1 ?n1.} } \\ 
 -3697.6 & 16 & \makecell{\texttt{\_:genid18 rdf:\_2 ?n1.} } \\ 
 -3733.0 & 15 & \makecell{\texttt{?n1 foaf:maker dsw:person/deborah-mcguinness.} } \\ 
 -3716.3 & 15 & \makecell{\texttt{?n1 ?p2 dsw:person/rolf-gruetter.} } \\ 
 -3787.8 & 14 & \makecell{\texttt{dsw:person/adrian-marte ?p2 ?n1.} } \\ 
 -3788.7 & 14 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 ?p2 ?n1.} } \\ 
 -3722.1 & 14 & \makecell{\texttt{dsw:conference/iswc/2011/paper/poster-demo/40 ?p2 ?n1.} } \\ 
 -3674.6 & 14 & \makecell{\texttt{\_:genid23 rdf:\_2 ?n1.} } \\ 
 -3729.2 & 13 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p2 ?n1.} } \\ 
 -3691.8 & 13 & \makecell{\texttt{\_:genid48 rdf:\_1 ?n1.} } \\ 
 -3673.6 & 13 & \makecell{\texttt{?n1 ?p2 \_:genid48.} } \\ 
 -3646.4 & 13 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid48.} } \\ 
 -3690.6 & 13 & \makecell{\texttt{\_:genid27 rdf:\_3 ?n1.} } \\ 
 -3717.4 & 12 & \makecell{\texttt{?n1 ?p2 dsw:person/samur-araujo.} } \\ 
 -3736.6 & 11 & \makecell{\texttt{?n1 ?p2 dsw:person/kai-eckert.} } \\ 
 -3703.9 & 11 & \makecell{\texttt{?n1 ?p2 dsw:person/james-michaelis.} } \\ 
 -3717.9 & 10 & \makecell{\texttt{?n1 ?p2 dsw:person/varish-mulwad.} } \\ 
 -3786.8 & 9 & \makecell{\texttt{dsw:person/kenneth-church ?p2 ?n1.} } \\ 
 -3704.2 & 9 & \makecell{\texttt{?n1 ?p2 dsw:person/jianqiang-li.} } \\ 
 -3714.2 & 9 & \makecell{\texttt{?n1 swc:heldBy dsw:person/matthew-rowe.} } \\ 
 -3764.5 & 8 & \makecell{\texttt{dsw:person/nematollaah-shiri ?p2 ?n1.} } \\ 
 -3704.0 & 8 & \makecell{\texttt{?n1 http://purl.org/dc/terms/creator dsw:person/fabio-ciravegna.} } \\ 
 -3672.5 & 8 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid72.} } \\ 
 -3694.1 & 8 & \makecell{\texttt{\_:genid78 rdf:\_2 ?n1.} } \\ 
 -3703.7 & 7 & \makecell{\texttt{dsw:organization/yahoo-inc rdfs:label ?n1.} } \\ 
 -3699.9 & 7 & \makecell{\texttt{dsw:organization/yahoo-inc foaf:name ?n1.} } \\ 
 -3744.7 & 6 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors ?p2 ?n1.} } \\ 
 -3705.6 & 6 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/willem-van-hage.} } \\ 
 -3742.8 & 5 & \makecell{\texttt{dsw:conference/iswc/2010/paper/515/authorlist ?p2 ?n1.} } \\ 
 -3726.0 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/lisa-lena-opas-haenninen.} } \\ 
 -3724.1 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/yi-chin-tu.} } \\ 
 -3729.5 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/yon-dourisboure.} } \\ 
 -3726.8 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/kukulenz-dirk.} } \\ 
 -3729.2 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/jan-zemanek.} } \\ 
 -3745.6 & 5 & \makecell{\texttt{\_:b5 ?p2 ?n1.} } \\ 
 -3694.8 & 5 & \makecell{\texttt{\_:genid72 rdf:\_3 ?n1.} } \\ 
 -3726.4 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/brett-cannon.} } \\ 
 -3711.8 & 5 & \makecell{\texttt{\_:genid1 rdf:\_1 ?n1.} } \\ 
 -3680.9 & 4 & \makecell{\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date "2010-05-31T13:00:00".} } \\ 
 -3705.2 & 4 & \makecell{\texttt{?n1 ?p2 "2010-05-31T13:00:00".} } \\ 
 -3663.1 & 4 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26 ?p1 dsw:person/james-michaelis.} } \\ 
 -3739.7 & 4 & \makecell{\texttt{dsw:conference/www/2012/paper/518/talk/speaker ?p2 ?n1.} } \\ 
 -3707.7 & 3 & \makecell{\texttt{\_:genid135 rdf:\_1 ?n1.} } \\ 
 -3725.5 & 3 & \makecell{\texttt{dsw:workshop/ores/2010/paper/main/9/authorlist ?p2 ?n1.} } \\ 
 -3703.2 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/gang-wang.} } \\ 
 -3703.3 & 3 & \makecell{\texttt{?n1 dc:creator dsw:person/samur-araujo.} } \\ 
 -3696.2 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/samur-araujo.} } \\ 
 -3667.3 & 3 & \makecell{\texttt{dsw:conference/iswc/2011/paper/semantic-web-in-use/68 ?p1 dsw:person/jianqiang-li.} } \\ 
 -3668.7 & 3 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p1 dsw:person/jianqiang-li.} } \\ 
 -3668.7 & 3 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p1 dsw:person/yu-zhao.} } \\ 
 -3682.9 & 3 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p1 dsw:person/jianqiang-li.} \\\texttt{dsw:workshop/LDOW/2008/paper/3 swc:isPartOf dsw:workshop/LDOW/2008/proceedings.} } \\ 
 -3684.5 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/varish-mulwad.} } \\ 
 -3711.5 & 3 & \makecell{\texttt{?n1 foaf:member dsw:person/xuanhui-wang.} } \\ 
 -3692.5 & 3 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid132.} } \\ 
 -3718.2 & 3 & \makecell{\texttt{\_:genid69 rdf:type ?n1.} } \\ 
 -3707.6 & 3 & \makecell{\texttt{\_:genid18 ?p2 "2010-05-31T13:00:00".} \\\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date "2010-05-31T13:00:00".} } \\ 
 -3684.5 & 3 & \makecell{\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date "2008-04-25T15:30:00".} } \\ 
 -3707.5 & 2 & \makecell{\texttt{dsw:person/rayid-ghani swc:holdsRole ?n1.} } \\ 
 -3704.5 & 2 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 foaf:maker ?n1.} } \\ 
 -3703.5 & 2 & \makecell{\texttt{?n1 foaf:maker dsw:person/yu-zhao.} } \\ 
 -3703.5 & 2 & \makecell{\texttt{?n1 foaf:maker dsw:person/jianqiang-li.} } \\ 
 -3706.2 & 2 & \makecell{\texttt{?n1 swc:heldBy dsw:person/kai-eckert.} } \\ 
 -3694.9 & 2 & \makecell{\texttt{dsw:organization/yahoo-inc ?p1 "Yahoo! Inc".} } \\ 
 -3732.0 & 2 & \makecell{\texttt{?n1 ?p2 "Yahoo! Inc".} } \\ 
 -3712.6 & 2 & \makecell{\texttt{\_:genid11 ?p2 ?n1.} \\\texttt{dsw:workshop/obi/2008/paper/main/7 swrs:author ?n1.} } \\ 
 -3706.3 & 2 & \makecell{\texttt{dsw:conference/iswc/2011/paper/poster-demo/40 swrs:author ?n1.} } \\ 
 -3701.4 & 2 & \makecell{\texttt{dsw:conference/iswc/2011/paper/poster-demo/40 http://purl.org/dc/terms/creator ?n1.} } \\ 
 -3708.3 & 2 & \makecell{\texttt{\_:genid10 http://www.w3.org/2002/12/cal/ical\_tzid ?n1.} } \\ 
 -3703.8 & 1 & \makecell{\texttt{dsw:person/kenneth-church swc:holdsRole ?n1.} } \\ 
 -3689.9 & 1 & \makecell{\texttt{dsw:person/lorenzino-vaccari ?p1 dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{dsw:person/rayid-ghani ?p1 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3667.9 & 1 & \makecell{\texttt{dsw:person/lorenzino-vaccari swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{dsw:person/rayid-ghani swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3692.5 & 1 & \makecell{\texttt{dsw:person/lorenzino-vaccari ?p1 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3691.0 & 1 & \makecell{\texttt{dsw:person/neil-ireson ?p1 dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{dsw:person/rayid-ghani ?p1 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3692.7 & 1 & \makecell{\texttt{dsw:person/rayid-ghani ?p1 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3670.7 & 1 & \makecell{\texttt{dsw:person/rayid-ghani swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3669.6 & 1 & \makecell{\texttt{dsw:workshop/swcs/2012/programme-committee-member swc:heldBy dsw:person/maria-esther-vidal.} } \\ 
 -3691.6 & 1 & \makecell{\texttt{dsw:workshop/swcs/2012/programme-committee-member ?p1 dsw:person/maria-esther-vidal.} } \\ 
 -3688.4 & 1 & \makecell{\texttt{dsw:organization/universidad-simon-bolivar ?p1 dsw:person/maria-esther-vidal.} } \\ 
 -3719.8 & 1 & \makecell{\texttt{dsw:workshop/ores/2010/paper/main/9/authorlist ?p2 ?n1.} \\\texttt{?n1 foaf:mbox\_sha1sum "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
 -3697.7 & 1 & \makecell{\texttt{dsw:workshop/ores/2010/paper/main/9/authorlist rdf:\_3 ?n1.} \\\texttt{?n1 foaf:mbox\_sha1sum "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
 -3707.2 & 1 & \makecell{\texttt{?n1 foaf:mbox\_sha1sum "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
 -3671.1 & 1 & \makecell{\texttt{dsw:person/adrian-marte foaf:mbox\_sha1sum "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
 -3729.2 & 1 & \makecell{\texttt{?n1 ?p2 "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
 -3693.1 & 1 & \makecell{\texttt{dsw:person/adrian-marte ?p1 "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
 -3685.9 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/515/authorlist ?p1 dsw:person/samur-araujo.} } \\ 
 -3663.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/515/authorlist rdf:\_1 dsw:person/samur-araujo.} } \\ 
 -3680.4 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/515 foaf:maker dsw:person/samur-araujo.} \\\texttt{dsw:conference/iswc/2010/paper/515/authorlist ?p1 dsw:person/samur-araujo.} } \\ 
 -3701.3 & 1 & \makecell{\texttt{?n1 http://purl.org/dc/terms/creator dsw:person/jianqiang-li.} } \\ 
 -3665.1 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/semantic-web-in-use/68 http://purl.org/dc/terms/creator dsw:person/jianqiang-li.} } \\ 
 -3666.5 & 1 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 foaf:maker dsw:person/jianqiang-li.} } \\ 
 -3666.5 & 1 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 foaf:maker dsw:person/yu-zhao.} } \\ 
 -3690.7 & 1 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p1 http://sunsite.informatik.rwth-aachen.de/Publications/CEUR-WS/Vol-369/paper04.pdf.} } \\ 
 -3666.3 & 1 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 dc:creator dsw:person/jianqiang-li.} } \\ 
 -3699.5 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/lisa-lena-opas-haenninen.} } \\ 
 -3663.3 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-775/authors rdf:\_1 dsw:person/lisa-lena-opas-haenninen.} } \\ 
 -3685.3 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-775/authors ?p1 dsw:person/lisa-lena-opas-haenninen.} } \\ 
 -3705.0 & 1 & \makecell{\texttt{dsw:person/nematollaah-shiri foaf:firstName ?n1.} } \\ 
 -3698.6 & 1 & \makecell{\texttt{dsw:person/nematollaah-shiri ?p1 foaf:Person.} } \\ 
 -3676.6 & 1 & \makecell{\texttt{dsw:person/nematollaah-shiri rdf:type foaf:Person.} } \\ 
 -3712.8 & 1 & \makecell{\texttt{dsw:person/nematollaah-shiri rdf:type ?n1.} } \\ 
 -3686.6 & 1 & \makecell{\texttt{\_:genid24 ?p1 dsw:person/yi-chin-tu.} } \\ 
 -3731.0 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 http://purl.org/ontology/bibo/authorList ?n1.} \\\texttt{?n1 rdf:\_1 ?n2.} } \\ 
 -3705.5 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 http://purl.org/ontology/bibo/authorList ?n1.} } \\ 
 -3691.4 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 ?p1 dsw:conference/dh/2010/abstracts/poster/ab-643/authors.} } \\ 
 -3727.5 & 1 & \makecell{\texttt{?n1 ?p2 dsw:conference/dh/2010/abstracts/poster/ab-643/authors.} } \\ 
 -3669.3 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 http://purl.org/ontology/bibo/authorList dsw:conference/dh/2010/abstracts/poster/ab-643/authors.} } \\ 
 -3705.5 & 1 & \makecell{\texttt{?n1 http://purl.org/ontology/bibo/authorList dsw:conference/dh/2010/abstracts/poster/ab-643/authors.} } \\ 
 -3670.1 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/research-track-committee-member swc:heldBy dsw:person/kai-eckert.} } \\ 
 -3692.1 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/research-track-committee-member ?p1 dsw:person/kai-eckert.} } \\ 
 -3750.6 & 1 & \makecell{\texttt{?n1 ?p3 ?n2.} \\\texttt{?n2 foaf:made dsw:conference/www/2010/paper/main/23.} } \\ 
 -3688.0 & 1 & \makecell{\texttt{\_:genid90 ?p1 dsw:person/jan-zemanek.} } \\ 
 -3675.7 & 1 & \makecell{\texttt{\_:genid15 rdf:\_1 dsw:person/c-maria-keet.} } \\ 
 -3671.7 & 1 & \makecell{\texttt{\_:genid15 rdf:\_1 dsw:person/daniel-bachlechner.} } \\ 
 -3699.6 & 1 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors rdf:\_1 ?n1.} } \\ 
 -3702.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors rdf:\_2 ?n1.} } \\ 
 -3666.6 & 1 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors rdf:\_2 dsw:person/deborah-mcguinness.} } \\ 
 -3688.7 & 1 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors ?p1 dsw:person/deborah-mcguinness.} } \\ 
 -3712.0 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/doctoral-consortium/14/authorlist ?p2 ?n1.} } \\ 
 -3675.9 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/doctoral-consortium/14/authorlist ?p1 dsw:person/varish-mulwad.} } \\ 
 -3653.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/doctoral-consortium/14/authorlist rdf:\_1 dsw:person/varish-mulwad.} } \\ 
 -3662.5 & 1 & \makecell{\texttt{\_:b3 rdf:\_1 dsw:person/varish-mulwad.} } \\ 
 -3684.5 & 1 & \makecell{\texttt{\_:b3 ?p1 dsw:person/varish-mulwad.} } \\ 
 -3698.7 & 1 & \makecell{\texttt{\_:b3 rdf:\_1 ?n1.} } \\ 
 -3667.4 & 1 & \makecell{\texttt{dsw:organization/yahoo-inc rdfs:label "Yahoo! Inc".} } \\ 
 -3670.4 & 1 & \makecell{\texttt{dsw:organization/yahoo-inc foaf:member dsw:person/xuanhui-wang.} } \\ 
 -3695.7 & 1 & \makecell{\texttt{\_:genid11 rdf:\_1 ?n1.} \\\texttt{dsw:workshop/obi/2008/paper/main/7 swrs:author ?n1.} } \\ 
 -3699.7 & 1 & \makecell{\texttt{\_:genid11 rdf:\_1 ?n1.} \\\texttt{dsw:conference/iswc/2011/paper/poster-demo/40 swrs:author ?n1.} } \\ 
 -3698.8 & 1 & \makecell{\texttt{?n1 http://purl.org/dc/terms/creator dsw:person/akira-mori.} } \\ 
 -3699.2 & 1 & \makecell{\texttt{?n1 swrs:editor dsw:person/denny-vrandecic.} } \\ 
 -3663.1 & 1 & \makecell{\texttt{dsw:workshop/insemtive/2008/proceedings swrs:editor dsw:person/denny-vrandecic.} } \\ 
 -3685.1 & 1 & \makecell{\texttt{dsw:workshop/insemtive/2008/proceedings ?p1 dsw:person/denny-vrandecic.} } \\ 
 -3690.3 & 1 & \makecell{\texttt{\_:genid48 ?p1 dsw:person/willem-van-hage.} } \\ 
 -3668.3 & 1 & \makecell{\texttt{\_:genid48 rdf:\_1 dsw:person/willem-van-hage.} } \\ 
 -3665.3 & 1 & \makecell{\texttt{dsw:conference/www/2008/paper/431 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid48.} } \\ 
 -3687.4 & 1 & \makecell{\texttt{dsw:conference/www/2008/paper/431 ?p1 \_:genid48.} } \\ 
 -3701.7 & 1 & \makecell{\texttt{dsw:conference/www/2007/paper/main/429 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n1.} } \\ 
 -3703.1 & 1 & \makecell{\texttt{?n1 foaf:maker dsw:person/brett-cannon.} } \\ 
 -3691.2 & 1 & \makecell{\texttt{\_:genid36 ?p1 dsw:person/vanja-josifovski.} } \\ 
 -3728.0 & 1 & \makecell{\texttt{dsw:workshop/om/2008/paper/main/21 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n1.} \\\texttt{?n1 rdf:\_7 ?n2.} } \\ 
 -3693.5 & 1 & \makecell{\texttt{\_:genid16 rdf:\_7 ?n1.} \\\texttt{dsw:workshop/om/2008/paper/main/21 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid16.} } \\ 
 -3690.1 & 1 & \makecell{\texttt{\_:genid23 ?p1 dsw:person/zheng-chen.} } \\ 
 -3697.6 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/semantic-web-in-use/76/authorlist rdf:\_9 ?n1.} } \\ 
 -3691.1 & 1 & \makecell{\texttt{dsw:workshop/sdow/2011/pcmember ?p1 dsw:person/matthew-rowe.} } \\ 
 -3669.0 & 1 & \makecell{\texttt{dsw:workshop/sdow/2011/pcmember swc:heldBy dsw:person/matthew-rowe.} } \\ 
 -3690.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/industry/21/authorlist rdf:\_1 ?n1.} } \\ 
 -3707.0 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/karl-aberer.} } \\ 
 -3670.8 & 1 & \makecell{\texttt{\_:genid1 rdf:\_1 dsw:person/karl-aberer.} } \\ 
 -3692.9 & 1 & \makecell{\texttt{\_:genid1 ?p1 dsw:person/karl-aberer.} } \\ 
 -3700.1 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/peyman-sazedj.} } \\ 
 -3700.1 & 1 & \makecell{\texttt{\_:genid18 http://www.w3.org/2002/12/cal/ical\_date ?n1.} } \\ 
 -3663.9 & 1 & \makecell{\texttt{\_:genid18 http://www.w3.org/2002/12/cal/ical\_date "2010-05-31T13:00:00".} } \\ 
 -3686.0 & 1 & \makecell{\texttt{\_:genid18 ?p1 "2010-05-31T13:00:00".} } \\ 
 -3671.5 & 1 & \makecell{\texttt{\_:genid11 http://www.w3.org/2002/12/cal/ical\_date "2010-05-31T13:00:00".} \\\texttt{\_:genid18 ?p1 "2010-05-31T13:00:00".} } \\ 
 -3707.6 & 1 & \makecell{\texttt{\_:genid11 http://www.w3.org/2002/12/cal/ical\_date ?n1.} \\\texttt{\_:genid18 ?p2 ?n1.} } \\ 
 -3693.5 & 1 & \makecell{\texttt{\_:genid11 ?p2 "2010-05-31T13:00:00".} \\\texttt{\_:genid18 ?p1 "2010-05-31T13:00:00".} } \\ 
 -3809.4 & 1 & \makecell{\texttt{dsw:conference/iswc-aswc/2007/chair\_12 swc:heldBy ?n1.} \\\texttt{?n2 ?p4 ?n3.} \\\texttt{?n2 ?p4 ?n1.} } \\ 
 -3707.5 & 1 & \makecell{\texttt{dsw:conference/eswc/2007/time-20070605T153000EuropeVienna rdf:type ?n1.} } \\ 
 -3695.5 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors ?p1 rdf:Seq.} } \\ 
 -3673.4 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:type rdf:Seq.} } \\ 
 -3709.6 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:type ?n1.} } \\ 
 -3699.6 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:type ?n1.} \\\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:\_4 dsw:person/stefan-beyer.} } \\ 
 -3735.8 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:type ?n2.} \\\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:\_4 ?n1.} } \\ 
 -3709.7 & 1 & \makecell{\texttt{dsw:conference/www/2012/paper/518/talk/speaker rdf:type ?n1.} } \\ 
 -3744.2 & 0 & \makecell{\texttt{?n1 ?p4 ?n2.} \\\texttt{?n2 rdf:\_7 ?n3.} } \\ 
\hline
\end{longtable}

\subsection{AIFB, top 100 by log factor}

\begin{longtable}{ r r p{19cm} }
 log factor & frequency & \\
\hline\endhead
 8919.0 & 2477 & \makecell{\texttt{?n1 swrs:isAbout ?n2.} } \\ 
 3857.8 & 4124 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
 923.9 & 1066 & \makecell{\texttt{?n1 ?p2 swrs:Person.} } \\ 
 827.3 & 1045 & \makecell{\texttt{?n1 rdf:type swrs:Person.} } \\ 
 -141.5 & 1240 & \makecell{\texttt{?n1 ?p2 swrs:Publication.} } \\ 
 -167.7 & 1222 & \makecell{\texttt{?n1 rdf:type swrs:Publication.} } \\ 
 -893.7 & 1529 & \makecell{\texttt{?n1 swrs:author ?n2.} } \\ 
 -907.8 & 3388 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
 -939.9 & 687 & \makecell{\texttt{?n1 rdf:type swrs:InProceedings.} } \\ 
 -975.2 & 687 & \makecell{\texttt{?n1 ?p2 swrs:InProceedings.} } \\ 
 -1495.3 & 31 & \makecell{\texttt{?n1 ?p2 owl:DatatypeProperty.} } \\ 
 -1536.6 & 146 & \makecell{\texttt{?n1 rdf:type swrs:ResearchTopic.} } \\ 
 -1582.5 & 152 & \makecell{\texttt{?n1 rdf:type owl:Restriction.} } \\ 
 -1583.0 & 147 & \makecell{\texttt{?n1 ?p2 swrs:ResearchTopic.} } \\ 
 -1609.0 & 152 & \makecell{\texttt{?n1 owl:onProperty ?n2.} } \\ 
 -1613.5 & 152 & \makecell{\texttt{?n1 ?p2 owl:Restriction.} } \\ 
 -1668.8 & 55 & \makecell{\texttt{?n1 rdf:type owl:Class.} } \\ 
 -1696.9 & 55 & \makecell{\texttt{?n1 ?p2 owl:Class.} } \\ 
 -1745.2 & 65 & \makecell{\texttt{?n1 rdf:type swrs:PhDStudent.} } \\ 
 -1747.9 & 40 & \makecell{\texttt{?n1 swrs:isAbout http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1767.8 & 78 & \makecell{\texttt{?n1 rdf:type swrs:Project.} } \\ 
 -1769.7 & 13 & \makecell{\texttt{swrs:InBook rdfs:subClassOf ?n1.} } \\ 
 -1775.0 & 87 & \makecell{\texttt{?n1 ?p2 swrs:Project.} } \\ 
 -1775.1 & 40 & \makecell{\texttt{?n1 ?p2 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1777.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance swrs:isAbout http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1779.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:publication http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
 -1780.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance rdf:type swrs:ResearchTopic.} } \\ 
 -1781.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance swrs:name "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1784.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type swrs:Article.} } \\ 
 -1784.3 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id655instance rdf:type swrs:Person.} } \\ 
 -1784.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id262instance rdf:type swrs:Person.} } \\ 
 -1785.0 & 1 & \makecell{\texttt{\_:ub1bL199C18 rdf:type owl:Restriction.} } \\ 
 -1785.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id583instance rdf:type swrs:Person.} } \\ 
 -1785.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id637instance rdf:type swrs:Person.} } \\ 
 -1785.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id594instance rdf:type swrs:InProceedings.} } \\ 
 -1786.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id684instance rdf:type swrs:InProceedings.} } \\ 
 -1786.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id499instance rdf:type swrs:InProceedings.} } \\ 
 -1786.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id841instance rdf:type swrs:InProceedings.} } \\ 
 -1787.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2008instance rdf:type swrs:Person.} } \\ 
 -1787.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type swrs:Publication.} } \\ 
 -1787.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id663instance rdf:type swrs:InProceedings.} } \\ 
 -1787.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type swrs:Person.} } \\ 
 -1787.3 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance rdf:type swrs:Publication.} } \\ 
 -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2118instance rdf:type swrs:Person.} } \\ 
 -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance rdf:type swrs:Publication.} } \\ 
 -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance rdf:type swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id709instance rdf:type swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance rdf:type swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance rdf:type swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance rdf:type swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id703instance rdf:type swrs:Person.} } \\ 
 -1791.2 & 1 & \makecell{\texttt{swrs:title ?p1 owl:DatatypeProperty.} } \\ 
 -1796.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance ?p1 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1796.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance ?p1 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1798.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance ?p1 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
 -1799.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance ?p1 swrs:ResearchTopic.} } \\ 
 -1800.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p1 "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1800.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p1 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id877instance.} } \\ 
 -1801.0 & 14 & \makecell{\texttt{swrs:InBook ?p2 ?n1.} } \\ 
 -1802.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id28instance ?p1 swrs:ResearchTopic.} } \\ 
 -1802.2 & 227 & \makecell{\texttt{?n1 swrs:phone ?n2.} } \\ 
 -1803.3 & 1 & \makecell{\texttt{swrs:Proceedings ?p1 owl:Class.} } \\ 
 -1803.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id655instance ?p1 swrs:Person.} } \\ 
 -1803.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id350instance ?p1 swrs:Person.} } \\ 
 -1803.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id262instance ?p1 swrs:Person.} } \\ 
 -1804.0 & 1 & \makecell{\texttt{\_:ub1bL474C18 ?p1 owl:Restriction.} } \\ 
 -1804.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id583instance ?p1 swrs:Person.} } \\ 
 -1804.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id810instance ?p1 swrs:Person.} } \\ 
 -1804.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id184instance ?p1 swrs:Person.} } \\ 
 -1805.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id594instance ?p1 swrs:InProceedings.} } \\ 
 -1805.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id275instance ?p1 swrs:InProceedings.} } \\ 
 -1805.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id841instance ?p1 swrs:InProceedings.} } \\ 
 -1806.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2008instance ?p1 swrs:Person.} } \\ 
 -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance ?p1 swrs:Publication.} } \\ 
 -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id663instance ?p1 swrs:InProceedings.} } \\ 
 -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance ?p1 swrs:Person.} } \\ 
 -1806.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance ?p1 swrs:Publication.} } \\ 
 -1806.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id773instance ?p1 swrs:Publication.} } \\ 
 -1806.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance ?p1 swrs:Publication.} } \\ 
 -1806.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance ?p1 swrs:Person.} } \\ 
 -1806.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id954instance ?p1 swrs:Publication.} } \\ 
 -1806.7 & 7 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:publication ?n1.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance ?p1 swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance ?p1 swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id820instance ?p1 swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance ?p1 swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id786instance ?p1 swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id848instance ?p1 swrs:Person.} } \\ 
 -1807.3 & 1 & \makecell{\texttt{\_:ub1bL560C18 ?p1 swrs:Person.} } \\ 
 -1808.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id165instance swrs:name ?n1.} } \\ 
 -1808.9 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:worksAtProject ?n1.} } \\ 
 -1811.1 & 1 & \makecell{\texttt{?n1 swrs:name "CKDD - Conceptual Knowledge Discovery"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1811.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance rdf:type ?n1.} } \\ 
 -1812.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance swrs:name ?n1.} } \\ 
 -1812.1 & 1 & \makecell{\texttt{?n1 swrs:name "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1813.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id68instance swrs:phone ?n1.} } \\ 
 -1813.3 & 9 & \makecell{\texttt{swrs:Organization ?p2 ?n1.} } \\ 
 -1813.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id79instance swrs:name ?n1.} } \\ 
 -1814.1 & 1 & \makecell{\texttt{\_:ub1bL474C18 owl:onProperty ?n1.} } \\ 
 -1814.8 & 1 & \makecell{\texttt{swrs:Organization rdf:type ?n1.} } \\ 
 -1815.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id18instance rdf:type ?n1.} } \\ 
 -1815.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id814instance rdf:type ?n1.} } \\ 
 -1815.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id20instance rdf:type ?n1.} } \\ 
 -1816.0 & 1 & \makecell{\texttt{\_:ub1bL422C18 rdf:type ?n1.} } \\ 
 -1816.0 & 1 & \makecell{\texttt{\_:ub1bL355C18 rdf:type ?n1.} } \\ 
 -1818.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance rdf:type ?n1.} } \\ 
 -1818.8 & 152 & \makecell{\texttt{?n1 swrs:author http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id20instance.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id593instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id439instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id804instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id305instance rdf:type ?n1.} } \\ 
 -1820.7 & 22 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance ?p2 ?n1.} } \\ 
 -1821.4 & 1 & \makecell{\texttt{swrs:member ?p2 ?n1.} } \\ 
 -1821.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type ?n1.} } \\ 
 -1822.2 & 1 & \makecell{\texttt{swrs:name ?p2 ?n1.} } \\ 
 -1822.2 & 1 & \makecell{\texttt{swrs:title ?p2 ?n1.} } \\ 
 -1824.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id19instance rdf:type ?n1.} } \\ 
 -1824.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type ?n1.} } \\ 
 -1824.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1089instance rdf:type ?n1.} } \\ 
 -1825.1 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance rdf:type ?n1.} } \\ 
 -1825.9 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id552instance swrs:author ?n1.} } \\ 
 -1826.2 & 3 & \makecell{\texttt{\_:ub1bL658C27 ?p2 ?n1.} } \\ 
 -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id617instance rdf:type ?n1.} } \\ 
 -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id499instance rdf:type ?n1.} } \\ 
 -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id516instance rdf:type ?n1.} } \\ 
 -1826.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance rdf:type ?n1.} } \\ 
 -1826.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1333instance rdf:type ?n1.} } \\ 
 -1826.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id512instance rdf:type ?n1.} } \\ 
 -1826.8 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id572instance rdf:type ?n1.} } \\ 
 -1826.8 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance rdf:type ?n1.} } \\ 
 -1827.0 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id64instance rdf:type ?n1.} } \\ 
 -1827.0 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id777instance rdf:type ?n1.} } \\ 
 -1827.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1306instance rdf:type ?n1.} } \\ 
 -1827.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance rdf:type ?n1.} } \\ 
 -1828.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1166instance rdf:type ?n1.} } \\ 
 -1830.1 & 1 & \makecell{\texttt{?n1 ?p2 "CKDD - Conceptual Knowledge Discovery"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1831.1 & 1 & \makecell{\texttt{?n1 ?p2 "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1833.9 & 11 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance swrs:isAbout ?n1.} } \\ 
 -1834.8 & 3 & \makecell{\texttt{?n1 ?p2 swrs:Graduate.} } \\ 
 -1837.2 & 3 & \makecell{\texttt{\_:ub1bL58C27 ?p2 ?n1.} } \\ 
 -1837.9 & 3 & \makecell{\texttt{\_:ub1bL474C18 ?p2 ?n1.} } \\ 
 -1839.0 & 3 & \makecell{\texttt{\_:ub1bL102C18 ?p2 ?n1.} } \\ 
 -1839.6 & 8 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance ?p2 ?n1.} } \\ 
 -1839.9 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance ?p2 ?n1.} } \\ 
 -1840.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id144instance ?p2 ?n1.} } \\ 
 -1840.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id190instance ?p2 ?n1.} } \\ 
 -1841.0 & 14 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Projekte/viewProjektOWL/id51instance swrs:member ?n1.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -1841.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance ?p2 ?n1.} } \\ 
 -1841.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance ?p2 ?n1.} } \\ 
 -1841.7 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id774instance ?p2 ?n1.} } \\ 
 -1842.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p2 ?n1.} } \\ 
 -1842.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id607instance ?p2 ?n1.} } \\ 
 -1843.9 & 6 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id530instance ?p2 ?n1.} } \\ 
 -1844.5 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id804instance ?p2 ?n1.} } \\ 
 -1850.2 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id165instance ?p2 ?n1.} } \\ 
 -1850.6 & 12 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2067instance ?p2 ?n1.} } \\ 
 -1860.0 & 1907 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance rdf:type ?n2.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -1869.8 & 17 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance ?p2 ?n1.} } \\ 
 -1877.4 & 10 & \makecell{\texttt{?n1 ?p2 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
 -1882.2 & 9 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id275instance ?p2 ?n1.} } \\ 
 -1883.4 & 18 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1272instance ?p2 ?n1.} } \\ 
 -1883.7 & 13 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance ?p2 ?n1.} } \\ 
 -1887.6 & 1130 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 -1888.2 & 12 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id228instance ?p2 ?n1.} } \\ 
 -1895.2 & 11 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id572instance ?p2 ?n1.} } \\ 
 -1903.8 & 16 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1305instance ?p2 ?n1.} } \\ 
 -1904.3 & 16 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1317instance ?p2 ?n1.} } \\ 
 -1908.7 & 13 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1152instance ?p2 ?n1.} } \\ 
 -1930.3 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1124instance ?p2 ?n1.} } \\ 
 -1931.5 & 2184 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance ?p3 ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 -1937.4 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id882instance ?p2 ?n1.} } \\ 
 -1942.5 & 25 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance ?p2 ?n1.} } \\ 
 -1942.7 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1025instance ?p2 ?n1.} } \\ 
 -2254.8 & 24274 & \makecell{\texttt{?n1 rdf:type ?n3.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
 -4014.6 & 494 & \makecell{\texttt{?n1 swrs:name ?n2.} } \\ 
\hline
\end{longtable}

\subsection{AIFB, top 100 by frequency}

\begin{longtable}{ r r p{19cm} }
 log factor & frequency & \\
\hline\endhead
 -2254.8 & 24274 & \makecell{\texttt{?n1 rdf:type ?n3.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
 3857.8 & 4124 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
 -907.8 & 3388 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
 8919.0 & 2477 & \makecell{\texttt{?n1 swrs:isAbout ?n2.} } \\ 
 -1931.5 & 2184 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance ?p3 ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 -1860.0 & 1907 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance rdf:type ?n2.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -893.7 & 1529 & \makecell{\texttt{?n1 swrs:author ?n2.} } \\ 
 -141.5 & 1240 & \makecell{\texttt{?n1 ?p2 swrs:Publication.} } \\ 
 -167.7 & 1222 & \makecell{\texttt{?n1 rdf:type swrs:Publication.} } \\ 
 -1887.6 & 1130 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 923.9 & 1066 & \makecell{\texttt{?n1 ?p2 swrs:Person.} } \\ 
 827.3 & 1045 & \makecell{\texttt{?n1 rdf:type swrs:Person.} } \\ 
 -975.2 & 687 & \makecell{\texttt{?n1 ?p2 swrs:InProceedings.} } \\ 
 -939.9 & 687 & \makecell{\texttt{?n1 rdf:type swrs:InProceedings.} } \\ 
 -4014.6 & 494 & \makecell{\texttt{?n1 swrs:name ?n2.} } \\ 
 -1802.2 & 227 & \makecell{\texttt{?n1 swrs:phone ?n2.} } \\ 
 -1613.5 & 152 & \makecell{\texttt{?n1 ?p2 owl:Restriction.} } \\ 
 -1582.5 & 152 & \makecell{\texttt{?n1 rdf:type owl:Restriction.} } \\ 
 -1609.0 & 152 & \makecell{\texttt{?n1 owl:onProperty ?n2.} } \\ 
 -1818.8 & 152 & \makecell{\texttt{?n1 swrs:author http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id20instance.} } \\ 
 -1583.0 & 147 & \makecell{\texttt{?n1 ?p2 swrs:ResearchTopic.} } \\ 
 -1536.6 & 146 & \makecell{\texttt{?n1 rdf:type swrs:ResearchTopic.} } \\ 
 -1775.0 & 87 & \makecell{\texttt{?n1 ?p2 swrs:Project.} } \\ 
 -1767.8 & 78 & \makecell{\texttt{?n1 rdf:type swrs:Project.} } \\ 
 -1745.2 & 65 & \makecell{\texttt{?n1 rdf:type swrs:PhDStudent.} } \\ 
 -1668.8 & 55 & \makecell{\texttt{?n1 rdf:type owl:Class.} } \\ 
 -1696.9 & 55 & \makecell{\texttt{?n1 ?p2 owl:Class.} } \\ 
 -1775.1 & 40 & \makecell{\texttt{?n1 ?p2 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1747.9 & 40 & \makecell{\texttt{?n1 swrs:isAbout http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1495.3 & 31 & \makecell{\texttt{?n1 ?p2 owl:DatatypeProperty.} } \\ 
 -1942.5 & 25 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance ?p2 ?n1.} } \\ 
 -1820.7 & 22 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance ?p2 ?n1.} } \\ 
 -1937.4 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id882instance ?p2 ?n1.} } \\ 
 -1942.7 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1025instance ?p2 ?n1.} } \\ 
 -1930.3 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1124instance ?p2 ?n1.} } \\ 
 -1883.4 & 18 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1272instance ?p2 ?n1.} } \\ 
 -1869.8 & 17 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance ?p2 ?n1.} } \\ 
 -1903.8 & 16 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1305instance ?p2 ?n1.} } \\ 
 -1904.3 & 16 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1317instance ?p2 ?n1.} } \\ 
 -1801.0 & 14 & \makecell{\texttt{swrs:InBook ?p2 ?n1.} } \\ 
 -1841.0 & 14 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Projekte/viewProjektOWL/id51instance swrs:member ?n1.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -1769.7 & 13 & \makecell{\texttt{swrs:InBook rdfs:subClassOf ?n1.} } \\ 
 -1908.7 & 13 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1152instance ?p2 ?n1.} } \\ 
 -1883.7 & 13 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance ?p2 ?n1.} } \\ 
 -1888.2 & 12 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id228instance ?p2 ?n1.} } \\ 
 -1850.6 & 12 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2067instance ?p2 ?n1.} } \\ 
 -1833.9 & 11 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance swrs:isAbout ?n1.} } \\ 
 -1895.2 & 11 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id572instance ?p2 ?n1.} } \\ 
 -1877.4 & 10 & \makecell{\texttt{?n1 ?p2 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
 -1882.2 & 9 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id275instance ?p2 ?n1.} } \\ 
 -1813.3 & 9 & \makecell{\texttt{swrs:Organization ?p2 ?n1.} } \\ 
 -1839.6 & 8 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance ?p2 ?n1.} } \\ 
 -1806.7 & 7 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:publication ?n1.} } \\ 
 -1843.9 & 6 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id530instance ?p2 ?n1.} } \\ 
 -1839.9 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance ?p2 ?n1.} } \\ 
 -1850.2 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id165instance ?p2 ?n1.} } \\ 
 -1825.9 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id552instance swrs:author ?n1.} } \\ 
 -1841.7 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id774instance ?p2 ?n1.} } \\ 
 -1837.9 & 3 & \makecell{\texttt{\_:ub1bL474C18 ?p2 ?n1.} } \\ 
 -1837.2 & 3 & \makecell{\texttt{\_:ub1bL58C27 ?p2 ?n1.} } \\ 
 -1840.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id190instance ?p2 ?n1.} } \\ 
 -1841.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance ?p2 ?n1.} } \\ 
 -1834.8 & 3 & \makecell{\texttt{?n1 ?p2 swrs:Graduate.} } \\ 
 -1841.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance ?p2 ?n1.} } \\ 
 -1842.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p2 ?n1.} } \\ 
 -1839.0 & 3 & \makecell{\texttt{\_:ub1bL102C18 ?p2 ?n1.} } \\ 
 -1844.5 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id804instance ?p2 ?n1.} } \\ 
 -1826.2 & 3 & \makecell{\texttt{\_:ub1bL658C27 ?p2 ?n1.} } \\ 
 -1840.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id144instance ?p2 ?n1.} } \\ 
 -1842.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id607instance ?p2 ?n1.} } \\ 
 -1808.9 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:worksAtProject ?n1.} } \\ 
 -1821.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type ?n1.} } \\ 
 -1825.1 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance rdf:type ?n1.} } \\ 
 -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id617instance rdf:type ?n1.} } \\ 
 -1827.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1306instance rdf:type ?n1.} } \\ 
 -1826.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id512instance rdf:type ?n1.} } \\ 
 -1826.8 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id572instance rdf:type ?n1.} } \\ 
 -1824.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type ?n1.} } \\ 
 -1827.0 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id64instance rdf:type ?n1.} } \\ 
 -1827.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance rdf:type ?n1.} } \\ 
 -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id499instance rdf:type ?n1.} } \\ 
 -1827.0 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id777instance rdf:type ?n1.} } \\ 
 -1826.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1333instance rdf:type ?n1.} } \\ 
 -1824.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1089instance rdf:type ?n1.} } \\ 
 -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id516instance rdf:type ?n1.} } \\ 
 -1824.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id19instance rdf:type ?n1.} } \\ 
 -1826.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance rdf:type ?n1.} } \\ 
 -1828.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1166instance rdf:type ?n1.} } \\ 
 -1826.8 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance rdf:type ?n1.} } \\ 
 -1798.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance ?p1 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
 -1779.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:publication http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
 -1804.0 & 1 & \makecell{\texttt{\_:ub1bL474C18 ?p1 owl:Restriction.} } \\ 
 -1814.1 & 1 & \makecell{\texttt{\_:ub1bL474C18 owl:onProperty ?n1.} } \\ 
 -1816.0 & 1 & \makecell{\texttt{\_:ub1bL422C18 rdf:type ?n1.} } \\ 
 -1802.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id28instance ?p1 swrs:ResearchTopic.} } \\ 
 -1780.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance rdf:type swrs:ResearchTopic.} } \\ 
 -1811.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance rdf:type ?n1.} } \\ 
 -1799.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance ?p1 swrs:ResearchTopic.} } \\ 
 -1804.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id583instance ?p1 swrs:Person.} } \\ 
 -1785.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id583instance rdf:type swrs:Person.} } \\ 
 -1787.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type swrs:Person.} } \\ 
 -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance ?p1 swrs:Person.} } \\ 
 -1796.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance ?p1 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1777.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance swrs:isAbout http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1796.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance ?p1 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1822.2 & 1 & \makecell{\texttt{swrs:name ?p2 ?n1.} } \\ 
 -1787.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id663instance rdf:type swrs:InProceedings.} } \\ 
 -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id663instance ?p1 swrs:InProceedings.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id709instance rdf:type swrs:Person.} } \\ 
 -1784.3 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id655instance rdf:type swrs:Person.} } \\ 
 -1803.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id655instance ?p1 swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id848instance ?p1 swrs:Person.} } \\ 
 -1816.0 & 1 & \makecell{\texttt{\_:ub1bL355C18 rdf:type ?n1.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance ?p1 swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance rdf:type swrs:Person.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance rdf:type ?n1.} } \\ 
 -1804.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id810instance ?p1 swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance ?p1 swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance rdf:type swrs:Person.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance rdf:type ?n1.} } \\ 
 -1805.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id594instance ?p1 swrs:InProceedings.} } \\ 
 -1785.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id594instance rdf:type swrs:InProceedings.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id820instance ?p1 swrs:Person.} } \\ 
 -1807.3 & 1 & \makecell{\texttt{\_:ub1bL560C18 ?p1 swrs:Person.} } \\ 
 -1804.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id184instance ?p1 swrs:Person.} } \\ 
 -1786.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id841instance rdf:type swrs:InProceedings.} } \\ 
 -1805.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id841instance ?p1 swrs:InProceedings.} } \\ 
 -1786.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id684instance rdf:type swrs:InProceedings.} } \\ 
 -1784.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type swrs:Article.} } \\ 
 -1787.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type swrs:Publication.} } \\ 
 -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance ?p1 swrs:Publication.} } \\ 
 -1806.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id954instance ?p1 swrs:Publication.} } \\ 
 -1806.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id773instance ?p1 swrs:Publication.} } \\ 
 -1815.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id20instance rdf:type ?n1.} } \\ 
 -1803.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id350instance ?p1 swrs:Person.} } \\ 
 -1803.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id262instance ?p1 swrs:Person.} } \\ 
 -1784.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id262instance rdf:type swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance ?p1 swrs:Person.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance rdf:type ?n1.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance rdf:type swrs:Person.} } \\ 
 -1813.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id68instance swrs:phone ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id305instance rdf:type ?n1.} } \\ 
 -1822.2 & 1 & \makecell{\texttt{swrs:title ?p2 ?n1.} } \\ 
 -1791.2 & 1 & \makecell{\texttt{swrs:title ?p1 owl:DatatypeProperty.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance rdf:type ?n1.} } \\ 
 -1800.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p1 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id877instance.} } \\ 
 -1812.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance swrs:name ?n1.} } \\ 
 -1781.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance swrs:name "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1800.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p1 "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1831.1 & 1 & \makecell{\texttt{?n1 ?p2 "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1812.1 & 1 & \makecell{\texttt{?n1 swrs:name "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1813.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id79instance swrs:name ?n1.} } \\ 
 -1811.1 & 1 & \makecell{\texttt{?n1 swrs:name "CKDD - Conceptual Knowledge Discovery"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1830.1 & 1 & \makecell{\texttt{?n1 ?p2 "CKDD - Conceptual Knowledge Discovery"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1808.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id165instance swrs:name ?n1.} } \\ 
 -1785.0 & 1 & \makecell{\texttt{\_:ub1bL199C18 rdf:type owl:Restriction.} } \\ 
 -1815.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id18instance rdf:type ?n1.} } \\ 
 -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance rdf:type swrs:Publication.} } \\ 
 -1806.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance ?p1 swrs:Publication.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id593instance rdf:type ?n1.} } \\ 
 -1786.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id499instance rdf:type swrs:InProceedings.} } \\ 
 -1805.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id275instance ?p1 swrs:InProceedings.} } \\ 
 -1821.4 & 1 & \makecell{\texttt{swrs:member ?p2 ?n1.} } \\ 
 -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2118instance rdf:type swrs:Person.} } \\ 
 -1814.8 & 1 & \makecell{\texttt{swrs:Organization rdf:type ?n1.} } \\ 
 -1815.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id814instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id439instance rdf:type ?n1.} } \\ 
 -1785.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id637instance rdf:type swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id703instance rdf:type swrs:Person.} } \\ 
 -1787.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2008instance rdf:type swrs:Person.} } \\ 
 -1806.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2008instance ?p1 swrs:Person.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id804instance rdf:type ?n1.} } \\ 
 -1803.3 & 1 & \makecell{\texttt{swrs:Proceedings ?p1 owl:Class.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id786instance ?p1 swrs:Person.} } \\ 
 -1806.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance ?p1 swrs:Person.} } \\ 
 -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance rdf:type swrs:Person.} } \\ 
 -1818.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance rdf:type ?n1.} } \\ 
 -1787.3 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance rdf:type swrs:Publication.} } \\ 
 -1806.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance ?p1 swrs:Publication.} } \\ 
\hline
\end{longtable}

\subsection{Mutag, top 100 by log factor}

\begin{longtable}{ r r p{19cm} }
 log factor & frequency & \\
\hline\endhead
 11178.4 & 5996 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_hasAtom ?n2.} } \\ 
 1014.6 & 1840 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 938.0 & 1850 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 577.6 & 1897 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
 -1250.9 & 138 & \makecell{\texttt{?n1 rdfs:subClassOf ?n2.} } \\ 
 -1422.9 & 320 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Ar\_halide.} } \\ 
 -1593.5 & 319 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_charge ?n2.} } \\ 
 -1689.4 & 181 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -1707.3 & 173 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -1719.3 & 181 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -1737.1 & 173 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -1830.8 & 110 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -1859.3 & 110 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -1865.4 & 142 & \makecell{\texttt{?n1 rdf:type owl:Class.} } \\ 
 -1879.4 & 22 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -1886.3 & 19 & \makecell{\texttt{?n1 rdfs:subClassOf http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -1894.6 & 142 & \makecell{\texttt{?n1 ?p2 owl:Class.} } \\ 
 -1911.7 & 67 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -1928.9 & 50 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -1938.7 & 67 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -1981.2 & 36 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d219 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -1988.7 & 16 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d84 http://dl-learner.org/carcinogenesis\_hasStructure ?n1.} \\\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -1996.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} \\\texttt{http://dl-learner.org/carcinogenesis\_bond5500 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d270\_13.} } \\ 
 -2000.2 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester rdfs:subClassOf http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -2000.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2001.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_6.} } \\ 
 -2002.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 rdf:type owl:Class.} \\\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond100 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2003.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3151 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2008.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2008.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-318 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1585 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2172 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2009.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2182 rdf:type http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2010.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Molybdenum rdf:type owl:Class.} } \\ 
 -2010.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Carbon-22 rdf:type owl:Class.} } \\ 
 -2011.7 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} \\\texttt{http://dl-learner.org/carcinogenesis\_bond32 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d10\_6.} } \\ 
 -2012.2 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond3442 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond16 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2014.7 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 rdf:type http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
 -2017.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester ?p1 http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -2018.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 ?p1 http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2018.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 ?p1 http://dl-learner.org/carcinogenesis\_d135\_23.} } \\ 
 -2019.5 & 25 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d216 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 ?p1 http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2020.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3218 ?p1 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2020.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3151 ?p1 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2020.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p1 owl:Class.} \\\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2022.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2025.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2026.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-120 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-980 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1585 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2172 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-537 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2182 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-939 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1155 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1260 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-946 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2275 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p1 owl:Class.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ar\_halide ?p1 owl:Class.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Alcohol ?p1 owl:Class.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Carbon-22 ?p1 owl:Class.} } \\ 
 -2030.2 & 4 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2031.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -2031.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di260 rdfs:subClassOf ?n1.} } \\ 
 -2032.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 ?p1 http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
 -2032.9 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2033.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester rdfs:subClassOf ?n1.} } \\ 
 -2034.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Iodine-95 rdfs:subClassOf ?n1.} } \\ 
 -2034.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Chlorine-93 rdfs:subClassOf ?n1.} } \\ 
 -2034.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type ?n1.} } \\ 
 -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2036.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2036.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2041.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type ?n1.} } \\ 
 -2042.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2042.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_di10-3291 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2491 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2916 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-222 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-1445 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1069 rdf:type ?n1.} } \\ 
 -2042.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 rdf:type ?n1.} } \\ 
 -2042.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2634 rdf:type ?n1.} } \\ 
 -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Nitrogen-499 rdf:type ?n1.} } \\ 
 -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 rdf:type ?n1.} } \\ 
 -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Non\_ar\_6c\_ring rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-1525 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-2912 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-3038 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-216 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-180 rdf:type ?n1.} } \\ 
 -2044.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d243\_16 http://dl-learner.org/carcinogenesis\_charge ?n1.} } \\ 
 -2045.1 & 1 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2045.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d82\_26 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1421 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4108 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond7147 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond566 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond3448 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond662 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond5306 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond5197 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond6533 rdf:type ?n1.} } \\ 
 -2048.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 rdf:type ?n1.} } \\ 
 -2052.2 & 6 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2053.6 & 66 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3169 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} \\\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2056.6 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p2 ?n1.} } \\ 
 -2059.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p2 ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p2 ?n1.} } \\ 
 -2059.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 ?p2 ?n1.} } \\ 
 -2059.9 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Non\_ar\_6c\_ring ?p2 ?n1.} } \\ 
 -2059.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p2 ?n1.} } \\ 
 -2059.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2916 ?p2 ?n1.} } \\ 
 -2060.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 ?p2 ?n1.} } \\ 
 -2060.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1069 ?p2 ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 ?p2 ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2634 ?p2 ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-939 ?p2 ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1260 ?p2 ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2275 ?p2 ?n1.} } \\ 
 -2061.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di51 ?p2 ?n1.} } \\ 
 -2061.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di260 ?p2 ?n1.} } \\ 
 -2061.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-180 ?p2 ?n1.} } \\ 
 -2062.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester ?p2 ?n1.} } \\ 
 -2063.1 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Chlorine-93 ?p2 ?n1.} } \\ 
 -2065.9 & 9 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d104 http://dl-learner.org/carcinogenesis\_hasStructure ?n1.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -2071.1 & 2 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d135\_23.} } \\ 
 -2071.4 & 5 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2071.9 & 5 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2073.9 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 ?p2 ?n1.} } \\ 
 -2075.7 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 ?p2 ?n1.} } \\ 
 -2075.7 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 ?p2 ?n1.} } \\ 
 -2077.4 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d243\_16 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 ?p2 ?n1.} } \\ 
 -2086.2 & 3 & \makecell{\texttt{?n1 rdf:type ?n2.} \\\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d11\_4.} } \\ 
 -2094.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_di10-3291 rdf:type ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
 -2094.6 & 541 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
 -2114.7 & 92 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond ?n2.} } \\ 
 -2139.2 & 4 & \makecell{\texttt{?n1 ?p3 http://dl-learner.org/carcinogenesis\_d11\_4.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -2151.0 & 62 & \makecell{\texttt{?n1 ?p4 http://dl-learner.org/carcinogenesis\_d11\_4.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 -2743.8 & 126 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
\hline
\end{longtable}

\subsection{Mutag, top 100 by frequency}

\begin{longtable}{ r r p{19cm} }
 log factor & frequency & \\
\hline\endhead
 11178.4 & 5996 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_hasAtom ?n2.} } \\ 
 577.6 & 1897 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
 938.0 & 1850 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 1014.6 & 1840 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2094.6 & 541 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
 -1422.9 & 320 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Ar\_halide.} } \\ 
 -1593.5 & 319 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_charge ?n2.} } \\ 
 -1719.3 & 181 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -1689.4 & 181 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -1707.3 & 173 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -1737.1 & 173 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -1894.6 & 142 & \makecell{\texttt{?n1 ?p2 owl:Class.} } \\ 
 -1865.4 & 142 & \makecell{\texttt{?n1 rdf:type owl:Class.} } \\ 
 -1250.9 & 138 & \makecell{\texttt{?n1 rdfs:subClassOf ?n2.} } \\ 
 -2743.8 & 126 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
 -1859.3 & 110 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -1830.8 & 110 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2114.7 & 92 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond ?n2.} } \\ 
 -1911.7 & 67 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -1938.7 & 67 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2053.6 & 66 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3169 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} \\\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2151.0 & 62 & \makecell{\texttt{?n1 ?p4 http://dl-learner.org/carcinogenesis\_d11\_4.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 -1928.9 & 50 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -1981.2 & 36 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d219 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -2019.5 & 25 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d216 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -1879.4 & 22 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -1886.3 & 19 & \makecell{\texttt{?n1 rdfs:subClassOf http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -1988.7 & 16 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d84 http://dl-learner.org/carcinogenesis\_hasStructure ?n1.} \\\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2065.9 & 9 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d104 http://dl-learner.org/carcinogenesis\_hasStructure ?n1.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -2052.2 & 6 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2071.4 & 5 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2071.9 & 5 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2030.2 & 4 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2139.2 & 4 & \makecell{\texttt{?n1 ?p3 http://dl-learner.org/carcinogenesis\_d11\_4.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -2075.7 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 ?p2 ?n1.} } \\ 
 -2032.9 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2073.9 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 ?p2 ?n1.} } \\ 
 -2086.2 & 3 & \makecell{\texttt{?n1 rdf:type ?n2.} \\\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d11\_4.} } \\ 
 -2075.7 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 ?p2 ?n1.} } \\ 
 -2036.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2036.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2031.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2071.1 & 2 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d135\_23.} } \\ 
 -2062.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester ?p2 ?n1.} } \\ 
 -2063.1 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Chlorine-93 ?p2 ?n1.} } \\ 
 -2061.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di51 ?p2 ?n1.} } \\ 
 -2056.6 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p2 ?n1.} } \\ 
 -2059.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 ?p2 ?n1.} } \\ 
 -2061.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di260 ?p2 ?n1.} } \\ 
 -2077.4 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d243\_16 ?p2 ?n1.} } \\ 
 -2042.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2094.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_di10-3291 rdf:type ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
 -2059.9 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Non\_ar\_6c\_ring ?p2 ?n1.} } \\ 
 -2045.1 & 1 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 ?p1 http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond100 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2001.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_6.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2012.2 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2026.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2018.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 ?p1 http://dl-learner.org/carcinogenesis\_d135\_23.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-980 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1585 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1585 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1421 rdf:type ?n1.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-120 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4108 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond7147 rdf:type ?n1.} } \\ 
 -2017.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester ?p1 http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -2000.2 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester rdfs:subClassOf http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -2033.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester rdfs:subClassOf ?n1.} } \\ 
 -2034.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Iodine-95 rdfs:subClassOf ?n1.} } \\ 
 -2034.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Chlorine-93 rdfs:subClassOf ?n1.} } \\ 
 -2045.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d82\_26 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-1525 rdf:type ?n1.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2172 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2172 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Nitrogen-499 rdf:type ?n1.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-537 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2060.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 ?p2 ?n1.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 rdf:type ?n1.} } \\ 
 -2042.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 rdf:type ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 ?p2 ?n1.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2182 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2009.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2182 rdf:type http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond566 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-2912 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type ?n1.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2008.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2025.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2022.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2026.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2008.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2041.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type ?n1.} } \\ 
 -2059.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p2 ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p2 ?n1.} } \\ 
 -2059.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p2 ?n1.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type ?n1.} } \\ 
 -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 rdf:type ?n1.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p1 owl:Class.} } \\ 
 -2020.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p1 owl:Class.} \\\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2002.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 rdf:type owl:Class.} \\\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2000.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -1996.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} \\\texttt{http://dl-learner.org/carcinogenesis\_bond5500 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d270\_13.} } \\ 
 -2018.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 ?p1 http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2034.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2634 ?p2 ?n1.} } \\ 
 -2042.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2634 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond6533 rdf:type ?n1.} } \\ 
 -2048.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 rdf:type ?n1.} } \\ 
 -2014.7 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 rdf:type http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
 -2032.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 ?p1 http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
 -2010.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Molybdenum rdf:type owl:Class.} } \\ 
 -2031.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di260 rdfs:subClassOf ?n1.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 rdf:type ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-939 ?p2 ?n1.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-939 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond3448 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2491 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-3038 rdf:type ?n1.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1155 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1260 ?p2 ?n1.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1260 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-946 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ar\_halide ?p1 owl:Class.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Alcohol ?p1 owl:Class.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Carbon-22 ?p1 owl:Class.} } \\ 
 -2010.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Carbon-22 rdf:type owl:Class.} } \\ 
 -2044.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d243\_16 http://dl-learner.org/carcinogenesis\_charge ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-216 rdf:type ?n1.} } \\ 
 -2059.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2916 ?p2 ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2916 rdf:type ?n1.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond3442 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 rdf:type ?n1.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-318 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2042.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_di10-3291 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-180 rdf:type ?n1.} } \\ 
 -2061.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-180 ?p2 ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-222 rdf:type ?n1.} } \\ 
 -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Non\_ar\_6c\_ring rdf:type ?n1.} } \\ 
 -2060.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1069 ?p2 ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1069 rdf:type ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2275 ?p2 ?n1.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2275 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-1445 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type ?n1.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2011.7 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} \\\texttt{http://dl-learner.org/carcinogenesis\_bond32 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d10\_6.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond16 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond662 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond5306 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond5197 rdf:type ?n1.} } \\ 
 -2020.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3218 ?p1 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2020.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3151 ?p1 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2003.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3151 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
\hline
\end{longtable}

\end{landscape}
\end{document}

### 2018-09-06 kgmotifs/kgmotifs.tex
deadline approaching

@@
 \setlength{\pdfpageheight}{11in}  % Required
+\usepackage{makecell}
+\usepackage{arydshln}
+\usepackage{float}
 
@@
 We extend the common definition of a network motif to coincide with a \emph{basic graph pattern}, commonly used in the knowledge graph literature.
-We introduce an approach, inspired by recent work for simple graphs, for inducing these for a given knowledge knowledge graph, and show that the motifs returned reflect the basic structure of the graph.
+We introduce an approach, inspired by recent work for simple graphs, for inducing these for a given knowledge knowledge graph, and show that the motifs returned, reflect the basic structure of the graph.
 Specifically we show that in random graphs, no motifs are found, and that when we insert a motif artificially, it can be detected, even if as few as 10 instances are inserted.
@@
 \noindent \emph{Knowledge graphs} are an extremely versatile way of storing knowledge.
-They allow knowledge to be encoded without a predefined schema, they allow different amounts of data to be encoded for each known entity and they allow different modalities to be integrated seamlessly \cite{wilcke2017knowledge}.
+They allow knowledge to be encoded without a predefined schema, they allow different amounts of data to be encoded for different entities and they allow different modalities to be integrated seamlessly \cite{wilcke2017knowledge}.
 This versatility comes at a price.
@@
 What are small, recurring patterns that might represent a novel insight into the data?
+Answering these questions could benefit problem domains like graph simplification, graph navigation and schema induction.
 
@@
 Network motifs are small subgraphs whose frequency in the graph is unexpected with respect to a \emph{null model}.
-That is, for a given graph dataset $G$ and small graph $M$, we count the frequency $f(M, G)$: how often $M$ occurs in $G$ as a subgraph.
+
+% That is, for a given graph dataset $G$ and small graph $M$, we count the frequency $f(M, G)$: how often $M$ occurs in $G$ as a subgraph.
 We define a probability distribution over graphs $p^\text{null}(\G)$, and estimate the probability that a graph sampled from $p^\text{null}$ contains more instances of $M$ than observed in our data: $p^\text{null}(f(M, \G) \geq f(M, G))$, where $\G$ is a random variable representing a graph.
@@
 To avoid this costly operation, \cite{bloem2017large} introduces  an alternative method, using \emph{compression} as a heuristic for motif relevance: the better a motif compresses the data, the more likely it is to be meaningful.
-It is shown that this principle can be implemented in a similar sort of hypothesis test for a null model, allowing a comparable workflow to classical motif analysis.
 
-\footnotetext{This procedure has the structure of a hypothesis test, but it is important not to interpret it as \emph{statistical evidence} for the meaningfulness of the motif.
+%It is shown that this principle can be implemented in a similar sort of hypothesis test for a null model, allowing a comparable workflow to traditional motif analysis.
+
+%\footnotetext{This procedure has the structure of a hypothesis test, but it is important not to interpret it as \emph{statistical evidence} for the meaningfulness of the motif.
 The only thing it shows (in a frequentist statistical sense) is that $p^\text{null}$ is not the true source of the data.
@@
 We perform several experiments to show that our method returns meaningful subgraphs.
-First we test the intuition that a random graph shoudl contain no motifs.
-We also show that when we artifically insert motifs into a random graph, we can then detect these as motifs, sometimes with as few as 10 instances inserted into the graph.
+First we test the intuition that a random graph should contain no motifs.
+We also show that when we artificially insert motifs into a random graph, we can then detect these as motifs, sometimes with as few as 10 instances inserted into the graph.
 Finally, we show the results of motif analysis on three real-world knowledge graphs.
@@
 
-When we need to encode a single choice from a finite set $S$ of options, we will often use the code with length $\log |S|$, corresponding to a uniform distribution on $S$.
+% When we need to encode a single choice from a finite set $S$ of options, we will often use the code with length $\log |S|$, corresponding to a uniform distribution on $S$.
 In some cases, we can allow codes with multiple codewords for a single object.
@@
 A simple but crucial result, known as the \emph{no-hypercompression inequality} \cite[p103]{grunwald2007minimum} tells us that the probability of sampling data $x$ from $p^\text{null}$ that can be described shorter than $L^\text{null}(x)$ by $k$ or more bits, \emph{using any code} is less than $2^{-k}$.
-Thus, a standard MDL hypothesis test would consist of stating the null hypothesis in terms of a code $L^\text{null}$, designing a code (before seeing the data) which compresses the data better than $L^\text{null}$ by, say, 10 bits and rejecting the null hypothesis with confidence $2^{-10}$.
+Thus, an MDL hypothesis test would consist of stating the null hypothesis in terms of a code $L^\text{null}$, designing a code (before seeing the data) which compresses the data better than $L^\text{null}$ by, say, 10 bits and rejecting the null hypothesis with confidence $2^{-10}$.
 For a longer, more intuitive explanation of this principle in pattern induction, we refer the reader to \cite[Section~2]{bloem2017large}.
 
-We must note again, that when we use this procedure to find motifs, we are not providing evidence for the hypothesis that the motif is ``correct.'' The only thing we are statistically proving is that the null-model is incorrect.
-Nevertheless, hypothesis testing has shown to be a valuable heuristic in motif analysis.
-
+Note that when we use this procedure to find motifs, we are not providing evidence for the hypothesis that the motif is ``correct'' \cite[Section~2]{bloem2017large}.
 \paragraph{Common codes}
@@
 \end{align*}
-See \cite{de2016names} for a more intuitive explanation. In all experiments we use $\alpha = 0.5$, $d=0.1$. We will refer to the total for a sequence  codelength under this procedure as $L^PY(S)$
+See \cite{de2016names} for a more intuitive explanation. In all experiments we use $\alpha = 0.5$, $d=0.1$. We will refer to the total codelength for a sequence under this procedure as $L^{PY}(S)$
 
 \section{Method}
-
-We will start with some formal definitions of our basic ingredients.
 To comfortably define probabilities and codes on graphs, we will slightly deviate from conventional definitions.
 Specifically, we will analyse the \emph{structure} of knowledge graphs only, ignoring any meaning they have outside the graph.
-For instance, we will distinguish between links with different labels, but we will ignore any internal structure of the label itself (such as shared IRI prefixes).
-Under this assumption we can assume that the node and link labels of our knowledge graphs are simply natural numbers.\footnotemark This leads to the following definition.
+Under this assumption we can assume that the node and link labels of our knowledge graphs are simply natural numbers.\footnotemark~This leads to the following definition.
 
@@
 The negative integers are always contiguous within a single graph, with the highest representing the node labels and the lowest representing the edge labels.
-
+\begin{table}[H]x
+\centering
+\tiny
+\begin{tabular}{r r p{0.5\linewidth}}
+\hline
+log-factor & frequency & \\
+\hline
+\multicolumn{3}{c}{Dogfood , top 5 by log-factor ($>100$ positive)}\\
+\hline	
+ 83119.1 & 9344 & \makecell{\texttt{?n1 dc:creator ?n2.} \\\texttt{?n2 rdfs:label ?n3.} \\\texttt{?n2 foaf:name ?n3.} } \\ 
+ \hdashline
+ 81075.1 & 10322 & \makecell{\texttt{?n1 foaf:maker ?n2.} \\\texttt{?n2 rdfs:label ?n3.} \\\texttt{?n2 foaf:name ?n3.} } \\
+  \hdashline 
+ 77692.6 & 5234 & \makecell{\texttt{?n2 rdf:type ?n3.} \\\texttt{?n2 rdfs:label ?n1.} \\\texttt{?n2 foaf:name ?n1.} } \\ 
+  \hdashline
+ 75632.3 & 3536 & \makecell{\texttt{?n1 swc:isPartOf ?n2.} \\\texttt{?n2 swc:hasPart ?n1.} } \\ 
+  \hdashline
+ 75592.4 & 3536 & \makecell{\texttt{?n1 swc:isPartOf ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
+\hline
+\multicolumn{3}{c}{Dogfood, top 3 by frequency}\\
+\hline	
+ -3032.5 & 137008 & \makecell{\texttt{?n1 rdf:\_1 ?n2.} \\\texttt{?n1 rdf:\_2 ?n4.} \\\texttt{?n1 rdf:\_3 ?n3.} } \\ 
+  \hdashline
+ -3251.4 & 109315 & \makecell{\texttt{?n1 swc:isPartOf ?n3.} \\\texttt{?n2 swc:isPartOf ?n3.} } \\ 
+ \hdashline
+ 2089.0 & 92165 & \makecell{\texttt{?n2 swc:isPartOf ?n3.} \\\texttt{?n3 swc:hasPart ?n1.} } \\ 
+\hline
+\multicolumn{3}{c}{AIFB, top 5 by log-factor ($>100$ positive)}\\
+\hline	
+ 79234.0 & 7728 & \makecell{\texttt{?n1 ?p3 ?n2.} \\\texttt{?n2 ?p4 ?n1.} } \\ \hdashline
+ 61310.4 & 4154 & \makecell{\texttt{?n1 swrs:publication ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ \hdashline
+ 57641.1 & 3965 & \makecell{\texttt{?n1 swrs:publication ?n2.} \\\texttt{?n2 swrs:author ?n1.} } \\ \hdashline
+ 57603.1 & 3965 & \makecell{\texttt{?n1 swrs:author ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ \hdashline
+ 42110.1 & 17512 & \makecell{\texttt{?n1 swrs:publication ?n3.} \\\texttt{?n3 swrs:author ?n2.} \\\texttt{?n3 swrs:author ?n1.} \\\texttt{?n2 swrs:publication ?n3.} } \\ 
+\hline
+\multicolumn{3}{c}{AIFB, top 3 by frequency}\\
+\hline
+ 2287.7 & 198668 & \makecell{\texttt{?n1 swrs:publication ?n3.} \\\texttt{?n1 swrs:publication ?n2.} } \\ \hdashline
+ 7822.1 & 191083 & \makecell{\texttt{?n1 swrs:author ?n2.} \\\texttt{?n2 swrs:publication ?n3.} } \\ \hdashline
+ 1264.0 & 190636 & \makecell{\texttt{?n1 swrs:isAbout ?n3.} \\\texttt{?n2 swrs:isAbout ?n3.} } \\ 
+\hline
+\multicolumn{3}{c}{Mutag, top 5 by log-factor ($87$ positive)}\\
+\hline
+ 178304.4 & 18634 & \makecell{\texttt{?n1 mtg:\_hasAtom ?n3.} \\\texttt{?n1 mtg:\_hasBond ?n2.} \\\texttt{?n2 mtg:\_inBond ?n3.} } \\ \hdashline
+ 97237.8 & 9189 & \makecell{\texttt{?n1 mtg:\_hasAtom ?n2.} \\\texttt{?n2 mtg:\_charge ?n3.} } \\ \hdashline
+ 90447.5 & 18634 & \makecell{\texttt{?n1 mtg:\_hasBond ?n2.} \\\texttt{?n2 mtg:\_inBond ?n4.} \\\texttt{?n2 mtg:\_inBond ?n3.} } \\ \hdashline
+ 79052.0 & 9317 & \makecell{\texttt{?n1 mtg:\_hasBond ?n2.} \\\texttt{?n2 rdf:type ?n3.} } \\ \hdashline
+ 58973.9 & 7892 & \makecell{\texttt{?n1 mtg:\_hasAtom ?n2.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
+\hline
+\multicolumn{3}{c}{Mutag, top 3 by frequency}\\
+\hline
+ 1025.7 & 341143 & \makecell{\texttt{?n2 mtg:\_hasAtom ?n3.} \\\texttt{?n2 mtg:\_hasBond ?n1.} } \\ \hdashline
+ -374.5 & 178837 & \makecell{\texttt{?n1 mtg:\_charge ?n3.} \\\texttt{?n2 mtg:\_charge ?n3.} } \\ \hdashline
+ -2088.0 & 52985 & \makecell{\texttt{?n1 rdf:type owl:Class.} \\\texttt{?n3 rdfs:subClassOf ?n2.} \\\texttt{?n4 ?p5 owl:Class.} \\\texttt{?n4 rdfs:subClassOf ?n2.} } \\ 
+\hline
+\end{tabular}
+\caption{Results of the experiment on real-world data. For each experiment we also report the number of motifs found with a positive log-factor.}
+\label{table:topk}
+\end{table}
 An \emph{instance} for $M$ in $G$ is a pair of sequences of integers: $I = (I^n, I^r)$.
@@
 Moreover, we require that all instances in $\I$ are mutually disjoint: no two subgraphs defined by member of $\I$ may share an edge, but nodes may be shared.
-In most settings we do not actually have a given $M$ and $\I$ and we will need to search for them.
-We describe a simple and search algorithm in Section~\ref{section:search}.
 
-\label{section:relevance-test}
-
-\section{Null model}
-
-\label{section:null-model}
-
+\paragraph{Null model}
 The most common null model for classical motif analysis is the degree-sequence model (also known a the configuration model \cite{newman2010networks}): a uniform distribution over all graphs with a particular degree sequence.
@@
 \footnote{Or, equivalently, to make $p^\text{EL}$ a complete distribution on all graphs, we must provide it with a prior on $D$.} For each of the three sequences in $D$ we use the following model:
-\begin{align*}
- p(D) &= \prod_i q^{\N}(D_i) \\
- L(D) &= - \sum_i \log q^{\N}(D_i)
-\end{align*}
+\[
+ p(D) = \prod_i q^{\N}(D_i)\;\;\;\;\;
+ L(D) = - \sum_i \log q^{\N}(D_i)
+\]
 
@@
 Specifically, when we economize, we will only do so in a way that makes the hypothesis test \emph{more conservative}.
+
+\begin{table}[tbp]
+\small
+\begin{tabular}{r r p{4cm} r}
+\hline
+log-factor & frequency & &  \\
+\hline
+220360.2 & 12138 & \makecell{\texttt{?n1 foaf:maker ?n2.} \\\texttt{?n2 foaf:made ?n1.}} & D  \\ \hdashline
+ 3157.0 & 1011 & \makecell{\texttt{?n1 ?p2 "false".}}  & M \\ \hdashline
+  3150.2 & 985 & \makecell{\texttt{?n1 ?p2 "true".}} & M \\ \hdashline
+   12871.8 & 8308 & \makecell{\texttt{?n1 rdf:type ?n2.} \\\texttt{?n1 swrs:year ?n3.} \\\texttt{?n4 swrs:publication ?n1.} }& A \\ 
+\hdashline
+\end{tabular}
+\caption{Selected motifs.The last column indicates the dataset.}
+\label{table:selected}
+\end{table}
 
@@
 
-We note a few aspects of this code.
-Firstly, it achieves its compression by reducing the number of links it has to store.
-For each additional instance, it pays the cost of storing the instance, and any variable nodes and links, but it saves $|E_M|$ links in storing the template.
-The precise trade-off depends on the choice of base code, and on the structure of the rest of the graph (some edges are cheaper to store than others, for some codes).
-
-Nevertheless, we can make a few broad observations: patterns with many variable nodes and links are more costly to store than fully labeled patterns.
-On the other hand, a pattern with many variable nodes and links is more likely to have many instances, so that \emph{if} it results in a small amount of compression per instance, the large number of instances may cause it to compress a lot.
-For the variable nodes and links, we use a DM model for each specific node and link.
-This means that the more structure there is to the label given to a particular part of the pattern, the better we compress.
-For instance, suppose a particular node of the pattern is always labeled with either a node corresponding to YES or NO, in equal proportion, we will require only a bit per instance to record those labels.
-If, however, all nodes in the graph are equally likely to appear in the position of the variable, we will pay approximately $\log |V_G|$ bits per instance.
-If only a single node appears in the position, we pay only a a constant number of bits to record the entire sequence of instances.
-Thus, to our code, non-variable nodes are just a highly regular type of variable node.
-The distinction is only meaningful in the search algorithm.
-
 \begin{pseudo}[tb]
@@
 
-\subsection{Encoding motif instances}
+\paragraph{Encoding motif instances} To encode a list of instances $\I$ of a given pattern $M$, we generalize  the idea of the edgelist model described above.
+To understand the following ,it may be instructive to think of a single triple as an instance for the pattern \texttt{?n1 ?rel ?n2}.
+We can think of the edgelist model as storing a sequence of instances for this pattern, which together make up the entire knowledge graph.
 
-To encode a list of instances $\I$ of a given pattern $M$, we generalize  the idea of the edgelist model described above.
-To understand the following ,it may be instructive to think of a single triple as an instance for the pattern
-\[
-\texttt{?n1 ?rel ?n2}.
-\]
- We can now think of the edgelist model as storing a sequence of instances for this pattern, which together make up the entire knowledge graph.
-
- To generalize this notion to arbitrary patterns, to be defined for a given template graph, we define the \emph{degree constraint} $D^\I$ of a list of instamces for a given pattern as follows: for each variable node $i$ in the pattern, the degree sequence provides an integer sequence $D^i$ of length $v_G$, indicating how often each node in the completed knowledge graph takes that position in the pattern.
+ To generalize this notion to arbitrary patterns, to be defined for a given template graph, we define the \emph{degree constraint} $D^\I$ of a list of instances for a given pattern as follows: for each variable node $i$ in the pattern, the degree constraint provides an integer sequence $D^i$ of length $v_G$, indicating how often each node in the completed knowledge graph takes that position in the pattern.
 Similarly, for each variable link $j$ in the pattern, the degree sequence provides an integer sequence $C^j$ of length $r_G$ indicating for each relation how often it takes that position in the pattern.
@@
 
-Note that if we store a graph with the pattern $\texttt{?n1 ?rel ?n2.}$, we obtain an empty template graph, and the motif code achieves the same codelength as the edgelist model, up to a small constant amount for storing the pattern.
+Note that if we store a graph with the pattern \texttt{?n1 ?rel ?n2} we obtain an empty template graph, and the motif code achieves the same codelength as the edgelist model, up to a small constant amount for storing the pattern.
 
-For a given graph and pattern, we can simply find the complete list of instances using a graph patter search.
+For a given graph and pattern, we can simply find the complete list of instances using a graph pattern search.
 Since we require a slightly different semantics than standard graph pattern matchers, we adapt the DualIso algorithm \cite{saltz2014dualiso} for knowledge graph matching.
 Before computing the motif code, we prune the list of instances provided by this search iterating over the instances and removing and instance that produces a triple also produced by an earlier instance.
-
-Since graph matching encompasses subgraph isomorphism, it is known to be NP-Complete.
 To guard against rare patterns that produce long-running searches we terminate all searches after 5 seconds, returning  only those matches that were found within the time limit.
@@
 
+\begin{figure}[tb]
+  \centering
+    \includegraphics[width=\linewidth]{synthetic-plot.pdf}
+    \caption{The result of the random graph experiment. We sort the motifs by their score in the $k=10$ experiment and plot their frequency and log-factor.}
+    \label{figure:synthetic}
+\end{figure}
+
 \label{section:search}
@@
 The starting pattern is always a single random triple from the graph, with its relation made a variable.
-We define seven possible transition from one pattern to another:
-\begin{description}
-	\item[Extend (0.1)] Choose an instance of the pattern and an adjacent triple not part of the instance. Add the triple to the pattern.
-	\item[Make a node a variable (3)] Choose a random constant node, and turn it in to a variable node.
-	\item[Make an edge a variable (3)] Choose a random constant edge label, and turn it in to a variable (always introducing a new variable).
-	\item[Make a variable node constant (2)] Choose a random variable node and turn it into a constant. Take the value from a random instance.
-	\item[Make a variable edge constant (2)] Choose a random variable edge and turn it into a constant. Take the value from a random instance.
-	\item[Remove edge (3)] Remove a random edge from the pattern, ensuring that it stays connected.
-	\item[Couple (1)] Take two distinct edge variables, which for at least one instance hold the same value and turn them into a single variable.
-\end{description}
+We define seven possible transition from one pattern to another: \\
+\textbf{Extend (2)} Choose an instance of the pattern and an adjacent triple not part of the instance. Add the triple to the pattern.\\
+\textbf{Make a node a variable (3)} Choose a random constant node, and turn it in to a variable node.\\
+\textbf{Make an edge a variable (3)} Choose a random constant edge label, and turn it in to a variable (always introducing a new variable).\\
+\textbf{Make a variable node constant (2)} Choose a random variable node and turn it into a constant. Take the value from a random instance.\\
+\textbf{Make a variable edge constant (2)} Choose a random variable edge and turn it into a constant. Take the value from a random instance.\\
+\textbf{Remove edge (3)} Remove a random edge from the pattern, ensuring that it stays connected.\\
+\textbf{Couple (1)} Take two distinct edge variables, which for at least one instance hold the same value and turn them into a single variable.
 
-The values in brackets are the relative weights of each transition type (chosen by trial and error to keep the algorithm stable).
-We choose a transition at random and compute a new pattern.
+The values in brackets are the relative weights of each transition type (chosen by trial and error).
+We choose a transition at random using the normalized weights as probabilities and compute a new pattern.
 If the transition cannot be made (for instance, there are no constant node to make variable) or if the resultant pattern is in some way invalid, we sample a new transition.
@@
 If the codelength is lower, we continue with the new pattern.
-If the codelentgh is longer, we continue with the new sample with probability $\alpha$ or return to the previous pattern otherwise.
+If the codelength is longer, we continue with the new sample with probability $\alpha$ or return to the previous pattern otherwise.
 We use $\alpha = 0.5$ in all experiments.
@@
 In order to exploit all available processor cores, we run several searches in parallel.
-After each the searches are finished, we tajke the top 1000 patterns from each and sort them by motif codelength.
-If two searches return different codelengths\footnotemark for a motif we take the lowest value.
+After each the searches are finished, we take the top 1000 patterns from each and sort them by motif codelength.
 
-\footnotemark{This is possible if the pattern search is cut off due to the 5 second time limit.}
-
+ \begin{figure}[tb]
+  \centering
+    \includegraphics[width=\linewidth]{synthrep-plot.pdf}
+    \caption{The result of the repeated random graph experiment. Color and size show the number of matches of the pattern after pruning. Plot titles show the graph dimensions before adding instances.}
+    \label{figure:synthrep}
+\end{figure}
 \section{Experiments}
 
-To validate the method, we first test it on random graphs.
+\paragraph{Random graphs} To validate the method, we first test it on random graphs.
 We sample a directed graph with a given number of nodes $n$ and edges $m$, with no self-connections and multiple edges (that is, we sample from from the $G(n, m)$ Erd\H{o}s-Renyi model).
 We then label the nodes uniformly at random with one of the relations in $0, \ldots, r$.
-To make the dimensions realistic we base them on thos of the Mutag dataset used later: $n=23644$, $m74567$, $r=24$.
+To make the dimensions realistic we base them on those of the Mutag dataset used later: $n=23644$, $m74567$, $r=24$.
 
-We then take one randomly chosen pattern, and insert $k$ instances of the pattern into the graph. We run the experiment for $k=0$, $k=10$ and $k=100$.
+We then take one randomly chosen pattern, and insert $k$ instances of the pattern into the graph.
+We run the experiment for $k=0$, $k=75$ and $k=150$.
+The results are shown in Figure~\ref{figure:synthetic}.
+For $k=0$, as expected, we find no patterns with positive compression.
+For $k=75$, we find ...such patterns and for $k=150$ we find ...
+.
+While we only inserted one pattern we can expect to find more than one, since its subpatterns may also become motifs.
+However, the pattern we inserted should receive the highest log-factor.
 
-\section{Acknowledgments}
+Of course, this experiment only tests a single pattern. To see the effect of multiple random patterns, we repeat the experiment many times, sampling both the pattern and the random graph. 
+To sample the pattern we first sample a random number of nodes $n$ from $U(3, 6)$, the uniform distribution over the integer range $(3, 6)$ (including both end points).
+We then sample a random number of links $m$ from $U(n, n^2 - n)$, and sample a random directed graph from $G(n, m)$.
+We make $U(0, n)$ nodes and $U(0, m)$ links into variables, choosing constants for the rest uniformly from the data.
+If the pattern is disconnected, we reject and sample again.
 
-Withheld for purposes of blind reviewing.
+We sample a random graph as in the previous experiments, using the dimensions from the three real world datasets used later.
+We then add $k$ instances of the motif to the graph and compute the log-factor of the sampled pattern (we do not conduct a search).
+
+We let $k$ range from 0 to 200, and repeat the experiment 25 times for each $k$, sampling a new graph and pattern each time.
+The results are shown in Figure~\ref{figure:synthrep}.
+We observe first that under this ad-hoc sampling regime, we produce some patterns that create only very few instances in the graph, after overlapping instances are pruned.
+Since it is no surprise that these don't allow significant compression, we plot these as small points so that they don't obscure the other points.
+
+We see that most of the other instances---those that generate enough non-overlapping instances---can be retrieved as motifs.
+As the number of added instances increases, the proportion of motifs with separate instance that cannot be retrieved diminishes.
+
+\paragraph{Real-world data} Finally, we will test our method on real-world dat, to confirm that the motifs found coincide with our intuition.
+We test three datasets: The semantic Web dogfood dataset \cite{moller2007recipes} ($n=7611$,$m=242256$,$r=170$) describing researchers and publications in the Semantic Web domain, the AIFB dataset \cite{bloehdorn2007kernel} ($n=8285$,$m=29226$,$r=47$) describing the structure of the AIFB institute, and the Mutag RDF dataset \footnotemark ($n=23644$,$m=74567$,$r=24$), describing a set of carcinogenic and non-carcinogenic molecules both in structure and properties.
+
+\footnotetext{Distributed as an example dataset with the DL-Learner framework. \cite{lehmann2009dl}}
+
+For all datasets we run 32 parallel searches, with 3125 iterations per search.
+For each dataset we report the top 5 motifs by log-factor, and the top 3 motifs by frequency.
+Note that these are the top 10 motifs for frequency in the set of 32000 high-log-factor motifs resulting from the parallel search.
+The supplemental materials show the top 100 motifs for both criteria.
+
+The patterns with negative score but high frequency show the pitfalls our code avoids.
+For instance, in the top motif by frequency for the Dogfood dataset, we see a four-node pattern with two edges fanning in to a single node.
+These are caused by a single node having the same ingoing relation to many other nodes.
+Clearly such a pattern is very frequent in the data, but capturing just three of these incoming links does not constitute a satisfying motif, and indeed such patterns receive a negative log-factor.
+
+Much of what the motif code picks up on is redundancy in the original data.
+For instance, in the AIFB data both the \texttt{swrs:publication} relation and its inverse \texttt{swrs:author} are always included.
+Extracting these into a motif is simple way of achieving compression.
+In fact, the AIFB data caontains so many of these relation pairs that the two-node loop with variable labels is the highest scoring motif.
+In the Dogfood data, we see similar patterns emerge.
+
+Table~\ref{table:selected} shows some interesting motifs from the top 100 for each dataset.
+We see, for instance that e=the assertions that something is true or false are both motifs.
+The example from the AIFB date shows
+\section{Discussion}
+
+We have presented a new method for inducing graph patterns from knowledge graphs.
+To our knowledge, this is the first method presented that can potentially find arbitrary graph patterns to describe the innate structure of a knowledge graph.
+
+\subsection{Limitations and future work}
+
+Currently, the greatest limitation of this method is scalability.
+In \cite{bloem2017large}, the original method on which this method is based was shown to scale to graphs with billions of links.
+However, this scalability did not translate directly to knowledge graphs, since the sampling approach, used there is exteremely unlikely to generate the patterns that are succesful for motifs.
+If a better sampling method could be divised, we could likely scale the method up to web-scale knowledge graphs.
+
+Our method currently produces a large number of motifs.
+We can show that worthwhile motifs are included, and that it performs better than a baseline like selecting by frequency, but it still takes some manual effort to sort through the suggestions to find the kind of motifs that fit a particular usecase.
+This is not surprising; it is the nature of knowledge graphs that many different and overlapping substructures can be seen as natural or meaningful.
+One promising avenue to reduce this manual effort is to search of a \emph{set} of motifs which together compress well, each motif claiming a certain part of the knowledge graph to represent.
+If such an approach were fruitful it would truly represent a versatile and powerful form of schema induction for knowledge graphs.
+
+% \section{Acknowledgments}
+
 % This publication was supported by the Amsterdam Academic Alliance Data Science (AAA-DS) Program Award to the UvA and VU Universities.

### 2021-04-16 kgmotifs/kgmotifs.tex
...

@@
-\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
+\documentclass[runningheads]{style/llncs}
 
-\usepackage{aaai19}    % Required
-\usepackage{times}     % Required
-\usepackage{helvet}    % Required
-\usepackage{courier}   % Required
-\usepackage{url}       % Required
-\usepackage{graphicx}  % Required
-\frenchspacing         % Required
-\setlength{\pdfpagewidth}{8.5in}  % Required
-\setlength{\pdfpageheight}{11in}  % Required
+\usepackage{url}       
+\usepackage{graphicx}  
+\usepackage{float}
+
 \usepackage{makecell}
 \usepackage{arydshln}
-\usepackage{float}
-
-% PDF Info Is Required:
-\pdfinfo{
-/Title (Detecting Motifs in Knowledge Graphs using Compression)
-/Author (...)}
-
-\setcounter{secnumdepth}{0}
 
@@
 \usepackage{float}
+
+\usepackage{booktabs}
 
@@
 
+\newcommand{\p}{\;\text{.}}
+
 \newcommand{\I}{{\cal I}}
@@
 
-\title{Detecting Motifs in Knowledge Graphs using Compression}
-\author{
-Author names withheld
-% Peter Bloem\\
-% Knowledge Representation \& Reasoning Group\\
-% VU University\\
-% vu@peterbloem.nl\\
-}
+\begin{document}
 
-\begin{document}
+\title{Finding Motifs in Knowledge Graphs using Compression}
+\titlerunning{Finding Motifs in Knowledge Graphs}
+
+\author{Peter Bloem\\ \orcidID{0000-0002-0189-5817}}
+% 
+%
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+\institute{Knowledge Representation and Reasoning group,\\ Vrije Universiteit Amsterdam \\
+\email{vu@peterbloem.nl} \\
+\url{peterbloem.nl}}
 
@@
 \begin{abstract}
-\noindent We introduce a method to detect \emph{network motifs} in knowledge graphs.
+\noindent We introduce a method to find \emph{network motifs} in knowledge graphs.
 Network motifs are useful patterns or meaningful subunits of the graph that recur frequently.
-We extend the common definition of a network motif to coincide with a \emph{basic graph pattern}, commonly used in the knowledge graph literature.
-We introduce an approach, inspired by recent work for simple graphs, for inducing these for a given knowledge knowledge graph, and show that the motifs returned, reflect the basic structure of the graph.
-Specifically we show that in random graphs, no motifs are found, and that when we insert a motif artificially, it can be detected, even if as few as 10 instances are inserted.
-Finally, we show the results of motif induction on three real-world knowledge graphs, and show that our method provides good suggestions for motifs that reflect the structure of the graph in an intuitive way.
+We extend the common definition of a network motif to coincide with a \emph{basic graph pattern}.
+We introduce an approach, inspired by recent work for simple graphs, to induce these from a given knowledge knowledge graph, and show that the motifs found reflect the basic structure of the graph.
+Specifically, we show that in random graphs, no motifs are found, and that when we insert a motif artificially, it can be detected.
+Finally, we show the results of motif induction on three real-world knowledge graphs.
 \end{abstract}
 
-\noindent \emph{Knowledge graphs} are an extremely versatile way of storing knowledge.
-They allow knowledge to be encoded without a predefined schema, they allow different amounts of data to be encoded for different entities and they allow different modalities to be integrated seamlessly \cite{wilcke2017knowledge}.
+\section{Introduction}
+\noindent \emph{Knowledge graphs} are an extremely versatile and flexible data model.
+They allow knowledge to be encoded without a predefined format and they are extremely robust in the face of missing data.
 This versatility comes at a price.
@@
 a) A basic knowledge graph.
-Note that in our analysis, we consider only the indices of the nodes and relations, not their labels.
+We consider only the indices of the nodes and relations, not their labels.
 b) A motif that occurs frequently.
-c) A compressed representation; we remove all edges that are part of an occurrence of the motif.
-We then store separately which nodes match the motif.
-The motif, together with the template graph and the matches can be used to reconstruct the graph.
-We take this as a heuristic: the better the motif compresses, the more characteristic it is likely to be for the graph.}
+c) A compressed representation; we remove all edges that are part of an occurrence of the motif and store separately which nodes match the motif.}
     \label{figure:example}
@@
 
-In this paper, we extend the compression-based motif analysis to knowledge graphs.
+In this paper, we extend this compression-based motif analysis to \emph{knowledge graphs}.
 For the purposes of this research we define knowledge graphs as labeled, directed multigraphs.
 Nodes are uniquely labeled with entity names, and links are non-uniquely labeled with relations.
-We extend the definition of a motif to that of a \emph{basic graph pattern}: that is, a motif is a small graph with \emph{some} of its nodes and links labeled with entities and relations as they occur in the graph.
-Other nodes and links are labeled as \emph{variables}.
-A pattern matches at a particular location in the graph, if the variables can be replaced with values from the graph, and the motif becomes a subgraph as a result.
-The intuition behind our method is that the more compactly a pattern compresses the graph, the better the motif.
-Figure~\ref{figure:example} shows an illustration of the basic principle.
-In the preliminaries section below, we justify this intuition more formally.
+We extend the definition of a motif to that of a \emph{basic graph pattern}: a small graph labeled with both variables and explicit entitites and relations.
+A pattern matches if the variables can be replaced with specific values from the graph so that the pattern becomes a subgraph as a result.
+
+The intuition behind our method is that we can use graph patterns to \emph{compress the graph}: we store the pattern, its instances, and the remainder of the graph.
+The better this representation compresses the graph, the more relevant the pattern.
+Figure~\ref{figure:example} illustrates the principle.
+In Section~\ref{section:preliminaries}, we justify this intuition more formally.
 
@@
 First we test the intuition that a random graph should contain no motifs.
-We also show that when we artificially insert motifs into a random graph, we can then detect these as motifs, sometimes with as few as 10 instances inserted into the graph.
-Finally, we show the results of motif analysis on three real-world knowledge graphs.
+We also show that when we artificially insert motifs into a random graph, we can then detect these as motifs.
+Finally, we show the results of motif analysis on three real-world knowledge graphs, compared to the baseline of selecting the most frequent graph patterns.
 
-All code and datasets used in this paper are available, under open licenses.\footnotemark
+All code and datasets used in this paper are available online.\footnotemark
 
-\footnotetext{URL withheld for purposes of blind reviewing.}
+\footnotetext{\url{https://github.com/MaestroGraph/motive-rdf}}
 
@@
 Network motifs for unlabeled simple graphs were introduced in \cite{milo2002network}.
-A more comprehensive overview of the related literature can be found in \cite[Section~1.1]{bloem2017large}.
+A more comprehensive overview of the related literature can be found in \cite{bloem2017large}.
 In \cite{bloem2017large}, the principle of Minimum Description Length (MDL) was first connected to motif analysis.
@@
 
-A few other methods have been proposed for inducing the structure of a given knowledge graph.
+A few other methods have been proposed for inducing the structure of a given knowledge graph in terms of subgraphs.
 In \cite{pham2015deriving}, the authors use the principle of characteristic sets to characterize a knowledge graph in terms of the star patterns it contains.
 In \cite{pham2016exploiting}, they show that the majority of the LOD cloud can be efficiently described using such principles, showing the highly tabular structure of many knowledge graphs.
-In \cite{volker2011statistical}, association rules mining is used to induce basic patterns in the graph.
+In \cite{volker2011statistical}, association rule mining is used to induce basic patterns in the graph.
 
@@
 
+\label{section:preliminaries}
+
 \paragraph{Minimum Description Length}
-Our method is based on the MDL principle: that we should favour models that compress the data.
+Our method is based on the MDL principle: we should favour models that compress the data.
 We will show briefly how this intuition can be made mathematically precise.
-For more details, we refer the reader to \cite{grunwald2007minimum} for MDL in general, and to \cite{bloem2017large}, for a more extensive discussion these principles in the domain of graph analysis.
+For more details, we refer the reader to \cite{grunwald2007minimum} for MDL in general, and to \cite{bloem2018tutorial}, for a more extensive discussion these principles in the domain of graph analysis.
 
@@
 Let $\log(x) = \log_2(x)$.
-A \emph{code} for a set of objects $\cal X$ (such as graphs) is an injective function $f: {\cal X} \to \B$, mapping objects to binary code words.
+A \emph{code} for a set of objects $\cal X$  is an injective function $f: {\cal X} \to \B$.
 All codes in this paper are \emph{prefix-free}: no code word is the prefix of another.
@@
 $L(x) = |f(x)|$.
-We commonly compute $L(x)$ directly, without first computing $f(x)$ and we refer to $L(x)$ as a code.
+We commonly compute $L(x)$ directly, without first computing $f(x)$.
 
-A well known result in information theory is the association between codes and probability distributions, implied by the \emph{Kraft inequality}: for each probability distribution $p$ on $\cal X$, there exists a prefix-free code $L$ such that for all $x \in \cal X$: $- \log p(x) \leq L(x) < -\log p(x) + 1$.
+There is a strong relation between codes and probability distributions: for each probability distribution $p$ on $\cal X$, there exists a prefix-free code $L$ such that for all $x \in \cal X$: $- \log p(x) \leq L(x) < -\log p(x) + 1$.
 Inversely, for every prefix-free code $L$ for $\cal X$, there exists a probability distribution $p$ such that for all $x \in \cal X$: $p(x) = 2^{-L(x)}$.
 For proofs, see \cite[Section~3.2.1]{grunwald2007minimum} or \cite[Theorem~5.2.1]{cover2006elements}.
-To explain the intuition, note that we can easily transform a code $L$ into a sampling algorithm for $p$ by feeding the decoding function random bits until it produces an output.
+
+% To explain the intuition, note that we can easily transform a code $L$ into a sampling algorithm for $p$ by feeding the decoding function random bits until it produces an output.
 To transform a probability distribution to a code, techniques like arithmetic coding \cite{rissanen1979arithmetic} can be used.\footnotemark
-\footnotetext{
-As explained in \cite[page 96]{grunwald2007minimum}, the fact that $-\log p^*(x)$ is real-valued and $L^*(x)$ is integer-valued can be safely ignored and we may \emph{identify} codes with probability distributions, allowing codes to take non-integer values.
-}
+%\footnotetext{
+%As explained in \cite[page 96]{grunwald2007minimum}, the fact that $-\log p^*(x)$ is real-valued and $L^*(x)$ is integer-valued can be safely ignored and we may \emph{identify} codes with probability distributions, allowing codes to take non-integer values.
+%}
 
@@
 \paragraph{Relevance testing} We will use the MDL principle to perform a hypothesis test.
-Assume we have some data $x$ and a null hypothesis that it was sampled from distribution $p^\text{null}$.
-Let $L^\text{null}$ be the corresponding codelength.
-A simple but crucial result, known as the \emph{no-hypercompression inequality} \cite[p103]{grunwald2007minimum} tells us that the probability of sampling data $x$ from $p^\text{null}$ that can be described shorter than $L^\text{null}(x)$ by $k$ or more bits, \emph{using any code} is less than $2^{-k}$.
-Thus, an MDL hypothesis test would consist of stating the null hypothesis in terms of a code $L^\text{null}$, designing a code (before seeing the data) which compresses the data better than $L^\text{null}$ by, say, 10 bits and rejecting the null hypothesis with confidence $2^{-10}$.
-For a longer, more intuitive explanation of this principle in pattern induction, we refer the reader to \cite[Section~2]{bloem2017large}.
+Assume we have some data $x \in \B$ and a null hypothesis stating that it was sampled from distribution $p^\text{null}$ (with corresponding code $L^\text{null}$).
+A simple but crucial result, known as the \emph{no-hypercompression inequality} \cite[p103]{grunwald2007minimum} tells us that the probability of sampling any data $x$ from $p^\text{null}$ that can be described in less than $L^\text{null}(x) - k$ or more bits, \emph{using any code} is less than $2^{-k}$.
+Thus, we can reject the hypothesis that the data was sampled from $p^\text{null}$ by designing a code which compresses the data better than $L^\text{null}$ by, say, 10 bits and rejecting the null hypothesis with confidence $2^{-10}$.
+For a longer, more intuitive explanation of this principle in pattern induction, we refer the reader to \cite{bloem2018tutorial}.
 
-Note that when we use this procedure to find motifs, we are not providing evidence for the hypothesis that the motif is ``correct'' \cite[Section~2]{bloem2017large}.
+Note that when we use this procedure to find motifs, we are not providing statistical evidence for the hypothesis that the motif is ``correct'' \cite[Section~4.1]{bloem2018tutorial}.
+We are simply using the principle of hypothesis testing as a \emph{heuristic} for pattern mining.
+The only assertion we are proving (in a statistical sense) is that the data did not come from the null model.
+
 \paragraph{Common codes}
-In the construction of our codes for graphs, we require some simpler codes as building blocks.
+In the construction of our graph codes, we require some simpler codes as building blocks.
 First, when we store any positive integer $n$, we do so with the code corresponding to the distribution $p^\N(n) = 1/(n(n+1))$, and denote it $L^\N(n)$.
 For nonnegative numbers we add 1 to the argument.
-for the full range of integers, we add an extra bit for the sign, and then use the first code for negative integers and the second for positive ones.
+For the full range of integers ($L^\Z$), we add an extra bit for the sign, and then use the first code for negative integers and the second for positive ones.
 
-We will often need to encode sequences of integers.
+We will often need to encode \emph{sequences} of integers as well.
 These will be highly skewed, with only a subset of integers occurring frequently, and others occurring infrequently or not at all.
@@
 Let $S = \langle S_1, ..., S_n\rangle$ be a sequence of integers of length $n$.
-We first store its members $m(s)$ in the order in which they first occur: we first store $n$ and the first member using the codes described above.
-The rest of the members are stored sequentially using the distance from the current to the next member.
+We first store its members $m(S)$ in the order in which they occur: we first store $n$ and the first member using $L^\N$ and then store each subsequent member by encoding the distance to the previous member using $L^\Z$.
+Having encoded the members of $S$ we can store the sequence itself using the Pitman-Yor model as follows.
 
-Let $f(A, B)$ be the frequency of symbol $A$ in sequence $B$.We then store the complete sequence using the code corresponding to the following distribution:
-\begin{align*}
-&p^\text{PY}_{\alpha,d}(S) = \prod_{i \in [1,k]} \text{PY}_{\alpha,d}(S_i\mid S_{1:i-1}) \\
-&\text{PY}_{\alpha,d}(S_i \mid S') = 
+Let $f(A, B)$ be the frequency of symbol $A$ in sequence $B$. We then store the complete sequence using the code corresponding to the following distribution:
+\[
+p(S) = \prod_{i \in [1,k]} p(S_i\mid S_{1:i-1})\;\;\text{with}\;\;
+p(S_i \mid S') = 
 \begin{cases}
-\frac{\alpha - d |m(S)|}{|m(S)| + \alpha} & \text{if} f(S_i, S') = 0 \\ 
-\frac{f(S_i, S') - d}{|m(S)| + \alpha} & \text{otherwise}
+\frac{\alpha - d |m(S')|}{|m(S')| + \alpha} & \text{if} f(S_i, S') = 0 \\ 
+\frac{f(S_i, S') - d}{|m(S')| + \alpha} & \text{otherwise}
 \end{cases}
-\end{align*}
-See \cite{de2016names} for a more intuitive explanation. In all experiments we use $\alpha = 0.5$, $d=0.1$. We will refer to the total codelength for a sequence under this procedure as $L^{PY}(S)$
+\]
+See \cite{de2016names} for a more intuitive explanation. In all experiments we use $\alpha = 0.5$, $d=0.1$. We will refer to the total resulting codelength as $L^{PY}(S)$.
 
 \section{Method}
-To comfortably define probabilities and codes on graphs, we will slightly deviate from conventional definitions.
-Specifically, we will analyse the \emph{structure} of knowledge graphs only, ignoring any meaning they have outside the graph.
-Under this assumption we can assume that the node and link labels of our knowledge graphs are simply natural numbers.\footnotemark~This leads to the following definition.
 
+We will first give a precise definition of a knowledge graph as used in this paper.
+We will then describe the null model which is used both as a point of comparison in our hypothesis test, and within the motif code to compress the remainder of the graph.
+Next, we describe how to compress a graph using a given motif, and a set of instances.
+Finally, we will describe how to search for likely motifs using simulated annealing.
+%
+%We analyse the \emph{structure} of knowledge graphs only, ignoring any meaning in relation to other graphs, encoded in the content of names or literals, or from ontology languages. 
+Specifically, we model a knowledge graph as a multigraph with nodes and edges labeled with integers that map to entities and relations.
+This mapping is stored, but only the integer-labeled graph is modelled.\footnotemark
 \footnotetext{For practitioners this restriction is not noticeable, as the indices can simply be mapped back to the original strings when the found motifs are presented.}
 
-A \emph{knowledge graph} $G$, hereafter simply a \emph{graph}, is a tuple $G = (v_G, r_G, E_G)$.
+A \emph{knowledge graph} $G$, is a tuple $G = (v_G, r_G, E_G)$.
 $v_G \in \N$ is the number of nodes in the graph, and $r_G \in \N$ is the number of relations.
-We define the nodeset as $V_G = \{0, \ldots, v_G-1\}$ and the relation-set as $R_G = \{0, \ldots, r_G\}$.
-The \emph{tripleset} $E_G \subset V_G \times R \times V_G$ determines the edges of the graph and their labels.
+We define the nodeset of $G$ as $V_G = \{0, \ldots, v_G-1\}$ and the relation-set as $R_G = \{0, \ldots, r_G\}$.
+The \emph{tripleset} $E_G \subset V_G \times R \times V_G$ determines the edges of the graph and their labels: each triple $(s, r, o) \in E_G$ encodes the originated node $s$, the target node $o$ and the relation $r$ of an edge in the graph.
 
-As shown in Figure~\ref{figure:example}, the patterns that we aim to find are partially labeled: some nodes and edges are labeled with as variables.
-We will represent these with negative integers.
-A pattern $M$ for graph $G$ is a tuple $(V_M, R_M, G, E_M)$.
-Let $v_M$ and $r_M$ indicate the number of variable nodes and variable links in $M$ respectively, then $V_M \subseteq \{-v'_M, \ldots, v_G-1\}$ and $R_M \subseteq \{-(r'_M+v'M), \ldots,-v'M, 0,\ldots, r_G-1\}$, with $E_M \subset V_M \times R_M \times V_M$, as for regular graphs.
+This definition is compatible with RDF data. We interpret literals as nodes, considered the same node if they are expressed by the same string.
+
+%As shown in Figure~\ref{figure:example}, the patterns that we aim to find are partially labeled: some nodes and edges are labeled as \emph{variables}. We will represent these with negative integers.
+
+A \emph{pattern} $M$ for graph $G$ is a tuple $(V_M, R_M, G, E_M)$.
+Let $v_M$ and $r_M$ indicate the number of variable nodes and variable links in $M$ respectively, then $V_M \subseteq \{-v'_M, \ldots, v_G-1\}$ and $R_M \subseteq \{-(r'_M+v'M), \ldots,-v'M, 0,\ldots, r_G-1\}$, with $E_M \subset V_M \times R_M \times V_M$ representing the edges as before.
 That is; nodes can be labeled either with nonnegative integers referring to $G$'s nodes or with negative integers representing a variable node, and similar for relations.
-The negative integers are always contiguous within a single graph, with the highest representing the node labels and the lowest representing the edge labels.
-\begin{table}[H]x
-\centering
-\tiny
-\begin{tabular}{r r p{0.5\linewidth}}
-\hline
-log-factor & frequency & \\
-\hline
-\multicolumn{3}{c}{Dogfood , top 5 by log-factor ($>100$ positive)}\\
-\hline	
- 83119.1 & 9344 & \makecell{\texttt{?n1 dc:creator ?n2.} \\\texttt{?n2 rdfs:label ?n3.} \\\texttt{?n2 foaf:name ?n3.} } \\ 
- \hdashline
- 81075.1 & 10322 & \makecell{\texttt{?n1 foaf:maker ?n2.} \\\texttt{?n2 rdfs:label ?n3.} \\\texttt{?n2 foaf:name ?n3.} } \\
-  \hdashline 
- 77692.6 & 5234 & \makecell{\texttt{?n2 rdf:type ?n3.} \\\texttt{?n2 rdfs:label ?n1.} \\\texttt{?n2 foaf:name ?n1.} } \\ 
-  \hdashline
- 75632.3 & 3536 & \makecell{\texttt{?n1 swc:isPartOf ?n2.} \\\texttt{?n2 swc:hasPart ?n1.} } \\ 
-  \hdashline
- 75592.4 & 3536 & \makecell{\texttt{?n1 swc:isPartOf ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
-\hline
-\multicolumn{3}{c}{Dogfood, top 3 by frequency}\\
-\hline	
- -3032.5 & 137008 & \makecell{\texttt{?n1 rdf:\_1 ?n2.} \\\texttt{?n1 rdf:\_2 ?n4.} \\\texttt{?n1 rdf:\_3 ?n3.} } \\ 
-  \hdashline
- -3251.4 & 109315 & \makecell{\texttt{?n1 swc:isPartOf ?n3.} \\\texttt{?n2 swc:isPartOf ?n3.} } \\ 
- \hdashline
- 2089.0 & 92165 & \makecell{\texttt{?n2 swc:isPartOf ?n3.} \\\texttt{?n3 swc:hasPart ?n1.} } \\ 
-\hline
-\multicolumn{3}{c}{AIFB, top 5 by log-factor ($>100$ positive)}\\
-\hline	
- 79234.0 & 7728 & \makecell{\texttt{?n1 ?p3 ?n2.} \\\texttt{?n2 ?p4 ?n1.} } \\ \hdashline
- 61310.4 & 4154 & \makecell{\texttt{?n1 swrs:publication ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ \hdashline
- 57641.1 & 3965 & \makecell{\texttt{?n1 swrs:publication ?n2.} \\\texttt{?n2 swrs:author ?n1.} } \\ \hdashline
- 57603.1 & 3965 & \makecell{\texttt{?n1 swrs:author ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ \hdashline
- 42110.1 & 17512 & \makecell{\texttt{?n1 swrs:publication ?n3.} \\\texttt{?n3 swrs:author ?n2.} \\\texttt{?n3 swrs:author ?n1.} \\\texttt{?n2 swrs:publication ?n3.} } \\ 
-\hline
-\multicolumn{3}{c}{AIFB, top 3 by frequency}\\
-\hline
- 2287.7 & 198668 & \makecell{\texttt{?n1 swrs:publication ?n3.} \\\texttt{?n1 swrs:publication ?n2.} } \\ \hdashline
- 7822.1 & 191083 & \makecell{\texttt{?n1 swrs:author ?n2.} \\\texttt{?n2 swrs:publication ?n3.} } \\ \hdashline
- 1264.0 & 190636 & \makecell{\texttt{?n1 swrs:isAbout ?n3.} \\\texttt{?n2 swrs:isAbout ?n3.} } \\ 
-\hline
-\multicolumn{3}{c}{Mutag, top 5 by log-factor ($87$ positive)}\\
-\hline
- 178304.4 & 18634 & \makecell{\texttt{?n1 mtg:\_hasAtom ?n3.} \\\texttt{?n1 mtg:\_hasBond ?n2.} \\\texttt{?n2 mtg:\_inBond ?n3.} } \\ \hdashline
- 97237.8 & 9189 & \makecell{\texttt{?n1 mtg:\_hasAtom ?n2.} \\\texttt{?n2 mtg:\_charge ?n3.} } \\ \hdashline
- 90447.5 & 18634 & \makecell{\texttt{?n1 mtg:\_hasBond ?n2.} \\\texttt{?n2 mtg:\_inBond ?n4.} \\\texttt{?n2 mtg:\_inBond ?n3.} } \\ \hdashline
- 79052.0 & 9317 & \makecell{\texttt{?n1 mtg:\_hasBond ?n2.} \\\texttt{?n2 rdf:type ?n3.} } \\ \hdashline
- 58973.9 & 7892 & \makecell{\texttt{?n1 mtg:\_hasAtom ?n2.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
-\hline
-\multicolumn{3}{c}{Mutag, top 3 by frequency}\\
-\hline
- 1025.7 & 341143 & \makecell{\texttt{?n2 mtg:\_hasAtom ?n3.} \\\texttt{?n2 mtg:\_hasBond ?n1.} } \\ \hdashline
- -374.5 & 178837 & \makecell{\texttt{?n1 mtg:\_charge ?n3.} \\\texttt{?n2 mtg:\_charge ?n3.} } \\ \hdashline
- -2088.0 & 52985 & \makecell{\texttt{?n1 rdf:type owl:Class.} \\\texttt{?n3 rdfs:subClassOf ?n2.} \\\texttt{?n4 ?p5 owl:Class.} \\\texttt{?n4 rdfs:subClassOf ?n2.} } \\ 
-\hline
-\end{tabular}
-\caption{Results of the experiment on real-world data. For each experiment we also report the number of motifs found with a positive log-factor.}
-\label{table:topk}
-\end{table}
+The negative integers are always contiguous within a single pattern, with the highest representing the node labels and the lowest representing the edge labels
+
 An \emph{instance} for $M$ in $G$ is a pair of sequences of integers: $I = (I^n, I^r)$.
@@
 $I^r$ is a sequence of non-distinct integers of length $r'_M$.
-For each edge $(s, p, o) \in E_M$ containing a negative $s$, $p$ or $o$, there is a corresponding link in $E_G$ with a negative $s$ replaced by $I^n_{-s}$, a negative $o$ replaced by $I^n_{-o}$, and a negative $p$ replaced by $I^r_{-p - v'_M}$.
+For each edge $(s, p, o) \in E_M$ with any or all of $s$, $p$ and $o$ negative, there is a corresponding link in $E_G$ with a negative $s$ replaced by $I^n_{-s}$, a negative $o$ replaced by $I^n_{-o}$, and a negative $p$ replaced by $I^r_{-p - v'_M}$.
 Put simply: for a pattern to match, variable edges marked with the same negative integer, must map to the same relation in order for the pattern to match, but variable links labeled with different negative integers \emph{may} map to the same relation.
@@
 Each edge in the motif may only match one edge in the graph.
-In other words, the occurrence of the motif in the graph must have as many edges as the motif itself.
+In other words, the occurrence of the motif in the graph must have as many edges as the motif itself.\footnotemark
 
-We will first assume that a target pattern $M$ is given for the data $G$ and that we have a set of instances $\I$.
-Moreover, we require that all instances in $\I$ are mutually disjoint: no two subgraphs defined by member of $\I$ may share an edge, but nodes may be shared.
+\footnotetext{In this aspect our definition differs from the SPARQL Basic Graph Pattern.
+Patterns for which this distinction is relevant are rare, and patterns returned by our method are still compatible with SPARQL.}
 
-\paragraph{Null model}
-The most common null model for classical motif analysis is the degree-sequence model (also known a the configuration model \cite{newman2010networks}): a uniform distribution over all graphs with a particular degree sequence.
-We extend this to knowledge graphs by also including the degree of each relation: that is, the frequency with which it occurs in the tripleset.
+\subsection{Null model}
+
+For a proper hypothesis test, we must compare the compression achieved by our motif code to the compression under a general model for knowledge graphs: a null model.
+
+The most common null model in classical motif analysis is the degree-sequence model (also known as the configuration model \cite{newman2010networks}): a uniform distribution over all graphs with a particular degree sequence.
+We extend this to knowledge graphs by also including the degree of each relation: that is, degree of a relation is the frequency with which it occurs in the tripleset.
 Let a \emph{degree sequence} $D$ of length $n$ be a triple of three integer sequences: $(D^\text{in}, D^\text{rel}, D^\text{out})$.
@@
 
-Let $\G_D$ be the set of all graphs with degree sequence $D$. Then the configuration model can be expressed simply as
+Let $\G_D$ be the set of all graphs with degree sequence $D$. Then the degree-sequence model can be expressed simply as
 \[
-p^C(G) = \frac{1}{|\G_D|}
+p^\text{DS}(G) = \frac{1}{|\G_D|}
 \]
@@
  {m \choose {D_1^\text{rel}, \ldots, D_{|R_G|}^\text{rel}} }
- {m \choose {D_1^\text{in}, \ldots, D_n^\text{in}} }} \leq p^C(G) \text{.}
+ {m \choose {D_1^\text{in}, \ldots, D_n^\text{in}} }} \leq p^\text{DS}(G) \text{.}
 \]
@@
 \begin{align*}
-- &\log p^\text{EL}_D(G) = 2 \log(m!) \\
-&- \sum_i \log(D_i^\text{in}!)- \sum_i \log(D_i^\text{rel}!)- \sum_i \log(D_i^\text{out}!)
+- \log p^\text{EL}_D(G) = 2 \log(m!)
+- \sum_i \log(D_i^\text{in}!)- \sum_i \log(D_i^\text{rel}!)- \sum_i \log(D_i^\text{out}!)
 \end{align*}
@@
 We call this the edgelist (EL) model.
-It gives a probability that always lower-bounds the configuration model, since it affords some probability mass to graphs that cannot exist.\footnotemark ~Experiments in the classical motif setting have shown that the EL model is an acceptable proxy for the DS model \cite{bloem2017large}, especially considering the extra scalability it affords.
+It gives a probability that always lower-bounds the configuration model, since it affords some probability mass to graphs that cannot exist.
+Experiments in the classical motif setting have shown that the EL model is an acceptable proxy for the DS model \cite{bloem2017large}, especially considering the extra scalability it affords.
 
-\footnotetext{Note that we cannot simply think of $p^\text{EL}$ as a uniform model for graphs containing multiple links, since we can only divide by $m!$ if we know that all triples are unique.}
+%\footnotetext{Note that we cannot simply think of $p^\text{EL}$ as a uniform model for graphs containing multiple links, since we can only divide by $m!$ if we know that all triples are unique.}
+% 
+ \paragraph{Encoding D} In order to encode a graph with $L^\text{EL}_D$, we must first encode $D$.\footnotemark~For each of the three sequences $D'$ in $D$ we use the following model:
+\[
+ p(D') = \prod_i q^{\N}(D'_i)\;\;\;\;\;
+ L(D') = - \sum_i \log q^{\N}(D'_i)
+\]
 
- \paragraph{Encoding D} In order to encode a graph with $L^\text{EL}_D$, we must first encode $D$.
-\footnote{Or, equivalently, to make $p^\text{EL}$ a complete distribution on all graphs, we must provide it with a prior on $D$.} For each of the three sequences in $D$ we use the following model:
-\[
- p(D) = \prod_i q^{\N}(D_i)\;\;\;\;\;
- L(D) = - \sum_i \log q^{\N}(D_i)
-\]
+\footnotetext{Or, equivalently, to make $p^\text{EL}$ a complete distribution on all graphs, we must provide it with a prior on $D$.}
 
@@
 When we use $p^\text{EL}$ as part of the motif code, we must use a fair encoding, so we use the Pitman-Yor code to store each sequence in $D$.
-
-In the design of our method, we will constantly aim to find a trade-off between completeness and efficiency that allows the method to scale to very large graphs.
+%
+%In the design of our method, we will constantly aim to find a trade-off between completeness and efficiency that allows the method to scale to very large graphs.
 Specifically, when we economize, we will only do so in a way that makes the hypothesis test \emph{more conservative}.
-
-\begin{table}[tbp]
-\small
-\begin{tabular}{r r p{4cm} r}
-\hline
-log-factor & frequency & &  \\
-\hline
-220360.2 & 12138 & \makecell{\texttt{?n1 foaf:maker ?n2.} \\\texttt{?n2 foaf:made ?n1.}} & D  \\ \hdashline
- 3157.0 & 1011 & \makecell{\texttt{?n1 ?p2 "false".}}  & M \\ \hdashline
-  3150.2 & 985 & \makecell{\texttt{?n1 ?p2 "true".}} & M \\ \hdashline
-   12871.8 & 8308 & \makecell{\texttt{?n1 rdf:type ?n2.} \\\texttt{?n1 swrs:year ?n3.} \\\texttt{?n4 swrs:publication ?n1.} }& A \\ 
-\hdashline
-\end{tabular}
-\caption{Selected motifs.The last column indicates the dataset.}
-\label{table:selected}
-\end{table}
 
@@
 
-As described above, we can perform our relevance test with any compression method which exploits the pattern $M$, and its instances $\I$ to store the graph efficiently.
+Having defined our representation of a knowledge graph, and a general null model for compressing such knowledge graphs, we can now define how we use a given motif (together with its instances) to compress a dataset.
+
+We will assume that a target pattern $M$ is given for the data $G$ and that we have a set of instances $\I$ of $M$ in $G$.
+Moreover, we require that all instances in $\I$ are mutually disjoint: no two subgraphs defined by a member of $\I$ may share an edge, but nodes may be shared.
+Given this information, we will define a motif code that will help us determine whether or not $M$ is a likely motif for $G$.
+In section Section~\ref{section:search}, we detail a method to search for pairs $(M, \I)$ to pass to the motif code.
+
+As described above, we can perform our relevance test with any compression method which exploits the pattern $M$ and its instances $\I$ to store the graph efficiently.
 The better our method, the more motifs we will find.
 Note that there is no need for our code to be optimal in any sense.
-So long as we accept that we won't find all motifs that exist, we are free to trade off compression performance against efficiency of computation.
+We know that we will not find all motifs that exist, and we will not use them optimally to represent the graph, but the test is still valid.
+This also means that we are free to trade off compression performance against efficiency of computation.
 
@@
 The information in all of these together is sufficient to reconstruct the graph.
-Note that everything is stored using prefix-free codes, so that we can simply concatenate the codewords (or rather, sum the codelengths we get for each aspect).
+Note that everything is stored using prefix-free codes, so that we can simply concatenate the codewords we get for each aspect, to get a codeword for the whole graph.
 
@@
 
-\emph{\# Turn the pattern into a normal knowledge graph}\\
+\emph{// Turn the pattern into a normal knowledge graph}\\
 $E_{M'} \leftarrow$ the edges of $M$ with contiguous integer labels \\
@@
 
-\emph{\# Store the template graph}\\
+\emph{// Store the template graph}\\
 $E'_G \leftarrow E_G - \cup_{\I \in {\cal I}} \text{triples}(I)$ \\
@@
 \paragraph{Encoding motif instances} To encode a list of instances $\I$ of a given pattern $M$, we generalize  the idea of the edgelist model described above.
-To understand the following ,it may be instructive to think of a single triple as an instance for the pattern \texttt{?n1 ?rel ?n2}.
+
+%To understand the following ,it may be instructive to think of a single triple as an instance for the pattern \texttt{?n1 ?rel ?n2}.
 We can think of the edgelist model as storing a sequence of instances for this pattern, which together make up the entire knowledge graph.
@@
 If this set of sequences satisfies the degree constraint, we know that node $q$ must occur $D^i_q$ times in sequence $N^i$, and similarly for the variable links.
-Let $\cS_\I$ be the set of all such integer sequences satisfying the degree constraint.
-We will follow the same logic as for the EdgeList model.
+Let $\cS_\I$ be the set of all such integer sequences satisfying the  constraint.
+We follow the same logic as for the EL model.
 Let $k$ be the number of matches of the pattern.
@@
  \begin{align*}
- - \log p_M(G) &= (w+l-1)\log(k!) \\
-  -&\sum_{j\in [1,w], i} \log(D^j_i!) -\sum_{j\in [1,l], i} \log(C^j_i!)
+ - \log p_M(G) &= (w+l-1)\log(k!)
+  -\sum_{j\in [1,w], i} \log(D^j_i!) -\sum_{j\in [1,l], i} \log(C^j_i!)
  \end{align*}
@@
 
-We express the strength of a motif by its log-factor: $-\log p^\text{null}(G) + \log p^\text{motif}(G; M, \I, L^\text{base})$.
-If this value is positive, the motif code compresses the graph better than the null model.
-If the log-factor is greater than 10 bits, it corresponds to a rejection of the null model at $p < 0.001$.
+We express the strength of a motif by its log-factor: 
+\[
+-\log p^\text{null}(G) + \log p^\text{motif}(G; M, \I, L^\text{base})\p
+\] 
+If this value is positive, the motif code compresses the graph better than the null model. If the log-factor is greater than 10 bits, it corresponds to a rejection of the null model at $p < 0.001$.
 
@@
     \includegraphics[width=\linewidth]{synthetic-plot.pdf}
-    \caption{The result of the random graph experiment. We sort the motifs by their score in the $k=10$ experiment and plot their frequency and log-factor.}
+    \caption{The result of the random graph experiment. We sort the motifs by their score in the $k=75$ experiment and plot their frequency and log-factor.}
     \label{figure:synthetic}
@@
 Since we can readily compute the log-factor, any black-box optimization algorithm can be used to search the space of all possible motifs.
-For the sake of simplicity, we will use basic simulated annealing here.
-The starting pattern is always a single random triple from the graph, with its relation made a variable.
-We define seven possible transition from one pattern to another: \\
-\textbf{Extend (2)} Choose an instance of the pattern and an adjacent triple not part of the instance. Add the triple to the pattern.\\
-\textbf{Make a node a variable (3)} Choose a random constant node, and turn it in to a variable node.\\
-\textbf{Make an edge a variable (3)} Choose a random constant edge label, and turn it in to a variable (always introducing a new variable).\\
-\textbf{Make a variable node constant (2)} Choose a random variable node and turn it into a constant. Take the value from a random instance.\\
-\textbf{Make a variable edge constant (2)} Choose a random variable edge and turn it into a constant. Take the value from a random instance.\\
-\textbf{Remove edge (3)} Remove a random edge from the pattern, ensuring that it stays connected.\\
-\textbf{Couple (1)} Take two distinct edge variables, which for at least one instance hold the same value and turn them into a single variable.
+For the sake of simplicity, we will use basic simulated annealing here: We start with a given pattern, and iterate by modifying the pattern in one of seven ways, chosen randomly.
+At each iteration, we search for instances of the pattern (limiting the time per search to 5 seconds) and compute the log-factor.
+If the log factor is better, we move to the new pattern, if it is worse, we move to the new pattern with probability 0.5.
 
-The values in brackets are the relative weights of each transition type (chosen by trial and error).
-We choose a transition at random using the normalized weights as probabilities and compute a new pattern.
-If the transition cannot be made (for instance, there are no constant node to make variable) or if the resultant pattern is in some way invalid, we sample a new transition.
+The starting pattern is always a single random triple from the graph, with its relation made a variable. We define seven possible transition from one pattern to another: \\
+\textbf{Extend} Choose an instance of the pattern and an adjacent triple not part of the instance. Add the triple to the pattern.\\
+\textbf{Make a node a variable} Choose a random constant node, and turn it in to a variable node.\\
+\textbf{Make an edge a variable } Choose a random constant edge label, and turn it in to a variable (always introducing a new variable).\\
+\textbf{Make a variable node constant} Choose a random variable node and turn it into a constant. Take the value from a random instance.\\
+\textbf{Make a variable edge constant} Choose a random variable edge and turn it into a constant. Take the value from a random instance.\\
+\textbf{Remove edge} Remove a random edge from the pattern, ensuring that it stays connected.\\
+\textbf{Couple} Take two distinct edge variables, which for at least one instance hold the same value and turn them into a single variable.
 
-Once a new pattern has been sampled, we compare its codelength under the motif model to that of the previous sample.
+All transitions are equally likely.
+If the transition cannot be made (for instance, there are no constant nodes to make variable) or if the resultant pattern is in some way invalid, we sample a new transition.
+
+%Once a new pattern has been sampled, we compare its codelength under the motif model to that of the previous sample.
 If the codelength is lower, we continue with the new pattern.
@@
 
-We store all patterns seen and their scores.
+We store all encountered patterns and their scores.
 In order to exploit all available processor cores, we run several searches in parallel.
-After each the searches are finished, we take the top 1000 patterns from each and sort them by motif codelength.
+We take the top 1000 patterns from each and sort them by motif codelength.
+Variables are re-ordered to a canonical ordering using the Nauty algorithm \cite{mckay1981practical}, so that isomorphic patterns are not tested twice.
 
@@
   \centering
-    \includegraphics[width=\linewidth]{synthrep-plot.pdf}
+    \includegraphics[width=0.9\linewidth]{synthrep-plot.pdf}
     \caption{The result of the repeated random graph experiment. Color and size show the number of matches of the pattern after pruning. Plot titles show the graph dimensions before adding instances.}
@@
 
-We then take one randomly chosen pattern, and insert $k$ instances of the pattern into the graph.
-We run the experiment for $k=0$, $k=75$ and $k=150$.
+We then take one randomly chosen pattern, and insert $k=75$ instances of the pattern into the graph.
+We run a search for 100000 iterations.
+And collect the 10 motifs with the best log-factor.
+We then sample two other graphs by the same method: one with $k=0$ and one with $k=150$.
+We also test each of the 10 motifs found on these two graphs.
+
  The results are shown in Figure~\ref{figure:synthetic}.
 For $k=0$, as expected, we find no patterns with positive compression.
-For $k=75$, we find ...such patterns and for $k=150$ we find ...
-.
-While we only inserted one pattern we can expect to find more than one, since its subpatterns may also become motifs.
-However, the pattern we inserted should receive the highest log-factor.
+We also ran a full search on this graph to verify that no motifs can be found unless they are explicitly added to the graph.
 
-Of course, this experiment only tests a single pattern. To see the effect of multiple random patterns, we repeat the experiment many times, sampling both the pattern and the random graph. 
+This experiment only tests a single pattern. To see the effect of multiple random patterns, we repeat the experiment many times, sampling both the pattern and the random graph.
+
 To sample the pattern we first sample a random number of nodes $n$ from $U(3, 6)$, the uniform distribution over the integer range $(3, 6)$ (including both end points).
@@
 We see that most of the other instances---those that generate enough non-overlapping instances---can be retrieved as motifs.
-As the number of added instances increases, the proportion of motifs with separate instance that cannot be retrieved diminishes.
 
-\paragraph{Real-world data} Finally, we will test our method on real-world dat, to confirm that the motifs found coincide with our intuition.
-We test three datasets: The semantic Web dogfood dataset \cite{moller2007recipes} ($n=7611$,$m=242256$,$r=170$) describing researchers and publications in the Semantic Web domain, the AIFB dataset \cite{bloehdorn2007kernel} ($n=8285$,$m=29226$,$r=47$) describing the structure of the AIFB institute, and the Mutag RDF dataset \footnotemark ($n=23644$,$m=74567$,$r=24$), describing a set of carcinogenic and non-carcinogenic molecules both in structure and properties.
+%\begin{table}[p]
+%\centering
+%\tiny
+%\begin{tabular}{r r p{0.5\linewidth}}
+%\hline
+%log-factor & frequency & \\
+%\hline
+%\multicolumn{3}{c}{Dogfood , top 5 by log-factor ($>100$ positive)}\\
+%\hline	
+% 83119.1 & 9344 & \makecell{\texttt{?n1 dc:creator ?n2.} \\\texttt{?n2 rdfs:label ?n3.} \\\texttt{?n2 foaf:name ?n3.} } \\ 
+% \hdashline
+% 81075.1 & 10322 & \makecell{\texttt{?n1 foaf:maker ?n2.} \\\texttt{?n2 rdfs:label ?n3.} \\\texttt{?n2 foaf:name ?n3.} } \\
+%  \hdashline 
+% 77692.6 & 5234 & \makecell{\texttt{?n2 rdf:type ?n3.} \\\texttt{?n2 rdfs:label ?n1.} \\\texttt{?n2 foaf:name ?n1.} } \\ 
+%  \hdashline
+% 75632.3 & 3536 & \makecell{\texttt{?n1 swc:isPartOf ?n2.} \\\texttt{?n2 swc:hasPart ?n1.} } \\ 
+%  \hdashline
+% 75592.4 & 3536 & \makecell{\texttt{?n1 swc:isPartOf ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
+%\hline
+%\multicolumn{3}{c}{Dogfood, top 3 by frequency}\\
+%\hline	
+% -3032.5 & 137008 & \makecell{\texttt{?n1 rdf:\_1 ?n2.} \\\texttt{?n1 rdf:\_2 ?n4.} \\\texttt{?n1 rdf:\_3 ?n3.} } \\ 
+%  \hdashline
+% -3251.4 & 109315 & \makecell{\texttt{?n1 swc:isPartOf ?n3.} \\\texttt{?n2 swc:isPartOf ?n3.} } \\ 
+% \hdashline
+% 2089.0 & 92165 & \makecell{\texttt{?n2 swc:isPartOf ?n3.} \\\texttt{?n3 swc:hasPart ?n1.} } \\ 
+%\hline
+%\multicolumn{3}{c}{AIFB, top 5 by log-factor ($>100$ positive)}\\
+%\hline	
+% 79234.0 & 7728 & \makecell{\texttt{?n1 ?p3 ?n2.} \\\texttt{?n2 ?p4 ?n1.} } \\ \hdashline
+% 61310.4 & 4154 & \makecell{\texttt{?n1 swrs:publication ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ \hdashline
+% 57641.1 & 3965 & \makecell{\texttt{?n1 swrs:publication ?n2.} \\\texttt{?n2 swrs:author ?n1.} } \\ \hdashline
+% 57603.1 & 3965 & \makecell{\texttt{?n1 swrs:author ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ \hdashline
+% 42110.1 & 17512 & \makecell{\texttt{?n1 swrs:publication ?n3.} \\\texttt{?n3 swrs:author ?n2.} \\\texttt{?n3 swrs:author ?n1.} \\\texttt{?n2 swrs:publication ?n3.} } \\ 
+%\hline
+%\multicolumn{3}{c}{AIFB, top 3 by frequency}\\
+%\hline
+% 2287.7 & 198668 & \makecell{\texttt{?n1 swrs:publication ?n3.} \\\texttt{?n1 swrs:publication ?n2.} } \\ \hdashline
+% 7822.1 & 191083 & \makecell{\texttt{?n1 swrs:author ?n2.} \\\texttt{?n2 swrs:publication ?n3.} } \\ \hdashline
+% 1264.0 & 190636 & \makecell{\texttt{?n1 swrs:isAbout ?n3.} \\\texttt{?n2 swrs:isAbout ?n3.} } \\ 
+%\hline
+%\multicolumn{3}{c}{Mutag, top 5 by log-factor ($87$ positive)}\\
+%\hline
+% 178304.4 & 18634 & \makecell{\texttt{?n1 mtg:\_hasAtom ?n3.} \\\texttt{?n1 mtg:\_hasBond ?n2.} \\\texttt{?n2 mtg:\_inBond ?n3.} } \\ \hdashline
+% 97237.8 & 9189 & \makecell{\texttt{?n1 mtg:\_hasAtom ?n2.} \\\texttt{?n2 mtg:\_charge ?n3.} } \\ \hdashline
+% 90447.5 & 18634 & \makecell{\texttt{?n1 mtg:\_hasBond ?n2.} \\\texttt{?n2 mtg:\_inBond ?n4.} \\\texttt{?n2 mtg:\_inBond ?n3.} } \\ \hdashline
+% 79052.0 & 9317 & \makecell{\texttt{?n1 mtg:\_hasBond ?n2.} \\\texttt{?n2 rdf:type ?n3.} } \\ \hdashline
+% 58973.9 & 7892 & \makecell{\texttt{?n1 mtg:\_hasAtom ?n2.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
+%\hline
+%\multicolumn{3}{c}{Mutag, top 3 by frequency}\\
+%\hline
+% 1025.7 & 341143 & \makecell{\texttt{?n2 mtg:\_hasAtom ?n3.} \\\texttt{?n2 mtg:\_hasBond ?n1.} } \\ \hdashline
+% -374.5 & 178837 & \makecell{\texttt{?n1 mtg:\_charge ?n3.} \\\texttt{?n2 mtg:\_charge ?n3.} } \\ \hdashline
+% -2088.0 & 52985 & \makecell{\texttt{?n1 rdf:type owl:Class.} \\\texttt{?n3 rdfs:subClassOf ?n2.} \\\texttt{?n4 ?p5 owl:Class.} \\\texttt{?n4 rdfs:subClassOf ?n2.} } \\ 
+%\hline
+%\end{tabular}
+%\vspace{2mm}
+%\caption{Results of the experiment on real-world data. For each experiment we also report the number of motifs found with a positive log-factor.}
+%\label{table:topk}
+%\end{table}
 
-\footnotetext{Distributed as an example dataset with the DL-Learner framework. \cite{lehmann2009dl}}
+\begin{table}[p]
+\centering
+\tiny
+\begin{tabular}{r r p{0.5\linewidth}}
+\hline
+log-factor & frequency & \\
+\hline
+\multicolumn{3}{c}{Dogfood , top 5 by log-factor ($>100$ positive)}\\
+\hline	
+ 361495.0 & 10475 & \makecell{\texttt{?n1 dc:creator ?n2.} \\\texttt{?n1 foaf:maker ?n2.} \\\texttt{?n2 foaf:made ?n1.} } \\ 
+ \hdashline 244579.5 & 7674 & \makecell{\texttt{?n1 dc:creator ?n2.} \\\texttt{?n1 foaf:maker ?n2.} \\\texttt{?n1 swrs:author ?n2.} } \\ 
+  \hdashline 220360.2 & 12138 & \makecell{\texttt{?n1 foaf:maker ?n2.} \\\texttt{?n2 foaf:made ?n1.} } \\ 
+ \hdashline  189627.3 & 9888 & \makecell{\texttt{?n1 foaf:member ?n2.} \\\texttt{?n2 swrs:affiliation ?n1.} } \\ 
+ \hdashline  187972.9 & 10475 & \makecell{\texttt{?n1 dc:creator ?n2.} \\\texttt{?n2 foaf:made ?n1.} } \\ 
+\hline
+\multicolumn{3}{c}{Dogfood, top 3 by frequency}\\
+\hline	
+ -3076.2 & 134853 & \makecell{\texttt{?n1 rdf:\_1 ?n2.} \\\texttt{?n1 rdf:\_2 ?n4.} \\\texttt{?n1 rdf:\_3 ?n3.} } \\ 
+ \hdashline  -3435.0 & 116074 & \makecell{\texttt{?n1 swc:heldBy ?n3.} \\\texttt{?n1 swc:heldBy ?n2.} } \\ 
+ \hdashline  -2379.9 & 110461 & \makecell{\texttt{?n1 rdf:type owl:Thing.} \\\texttt{?n2 rdf:type owl:Thing.} } \\ 
+\hline
+\multicolumn{3}{c}{AIFB, top 5 by log-factor ($>100$ positive)}\\
+\hline	
+ 79234.0 & 7549 & \makecell{\texttt{?n1 ?p3 ?n2.} \\\texttt{?n2 ?p4 ?n1.} } \\ 
+ \hdashline 61310.4 & 4154 & \makecell{\texttt{?n1 swrs:publication ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
+ \hdashline 57641.1 & 3965 & \makecell{\texttt{?n1 swrs:publication ?n2.} \\\texttt{?n2 swrs:author ?n1.} } \\ 
+ \hdashline 57603.1 & 3965 & \makecell{\texttt{?n1 swrs:author ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
+ \hdashline 33168.0 & 7930 & \makecell{\texttt{?n1 swrs:publication ?n2.} \\\texttt{?n2 rdf:type ?n3.} \\\texttt{?n2 swrs:author ?n1.} } \\ 
+\hline
+\multicolumn{3}{c}{AIFB, top 3 by frequency}\\
+\hline
+ -908.2 & 181246 & \makecell{\texttt{?n1 swrs:year ?n3.} \\\texttt{?n2 swrs:year ?n3.} } \\ 
+ \hdashline -1524.3 & 173059 & \makecell{\texttt{?n1 swrs:publication ?n3.} \\\texttt{?n1 swrs:publication ?n2.} } \\ 
+ \hdashline -1667.9 & 103434 & \makecell{\texttt{?n1 swrs:member ?n2.} \\\texttt{?n3 ?p5 ?n1.} \\\texttt{?n4 swrs:author ?n2.} } \\  
+\hline
+\multicolumn{3}{c}{Mutag, top 5 by log-factor ($87$ positive)}\\
+\hline
+ 178304.4 & 18634 & \makecell{\texttt{?n1 mtg:\_hasAtom ?n3.} \\\texttt{?n1 mtg:\_hasBond ?n2.} \\\texttt{?n2 mtg:\_inBond ?n3.} } \\ 
+ \hdashline 97237.8 & 9189 & \makecell{\texttt{?n1 mtg:\_hasAtom ?n2.} \\\texttt{?n2 mtg:\_charge ?n3.} } \\ 
+ \hdashline 93819.3 & 8924 & \makecell{\texttt{?n2 rdf:type ?n3.} \\\texttt{?n2 mtg:\_charge ?n1.} } \\ 
+ \hdashline 90447.5 & 18634 & \makecell{\texttt{?n1 mtg:\_hasBond ?n2.} \\\texttt{?n2 mtg:\_inBond ?n4.} \\\texttt{?n2 mtg:\_inBond ?n3.} } \\ 
+ \hdashline 79027.5 & 8924 & \makecell{\texttt{?n1 mtg:\_hasAtom ?n2.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
+\hline
+\multicolumn{3}{c}{Mutag, top 3 by frequency}\\
+\hline
+   -2040.6 & 68514 & \makecell{\texttt{?n1 rdfs:subClassOf ?n2.} \\\texttt{?n3 rdf:type owl:Class.} \\\texttt{?n4 rdf:type owl:Class.} \\\texttt{?n4 rdfs:subClassOf ?n2.} } \\ 
+ \hdashline -2077.8 & 60832 & \makecell{\texttt{?n1 ?p5 owl:Class.} \\\texttt{?n3 rdfs:subClassOf ?n2.} \\\texttt{?n4 ?p5 owl:Class.} \\\texttt{?n4 rdfs:subClassOf ?n2.} } \\ 
+ \hdashline -1532.6 & 32009 & \makecell{\texttt{?n1 mtg:\_cytogen\_sce "true".} \\\texttt{?n1 mtg:\_salmonella ?n3.} \\\texttt{?n2 mtg:\_amesTestPositive ?n3.} } \\ 
+\hline
+\end{tabular}
+\vspace{2mm}
+\caption{Results of the experiment on real-world data. For each experiment we also report the number of motifs found with a positive log-factor.}
+\label{table:topk}
+\end{table}
 
-For all datasets we run 32 parallel searches, with 3125 iterations per search.
-For each dataset we report the top 5 motifs by log-factor, and the top 3 motifs by frequency.
-Note that these are the top 10 motifs for frequency in the set of 32000 high-log-factor motifs resulting from the parallel search.
-The supplemental materials show the top 100 motifs for both criteria.
+\paragraph{Real data} Finally, we will test our method on real data, to confirm that the motifs found coincide with our intuition.
+We test three datasets: The Semantic Web dogfood dataset \cite{moller2007recipes} ($n=7611, m=242256, r=170$) describing researchers and publications in the Semantic Web domain, the AIFB dataset \cite{bloehdorn2007kernel} ($n=8285, m=29226, r=47$) describing the structure of the AIFB institute, and the Mutag RDF dataset\footnotemark~ ($n=23644, m=74567, r=24$), describing a set of carcinogenic and non-carcinogenic molecules both in structure and properties.
 
-The patterns with negative score but high frequency show the pitfalls our code avoids.
-For instance, in the top motif by frequency for the Dogfood dataset, we see a four-node pattern with two edges fanning in to a single node.
-These are caused by a single node having the same ingoing relation to many other nodes.
-Clearly such a pattern is very frequent in the data, but capturing just three of these incoming links does not constitute a satisfying motif, and indeed such patterns receive a negative log-factor.
+\footnotetext{Originally distributed as an example dataset with the DL-Learner framework \cite{lehmann2009dl}.}
+
+For all datasets, we run 32 parallel searches, with 3125 iterations per search.
+Table~\ref{table:topk} reports the top 5 motifs by log-factor, and the top 3 motifs by frequency.
+We provide the top 100 motifs under both criteria online.
+\footnotemark
+
+\footnotetext{\url{https://github.com/MaestroGraph/motive-rdf}}
+
+\begin{table}[bt]
+\centering
+\begin{tabular}{r r c r}
+\hline
+log-factor & frequency & &  \\
+\hline
+220360.2 & 12138 & \makecell{\texttt{?n1 foaf:maker ?n2.} \\\texttt{?n2 foaf:made ?n1.}} & Dogfood  \\ \hdashline
+ 3157.0 & 1011 & \makecell{\texttt{?n1 ?p2 "false".}}  & Mutag \\ \hdashline
+  3150.2 & 985 & \makecell{\texttt{?n1 ?p2 "true".}} & Mutag \\ \hdashline
+   12871.8 & 8308 & \makecell{\texttt{?n1 rdf:type ?n2.} \\\texttt{?n1 swrs:year ?n3.} \\\texttt{?n4 swrs:publication ?n1.} }& AIFB \\ 
+\hdashline
+\end{tabular}
+\vspace{2mm}
+\caption{Selected motifs. The frequency is the number of matches found in the set time limit. The last column indicates the dataset.}
+\label{table:selected}
+\end{table}
+
+The method provides many positives.
+To see that these are not just random noise, consider those patterns that have high frequency, but a negative log-factor.
+For instance, the most frequent pattern in the AIFB data describes entities having the same ``year'' property.
+Clearly, such a pattern can be matched often, and in many different ways, but it does not provide a satisfying explanation of the the structure of the graph.
 
@@
 Table~\ref{table:selected} shows some interesting motifs from the top 100 for each dataset.
-We see, for instance that e=the assertions that something is true or false are both motifs.
-The example from the AIFB date shows
+We see, for instance that the assertions that something is true or false are both motifs.
+While these are single triples with only one variable, they occur often enough, that encoding them separately provides a positive compression.
+The example from the AIFB date shows a typical "star" pattern likely to emerge from relational data: a single entity, surrounded by a set of attributes.
+
 \section{Discussion}
 
-We have presented a new method for inducing graph patterns from knowledge graphs.
-To our knowledge, this is the first method presented that can potentially find arbitrary graph patterns to describe the innate structure of a knowledge graph.
+We have presented a new method for mining graph patterns from knowledge graphs.
+To our knowledge, this is the first method presented that can potentially find arbitrary basic graph patterns to describe the innate structure of a knowledge graph.
 
-\subsection{Limitations and future work}
+\paragraph{Limitations and future work}
 
@@
 In \cite{bloem2017large}, the original method on which this method is based was shown to scale to graphs with billions of links.
-However, this scalability did not translate directly to knowledge graphs, since the sampling approach, used there is exteremely unlikely to generate the patterns that are succesful for motifs.
-If a better sampling method could be divised, we could likely scale the method up to web-scale knowledge graphs.
+However, this scalability does not translate directly to knowledge graphs, since the sampling approach used there is exteremely unlikely to generate the patterns that are succesful for knowledge graphs (requiring the use of simulated annealing istead).
+Since the motif test is just as scalable as it simple-graph equivalent, the only thing that is required to make the full motif-finding method scalable, is a faster search algorithm, perhaps based on an adaptation of the sampling approach from \cite{bloem2017large}.
 
 Our method currently produces a large number of motifs.
-We can show that worthwhile motifs are included, and that it performs better than a baseline like selecting by frequency, but it still takes some manual effort to sort through the suggestions to find the kind of motifs that fit a particular usecase.
+We can show that worthwhile motifs are included, and that it performs better than a frequency baseline, but it still takes some manual effort to sort through the suggestions to find the kind of motifs that fit a particular use case.
 This is not surprising; it is the nature of knowledge graphs that many different and overlapping substructures can be seen as natural or meaningful.
-One promising avenue to reduce this manual effort is to search of a \emph{set} of motifs which together compress well, each motif claiming a certain part of the knowledge graph to represent.
-If such an approach were fruitful it would truly represent a versatile and powerful form of schema induction for knowledge graphs.
+One promising avenue to reduce this manual effort is to search for a \emph{set} of motifs which together compress well, each motif claiming a certain part of the knowledge graph to represent.
 
-% \section{Acknowledgments}
+\paragraph{Acknowledgements}
 
-% This publication was supported by the Amsterdam Academic Alliance Data Science (AAA-DS) Program Award to the UvA and VU Universities.
+This publication was supported by the Amsterdam Academic Alliance Data Science (AAA-DS) Program Award to the UvA and VU Universities.
 
@@
 \bibliography{kgmotifs}
-\bibliographystyle{aaai}
+\bibliographystyle{style/splncs04}
 

### 2021-04-16 kgmotifs/supplement.tex
...

@@
 \usepackage{makecell}
-
-% PDF Info Is Required:
-\pdfinfo{
-/Title (Supplement: Detecting Motifs in Knowledge Graphs using Compression)
-/Author (...)}
 
@@
 \maketitle 
-\section{The canonical isomorphism algorithm for patterns}
-
-During search, we want to re-order the nodes and links in an encountered pattern to some canonical ordering.
+%\section{The canonical isomorphism algorithm for patterns}
+%
+%During search, we want to re-order the nodes and links in an encountered pattern to some canonical ordering.
 This way, when we later return an isomorphic pattern, we recognize it, and will not need to compute its logfactor again.
-The \emph{nauty} algorithm is the standard solution for findings such \emph{canonical isomorphs}.
+%The \emph{nauty} algorithm is the standard solution for findings such \emph{canonical isomorphs}.
 As we are sampling partially labeled, directed graphs, we have a lot of information that can speed up canonicalization.
@@
 Below, we provide a very succinct description of how nauty works, when applied to our patterns.
-
-Let $V'_M$ and $R'_M$ represent the labels of the variable nodes and links in $M$.
+%
+%Let $V'_M$ and $R'_M$ represent the labels of the variable nodes and links in $M$.
 Pattern $M$ is an isomorphism of pattern $M'$ if there exist bijections $\sigma_v : V'_M \times V'M$ and $\sigma_r: \R'_M \times \R'_M$ such that applying these to the variable elements of $M$ yields $M'$ (under graph equivalence).
@@
 A function $C(P)$ is a canonicalization function if it produces for all any pattern $M$ in $[M]$ the same, canonical member of $M$.
-
-Note that a very simple canonicalization algorithm would be to pass the unlabeled graph to the Nauty algorithm, and use its canonical ordering on the labeled graph.
+%
+%Note that a very simple canonicalization algorithm would be to pass the unlabeled graph to the Nauty algorithm, and use its canonical ordering on the labeled graph.
 However, the labels can make the canonicalization procedure significantly faster.
-
-The key principle behind the Nauty algorithm is the use of a \emph{colouring}: a partition of the nodes of a graph.
-A \emph{safe colouring} is defined as a partition $\pi$ of the nodes of the graph such that any permutation that maps nodes from different partitions to one another is guaranteed \emph{not} to be an automorphism.
-No guarantees are made about nodes in the same cell.
+%
+%The key principle behind the Nauty algorithm is the use of a \emph{colouring}: a partition of the nodes of a graph. 
+% A \emph{safe colouring} is defined as a partition $\pi$ of the nodes of the graph such that any permutation that maps nodes from different partitions to one another is guaranteed \emph{not} to be an automorphism.
+(Two nodes with the same node cannot be mapped to one another to create an automorphism.) No guarantees are made about nodes in the same cell.
 For instance, the partition that puts all nodes in the same cell is always safe.
@@
 Given one safe colouring $\pi$, if node $a$ from partition $i$ has a neighbor in partition $j$ and node $b$ from partition $j$ has no such node, we can \emph{refine} the partition $\pi$ by separating $i$ into a new partition (one contain $a$ and one containing $b$).
-
-The nauty algorithm operates on a graph with a given colouring (with the unit colouring used by default).
+%
+%The Nauty algorithm operates on a graph with a given colouring (with the unit colouring used by default).
 It first refines the colouring as much as possible, using degree information, and propagating the distinguishing features through the graph.
 Once the most refined colouring has been achieved, it searches through all autorphisms of the graph to produce a canonical example (using various optimizations to prune the search tree).
+%
+%To adapt Nauty to our situation we make two major modifications.
+Firstly, we have node labels as well as edge labels.
+We can permute both the node variables and the edge variables to create an automorphism of a given BGP.
+Secondly, we have much more information with which to create the basic coloring (which Nauty then refines).
+We can treat labeled nodes as node in a single cellin their partition.
+This ensures not only that they are not mapped to other nodes by any automorphism, but also that their information propagates to the neighboring nodes during refinement.
 
-In our case, we use the colouring to indicate what our labeling tells us about the colouring of the nodes.
-Note the following:
-\begin{easylist}
-§ Labeled nodes in the pattern can be placed in a discrete cell of the colouring.
-§ Variable nodes should be placed into a single cell in the partition.
-§ We define a partition over nodes \emph{and} relation labels. 
-\end{easylist}
-
-\begin{landscape}
-
-\section*{Motifs found for the three datasets}
-
+\tiny
 \subsection{Dogfood, top 100 by log factor}
 
-\begin{longtable}{ r r p{18cm} }
+\begin{longtable}{ r r p{10cm} }
  log factor & frequency & \\
 \hline\endhead
- 28232.1 & 5807 & \makecell{\texttt{?n1 swc:heldBy ?n2.} } \\ 
- 5125.5 & 2538 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
- 3997.0 & 2180 & \makecell{\texttt{?n1 swc:isPartOf ?n2.} } \\ 
- 1755.1 & 566 & \parbox{2cm}{\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_tzid ?n2.} } \\ 
- 378.1 & 16651 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 rdf:\_1 ?n3.} } \\ 
- 201.2 & 539 & \makecell{\texttt{?n1 swc:hasRole ?n2.} \\\texttt{?n2 rdf:type swc:Presenter.} } \\ 
- -2723.1 & 466 & \makecell{\texttt{?n1 rdf:type swc:Presenter.} } \\ 
- -2994.2 & 2031 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
- -3191.3 & 30 & \makecell{dsw:conference/eswc/2006/sessions/poster-session swc:hasRole ?n1. \\\texttt{?n1 rdf:type swc:Presenter.} } \\ 
- -3227.7 & 154 & \makecell{\texttt{dsw:conference/iswc/2008/research/pcmember swc:heldBy ?n1.} } \\ 
- -3271.9 & 156 & \makecell{\texttt{dsw:conference/iswc/2008/research/pcmember ?p2 ?n1.} } \\ 
- -3415.6 & 49 & \texttt{dsw:conference/dc/2010/programme-committee-member swc:heldBy ?n1.} \\ 
- -3514.0 & 138 & \makecell{\texttt{?n1 ?p2 dsw:person/eero-hyvoenen.} } \\ 
- -3596.0 & 29 & \makecell{\texttt{dsw:conference/iswc/2008/in\_use/pcmember swc:heldBy ?n1.} } \\ 
- -3605.8 & 33 & \makecell{\texttt{?n1 ?p2 dsw:person/heiko-paulheim.} } \\ 
- -3608.1 & 31 & \makecell{\texttt{dsw:conference/eswc/2012/poster-track-committee-member swc:heldBy ?n1.} } \\ 
- -3608.5 & 96 & \makecell{\texttt{\_:genid8 ?p2 ?n1.} } \\ 
- -3619.3 & 81 & \makecell{\texttt{\_:genid14 ?p2 ?n1.} } \\ 
- -3637.2 & 55 & \makecell{\texttt{\_:genid37 ?p2 ?n1.} } \\ 
- -3638.1 & 60 & \makecell{\texttt{\_:genid17 ?p2 ?n1.} } \\ 
- -3642.2 & 36 & \makecell{\texttt{dsw:conference/iswc/2010/dc/pcmember swc:heldBy ?n1.} } \\ 
- -3643.8 & 47 & \makecell{\texttt{?n1 ?p2 dsw:person/kim-viljanen.} } \\ 
- -3644.5 & 46 & \makecell{\texttt{?n1 ?p2 dsw:person/bernhard-schandl.} } \\ 
- -3656.0 & 1 & \makecell{\texttt{dsw:conference/eswc/2010/paper/services/3/authorlist rdf:\_1 dsw:person/heiko-paulheim.} } \\ 
- -3658.5 & 1 & \makecell{\texttt{\_:genid12 rdf:\_10 dsw:person/michael-pool.} } \\ 
- -3660.9 & 1 & \makecell{\texttt{\_:genid14 rdf:\_1 dsw:person/siegfried-handschuh.} } \\ 
- -3660.9 & 19 & \makecell{\texttt{?n1 ?p2 dsw:person/xi-bai.} } \\ 
- -3662.3 & 1 & \makecell{dsw:person/doroteo-t-toledano foaf:name "Doroteo T. Toledano". } \\ 
- -3662.8 & 34 & \makecell{\texttt{dsw:conference/eswc/2012/poster-track-committee-member ?p2 ?n1.} } \\ 
- -3662.9 & 1 & \makecell{\texttt{dsw:person/doroteo-t-toledano rdfs:label "Doroteo T. Toledano".} } \\ 
- -3663.6 & 1 & \makecell{\texttt{dsw:conference/www/2012/demo/demo-chair swc:heldBy dsw:person/bebo-white.} } \\ 
- -3664.9 & 1 & \makecell{\texttt{\_:genid17 rdf:\_1 dsw:person/john-barton.} } \\ 
- -3665.5 & 1 & \makecell{dsw:organization/czech-technical-university-in-prague foaf:member dsw:person/petr-pollak. } \\ 
- -3665.9 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/431/authorlist rdf:\_1 dsw:person/alexandre-passant.} } \\ 
- -3666.2 & 1 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 dc:creator dsw:person/violeta-damjanovic.} } \\ 
- -3666.4 & 1 & \makecell{\texttt{\_:genid37 rdf:\_2 dsw:person/stephen-larson.} } \\ 
- -3666.7 & 1 & \makecell{\texttt{\_:genid117 rdf:\_1 dsw:person/tomas-vitvar.} } \\ 
- -3666.8 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember rdf:type swc:OrganisingCommitteeMember.} } \\ 
- -3666.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/468/authorlist rdf:\_1 dsw:person/jerome-euzenat.} } \\ 
- -3666.8 & 1 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 dc:creator dsw:person/raluca-zaharia.} } \\ 
- -3666.8 & 1 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 dc:creator dsw:person/laurentiu-vasiliu.} } \\ 
- -3667.2 & 1 & \makecell{\texttt{dsw:conference/iswc/2008/paper/poster\_demo/43 foaf:maker dsw:person/malte-kiesel.} } \\ 
- -3667.2 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember swc:heldBy dsw:person/achille-fokoue.} } \\ 
- -3667.6 & 3 & \makecell{\texttt{dsw:conference/lrec/2008/papers/421 ?p1 dsw:person/petr-pollak.} } \\ 
- -3667.6 & 1 & \makecell{\texttt{dsw:conference/lrec/2008/papers/421 dc:creator dsw:person/petr-pollak.} } \\ 
- -3667.6 & 1 & \makecell{\texttt{dsw:workshop/iwod/2008/main/chair/organisingcommitteemember rdf:type swc:OrganisingCommitteeMember.} } \\ 
- -3667.7 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/poster-track-committee-member swc:heldBy dsw:person/tomas-vitvar.} } \\ 
- -3667.8 & 1 & \makecell{\texttt{\_:genid39 rdf:\_1 dsw:person/tomas-vitvar.} } \\ 
- -3667.9 & 3 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 ?p1 dsw:person/raluca-zaharia.} } \\ 
- -3667.9 & 3 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 ?p1 dsw:person/laurentiu-vasiliu.} } \\ 
- -3668.0 & 1 & \makecell{\texttt{dsw:workshop/swsc/2008/main/pcmember swc:heldBy dsw:person/tomas-vitvar.} } \\ 
- -3668.1 & 59 & \makecell{\texttt{?n1 ?p2 dsw:person/achille-fokoue.} } \\ 
- -3668.3 & 1 & \makecell{\texttt{\_:genid38 rdf:\_1 dsw:person/nicola-fanizzi.} } \\ 
- -3668.3 & 1 & \makecell{\texttt{dsw:conference/lrec/2008/papers/301 swrs:author dsw:person/hiroyuki-shinnou.} } \\ 
- -3669.0 & 1 & \makecell{\texttt{\_:genid37 rdf:\_2 dsw:person/hanmin-jung.} } \\ 
- -3670.3 & 1 & \makecell{\texttt{dsw:conference/iswc/2008/research/pcmember swc:heldBy dsw:person/chiara-ghidini.} } \\ 
- -3670.6 & 13 & \makecell{\texttt{?n1 rdf:type swc:OrganisingCommitteeMember.} } \\ 
- -3670.9 & 80 & \makecell{\texttt{\_:genid12 ?p2 ?n1.} } \\ 
- -3671.1 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/dc/pcmember swc:heldBy dsw:person/andreas-harth.} } \\ 
- -3671.2 & 18 & \makecell{\texttt{\_:genid17 rdf:\_1 ?n1.} } \\ 
- -3673.4 & 3 & \makecell{\texttt{dsw:conference/iswc/2008/paper/poster\_demo/43 ?p1 dsw:person/malte-kiesel.} } \\ 
- -3673.7 & 13 & \makecell{\texttt{?n1 ?p2 dsw:person/hiroyuki-shinnou.} } \\ 
- -3674.2 & 3 & \makecell{\texttt{dsw:conference/lrec/2008/papers/301 ?p1 dsw:person/hiroyuki-shinnou.} } \\ 
- -3674.7 & 69 & \makecell{\texttt{?n1 ?p2 dsw:person/andreas-harth.} } \\ 
- -3675.6 & 20 & \makecell{\texttt{?n1 ?p2 dsw:person/jianfeng-du.} } \\ 
- -3676.1 & 1 & \makecell{\texttt{dsw:person/akira-mori rdf:type foaf:Person.} } \\ 
- -3677.1 & 1 & \makecell{\texttt{dsw:person/alberto-musetti rdf:type foaf:Person.} } \\ 
- -3677.9 & 1 & \makecell{\texttt{dsw:person/alejandro-jaimes rdf:type foaf:Person.} } \\ 
- -3678.1 & 1 & \makecell{\texttt{dsw:conference/eswc/2010/paper/services/3/authorlist ?p1 dsw:person/heiko-paulheim.} } \\ 
- -3678.7 & 13 & \makecell{\texttt{\_:genid38 rdf:\_1 ?n1.} } \\ 
- -3678.8 & 1 & \makecell{\texttt{dsw:person/georg-lausen rdf:type foaf:Person.} } \\ 
- -3679.2 & 2 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 ?p1 dsw:person/violeta-damjanovic.} } \\ 
- -3679.5 & 10 & \makecell{\texttt{\_:genid8 rdf:\_4 ?n1.} } \\ 
- -3679.7 & 10 & \makecell{\texttt{\_:genid59 rdf:\_2 ?n1.} } \\ 
- -3680.2 & 2 & \makecell{\texttt{dsw:person/doroteo-t-toledano ?p1 "Doroteo T. Toledano".} } \\ 
- -3680.3 & 13 & \makecell{\texttt{\_:genid39 rdf:\_1 ?n1.} } \\ 
- -3680.5 & 1 & \makecell{\texttt{\_:genid12 ?p1 dsw:person/michael-pool.} } \\ 
- -3680.7 & 3 & \makecell{\texttt{dsw:conference/iswc/2009/paper/research/347 ?p1 dsw:person/bernhard-schandl.} } \\ 
- -3682.2 & 33 & \makecell{\texttt{\_:genid38 ?p2 ?n1.} } \\ 
- -3682.9 & 1 & \makecell{\texttt{\_:genid14 ?p1 dsw:person/siegfried-handschuh.} } \\ 
- -3683.8 & 13 & \makecell{\texttt{\_:genid37 rdf:\_2 ?n1.} } \\ 
- -3685.6 & 1 & \makecell{\texttt{dsw:conference/www/2012/demo/demo-chair ?p1 dsw:person/bebo-white.} } \\ 
- -3685.7 & 1 & \makecell{\texttt{\_:genid150 ?p1 dsw:person/bettina-bauer-messmer.} } \\ 
- -3686.3 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/semantic-web-in-use/60/authorlist ?p1 dsw:person/xi-bai.} } \\ 
- -3686.8 & 169 & \makecell{\texttt{\_:genid38 rdf:\_1 ?n2.} \\\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid38.} } \\ 
- -3687.0 & 43 & \makecell{\texttt{\_:genid42 ?p2 ?n1.} } \\ 
- -3687.1 & 1 & \makecell{\texttt{dsw:conference/eswc/2010/social\_web/pcmember ?p1 dsw:person/eero-hyvoenen.} } \\ 
- -3687.2 & 3 & \makecell{\texttt{dsw:conference/eswc/2008/paper/54 ?p1 dsw:person/eero-hyvoenen.} } \\ 
- -3687.5 & 1 & \makecell{\texttt{dsw:organization/czech-technical-university-in-prague ?p1 dsw:person/petr-pollak.} } \\ 
- -3688.1 & 1 & \makecell{\texttt{dsw:conference/rr/2008/full/4/authors ?p1 dsw:person/jianfeng-du.} } \\ 
- -3688.4 & 1 & \makecell{\texttt{\_:genid37 ?p1 dsw:person/stephen-larson.} } \\ 
- -3688.7 & 1 & \makecell{\texttt{\_:genid117 ?p1 dsw:person/tomas-vitvar.} } \\ 
- -3688.7 & 2 & \makecell{\texttt{\_:genid8 ?p1 dsw:person/irene-celino.} } \\ 
- -3688.8 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember ?p1 swc:OrganisingCommitteeMember.} } \\ 
- -3688.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/468/authorlist ?p1 dsw:person/jerome-euzenat.} } \\ 
- -3689.0 & 13 & \makecell{\texttt{\_:genid37 rdf:\_1 ?n1.} } \\ 
- -3689.1 & 1 & \makecell{\texttt{\_:genid36 ?p1 dsw:person/benjamin-coe.} } \\ 
- -3689.2 & 2 & \makecell{\texttt{dsw:conference/eswc/2011/paper/in-use/52 ?p1 dsw:person/eero-hyvoenen.} } \\ 
- -3689.3 & 1 & \makecell{\texttt{dsw:conference/eswc/2006/demo-papers/FD01-hitzler ?p1 dsw:person/malte-kiesel.} } \\ 
- -3689.3 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember ?p1 dsw:person/achille-fokoue.} } \\ 
- -3689.7 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/poster-track-committee-member ?p1 dsw:person/tomas-vitvar.} } \\ 
- -3689.8 & 8 & \makecell{\texttt{\_:genid96 rdf:\_1 ?n1.} } \\ 
- -3690.1 & 1 & \makecell{\texttt{dsw:workshop/swsc/2008/main/pcmember ?p1 dsw:person/tomas-vitvar.} } \\ 
- -3690.3 & 1 & \makecell{\texttt{\_:genid38 ?p1 dsw:person/nicola-fanizzi.} } \\ 
- -3691.0 & 1 & \makecell{\texttt{\_:genid37 ?p1 dsw:person/hanmin-jung.} } \\ 
- -3692.3 & 3 & \makecell{\texttt{?n1 swc:heldBy dsw:person/bebo-white.} } \\ 
- -3692.9 & 47 & \makecell{\texttt{\_:genid36 ?p2 ?n1.} } \\ 
- -3693.1 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/dc/pcmember ?p1 dsw:person/andreas-harth.} } \\ 
- -3694.6 & 1 & \makecell{\texttt{?n1 rdf:\_10 dsw:person/michael-pool.} } \\ 
- -3694.6 & 1 & \makecell{\texttt{\_:genid12 rdf:\_10 ?n1.} } \\ 
- -3695.0 & 2 & \makecell{\texttt{?n1 swc:heldBy dsw:person/marco-neumann.} } \\ 
- -3695.4 & 13 & \makecell{\texttt{\_:genid38 rdf:\_1 ?n1.} \\\texttt{dsw:conference/iswc/2008/paper/poster\_demo/18 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid38.} } \\ 
- -3695.9 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/hiroyuki-shinnou.} } \\ 
- -3696.4 & 39 & \makecell{\texttt{\_:genid39 ?p2 ?n1.} } \\ 
- -3696.7 & 6 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/45 foaf:maker ?n1.} } \\ 
- -3697.1 & 1 & \makecell{\texttt{dsw:conference/www/2012/dev/16/talk swc:hasRole ?n1.} \\\texttt{?n1 rdf:type swc:Presenter.} } \\ 
- -3697.1 & 2 & \makecell{\texttt{dsw:person/doroteo-t-toledano foaf:made ?n1.} } \\ 
- -3697.6 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/bruno-alves.} } \\ 
- -3698.0 & 1 & \makecell{\texttt{dsw:conference/dc/2010/tutorial/2/tutorial-presenter rdf:type swc:Presenter.} \\\texttt{?n1 swc:hasRole dsw:conference/dc/2010/tutorial/2/tutorial-presenter.} } \\ 
- -3698.1 & 1 & \makecell{\texttt{dsw:person/akira-mori ?p1 foaf:Person.} } \\ 
- -3698.5 & 1 & \makecell{\texttt{?n1 foaf:name "Doroteo T. Toledano".} } \\ 
- -3698.8 & 1 & \makecell{\texttt{dsw:person/michalis-foukarakis ?p1 foaf:Person.} } \\ 
- -3699.0 & 1 & \makecell{\texttt{?n1 rdfs:label "Doroteo T. Toledano".} } \\ 
- -3699.0 & 1 & \makecell{\texttt{dsw:person/doroteo-t-toledano rdfs:label ?n1.} } \\ 
- -3699.2 & 1 & \makecell{\texttt{dsw:person/alberto-musetti ?p1 foaf:Person.} } \\ 
- -3699.2 & 2 & \makecell{\texttt{\_:genid152 rdf:\_1 ?n1.} } \\ 
- -3699.3 & 1 & \makecell{\texttt{dsw:person/michalis-foukarakis foaf:name ?n1.} } \\ 
- -3699.7 & 1 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 rdfs:label ?n1.} } \\ 
- -3699.8 & 1 & \makecell{\texttt{dsw:person/michalis-foukarakis rdfs:label ?n1.} } \\ 
- -3699.9 & 1 & \makecell{\texttt{dsw:person/alejandro-jaimes ?p1 foaf:Person.} } \\ 
- -3699.9 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/paper/in-use-industrial/138/authorlist rdf:\_1 ?n1.} } \\ 
- -3700.1 & 1 & \makecell{\texttt{dsw:person/alain-leger ?p1 foaf:Person.} } \\ 
- -3700.2 & 4 & \makecell{\texttt{\_:genid117 rdf:\_1 ?n1.} } \\ 
- -3700.4 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/nikos-loutas.} } \\ 
- -3700.5 & 4 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember swc:heldBy ?n1.} } \\ 
- -3700.8 & 1 & \makecell{\texttt{dsw:person/georg-lausen ?p1 foaf:Person.} } \\ 
- -3700.9 & 1 & \makecell{\texttt{?n1 rdfs:label "Graph Kernels for RDF data".} } \\ 
- -3700.9 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/petr-pollak.} } \\ 
- -3701.1 & 2 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/aidan-boran.} } \\ 
- -3701.8 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-633 rdfs:label ?n1.} } \\ 
- -3702.1 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/431/authorlist rdf:\_1 ?n1.} } \\ 
- -3702.1 & 2 & \makecell{\texttt{?n1 foaf:maker dsw:person/bettina-bauer-messmer.} } \\ 
- -3702.1 & 2 & \makecell{\texttt{dsw:conference/www/2012/eu/10 swc:isPartOf ?n1.} } \\ 
- -3702.3 & 1 & \makecell{\texttt{?n1 dc:creator dsw:person/violeta-damjanovic.} } \\ 
- -3702.3 & 1 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 dc:creator ?n1.} } \\ 
- -3702.5 & 37 & \makecell{\texttt{?n1 ?p2 dsw:person/chiara-ghidini.} } \\ 
- -3702.5 & 1 & \makecell{\texttt{?n1 rdf:\_2 dsw:person/stephen-larson.} } \\ 
- -3702.8 & 2 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/siegfried-handschuh.} } \\ 
- -3702.9 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember rdf:type ?n1.} } \\ 
- -3702.9 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/468/authorlist rdf:\_1 ?n1.} } \\ 
- -3703.0 & 1 & \makecell{\texttt{?n1 dc:creator dsw:person/raluca-zaharia.} } \\ 
- -3703.4 & 4 & \makecell{\texttt{?n1 foaf:maker dsw:person/malte-kiesel.} } \\ 
- -3703.5 & 1 & \makecell{\texttt{?n1 dc:creator dsw:person/naimul-basher.} } \\ 
- -3703.6 & 37 & \makecell{\texttt{\_:genid94 ?p2 ?n1.} } \\ 
- -3703.8 & 4 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 dc:creator ?n1.} } \\ 
- -3704.0 & 1 & \makecell{\texttt{?n1 dc:creator dsw:person/vinhtuan-thai.} } \\ 
- -3704.3 & 2 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/jerome-euzenat.} } \\ 
- -3704.7 & 39 & \makecell{\texttt{\_:genid62 ?p2 ?n1.} } \\ 
- -3705.4 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/tomas-vitvar.} } \\ 
- -3705.7 & 3 & \makecell{\texttt{?n1 swc:heldBy dsw:person/tomas-vitvar.} } \\ 
- -3706.6 & 2 & \makecell{\texttt{?n1 foaf:maker dsw:person/aidan-boran.} } \\ 
- -3706.9 & 3 & \makecell{\texttt{?n1 swrs:author dsw:person/hiroyuki-shinnou.} } \\ 
- -3708.8 & 2 & \makecell{\texttt{\_:genid65 http://www.w3.org/2002/12/cal/ical\_tzid ?n1.} } \\ 
- -3709.0 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-766/authors rdf:type ?n1.} } \\ 
- -3710.4 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/achille-fokoue.} } \\ 
- -3710.8 & 2 & \makecell{\texttt{?n1 swc:heldBy dsw:person/aidan-boran.} } \\ 
- -3711.7 & 6 & \makecell{\texttt{?n1 ?p2 dsw:person/bruno-alves.} } \\ 
- -3714.2 & 1 & \makecell{\texttt{dsw:conference/eswc/2010/paper/services/3/authorlist ?p2 ?n1.} } \\ 
- -3714.3 & 5 & \makecell{\texttt{?n1 swc:heldBy dsw:person/chiara-ghidini.} } \\ 
- -3714.9 & 1 & \makecell{\texttt{dsw:person/georg-lausen rdf:type ?n1.} } \\ 
- -3715.9 & 12 & \makecell{\texttt{?n1 rdf:\_3 dsw:person/eero-hyvoenen.} } \\ 
- -3716.7 & 5 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/andreas-harth.} } \\ 
- -3717.0 & 9 & \makecell{\texttt{dsw:person/georg-lausen swc:holdsRole ?n1.} } \\ 
- -3717.4 & 13 & \makecell{\texttt{\_:genid38 rdf:\_1 ?n1.} \\\texttt{dsw:conference/iswc/2008/paper/poster\_demo/18 ?p2 \_:genid38.} } \\ 
- -3718.4 & 8 & \makecell{\texttt{?n1 swc:heldBy dsw:person/achille-fokoue.} } \\ 
- -3719.7 & 37 & \makecell{\texttt{\_:genid59 ?p2 ?n1.} } \\ 
- -3719.9 & 13 & \makecell{\texttt{?n1 ?p2 dsw:person/aidan-boran.} } \\ 
- -3720.5 & 20 & \makecell{\texttt{?n1 ?p2 dsw:person/tomas-vitvar.} } \\ 
- -3721.7 & 4 & \makecell{\texttt{dsw:conference/eswc/2012/paper/in-use-industrial/138/authorlist ?p2 ?n1.} } \\ 
- -3723.0 & 9 & \makecell{\texttt{?n1 dc:creator dsw:person/andreas-harth.} } \\ 
- -3724.0 & 12 & \makecell{\texttt{?n1 swc:heldBy dsw:person/bernhard-schandl.} } \\ 
- -3726.0 & 58 & \makecell{\texttt{dsw:conference/iswc/2009/paper/inuse/143 dc:creator dsw:person/achille-fokoue.} \\\texttt{?n1 ?p2 dsw:person/achille-fokoue.} } \\ 
- -3726.2 & 26 & \makecell{\texttt{?n1 swrs:author dsw:person/eero-hyvoenen.} } \\ 
- -3726.4 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/petr-pollak.} } \\ 
- -3727.7 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/benjamin-coe.} } \\ 
- -3728.1 & 24 & \makecell{\texttt{?n1 swc:heldBy dsw:person/mathieu-daquin.} } \\ 
- -3728.9 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/naimul-basher.} } \\ 
- -3729.6 & 0 & \makecell{\texttt{?n1 ?p3 ?n2.} \\\texttt{?n2 rdf:type swc:Presenter.} } \\ 
- -3730.0 & 2 & \makecell{\texttt{dsw:conference/iswc/2010/paper/445/authorlist ?p2 ?n1.} } \\ 
- -3731.5 & 30 & \makecell{\texttt{\_:genid96 ?p2 ?n1.} } \\ 
- -3731.5 & 4 & \makecell{\texttt{?n1 ?p2 dsw:person/michael-pool.} } \\ 
- -3731.7 & 2 & \makecell{\texttt{dsw:conference/iswc/2010/paper/431/authorlist ?p2 ?n1.} } \\ 
- -3731.9 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/raluca-zaharia.} } \\ 
- -3732.4 & 13 & \makecell{\texttt{dsw:conference/iswc/2008/paper/poster\_demo/18 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n1.} \\\texttt{?n1 rdf:\_1 ?n2.} } \\ 
- -3732.6 & 6 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember ?p2 ?n1.} } \\ 
- -3733.0 & 3 & \makecell{\texttt{?n2 rdfs:label "Graph Kernels for RDF data".} \\\texttt{?n2 dc:creator ?n1.} } \\ 
- -3735.3 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/julien-gaugaz.} } \\ 
- -3738.0 & 45 & \makecell{\texttt{http://www.eswc2007.org/pdf/eswc07-dellavalle.pdf rdf:type ?n2.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
- -3738.4 & 2 & \makecell{\texttt{?n1 ?p2 dsw:person/sukwon-chung.} } \\ 
- -3739.9 & 5 & \makecell{\texttt{\_:bNode14 ?p2 ?n1.} } \\ 
- -3740.6 & 7 & \makecell{\texttt{?n1 ?p2 dsw:person/bettina-bauer-messmer.} } \\ 
- -3741.2 & 5 & \makecell{\texttt{dsw:conference/iswc/2010/paper/489/authorlist ?p2 ?n1.} } \\ 
- -3741.4 & 6 & \makecell{\texttt{dsw:conference/iswc/2010/paper/261/authorlist ?p2 ?n1.} } \\ 
- -3742.5 & 36 & \makecell{\texttt{\_:genid99 ?p2 ?n1.} } \\ 
- -3743.8 & 16 & \makecell{\texttt{\_:genid121 ?p2 ?n1.} } \\ 
- -3744.2 & 0 & \makecell{\texttt{?n1 ?p4 ?n2.} \\\texttt{?n2 rdf:\_1 ?n3.} } \\ 
- -3746.9 & 4 & \makecell{\texttt{?n1 ?p2 dsw:person/stephen-larson.} } \\ 
- -3749.0 & 7 & \makecell{\texttt{dsw:conference/www/2012/demo/demo-chair ?p2 ?n1.} } \\ 
- -3749.7 & 20 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-633 ?p2 ?n1.} } \\ 
- -3750.6 & 13 & \makecell{\texttt{\_:genid152 ?p2 ?n1.} } \\ 
- -3751.7 & 29 & \makecell{\texttt{\_:genid65 ?p2 ?n1.} } \\ 
- -3752.6 & 12 & \makecell{\texttt{?n1 ?p2 dsw:person/vinhtuan-thai.} } \\ 
- -3753.1 & 6 & \makecell{\texttt{dsw:person/doroteo-t-toledano ?p2 ?n1.} } \\ 
- -3755.1 & 59 & \makecell{\texttt{dsw:conference/iswc-aswc/2007/tracks/posters-demos/papers/375 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
- -3755.4 & 9 & \makecell{\texttt{?n1 ?p2 dsw:person/violeta-damjanovic.} } \\ 
- -3759.6 & 34 & \makecell{\texttt{\_:genid69 ?p2 ?n1.} } \\ 
- -3761.7 & 9 & \makecell{\texttt{dsw:person/alberto-musetti ?p2 ?n1.} } \\ 
- -3763.9 & 7 & \makecell{\texttt{dsw:person/michalis-foukarakis ?p2 ?n1.} } \\ 
- -3764.3 & 17 & \makecell{\texttt{\_:genid150 ?p2 ?n1.} } \\ 
- -3769.0 & 15 & \makecell{\texttt{?n1 ?p2 dsw:person/malte-kiesel.} } \\ 
- -3770.1 & 10 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 ?p2 ?n1.} } \\ 
- -3772.9 & 45 & \makecell{\texttt{?n1 ?p2 dsw:person/elena-paslaru-bontas-simperl.} } \\ 
- -3780.1 & 18 & \makecell{\texttt{dsw:conference/iswc/2008/paper/poster\_demo/43 ?p2 ?n1.} } \\ 
- -3780.2 & 18 & \makecell{\texttt{dsw:conference/eswc/2006/demo-papers/FD01-hitzler ?p2 ?n1.} } \\ 
- -3784.0 & 1290 & \makecell{\texttt{?n1 rdf:type ?n3.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
- -3797.8 & 12 & \makecell{\texttt{dsw:person/alejandro-jaimes ?p2 ?n1.} } \\ 
- -3800.4 & 26 & \makecell{\texttt{dsw:person/georg-lausen ?p2 ?n1.} } \\ 
- -3833.3 & 233 & \makecell{\texttt{?n1 ?p5 ?n3.} \\\texttt{?n3 ?p4 ?n2.} } \\ 
- -3902.5 & 47 & \makecell{\texttt{?n1 rdf:\_10 ?n2.} } \\ 
- -3993.4 & 405 & \makecell{\texttt{?n1 ?p4 ?n2.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
- -4021.2 & 2102 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
- -4106.8 & 156 & \makecell{\texttt{?n1 rdfs:label ?n2.} } \\ 
- -4204.5 & 373 & \makecell{\texttt{?n1 ?p4 ?n3.} \\\texttt{?n2 ?p4 ?n3.} } \\ 
- -4750.2 & 2914 & \makecell{\texttt{?n1 rdf:\_1 ?n2.} } \\ 
- -4761.6 & 248 & \makecell{\texttt{?n1 ?p2 foaf:Person.} } \\ 
- -4802.7 & 293 & \makecell{\texttt{?n1 rdf:type foaf:Person.} } \\ 
- -5632.4 & 301 & \makecell{\texttt{?n1 foaf:name ?n2.} } \\ 
- -6987.6 & 3047 & \makecell{\texttt{?n1 rdf:\_2 ?n2.} } \\ 
- -9429.5 & 2385 & \makecell{\texttt{?n1 foaf:maker ?n2.} } \\ 
- -10493.1 & 2528 & \makecell{\texttt{?n1 dc:creator ?n2.} } \\ 
+\input{results/dogfood/motifs-byscore.latex}
  \hline
 \end{longtable}
+\subsection{Dogfood, top 100 by frequency}
 
-\subsection{Dogfood, top 500 by frequency}
-
-\begin{longtable}{ r r p{19cm} }
+\begin{longtable}{ r r p{10cm} }
  log factor & frequency & \\
 \hline\endhead
- -5052.5 & 16582 & \makecell{\texttt{?n1 rdf:\_1 ?n3.} \\\texttt{?n1 rdf:\_1 ?n2.} } \\ 
- -4021.9 & 6804 & \makecell{\texttt{\_:genid9 ?p4 ?n1.} \\\texttt{\_:genid9 ?p3 ?n2.} } \\ 
- 28232.1 & 5807 & \makecell{\texttt{?n1 swc:heldBy ?n2.} } \\ 
- 122602.4 & 5709 & \makecell{\texttt{?n1 swc:heldBy ?n2.} \\\texttt{?n2 swc:holdsRole ?n1.} } \\ 
- -3801.0 & 3734 & \makecell{\texttt{?n1 ?p3 dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{?n2 ?p3 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
- -4178.4 & 3157 & \makecell{\texttt{?n2 ?p4 ?n1.} \\\texttt{?n2 rdf:\_4 ?n3.} } \\ 
- -7299.7 & 2799 & \makecell{\texttt{?n1 rdf:\_2 ?n2.} } \\ 
- -4442.9 & 2449 & \makecell{\texttt{?n1 rdf:\_1 ?n2.} } \\ 
- -4599.6 & 2393 & \makecell{\texttt{?n1 rdf:\_1 ?n4.} \\\texttt{?n2 rdf:\_1 ?n4.} \\\texttt{?n2 rdf:\_1 ?n3.} } \\ 
- -9551.9 & 2193 & \makecell{\texttt{?n1 foaf:maker ?n2.} } \\ 
- 6218.0 & 1939 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
- -3804.8 & 1938 & \makecell{\texttt{dsw:conference/iswc-aswc/2007/chair\_12 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n2.} } \\ 
- -1305.2 & 1713 & \makecell{\texttt{?n1 http://purl.org/dc/terms/creator ?n2.} } \\ 
- -15966.3 & 1672 & \makecell{\texttt{?n1 http://purl.org/ontology/bibo/authorList ?n2.} } \\ 
- 3849.2 & 1663 & \makecell{\texttt{?n1 swc:holdsRole ?n2.} } \\ 
- -1446.4 & 1587 & \makecell{\texttt{?n1 rdf:type rdf:Seq.} } \\ 
- 271.9 & 1587 & \makecell{\texttt{?n1 ?p2 rdf:Seq.} } \\ 
- -2977.3 & 1541 & \makecell{\texttt{?n2 rdf:\_3 ?n3.} \\\texttt{?n2 http://www.w3.org/2002/12/cal/ical\_date ?n1.} } \\ 
- -197.2 & 1466 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} } \\ 
- -1595.5 & 1463 & \makecell{\texttt{?n2 rdf:type ?n3.} \\\texttt{?n2 rdf:\_4 ?n1.} } \\ 
- -3974.4 & 1292 & \makecell{\texttt{?n1 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n2.} } \\ 
- -3829.0 & 1216 & \makecell{\texttt{\_:genid9 ?p3 ?n2.} \\\texttt{\_:genid9 ?p3 ?n1.} } \\ 
- -3828.3 & 1209 & \makecell{\texttt{dsw:conference/iswc-aswc/2007/chair\_12 ?p5 ?n3.} \\\texttt{?n1 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n2.} } \\ 
- -3082.3 & 1044 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
- -4012.3 & 670 & \makecell{\texttt{?n1 ?p5 ?n4.} \\\texttt{?n2 ?p5 ?n4.} \\\texttt{?n2 ?p5 ?n3.} } \\ 
- -4179.5 & 613 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 rdf:\_7 ?n3.} } \\ 
- -3790.2 & 479 & \makecell{\texttt{?n1 ?p3 ?n2.} \\\texttt{?n1 ?p3 dsw:person/jason-jung.} } \\ 
- -3067.7 & 421 & \makecell{\texttt{?n1 rdf:type ?n3.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
- -3808.7 & 398 & \makecell{\texttt{dsw:person/rayid-ghani ?p3 ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
- -1939.1 & 396 & \makecell{\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date ?n2.} } \\ 
- -3745.9 & 359 & \makecell{\texttt{dsw:person/rayid-ghani swc:holdsRole ?n2.} \\\texttt{?n1 swc:holdsRole ?n2.} } \\ 
- -2779.2 & 335 & \makecell{\texttt{?n1 ?p2 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
- -2727.6 & 334 & \makecell{\texttt{?n1 swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\ 
- 1626.1 & 334 & \makecell{\texttt{dsw:conference/eswc/2011/programme-committee-member swc:heldBy ?n1.} \\\texttt{?n1 swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\ 
- -3707.5 & 333 & \makecell{\texttt{dsw:person/rayid-ghani swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{?n1 swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\
- -3729.5 & 333 & \makecell{\texttt{dsw:person/rayid-ghani ?p2 dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{?n1 ?p2 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
- -3171.2 & 309 & \makecell{\texttt{?n1 rdf:type geo:Point.} } \\ 
- -2738.5 & 277 & \makecell{\texttt{dsw:conference/eswc/2012/research-track-committee-member ?p2 ?n1.} } \\ 
- -2678.0 & 274 & \makecell{\texttt{dsw:conference/eswc/2012/research-track-committee-member swc:heldBy ?n1.} } \\ 
- -3887.4 & 271 & \makecell{\texttt{\_:genid95 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n2.} } \\ 
- -3754.4 & 164 & \makecell{\texttt{\_:genid9 ?p3 ?n1.} \\\texttt{\_:genid9 http://www.w3.org/2002/12/cal/ical\_date ?n2.} } \\ 
- -3817.4 & 147 & \makecell{\texttt{\_:genid95 rdf:type ?n3.} \\\texttt{?n1 rdf:type ?n3.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
- -3708.6 & 126 & \makecell{\texttt{?n2 ?p3 ?n1.} \\\texttt{?n2 http://www.w3.org/2002/12/cal/ical\_date "2009-05-31T18:00:00".} } \\ 
- -3550.9 & 97 & \makecell{\texttt{\_:genid10 ?p2 ?n1.} } \\ 
- -4294.2 & 95 & \makecell{\texttt{?n1 ?p2 foaf:Person.} } \\ 
- -3625.0 & 89 & \makecell{\texttt{\_:genid5 ?p2 ?n1.} } \\ 
- -3505.4 & 89 & \makecell{\texttt{dsw:organization/yahoo-inc ?p2 ?n1.} } \\ 
- -3698.7 & 86 & \makecell{\texttt{\_:genid6 ?p2 ?n1.} } \\ 
- -3633.7 & 86 & \makecell{\texttt{\_:genid11 ?p2 ?n1.} } \\ 
- -3757.9 & 82 & \makecell{\texttt{?n2 ?p3 ?n1.} \\\texttt{?n2 rdf:\_2 dsw:person/jan-hettenhausen.} } \\ 
- -3468.0 & 73 & \makecell{\texttt{dsw:organization/yahoo-inc foaf:member ?n1.} } \\ 
- -3815.9 & 72 & \makecell{\texttt{\_:genid11 rdf:\_1 ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
- -3755.1 & 70 & \makecell{\texttt{dsw:workshop/om/2008/paper/main/21 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
- -3596.9 & 67 & \makecell{\texttt{?n1 ?p2 dsw:person/haofen-wang.} } \\ 
- -3660.9 & 67 & \makecell{\texttt{\_:genid18 ?p2 ?n1.} } \\ 
- -3721.2 & 66 & \makecell{\texttt{\_:genid18 ?p2 ?n1.} \\\texttt{\_:genid18 rdf:\_1 dsw:person/gary-wills.} } \\ 
- -3576.6 & 65 & \makecell{\texttt{?n1 ?p2 dsw:person/deborah-mcguinness.} } \\ 
- -3719.8 & 65 & \makecell{\texttt{\_:genid1 ?p2 ?n1.} } \\ 
- -3970.0 & 64 & \makecell{\texttt{?n1 rdf:\_9 ?n2.} } \\ 
- -3695.6 & 59 & \makecell{\texttt{\_:genid27 ?p2 ?n1.} } \\ 
- -3696.5 & 59 & \makecell{\texttt{\_:genid50 ?p2 ?n1.} } \\ 
- -3688.2 & 52 & \makecell{\texttt{\_:genid24 ?p2 ?n1.} } \\ 
- -3740.6 & 52 & \makecell{\texttt{?n1 ?p2 dsw:person/denny-vrandecic.} } \\ 
- -3737.8 & 51 & \makecell{\texttt{\_:genid11 rdf:\_1 ?n2.} \\\texttt{?n1 swrs:author ?n2.} } \\ 
- -3679.5 & 49 & \makecell{\texttt{\_:genid23 ?p2 ?n1.} } \\ 
- -3706.1 & 48 & \makecell{\texttt{\_:genid48 ?p2 ?n1.} } \\ 
- -3716.6 & 48 & \makecell{\texttt{\_:genid48 ?p2 ?n1.} \\\texttt{dsw:conference/www/2008/paper/431 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid48.} } \\ 
- -3661.4 & 47 & \makecell{\texttt{?n1 ?p2 dsw:person/maria-esther-vidal.} } \\ 
- -3681.1 & 47 & \makecell{\texttt{\_:genid46 ?p2 ?n1.} } \\ 
- -3692.9 & 47 & \makecell{\texttt{\_:genid36 ?p2 ?n1.} } \\ 
- -3609.9 & 43 & \makecell{\texttt{dsw:workshop/swcs/2012/programme-committee-member ?p2 ?n1.} } \\ 
- -3729.1 & 43 & \makecell{\texttt{\_:genid56 ?p2 ?n1.} } \\ 
- -3677.4 & 41 & \makecell{\texttt{\_:genid45 ?p2 ?n1.} } \\ 
- -3564.6 & 40 & \makecell{\texttt{dsw:workshop/swcs/2012/programme-committee-member swc:heldBy ?n1.} } \\ 
- -3611.2 & 39 & \makecell{\texttt{?n1 ?p2 dsw:person/gong-cheng.} } \\ 
- -3698.4 & 39 & \makecell{\texttt{?n1 ?p2 dsw:person/matthew-rowe.} } \\ 
- -3743.1 & 38 & \makecell{\texttt{?n1 ?p2 dsw:person/giorgos-stoilos.} } \\ 
- -3725.6 & 37 & \makecell{\texttt{\_:genid82 ?p2 ?n1.} } \\ 
- -3597.5 & 37 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26 ?p2 ?n1.} } \\ 
- -3716.3 & 36 & \makecell{\texttt{?n1 ?p2 dsw:person/katharina-siorpaes.} } \\ 
- -3758.7 & 36 & \makecell{\texttt{\_:genid66 ?p2 ?n1.} } \\ 
- -3730.4 & 36 & \makecell{\texttt{\_:genid78 ?p2 ?n1.} } \\ 
- -3649.8 & 36 & \makecell{\texttt{?n1 ?p2 dsw:person/vanja-josifovski.} } \\ 
- -3759.6 & 34 & \makecell{\texttt{\_:genid69 ?p2 ?n1.} } \\ 
- -3731.8 & 33 & \makecell{\texttt{\_:genid72 ?p2 ?n1.} } \\ 
- -3693.2 & 32 & \makecell{\texttt{dsw:workshop/sdow/2011/pcmember ?p2 ?n1.} } \\ 
- -3745.8 & 31 & \makecell{\texttt{\_:genid90 ?p2 ?n1.} } \\ 
- -3732.2 & 30 & \makecell{\texttt{\_:genid9 rdf:\_3 ?n2.} \\\texttt{\_:genid9 http://www.w3.org/2002/12/cal/ical\_date ?n1.} } \\ 
- -3730.5 & 30 & \makecell{\texttt{?n1 rdf:\_1 ?n2.} \\\texttt{?n1 rdf:\_1 dsw:person/jason-jung.} } \\ 
- -3751.7 & 29 & \makecell{\texttt{\_:genid65 ?p2 ?n1.} } \\ 
- -3737.3 & 26 & \makecell{\texttt{?n1 ?p2 dsw:person/karl-aberer.} } \\ 
- -3665.6 & 24 & \makecell{\texttt{\_:genid5 rdf:\_1 ?n1.} } \\ 
- -3737.1 & 23 & \makecell{\texttt{?n1 ?p2 dsw:person/gang-wang.} } \\ 
- -3703.3 & 21 & \makecell{\texttt{?n1 rdf:\_3 ?n2.} \\\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date "2009-05-31T18:00:00".} } \\ 
- -3728.1 & 21 & \makecell{\texttt{?n1 ?p3 "2009-05-31T18:00:00".} \\\texttt{?n1 rdf:\_3 ?n2.} } \\ 
- -3769.4 & 20 & \makecell{\texttt{\_:genid135 ?p2 ?n1.} } \\ 
- -3670.9 & 19 & \makecell{\texttt{\_:genid15 rdf:\_1 ?n1.} } \\ 
- -3697.6 & 16 & \makecell{\texttt{\_:genid18 rdf:\_2 ?n1.} } \\ 
- -3733.0 & 15 & \makecell{\texttt{?n1 foaf:maker dsw:person/deborah-mcguinness.} } \\ 
- -3716.3 & 15 & \makecell{\texttt{?n1 ?p2 dsw:person/rolf-gruetter.} } \\ 
- -3787.8 & 14 & \makecell{\texttt{dsw:person/adrian-marte ?p2 ?n1.} } \\ 
- -3788.7 & 14 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 ?p2 ?n1.} } \\ 
- -3722.1 & 14 & \makecell{\texttt{dsw:conference/iswc/2011/paper/poster-demo/40 ?p2 ?n1.} } \\ 
- -3674.6 & 14 & \makecell{\texttt{\_:genid23 rdf:\_2 ?n1.} } \\ 
- -3729.2 & 13 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p2 ?n1.} } \\ 
- -3691.8 & 13 & \makecell{\texttt{\_:genid48 rdf:\_1 ?n1.} } \\ 
- -3673.6 & 13 & \makecell{\texttt{?n1 ?p2 \_:genid48.} } \\ 
- -3646.4 & 13 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid48.} } \\ 
- -3690.6 & 13 & \makecell{\texttt{\_:genid27 rdf:\_3 ?n1.} } \\ 
- -3717.4 & 12 & \makecell{\texttt{?n1 ?p2 dsw:person/samur-araujo.} } \\ 
- -3736.6 & 11 & \makecell{\texttt{?n1 ?p2 dsw:person/kai-eckert.} } \\ 
- -3703.9 & 11 & \makecell{\texttt{?n1 ?p2 dsw:person/james-michaelis.} } \\ 
- -3717.9 & 10 & \makecell{\texttt{?n1 ?p2 dsw:person/varish-mulwad.} } \\ 
- -3786.8 & 9 & \makecell{\texttt{dsw:person/kenneth-church ?p2 ?n1.} } \\ 
- -3704.2 & 9 & \makecell{\texttt{?n1 ?p2 dsw:person/jianqiang-li.} } \\ 
- -3714.2 & 9 & \makecell{\texttt{?n1 swc:heldBy dsw:person/matthew-rowe.} } \\ 
- -3764.5 & 8 & \makecell{\texttt{dsw:person/nematollaah-shiri ?p2 ?n1.} } \\ 
- -3704.0 & 8 & \makecell{\texttt{?n1 http://purl.org/dc/terms/creator dsw:person/fabio-ciravegna.} } \\ 
- -3672.5 & 8 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid72.} } \\ 
- -3694.1 & 8 & \makecell{\texttt{\_:genid78 rdf:\_2 ?n1.} } \\ 
- -3703.7 & 7 & \makecell{\texttt{dsw:organization/yahoo-inc rdfs:label ?n1.} } \\ 
- -3699.9 & 7 & \makecell{\texttt{dsw:organization/yahoo-inc foaf:name ?n1.} } \\ 
- -3744.7 & 6 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors ?p2 ?n1.} } \\ 
- -3705.6 & 6 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/willem-van-hage.} } \\ 
- -3742.8 & 5 & \makecell{\texttt{dsw:conference/iswc/2010/paper/515/authorlist ?p2 ?n1.} } \\ 
- -3726.0 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/lisa-lena-opas-haenninen.} } \\ 
- -3724.1 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/yi-chin-tu.} } \\ 
- -3729.5 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/yon-dourisboure.} } \\ 
- -3726.8 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/kukulenz-dirk.} } \\ 
- -3729.2 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/jan-zemanek.} } \\ 
- -3745.6 & 5 & \makecell{\texttt{\_:b5 ?p2 ?n1.} } \\ 
- -3694.8 & 5 & \makecell{\texttt{\_:genid72 rdf:\_3 ?n1.} } \\ 
- -3726.4 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/brett-cannon.} } \\ 
- -3711.8 & 5 & \makecell{\texttt{\_:genid1 rdf:\_1 ?n1.} } \\ 
- -3680.9 & 4 & \makecell{\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date "2010-05-31T13:00:00".} } \\ 
- -3705.2 & 4 & \makecell{\texttt{?n1 ?p2 "2010-05-31T13:00:00".} } \\ 
- -3663.1 & 4 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26 ?p1 dsw:person/james-michaelis.} } \\ 
- -3739.7 & 4 & \makecell{\texttt{dsw:conference/www/2012/paper/518/talk/speaker ?p2 ?n1.} } \\ 
- -3707.7 & 3 & \makecell{\texttt{\_:genid135 rdf:\_1 ?n1.} } \\ 
- -3725.5 & 3 & \makecell{\texttt{dsw:workshop/ores/2010/paper/main/9/authorlist ?p2 ?n1.} } \\ 
- -3703.2 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/gang-wang.} } \\ 
- -3703.3 & 3 & \makecell{\texttt{?n1 dc:creator dsw:person/samur-araujo.} } \\ 
- -3696.2 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/samur-araujo.} } \\ 
- -3667.3 & 3 & \makecell{\texttt{dsw:conference/iswc/2011/paper/semantic-web-in-use/68 ?p1 dsw:person/jianqiang-li.} } \\ 
- -3668.7 & 3 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p1 dsw:person/jianqiang-li.} } \\ 
- -3668.7 & 3 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p1 dsw:person/yu-zhao.} } \\ 
- -3682.9 & 3 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p1 dsw:person/jianqiang-li.} \\\texttt{dsw:workshop/LDOW/2008/paper/3 swc:isPartOf dsw:workshop/LDOW/2008/proceedings.} } \\ 
- -3684.5 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/varish-mulwad.} } \\ 
- -3711.5 & 3 & \makecell{\texttt{?n1 foaf:member dsw:person/xuanhui-wang.} } \\ 
- -3692.5 & 3 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid132.} } \\ 
- -3718.2 & 3 & \makecell{\texttt{\_:genid69 rdf:type ?n1.} } \\ 
- -3707.6 & 3 & \makecell{\texttt{\_:genid18 ?p2 "2010-05-31T13:00:00".} \\\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date "2010-05-31T13:00:00".} } \\ 
- -3684.5 & 3 & \makecell{\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date "2008-04-25T15:30:00".} } \\ 
- -3707.5 & 2 & \makecell{\texttt{dsw:person/rayid-ghani swc:holdsRole ?n1.} } \\ 
- -3704.5 & 2 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 foaf:maker ?n1.} } \\ 
- -3703.5 & 2 & \makecell{\texttt{?n1 foaf:maker dsw:person/yu-zhao.} } \\ 
- -3703.5 & 2 & \makecell{\texttt{?n1 foaf:maker dsw:person/jianqiang-li.} } \\ 
- -3706.2 & 2 & \makecell{\texttt{?n1 swc:heldBy dsw:person/kai-eckert.} } \\ 
- -3694.9 & 2 & \makecell{\texttt{dsw:organization/yahoo-inc ?p1 "Yahoo! Inc".} } \\ 
- -3732.0 & 2 & \makecell{\texttt{?n1 ?p2 "Yahoo! Inc".} } \\ 
- -3712.6 & 2 & \makecell{\texttt{\_:genid11 ?p2 ?n1.} \\\texttt{dsw:workshop/obi/2008/paper/main/7 swrs:author ?n1.} } \\ 
- -3706.3 & 2 & \makecell{\texttt{dsw:conference/iswc/2011/paper/poster-demo/40 swrs:author ?n1.} } \\ 
- -3701.4 & 2 & \makecell{\texttt{dsw:conference/iswc/2011/paper/poster-demo/40 http://purl.org/dc/terms/creator ?n1.} } \\ 
- -3708.3 & 2 & \makecell{\texttt{\_:genid10 http://www.w3.org/2002/12/cal/ical\_tzid ?n1.} } \\ 
- -3703.8 & 1 & \makecell{\texttt{dsw:person/kenneth-church swc:holdsRole ?n1.} } \\ 
- -3689.9 & 1 & \makecell{\texttt{dsw:person/lorenzino-vaccari ?p1 dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{dsw:person/rayid-ghani ?p1 dsw:conference/eswc/2011/programme-committee-member.} } \\
- -3667.9 & 1 & \makecell{\texttt{dsw:person/lorenzino-vaccari swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{dsw:person/rayid-ghani swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\
- -3692.5 & 1 & \makecell{\texttt{dsw:person/lorenzino-vaccari ?p1 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
- -3691.0 & 1 & \makecell{\texttt{dsw:person/neil-ireson ?p1 dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{dsw:person/rayid-ghani ?p1 dsw:conference/eswc/2011/programme-committee-member.} } \\
- -3692.7 & 1 & \makecell{\texttt{dsw:person/rayid-ghani ?p1 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
- -3670.7 & 1 & \makecell{\texttt{dsw:person/rayid-ghani swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\ 
- -3669.6 & 1 & \makecell{\texttt{dsw:workshop/swcs/2012/programme-committee-member swc:heldBy dsw:person/maria-esther-vidal.} } \\ 
- -3691.6 & 1 & \makecell{\texttt{dsw:workshop/swcs/2012/programme-committee-member ?p1 dsw:person/maria-esther-vidal.} } \\ 
- -3688.4 & 1 & \makecell{\texttt{dsw:organization/universidad-simon-bolivar ?p1 dsw:person/maria-esther-vidal.} } \\ 
- -3719.8 & 1 & \makecell{\texttt{dsw:workshop/ores/2010/paper/main/9/authorlist ?p2 ?n1.} \\\texttt{?n1 foaf:mbox\_sha1sum "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
- -3697.7 & 1 & \makecell{\texttt{dsw:workshop/ores/2010/paper/main/9/authorlist rdf:\_3 ?n1.} \\\texttt{?n1 foaf:mbox\_sha1sum "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
- -3707.2 & 1 & \makecell{\texttt{?n1 foaf:mbox\_sha1sum "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
- -3671.1 & 1 & \makecell{\texttt{dsw:person/adrian-marte foaf:mbox\_sha1sum "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
- -3729.2 & 1 & \makecell{\texttt{?n1 ?p2 "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
- -3693.1 & 1 & \makecell{\texttt{dsw:person/adrian-marte ?p1 "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
- -3685.9 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/515/authorlist ?p1 dsw:person/samur-araujo.} } \\ 
- -3663.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/515/authorlist rdf:\_1 dsw:person/samur-araujo.} } \\ 
- -3680.4 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/515 foaf:maker dsw:person/samur-araujo.} \\\texttt{dsw:conference/iswc/2010/paper/515/authorlist ?p1 dsw:person/samur-araujo.} } \\ 
- -3701.3 & 1 & \makecell{\texttt{?n1 http://purl.org/dc/terms/creator dsw:person/jianqiang-li.} } \\ 
- -3665.1 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/semantic-web-in-use/68 http://purl.org/dc/terms/creator dsw:person/jianqiang-li.} } \\ 
- -3666.5 & 1 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 foaf:maker dsw:person/jianqiang-li.} } \\ 
- -3666.5 & 1 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 foaf:maker dsw:person/yu-zhao.} } \\ 
- -3690.7 & 1 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p1 http://sunsite.informatik.rwth-aachen.de/Publications/CEUR-WS/Vol-369/paper04.pdf.} } \\ 
- -3666.3 & 1 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 dc:creator dsw:person/jianqiang-li.} } \\ 
- -3699.5 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/lisa-lena-opas-haenninen.} } \\ 
- -3663.3 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-775/authors rdf:\_1 dsw:person/lisa-lena-opas-haenninen.} } \\ 
- -3685.3 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-775/authors ?p1 dsw:person/lisa-lena-opas-haenninen.} } \\ 
- -3705.0 & 1 & \makecell{\texttt{dsw:person/nematollaah-shiri foaf:firstName ?n1.} } \\ 
- -3698.6 & 1 & \makecell{\texttt{dsw:person/nematollaah-shiri ?p1 foaf:Person.} } \\ 
- -3676.6 & 1 & \makecell{\texttt{dsw:person/nematollaah-shiri rdf:type foaf:Person.} } \\ 
- -3712.8 & 1 & \makecell{\texttt{dsw:person/nematollaah-shiri rdf:type ?n1.} } \\ 
- -3686.6 & 1 & \makecell{\texttt{\_:genid24 ?p1 dsw:person/yi-chin-tu.} } \\ 
- -3731.0 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 http://purl.org/ontology/bibo/authorList ?n1.} \\\texttt{?n1 rdf:\_1 ?n2.} } \\ 
- -3705.5 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 http://purl.org/ontology/bibo/authorList ?n1.} } \\ 
- -3691.4 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 ?p1 dsw:conference/dh/2010/abstracts/poster/ab-643/authors.} } \\ 
- -3727.5 & 1 & \makecell{\texttt{?n1 ?p2 dsw:conference/dh/2010/abstracts/poster/ab-643/authors.} } \\ 
- -3669.3 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 http://purl.org/ontology/bibo/authorList dsw:conference/dh/2010/abstracts/poster/ab-643/authors.} } \\ 
- -3705.5 & 1 & \makecell{\texttt{?n1 http://purl.org/ontology/bibo/authorList dsw:conference/dh/2010/abstracts/poster/ab-643/authors.} } \\ 
- -3670.1 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/research-track-committee-member swc:heldBy dsw:person/kai-eckert.} } \\ 
- -3692.1 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/research-track-committee-member ?p1 dsw:person/kai-eckert.} } \\ 
- -3750.6 & 1 & \makecell{\texttt{?n1 ?p3 ?n2.} \\\texttt{?n2 foaf:made dsw:conference/www/2010/paper/main/23.} } \\ 
- -3688.0 & 1 & \makecell{\texttt{\_:genid90 ?p1 dsw:person/jan-zemanek.} } \\ 
- -3675.7 & 1 & \makecell{\texttt{\_:genid15 rdf:\_1 dsw:person/c-maria-keet.} } \\ 
- -3671.7 & 1 & \makecell{\texttt{\_:genid15 rdf:\_1 dsw:person/daniel-bachlechner.} } \\ 
- -3699.6 & 1 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors rdf:\_1 ?n1.} } \\ 
- -3702.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors rdf:\_2 ?n1.} } \\ 
- -3666.6 & 1 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors rdf:\_2 dsw:person/deborah-mcguinness.} } \\ 
- -3688.7 & 1 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors ?p1 dsw:person/deborah-mcguinness.} } \\ 
- -3712.0 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/doctoral-consortium/14/authorlist ?p2 ?n1.} } \\ 
- -3675.9 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/doctoral-consortium/14/authorlist ?p1 dsw:person/varish-mulwad.} } \\ 
- -3653.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/doctoral-consortium/14/authorlist rdf:\_1 dsw:person/varish-mulwad.} } \\ 
- -3662.5 & 1 & \makecell{\texttt{\_:b3 rdf:\_1 dsw:person/varish-mulwad.} } \\ 
- -3684.5 & 1 & \makecell{\texttt{\_:b3 ?p1 dsw:person/varish-mulwad.} } \\ 
- -3698.7 & 1 & \makecell{\texttt{\_:b3 rdf:\_1 ?n1.} } \\ 
- -3667.4 & 1 & \makecell{\texttt{dsw:organization/yahoo-inc rdfs:label "Yahoo! Inc".} } \\ 
- -3670.4 & 1 & \makecell{\texttt{dsw:organization/yahoo-inc foaf:member dsw:person/xuanhui-wang.} } \\ 
- -3695.7 & 1 & \makecell{\texttt{\_:genid11 rdf:\_1 ?n1.} \\\texttt{dsw:workshop/obi/2008/paper/main/7 swrs:author ?n1.} } \\ 
- -3699.7 & 1 & \makecell{\texttt{\_:genid11 rdf:\_1 ?n1.} \\\texttt{dsw:conference/iswc/2011/paper/poster-demo/40 swrs:author ?n1.} } \\ 
- -3698.8 & 1 & \makecell{\texttt{?n1 http://purl.org/dc/terms/creator dsw:person/akira-mori.} } \\ 
- -3699.2 & 1 & \makecell{\texttt{?n1 swrs:editor dsw:person/denny-vrandecic.} } \\ 
- -3663.1 & 1 & \makecell{\texttt{dsw:workshop/insemtive/2008/proceedings swrs:editor dsw:person/denny-vrandecic.} } \\ 
- -3685.1 & 1 & \makecell{\texttt{dsw:workshop/insemtive/2008/proceedings ?p1 dsw:person/denny-vrandecic.} } \\ 
- -3690.3 & 1 & \makecell{\texttt{\_:genid48 ?p1 dsw:person/willem-van-hage.} } \\ 
- -3668.3 & 1 & \makecell{\texttt{\_:genid48 rdf:\_1 dsw:person/willem-van-hage.} } \\ 
- -3665.3 & 1 & \makecell{\texttt{dsw:conference/www/2008/paper/431 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid48.} } \\ 
- -3687.4 & 1 & \makecell{\texttt{dsw:conference/www/2008/paper/431 ?p1 \_:genid48.} } \\ 
- -3701.7 & 1 & \makecell{\texttt{dsw:conference/www/2007/paper/main/429 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n1.} } \\ 
- -3703.1 & 1 & \makecell{\texttt{?n1 foaf:maker dsw:person/brett-cannon.} } \\ 
- -3691.2 & 1 & \makecell{\texttt{\_:genid36 ?p1 dsw:person/vanja-josifovski.} } \\ 
- -3728.0 & 1 & \makecell{\texttt{dsw:workshop/om/2008/paper/main/21 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n1.} \\\texttt{?n1 rdf:\_7 ?n2.} } \\ 
- -3693.5 & 1 & \makecell{\texttt{\_:genid16 rdf:\_7 ?n1.} \\\texttt{dsw:workshop/om/2008/paper/main/21 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid16.} } \\ 
- -3690.1 & 1 & \makecell{\texttt{\_:genid23 ?p1 dsw:person/zheng-chen.} } \\ 
- -3697.6 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/semantic-web-in-use/76/authorlist rdf:\_9 ?n1.} } \\ 
- -3691.1 & 1 & \makecell{\texttt{dsw:workshop/sdow/2011/pcmember ?p1 dsw:person/matthew-rowe.} } \\ 
- -3669.0 & 1 & \makecell{\texttt{dsw:workshop/sdow/2011/pcmember swc:heldBy dsw:person/matthew-rowe.} } \\ 
- -3690.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/industry/21/authorlist rdf:\_1 ?n1.} } \\ 
- -3707.0 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/karl-aberer.} } \\ 
- -3670.8 & 1 & \makecell{\texttt{\_:genid1 rdf:\_1 dsw:person/karl-aberer.} } \\ 
- -3692.9 & 1 & \makecell{\texttt{\_:genid1 ?p1 dsw:person/karl-aberer.} } \\ 
- -3700.1 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/peyman-sazedj.} } \\ 
- -3700.1 & 1 & \makecell{\texttt{\_:genid18 http://www.w3.org/2002/12/cal/ical\_date ?n1.} } \\ 
- -3663.9 & 1 & \makecell{\texttt{\_:genid18 http://www.w3.org/2002/12/cal/ical\_date "2010-05-31T13:00:00".} } \\ 
- -3686.0 & 1 & \makecell{\texttt{\_:genid18 ?p1 "2010-05-31T13:00:00".} } \\ 
- -3671.5 & 1 & \makecell{\texttt{\_:genid11 http://www.w3.org/2002/12/cal/ical\_date "2010-05-31T13:00:00".} \\\texttt{\_:genid18 ?p1 "2010-05-31T13:00:00".} } \\ 
- -3707.6 & 1 & \makecell{\texttt{\_:genid11 http://www.w3.org/2002/12/cal/ical\_date ?n1.} \\\texttt{\_:genid18 ?p2 ?n1.} } \\ 
- -3693.5 & 1 & \makecell{\texttt{\_:genid11 ?p2 "2010-05-31T13:00:00".} \\\texttt{\_:genid18 ?p1 "2010-05-31T13:00:00".} } \\ 
- -3809.4 & 1 & \makecell{\texttt{dsw:conference/iswc-aswc/2007/chair\_12 swc:heldBy ?n1.} \\\texttt{?n2 ?p4 ?n3.} \\\texttt{?n2 ?p4 ?n1.} } \\ 
- -3707.5 & 1 & \makecell{\texttt{dsw:conference/eswc/2007/time-20070605T153000EuropeVienna rdf:type ?n1.} } \\ 
- -3695.5 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors ?p1 rdf:Seq.} } \\ 
- -3673.4 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:type rdf:Seq.} } \\ 
- -3709.6 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:type ?n1.} } \\ 
- -3699.6 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:type ?n1.} \\\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:\_4 dsw:person/stefan-beyer.} } \\
- -3735.8 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:type ?n2.} \\\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:\_4 ?n1.} } \\ 
- -3709.7 & 1 & \makecell{\texttt{dsw:conference/www/2012/paper/518/talk/speaker rdf:type ?n1.} } \\ 
- -3744.2 & 0 & \makecell{\texttt{?n1 ?p4 ?n2.} \\\texttt{?n2 rdf:\_7 ?n3.} } \\ 
+\input{results/dogfood/motifs-byfreq.latex}
  \hline
@@
 
-\begin{longtable}{ r r p{19cm} }
+\begin{longtable}{ r r p{10cm} }
  log factor & frequency & \\
 \hline\endhead
- 8919.0 & 2477 & \makecell{\texttt{?n1 swrs:isAbout ?n2.} } \\ 
- 3857.8 & 4124 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
- 923.9 & 1066 & \makecell{\texttt{?n1 ?p2 swrs:Person.} } \\ 
- 827.3 & 1045 & \makecell{\texttt{?n1 rdf:type swrs:Person.} } \\ 
- -141.5 & 1240 & \makecell{\texttt{?n1 ?p2 swrs:Publication.} } \\ 
- -167.7 & 1222 & \makecell{\texttt{?n1 rdf:type swrs:Publication.} } \\ 
- -893.7 & 1529 & \makecell{\texttt{?n1 swrs:author ?n2.} } \\ 
- -907.8 & 3388 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
- -939.9 & 687 & \makecell{\texttt{?n1 rdf:type swrs:InProceedings.} } \\ 
- -975.2 & 687 & \makecell{\texttt{?n1 ?p2 swrs:InProceedings.} } \\ 
- -1495.3 & 31 & \makecell{\texttt{?n1 ?p2 owl:DatatypeProperty.} } \\ 
- -1536.6 & 146 & \makecell{\texttt{?n1 rdf:type swrs:ResearchTopic.} } \\ 
- -1582.5 & 152 & \makecell{\texttt{?n1 rdf:type owl:Restriction.} } \\ 
- -1583.0 & 147 & \makecell{\texttt{?n1 ?p2 swrs:ResearchTopic.} } \\ 
- -1609.0 & 152 & \makecell{\texttt{?n1 owl:onProperty ?n2.} } \\ 
- -1613.5 & 152 & \makecell{\texttt{?n1 ?p2 owl:Restriction.} } \\ 
- -1668.8 & 55 & \makecell{\texttt{?n1 rdf:type owl:Class.} } \\ 
- -1696.9 & 55 & \makecell{\texttt{?n1 ?p2 owl:Class.} } \\ 
- -1745.2 & 65 & \makecell{\texttt{?n1 rdf:type swrs:PhDStudent.} } \\ 
- -1747.9 & 40 & \makecell{\texttt{?n1 swrs:isAbout http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
- -1767.8 & 78 & \makecell{\texttt{?n1 rdf:type swrs:Project.} } \\ 
- -1769.7 & 13 & \makecell{\texttt{swrs:InBook rdfs:subClassOf ?n1.} } \\ 
- -1775.0 & 87 & \makecell{\texttt{?n1 ?p2 swrs:Project.} } \\ 
- -1775.1 & 40 & \makecell{\texttt{?n1 ?p2 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
- -1777.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance swrs:isAbout http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\
- -1779.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:publication http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\
- -1780.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance rdf:type swrs:ResearchTopic.} } \\ 
- -1781.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance swrs:name "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
- -1784.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type swrs:Article.} } \\ 
- -1784.3 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id655instance rdf:type swrs:Person.} } \\ 
- -1784.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id262instance rdf:type swrs:Person.} } \\ 
- -1785.0 & 1 & \makecell{\texttt{\_:ub1bL199C18 rdf:type owl:Restriction.} } \\ 
- -1785.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id583instance rdf:type swrs:Person.} } \\ 
- -1785.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id637instance rdf:type swrs:Person.} } \\ 
- -1785.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id594instance rdf:type swrs:InProceedings.} } \\ 
- -1786.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id684instance rdf:type swrs:InProceedings.} } \\ 
- -1786.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id499instance rdf:type swrs:InProceedings.} } \\ 
- -1786.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id841instance rdf:type swrs:InProceedings.} } \\ 
- -1787.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2008instance rdf:type swrs:Person.} } \\ 
- -1787.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type swrs:Publication.} } \\ 
- -1787.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id663instance rdf:type swrs:InProceedings.} } \\ 
- -1787.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type swrs:Person.} } \\ 
- -1787.3 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance rdf:type swrs:Publication.} } \\ 
- -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2118instance rdf:type swrs:Person.} } \\ 
- -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance rdf:type swrs:Publication.} } \\ 
- -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance rdf:type swrs:Person.} } \\ 
- -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id709instance rdf:type swrs:Person.} } \\ 
- -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance rdf:type swrs:Person.} } \\ 
- -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance rdf:type swrs:Person.} } \\ 
- -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance rdf:type swrs:Person.} } \\ 
- -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id703instance rdf:type swrs:Person.} } \\ 
- -1791.2 & 1 & \makecell{\texttt{swrs:title ?p1 owl:DatatypeProperty.} } \\ 
- -1796.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance ?p1 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\
- -1796.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance ?p1 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\
- -1798.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance ?p1 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
- -1799.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance ?p1 swrs:ResearchTopic.} } \\ 
- -1800.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p1 "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
- -1800.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p1 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id877instance.} } \\
- -1801.0 & 14 & \makecell{\texttt{swrs:InBook ?p2 ?n1.} } \\ 
- -1802.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id28instance ?p1 swrs:ResearchTopic.} } \\ 
- -1802.2 & 227 & \makecell{\texttt{?n1 swrs:phone ?n2.} } \\ 
- -1803.3 & 1 & \makecell{\texttt{swrs:Proceedings ?p1 owl:Class.} } \\ 
- -1803.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id655instance ?p1 swrs:Person.} } \\ 
- -1803.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id350instance ?p1 swrs:Person.} } \\ 
- -1803.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id262instance ?p1 swrs:Person.} } \\ 
- -1804.0 & 1 & \makecell{\texttt{\_:ub1bL474C18 ?p1 owl:Restriction.} } \\ 
- -1804.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id583instance ?p1 swrs:Person.} } \\ 
- -1804.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id810instance ?p1 swrs:Person.} } \\ 
- -1804.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id184instance ?p1 swrs:Person.} } \\ 
- -1805.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id594instance ?p1 swrs:InProceedings.} } \\ 
- -1805.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id275instance ?p1 swrs:InProceedings.} } \\ 
- -1805.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id841instance ?p1 swrs:InProceedings.} } \\ 
- -1806.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2008instance ?p1 swrs:Person.} } \\ 
- -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance ?p1 swrs:Publication.} } \\ 
- -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id663instance ?p1 swrs:InProceedings.} } \\ 
- -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance ?p1 swrs:Person.} } \\ 
- -1806.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance ?p1 swrs:Publication.} } \\ 
- -1806.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id773instance ?p1 swrs:Publication.} } \\ 
- -1806.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance ?p1 swrs:Publication.} } \\ 
- -1806.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance ?p1 swrs:Person.} } \\ 
- -1806.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id954instance ?p1 swrs:Publication.} } \\ 
- -1806.7 & 7 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:publication ?n1.} } \\ 
- -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance ?p1 swrs:Person.} } \\ 
- -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance ?p1 swrs:Person.} } \\ 
- -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id820instance ?p1 swrs:Person.} } \\ 
- -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance ?p1 swrs:Person.} } \\ 
- -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id786instance ?p1 swrs:Person.} } \\ 
- -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id848instance ?p1 swrs:Person.} } \\ 
- -1807.3 & 1 & \makecell{\texttt{\_:ub1bL560C18 ?p1 swrs:Person.} } \\ 
- -1808.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id165instance swrs:name ?n1.} } \\ 
- -1808.9 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:worksAtProject ?n1.} } \\ 
- -1811.1 & 1 & \makecell{\texttt{?n1 swrs:name "CKDD - Conceptual Knowledge Discovery"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
- -1811.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance rdf:type ?n1.} } \\ 
- -1812.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance swrs:name ?n1.} } \\ 
- -1812.1 & 1 & \makecell{\texttt{?n1 swrs:name "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
- -1813.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id68instance swrs:phone ?n1.} } \\ 
- -1813.3 & 9 & \makecell{\texttt{swrs:Organization ?p2 ?n1.} } \\ 
- -1813.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id79instance swrs:name ?n1.} } \\ 
- -1814.1 & 1 & \makecell{\texttt{\_:ub1bL474C18 owl:onProperty ?n1.} } \\ 
- -1814.8 & 1 & \makecell{\texttt{swrs:Organization rdf:type ?n1.} } \\ 
- -1815.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id18instance rdf:type ?n1.} } \\ 
- -1815.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id814instance rdf:type ?n1.} } \\ 
- -1815.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id20instance rdf:type ?n1.} } \\ 
- -1816.0 & 1 & \makecell{\texttt{\_:ub1bL422C18 rdf:type ?n1.} } \\ 
- -1816.0 & 1 & \makecell{\texttt{\_:ub1bL355C18 rdf:type ?n1.} } \\ 
- -1818.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance rdf:type ?n1.} } \\ 
- -1818.8 & 152 & \makecell{\texttt{?n1 swrs:author http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id20instance.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance rdf:type ?n1.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance rdf:type ?n1.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance rdf:type ?n1.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance rdf:type ?n1.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id593instance rdf:type ?n1.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id439instance rdf:type ?n1.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id804instance rdf:type ?n1.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id305instance rdf:type ?n1.} } \\ 
- -1820.7 & 22 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance ?p2 ?n1.} } \\ 
- -1821.4 & 1 & \makecell{\texttt{swrs:member ?p2 ?n1.} } \\ 
- -1821.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type ?n1.} } \\ 
- -1822.2 & 1 & \makecell{\texttt{swrs:name ?p2 ?n1.} } \\ 
- -1822.2 & 1 & \makecell{\texttt{swrs:title ?p2 ?n1.} } \\ 
- -1824.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id19instance rdf:type ?n1.} } \\ 
- -1824.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type ?n1.} } \\ 
- -1824.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1089instance rdf:type ?n1.} } \\ 
- -1825.1 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance rdf:type ?n1.} } \\ 
- -1825.9 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id552instance swrs:author ?n1.} } \\ 
- -1826.2 & 3 & \makecell{\texttt{\_:ub1bL658C27 ?p2 ?n1.} } \\ 
- -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id617instance rdf:type ?n1.} } \\ 
- -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id499instance rdf:type ?n1.} } \\ 
- -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id516instance rdf:type ?n1.} } \\ 
- -1826.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance rdf:type ?n1.} } \\ 
- -1826.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1333instance rdf:type ?n1.} } \\ 
- -1826.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id512instance rdf:type ?n1.} } \\ 
- -1826.8 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id572instance rdf:type ?n1.} } \\ 
- -1826.8 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance rdf:type ?n1.} } \\ 
- -1827.0 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id64instance rdf:type ?n1.} } \\ 
- -1827.0 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id777instance rdf:type ?n1.} } \\ 
- -1827.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1306instance rdf:type ?n1.} } \\ 
- -1827.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance rdf:type ?n1.} } \\ 
- -1828.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1166instance rdf:type ?n1.} } \\ 
- -1830.1 & 1 & \makecell{\texttt{?n1 ?p2 "CKDD - Conceptual Knowledge Discovery"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
- -1831.1 & 1 & \makecell{\texttt{?n1 ?p2 "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
- -1833.9 & 11 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance swrs:isAbout ?n1.} } \\ 
- -1834.8 & 3 & \makecell{\texttt{?n1 ?p2 swrs:Graduate.} } \\ 
- -1837.2 & 3 & \makecell{\texttt{\_:ub1bL58C27 ?p2 ?n1.} } \\ 
- -1837.9 & 3 & \makecell{\texttt{\_:ub1bL474C18 ?p2 ?n1.} } \\ 
- -1839.0 & 3 & \makecell{\texttt{\_:ub1bL102C18 ?p2 ?n1.} } \\ 
- -1839.6 & 8 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance ?p2 ?n1.} } \\ 
- -1839.9 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance ?p2 ?n1.} } \\ 
- -1840.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id144instance ?p2 ?n1.} } \\ 
- -1840.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id190instance ?p2 ?n1.} } \\ 
- -1841.0 & 14 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Projekte/viewProjektOWL/id51instance swrs:member ?n1.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
- -1841.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance ?p2 ?n1.} } \\ 
- -1841.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance ?p2 ?n1.} } \\ 
- -1841.7 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id774instance ?p2 ?n1.} } \\ 
- -1842.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p2 ?n1.} } \\ 
- -1842.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id607instance ?p2 ?n1.} } \\ 
- -1843.9 & 6 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id530instance ?p2 ?n1.} } \\ 
- -1844.5 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id804instance ?p2 ?n1.} } \\ 
- -1850.2 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id165instance ?p2 ?n1.} } \\ 
- -1850.6 & 12 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2067instance ?p2 ?n1.} } \\ 
- -1860.0 & 1907 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance rdf:type ?n2.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
- -1869.8 & 17 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance ?p2 ?n1.} } \\ 
- -1877.4 & 10 & \makecell{\texttt{?n1 ?p2 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
- -1882.2 & 9 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id275instance ?p2 ?n1.} } \\ 
- -1883.4 & 18 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1272instance ?p2 ?n1.} } \\ 
- -1883.7 & 13 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance ?p2 ?n1.} } \\ 
- -1887.6 & 1130 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
- -1888.2 & 12 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id228instance ?p2 ?n1.} } \\ 
- -1895.2 & 11 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id572instance ?p2 ?n1.} } \\ 
- -1903.8 & 16 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1305instance ?p2 ?n1.} } \\ 
- -1904.3 & 16 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1317instance ?p2 ?n1.} } \\ 
- -1908.7 & 13 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1152instance ?p2 ?n1.} } \\ 
- -1930.3 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1124instance ?p2 ?n1.} } \\ 
- -1931.5 & 2184 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance ?p3 ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
- -1937.4 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id882instance ?p2 ?n1.} } \\ 
- -1942.5 & 25 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance ?p2 ?n1.} } \\ 
- -1942.7 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1025instance ?p2 ?n1.} } \\ 
- -2254.8 & 24274 & \makecell{\texttt{?n1 rdf:type ?n3.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
- -4014.6 & 494 & \makecell{\texttt{?n1 swrs:name ?n2.} } \\ 
+\input{results/aifb/motifs-byscore.latex}
 \hline
@@
 
-\begin{longtable}{ r r p{19cm} }
+\begin{longtable}{ r r p{10cm} }
  log factor & frequency & \\
 \hline\endhead
- -2254.8 & 24274 & \makecell{\texttt{?n1 rdf:type ?n3.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
- 3857.8 & 4124 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
- -907.8 & 3388 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
- 8919.0 & 2477 & \makecell{\texttt{?n1 swrs:isAbout ?n2.} } \\ 
- -1931.5 & 2184 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance ?p3 ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
- -1860.0 & 1907 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance rdf:type ?n2.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
- -893.7 & 1529 & \makecell{\texttt{?n1 swrs:author ?n2.} } \\ 
- -141.5 & 1240 & \makecell{\texttt{?n1 ?p2 swrs:Publication.} } \\ 
- -167.7 & 1222 & \makecell{\texttt{?n1 rdf:type swrs:Publication.} } \\ 
- -1887.6 & 1130 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
- 923.9 & 1066 & \makecell{\texttt{?n1 ?p2 swrs:Person.} } \\ 
- 827.3 & 1045 & \makecell{\texttt{?n1 rdf:type swrs:Person.} } \\ 
- -975.2 & 687 & \makecell{\texttt{?n1 ?p2 swrs:InProceedings.} } \\ 
- -939.9 & 687 & \makecell{\texttt{?n1 rdf:type swrs:InProceedings.} } \\ 
- -4014.6 & 494 & \makecell{\texttt{?n1 swrs:name ?n2.} } \\ 
- -1802.2 & 227 & \makecell{\texttt{?n1 swrs:phone ?n2.} } \\ 
- -1613.5 & 152 & \makecell{\texttt{?n1 ?p2 owl:Restriction.} } \\ 
- -1582.5 & 152 & \makecell{\texttt{?n1 rdf:type owl:Restriction.} } \\ 
- -1609.0 & 152 & \makecell{\texttt{?n1 owl:onProperty ?n2.} } \\ 
- -1818.8 & 152 & \makecell{\texttt{?n1 swrs:author http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id20instance.} } \\ 
- -1583.0 & 147 & \makecell{\texttt{?n1 ?p2 swrs:ResearchTopic.} } \\ 
- -1536.6 & 146 & \makecell{\texttt{?n1 rdf:type swrs:ResearchTopic.} } \\ 
- -1775.0 & 87 & \makecell{\texttt{?n1 ?p2 swrs:Project.} } \\ 
- -1767.8 & 78 & \makecell{\texttt{?n1 rdf:type swrs:Project.} } \\ 
- -1745.2 & 65 & \makecell{\texttt{?n1 rdf:type swrs:PhDStudent.} } \\ 
- -1668.8 & 55 & \makecell{\texttt{?n1 rdf:type owl:Class.} } \\ 
- -1696.9 & 55 & \makecell{\texttt{?n1 ?p2 owl:Class.} } \\ 
- -1775.1 & 40 & \makecell{\texttt{?n1 ?p2 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
- -1747.9 & 40 & \makecell{\texttt{?n1 swrs:isAbout http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
- -1495.3 & 31 & \makecell{\texttt{?n1 ?p2 owl:DatatypeProperty.} } \\ 
- -1942.5 & 25 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance ?p2 ?n1.} } \\ 
- -1820.7 & 22 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance ?p2 ?n1.} } \\ 
- -1937.4 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id882instance ?p2 ?n1.} } \\ 
- -1942.7 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1025instance ?p2 ?n1.} } \\ 
- -1930.3 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1124instance ?p2 ?n1.} } \\ 
- -1883.4 & 18 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1272instance ?p2 ?n1.} } \\ 
- -1869.8 & 17 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance ?p2 ?n1.} } \\ 
- -1903.8 & 16 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1305instance ?p2 ?n1.} } \\ 
- -1904.3 & 16 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1317instance ?p2 ?n1.} } \\ 
- -1801.0 & 14 & \makecell{\texttt{swrs:InBook ?p2 ?n1.} } \\ 
- -1841.0 & 14 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Projekte/viewProjektOWL/id51instance swrs:member ?n1.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
- -1769.7 & 13 & \makecell{\texttt{swrs:InBook rdfs:subClassOf ?n1.} } \\ 
- -1908.7 & 13 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1152instance ?p2 ?n1.} } \\ 
- -1883.7 & 13 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance ?p2 ?n1.} } \\ 
- -1888.2 & 12 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id228instance ?p2 ?n1.} } \\ 
- -1850.6 & 12 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2067instance ?p2 ?n1.} } \\ 
- -1833.9 & 11 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance swrs:isAbout ?n1.} } \\ 
- -1895.2 & 11 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id572instance ?p2 ?n1.} } \\ 
- -1877.4 & 10 & \makecell{\texttt{?n1 ?p2 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
- -1882.2 & 9 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id275instance ?p2 ?n1.} } \\ 
- -1813.3 & 9 & \makecell{\texttt{swrs:Organization ?p2 ?n1.} } \\ 
- -1839.6 & 8 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance ?p2 ?n1.} } \\ 
- -1806.7 & 7 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:publication ?n1.} } \\ 
- -1843.9 & 6 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id530instance ?p2 ?n1.} } \\ 
- -1839.9 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance ?p2 ?n1.} } \\ 
- -1850.2 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id165instance ?p2 ?n1.} } \\ 
- -1825.9 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id552instance swrs:author ?n1.} } \\ 
- -1841.7 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id774instance ?p2 ?n1.} } \\ 
- -1837.9 & 3 & \makecell{\texttt{\_:ub1bL474C18 ?p2 ?n1.} } \\ 
- -1837.2 & 3 & \makecell{\texttt{\_:ub1bL58C27 ?p2 ?n1.} } \\ 
- -1840.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id190instance ?p2 ?n1.} } \\ 
- -1841.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance ?p2 ?n1.} } \\ 
- -1834.8 & 3 & \makecell{\texttt{?n1 ?p2 swrs:Graduate.} } \\ 
- -1841.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance ?p2 ?n1.} } \\ 
- -1842.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p2 ?n1.} } \\ 
- -1839.0 & 3 & \makecell{\texttt{\_:ub1bL102C18 ?p2 ?n1.} } \\ 
- -1844.5 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id804instance ?p2 ?n1.} } \\ 
- -1826.2 & 3 & \makecell{\texttt{\_:ub1bL658C27 ?p2 ?n1.} } \\ 
- -1840.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id144instance ?p2 ?n1.} } \\ 
- -1842.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id607instance ?p2 ?n1.} } \\ 
- -1808.9 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:worksAtProject ?n1.} } \\ 
- -1821.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type ?n1.} } \\ 
- -1825.1 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance rdf:type ?n1.} } \\ 
- -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id617instance rdf:type ?n1.} } \\ 
- -1827.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1306instance rdf:type ?n1.} } \\ 
- -1826.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id512instance rdf:type ?n1.} } \\ 
- -1826.8 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id572instance rdf:type ?n1.} } \\ 
- -1824.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type ?n1.} } \\ 
- -1827.0 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id64instance rdf:type ?n1.} } \\ 
- -1827.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance rdf:type ?n1.} } \\ 
- -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id499instance rdf:type ?n1.} } \\ 
- -1827.0 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id777instance rdf:type ?n1.} } \\ 
- -1826.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1333instance rdf:type ?n1.} } \\ 
- -1824.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1089instance rdf:type ?n1.} } \\ 
- -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id516instance rdf:type ?n1.} } \\ 
- -1824.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id19instance rdf:type ?n1.} } \\ 
- -1826.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance rdf:type ?n1.} } \\ 
- -1828.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1166instance rdf:type ?n1.} } \\ 
- -1826.8 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance rdf:type ?n1.} } \\ 
- -1798.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance ?p1 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
- -1779.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:publication http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\
- -1804.0 & 1 & \makecell{\texttt{\_:ub1bL474C18 ?p1 owl:Restriction.} } \\ 
- -1814.1 & 1 & \makecell{\texttt{\_:ub1bL474C18 owl:onProperty ?n1.} } \\ 
- -1816.0 & 1 & \makecell{\texttt{\_:ub1bL422C18 rdf:type ?n1.} } \\ 
- -1802.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id28instance ?p1 swrs:ResearchTopic.} } \\ 
- -1780.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance rdf:type swrs:ResearchTopic.} } \\ 
- -1811.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance rdf:type ?n1.} } \\ 
- -1799.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance ?p1 swrs:ResearchTopic.} } \\ 
- -1804.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id583instance ?p1 swrs:Person.} } \\ 
- -1785.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id583instance rdf:type swrs:Person.} } \\ 
- -1787.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type swrs:Person.} } \\ 
- -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance ?p1 swrs:Person.} } \\ 
- -1796.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance ?p1 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\
- -1777.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance swrs:isAbout http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\
- -1796.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance ?p1 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\
- -1822.2 & 1 & \makecell{\texttt{swrs:name ?p2 ?n1.} } \\ 
- -1787.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id663instance rdf:type swrs:InProceedings.} } \\ 
- -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id663instance ?p1 swrs:InProceedings.} } \\ 
- -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id709instance rdf:type swrs:Person.} } \\ 
- -1784.3 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id655instance rdf:type swrs:Person.} } \\ 
- -1803.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id655instance ?p1 swrs:Person.} } \\ 
- -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id848instance ?p1 swrs:Person.} } \\ 
- -1816.0 & 1 & \makecell{\texttt{\_:ub1bL355C18 rdf:type ?n1.} } \\ 
- -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance ?p1 swrs:Person.} } \\ 
- -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance rdf:type swrs:Person.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance rdf:type ?n1.} } \\ 
- -1804.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id810instance ?p1 swrs:Person.} } \\ 
- -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance ?p1 swrs:Person.} } \\ 
- -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance rdf:type swrs:Person.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance rdf:type ?n1.} } \\ 
- -1805.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id594instance ?p1 swrs:InProceedings.} } \\ 
- -1785.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id594instance rdf:type swrs:InProceedings.} } \\ 
- -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id820instance ?p1 swrs:Person.} } \\ 
- -1807.3 & 1 & \makecell{\texttt{\_:ub1bL560C18 ?p1 swrs:Person.} } \\ 
- -1804.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id184instance ?p1 swrs:Person.} } \\ 
- -1786.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id841instance rdf:type swrs:InProceedings.} } \\ 
- -1805.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id841instance ?p1 swrs:InProceedings.} } \\ 
- -1786.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id684instance rdf:type swrs:InProceedings.} } \\ 
- -1784.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type swrs:Article.} } \\ 
- -1787.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type swrs:Publication.} } \\ 
- -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance ?p1 swrs:Publication.} } \\ 
- -1806.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id954instance ?p1 swrs:Publication.} } \\ 
- -1806.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id773instance ?p1 swrs:Publication.} } \\ 
- -1815.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id20instance rdf:type ?n1.} } \\ 
- -1803.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id350instance ?p1 swrs:Person.} } \\ 
- -1803.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id262instance ?p1 swrs:Person.} } \\ 
- -1784.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id262instance rdf:type swrs:Person.} } \\ 
- -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance ?p1 swrs:Person.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance rdf:type ?n1.} } \\ 
- -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance rdf:type swrs:Person.} } \\ 
- -1813.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id68instance swrs:phone ?n1.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id305instance rdf:type ?n1.} } \\ 
- -1822.2 & 1 & \makecell{\texttt{swrs:title ?p2 ?n1.} } \\ 
- -1791.2 & 1 & \makecell{\texttt{swrs:title ?p1 owl:DatatypeProperty.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance rdf:type ?n1.} } \\ 
- -1800.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p1 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id877instance.} } \\
- -1812.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance swrs:name ?n1.} } \\ 
- -1781.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance swrs:name "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
- -1800.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p1 "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
- -1831.1 & 1 & \makecell{\texttt{?n1 ?p2 "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
- -1812.1 & 1 & \makecell{\texttt{?n1 swrs:name "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
- -1813.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id79instance swrs:name ?n1.} } \\ 
- -1811.1 & 1 & \makecell{\texttt{?n1 swrs:name "CKDD - Conceptual Knowledge Discovery"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
- -1830.1 & 1 & \makecell{\texttt{?n1 ?p2 "CKDD - Conceptual Knowledge Discovery"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
- -1808.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id165instance swrs:name ?n1.} } \\ 
- -1785.0 & 1 & \makecell{\texttt{\_:ub1bL199C18 rdf:type owl:Restriction.} } \\ 
- -1815.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id18instance rdf:type ?n1.} } \\ 
- -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance rdf:type swrs:Publication.} } \\ 
- -1806.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance ?p1 swrs:Publication.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id593instance rdf:type ?n1.} } \\ 
- -1786.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id499instance rdf:type swrs:InProceedings.} } \\ 
- -1805.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id275instance ?p1 swrs:InProceedings.} } \\ 
- -1821.4 & 1 & \makecell{\texttt{swrs:member ?p2 ?n1.} } \\ 
- -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2118instance rdf:type swrs:Person.} } \\ 
- -1814.8 & 1 & \makecell{\texttt{swrs:Organization rdf:type ?n1.} } \\ 
- -1815.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id814instance rdf:type ?n1.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id439instance rdf:type ?n1.} } \\ 
- -1785.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id637instance rdf:type swrs:Person.} } \\ 
- -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id703instance rdf:type swrs:Person.} } \\ 
- -1787.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2008instance rdf:type swrs:Person.} } \\ 
- -1806.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2008instance ?p1 swrs:Person.} } \\ 
- -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id804instance rdf:type ?n1.} } \\ 
- -1803.3 & 1 & \makecell{\texttt{swrs:Proceedings ?p1 owl:Class.} } \\ 
- -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id786instance ?p1 swrs:Person.} } \\ 
- -1806.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance ?p1 swrs:Person.} } \\ 
- -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance rdf:type swrs:Person.} } \\ 
- -1818.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance rdf:type ?n1.} } \\ 
- -1787.3 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance rdf:type swrs:Publication.} } \\ 
- -1806.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance ?p1 swrs:Publication.} } \\ 
+ \input{results/aifb/motifs-byfreq.latex}
 \hline
@@
 
-\begin{longtable}{ r r p{19cm} }
+\begin{longtable}{ r r p{10cm} }
  log factor & frequency & \\
 \hline\endhead
- 11178.4 & 5996 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_hasAtom ?n2.} } \\ 
- 1014.6 & 1840 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- 938.0 & 1850 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- 577.6 & 1897 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
- -1250.9 & 138 & \makecell{\texttt{?n1 rdfs:subClassOf ?n2.} } \\ 
- -1422.9 & 320 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Ar\_halide.} } \\ 
- -1593.5 & 319 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_charge ?n2.} } \\ 
- -1689.4 & 181 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -1707.3 & 173 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
- -1719.3 & 181 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -1737.1 & 173 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
- -1830.8 & 110 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -1859.3 & 110 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -1865.4 & 142 & \makecell{\texttt{?n1 rdf:type owl:Class.} } \\ 
- -1879.4 & 22 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
- -1886.3 & 19 & \makecell{\texttt{?n1 rdfs:subClassOf http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
- -1894.6 & 142 & \makecell{\texttt{?n1 ?p2 owl:Class.} } \\ 
- -1911.7 & 67 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
- -1928.9 & 50 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
- -1938.7 & 67 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
- -1981.2 & 36 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d219 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
- -1988.7 & 16 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d84 http://dl-learner.org/carcinogenesis\_hasStructure ?n1.} \\\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\
- -1996.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} \\\texttt{http://dl-learner.org/carcinogenesis\_bond5500 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d270\_13.} } \\
- -2000.2 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester rdfs:subClassOf http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
- -2000.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
- -2001.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_6.} } \\ 
- -2002.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 rdf:type owl:Class.} \\\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\
- -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
- -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond100 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2003.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3151 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
- -2008.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\
- -2008.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\
- -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
- -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-318 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
- -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1585 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2172 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\
- -2009.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2182 rdf:type http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2010.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Molybdenum rdf:type owl:Class.} } \\ 
- -2010.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Carbon-22 rdf:type owl:Class.} } \\ 
- -2011.7 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} \\\texttt{http://dl-learner.org/carcinogenesis\_bond32 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d10\_6.} } \\
- -2012.2 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond3442 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond16 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2014.7 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 rdf:type http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
- -2017.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester ?p1 http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
- -2018.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 ?p1 http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
- -2018.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 ?p1 http://dl-learner.org/carcinogenesis\_d135\_23.} } \\ 
- -2019.5 & 25 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d216 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
- -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 ?p1 http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
- -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2020.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3218 ?p1 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
- -2020.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3151 ?p1 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
- -2020.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p1 owl:Class.} \\\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\
- -2022.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\
- -2025.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\
- -2026.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-120 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-980 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1585 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2172 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-537 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2026.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\
- -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2182 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-939 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1155 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1260 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-946 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2275 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p1 owl:Class.} } \\ 
- -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ar\_halide ?p1 owl:Class.} } \\ 
- -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Alcohol ?p1 owl:Class.} } \\ 
- -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Carbon-22 ?p1 owl:Class.} } \\ 
- -2030.2 & 4 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2031.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\
- -2031.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di260 rdfs:subClassOf ?n1.} } \\ 
- -2032.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 ?p1 http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
- -2032.9 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\
- -2033.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester rdfs:subClassOf ?n1.} } \\ 
- -2034.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Iodine-95 rdfs:subClassOf ?n1.} } \\ 
- -2034.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Chlorine-93 rdfs:subClassOf ?n1.} } \\ 
- -2034.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type ?n1.} } \\ 
- -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
- -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
- -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
- -2036.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
- -2036.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
- -2041.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type ?n1.} } \\ 
- -2042.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
- -2042.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_di10-3291 rdf:type ?n1.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type ?n1.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type ?n1.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2491 rdf:type ?n1.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2916 rdf:type ?n1.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-222 rdf:type ?n1.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-1445 rdf:type ?n1.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 rdf:type ?n1.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1069 rdf:type ?n1.} } \\ 
- -2042.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 rdf:type ?n1.} } \\ 
- -2042.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2634 rdf:type ?n1.} } \\ 
- -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Nitrogen-499 rdf:type ?n1.} } \\ 
- -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 rdf:type ?n1.} } \\ 
- -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Non\_ar\_6c\_ring rdf:type ?n1.} } \\ 
- -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-1525 rdf:type ?n1.} } \\ 
- -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-2912 rdf:type ?n1.} } \\ 
- -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-3038 rdf:type ?n1.} } \\ 
- -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-216 rdf:type ?n1.} } \\ 
- -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-180 rdf:type ?n1.} } \\ 
- -2044.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d243\_16 http://dl-learner.org/carcinogenesis\_charge ?n1.} } \\ 
- -2045.1 & 1 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
- -2045.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d82\_26 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1421 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4108 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond7147 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond566 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond3448 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond662 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond5306 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond5197 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond6533 rdf:type ?n1.} } \\ 
- -2048.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 rdf:type ?n1.} } \\ 
- -2052.2 & 6 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
- -2053.6 & 66 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3169 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} \\\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amino.} } \\
- -2056.6 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p2 ?n1.} } \\ 
- -2059.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p2 ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p2 ?n1.} } \\ 
- -2059.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 ?p2 ?n1.} } \\ 
- -2059.9 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Non\_ar\_6c\_ring ?p2 ?n1.} } \\ 
- -2059.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p2 ?n1.} } \\ 
- -2059.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2916 ?p2 ?n1.} } \\ 
- -2060.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 ?p2 ?n1.} } \\ 
- -2060.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1069 ?p2 ?n1.} } \\ 
- -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 ?p2 ?n1.} } \\ 
- -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2634 ?p2 ?n1.} } \\ 
- -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-939 ?p2 ?n1.} } \\ 
- -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1260 ?p2 ?n1.} } \\ 
- -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2275 ?p2 ?n1.} } \\ 
- -2061.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di51 ?p2 ?n1.} } \\ 
- -2061.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di260 ?p2 ?n1.} } \\ 
- -2061.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-180 ?p2 ?n1.} } \\ 
- -2062.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester ?p2 ?n1.} } \\ 
- -2063.1 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Chlorine-93 ?p2 ?n1.} } \\ 
- -2065.9 & 9 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d104 http://dl-learner.org/carcinogenesis\_hasStructure ?n1.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
- -2071.1 & 2 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d135\_23.} } \\ 
- -2071.4 & 5 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
- -2071.9 & 5 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2073.9 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 ?p2 ?n1.} } \\ 
- -2075.7 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 ?p2 ?n1.} } \\ 
- -2075.7 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 ?p2 ?n1.} } \\ 
- -2077.4 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d243\_16 ?p2 ?n1.} } \\ 
- -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 ?p2 ?n1.} } \\ 
- -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 ?p2 ?n1.} } \\ 
- -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 ?p2 ?n1.} } \\ 
- -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 ?p2 ?n1.} } \\ 
- -2086.2 & 3 & \makecell{\texttt{?n1 rdf:type ?n2.} \\\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d11\_4.} } \\ 
- -2094.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_di10-3291 rdf:type ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
- -2094.6 & 541 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
- -2114.7 & 92 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond ?n2.} } \\ 
- -2139.2 & 4 & \makecell{\texttt{?n1 ?p3 http://dl-learner.org/carcinogenesis\_d11\_4.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
- -2151.0 & 62 & \makecell{\texttt{?n1 ?p4 http://dl-learner.org/carcinogenesis\_d11\_4.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
- -2743.8 & 126 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
+\input{results/mutag/motifs-byscore.latex}
 \hline
@@
 
-\begin{longtable}{ r r p{19cm} }
+\begin{longtable}{ r r p{10cm} }
  log factor & frequency & \\
 \hline\endhead
- 11178.4 & 5996 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_hasAtom ?n2.} } \\ 
- 577.6 & 1897 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
- 938.0 & 1850 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- 1014.6 & 1840 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2094.6 & 541 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
- -1422.9 & 320 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Ar\_halide.} } \\ 
- -1593.5 & 319 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_charge ?n2.} } \\ 
- -1719.3 & 181 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -1689.4 & 181 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -1707.3 & 173 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
- -1737.1 & 173 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
- -1894.6 & 142 & \makecell{\texttt{?n1 ?p2 owl:Class.} } \\ 
- -1865.4 & 142 & \makecell{\texttt{?n1 rdf:type owl:Class.} } \\ 
- -1250.9 & 138 & \makecell{\texttt{?n1 rdfs:subClassOf ?n2.} } \\ 
- -2743.8 & 126 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
- -1859.3 & 110 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -1830.8 & 110 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2114.7 & 92 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond ?n2.} } \\ 
- -1911.7 & 67 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
- -1938.7 & 67 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
- -2053.6 & 66 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3169 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} \\\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amino.} } \\
- -2151.0 & 62 & \makecell{\texttt{?n1 ?p4 http://dl-learner.org/carcinogenesis\_d11\_4.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
- -1928.9 & 50 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
- -1981.2 & 36 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d219 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
- -2019.5 & 25 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d216 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
- -1879.4 & 22 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
- -1886.3 & 19 & \makecell{\texttt{?n1 rdfs:subClassOf http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
- -1988.7 & 16 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d84 http://dl-learner.org/carcinogenesis\_hasStructure ?n1.} \\\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\
- -2065.9 & 9 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d104 http://dl-learner.org/carcinogenesis\_hasStructure ?n1.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
- -2052.2 & 6 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
- -2071.4 & 5 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
- -2071.9 & 5 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2030.2 & 4 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2139.2 & 4 & \makecell{\texttt{?n1 ?p3 http://dl-learner.org/carcinogenesis\_d11\_4.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
- -2075.7 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 ?p2 ?n1.} } \\ 
- -2032.9 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\
- -2073.9 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 ?p2 ?n1.} } \\ 
- -2086.2 & 3 & \makecell{\texttt{?n1 rdf:type ?n2.} \\\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d11\_4.} } \\ 
- -2075.7 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 ?p2 ?n1.} } \\ 
- -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 ?p2 ?n1.} } \\ 
- -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 ?p2 ?n1.} } \\ 
- -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 ?p2 ?n1.} } \\ 
- -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 ?p2 ?n1.} } \\ 
- -2036.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
- -2036.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
- -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
- -2031.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\
- -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
- -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
- -2071.1 & 2 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d135\_23.} } \\ 
- -2062.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester ?p2 ?n1.} } \\ 
- -2063.1 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Chlorine-93 ?p2 ?n1.} } \\ 
- -2061.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di51 ?p2 ?n1.} } \\ 
- -2056.6 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p2 ?n1.} } \\ 
- -2059.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 ?p2 ?n1.} } \\ 
- -2061.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di260 ?p2 ?n1.} } \\ 
- -2077.4 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d243\_16 ?p2 ?n1.} } \\ 
- -2042.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
- -2094.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_di10-3291 rdf:type ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
- -2059.9 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Non\_ar\_6c\_ring ?p2 ?n1.} } \\ 
- -2045.1 & 1 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
- -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 ?p1 http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
- -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
- -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond100 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2001.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_6.} } \\ 
- -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\
- -2012.2 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2026.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\
- -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
- -2018.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 ?p1 http://dl-learner.org/carcinogenesis\_d135\_23.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-980 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1585 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1585 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1421 rdf:type ?n1.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-120 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4108 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond7147 rdf:type ?n1.} } \\ 
- -2017.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester ?p1 http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
- -2000.2 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester rdfs:subClassOf http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
- -2033.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester rdfs:subClassOf ?n1.} } \\ 
- -2034.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Iodine-95 rdfs:subClassOf ?n1.} } \\ 
- -2034.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Chlorine-93 rdfs:subClassOf ?n1.} } \\ 
- -2045.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d82\_26 rdf:type ?n1.} } \\ 
- -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-1525 rdf:type ?n1.} } \\ 
- -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2172 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2172 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Nitrogen-499 rdf:type ?n1.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-537 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2060.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 ?p2 ?n1.} } \\ 
- -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 rdf:type ?n1.} } \\ 
- -2042.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 rdf:type ?n1.} } \\ 
- -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 ?p2 ?n1.} } \\ 
- -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2182 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2009.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2182 rdf:type http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond566 rdf:type ?n1.} } \\ 
- -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-2912 rdf:type ?n1.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type ?n1.} } \\ 
- -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
- -2008.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\
- -2025.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\
- -2022.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\
- -2026.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\
- -2008.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\
- -2041.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type ?n1.} } \\ 
- -2059.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p2 ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p2 ?n1.} } \\ 
- -2059.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p2 ?n1.} } \\ 
- -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type ?n1.} } \\ 
- -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 rdf:type ?n1.} } \\ 
- -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p1 owl:Class.} } \\ 
- -2020.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p1 owl:Class.} \\\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\
- -2002.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 rdf:type owl:Class.} \\\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\
- -2000.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
- -1996.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} \\\texttt{http://dl-learner.org/carcinogenesis\_bond5500 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d270\_13.} } \\
- -2018.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 ?p1 http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
- -2034.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type ?n1.} } \\ 
- -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2634 ?p2 ?n1.} } \\ 
- -2042.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2634 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond6533 rdf:type ?n1.} } \\ 
- -2048.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 rdf:type ?n1.} } \\ 
- -2014.7 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 rdf:type http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
- -2032.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 ?p1 http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
- -2010.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Molybdenum rdf:type owl:Class.} } \\ 
- -2031.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di260 rdfs:subClassOf ?n1.} } \\ 
- -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 rdf:type ?n1.} } \\ 
- -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-939 ?p2 ?n1.} } \\ 
- -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-939 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond3448 rdf:type ?n1.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2491 rdf:type ?n1.} } \\ 
- -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-3038 rdf:type ?n1.} } \\ 
- -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1155 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1260 ?p2 ?n1.} } \\ 
- -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1260 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-946 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ar\_halide ?p1 owl:Class.} } \\ 
- -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Alcohol ?p1 owl:Class.} } \\ 
- -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Carbon-22 ?p1 owl:Class.} } \\ 
- -2010.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Carbon-22 rdf:type owl:Class.} } \\ 
- -2044.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d243\_16 http://dl-learner.org/carcinogenesis\_charge ?n1.} } \\ 
- -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-216 rdf:type ?n1.} } \\ 
- -2059.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2916 ?p2 ?n1.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2916 rdf:type ?n1.} } \\ 
- -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond3442 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 rdf:type ?n1.} } \\ 
- -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-318 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
- -2042.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_di10-3291 rdf:type ?n1.} } \\ 
- -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-180 rdf:type ?n1.} } \\ 
- -2061.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-180 ?p2 ?n1.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-222 rdf:type ?n1.} } \\ 
- -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Non\_ar\_6c\_ring rdf:type ?n1.} } \\ 
- -2060.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1069 ?p2 ?n1.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1069 rdf:type ?n1.} } \\ 
- -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2275 ?p2 ?n1.} } \\ 
- -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2275 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
- -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-1445 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type ?n1.} } \\ 
- -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2011.7 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} \\\texttt{http://dl-learner.org/carcinogenesis\_bond32 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d10\_6.} } \\
- -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond16 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond662 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond5306 rdf:type ?n1.} } \\ 
- -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond5197 rdf:type ?n1.} } \\ 
- -2020.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3218 ?p1 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
- -2020.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3151 ?p1 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
- -2003.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3151 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
+ \input{results/mutag/motifs-byfreq.latex} 
 \hline
@@
 
-\end{landscape}
 \end{document}

### 2021-04-16 kgmotifs/results/aifb/aifb.sh

@@ -1,0 +1,18 @@
+#!/bin/bash
+#SBATCH --time=72:00:00
+#SBATCH -N 1
+#SBATCH --ntasks-per-node=1
+
+module load java/jre-1.8.0
+
+source /home/pbloem/.bashrc
+
+mkdir -p /local/pbloem
+cp /home/pbloem/java/motive-rdf.jar /local/pbloem
+
+cd /home/pbloem/experiments/kgmotifs/aifb
+
+java -jar /local/pbloem/motive-rdf.jar --experiment real-world --iterations 100000 --topk 100 --data aifb --max-time 5
+wait          # wait until programs are finished
+
+rm -rf /local/pbloem

### 2021-04-16 kgmotifs/results/dogfood/dogfood.sh

@@ -1,0 +1,18 @@
+#!/bin/bash
+#SBATCH --time=72:00:00
+#SBATCH -N 1
+#SBATCH --ntasks-per-node=1
+
+module load java/jre-1.8.0
+
+source /home/pbloem/.bashrc
+
+mkdir -p /local/pbloem
+cp /home/pbloem/java/motive-rdf.jar /local/pbloem
+
+cd /home/pbloem/experiments/kgmotifs/dogfood
+
+java -jar /local/pbloem/motive-rdf.jar --experiment real-world --iterations 100000 --topk 100 --data dogfood --max-time 5
+wait          # wait until programs are finished
+
+rm -rf /local/pbloem

### 2021-04-16 kgmotifs/results/mutag/mutag.sh

@@ -1,0 +1,18 @@
+#!/bin/bash
+#SBATCH --time=72:00:00
+#SBATCH -N 1
+#SBATCH --ntasks-per-node=1
+
+module load java/jre-1.8.0
+
+source /home/pbloem/.bashrc
+
+mkdir -p /local/pbloem
+cp /home/pbloem/java/motive-rdf.jar /local/pbloem
+
+cd /home/pbloem/experiments/kgmotifs/mutag
+
+java -jar /local/pbloem/motive-rdf.jar --experiment real-world --iterations 100000 --topk 100 --data mutag --max-time 5
+wait          # wait until programs are finished
+
+rm -rf /local/pbloem
